### 2021-02-07 working notes.Rmd
Init

---
title: "JI manipulation and protest notes"
output: html_notebook
---
(Another approach to this could make use of Lankina data on Russian protest, with some indicator of judicial independence)
Thinking that the first step is to link NELDA with V-DEM and selected lag variables from V-DEM, plus LJI.

```{r setup}
library(tidyverse)
library(countrycode)
library(readxl)
library(vdemdata)
library(lme4)
library(interplot)
library(lubridate)
```

Loading data

Note: check on how massmob data merges with nelda. We want cases where this is no protest after the election to score as 0.

```{r}
nelda <- read_xls("NELDA.xls")
vdem.data <- vdem
lji <- read.csv("LJI-estimates-20140422.csv")
massmob <- read.csv("mmALL_073119_csv.csv")
```

Cleaning nelda variables--this will have to be done for all used variables
```{r}
nelda <- nelda %>% rename(first.mp.election = nelda2)
nelda <- nelda %>% rename(protests = nelda29)
nelda <- nelda %>% rename(inc.fav.cancel = nelda34)

nelda <- nelda %>% mutate(protests = ifelse(protests == "yes", 1, ifelse(protests == "no", 0, NA)))
nelda <- nelda %>% mutate(inc.fav.cancel = ifelse(inc.fav.cancel == "yes", 1, ifelse(inc.fav.cancel == "no", 0, NA)))
nelda <- nelda %>% mutate(inc.fav.cancel.protests = ifelse(nelda35 == "yes", 1, 0))
nelda <- nelda %>% mutate(opp.fav.cancel = ifelse(nelda32 == "yes", 1, 0))

```

Adding a vdem ID to the other packages

Note: the missing values from NELDA are cases not included in VDEM (e.g. Bahamas, Dominica.)
```{r}
nelda <- nelda %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "gwn", destination = "vdem"))

lji <- lji %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "cown", 
                                               destination = "vdem"))
massmob <- massmob %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "p4n", 
                                               destination = "vdem"))

nelda <- nelda %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
lji <- lji %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
vdem.data <- vdem.data %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
massmob <- massmob %>% mutate(country_id_year = paste(country_id, year, sep = "_"))

```

VDEM variables to lag
v2x_polyarchy, v2x_libdem, v2xpartipdem, v2x_delibdem, v2x_egaldem, v2x_freexp_altinf, v2x_api, v2x_mpi, v2x_frassoc_thick, v2x_suffr, v2xel_frefair, v2x_elecoff, v2x_liberal, v2x_jucon, v2x_partip, v2x_cspart, v2xeg_eqdr, v2psoppaut, v2jureform, v2jupurge, v2jupoatck, v2jupack, v2juaccnt, v2juhcind, v2juncind, v2mecenefm, e_migdppcln, e_migdpgro, e_miinflat

The approach below works for lagging variables by group. Next step is to lag all of the above variables, lag LJI variables, and then merge with nelda.

```{r}
vdem.recent <- vdem.data %>% filter(year >= 1944)

vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(polyarchy.1lag = lag(v2x_polyarchy))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(libdem.1lag = lag(v2x_libdem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(partipdem.1lag = lag(v2x_partipdem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(delibdem.1lag = lag(v2x_delibdem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(egaldem.1lag = lag(v2x_egaldem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(altinf.1lag = lag(v2x_freexp_altinf))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(api.1lag = lag(v2x_api))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(mpi.1lag = lag(v2x_mpi))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(frassoc.1lag = lag(v2x_frassoc_thick))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(suffr.1lag = lag(v2x_suffr))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(frefair.1lag = lag(v2xel_frefair))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(elecoff.1lag = lag(v2x_elecoff))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(liberal.1lag = lag(v2x_liberal))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jucon.1lag = lag(v2x_jucon))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(partip.1lag = lag(v2x_partip))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(cspart.1lag = lag(v2x_cspart))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(eqdr.1lag = lag(v2xeg_eqdr))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(oppaut.1lag = lag(v2psoppaut))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jureform.1lag = lag(v2jureform))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jupurge.1lag = lag(v2jupurge))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jupoatck.1lag = lag(v2jupoatck))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jupack.1lag = lag(v2jupack))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(juaccnt.1lag = lag(v2juaccnt))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(juhcind.1lag = lag(v2juhcind))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(juncind.1lag = lag(v2juncind))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(mecenefm.1lag = lag(v2mecenefm))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(gdppcln.1lag = lag(e_migdppcln))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(gdpgro.1lag = lag(e_migdpgro))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(inflat.1lag = lag(e_miinflat))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(legcon.1lag = lag(v2xlg_legcon))

#test.set.samp <- test.set %>% select(country_id, year, v2x_polyarchy, polyarchy.1lag)
```

Imporant variables to include but not lag: v2eltvrexo, v2elvotlrg, v2elvotsml, v2ellostlg, v2ellostsl, v2ellostsm, v2ellostss, v2ellovtlg, v2ellovtsm

```{r}
lji <- lji %>% group_by(ccode) %>% mutate(lji.lag = lag(LJI))
```

Cleaning nelda date variable and massmob date
```{r}
nelda <- nelda %>% mutate(mmdd = ifelse(mmdd < 1000, paste0(0, mmdd), mmdd))
nelda <- nelda %>% mutate(election_mdy = paste0(mmdd, year))
nelda <- nelda %>% mutate(election_mdy = mdy(election_mdy))

massmob <- massmob %>% mutate(startday = ifelse(startday < 10, paste0(0, startday), startday))
massmob <- massmob %>% mutate(startmonth = ifelse(startmonth < 10, paste0(0, startmonth), startmonth))
massmob <- massmob %>% mutate(protest_mdy = ifelse(is.na(startday) == T | is.na(startmonth) == T, NA,  paste0(startmonth, startday, year)))

massmob <- massmob %>% mutate(protest_mdy = mdy(protest_mdy))
massmob <- massmob %>% filter(is.na(protest_mdy) == F)
```

Merging

```{r}
lji <- lji %>% filter(is.na(country_id) == FALSE)
nelda <- nelda %>% filter(is.na(country_id) == FALSE)
massmob <- massmob %>% filter(is.na(country_id) == FALSE)

nelda2 <- nelda %>% merge(lji %>% dplyr::select(LJI, lji.lag, country_id_year), by = "country_id_year")

nelda2 <- nelda2 %>% merge(vdem.recent %>% dplyr::select(v2eltvrexo, v2elvotlrg, v2elvotsml, v2ellostlg, v2ellostsl, v2ellostsm, v2ellostss, v2ellovtlg, v2ellovtsm, v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, v2x_egaldem, v2x_freexp_altinf, v2x_api, v2x_mpi, v2x_frassoc_thick, v2x_suffr, v2xel_frefair, v2x_elecoff, v2x_liberal, v2x_jucon, v2x_partip, v2x_cspart, v2xeg_eqdr, v2psoppaut, v2jureform, v2jupurge, v2jupoatck, v2jupack, v2juaccnt, v2juhcind, v2juncind, v2mecenefm, e_migdppcln, e_migdpgro, e_miinflat, v2xlg_legcon, legcon.1lag, polyarchy.1lag, libdem.1lag, partipdem.1lag, 
                                                  delibdem.1lag,
egaldem.1lag ,
altinf.1lag,
api.1lag,
mpi.1lag,
frassoc.1lag,
suffr.1lag,
frefair.1lag,
elecoff.1lag,
liberal.1lag,
jucon.1lag,
partip.1lag,
cspart.1lag,
eqdr.1lag,
oppaut.1lag,
jureform.1lag,
jupurge.1lag,
jupoatck.1lag,
jupack.1lag,
juaccnt.1lag,
juhcind.1lag,
juncind.1lag,
mecenefm.1lag,
gdppcln.1lag,
gdpgro.1lag,
inflat.1lag,
v2elvotbuy,
v2elirreg,
v2elintim, v2elfrfair,
                                                  country_id_year), by = "country_id_year")

nelda3 <- nelda2 %>% merge(massmob, by = "country_id_year")
```

Cutting out protests before elections
```{r}
nelda3 <- nelda3 %>% filter(protest_mdy >= election_mdy)
```

##Nelda2 looks at NELDA variables as DV

```{r}
rm(vdem.data)
rm(vdem.recent)
rm(lji)
rm(nelda)
rm(massmob)
```

Correlations

```{r}
df <- nelda2 %>% select(protests, lji.lag, v2xel_frefair)
cor(na.omit(df))
```

These models include LJI and freefair, with controls for largest party vote-share, civil society openness, and economic growth
```{r}
model.base.pres <- lmer(protests ~ lji.lag*v2xel_frefair + v2elvotlrg + cspart.1lag + gdpgro.1lag + (1 | country_id.x),  data = nelda2 %>% filter(opp.fav.cancel == 0))
summary(model.base.pres)
interplot(model.base.pres, var2 = "v2xel_frefair", var1 = "lji.lag")

model.base.leg <- lmer(protests ~ lji.lag*v2xel_frefair + v2ellostlg + cspart.1lag + gdpgro.1lag + (1 | country_id.x), data = nelda2 %>% filter(opp.fav.cancel == 0))
summary(model.base.leg)

interplot(model.base.leg, var2 = "v2xel_frefair", var1 = "lji.lag")

```

Preliminary results show that for legislative elections (the larger dataset), greater judicial independence is associated with a higher risk of protest after manipulated elections.

One thing to consider is that opposition may use protest when there is some judicial independence (i.e., it is not that courts give information to protesters, but that protesters protest to put pressure on elites including judges/justices)

Models below regress favorable results for incumbent being canceled on protests*lji.
These are surprising, and against my first cut hypothesis. I would expect that higher JI would lead to more annulments in the face of protest, but here we see less. In fact, low LJI is when protest is most likely to lead to canceled election results.
```{r}
model.cancel.pres <- lmer(inc.fav.cancel ~ protests*lji.lag + legcon.1lag + v2xel_frefair + v2elvotlrg + cspart.1lag + (1 | country_id.x), data = nelda2 %>% filter(opp.fav.cancel == 0))
summary(model.cancel.pres)
interplot(model.cancel.pres, var1 = "protests", var2 = "lji.lag")
interplot(model.cancel.pres, var2 = "protests", var1 = "lji.lag")

model.cancel.leg <- lmer(inc.fav.cancel ~ protests*lji.lag + legcon.1lag + v2xel_frefair + v2elvotlrg + cspart.1lag +(1 | country_id.x), data = nelda2 %>% filter(opp.fav.cancel == 0))
summary(model.cancel.leg)

interplot(model.cancel.leg, var1 = "protests", var2 = "lji.lag")
interplot(model.cancel.leg, var2 = "protests", var1 = "lji.lag")

```

##Nelda3 looks at MMD variables as DV

```{r}
nelda3 <- nelda3 %>% mutate(protests.postelex = ifelse(protests == 1 & protesterdemand1 == "political behavior, process", 1, 0))
```

Models

```{r}
model.base.pres <- lmer(protests.postelex ~ lji.lag*v2xel_frefair + v2elvotlrg + cspart.1lag + gdpgro.1lag + (1 | country_id.x),  data = nelda3 %>% filter(opp.fav.cancel == 0))
summary(model.base.pres)
interplot(model.base.pres, var2 = "v2xel_frefair", var1 = "lji.lag")

model.base.leg <- lmer(protests.postelex ~ lji.lag*v2xel_frefair + v2ellostlg + cspart.1lag + gdpgro.1lag + (1 | country_id.x), data = nelda3 %>% filter(opp.fav.cancel == 0))
summary(model.base.leg)

interplot(model.base.leg, var2 = "v2xel_frefair", var1 = "lji.lag")
```

```{r}
model.cancel.pres <- lmer(inc.fav.cancel ~ protests*lji.lag + legcon.1lag + v2xel_frefair + v2elvotlrg + cspart.1lag + (1 | country_id.x), data = nelda3 %>% filter(opp.fav.cancel == 0))
summary(model.cancel.pres)
interplot(model.cancel.pres, var1 = "protests", var2 = "lji.lag")
interplot(model.cancel.pres, var2 = "protests", var1 = "lji.lag")

model.cancel.leg <- lmer(inc.fav.cancel ~ protests*lji.lag + legcon.1lag + v2xel_frefair + v2elvotlrg + cspart.1lag +(1 | country_id.x), data = nelda3 %>% filter(opp.fav.cancel == 0))
summary(model.cancel.leg)

interplot(model.cancel.leg, var1 = "protests", var2 = "lji.lag")
interplot(model.cancel.leg, var2 = "protests", var1 = "lji.lag")

```

Should also look into additional DVs: protest size and protest violence

```{r}
model.size.pres <- lmer( ~ lji.lag*v2xel_frefair + v2elvotlrg + cspart.1lag + gdpgro.1lag + (1 | country_id.x),  data = nelda3 %>% filter(opp.fav.cancel == 0))
summary(model.base.pres)
```

### 2022-02-21 Output/manipulation_protest_draft.Rmd
Updates to draft

---
title: "Draft: Judicial independence and post-election protest"
author: "Cole J. Harvey"
date: "`r Sys.Date()'"
output:
    bookdown::pdf_document2: default
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
header-includes: \usepackage{booktabs}
---

# Literature review

Several prominent models of electoral manipulation argue that the risk of mass post-election protest helps restrain incumbents who might otherwise prefer larger levels of manipulation [@magaloni2010; @Little2015; @Little2012; @birch2011; @Fearon2011]. Building on this logic, researchers have proposed that third-party institutions---like election monitors [@Hyde2014; @daxecker2012], civil society groups [@birch2017getting], and courts [@Chernykh2015]---can reduce manipulation and increase protest risk by revealing information about electoral malfeasance. Chernykh and Svolik, in particular, argue that by revealing information about manipulation, even moderately independent courts can increase the risk of protest against manipulated elections (thus deterring manipulation), while also reducing the risk that opposition groups mistakenly protest clean elections. This result is in keeping with the argument that electoral protests are the outcome of bargaining failure among incumbents and opposition groups, which are worsened in low-information environments [@Beaulieu2014a].

Despite courts' connection to one of the most prominent theorized mechanisms for limiting electoral manipulation---protest risk---the relationship between judicial independence and post-election protest is conceptually and empirically muddled. Chernykh [-@chernykh2014] and Eisenstadt [-@eisenstadt_courting_2003] both find no effect for judicial independence on the likelihood of protest. This at least partially calls into question the prediction that more independent courts make protest more likely in the presence of election manipulation.

Judicial independence is understood as courts' ability to come to their own conclusions on the cases before them, without undue influence from outside actors [@Linzer2015]. At the same time that many non-democratic regimes have adopted the institution of (unfairly biased) multiparty elections [@magaloni2010political], many have also made efforts to improve the autonomy of their courts under some circumstances [@Staton2020]. Non-democratic governments do this due to the variety of benefits that independent courts can provide to such regimes. These include enforcing contracts and protecting investments [@Moustafa2008], gathering information on citizen grievances [@rios-figueroa_justice_2017], opposition groups [@Sievert2018], and on lower-level officials [@Ginsburg2008], police rival elite groups [@Ma2017], and---crucially---help legitimate the regime and its other institutions [@moustafa_tamir_law_2008; @Whiting2017].

# Theory

Judicial independence is thought to influence the decision to protest in two broad ways: by affecting citizens' grievances, and by shaping opposition parties' incentives to call their supporters to the streets. Within each of these general effects, at least two channels of influence have been hypothesized. On the strategic side, independent courts may have two effects.  First, greater judicial independence should lead to reduced election manipulation in less competitive settings [@Harvey2022]. Reduced manipulation should result in a smaller margin of victory for incumbents, which in turn should increase the likelihood of protest [@simpser2013; @Brancati2016; @Chernykh2015]. This can be called the margin-of-victory channel.

Second, there is the mechanism proposed by Chernykh and Svolik [-@Chernykh2015], in which independent courts reveal more information about electoral manipulation. Opposition parties use information from the courts to estimate their true popularity, increasing the risk of protest by opposition actors when manipulation is high. This might be called the information-strategy channel. The effects of these channels point in the same direction, though under different circumstances. On the one hand, independent courts should improve opposition parties' electoral results (without conditioning on manipulation level), increasing the odds of opposition protest. On the other hand, when manipulation is high, courts reveal this information to opposition parties, making those parties more likely to challenge election outcomes in the streets. Chernykh and Svolik note that other institutions, like election monitors, can have this same informational effect, so I expect that courts' informational role will be most noticeable when election observers are absent.

On the grievance side, courts also perform two functions that may influence protest likelihood. First, researchers have argued that election manipulation acts as a focal point for citizens' existing grievances with the regime, creating an opportunity for individuals with discrete complaints to register their displeasure at the same time [@tucker2007; @Brancati2016]. In other words, citizens with diverse complaints may be united in the post-electoral moment by information that the election was rigged---a unifying grievance. To the extent that courts reveal information about electoral malfeasance, independent courts may contribute to this sense of post-election grievance---increasing the risk of protest. This effect could be called the information-grievance channel.

However, one of the central functions of independent judiciaries outside of liberal democracies is to create mechanism through which citizens grievances can be addressed [@Whiting2017; @Solomon2007; @Hilbink2007d]. Relatively impartial courts create an opportunity for citizens to resolve grievances against the state and against other individuals who may be more powerful or influential. For example, survey-experimental data show that citizens of several Middle Eastern non-democracies were more willing to engage in legal action to address their grievances when they believe courts are more independent and assertive [@Schaaf2021]. Moreover, the belief that procedures are generally followed impartially can bolster the state's claim to legitimacy [@Gerschewski2013], increasing citizens' trust in authority and willingness to abide by political outcomes [@Grimes2006]. Evidence from European democracies suggests that that de facto independence is key for such judgments [@Buhlmann2011]. Taken together, greater de facto judicial independence ought to reduce protest risk, both by creating a channel for the state to address citizens' general grievances and by offering an alternative venue to the street for addressing specific electoral concerns. I call this last effect the legitimation channel.

![Directed acyclic graph of the relationship between de facto judicial independence (DFJI) and post-election protest](dagitty-model.png){# .class width=75% height=75%}

Figure 1 illustrates the relationship between de facto judicial independence and post-election protest as a directed acyclic graph [@pearl2000causality]. Plus- and minus signs in the figure indicate the theorized direction of specific channels of the relationship. For example, greater judicial independence is predicted to improve the regime's legitimacy, reducing citizens' grievances, which in turn have a positive effect on the risk of protest. Taken together, this channel suggests that greater independence will reduce protest risk by ameliorating grievances. As the figure shows, full effect of judicial independence is multi-faceted, and the the direction and size of overall effect is likely to be dependent on factors that vary across cases and over time. De facto judicial independence may increase the risk of protest by exposing information on manipulation, for example, but this effect is conditional on other sources of such information (e.g. election monitors) and on the broader media environment to broadcast such claims. Court cases may indeed be a very prominent source of information on election integrity for opposition parties, for example, but the widespread presence of election monitors should reduce courts' importance in this regard. Likewise, a closed media environment makes it difficult for citizens to access information about low-quality elections.

The relationship between judicial independence and electoral protest is likely to be unclear, in part, because there are at least four distinct channels through which the former can effect the latter. While three of the four effects ought to make protest more likely, they are predicted to do so under distinct circumstances. Moreover, one of these channels is expected to reduce the likelihood of electoral protest. There should be, as a result, no *a priori* expectation as to the direction of the overall effect. If, for example, the grievance-based channels do most of the work, we might expect judicial independence to reduce the overall risk to protest: reduced manipulation at the front end limits the unifying grievance of a flawed election, while the legitimation channel helps defuse citizens' more diverse grievances. Nevertheless, given the conventional wisdom that independent courts should increase protest risk, I phrase Hypothesis 1 as follows.

>H1 (Overall effect): Without conditioning on the level of manipulation, increased judicial independence is expected to increase the risk of post-election protest.

>H2 (Information-strategic channel): When manipulation is high and alternative sources of information are scarce, increased judicial independence should increase the risk of protest.

>H3 (Margin of victory channel): Controlling for legitimation and grievances, judicial independence should increase the risk of protest.

>H4 (Legitimation channel): Controlling for the level of electoral manipulation, greater judicial independence should reduce the risk of protest when overall grievances are high.

>H5 (Information-grievance channel): Conditional on the level of manipulation, increased judicial independence should increase the risk of protest when greivances are high.

# Data and methods

These hypotheses are tested using data from three datasets: NELDA [@hyde2012], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence (ECAV) dataset [@Daxecker2019]. The unit of observation for this study is an election period as recorded in NELDA, which counts all national elections between 1945 and 2015. This framework prevents selecting on the dependent variable, by capturing both protested and non-protested elections. NELDA also includes a variable indicating elections in which incumbents lost; as this study is focused on protest against ostensibly victorious incumbents, I use this variable to filter out elections which the opposition won. I further limit the sample to regimes coded as electoral autocracies or electoral democracies in V-Dem. This excludes closed autocracies (where no national elections are held) and liberal democracies (where electoral integrity and rule of law are both high). Finally, the date range is limited by the availability of data on protest from ECAV, as discussed below, which runs from 1990 to 2012. The resulting dataset includes 647 election-period observations.

The dependent variable for this study is taken from ECAV, which records protests and other contentious actions from six months before to three months after an election. For this study, I include protests only, excluding violent acts like shootings, bombings, and coups. ECAV data is in event-day format. To capture this variation, I take the sum of all post-election protest-days as the dependent variable. The result is a count variable ranging from 0 to 90, with the large majority of elections going unprotested. To be specific, no protests are recorded after 460 of the 647 elections in the dataset. As a result of this distribution, I use negative binomial regression to model the causes of post-election protest.

Several control variables are included, depending on the hypothesis being tested. First, a binary variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest. To help control for the underlying mobilizational capacity, which likely affects incumbents' decision to manipulate and the latent risk of protest, I include the total number of *pre-election protests*, using ECAV data. Likewise, the ability of the state to engage in repression is likely to embolden the ruling party and deter protests; I thus control for *physical integrity* using the 'physical violence index' from V-Dem lagged one year. Higher values for this variable indicate greater freedom from torture and political killing. Next, all models include two measures of the degree of political openness in the country. The first of these is an indicator variable for *electoral democracy*; the reference category is electoral autocracy. The second is *legislative oversight*, a V-Dem variable which captures the extent to which opposition parties in the legislature exercise oversight on the executive in practice.

Additional controls are used in specific models. The extent of electoral manipulation is captured by *intentional voting irregularities*, a V-Dem variable that measures techniques like voter fraud, ballot stuffing, and falsification of results. In the original coding, lower values of this variable indicate worse manipulation, so I multiply the variable by -1 to avoid this reverse coding. *International observers* is a binary variable indicating the presence of international election monitors. *Media freedom* measures the degree to which the media is able to cover the opposition, criticize the government, and report a variety of perspectives. A final control variable serves as a proxy for citizens' grievances: the *equal distribution of resources* index from V-Dem. This variable reflects the degree to which public funds support public (rather than particularistic) goods, access to the welfare state, educational equality, and health equality. I expect that higher levels of equality translate to reduced social grievances.

Lastly, the explanatory variable for this study if the measure of *latent judicial independence* (LJI) developed by Linzer and Staton [-@Linzer2015]. LJI is a de facto measure of judicial independence that combines data from expert surveys, economic indicators, and components from additional datasets to estimate the degree to which courts are both autonomous and influential.

# Results

```{r setup, echo=FALSE, message = FALSE}
library(tidyverse)
library(countrycode)
library(readxl)
library(vdemdata)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
library(stargazer)

nelda2.sub <- read.csv(here::here("Data", "nelda2_sub.csv"))
nelda3.sub <- read.csv(here::here("Data", "nelda3_sub.csv"))

nelda2.sub <- nelda2.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda2.sub <- nelda2.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))

nelda3.sub <- nelda3.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda3.sub <- nelda3.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))

##Creating economic crisis variable

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = as.factor(econ.crisis))

nelda2.sub <- nelda2.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))

##Removing cases before 1990, when ECAV starts
nelda2.sub <- nelda2.sub %>% filter(year >= 1990)

## Removing cases where inc. loses
nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)

##Intmon as a factor
nelda2.sub <- nelda2.sub %>% mutate(v2elintmon = as.factor(v2elintmon))

```

The results are mixed promising for the general theory that judicial independence, by raising the risk of protest, can help deter election manipulation. Models 1 and 3, where latent judicial independence is a standalone variable, show no statistically significant effect on protest. While this shows that Hypothesis 1 is not supported, the result is in keeping with the theoretical discussion above. The existence of multiple competing channels through which independent courts might affect protest risk means that only in specific circumstances are we likely to see consistently positive or negative effects. Likewise, Hypothesis 3 is not supported; the effect is negative, as predicted, but not statistically significant. It may be the case that, as Harvey [-@Harvey2022] suggests, the margin of victory channel substantially increases protest risk only under distinct circumstances (such as when an incumbent president is not running for re-election).

```{r, results = FALSE, echo=FALSE, warning=FALSE}
##Hypothesis 1
model.ecav.base <- glm.nb(n.events.post  ~ lji.lag  + presidential + physinteg.1lag +  v2xlg_legcon  +  n.events.pre + as.factor(v2x_regime),  data = nelda2.sub)
#summary(model.ecav.base)
```
```{r,  echo=FALSE, warning=FALSE}
# Hypothesis 2
model.ecav.base2 <- glm.nb(n.events.post  ~ lji.lag*v2elintmon + v2elirreg.inv + presidential  + physinteg.1lag +   v2xlg_legcon  +  n.events.pre + as.factor(v2x_regime),  data = nelda2.sub %>% filter(inc.lose == 0))
#summary(model.ecav.base2)
sjPlot::plot_model(model.ecav.base2, type = "pred", terms =  c("v2elintmon", "lji.lag [.1, .75]"))

```

```{r, results = FALSE, echo=FALSE, warning=FALSE}
##Hypothesis 3
model.ecav.marg <- glm.nb(n.events.post  ~ lji.lag*v2xlg_legcon  +  legit_total + v2xeg_eqdr + presidential + physinteg.1lag +  v2xlg_legcon  +  n.events.pre + as.factor(v2x_regime),  data = nelda2.sub)
#summary(model.ecav.marg)
sjPlot::plot_model(model.ecav.marg, type = "pred", terms =  c("v2xlg_legcon", "lji.lag [.1, .75]"))

```

```{r, echo=FALSE, warning=FALSE}
# Hypothesis 4

model.ecav.griev <- glm.nb(n.events.post  ~ lji.lag*v2xeg_eqdr + v2elirreg.inv +  presidential  + physinteg.1lag +  v2xlg_legcon  +  n.events.pre + as.factor(v2x_regime),  data = nelda2.sub %>% filter(inc.lose == 0))
#summary(model.ecav.griev)
sjPlot::plot_model(model.ecav.griev, type = "pred", terms =  c("v2xeg_eqdr", "lji.lag [.1, .7]")) + coord_cartesian(ylim = c(0, 2))

```

```{r,  echo=FALSE, warning=FALSE}
##Hypothesis 5

model.ecav.infog <- glm.nb(n.events.post  ~ lji.lag*v2xeg_eqdr + v2elirreg.inv + presidential + physinteg.1lag +  v2xlg_legcon  +  n.events.pre + as.factor(v2x_regime),  data = nelda2.sub)
#summary(model.ecav.infog)
sjPlot::plot_model(model.ecav.infog, type = "pred", terms =  c("v2xeg_eqdr", "lji.lag [.1, .75]")) + coord_cartesian(ylim = c(0,3))

model.ecav.infog2 <- glm.nb(n.events.post  ~ lji.lag*v2xeg_eqdr*v2elintmon + v2elirreg.inv + presidential + physinteg.1lag +  v2xlg_legcon  +  n.events.pre + as.factor(v2x_regime),  data = nelda2.sub)
#summary(model.ecav.infog)
p.infog2 <- sjPlot::plot_model(model.ecav.infog2, type = "pred", terms =  c("v2xeg_eqdr", "lji.lag [.1, .75]", "v2elintmon"))
p.infog2 <- p.infog2 + coord_cartesian(ylim = c(0,3))

```

```{r, results='asis'}
stargazer(model.ecav.base, model.ecav.base2, model.ecav.marg,  header=FALSE, type='latex', omit.stat = "theta")
```

```{r, results='asis'}
stargazer(model.ecav.griev, model.ecav.infog,  header=FALSE, type='latex', omit.stat = "theta", no.space = T)
```

Hypothesis 5 predicts that high levels of judicial independence will increase the risk of protest when pre-existing grievances are high, holding the level of manipulation constant. This hypothesis is supported only in the most unequal societies. In socities with the most unequal access to resources, a high level of judicial independence is associated with two or three post-election protests, while low judicial independence predicts approximately zero.

Hypothesis 4, on the legitimacy channel, is also weakly supported. When judicial independence is low, the risk of protest is essentially constant across ordinary economic times and crises. When judicial independence is high, however, the risk of protest is lower in crisis times than in ordinary times. The strength of this evidence is relatively modest, however, since conditional on an economic crisis, judicial independence has no significant effect.

Finally, Hypothesis 2---the information-strategic channel---is supported. When information on election integrity is otherwise scarce due to the absence of credible monitors, a country with more independent courts could expect three post-election protests. In the same circumstances with less independent courts, the predicted number of protests is zero.

```{r, include=FALSE}
## Density plot

p.dens <- ggplot(nelda2.sub, aes(x=n.events.post)) + 
  geom_density()
p.dens
```

# Discussion

These findings make two primary contributions to the study of election integrity and protest. Most convincingly, they help validate the assumption that more independent courts can make protest more likely by revealing information. This assumption was not tested empirically in the work by Chernykh and Svolik [-@Chernykh2015]; these results lend credence to their proposed mechanism. However, and in keeping with the broader implications of this study, independent courts only serve this role in relatively narrow circumstances.

```{r}
model.margin <- lm(winner.margin ~ lji.lag + v2xlg_legcon + physinteg.1lag + v2x_regime + gdpgro.1lag, data = nelda2.sub)
summary(model.margin)

model.manip <- lm(v2elirreg.inv ~ lji.lag + v2xlg_legcon + physinteg.1lag + v2x_regime + gdpgro.1lag, data = nelda2.sub)
summary(model.manip)
```

Code for dagitty web browser

dag {
bb="0,0,1,1"
"Citizen grievances" [pos="0.586,0.365"]
"Economic problems" [pos="0.586,0.246"]
"Incumbent share" [pos="0.221,0.523"]
"Information on manipulation" [pos="0.337,0.361"]
"Manipulation level" [pos="0.100,0.359"]
"Media freedom" [pos="0.375,0.245"]
"Political openness" [adjusted,pos="0.097,0.123"]
"Regime legitimation" [pos="0.589,0.542"]
"Strategic choice to protest" [pos="0.345,0.638"]
DFJI [exposure,pos="0.181,0.208"]
Monitors [pos="0.293,0.246"]
Protest [outcome,pos="0.669,0.771"]
Repression [pos="0.103,0.772"]
"Citizen grievances" -> "Regime legitimation"
"Economic problems" -> "Citizen grievances"
"Economic problems" -> "Media freedom"
"Incumbent share" -> "Strategic choice to protest"
"Information on manipulation" -> "Citizen grievances"
"Information on manipulation" -> "Strategic choice to protest"
"Manipulation level" -> "Incumbent share"
"Manipulation level" -> "Information on manipulation"
"Media freedom" -> "Information on manipulation"
"Political openness" -> "Manipulation level"
"Political openness" -> "Media freedom" [pos="0.298,0.092"]
"Political openness" -> DFJI
"Regime legitimation" -> "Incumbent share"
"Regime legitimation" -> Protest
"Strategic choice to protest" -> Protest
DFJI -> "Information on manipulation"
DFJI -> "Manipulation level"
DFJI -> "Regime legitimation" [pos="0.203,0.517"]
Monitors -> "Information on manipulation"
Repression -> "Manipulation level"
Repression -> Protest
}

# References

### 2022-02-21 Scripts/data_cleaning.Rmd
Updates to draft

---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(countrycode)
library(readxl)
library(vdemdata)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
```

Loading data

Note: check on how massmob data merges with nelda. We want cases where this is no protest after the election to score as 0.

```{r}
nelda <- read_xls(here::here("Data", "NELDA.xls"))
ecav <- read_xls(here::here("Data", "ECAV datatset_Version 1.2.xls"))
vdem.data <- vdem #At the time of coding, this was v. 11.1
lji <- read.csv(here::here("Data", "LJI-estimates-20140422.csv"))
massmob <- read.csv(here::here("Data", "mmALL_073119_csv.csv"))
```

Cleaning nelda variables--this will have to be done for all used variables
```{r}
nelda <- nelda %>% rename(first.mp.election = nelda2)
nelda <- nelda %>% rename(protests = nelda29)
nelda <- nelda %>% rename(inc.fav.cancel = nelda34)
nelda <- nelda %>% rename(inc.lose = nelda24)

nelda <- nelda %>% mutate(protests = ifelse(protests == "yes", 1, ifelse(protests == "no", 0, NA)))
nelda <- nelda %>% mutate(inc.fav.cancel = ifelse(inc.fav.cancel == "yes", 1, ifelse(inc.fav.cancel == "no", 0, NA)))
nelda <- nelda %>% mutate(inc.fav.cancel.protests = ifelse(nelda35 == "yes", 1, 0))
nelda <- nelda %>% mutate(opp.fav.cancel = ifelse(nelda32 == "yes", 1, 0))
nelda <- nelda %>% mutate(inc.lose = ifelse(inc.lose == "yes", 1, ifelse(inc.lose == "no", 0, NA)))

```

Adding a vdem ID to the other packages

Note: the missing values from NELDA are cases not included in VDEM (e.g. Bahamas, Dominica.)
```{r}
nelda <- nelda %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "gwn", destination = "vdem"))

lji <- lji %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "cown", 
                                               destination = "vdem"))
massmob <- massmob %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "p4n", 
                                               destination = "vdem"))

nelda <- nelda %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
lji <- lji %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
vdem.data <- vdem.data %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
massmob <- massmob %>% mutate(country_id_year = paste(country_id, year, sep = "_"))

```

VDEM variables to lag
v2x_polyarchy, v2x_libdem, v2xpartipdem, v2x_delibdem, v2x_egaldem, v2x_freexp_altinf, v2x_api, v2x_mpi, v2x_frassoc_thick, v2x_suffr, v2xel_frefair, v2x_elecoff, v2x_liberal, v2x_jucon, v2x_partip, v2x_cspart, v2xeg_eqdr, v2psoppaut, v2jureform, v2jupurge, v2jupoatck, v2jupack, v2juaccnt, v2juhcind, v2juncind, v2mecenefm, e_migdppcln, e_migdpgro, e_miinflat

The approach below works for lagging variables by group. Next step is to lag all of the above variables, lag LJI variables, and then merge with nelda.

```{r}
vdem.recent <- vdem.data %>% filter(year >= 1944)

vdem.recent <- vdem.recent %>% mutate(elcompindex.m = ifelse(v2xel_frefair == 0, 0, v2x_mpi / v2xel_frefair))
vdem.recent$v2elirreg.inv <- vdem.recent$v2elirreg * -1

vdem.recent$reform_positive <- NA
vdem.recent$reform_positive[vdem.recent$v2jureform_ord <= 1] <- 0
vdem.recent$reform_positive[vdem.recent$v2jureform_ord == 2 ] <- 1

vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(v2elintmon.lag = lag(v2elintmon))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(reform_positive.lag = lag(reform_positive))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(v2elirreg.inv.lag = lag(v2elirreg.inv))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(elcompindex.m.1lag = lag(elcompindex.m))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(polyarchy.1lag = lag(v2x_polyarchy))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(libdem.1lag = lag(v2x_libdem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(partipdem.1lag = lag(v2x_partipdem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(delibdem.1lag = lag(v2x_delibdem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(egaldem.1lag = lag(v2x_egaldem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(altinf.1lag = lag(v2x_freexp_altinf))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(api.1lag = lag(v2x_api))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(mpi.1lag = lag(v2x_mpi))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(frassoc.1lag = lag(v2x_frassoc_thick))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(suffr.1lag = lag(v2x_suffr))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(frefair.1lag = lag(v2xel_frefair))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(elecoff.1lag = lag(v2x_elecoff))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(liberal.1lag = lag(v2x_liberal))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jucon.1lag = lag(v2x_jucon))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(partip.1lag = lag(v2x_partip))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(cspart.1lag = lag(v2x_cspart))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(eqdr.1lag = lag(v2xeg_eqdr))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(oppaut.1lag = lag(v2psoppaut))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jureform.1lag = lag(v2jureform))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jupurge.1lag = lag(v2jupurge))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jupoatck.1lag = lag(v2jupoatck))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jupack.1lag = lag(v2jupack))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(juaccnt.1lag = lag(v2juaccnt))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(juhcind.1lag = lag(v2juhcind))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(juncind.1lag = lag(v2juncind))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(mecenefm.1lag = lag(v2mecenefm))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(gdppcln.1lag = lag(e_migdppcln))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(gdpgro.1lag = lag(e_migdpgro))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(inflat.1lag = lag(e_miinflat))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(legcon.1lag = lag(v2xlg_legcon))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(physinteg.1lag = lag(v2x_clphy))
vdem.recent <- vdem.recent %>% group_by(COWcode) %>% mutate(opposition.oversight.lag = lag(v2lgoppart))

##Leading variables for protest after election years
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(prodem.protest.1lead = lead(v2cademmob))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(proauth.protest.1lead = lead(v2caautmob))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(protest.scale.1lead = lead(v2cagenmob))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(concen.protest.1lead = lead(v2caconmob))

```
Subtracting election integrity variable from additive polyarchy index and judicial constraints from liberal component index
```{r}
vdem.recent <- vdem.recent %>% mutate(polyarchy_no_ei = v2x_api - (.25*v2xel_frefair))

vdem.recent <- vdem.recent %>% mutate(liberal_no_jucon = (v2xcl_rol + v2xlg_legcon)/2)
```

Imporant variables to include but not lag: v2eltvrexo, v2elvotlrg, v2elvotsml, v2ellostlg, v2ellostsl, v2ellostsm, v2ellostss, v2ellovtlg, v2ellovtsm

```{r}
lji <- lji %>% group_by(ccode) %>% mutate(lji.lag = lag(LJI))
```

Cleaning nelda date variable and massmob date
```{r}
nelda <- nelda %>% mutate(mmdd = ifelse(mmdd < 1000, paste0(0, mmdd), mmdd))
nelda <- nelda %>% mutate(election_mdy = paste0(mmdd, year))
nelda <- nelda %>% mutate(election_mdy = mdy(election_mdy))

massmob <- massmob %>% mutate(startday = ifelse(startday < 10, paste0(0, startday), startday))
massmob <- massmob %>% mutate(startmonth = ifelse(startmonth < 10, paste0(0, startmonth), startmonth))
massmob <- massmob %>% mutate(protest_mdy = ifelse(is.na(startday) == T | is.na(startmonth) == T, NA,  paste0(startmonth, startday, year)))

massmob <- massmob %>% mutate(protest_mdy = mdy(protest_mdy))
massmob <- massmob %>% filter(is.na(protest_mdy) == F)
```

Merging

```{r}
lji <- lji %>% filter(is.na(country_id) == FALSE)
nelda <- nelda %>% filter(is.na(country_id) == FALSE)
massmob <- massmob %>% filter(is.na(country_id) == FALSE)

nelda2 <- nelda %>% merge(lji %>% dplyr::select(LJI, lji.lag, country_id_year), by = "country_id_year")

nelda2 <- nelda2 %>% merge(vdem.recent %>% dplyr::select(v2elintmon.lag, v2elintmon, v2eltvrexo, v2elvotlrg, v2elvotsml, v2ellostlg, v2ellostsl, v2ellostsm, v2ellostss, v2ellovtlg, v2ellovtsm, v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, v2x_egaldem, v2x_freexp_altinf, v2x_api, v2x_mpi, v2x_frassoc_thick, v2x_suffr, v2xel_frefair, v2x_elecoff, v2x_liberal, v2x_jucon, v2x_partip, v2x_cspart, v2xeg_eqdr, v2psoppaut, v2jureform, v2jupurge, v2jupoatck, v2jupack, v2juaccnt, v2juhcind, v2juncind, v2mecenefm, e_migdppcln, e_migdpgro, e_miinflat, v2xlg_legcon, legcon.1lag, polyarchy.1lag, libdem.1lag, partipdem.1lag, 
                                                  delibdem.1lag,
                                                  elcompindex.m.1lag,
                                                  opposition.oversight.lag,
                                                  v2elirreg.inv,
                                                  reform_positive.lag,
                                                  e_mipopula,
                                                  e_miurbani,
                                                  e_miurbpop,
                                                  e_wb_pop, #Population
                                                  v2xps_party, #Party institutionalization
                                                  #v2regoppgroupssize, #Opp group size (categorical)
                                                  v2pscohesv, #Legislative party cohesion
                                                  v2exl_legitratio, 
                                                  v2exl_legitperf,
                                                  v2exl_legitlead,
                                                  v2exl_legitideol,
                                                  v2xnp_regcorr,
                                                  v2xeg_eqprotec,
                                                  v2xeg_eqdr,
                                                  v2xeg_eqaccess,

egaldem.1lag ,
altinf.1lag,
api.1lag,
mpi.1lag,
frassoc.1lag,
suffr.1lag,
frefair.1lag,
elecoff.1lag,
liberal.1lag,
jucon.1lag,
partip.1lag,
cspart.1lag,
eqdr.1lag,
oppaut.1lag,
jureform.1lag,
jupurge.1lag,
jupoatck.1lag,
jupack.1lag,
juaccnt.1lag,
juhcind.1lag,
juncind.1lag,
mecenefm.1lag,
gdppcln.1lag,
gdpgro.1lag,
inflat.1lag,
v2elvotbuy,
v2elirreg,
v2elintim, v2elfrfair,
physinteg.1lag, v2x_clphy,
polyarchy_no_ei,
liberal_no_jucon, v2elaccept,
v2cademmob, #pro-democracy mass mob
v2caautmob, #pro-autoc mass mob
v2cagenmob, #scale of mobilization
v2caconmob, #concentration of mobilization in the capital
v2x_regime,
prodem.protest.1lead,
proauth.protest.1lead,
protest.scale.1lead,
concen.protest.1lead,
country_id_year), by = "country_id_year")

nelda2 <- nelda2 %>% mutate(winner.margin = ifelse(types == "Executive", v2elvotlrg/100, v2ellovtlg/100))
nelda2 <- nelda2 %>% mutate(presidential = ifelse(types == "Executive", 1, 0))

nelda3 <- nelda2 %>% merge(massmob, by = "country_id_year")
```

Cutting out protests before elections for nelda 3, 
```{r}
nelda3 <- nelda3 %>% filter(protest_mdy >= election_mdy)
```

##Nelda2 looks at NELDA variables as DV

```{r}
rm(vdem.data)
rm(vdem.recent)
rm(lji)
rm(nelda)
rm(massmob)
```

```{r}
nelda2.sub <- nelda2 %>% filter(v2x_regime == 1 | v2x_regime == 2)
```

```{r}
ecav.pre <- ecav %>% filter(Date < Electiondate)
ecav.pre <- ecav.pre %>% filter(str_detect(EventName, "rotest") == T) #Only captures protests

ecav <- ecav %>% filter(Date >= Electiondate)
ecav <- ecav %>% filter(str_detect(EventName, "rotest") == T) #Only captures protests
```

Grouped summaries
```{r}
count_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(n.events.pre = n())
count_data_post <- ecav %>% group_by(NeldaID) %>% summarize(n.events.post = n())

##Duration
max_date <- ecav %>% group_by(NeldaID) %>% summarize(latest.event = max(Date))
min_date <- ecav %>% group_by(NeldaID) %>% summarize(earliest.event = min(Date))
duration_data <- left_join(min_date, max_date, by = "NeldaID")
duration_data <- duration_data %>% mutate(d.events = as.duration(interval(earliest.event, latest.event)))
duration_data <- duration_data %>% mutate(d.events.num = as.numeric(d.events, "days"))
duration_data <- duration_data %>% mutate(d.events.num = ifelse(d.events.num == 0, 1, d.events.num)) #1-day events

##Size
ecav <- ecav %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
size_data <- ecav %>% group_by(NeldaID) %>% summarize(median.participants = median(ParticipantNumber))

ecav.pre <- ecav.pre %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
size_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(median.participants.pre = median(ParticipantNumber))

##Combined
ecav_data <- left_join(count_data_post, duration_data, by = "NeldaID")
ecav_data <- left_join(ecav_data, size_data, by = "NeldaID")

ecav_data_pre <- left_join(count_data_pre, size_data_pre, by = "NeldaID")
ecav_data <- left_join(ecav_data, ecav_data_pre, by = "NeldaID")

## Merge back with NELDA 2
```

Merge ECAV with NELDA

```{r}
nelda2.sub <- nelda2.sub %>% rename(NeldaID = electionid)
nelda2.sub <- left_join(nelda2.sub, ecav_data, by = "NeldaID")

##Cleaning

nelda2.sub <- nelda2.sub %>% mutate(n.events.post = ifelse(is.na(n.events.post)==T, 0, n.events.post))
nelda2.sub <- nelda2.sub %>% mutate(d.events.num = ifelse(is.na(d.events.num)==T, 0, d.events.num))
nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(is.na(median.participants)==T, 0, median.participants))

nelda2.sub <- nelda2.sub %>% mutate(any.ecav.protest = ifelse(n.events.post >= 1, 1, 0))

##Making the assumption that NAs for pre-election protest are zero
nelda2.sub <- nelda2.sub %>% mutate(n.events.pre = ifelse(is.na(n.events.pre) == T, 0, n.events.pre))

```

```{r}
nelda3 <- nelda3 %>% mutate(protests.postelex = ifelse(protests == 1 & protesterdemand1 == "political behavior, process", 1, 0))

nelda3 <- nelda3 %>% mutate(protests.postelex.quant = ifelse(protests.postelex == 1, protestnumber, 0))
```

Cutting out closed autocracies and liberal democracies

```{r}
nelda3.sub <- nelda3 %>% filter(v2x_regime == 1 | v2x_regime == 2)
```

## Writing

```{r}
write.csv(nelda2.sub, here::here("Data", "nelda2_sub.csv"))
```

```{r}
write.csv(nelda3.sub, here::here("Data", "nelda3_sub.csv"))

```

### 2022-02-21 working notes.Rmd
Updates to draft

@@
 library(lubridate)
+library(summarytools)
 ```
@@
 ```{r}
-nelda <- read_xls("NELDA.xls")
+nelda <- read_xls("./Data/NELDA.xls")
+ecav <- read_xls("./Data/ECAV datatset_Version 1.2.xls")
 vdem.data <- vdem
-lji <- read.csv("LJI-estimates-20140422.csv")
-massmob <- read.csv("mmALL_073119_csv.csv")
+lji <- read.csv("./Data/LJI-estimates-20140422.csv")
+massmob <- read.csv("./Data/mmALL_073119_csv.csv")
 ```
@@
 
+vdem.recent <- vdem.recent %>% mutate(elcompindex.m = ifelse(v2xel_frefair == 0, 0, v2x_mpi / v2xel_frefair))
+vdem.recent$v2elirreg.inv <- vdem.recent$v2elirreg * -1
+
+vdem.recent$reform_positive <- NA
+vdem.recent$reform_positive[vdem.recent$v2jureform_ord <= 1] <- 0
+vdem.recent$reform_positive[vdem.recent$v2jureform_ord == 2 ] <- 1
+
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(v2elintmon.lag = lag(v2elintmon))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(reform_positive.lag = lag(reform_positive))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(v2elirreg.inv.lag = lag(v2elirreg.inv))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(elcompindex.m.1lag = lag(elcompindex.m))
 vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(polyarchy.1lag = lag(v2x_polyarchy))
@@
 vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(legcon.1lag = lag(v2xlg_legcon))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(physinteg.1lag = lag(v2x_clphy))
+vdem.recent <- vdem.recent %>% group_by(COWcode) %>% mutate(opposition.oversight.lag = lag(v2lgoppart))
 
-#test.set.samp <- test.set %>% select(country_id, year, v2x_polyarchy, polyarchy.1lag)
+##Leading variables for protest after election years
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(prodem.protest.1lead = lead(v2cademmob))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(proauth.protest.1lead = lead(v2caautmob))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(protest.scale.1lead = lead(v2cagenmob))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(concen.protest.1lead = lead(v2caconmob))
+
+```
+Subtracting election integrity variable from additive polyarchy index and judicial constraints from liberal component index
+```{r}
+vdem.recent <- vdem.recent %>% mutate(polyarchy_no_ei = v2x_api - (.25*v2xel_frefair))
+
+vdem.recent <- vdem.recent %>% mutate(liberal_no_jucon = (v2xcl_rol + v2xlg_legcon)/2)
 ```
@@
 
-nelda2 <- nelda2 %>% merge(vdem.recent %>% dplyr::select(v2eltvrexo, v2elvotlrg, v2elvotsml, v2ellostlg, v2ellostsl, v2ellostsm, v2ellostss, v2ellovtlg, v2ellovtsm, v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, v2x_egaldem, v2x_freexp_altinf, v2x_api, v2x_mpi, v2x_frassoc_thick, v2x_suffr, v2xel_frefair, v2x_elecoff, v2x_liberal, v2x_jucon, v2x_partip, v2x_cspart, v2xeg_eqdr, v2psoppaut, v2jureform, v2jupurge, v2jupoatck, v2jupack, v2juaccnt, v2juhcind, v2juncind, v2mecenefm, e_migdppcln, e_migdpgro, e_miinflat, v2xlg_legcon, legcon.1lag, polyarchy.1lag, libdem.1lag, partipdem.1lag,
+nelda2 <- nelda2 %>% merge(vdem.recent %>% dplyr::select(v2elintmon.lag, v2elintmon, v2eltvrexo, v2elvotlrg, v2elvotsml, v2ellostlg, v2ellostsl, v2ellostsm, v2ellostss, v2ellovtlg, v2ellovtsm, v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, v2x_egaldem, v2x_freexp_altinf, v2x_api, v2x_mpi, v2x_frassoc_thick, v2x_suffr, v2xel_frefair, v2x_elecoff, v2x_liberal, v2x_jucon, v2x_partip, v2x_cspart, v2xeg_eqdr, v2psoppaut, v2jureform, v2jupurge, v2jupoatck, v2jupack, v2juaccnt, v2juhcind, v2juncind, v2mecenefm, e_migdppcln, e_migdpgro, e_miinflat, v2xlg_legcon, legcon.1lag, polyarchy.1lag, libdem.1lag, partipdem.1lag,
                                                   delibdem.1lag,
+                                                  elcompindex.m.1lag,
+                                                  opposition.oversight.lag,
+                                                  v2elirreg.inv,
+                                                  reform_positive.lag,
+                                                  e_mipopula,
+                                                  e_miurbani,
+                                                  e_miurbpop,
+                                                  e_wb_pop, #Population
+                                                  v2xps_party, #Party institutionalization
+                                                  #v2regoppgroupssize, #Opp group size (categorical)
+                                                  v2pscohesv, #Legislative party cohesion
+
 egaldem.1lag ,
@@
 v2elintim, v2elfrfair,
+physinteg.1lag, v2x_clphy,
+polyarchy_no_ei,
+liberal_no_jucon, v2elaccept,
+v2cademmob, #pro-democracy mass mob
+v2caautmob, #pro-autoc mass mob
+v2cagenmob, #scale of mobilization
+v2caconmob, #concentration of mobilization in the capital
+v2x_regime,
+prodem.protest.1lead,
+proauth.protest.1lead,
+protest.scale.1lead,
+concen.protest.1lead,
 country_id_year), by = "country_id_year")
+
+nelda2 <- nelda2 %>% mutate(winner.margin = ifelse(types == "Executive", v2elvotlrg/100, v2ellovtlg/100))
+nelda2 <- nelda2 %>% mutate(presidential = ifelse(types == "Executive", 1, 0))
 
@@
 
-Cutting out protests before elections
+Cutting out protests before elections for nelda 3, 
 ```{r}
@@
 
-Correlations
+##Data exploration
 
 ```{r}
-df <- nelda2 %>% select(protests, lji.lag, v2xel_frefair)
-cor(na.omit(df))
+nelda2.small <- nelda2 %>% dplyr::select(protests, lji.lag, v2xel_frefair, v2elvotlrg, v2ellostlg, cspart.1lag, gdpgro.1lag, country_id.x, opp.fav.cancel, winner.margin, presidential, inc.fav.cancel, first.mp.election, altinf.1lag, v2elvotbuy, v2elirreg, v2elintim, frassoc.1lag, physinteg.1lag, v2x_clphy, polyarchy_no_ei,
+liberal_no_jucon, v2elaccept, v2cademmob, #pro-democracy mass mob
+v2caautmob, #pro-autoc mass mob
+v2cagenmob, #scale of mobilization
+v2caconmob, #concentration of mobilization in the capital
+v2x_regime,
+prodem.protest.1lead,
+proauth.protest.1lead,
+protest.scale.1lead,
+concen.protest.1lead)
+
+view(dfSummary(nelda2.small))
 ```
 
-These models include LJI and freefair, with controls for largest party vote-share, civil society openness, and economic growth
+##Correlations
+
+Show that protests are negatively associated with LJI and with free and fair elections, and with less vote-buying, intimidation, and fraud (note these are reverse coded; higher values means less of the thing)
+
 ```{r}
-model.base.pres <- lmer(protests ~ lji.lag*v2xel_frefair + v2elvotlrg + cspart.1lag + gdpgro.1lag + (1 | country_id.x),  data = nelda2 %>% filter(opp.fav.cancel == 0))
-summary(model.base.pres)
-interplot(model.base.pres, var2 = "v2xel_frefair", var1 = "lji.lag")
+DataExplorer::plot_correlation(nelda2.small, cor_args = list("use" = "pairwise.complete.obs"))
+```
 
-model.base.leg <- lmer(protests ~ lji.lag*v2xel_frefair + v2ellostlg + cspart.1lag + gdpgro.1lag + (1 | country_id.x), data = nelda2 %>% filter(opp.fav.cancel == 0))
-summary(model.base.leg)
+##Models
 
-interplot(model.base.leg, var2 = "v2xel_frefair", var1 = "lji.lag")
+```{r}
+nelda2.sub <- nelda2 %>% filter(v2x_regime == 1 | v2x_regime == 2)
+```
+
+## ECAV
+
+Getting events after the election, applying other filters
+
+```{r}
+ecav.pre <- ecav %>% filter(Date < Electiondate)
+ecav.pre <- ecav.pre %>% filter(str_detect(EventName, "rotest") == T) #Only captures protests
+
+ecav <- ecav %>% filter(Date >= Electiondate)
+ecav <- ecav %>% filter(str_detect(EventName, "rotest") == T) #Only captures protests
+```
+
+Grouped summaries
+```{r}
+count_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(n.events.pre = n())
+count_data_post <- ecav %>% group_by(NeldaID) %>% summarize(n.events.post = n())
+
+##Duration
+max_date <- ecav %>% group_by(NeldaID) %>% summarize(latest.event = max(Date))
+min_date <- ecav %>% group_by(NeldaID) %>% summarize(earliest.event = min(Date))
+duration_data <- left_join(min_date, max_date, by = "NeldaID")
+duration_data <- duration_data %>% mutate(d.events = as.duration(interval(earliest.event, latest.event)))
+duration_data <- duration_data %>% mutate(d.events.num = as.numeric(d.events, "days"))
+duration_data <- duration_data %>% mutate(d.events.num = ifelse(d.events.num == 0, 1, d.events.num)) #1-day events
+
+##Size
+ecav <- ecav %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
+size_data <- ecav %>% group_by(NeldaID) %>% summarize(median.participants = median(ParticipantNumber))
+
+ecav.pre <- ecav.pre %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
+size_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(median.participants.pre = median(ParticipantNumber))
+
+##Combined
+ecav_data <- left_join(count_data_post, duration_data, by = "NeldaID")
+ecav_data <- left_join(ecav_data, size_data, by = "NeldaID")
+
+ecav_data_pre <- left_join(count_data_pre, size_data_pre, by = "NeldaID")
+ecav_data <- left_join(ecav_data, ecav_data_pre, by = "NeldaID")
+
+## Merge back with NELDA 2
+```
+
+Merge ECAV with NELDA
+
+```{r}
+nelda2.sub <- nelda2.sub %>% rename(NeldaID = electionid)
+nelda2.sub <- left_join(nelda2.sub, ecav_data, by = "NeldaID")
+
+##Cleaning
+
+nelda2.sub <- nelda2.sub %>% mutate(n.events.post = ifelse(is.na(n.events.post)==T, 0, n.events.post))
+nelda2.sub <- nelda2.sub %>% mutate(d.events.num = ifelse(is.na(d.events.num)==T, 0, d.events.num))
+nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(is.na(median.participants)==T, 0, median.participants))
+
+nelda2.sub <- nelda2.sub %>% mutate(any.ecav.protest = ifelse(n.events.post >= 1, 1, 0))
+
+##Making the assumption that NAs for pre-election protest are zero
+nelda2.sub <- nelda2.sub %>% mutate(n.events.pre = ifelse(is.na(n.events.pre) == T, 0, n.events.pre))
 
@@
 
-Preliminary results show that for legislative elections (the larger dataset), greater judicial independence is associated with a higher risk of protest after manipulated elections.
+Models
 
-One thing to consider is that opposition may use protest when there is some judicial independence (i.e., it is not that courts give information to protesters, but that protesters protest to put pressure on elites including judges/justices)
+Test dispersion
 
+```{r}
+modelnb <- MASS::glm.nb(d.events.num ~ lji.lag*v2xel_frefair*altinf.1lag + v2elintmon + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag +  liberal_no_jucon + e_wb_pop + e_migdppcln + n.events.pre + v2psoppaut + as.factor(v2x_regime),  data = nelda2.sub %>% filter(opp.fav.cancel == 0),
+trace=TRUE)
+pscl::odTest(modelnb)
+```
+
+```{r}
+
+model.evac.num <- glm.nb(n.events.post  ~ lji.lag*v2xel_frefair + altinf.1lag + v2elintmon + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag +  liberal_no_jucon + e_wb_pop + e_migdppcln + n.events.pre + v2pscohesv+ #v2psoppaut +
+                             as.factor(v2x_regime),  data = nelda2.sub %>% filter(opp.fav.cancel == 0))
+summary(model.evac.num)
+sjPlot::plot_model(model.evac.num, type = "pred", terms =  c("v2xel_frefair", "lji.lag [.17, .61]"))
+
+model.evac.dur <- glm.nb(d.events.num ~ lji.lag*v2xel_frefair + altinf.1lag + v2elintmon + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag +  liberal_no_jucon + e_wb_pop + e_migdppcln + v2pscohesv+ #v2psoppaut + #n.events.pre +
+                           as.factor(v2x_regime),  data = nelda2.sub %>% filter(opp.fav.cancel == 0)) #Does not include prior mobilization control due to error fitting model
+summary(model.evac.dur)
+sjPlot::plot_model(model.evac.dur, type = "pred", terms =  c("v2xel_frefair", "lji.lag [.17, .61]"))
+
+#Not many observations have participant size
+
+##Not run: inlcuding control for prior protest results in perfect separation
+#model.ecav.bin <- glm(any.ecav.protest ~ lji.lag*v2xel_frefair + altinf.1lag  + v2elintmon + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag +  liberal_no_jucon + e_wb_pop + e_migdppcln + n.events.pre +
+  #                       as.factor(v2x_regime),  data = nelda2.sub %>% filter(opp.fav.cancel == 0), family = #binomial(link="logit"))
+#summary(model.ecav.bin)
+
+#sjPlot::plot_model(model.ecav.bin, type = "pred", terms =  c("v2xel_frefair", "lji.lag [.19, .63]"))
+```
+
+3-way evac models
+
+```{r}
+model.evac.num3 <- glm.nb(n.events.post  ~ lji.lag*v2xel_frefair*altinf.1lag + v2elintmon + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag +  liberal_no_jucon + e_wb_pop + e_migdppcln + n.events.pre + v2pscohesv + as.factor(v2x_regime),  data = nelda2.sub %>% filter(opp.fav.cancel == 0))
+summary(model.evac.num3)
+p.quant.3 <- sjPlot::plot_model(model.evac.num3, type = "pred", terms =  c("v2xel_frefair", "lji.lag [.17, .61]", "altinf.1lag"))
+p.quant.3 + coord_cartesian(xlim =c(0, 1), ylim = c(0, 10))
+
+model.evac.dur3 <- glm.nb(d.events.num ~ lji.lag*v2xel_frefair*altinf.1lag + v2elintmon + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag +  liberal_no_jucon + e_wb_pop + e_migdppcln + n.events.pre + v2pscohesv+ as.factor(v2x_regime),  data = nelda2.sub %>% filter(opp.fav.cancel == 0))
+summary(model.evac.dur3)
+p.dur.3 <- sjPlot::plot_model(model.evac.dur3, type = "pred", terms =  c("v2xel_frefair", "lji.lag [.17, .61]", "altinf.1lag"))
+p.dur.3 + coord_cartesian(xlim =c(0, 1), ylim = c(0, 150))
+
+```
+
+```{r}
+##Winner margin, combines pres and leg elections
+
+model.3way <- glm(as.factor(protests) ~ lji.lag*v2xel_frefair*altinf.1lag + v2elintmon + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag + e_wb_pop + e_migdppcln + n.events.pre + v2pscohesv+ as.factor(v2x_regime) + liberal_no_jucon, data = nelda2.sub, family = binomial(link="logit"))
+summary(model.3way)
+
+sjPlot::plot_model(model.3way, type = "pred", terms =  c("v2xel_frefair", "lji.lag [.17, .61]", "altinf.1lag [.32, .70, .89]")) #There is an interaction effect under extreme conditions: altinfo low, comparing min-max of LJI--highest LJI is significantly higher pp of protest than lowest.
+No effect when altinfo is medium or high.
+This only emerges in models without country fixed effects.
+With FE, no effect
+
+model.2way <- lmer(protests ~ lji.lag*v2xel_frefair + altinf.1lag + v2elintmon + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag + liberal_no_jucon + e_wb_pop + e_migdppcln + n.events.pre + v2pscohesv+ (1 | country_id.x), data = nelda2.sub)
+summary(model.2way)
+
+#interplot(model.2way, var1 = "lji.lag", var2 = "v2xel_frefair")
+sjPlot::plot_model(model.2way, type = "eff", terms =  c("v2xel_frefair", "lji.lag [.17, .61]")) #1st and 3rd quartiles
+#interactions::interact_plot(model.2way, pred = v2xel_frefair, modx = lji.lag, interval = T)
+```
+
+##Models using vdem protest data
+Uses lmer since DV is not binary
+
+Current-year protest
+```{r}
+##Overall
+model.3way.overall <- lmer(v2cagenmob ~ lji.lag*v2xel_frefair*altinf.1lag + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag + as.factor(v2x_regime) + liberal_no_jucon +  e_wb_pop + e_migdppcln + n.events.pre  + v2pscohesv+ (1 | country_id.x), data = nelda2.sub)
+summary(model.3way.overall)
+
+sjPlot::plot_model(model.3way.overall, type = "pred", terms =  c("lji.lag", "v2xel_frefair [.23, .87]",  "altinf.1lag [.32, .89]"))  #Here we see that low LJI results in more protest in higher altinf environments
+
+##Pro-dem
+
+model.3way.prodem <- lmer(v2cademmob ~ lji.lag*v2xel_frefair*altinf.1lag + v2elintmon + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag + as.factor(v2x_regime) + liberal_no_jucon +  e_wb_pop + e_migdppcln + n.events.pre  + v2pscohesv+ (1 | country_id.x), data = nelda2.sub)
+summary(model.3way.prodem)
+
+sjPlot::plot_model(model.3way.prodem, type = "pred", terms = c("v2xel_frefair",  "lji.lag [.24, .72]", "altinf.1lag [.37, .81]" ))  #Here we see low LJI is associated with more protest in higher alt-inf environments (same as above)
+#interactions::interact_plot(model.3way.prodem, pred = lji.lag, modx = v2xel_frefair, mod2 = altinf.1lag,
+#                            data = nelda2 %>% filter(opp.fav.cancel == 0 & inc.fav.cancel == 0))
+
+model.2way.prodem <- lmer(v2cademmob ~ lji.lag*v2xel_frefair + altinf.1lag + v2elintmon + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag + as.factor(v2x_regime) + liberal_no_jucon + e_wb_pop + e_migdppcln + n.events.pre  + v2psoppaut + (1 | country_id.x), data = nelda2.sub)
+summary(model.2way.prodem)
+
+sjPlot::plot_model(model.2way.prodem, type = "pred", terms = c("v2xel_frefair",  "lji.lag [.19, .61]" ))  #Here we see low LJI is associated with more protest in higher alt-inf environments (same as above)
+#interactions::interact_plot(model.3way.prodem, pred = lji.lag, modx = v2xel_frefair, mod2 = altinf.1lag,
+#                            data = nelda2 %>% filter(opp.fav.cancel == 0 & inc.fav.cancel == 0))
+
+##Pro-authoritarian
+
+model.3way.proauth <- lmer(v2caautmob ~ lji.lag*v2xel_frefair*altinf.1lag + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag + as.factor(v2x_regime) + liberal_no_jucon + e_wb_pop + e_migdppcln + n.events.pre  + v2pscohesv+(1 | country_id.x), data = nelda2.sub)
+summary(model.3way.proauth)
+
+sjPlot::plot_model(model.3way.proauth, type = "pred", terms =  c("v2xel_frefair", "lji.lag [.01, 1.0]", "altinf.1lag [.32, .70, .89]"))  #Here we see a similar relationship, but only at highest levels of altinf
+
+```
+Summary of findings of negative binomial models:
+
+-ECAV count data: LJI has no effect except in a very small number of high-quality elections in low-info environments
+-ECAV duration data: LJI has positive effect w/r/t duration in high-quality, low-info elections
+-NELDA data: LJI increases risk of protest in low-quality, high-info elections
+
+Summary of LMER models:
+
+-Pro-dem mobilization (VDEM): LJI has no effect in low-info environments, may reduce protest risk in high-info cases
+
+This seems to show that high LJI does not increase protest risk in a low-quality election when media freedom is low.
+This would counter the information-revelation argument about courts (or at least complicate it--courts can only reveal so much info by themselves).
+
+Things to consider:
+- Free media can reveal the fact that low-independence courts are upholding tainted elections
+- Opposition choice to press cases is strategic
+- These models do not include CCP data on electoral courts
+
+One thing to consider is that opposition choose not to protest when LJI is high, but when media is able to provide information on manipulation.
+
+Idea:
+There are two proposed channels for independent courts to influence electoral manipulation.
+First, and most traditionally, through provision of information leading to an increased risk of protest; second, by raising the potential penalties for engaging in manipulation.
+
+##Models of canceling election results after the vote
 Models below regress favorable results for incumbent being canceled on protests*lji.
@@
 
+There are several ways to do this:
+
+1) Binary protest or not
+2) Number of protests in the wave
+3) Length of wave
+4) Number of participants
+5) Binary violence or not
+
 ```{r}
 nelda3 <- nelda3 %>% mutate(protests.postelex = ifelse(protests == 1 & protesterdemand1 == "political behavior, process", 1, 0))
+
+nelda3 <- nelda3 %>% mutate(protests.postelex.quant = ifelse(protests.postelex == 1, protestnumber, 0))
+```
+
+Cutting out closed autocracies and liberal democracies
+
+```{r}
+nelda3.sub <- nelda3 %>% filter(v2x_regime == 1 | v2x_regime == 2)
 ```
@@
 ```{r}
-model.base.pres <- lmer(protests.postelex ~ lji.lag*v2xel_frefair + v2elvotlrg + cspart.1lag + gdpgro.1lag + (1 | country_id.x),  data = nelda3 %>% filter(opp.fav.cancel == 0))
-summary(model.base.pres)
-interplot(model.base.pres, var2 = "v2xel_frefair", var1 = "lji.lag")
+model.mmd.binary <- glm(protests.postelex ~ lji.lag*v2xel_frefair*altinf.1lag + v2elintmon + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag +  liberal_no_jucon + as.factor(country_id.x),  data = nelda3.sub, family = binomial(link="logit"))  #as.factor(v2x_regime) excluded due to multico.
+summary(model.mmd.binary)
+sjPlot::plot_model(model.mmd.binary, type = "pred", terms =  c("v2xel_frefair", "lji.lag [.26, .49]", "altinf.1lag"))
 
-model.base.leg <- lmer(protests.postelex ~ lji.lag*v2xel_frefair + v2ellostlg + cspart.1lag + gdpgro.1lag + (1 | country_id.x), data = nelda3 %>% filter(opp.fav.cancel == 0))
-summary(model.base.leg)
+model.mmd.binary2 <- glmer(protests.postelex ~ lji.lag*v2xel_frefair*altinf.1lag + v2elintmon + winner.margin  + physinteg.1lag +  (1 | country_id.x), data = nelda3 %>% filter(opp.fav.cancel == 0), family = binomial(link = "logit"))
+summary(model.mmd.binary2)  # Various variables when included causes the model to fail to converge
 
-interplot(model.base.leg, var2 = "v2xel_frefair", var1 = "lji.lag")
+interplot(model.mmd.binary2, var2 = "v2xel_frefair", var1 = "lji.lag")
+
+sjPlot::plot_model(model.mmd.binary2, type = "pred", terms =  c("v2xel_frefair", "lji.lag [.26, .49]"))
+
+```
+
+Protest quantity
+```{r}
+model.base.num <- glm(protests.postelex.quant ~ lji.lag*v2xel_frefair + altinf.1lag + v2elintmon + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag +  liberal_no_jucon + as.factor(v2x_regime) + as.factor(country_id.x),  data = nelda3.sub %>% filter(opp.fav.cancel == 0), family = poisson(link="log"))
+summary(model.base.num)
+interplot(model.base.num, var2 = "v2xel_frefair", var1 = "lji.lag")
+sjPlot::plot_model(model.base.num, type = "pred", terms =  c("v2xel_frefair", "lji.lag [.24, .72]"))
+
+model.mmd.num2 <- glmer(protests.postelex.quant ~ lji.lag*v2xel_frefair*altinf.1lag + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag +  liberal_no_jucon + as.factor(v2x_regime) +  (1 | country_id.x), data = nelda3.sub %>% filter(opp.fav.cancel == 0), family = poisson(link = "log"))
+summary(model.mmd.num2)  
+sjPlot::plot_model(model.mmd.num2, type = "pred", terms =  c("v2xel_frefair", "lji.lag [.24, .72]", "altinf.1lag"))
+
 ```
@@
 
+Violence
+
+```{r}
+model.viol.binary <- glm(protesterviolence ~ lji.lag*v2xel_frefair + altinf.1lag + v2elintmon + v2elintmon + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag +  liberal_no_jucon + as.factor(country_id.x),  data = nelda3.sub, family = binomial(link="logit"))  #as.factor(v2x_regime) excluded due to multico.
+summary(model.viol.binary)
+sjPlot::plot_model(model.viol.binary, type = "pred", terms =  c("v2xel_frefair", "lji.lag [.26, .49]"))
+
+model.mmd.binary2 <- glmer(protesterviolence ~ lji.lag*v2xel_frefair*altinf.1lag + winner.margin  + physinteg.1lag +  (1 | country_id.x), data = nelda3 %>% filter(opp.fav.cancel == 0), family = binomial(link = "logit"))
+summary(model.mmd.binary2)  # Various variables when included causes the model to fail to converge
+
+interplot(model.mmd.binary2, var2 = "v2xel_frefair", var1 = "lji.lag")
+
+sjPlot::plot_model(model.mmd.binary2, type = "pred", terms =  c("v2xel_frefair", "lji.lag [.26, .49]"))
+
+```
+
 ```{r}
@@
 ```
+
+# Looking at protest and competitiveness
+
+```{r}
+#Same pattern here occurs with LJI
+
+model.3way <- glm(protests.postelex ~ reform_positive.lag*v2elirreg.inv*opposition.oversight.lag + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag + as.factor(v2x_regime) +  factor(country_id.x), data = nelda3)
+summary(model.3way)
+
+sjPlot::plot_model(model.3way, type = "pred", terms =  c( "opposition.oversight.lag",  "reform_positive.lag", "v2elirreg.inv [-2.87, 2.63]")) #There is an interaction effect under extreme conditions:
+
+model.3way <- glm(protests.postelex ~ reform_positive.lag*v2elirreg.inv*elcompindex.m.1lag + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag + as.factor(v2x_regime) +  factor(country_id.x), data = nelda3)
+summary(model.3way)
+
+sjPlot::plot_model(model.3way, type = "pred", terms =  c( "elcompindex.m.1lag ", "reform_positive.lag",   "v2elirreg.inv [-2.87, 2.63]")) #There is an interaction effect under extreme condi
+
+model.3way <- glm(protests.postelex ~ reform_positive.lag*v2elirreg.inv*elcompindex.m.1lag + winner.margin + cspart.1lag + gdpgro.1lag + presidential + physinteg.1lag + as.factor(v2x_regime) +  factor(country_id.x), data = nelda3)
+summary(model.3way)
+
+sjPlot::plot_model(model.3way, type = "pred", terms =  c( "elcompindex.m.1lag ", "reform_positive.lag",   "v2elirreg.inv [-2.87, 2.63]")) #There is an interaction effect under extreme condi
+```
+
+Looking at the ECAV data, it suggests that LJI does not increase the risk that any protest will occur.
+However, it does appear to increase the risk that protests will last longer when elections are low quality

### 2022-02-21 Output/manipulation_protest_draft.Rmd
Updates to intro

@@
 ---
+
+Just as many less democratic states have enshrined multiparty elections into their political systems in the post-Cold War period, many have simultaneously experimented with varying degrees of judicial independence.
+A growing research tradition identifies various benefits that more independent courts can provide for non-democratic ruling parties, including regime legitimation, monitoring of discontent and of abuses by low-level officials, protecting investments, and more [@Moustafa2008].
+As a result of these benefits, even many of the most authoritarian states have added de jure protections for courts's autonomy to their laws and constitutions [Staton2020].
+At the same time, just like elections, more independent courts create an unpredictable connection between state and society; more open courts create opportunities for legal mobilization against the regime, with uncertain consequences.
+Consequently, de-democratizing states may act to restrain previously independent courts.
+
+These dynamics can be seen across Europe, from legal contestation in Ukraine and Russia [@popova2020politicized] to the erosion of judicial independence in Poland and Hungary.
+In the latter case, the ruling Law and Justice Party (PiS) has taken significant steps to weaken judicial independence since 2015.
+These include increasing the legislature's influence over judicial appointments, stacking the judicial-nomination council with ruling-party allies, and the creation of a disciplinary chamber under the control of the Ministry of Justice.
+This disciplinary body can impose salary cuts, or even dismissal, on judges found to have in vaguely-defined "political activity"---which includes discussions of judicial independence and the PiS reforms themselves [@Duncan2020].
+The situation has raised sufficient alarm among other EU member states that the European Commission has announced an infringement procedure against Poland, for flouting EU law while undermining its national courts' autonomy (https://www.dw.com/en/eu-starts-new-legal-action-against-poland-over-rule-of-law/a-60220102).
+Similarly, in Hungary, the parliamentary majority created a new National Office for the Judiciary in 2012, with broad powers to appoint and promote judges (https://www.venice.coe.int/Newsletter/NEWSLETTER_2012_02/1_HUN_EN.html).
 

### 2022-07-08 Output/manipulation_protest_draft_v2.Rmd
Major revisions to focus of paper; away from LJI toward protest generally

---
title: "Draft: Judicial independence and post-election protest"
author: "Cole J. Harvey (Oklahoma State University)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    bookdown::pdf_document2: default
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
header-includes: \usepackage{booktabs}
abstract: "Multiparty elections and courts with a degree of independence are common features of non-democratic politics. Previous research posits a connection between these two institutions, arguing that judicial independence increases the risk of post-election protest, thus reducing electoral manipulation. This paper argues that the relationship is not straightforward, identifying four distinct (and competing) channels through which courts may affect protest risk. Using cross-national data from 1990-2012, the paper shows that independent courts only increase protest risk under limited circumstances. The results show that judicial independence increases protest risk where economic grievances are high, while reducing protest risk where grievances are low. As a result, creates judicial independence an underappreciated risk of democratic backsliding in well-off societies, while improving the odds of democratization elsewhere."
---

Headlines are often captured by scenes of large-scale post-election protest, usually with accompanying allegations of electoral malfeasance. On some occasions, these protests can result in new elections and the fall of an authoritarian regime, heightening the sense of drama. An extended election protest movement creates a moment of heightened risk for the regime; the uncertainty of the outcome creates opportunities for elite defection, splits in the security services, and other negative outcomes. For this reason, many scholars---especially those working in the formal modeling tradition---have identified the risk that election manipulation will spark protest as a major deterrent to regimes wishing to engineer election outcomes. Most often, though, non-democratic states can rely on superior resources, elite solidarity, and repressive capacity to quash electoral protests. How threatening, then, are electoral protests really, and can that threat constrain electoral manipulation by governments?

Research on the risk posed to non-democratic governments by electoral protest often acknowledges---whether explicitly or implicitly---the importance of sustained protest campaigns, rather than one-off events [@Bunce2006;@tucker2007;@DAnieri2006;@Carnaghan2016;@bunce2010]. As Brancati [-@Brancati2016] finds, approximately two-thirds of pro-democracy electoral protests ended within three days, posing little threat to incumbents. Only about ten percent of such protests last long enough to see incumbents fall. Classic examples of such outcomes include Ukraine's Orange Revolution, which lasted roughly two months, the months-long series of protests and strikes that followed the 2001 Madagascar presidential election, or the three-week-long Rose Revolution in Georgia. Other major protest movements do not succeed in ousting ruling parties, but do win some concessions. Russia's 'For Fair Elections' movement in 2011 did not oust the Putin regime, for example, but did see the reintroduction of gubernatorial elections. In sum---though with notable exceptions, like the People Power Revolution in the Philippines---it is generally protest *mobilization* that poses challenges to incumbents, rather than the occurrence of protest per se. Here, I integrate this understanding of protest's risk to regimes by using duration data, rather than binary indicators of protest.

This implicit assumption has largely been overlooked in theories and empirical studies of the relationship between election manipulation and protest, however. In most cases, the risk of protest has both been conceptualized and operationalized as a binary outcome: protests occur or they do not. This approach obscures the importance of the number and duration of electoral protests in a ruling party's calculation of the risk involved in manipulation an election.\footnote{One exception can be found in Hyde and Marinov's [-@hyde2014] paper, which considers protest duration in a secondary analysis. However, their duration analysis is a two-by-two table that does not control for possible confounding variables, and relies on a dataset that excludes un-protested elections that received positive observer reports.} Small, short-lived protests may be easily contained by a well-resourced government, while a long-lasting movement with a large number of protest events poses a more significant challenge.

Using cross-national data on 467 elections from 1990 to 2012, adapted from the recently-released Electoral Contention and Violence (ECAV) dataset [@Daxecker2019], this paper finds that more fraudulent elections are associated with a higher risk any post-election protest event occurring. This is in keeping with the formal-modeling literature, which predicts that overreach in the form of election fraud indicates government weakness, making opposition groups confident of winning a conflict in the streets. However, there is no statistically significant association between the level of election fraud and the duration of protest. This suggests a more complex relationship between election fraud and protest. One the one hand, the result is consistent with the prediction of some formal models that fraud is a noisy signal---in some cases, opposition groups may misperceive a strong government as weak. On the other hand, since fraud is not associated with the overall duration of a protest wave, the risk to governments has likely been overemphasized. That is, even if protests do break out in response to large-scale fraud, the government has many tools it may deploy to curtail them. To the extent that longer protest waves are more damaging to the government than short ones, protest risk does not appear to be a substantial constraint on the severity of election fraud.

However, once protests begin, the noisy signal of regime strength provided by the election is replaced by a very clear signal: the force of the regime's response to the emerging protest movement in contest with the mobilizational capacity of the opposition.

This suggests that the risk to incumbents generated by election manipulation is primarily informational and strategic, rather than the result of grievance and legitimacy cost. In turn, this implies that election manipulation is less costly than many models assume; many non-democratic states have the tools needed to curtail a protest movement. In addition, it indicates that ruling parties should follow an 'all-in' strategy on election manipulation. More manipulation is not associated with more protest mobilization, but a large margin of victory is associated with both smaller and shorter protest waves. These results help explain both the relative paucity of electoral protests, as well as the rarity of their success in overturning incumbent regimes.

# Literature review

Broadly speaking, there are thought to be two mechanisms by which election manipulation leads to protest. The first is through revealing information about the strength of the government or ruling party. This information, in turn, feeds into strategic decision-making by opposition groups.  This logic is often seen in formal theory models [@Egorov2021;@chernykh2015third]. Several prominent models of electoral manipulation argue that the risk of mass post-election protest helps restrain incumbents who might otherwise prefer larger levels of manipulation [@magaloni2010; @Little2015; @Little2012; @birch2011; @Fearon2011;@Luo2020;@Rozenas2016b]. The risk of mass protest in these models is assumed to constitute a significant guardrail for election integrity in democracies, and a significant challenge for ruling parties in non-democracies. Building on this logic, researchers have proposed that third-party institutions---like election monitors [@Little2012;@Hyde2014; @daxecker2012], civil society groups [@birch2017getting], and courts [@Chernykh2015]---can reduce manipulation and increase protest risk by revealing information about electoral malfeasance.

Despite this consensus in the formal literature, there is relatively little empirical evidence to support the claim that manipulation per se indicates weakness, and thus increases protest risk. On the contrary, Simpser [-@Simpser2011] shows support for the alternative (though still informational) view that excessive manipulation can induce cooperation from opposition groups, elites, and other actors by revealing information about regime strength. Similarly, the type of manipulation employed can convey information about the ruling party's capacity to mobilize [@Harvey2020]. In this way, large-scale manipulation can contribute to a ruling party's ability to produce a large, stabilizing majority vote-share [@magaloni2006]. Similarly, Luo and Rozenas [-@Luo2020] posit that ex post fraud is more likely to result in protest than ex ante manipulation, though this can be contrasted with the finding that parties are more likely to protest elections after ex ante changes to the rules [@chernykh2014]. In sum, it is likely that manipulation *may* signal regime weakness, but this is more likely when margins of victory are narrow and when falsification is a major tool.

<!--  One critique of this view, although still within the information/strategic framework, comes from Simpser's [-@Simpser2011] pathbreaking book. Simpser argues that when wanipulation efforts are excessive---well beyond what is needed simply to win---they demonstrate the organizational capacity, resources, and manpower of the ruling party [@Simpser2013]. Large-scale manipulation is, thus, a signal of strength rather than weakness, and can deter protest. By contrast, partial manipulation efforts that fail to secure a large majority vote-share signals the government lacks both popular support and mobilizational capacity, making opposition protest more likely. -->

Without disputing that election manipulation conveys information, a second school of thought has emphasized the ways in which biased elections may predispose individuals to engage in protest [@kuntz2009]. In this view, election manipulation acts as a grievance shared by individuals who prefer to see the ruling party lose the election---especially when it appears that fraud may have been decisive [@kuntz_more_2009]. Belief in a stolen election is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007], possibly initiating a protest cascade [@Kuran1991]. This conception is closely connected to the idea that manipulated elections carry legitimacy costs for the ruling party [@birch2011].

For example, ballot-stuffing and voter coercion have been shown to generate stronger feelings of anger than pre-election manipulation of ballot access [@Szakonyi2021], and access to information about manipulation appears to reduce support for the ruling party [@Reuter2021], increase support for protest and reduce the perceived legitimacy of the regime [@Williamson2021]. This argument has found support in studies of the relationship between economic grievances, manipulated elections, and protest [@Rod2019;@Brancati2016]. A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b], which is suggests that emotions, moral condemnation, or grievances take precedence over strategic calculations for individuals. Subjective belief in fraud was also shown to be correlated with approval of mass protests after Russia's 2011 election [@Chaisty2013]. Furthermore, simply being on the losing side in an election is associated with reduced belief that the election was fair [@Cantu2015], and with increased support for electoral protest [@Anderson2006]. [Add procedural justice cites to this graf.]

These two explanations are, of course, interconnected. As Schedler writes, "Grievances provide moral justifications for collective protest: we need to do something. Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304]. Several empirical studies have, accordingly, found links between election manipulation and binary measures of protest [@Brancati2016;@Hyde2014a;@Harvey2020;@Rod2019].

Sustained, sizeable protests may also contribute to elites' decisions to stand by the incumbent regime, or to side with "the people" in their own pursuit of power [@hale_patronal_2014;@svolik_politics_2012;@robertson_protest_2010].

The literature thus expects that manipulated elections increase the risk of protest by 1) revealing regime weakness, especially when the margin of victory is low, and 2) exacerbating existing grievances and upsetting citizens' sense of procedural fairness. However, as discussed below, electoral protest is relatively rare, and even in cases where it does occur, most efforts to overturn fraudulent election results through protest have failed [@kalandadze2009electoral]. The apparent failure of protest risk to deter election fraud underlines our limited understanding of non-democratic ruling parties' decision-making. Does the extent of fraud, or the severity of underlying social grievances, make long-lasting protest more likely (and thus restrain fraud in anticipation)?

# Theory

The theory proposed here is straightforward. First, it follows previous work in acknowledging that election manipulation is most likely to spark a protest movement under two conditions: when narrow margins of victory signal weakness [@Simpser2011], and when general social grievances are high [@Brancati2016;@tucker2007].

>H1: Post-election protest will be more likely to occur when fraud is high and ruling-party margin of victory is small, and less likely when fraud is low or the margin of victory large.

>H2: Post-election protest will be more likely to occur when fraud is high and economic and social grievances are elevated.

However, once protests begin, the extent of electoral fraud is unlikely to affect their duration. The initial emerge of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@Tilly1977;@Lipsky1968;@Hummel2019;@Rasler1996;@Bishara2015;@Frantz2014]. (**See others**.) Given that post-electoral protests are a direct challenge to the regime, toleration is unlikely in these cases. Instead, incumbents usually begin a process of suppressing protests, repressing activists and movement leaders , and offering 'sideways concessions' to keep other citizens off the streets [@Hummel2019].\footnote{Of course, much repression is pre-emptive [@Sullivan2016]. For example, non-democratic governments typically repress opposition activists in the run-up to elections[@Bhasin2013].} Because the emergence of protest indicates that the regime's pre-emptive efforts have failed, it is not clear a prior how the regime should handle protesters it failed to deter [@Ritter2016]. In many cases, the government doubles down on repression, calling upon the security services to disperse protests, arrest leaders, and perhaps employ harsher methods including disappearance, torture, and execution [@Escriba_Folch2013;@Davenport2007]. Though such crackdowns risk provoking a backlash, they are aimed at demobilizing the opposition by increasing the costs of participation in overt dissent [@DeVogel2022]. In other cases, the government may attempt to accommodate the opposition, by making concessions on policy or institutions [@Moore2000]. Alternatively, concessions may be aimed at the more moderate public, in order to keep them on the sidelines of the conflict [@Hummel2019].

What results is a test of strength between the anti-incumbent protest coalition and the regime with its supporters (Tilly 2006). In essence, as Beaulieu [@Beaulieu2014a] notes, the onset of electoral protest marks a case of bargaining failure. Just as in the international context, the ensuing conflict acts to reveal information on the strength and resolve of the two sides.  In such a contest, the incumbent enjoys multiple advantages, though it may lack the support of a majority of the population. It controls the resources of the state, and can call upon the support of those elites who benefit from the status quo. High-capacity states, as a result, are more likely to be able to engage in successful repression (Tarrow 2011). Though a biased election may temporarily allow diverse opposition groups to set aide their differences [@trejo_ballot_2014], this internal diversity can enable the incumbent to divide the opposition with selective repression or concessions. For example, survey research from Russia's 'For Fair Elections' movement found that many protesters were motivated as much by... [Find citation]. While the regime may have its own splits---between reformists and hardliners, for example---the emergence of protest can act to solidify support for the ruling party among its supporters.

After the initial moment of opportunity and heightened grievance provided by the flawed election, I argue that protesters and potential movement supporters return to a typical status as "passionate economists" [@VanZomeren2012]. In other words, protesters and potential protesters are motivated both by emotion in response to grievance, and by a more rational calculation of individual-level risk and group-level efficacy. Individuals evaluate whether their collective action is likely to bring about changes in response to their sense of grievance, and whether that likelihood outweighs the danger they may expose themselves to personally by protesting. Crucially, this process is dynamic, and can be updated over time in response to new information and new events [@VanZomeren2012]. If protest movements have sufficient numbers and organizational resources to put the regime on the back foot, this can lead to longer-lasting and more numerous protest actions by increasing individuals' sense of group efficacy. On the other hand, if the regime is well-resourced, capable, and unified, its actions to put down protest can lead to a negative re-evaluation of the movement's odds of success. This does not, of course, eliminate citizens' sense of grievance. Instead, it leads them to abandon 'approach-oriented' forms of coping, to avoidance-oriented forms: demobilization and disengagement.

>H3: Once protests begin, the severity of election fraud will not be correlated with the duration of protest.

Since the level of manipulation will no longer serve as an important indicator of regime strength once a challenge is taken to the streets, the duration of protests will instead be related to other indicators of oppostion and incumbent capacity. As identified in prior research, these factors include the previous repressive history of the government, the institutionalization of opposition parties, the margin of victory in the election, and whether or not the chief executive faces term limits.

-Democracy leads to less repression: Bueno de Mesquita et al., 2005). Davenport & Armstrong (2004), 
-Prior repression makes repression more likely (Davenport 2007 (Tyrannical peace))
-Resource wealth makes authoritarian survival more likely (Girod et al 2016)

From E-F 2013: Nevertheless, violent repression may have pernicious consequences, since it may bring about a
violent escalation or a radicalization of positions, thereby unleashing a violent response from some
groups and even potentially provoking a coup (Gurr, 1970; Ortiz, 2007; Pierskalla, 2010). Thus,
Opp and Ruehl (1990) suggest that repression may lead to micro-mobilization processes leading to
further protest activities and unrest. Likewise, Ginkel and Smith suggest that ‘dissident activity is
more likely to be effective in motivating rebellious action under highly repressive conditions’
(1999: 292). Some argue that when nonviolent actions are violently repressed, protesters may turn
to violent activities (Lichbach, 1987; Moore, 1998), as the Syrian case illustrates. Similarly,
Goodwin (2001) argues that revolutions are more likely in patrimonial and very repressive states.
For instance, in Iran on Black Friday in September 1978, the regime’s security forces killed hun-
dreds of anti-shah protesters in Tehran. After a brief decrease in activity, public dissent and strikes
quickly resumed and escalated. The mobilizations that took place in December gathered hundreds
of thousands of people calling for the shah’s rule to end. In January 1979, the shah left the
country

...restrictions on civil rights can thwart the ability of some
actors (especially the opposition) to organize, coordinate, and mobilize in the short run

From de Vogel:  The strategic use of repression is founded in the fact that protests in general are undesirable
to authoritarian regimes. Frantz and Kendall-Taylor (2014) demonstrate that popular revolution is
an increasingly common avenue by which autocrats leave office, and some of these originate as
everyday protests. Repression is often seen as the first line of defense against protest, given that
it is inherent to autocracies (Wintrobe, 1998; Escrib`a-Folch, 2013), and it is consistently observed
to increase in response to dissent (Davenport, 2007a). But cracking down on less threatening
protest can create the appearance of weakness on the part of the regime, and repressing “average
citizens”—like those who often make up the ranks of everyday protests—can increase dissent
(Francisco, 1995; Rasler, 1996). On the local level where everyday protests often originate, control
over the repressive apparatus may be less uncertain (Earl, 2011). That can lead to the use of “softer”
forms of repression that are easier to control (Fu, 2018), or the use of non-state actors to execute
it, which can undermine state legitimacy (Ong, 2018).
On the other hand, the regime can respond to protests with concessions, but the costs and
benefits of doing so are unclear. Concessions are conventionally associated with the cooptation
of regime challengers, most often through sharing spoils, redistribution or patronage (Bueno de
Mesquita and Smith, 2010), or via incorporation into authoritarian institutions such as legislatures
(Lust-Okar, 2005; Gandhi and Przeworski, 2007). Lipsky (1968) classifies such concessions as
protest bargaining, and underscores that they can only occur when the protesters have political
resources the government wants. Everyday protest campaigns would therefore appear poor can-
didates for such bargaining, given that they are may have few political resources at their disposal
and are only sometimes organized by political parties

On uncertainty of repression and protest, from Girod et al: Yet the differences in demobilization of protest movements following repression remain largely unexplained. Government repression of protesters is strongly associated with quelling protest activity in some cases (Lyall, 2009; Schatzman, 2005), and triggering increased mass mobilization in others (Francisco, 1995; Khawaja, 1993; Mason and Krane, 1989). Complicating matters further, in some cases the relationship between repression and protest appears to change over time (Francisco, 1995; Inclan, 2009; Mason and Krane, 1989; Rasler, 1996; Tsebelis and Sprague, 1989), across levels of repression (DeNardo, 1985; Levitsky and Way, 2010; Muller and Weede, 1990), and based on the complex ways in which “threat” and “opportunity” combine to influence protest dynamics (Goldstone and Tilly, 2001). Because fine-grained global data on repression of mass movements have not been available until now (Davenport, 2005: viii; 2007a: 10), extant studies depend largely on only a few cases. In short, the literature has yet to identify cross-national patterns to explain when publics in an authoritarian regime respond to repression with greater collective action vs. disbanding.

 <!-- Chernykh and Svolik, in particular, argue that by revealing information about manipulation, even moderately independent courts can increase the risk of protest against manipulated elections (thus deterring manipulation), while also reducing the risk that opposition groups mistakenly protest clean elections. This result is in keeping with the argument that electoral protests are the outcome of bargaining failure among incumbents and opposition groups, which are worsened in low-information environments [@Beaulieu2014a]. -->

<!-- However, one of the central functions of independent judiciaries outside of liberal democracies is to create mechanisms through which citizens grievances can be addressed [@Whiting2017; @Solomon2007; @Hilbink2007d]. Relatively impartial courts create an opportunity for citizens to resolve grievances against the state and against other individuals who may be more powerful or influential. For example, survey-experimental data show that citizens of several Middle Eastern non-democracies were more willing to engage in legal action to address their grievances when they believe courts are more independent and assertive [@Schaaf2021]. Moreover, the belief that procedures are generally followed impartially can bolster the state's claim to legitimacy [@Gerschewski2013], increasing citizens' trust in authority and willingness to abide by political outcomes [@Grimes2006]. Evidence from European democracies suggests that that de facto independence is key for such judgments [@Buhlmann2011]. Taken together, greater de facto judicial independence ought to reduce protest risk, both by creating a channel for the state to address citizens' general grievances and by offering an alternative venue to the street for addressing specific electoral concerns.  -->

<!-- These two functions for courts are at odds with regard to electoral disputes. As a result, the relationship between judicial independence and post-election protest is conceptually and empirically muddled---despite the centrality of protest risk to predominant models of electoral manipulation. The formal model proposed by Chernykh and Svolik [-@Chernykh2015] is not tested empirically, while Chernykh [-@chernykh2014] and Eisenstadt [-@eisenstadt_courting_2003] both find no effect for judicial independence on the likelihood of protest in post-communist Europe and Mexico, respectively. This calls into question the prediction that more independent courts make protest more likely in the presence of election manipulation, though to my knowledge the question has not been studied in a large-n cross-national context. -->

# Data and methods

These hypotheses are tested using data from three datasets: NELDA [@hyde2012], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence (ECAV) dataset [@Daxecker2019]. The unit of observation for this study is an election period as recorded in NELDA, which counts all national elections between 1945 and 2015. This framework prevents selecting on the dependent variable, by capturing both protested and non-protested elections. NELDA also includes a variable indicating elections in which incumbents lost; as this study is focused on protest against ostensibly victorious incumbents, I use this variable to filter out elections which the opposition won outright. I further limit the sample to regimes coded as electoral autocracies or electoral democracies in V-Dem. This excludes closed autocracies (where no national elections are held) and liberal democracies (where electoral integrity and rule of law are both high). Finally, the date range is limited by the availability of data on protest from ECAV, as discussed below, which runs from 1990 to 2012. The resulting dataset includes 647 election-period observations.

ECAV, from which study's dependent variables are drawn, records protests and other contentious actions from six months before to three months after an election. For this study, I include protests only, excluding acts like shootings, bombings, and coups. ECAV data is in event-day format. To capture *protest duration*, I take the number of days between the first electoral protest and the last. The result is a count variable ranging from 0 (no protests) to 90. The large majority of elections experience no post-election protest: in 460 of the 647 elections, no protests are recorded.

This distribution is in line with the theoretical proposition that two data-generating processes are at work: the factors that lead to the occurrence of protest may not be the same as those that sustain or instensify it. Consequently, I model the results in two stages. The first stage uses logit models to identify the factors that make *any protest* more likely to occur---this is a binary variable that is coded as 1 if any protest occurred after the election. Since election-year observations are nested within countries here, I use multilevel models with random intercepts by country in. As a second stage of analysis, I use negative binomial regressions to model the duration of protests given that protest has begun. That is, these latter models exclude cases where no protest occurred. For ease of interpretation, these processes are modeled and depicted separately in the main text, while hurdle regression models (which use the full dataset in one stage) are presented in the appendix, with similar results.

Each model includes the explanatory variables proposed above. First, three proxy measures of economic grievance include an indicator of *public sector corruption* and the *lagged GDP growth rate*, both taken from V-Dem, and a binary measure of *economic crisis* from the NELDA dataset. Higher levels of corruption, negative GDP growth rates, and a positive value for *economic crisis* should each be expected to increase citizens' sense of grievance. To capture the *severity of election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than aggregate measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest---it is often incumbents' tool of last resort [@Sjoberg2016], it can fail to signal strength [@Harvey2020], and it can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021]. Finally, the winning party's *margin of victory* is also taken from the V-Dem dataset.

Several control variables are included, which could be common causes for both the severity of election manipulation and protest. To help control for the underlying mobilizational capacity, which likely affects incumbents' decision to manipulate and the latent risk of protest, I include indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem) and the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'). Likewise, the ability of the state to engage in repression is likely to embolden the ruling party and deter protests; I thus control for *physical integrity* using the 'physical violence index' from V-Dem lagged one year. Higher values for this variable indicate greater freedom from torture and political killing. Since intense repression may itself become a grievance that triggers further protest, I also include the variable's square term. Next, all models include three measures of the degree of political contestation in the country. The first of these is an indicator variable for *electoral democracy*; the reference category is electoral autocracy. The second and third are *judicial constraints* and  *legislative constraints* on the executive, V-Dem variables which captures the extent to which the judicial and legislative bodies in the country act independently in practice to limit executive power.Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019].

In addition, several binary control variables are included. First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@Simpser2011] and a greater risk of protest. Next, a variable from NELDA indicates if the chief executive is *term limited* during the election at hand. This variable is included since incumbents running up against constitutional term limits can pose major challenges for the unity of the regime, possibly leading to fractured manipulation efforts and post-election protest [@hale_patronal_2014]. Finally, I include dummy variables that provide a categorical evalution of state capacity. These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap). If the V-Dem data codes a state as not capable of raising revenue or primarily relying on loans and foreign aid, I code it as a *weak state*. If the state is described as primarily relying on resource rents and expropriation, I code it is a *rentier state*. Finally, states that rely primarily on taxes on property or economic transactions are coded as *tax states*. This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018].

```{r setup, include = FALSE}
library(tidyverse)
library(countrycode)
library(readxl)
library(vdemdata)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
library(stargazer)
library(modelsummary)

nelda2.sub <- read.csv(here::here("Data", "nelda2_sub.csv"))

nelda2.sub <- nelda2.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda2.sub <- nelda2.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))

##Naming nelda variables

nelda2.sub <- nelda2.sub %>% rename(boycott = nelda14)
nelda2.sub <- nelda2.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                 NA)))

##Creating economic crisis variable

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = as.factor(econ.crisis))

nelda2.sub <- nelda2.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))

nelda2.sub <- nelda2.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))

##Removing cases before 1990, when ECAV starts
nelda2.sub <- nelda2.sub %>% filter(year >= 1990)

## Removing cases where inc. loses
nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)

##Intmon as a factor
nelda2.sub <- nelda2.sub %>% mutate(v2elintmon = as.factor(v2elintmon))

nelda2.sub <- nelda2.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))

#Square of physint

nelda2.sub <- nelda2.sub %>% mutate(physintlag.sq = physinteg.1lag^2)

#Crisis as factor

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))

#Margins with updates

nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))

#Protest cases only

protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1)

```

```{r data-summary, echo=FALSE}
incl.vars <- nelda2.sub %>% dplyr::select(jucon.1lag, legcon.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, any.ecav.protest, d.events.num)

datasummary_skim(incl.vars)

```

```{r include=F}

datasummary_correlation(incl.vars)
```

```{r binary-models, warning = FALSE, message = FALSE, include = F}

model.ecavbin.mm <- glm(inc.replaced  ~ n.events.post,  data = nelda2.sub)
#summary(model.ecavbin.mm)

model.ecavbin.mm <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    osorg.1lag +   winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
#summary(model.ecavbin.mm)

model.ecavbin.mm2 <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd + pubseccorrup.lag  +   osorg.1lag +  winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
#summary(model.ecavbin.mm2)
p1 <- sjPlot::plot_model(model.ecavbin.mm2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.41, 1.79]"))

model.ecavbin.mm.type <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd +  pubseccorrup.lag+  osorg.1lag +  winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate  + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
#summary(model.ecavbin.mm.type)
#sjPlot::plot_model(model.ecavbin.mm.type, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.41, 1.79]"))

model.ecavbin.mm.crisis <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag +     econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
#summary(model.ecavbin.mm.crisis)
#sjPlot::plot_model(model.ecavbin.mm.crisis, type = "pred", terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-1, 2]"))

model.ecavbin.mm.corr <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag +  osorg.1lag + pubseccorrup.lag +  natlpartyorg.lag +  pubseccorrup.lag +     pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
#summary(model.ecavbin.mm.corr)
#sjPlot::plot_model(model.ecavbin.mm.corr, type = "pred", terms = c("v2elirreg.inv", "pubseccorrup.lag"))

model.ecavbin.mm.gdp <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag + osorg.1lag +   natlpartyorg.lag +     gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub %>% filter(gdpgro.1lag < 1))
#summary(model.ecavbin.mm.gdp)
#sjPlot::plot_model(model.ecavbin.mm.gdp, type = "pred", terms = c("gdpgro.1lag", "v2elirreg.inv [-2, 2]"))

#model.ecavbin.mm3 <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag +  pubseccorrup.lag +  electoral.dem +   winner.share.upd*v2elirreg.inv*jucon.1lag + legcon.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq +  urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub %>% filter(inc.lose == 0))
#summary(model.ecavbin.mm3)
#sjPlot::plot_model(model.ecavbin.mm3, type = "pred", terms = c("jucon.1lag + legcon.1lag [all]", "v2elirreg.inv [-.4, 1.79]", "winner.share.upd"))

```
```{r}
p1 + theme_bw()
```

```{r count-models-base, warning = FALSE, message = FALSE, include = F}
model.d.base <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
#summary(model.d.base)

model.n.base <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
#summary(model.n.base)
```

```{r count-models-interact, warning = FALSE, message = FALSE, include = F}
model.d.base2 <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
#summary(model.d.base2)
p2 <- sjPlot::plot_model(model.d.base2, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.36, 1.85]")) + coord_cartesian(ylim = c(0, 200))

model.n.base2 <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
#summary(model.n.base2)
sjPlot::plot_model(model.n.base2, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-1, 1.79]"))

```

```{r, count-models-interact-griev, warning = FALSE, message = FALSE, include = F}
model.d.griev <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +    gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
#summary(model.d.griev)
sjPlot::plot_model(model.d.griev, type = "pred", terms = c("gdpgro.1lag", "v2elirreg.inv")) + coord_cartesian(ylim = c(0, 200))

model.n.griev <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
#summary(model.n.griev)
sjPlot::plot_model(model.n.griev, type = "pred", terms = c("gdpgro.1lag", "v2elirreg.inv"))

```

```{r, count-models-interact-griev2, warning = FALSE, message = FALSE, include=F}
model.d.griev2 <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv +  econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
#summary(model.d.griev2)
p3 <- sjPlot::plot_model(model.d.griev2, type = "pred", terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-2, 4]")) + coord_cartesian(ylim = c(0, 200))

model.n.griev2 <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
#summary(model.n.griev2)
sjPlot::plot_model(model.n.griev2, type = "pred", terms = c("v2elirreg.inv", "econ.crisis.nelda.1lag"))

```

```{r, count-models-interact-griev3, warning = FALSE, message = FALSE, include=F}
model.d.griev3 <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +   natlpartyorg.lag + v2elirreg.inv +  pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
#summary(model.d.griev3)
sjPlot::plot_model(model.d.griev3, type = "pred", terms = c("v2elirreg.inv", "pubseccorrup.lag")) + coord_cartesian(ylim = c(0, 200))

model.n.griev3 <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
#summary(model.n.griev3)
sjPlot::plot_model(model.n.griev3, type = "pred", terms = c("v2elirreg.inv", "pubseccorrup.lag"))

```

```{r tables-binary, echo=FALSE}
models.binary <- c(model.ecavbin.mm, model.ecavbin.mm2, model.ecavbin.mm.type, model.ecavbin.mm.corr, model.ecavbin.mm.crisis, model.ecavbin.mm.gdp)

modelsummary(models.binary, stars = T, coef_rename = c("v2elirreg.inv:winner.share.upd" = "Fraud x winner share",
                                                       "pubseccorrup.lag:v2elirreg.inv" = "Fraud x corruption",
                                                       "econ.crisis.nelda.1lag1:v2elirreg.inv" = "Fraud x econ. crisis"))
```

```{r tables-duration, results='asis', echo=FALSE}
models.duration <- list("Model 8" = model.d.base, "Model 9" = model.d.base2, "Model 10" = model.d.griev, "Model 11" = model.d.griev2, "Model 12" = model.d.griev3)
modelsummary(models.duration, stars = T)

```

```{r tables-quantity, include=FALSE}
models.quantity <- list("Model 13" = model.n.base, "Model 14" = model.n.base2, "Model 15" = model.n.griev, "Model 16" = model.n.griev2, "Model 17" = model.n.griev3)
#modelsummary(tidytest)
modelsummary(models.quantity, stars = T)
```

# Results

The results of these models clearly indicate that more severe electoral manipulation is associated with an increased risk that some degree of post-election protest will occur. Models 1 through 6 suggest that this relationship is driven more by parties' strategic calculations than by economic grievance. As Figure 1 shows, the predicted probability of any protest occurring is highest when fraud is substantial and the winner's vote-share low, in keeping with the strategic logic of protest. However, the evidence does not support the argument that fraud makes protest more likely when economic grievances are more widespread. Neither GDP growth, the economic crisis indicator, or the measure of public sector corruption are associated with higher protest risk when election fraud is more widespread. These models make clear that...

Turning to the results in Table 2, the models suggest that the severity of election manipulation has a significant relationship with the duration of post-election protest only in extreme cases (conditional on protests having begun). The election fraud coefficient is significant in Model 8; however, this relationship only holds when fraud is severe, and the winner's vote-share is very large---that is, in the most implausibly manipulated elections. The values of *intentional voting irregularities* used to calculate the marginal effects represent one standard deviation above and below the mean for that variable. For nearly all cases, there would be no difference in protest duration between a relatively clean and highly fraudulent election. One implication of this result could be that election results close to 100\% where there is evidence of fraud are so implausible as to be almost completely uninformative about the strength of the 'winning' party, leading to more uncertainty on all sides and longer protests (compared to an election with less fraud but similar margins). At extremely low winning vote-shares, the model suggests that clean elections are more likely to be protested. However, this result should be interpreted cautiously, as there are very few cases with such low winning vote-shares in this dataset, leading the model to engage in extrapolation.

```{r, echo=FALSE}
p2 + theme_bw()
hist(nelda2.sub$winner.share.upd)
```

The results from Model 8 do not comport with the standard view of the information / strategic model of election fraud, in which fraud in combination with low winning vote-shares leads to higher risk of damaging protest for the manipulator. Indeed, we observe the opposite relationship: when winning margins are low, *clean* elections are associated with lengthier protest. This suggests that the logic of power struggle and effectiveness are at work here---a low margin of victory and low level of fraud could indicate a genuinely weak incumbent, allowing for a longer struggle between the opposition and the government. At the same time, the significant effect at the high end of the winner's vote-share suggests a possible corrolary to the information / strategic view: when vote shares are implausibly high in the presence of extreme fraud, regime opponents may perceive genuine weakness as well.

On the other hand, just as with the results for the binary models, economic grievances do not appear associated with increased protest duration. There is no significant relationship between corruption, economic crisis, or GDP growth and protest at any level of election fraud. This further undermines the logic of the grievance model, which holds that fraud should help unite citizens with diverse grievances in a single campaign against the incumbent.

We do observe support for the more structural explanations of durable protest. The coefficient for the *legislative constraint* variable, which measures the size and oppositional character of the legislative opposition, is consistently positive and  statistically significant. In the baseline Model 7, the coefficient on *national party organization* is also positive and significant. The implication is that a larger, more institutionalized opposition party or coalition prolongs protest, in line with the 'test of strength' model. Similarly, the *incumbent term limit* variable is positive and consistently significant. This supports Hale's [-@hale_patronal_2014] argument that uncertainty around succession in non-democracies makes mobilization more likely. Finally, there is some evidence in Models 2 and 4 that greater degrees of judicial constraint on the executive are associated with reduced protest duration. While not conclusive, this suggests that courts may play a role in diffusing protest, perhaps by providing parties with an alternative venue to press their claims. ADD CITATIONS, THINK ABOUT ADDING GRIEVANCES LIT TOO (PROBLEM IS THAT GRIEVANCES DON'T MATTER).

<!-- Finally, the results are broadly similar for the number of protests that occur after the election. There is no significant relationship between the extent of fraud and the number of protests, though a smaller vote-share for the winner does appear to be associated with less protest. Neither GDP growth, corruption, or an economic crisis are significantly associated with protest number. Instead, once again, structural forces matter most. A large, active legislative opposition has a significant and positive relationship with protest size across all models, as does a term-limited chief executive. In two of the models based on economic conditions---Models 15 and 17---repression capability also appears to matter. A more permissive physical security environment is first associated with an increase in the number of protest, then a decrease. Finally, and perhaps surprisingly, a less restrictive environment for civil society or organizations is associated with a smaller number of post-election protests in the same models. -->

# Discussion

This is a pessimistic finding for proponents of democracy. While the results do find that there are circumstances in which more fraudulent elections are associated with a higher risk of *any* post-election protest occurring, this risk is limited primarily to cases in which the incumbent party wins by a narrow margin. With a large margin of victory, perhaps generated by extensive manipulation, protest appears much less likely. Moreover, the risk of protest does not appear to be heightened by poor economic conditions or intensive corruption in the public sector. As a result, circumstances that are to some extent outside of leaders' control, like a worsening macroeconomic climate, do not appear to make engaging in fraud any riskier. Of course, such conditions are likely to have an independent effect on incumbent's vote-share, but to an extent this is already 'priced in,' and incumbents may be able to effectively blunt their impact by artificially inflating their vote-share using fraud and other techniques.

More importantly, the severity of fraud appears almost wholly unrelated to the duration or number of post-election protests. Given that more widespread, frequent, or long-last protests pose a more significant challenge to non-democratic governments than isolated, rare, or brief events, these results suggest that the political cost of fraudulent elections is considerably lower than often assumed. Likewise, corruption and poor economic conditions do not appear associated with longer or more numerous protest.

Instead, the conditions that are correlated with more damaging protest are generally institutional factors, which are either more amenable to control or can be anticipated by non-democratic incumbents. For example, presidential elections appear to be associated with a greater risk of protest than legislative elections. Non-democratic leaders in presidential systems can mitigate this risk by adjusting the electoral calendar (i.e. to separate legislative and presidential elections, or by holding presidential elections early), or by ensuring that the state's coercive apparatus is fully in place prior to elections. Legislative constraints, an important factor for the duration and number of protests, can be adjusted through institutional reforms and cooptation of the opposition, over time. Finally, incumbent term limits can be managed through constitutional reforms, where possible, or through the appointment of (presumably) loyal successors. In other words, election fraud only has a minor cost in terms of protest risk, a risk which can be further blunted to the extent that elections can often be held on incumbents' terms.

None of this is to say that election fraud cannot, in theory, pose a cost to incumbents in the form of protest. This study is not a causal one; the decision by the incumbent to engage in fraud and the decision by opposition groups to protest are endogenous to one another. Instead, it shows that in practice, manipulated elections are not associated with longer or more numerous protests than clean ones. In typical cases, as a result, protest risk is likely not the constraining factor on the level of election fraud, contrary to the assumption in much of the formal theoretical literature. Rather, the constraints on manipulation are more likely to be driven by factors like resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022].

Consider also interacting fraud with rentier state status

# References

### 2022-07-20 Output/manipulation_protest_draft_v2.Rmd
Version prior to shift to quantity as DV

@@
 In sum---though with notable exceptions, like the People Power Revolution in the Philippines---it is generally protest *mobilization* that poses challenges to incumbents, rather than the occurrence of protest per se.
-Here, I integrate this understanding of protest's risk to regimes by using duration data, rather than binary indicators of protest.
+Here, I integrate this understanding of protest's risk to regimes by using count data of post-election protest events, rather than binary indicators of protest.
 
@@
 In most cases, the risk of protest has both been conceptualized and operationalized as a binary outcome: protests occur or they do not.
-This approach obscures the importance of the number and duration of electoral protests in a ruling party's calculation of the risk involved in manipulation an election.\footnote{One exception can be found in Hyde and Marinov's [-@hyde2014] paper, which considers protest duration in a secondary analysis.
-However, their duration analysis is a two-by-two table that does not control for possible confounding variables, and relies on a dataset that excludes un-protested elections that received positive observer reports.} Small, short-lived protests may be easily contained by a well-resourced government, while a long-lasting movement with a large number of protest events poses a more significant challenge.
+This approach obscures the importance of the number of electoral protests in a ruling party's calculation of the risk involved in manipulation an election.
+Small, short-lived protests may be easily contained by a well-resourced government, while a long-lasting movement with a large number of protest events poses a more significant challenge.
 
@@
 
-Sustained, sizeable protests may also contribute to elites' decisions to stand by the incumbent regime, or to side with "the people" in their own pursuit of power [@hale_patronal_2014;@svolik_politics_2012;@robertson_protest_2010].
+Sustained protests may also contribute to elites' decisions to stand by the incumbent regime, or to side with "the people" in their own pursuit of power [@hale_patronal_2014;@svolik_politics_2012;@robertson_protest_2010].
 
@@
 
-The theory proposed here is straightforward.
-First, it follows previous work in acknowledging that election manipulation is most likely to spark a protest movement under two conditions: when narrow margins of victory signal weakness [@Simpser2011], and when general social grievances are high [@Brancati2016;@tucker2007].
+Research on the duration of protest waves is limited [@Davenport2019], but the foregoing discussion suggests several baseline hypotheses to be tested in the the occurrence of post-election protest as a binary variable.
+These include, first, the prediction that more severe election fraud is a signal of deeper regime weakness (i.e.
+unpopularity), which makes protest more likely.
+Second, there is the alternative view that election manipulation is most likely to spark a protest movement when margins of victory are narrow [@Simpser2011].
+Finally, the grievances approach suggests that fraud will most likely lead to protest when social grievances are high [@Brancati2016;@tucker2007].
 
->H1: Post-election protest will be more likely to occur when fraud is high and ruling-party margin of victory is small, and less likely when fraud is low or the margin of victory large.
+>H1: Post-election protest will be more likely to occur when fraud is severe.
 
->H2: Post-election protest will be more likely to occur when fraud is high and economic and social grievances are elevated.
+>H2: Post-election protest will be more likely to occur when fraud is severe and ruling-party margin of victory is small, and less likely when fraud is low or the margin of victory large.
 
-However, once protests begin, the extent of electoral fraud is unlikely to affect their duration.
-The initial emerge of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@Tilly1977;@Lipsky1968;@Hummel2019;@Rasler1996;@Bishara2015;@Frantz2014].
+>H3: Post-election protest will be more likely to occur when fraud is severe and economic and social grievances are elevated.
+
+However, these predictions may not apply to the duration of a protest wave.
+Once protests begin, the extent of electoral fraud is unlikely to affect their duration.
+The initial emergence of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@Tilly1977;@Lipsky1968;@Hummel2019;@Rasler1996;@Bishara2015;@Frantz2014].
 (**See others**.) Given that post-electoral protests are a direct challenge to the regime, toleration is unlikely in these cases.
-Instead, incumbents usually begin a process of suppressing protests, repressing activists and movement leaders , and offering 'sideways concessions' to keep other citizens off the streets [@Hummel2019].\footnote{Of course, much repression is pre-emptive [@Sullivan2016].
-For example, non-democratic governments typically repress opposition activists in the run-up to elections[@Bhasin2013].} Because the emergence of protest indicates that the regime's pre-emptive efforts have failed, it is not clear a prior how the regime should handle protesters it failed to deter [@Ritter2016].
+That election dates are known in advance creates opportunities for opposition mobilization; anticipating this, regimes can more easily engage in pre-emptive repression [@Truex2019;@Sullivan2016;@Bhasin2013].
+Because the emergence of protest indicates that the regime's pre-emptive efforts have failed, it is not clear a priori how the regime should handle protesters it failed to deter [@Ritter2016].
 In many cases, the government doubles down on repression, calling upon the security services to disperse protests, arrest leaders, and perhaps employ harsher methods including disappearance, torture, and execution [@Escriba_Folch2013;@Davenport2007].
 Though such crackdowns risk provoking a backlash, they are aimed at demobilizing the opposition by increasing the costs of participation in overt dissent [@DeVogel2022].
-In other cases, the government may attempt to accommodate the opposition, by making concessions on policy or institutions [@Moore2000].
+In other cases, the government may attempt to accommodate the opposition, by making concessions on policy or institutions [@Moore2000;@Hafner-Burton2016;@Piven1979].
 Alternatively, concessions may be aimed at the more moderate public, in order to keep them on the sidelines of the conflict [@Hummel2019].
@@
 
-After the initial moment of opportunity and heightened grievance provided by the flawed election, I argue that protesters and potential movement supporters return to a typical status as "passionate economists" [@VanZomeren2012].
-In other words, protesters and potential protesters are motivated both by emotion in response to grievance, and by a more rational calculation of individual-level risk and group-level efficacy.
-Individuals evaluate whether their collective action is likely to bring about changes in response to their sense of grievance, and whether that likelihood outweighs the danger they may expose themselves to personally by protesting.
-Crucially, this process is dynamic, and can be updated over time in response to new information and new events [@VanZomeren2012].
-If protest movements have sufficient numbers and organizational resources to put the regime on the back foot, this can lead to longer-lasting and more numerous protest actions by increasing individuals' sense of group efficacy.
-On the other hand, if the regime is well-resourced, capable, and unified, its actions to put down protest can lead to a negative re-evaluation of the movement's odds of success.
+In this struggle between the potential protest coalition and the incumbent government, protesters and their possible allies behave as "passionate economists" [@VanZomeren2012]---they are motivated both by emotional responses to grievance, and by a more rational calculation of individual-level risk and group-level efficacy.
+What is the likelihood that protest action can help resolve the grievance, and does the benefit of participating outweigh the risk of personal consequences to the individual for participating?
+A fraudulent election may heighten citizens' grievances and generate uncertainty around the underlying popularity of the regime, creating a moment of elevated protest risk.
+Crucially, this evaluation is dynamic, and can be updated over time in response to new information and new events [@VanZomeren2012].
+If protest movements have sufficient numbers and organizational resources to put the regime on the back foot, this can lead to more durable protest movements by increasing individuals' sense of group efficacy and decreasing their sense of risk [@Chenoweth_Stephan2011].
+On the other hand, if the regime is well-resourced, capable, and unified, its actions to end a protest movement (including repression, concessions, and co-optation) can lead to a negative re-evaluation of the movement's odds of success.
 This does not, of course, eliminate citizens' sense of grievance.
-Instead, it leads them to abandon 'approach-oriented' forms of coping, to avoidance-oriented forms: demobilization and disengagement.
+Instead, it leads them to abandon 'approach-oriented' forms of coping, to avoidance-oriented forms: demobilization and disengagement [@van_zomeren_put_2004;@Young2019].
+For example, while ruling-party supporters in Russia disapprove of election fraud [@Reuter2021], they are more likely to withdraw from politics after exposure to electoral protest than to support the opposition [@Tertytchnaya2020].
 
->H3: Once protests begin, the severity of election fraud will not be correlated with the duration of protest.
+In summary, though election fraud may be taken as a signal by opposition groups, it is a noisy one: weak and strong governments alike may engage in election manipulation.
+However, once a protest movement begins, the level of election manipulation is no longer a useful signal of incumbent strength.
+Instead, protesters and bystanders can update their estimation of the relative strength of the protest coalition and the incumbent government by observing the size of crowds, the number of protests, and the government's reaction.
+Since the level of manipulation will no longer serve as an important indicator of regime strength once a challenge is taken to the streets, the duration of protests will instead be related to other indicators of opposition and incumbent capacity.
+As identified in prior research, these factors include the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], the margin of victory in the election [@simpser2013], and whether or not the chief executive faces term limits [@hale_patronal_2014].
 
-Since the level of manipulation will no longer serve as an important indicator of regime strength once a challenge is taken to the streets, the duration of protests will instead be related to other indicators of oppostion and incumbent capacity.
-As identified in prior research, these factors include the previous repressive history of the government, the institutionalization of opposition parties, the margin of victory in the election, and whether or not the chief executive faces term limits.
+>H4: Once protests begin, the severity of election fraud will not be correlated with the duration of protest.
 
--Democracy leads to less repression: Bueno de Mesquita et al., 2005). Davenport & Armstrong (2004), 
--Prior repression makes repression more likely (Davenport 2007 (Tyrannical peace))
--Resource wealth makes authoritarian survival more likely (Girod et al 2016)
+<!-- -Democracy leads to less repression: Bueno de Mesquita et al., 2005). Davenport & Armstrong (2004),  -->
+<!-- -Prior repression makes repression more likely (Davenport 2007 (Tyrannical peace)) -->
+<!-- -Resource wealth makes authoritarian survival more likely (Girod et al 2016) -->
 
-From E-F 2013: Nevertheless, violent repression may have pernicious consequences, since it may bring about a
-violent escalation or a radicalization of positions, thereby unleashing a violent response from some
-groups and even potentially provoking a coup (Gurr, 1970; Ortiz, 2007; Pierskalla, 2010). Thus,
-Opp and Ruehl (1990) suggest that repression may lead to micro-mobilization processes leading to
-further protest activities and unrest. Likewise, Ginkel and Smith suggest that ‘dissident activity is
-more likely to be effective in motivating rebellious action under highly repressive conditions’
-(1999: 292). Some argue that when nonviolent actions are violently repressed, protesters may turn
-to violent activities (Lichbach, 1987; Moore, 1998), as the Syrian case illustrates. Similarly,
-Goodwin (2001) argues that revolutions are more likely in patrimonial and very repressive states.
-For instance, in Iran on Black Friday in September 1978, the regime’s security forces killed hun-
-dreds of anti-shah protesters in Tehran. After a brief decrease in activity, public dissent and strikes
-quickly resumed and escalated. The mobilizations that took place in December gathered hundreds
-of thousands of people calling for the shah’s rule to end. In January 1979, the shah left the
-country
+<!-- From E-F 2013: Nevertheless, violent repression may have pernicious consequences, since it may bring about a -->
+<!-- violent escalation or a radicalization of positions, thereby unleashing a violent response from some -->
+<!-- groups and even potentially provoking a coup (Gurr, 1970; Ortiz, 2007; Pierskalla, 2010). Thus, -->
+<!-- Opp and Ruehl (1990) suggest that repression may lead to micro-mobilization processes leading to -->
+<!-- further protest activities and unrest. Likewise, Ginkel and Smith suggest that ‘dissident activity is -->
+<!-- more likely to be effective in motivating rebellious action under highly repressive conditions’ -->
+<!-- (1999: 292). Some argue that when nonviolent actions are violently repressed, protesters may turn -->
+<!-- to violent activities (Lichbach, 1987; Moore, 1998), as the Syrian case illustrates. Similarly, -->
+<!-- Goodwin (2001) argues that revolutions are more likely in patrimonial and very repressive states. -->
+<!-- For instance, in Iran on Black Friday in September 1978, the regime’s security forces killed hun- -->
+<!-- dreds of anti-shah protesters in Tehran. After a brief decrease in activity, public dissent and strikes -->
+<!-- quickly resumed and escalated. The mobilizations that took place in December gathered hundreds -->
+<!-- of thousands of people calling for the shah’s rule to end. In January 1979, the shah left the -->
+<!-- country -->
 
-...restrictions on civil rights can thwart the ability of some
-actors (especially the opposition) to organize, coordinate, and mobilize in the short run
+<!-- ...restrictions on civil rights can thwart the ability of some -->
+<!-- actors (especially the opposition) to organize, coordinate, and mobilize in the short run -->
 
-From de Vogel:  The strategic use of repression is founded in the fact that protests in general are undesirable
-to authoritarian regimes. Frantz and Kendall-Taylor (2014) demonstrate that popular revolution is
-an increasingly common avenue by which autocrats leave office, and some of these originate as
-everyday protests. Repression is often seen as the first line of defense against protest, given that
-it is inherent to autocracies (Wintrobe, 1998; Escrib`a-Folch, 2013), and it is consistently observed
-to increase in response to dissent (Davenport, 2007a). But cracking down on less threatening
-protest can create the appearance of weakness on the part of the regime, and repressing “average
-citizens”—like those who often make up the ranks of everyday protests—can increase dissent
-(Francisco, 1995; Rasler, 1996). On the local level where everyday protests often originate, control
-over the repressive apparatus may be less uncertain (Earl, 2011). That can lead to the use of “softer”
-forms of repression that are easier to control (Fu, 2018), or the use of non-state actors to execute
-it, which can undermine state legitimacy (Ong, 2018).
-On the other hand, the regime can respond to protests with concessions, but the costs and
-benefits of doing so are unclear. Concessions are conventionally associated with the cooptation
-of regime challengers, most often through sharing spoils, redistribution or patronage (Bueno de
-Mesquita and Smith, 2010), or via incorporation into authoritarian institutions such as legislatures
-(Lust-Okar, 2005; Gandhi and Przeworski, 2007). Lipsky (1968) classifies such concessions as
-protest bargaining, and underscores that they can only occur when the protesters have political
-resources the government wants. Everyday protest campaigns would therefore appear poor can-
-didates for such bargaining, given that they are may have few political resources at their disposal
-and are only sometimes organized by political parties
-
-On uncertainty of repression and protest, from Girod et al: Yet the differences in demobilization of protest movements following repression remain largely unexplained.
+<!-- On uncertainty of repression and protest, from Girod et al: Yet the differences in demobilization of protest movements following repression remain largely unexplained.
 Government repression of protesters is strongly associated with quelling protest activity in some cases (Lyall, 2009; Schatzman, 2005), and triggering increased mass mobilization in others (Francisco, 1995; Khawaja, 1993; Mason and Krane, 1989).
@@
 disbanding.
+-->
 
@@
 # Data and methods
-
-These hypotheses are tested using data from three datasets: NELDA [@hyde2012], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence (ECAV) dataset [@Daxecker2019].
-The unit of observation for this study is an election period as recorded in NELDA, which counts all national elections between 1945 and 2015.
-This framework prevents selecting on the dependent variable, by capturing both protested and non-protested elections.
-NELDA also includes a variable indicating elections in which incumbents lost; as this study is focused on protest against ostensibly victorious incumbents, I use this variable to filter out elections which the opposition won outright.
-I further limit the sample to regimes coded as electoral autocracies or electoral democracies in V-Dem.
-This excludes closed autocracies (where no national elections are held) and liberal democracies (where electoral integrity and rule of law are both high).
-Finally, the date range is limited by the availability of data on protest from ECAV, as discussed below, which runs from 1990 to 2012.
-The resulting dataset includes 647 election-period observations.
-
-ECAV, from which study's dependent variables are drawn, records protests and other contentious actions from six months before to three months after an election.
-For this study, I include protests only, excluding acts like shootings, bombings, and coups.
-ECAV data is in event-day format.
-To capture *protest duration*, I take the number of days between the first electoral protest and the last.
-The result is a count variable ranging from 0 (no protests) to 90.
-The large majority of elections experience no post-election protest: in 460 of the 647 elections, no protests are recorded.
-
-This distribution is in line with the theoretical proposition that two data-generating processes are at work: the factors that lead to the occurrence of protest may not be the same as those that sustain or instensify it.
-Consequently, I model the results in two stages.
-The first stage uses logit models to identify the factors that make *any protest* more likely to occur---this is a binary variable that is coded as 1 if any protest occurred after the election.
-Since election-year observations are nested within countries here, I use multilevel models with random intercepts by country in.
-As a second stage of analysis, I use negative binomial regressions to model the duration of protests given that protest has begun.
-That is, these latter models exclude cases where no protest occurred.
-For ease of interpretation, these processes are modeled and depicted separately in the main text, while hurdle regression models (which use the full dataset in one stage) are presented in the appendix, with similar results.
-
-Each model includes the explanatory variables proposed above.
-First, three proxy measures of economic grievance include an indicator of *public sector corruption* and the *lagged GDP growth rate*, both taken from V-Dem, and a binary measure of *economic crisis* from the NELDA dataset.
-Higher levels of corruption, negative GDP growth rates, and a positive value for *economic crisis* should each be expected to increase citizens' sense of grievance.
-To capture the *severity of election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than aggregate measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest---it is often incumbents' tool of last resort [@Sjoberg2016], it can fail to signal strength [@Harvey2020], and it can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021].
-Finally, the winning party's *margin of victory* is also taken from the V-Dem dataset.
-
-Several control variables are included, which could be common causes for both the severity of election manipulation and protest.
-To help control for the underlying mobilizational capacity, which likely affects incumbents' decision to manipulate and the latent risk of protest, I include indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem) and the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs').
-Likewise, the ability of the state to engage in repression is likely to embolden the ruling party and deter protests; I thus control for *physical integrity* using the 'physical violence index' from V-Dem lagged one year.
-Higher values for this variable indicate greater freedom from torture and political killing.
-Since intense repression may itself become a grievance that triggers further protest, I also include the variable's square term.
-Next, all models include three measures of the degree of political contestation in the country.
-The first of these is an indicator variable for *electoral democracy*; the reference category is electoral autocracy.
-The second and third are *judicial constraints* and  *legislative constraints* on the executive, V-Dem variables which captures the extent to which the judicial and legislative bodies in the country act independently in practice to limit executive power.Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019].
-
-In addition, several binary control variables are included.
-First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@Simpser2011] and a greater risk of protest.
-Next, a variable from NELDA indicates if the chief executive is *term limited* during the election at hand.
-This variable is included since incumbents running up against constitutional term limits can pose major challenges for the unity of the regime, possibly leading to fractured manipulation efforts and post-election protest [@hale_patronal_2014].
-Finally, I include dummy variables that provide a categorical evalution of state capacity.
-These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap).
-If the V-Dem data codes a state as not capable of raising revenue or primarily relying on loans and foreign aid, I code it as a *weak state*.
-If the state is described as primarily relying on resource rents and expropriation, I code it is a *rentier state*.
-Finally, states that rely primarily on taxes on property or economic transactions are coded as *tax states*.
-This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018].
 
@@
 
+#Successful protest = new elections or new incumbent
+
+nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))
+
 #Protest cases only
@@
 
+These hypotheses are tested using data from three datasets: NELDA [@hyde2012], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence (ECAV) dataset [@Daxecker2019].
+The unit of observation for this study is an election period as recorded in NELDA, which counts all national elections between 1945 and 2015.
+This framework prevents selecting on the dependent variable, by capturing both protested and non-protested elections.
+NELDA also includes a variable indicating elections in which incumbents lost; as this study is focused on protest against ostensibly victorious incumbents, I use this variable to filter out elections which the opposition won outright.
+I further limit the sample to regimes coded as electoral autocracies or electoral democracies in V-Dem.
+This excludes closed autocracies (where no national elections are held) and liberal democracies (where electoral integrity and rule of law are both high).
+Finally, the date range is limited by the availability of data on protest from ECAV, as discussed below, which runs from 1990 to 2012.
+The resulting dataset includes 647 election-period observations.
+
+ECAV, from which study's dependent variable is drawn, records protests and other contentious actions from six months before to three months after an election.
+For this study, I include protests only, excluding acts like shootings, bombings, and coups.
+ECAV data is in event-day format.
+To capture *protest duration*, I take the number of days between the first electoral protest and the last.
+The result is a count variable ranging from 0 (no protests) to 90.
+The large majority of elections experience no post-election protest: in 460 of the 647 elections, no protests are recorded.
+Figure 1 reports the distribution of protest duration across the dataset, as well as winners' vote-shares.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+
+ggplot(nelda2.sub, aes(x=n.events.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()
+
+ggplot(nelda2.sub, aes(x=winner.share.upd)) + geom_histogram(bins=100) + xlab("Incumbent vote-share") + theme_bw()
+
+#nelda2.sub.success <- nelda2.sub %>% filter(protest.success == 1)
+mu <- plyr::ddply(nelda2.sub, "protest.success", summarise, grp.mean=mean(n.events.post))
+head(mu)
+```
+
+```{r, include=FALSE}
+p.success <- ggplot(nelda2.sub, aes(x=n.events.post, fill = as.factor(protest.success), color=as.factor(protest.success)))+
+  geom_histogram(position="identity", alpha=0.5) + 
+  geom_vline(data=mu, aes(xintercept=grp.mean, color=as.factor(protest.success)),
+             linetype="dashed")
+p.success
+```
+
+This distribution is in line with the theoretical proposition that two data-generating processes are at work: the factors that lead to the occurrence of protest may not be the same as those that sustain or intensify it.
+Consequently, I model the results in two ways.
+The first approach uses logit models to identify the factors that make *any protest* more likely to occur---this is a binary variable that is coded as 1 if any protest occurred after the election.
+Since election-year observations are nested within countries here, I use multilevel models with random intercepts by country in.
+In the second analysis, I use negative binomial regressions to model the duration of protests given that protest has begun.
+That is, these latter models exclude cases where no protest occurred.
+For ease of interpretation, these processes are modeled and depicted separately in the main text, while hurdle regression models (which use the full dataset in one stage) are presented in the appendix, with similar results.
+
+Each model includes the explanatory variables proposed above.
+First, three proxy measures of economic grievance include an indicator of *public sector corruption* and the *lagged GDP growth rate*, both taken from V-Dem, and a binary measure of *economic crisis* from the NELDA dataset.
+Higher levels of corruption, negative GDP growth rates, and a positive value for *economic crisis* should each be expected to increase citizens' sense of grievance.
+To capture the *severity of election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than aggregate measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], it can fail to signal strength [@Harvey2020], and it can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021].
+Finally, the winning party's *vote-share* is also taken from the V-Dem dataset.
+
+Several control variables are included, which could be common causes for both the severity of election manipulation and protest.
+To help control for the underlying mobilizational capacity, which likely affects incumbents' decision to manipulate and the latent risk of protest, I include lagged indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem) and the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs').
+Likewise, the ability of the state to engage in repression is likely to embolden the ruling party and deter protests; I thus control for *physical integrity* using the 'physical violence index' from V-Dem lagged one year.
+Higher values for this variable indicate greater freedom from torture and political killing.
+Since intense repression may itself become a grievance that triggers further protest, I also include the variable's square term.
+
+Next, all models include three measures of the degree of political contestation in the country.
+The first of these is an indicator variable for *electoral democracy*; the reference category is electoral autocracy.
+The second and third are *judicial constraints* and  *legislative constraints* on the executive, V-Dem variables which captures the extent to which the judicial and legislative bodies in the country act independently in practice to limit executive power.
+Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019].
+
+Finally, several binary control variables are included.
+First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@Simpser2011] and a greater risk of protest.
+Next, a variable from NELDA indicates if the chief executive is *term limited* during the election at hand.
+This variable is included since incumbents running up against constitutional term limits can pose major challenges for the unity of the regime, possibly leading to fractured manipulation efforts and post-election protest [@hale_patronal_2014].
+Finally, I include dummy variables that provide a categorical evalution of state capacity.
+These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap).
+If the V-Dem data codes a state as not capable of raising revenue or primarily relying on loans and foreign aid, I code it as a *weak state*.
+If the state is described as primarily relying on resource rents and expropriation, I code it is a *rentier state*.
+Finally, states that rely primarily on taxes on property or economic transactions are coded as *tax states*.
+This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018].
+Figure 2 shows summary statistics for all of these variables.
+
 ```{r data-summary, echo=FALSE}
-incl.vars <- nelda2.sub %>% dplyr::select(jucon.1lag, legcon.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, any.ecav.protest, d.events.num)
+incl.vars <- nelda2.sub %>% dplyr::select(jucon.1lag, legcon.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, any.ecav.protest, n.events.post)
 
@@
 
-model.ecavbin.mm <- glm(inc.replaced  ~ n.events.post,  data = nelda2.sub)
+model.ecavbin.mm <- glm(inc.replaced.prot  ~ n.events.post,  data = nelda2.sub)
 #summary(model.ecavbin.mm)
@@
 ```
-```{r}
-p1 + theme_bw()
-```
 
@@
 #summary(model.d.griev)
-sjPlot::plot_model(model.d.griev, type = "pred", terms = c("gdpgro.1lag", "v2elirreg.inv")) + coord_cartesian(ylim = c(0, 200))
+sjPlot::plot_model(model.d.griev, type = "pred", terms = c("gdpgro.1lag", "v2elirreg.inv [.23, 2.12]")) + coord_cartesian(ylim = c(0, 200))
 
@@
 
+```{r tables-quantity, include=FALSE}
+models.quantity <- list("Model 13" = model.n.base, "Model 14" = model.n.base2, "Model 15" = model.n.griev, "Model 16" = model.n.griev2, "Model 17" = model.n.griev3)
+#modelsummary(tidytest)
+modelsummary(models.quantity, stars = T)
+```
+
+# Results
+
+The results of these models generally support the conventional wisdom that more severe electoral manipulation is associated with an increased risk that *any* post-election protest will occur.
+Model 1, which includes no interaction terms, shows a positive relationship between fraud and protest which approaches the conventional p < .05 level of statistical significance.
+Models 2 and 3 indicate that this relationship is driven more by parties' strategic calculations than by economic grievance.
+As Figure 3 shows, the predicted probability of any protest occurring is highest when fraud is substantial and the winner's vote-share low, in keeping with the strategic logic of protest advanced by Simpser [-@simpser2013].
+However, the evidence does not support the argument that fraud makes protest more likely when economic grievances are more widespread.
+Neither GDP growth, the economic crisis indicator, or the measure of public sector corruption are associated with higher protest risk when election fraud is more widespread.
+Figure 4 shows the marginal effect of the interaction of *economic crisis* and election fraud, demonstrating that there is no significant increase in the likelihood of protest associated with  higher levels of fraud during bad economic times.
+Similar, insignificant results obtain for measures of corruption and GDP growth.
+
 ```{r tables-binary, echo=FALSE}
@@
 
-modelsummary(models.binary, stars = T, coef_rename = c("v2elirreg.inv:winner.share.upd" = "Fraud x winner share",
+modelsummary(models.binary, stars = T, gof_omit = c("R2"), coef_rename = c("v2elirreg.inv:winner.share.upd" = "Fraud x winner share",
                                                        "pubseccorrup.lag:v2elirreg.inv" = "Fraud x corruption",
                                                        "econ.crisis.nelda.1lag1:v2elirreg.inv" = "Fraud x econ. crisis"))
+```
+
+```{r, echo=FALSE}
+p1 + theme_bw()
+```
+
+```{r, echo=FALSE}
+sjPlot::plot_model(model.ecavbin.mm.crisis, type = "pred", terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-.17, 2.01]")) + theme_bw()
 ```
@@
 
-```{r tables-quantity, include=FALSE}
-models.quantity <- list("Model 13" = model.n.base, "Model 14" = model.n.base2, "Model 15" = model.n.griev, "Model 16" = model.n.griev2, "Model 17" = model.n.griev3)
-#modelsummary(tidytest)
+```{r}
+models.quantity <- list("Model 8" = model.n.base, "Model 9" = model.n.base2, "Model 10" = model.n.griev, "Model 11" = model.n.griev2, "Model 12" = model.n.griev3)
 modelsummary(models.quantity, stars = T)
@@
 
-# Results
+Turning to models of protest duration presented in Table 3, there is no statistically significant relationship between fraud and duration in the models of economic grievance (Models 10 through 12) or in the model without any interaction term (Model 8).
+Only in Model 9, where the severity of fraud is interacted with the winner's vote-share, is a statistically significant relationship observed.
+The models suggest that the severity of election manipulation has a significant relationship with the duration of post-election protest only in extreme cases (conditional on protests having begun).
+However, this relationship only holds for the most extreme cases---when winning vote-shares are either very small or very large.
+The values of *intentional voting irregularities* used to calculate the marginal effects in Figure 4 represent one standard deviation above and below the mean for that variable.
+For nearly all cases, there is no expected difference in protest duration between a relatively clean and highly fraudulent election.
 
-The results of these models clearly indicate that more severe electoral manipulation is associated with an increased risk that some degree of post-election protest will occur.
-Models 1 through 6 suggest that this relationship is driven more by parties' strategic calculations than by economic grievance.
-As Figure 1 shows, the predicted probability of any protest occurring is highest when fraud is substantial and the winner's vote-share low, in keeping with the strategic logic of protest.
-However, the evidence does not support the argument that fraud makes protest more likely when economic grievances are more widespread.
-Neither GDP growth, the economic crisis indicator, or the measure of public sector corruption are associated with higher protest risk when election fraud is more widespread.
-These models make clear that...
-
-Turning to the results in Table 2, the models suggest that the severity of election manipulation has a significant relationship with the duration of post-election protest only in extreme cases (conditional on protests having begun).
-The election fraud coefficient is significant in Model 8; however, this relationship only holds when fraud is severe, and the winner's vote-share is very large---that is, in the most implausibly manipulated elections.
-The values of *intentional voting irregularities* used to calculate the marginal effects represent one standard deviation above and below the mean for that variable.
-For nearly all cases, there would be no difference in protest duration between a relatively clean and highly fraudulent election.
-One implication of this result could be that election results close to 100\% where there is evidence of fraud are so implausible as to be almost completely uninformative about the strength of the 'winning' party, leading to more uncertainty on all sides and longer protests (compared to an election with less fraud but similar margins).
-At extremely low winning vote-shares, the model suggests that clean elections are more likely to be protested.
-However, this result should be interpreted cautiously, as there are very few cases with such low winning vote-shares in this dataset, leading the model to engage in extrapolation.
+Three other salient observations can be taken from the marginal effects.
+First, the expected duration of protest remains essentially constant for a relatively fraudulent election, regardless of the margin of victory.
+Second, with the caveat that the result is based on a small number of observations, it appears that the expected duration of a protest movement following a narrow incumbent victory in a relatively clean election is substantially longer than would be expected in a fraudulent election.
+Lastly, while there is a positive expected protest duration in high-margin, high-fraud elections, the expected duration approaches zero when a high-margin election is relatively clean.
+Taken together, all of this supports an understanding of fraud as a noisy signal.
+When an incumbent soundly defeats the opposition without resorting to fraud, the incumbent is unambiguously strong---a deterrent to protest \footnote{This is not to say that the regime is unambiguously \textit{popular}, since a large body of citizens may oppose the regime but sit out the election by choice or by force.}.
+Likewise, when fraud is absent and the incumbent's vote-share is small, the regime is unambiguously weak.
+Fraud muddles the picture, however, leading to instances where organizations and citizens are uncertain of the incumbent's true strength.
+As a result, strong incumbents sometimes face protest movements that they can quickly demobilize.
+It is also possible that fraud has a protective function for incumbents in low-margin elections, by making the sources of the regime's strength uncertain, though this result has a limited base of support.
 
 ```{r, echo=FALSE}
-p2 + theme_bw()
-hist(nelda2.sub$winner.share.upd)
+p2 + geom_rug(sides = "b")  #Adding rug plot breaks theme_bw
+
 ```
-
-The results from Model 8 do not comport with the standard view of the information / strategic model of election fraud, in which fraud in combination with low winning vote-shares leads to higher risk of damaging protest for the manipulator.
-Indeed, we observe the opposite relationship: when winning margins are low, *clean* elections are associated with lengthier protest.
-This suggests that the logic of power struggle and effectiveness are at work here---a low margin of victory and low level of fraud could indicate a genuinely weak incumbent, allowing for a longer struggle between the opposition and the government.
-At the same time, the significant effect at the high end of the winner's vote-share suggests a possible corrolary to the information / strategic view: when vote shares are implausibly high in the presence of extreme fraud, regime opponents may perceive genuine weakness as well.
-
-On the other hand, just as with the results for the binary models, economic grievances do not appear associated with increased protest duration.
-There is no significant relationship between corruption, economic crisis, or GDP growth and protest at any level of election fraud.
-This further undermines the logic of the grievance model, which holds that fraud should help unite citizens with diverse grievances in a single campaign against the incumbent.
 
@@
 The coefficient for the *legislative constraint* variable, which measures the size and oppositional character of the legislative opposition, is consistently positive and  statistically significant.
-In the baseline Model 7, the coefficient on *national party organization* is also positive and significant.
+In Models 8 and 9 (those without a grievance interaction), the coefficient on *national party organization* is also positive and significant.
 The implication is that a larger, more institutionalized opposition party or coalition prolongs protest, in line with the 'test of strength' model.
 Similarly, the *incumbent term limit* variable is positive and consistently significant.
-This supports Hale's [-@hale_patronal_2014] argument that uncertainty around succession in non-democracies makes mobilization more likely.
-Finally, there is some evidence in Models 2 and 4 that greater degrees of judicial constraint on the executive are associated with reduced protest duration.
-While not conclusive, this suggests that courts may play a role in diffusing protest, perhaps by providing parties with an alternative venue to press their claims.
-ADD CITATIONS, THINK ABOUT ADDING GRIEVANCES LIT TOO (PROBLEM IS THAT GRIEVANCES DON'T MATTER).
+This supports Hale's [-@hale_patronal_2014] argument that uncertainty around succession in non-democracies makes mobilization more intense as competing power centers emerge.
+Finally, there is some evidence in Models 9 and 11 that greater degrees of judicial constraint on the executive are associated with reduced protest duration.
+While not conclusive, this suggests that courts may play a role in diffusing protest, perhaps by providing parties with an alternative venue to press their claims [@Moustafa2014a;@Epperly2019c].
 
@@
 
-This is a pessimistic finding for proponents of democracy.
-While the results do find that there are circumstances in which more fraudulent elections are associated with a higher risk of *any* post-election protest occurring, this risk is limited primarily to cases in which the incumbent party wins by a narrow margin.
-With a large margin of victory, perhaps generated by extensive manipulation, protest appears much less likely.
+The above results make several contributions.
+First, there is some evidence that worsening fraud is associated with a greater likelihood that some post-election protest will occur.
+In particular, there is support for the argument that fraudulent elections are associated with a greater risk of protest when the incumbent wins by a narrow margin.
+However, the evidence does not support the claim that fraud is more likely to catalyze protest when economic conditions are poor.
+This is consistent with a strategic, rather than grievance-based, model of electoral protest.
+In other words, the results in Table 2 support a model of electoral protest in which organizations and potential protesters use the extent of fraud as a signal of the political strength of the incumbent, an evaluation which occurs largely independent of economic grievances.
+
+However, the results from the models of duration support the argument that election fraud is a noisy and short-lived signal of incumbent strength.
+In most models, the extent of election fraud has no significant association with the duration of protest.
+Instead, duration appears to be driven largely by institutional indicators of the balance of power between the opposition and incumbent---in particular, the size of the legislative opposition, and whether or not the chief executive is term limited.
+Just as in the binary models, grievances do not appear to make longer-lasting protest more likely in fraudulent elections.
+The results from Model 9 complicate the support for the strategic model found in Table 2, since fraud is not associated with longer protest in a low-margin election.
+The combination of the two results suggests that much strategic electoral protest is a gamble that opposition groups lose; anticipating a weak incumbent, they instead find one that is capable of repressing, coopting, or conciliating an emerging protest movement.
+Indeed, while there are 187 cases of protest out of the 647 elections in this dataset, only 15 elections resulted in either the incumbent being replaced due to protest or the election being re-run.\footnote{This information is determined using the NELDA variables NELDA37 and NELDA41.
+The cases of successful protest are: Ukraine 2004, Azerbaijan 2005, Georgia 2003, Kazakhstan 1994, Kyrgyzstan 2005, Bulgaria 1990, Malaysia 2004, Bangladesh 1996, Mali 1997, Peru 2000, Burkina Faso 1997, and Cote d'Ivoire 2000.}
+
+These are pessimistic findings for proponents of democracy.
+While the results do find that there are circumstances in which more fraudulent elections are associated with a higher risk of *some* post-election protest occurring, this risk is limited primarily to cases in which the incumbent party wins by a narrow margin.
+Though protests may occur in such settings, they are no more likely to persist than in  With a large margin of victory, perhaps generated by extensive manipulation, protest appears much less likely.
 Moreover, the risk of protest does not appear to be heightened by poor economic conditions or intensive corruption in the public sector.

### 2022-07-20 Scripts/data_cleaning.Rmd
Version prior to shift to quantity as DV

@@
 nelda <- nelda %>% rename(inc.lose = nelda24)
+nelda <- nelda %>% rename(inc.termlimit = nelda8)
+nelda <- nelda %>% rename(econ.crisis.nelda = nelda18)
+nelda <- nelda %>% rename(opp.votegain = nelda27)
+nelda <- nelda %>% rename(nelda.protests.fraud = nelda30)
+nelda <- nelda %>% rename(new.elections.held = nelda37)
+nelda <- nelda %>% rename(inc.replaced.prot = nelda41)
 
+nelda <- nelda %>% mutate(first.mp.election = ifelse(first.mp.election == "yes", 1, ifelse(first.mp.election == "no", 0, NA)))
 nelda <- nelda %>% mutate(protests = ifelse(protests == "yes", 1, ifelse(protests == "no", 0, NA)))
@@
 
+nelda <- nelda %>% mutate(inc.termlimit.num = ifelse(types == "Legislative/Parliamentary" & inc.termlimit == "N/A", 0, ifelse(types == "Executive" & inc.termlimit == "yes", 1,                                                                                                                       ifelse(types == "Executive" & inc.termlimit == "no", 0, NA))))
+
+nelda <- nelda %>% mutate(econ.crisis.nelda = ifelse(econ.crisis.nelda == "yes", 1, ifelse(econ.crisis.nelda == "no", 0, NA)))
+nelda <- nelda %>% mutate(opp.votegain = ifelse(opp.votegain == "yes", 1, ifelse(opp.votegain == "no", 0, NA)))
+nelda <- nelda %>% mutate(nelda.protests.fraud = ifelse(protests == 0, 0, ifelse(nelda.protests.fraud == "yes", 1, ifelse(nelda.protests.fraud == "no", 0, NA))))
+nelda <- nelda %>% mutate(inc.replaced.prot = ifelse(inc.replaced.prot == "yes", 1, 0))
+nelda <- nelda %>% mutate(new.elections.held = ifelse(new.elections.held == "yes", 1, 0))
+
+```
+
+Coding anti-gov violence from ecav
+
+```{r}
+## Coding anti-government initiation of violence
+
+ecav <- ecav %>% mutate(antigov.violence = 
+                          ifelse(EventViolence == 0, 0,
+                                 ifelse(ViolenceInitiator == "-99", NA,
+                                                  ifelse(ViolenceInitiator == ".1" & Actor1Side == 1, 1, 
+                                                  ifelse(ViolenceInitiator == ".2" & Actor2Side == 1, 1,
+                                                         ifelse(ViolenceInitiator == "1.1" & Target1Side == 1, 1,
+                                                                ifelse(ViolenceInitiator == "2.1" & Target2Side == 1, 1, 0)))))))
 ```
@@
 vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(physinteg.1lag = lag(v2x_clphy))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(legcorrupt.lag = lag(v3lgcrrpt))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(exbribe.lag = lag(v2exbribe))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(exembez.lag = lag(v2exembez))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(pubseccorrup.lag = lag(v2excrptps))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(partybarriers.lag = lag(v2psbars))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(natlpartyorg.lag = lag(v2psorgs))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(partybranch.lag = lag(v2psprbrch))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(csrepress.1lag = lag(v2csreprss))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(osorg.1lag = lag(v2cseeorgs))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(civlib.1lag = lag(v2x_clpol))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jucon.1lag = lag(v2x_jucon))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(legcon.1lag = lag(v2xlg_legcon))
+
 vdem.recent <- vdem.recent %>% group_by(COWcode) %>% mutate(opposition.oversight.lag = lag(v2lgoppart))
@@
 
+#Lagging econ crisis
+
+nelda <- nelda %>% group_by(country_id) %>% mutate(econ.crisis.nelda.1lag = lag(econ.crisis.nelda))
+
+#Merging
 nelda2 <- nelda %>% merge(lji %>% dplyr::select(LJI, lji.lag, country_id_year), by = "country_id_year")
@@
                                                   v2xeg_eqaccess,
-
+                                                  v3lgcrrpt,
+                                                  legcorrupt.lag,
+                                                  exbribe.lag,
+                                                  v2exbribe,
+                                                  exembez.lag,
+                                                  v2exembez,
+                                                  pubseccorrup.lag,
+                                                  v2excrptps,
+                                                  partybarriers.lag,
+                                                  v2psbars,
+                                                  natlpartyorg.lag,
+                                                  v2psorgs,
+                                                  partybranch.lag,
+                                                  v2psprbrch,
+                                                  v2stfisccap_ord,
+                                                  v2stfisccap,
+                                                  v2pepwrses, #Power distribution by wealth
+                                                  v2pepwrsoc, #Power distribution by social group
+                                                  v2csreprss,
+                                                  csrepress.1lag,
+                                                  osorg.1lag,
+                                                  v2cseeorgs,
+                                                  v2x_clpol,
+                                                  civlib.1lag,
+                                                  jucon.1lag,
+                                                  legcon.1lag,
 egaldem.1lag ,
@@
 nelda2 <- nelda2 %>% mutate(presidential = ifelse(types == "Executive", 1, 0))
+
+#
+
+nelda2 <- nelda2 %>% mutate(weakstate = ifelse(is.na(v2stfisccap_ord) == T, NA, ifelse(v2stfisccap_ord == 0 | v2stfisccap_ord == 1, 1, 0)))
+nelda2 <- nelda2 %>% mutate(rentierstate = ifelse(is.na(v2stfisccap_ord) == T, NA, ifelse(v2stfisccap_ord == 2, 1, 0)))
+nelda2 <- nelda2 %>% mutate(taxstate = ifelse(is.na(v2stfisccap_ord) == T, NA, ifelse(v2stfisccap_ord == 3 | v2stfisccap_ord == 4, 1, 0)))
 
@@
 
+## Violence
+violence_data <- ecav %>% group_by(NeldaID) %>% summarize(antigov.violence = sum(antigov.violence, na.rm = T))
+violence_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(antigov.violence.pre = sum(antigov.violence, na.rm = T))
+
 ##Combined
@@
 ecav_data <- left_join(ecav_data, size_data, by = "NeldaID")
+ecav_data <- left_join(ecav_data, violence_data, by = "NeldaID")
 
 ecav_data_pre <- left_join(count_data_pre, size_data_pre, by = "NeldaID")
+ecav_data_pre <- left_join(ecav_data_pre, violence_data_pre, by = "NeldaID")
+
 ecav_data <- left_join(ecav_data, ecav_data_pre, by = "NeldaID")
@@
 
+nelda2.sub <- nelda2.sub %>% mutate(antigov.violence = ifelse(is.na(antigov.violence)==T, 0, antigov.violence))
 nelda2.sub <- nelda2.sub %>% mutate(n.events.post = ifelse(is.na(n.events.post)==T, 0, n.events.post))
@@
 nelda2.sub <- nelda2.sub %>% mutate(n.events.pre = ifelse(is.na(n.events.pre) == T, 0, n.events.pre))
+nelda2.sub <- nelda2.sub %>% mutate(antigov.violence.pre = ifelse(is.na(antigov.violence.pre) == T, 0, antigov.violence.pre))
 
@@
 
+## World Bank urbanization data
+
+```{r}
+urban <- read.csv(here::here("Data", "urbanization_worldbank.csv"))
+test <- urban %>%
+  pivot_longer(cols = starts_with("X"),
+               names_to = "year")
+test <- test %>% mutate(year = gsub("X", "", year))
+
+#Next is line up country codes in test and nelda 2, then merge
+
+library(countrycode)
+
+test <- test %>% mutate(stateid = countrycode(sourcevar = Country.Code, origin = 'iso3c', destination = 'gwc'))
+test <- test %>% rename(urban.pct = value)
+test <- test %>% dplyr::select(year, urban.pct, stateid)
+
+nelda2.sub <- nelda2.sub %>% mutate(year = as.character(year))
+
+nelda2.sub <- nelda2.sub %>% left_join(test, by = c("stateid", "year"))
+
+```
+
+## Merge with winner share updated
+
+```{r}
+winner.shares <- read.csv(here::here("Data", "amended_vote_shares.csv"))
+nelda2.sub <- nelda2.sub %>% left_join(winner.shares, by = "NeldaID")
+```
+
 ## Writing

### 2022-07-20 Output/manipulation_protest_draft_v2.Rmd
Finished rough draft of v3 of paper (DV: quantity of protest)

@@
 ---
-title: "Draft: Judicial independence and post-election protest"
+title: "Draft: It's Not So Bad---The minimal effect of election fraud on protest intensity in non-democracies"
 author: "Cole J. Harvey (Oklahoma State University)"
@@
 This is in keeping with the formal-modeling literature, which predicts that overreach in the form of election fraud indicates government weakness, making opposition groups confident of winning a conflict in the streets.
-However, there is no statistically significant association between the level of election fraud and the duration of protest.
+However, there is no statistically significant association between the level of election fraud and the number of protest events.
 This suggests a more complex relationship between election fraud and protest.
 One the one hand, the result is consistent with the prediction of some formal models that fraud is a noisy signal---in some cases, opposition groups may misperceive a strong government as weak.
-On the other hand, since fraud is not associated with the overall duration of a protest wave, the risk to governments has likely been overemphasized.
+On the other hand, since fraud is not associated with the overall number of events in a protest wave, the risk to governments has likely been overemphasized.
 That is, even if protests do break out in response to large-scale fraud, the government has many tools it may deploy to curtail them.
-To the extent that longer protest waves are more damaging to the government than short ones, protest risk does not appear to be a substantial constraint on the severity of election fraud.
+To the extent that larger protest waves are more damaging to the government than small ones, protest risk does not appear to be a substantial constraint on the severity of election fraud.
 
@@
 
-Sustained protests may also contribute to elites' decisions to stand by the incumbent regime, or to side with "the people" in their own pursuit of power [@hale_patronal_2014;@svolik_politics_2012;@robertson_protest_2010].
+Large, sustained protest waves may also contribute to elites' decisions to stand by the incumbent regime, or to side with "the people" in their own pursuit of power [@hale_patronal_2014;@svolik_politics_2012;@robertson_protest_2010].
 
@@
 The apparent failure of protest risk to deter election fraud underlines our limited understanding of non-democratic ruling parties' decision-making.
-Does the extent of fraud, or the severity of underlying social grievances, make long-lasting protest more likely (and thus restrain fraud in anticipation)?
+Does the extent of fraud, or the severity of underlying social grievances, make more numerous (and thus more costly) protest more likely?
 
@@
 
-Research on the duration of protest waves is limited [@Davenport2019], but the foregoing discussion suggests several baseline hypotheses to be tested in the the occurrence of post-election protest as a binary variable.
+With some exceptions, such as the study by Lankina and Skovoroda [-@Lankina2017], most research on election fraud and protest treats the latter as a binary outcome.
+The foregoing discussion of the literature suggests several baseline hypotheses to be tested regarding post-election protest as a binary variable.
 These include, first, the prediction that more severe election fraud is a signal of deeper regime weakness (i.e.
@@
 
-However, these predictions may not apply to the duration of a protest wave.
-Once protests begin, the extent of electoral fraud is unlikely to affect their duration.
+However, these predictions may not apply to the size or intensity of a protest wave.
+Once protests begin, the extent of electoral fraud is unlikely to affect their scope.
 The initial emergence of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@Tilly1977;@Lipsky1968;@Hummel2019;@Rasler1996;@Bishara2015;@Frantz2014].
@@
 A fraudulent election may heighten citizens' grievances and generate uncertainty around the underlying popularity of the regime, creating a moment of elevated protest risk.
-Crucially, this evaluation is dynamic, and can be updated over time in response to new information and new events [@VanZomeren2012].
-If protest movements have sufficient numbers and organizational resources to put the regime on the back foot, this can lead to more durable protest movements by increasing individuals' sense of group efficacy and decreasing their sense of risk [@Chenoweth_Stephan2011].
+Crucially, though, this evaluation is dynamic and can be updated over time in response to new information and new events [@VanZomeren2012].
+If protest movements have sufficient numbers and organizational resources to put the regime on the back foot, this can lead to more numerous protests as individuals' sense of group efficacy grows and their sense of risk declines [@Chenoweth_Stephan2011].
 On the other hand, if the regime is well-resourced, capable, and unified, its actions to end a protest movement (including repression, concessions, and co-optation) can lead to a negative re-evaluation of the movement's odds of success.
@@
 Instead, protesters and bystanders can update their estimation of the relative strength of the protest coalition and the incumbent government by observing the size of crowds, the number of protests, and the government's reaction.
-Since the level of manipulation will no longer serve as an important indicator of regime strength once a challenge is taken to the streets, the duration of protests will instead be related to other indicators of opposition and incumbent capacity.
+Since the level of manipulation will no longer serve as an important indicator of regime strength once a challenge is taken to the streets, the number of protests will instead be related to other indicators of opposition and incumbent capacity.
 As identified in prior research, these factors include the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], the margin of victory in the election [@simpser2013], and whether or not the chief executive faces term limits [@hale_patronal_2014].
+Generally speaking, these structural variables indicate aspects of the political opportunity structure that make protest more likely to occur [@meyer_protest_2004].
 
->H4: Once protests begin, the severity of election fraud will not be correlated with the duration of protest.
+>H4: Once protests begin, the severity of election fraud will not be correlated with the number of protest events.
 
@@
 ECAV data is in event-day format.
-To capture *protest duration*, I take the number of days between the first electoral protest and the last.
-The result is a count variable ranging from 0 (no protests) to 90.
-The large majority of elections experience no post-election protest: in 460 of the 647 elections, no protests are recorded.
-Figure 1 reports the distribution of protest duration across the dataset, as well as winners' vote-shares.
+To capture the *number of protests*, I take the sum number of events recorded in ECAV from election day to the end of the dataset's coverage period ninety days later.
+The result is a count variable ranging from zero to 90.\footnote{That the max value for the quantity of protests matches the number of days in the recording window is coincidental.} The large majority of elections experience no post-election protest: in 460 of the 647 elections, no protests are recorded.
+Figure 1 reports the distribution of protest counts across the dataset, as well as winners' vote-shares.
 
@@
 Since election-year observations are nested within countries here, I use multilevel models with random intercepts by country in.
-In the second analysis, I use negative binomial regressions to model the duration of protests given that protest has begun.
+In the second analysis, I use negative binomial regressions to model the count of protests given that protest has begun.
 That is, these latter models exclude cases where no protest occurred.
@@
 
-Next, all models include three measures of the degree of political contestation in the country.
+Next, all models include three measures of the degree of political openness in the country.
 The first of these is an indicator variable for *electoral democracy*; the reference category is electoral autocracy.
@@
 #summary(model.d.base2)
-p2 <- sjPlot::plot_model(model.d.base2, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.36, 1.85]")) + coord_cartesian(ylim = c(0, 200))
+#p2 <- sjPlot::plot_model(model.d.base2, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.36, 1.85]")) + coord_cartesian(ylim = c(0, 200))
 
@@
 #summary(model.n.base2)
-sjPlot::plot_model(model.n.base2, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-1, 1.79]"))
+p2 <- sjPlot::plot_model(model.n.base2, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.36, 1.85]")) + coord_cartesian(ylim = c(0, 25))
 
@@
 
-```{r tables-quantity, include=FALSE}
-models.quantity <- list("Model 13" = model.n.base, "Model 14" = model.n.base2, "Model 15" = model.n.griev, "Model 16" = model.n.griev2, "Model 17" = model.n.griev3)
+```{r, include=FALSE}
+#models.quantity <- list("Model 13" = model.n.base, "Model 14" = model.n.base2, "Model 15" = model.n.griev, "Model 16" = model.n.griev2, "Model 17" = model.n.griev3)
 #modelsummary(tidytest)
-modelsummary(models.quantity, stars = T)
+#modelsummary(models.quantity, stars = T)
 ```
@@
 
-```{r tables-duration, results='asis', echo=FALSE}
+```{r tables-duration, results='asis', echo=FALSE, include=FALSE}
 models.duration <- list("Model 8" = model.d.base, "Model 9" = model.d.base2, "Model 10" = model.d.griev, "Model 11" = model.d.griev2, "Model 12" = model.d.griev3)
@@
 
-```{r}
+```{r tables-quantityr}
 models.quantity <- list("Model 8" = model.n.base, "Model 9" = model.n.base2, "Model 10" = model.n.griev, "Model 11" = model.n.griev2, "Model 12" = model.n.griev3)
@@
 
-Turning to models of protest duration presented in Table 3, there is no statistically significant relationship between fraud and duration in the models of economic grievance (Models 10 through 12) or in the model without any interaction term (Model 8).
-Only in Model 9, where the severity of fraud is interacted with the winner's vote-share, is a statistically significant relationship observed.
-The models suggest that the severity of election manipulation has a significant relationship with the duration of post-election protest only in extreme cases (conditional on protests having begun).
-However, this relationship only holds for the most extreme cases---when winning vote-shares are either very small or very large.
-The values of *intentional voting irregularities* used to calculate the marginal effects in Figure 4 represent one standard deviation above and below the mean for that variable.
-For nearly all cases, there is no expected difference in protest duration between a relatively clean and highly fraudulent election.
-
-Three other salient observations can be taken from the marginal effects.
-First, the expected duration of protest remains essentially constant for a relatively fraudulent election, regardless of the margin of victory.
-Second, with the caveat that the result is based on a small number of observations, it appears that the expected duration of a protest movement following a narrow incumbent victory in a relatively clean election is substantially longer than would be expected in a fraudulent election.
-Lastly, while there is a positive expected protest duration in high-margin, high-fraud elections, the expected duration approaches zero when a high-margin election is relatively clean.
-Taken together, all of this supports an understanding of fraud as a noisy signal.
-When an incumbent soundly defeats the opposition without resorting to fraud, the incumbent is unambiguously strong---a deterrent to protest \footnote{This is not to say that the regime is unambiguously \textit{popular}, since a large body of citizens may oppose the regime but sit out the election by choice or by force.}.
-Likewise, when fraud is absent and the incumbent's vote-share is small, the regime is unambiguously weak.
-Fraud muddles the picture, however, leading to instances where organizations and citizens are uncertain of the incumbent's true strength.
-As a result, strong incumbents sometimes face protest movements that they can quickly demobilize.
-It is also possible that fraud has a protective function for incumbents in low-margin elections, by making the sources of the regime's strength uncertain, though this result has a limited base of support.
-
-```{r, echo=FALSE}
-p2 + geom_rug(sides = "b")  #Adding rug plot breaks theme_bw
-
-```
+Turning to models of protest quantity presented in Table 3, there is no statistically significant relationship between fraud and protest in any of the models.
+Figure 4, using results from Model 9, presents the marginal effects for fraud conditional on the incumbent's vote-share.
+As the figure shows, there is no statistically significant difference in the expected number of protest following a relatively clean election compared to a relatively fraudulent one, for any margin of incumbent victory.
+Taking the results from Tables 2 and 3 together, there is support for Hypothesis 1 (an unconditional increase in the risk of protest as fraud increases), as well as strong support for Hypothesis 2 (an increase in protest risk in fraudulent, low-margin elections) and Hypothesis 4 (no relationship between fraud and protest quantity).
+Hypothesis 3, on the connection between fraud, grievance, and protest, is not supported.
 
@@
 This supports Hale's [-@hale_patronal_2014] argument that uncertainty around succession in non-democracies makes mobilization more intense as competing power centers emerge.
-Finally, there is some evidence in Models 9 and 11 that greater degrees of judicial constraint on the executive are associated with reduced protest duration.
-While not conclusive, this suggests that courts may play a role in diffusing protest, perhaps by providing parties with an alternative venue to press their claims [@Moustafa2014a;@Epperly2019c].
+
+```{r, echo=FALSE}
+p2 + geom_rug(sides = "b")  #Adding rug plot breaks theme_bw
+
+```
 
@@
 
-The above results make several contributions.
-First, there is some evidence that worsening fraud is associated with a greater likelihood that some post-election protest will occur.
-In particular, there is support for the argument that fraudulent elections are associated with a greater risk of protest when the incumbent wins by a narrow margin.
-However, the evidence does not support the claim that fraud is more likely to catalyze protest when economic conditions are poor.
-This is consistent with a strategic, rather than grievance-based, model of electoral protest.
-In other words, the results in Table 2 support a model of electoral protest in which organizations and potential protesters use the extent of fraud as a signal of the political strength of the incumbent, an evaluation which occurs largely independent of economic grievances.
+Altogether, the findings support an understanding of fraud as a signal of incumbent strength, but a noisy and unreliable one.
+When an incumbent party wins by a narrow margin in a fraudulent election, the likelihood of protest initiation is substantially higher than when it wins more cleanly: roughly a 75\% chance in the former case versus 15\% in the latter as shown in Figure 2.
+However, since even strong incumbents commit fraud, the decision to take to the streets is often a losing gamble---anticipating a weak incumbent, they instead find one that is capable of repressing, coopting, or conciliating an emerging protest movement.
+Indeed, while there are 187 cases of protest out of the 647 elections in this dataset, only 15 elections resulted in either the incumbent being replaced due to protest or the election being re-run.\footnote{This information is determined using the NELDA variables NELDA37 and NELDA41.
+The cases of successful protest are: Ukraine 2004, Azerbaijan 2005, Georgia 2003, Kazakhstan 1994, Kyrgyzstan 2005, Bulgaria 1990, Malaysia 2004, Bangladesh 1996, Mali 1997, Peru 2000, Burkina Faso 1997, and Cote d'Ivoire 2000.} This effect can be seen in the results from the quantity models.
 
-However, the results from the models of duration support the argument that election fraud is a noisy and short-lived signal of incumbent strength.
-In most models, the extent of election fraud has no significant association with the duration of protest.
-Instead, duration appears to be driven largely by institutional indicators of the balance of power between the opposition and incumbent---in particular, the size of the legislative opposition, and whether or not the chief executive is term limited.
-Just as in the binary models, grievances do not appear to make longer-lasting protest more likely in fraudulent elections.
-The results from Model 9 complicate the support for the strategic model found in Table 2, since fraud is not associated with longer protest in a low-margin election.
-The combination of the two results suggests that much strategic electoral protest is a gamble that opposition groups lose; anticipating a weak incumbent, they instead find one that is capable of repressing, coopting, or conciliating an emerging protest movement.
-Indeed, while there are 187 cases of protest out of the 647 elections in this dataset, only 15 elections resulted in either the incumbent being replaced due to protest or the election being re-run.\footnote{This information is determined using the NELDA variables NELDA37 and NELDA41.
-The cases of successful protest are: Ukraine 2004, Azerbaijan 2005, Georgia 2003, Kazakhstan 1994, Kyrgyzstan 2005, Bulgaria 1990, Malaysia 2004, Bangladesh 1996, Mali 1997, Peru 2000, Burkina Faso 1997, and Cote d'Ivoire 2000.}
-
-These are pessimistic findings for proponents of democracy.
-While the results do find that there are circumstances in which more fraudulent elections are associated with a higher risk of *some* post-election protest occurring, this risk is limited primarily to cases in which the incumbent party wins by a narrow margin.
-Though protests may occur in such settings, they are no more likely to persist than in  With a large margin of victory, perhaps generated by extensive manipulation, protest appears much less likely.
-Moreover, the risk of protest does not appear to be heightened by poor economic conditions or intensive corruption in the public sector.
+The above results have several implications for understanding election manipulation and regime resilience.
+Most importantly, by treating protest primarily as a binary event rather, the existing literature on electoral protest has overemphasized the risk to incumbents of engaging in electoral manipulation.
+This results presented here show that most electoral protests quickly dwindle, and that fraud has no bearing on how numerous protest events are.
+Moreover, contrary to some prior research, the risk of protest does not appear to be heightened by poor economic conditions [@Brancati2014a] or intensive corruption in the public sector.
 As a result, circumstances that are to some extent outside of leaders' control, like a worsening macroeconomic climate, do not appear to make engaging in fraud any riskier.
-Of course, such conditions are likely to have an independent effect on incumbent's vote-share, but to an extent this is already 'priced in,' and incumbents may be able to effectively blunt their impact by artificially inflating their vote-share using fraud and other techniques.
-
-More importantly, the severity of fraud appears almost wholly unrelated to the duration or number of post-election protests.
-Given that more widespread, frequent, or long-last protests pose a more significant challenge to non-democratic governments than isolated, rare, or brief events, these results suggest that the political cost of fraudulent elections is considerably lower than often assumed.
-Likewise, corruption and poor economic conditions do not appear associated with longer or more numerous protest.
-
 Instead, the conditions that are correlated with more damaging protest are generally institutional factors, which are either more amenable to control or can be anticipated by non-democratic incumbents.
-For example, presidential elections appear to be associated with a greater risk of protest than legislative elections.
-Non-democratic leaders in presidential systems can mitigate this risk by adjusting the electoral calendar (i.e.
-to separate legislative and presidential elections, or by holding presidential elections early), or by ensuring that the state's coercive apparatus is fully in place prior to elections.
 Legislative constraints, an important factor for the duration and number of protests, can be adjusted through institutional reforms and cooptation of the opposition, over time.
@@
 In other words, election fraud only has a minor cost in terms of protest risk, a risk which can be further blunted to the extent that elections can often be held on incumbents' terms.
+These are pessimistic findings for proponents of democracy.
+
+That electoral protest is less risky than is often assumed, in turn, implies that governments and ruling parties have a freer hand to engage in electoral manipulation than many models posit.
+Since the severity of fraud has no apparent bearing on whether or not a major protest wave will emerge as a serious threat to the regime, as shown in Figure 3, government have an incentive to manipulate as much as possible to secure a large winning vote-share.
+This perverse incentive undermines the case that democracy, and fair elections generally, are a self-enforcing equilibrium backed up by the risk of citizen protest [@fearon_self-enforcing_2011].
 
@@
 This study is not a causal one; the decision by the incumbent to engage in fraud and the decision by opposition groups to protest are endogenous to one another.
-Instead, it shows that in practice, manipulated elections are not associated with longer or more numerous protests than clean ones.
+Instead, it shows that in practice, manipulated elections are not associated with more numerous protests than clean ones.
 In typical cases, as a result, protest risk is likely not the constraining factor on the level of election fraud, contrary to the assumption in much of the formal theoretical literature.
@@
 
-Consider also interacting fraud with rentier state status
+# Conclusion
+
+Protest risk is often considered to be a major risk to incumbent governments that work to undermine election integrity.
+This risk has a central role in several prominent models of electoral manipulation, serving to deter manipulation and to incentivize the toleration of election monitors.
+However, empirical research on the subject has largely treated protest as a binary variable, which obscures the degree to which post-election protest imposes a real risk on incumbents in non-democracies.
+One protest may be easily contained, while dozens may pose a more significant risk to the regime.
+
+This paper addresses this concern by testing the relationship between election fraud, the risk that any protest occurs, and the number of protest events that occur after an election.
+Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 647 elections to show that---while there is an elevated risk of some protest occurring in a fraudulent election with narrow margins of victory---election fraud has no significant relationship with the number of protests that occur in a protest wave.
+These results imply that protest risk should not be considered a central deterrent to election fraud.
+Instead, most post-election protests are strategic gambles gone wrong; the initiation of protest more often reveals the incumbent to be strong enough to manage protest through repression, cooptation, or sideways concessions.
+These findings indicate that, rather than being deterred by protest risk, governments have an incentive to manipulate as much as possible in order to achieve a large official vote-share.
+The evidence suggests that democratic breakthoughs in non-democratic elections are rare and unlikely to be driven by the severity of manipulation---a gloomy picture for supporters of democratization---and that researchers should consider other mechanisms beyond mass protest in their models of electoral manipulation.
 

### 2022-07-22 Output/manipulation_protest_draft_v2.Rmd
First shareable version

@@
 header-includes: \usepackage{booktabs}
-abstract: "Multiparty elections and courts with a degree of independence are common features of non-democratic politics.
-Previous research posits a connection between these two institutions, arguing that judicial independence increases the risk of post-election protest, thus reducing electoral manipulation.
-This paper argues that the relationship is not straightforward, identifying four distinct (and competing) channels through which courts may affect protest risk.
-Using cross-national data from 1990-2012, the paper shows that independent courts only increase protest risk under limited circumstances.
-The results show that judicial independence increases protest risk where economic grievances are high, while reducing protest risk where grievances are low.
-As a result, creates judicial independence an underappreciated risk of democratic backsliding in well-off societies, while improving the odds of democratization elsewhere."
+abstract: "The risk of mass protest figures prominently in many models of election manipulation, as a deterrent to tampering with elections.
+Yet election manipulation is common in non-democracies, while protest is rare, and successful protest rarer still.
+Using count data on protest events from 647 elections from 1990 to 2012, this paper argues that the risk to incumbents has been overstated in prior work.
+Specifically, by treating protest as a binary event, previous research has overestimated the strength of protest movements relative to the incumbent's institutional resources.
+The results show that the level of fraud has no bearing on the number of protests after an election, though protest initiation is more likely in high-fraud elections with low margins of victory.
+This pessimistic finding implies that, rather than serving as a constraint, the risk of protest creates a perverse incentive to manipulate as much as possible in order to achieve a large official vote-share."
 ---
 
-Headlines are often captured by scenes of large-scale post-election protest, usually with accompanying allegations of electoral malfeasance.
-On some occasions, these protests can result in new elections and the fall of an authoritarian regime, heightening the sense of drama.
-An extended election protest movement creates a moment of heightened risk for the regime; the uncertainty of the outcome creates opportunities for elite defection, splits in the security services, and other negative outcomes.
+On August 9, 2020, Belarus held an election in which President Aleksander Lukashenko ran for a sixth term.
+The official results credited Lukashenko with over 80\% of the vote, a result which was regarded by observers as wildly inflated [@GolosPlatformPresents2020].
+In response, hundreds of thousands of Belarusians took to the streets over the course of several months.
+Though the protests put significant strain on the regime, a harsh campaign of repression ultimately quelled the protests and President Lukashenko was inaugurated into a new five-year term.
+The Belarusian case offers a useful illustration of a central puzzle in the study of election manipulation: why did the risk of costly protest fail to deter the incumbent from engaging in such widespread fraud?
+And why did the regime use fraud to claim such an implausibly high margin of victory?
+
+Events like those in Belarus are puzzling because many theories of election manipulation expect that incumbents will work hard to avoid them.
+A large, sustained protest movement creates a moment of heightened risk for the regime; the uncertainty of the outcome creates opportunities for elite defection, splits in the security services, and other negative outcomes.
+On some occasions, like Ukraine's Orange Revolution or the People Power Revolution in the Philippines, these protests can result in the fall of an incumbent and a political opening.
 For this reason, many scholars---especially those working in the formal modeling tradition---have identified the risk that election manipulation will spark protest as a major deterrent to regimes wishing to engineer election outcomes.
@@
 Only about ten percent of such protests last long enough to see incumbents fall.
-Classic examples of such outcomes include Ukraine's Orange Revolution, which lasted roughly two months, the months-long series of protests and strikes that followed the 2001 Madagascar presidential election, or the three-week-long Rose Revolution in Georgia.
-Other major protest movements do not succeed in ousting ruling parties, but do win some concessions.
-Russia's 'For Fair Elections' movement in 2011 did not oust the Putin regime, for example, but did see the reintroduction of gubernatorial elections.
-In sum---though with notable exceptions, like the People Power Revolution in the Philippines---it is generally protest *mobilization* that poses challenges to incumbents, rather than the occurrence of protest per se.
-Here, I integrate this understanding of protest's risk to regimes by using count data of post-election protest events, rather than binary indicators of protest.
-
 This implicit assumption has largely been overlooked in theories and empirical studies of the relationship between election manipulation and protest, however.
@@
 
-Using cross-national data on 467 elections from 1990 to 2012, adapted from the recently-released Electoral Contention and Violence (ECAV) dataset [@Daxecker2019], this paper finds that more fraudulent elections are associated with a higher risk any post-election protest event occurring.
+Using cross-national data on 647 elections from 1990 to 2012, adapted from the recently-released Electoral Contention and Violence (ECAV) dataset [@Daxecker2019], this paper finds that more fraudulent elections are associated with a higher risk that any post-election protest event occurs.
 This is in keeping with the formal-modeling literature, which predicts that overreach in the form of election fraud indicates government weakness, making opposition groups confident of winning a conflict in the streets.
@@
 This suggests a more complex relationship between election fraud and protest.
-One the one hand, the result is consistent with the prediction of some formal models that fraud is a noisy signal---in some cases, opposition groups may misperceive a strong government as weak.
-On the other hand, since fraud is not associated with the overall number of events in a protest wave, the risk to governments has likely been overemphasized.
+
+Since fraud is not associated with the overall number of events in a protest wave, the risk to governments has likely been overemphasized.
 That is, even if protests do break out in response to large-scale fraud, the government has many tools it may deploy to curtail them.
+Once protests begin, the noisy signal of regime strength provided by the election is replaced by a very clear signal: the force of the regime's response to the emerging protest movement in contest with the mobilizational capacity of the opposition.
 To the extent that larger protest waves are more damaging to the government than small ones, protest risk does not appear to be a substantial constraint on the severity of election fraud.
-
-However, once protests begin, the noisy signal of regime strength provided by the election is replaced by a very clear signal: the force of the regime's response to the emerging protest movement in contest with the mobilizational capacity of the opposition.
-
-This suggests that the risk to incumbents generated by election manipulation is primarily informational and strategic, rather than the result of grievance and legitimacy cost.
-In turn, this implies that election manipulation is less costly than many models assume; many non-democratic states have the tools needed to curtail a protest movement.
-In addition, it indicates that ruling parties should follow an 'all-in' strategy on election manipulation.
-More manipulation is not associated with more protest mobilization, but a large margin of victory is associated with both smaller and shorter protest waves.
+Troublingly, these findings indicate that ruling parties should follow an 'all-in' strategy on election manipulation.
+More manipulation is not associated with more protest mobilization, but a large margin of victory is associated with less risk of protest initiation.
 These results help explain both the relative paucity of electoral protests, as well as the rarity of their success in overturning incumbent regimes.
@@
 Furthermore, simply being on the losing side in an election is associated with reduced belief that the election was fair [@Cantu2015], and with increased support for electoral protest [@Anderson2006].
-[Add procedural justice cites to this graf.]
+[Note: may add procedural justice cites to this graf.]
 
@@
 Several empirical studies have, accordingly, found links between election manipulation and binary measures of protest [@Brancati2016;@Hyde2014a;@Harvey2020;@Rod2019].
-
-Large, sustained protest waves may also contribute to elites' decisions to stand by the incumbent regime, or to side with "the people" in their own pursuit of power [@hale_patronal_2014;@svolik_politics_2012;@robertson_protest_2010].
 
@@
 library(stargazer)
+library(pscl)
 library(modelsummary)
@@
 ggplot(nelda2.sub, aes(x=winner.share.upd)) + geom_histogram(bins=100) + xlab("Incumbent vote-share") + theme_bw()
+```
 
+```{r, include=FALSE}
 #nelda2.sub.success <- nelda2.sub %>% filter(protest.success == 1)
@@
 #summary(model.n.base)
+
+model.n.base.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
+summary(model.n.base.hurd)
 ```
@@
 
+model.n.base2.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
+summary(model.n.base2.hurd)
 ```
@@
 
+model.n.griev.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, dist = "negbin", data = nelda2.sub %>% filter(gdpgro.1lag < 1))
+summary(model.n.griev.hurd)
 ```
@@
 
+model.n.griev2.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub, dist = "negbin")
+summary(model.n.griev2.hurd)
 ```
@@
 
+model.n.griev3.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub, dist = "negbin")
+summary(model.n.griev3.hurd)
 ```
@@
 
-```{r tables-quantityr}
+```{r tables-quantity, echo=FALSE}
 models.quantity <- list("Model 8" = model.n.base, "Model 9" = model.n.base2, "Model 10" = model.n.griev, "Model 11" = model.n.griev2, "Model 12" = model.n.griev3)

### 2022-08-26 Output/appendix.Rmd
First draft appendix

---
title: "Appendix: Not So Bad---The minimal effect of election fraud on protest intensity in non-democracies"
author: "Cole J. Harvey (Oklahoma State University)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    bookdown::pdf_document2: default
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
header-includes: \usepackage{booktabs}
---

## Hurdle models

As noted in the main text, the distribution of the count data on protest events is highly skewed; the large majority of elections record zero protest events. This distribution requires an appropriate modeling strategy in order to avoid biased coefficient estimates. Hurdle models are one strategy for analysis of this kind of data [@mullahy1986;@feng2021]. The hurdle approach models the data generating process as a mixture model. One component of the model estimates the probability that each observation will belong to the zero component of the data; this is typically a binominal model, and the logit function is used in this case. The second component, for observations that take on a positive value, is modeled as a standard count model. In this case, due to the distribution of the positive counts, a negative binomial model is used.

```{r setup, include = FALSE}
library(tidyverse)
library(countrycode)
library(readxl)
library(vdemdata)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
library(stargazer)
library(pscl)
library(modelsummary)

nelda2.sub <- read.csv(here::here("Data", "nelda2_sub.csv"))

nelda2.sub <- nelda2.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda2.sub <- nelda2.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))

##Naming nelda variables

nelda2.sub <- nelda2.sub %>% rename(boycott = nelda14)
nelda2.sub <- nelda2.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                 NA)))

##Creating economic crisis variable

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = as.factor(econ.crisis))

nelda2.sub <- nelda2.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))

nelda2.sub <- nelda2.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))

##Removing cases before 1990, when ECAV starts
nelda2.sub <- nelda2.sub %>% filter(year >= 1990)

## Removing cases where inc. loses
nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)

##Intmon as a factor
nelda2.sub <- nelda2.sub %>% mutate(v2elintmon = as.factor(v2elintmon))

nelda2.sub <- nelda2.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))

#Square of physint

nelda2.sub <- nelda2.sub %>% mutate(physintlag.sq = physinteg.1lag^2)

#Crisis as factor

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))

#Margins with updates

nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))

#Successful protest = new elections or new incumbent

nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))

#Protest cases only

protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1)

```

```{r models, warning = FALSE, message = FALSE, include = F}
model.n.base.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
summary(model.n.base.hurd)

model.n.base2.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
summary(model.n.base2.hurd)

model.n.griev.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, dist = "negbin", data = nelda2.sub %>% filter(gdpgro.1lag < 1))
summary(model.n.griev.hurd)

model.n.griev2.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub, dist = "negbin")

model.n.griev3.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub, dist = "negbin")
summary(model.n.griev3.hurd)
```

```{r custom-tidiers, include=FALSE}
tidy.hurdle.count <- function(x, conf.int = FALSE, conf.level = 0.95, ...) {

  result <- summary(x)$coefficients$count %>%
    tibble::as_tibble(rownames = "term") #%>%
    #dplyr::rename(estimate = Estimate,
    #              std.error = `Std. Error`,
    #              statistic = `t value`,
     #             p.value = `Pr(>|t|)`)

  if (conf.int) {
    ci <- confint(x, level = conf.level)
    result <- dplyr::left_join(result, ci, by = "term")
  }

  result
}

#test <- summary(model.n.base.hurd)$coefficients$count

test <- tidy.hurdle.count(model.n.base.hurd)

#tbl_regression(test)
```

```{r zero-model-table, echo=FALSE}

models <- list()
models[['Model 1']] <- model.n.base.hurd
models[['Model 2']] <- model.n.base2.hurd
models[['Model 3']] <- model.n.griev.hurd
models[['Model 4']] <- model.n.griev2.hurd
models[['Model 5']] <- model.n.griev3.hurd

modelsummary::modelsummary(models, coef_omit = "count", gof_omit = "R2", stars = T)

```

Table 1 shows the results of the binomial portion of the model; that is, it shows which coefficients are associated with any protest occurring after the election. The results are very similar to those in Table 1 in the main text. Presidential elections are the main influence associated with a higher probability of protest initiation in both tables. In both, the effect of electoral manipulation is most strongly felt in elections with low margins of victory (Model 2). In the other hurdle models, the severity of election fraud is not associated with increased protest risk. While the main findings are consistent between the hurdle models and the logit models presented in the main text, the hurdle models are even more pessimistic regarding the relationship between manipulation and protest: the coefficient is not statistically significant in any of the other four models in Table 1.

```{r count-model-table, echo=FALSE}
#knitr::kable(test)

modelsummary::modelsummary(models, coef_omit = "zero", gof_omit = "R2", stars = T)
```

Table 2 presents the results from the count portion of the same hurdle models. These results are, likewise, very similar to those presented for the negative binomial models in the main text. Legislative constraints on the executive and presidential term limits show the strongest association with the number of post-election protests. The severity of electoral manipulation has no statistically significant association with the number of events in any model.

As a whole. the hurdle models tell the same story as the separate models in the main text. Election fraud may have a positive association with protest onset in a close election, but rarely otherwise. The number of post-election protests appears to have little to no association with election fraud. Economic grievances, likewise, appear to play little role. The factors that appear to matter most are political and institutional: term limits, election type, and the size and nature of the legislative opposition.

```{r, include=FALSE}
(-.682*.04) + (-4.3*.55) + (1.625*.55*.04)  # Low fraud, mean margin

(-.682*1.67) + (-4.3*.55) + (1.625*.55*1.67)  # High fraud, mean margin

#Expected quantity is lower when fraud is high
```

## References

### 2022-08-26 Output/manipulation_protest_draft_v2.Rmd
First draft appendix

@@
 ---
-title: "Draft: It's Not So Bad---The minimal effect of election fraud on protest intensity in non-democracies"
+title: "Draft: Not So Bad---The minimal effect of election fraud on protest intensity in non-democracies"
 author: "Cole J. Harvey (Oklahoma State University)"
@@
 
-Using cross-national data on 647 elections from 1990 to 2012, adapted from the recently-released Electoral Contention and Violence (ECAV) dataset [@Daxecker2019], this paper finds that more fraudulent elections are associated with a higher risk that any post-election protest event occurs.
+Using cross-national data on 647 elections from 1990 to 2012 where incumbents claimed victory, adapted from the recently-released Electoral Contention and Violence (ECAV) dataset [@Daxecker2019], this paper finds support for the hypothesis that more fraudulent elections are associated with a higher risk that any post-election protest event occurs.
 This is in keeping with the formal-modeling literature, which predicts that overreach in the form of election fraud indicates government weakness, making opposition groups confident of winning a conflict in the streets.
 However, there is no statistically significant association between the level of election fraud and the number of protest events.
-This suggests a more complex relationship between election fraud and protest.
+This suggests a more complex relationship between election fraud and protest, undermining the argument that protest risk is a major deterrent to electoral malfeasance.
 
@@
 Troublingly, these findings indicate that ruling parties should follow an 'all-in' strategy on election manipulation.
-More manipulation is not associated with more protest mobilization, but a large margin of victory is associated with less risk of protest initiation.
-These results help explain both the relative paucity of electoral protests, as well as the rarity of their success in overturning incumbent regimes.
+More manipulation is not associated with a larger number of protests, but a large margin of victory is associated with a lower risk of protest initiation.
+These results help explain both the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory.
 
@@
 Despite this consensus in the formal literature, there is relatively little empirical evidence to support the claim that manipulation per se indicates weakness, and thus increases protest risk.
-On the contrary, Simpser [-@Simpser2011] shows support for the alternative (though still informational) view that excessive manipulation can induce cooperation from opposition groups, elites, and other actors by revealing information about regime strength.
-Similarly, the type of manipulation employed can convey information about the ruling party's capacity to mobilize [@Harvey2020].
+On the contrary, Simpser [-@Simpser2011] shows support for the alternative view that excessive manipulation can induce cooperation from opposition groups, elites, and other actors by revealing information about regime strength.
+Similarly, the type of manipulation employed can convey information about the ruling party's capacity to mobilize, with falsification providing the least credible information [@Harvey2020].
 In this way, large-scale manipulation can contribute to a ruling party's ability to produce a large, stabilizing majority vote-share [@magaloni2006].
@@
 This argument has found support in studies of the relationship between economic grievances, manipulated elections, and protest [@Rod2019;@Brancati2016].
-A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b], which is suggests that emotions, moral condemnation, or grievances take precedence over strategic calculations for individuals.
+A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b], which suggests that emotions, moral condemnation, or grievances take precedence over strategic calculations for individuals.
 Subjective belief in fraud was also shown to be correlated with approval of mass protests after Russia's 2011 election [@Chaisty2013].
@@
 
-These two explanations are, of course, interconnected.
+The strategic and social-psychological explanations are, of course, interconnected.
 As Schedler writes, "Grievances provide moral justifications for collective protest: we need to do something.
@@
 The literature thus expects that manipulated elections increase the risk of protest by 1) revealing regime weakness, especially when the margin of victory is low, and 2) exacerbating existing grievances and upsetting citizens' sense of procedural fairness.
+In turn, governments should be reluctant to manipulate.
 However, as discussed below, electoral protest is relatively rare, and even in cases where it does occur, most efforts to overturn fraudulent election results through protest have failed [@kalandadze2009electoral].
 The apparent failure of protest risk to deter election fraud underlines our limited understanding of non-democratic ruling parties' decision-making.
-Does the extent of fraud, or the severity of underlying social grievances, make more numerous (and thus more costly) protest more likely?
+Does the extent of fraud, in fact, make more numerous (and thus more costly) protest more likely?
 
@@
 
->H2: Post-election protest will be more likely to occur when fraud is severe and ruling-party margin of victory is small, and less likely when fraud is low or the margin of victory large.
+>H2: More sever election fraud is associated with a higher likelihood of protest when winning-party vote-share is low.
 
@@
 Once protests begin, the extent of electoral fraud is unlikely to affect their scope.
-The initial emergence of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@Tilly1977;@Lipsky1968;@Hummel2019;@Rasler1996;@Bishara2015;@Frantz2014].
-(**See others**.) Given that post-electoral protests are a direct challenge to the regime, toleration is unlikely in these cases.
+The initial emergence of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014].
+Given that post-electoral protests are a direct challenge to the regime, toleration is unlikely in these cases.
 That election dates are known in advance creates opportunities for opposition mobilization; anticipating this, regimes can more easily engage in pre-emptive repression [@Truex2019;@Sullivan2016;@Bhasin2013].
@@
 
-What results is a test of strength between the anti-incumbent protest coalition and the regime with its supporters (Tilly 2006).
+What results is a test of strength between the anti-incumbent protest coalition and the regime with its supporters [@mcadamDynamicsContention2001;@tillyRegimesRepertoires2010].
 In essence, as Beaulieu [@Beaulieu2014a] notes, the onset of electoral protest marks a case of bargaining failure.
@@
 It controls the resources of the state, and can call upon the support of those elites who benefit from the status quo.
-High-capacity states, as a result, are more likely to be able to engage in successful repression (Tarrow 2011).
+High-capacity states, as a result, are more likely to be able to engage in successful repression [@tarrowPowerMovementSocial2011].
 Though a biased election may temporarily allow diverse opposition groups to set aide their differences [@trejo_ballot_2014], this internal diversity can enable the incumbent to divide the opposition with selective repression or concessions.
-For example, survey research from Russia's 'For Fair Elections' movement found that many protesters were motivated as much by...
-[Find citation].
 While the regime may have its own splits---between reformists and hardliners, for example---the emergence of protest can act to solidify support for the ruling party among its supporters.
@@
 These hypotheses are tested using data from three datasets: NELDA [@hyde2012], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence (ECAV) dataset [@Daxecker2019].
-The unit of observation for this study is an election period as recorded in NELDA, which counts all national elections between 1945 and 2015.
+The unit of observation for this study is an election period as recorded in NELDA.
 This framework prevents selecting on the dependent variable, by capturing both protested and non-protested elections.
@@
 Finally, the date range is limited by the availability of data on protest from ECAV, as discussed below, which runs from 1990 to 2012.
-The resulting dataset includes 647 election-period observations.
+The resulting dataset includes 647 election-period observations across 109 countries.
 
@@
 To capture the *severity of election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than aggregate measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], it can fail to signal strength [@Harvey2020], and it can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021].
+This measure of election integrity should thus pose a strong test of the protest-oriented model.
 Finally, the winning party's *vote-share* is also taken from the V-Dem dataset.
@@
 As the figure shows, there is no statistically significant difference in the expected number of protest following a relatively clean election compared to a relatively fraudulent one, for any margin of incumbent victory.
-Taking the results from Tables 2 and 3 together, there is support for Hypothesis 1 (an unconditional increase in the risk of protest as fraud increases), as well as strong support for Hypothesis 2 (an increase in protest risk in fraudulent, low-margin elections) and Hypothesis 4 (no relationship between fraud and protest quantity).
+Taking the results from Tables 2 and 3 together, there is support for Hypothesis 1 (an unconditional increase in the risk of any protest as fraud increases), as well as strong support for Hypothesis 2 (an increase in protest risk in fraudulent, low-margin elections) and Hypothesis 4 (no relationship between fraud and protest quantity).
 Hypothesis 3, on the connection between fraud, grievance, and protest, is not supported.
@@
 
-Altogether, the findings support an understanding of fraud as a signal of incumbent strength, but a noisy and unreliable one.
+Altogether, the findings challenge the idea that the expectation of protest will deter election-manipulation by ruling parties.
+To the extent that fraud is a signal of incumbent strength, it is a noisy and unreliable one.
 When an incumbent party wins by a narrow margin in a fraudulent election, the likelihood of protest initiation is substantially higher than when it wins more cleanly: roughly a 75\% chance in the former case versus 15\% in the latter as shown in Figure 2.
@@
 Legislative constraints, an important factor for the duration and number of protests, can be adjusted through institutional reforms and cooptation of the opposition, over time.
-Finally, incumbent term limits can be managed through constitutional reforms, where possible, or through the appointment of (presumably) loyal successors.
+Incumbent term limits can be managed through constitutional reforms, where possible, or through the appointment of (presumably) loyal successors.
 In other words, election fraud only has a minor cost in terms of protest risk, a risk which can be further blunted to the extent that elections can often be held on incumbents' terms.
@@
 
-None of this is to say that election fraud cannot, in theory, pose a cost to incumbents in the form of protest.
-This study is not a causal one; the decision by the incumbent to engage in fraud and the decision by opposition groups to protest are endogenous to one another.
-Instead, it shows that in practice, manipulated elections are not associated with more numerous protests than clean ones.
-In typical cases, as a result, protest risk is likely not the constraining factor on the level of election fraud, contrary to the assumption in much of the formal theoretical literature.
+One objection to this framing may be that, if the level of manipulation and the risk of protest are endogenous to one another, perhaps the null result found for the number of protests is to be expected.
+That is, if incumbents choose a level of manipulation calculated to avoid large-scale protest given other contextual factors, no correlation between fraud and protest quantity would be found.
+This paper does not dispute that decisions about manipulation and protest may be endogenous, but it does argue that the resulting equilibrium favors governments.
+In the dataset used here, only ten elections in four countries are coded as having an *intentional irregularities* score of less than -2 (indicating extremely fraud-free elections).
+By contrast, 103 elections in 27 countries score higher than 2 (indicating highly fraudulent elections).
+The frequency of fraud, the rarity of protest, and the finding that fraud severity is not correlated with the scope of protest events all point to the same conclusion: protest risk is not a substantial constraint on election fraud in non-democracies.
 Rather, the constraints on manipulation are more likely to be driven by factors like resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022].

### 2022-11-04 Output/manipulation_protest_draft_v2.Rmd
Minor text edits + fixing NAs in inc.termlimit.num (0s for parliamentary elections)

@@
 
-Research on the risk posed to non-democratic governments by electoral protest often acknowledges---whether explicitly or implicitly---the importance of sustained protest campaigns, rather than one-off events [@Bunce2006;@tucker2007;@DAnieri2006;@Carnaghan2016;@bunce2010].
+Research on the risk posed to non-democratic governments by electoral protest often acknowledges---whether explicitly or implicitly---the importance of broad-based protest campaigns, rather than one-off events [@Bunce2006;@tucker2007;@DAnieri2006;@Carnaghan2016;@bunce2010].
 As Brancati [-@Brancati2016] finds, approximately two-thirds of pro-democracy electoral protests ended within three days, posing little threat to incumbents.
-Only about ten percent of such protests last long enough to see incumbents fall.
+Only about ten percent of such protests are widespread and long-lasting enough to see incumbents fall.
 This implicit assumption has largely been overlooked in theories and empirical studies of the relationship between election manipulation and protest, however.
@@
 In this way, large-scale manipulation can contribute to a ruling party's ability to produce a large, stabilizing majority vote-share [@magaloni2006].
-Similarly, Luo and Rozenas [-@Luo2020] posit that ex post fraud is more likely to result in protest than ex ante manipulation, though this can be contrasted with the finding that parties are more likely to protest elections after ex ante changes to the rules [@chernykh2014].
-In sum, it is likely that manipulation *may* signal regime weakness, but this is more likely when margins of victory are narrow and when falsification is a major tool.
+Similarly, Luo and Rozenas [-@Luo2020] posit that ex post fraud is more likely to result in protest than ex ante manipulation, though this stands in contrast with the finding that parties are more likely to protest elections after ex ante changes to the rules [@chernykh2014].
+In sum, it is likely that manipulation *may* signal regime weakness, but empirical works suggests that this is more likely when margins of victory are narrow and when falsification is a major tool.
 
@@
 Furthermore, simply being on the losing side in an election is associated with reduced belief that the election was fair [@Cantu2015], and with increased support for electoral protest [@Anderson2006].
-[Note: may add procedural justice cites to this graf.]
 
@@
 protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1)
+
+# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections
+
+nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))
 
@@
 One objection to this framing may be that, if the level of manipulation and the risk of protest are endogenous to one another, perhaps the null result found for the number of protests is to be expected.
-That is, if incumbents choose a level of manipulation calculated to avoid large-scale protest given other contextual factors, no correlation between fraud and protest quantity would be found.
-This paper does not dispute that decisions about manipulation and protest may be endogenous, but it does argue that the resulting equilibrium favors governments.
+That is, if incumbents choose a level of manipulation calculated to avoid large-scale protest---sometimes high and sometimes low, conditional on contextual factors---no correlation between fraud and protest quantity would be found.
+This paper does not dispute that decisions about manipulation and protest may be endogenous to one another, but it does argue that the resulting equilibrium favors governments.
 In the dataset used here, only ten elections in four countries are coded as having an *intentional irregularities* score of less than -2 (indicating extremely fraud-free elections).

### 2022-11-07 Output/appendix.Rmd
Update appendix to remove NAs for inc.termlimit.num from parliamentary cases
Add figures to appendix + change title

@@
 ---
-title: "Appendix: Not So Bad---The minimal effect of election fraud on protest intensity in non-democracies"
+title: "Appendix: A missing guardrail---The minimal effect of election fraud on protest intensity in non-democracies"
 author: "Cole J. Harvey (Oklahoma State University)"
@@
 library(stargazer)
+library(cowplot)
 library(pscl)
@@
 
+## Marginal effects plots for grievance models
+
+The figures presented in this section illustrate the marginal effects for the grievance models in the main text (for both the binomial and count models of protest).
+Figure \@ref(fig:binary-plots) shows the marginal effects from Models 4 - 6 from the main text, while Figure \@ref(fig:count-plots) shows the results from Models 10 - 12.
+They demonstrate that the poor GDP growth, an economic crisis, and public corruption are not significantly associated with a changes in the marginal effect of election fraud on protest risk.
+
+```{r, include=FALSE}
+model.ecavbin.mm.crisis <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag +     econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
+#summary(model.ecavbin.mm.crisis)
+p.bin.crisis <- sjPlot::plot_model(model.ecavbin.mm.crisis, type = "pred", terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv"), title = "")
+
+model.ecavbin.mm.corr <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag +  osorg.1lag + pubseccorrup.lag +  natlpartyorg.lag +  pubseccorrup.lag +     pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
+#summary(model.ecavbin.mm.corr)
+p.bin.corr <- sjPlot::plot_model(model.ecavbin.mm.corr, type = "pred", terms = c("pubseccorrup.lag", "v2elirreg.inv"), title = "")
+
+model.ecavbin.mm.gdp <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag + osorg.1lag +   natlpartyorg.lag +     gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub %>% filter(gdpgro.1lag < 1))
+p.bin.gdp <- sjPlot::plot_model(model.ecavbin.mm.gdp, type = "pred", terms = c("gdpgro.1lag", "v2elirreg.inv"), title = "")
+
+```
+
+```{r count_models, include=FALSE}
+model.n.griev <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
+#summary(model.n.griev)
+
+p.count.gdp <-  sjPlot::plot_model(model.n.griev, type = "pred", terms = c("gdpgro.1lag", "v2elirreg.inv [.23, 2.12]"), title = "")
+
+model.n.griev2 <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
+#summary(model.n.griev2)
+p.count.crisis <- sjPlot::plot_model(model.n.griev2, type = "pred", terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [.21, 2.13]"), title = "")
+
+model.n.griev3 <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
+#summary(model.n.griev3)
+p.count.corr <- sjPlot::plot_model(model.n.griev3, type = "pred", terms = c("pubseccorrup.lag", "v2elirreg.inv [.22, 2.11]"), title = "")
+```
+
+```{r binary-plots, fig.cap = "Economic grievances and any protest",  echo=FALSE, message=FALSE, warning=FALSE}
+plot_rows <-  plot_grid(p.bin.gdp, p.bin.crisis, p.bin.corr)
+
+title <- ggdraw() + 
+  draw_label(
+    "Predicted probabilities of any protest",
+    fontface = 'bold',
+    x = 0,
+    hjust = 0
+  ) +
+  theme(
+    # add margin on the left of the drawing canvas,
+    # so title is aligned with left edge of first plot
+    plot.margin = margin(0, 0, 0, 7)
+  )
+plot_grid(
+  title, plot_rows,
+  ncol = 1,
+  # rel_heights values control vertical title margins
+  rel_heights = c(0.1, 1)
+)
+
+```
+
+```{r count-plots, fig.cap = "Economic grievances and number of protests", echo=FALSE, message=FALSE, warning=FALSE}
+plot_rows <-  plot_grid(p.count.gdp, p.count.crisis, p.count.corr)
+
+title <- ggdraw() + 
+  draw_label(
+    "Predicted number of protest events",
+    fontface = 'bold',
+    x = 0,
+    hjust = 0
+  ) +
+  theme(
+    # add margin on the left of the drawing canvas,
+    # so title is aligned with left edge of first plot
+    plot.margin = margin(0, 0, 0, 7)
+  )
+plot_grid(
+  title, plot_rows,
+  ncol = 1,
+  # rel_heights values control vertical title margins
+  rel_heights = c(0.1, 1)
+)
+```
+
+\newpage
+
 ## References

### 2022-11-07 Output/manipulation_protest_draft_v2.Rmd
Add figures to appendix + change title

@@
 ---
-title: "Draft: Not So Bad---The minimal effect of election fraud on protest intensity in non-democracies"
+title: "Draft: A missing guardrail---The minimal effect of election fraud on protest intensity in non-democracies"
 author: "Cole J. Harvey (Oklahoma State University)"
@@
 #summary(model.n.griev)
-sjPlot::plot_model(model.n.griev, type = "pred", terms = c("gdpgro.1lag", "v2elirreg.inv"))
+sjPlot::plot_model(model.n.griev, type = "pred", terms = c("gdpgro.1lag", "v2elirreg.inv [.23, 2.12]"))
 
@@
 
-We do observe support for the more structural explanations of durable protest.
+We do observe support for the more structural explanations of more extensive protest.
 The coefficient for the *legislative constraint* variable, which measures the size and oppositional character of the legislative opposition, is consistently positive and  statistically significant.
@@
 To the extent that fraud is a signal of incumbent strength, it is a noisy and unreliable one.
-When an incumbent party wins by a narrow margin in a fraudulent election, the likelihood of protest initiation is substantially higher than when it wins more cleanly: roughly a 75\% chance in the former case versus 15\% in the latter as shown in Figure 2.
+When an incumbent party wins by a narrow margin in a fraudulent election, the likelihood of protest initiation is substantially higher than when it wins more cleanly: roughly a 50\% chance in the former case versus 20\% in the latter as shown in Figure 2.
 However, since even strong incumbents commit fraud, the decision to take to the streets is often a losing gamble---anticipating a weak incumbent, they instead find one that is capable of repressing, coopting, or conciliating an emerging protest movement.
@@
 Most importantly, by treating protest primarily as a binary event rather, the existing literature on electoral protest has overemphasized the risk to incumbents of engaging in electoral manipulation.
-This results presented here show that most electoral protests quickly dwindle, and that fraud has no bearing on how numerous protest events are.
+This results presented here show that most electoral protests are narrow, and that fraud has no bearing on how numerous protest events are.
 Moreover, contrary to some prior research, the risk of protest does not appear to be heightened by poor economic conditions [@Brancati2014a] or intensive corruption in the public sector.
@@
 Legislative constraints, an important factor for the duration and number of protests, can be adjusted through institutional reforms and cooptation of the opposition, over time.
-Incumbent term limits can be managed through constitutional reforms, where possible, or through the appointment of (presumably) loyal successors.
+Incumbents may attempt to manage term limit through constitutional reforms, where possible, or through the appointment of (presumably) loyal successors.
 In other words, election fraud only has a minor cost in terms of protest risk, a risk which can be further blunted to the extent that elections can often be held on incumbents' terms.
+
 These are pessimistic findings for proponents of democracy.
-
 That electoral protest is less risky than is often assumed, in turn, implies that governments and ruling parties have a freer hand to engage in electoral manipulation than many models posit.

### 2022-11-15 Scripts/data_cleaning.Rmd
Minor text edits + fixing NAs in inc.termlimit.num (0s for parliamentary elections)
Adding event counts and protest sizes from Mass Mobilization dataset (note: quality checks still needed)

@@
                                                   legcon.1lag,
+                                                  v2ex_elechog,
 egaldem.1lag ,
@@
 
-nelda3 <- nelda2 %>% merge(massmob, by = "country_id_year")
+#nelda3 <- nelda2 %>% merge(massmob, by = "country_id_year")
+
+nelda3 <- nelda %>% dplyr::select(ccode, country, year, electionid, election_mdy)
+nelda3 <- nelda3 %>% mutate(election_year = year)
+nelda3 <- massmob %>% left_join(nelda3, by = c("ccode", "year")) #Nelda3 is in progress, need to figure out a) if needed, b) how to include pre-election and post-election mobilization
 ```
@@
 ```{r}
-nelda3 <- nelda3 %>% filter(protest_mdy >= election_mdy)
+#Add electionid to the year before the election and year after (so that grouping variable can capture those protests if needed)
+nelda3 <- nelda3 %>% group_by(ccode) %>% mutate(election_date_upcoming = as_date(ifelse(is.na(election_year)==F, election_mdy, 
+                                                  ifelse(is.na(election_year)==T & is.na(lead(election_mdy))==F,
+                                                                          lead(election_mdy), NA))))
+
+nelda3 <- nelda3 %>% group_by(ccode) %>% mutate(election_date_lagged = as_date(ifelse(is.na(election_year)==F, election_mdy,                                                         ifelse(is.na(election_year)==T & is.na(lag(election_mdy))==F, lag(election_mdy), NA))))
+
+## Now filter by date of protest
+
+# Steps:
+  #1) Remove rows where election_upcoming and election_lagged are both NA. 
+  #2) This results in a dataframe with only election years +-1 year; create interval variables (lagged and leading days between protest and election)
+  #3 Filter to cases that are six months before to 3 months after election (i.e 180 days and 90 days)
+
+nelda3.sub <- nelda3 %>% filter(is.na(election_date_upcoming)==F | is.na(election_date_lagged) == F)
+
+nelda3.sub <- nelda3.sub %>% mutate(protest_preelex_interval = ifelse((election_mdy - protest_mdy) >0, (election_mdy-protest_mdy), NA))
+
+nelda3.sub <- nelda3.sub %>% mutate(protest_postelex_interval = ifelse((election_mdy - protest_mdy) <= 0, (protest_mdy - election_mdy), NA))
+
+nelda3.sub <- nelda3.sub %>% group_by(ccode) %>% filter(protest_preelex_interval <= 180 | protest_postelex_interval <= 90)
+
+# Next step is to aggregate needed variables (participants) by electionid
+```
+
+Aggregating participant variables
+```{r}
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(participants_category == "", NA, 
+                                            ifelse(participants_category == "50-99", 1, 
+                                            ifelse(participants_category == "100-999", 2,
+                                            ifelse(participants_category == "1000-1999", 3,
+                                            ifelse(participants_category == "2000-4999", 4,
+                                            ifelse(participants_category == "5000-10000", 5,
+                                                   6)))))))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100s", 2, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000s", 4, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">50", 1, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "50", 1, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000", 3, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "50+", 1, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100", 2, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "10000", 5, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "2000", 4, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "200", 2, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "300", 2, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "500", 2, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "3000", 4, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100+", 2, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "5000", 5, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "20000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "10000s", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "150", 2, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "400", 2, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "hundreds", 2, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "30000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "15000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1500", 3, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 50", 1, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "4000", 4, participants_category_ordered))  ## Stopping here; all other categories have fewer than 20 observations.
+Beginning second stopping rule:
+
+## Second rule: collect all observations greater than 10,000
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "hundreds of thousands", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "<20000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 40000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 42000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">35000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">40000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">50000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "110000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "14059", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "16000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "18000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "24000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "250000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "30000+", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "35000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "36000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "5000-50000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "55000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "650000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 10000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "13000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "500000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">15000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">20000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "14000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "400000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "Tens of thousands", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">100000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "150000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "20000+", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "60000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "90000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "10000+", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100000s", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "200000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "80000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">10000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "70000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "80000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "40000", 6, participants_category_ordered))
+
+nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "25000", 6, participants_category_ordered))
+
+#nelda3.sub <- nelda3.sub %>% group_by(electionid) %>% 
+```
+
+Remove post-election protests unrelated to election
+
+```{r}
+nelda3.sub <- nelda3.sub %>% filter(protest_mdy < election_mdy | (protest_mdy >= election_mdy & str_detect(notes, regex(pattern = "election*|vot*|ballot*", ignore_case = T))))
+```
+
+Getting grouped summarize for number and size
+```{r}
+nelda3.sub.pre <- nelda3.sub %>% filter(protest_mdy < election_mdy)
+nelda3.sub.post <- nelda3.sub %>% filter(protest_mdy >= election_mdy)
+
+count_data_mm_pre <- nelda3.sub.pre %>% group_by(electionid) %>% summarize(n.events.massmob.pre = sum(protestnumber))
+count_data_mm_post <- nelda3.sub.post %>% group_by(electionid) %>% summarize(n.events.massmob.post = sum(protestnumber))
+
+size_data_mm_pre <- nelda3.sub.pre %>% group_by(electionid) %>% summarize(median.participants.pre = median(participants_category_ordered, na.rm = T))
+size_data_mm_post <- nelda3.sub.post %>% group_by(electionid) %>% summarize(median.participants.post = median(participants_category_ordered, na.rm = T))
+
 ```
@@
 ```{r}
-nelda2.sub <- nelda2 %>% filter(v2x_regime == 1 | v2x_regime == 2)
+nelda2.sub <- nelda2 %>% filter(v2x_regime == 1 | v2x_regime == 2) #Note that coding scheme for electoral democracies allows for "some degree of fraud and irregularities but these did not in the end affect the outcome" (v2elfrfair, which is a component of v2x_regime)
 ```
@@
 ecav <- ecav %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
-size_data <- ecav %>% group_by(NeldaID) %>% summarize(median.participants = median(ParticipantNumber))
+size_data <- ecav %>% group_by(NeldaID) %>% summarize(median.participants = median(ParticipantNumber, na.rm = T))
 
 ecav.pre <- ecav.pre %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
-size_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(median.participants.pre = median(ParticipantNumber))
+size_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(median.participants.pre = median(ParticipantNumber, na.rm = T))
 
@@
 nelda2.sub <- nelda2.sub %>% mutate(d.events.num = ifelse(is.na(d.events.num)==T, 0, d.events.num))
-nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(is.na(median.participants)==T, 0, median.participants))
+#nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(is.na(median.participants)==T, 0, median.participants))
 
@@
 ```{r}
-nelda3 <- nelda3 %>% mutate(protests.postelex = ifelse(protests == 1 & protesterdemand1 == "political behavior, process", 1, 0))
+#nelda3 <- nelda3 %>% mutate(protests.postelex = ifelse(protests == 1 & protesterdemand1 == "political behavior, process", 1, 0))
 
-nelda3 <- nelda3 %>% mutate(protests.postelex.quant = ifelse(protests.postelex == 1, protestnumber, 0))
-```
-
-Cutting out closed autocracies and liberal democracies
-
-```{r}
-nelda3.sub <- nelda3 %>% filter(v2x_regime == 1 | v2x_regime == 2)
+#nelda3 <- nelda3 %>% mutate(protests.postelex.quant = ifelse(protests.postelex == 1, protestnumber, 0))
 ```
@@
 
+## Merge data from nelda3.sub with nelda2.sub
+
+```{r}
+##Combined post data together, then pre data, then both
+mm_data_post_comb <- left_join(count_data_mm_post, size_data_mm_post, by = "electionid")
+mm_data_pre_comb <- left_join(count_data_mm_pre, size_data_mm_pre, by = "electionid")
+
+mm_data_comb <- left_join(mm_data_pre_comb, mm_data_post_comb, by = "electionid")
+
+mm_data_comb <- mm_data_comb %>% rename(NeldaID = electionid)
+
+nelda2.sub <- nelda2.sub %>% left_join(mm_data_comb, by = "NeldaID")  ## Double check to make sure this joining works
+
+```
+
 ## Writing

### 2022-11-16 Output/appendix.Rmd
Various changes to add new Nelda and MassMob variables; moved MassMob models to appendix 2

@@
 library(modelsummary)
+library(MASS)
 
@@
 nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))
+
+## Changing participants from numeric category to text
+nelda2.sub <- nelda2.sub %>% mutate(median.participants = round(median.participants))
+nelda2.sub <- nelda2.sub %>% mutate(particpants_categorical = ifelse(is.na(median.participants)==T, NA, ifelse(median.participants == 1, "Less than 10",
+                     ifelse(2, "10 to 99", ifelse(3, "100 to 999",
+                                                  ifelse(4, "1000 to 9,999", ifelse(5, "Greater than 10,000", NA)))))))
+nelda2.sub <- nelda2.sub %>% mutate(particpants_categorical = as.factor(particpants_categorical))
+
+## Median particiants (MM) as factor
+
+nelda2.sub <- nelda2.sub %>% mutate(median.participants.mm.post = as.factor(median.participants.mm.post))
 
@@
 
+## Protest size (categories) using mass mobilization data
+
+```{r}
+## Econ grievance
+dat <- ggpredict(model.size.griev2, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv"))
+dat.sub <- dat %>% dplyr::filter(response.level == "0" | response.level == "2" | response.level == "4" | response.level == "6")
+plot(dat.sub)
+
+dat_max <- ggpredict(model.size.griev2, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-2.3, 3.2]"))
+dat.sub <- dat %>% dplyr::filter(response.level == "0" | response.level == "3" | response.level == "6")
+plot(dat.sub)
+```
+
+## Protest event quantity, MM version
+
+```{r}
+protest.cases.mm <- nelda2.sub %>% filter(n.events.massmob.post >= 1)
+
+ ## This yields only 43 observations, do not run further.
+
+model.n.base.mm <- glm.nb(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  + v2elintmon +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
+summary(model.n.base.mm)
+```
+
 \newpage

### 2022-11-16 Output/appendix2.Rmd
Various changes to add new Nelda and MassMob variables; moved MassMob models to appendix 2

---
title: "R Notebook"
output: html_notebook
---

```{r setup, include = FALSE}
library(tidyverse)
library(countrycode)
library(readxl)
library(vdemdata)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
library(stargazer)
library(cowplot)
library(pscl)
library(modelsummary)
library(ggeffects)
library(MASS)

nelda3.sub <- read.csv(here::here("Data", "nelda3_sub.csv"))

nelda3.sub <- nelda3.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda3.sub <- nelda3.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))

##Naming nelda variables

nelda3.sub <- nelda3.sub %>% rename(boycott = nelda14)
nelda3.sub <- nelda3.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                 NA)))

##Creating economic crisis variable

nelda3.sub <- nelda3.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
nelda3.sub <- nelda3.sub %>% mutate(econ.crisis = as.factor(econ.crisis))

nelda3.sub <- nelda3.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
nelda3.sub <- nelda3.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))

nelda3.sub <- nelda3.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))

##Removing cases before 1990, when ECAV starts
nelda3.sub <- nelda3.sub %>% filter(year >= 1990)

## Removing cases where inc. loses
nelda3.sub <- nelda3.sub %>% filter(inc.lose == 0)

##Intmon as a factor
nelda3.sub <- nelda3.sub %>% mutate(v2elintmon = as.factor(v2elintmon))

nelda3.sub <- nelda3.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))

#Square of physint

nelda3.sub <- nelda3.sub %>% mutate(physintlag.sq = physinteg.1lag^2)

#Crisis as factor

nelda3.sub <- nelda3.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))

#Margins with updates

nelda3.sub <- nelda3.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))

#Successful protest = new elections or new incumbent

nelda3.sub <- nelda3.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))

#Protest cases only

protest.cases <- nelda3.sub %>% filter(any.ecav.protest == 1)

# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections

nelda3.sub <- nelda3.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))

## Changing participants from numeric category to text
nelda3.sub <- nelda3.sub %>% mutate(median.participants = round(median.participants))
nelda3.sub <- nelda3.sub %>% mutate(particpants_categorical = ifelse(is.na(median.participants)==T, NA, ifelse(median.participants == 1, "Less than 10",
                     ifelse(2, "10 to 99", ifelse(3, "100 to 999",
                                                  ifelse(4, "1000 to 9,999", ifelse(5, "Greater than 10,000", NA)))))))
nelda3.sub <- nelda3.sub %>% mutate(particpants_categorical = as.factor(particpants_categorical))

## Median particiants (MM) as factor

nelda3.sub <- nelda3.sub %>% mutate(median.participants.mm.post = as.factor(median.participants.mm.post))

```

## Protest size (categories) using mass mobilization data

```{r}
model.size.base <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp +  pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre, data = nelda3.sub %>% filter(gdpgro.1lag < 1))
summary(model.size.base)

model.size.base2 <- polr(median.participants.mm.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre, data = nelda3.sub %>% filter(gdpgro.1lag < 1))
summary(model.size.base2)

model.size.griev <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre, data = nelda3.sub %>% filter(gdpgro.1lag < 1))
summary(model.size.griev)

model.size.griev2 <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + + harass_opp + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre, data = nelda3.sub)
summary(model.size.griev2)

model.size.griev3 <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + altinf.1lag + harass_opp +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre, data = nelda3.sub)
summary(model.size.griev3)
```

Plotting above models

```{r}
##Winning margin
dat <- ggpredict(model.size.base2, terms = c("winner.share.upd", "v2elirreg.inv [-.47, 1.77]"))
dat.sub <- dat %>% dplyr::filter(response.level == "0" | response.level == "2" | response.level == "4" | response.level == "6")
plot(dat.sub)

dat_max <- ggpredict(model.size.base2, terms = c("winner.share.upd", "v2elirreg.inv [-2.3, 3.2]"))
dat.sub <- dat_max %>% dplyr::filter(response.level == "0" | response.level == "3" | response.level == "6")
plot(dat.sub)
```

```{r}
## Econ grievance
dat <- ggpredict(model.size.griev2, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv"))
dat.sub <- dat %>% dplyr::filter(response.level == "0" | response.level == "2" | response.level == "4" | response.level == "6")
plot(dat.sub)

dat_max <- ggpredict(model.size.griev2, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-2.3, 3.2]"))
dat.sub <- dat %>% dplyr::filter(response.level == "0" | response.level == "3" | response.level == "6")
plot(dat.sub)
```

## Hurdle models with MM quantity

```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
model.n.base.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda3.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
summary(model.n.base.hurd.mm)

model.n.base2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda3.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
summary(model.n.base2.hurd.mm)

model.n.griev.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, dist = "negbin", data = nelda3.sub %>% filter(gdpgro.1lag < 1))
summary(model.n.griev.hurd.mm)

model.n.griev2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda3.sub, dist = "negbin")
summary(model.n.griev2.hurd.mm)

model.n.griev3.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda3.sub, dist = "negbin")
summary(model.n.griev3.hurd.mm)
```

## Plots for hurdles

```{r}
dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd", "v2elirreg.inv [-.45, 1.78]")) 
p.mm.margin <- plot(dat)
p.mm.margin + coord_cartesian(ylim = c(0, 100))
```

```{r custom-tidiers, include=FALSE}
tidy.hurdle.count <- function(x, conf.int = FALSE, conf.level = 0.95, ...) {

  result <- summary(x)$coefficients$count %>%
    tibble::as_tibble(rownames = "term") #%>%
    #dplyr::rename(estimate = Estimate,
    #              std.error = `Std. Error`,
    #              statistic = `t value`,
     #             p.value = `Pr(>|t|)`)

  if (conf.int) {
    ci <- confint(x, level = conf.level)
    result <- dplyr::left_join(result, ci, by = "term")
  }

  result
}

#test <- summary(model.n.base.hurd)$coefficients$count

test <- tidy.hurdle.count(model.n.base.hurd)

#tbl_regression(test)
```

```{r zero-model-table, echo=FALSE}

models <- list()
models[['Model 1']] <- model.n.base.hurd
models[['Model 2']] <- model.n.base2.hurd
models[['Model 3']] <- model.n.griev.hurd
models[['Model 4']] <- model.n.griev2.hurd
models[['Model 5']] <- model.n.griev3.hurd

modelsummary::modelsummary(models, coef_omit = "count", gof_omit = "R2", stars = T)

```

### 2022-11-16 Output/manipulation_protest_draft_v2.Rmd
Various changes to add new Nelda and MassMob variables; moved MassMob models to appendix 2

@@
 ```{r data-summary, echo=FALSE}
-incl.vars <- nelda2.sub %>% dplyr::select(jucon.1lag, legcon.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, any.ecav.protest, n.events.post)
+incl.vars <- nelda2.sub %>% dplyr::select(jucon.1lag, legcon.1lag, altinf.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, harass_opp, any.ecav.protest, n.events.post)
+#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))
 
@@
 
-model.ecavbin.mm <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    osorg.1lag +   winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
+model.ecavbin.mm <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    osorg.1lag +   winner.share.upd   + presidential + inc.termlimit.num +  harass_opp  + physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
 #summary(model.ecavbin.mm)
 
-model.ecavbin.mm2 <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd + pubseccorrup.lag  +   osorg.1lag +  winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
+model.ecavbin.mm2 <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag +  econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd + pubseccorrup.lag  +   osorg.1lag +  winner.share.upd   + presidential + inc.termlimit.num  + harass_opp +  physinteg.1lag + physintlag.sq + urban.pct  + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
 #summary(model.ecavbin.mm2)
-p1 <- sjPlot::plot_model(model.ecavbin.mm2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.41, 1.79]"))
+p1 <- sjPlot::plot_model(model.ecavbin.mm2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.44, 2.23]")) #Note extreme values here: 10th and 90th percentiles
 
-model.ecavbin.mm.type <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd +  pubseccorrup.lag+  osorg.1lag +  winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate  + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
+model.ecavbin.mm.type <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag +  econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd +  pubseccorrup.lag+  osorg.1lag +  winner.share.upd   + presidential + inc.termlimit.num +   physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate  + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
 #summary(model.ecavbin.mm.type)
@@
 
-model.ecavbin.mm.crisis <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag +     econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
+model.ecavbin.mm.crisis <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag +     econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num +  harass_opp +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
 #summary(model.ecavbin.mm.crisis)
@@
 
-model.ecavbin.mm.corr <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag +  osorg.1lag + pubseccorrup.lag +  natlpartyorg.lag +  pubseccorrup.lag +     pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
+model.ecavbin.mm.corr <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + pubseccorrup.lag +  natlpartyorg.lag +  pubseccorrup.lag +     pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num + harass_opp +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
 #summary(model.ecavbin.mm.corr)
@@
 
-model.ecavbin.mm.gdp <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag + osorg.1lag +   natlpartyorg.lag +     gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub %>% filter(gdpgro.1lag < 1))
+model.ecavbin.mm.gdp <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + osorg.1lag +   natlpartyorg.lag +     gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num + harass_opp +  physinteg.1lag +  physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub %>% filter(gdpgro.1lag < 1))
 #summary(model.ecavbin.mm.gdp)
@@
 
-#model.ecavbin.mm3 <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag +  pubseccorrup.lag +  electoral.dem +   winner.share.upd*v2elirreg.inv*jucon.1lag + legcon.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq +  urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub %>% filter(inc.lose == 0))
+#model.ecavbin.mm3 <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag +  pubseccorrup.lag +  electoral.dem +   winner.share.upd*v2elirreg.inv*jucon.1lag + legcon.1lag + altinf.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq +  urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub %>% filter(inc.lose == 0))
 #summary(model.ecavbin.mm3)
@@
 ```{r count-models-base, warning = FALSE, message = FALSE, include = F}
-model.d.base <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
+model.d.base <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  + harass_opp +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
 #summary(model.d.base)
 
-model.n.base <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
+model.n.base <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  + harass_opp +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
 #summary(model.n.base)
 
-model.n.base.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
+model.n.base.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  + harass_opp +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  + harass_opp + physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
 summary(model.n.base.hurd)
@@
 ```{r count-models-interact, warning = FALSE, message = FALSE, include = F}
-model.d.base2 <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
+model.d.base2 <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num + harass_opp +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
 #summary(model.d.base2)
@@
 
-model.n.base2 <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
+model.n.base2 <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num + harass_opp  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
 #summary(model.n.base2)
@@
 
-model.n.base2.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
+model.n.base2.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  + harass_opp +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  + harass_opp + physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
 summary(model.n.base2.hurd)
@@
 ```{r, count-models-interact-griev, warning = FALSE, message = FALSE, include = F}
-model.d.griev <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +    gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
+model.d.griev <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +    gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  + harass_opp +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
 #summary(model.d.griev)
@@
 
-model.n.griev <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
+model.n.griev <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num + harass_opp  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
 #summary(model.n.griev)
@@
 
-model.n.griev.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, dist = "negbin", data = nelda2.sub %>% filter(gdpgro.1lag < 1))
+model.n.griev.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num + harass_opp  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num + harass_opp  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, dist = "negbin", data = nelda2.sub %>% filter(gdpgro.1lag < 1))
 summary(model.n.griev.hurd)
@@
 ```{r, count-models-interact-griev2, warning = FALSE, message = FALSE, include=F}
-model.d.griev2 <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv +  econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
+model.d.griev2 <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv +  econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num + harass_opp  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
 #summary(model.d.griev2)
@@
 
-model.n.griev2 <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
+model.n.griev2 <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num + harass_opp +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
 #summary(model.n.griev2)
@@
 
-model.n.griev2.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub, dist = "negbin")
+model.n.griev2.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  + harass_opp + physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num + harass_opp  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub, dist = "negbin")
 summary(model.n.griev2.hurd)
@@
 ```{r, count-models-interact-griev3, warning = FALSE, message = FALSE, include=F}
-model.d.griev3 <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +   natlpartyorg.lag + v2elirreg.inv +  pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
+model.d.griev3 <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +   natlpartyorg.lag + v2elirreg.inv +  pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num + harass_opp  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
 #summary(model.d.griev3)
@@
 
-model.n.griev3 <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
+model.n.griev3 <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num + harass_opp  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
 #summary(model.n.griev3)
@@
 
-model.n.griev3.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub, dist = "negbin")
+model.n.griev3.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  + harass_opp + physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  + harass_opp +   physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub, dist = "negbin")
 summary(model.n.griev3.hurd)
@@
 ```{r tables-binary, echo=FALSE}
-models.binary <- c(model.ecavbin.mm, model.ecavbin.mm2, model.ecavbin.mm.type, model.ecavbin.mm.corr, model.ecavbin.mm.crisis, model.ecavbin.mm.gdp)
+models.binary <- c(model.ecavbin.mm, model.ecavbin.mm2, model.ecavbin.mm.corr, model.ecavbin.mm.crisis, model.ecavbin.mm.gdp)
 

### 2022-11-16 Scripts/data_cleaning.Rmd
Various changes to add new Nelda and MassMob variables; moved MassMob models to appendix 2

@@
 nelda <- nelda %>% rename(inc.replaced.prot = nelda41)
+nelda <- nelda %>% rename(preelection_concerns = nelda11)
+nelda <- nelda %>% rename(harass_opp = nelda15)
+nelda <- nelda %>% rename(opp_leaders_prevented = nelda13)
 
@@
 nelda <- nelda %>% mutate(inc.termlimit.num = ifelse(types == "Legislative/Parliamentary" & inc.termlimit == "N/A", 0, ifelse(types == "Executive" & inc.termlimit == "yes", 1,                                                                                                                       ifelse(types == "Executive" & inc.termlimit == "no", 0, NA))))
+nelda <- nelda %>% mutate(preelection_concerns = ifelse(preelection_concerns == "yes", 1, ifelse(preelection_concerns == "no", 0, NA)))
+nelda <- nelda %>% mutate(harass_opp = ifelse(nelda3 == "No", 1, ifelse(harass_opp == "yes", 1, ifelse(harass_opp == "no", 0, NA))))
+nelda <- nelda %>% mutate(opp_leaders_prevented = ifelse(nelda3 == "No", "In general", ifelse(opp_leaders_prevented == "yes", "Specifically", ifelse(opp_leaders_prevented == "no", "No", NA))))
 
@@
 nelda3 <- nelda3 %>% mutate(election_year = year)
-nelda3 <- massmob %>% left_join(nelda3, by = c("ccode", "year")) #Nelda3 is in progress, need to figure out a) if needed, b) how to include pre-election and post-election mobilization
+nelda3 <- massmob %>% left_join(nelda3, by = c("ccode", "year"))
+```
+
+##Nelda2 looks at NELDA variables as DV
+
+```{r}
+rm(vdem.data)
+rm(vdem.recent)
+rm(lji)
+rm(nelda)
+rm(massmob)
+```
+
+```{r}
+nelda2.sub <- nelda2 %>% filter(v2x_regime == 1 | v2x_regime == 2) #Note that coding scheme for electoral democracies allows for "some degree of fraud and irregularities but these did not in the end affect the outcome" (v2elfrfair, which is a component of v2x_regime)
+```
+
+```{r}
+ecav.pre <- ecav %>% filter(Date < Electiondate)
+ecav.pre <- ecav.pre %>% filter(str_detect(EventName, "rotest") == T) #Only captures protests
+
+ecav <- ecav %>% filter(Date >= Electiondate)
+ecav <- ecav %>% filter(str_detect(EventName, "rotest") == T) #Only captures protests
+```
+
+Grouped summaries
+```{r}
+count_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(n.events.pre = n())
+count_data_post <- ecav %>% group_by(NeldaID) %>% summarize(n.events.post = n())
+
+##Duration
+max_date <- ecav %>% group_by(NeldaID) %>% summarize(latest.event = max(Date))
+min_date <- ecav %>% group_by(NeldaID) %>% summarize(earliest.event = min(Date))
+duration_data <- left_join(min_date, max_date, by = "NeldaID")
+duration_data <- duration_data %>% mutate(d.events = as.duration(interval(earliest.event, latest.event)))
+duration_data <- duration_data %>% mutate(d.events.num = as.numeric(d.events, "days"))
+duration_data <- duration_data %>% mutate(d.events.num = ifelse(d.events.num == 0, 1, d.events.num)) #1-day events
+
+##Size
+ecav <- ecav %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
+size_data <- ecav %>% group_by(NeldaID) %>% summarize(median.participants = median(ParticipantNumber, na.rm = T))
+
+ecav.pre <- ecav.pre %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
+size_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(median.participants.pre = median(ParticipantNumber, na.rm = T))
+
+## Violence
+violence_data <- ecav %>% group_by(NeldaID) %>% summarize(antigov.violence = sum(antigov.violence, na.rm = T))
+violence_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(antigov.violence.pre = sum(antigov.violence, na.rm = T))
+
+##Combined
+ecav_data <- left_join(count_data_post, duration_data, by = "NeldaID")
+ecav_data <- left_join(ecav_data, size_data, by = "NeldaID")
+ecav_data <- left_join(ecav_data, violence_data, by = "NeldaID")
+
+ecav_data_pre <- left_join(count_data_pre, size_data_pre, by = "NeldaID")
+ecav_data_pre <- left_join(ecav_data_pre, violence_data_pre, by = "NeldaID")
+
+ecav_data <- left_join(ecav_data, ecav_data_pre, by = "NeldaID")
+
+## Merge back with NELDA 2
+```
+
+Merge ECAV with NELDA
+
+```{r}
+nelda2.sub <- nelda2.sub %>% rename(NeldaID = electionid)
+nelda2.sub <- left_join(nelda2.sub, ecav_data, by = "NeldaID")
+
+##Cleaning
+
+nelda2.sub <- nelda2.sub %>% mutate(antigov.violence = ifelse(is.na(antigov.violence)==T, 0, antigov.violence))
+nelda2.sub <- nelda2.sub %>% mutate(n.events.post = ifelse(is.na(n.events.post)==T, 0, n.events.post))
+nelda2.sub <- nelda2.sub %>% mutate(d.events.num = ifelse(is.na(d.events.num)==T, 0, d.events.num))
+#nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(is.na(median.participants)==T, 0, median.participants))
+
+nelda2.sub <- nelda2.sub %>% mutate(any.ecav.protest = ifelse(n.events.post >= 1, 1, 0))
+
+##Making the assumption that NAs for pre-election protest are zero
+nelda2.sub <- nelda2.sub %>% mutate(n.events.pre = ifelse(is.na(n.events.pre) == T, 0, n.events.pre))
+nelda2.sub <- nelda2.sub %>% mutate(antigov.violence.pre = ifelse(is.na(antigov.violence.pre) == T, 0, antigov.violence.pre))
+
+```
+
+```{r}
+#nelda3 <- nelda3 %>% mutate(protests.postelex = ifelse(protests == 1 & protesterdemand1 == "political behavior, process", 1, 0))
+
+#nelda3 <- nelda3 %>% mutate(protests.postelex.quant = ifelse(protests.postelex == 1, protestnumber, 0))
+```
+
+## World Bank urbanization data
+
+```{r}
+urban <- read.csv(here::here("Data", "urbanization_worldbank.csv"))
+test <- urban %>%
+  pivot_longer(cols = starts_with("X"),
+               names_to = "year")
+test <- test %>% mutate(year = gsub("X", "", year))
+
+#Next is line up country codes in test and nelda 2, then merge
+
+library(countrycode)
+
+test <- test %>% mutate(stateid = countrycode(sourcevar = Country.Code, origin = 'iso3c', destination = 'gwc'))
+test <- test %>% rename(urban.pct = value)
+test <- test %>% dplyr::select(year, urban.pct, stateid)
+
+nelda2.sub <- nelda2.sub %>% mutate(year = as.character(year))
+
+nelda2.sub <- nelda2.sub %>% left_join(test, by = c("stateid", "year"))
+
 ```
@@
 
-size_data_mm_pre <- nelda3.sub.pre %>% group_by(electionid) %>% summarize(median.participants.pre = median(participants_category_ordered, na.rm = T))
-size_data_mm_post <- nelda3.sub.post %>% group_by(electionid) %>% summarize(median.participants.post = median(participants_category_ordered, na.rm = T))
+size_data_mm_pre <- nelda3.sub.pre %>% group_by(electionid) %>% summarize(median.participants.mm.pre = median(participants_category_ordered, na.rm = T))
+size_data_mm_post <- nelda3.sub.post %>% group_by(electionid) %>% summarize(median.participants.mm.post = median(participants_category_ordered, na.rm = T))
 
-```
-
-##Nelda2 looks at NELDA variables as DV
-
-```{r}
-rm(vdem.data)
-rm(vdem.recent)
-rm(lji)
-rm(nelda)
-rm(massmob)
-```
-
-```{r}
-nelda2.sub <- nelda2 %>% filter(v2x_regime == 1 | v2x_regime == 2) #Note that coding scheme for electoral democracies allows for "some degree of fraud and irregularities but these did not in the end affect the outcome" (v2elfrfair, which is a component of v2x_regime)
-```
-
-```{r}
-ecav.pre <- ecav %>% filter(Date < Electiondate)
-ecav.pre <- ecav.pre %>% filter(str_detect(EventName, "rotest") == T) #Only captures protests
-
-ecav <- ecav %>% filter(Date >= Electiondate)
-ecav <- ecav %>% filter(str_detect(EventName, "rotest") == T) #Only captures protests
-```
-
-Grouped summaries
-```{r}
-count_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(n.events.pre = n())
-count_data_post <- ecav %>% group_by(NeldaID) %>% summarize(n.events.post = n())
-
-##Duration
-max_date <- ecav %>% group_by(NeldaID) %>% summarize(latest.event = max(Date))
-min_date <- ecav %>% group_by(NeldaID) %>% summarize(earliest.event = min(Date))
-duration_data <- left_join(min_date, max_date, by = "NeldaID")
-duration_data <- duration_data %>% mutate(d.events = as.duration(interval(earliest.event, latest.event)))
-duration_data <- duration_data %>% mutate(d.events.num = as.numeric(d.events, "days"))
-duration_data <- duration_data %>% mutate(d.events.num = ifelse(d.events.num == 0, 1, d.events.num)) #1-day events
-
-##Size
-ecav <- ecav %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
-size_data <- ecav %>% group_by(NeldaID) %>% summarize(median.participants = median(ParticipantNumber, na.rm = T))
-
-ecav.pre <- ecav.pre %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
-size_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(median.participants.pre = median(ParticipantNumber, na.rm = T))
-
-## Violence
-violence_data <- ecav %>% group_by(NeldaID) %>% summarize(antigov.violence = sum(antigov.violence, na.rm = T))
-violence_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(antigov.violence.pre = sum(antigov.violence, na.rm = T))
-
-##Combined
-ecav_data <- left_join(count_data_post, duration_data, by = "NeldaID")
-ecav_data <- left_join(ecav_data, size_data, by = "NeldaID")
-ecav_data <- left_join(ecav_data, violence_data, by = "NeldaID")
-
-ecav_data_pre <- left_join(count_data_pre, size_data_pre, by = "NeldaID")
-ecav_data_pre <- left_join(ecav_data_pre, violence_data_pre, by = "NeldaID")
-
-ecav_data <- left_join(ecav_data, ecav_data_pre, by = "NeldaID")
-
-## Merge back with NELDA 2
-```
-
-Merge ECAV with NELDA
-
-```{r}
-nelda2.sub <- nelda2.sub %>% rename(NeldaID = electionid)
-nelda2.sub <- left_join(nelda2.sub, ecav_data, by = "NeldaID")
-
-##Cleaning
-
-nelda2.sub <- nelda2.sub %>% mutate(antigov.violence = ifelse(is.na(antigov.violence)==T, 0, antigov.violence))
-nelda2.sub <- nelda2.sub %>% mutate(n.events.post = ifelse(is.na(n.events.post)==T, 0, n.events.post))
-nelda2.sub <- nelda2.sub %>% mutate(d.events.num = ifelse(is.na(d.events.num)==T, 0, d.events.num))
-#nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(is.na(median.participants)==T, 0, median.participants))
-
-nelda2.sub <- nelda2.sub %>% mutate(any.ecav.protest = ifelse(n.events.post >= 1, 1, 0))
-
-##Making the assumption that NAs for pre-election protest are zero
-nelda2.sub <- nelda2.sub %>% mutate(n.events.pre = ifelse(is.na(n.events.pre) == T, 0, n.events.pre))
-nelda2.sub <- nelda2.sub %>% mutate(antigov.violence.pre = ifelse(is.na(antigov.violence.pre) == T, 0, antigov.violence.pre))
-
-```
-
-```{r}
-#nelda3 <- nelda3 %>% mutate(protests.postelex = ifelse(protests == 1 & protesterdemand1 == "political behavior, process", 1, 0))
-
-#nelda3 <- nelda3 %>% mutate(protests.postelex.quant = ifelse(protests.postelex == 1, protestnumber, 0))
-```
-
-## World Bank urbanization data
-
-```{r}
-urban <- read.csv(here::here("Data", "urbanization_worldbank.csv"))
-test <- urban %>%
-  pivot_longer(cols = starts_with("X"),
-               names_to = "year")
-test <- test %>% mutate(year = gsub("X", "", year))
-
-#Next is line up country codes in test and nelda 2, then merge
-
-library(countrycode)
-
-test <- test %>% mutate(stateid = countrycode(sourcevar = Country.Code, origin = 'iso3c', destination = 'gwc'))
-test <- test %>% rename(urban.pct = value)
-test <- test %>% dplyr::select(year, urban.pct, stateid)
-
-nelda2.sub <- nelda2.sub %>% mutate(year = as.character(year))
-
-nelda2.sub <- nelda2.sub %>% left_join(test, by = c("stateid", "year"))
-
-```
-
-## Merge with winner share updated
-
-```{r}
-winner.shares <- read.csv(here::here("Data", "amended_vote_shares.csv"))
-nelda2.sub <- nelda2.sub %>% left_join(winner.shares, by = "NeldaID")
 ```
@@
 
-nelda2.sub <- nelda2.sub %>% left_join(mm_data_comb, by = "NeldaID")  ## Double check to make sure this joining works
+nelda3.sub <- nelda2.sub %>% left_join(mm_data_comb, by = "NeldaID")  ## Double check to make sure this joining works
+
+## Code NAs for number of protests as 0s (these reflect election years recorded in NELDA where there were no protests recorded by MM)
+
+nelda3.sub <- nelda3.sub %>% mutate(n.events.massmob.pre = ifelse(is.na(n.events.massmob.pre)==T, 0, n.events.massmob.pre))
+nelda3.sub <- nelda3.sub %>% mutate(n.events.massmob.post = ifelse(is.na(n.events.massmob.post)==T, 0, n.events.massmob.post))
+
+nelda3.sub <- nelda3.sub %>% mutate(median.participants.mm.pre = ifelse(is.na(median.participants.mm.pre)==T & n.events.massmob.pre == 0, 0, median.participants.mm.pre))
+nelda3.sub <- nelda3.sub %>% mutate(median.participants.mm.post = ifelse(is.na(median.participants.mm.post)==T & n.events.massmob.post == 0, 0, median.participants.mm.pre))
+
+## Rounding participant median numbers down (removes half-steps)
+
+nelda3.sub <- nelda3.sub %>% mutate(median.participants.mm.post = floor(median.participants.mm.post))
+nelda3.sub <- nelda3.sub %>% mutate(median.participants.mm.pre = floor(median.participants.mm.pre))
+
+```
+
+## Merge with winner share updated
+
+```{r}
+winner.shares <- read.csv(here::here("Data", "amended_vote_shares.csv"))
+nelda2.sub <- nelda2.sub %>% left_join(winner.shares, by = "NeldaID")
+nelda3.sub <- nelda3.sub %>% left_join(winner.shares, by = "NeldaID")
 

### 2022-11-17 Output/appendix2.Rmd
Working to build a v3 draft that highlights MM data

@@
 
+## Inverting the vote buying measure so it is not reverse coded
+
+nelda3.sub <- nelda3.sub %>% mutate(v2elvotbuy_inv = -1*v2elvotbuy)
+
+```
+
+```{r}
+incl.vars <- nelda3.sub %>% dplyr::select(jucon.1lag, legcon.1lag, altinf.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, harass_opp, any.ecav.protest, median.participants.mm.pre, median.participants.mm.post)
+#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))
+
+datasummary_skim(incl.vars)
 ```
@@
 ```{r}
-model.size.base <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp +  pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre, data = nelda3.sub %>% filter(gdpgro.1lag < 1))
+model.size.base <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp +  pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda3.sub %>% filter(gdpgro.1lag < 1))
 summary(model.size.base)
 
-model.size.base2 <- polr(median.participants.mm.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre, data = nelda3.sub %>% filter(gdpgro.1lag < 1))
+model.size.base2 <- polr(median.participants.mm.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda3.sub %>% filter(gdpgro.1lag < 1))
 summary(model.size.base2)
 
-model.size.griev <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre, data = nelda3.sub %>% filter(gdpgro.1lag < 1))
+model.size.griev <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda3.sub %>% filter(gdpgro.1lag < 1))
 summary(model.size.griev)
 
-model.size.griev2 <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + + harass_opp + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre, data = nelda3.sub)
+model.size.griev2 <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + + harass_opp + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda3.sub)
 summary(model.size.griev2)
 
-model.size.griev3 <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + altinf.1lag + harass_opp +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre, data = nelda3.sub)
+model.size.griev3 <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + altinf.1lag + harass_opp +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda3.sub)
 summary(model.size.griev3)
@@
 dat <- ggpredict(model.size.base2, terms = c("winner.share.upd", "v2elirreg.inv [-.47, 1.77]"))
-dat.sub <- dat %>% dplyr::filter(response.level == "0" | response.level == "2" | response.level == "4" | response.level == "6")
+dat.sub <- dat %>% dplyr::filter(response.level == "0" | response.level == "1" | response.level == "2" | response.level == "3")
+plot(dat.sub)
+
+dat.sub <- dat %>% dplyr::filter(response.level == "4" | response.level == "5" | response.level == "6")
 plot(dat.sub)
@@
 dat_max <- ggpredict(model.size.base2, terms = c("winner.share.upd", "v2elirreg.inv [-2.3, 3.2]"))
-dat.sub <- dat_max %>% dplyr::filter(response.level == "0" | response.level == "3" | response.level == "6")
-plot(dat.sub)
+dat.sub <- dat_max %>% dplyr::filter(response.level == "2" | response.level == "4" | response.level == "6")
+plot(dat.sub) + coord_cartesian(ylim = c(0, .25))
 ```
@@
 ```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
-model.n.base.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda3.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
+model.n.base.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, data = nelda3.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
 summary(model.n.base.hurd.mm)
 
-model.n.base2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda3.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
+model.n.base2.hurd.mm <- hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre  + v2elvotbuy_inv|  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, data = nelda3.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
 summary(model.n.base2.hurd.mm)
 
-model.n.griev.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, dist = "negbin", data = nelda3.sub %>% filter(gdpgro.1lag < 1))
+model.n.griev.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv| jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, dist = "negbin", data = nelda3.sub %>% filter(gdpgro.1lag < 1))
 summary(model.n.griev.hurd.mm)
 
-model.n.griev2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda3.sub, dist = "negbin")
+model.n.griev2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv | jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, data = nelda3.sub, dist = "negbin")
 summary(model.n.griev2.hurd.mm)
 
-model.n.griev3.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda3.sub, dist = "negbin")
+model.n.griev3.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv | jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, data = nelda3.sub, dist = "negbin")
 summary(model.n.griev3.hurd.mm)

### 2022-11-17 Output/manipulation_protest_draft_v3.Rmd
Working to build a v3 draft that highlights MM data

--- 2022-11-07 Output/manipulation_protest_draft_v2.Rmd
@@
 library(stargazer)
+library(cowplot)
 library(pscl)
 library(modelsummary)
+library(ggeffects)
+library(MASS)
 
-nelda2.sub <- read.csv(here::here("Data", "nelda2_sub.csv"))
+nelda3.sub <- read.csv(here::here("Data", "nelda3_sub.csv"))
 
-nelda2.sub <- nelda2.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
-nelda2.sub <- nelda2.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))
+nelda3.sub <- nelda3.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
+nelda3.sub <- nelda3.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))
 
@@
 
-nelda2.sub <- nelda2.sub %>% rename(boycott = nelda14)
-nelda2.sub <- nelda2.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
+nelda3.sub <- nelda3.sub %>% rename(boycott = nelda14)
+nelda3.sub <- nelda3.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                  NA)))
@@
 
-nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
-nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = as.factor(econ.crisis))
+nelda3.sub <- nelda3.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
+nelda3.sub <- nelda3.sub %>% mutate(econ.crisis = as.factor(econ.crisis))
 
-nelda2.sub <- nelda2.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
-nelda2.sub <- nelda2.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))
+nelda3.sub <- nelda3.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
+nelda3.sub <- nelda3.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))
 
-nelda2.sub <- nelda2.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))
+nelda3.sub <- nelda3.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))
 
 ##Removing cases before 1990, when ECAV starts
-nelda2.sub <- nelda2.sub %>% filter(year >= 1990)
+nelda3.sub <- nelda3.sub %>% filter(year >= 1990)
 
 ## Removing cases where inc. loses
-nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)
+nelda3.sub <- nelda3.sub %>% filter(inc.lose == 0)
 
 ##Intmon as a factor
-nelda2.sub <- nelda2.sub %>% mutate(v2elintmon = as.factor(v2elintmon))
+nelda3.sub <- nelda3.sub %>% mutate(v2elintmon = as.factor(v2elintmon))
 
-nelda2.sub <- nelda2.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))
+nelda3.sub <- nelda3.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))
 
@@
 
-nelda2.sub <- nelda2.sub %>% mutate(physintlag.sq = physinteg.1lag^2)
+nelda3.sub <- nelda3.sub %>% mutate(physintlag.sq = physinteg.1lag^2)
 
@@
 
-nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))
+nelda3.sub <- nelda3.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))
 
@@
 
-nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))
+nelda3.sub <- nelda3.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))
 
@@
 
-nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))
+nelda3.sub <- nelda3.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | elections.cancel.prot, 1, 0))
 
@@
 
-protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1)
+protest.cases <- nelda3.sub %>% filter(any.ecav.protest == 1)
 
@@
 
-nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))
+nelda3.sub <- nelda3.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))
+
+## Changing participants from numeric category to text
+nelda3.sub <- nelda3.sub %>% mutate(median.participants = round(median.participants))
+nelda3.sub <- nelda3.sub %>% mutate(particpants_categorical = ifelse(is.na(median.participants)==T, NA, ifelse(median.participants == 1, "Less than 10",
+                     ifelse(2, "10 to 99", ifelse(3, "100 to 999",
+                                                  ifelse(4, "1000 to 9,999", ifelse(5, "Greater than 10,000", NA)))))))
+nelda3.sub <- nelda3.sub %>% mutate(particpants_categorical = as.factor(particpants_categorical))
+
+## Median particiants (MM) as factor
+
+nelda3.sub <- nelda3.sub %>% mutate(median.participants.mm.post = as.factor(median.participants.mm.post))
+
+## Inverting the vote buying measure so it is not reverse coded
+
+nelda3.sub <- nelda3.sub %>% mutate(v2elvotbuy_inv = -1*v2elvotbuy)
 
@@
 
-ggplot(nelda2.sub, aes(x=n.events.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()
+ggplot(nelda3.sub, aes(x=n.events.massmob.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()
 
-ggplot(nelda2.sub, aes(x=winner.share.upd)) + geom_histogram(bins=100) + xlab("Incumbent vote-share") + theme_bw()
+ggplot(nelda3.sub, aes(x=winner.share.upd)) + geom_histogram(bins=100) + xlab("Incumbent vote-share") + theme_bw()
 ```
@@
 ```{r, include=FALSE}
-#nelda2.sub.success <- nelda2.sub %>% filter(protest.success == 1)
-mu <- plyr::ddply(nelda2.sub, "protest.success", summarise, grp.mean=mean(n.events.post))
-head(mu)
+
+p.success <- ggplot(nelda3.sub, aes(x=as.numeric(median.participants.mm.post), fill = as.factor(protest.success), color=as.factor(protest.success)))+
+  geom_histogram(position="identity", alpha=0.5, binwidth = 1) 
+p.success + coord_cartesian(ylim = c(0, 100))
 ```
@@
 ```{r, include=FALSE}
-p.success <- ggplot(nelda2.sub, aes(x=n.events.post, fill = as.factor(protest.success), color=as.factor(protest.success)))+
-  geom_histogram(position="identity", alpha=0.5) + 
+mu <- plyr::ddply(nelda3.sub, "protest.success", summarise, grp.mean=mean(n.events.massmob.post))
+head(mu)
+
+p.success <- ggplot(nelda3.sub, aes(x=n.events.massmob.post, fill = as.factor(protest.success), color=as.factor(protest.success)))+
+  geom_histogram(position="identity", alpha=0.5, binwidth = 1) + 
   geom_vline(data=mu, aes(xintercept=grp.mean, color=as.factor(protest.success)),
@@
 p.success
+```
+
+```{r}
+incl.vars <- nelda3.sub %>% dplyr::select(jucon.1lag, legcon.1lag, altinf.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, harass_opp, any.ecav.protest, median.participants.mm.pre, median.participants.mm.post)
+#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))
+
+datasummary_skim(incl.vars)
 ```
@@
 Figure 2 shows summary statistics for all of these variables.
-
-```{r data-summary, echo=FALSE}
-incl.vars <- nelda2.sub %>% dplyr::select(jucon.1lag, legcon.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, any.ecav.protest, n.events.post)
-
-datasummary_skim(incl.vars)
-
-```
-
-```{r include=F}
-
-datasummary_correlation(incl.vars)
-```
-
-```{r binary-models, warning = FALSE, message = FALSE, include = F}
-
-model.ecavbin.mm <- glm(inc.replaced.prot  ~ n.events.post,  data = nelda2.sub)
-#summary(model.ecavbin.mm)
-
-model.ecavbin.mm <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    osorg.1lag +   winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
-#summary(model.ecavbin.mm)
-
-model.ecavbin.mm2 <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd + pubseccorrup.lag  +   osorg.1lag +  winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
-#summary(model.ecavbin.mm2)
-p1 <- sjPlot::plot_model(model.ecavbin.mm2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.41, 1.79]"))
-
-model.ecavbin.mm.type <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd +  pubseccorrup.lag+  osorg.1lag +  winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate  + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
-#summary(model.ecavbin.mm.type)
-#sjPlot::plot_model(model.ecavbin.mm.type, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.41, 1.79]"))
-
-model.ecavbin.mm.crisis <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag +     econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
-#summary(model.ecavbin.mm.crisis)
-#sjPlot::plot_model(model.ecavbin.mm.crisis, type = "pred", terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-1, 2]"))
-
-model.ecavbin.mm.corr <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag +  osorg.1lag + pubseccorrup.lag +  natlpartyorg.lag +  pubseccorrup.lag +     pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub)
-#summary(model.ecavbin.mm.corr)
-#sjPlot::plot_model(model.ecavbin.mm.corr, type = "pred", terms = c("v2elirreg.inv", "pubseccorrup.lag"))
-
-model.ecavbin.mm.gdp <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag + osorg.1lag +   natlpartyorg.lag +     gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub %>% filter(gdpgro.1lag < 1))
-#summary(model.ecavbin.mm.gdp)
-#sjPlot::plot_model(model.ecavbin.mm.gdp, type = "pred", terms = c("gdpgro.1lag", "v2elirreg.inv [-2, 2]"))
-
-#model.ecavbin.mm3 <- glmer(any.ecav.protest  ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag +  pubseccorrup.lag +  electoral.dem +   winner.share.upd*v2elirreg.inv*jucon.1lag + legcon.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq +  urban.pct + rentierstate + taxstate + (1|country), nAGQ = 0, family = binomial,  data = nelda2.sub %>% filter(inc.lose == 0))
-#summary(model.ecavbin.mm3)
-#sjPlot::plot_model(model.ecavbin.mm3, type = "pred", terms = c("jucon.1lag + legcon.1lag [all]", "v2elirreg.inv [-.4, 1.79]", "winner.share.upd"))
-
-```
-
-```{r count-models-base, warning = FALSE, message = FALSE, include = F}
-model.d.base <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
-#summary(model.d.base)
-
-model.n.base <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
-#summary(model.n.base)
-
-model.n.base.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
-summary(model.n.base.hurd)
-```
-
-```{r count-models-interact, warning = FALSE, message = FALSE, include = F}
-model.d.base2 <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
-#summary(model.d.base2)
-#p2 <- sjPlot::plot_model(model.d.base2, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.36, 1.85]")) + coord_cartesian(ylim = c(0, 200))
-
-model.n.base2 <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
-#summary(model.n.base2)
-p2 <- sjPlot::plot_model(model.n.base2, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.36, 1.85]")) + coord_cartesian(ylim = c(0, 25))
-
-model.n.base2.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
-summary(model.n.base2.hurd)
-```
-
-```{r, count-models-interact-griev, warning = FALSE, message = FALSE, include = F}
-model.d.griev <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +    gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
-#summary(model.d.griev)
-sjPlot::plot_model(model.d.griev, type = "pred", terms = c("gdpgro.1lag", "v2elirreg.inv [.23, 2.12]")) + coord_cartesian(ylim = c(0, 200))
-
-model.n.griev <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases %>% filter(gdpgro.1lag < 1))
-#summary(model.n.griev)
-sjPlot::plot_model(model.n.griev, type = "pred", terms = c("gdpgro.1lag", "v2elirreg.inv [.23, 2.12]"))
-
-model.n.griev.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, dist = "negbin", data = nelda2.sub %>% filter(gdpgro.1lag < 1))
-summary(model.n.griev.hurd)
-```
-
-```{r, count-models-interact-griev2, warning = FALSE, message = FALSE, include=F}
-model.d.griev2 <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv +  econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
-#summary(model.d.griev2)
-p3 <- sjPlot::plot_model(model.d.griev2, type = "pred", terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-2, 4]")) + coord_cartesian(ylim = c(0, 200))
-
-model.n.griev2 <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
-#summary(model.n.griev2)
-sjPlot::plot_model(model.n.griev2, type = "pred", terms = c("v2elirreg.inv", "econ.crisis.nelda.1lag"))
-
-model.n.griev2.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub, dist = "negbin")
-summary(model.n.griev2.hurd)
-```
-
-```{r, count-models-interact-griev3, warning = FALSE, message = FALSE, include=F}
-model.d.griev3 <- glm.nb(d.events.num ~ jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +   natlpartyorg.lag + v2elirreg.inv +  pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
-#summary(model.d.griev3)
-sjPlot::plot_model(model.d.griev3, type = "pred", terms = c("v2elirreg.inv", "pubseccorrup.lag")) + coord_cartesian(ylim = c(0, 200))
-
-model.n.griev3 <- glm.nb(n.events.post ~ jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = protest.cases)
-#summary(model.n.griev3)
-sjPlot::plot_model(model.n.griev3, type = "pred", terms = c("v2elirreg.inv", "pubseccorrup.lag"))
-
-model.n.griev3.hurd <- hurdle(n.events.post ~ jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate | jucon.1lag + legcon.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate, data = nelda2.sub, dist = "negbin")
-summary(model.n.griev3.hurd)
-```
-
-```{r, include=FALSE}
-#models.quantity <- list("Model 13" = model.n.base, "Model 14" = model.n.base2, "Model 15" = model.n.griev, "Model 16" = model.n.griev2, "Model 17" = model.n.griev3)
-#modelsummary(tidytest)
-#modelsummary(models.quantity, stars = T)
-```
-
-# Results
-
-The results of these models generally support the conventional wisdom that more severe electoral manipulation is associated with an increased risk that *any* post-election protest will occur.
-Model 1, which includes no interaction terms, shows a positive relationship between fraud and protest which approaches the conventional p < .05 level of statistical significance.
-Models 2 and 3 indicate that this relationship is driven more by parties' strategic calculations than by economic grievance.
-As Figure 3 shows, the predicted probability of any protest occurring is highest when fraud is substantial and the winner's vote-share low, in keeping with the strategic logic of protest advanced by Simpser [-@simpser2013].
-However, the evidence does not support the argument that fraud makes protest more likely when economic grievances are more widespread.
-Neither GDP growth, the economic crisis indicator, or the measure of public sector corruption are associated with higher protest risk when election fraud is more widespread.
-Figure 4 shows the marginal effect of the interaction of *economic crisis* and election fraud, demonstrating that there is no significant increase in the likelihood of protest associated with  higher levels of fraud during bad economic times.
-Similar, insignificant results obtain for measures of corruption and GDP growth.
-
-```{r tables-binary, echo=FALSE}
-models.binary <- c(model.ecavbin.mm, model.ecavbin.mm2, model.ecavbin.mm.type, model.ecavbin.mm.corr, model.ecavbin.mm.crisis, model.ecavbin.mm.gdp)
-
-modelsummary(models.binary, stars = T, gof_omit = c("R2"), coef_rename = c("v2elirreg.inv:winner.share.upd" = "Fraud x winner share",
-                                                       "pubseccorrup.lag:v2elirreg.inv" = "Fraud x corruption",
-                                                       "econ.crisis.nelda.1lag1:v2elirreg.inv" = "Fraud x econ. crisis"))
-```
-
-```{r, echo=FALSE}
-p1 + theme_bw()
-```
-
-```{r, echo=FALSE}
-sjPlot::plot_model(model.ecavbin.mm.crisis, type = "pred", terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-.17, 2.01]")) + theme_bw()
-```
-
-```{r tables-duration, results='asis', echo=FALSE, include=FALSE}
-models.duration <- list("Model 8" = model.d.base, "Model 9" = model.d.base2, "Model 10" = model.d.griev, "Model 11" = model.d.griev2, "Model 12" = model.d.griev3)
-modelsummary(models.duration, stars = T)
-
-```
-
-```{r tables-quantity, echo=FALSE}
-models.quantity <- list("Model 8" = model.n.base, "Model 9" = model.n.base2, "Model 10" = model.n.griev, "Model 11" = model.n.griev2, "Model 12" = model.n.griev3)
-modelsummary(models.quantity, stars = T)
-```
-
-Turning to models of protest quantity presented in Table 3, there is no statistically significant relationship between fraud and protest in any of the models.
-Figure 4, using results from Model 9, presents the marginal effects for fraud conditional on the incumbent's vote-share.
-As the figure shows, there is no statistically significant difference in the expected number of protest following a relatively clean election compared to a relatively fraudulent one, for any margin of incumbent victory.
-Taking the results from Tables 2 and 3 together, there is support for Hypothesis 1 (an unconditional increase in the risk of any protest as fraud increases), as well as strong support for Hypothesis 2 (an increase in protest risk in fraudulent, low-margin elections) and Hypothesis 4 (no relationship between fraud and protest quantity).
-Hypothesis 3, on the connection between fraud, grievance, and protest, is not supported.
-
-We do observe support for the more structural explanations of more extensive protest.
-The coefficient for the *legislative constraint* variable, which measures the size and oppositional character of the legislative opposition, is consistently positive and  statistically significant.
-In Models 8 and 9 (those without a grievance interaction), the coefficient on *national party organization* is also positive and significant.
-The implication is that a larger, more institutionalized opposition party or coalition prolongs protest, in line with the 'test of strength' model.
-Similarly, the *incumbent term limit* variable is positive and consistently significant.
-This supports Hale's [-@hale_patronal_2014] argument that uncertainty around succession in non-democracies makes mobilization more intense as competing power centers emerge.
-
-```{r, echo=FALSE}
-p2 + geom_rug(sides = "b")  #Adding rug plot breaks theme_bw
-
-```
-
-<!-- Finally, the results are broadly similar for the number of protests that occur after the election.
-There is no significant relationship between the extent of fraud and the number of protests, though a smaller vote-share for the winner does appear to be associated with less protest.
-Neither GDP growth, corruption, or an economic crisis are significantly associated with protest number.
-Instead, once again, structural forces matter most.
-A large, active legislative opposition has a significant and positive relationship with protest size across all models, as does a term-limited chief executive.
-In two of the models based on economic conditions---Models 15 and 17---repression capability also appears to matter.
-A more permissive physical security environment is first associated with an increase in the number of protest, then a decrease.
-Finally, and perhaps surprisingly, a less restrictive environment for civil society or organizations is associated with a smaller number of post-election protests in the same models.
--->
-
-# Discussion
-
-Altogether, the findings challenge the idea that the expectation of protest will deter election-manipulation by ruling parties.
-To the extent that fraud is a signal of incumbent strength, it is a noisy and unreliable one.
-When an incumbent party wins by a narrow margin in a fraudulent election, the likelihood of protest initiation is substantially higher than when it wins more cleanly: roughly a 50\% chance in the former case versus 20\% in the latter as shown in Figure 2.
-However, since even strong incumbents commit fraud, the decision to take to the streets is often a losing gamble---anticipating a weak incumbent, they instead find one that is capable of repressing, coopting, or conciliating an emerging protest movement.
-Indeed, while there are 187 cases of protest out of the 647 elections in this dataset, only 15 elections resulted in either the incumbent being replaced due to protest or the election being re-run.\footnote{This information is determined using the NELDA variables NELDA37 and NELDA41.
-The cases of successful protest are: Ukraine 2004, Azerbaijan 2005, Georgia 2003, Kazakhstan 1994, Kyrgyzstan 2005, Bulgaria 1990, Malaysia 2004, Bangladesh 1996, Mali 1997, Peru 2000, Burkina Faso 1997, and Cote d'Ivoire 2000.} This effect can be seen in the results from the quantity models.
-
-The above results have several implications for understanding election manipulation and regime resilience.
-Most importantly, by treating protest primarily as a binary event rather, the existing literature on electoral protest has overemphasized the risk to incumbents of engaging in electoral manipulation.
-This results presented here show that most electoral protests are narrow, and that fraud has no bearing on how numerous protest events are.
-Moreover, contrary to some prior research, the risk of protest does not appear to be heightened by poor economic conditions [@Brancati2014a] or intensive corruption in the public sector.
-As a result, circumstances that are to some extent outside of leaders' control, like a worsening macroeconomic climate, do not appear to make engaging in fraud any riskier.
-Instead, the conditions that are correlated with more damaging protest are generally institutional factors, which are either more amenable to control or can be anticipated by non-democratic incumbents.
-Legislative constraints, an important factor for the duration and number of protests, can be adjusted through institutional reforms and cooptation of the opposition, over time.
-Incumbents may attempt to manage term limit through constitutional reforms, where possible, or through the appointment of (presumably) loyal successors.
-In other words, election fraud only has a minor cost in terms of protest risk, a risk which can be further blunted to the extent that elections can often be held on incumbents' terms.
-
-These are pessimistic findings for proponents of democracy.
-That electoral protest is less risky than is often assumed, in turn, implies that governments and ruling parties have a freer hand to engage in electoral manipulation than many models posit.
-Since the severity of fraud has no apparent bearing on whether or not a major protest wave will emerge as a serious threat to the regime, as shown in Figure 3, government have an incentive to manipulate as much as possible to secure a large winning vote-share.
-This perverse incentive undermines the case that democracy, and fair elections generally, are a self-enforcing equilibrium backed up by the risk of citizen protest [@fearon_self-enforcing_2011].
-
-One objection to this framing may be that, if the level of manipulation and the risk of protest are endogenous to one another, perhaps the null result found for the number of protests is to be expected.
-That is, if incumbents choose a level of manipulation calculated to avoid large-scale protest---sometimes high and sometimes low, conditional on contextual factors---no correlation between fraud and protest quantity would be found.
-This paper does not dispute that decisions about manipulation and protest may be endogenous to one another, but it does argue that the resulting equilibrium favors governments.
-In the dataset used here, only ten elections in four countries are coded as having an *intentional irregularities* score of less than -2 (indicating extremely fraud-free elections).
-By contrast, 103 elections in 27 countries score higher than 2 (indicating highly fraudulent elections).
-The frequency of fraud, the rarity of protest, and the finding that fraud severity is not correlated with the scope of protest events all point to the same conclusion: protest risk is not a substantial constraint on election fraud in non-democracies.
-Rather, the constraints on manipulation are more likely to be driven by factors like resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022].
-
-# Conclusion
-
-Protest risk is often considered to be a major risk to incumbent governments that work to undermine election integrity.
-This risk has a central role in several prominent models of electoral manipulation, serving to deter manipulation and to incentivize the toleration of election monitors.
-However, empirical research on the subject has largely treated protest as a binary variable, which obscures the degree to which post-election protest imposes a real risk on incumbents in non-democracies.
-One protest may be easily contained, while dozens may pose a more significant risk to the regime.
-
-This paper addresses this concern by testing the relationship between election fraud, the risk that any protest occurs, and the number of protest events that occur after an election.
-Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 647 elections to show that---while there is an elevated risk of some protest occurring in a fraudulent election with narrow margins of victory---election fraud has no significant relationship with the number of protests that occur in a protest wave.
-These results imply that protest risk should not be considered a central deterrent to election fraud.
-Instead, most post-election protests are strategic gambles gone wrong; the initiation of protest more often reveals the incumbent to be strong enough to manage protest through repression, cooptation, or sideways concessions.
-These findings indicate that, rather than being deterred by protest risk, governments have an incentive to manipulate as much as possible in order to achieve a large official vote-share.
-The evidence suggests that democratic breakthoughs in non-democratic elections are rare and unlikely to be driven by the severity of manipulation---a gloomy picture for supporters of democratization---and that researchers should consider other mechanisms beyond mass protest in their models of electoral manipulation.
-
-# References

### 2022-11-17 Scripts/data_cleaning.Rmd
Working to build a v3 draft that highlights MM data

@@
 
+nelda <- nelda %>% mutate(elections.cancel.prot = ifelse(is.na(new.elections.held)==T, NA, ifelse(nelda35 == "yes" & new.elections.held == "yes", 1, 0)))
 nelda <- nelda %>% mutate(first.mp.election = ifelse(first.mp.election == "yes", 1, ifelse(first.mp.election == "no", 0, NA)))

### 2022-11-17 Scripts/data_cleaning.Rmd
Reworking to ensure all pre-election and post-election protests from MM data are captured. Using nested for loops

@@
 library(readxl)
-library(vdemdata)
 library(lme4)
@@
 ecav <- read_xls(here::here("Data", "ECAV datatset_Version 1.2.xls"))
-vdem.data <- vdem #At the time of coding, this was v. 11.1
+vdem.data <- read_rds("C:/Users/colharv/Documents/Research projects/Country_Year_V-Dem_Full_others_R_v11.1/Country_Year_V-Dem_Full+others_R_v11.1/V-Dem-CY-Full+Others-v11.1.rds") #Switch between these two for work desktop vs.
+laptop
+#vdem.data <- vdem
 lji <- read.csv(here::here("Data", "LJI-estimates-20140422.csv"))
@@
 ```{r}
-nelda <- nelda %>% rename(first.mp.election = nelda2)
+nelda <- nelda %>% dplyr::rename(first.mp.election = nelda2)
 nelda <- nelda %>% rename(protests = nelda29)
@@
 nelda <- nelda %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "gwn", destination = "vdem"))
-
+ #Missing matches here look to be countries below the population threshold: some Carib. islands, Monaco, etc.
 lji <- lji %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "cown", 
@@
 massmob <- massmob %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "p4n", 
-                                               destination = "vdem"))
+                                               destination = "vdem")) #Missing here tend to be communist states that no longer exist
 
@@
 
-nelda3 <- nelda %>% dplyr::select(ccode, country, year, electionid, election_mdy)
+nelda3 <- nelda %>% dplyr::select(ccode, country, country_id, year, electionid, election_mdy)
 nelda3 <- nelda3 %>% mutate(election_year = year)
-nelda3 <- massmob %>% left_join(nelda3, by = c("ccode", "year"))
+nelda3 <- massmob %>% left_join(nelda3, by = c("country_id", "year"))
 ```
@@
 
+Excluding protest events in MM that take place more than 180 days before an election (events_pre) or more than 90 days after an election (events_post)
+
+```{r}
+elections_db <- nelda %>% dplyr::select(country_id, election_mdy) # A second dataframe with election dates by country
+
+nelda3.sub <- slice(nelda3, 0)
+
+for (i in 1:nrow(nelda3)) {
+  country_id_temp <- nelda3$country_id[i]
+  elections_db_temp <- elections_db %>% filter(country_id == country_id_temp)
+  protest_event_slice <- nelda3[i,]
+  for (j in 1:nrow(elections_db_temp)) {
+    elections_db_temp <- elections_db_temp %>% mutate(days_to_election_all =  election_mdy - protest_event_slice$protest_mdy)
+    #Negative numbers mean the protest is after the election, positive numbers mean the protest is before the election
+    elections_db_temp <- elections_db_temp %>% mutate(days_to_election_all_num = as.numeric(days_to_election_all))
+    elections_db_temp <- elections_db_temp %>% filter(days_to_election_all_num <= 180 & days_to_election_all_num >= -90)
+  }
+}
+```
+
 Cutting out protests before elections for nelda 3, 
@@
 #Add electionid to the year before the election and year after (so that grouping variable can capture those protests if needed)
-nelda3 <- nelda3 %>% group_by(ccode) %>% mutate(election_date_upcoming = as_date(ifelse(is.na(election_year)==F, election_mdy, 
+nelda3 <- nelda3 %>% group_by(country_id) %>% mutate(election_date_upcoming = as_date(ifelse(is.na(election_year)==F, election_mdy, 
                                                   ifelse(is.na(election_year)==T & is.na(lead(election_mdy))==F,
@@
 
-nelda3 <- nelda3 %>% group_by(ccode) %>% mutate(election_date_lagged = as_date(ifelse(is.na(election_year)==F, election_mdy,                                                         ifelse(is.na(election_year)==T & is.na(lag(election_mdy))==F, lag(election_mdy), NA))))
+nelda3 <- nelda3 %>% group_by(country_id) %>% mutate(election_date_lagged = as_date(ifelse(is.na(election_year)==F, election_mdy,                                                         ifelse(is.na(election_year)==T & is.na(lag(election_mdy))==F, lag(election_mdy), NA))))
 

### 2022-11-22 Scripts/data_cleaning_massmob.Rmd
Fixing a bug that introduced duplicates into the mass mob data

---
title: "Data cleaning for mass mobiliziation data version"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(countrycode)
library(readxl)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
```

Loading data

Note: check on how massmob data merges with nelda. We want cases where this is no protest after the election to score as 0.

```{r}
nelda <- read_xls(here::here("Data", "NELDA.xls"))
#ecav <- read_xls(here::here("Data", "ECAV datatset_Version 1.2.xls"))
#vdem.data <- read_rds("C:/Users/colharv/Documents/Research projects/Country_Year_V-Dem_Full_others_R_v11.1/Country_Year_V-Dem_Full+others_R_v11.1/V-Dem-CY-Full+Others-v11.1.rds") #Switch between these two for work desktop vs. laptop
vdem.data <- vdemdata::vdem
lji <- read.csv(here::here("Data", "LJI-estimates-20140422.csv"))
massmob <- read.csv(here::here("Data", "mmALL_073119_csv.csv"))
```

Cleaning nelda variables--this will have to be done for all used variables
```{r}
nelda <- nelda %>% dplyr::rename(first.mp.election = nelda2)
nelda <- nelda %>% rename(protests = nelda9)
nelda <- nelda %>% rename(inc.fav.cancel = nelda34)
nelda <- nelda %>% rename(inc.lose = nelda24)
nelda <- nelda %>% rename(inc.termlimit = nelda8)
nelda <- nelda %>% rename(econ.crisis.nelda = nelda18)
nelda <- nelda %>% rename(opp.votegain = nelda7)
nelda <- nelda %>% rename(nelda.protests.fraud = nelda30)
nelda <- nelda %>% rename(new.elections.held = nelda37)
nelda <- nelda %>% rename(inc.replaced.prot = nelda41)
nelda <- nelda %>% rename(preelection_concerns = nelda11)
nelda <- nelda %>% rename(harass_opp = nelda15)
nelda <- nelda %>% rename(opp_leaders_prevented = nelda13)

nelda <- nelda %>% mutate(elections.cancel.prot = ifelse(is.na(new.elections.held)==T, NA, ifelse(nelda35 == "yes" & new.elections.held == "yes", 1, 0)))
nelda <- nelda %>% mutate(first.mp.election = ifelse(first.mp.election == "yes", 1, ifelse(first.mp.election == "no", 0, NA)))
nelda <- nelda %>% mutate(protests = ifelse(protests == "yes", 1, ifelse(protests == "no", 0, NA)))
nelda <- nelda %>% mutate(inc.fav.cancel = ifelse(inc.fav.cancel == "yes", 1, ifelse(inc.fav.cancel == "no", 0, NA)))
nelda <- nelda %>% mutate(inc.fav.cancel.protests = ifelse(nelda35 == "yes", 1, 0))
nelda <- nelda %>% mutate(opp.fav.cancel = ifelse(nelda32 == "yes", 1, 0))
nelda <- nelda %>% mutate(inc.lose = ifelse(inc.lose == "yes", 1, ifelse(inc.lose == "no", 0, NA)))

nelda <- nelda %>% mutate(inc.termlimit.num = ifelse(types == "Legislative/Parliamentary" & inc.termlimit == "N/A", 0, ifelse(types == "Executive" & inc.termlimit == "yes", 1,                                                                                                                       ifelse(types == "Executive" & inc.termlimit == "no", 0, NA))))
nelda <- nelda %>% mutate(preelection_concerns = ifelse(preelection_concerns == "yes", 1, ifelse(preelection_concerns == "no", 0, NA)))
nelda <- nelda %>% mutate(harass_opp = ifelse(nelda3 == "No", 1, ifelse(harass_opp == "yes", 1, ifelse(harass_opp == "no", 0, NA))))
nelda <- nelda %>% mutate(opp_leaders_prevented = ifelse(nelda3 == "No", "In general", ifelse(opp_leaders_prevented == "yes", "Specifically", ifelse(opp_leaders_prevented == "no", "No", NA))))

nelda <- nelda %>% mutate(econ.crisis.nelda = ifelse(econ.crisis.nelda == "yes", 1, ifelse(econ.crisis.nelda == "no", 0, NA)))
nelda <- nelda %>% mutate(opp.votegain = ifelse(opp.votegain == "yes", 1, ifelse(opp.votegain == "no", 0, NA)))
nelda <- nelda %>% mutate(nelda.protests.fraud = ifelse(protests == 0, 0, ifelse(nelda.protests.fraud == "yes", 1, ifelse(nelda.protests.fraud == "no", 0, NA))))
nelda <- nelda %>% mutate(inc.replaced.prot = ifelse(inc.replaced.prot == "yes", 1, 0))
nelda <- nelda %>% mutate(new.elections.held = ifelse(new.elections.held == "yes", 1, 0))

```

Adding a vdem ID to the other packages

Note: the missing values from NELDA are cases not included in VDEM (e.g. Bahamas, Dominica.)
```{r}
nelda <- nelda %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "gwn", destination = "vdem"))
 #Missing matches here look to be countries below the population threshold: some Carib. islands, Monaco, etc.
lji <- lji %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "cown", 
                                               destination = "vdem"))
massmob <- massmob %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "p4n", 
                                               destination = "vdem")) #Missing here tend to be communist states that no longer exist

nelda <- nelda %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
lji <- lji %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
vdem.data <- vdem.data %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
massmob <- massmob %>% mutate(country_id_year = paste(country_id, year, sep = "_"))

```

```{r}
vdem.recent <- vdem.data %>% filter(year >= 1944)

vdem.recent <- vdem.recent %>% mutate(elcompindex.m = ifelse(v2xel_frefair == 0, 0, v2x_mpi / v2xel_frefair))
vdem.recent$v2elirreg.inv <- vdem.recent$v2elirreg * -1

vdem.recent$reform_positive <- NA
vdem.recent$reform_positive[vdem.recent$v2jureform_ord <= 1] <- 0
vdem.recent$reform_positive[vdem.recent$v2jureform_ord == 2 ] <- 1

vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(v2elintmon.lag = lag(v2elintmon))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(reform_positive.lag = lag(reform_positive))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(v2elirreg.inv.lag = lag(v2elirreg.inv))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(elcompindex.m.1lag = lag(elcompindex.m))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(polyarchy.1lag = lag(v2x_polyarchy))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(libdem.1lag = lag(v2x_libdem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(partipdem.1lag = lag(v2x_partipdem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(delibdem.1lag = lag(v2x_delibdem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(egaldem.1lag = lag(v2x_egaldem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(altinf.1lag = lag(v2x_freexp_altinf))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(api.1lag = lag(v2x_api))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(mpi.1lag = lag(v2x_mpi))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(frassoc.1lag = lag(v2x_frassoc_thick))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(suffr.1lag = lag(v2x_suffr))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(frefair.1lag = lag(v2xel_frefair))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(elecoff.1lag = lag(v2x_elecoff))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(liberal.1lag = lag(v2x_liberal))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jucon.1lag = lag(v2x_jucon))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(partip.1lag = lag(v2x_partip))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(cspart.1lag = lag(v2x_cspart))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(eqdr.1lag = lag(v2xeg_eqdr))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(oppaut.1lag = lag(v2psoppaut))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jureform.1lag = lag(v2jureform))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jupurge.1lag = lag(v2jupurge))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jupoatck.1lag = lag(v2jupoatck))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jupack.1lag = lag(v2jupack))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(juaccnt.1lag = lag(v2juaccnt))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(juhcind.1lag = lag(v2juhcind))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(juncind.1lag = lag(v2juncind))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(mecenefm.1lag = lag(v2mecenefm))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(gdppcln.1lag = lag(e_migdppcln))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(gdpgro.1lag = lag(e_migdpgro))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(inflat.1lag = lag(e_miinflat))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(legcon.1lag = lag(v2xlg_legcon))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(physinteg.1lag = lag(v2x_clphy))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(legcorrupt.lag = lag(v3lgcrrpt))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(exbribe.lag = lag(v2exbribe))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(exembez.lag = lag(v2exembez))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(pubseccorrup.lag = lag(v2excrptps))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(partybarriers.lag = lag(v2psbars))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(natlpartyorg.lag = lag(v2psorgs))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(partybranch.lag = lag(v2psprbrch))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(csrepress.1lag = lag(v2csreprss))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(osorg.1lag = lag(v2cseeorgs))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(civlib.1lag = lag(v2x_clpol))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jucon.1lag = lag(v2x_jucon))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(legcon.1lag = lag(v2xlg_legcon))

vdem.recent <- vdem.recent %>% group_by(COWcode) %>% mutate(opposition.oversight.lag = lag(v2lgoppart))

##Leading variables for protest after election years
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(prodem.protest.1lead = lead(v2cademmob))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(proauth.protest.1lead = lead(v2caautmob))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(protest.scale.1lead = lead(v2cagenmob))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(concen.protest.1lead = lead(v2caconmob))

```
Subtracting election integrity variable from additive polyarchy index and judicial constraints from liberal component index
```{r}
vdem.recent <- vdem.recent %>% mutate(polyarchy_no_ei = v2x_api - (.25*v2xel_frefair))

vdem.recent <- vdem.recent %>% mutate(liberal_no_jucon = (v2xcl_rol + v2xlg_legcon)/2)
```

Imporant variables to include but not lag: v2eltvrexo, v2elvotlrg, v2elvotsml, v2ellostlg, v2ellostsl, v2ellostsm, v2ellostss, v2ellovtlg, v2ellovtsm

```{r}
lji <- lji %>% group_by(ccode) %>% mutate(lji.lag = lag(LJI))
```

Cleaning nelda date variable and massmob date
```{r}
nelda <- nelda %>% mutate(mmdd = ifelse(mmdd < 1000, paste0(0, mmdd), mmdd))

nelda <- nelda %>% mutate(election_mdy = paste0(mmdd, year))
nelda <- nelda %>% mutate(election_mdy = mdy(election_mdy))

massmob <- massmob %>% mutate(startday = ifelse(startday < 10, paste0(0, startday), startday))
massmob <- massmob %>% mutate(startmonth = ifelse(startmonth < 10, paste0(0, startmonth), startmonth))
massmob <- massmob %>% mutate(protest_mdy = ifelse(is.na(startday) == T | is.na(startmonth) == T, NA,  paste0(startmonth, startday, year)))

massmob <- massmob %>% mutate(protest_mdy = mdy(protest_mdy))
massmob <- massmob %>% filter(is.na(protest_mdy) == F)
```

```{r}
lji <- lji %>% filter(is.na(country_id) == FALSE)
nelda <- nelda %>% filter(is.na(country_id) == FALSE)
massmob <- massmob %>% filter(is.na(country_id) == FALSE)

#Lagging econ crisis

nelda <- nelda %>% group_by(country_id) %>% mutate(econ.crisis.nelda.1lag = lag(econ.crisis.nelda))

#Merging
nelda <- nelda %>% merge(lji %>% dplyr::select(LJI, lji.lag, country_id_year), by = "country_id_year")

nelda <- nelda %>% merge(vdem.recent %>% dplyr::select(v2elintmon.lag, v2elintmon, v2eltvrexo, v2elvotlrg, v2elvotsml, v2ellostlg, v2ellostsl, v2ellostsm, v2ellostss, v2ellovtlg, v2ellovtsm, v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, v2x_egaldem, v2x_freexp_altinf, v2x_api, v2x_mpi, v2x_frassoc_thick, v2x_suffr, v2xel_frefair, v2x_elecoff, v2x_liberal, v2x_jucon, v2x_partip, v2x_cspart, v2xeg_eqdr, v2psoppaut, v2jureform, v2jupurge, v2jupoatck, v2jupack, v2juaccnt, v2juhcind, v2juncind, v2mecenefm, e_migdppcln, e_migdpgro, e_miinflat, v2xlg_legcon, legcon.1lag, polyarchy.1lag, libdem.1lag, partipdem.1lag, 
                                                  delibdem.1lag,
                                                  elcompindex.m.1lag,
                                                  opposition.oversight.lag,
                                                  v2elirreg.inv,
                                                  reform_positive.lag,
                                                  e_mipopula,
                                                  e_miurbani,
                                                  e_miurbpop,
                                                  e_wb_pop, #Population
                                                  v2xps_party, #Party institutionalization
                                                  #v2regoppgroupssize, #Opp group size (categorical)
                                                  v2pscohesv, #Legislative party cohesion
                                                  v2exl_legitratio, 
                                                  v2exl_legitperf,
                                                  v2exl_legitlead,
                                                  v2exl_legitideol,
                                                  v2xnp_regcorr,
                                                  v2xeg_eqprotec,
                                                  v2xeg_eqdr,
                                                  v2xeg_eqaccess,
                                                  v3lgcrrpt,
                                                  legcorrupt.lag,
                                                  exbribe.lag,
                                                  v2exbribe,
                                                  exembez.lag,
                                                  v2exembez,
                                                  pubseccorrup.lag,
                                                  v2excrptps,
                                                  partybarriers.lag,
                                                  v2psbars,
                                                  natlpartyorg.lag,
                                                  v2psorgs,
                                                  partybranch.lag,
                                                  v2psprbrch,
                                                  v2stfisccap_ord,
                                                  v2stfisccap,
                                                  v2pepwrses, #Power distribution by wealth
                                                  v2pepwrsoc, #Power distribution by social group
                                                  v2csreprss,
                                                  csrepress.1lag,
                                                  osorg.1lag,
                                                  v2cseeorgs,
                                                  v2x_clpol,
                                                  civlib.1lag,
                                                  jucon.1lag,
                                                  legcon.1lag,
                                                  v2ex_elechog,
egaldem.1lag ,
altinf.1lag,
api.1lag,
mpi.1lag,
frassoc.1lag,
suffr.1lag,
frefair.1lag,
elecoff.1lag,
liberal.1lag,
jucon.1lag,
partip.1lag,
cspart.1lag,
eqdr.1lag,
oppaut.1lag,
jureform.1lag,
jupurge.1lag,
jupoatck.1lag,
jupack.1lag,
juaccnt.1lag,
juhcind.1lag,
juncind.1lag,
mecenefm.1lag,
gdppcln.1lag,
gdpgro.1lag,
inflat.1lag,
v2elvotbuy,
v2elirreg,
v2elintim, v2elfrfair,
physinteg.1lag, v2x_clphy,
polyarchy_no_ei,
liberal_no_jucon, v2elaccept,
v2cademmob, #pro-democracy mass mob
v2caautmob, #pro-autoc mass mob
v2cagenmob, #scale of mobilization
v2caconmob, #concentration of mobilization in the capital
v2x_regime,
prodem.protest.1lead,
proauth.protest.1lead,
protest.scale.1lead,
concen.protest.1lead,
country_id_year), by = "country_id_year")

nelda <- nelda %>% mutate(winner.margin = ifelse(types == "Executive", v2elvotlrg/100, v2ellovtlg/100))
nelda <- nelda %>% mutate(presidential = ifelse(types == "Executive", 1, 0))

#

nelda <- nelda %>% mutate(weakstate = ifelse(is.na(v2stfisccap_ord) == T, NA, ifelse(v2stfisccap_ord == 0 | v2stfisccap_ord == 1, 1, 0)))
nelda <- nelda %>% mutate(rentierstate = ifelse(is.na(v2stfisccap_ord) == T, NA, ifelse(v2stfisccap_ord == 2, 1, 0)))
nelda <- nelda %>% mutate(taxstate = ifelse(is.na(v2stfisccap_ord) == T, NA, ifelse(v2stfisccap_ord == 3 | v2stfisccap_ord == 4, 1, 0)))

```

```{r}
rm(vdem.data)
rm(vdem.recent)
```

## World Bank urbanization data

```{r}
urban <- read.csv(here::here("Data", "urbanization_worldbank.csv"))
test <- urban %>%
  pivot_longer(cols = starts_with("X"),
               names_to = "year")
test <- test %>% mutate(year = gsub("X", "", year))

#Next is line up country codes in test and nelda 2, then merge

#nelda <- nelda %>% rename(year = year.x)

library(countrycode)

test <- test %>% mutate(stateid = countrycode(sourcevar = Country.Code, origin = 'iso3c', destination = 'gwc'))
test <- test %>% rename(urban.pct = value)
test <- test %>% dplyr::select(year, urban.pct, stateid)
test <- test %>% mutate(year = as.numeric(year))

nelda <- nelda %>% left_join(test, by = c("stateid", "year"))

```

Starting with nelda from above, and merging with MassMob. This should yield a dataset with nelda elections, vdem data, and massmob protest data

```{r}
nelda <- nelda %>% rename(country_id = country_id.x)
nelda_selection <- nelda %>% dplyr::select(electionid, country_id, election_mdy)

  election_id_temp <- nelda_selection$electionid[1]
  country_temp <- nelda_selection$country_id[1]
  election_date_temp <- nelda_selection$election_mdy[1] # Get the election date
  election_row <- nelda_selection[1,]  # Get the election at hand (all Nelda data)
  massmob_temp <- massmob %>% filter(country_id == country_temp) #Get all protests from the country at hand
  massmob_temp <- massmob_temp %>% mutate(days_to_election = election_date_temp - protest_mdy) # Get the days between each protest and the election
  massmob_temp <- massmob_temp %>% filter(days_to_election <= 180 & days_to_election >= -90) #Remove elections outside the window
  massmob_temp <- massmob_temp %>% mutate(electionid = election_id_temp)
  election_row <- election_row %>% left_join(massmob_temp, by = "electionid") #Merge Nelda data with remaining protests

nelda_massmob <- election_row #Start the dataset

for (i in 201:nrow(nelda)) {  #
  election_id_temp <- nelda_selection$electionid[i]
  country_temp <- nelda_selection$country_id[i]
  election_date_temp <- nelda_selection$election_mdy[i]
  election_row <- nelda_selection[i,]
  massmob_temp <- massmob %>% filter(country_id == country_temp)
  massmob_temp <- massmob_temp %>% mutate(days_to_election = election_date_temp - protest_mdy)
  massmob_temp <- massmob_temp %>% filter(days_to_election <= 180 & days_to_election >= -90)

  massmob_temp <- massmob_temp %>% mutate(electionid = election_id_temp)
  election_row <- election_row %>% left_join(massmob_temp, by = "electionid")
  nelda_massmob <- bind_rows(nelda_massmob, election_row)
}

##This seems to have worked. Next step should be quality control, as well as making sure all relevant nelda changes happen before this chunk (i.e. the urbanization data below.)
```

Aggregating participant variables
```{r}

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(participants_category == "", NA, 
                                            ifelse(participants_category == "50-99", 1, 
                                            ifelse(participants_category == "100-999", 2,
                                            ifelse(participants_category == "1000-1999", 3,
                                            ifelse(participants_category == "2000-4999", 4,
                                            ifelse(participants_category == "5000-10000", 5,
                                                   6)))))))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100s", 2, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000s", 4, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">50", 1, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "50", 1, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000", 3, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "50+", 1, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100", 2, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "10000", 5, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "2000", 4, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "200", 2, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "300", 2, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "500", 2, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "3000", 4, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100+", 2, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "5000", 5, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "20000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "10000s", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "150", 2, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "400", 2, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "hundreds", 2, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "30000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "15000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1500", 3, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 50", 1, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "4000", 4, participants_category_ordered))  ## Stopping here; all other categories have fewer than 20 observations. Beginning second stopping rule:

## Second rule: collect all observations greater than 10,000

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "hundreds of thousands", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "<20000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 40000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 42000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">35000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">40000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">50000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "110000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "14059", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "16000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "18000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "24000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "250000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "30000+", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "35000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "36000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "5000-50000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "55000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "650000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 10000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "13000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "500000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">15000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">20000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "14000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "400000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "Tens of thousands", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">100000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "150000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "20000+", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "60000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "90000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "10000+", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100000s", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "200000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "80000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">10000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "70000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "80000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "40000", 6, participants_category_ordered))

nelda_massmob <- nelda_massmob %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "25000", 6, participants_category_ordered))

#nelda_massmob <- nelda_massmob %>% group_by(electionid) %>% 
```

Remove post-election protests unrelated to election

```{r}
nelda_massmob <- nelda_massmob %>% rename(notes_massmob = notes)

nelda_massmob <- nelda_massmob %>% filter(is.na(protest_mdy) == T | protest_mdy < election_mdy | (protest_mdy >= election_mdy & str_detect(notes_massmob, regex(pattern = "election*|vot*|ballot*", ignore_case = T))))
```

Getting grouped summarize for number and size
```{r}
nelda_massmob.pre <- nelda_massmob %>% filter(as.numeric(days_to_election) > 0) #protest_mdy < election_mdy
nelda_massmob.post <- nelda_massmob %>% filter(as.numeric(days_to_election) <= 0)

count_data_mm_pre <- nelda_massmob.pre %>% group_by(electionid) %>% summarize(n.events.massmob.pre = sum(protestnumber))
count_data_mm_post <-nelda_massmob.post %>% group_by(electionid) %>% summarize(n.events.massmob.post = sum(protestnumber))

size_data_mm_pre <- nelda_massmob.pre %>% group_by(electionid) %>% summarize(median.participants.mm.pre = median(participants_category_ordered, na.rm = T))
size_data_mm_post <- nelda_massmob.post %>% group_by(electionid) %>% summarize(median.participants.mm.post = median(participants_category_ordered, na.rm = T))

```

## Merge data from nelda_massmob back to edited nelda

```{r}
##Problem here: Do not merge back into nelda_massmob. Merge back into a version of NELDA without massmob. (But still include V-Dem variables etc, and exclude the necessary time periods and cases.)

##Combined post data together, then pre data, then both
mm_data_post_comb <- left_join(count_data_mm_post, size_data_mm_post, by = "electionid")
mm_data_pre_comb <- left_join(count_data_mm_pre, size_data_mm_pre, by = "electionid")

mm_data_comb <- left_join(mm_data_pre_comb, mm_data_post_comb, by = "electionid")

#mm_data_comb <- mm_data_comb %>% rename(electionid = NeldaID)

## Merge back to Nelda classic
nelda_complete <- nelda %>% left_join(mm_data_comb, by = "electionid")

#nelda_massmob <- nelda_massmob %>% left_join(mm_data_comb, by = "electionid")  ## Double check to make sure this joining works

## Code NAs for number of protests as 0s (these reflect election years recorded in NELDA where there were no protests recorded by MM)

nelda_complete <- nelda_complete %>% mutate(n.events.massmob.pre = ifelse(is.na(n.events.massmob.pre)==T, 0, n.events.massmob.pre))
nelda_complete <- nelda_complete %>% mutate(n.events.massmob.post = ifelse(is.na(n.events.massmob.post)==T, 0, n.events.massmob.post))

nelda_complete <- nelda_complete %>% mutate(median.participants.mm.pre = ifelse(is.na(median.participants.mm.pre)==T & n.events.massmob.pre == 0, 0, median.participants.mm.pre))
nelda_complete <- nelda_complete %>% mutate(median.participants.mm.post = ifelse(is.na(median.participants.mm.post)==T & n.events.massmob.post == 0, 0, median.participants.mm.pre))

## Rounding participant median numbers down (removes half-steps)

nelda_complete <- nelda_complete %>% mutate(median.participants.mm.post = floor(median.participants.mm.post))
nelda_complete <- nelda_complete %>% mutate(median.participants.mm.pre = floor(median.participants.mm.pre))

```

## Subset to limit to electoral auth and dem

```{r}
nelda_complete <- nelda_complete %>% filter(v2x_regime == 1 | v2x_regime == 2) #Note that coding scheme for electoral democracies allows for "some degree of fraud and irregularities but these did not in the end affect the outcome" (v2elfrfair, which is a component of v2x_regime) 
```

## Merge with winner share updated

```{r}
winner.shares <- read.csv(here::here("Data", "amended_vote_shares.csv"))
winner.shares <- winner.shares %>% rename(electionid = NeldaID)
nelda_complete <- nelda_complete %>% left_join(winner.shares, by = "electionid")

```

## Writing

```{r}
write.csv(nelda_complete, here::here("Data", "nelda_complete.csv"))
```

### 2022-11-23 Output/manipulation_protest_draft_v3.Rmd
Updates on draft; beginning results section. Winners vote share section finished.

@@
 
-nelda3.sub <- read.csv(here::here("Data", "nelda3_sub.csv"))
+nelda_massmob <- read.csv(here::here("Data", "nelda_complete.csv"))
 
-nelda3.sub <- nelda3.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
-nelda3.sub <- nelda3.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))
+nelda_massmob <- nelda_massmob %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
+nelda_massmob <- nelda_massmob %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))
 
@@
 
-nelda3.sub <- nelda3.sub %>% rename(boycott = nelda14)
-nelda3.sub <- nelda3.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
+nelda_massmob <- nelda_massmob %>% rename(boycott = nelda14)
+nelda_massmob <- nelda_massmob %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                  NA)))
@@
 
-nelda3.sub <- nelda3.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
-nelda3.sub <- nelda3.sub %>% mutate(econ.crisis = as.factor(econ.crisis))
+nelda_massmob <- nelda_massmob %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
+nelda_massmob <- nelda_massmob %>% mutate(econ.crisis = as.factor(econ.crisis))
 
-nelda3.sub <- nelda3.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
-nelda3.sub <- nelda3.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))
+nelda_massmob <- nelda_massmob %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
+nelda_massmob <- nelda_massmob %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))
 
-nelda3.sub <- nelda3.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))
+nelda_massmob <- nelda_massmob %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))
 
 ##Removing cases before 1990, when ECAV starts
-nelda3.sub <- nelda3.sub %>% filter(year >= 1990)
+nelda_massmob <- nelda_massmob %>% filter(year >= 1990)
 
 ## Removing cases where inc. loses
-nelda3.sub <- nelda3.sub %>% filter(inc.lose == 0)
+nelda_massmob <- nelda_massmob %>% filter(inc.lose == 0)
 
 ##Intmon as a factor
-nelda3.sub <- nelda3.sub %>% mutate(v2elintmon = as.factor(v2elintmon))
+nelda_massmob <- nelda_massmob %>% mutate(v2elintmon = as.factor(v2elintmon))
 
-nelda3.sub <- nelda3.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))
+nelda_massmob <- nelda_massmob %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))
 
@@
 
-nelda3.sub <- nelda3.sub %>% mutate(physintlag.sq = physinteg.1lag^2)
+nelda_massmob <- nelda_massmob %>% mutate(physintlag.sq = physinteg.1lag^2)
 
@@
 
-nelda3.sub <- nelda3.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))
+nelda_massmob <- nelda_massmob %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))
 
@@
 
-nelda3.sub <- nelda3.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))
+nelda_massmob <- nelda_massmob %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))
 
@@
 
-nelda3.sub <- nelda3.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | elections.cancel.prot, 1, 0))
+nelda_massmob <- nelda_massmob %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | elections.cancel.prot, 1, 0))
 
@@
 
-protest.cases <- nelda3.sub %>% filter(any.ecav.protest == 1)
+#protest.cases <- nelda_massmob %>% filter(any.ecav.protest == 1)
 
@@
 
-nelda3.sub <- nelda3.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))
+nelda_massmob <- nelda_massmob %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))
 
 ## Changing participants from numeric category to text
-nelda3.sub <- nelda3.sub %>% mutate(median.participants = round(median.participants))
-nelda3.sub <- nelda3.sub %>% mutate(particpants_categorical = ifelse(is.na(median.participants)==T, NA, ifelse(median.participants == 1, "Less than 10",
-                     ifelse(2, "10 to 99", ifelse(3, "100 to 999",
-                                                  ifelse(4, "1000 to 9,999", ifelse(5, "Greater than 10,000", NA)))))))
-nelda3.sub <- nelda3.sub %>% mutate(particpants_categorical = as.factor(particpants_categorical))
+#nelda_massmob <- nelda_massmob %>% mutate(median.participants = round(median.participants))
+#nelda_massmob <- nelda_massmob %>% mutate(particpants_categorical = ifelse(is.na(median.participants)==T, NA, ifelse(median.participants == 1, "Less than 10",
+#                     ifelse(2, "10 to 99", ifelse(3, "100 to 999",
+#                                                  ifelse(4, "1000 to 9,999", ifelse(5, "Greater than 10,000", NA)))))))
+#nelda_massmob <- nelda_massmob %>% mutate(particpants_categorical = as.factor(particpants_categorical))
 
@@
 
-nelda3.sub <- nelda3.sub %>% mutate(median.participants.mm.post = as.factor(median.participants.mm.post))
+nelda_massmob <- nelda_massmob %>% mutate(median.participants.mm.post = as.factor(median.participants.mm.post))
 
@@
 
-nelda3.sub <- nelda3.sub %>% mutate(v2elvotbuy_inv = -1*v2elvotbuy)
+nelda_massmob <- nelda_massmob %>% mutate(v2elvotbuy_inv = -1*v2elvotbuy)
 
@@
 
-These hypotheses are tested using data from three datasets: NELDA [@hyde2012], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence (ECAV) dataset [@Daxecker2019].
+These hypotheses are tested using data from three datasets: NELDA [@hyde2012], V-Dem version 11.1 [@vdem11.1], and the Mass Mobilization Project dataset [@massmob].
 The unit of observation for this study is an election period as recorded in NELDA.
@@
 I further limit the sample to regimes coded as electoral autocracies or electoral democracies in V-Dem.
-This excludes closed autocracies (where no national elections are held) and liberal democracies (where electoral integrity and rule of law are both high).
-Finally, the date range is limited by the availability of data on protest from ECAV, as discussed below, which runs from 1990 to 2012.
-The resulting dataset includes 647 election-period observations across 109 countries.
+This excludes closed autocracies (where no national elections are held) and liberal democracies (where electoral integrity and rule of law are both high).\footnote{The specific variable used is 'v2x_regime' from V-Dem.
+This variable codes an electoral democracy, in part, if the country scores above a 2 on the measure of free and fair elections ('v2elfrfair_osp').
+This still permits election fraud to be detected in the electoral democracies; v2elfrfair can take on a value of 3 when "deficiencies and some degree of fraud and irregularities" occur.} Finally, the beginning of date range is limited by the availability of data on protest from the Mass Mobilization dataset, as discussed below, which begins in 1990.
+The inclusion of a control variable from the Latent Judicial Independence dataset [@Linzer2015] limits the end of the time range to 2012.
+The resulting dataset includes 647 total election-period observations across 109 countries.
 
-ECAV, from which study's dependent variable is drawn, records protests and other contentious actions from six months before to three months after an election.
-For this study, I include protests only, excluding acts like shootings, bombings, and coups.
-ECAV data is in event-day format.
-To capture the *number of protests*, I take the sum number of events recorded in ECAV from election day to the end of the dataset's coverage period ninety days later.
-The result is a count variable ranging from zero to 90.\footnote{That the max value for the quantity of protests matches the number of days in the recording window is coincidental.} The large majority of elections experience no post-election protest: in 460 of the 647 elections, no protests are recorded.
+The Mass Mobilization dataset, from which study's dependent variables are drawn, attempts to record all anti-government protests drawing 50 or more participants.
+The data is in event-day format.
+I divide this data into a pre-election period (180 days prior to the election) to capture control variables for the number and size of pre-election protest, and a post-election period (90 days after the election) for the dependent variables.
+For the post-election period, I include only those protests where the summary notes for the protest include some form of any of the words "election," "vote," or "ballot." This excludes post-election protests that do not target the election itself.
+
+Two dependent variables are derived from the resulting post-election data.
+First, I collect the *median size* of post-election protest.
+The Mass Moblization data records protest size using six categories.\footnote{This are: 50-99 participants, 100-999, 1,000-1,999, 2,000-4,999, 5,000-9,999, and greater than 10,000 participants.} I then take the median size of the full set of post-election protest per election period; the result is an ordered categorical variable.
+Second, I capture the post-election *number of protests* by taking the sum number of events recorded in the filtered post-election data.
+The result is a count variable ranging from zero to 238.
+The large majority of elections experience no post-election protest: in 575 of the 647 elections, no protests are recorded.
 Figure 1 reports the distribution of protest counts across the dataset, as well as winners' vote-shares.
@@
 
-ggplot(nelda3.sub, aes(x=n.events.massmob.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()
+ggplot(nelda_massmob, aes(x=n.events.massmob.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()
 
-ggplot(nelda3.sub, aes(x=winner.share.upd)) + geom_histogram(bins=100) + xlab("Incumbent vote-share") + theme_bw()
+ggplot(nelda_massmob, aes(x=winner.share.upd)) + geom_histogram(bins=100) + xlab("Incumbent vote-share") + theme_bw()
 ```
@@
 
-p.success <- ggplot(nelda3.sub, aes(x=as.numeric(median.participants.mm.post), fill = as.factor(protest.success), color=as.factor(protest.success)))+
+p.success <- ggplot(nelda_massmob, aes(x=as.numeric(median.participants.mm.post), fill = as.factor(protest.success), color=as.factor(protest.success)))+
   geom_histogram(position="identity", alpha=0.5, binwidth = 1) 
@@
 ```{r, include=FALSE}
-mu <- plyr::ddply(nelda3.sub, "protest.success", summarise, grp.mean=mean(n.events.massmob.post))
+mu <- plyr::ddply(nelda_massmob, "protest.success", summarise, grp.mean=mean(n.events.massmob.post))
 head(mu)
 
-p.success <- ggplot(nelda3.sub, aes(x=n.events.massmob.post, fill = as.factor(protest.success), color=as.factor(protest.success)))+
+p.success <- ggplot(nelda_massmob, aes(x=n.events.massmob.post, fill = as.factor(protest.success), color=as.factor(protest.success)))+
   geom_histogram(position="identity", alpha=0.5, binwidth = 1) + 
@@
 ```{r}
-incl.vars <- nelda3.sub %>% dplyr::select(jucon.1lag, legcon.1lag, altinf.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, harass_opp, any.ecav.protest, median.participants.mm.pre, median.participants.mm.post)
+incl.vars <- nelda_massmob %>% dplyr::select(jucon.1lag, legcon.1lag, altinf.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, harass_opp,  median.participants.mm.pre, median.participants.mm.post)
 #incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))
@@
 This distribution is in line with the theoretical proposition that two data-generating processes are at work: the factors that lead to the occurrence of protest may not be the same as those that sustain or intensify it.
-Consequently, I model the results in two ways.
-The first approach uses logit models to identify the factors that make *any protest* more likely to occur---this is a binary variable that is coded as 1 if any protest occurred after the election.
-Since election-year observations are nested within countries here, I use multilevel models with random intercepts by country in.
-In the second analysis, I use negative binomial regressions to model the count of protests given that protest has begun.
-That is, these latter models exclude cases where no protest occurred.
-For ease of interpretation, these processes are modeled and depicted separately in the main text, while hurdle regression models (which use the full dataset in one stage) are presented in the appendix, with similar results.
+Consequently, I model the results using negative binomial hurdle regressions.
+Hurdle models are two-part regressions in which the variation in the dependent variable is assumed to follow two distributions: the probability of obtaining a zero or non-zero outcome is modeled as binomial logic function, while the distribution of the non-zero count variable is modeled (in this case) using the negative binomial function.
+The logistic portion of the model identifies the factors that make *any protest* more likely to occur; the negative binomial regressions to model the count of protests given that protest has begun.
 
@@
 This measure of election integrity should thus pose a strong test of the protest-oriented model.
-Finally, the winning party's *vote-share* is also taken from the V-Dem dataset.
+Finally, the winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia and Psephos.}
 
@@
 This variable is included since incumbents running up against constitutional term limits can pose major challenges for the unity of the regime, possibly leading to fractured manipulation efforts and post-election protest [@hale_patronal_2014].
-Finally, I include dummy variables that provide a categorical evalution of state capacity.
+Finally, I include dummy variables that provide a categorical evaluation of state capacity.
 These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap).
@@
 Figure 2 shows summary statistics for all of these variables.
+
+# Results
+
+The results of these models fail to support the conventional wisdom that more severe electoral manipulation is associated with an increased risk of anti-government protest, under ordinary circumstances.
+Beginning with the models of number of protest events, Model 1\textemdash which includes no interaction terms\textemdash shows no significant relationship between fraud and protest in either the binary or count portions of the hurdle regression.
+Likewise, Model 2 shows no significant interaction effect for fraud and winning vote-share on quantity of protest.
+Figure \@ref(fig:vote-share-fig) shows the marginal effect of election fraud at varying levels of winning vote-share, holding other variables constant at their means (continuous variables) or modes (categorical variables).
+The high and low levels of fraud in the plot represent one standard deviation above and below the mean, respectively.
+As the figure shows, the predicted number of post-election protests is higher in the high fraud condition when winning margin is low, as predicted, but this relationship is not statistically significant.
+Moreover, there is no winning margin where the difference in predicted number of protests is significant.
+
+In the same way, there is no statistically significant relationship between the level of fraud, the winning margin of victory, and the median size of post-election protest.
+This is true both for Model 6, where fraud is a standalone coefficient, and for the interaction term in Model 7.
+Figure \@ref(fig:vote-share-size-fig) shows no significant difference between the low-fraud and high-fraud conditions in any of the seven protest size categories, at any level of winning vote-share.
+Neither the models of protest size nor protest number support the strategic logic of protest advanced in some models, in which fraud sends a signal of weakness especially in the context of a weak ruling-party performance.
+
+Turning to models of protest quantity presented in Table 3, there is no statistically significant relationship between fraud and protest in any of the models.
+Figure 4, using results from Model 9, presents the marginal effects for fraud conditional on the incumbent's vote-share.
+As the figure shows, there is no statistically significant difference in the expected number of protest following a relatively clean election compared to a relatively fraudulent one, for any margin of incumbent victory.
+Taking the results from Tables 2 and 3 together, there is support for Hypothesis 1 (an unconditional increase in the risk of any protest as fraud increases), as well as strong support for Hypothesis 2 (an increase in protest risk in fraudulent, low-margin elections) and Hypothesis 4 (no relationship between fraud and protest quantity).
+Hypothesis 3, on the connection between fraud, grievance, and protest, is not supported.
+
+We do observe support for the more structural explanations of more extensive protest.
+The coefficient for the *legislative constraint* variable, which measures the size and oppositional character of the legislative opposition, is consistently positive and  statistically significant.
+In Models 8 and 9 (those without a grievance interaction), the coefficient on *national party organization* is also positive and significant.
+The implication is that a larger, more institutionalized opposition party or coalition prolongs protest, in line with the 'test of strength' model.
+Similarly, the *incumbent term limit* variable is positive and consistently significant.
+This supports Hale's [-@hale_patronal_2014] argument that uncertainty around succession in non-democracies makes mobilization more intense as competing power centers emerge.
+
+```{r, include=FALSE}
+model.novoteshare.size <-  polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp +  pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag  + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+summary(model.novoteshare.size)
+
+model.size.base <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp +  pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+summary(model.size.base)
+
+model.size.base2 <- polr(median.participants.mm.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+summary(model.size.base2)
+
+model.size.griev <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+summary(model.size.griev)
+
+model.size.griev2 <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + + harass_opp + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda_massmob)
+summary(model.size.griev2)
+
+model.size.griev3 <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + altinf.1lag + harass_opp +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda_massmob)
+summary(model.size.griev3)
+```
+
+```{r include=FALSE}
+##Winning margin
+dat <- ggpredict(model.size.base2, terms = c("winner.share.upd", "v2elirreg.inv [-.47, 1.63]"))
+dat.sub <- dat %>% dplyr::filter(response.level == "0")
+p.win.zero <- plot(dat.sub)
+p.win.zero <- p.win.zero  + theme_bw() + xlab(NULL) + ylab("Predicted probability") + labs(title = "Predicted size of post-election protest", color = "Election fraud")
+
+dat.sub <- dat %>% dplyr::filter(response.level == "1" | response.level == "2" | response.level == "3")
+p.win.low <- plot(dat.sub)
+p.win.low <- p.win.low  + theme_bw() + xlab(NULL) + ylab("Predicted probability") + labs(title = NULL, color = "Election fraud")
+
+dat.sub <- dat %>% dplyr::filter(response.level == "4" | response.level == "5" | response.level == "6")
+p.win.high <- plot(dat.sub)
+p.win.high <- p.win.high  + theme_bw() + xlab("Winning vote-share") + ylab("Predicted probability") + labs(title = NULL, color = "Election fraud")
+
+#dat_max <- ggpredict(model.size.base2, terms = c("winner.share.upd", "v2elirreg.inv [-2.3, 3.2]"))
+#dat.sub <- dat_max %>% dplyr::filter(response.level == "0")
+#plot(dat.sub)
+
+#dat.sub <- dat_max %>% dplyr::filter(response.level == "1" | response.level == "2" | response.level == "3")
+#plot(dat.sub) + coord_cartesian(ylim = c(0, .5))
+#dat.sub <- dat_max %>% dplyr::filter(response.level == "4" | response.level == "5" | response.level == "6")
+#plot(dat.sub) + coord_cartesian(ylim = c(0, .5))
+```
+
+```{r vote-share-size-fig, fig.cap="Fraud, winning vote-share, and median protest size", fig.height=7}
+plot.combined.size <- cowplot::plot_grid(p.win.zero, p.win.low, p.win.high, ncol=1)
+plot.combined.size
+```
+
+```{r include=FALSE}
+## Econ grievance
+dat <- ggpredict(model.size.griev2, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv"))
+dat.sub <- dat %>% dplyr::filter(response.level == "0" | response.level == "1" | response.level == "2") 
+plot(dat.sub)
+
+dat.sub <- dat %>% dplyr::filter(response.level == "3" | response.level == "4" | response.level == "5" | response.level == "6")
+plot(dat.sub)
+
+dat_max <- ggpredict(model.size.griev2, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-2.3, 3.2]"))
+dat.sub <- dat_max %>% dplyr::filter(response.level == "0" | response.level == "1" | response.level == "2") 
+plot(dat.sub)
+
+dat.sub <- dat_max %>% dplyr::filter(response.level == "3" | response.level == "4" | response.level == "5" | response.level == "6")
+plot(dat.sub)
+```
+
+```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
+model.n.novoteshare.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +     + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+summary(model.n.novoteshare.hurd.mm)
+
+model.n.base.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+summary(model.n.base.hurd.mm)
+
+model.n.base2.hurd.mm <- hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre  + v2elvotbuy_inv|  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+summary(model.n.base2.hurd.mm)
+
+model.n.griev.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv| jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, dist = "negbin", data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+summary(model.n.griev.hurd.mm)
+
+model.n.griev2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv | jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, data = nelda_massmob, dist = "negbin")
+summary(model.n.griev2.hurd.mm)
+
+model.n.griev3.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv | jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, data = nelda_massmob, dist = "negbin")
+summary(model.n.griev3.hurd.mm)
+```
+
+```{r include=FALSE}
+dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd", "v2elirreg.inv [-.45, 1.64]")) 
+p.mm.margin <- plot(dat)
+p.mm.margin <- p.mm.margin + coord_cartesian(ylim = c(0, 70))
+
+#dat_max <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd", "v2elirreg.inv [-2.3, 3.2]")) 
+#p.mm.margin <- plot(dat_max)
+#p.mm.margin + coord_cartesian(ylim = c(0, 150))
+```
+
+```{r vote-share-fig, fig.cap="Fraud, winning vote-share, and number of protest events"}
+p.mm.margin + theme_bw() + xlab("Winning vote-share") + ylab("Predicted number") + labs(title = "Predicted number of post-election protest", color = "Election fraud")  
+```
+
+```{r include=FALSE}
+dat <- ggpredict(model.n.griev.hurd.mm, terms = c("gdpgro.1lag", "v2elirreg.inv [-.45, 1.64]")) 
+p.mm.margin <- plot(dat)
+p.mm.margin + coord_cartesian(ylim = c(0, 250))
+
+```
+
+```{r include=FALSE}
+dat <- ggpredict(model.n.griev2.hurd.mm, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-.45, 1.64]")) 
+p.mm.margin <- plot(dat)
+p.mm.margin
+
+```
+
+```{r include=FALSE}
+dat <- ggpredict(model.n.griev3.hurd.mm, terms = c("pubseccorrup.lag", "v2elirreg.inv [-.45, 1.64]")) 
+p.mm.margin <- plot(dat)
+p.mm.margin
+
+```
+```{r zero-model-table, echo=FALSE}
+
+models <- list()
+models[['Model 1']] <- model.n.base.hurd.mm
+models[['Model 2']] <- model.n.base2.hurd.mm
+models[['Model 3']] <- model.n.griev.hurd.mm
+models[['Model 4']] <- model.n.griev2.hurd.mm
+models[['Model 5']] <- model.n.griev3.hurd.mm
+
+modelsummary::modelsummary(models, coef_omit = "count", gof_omit = "R2", stars = T)
+
+```
+
+```{r zero-model-table-count, echo=FALSE}
+modelsummary::modelsummary(models, coef_omit = "zero", gof_omit = "R2", stars = T)
+
+```
+
+```{r size-model-table, echo=FALSE}
+
+models <- list()
+models[['Model 6']] <- model.size.base
+models[['Model 7']] <- model.size.base2
+models[['Model 8']] <- model.size.griev
+models[['Model 9']] <- model.size.griev2
+models[['Model 10']] <- model.size.griev3
+
+modelsummary::modelsummary(models, longtable = T)
+
+```
+
+\newpage
+
+## References

### 2022-11-30 Output/appendix2.Rmd
Edits to update v3 draft to MM instead of ECAV

@@
 models <- list()
-models[['Model 1']] <- model.n.base.hurd
-models[['Model 2']] <- model.n.base2.hurd
-models[['Model 3']] <- model.n.griev.hurd
-models[['Model 4']] <- model.n.griev2.hurd
-models[['Model 5']] <- model.n.griev3.hurd
+models[['Model 1']] <- model.n.base.hurd.mm
+models[['Model 2']] <- model.n.base2.hurd.mm
+models[['Model 3']] <- model.n.griev.hurd.mm
+models[['Model 4']] <- model.n.griev2.hurd.mm
+models[['Model 5']] <- model.n.griev3.hurd.mm
 

### 2022-11-30 Output/manipulation_protest_draft_v3.Rmd
Edits to update v3 draft to MM instead of ECAV

@@
 
-Using cross-national data on 647 elections from 1990 to 2012 where incumbents claimed victory, adapted from the recently-released Electoral Contention and Violence (ECAV) dataset [@Daxecker2019], this paper finds support for the hypothesis that more fraudulent elections are associated with a higher risk that any post-election protest event occurs.
+Using cross-national data on 647 elections from 1990 to 2012 where incumbents claimed victory, combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Mass Mobilization dataset [@massmob], this paper finds support for the hypothesis that more fraudulent elections are associated with a higher risk that any post-election protest event occurs.
 This is in keeping with the formal-modeling literature, which predicts that overreach in the form of election fraud indicates government weakness, making opposition groups confident of winning a conflict in the streets.
@@
 
-```{r}
+```{r, include=F}
 incl.vars <- nelda_massmob %>% dplyr::select(jucon.1lag, legcon.1lag, altinf.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, harass_opp,  median.participants.mm.pre, median.participants.mm.post)
 #incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))
+```
 
+```{r}
 datasummary_skim(incl.vars)
@@
 Figure \@ref(fig:vote-share-size-fig) shows no significant difference between the low-fraud and high-fraud conditions in any of the seven protest size categories, at any level of winning vote-share.
-Neither the models of protest size nor protest number support the strategic logic of protest advanced in some models, in which fraud sends a signal of weakness especially in the context of a weak ruling-party performance.
+That is, the severity of fraud does not appear associated with changes in protest size.
+As a whole, neither the models of protest size nor protest number support the strategic logic of protest advanced in some models, in which fraud sends a signal of weakness especially in the context of a weak ruling-party performance.
 
-Turning to models of protest quantity presented in Table 3, there is no statistically significant relationship between fraud and protest in any of the models.
-Figure 4, using results from Model 9, presents the marginal effects for fraud conditional on the incumbent's vote-share.
+Turning to models presented in Table 3, there is no statistically significant relationship between fraud and the number of protest events in any of the models.
+Figure \@ref(vote-share-fig), using results from Model 9, presents the marginal effects for fraud conditional on the incumbent's vote-share.
 As the figure shows, there is no statistically significant difference in the expected number of protest following a relatively clean election compared to a relatively fraudulent one, for any margin of incumbent victory.
-Taking the results from Tables 2 and 3 together, there is support for Hypothesis 1 (an unconditional increase in the risk of any protest as fraud increases), as well as strong support for Hypothesis 2 (an increase in protest risk in fraudulent, low-margin elections) and Hypothesis 4 (no relationship between fraud and protest quantity).
+
+Taking the results from Tables 2 and 3 together, Hypothesis 1\textemdash on the unconditional relationship between increasingly severe fraud and the likelihood of protest\textemdash is not supported.
+Alone, fraud severity appears to have no relationship with either the size or quantity of protest.
+Similarly, there is no support for Hypothesis 2, which predicts an increase in protest risk in fraudulent, low-margin elections.
+and Hypothesis 4 (no relationship between fraud and protest quantity).
 Hypothesis 3, on the connection between fraud, grievance, and protest, is not supported.
@@
 
-```{r vote-share-size-fig, fig.cap="Fraud, winning vote-share, and median protest size", fig.height=7}
+```{r vote-share-size-fig, fig.cap="Fraud, winning vote-share, and median protest size", fig.height=7, echo=FALSE, message=FALSE}
 plot.combined.size <- cowplot::plot_grid(p.win.zero, p.win.low, p.win.high, ncol=1)
@@
 
-```{r vote-share-fig, fig.cap="Fraud, winning vote-share, and number of protest events"}
+```{r vote-share-fig, fig.cap="Fraud, winning vote-share, and number of protest events",  echo=FALSE, message=FALSE}
 p.mm.margin + theme_bw() + xlab("Winning vote-share") + ylab("Predicted number") + labs(title = "Predicted number of post-election protest", color = "Election fraud")  

### 2022-11-30 Scripts/data_cleaning.Rmd
Cleaning ECAV data cleaning script

@@
 lji <- read.csv(here::here("Data", "LJI-estimates-20140422.csv"))
-massmob <- read.csv(here::here("Data", "mmALL_073119_csv.csv"))
 ```
@@
 nelda <- nelda %>% dplyr::rename(first.mp.election = nelda2)
-nelda <- nelda %>% rename(protests = nelda29)
+nelda <- nelda %>% rename(protests = nelda9)
 nelda <- nelda %>% rename(inc.fav.cancel = nelda34)
@@
 nelda <- nelda %>% rename(econ.crisis.nelda = nelda18)
-nelda <- nelda %>% rename(opp.votegain = nelda27)
+nelda <- nelda %>% rename(opp.votegain = nelda7)
 nelda <- nelda %>% rename(nelda.protests.fraud = nelda30)
@@
 
-Coding anti-gov violence from ecav
-
-```{r}
-## Coding anti-government initiation of violence
-
-ecav <- ecav %>% mutate(antigov.violence = 
-                          ifelse(EventViolence == 0, 0,
-                                 ifelse(ViolenceInitiator == "-99", NA,
-                                                  ifelse(ViolenceInitiator == ".1" & Actor1Side == 1, 1, 
-                                                  ifelse(ViolenceInitiator == ".2" & Actor2Side == 1, 1,
-                                                         ifelse(ViolenceInitiator == "1.1" & Target1Side == 1, 1,
-                                                                ifelse(ViolenceInitiator == "2.1" & Target2Side == 1, 1, 0)))))))
-```
-
 Adding a vdem ID to the other packages
@@
                                                destination = "vdem"))
-massmob <- massmob %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "p4n", 
-                                               destination = "vdem")) #Missing here tend to be communist states that no longer exist
 
@@
 vdem.data <- vdem.data %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
-massmob <- massmob %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
+
+ecav <- ecav %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "cown", destination = "vdem"))
 
@@
 
-Cleaning nelda date variable and massmob date
+Cleaning nelda date variable
 ```{r}
@@
 
-massmob <- massmob %>% mutate(startday = ifelse(startday < 10, paste0(0, startday), startday))
-massmob <- massmob %>% mutate(startmonth = ifelse(startmonth < 10, paste0(0, startmonth), startmonth))
-massmob <- massmob %>% mutate(protest_mdy = ifelse(is.na(startday) == T | is.na(startmonth) == T, NA,  paste0(startmonth, startday, year)))
-
-massmob <- massmob %>% mutate(protest_mdy = mdy(protest_mdy))
-massmob <- massmob %>% filter(is.na(protest_mdy) == F)
 ```
 
-Merging
+Merging Nelda with LJI and Vdem
 
@@
 nelda <- nelda %>% filter(is.na(country_id) == FALSE)
-massmob <- massmob %>% filter(is.na(country_id) == FALSE)
 
@@
 
-#nelda3 <- nelda2 %>% merge(massmob, by = "country_id_year")
-
-nelda3 <- nelda %>% dplyr::select(ccode, country, country_id, year, electionid, election_mdy)
-nelda3 <- nelda3 %>% mutate(election_year = year)
-nelda3 <- massmob %>% left_join(nelda3, by = c("country_id", "year"))
 ```
 
-##Nelda2 looks at NELDA variables as DV
+##Nelda2 looks at ECAV variables as DV
 
@@
 rm(lji)
-rm(nelda)
-rm(massmob)
+#rm(nelda)
 ```
 
+Coding anti-gov violence from ecav
+
 ```{r}
-nelda2.sub <- nelda2 %>% filter(v2x_regime == 1 | v2x_regime == 2) #Note that coding scheme for electoral democracies allows for "some degree of fraud and irregularities but these did not in the end affect the outcome" (v2elfrfair, which is a component of v2x_regime)
+## Coding anti-government initiation of violence
+
+ecav <- ecav %>% mutate(antigov.violence = 
+                          ifelse(EventViolence == 0, 0,
+                                 ifelse(ViolenceInitiator == "-99", NA,
+                                                  ifelse(ViolenceInitiator == ".1" & Actor1Side == 1, 1, 
+                                                  ifelse(ViolenceInitiator == ".2" & Actor2Side == 1, 1,
+                                                         ifelse(ViolenceInitiator == "1.1" & Target1Side == 1, 1,
+                                                                ifelse(ViolenceInitiator == "2.1" & Target2Side == 1, 1, 0)))))))
 ```
@@
 ecav.pre <- ecav %>% filter(Date < Electiondate)
-ecav.pre <- ecav.pre %>% filter(str_detect(EventName, "rotest") == T) #Only captures protests
+ecav.pre <- ecav.pre %>% filter(str_detect(EventName, "rotest") == T | str_detect(EventName, "ccupation") == T |
+                                  str_detect(EventName, "lockade") == T) #Only captures protests
 
-ecav <- ecav %>% filter(Date >= Electiondate)
-ecav <- ecav %>% filter(str_detect(EventName, "rotest") == T) #Only captures protests
+ecav.post <- ecav %>% filter(Date >= Electiondate)
+ecav.post <- ecav.post %>% filter(str_detect(EventName, "rotest") == T | str_detect(EventName, "ccupation") == T |
+                                  str_detect(EventName, "lockade") == T) #Only captures protests
 ```
+
+--Next step is to create and double-check the summaries, make sure that when merged back in with Nelda2 that we get events with zero protests
 
@@
 
+## Remove closed autoc and lib dem
+
+```{r}
+nelda2.sub <- nelda2 %>% filter(v2x_regime == 1 | v2x_regime == 2) #Note that coding scheme for electoral democracies allows for "some degree of fraud and irregularities but these did not in the end affect the outcome" (v2elfrfair, which is a component of v2x_regime)
+```
+
 ## Merge with winner share updated

### 2022-11-30 Scripts/data_cleaning.Rmd
Finalized new ECAV data cleaning procedure (includes occupations and blockades)

@@
 
+## World Bank urbanization data
+
+```{r}
+urban <- read.csv(here::here("Data", "urbanization_worldbank.csv"))
+test <- urban %>%
+  pivot_longer(cols = starts_with("X"),
+               names_to = "year")
+test <- test %>% mutate(year = gsub("X", "", year))
+
+#Next is line up country codes in test and nelda 2, then merge
+
+library(countrycode)
+
+test <- test %>% mutate(stateid = countrycode(sourcevar = Country.Code, origin = 'iso3c', destination = 'gwc'))
+test <- test %>% rename(urban.pct = value)
+test <- test %>% dplyr::select(year, urban.pct, stateid)
+
+nelda2 <- nelda2 %>% mutate(year = as.character(year))
+
+nelda2 <- nelda2 %>% left_join(test, by = c("stateid", "year"))
+
+```
+
 ##Nelda2 looks at ECAV variables as DV
@@
 count_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(n.events.pre = n())
-count_data_post <- ecav %>% group_by(NeldaID) %>% summarize(n.events.post = n())
+count_data_post <- ecav.post %>% group_by(NeldaID) %>% summarize(n.events.post = n())
 
 ##Duration
-max_date <- ecav %>% group_by(NeldaID) %>% summarize(latest.event = max(Date))
-min_date <- ecav %>% group_by(NeldaID) %>% summarize(earliest.event = min(Date))
+max_date <- ecav.post %>% group_by(NeldaID) %>% summarize(latest.event = max(Date))
+min_date <- ecav.post %>% group_by(NeldaID) %>% summarize(earliest.event = min(Date))
 duration_data <- left_join(min_date, max_date, by = "NeldaID")
@@
 ##Size
-ecav <- ecav %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
-size_data <- ecav %>% group_by(NeldaID) %>% summarize(median.participants = median(ParticipantNumber, na.rm = T))
+ecav.post <- ecav.post %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
+size_data_post <- ecav.post %>% group_by(NeldaID) %>% summarize(median.participants = median(ParticipantNumber, na.rm = T))
 
@@
 ## Violence
-violence_data <- ecav %>% group_by(NeldaID) %>% summarize(antigov.violence = sum(antigov.violence, na.rm = T))
+violence_data_post <- ecav.post %>% group_by(NeldaID) %>% summarize(antigov.violence = sum(antigov.violence, na.rm = T))
 violence_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(antigov.violence.pre = sum(antigov.violence, na.rm = T))
@@
 ##Combined
-ecav_data <- left_join(count_data_post, duration_data, by = "NeldaID")
-ecav_data <- left_join(ecav_data, size_data, by = "NeldaID")
-ecav_data <- left_join(ecav_data, violence_data, by = "NeldaID")
+ecav_data_post <- left_join(count_data_post, duration_data, by = "NeldaID")
+ecav_data_post <- left_join(ecav_data_post, size_data_post, by = "NeldaID")
+ecav_data_post <- left_join(ecav_data_post, violence_data_post, by = "NeldaID")
 
@@
 
-ecav_data <- left_join(ecav_data, ecav_data_pre, by = "NeldaID")
+ecav_data <- left_join(ecav_data_post, ecav_data_pre, by = "NeldaID")
 
@@
 ```{r}
-nelda2.sub <- nelda2.sub %>% rename(NeldaID = electionid)
-nelda2.sub <- left_join(nelda2.sub, ecav_data, by = "NeldaID")
+nelda2 <- nelda2 %>% rename(NeldaID = electionid)
+nelda2 <- left_join(nelda2, ecav_data, by = "NeldaID")
+```
+
+## Remove closed autoc and lib dem
+
+```{r}
+nelda2.sub <- nelda2 %>% filter(v2x_regime == 1 | v2x_regime == 2) #Note that coding scheme for electoral democracies allows for "some degree of fraud and irregularities but these did not in the end affect the outcome" (v2elfrfair, which is a component of v2x_regime)
+```
+
+## Remove elections before 1989-10-04 and after 2013-1-13 (ecav range)
+
+```{r}
+nelda2.sub <- nelda2.sub %>% filter(election_mdy >= "1989-10-04")
+nelda2.test <- nelda2.sub %>% filter(election_mdy <= "2013-01-13")
+```
+
+```{r}
 
@@
 
-```
+##Making the assumption that NAs for post-election protest are zero
 
-```{r}
-#nelda3 <- nelda3 %>% mutate(protests.postelex = ifelse(protests == 1 & protesterdemand1 == "political behavior, process", 1, 0))
+nelda2.sub <- nelda2.sub %>% mutate(n.events.post = ifelse(is.na(n.events.post) == T, 0, n.events.post))
+nelda2.sub <- nelda2.sub %>% mutate(antigov.violence = ifelse(is.na(antigov.violence) == T, 0, antigov.violence))
 
-#nelda3 <- nelda3 %>% mutate(protests.postelex.quant = ifelse(protests.postelex == 1, protestnumber, 0))
-```
-
-## World Bank urbanization data
-
-```{r}
-urban <- read.csv(here::here("Data", "urbanization_worldbank.csv"))
-test <- urban %>%
-  pivot_longer(cols = starts_with("X"),
-               names_to = "year")
-test <- test %>% mutate(year = gsub("X", "", year))
-
-#Next is line up country codes in test and nelda 2, then merge
-
-library(countrycode)
-
-test <- test %>% mutate(stateid = countrycode(sourcevar = Country.Code, origin = 'iso3c', destination = 'gwc'))
-test <- test %>% rename(urban.pct = value)
-test <- test %>% dplyr::select(year, urban.pct, stateid)
-
-nelda2.sub <- nelda2.sub %>% mutate(year = as.character(year))
-
-nelda2.sub <- nelda2.sub %>% left_join(test, by = c("stateid", "year"))
-
-```
-
-Excluding protest events in MM that take place more than 180 days before an election (events_pre) or more than 90 days after an election (events_post)
-
-```{r}
-elections_db <- nelda %>% dplyr::select(country_id, election_mdy) # A second dataframe with election dates by country
-
-nelda3.sub <- slice(nelda3, 0)
-
-for (i in 1:nrow(nelda3)) {
-  country_id_temp <- nelda3$country_id[i]
-  elections_db_temp <- elections_db %>% filter(country_id == country_id_temp)
-  protest_event_slice <- nelda3[i,]
-  for (j in 1:nrow(elections_db_temp)) {
-    elections_db_temp <- elections_db_temp %>% mutate(days_to_election_all =  election_mdy - protest_event_slice$protest_mdy)
-    #Negative numbers mean the protest is after the election, positive numbers mean the protest is before the election
-    elections_db_temp <- elections_db_temp %>% mutate(days_to_election_all_num = as.numeric(days_to_election_all))
-    elections_db_temp <- elections_db_temp %>% filter(days_to_election_all_num <= 180 & days_to_election_all_num >= -90)
-  }
-}
-```
-
-Cutting out protests before elections for nelda 3, 
-```{r}
-#Add electionid to the year before the election and year after (so that grouping variable can capture those protests if needed)
-nelda3 <- nelda3 %>% group_by(country_id) %>% mutate(election_date_upcoming = as_date(ifelse(is.na(election_year)==F, election_mdy, 
-                                                  ifelse(is.na(election_year)==T & is.na(lead(election_mdy))==F,
-                                                                          lead(election_mdy), NA))))
-
-nelda3 <- nelda3 %>% group_by(country_id) %>% mutate(election_date_lagged = as_date(ifelse(is.na(election_year)==F, election_mdy,                                                         ifelse(is.na(election_year)==T & is.na(lag(election_mdy))==F, lag(election_mdy), NA))))
-
-## Now filter by date of protest
-
-# Steps:
-  #1) Remove rows where election_upcoming and election_lagged are both NA. 
-  #2) This results in a dataframe with only election years +-1 year; create interval variables (lagged and leading days between protest and election)
-  #3 Filter to cases that are six months before to 3 months after election (i.e 180 days and 90 days)
-
-nelda3.sub <- nelda3 %>% filter(is.na(election_date_upcoming)==F | is.na(election_date_lagged) == F)
-
-nelda3.sub <- nelda3.sub %>% mutate(protest_preelex_interval = ifelse((election_mdy - protest_mdy) >0, (election_mdy-protest_mdy), NA))
-
-nelda3.sub <- nelda3.sub %>% mutate(protest_postelex_interval = ifelse((election_mdy - protest_mdy) <= 0, (protest_mdy - election_mdy), NA))
-
-nelda3.sub <- nelda3.sub %>% group_by(ccode) %>% filter(protest_preelex_interval <= 180 | protest_postelex_interval <= 90)
-
-# Next step is to aggregate needed variables (participants) by electionid
-```
-
-Aggregating participant variables
-```{r}
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(participants_category == "", NA, 
-                                            ifelse(participants_category == "50-99", 1, 
-                                            ifelse(participants_category == "100-999", 2,
-                                            ifelse(participants_category == "1000-1999", 3,
-                                            ifelse(participants_category == "2000-4999", 4,
-                                            ifelse(participants_category == "5000-10000", 5,
-                                                   6)))))))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100s", 2, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000s", 4, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">50", 1, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "50", 1, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000", 3, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "50+", 1, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100", 2, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "10000", 5, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "2000", 4, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "200", 2, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "300", 2, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "500", 2, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "3000", 4, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100+", 2, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "5000", 5, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "20000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "10000s", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "150", 2, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "400", 2, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "hundreds", 2, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "30000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "15000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1500", 3, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 50", 1, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "4000", 4, participants_category_ordered))  ## Stopping here; all other categories have fewer than 20 observations.
-Beginning second stopping rule:
-
-## Second rule: collect all observations greater than 10,000
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "hundreds of thousands", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "<20000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 40000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 42000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">35000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">40000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">50000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "110000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "14059", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "16000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "18000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "24000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "250000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "30000+", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "35000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "36000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "5000-50000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "55000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "650000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 10000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "13000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "500000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">15000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">20000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "14000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "400000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "Tens of thousands", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">100000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "150000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "20000+", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "60000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "90000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "10000+", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100000s", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "200000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "80000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">10000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "70000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "80000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "40000", 6, participants_category_ordered))
-
-nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "25000", 6, participants_category_ordered))
-
-#nelda3.sub <- nelda3.sub %>% group_by(electionid) %>% 
-```
-
-Remove post-election protests unrelated to election
-
-```{r}
-nelda3.sub <- nelda3.sub %>% filter(protest_mdy < election_mdy | (protest_mdy >= election_mdy & str_detect(notes, regex(pattern = "election*|vot*|ballot*", ignore_case = T))))
-```
-
-Getting grouped summarize for number and size
-```{r}
-nelda3.sub.pre <- nelda3.sub %>% filter(protest_mdy < election_mdy)
-nelda3.sub.post <- nelda3.sub %>% filter(protest_mdy >= election_mdy)
-
-count_data_mm_pre <- nelda3.sub.pre %>% group_by(electionid) %>% summarize(n.events.massmob.pre = sum(protestnumber))
-count_data_mm_post <- nelda3.sub.post %>% group_by(electionid) %>% summarize(n.events.massmob.post = sum(protestnumber))
-
-size_data_mm_pre <- nelda3.sub.pre %>% group_by(electionid) %>% summarize(median.participants.mm.pre = median(participants_category_ordered, na.rm = T))
-size_data_mm_post <- nelda3.sub.post %>% group_by(electionid) %>% summarize(median.participants.mm.post = median(participants_category_ordered, na.rm = T))
-
-```
-
-## Merge data from nelda3.sub with nelda2.sub
-
-```{r}
-##Combined post data together, then pre data, then both
-mm_data_post_comb <- left_join(count_data_mm_post, size_data_mm_post, by = "electionid")
-mm_data_pre_comb <- left_join(count_data_mm_pre, size_data_mm_pre, by = "electionid")
-
-mm_data_comb <- left_join(mm_data_pre_comb, mm_data_post_comb, by = "electionid")
-
-mm_data_comb <- mm_data_comb %>% rename(NeldaID = electionid)
-
-nelda3.sub <- nelda2.sub %>% left_join(mm_data_comb, by = "NeldaID")  ## Double check to make sure this joining works
-
-## Code NAs for number of protests as 0s (these reflect election years recorded in NELDA where there were no protests recorded by MM)
-
-nelda3.sub <- nelda3.sub %>% mutate(n.events.massmob.pre = ifelse(is.na(n.events.massmob.pre)==T, 0, n.events.massmob.pre))
-nelda3.sub <- nelda3.sub %>% mutate(n.events.massmob.post = ifelse(is.na(n.events.massmob.post)==T, 0, n.events.massmob.post))
-
-nelda3.sub <- nelda3.sub %>% mutate(median.participants.mm.pre = ifelse(is.na(median.participants.mm.pre)==T & n.events.massmob.pre == 0, 0, median.participants.mm.pre))
-nelda3.sub <- nelda3.sub %>% mutate(median.participants.mm.post = ifelse(is.na(median.participants.mm.post)==T & n.events.massmob.post == 0, 0, median.participants.mm.pre))
-
-## Rounding participant median numbers down (removes half-steps)
-
-nelda3.sub <- nelda3.sub %>% mutate(median.participants.mm.post = floor(median.participants.mm.post))
-nelda3.sub <- nelda3.sub %>% mutate(median.participants.mm.pre = floor(median.participants.mm.pre))
-
-```
-
-## Remove closed autoc and lib dem
-
-```{r}
-nelda2.sub <- nelda2 %>% filter(v2x_regime == 1 | v2x_regime == 2) #Note that coding scheme for electoral democracies allows for "some degree of fraud and irregularities but these did not in the end affect the outcome" (v2elfrfair, which is a component of v2x_regime)
 ```
@@
 nelda2.sub <- nelda2.sub %>% left_join(winner.shares, by = "NeldaID")
-nelda3.sub <- nelda3.sub %>% left_join(winner.shares, by = "NeldaID")
 
@@
 ```
-
-```{r}
-write.csv(nelda3.sub, here::here("Data", "nelda3_sub.csv"))
-
-```

### 2022-11-30 Scripts/data_cleaning.Rmd
Laptop version to overwrite

--- 2022-11-17 Scripts/data_cleaning.Rmd
@@
 ecav <- read_xls(here::here("Data", "ECAV datatset_Version 1.2.xls"))
-vdem.data <- read_rds("C:/Users/colharv/Documents/Research projects/Country_Year_V-Dem_Full_others_R_v11.1/Country_Year_V-Dem_Full+others_R_v11.1/V-Dem-CY-Full+Others-v11.1.rds") #Switch between these two for work desktop vs.
+#vdem.data <- read_rds("C:/Users/colharv/Documents/Research projects/Country_Year_V-Dem_Full_others_R_v11.1/Country_Year_V-Dem_Full+others_R_v11.1/V-Dem-CY-Full+Others-v11.1.rds") #Switch between these two for work desktop vs.
 laptop
-#vdem.data <- vdem
+vdem.data <- vdemdata::vdem
 lji <- read.csv(here::here("Data", "LJI-estimates-20140422.csv"))
@@
 nelda <- nelda %>% mutate(mmdd = ifelse(mmdd < 1000, paste0(0, mmdd), mmdd))
+
 nelda <- nelda %>% mutate(election_mdy = paste0(mmdd, year))
@@
 
+Starting with nelda2 from above, and merging with MassMob. This should yield a dataset with nelda elections, vdem data, and massmob protest data
+
+```{r}
+nelda2 <- nelda2 %>% rename(country_id = country_id.x)
+
+  election_id_temp <- nelda2$electionid[i]
+  country_temp <- nelda2$country_id[i]
+  election_date_temp <- nelda2$election_mdy[i]
+  election_row <- nelda2[i,]
+  massmob_temp <- massmob %>% filter(country_id == country_temp)
+  massmob_temp <- massmob_temp %>% mutate(days_to_election = election_date_temp - protest_mdy)
+  massmob_temp <- massmob_temp %>% filter(days_to_election <= 180 & days_to_election >= -90)
+  massmob_temp <- massmob_temp %>% mutate(electionid = election_id_temp)
+  election_row <- election_row %>% left_join(massmob_temp, by = "electionid")
+
+nelda_massmob <- election_row
+
+for (i in 101:nrow(nelda2)) {
+  election_id_temp <- nelda2$electionid[i]
+  country_temp <- nelda2$country_id[i]
+  election_date_temp <- nelda2$election_mdy[i]
+  election_row <- nelda2[i,]
+  massmob_temp <- massmob %>% filter(country_id == country_temp)
+  massmob_temp <- massmob_temp %>% mutate(days_to_election = election_date_temp - protest_mdy)
+  massmob_temp <- massmob_temp %>% filter(days_to_election <= 180 & days_to_election >= -90)
+  massmob_temp <- massmob_temp %>% mutate(electionid = election_id_temp)
+  election_row <- election_row %>% left_join(massmob_temp, by = "electionid")
+  nelda_massmob <- bind_rows(nelda_massmob, election_row)
+}
+
+##This seems to have worked. Next step should be quality control, as well as making sure all relevant nelda2 changes happen before this chunk (i.e. the urbanization data below.)
+```
+
 ##Nelda2 looks at NELDA variables as DV
@@
 
-```
-
-Excluding protest events in MM that take place more than 180 days before an election (events_pre) or more than 90 days after an election (events_post)
-
-```{r}
-elections_db <- nelda %>% dplyr::select(country_id, election_mdy) # A second dataframe with election dates by country
-
-nelda3.sub <- slice(nelda3, 0)
-
-for (i in 1:nrow(nelda3)) {
-  country_id_temp <- nelda3$country_id[i]
-  elections_db_temp <- elections_db %>% filter(country_id == country_id_temp)
-  protest_event_slice <- nelda3[i,]
-  for (j in 1:nrow(elections_db_temp)) {
-    elections_db_temp <- elections_db_temp %>% mutate(days_to_election_all =  election_mdy - protest_event_slice$protest_mdy)
-    #Negative numbers mean the protest is after the election, positive numbers mean the protest is before the election
-    elections_db_temp <- elections_db_temp %>% mutate(days_to_election_all_num = as.numeric(days_to_election_all))
-    elections_db_temp <- elections_db_temp %>% filter(days_to_election_all_num <= 180 & days_to_election_all_num >= -90)
-  }
-}
-```
-
-Cutting out protests before elections for nelda 3, 
-```{r}
-#Add electionid to the year before the election and year after (so that grouping variable can capture those protests if needed)
-nelda3 <- nelda3 %>% group_by(country_id) %>% mutate(election_date_upcoming = as_date(ifelse(is.na(election_year)==F, election_mdy, 
-                                                  ifelse(is.na(election_year)==T & is.na(lead(election_mdy))==F,
-                                                                          lead(election_mdy), NA))))
-
-nelda3 <- nelda3 %>% group_by(country_id) %>% mutate(election_date_lagged = as_date(ifelse(is.na(election_year)==F, election_mdy,                                                         ifelse(is.na(election_year)==T & is.na(lag(election_mdy))==F, lag(election_mdy), NA))))
-
-## Now filter by date of protest
-
-# Steps:
-  #1) Remove rows where election_upcoming and election_lagged are both NA. 
-  #2) This results in a dataframe with only election years +-1 year; create interval variables (lagged and leading days between protest and election)
-  #3 Filter to cases that are six months before to 3 months after election (i.e 180 days and 90 days)
-
-nelda3.sub <- nelda3 %>% filter(is.na(election_date_upcoming)==F | is.na(election_date_lagged) == F)
-
-nelda3.sub <- nelda3.sub %>% mutate(protest_preelex_interval = ifelse((election_mdy - protest_mdy) >0, (election_mdy-protest_mdy), NA))
-
-nelda3.sub <- nelda3.sub %>% mutate(protest_postelex_interval = ifelse((election_mdy - protest_mdy) <= 0, (protest_mdy - election_mdy), NA))
-
-nelda3.sub <- nelda3.sub %>% group_by(ccode) %>% filter(protest_preelex_interval <= 180 | protest_postelex_interval <= 90)
-
-# Next step is to aggregate needed variables (participants) by electionid
 ```

### 2022-12-30 Output/appendix_v4_part2.Rmd

---
title: "Appendix part 2"
author: "Cole J. Harvey (Oklahoma State University)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    bookdown::pdf_document2: default
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
header-includes: \usepackage{booktabs}
---
```{r setup, include = FALSE}
library(tidyverse)
library(countrycode)
library(readxl)
library(vdemdata)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
library(stargazer)
library(cowplot)
library(pscl)
library(modelsummary)
library(ggeffects)
library(MASS)

nelda_massmob <- read.csv(here::here("Data", "nelda_massmob_all.csv")) #nelda_complete.csv for only elex-related

nelda_massmob <- nelda_massmob %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda_massmob <- nelda_massmob %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))

##Naming nelda variables

nelda_massmob <- nelda_massmob %>% rename(boycott = nelda14)
nelda_massmob <- nelda_massmob %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                 NA)))

##Creating economic crisis variable

nelda_massmob <- nelda_massmob %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
nelda_massmob <- nelda_massmob %>% mutate(econ.crisis = as.factor(econ.crisis))

nelda_massmob <- nelda_massmob %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
nelda_massmob <- nelda_massmob %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))

nelda_massmob <- nelda_massmob %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))

##Removing cases before 1990, when ECAV starts
nelda_massmob <- nelda_massmob %>% filter(year >= 1990)

## Removing cases where inc. loses
nelda_massmob <- nelda_massmob %>% filter(inc.lose == 0)

##Nelda crisis as a factor

nelda_massmob <- nelda_massmob %>% mutate(econ.crisis.nelda.1lag = as.factor(econ.crisis.nelda.1lag))

##Intmon as a factor
nelda_massmob <- nelda_massmob %>% mutate(v2elintmon = as.factor(v2elintmon))

##Presidential as a facotr
nelda_massmob <- nelda_massmob %>% mutate(presidential = as.factor(presidential))

nelda_massmob <- nelda_massmob %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))

#Square of physint

nelda_massmob <- nelda_massmob %>% mutate(physintlag.sq = physinteg.1lag^2)

#Crisis as factor

nelda_massmob <- nelda_massmob %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))

#Margins with updates

nelda_massmob <- nelda_massmob %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))

#Successful protest = new elections or new incumbent

nelda_massmob <- nelda_massmob %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | elections.cancel.prot, 1, 0))

#Protest cases only

#protest.cases <- nelda_massmob %>% filter(any.ecav.protest == 1)

# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections

nelda_massmob <- nelda_massmob %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))

## Changing participants from numeric category to text
#nelda_massmob <- nelda_massmob %>% mutate(median.participants = round(median.participants))
#nelda_massmob <- nelda_massmob %>% mutate(particpants_categorical = ifelse(is.na(median.participants)==T, NA, ifelse(median.participants == 1, "Less than 10",
#                     ifelse(2, "10 to 99", ifelse(3, "100 to 999",
#                                                  ifelse(4, "1000 to 9,999", ifelse(5, "Greater than 10,000", NA)))))))
#nelda_massmob <- nelda_massmob %>% mutate(particpants_categorical = as.factor(particpants_categorical))

## Median particiants (MM) as factor

nelda_massmob <- nelda_massmob %>% mutate(median.participants.mm.post = as.factor(median.participants.mm.post))

## Inverting the vote buying measure so it is not reverse coded

nelda_massmob <- nelda_massmob %>% mutate(v2elvotbuy_inv = -1*v2elvotbuy)

```

## Alternative measures of election manipulation

To ensure that the results presented in the main text are not biased by the chosen measure of election fraud, this section replaces that measure with a broader indicator of election integrity from V-Dem. This variable is `v2xel_frefair,' which is a comprehensive indicator of quality of the electoral process, including measures of electoral management body autonomy and capacity, vote-buying, intimidation, fraud, and an overall assessment of the fairness of the election. This variable ranges from 0 to 1, with higher values indicating freer and fairer elections. The models presented below utilize this variable in place of the main measure of election fraud, while retaining all other control and explanatory variables, using the Mass Mobilization data on protest.

```{r manipdv, echo=FALSE}
m1 <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   + presidential + inc.termlimit.num  +  libdem.1lag + urban.pct + rentierstate + taxstate + n.events.massmob.pre + median.participants.mm.pre  + (1 | stateid), data = nelda_massmob %>% filter(n.events.massmob.pre < 200)) #There are two extreme observations removed by this filter

#summary(m1)

modelsummary(m1, stars = T,  title = 'Random-intercept multilevel model of overall election integrity')
```

The first model shows that the number of pre-election protests is weakly associated with improved election integrity, but just as in the main text this effect is extremely small. Each additional protest is associated with an improvement in election integrity equivalent to 0.2\% of the standard deviation of the dependent variable. Given that the mean number of pre-election protests is six, this is an extremely small influence on the overall degree of election integrity.

```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
model.n.novoteshare.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    + v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +     + v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
summary(model.n.novoteshare.hurd.mm)

model.n.base.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    winner.share.upd + v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    winner.share.upd + v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
summary(model.n.base.hurd.mm)

model.n.base2.hurd.mm <- hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre |  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
summary(model.n.base2.hurd.mm)

model.n.griev.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   gdpgro.1lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   gdpgro.1lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, dist = "negbin", data = nelda_massmob %>% filter(gdpgro.1lag < 1))
summary(model.n.griev.hurd.mm)

model.n.griev2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob, dist = "negbin")
summary(model.n.griev2.hurd.mm)

model.n.griev3.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   pubseccorrup.lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre  | jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   pubseccorrup.lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, data = nelda_massmob, dist = "negbin")
summary(model.n.griev3.hurd.mm)
```

Moving to the hurdle models of post-election protest, we do not observe much of a relationship between overall election integrity and protest. There is no statistically significant relationship between election integrity and protest occurrence in Model 2. Neither is there any significant relationship in interaction with the winner's vote-share, as shown in Figure \@ref(fig:b-protests-margin), or with economic crisis, as in Figure \@ref(fig:b-protests-crisis). The coefficient on the interaction term for public sector corruption and overall election integrity is statistically significant, though there is no significant effect at typical values for covariates and election integrity (as shown in Figure \@ref(fig:b-protests-corrupt)).

In the corresponding results for the number of post-election protests, there is similarly muted evidence for a connection between the overall level of manipulation and protest risk to incumbents. There is no significant difference between the marginal effect of common levels of election integrity on the number of post-election protests for typical cases across any margin of victory (Figure \@ref(fig:n-protests-margin)). During an economic crisis, *cleaner* elections are associated with an increase in the number of protests (\@ref(fig:n-protests-crisis)). This is plausibly due to the crisis weakening both the government's ability to generate manipulation [@greene_why_2007] and its ability to co-opt or put down the protests. Finally, public sector corruption is associated with a significant increase in the marginal effect of poor-integrity elections on protest\textemdash but only when the corruption measure is at its maximum (Figure \@ref(fig:n-protests-corrupt)).

As a whole, the results are in accordance with the theory proposed in the main paper. There is no consistent evidence that lower-quality elections are associated with either a higher risk of protest initiation, or with a larger number of post-election protests.

```{r zero-model-table, echo=FALSE}

models <- list()
models[['Model 2']] <- model.n.base.hurd.mm
models[['Model 3']] <- model.n.base2.hurd.mm
#models[['Model 4']] <- model.n.griev.hurd.mm
models[['Model 4']] <- model.n.griev2.hurd.mm
models[['Model 5']] <- model.n.griev3.hurd.mm

modelsummary::modelsummary(models, coef_omit = "count", gof_omit = "R2", stars = T, longtable=T,  title = 'Binary portion of hurdle models of post-election protest')

```

```{r zero-model-table-count, echo=FALSE}
modelsummary::modelsummary(models, coef_omit = "zero", gof_omit = "R2", longtable=T, stars = T, title = 'Hurdle models of post-election protest (counts)')

```

```{r, echo=FALSE, message=FALSE}
dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd", "v2xel_frefair [.2, .7]")) 
p.n.margin <- plot(dat)
p.n.margin <- p.n.margin  + theme_bw()

dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd", "v2xel_frefair [.2, .7]"), type = "zi_prob") 
p.b.margin <- plot(dat)
p.b.margin <- p.b.margin  + theme_bw()

```

```{r n-protests-margin, fig.cap="Fraud, public sector corruption, and number of protest events"}
p.n.margin

```

```{r b-protests-margin, fig.cap="Fraud, public sector corruption, and likelihood of protest"}
p.b.margin

```

```{r, echo=FALSE, message=FALSE}
dat <- ggpredict(model.n.griev2.hurd.mm, terms = c("econ.crisis.nelda.1lag", "v2xel_frefair [.15, .66]")) 
p.n.crisis <- plot(dat)
p.n.crisis <- p.n.crisis  + theme_bw()

dat <- ggpredict(model.n.griev2.hurd.mm, terms = c("econ.crisis.nelda.1lag", "v2xel_frefair [.15, .66]"), type = "zi_prob") 
p.b.crisis <- plot(dat)
p.b.crisis <- p.b.crisis +  theme_bw()

```
```{r n-protests-crisis, fig.cap="Fraud, economic crisis, and number of protest events"}
p.n.crisis
```

```{r b-protests-crisis, fig.cap="Fraud, economic crisis, and likelihood of protest"}
p.b.crisis
```

```{r, echo=FALSE, message=FALSE}
dat <- ggpredict(model.n.griev3.hurd.mm, terms = c("pubseccorrup.lag", "v2xel_frefair [.29, .66]")) 
p.corr.n.margin <- plot(dat)
p.corr.n.margin <- p.corr.n.margin + coord_cartesian(ylim = c(0, 100)) + theme_bw()

dat <- ggpredict(model.n.griev3.hurd.mm, terms = c("pubseccorrup.lag", "v2xel_frefair [.29, .66]"), type = "zi_prob") 
p.corr.b.margin <- plot(dat)
p.corr.b.margin <- p.corr.b.margin  + theme_bw()

```

```{r n-protests-corrupt, fig.cap="Fraud, public sector corruption, and number of protest events"}
p.corr.n.margin
```

```{r b-protests-corrupt, fig.cap="Fraud, public sector corruption, and likelihood of protest"}
p.corr.b.margin
```

### 2022-12-30 Output/manipulation_protest_draft_v4.Rmd

--- 2022-11-17 Output/manipulation_protest_draft_v3.Rmd
@@
 ---
-title: "Draft: A missing guardrail---The minimal effect of election fraud on protest intensity in non-democracies"
+title: "Draft v4: A missing guardrail---The minimal effect of election fraud on protest intensity in non-democracies"
 author: "Cole J. Harvey (Oklahoma State University)"
@@
 ---
+
+TODO: Fix git merge issue on data cleaning file. Plan to stick with massmob version--Show economic marginal effects plots. present plots for both the zero-inflated and count versions of the models.
 
@@
 
-With some exceptions, such as the study by Lankina and Skovoroda [-@Lankina2017], most research on election fraud and protest treats the latter as a binary outcome.
-The foregoing discussion of the literature suggests several baseline hypotheses to be tested regarding post-election protest as a binary variable.
-These include, first, the prediction that more severe election fraud is a signal of deeper regime weakness (i.e.
-unpopularity), which makes protest more likely.
-Second, there is the alternative view that election manipulation is most likely to spark a protest movement when margins of victory are narrow [@Simpser2011].
-Finally, the grievances approach suggests that fraud will most likely lead to protest when social grievances are high [@Brancati2016;@tucker2007].
+As noted above, much of the formal literature on election manipulation assumes that the risk of protest serves as a deterrent to election manipulation.
+Given that pre-election protest is likely to be a strong predictor of post-election protest\textemdash by indicating strong social networks [@mateo2022], pre-existing opposition groups and civil society capacity [@Bunce2006], and/or popular grievances [@Brancati2016]\textemdash the deterrent logic suggests that non-democratic governments will pull back on election-day fraud when the pre-election environment is characterized by protest [@Rozenas2016b].
+This implies the following hypothesis.
 
->H1: Post-election protest will be more likely to occur when fraud is severe.
+>H1: Pre-election protest will be associated with lower levels of election-day manipulation.
 
->H2: More sever election fraud is associated with a higher likelihood of protest when winning-party vote-share is low.
+In addition, the logic of deterrence suggests that increasingly severe election fraud should make post-election more likely.
+More extensive election-day manipulation, in these models, is generally taken as a signal of regime unpopularity.
+This in turn makes citizens [@Egorov2021;@Little2015] or opposition parties [@Luo2020;@Rozenas2016b;@magaloni2010] more likely to stage protests.
+
+>H2: Post-election protest will be more likely to occur when fraud is severe.
+
+In addition to the deterrence mechanism, previous research has also suggested conditional mechanisms, whereby a fraudulent election is more likely to trigger protest under specific circumstances.
+These are: when the incumbent's official margin of victory is low [@Simpser2011], and when social grievances are high [@Brancati2016;@tucker2007].
+
+>H2: More severe election fraud is associated with a higher likelihood of protest when winning-party vote-share is low.
 
@@
 
-However, these predictions may not apply to the size or intensity of a protest wave.
-Once protests begin, the extent of electoral fraud is unlikely to affect their scope.
+With some exceptions, such as the study by Lankina and Skovoroda [-@Lankina2017], most research on election fraud and protest treats the latter as a binary outcome.
+Such an approach does not consider how the protest environment might evolve after the election.
+Here I argue that, once protests begin, the extent of electoral fraud is unlikely to affect their scope.
 The initial emergence of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014].
@@
 
-nelda3.sub <- read.csv(here::here("Data", "nelda3_sub.csv"))
+nelda_massmob <- read.csv(here::here("Data", "nelda_massmob_all.csv")) #nelda_complete.csv for only elex-related
 
-nelda3.sub <- nelda3.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
-nelda3.sub <- nelda3.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))
+nelda_massmob <- nelda_massmob %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
+nelda_massmob <- nelda_massmob %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))
 
@@
 
-nelda3.sub <- nelda3.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
-nelda3.sub <- nelda3.sub %>% mutate(econ.crisis = as.factor(econ.crisis))
+nelda_massmob <- nelda_massmob %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
+nelda_massmob <- nelda_massmob %>% mutate(econ.crisis = as.factor(econ.crisis))
 
-nelda3.sub <- nelda3.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
-nelda3.sub <- nelda3.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))
+nelda_massmob <- nelda_massmob %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
+nelda_massmob <- nelda_massmob %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))
 
-nelda3.sub <- nelda3.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))
+nelda_massmob <- nelda_massmob %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))
 
 ##Removing cases before 1990, when ECAV starts
-nelda3.sub <- nelda3.sub %>% filter(year >= 1990)
+nelda_massmob <- nelda_massmob %>% filter(year >= 1990)
 
 ## Removing cases where inc. loses
-nelda3.sub <- nelda3.sub %>% filter(inc.lose == 0)
+nelda_massmob <- nelda_massmob %>% filter(inc.lose == 0)
+
+##Nelda crisis as a factor
+
+nelda_massmob <- nelda_massmob %>% mutate(econ.crisis.nelda.1lag = as.factor(econ.crisis.nelda.1lag))
 
 ##Intmon as a factor
-nelda3.sub <- nelda3.sub %>% mutate(v2elintmon = as.factor(v2elintmon))
+nelda_massmob <- nelda_massmob %>% mutate(v2elintmon = as.factor(v2elintmon))
 
-nelda3.sub <- nelda3.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))
+##Presidential as a facotr
+nelda_massmob <- nelda_massmob %>% mutate(presidential = as.factor(presidential))
+
+nelda_massmob <- nelda_massmob %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))
 
@@
 I further limit the sample to regimes coded as electoral autocracies or electoral democracies in V-Dem.
-This excludes closed autocracies (where no national elections are held) and liberal democracies (where electoral integrity and rule of law are both high).
-Finally, the date range is limited by the availability of data on protest from ECAV, as discussed below, which runs from 1990 to 2012.
-The resulting dataset includes 647 election-period observations across 109 countries.
+This excludes closed autocracies (where no national elections are held) and liberal democracies (where electoral integrity and rule of law are both high).\footnote{The specific variable used is 'v2x_regime' from V-Dem.
+This variable codes an electoral democracy, in part, if the country scores above a 2 on the measure of free and fair elections ('v2elfrfair_osp').
+This still permits election fraud to be detected in the electoral democracies; v2elfrfair can take on a value of 3 when "deficiencies and some degree of fraud and irregularities" occur.} Finally, the beginning of the date range is limited by the availability of data on protest from the Mass Mobilization dataset, as discussed below, which begins in 1990.
+The inclusion of a control variable from the Latent Judicial Independence dataset [@Linzer2015] limits the end of the time range to 2012.
+The resulting dataset includes 647 total election-period observations across 109 countries.
 
-ECAV, from which study's dependent variable is drawn, records protests and other contentious actions from six months before to three months after an election.
-For this study, I include protests only, excluding acts like shootings, bombings, and coups.
-ECAV data is in event-day format.
-To capture the *number of protests*, I take the sum number of events recorded in ECAV from election day to the end of the dataset's coverage period ninety days later.
-The result is a count variable ranging from zero to 90.\footnote{That the max value for the quantity of protests matches the number of days in the recording window is coincidental.} The large majority of elections experience no post-election protest: in 460 of the 647 elections, no protests are recorded.
-Figure 1 reports the distribution of protest counts across the dataset, as well as winners' vote-shares.
+The Mass Mobilization dataset, from which study's measures of protest are drawn, attempts to record all anti-government protests drawing 50 or more participants.
+The data is in event-day format.
+I divide this data into a pre-election period (180 days prior to an election) to capture the number and size of pre-election protest, and a post-election period (90 days after the election) for the dependent variables.
+I include all protest events in the post-election period, rather than filtering based on key terms like "election" or "vote," since the conduct and outcome of the election may spark protests that do not explicitly refer back to the election in protesters' demands.
+This is an extension of both the strategic and grievance logics of post-election protest.
+An electoral outcome that signals weakness for the incumbent regime may lead to strategic protests by outsider groups pressing their demands, even if they do not specifically focus on the integrity of the election.
+Likewise, disaffected groups may use the election as a focal point for their socioeconomic demands, irrespective of the nature of the election.
+Excluding such cases would like bias the results as a consequence.
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
+Two variables are derived from the Mass Mobilization data.
+First, I collect the *median size* of post-election protest.
+The Mass Moblization data records protest size using six categories.\footnote{This are: 50-99 participants, 100-999, 1,000-1,999, 2,000-4,999, 5,000-9,999, and greater than 10,000 participants.} I then take the median size of the pre-election set of protests per election period; the result is an ordered categorical variable.
+Second, I capture the  *number of protests* by taking the sum number of events recorded in the pre- and post-election data.
+The result is a count variable ranging from zero to 435 for the pre-election period and 0 to 385 for the post-election period.
+The large majority of elections experience no post-election protest: in 504 of the 647 elections, no protests are recorded.
+Figure \@ref(fig:elex-hist) reports the distribution of protest counts across the dataset, as well as winners' vote-shares.
 
-ggplot(nelda3.sub, aes(x=n.events.massmob.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()
+```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
 
-ggplot(nelda3.sub, aes(x=winner.share.upd)) + geom_histogram(bins=100) + xlab("Incumbent vote-share") + theme_bw()
+p.hist.events <- ggplot(nelda_massmob, aes(x=n.events.massmob.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()
+
+p.hist.fraud <- ggplot(nelda_massmob, aes(x=v2elirreg.inv)) + geom_histogram(bins=100) + xlab("Election fraud") + theme_bw()
 ```
 
-```{r, include=FALSE}
-
-p.success <- ggplot(nelda3.sub, aes(x=as.numeric(median.participants.mm.post), fill = as.factor(protest.success), color=as.factor(protest.success)))+
-  geom_histogram(position="identity", alpha=0.5, binwidth = 1) 
-p.success + coord_cartesian(ylim = c(0, 100))
+```{r elex-hist, fig.cap="Distributions of election fraud and post-election protest", fig.height=7, echo=FALSE, message=FALSE, warning = FALSE}
+plot.combined.hist <- cowplot::plot_grid(p.hist.fraud, p.hist.events, ncol=1)
+plot.combined.hist
 ```
 
-```{r, include=FALSE}
-mu <- plyr::ddply(nelda3.sub, "protest.success", summarise, grp.mean=mean(n.events.massmob.post))
-head(mu)
-
-p.success <- ggplot(nelda3.sub, aes(x=n.events.massmob.post, fill = as.factor(protest.success), color=as.factor(protest.success)))+
-  geom_histogram(position="identity", alpha=0.5, binwidth = 1) + 
-  geom_vline(data=mu, aes(xintercept=grp.mean, color=as.factor(protest.success)),
-             linetype="dashed")
-p.success
+```{r, include=F}
+incl.vars <- nelda_massmob %>% dplyr::select(jucon.1lag, legcon.1lag, altinf.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, harass_opp,  median.participants.mm.pre, median.participants.mm.post)
+#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))
 ```
 
-```{r}
-incl.vars <- nelda3.sub %>% dplyr::select(jucon.1lag, legcon.1lag, altinf.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, harass_opp, any.ecav.protest, median.participants.mm.pre, median.participants.mm.post)
-#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))
-
+```{r, echo=FALSE}
 datasummary_skim(incl.vars)
@@
 Figure 2 shows summary statistics for all of these variables.
+
+# Results
+
+## Pre-election protest and election-day fraud
+
+The results of these models are not promising for argument that protest risk deters election manipulation.
+First, while Hypothesis 1 is supported in Model 1 in Table  \@ref(tab:manipdv) \textemdash the coefficient on *pre-election protest* is negative and significant\textemdash the substantive effect of the relationship is minuscule.
+The model predicts that each additional pre-election protest event is associated with a reduction in the fraud measure of 0.007.
+This is a very small effect, given that the dependent variable ranges from `r min(nelda_massmob$v2elirreg.inv, na.rm = T)` to `r max(nelda_massmob$v2elirreg.inv, na.rm = T)`.
+A very large increase in the degree of protest, from zero to the ninetieth percentile (`r as.numeric(quantile(nelda_massmob$n.events.massmob.pre, probs = .9))` events) is associated with a `r (as.numeric(quantile(nelda_massmob$n.events.massmob.pre, probs = .9)))*.007` reduction in election-day fraud.
+This represents less that ten percent of a standard deviation of the election fraud variable.
+So, while pre-election protest (an indicator of post-election protest risk) is indeed associated with reduced election-day fraud, this reduction is only likely to be noticeable in extreme circumstances.
+For example, pre-election protest in the 99th percentile is associated with a reduction of `r (as.numeric(quantile(nelda_massmob$n.events.massmob.pre, probs = .99)))*.007`, a bit less than half of a standard deviation.
+This suggests that the level of election manipulation is not highly responsive to changes in the pre-election protest environment.
+
+```{r manipdv, echo=FALSE}
+m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   + presidential + inc.termlimit.num  +  libdem.1lag + urban.pct + rentierstate + taxstate + n.events.massmob.pre + median.participants.mm.pre  + (1 | stateid), data = nelda_massmob %>% filter(n.events.massmob.pre < 200)) #There are two extreme observations removed by this filter
+
+#summary(m1)
+
+modelsummary(m1, stars = T,  title = 'Random-intercept multilevel model of election-day fraud')
+
+```
+
+## Election-day fraud and protest risk
+
+```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
+model.n.novoteshare.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +     + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+summary(model.n.novoteshare.hurd.mm)
+
+model.n.base.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+summary(model.n.base.hurd.mm)
+
+model.n.base2.hurd.mm <- hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre |  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+summary(model.n.base2.hurd.mm)
+
+model.n.griev.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, dist = "negbin", data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+summary(model.n.griev.hurd.mm)
+
+model.n.griev2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob, dist = "negbin")
+summary(model.n.griev2.hurd.mm)
+
+model.n.griev3.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre  | jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, data = nelda_massmob, dist = "negbin")
+summary(model.n.griev3.hurd.mm)
+```
+
+The second set of models also challenge the conventional wisdom that more severe electoral manipulation is associated with an increased risk of anti-government protest, except under extraordinary circumstances.
+Beginning with the binary portion of the hurdle models, which model the likelihood of any protest occurring after the election, the standalone coefficient for election fraud in Model 2 is not statistically significant.
+Model 3 interacts election fraud with the incumbent's official vote-share; here we do observe statistically significant relationships at the p < .05 level.
+However, as the marginal effects plot in Figure \@ref(fig:vote-share-prob-fig) shows, this relationship is only statistically significant when comparing extreme scenarios.
+
+While more fraudulent elections are associated with an increasing risk of protest as margins of victory shrink, this increase is not statistically  significant for typical values of manipulation.
+In the bottom panel of Figure  \@ref(fig:vote-share-prob-fig), typical values for low and high levels of fraud are used (one standard deviation below and above the mean, respectively).
+In this lower panel, the increased risk of protest in the high-fraud scenario is not statistically distinguishable from the low-fraud scenario.
+It is only when comparing extreme values, as in the top panel where the minimum and maximum values for election fraud are used, that the increase in risk becomes significant.
+It is also worth noting that the inclusion of the interaction term does not improve the fit of the model compared to the baseline Model 2.
+In summary, it appears that though the strategic logic of protest can be detected\textemdash protest appears more likely when fraud is high and winning margins low\textemdash this pattern only emerges in unusual scenarios.
+This finding, where the relationship become fraud and protest is only visible at the margins, is similar to that observed in Model 1 above.
+
+A similar, but somewhat stronger effect, can also be seen in the relationship between fraud, public sector corruption, and protest.
+As shown in Figure \@ref(fig:corrupt-prob-fig), higher levels of fraud are associated with higher risk of protest when corruption is high.
+This effect is not statistically significant for typical values of fraud (one SD above and below the mean), but quickly becomes so when comparing more disparate levels of fraud.
+At the maximum values for fraud, as the top panel shows, the difference in risk is significant even for modestly elevated levels of corruption.
+This is at least partially in line with the predictions of the literature following Tucker [-@tucker2007], though the relationship is not significant when comparing more typical levels of fraud.
+Finally, there is no significant effect for *economic crisis*, at least under the typical covariate values used to create the marginal effects plot in Figure \@ref(fig:crisis-prob-fig).
+
+```{r zero-model-table, echo=FALSE}
+
+models <- list()
+models[['Model 2']] <- model.n.base.hurd.mm
+models[['Model 3']] <- model.n.base2.hurd.mm
+#models[['Model 4']] <- model.n.griev.hurd.mm
+models[['Model 4']] <- model.n.griev2.hurd.mm
+models[['Model 5']] <- model.n.griev3.hurd.mm
+
+modelsummary::modelsummary(models, coef_omit = "count", gof_omit = "R2", stars = T, longtable=T,  title = 'Binary portion of hurdle models of post-election protest')
+
+```
+
+```{r include=FALSE}
+nelda_massmob <- nelda_massmob %>% mutate(any_protest = ifelse(n.events.massmob.post > 0, 1, 0))
+
+model.noint.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, family = binomial(link = "logit"), data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+summary(model.noint.binary)
+
+model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, family = binomial(link = "logit"), data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+summary(model.margin.binary)
+
+model.binary.griev1 <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, family = binomial(link = "logit"), data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+summary(model.binary.griev1)
+
+model.binary.griev2 <- glm(any_protest ~  jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob, family = binomial(link = "logit"))
+summary(model.binary.griev2)
+
+model.binary.griev3 <- glm(any_protest ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, data = nelda_massmob, family = binomial(link = "logit"))
+summary(model.binary.griev3)
+
+## Plots
+
+####Margin
+
+plot_binary_margin_max <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-2.3, 3.2]"))
+
+plot_binary_margin_typ <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]"))
+
+###GDP
+
+plot_binary_gdp_max <- sjPlot::plot_model(model.binary.griev1, type = "pred", terms = c("gdpgro.1lag [all]", "v2elirreg.inv [-2.3, 3.2]"))
+
+plot_binary_gdp_typ <- sjPlot::plot_model(model.binary.griev1, type = "pred", terms = c("gdpgro.1lag [all]", "v2elirreg.inv [-.45, 1.78]"))
+
+###Crisis
+
+plot_binary_crisis_max <- sjPlot::plot_model(model.binary.griev2, type = "pred", terms = c("v2elirreg.inv [all]", "econ.crisis.nelda.1lag"))
+
+plot_binary_crisis_typ <- sjPlot::plot_model(model.binary.griev2, type = "pred", terms = c("v2elirreg.inv [all]", "econ.crisis.nelda.1lag"))
+
+##Corruption
+
+plot_binary_corruption_max  <- sjPlot::plot_model(model.binary.griev3, type = "pred", terms = c("pubseccorrup.lag", "v2elirreg.inv [-2.3, 3.2]"))
+
+plot_binary_corruption_typ <- sjPlot::plot_model(model.binary.griev3, type = "pred", terms = c("pubseccorrup.lag", "v2elirreg.inv [-.45, 1.78]"))
+```
+
+```{r vote-share-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE}
+plot.combined.prob <- cowplot::plot_grid(plot_binary_margin_max, plot_binary_margin_typ, ncol=1)
+plot.combined.prob
+```
+
+```{r crisis-prob-fig, fig.cap="Fraud, economic crisis, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE}
+plot.combined.crisis.prob <- cowplot::plot_grid(plot_binary_crisis_max, plot_binary_crisis_typ, ncol=1)
+plot.combined.crisis.prob
+```
+
+```{r corrupt-prob-fig, fig.cap="Fraud, public sector corruption, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE}
+plot.combined.corrupt.prob <- cowplot::plot_grid(plot_binary_corruption_max, plot_binary_corruption_typ, ncol=1)
+plot.combined.corrupt.prob
+```
+
+## Election-day fraud and number of post-election protests
+
+```{r n-protests-margin, fig.cap="Fraud, public sector corruption, and number of protest events", echo=FALSE, message=FALSE}
+dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd", "v2elirreg.inv [-.45, 1.64]")) 
+p.n.margin <- plot(dat)
+p.n.margin + coord_cartesian(ylim = c(0, 70)) + theme_bw()
+
+```
+
+```{r n-protests-crisis, fig.cap="Fraud, public sector corruption, and number of protest events", echo=FALSE, message=FALSE}
+dat <- ggpredict(model.n.griev2.hurd.mm, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-.45, 1.64]")) 
+p.n.crisis <- plot(dat)
+p.n.crisis + coord_cartesian(ylim = c(0, 75)) + theme_bw()
+
+```
+
+```{r n-protests-corrupt, fig.cap="Fraud, public sector corruption, and number of protest events", echo=FALSE, message=FALSE}
+dat <- ggpredict(model.n.griev3.hurd.mm, terms = c("pubseccorrup.lag", "v2elirreg.inv [-.45, 1.64]")) 
+p.corr.n.margin <- plot(dat)
+p.corr.n.margin + coord_cartesian(ylim = c(0, 120)) + theme_bw()
+
+```
+
+```{r zero-model-table-count, echo=FALSE}
+modelsummary::modelsummary(models, coef_omit = "zero", gof_omit = "R2", longtable=T, stars = T, title = 'Hurdle models of post-election protest (counts)')
+
+```
+
+Turning to the models of number of protest events shown in Table \@ref(tab:zero-model-table-count), Model 2\textemdash which includes no interaction terms\textemdash shows no significant relationship between fraud and protest in the count portions of the hurdle regression.
+Similarly, none of the three models that include interaction terms yield a significant relationship between fraud, the conditional variable, and number of protest events.
+Model 3 shows no significant interaction effect for fraud and winning vote-share on quantity of protest.
+Figure \@ref(fig:n-protests-margin) shows the marginal effect of election fraud at varying levels of winning vote-share, holding other variables constant at their means (continuous variables) or modes (categorical variables).
+The high and low levels of fraud in the plot represent one standard deviation above and below the mean, respectively.
+As the figure shows, the predicted number of post-election protests is somewhat higher in the high fraud condition when winning margin is low, but this relationship is not statistically significant.
+
+Neither economic crises nor public sector corruption is associated with a significant marginal effect for typical levels of election fraud on protest quantity, as Figures \@ref(fig:n-protests-crisis) and \@ref(fig:n-protests-corrupt) show.\footnote{Figures demonstrating minimum and maximum values for fraud are shown in the appendix; they also indicate no significant relationship}.
+In fact, more fraudulent elections appear to be associated with *fewer* incidents of protest during periods of economic crisis, though this distinction is not statistically significant at the p < .05 level.
+
+Altogether, the argument that election fraud increases the risk of protest initiation is mixed.
+Hypothesis 1 is weakly supported: pre-election protest is associated with reduced election-day fraud, but the substantive effect of this relationship is minimal.
+Hypothesis 2 is also weakly supported; there is a significant and sizable association between election fraud and protest initiation when the ruling party's winning margin is low, but only at extremely low and high values of fraud.
+Support for Hypothesis 3 is mixed\textemdash fraud severity is not associated with increased protest risk during economic crises, though there is a significant and sizable increase in the marginal effect as public sector corruption increases.
+Election fraud appears to be associated with a higher risk of protest initiation, but generally in unusual scenarios\textemdash when comparing a highly fraudulent to a largely clean elections when the incumbent's margin of victory is quite low, for example\textemdash or with meager substantive effect.
+
+However, even this mixed finding likely presents too optimistic a picture for the protest-deterrent model of election.
+fraud.
+The count results from the hurdle regressions show no effect for election fraud across any of the models, alone or in interaction.
+Elections that were more fraudulent are not associated with more numerous protest.
+Even when protests are initiated\textemdash a relatively rare event in the first place\textemdash the size of a protest wave is not influenced by the degree of fraud perpetrated by the ruling party.
+In turn, this suggests two pessimistic interpretations of fraud-protest dynamic.
+First, protest may simply be less damaging to ruling parties than is regularly assumed in theories of election manipulation.
+Faced with a nascent protest movement, ruling parties are generally able to contain, co-opt, or repress them\textemdash regardless of the severity of election fraud.
+An even more troubling implication would suggest that the protest-manipulation logic, rather than serving as a deterrent, creates perverse incentives to manipulate as much as possible.
+If ruling parties do indeed seek to avoid protest, but manipulating heavily does not come with a cost in the number of post-election protests, ruling parties have an incentive fraudulently inflate their official margins of victory in order to reduce the risk of protest initiation.
+
+# Discussion
+
+Altogether, the findings challenge the idea that the expectation of protest will deter election-manipulation by ruling parties.
+To the extent that fraud is a signal of incumbent strength, it is a noisy and unreliable one.
+When an incumbent party wins by a narrow margin in a highly fraudulent election, the likelihood of protest initiation is substantially higher than when it wins more cleanly: roughly a 50\% chance in the former case versus 5\% in the latter as shown in Figure 2.
+However, since even strong incumbents commit fraud, the decision to take to the streets is often a losing gamble\textemdash anticipating a weak incumbent, they instead find one that is capable of repressing, coopting, or conciliating an emerging protest movement.
+Indeed, while there are 143 cases of protest out of the 647 elections in this dataset, only 9 elections resulted in either the incumbent being replaced or the election being re-run due to mass protests (according to NELDA).\footnote{This information is determined using the NELDA variables NELDA37 and NELDA41.
+The cases of successful protest are: Ukraine 2004, Azerbaijan 2005, Georgia 2003, Kyrgyzstan 2005, Bulgaria 1990, Bangladesh 1996, Mali 1997, Peru 2000, and Cote d'Ivoire 2000.} This suggests a `success rate' for post-election protest of about six percent *conditional on protest initiation*, and about one percent for all elections in electoral authoritarian and electoral democratic regimes.
+
+The above results have several implications for understanding election manipulation and regime resilience.
+Most importantly, by treating protest primarily as a binary event rather, the existing literature on electoral protest has overemphasized the risk to incumbents of engaging in electoral manipulation.
+This results presented here show that most electoral protests are narrow, and that fraud has no bearing on how numerous protest events are.
+Moreover, contrary to some prior research, the risk of protest does not appear to be heightened by poor economic conditions [@Brancati2014a] or intensive corruption in the public sector.
+As a result, circumstances that are to some extent outside of leaders' control, like a worsening macroeconomic climate, do not appear to make engaging in fraud any riskier.
+Instead, the conditions that are correlated with more damaging protest are generally institutional factors, which are either more amenable to control or can be anticipated by non-democratic incumbents.
+Legislative constraints, an important factor for the duration and number of protests, can be adjusted through institutional reforms and cooptation of the opposition, over time.
+Incumbents may attempt to manage term limit through constitutional reforms, where possible, or through the appointment of (presumably) loyal successors.
+In other words, election fraud only has a minor cost in terms of protest risk, a risk which can be further blunted to the extent that elections can often be held on incumbents' terms.
+
+These are pessimistic findings for proponents of democracy.
+That electoral protest is less risky than is often assumed, in turn, implies that governments and ruling parties have a freer hand to engage in electoral manipulation than many models posit.
+Since the severity of fraud has no apparent bearing on whether or not a major protest wave will emerge as a serious threat to the regime, as shown in Figure 3, government have an incentive to manipulate as much as possible to secure a large winning vote-share.
+This perverse incentive undermines the case that democracy, and fair elections generally, are a self-enforcing equilibrium backed up by the risk of citizen protest [@fearon_self-enforcing_2011].
+
+One objection to this framing may be that, if the level of manipulation and the risk of protest are endogenous to one another, perhaps the null result found for the number of protests is to be expected.
+That is, if incumbents choose a level of manipulation calculated to avoid large-scale protest---sometimes high and sometimes low, conditional on contextual factors---no correlation between fraud and protest quantity would be found.
+This paper does not dispute that decisions about manipulation and protest may be endogenous to one another, but it does argue that the resulting equilibrium favors governments.
+In the dataset used here, only ten elections in four countries are coded as having an *intentional irregularities* score of less than -2 (indicating extremely fraud-free elections).
+By contrast, 103 elections in 27 countries score higher than 2 (indicating highly fraudulent elections).
+The frequency of fraud, the rarity of protest, and the finding that fraud severity is not correlated with the scope of protest events all point to the same conclusion: protest risk is not a substantial constraint on election fraud in non-democracies.
+Rather, the constraints on manipulation are more likely to be driven by factors like resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022].
+
+These results are robust to different specifications presented in an online appendix.
+In particular, the main results are supported when data from the Electoral Contention and Violence (ECAV) dataset [@Daxecker2019] are used in place of the Mass Mobilization data.
+Other robustness checks include...
+
+# Conclusion
+
+Protest risk is often considered to be a major risk to incumbent governments that work to undermine election integrity.
+This risk has a central role in several prominent models of electoral manipulation, serving to deter manipulation and to incentivize the toleration of election monitors.
+However, empirical research on the subject has largely treated protest as a binary variable, which obscures the degree to which post-election protest imposes a real risk on incumbents in non-democracies.
+One protest may be easily contained, while dozens may pose a more significant risk to the regime.
+
+This paper addresses this concern by testing the relationship between election fraud, the risk that any protest occurs, and the number of protest events that occur after an election.
+Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 647 elections to show that---while there is an elevated risk of some protest occurring in a fraudulent election with narrow margins of victory---election fraud has no significant relationship with the number of protests that occur in a protest wave.
+These results imply that protest risk should not be considered a central deterrent to election fraud.
+Instead, most post-election protests are strategic gambles gone wrong; the initiation of protest more often reveals the incumbent to be strong enough to manage protest through repression, cooptation, or sideways concessions.
+These findings indicate that, rather than being deterred by protest risk, governments have an incentive to manipulate as much as possible in order to achieve a large official vote-share.
+The evidence suggests that democratic breakthoughs in non-democratic elections are rare and unlikely to be driven by the severity of manipulation---a gloomy picture for supporters of democratization---and that researchers should consider other mechanisms beyond mass protest in their models of electoral manipulation.
+
+\newpage
+##Bibliography

### 2023-01-05 Output/manipulation_protest_draft_v4.Rmd
Updates to theory and control vars (unscheduled election)

@@
 Yet election manipulation is common in non-democracies, while protest is rare, and successful protest rarer still.
-Using count data on protest events from 647 elections from 1990 to 2012, this paper argues that the risk to incumbents has been overstated in prior work.
-Specifically, by treating protest as a binary event, previous research has overestimated the strength of protest movements relative to the incumbent's institutional resources.
-The results show that the level of fraud has no bearing on the number of protests after an election, though protest initiation is more likely in high-fraud elections with low margins of victory.
+Using count data on protest events from 647 elections from 1990 to 2012, this paper argues that the risk to incumbents has been overstated in prior work for two reasons.
+First, information revealed by election fraud is amiguous; second, any such information quickly decays once protests do begin.
+The results show that the level of fraud has no bearing on the number of protests after an election, and is associated with protest initiation only in rare circumstances.
 This pessimistic finding implies that, rather than serving as a constraint, the risk of protest creates a perverse incentive to manipulate as much as possible in order to achieve a large official vote-share."
 ---
-
-TODO: Fix git merge issue on data cleaning file. Plan to stick with massmob version--Show economic marginal effects plots. present plots for both the zero-inflated and count versions of the models.
 
@@
 Despite this consensus in the formal literature, there is relatively little empirical evidence to support the claim that manipulation per se indicates weakness, and thus increases protest risk.
-On the contrary, Simpser [-@Simpser2011] shows support for the alternative view that excessive manipulation can induce cooperation from opposition groups, elites, and other actors by revealing information about regime strength.
+For example, looking at the 2011 election and protest cycle in Russia, Ananyev and Poyker [-@ananyevDictatorsSignalStrength2022] show that, while several measures of fraud were lower in areas where pre-election discontent was high, actual post-election protest was not associated with fraud.^[Though, it should be noted that Lankina and Skovoroda [-@Lankina2017] show that a different measure of fraud was associated with post-election protest in the subsequent presidential election.] In a cross-national study, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e.
+whether opponents are allowed to run, whether the media environment is pro-regime), rather than by the election-day manipulation that usually figures in theories of election integrity and protest.
+
+On the other hand, Simpser [-@Simpser2011] shows support for the alternative view that excessive manipulation can induce cooperation from opposition groups, elites, and other actors by revealing information about regime strength.
 Similarly, the type of manipulation employed can convey information about the ruling party's capacity to mobilize, with falsification providing the least credible information [@Harvey2020].
@@
 In turn, governments should be reluctant to manipulate.
-However, as discussed below, electoral protest is relatively rare, and even in cases where it does occur, most efforts to overturn fraudulent election results through protest have failed [@kalandadze2009electoral].
+However, elections in non-democracies are *generally* risky for incumbents, regardless of how free and fair they may be, as they create an opportunity for opposition collective action [@Knutsen2017].
+Furthermore, electoral protest is relatively rare; even in cases where it does occur, most efforts to overturn fraudulent election results through protest have failed [@kalandadze2009electoral].
 The apparent failure of protest risk to deter election fraud underlines our limited understanding of non-democratic ruling parties' decision-making.
@@
 >H1: Pre-election protest will be associated with lower levels of election-day manipulation.
+
+However, there are reasons to doubt the veracity of Hypothesis 1, as described above.
+Little [-@Little2012] argues that citizens\textemdash knowing the regime is capable of manipulating the results\textemdash discount an expected level of fraud from the results when deciding whether or not to protest.
+The severity of fraud would thus be untethered from protest risk; what matters is citizens' underlying estimate of the regime's strength.
+Of course, election fraud can itself be a signal of strength, albeit an ambiguous one.
+Fraud is not solely a weapon of weak incumbents\textemdash in fact, Seeberg [-@Seeberg2019] finds that higher-capacity states are weakly associated with more election fraud.
+Rather than acting out of desperation, strong ruling parties may engage in election fraud in order to pad their margins [@simpser2013] or because of a bandwagoning effect among front-line agents eager to please an incumbent patron [@rundlett2016deliver].
 
@@
 The initial emergence of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014].
-Given that post-electoral protests are a direct challenge to the regime, toleration is unlikely in these cases.
-That election dates are known in advance creates opportunities for opposition mobilization; anticipating this, regimes can more easily engage in pre-emptive repression [@Truex2019;@Sullivan2016;@Bhasin2013].
+That election dates are known in advance creates opportunities for opposition mobilization; anticipating this, regimes can more easily engage in pre-emptive repression [@Truex2019;@Sullivan2016;@Bhasin2013] \textemdash [-@Knutsen2017] argue that regimes become more effective at these sorts of behaviors over time.
+
 Because the emergence of protest indicates that the regime's pre-emptive efforts have failed, it is not clear a priori how the regime should handle protesters it failed to deter [@Ritter2016].
@@
 
-In summary, though election fraud may be taken as a signal by opposition groups, it is a noisy one: weak and strong governments alike may engage in election manipulation.
-However, once a protest movement begins, the level of election manipulation is no longer a useful signal of incumbent strength.
-Instead, protesters and bystanders can update their estimation of the relative strength of the protest coalition and the incumbent government by observing the size of crowds, the number of protests, and the government's reaction.
+Finally, protest waves can have self-defeating and regime-sustaining effects.
+When unexpected post-election protests occur, trust in the government can increase,  especially among opposition supporters [@Frye2019].
+As the number of protests increases, support for the protesters' aims may fall; this effect is even larger for those who consume pro-government media [@tertytchnayaElectoralProtestsPolitical2020].
+These empirical results are in line with models proposed by Lohmann [-@lohmannDynamicsInformationalCascades1994] and Shadmehr and Bernhardt [-@shadmehrCollectiveActionUncertain2011]; they argue that when some citizens are too predisposed to protest, others become more inclined to stay home\textemdash dooming the first group's efforts to failure.
+These effects give reason to doubt that protest risk serves as a guardrail against election manipulation; rather, post-election protests may burn themselves out while leaving the public more approving of the government.
+
+In summary, fraudulent elections are not likely to pose a significant protest risk to incumbents for two reasons.
+First, protest initiation is unlikely even in fraudulent elections because it is a noisy signal of regime strength.
+Weak and strong governments alike may engage in election manipulation [@simpser2013], and acts of manipulation can be interpreted either as signs of strength or weakness [@Pop-Eleches2015].
+Second, even if a protest movement begins after the election, any useful information provided by the level of election manipulation quickly decays and is replaced by signals from the protests themselves [@lohmannDynamicsInformationalCascades1994].
+Protesters, bystanders, and elites can update their estimation of the relative strength of the protest coalition and the incumbent government by observing the size of crowds, the number of protests, and the government's reaction.
 Since the level of manipulation will no longer serve as an important indicator of regime strength once a challenge is taken to the streets, the number of protests will instead be related to other indicators of opposition and incumbent capacity.
@@
 ```{r, include=F}
-incl.vars <- nelda_massmob %>% dplyr::select(jucon.1lag, legcon.1lag, altinf.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, harass_opp,  median.participants.mm.pre, median.participants.mm.post)
+incl.vars <- nelda_massmob %>% dplyr::select('Judicial independence' = jucon.1lag, 
+                                            'Legislative constraints' = legcon.1lag, 
+                                            'Alternative information' = altinf.1lag, 
+                                            'GDP growth rate' = gdpgro.1lag, 
+                                            'Civil society openness' = osorg.1lag, 
+                                            'Public sector corruption' = pubseccorrup.lag,
+                                            'Economic crisis' = econ.crisis.nelda.1lag,
+                                            'National party organization' = natlpartyorg.lag, 
+                                             'Election fraud' = v2elirreg.inv, 
+                                            'Presidential election' = presidential,
+                                            'Incumbent term-limited' = inc.termlimit.num,
+                                            'Physical integrity index' = physinteg.1lag,
+                                            'Urbanization' = urban.pct, 
+                                            'Rentier state' = rentierstate,
+                                            'Tax state' =  taxstate,
+                                            'Incumbent vote-share' = winner.share.upd, 
+                                            'Median pre-election protest size' = median.participants.mm.pre,
+                                            'Off-schedule election' = unscheduled_election)
 #incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))
+
 ```
 
-```{r, echo=FALSE}
-datasummary_skim(incl.vars)
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+library(kableExtra)
+datasummary_skim(incl.vars, output = "kableExtra")  %>%
+    kable_styling(font_size = 9, full_width = FALSE)
 ```
@@
 If the state is described as primarily relying on resource rents and expropriation, I code it is a *rentier state*.
-Finally, states that rely primarily on taxes on property or economic transactions are coded as *tax states*.
+States that rely primarily on taxes on property or economic transactions are coded as *tax states*.
 This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018].
+Lastly, since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.
 Figure 2 shows summary statistics for all of these variables.
@@
 ```{r manipdv, echo=FALSE}
-m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   + presidential + inc.termlimit.num  +  libdem.1lag + urban.pct + rentierstate + taxstate + n.events.massmob.pre + median.participants.mm.pre  + (1 | stateid), data = nelda_massmob %>% filter(n.events.massmob.pre < 200)) #There are two extreme observations removed by this filter
+m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   + presidential + inc.termlimit.num  +  libdem.1lag + urban.pct + rentierstate + taxstate + n.events.massmob.pre + median.participants.mm.pre + unscheduled_election + (1 | stateid), data = nelda_massmob %>% filter(n.events.massmob.pre < 200)) #There are two extreme observations removed by this filter
 
@@
 
-modelsummary(m1, stars = T,  title = 'Random-intercept multilevel model of election-day fraud')
+modelsummary(m1, stars = T,  title = 'Random-intercept multilevel model of election-day fraud', coef_rename = 
+               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
+                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
+                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election"))
 
@@
 ```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
-model.n.novoteshare.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +     + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+model.n.novoteshare.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +     + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election , data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
 summary(model.n.novoteshare.hurd.mm)
 
-model.n.base.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+model.n.base.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election| jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
 summary(model.n.base.hurd.mm)
 
-model.n.base2.hurd.mm <- hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre |  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+model.n.base2.hurd.mm <- hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election |  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
 summary(model.n.base2.hurd.mm)
 
-model.n.griev.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, dist = "negbin", data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+model.n.griev.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election| jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, dist = "negbin", data = nelda_massmob %>% filter(gdpgro.1lag < 1))
 summary(model.n.griev.hurd.mm)
 
-model.n.griev2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob, dist = "negbin")
+model.n.griev2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election | jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, data = nelda_massmob, dist = "negbin")
 summary(model.n.griev2.hurd.mm)
 
-model.n.griev3.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre  | jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, data = nelda_massmob, dist = "negbin")
+model.n.griev3.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election  | jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, data = nelda_massmob, dist = "negbin")
 summary(model.n.griev3.hurd.mm)
@@
 
-modelsummary::modelsummary(models, coef_omit = "count", gof_omit = "R2", stars = T, longtable=T,  title = 'Binary portion of hurdle models of post-election protest')
+modelsummary::modelsummary(models, coef_omit = "count", gof_omit = "R2", stars = T, longtable=T,  title = 'Binary portion of hurdle models of post-election protest', coef_rename = 
+               c("zero_(Intercept)" = "Intercept", "zero_jucon.1lag" = "Judical independence (lag)", "zero_legcon.1lag" = "Leg. constraints (lag)", 
+                 "zero_gdpgro.1lag" = "GDP growth rate (lag)", "zero_altinf.1lag" = "Alternative info. (lag)", 
+                 "zero_osorg.1lag" = "Civil soc. openness (lag)", "zero_econ.crisis.nelda.1lag1" = "Economic crisis", 
+                 "zero_natlpartyorg.lag" = "National party organization (lag)", "zero_presidential1" = "Presidential election",
+                 "zero_inc.termlimit.num" = "Incumbent term-limited",  
+                 "zero_urban.pct" = "Urbanization", "zero_rentierstate" = "Rentier state", "zero_taxstate" = "Tax state",
+                 "zero_n.events.massmob.pre" = "Number of pre-election protests", "zero_pubseccorrup.lag" = "Public sector corruption (lag)", "zero_v2elirreg.inv" = "Election fraud",
+                 "zero_physinteg.1lag" = "Physical integrity (lag)", "zero_physintlag.sq" = "Physical integrity (lag), squared", "zero_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "zero_unscheduled_election" = "Off-schedule election"))
 
@@
 
-model.noint.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, family = binomial(link = "logit"), data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+model.noint.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda_massmob %>% filter(gdpgro.1lag < 1))
 summary(model.noint.binary)
 
-model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, family = binomial(link = "logit"), data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda_massmob %>% filter(gdpgro.1lag < 1))
 summary(model.margin.binary)
 
-model.binary.griev1 <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, family = binomial(link = "logit"), data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+model.binary.griev1 <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda_massmob %>% filter(gdpgro.1lag < 1))
 summary(model.binary.griev1)
 
-model.binary.griev2 <- glm(any_protest ~  jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob, family = binomial(link = "logit"))
+model.binary.griev2 <- glm(any_protest ~  jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, data = nelda_massmob, family = binomial(link = "logit"))
 summary(model.binary.griev2)
 
-model.binary.griev3 <- glm(any_protest ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, data = nelda_massmob, family = binomial(link = "logit"))
+model.binary.griev3 <- glm(any_protest ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, data = nelda_massmob, family = binomial(link = "logit"))
 summary(model.binary.griev3)
@@
 plot_binary_margin_max <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-2.3, 3.2]"))
+plot_binary_margin_max <- plot_binary_margin_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()
 
 plot_binary_margin_typ <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]"))
+plot_binary_margin_typ <- plot_binary_margin_typ +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "", color = "Election fraud") + theme_bw()
 
@@
 plot_binary_gdp_max <- sjPlot::plot_model(model.binary.griev1, type = "pred", terms = c("gdpgro.1lag [all]", "v2elirreg.inv [-2.3, 3.2]"))
+plot_binary_gdp_max <- plot_binary_gdp_max
 
 plot_binary_gdp_typ <- sjPlot::plot_model(model.binary.griev1, type = "pred", terms = c("gdpgro.1lag [all]", "v2elirreg.inv [-.45, 1.78]"))
+plot_binary_gdp_typ <- plot_binary_gdp_typ
 
@@
 plot_binary_crisis_max <- sjPlot::plot_model(model.binary.griev2, type = "pred", terms = c("v2elirreg.inv [all]", "econ.crisis.nelda.1lag"))
+plot_binary_crisis_max <- plot_binary_crisis_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()
 
 plot_binary_crisis_typ <- sjPlot::plot_model(model.binary.griev2, type = "pred", terms = c("v2elirreg.inv [all]", "econ.crisis.nelda.1lag"))
+plot_binary_crisis_typ <- plot_binary_crisis_typ +  labs(x = "Election fraud", y = "Predicted probability", title = "", color = "Economic crisis") + theme_bw()
 
@@
 plot_binary_corruption_max  <- sjPlot::plot_model(model.binary.griev3, type = "pred", terms = c("pubseccorrup.lag", "v2elirreg.inv [-2.3, 3.2]"))
+plot_binary_corruption_max <- plot_binary_corruption_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()
 
 plot_binary_corruption_typ <- sjPlot::plot_model(model.binary.griev3, type = "pred", terms = c("pubseccorrup.lag", "v2elirreg.inv [-.45, 1.78]"))
+plot_binary_corruption_typ <- plot_binary_corruption_typ +  labs(x = "Public sector corruption", y = "Predicted probability", title = "", color = "Election fraud") + theme_bw()
 ```
@@
 
-```{r n-protests-margin, fig.cap="Fraud, public sector corruption, and number of protest events", echo=FALSE, message=FALSE}
+```{r n-protests-margin, fig.cap="Fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE}
 dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd", "v2elirreg.inv [-.45, 1.64]")) 
 p.n.margin <- plot(dat)
-p.n.margin + coord_cartesian(ylim = c(0, 70)) + theme_bw()
+p.n.margin + coord_cartesian(ylim = c(0, 70)) + labs(x = "Incumbent vote-share", y = "Number of events", title = "Predicted counts of protest events", color = "Election fraud") + theme_bw()
 
@@
 
-```{r n-protests-crisis, fig.cap="Fraud, public sector corruption, and number of protest events", echo=FALSE, message=FALSE}
+```{r n-protests-crisis, fig.cap="Fraud, economic crisis, and number of protest events", echo=FALSE, message=FALSE}
 dat <- ggpredict(model.n.griev2.hurd.mm, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-.45, 1.64]")) 
 p.n.crisis <- plot(dat)
-p.n.crisis + coord_cartesian(ylim = c(0, 75)) + theme_bw()
+p.n.crisis + coord_cartesian(ylim = c(0, 75)) + labs(x = "Economic crisis", y = "Number of events", title = "Predicted counts of protest events", color = "Election fraud") + theme_bw()
 
@@
 p.corr.n.margin <- plot(dat)
-p.corr.n.margin + coord_cartesian(ylim = c(0, 120)) + theme_bw()
+p.corr.n.margin + coord_cartesian(ylim = c(0, 120)) + labs(x = "Public sector corruption", y = "Number of events", title = "Predicted counts of protest events", color = "Election fraud") + theme_bw()
 
@@
 ```{r zero-model-table-count, echo=FALSE}
-modelsummary::modelsummary(models, coef_omit = "zero", gof_omit = "R2", longtable=T, stars = T, title = 'Hurdle models of post-election protest (counts)')
+modelsummary::modelsummary(models, coef_omit = "zero", gof_omit = "R2", longtable=T, stars = T, title = 'Hurdle models of post-election protest (counts)',
+                           coef_rename = 
+               c("count_(Intercept)" = "Intercept", "count_jucon.1lag" = "Judical independence (lag)", "count_legcon.1lag" = "Leg. constraints (lag)", 
+                 "count_gdpgro.1lag" = "GDP growth rate (lag)", "count_altinf.1lag" = "Alternative info. (lag)", 
+                 "count_osorg.1lag" = "Civil soc. openness (lag)", "count_econ.crisis.nelda.1lag1" = "Economic crisis", 
+                 "count_natlpartyorg.lag" = "National party organization (lag)", "count_presidential1" = "Presidential election",
+                 "count_inc.termlimit.num" = "Incumbent term-limited",  
+                 "count_urban.pct" = "Urbanization", "count_rentierstate" = "Rentier state", "count_taxstate" = "Tax state",
+                 "count_n.events.massmob.pre" = "Number of pre-election protests", "count_pubseccorrup.lag" = "Public sector corruption (lag)", "count_v2elirreg.inv" = "Election fraud",
+                 "count_physinteg.1lag" = "Physical integrity (lag)", "count_physintlag.sq" = "Physical integrity (lag), squared", "count_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "count_unscheduled_election" = "Off-schedule election"))
 
@@
 
-The above results have several implications for understanding election manipulation and regime resilience.
+The results indicate that we should be cautious when building and interpreting theories of non-democratic politics that emphasize the risk of mass protest as a trip-wire to be avoided, and add empirical evidence to Przeworski's [-@przeworskiFormalModelsAuthoritarian2022] critique of such models.
+The rarity of post-election protest, especially successful protest, suggest that such events are highly contingent; they hinge on a multitude of factors that go beyond the regime's efforts to bias election outcomes.
+Election fraud is a complex phenomenon with multiple drivers, including resource costs, signaling, information-gathering, principal-agent dynamics, and more; these findings indicate that protest risk is not a major consideration under most circumstances.
+
+This has several implications for understanding election manipulation and regime resilience.
 Most importantly, by treating protest primarily as a binary event rather, the existing literature on electoral protest has overemphasized the risk to incumbents of engaging in electoral manipulation.
@@
 
-These results are robust to different specifications presented in an online appendix.
-In particular, the main results are supported when data from the Electoral Contention and Violence (ECAV) dataset [@Daxecker2019] are used in place of the Mass Mobilization data.
-Other robustness checks include...
+These results are consistent with different specifications presented in an online appendix.
+In particular, the main results are supported when data from the Electoral Contention and Violence (ECAV) dataset [@Daxecker2019] are used in place of the Mass Mobilization data, and when a measure of overall election integrity is used in place of the measure of election fraud.
 
@@
 \newpage
-##Bibliography
+# Bibliography

### 2023-04-19 Output/manipulation_protest_draft_v4.Rmd
Creating new variable for rate and size of protest size increase / decrease

@@
 csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
-header-includes: \usepackage{booktabs}
+indent: true
+header-includes: 
+  - \usepackage{booktabs}
+  - \usepackage{setspace}
 abstract: "The risk of mass protest figures prominently in many models of election manipulation, as a deterrent to tampering with elections.
-Yet election manipulation is common in non-democracies, while protest is rare, and successful protest rarer still.
+Yet election manipulation is common outside liberal democracies, while protest is rare (and successful protest rarer still).
 Using count data on protest events from 647 elections from 1990 to 2012, this paper argues that the risk to incumbents has been overstated in prior work for two reasons.
@@
 ---
+
+\doublespace
 
@@
 On some occasions, like Ukraine's Orange Revolution or the People Power Revolution in the Philippines, these protests can result in the fall of an incumbent and a political opening.
-For this reason, many scholars---especially those working in the formal modeling tradition---have identified the risk that election manipulation will spark protest as a major deterrent to regimes wishing to engineer election outcomes.
+For this reason, many scholars---especially those working in the formal modeling tradition---have identified the risk that election manipulation will spark protest as a major deterrent to regimes working to engineer election outcomes.
 Most often, though, non-democratic states can rely on superior resources, elite solidarity, and repressive capacity to quash electoral protests.
@@
 
+In this paper, I argue that protest risk rarely serves as a deterrent to election manipulation for two reasons.
+First, election fraud is carried out both by strong, cohesive ruling parties, and by weakened parties seeking to cling to power.
+The ambiguity of the electoral environment in a non-democracy makes it difficult for opposition actors to correctly distinguish between the two.
+As a result, we observe a high proportion of 'failed' protest movements, and\textemdash most likely\textemdash instances where weak governments go un-protested.
+Second, if electoral protests do materialize, their success or failure quickly becomes untethered from the severity of manipulation itself.
+Instead, the noisy election-day signal of regime strength is replaced by much clearer signals in the contest with protesters.
+Consequently, there is little penalty for severe manipulation in terms of protest risk.
+
+Testing this proposition requires looking beyond binary measures of protest, and toward measures of its scope.
 Research on the risk posed to non-democratic governments by electoral protest often acknowledges---whether explicitly or implicitly---the importance of broad-based protest campaigns, rather than one-off events [@Bunce2006;@tucker2007;@DAnieri2006;@Carnaghan2016;@bunce2010].
 As Brancati [-@Brancati2016] finds, approximately two-thirds of pro-democracy electoral protests ended within three days, posing little threat to incumbents.
-Only about ten percent of such protests are widespread and long-lasting enough to see incumbents fall.
 This implicit assumption has largely been overlooked in theories and empirical studies of the relationship between election manipulation and protest, however.
@@
 
-Using cross-national data on 647 elections from 1990 to 2012 where incumbents claimed victory, combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Mass Mobilization dataset [@massmob], this paper finds support for the hypothesis that more fraudulent elections are associated with a higher risk that any post-election protest event occurs.
-This is in keeping with the formal-modeling literature, which predicts that overreach in the form of election fraud indicates government weakness, making opposition groups confident of winning a conflict in the streets.
-However, there is no statistically significant association between the level of election fraud and the number of protest events.
+Using cross-national data on 647 elections from 1990 to 2012 where incumbents claimed victory, combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Mass Mobilization dataset [@massmob], this paper finds little support for the hypothesis that more fraudulent elections are associated with a higher risk that any post-election protest event occurs.
+Very highly fraudulent elections are associated with increased risk of protest initiation when incumbent vote-shares are small, in keeping with the formal-modeling literature, which predicts that overreach in the form of election fraud indicates government weakness and makes opposition groups confident of winning a conflict in the streets.
+And, in keeping with the social-grievance literature, fraud is associated with protest initiation in socities with worse public sector corruption.
+However, these effects are largely confined to extreme scenarios, and there is no statistically significant association between the level of election fraud and the number of protest events that occur.
 This suggests a more complex relationship between election fraud and protest, undermining the argument that protest risk is a major deterrent to electoral malfeasance.
@@
 This information, in turn, feeds into strategic decision-making by opposition groups.
-This logic is often seen in formal theory models [@Egorov2021;@chernykh2015third].
+This logic is often seen in formal models [@Egorov2021;@chernykh2015third].
 Several prominent models of electoral manipulation argue that the risk of mass post-election protest helps restrain incumbents who might otherwise prefer larger levels of manipulation [@magaloni2010; @Little2015; @Little2012; @birch2011; @Fearon2011;@Luo2020;@Rozenas2016b].
@@
 Despite this consensus in the formal literature, there is relatively little empirical evidence to support the claim that manipulation per se indicates weakness, and thus increases protest risk.
-For example, looking at the 2011 election and protest cycle in Russia, Ananyev and Poyker [-@ananyevDictatorsSignalStrength2022] show that, while several measures of fraud were lower in areas where pre-election discontent was high, actual post-election protest was not associated with fraud.^[Though, it should be noted that Lankina and Skovoroda [-@Lankina2017] show that a different measure of fraud was associated with post-election protest in the subsequent presidential election.] In a cross-national study, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e.
+For example, looking at the 2011 election and protest cycle in Russia, Ananyev and Poyker [-@ananyevDictatorsSignalStrength2022] show that post-election protest was not associated with the severity of fraud using several measures.^[Though, it should be noted that Lankina and Skovoroda [-@Lankina2017] show that a different measure of fraud was associated with post-election protest in the subsequent presidential election.] In a cross-national study, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e.
 whether opponents are allowed to run, whether the media environment is pro-regime), rather than by the election-day manipulation that usually figures in theories of election integrity and protest.
@@
 On the other hand, Simpser [-@Simpser2011] shows support for the alternative view that excessive manipulation can induce cooperation from opposition groups, elites, and other actors by revealing information about regime strength.
-Similarly, the type of manipulation employed can convey information about the ruling party's capacity to mobilize, with falsification providing the least credible information [@Harvey2020].
+Similarly, the type of manipulation employed can convey information about the ruling party's capacity to mobilize, with costlier forms of manipulation associated with less protest [@Harvey2020].
 In this way, large-scale manipulation can contribute to a ruling party's ability to produce a large, stabilizing majority vote-share [@magaloni2006].
@@
 Without disputing that election manipulation conveys information, a second school of thought has emphasized the ways in which biased elections may predispose individuals to engage in protest [@kuntz2009].
-In this view, election manipulation acts as a grievance shared by individuals who prefer to see the ruling party lose the election---especially when it appears that fraud may have been decisive [@kuntz_more_2009].
+In this view, election manipulation acts as a shared grievance among individuals who prefer to see the ruling party lose---especially when it appears that fraud may have been decisive [@kuntz_more_2009].
 Belief in a stolen election is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007], possibly initiating a protest cascade [@Kuran1991].
@@
 This argument has found support in studies of the relationship between economic grievances, manipulated elections, and protest [@Rod2019;@Brancati2016].
+
+Subjective, grievance-based approaches hinge more on citizens' beliefs about fraud than on the actual level of malfeasance.
 A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b], which suggests that emotions, moral condemnation, or grievances take precedence over strategic calculations for individuals.
@@
 Furthermore, simply being on the losing side in an election is associated with reduced belief that the election was fair [@Cantu2015], and with increased support for electoral protest [@Anderson2006].
+These studies suggest that electoral protest can occur at low or high levels of fraud, contingent on other factors.
 
@@
 In turn, governments should be reluctant to manipulate.
-However, elections in non-democracies are *generally* risky for incumbents, regardless of how free and fair they may be, as they create an opportunity for opposition collective action [@Knutsen2017].
+However, elections in non-democracies are *generally* risky for incumbents, regardless of how free and fair they may be, as they create an opportunity for opposition collective action [@Knutsen2017] and reveal information [@Pop-Eleches2015].
 Furthermore, electoral protest is relatively rare; even in cases where it does occur, most efforts to overturn fraudulent election results through protest have failed [@kalandadze2009electoral].
@@
 
-However, there are reasons to doubt the veracity of Hypothesis 1, as described above.
-Little [-@Little2012] argues that citizens\textemdash knowing the regime is capable of manipulating the results\textemdash discount an expected level of fraud from the results when deciding whether or not to protest.
-The severity of fraud would thus be untethered from protest risk; what matters is citizens' underlying estimate of the regime's strength.
-Of course, election fraud can itself be a signal of strength, albeit an ambiguous one.
-Fraud is not solely a weapon of weak incumbents\textemdash in fact, Seeberg [-@Seeberg2019] finds that higher-capacity states are weakly associated with more election fraud.
-Rather than acting out of desperation, strong ruling parties may engage in election fraud in order to pad their margins [@simpser2013] or because of a bandwagoning effect among front-line agents eager to please an incumbent patron [@rundlett2016deliver].
-
-In addition, the logic of deterrence suggests that increasingly severe election fraud should make post-election more likely.
+In addition, the logic of deterrence suggests that increasingly severe election fraud should make post-election more likely, either unconditionally or when the incumbent's margin of victory is low.
 More extensive election-day manipulation, in these models, is generally taken as a signal of regime unpopularity.
@@
 
-In addition to the deterrence mechanism, previous research has also suggested conditional mechanisms, whereby a fraudulent election is more likely to trigger protest under specific circumstances.
-These are: when the incumbent's official margin of victory is low [@Simpser2011], and when social grievances are high [@Brancati2016;@tucker2007].
+>H3: More severe election fraud is associated with a higher likelihood of protest when winning-party vote-share is low.
 
->H2: More severe election fraud is associated with a higher likelihood of protest when winning-party vote-share is low.
+Previous research has also suggested that a fraudulent election is more likely to trigger protest when social grievances are high [@Brancati2016;@tucker2007].
 
->H3: Post-election protest will be more likely to occur when fraud is severe and economic and social grievances are elevated.
+>H4: Post-election protest will be more likely to occur when fraud is severe and economic and social grievances are elevated.
+
+However, there are reasons to doubt the veracity of the deterrence logic.
+Little [-@Little2012] argues that citizens\textemdash knowing the regime is capable of manipulating the results\textemdash discount an expected level of fraud from the results when deciding whether or not to protest.
+The severity of fraud would thus be untethered from protest risk; what matters is citizens' underlying estimate of the regime's strength.
+Election fraud can itself be a signal of strength, rather than weakness.
+Fraud is not solely a weapon of floundering incumbents\textemdash in fact, Seeberg [-@Seeberg2019] finds that higher-capacity states are weakly associated with more election fraud.
+Rather than acting out of desperation, strong ruling parties may engage in election fraud in order to pad their margins and secure cooperation from other oactors [@simpser2013] or because of a bandwagoning effect among front-line agents eager to please an incumbent patron [@rundlett2016deliver].
+Fraud is thus an ambiguous signal of regime strength, which cannot conclusively guide strategic decisions to protest.
 
@@
 
-Because the emergence of protest indicates that the regime's pre-emptive efforts have failed, it is not clear a priori how the regime should handle protesters it failed to deter [@Ritter2016].
 In many cases, the government doubles down on repression, calling upon the security services to disperse protests, arrest leaders, and perhaps employ harsher methods including disappearance, torture, and execution [@Escriba_Folch2013;@Davenport2007].
@@
 
->H4: Once protests begin, the severity of election fraud will not be correlated with the number of protest events.
+>H5: Once protests begin, the severity of election fraud will not be correlated with the number of protest events.
 
@@
 
+\singlespace
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 ```
+\doublespace
 
@@
 
+\singlespace
+
 ```{r manipdv, echo=FALSE}
@@
 
+\doublespace
+
 The second set of models also challenge the conventional wisdom that more severe electoral manipulation is associated with an increased risk of anti-government protest, except under extraordinary circumstances.
@@
 Finally, there is no significant effect for *economic crisis*, at least under the typical covariate values used to create the marginal effects plot in Figure \@ref(fig:crisis-prob-fig).
+
+\singlespace
 
@@
 
+\doublespace
+
 Turning to the models of number of protest events shown in Table \@ref(tab:zero-model-table-count), Model 2\textemdash which includes no interaction terms\textemdash shows no significant relationship between fraud and protest in the count portions of the hurdle regression.
@@
 \newpage
-# Bibliography
+# References

### 2023-04-19 Scripts/data_cleaning.Rmd
Creating new variable for rate and size of protest size increase / decrease

--- 2022-11-30 Scripts/data_cleaning.Rmd
@@
 library(summarytools)
+library(vdemdata)
 ```
@@
 ecav <- read_xls(here::here("Data", "ECAV datatset_Version 1.2.xls"))
-vdem.data <- read_rds("C:/Users/colharv/Documents/Research projects/Country_Year_V-Dem_Full_others_R_v11.1/Country_Year_V-Dem_Full+others_R_v11.1/V-Dem-CY-Full+Others-v11.1.rds") #Switch between these two for work desktop vs.
+#vdem.data <- read_rds("C:/Users/colharv/Documents/Research projects/Country_Year_V-Dem_Full_others_R_v11.1/Country_Year_V-Dem_Full+others_R_v11.1/V-Dem-CY-Full+Others-v11.1.rds") #Switch between these two for work desktop vs.
 laptop
-#vdem.data <- vdem
+vdem.data <- vdem
 lji <- read.csv(here::here("Data", "LJI-estimates-20140422.csv"))
@@
 
+## Ideas for adding rate of growth and rate of decay to be used as DVs
+
+ - Quick growth after fraud indicates opp believes state is weak
+ - Slow or no growth after fraud indicates opp believes state is strong
+ - Quick decay after peak indicates state is strong / believed strong
+ - Slow decay after peak indicates state is weak / believed weak
+
+The idea here would be:
+  - Replace categorical measures with minimum number (average won't work because top category is top-less)
+  - Replace -99 with...
+  - Filter by election
+  - Group by date and sum
+  - Fit exponential function regression to find coefficient for 'r' (rate of growth)
+  - This becomes the DV for a regression on fraud
+
+Instead, because protests don't always follow the exponential distribution, going to create a simple measure of protest growth / decay intensity.
+
+The idea here would be:
+  - Replace categorical measures with minimum number (average won't work because top category is top-less)
+  - Replace -99 with...
+  - Filter by election
+  - Group by date and sum
+  - Find date of largest protest
+  - Calculate max_size / max_date - 0 and max_size / end_date - max_date
+  - Merge these values into main dataset by election (NeldaID for example)
+  - This becomes the DV for a regression on fraud
+
+```{r}
+# Example
+
+## Set up holding df
+
+dv_df <- data.frame(NeldaID = NA, prot_intensity = NA, decay_intensity = NA)
+
+## Replace categorical with min value, replace -99s
+
+#ecav_rus <- ecav %>% filter(country == "Russia")   # Filter by country
+ecav <- ecav %>% mutate(ParticipantNumber_num_min = ifelse(ParticipantNumber == 1, 2, 
+                                ifelse(ParticipantNumber == 2, 10, 
+                                       ifelse(ParticipantNumber == 3, 100, 
+                                              ifelse(ParticipantNumber == 4, 1000,
+                                                     ifelse(ParticipantNumber == 5, 10000, 10)))))) #Last 10 here hard-codes any missing participant size as 10
+
+## Filter by election
+
+ election_ids <- unique(ecav$NeldaID)
+
+ for(i in 1:length(election_ids)){
+ ecav_temp <- ecav %>% filter(NeldaID == election_ids[i])  #There is a loop here
+
+## Group by date and sum number of participants by date:
+
+ ecav_temp_comb <- ecav_temp %>% mutate(date_col = date(Date)) %>%
+    group_by(date_col) %>%
+     summarize(total = sum(ParticipantNumber_num_min))
+
+ ## Get eleection date
+
+ election_date_temp <- unique(ecav_temp$Electiondate)
+
+ ## Remove protests before election
+
+ ecav_temp_comb <- ecav_temp_comb %>% filter(date_col >= election_date_temp)
+  if (nrow(ecav_temp_comb) == 0) {
+    new_row <- data.frame(election_ids[i], NA, NA)
+    dv_df[i,] <- new_row
+    print(paste0("No post-election protests for obseveration ", i, ", election ID", election_ids[i], ","))
+    next}  #End loop if zero post-election protests
+
+ ## Get date of largest protest
+
+ max_data <-  ecav_temp_comb %>% filter(total == max(total, na.rm = T))
+
+  if (nrow(max_data) > 1) {
+    prot_intensity <- max_data$total[1] / (1 + as.numeric(ymd(max_data$date_col[1]) - ymd(election_date_temp))) #[1] gets the first peak
+    last_peak <- tail(max_data, n = 1)
+    decay_intensity <-  last_peak$total / (1 + as.numeric(max(ecav_temp_comb$date_col) - (ymd(last_peak$date_col))))
+    new_row <- data.frame(election_ids[i], prot_intensity, decay_intensity)
+  #new_row <- new_row %>% rename(NeldaID = election_ids.i.)
+    dv_df[i,] <- new_row
+    print(paste0("Election observation ", i, " completed. (Note: multiple peaks."))
+    next
+  }
+
+ ## Calculate DVs
+
+ prot_intensity <- max_data$total / (1 + as.numeric(ymd(max_data$date_col) - ymd(election_date_temp)))
+ decay_intensity <-  max_data$total / (1 + as.numeric(max(ecav_temp_comb$date_col) - (ymd(max_data$date_col)))) #The +1 is a kludge to prevent Inf values
+
+ ##Add to holding df
+
+  new_row <- data.frame(election_ids[i], prot_intensity, decay_intensity)
+  #new_row <- new_row %>% rename(NeldaID = election_ids.i.)
+  dv_df[i,] <- new_row
+  print(paste0("Election observation ", i, " completed."))
+
+ }
+
+  ## Merge dv_df back into main dataset
+
+nelda2.sub.b <- nelda2.sub %>% left_join( dv_df, by = "NeldaID")
+
+```
+
 ## Writing
@@
 ```{r}
-write.csv(nelda2.sub, here::here("Data", "nelda2_sub.csv"))
+write.csv(nelda2.sub, here::here("Data", "nelda2_sub_2023-04-19.csv"))
 ```
+
+```{r}  
+  ## Plotting for visualizing election protest size
+
+  ecav_rus_comb <- ecav_rus_comb %>% filter(date_col > "2012-03-04")
+ ecav_rus_comb <- ecav_rus_comb %>% mutate(date_index = date(date_col) - date("2012-03-04"))
+
+ ggplot(ecav_rus_comb, aes(x=date_col, y = total)) +
+     geom_col()
+
+```

### 2023-05-30 Output/Submission files/manuscript_anon.Rmd
Submission round two updates

--- 2023-04-19 Output/manipulation_protest_draft_v4.Rmd
@@
 ---
-title: "Draft v4: A missing guardrail---The minimal effect of election fraud on protest intensity in non-democracies"
-author: "Cole J. Harvey (Oklahoma State University)"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
 output:
@@
   - \usepackage{setspace}
-abstract: "The risk of mass protest figures prominently in many models of election manipulation, as a deterrent to tampering with elections.
-Yet election manipulation is common outside liberal democracies, while protest is rare (and successful protest rarer still).
-Using count data on protest events from 647 elections from 1990 to 2012, this paper argues that the risk to incumbents has been overstated in prior work for two reasons.
-First, information revealed by election fraud is amiguous; second, any such information quickly decays once protests do begin.
-The results show that the level of fraud has no bearing on the number of protests after an election, and is associated with protest initiation only in rare circumstances.
-This pessimistic finding implies that, rather than serving as a constraint, the risk of protest creates a perverse incentive to manipulate as much as possible in order to achieve a large official vote-share."
+  - \usepackage{float}
 ---
@@
 First, election fraud is carried out both by strong, cohesive ruling parties, and by weakened parties seeking to cling to power.
-The ambiguity of the electoral environment in a non-democracy makes it difficult for opposition actors to correctly distinguish between the two.
+The ambiguity of the electoral environment outside liberal democracies, where political and civil liberties may not be fully protected, makes it difficult for opposition actors to correctly distinguish between the two.
 As a result, we observe a high proportion of 'failed' protest movements, and\textemdash most likely\textemdash instances where weak governments go un-protested.
@@
 Very highly fraudulent elections are associated with increased risk of protest initiation when incumbent vote-shares are small, in keeping with the formal-modeling literature, which predicts that overreach in the form of election fraud indicates government weakness and makes opposition groups confident of winning a conflict in the streets.
-And, in keeping with the social-grievance literature, fraud is associated with protest initiation in socities with worse public sector corruption.
+And, in keeping with the social-grievance literature, fraud is associated with protest initiation in societies with worse public sector corruption.
 However, these effects are largely confined to extreme scenarios, and there is no statistically significant association between the level of election fraud and the number of protest events that occur.
@@
 These results help explain both the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory.
+They also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations.
 
@@
 
-Broadly speaking, there are thought to be two mechanisms by which election manipulation leads to protest.
+Broadly speaking, prior work has considered two major explanations for the severity of manipulated elections: the costs and benefits of manipulation to the ruling party and government (a demand-side framework) and the costs and benefits for the front-line actors who do the work of manipulating elections (a supply-side framework).
+The risk of costly protest is the central, though not sole^[In some studies, like [@birch2011], the costs of manipulation are attributed to a more diffuse loss of legitimacy; such costs **may** manifiest as protest, but may also manifest as a reduction in general cooperation with the regime.
+However, it is worth noting the argument of [@gehlbach2015electoral] that election manipulation can **induce** cooperative behavior from some agents], factor which affects the demand-side cost of manipulation in most models.
+In many formal models, the decision to protest or consent to the election is the play available to citizens or opposition parties; protesting imposes a cost on the regime and/or carries a risk that the incumbent is overthrown.
+This risk is understood to restrain incumbents' preferred level of manipulation.
+This logic has been applied to democracies [@fearon_self-enforcing_2011;@millerRepublicIfYou2021], non-democracies [@Egorov2021], and any regime where elections are held but may be manipulated [@Little2015;@Luo2020].
+
+There are thought to be two mechanisms by which election manipulation leads to protest, and thereby restrains manipulation.
 The first is through revealing information about the strength of the government or ruling party.
@@
 The apparent failure of protest risk to deter election fraud underlines our limited understanding of non-democratic ruling parties' decision-making.
-Does the extent of fraud, in fact, make more numerous (and thus more costly) protest more likely?
+Does the severity of fraud, in fact, make more numerous (and thus more costly) protest more likely?
+If not, scholarly attention might turn more profitably toward supply-side explanations for manipulation, such as the organizational capacity of the state [@Simpser2011], the popularity of the government [@rundlett2016deliver], or the risks faced by front-line agents [@Harvey2022].
 
@@
 
+Additionally, in-depth investigations of the causes of the Color Revolutions (in particular) have emphasized the role of opposition agency and learning in staging post-election protest.
+In this view, the likelihood of mass electoral protest is increased by deliberate efforts by opposition parties to form a unified opposition coalition, to raise popular expectation of a stolen election, to prepare logistics, and to actively monitor the vote and publicize discrepancies [@bunce2010].
+Others have noted structural factors like the importance of ties to democratic Western states and the size and strength of the ruling party and security apparatus in influencing the course of the Color Revolutions [@LucanWay2008].
+In either case, these factors create an incentive for opposition protest whether electoral fraud is minor or severe\textemdash a kind of case of 'crying wolf'\textemdash an an outcome that is also posited in some of the formal modeling literature [@Little2015].
+As a result, it is likely that indicators of regime weakness or opposition strength will be correlated with post-election protest, while the severity of fraud will not.
+
 With some exceptions, such as the study by Lankina and Skovoroda [-@Lankina2017], most research on election fraud and protest treats the latter as a binary outcome.
@@
 The initial emergence of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014].
-That election dates are known in advance creates opportunities for opposition mobilization; anticipating this, regimes can more easily engage in pre-emptive repression [@Truex2019;@Sullivan2016;@Bhasin2013] \textemdash [-@Knutsen2017] argue that regimes become more effective at these sorts of behaviors over time.
+That election dates are known in advance creates opportunities for opposition mobilization; anticipating this, regimes can more easily engage in pre-emptive repression [@Truex2019;@Sullivan2016;@Bhasin2013] \textemdash Knutsen et al [-@Knutsen2017] argue that regimes become more effective at these sorts of behaviors over time.
 
@@
 While the regime may have its own splits---between reformists and hardliners, for example---the emergence of protest can act to solidify support for the ruling party among its supporters.
+This can be especially true non-democracies where elites have strong incentives to hold to power, in the form of access to rents and preservation of immunity for prosecution   [@baturoDemocracyDictatorshipTerm2014].
 
@@
 I further limit the sample to regimes coded as electoral autocracies or electoral democracies in V-Dem.
-This excludes closed autocracies (where no national elections are held) and liberal democracies (where electoral integrity and rule of law are both high).\footnote{The specific variable used is 'v2x_regime' from V-Dem.
+This excludes closed autocracies (where no national elections are held) and liberal democracies (where electoral integrity and rule of law are both high).^[The specific variable used is 'v2x_regime' from V-Dem.
 This variable codes an electoral democracy, in part, if the country scores above a 2 on the measure of free and fair elections ('v2elfrfair_osp').
-This still permits election fraud to be detected in the electoral democracies; v2elfrfair can take on a value of 3 when "deficiencies and some degree of fraud and irregularities" occur.} Finally, the beginning of the date range is limited by the availability of data on protest from the Mass Mobilization dataset, as discussed below, which begins in 1990.
+This still permits election fraud to be detected in the electoral democracies; v2elfrfair can take on a value of 3 when "deficiencies and some degree of fraud and irregularities" occur.] Finally, the beginning of the date range is limited by the availability of data on protest from the Mass Mobilization dataset, as discussed below, which begins in 1990.
 The inclusion of a control variable from the Latent Judicial Independence dataset [@Linzer2015] limits the end of the time range to 2012.
@@
 Figure 2 shows summary statistics for all of these variables.
+To address concerns about multicollinearity, I present models in the appendix (page 25) that exclude control variables associated with pre-election protest.
 
@@
 
+\newpage
+
 \singlespace
@@
 ```{r manipdv, echo=FALSE}
-m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   + presidential + inc.termlimit.num  +  libdem.1lag + urban.pct + rentierstate + taxstate + n.events.massmob.pre + median.participants.mm.pre + unscheduled_election + (1 | stateid), data = nelda_massmob %>% filter(n.events.massmob.pre < 200)) #There are two extreme observations removed by this filter
+m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   + presidential + inc.termlimit.num  +  libdem.1lag + urban.pct + rentierstate + taxstate + n.events.massmob.pre +  physinteg.1lag + physintlag.sq  + median.participants.mm.pre + unscheduled_election + (1 | stateid), data = nelda_massmob %>% filter(n.events.massmob.pre < 200)) #There are two extreme observations removed by this filter
 
@@
 
-modelsummary(m1, stars = T,  title = 'Random-intercept multilevel model of election-day fraud', coef_rename = 
+modelsummary(m1, output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Random-intercept multilevel model of election-day fraud', coef_rename = 
                c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
@@
                  "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
-                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election"))
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election")) %>% kable_styling(latex_options = "HOLD_position")
 
 ```
+
+\newpage
 
@@
 Finally, there is no significant effect for *economic crisis*, at least under the typical covariate values used to create the marginal effects plot in Figure \@ref(fig:crisis-prob-fig).
+
+\newpage
 
@@
 
-```{r vote-share-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE}
+```{r vote-share-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
 plot.combined.prob <- cowplot::plot_grid(plot_binary_margin_max, plot_binary_margin_typ, ncol=1)
@@
 
-```{r crisis-prob-fig, fig.cap="Fraud, economic crisis, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE}
+```{r crisis-prob-fig, fig.cap="Fraud, economic crisis, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
 plot.combined.crisis.prob <- cowplot::plot_grid(plot_binary_crisis_max, plot_binary_crisis_typ, ncol=1)
@@
 
-```{r corrupt-prob-fig, fig.cap="Fraud, public sector corruption, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE}
+```{r corrupt-prob-fig, fig.cap="Fraud, public sector corruption, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
 plot.combined.corrupt.prob <- cowplot::plot_grid(plot_binary_corruption_max, plot_binary_corruption_typ, ncol=1)
@@
 
+\newpage
+
 ## Election-day fraud and number of post-election protests
+
+\doublespace
+
+Turning to the models of number of protest events shown in Table \@ref(tab:zero-model-table-count), Model 2\textemdash which includes no interaction terms\textemdash shows no significant relationship between fraud and protest in the count portions of the hurdle regression.
+Similarly, none of the three models that include interaction terms yield a significant relationship between fraud, the conditional variable, and number of protest events.
+Model 3 shows no significant interaction effect for fraud and winning vote-share on quantity of protest.
+Figure \@ref(fig:n-protests-margin) shows the marginal effect of election fraud at varying levels of winning vote-share, holding other variables constant at their means (continuous variables) or modes (categorical variables).
+The high and low levels of fraud in the plot represent one standard deviation above and below the mean, respectively.
+As the figure shows, the predicted number of post-election protests is somewhat higher in the high fraud condition when winning margin is low, but this relationship is not statistically significant.
+
+Neither economic crises nor public sector corruption is associated with a significant marginal effect for typical levels of election fraud on protest quantity, as Figures \@ref(fig:n-protests-crisis) and \@ref(fig:n-protests-corrupt) show.\footnote{Figures demonstrating minimum and maximum values for fraud are shown in the appendix (page 5); they also indicate no significant relationship.} In fact, more fraudulent elections appear to be associated with *fewer* incidents of protest during periods of economic crisis, though this distinction is not statistically significant at the p < .05 level.
+
+Altogether, the argument that election fraud increases the risk of protest initiation is mixed.
+Hypothesis 1 is weakly supported: pre-election protest is associated with reduced election-day fraud, but the substantive effect of this relationship is minimal.
+Hypothesis 2 is also weakly supported; there is a significant and sizable association between election fraud and protest initiation when the ruling party's winning margin is low, but only at extremely low and high values of fraud.
+Support for Hypothesis 3 is mixed\textemdash fraud severity is not associated with increased protest risk during economic crises, though there is a significant and sizable increase in the marginal effect as public sector corruption increases.
+Election fraud appears to be associated with a higher risk of protest initiation, but generally in unusual scenarios\textemdash when comparing a highly fraudulent to a largely clean elections when the incumbent's margin of victory is quite low, for example\textemdash or with meager substantive effect.
+
+However, even this mixed finding likely presents too optimistic a picture for the protest-deterrent model of election.
+fraud.
+The count results from the hurdle regressions show no effect for election fraud across any of the models, alone or in interaction.
+Elections that were more fraudulent are not associated with more numerous protest.
+Even when protests are initiated\textemdash a relatively rare event in the first place\textemdash the size of a protest wave is not influenced by the degree of fraud perpetrated by the ruling party.
+In turn, this suggests two pessimistic interpretations of fraud-protest dynamic.
+First, protest may simply be less damaging to ruling parties than is regularly assumed in theories of election manipulation.
+Faced with a nascent protest movement, ruling parties are generally able to contain, co-opt, or repress them\textemdash regardless of the severity of election fraud.
+An even more troubling implication would suggest that the protest-manipulation logic, rather than serving as a deterrent, creates perverse incentives to manipulate as much as possible.
+If ruling parties do indeed seek to avoid protest, but manipulating heavily does not come with a cost in the number of post-election protests, ruling parties have an incentive fraudulently inflate their official margins of victory in order to reduce the risk of protest initiation.
 
@@
 
+\singlespace
 ```{r zero-model-table-count, echo=FALSE}
@@
 
-Turning to the models of number of protest events shown in Table \@ref(tab:zero-model-table-count), Model 2\textemdash which includes no interaction terms\textemdash shows no significant relationship between fraud and protest in the count portions of the hurdle regression.
-Similarly, none of the three models that include interaction terms yield a significant relationship between fraud, the conditional variable, and number of protest events.
-Model 3 shows no significant interaction effect for fraud and winning vote-share on quantity of protest.
-Figure \@ref(fig:n-protests-margin) shows the marginal effect of election fraud at varying levels of winning vote-share, holding other variables constant at their means (continuous variables) or modes (categorical variables).
-The high and low levels of fraud in the plot represent one standard deviation above and below the mean, respectively.
-As the figure shows, the predicted number of post-election protests is somewhat higher in the high fraud condition when winning margin is low, but this relationship is not statistically significant.
-
-Neither economic crises nor public sector corruption is associated with a significant marginal effect for typical levels of election fraud on protest quantity, as Figures \@ref(fig:n-protests-crisis) and \@ref(fig:n-protests-corrupt) show.\footnote{Figures demonstrating minimum and maximum values for fraud are shown in the appendix; they also indicate no significant relationship}.
-In fact, more fraudulent elections appear to be associated with *fewer* incidents of protest during periods of economic crisis, though this distinction is not statistically significant at the p < .05 level.
-
-Altogether, the argument that election fraud increases the risk of protest initiation is mixed.
-Hypothesis 1 is weakly supported: pre-election protest is associated with reduced election-day fraud, but the substantive effect of this relationship is minimal.
-Hypothesis 2 is also weakly supported; there is a significant and sizable association between election fraud and protest initiation when the ruling party's winning margin is low, but only at extremely low and high values of fraud.
-Support for Hypothesis 3 is mixed\textemdash fraud severity is not associated with increased protest risk during economic crises, though there is a significant and sizable increase in the marginal effect as public sector corruption increases.
-Election fraud appears to be associated with a higher risk of protest initiation, but generally in unusual scenarios\textemdash when comparing a highly fraudulent to a largely clean elections when the incumbent's margin of victory is quite low, for example\textemdash or with meager substantive effect.
-
-However, even this mixed finding likely presents too optimistic a picture for the protest-deterrent model of election.
-fraud.
-The count results from the hurdle regressions show no effect for election fraud across any of the models, alone or in interaction.
-Elections that were more fraudulent are not associated with more numerous protest.
-Even when protests are initiated\textemdash a relatively rare event in the first place\textemdash the size of a protest wave is not influenced by the degree of fraud perpetrated by the ruling party.
-In turn, this suggests two pessimistic interpretations of fraud-protest dynamic.
-First, protest may simply be less damaging to ruling parties than is regularly assumed in theories of election manipulation.
-Faced with a nascent protest movement, ruling parties are generally able to contain, co-opt, or repress them\textemdash regardless of the severity of election fraud.
-An even more troubling implication would suggest that the protest-manipulation logic, rather than serving as a deterrent, creates perverse incentives to manipulate as much as possible.
-If ruling parties do indeed seek to avoid protest, but manipulating heavily does not come with a cost in the number of post-election protests, ruling parties have an incentive fraudulently inflate their official margins of victory in order to reduce the risk of protest initiation.
-
 # Discussion
 
-Altogether, the findings challenge the idea that the expectation of protest will deter election-manipulation by ruling parties.
+Altogether, the findings challenge the idea that the fear of protest will deter election-manipulation by ruling parties.
 To the extent that fraud is a signal of incumbent strength, it is a noisy and unreliable one.
-When an incumbent party wins by a narrow margin in a highly fraudulent election, the likelihood of protest initiation is substantially higher than when it wins more cleanly: roughly a 50\% chance in the former case versus 5\% in the latter as shown in Figure 2.
+When an incumbent party wins by a narrow margin in a highly fraudulent election, the likelihood of protest initiation is substantially higher than when it wins more cleanly: roughly a 50 to 60\% chance in the former case versus 5 to 10\% in the latter as shown in Figure 2.
 However, since even strong incumbents commit fraud, the decision to take to the streets is often a losing gamble\textemdash anticipating a weak incumbent, they instead find one that is capable of repressing, coopting, or conciliating an emerging protest movement.
@@
 The results indicate that we should be cautious when building and interpreting theories of non-democratic politics that emphasize the risk of mass protest as a trip-wire to be avoided, and add empirical evidence to Przeworski's [-@przeworskiFormalModelsAuthoritarian2022] critique of such models.
-The rarity of post-election protest, especially successful protest, suggest that such events are highly contingent; they hinge on a multitude of factors that go beyond the regime's efforts to bias election outcomes.
-Election fraud is a complex phenomenon with multiple drivers, including resource costs, signaling, information-gathering, principal-agent dynamics, and more; these findings indicate that protest risk is not a major consideration under most circumstances.
+The rarity of post-election protest, especially successful protest, suggests that such events are highly contingent; they hinge on a multitude of factors that go beyond the regime's efforts to bias election outcomes.
+Election fraud is a complex phenomenon with multiple drivers, including resource costs, signaling, information-gathering, principal-agent dynamics, and more; these findings indicate that protest risk is not a significant factor under most circumstances.
 
 This has several implications for understanding election manipulation and regime resilience.
-Most importantly, by treating protest primarily as a binary event rather, the existing literature on electoral protest has overemphasized the risk to incumbents of engaging in electoral manipulation.
-This results presented here show that most electoral protests are narrow, and that fraud has no bearing on how numerous protest events are.
-Moreover, contrary to some prior research, the risk of protest does not appear to be heightened by poor economic conditions [@Brancati2014a] or intensive corruption in the public sector.
+By overlooking the poor quality of election fraud as a signal, or by treating protest primarily as a binary event, the existing literature on electoral protest has overemphasized the risk to incumbents of engaging in electoral manipulation.
+This results presented here show that fraud has no bearing on how numerous protest events are, and only little influence on protest initiation.
+While pre-election protest is associated with reduced fraud, the substantive size of this relationship is very small.
+Moreover, contrary to some prior research, the risk of a large protest wave does not appear to be heightened by poor economic conditions [@Brancati2014a] or intensive corruption in the public sector.
 As a result, circumstances that are to some extent outside of leaders' control, like a worsening macroeconomic climate, do not appear to make engaging in fraud any riskier.
 Instead, the conditions that are correlated with more damaging protest are generally institutional factors, which are either more amenable to control or can be anticipated by non-democratic incumbents.
-Legislative constraints, an important factor for the duration and number of protests, can be adjusted through institutional reforms and cooptation of the opposition, over time.
-Incumbents may attempt to manage term limit through constitutional reforms, where possible, or through the appointment of (presumably) loyal successors.
+Legislative constraints, an important predictor of the number of protests in all four models, can be adjusted through institutional reforms and cooptation of the opposition over time.
 In other words, election fraud only has a minor cost in terms of protest risk, a risk which can be further blunted to the extent that elections can often be held on incumbents' terms.
@@
 Since the severity of fraud has no apparent bearing on whether or not a major protest wave will emerge as a serious threat to the regime, as shown in Figure 3, government have an incentive to manipulate as much as possible to secure a large winning vote-share.
-This perverse incentive undermines the case that democracy, and fair elections generally, are a self-enforcing equilibrium backed up by the risk of citizen protest [@fearon_self-enforcing_2011].
+This perverse incentive undermines the case that free and fair elections in an electoral democracy represent a self-enforcing equilibrium backed up by the risk of citizen protest [@fearon_self-enforcing_2011], or that risk of popular mobilization can substantially moderate election manipulation in non-democracies.
+In effect, it is the existence of an election itself [@lucardiStrengthExpectationElections2019], and the opportunity it offers for opposition mobilization [@howard2006], that creates protest risk for incumbents, giving them a relatively free hand to manipulate to the extent they can recruit agents to do the work.
+
+This study also suggests implications for the study of election manipulation.
+First, it is clear that electoral protests charged by allegations of manipulation do occur at times.
+While the main argument of this paper is that the overall level of fraud does not predict these protests, there may still be connections between manipulation and protest to uncover.
+For example, while fraud was used as the main variable in this study, it could be that other forms of manipulation--perhaps intimidation or violence--are associated with higher risk of escalating protest.
+Perhaps some forms of electoral manipulation are particularly offensive to certain groups of people\textemdash young people, highly educated people, people with ideological commitments to democracy, etc.
+To the extent that countries vary according to the size of such populations, they may vary in their underlying protest risk conditional on the type of manipulation employed.
+Second, the study implies that supply-side theories of electoral manipulation\textemdash such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]\textemdash may be more important than protest risk for understanding election integrity, which may drive future research as well as policy interventions.
+Third, and relatedly, it undermines some current understandings of the role of election monitors, courts, and other third parties in upholding election integrity.
+As noted earlier, prior studies argue that these actors restrain manipulation by revealing information about fraud and intensifying protest risk.
+But if protest risk is largely untethered from the severity of fraud, it is likely that other mechanisms are at work.
 
@@
 That is, if incumbents choose a level of manipulation calculated to avoid large-scale protest---sometimes high and sometimes low, conditional on contextual factors---no correlation between fraud and protest quantity would be found.
-This paper does not dispute that decisions about manipulation and protest may be endogenous to one another, but it does argue that the resulting equilibrium favors governments.
+This paper does not dispute that decisions about manipulation and protest may be endogenous to one another, but it does argue that the resulting equilibrium is weighted heavily in favor of governments.
 In the dataset used here, only ten elections in four countries are coded as having an *intentional irregularities* score of less than -2 (indicating extremely fraud-free elections).
@@
 The frequency of fraud, the rarity of protest, and the finding that fraud severity is not correlated with the scope of protest events all point to the same conclusion: protest risk is not a substantial constraint on election fraud in non-democracies.
-Rather, the constraints on manipulation are more likely to be driven by factors like resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022].
 
 These results are consistent with different specifications presented in an online appendix.
-In particular, the main results are supported when data from the Electoral Contention and Violence (ECAV) dataset [@Daxecker2019] are used in place of the Mass Mobilization data, and when a measure of overall election integrity is used in place of the measure of election fraud.
+In particular, the main results are supported when data from the Electoral Contention and Violence (ECAV) dataset [@Daxecker2019] are used in place of the Mass Mobilization data (appendix page 1), and when a measure of overall election integrity is used in place of the measure of election fraud (appendix page 15).
 
@@
 
-Protest risk is often considered to be a major risk to incumbent governments that work to undermine election integrity.
-This risk has a central role in several prominent models of electoral manipulation, serving to deter manipulation and to incentivize the toleration of election monitors.
-However, empirical research on the subject has largely treated protest as a binary variable, which obscures the degree to which post-election protest imposes a real risk on incumbents in non-democracies.
-One protest may be easily contained, while dozens may pose a more significant risk to the regime.
+Protest risk is often considered to be a major risk to incumbent governments aiming to rig elections.
+This risk has a central role in several prominent models of electoral manipulation, serving to deter manipulation and to incentivize the toleration of election monitors and other restraints on the ruling party.
+However, I argue that such models overlook the ambiguity inherent in election manipulation efforts, as well as the quick decay of even such limited information once protests begin.
+Empirical research on the subject has largely treated protest as a binary variable, an approach which obscures the distinction between small, one-off protests and much more costly protest waves.
 
-This paper addresses this concern by testing the relationship between election fraud, the risk that any protest occurs, and the number of protest events that occur after an election.
-Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 647 elections to show that---while there is an elevated risk of some protest occurring in a fraudulent election with narrow margins of victory---election fraud has no significant relationship with the number of protests that occur in a protest wave.
+This paper addresses these concerns by testing the relationship between election fraud, the risk that any protest occurs, and the number of protest events that occur after an election.
+Drawing on data from the V-Dem, NELDA, and Mass Mobilization datasets, it uses cross-national data on 647 elections to show that---while there is an elevated risk of some protest occurring in a fraudulent election with narrow margins of victory or in highly corrupt societies---election fraud has no significant relationship with the number of protests that occur in a protest wave.
 These results imply that protest risk should not be considered a central deterrent to election fraud.
-Instead, most post-election protests are strategic gambles gone wrong; the initiation of protest more often reveals the incumbent to be strong enough to manage protest through repression, cooptation, or sideways concessions.
+Instead, most fraudulent elections go un-protested, and most post-election protests are strategic gambles gone wrong.
+The initiation of protest more often reveals the incumbent to be strong enough to manage protest through repression, cooptation, or other concessions.
 These findings indicate that, rather than being deterred by protest risk, governments have an incentive to manipulate as much as possible in order to achieve a large official vote-share.
-The evidence suggests that democratic breakthoughs in non-democratic elections are rare and unlikely to be driven by the severity of manipulation---a gloomy picture for supporters of democratization---and that researchers should consider other mechanisms beyond mass protest in their models of electoral manipulation.
+The evidence suggests that democratic breakthroughs in non-democratic elections are rare and unlikely to be driven by the severity of manipulation---a gloomy picture for supporters of democratization---and that researchers should consider other mechanisms beyond mass protest in their models of electoral manipulation.
 

### 2023-05-30 Output/Submission files/manuscript_anon_v2.Rmd
Submission round two updates

--- 2023-05-30 Output/Submission files/manuscript_anon.Rmd
@@
   - \usepackage{float}
+  - \usepackage{caption}
+  - \captionsetup[table]{labelformat=empty}
 ---
@@
 Broadly speaking, prior work has considered two major explanations for the severity of manipulated elections: the costs and benefits of manipulation to the ruling party and government (a demand-side framework) and the costs and benefits for the front-line actors who do the work of manipulating elections (a supply-side framework).
-The risk of costly protest is the central, though not sole^[In some studies, like [@birch2011], the costs of manipulation are attributed to a more diffuse loss of legitimacy; such costs **may** manifiest as protest, but may also manifest as a reduction in general cooperation with the regime.
+The risk of costly protest is the central, though not sole^[In some studies, like [@birch2011], the costs of manipulation are attributed to a more diffuse loss of legitimacy; such costs **may** manifest as protest, but may also manifest as a reduction in general cooperation with the regime.
 However, it is worth noting the argument of [@gehlbach2015electoral] that election manipulation can **induce** cooperative behavior from some agents], factor which affects the demand-side cost of manipulation in most models.
-In many formal models, the decision to protest or consent to the election is the play available to citizens or opposition parties; protesting imposes a cost on the regime and/or carries a risk that the incumbent is overthrown.
-This risk is understood to restrain incumbents' preferred level of manipulation.
-This logic has been applied to democracies [@fearon_self-enforcing_2011;@millerRepublicIfYou2021], non-democracies [@Egorov2021], and any regime where elections are held but may be manipulated [@Little2015;@Luo2020].
+This insight can be found, in related forms, in both the literature on self-enforcing democracy and on the conduct of non-democratic elections.
+In the former literature, generally free and fair elections are thought to be a self-enforcing equilibrium through the threat of mass protest by wronged groups if the incumbent government rigs the results.
+In many formal models, including those discussed below, the decision to protest or consent to the election is the play available to citizens or opposition parties; protesting imposes a cost on the regime and/or carries a risk that the incumbent is overthrown.
+
+In the major formulations of this argument, such as Fearon  [-@fearon_self-enforcing_2011] and Przeworski [-@przeworskiSelfenforcingDemocracy2006], 'democracy' is understood in procedural terms.
+For Przeworski, democracy requires that "authorization to rule is derived from elections" which designate winners and losers according to a set of rules [@przeworskiSelfenforcingDemocracy2006, p.
+312].
+Fearon adopts a similar definition for electoral democracy [@fearon_self-enforcing_2011, p.
+1662].
+Such arrangements constitute the large majority of regimes today, ranging from the electoral authoritarian to the liberal democratic.
+While many previous works argue that the risk of protest can mitigate election manipulation in regimes where elections are held\textemdash including established democracies [@millerRepublicIfYou2021], electoral authoritarian regimes [@Egorov2021], and any regime where elections are held but may be manipulated [@Little2015;@Luo2020]\textemdash, this body of work has yet to grapple with the empirical finding that election integrity is relatively poor across these regimes outside of a narrow club of 'consolidated' democracies.
 
@@
 Building on this logic, researchers have proposed that third-party institutions---like election monitors [@Little2012;@Hyde2014; @daxecker2012], civil society groups [@birch2017getting], and courts [@Chernykh2015]---can reduce manipulation and increase protest risk by revealing information about electoral malfeasance.
+Of course, electoral manipulation is not understood to be the sole cause of protest\textemdash protests occur for all sorts of reasons and across time\textemdash but the literature has drawn a close connection between the risk of specifically electoral protest and the conduct of elections.
 
@@
 
-On the other hand, Simpser [-@Simpser2011] shows support for the alternative view that excessive manipulation can induce cooperation from opposition groups, elites, and other actors by revealing information about regime strength.
+The literature on protest as a deterrent to manipulation also often overlooks the functions that elections, even highly controlled elections, may serve in non-democratic settings.
+For instance, Simpser [-@Simpser2011] shows that high levels of manipulation can actually induce cooperation from opposition groups, elites, and other actors by revealing information about regime strength.
 Similarly, the type of manipulation employed can convey information about the ruling party's capacity to mobilize, with costlier forms of manipulation associated with less protest [@Harvey2020].
@@
 Similarly, Luo and Rozenas [-@Luo2020] posit that ex post fraud is more likely to result in protest than ex ante manipulation, though this stands in contrast with the finding that parties are more likely to protest elections after ex ante changes to the rules [@chernykh2014].
-In sum, it is likely that manipulation *may* signal regime weakness, but empirical works suggests that this is more likely when margins of victory are narrow and when falsification is a major tool.
+In sum, it is likely that manipulation *may* signal regime weakness, but empirical works suggests that this is more likely when margins of victory are narrow and when falsification is a major tool; by contrast, other work suggests that manipulation may at times signal strength.
 
@@
 
-Subjective, grievance-based approaches hinge more on citizens' beliefs about fraud than on the actual level of malfeasance.
+The grievance-based approach poses a challenge for the deterrence logic, in that protests hinge more on citizens' subjective beliefs about fraud than on the actual level of malfeasance.
 A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b], which suggests that emotions, moral condemnation, or grievances take precedence over strategic calculations for individuals.
@@
 
-The literature thus expects that manipulated elections increase the risk of protest by 1) revealing regime weakness, especially when the margin of victory is low, and 2) exacerbating existing grievances and upsetting citizens' sense of procedural fairness.
-In turn, governments should be reluctant to manipulate.
-However, elections in non-democracies are *generally* risky for incumbents, regardless of how free and fair they may be, as they create an opportunity for opposition collective action [@Knutsen2017] and reveal information [@Pop-Eleches2015].
-Furthermore, electoral protest is relatively rare; even in cases where it does occur, most efforts to overturn fraudulent election results through protest have failed [@kalandadze2009electoral].
+There are thus unresolved tensions in the literature on the relationship between electoral manipulation and electoral protest.
+One branch of the literature argues that the risk of mass protest reduces incumbents' incentive to engage in manipulation, resulting in cleaner elections.
+A second branch argues that mass protest hinges on citizens' subjective assessments of election integrity, their individual grievances, and whether or not the election is seen to be stolen.
+This view complicates the deterrence logic, since incumbents may minimize protest risk (and thus more freely manipulate) by reducing citizen grievances (i.e.
+intervening in the business cycle at election time, distributing club goods, etc.) or by controlling the information environment.
+Since citizens' subjective views of manipulation may diverge from the true level, incumbents may also face an incentive to manipulate the election severely in the hopes of achieving a large margin of victory, under the assumption that protest is likely even at low levels of true manipulation.
+Finally, while a few empirical studies do show evidence for the deterrence model, overall empirical support for the proposition is mixed at best and lacks a general test.
+
+Despite widespread election manipulation, electoral protest is relatively rare; even in cases where it does occur, most efforts to overturn fraudulent election results through protest have failed [@kalandadze2009electoral].
 The apparent failure of protest risk to deter election fraud underlines our limited understanding of non-democratic ruling parties' decision-making.
-Does the severity of fraud, in fact, make more numerous (and thus more costly) protest more likely?
-If not, scholarly attention might turn more profitably toward supply-side explanations for manipulation, such as the organizational capacity of the state [@Simpser2011], the popularity of the government [@rundlett2016deliver], or the risks faced by front-line agents [@Harvey2022].
+Below, it will be argued that election manipulation by its nature will rarely be associated with mass post-election protest.
+Elections in non-democracies are *generally* risky for incumbents, regardless of how free and fair they may be, as they create an opportunity for opposition collective action [@Knutsen2017] and reveal information about the government and society [@Pop-Eleches2015].
+Scholarly attention might turn more profitably toward supply-side explanations for manipulation, such as the organizational capacity of the state [@Simpser2011], the popularity of the government [@rundlett2016deliver], the risks faced by front-line agents [@Harvey2022].
 
@@
 
+The foregoing literature review sketches out prior models of the relationship between election manipulation and protest.
+That existing work has several implications that will be tested here.
+First, does a higher risk of protest deter subsequent election fraud?
+And second, do high levels of fraud increase the risk of subsequent protest?
+
+## Predictions from prior work: Fraud and the risk of protest
+
 As noted above, much of the formal literature on election manipulation assumes that the risk of protest serves as a deterrent to election manipulation.
 Given that pre-election protest is likely to be a strong predictor of post-election protest\textemdash by indicating strong social networks [@mateo2022], pre-existing opposition groups and civil society capacity [@Bunce2006], and/or popular grievances [@Brancati2016]\textemdash the deterrent logic suggests that non-democratic governments will pull back on election-day fraud when the pre-election environment is characterized by protest [@Rozenas2016b].
-This implies the following hypothesis.
+While it is possible that pre-election protest may provide an opportunity for incumbents to crack down, creating more space for manipulating the election, more active pre-election protest suggests that opposition capacity is more robust, all else held equal.
+Moreover, in general, incumbents tend to avoid mass repression in the pre-election phase in favor of targeted harassment of opposition activists [@Bhasin2013].
+The prior literature thus implies the following hypothesis.
 
->H1: Pre-election protest will be associated with lower levels of election-day manipulation.
+>H1: Great pre-election protest risk will be associated with lower levels of election-day manipulation.
 
@@
 This in turn makes citizens [@Egorov2021;@Little2015] or opposition parties [@Luo2020;@Rozenas2016b;@magaloni2010] more likely to stage protests.
+Previous research has also suggested that a fraudulent election is more likely to trigger protest when social grievances are high [@Brancati2016;@tucker2007].
+With some exceptions, such as the study by Lankina and Skovoroda [-@Lankina2017], most research on election fraud and protest treats the latter as a binary outcome\textemdash protests either occur or not.
+This leads to the following hypotheses:
 
->H2: Post-election protest will be more likely to occur when fraud is severe.
+>H2: The likelihood of protest initiation will increase as the severity of election fraud increases.
 
->H3: More severe election fraud is associated with a higher likelihood of protest when winning-party vote-share is low.
+>H3: The effect of election manipulation on the likelihood of protest initiation will increase as the incumbent's margin of victory falls.
 
-Previous research has also suggested that a fraudulent election is more likely to trigger protest when social grievances are high [@Brancati2016;@tucker2007].
+>H4: The effect of election manipulation on the likelihood of protest initiation will increase as economic grievances rise.
 
->H4: Post-election protest will be more likely to occur when fraud is severe and economic and social grievances are elevated.
+[My theory here:  Election manipulation (fraud?) cannot serve as a reliable indicator of strength or weakness, meaning that opposition groups cannot use it to behave strategically.
+Strong and weak alike engage in it (noted already).
+Also, election manipulation or its absence can be deployed by regimes for varying reasons.
+A strong regime may feel comfortable dialing back on fraud to gain more information (find cites for this); a weak one may not be able to muster the resources.
+A strong regime may generate lots of manipulation in order to flaunt its capacity; a weak regime may need to rely on fraud to boost otherwise anemic public support.
+So because it doesn't communicate information about strength, it doesn't influence protest; protests happen or not--and live or die--based on the balance of resources on either side.]
 
 However, there are reasons to doubt the veracity of the deterrence logic.
-Little [-@Little2012] argues that citizens\textemdash knowing the regime is capable of manipulating the results\textemdash discount an expected level of fraud from the results when deciding whether or not to protest.
-The severity of fraud would thus be untethered from protest risk; what matters is citizens' underlying estimate of the regime's strength.
-Election fraud can itself be a signal of strength, rather than weakness.
-Fraud is not solely a weapon of floundering incumbents\textemdash in fact, Seeberg [-@Seeberg2019] finds that higher-capacity states are weakly associated with more election fraud.
-Rather than acting out of desperation, strong ruling parties may engage in election fraud in order to pad their margins and secure cooperation from other oactors [@simpser2013] or because of a bandwagoning effect among front-line agents eager to please an incumbent patron [@rundlett2016deliver].
-Fraud is thus an ambiguous signal of regime strength, which cannot conclusively guide strategic decisions to protest.
+Election fraud and other forms of manipulation that do not directly mobilize voters (such as vote-buying) cannot be a reliable indicator of the strength or weakness of the regime; consequently the 'signal' provided by a fraudulent election will not be able to reliably generate strategic behavior by the opposition.
+The central problem is that election manipulation is a tool that can be used by weak and strong governments alike [@simpser2013].
+A strong government may allow for relatively clean elections, in order to gather more information about the opposition and society [@ananyevDictatorsSignalStrength2022;@trinhTeaLeafElections2022]; a weak government may be unable to muster the necessary resources [@greene_why_2007].
+The observation of a relatively clean election alone does not discriminate between the two.
+Likewise, a strong government may manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver].
+This is all in contrast to the typical deterrence model, in which a weaker government attempts to manipulate in order to obscure its weakness, but in keeping with Seeberg's [-@Seeberg2019] finding that higher-capacity states are weakly associated with more election fraud.
+For these reasons, acts of manipulation can be interpreted either as signs of regime strength or weakness [@Pop-Eleches2015].
+It is also in line with the formal model proposed by Little [-@Little2012], in which citizens\textemdash knowing the regime is capable of manipulating the results\textemdash discount an expected level of fraud from the results when deciding whether or not to protest.
 
@@
 Others have noted structural factors like the importance of ties to democratic Western states and the size and strength of the ruling party and security apparatus in influencing the course of the Color Revolutions [@LucanWay2008].
-In either case, these factors create an incentive for opposition protest whether electoral fraud is minor or severe\textemdash a kind of case of 'crying wolf'\textemdash an an outcome that is also posited in some of the formal modeling literature [@Little2015].
-As a result, it is likely that indicators of regime weakness or opposition strength will be correlated with post-election protest, while the severity of fraud will not.
+In either case, these factors create an incentive for opposition protest whether electoral fraud is minor or severe\textemdash a kind of case of 'crying wolf'\textemdash an outcome that is also posited in some of the formal modeling literature [@Little2015].
+As a result, it is likely that the severity of fraud is thus untethered from protest risk; what matters is citizens' underlying estimate of the regime's strength.
 
-With some exceptions, such as the study by Lankina and Skovoroda [-@Lankina2017], most research on election fraud and protest treats the latter as a binary outcome.
-Such an approach does not consider how the protest environment might evolve after the election.
+## The course of protest: do higher levels of fraud make protest waves more significant?
+
+Finally, and perhaps most importantly, by emphasizing binary protest outcomes at the expense of trends over time, prior work has overemphasized the risk of popular backlash to incumbents.
+While many incumbents may indeed prefer a quiet post-election period, the occurrence of protest can also have regime-sustaining benefits.
+When unexpected post-election protests occur, trust in the government can increase, especially among opposition supporters [@Frye2019].
+As the number of protests increases, support for the protesters' aims may fall; this effect is even larger for those who consume pro-government media [@tertytchnayaElectoralProtestsPolitical2020].
+These empirical results are in line with models proposed by Lohmann [-@lohmannDynamicsInformationalCascades1994] and Shadmehr and Bernhardt [-@shadmehrCollectiveActionUncertain2011]; they argue that when some citizens are too predisposed to protest, others become more inclined to stay home\textemdash dooming the first group's efforts to failure.
+Such influences help offset the 'risk' of protest to incumbents, especially when protest movements are relatively small and/or isolated.
+
+What incumbents likely *do* seek to avoid are large, widespread, and long-lasting protest movements that strain the government's resources and put pressure on elites [@chenowethPhysicsDissentEffects2019].
 Here I argue that, once protests begin, the extent of electoral fraud is unlikely to affect their scope.
@@
 
+In addition, governments may learn from the experiences of other governments, just as opposition groups may learn from opposition.
+Anticipating threats from below may encourage less democratic governments to invest more in the security services, or to reform political and legal institutions in order to stifle potential protests before they take shape [@hallAuthoritarianLearningConceptual2017a;@Morgenbesser2020].
+
 What results is a test of strength between the anti-incumbent protest coalition and the regime with its supporters [@mcadamDynamicsContention2001;@tillyRegimesRepertoires2010].
@@
 High-capacity states, as a result, are more likely to be able to engage in successful repression [@tarrowPowerMovementSocial2011].
+Here I expect to see support for the so-called 'more murder in the middle' framework [@feinLifeIntegrityViolationsDemocracy1995], in which the regimes with the most demonstrated repressive capacity deter potential challengers from taking to the streets in the first place, and the least repressive states defuse protest through accommodation [@pierskallaProtestDeterrenceEscalation2010;@mooreRepressionDissentSubstitution2000].
+
 Though a biased election may temporarily allow diverse opposition groups to set aide their differences [@trejo_ballot_2014], this internal diversity can enable the incumbent to divide the opposition with selective repression or concessions.
@@
 
-In this struggle between the potential protest coalition and the incumbent government, protesters and their possible allies behave as "passionate economists" [@VanZomeren2012]---they are motivated both by emotional responses to grievance, and by a more rational calculation of individual-level risk and group-level efficacy.
+In this struggle between the potential protest coalition and the incumbent government, protesters and their possible allies behave as "passionate economists" [@VanZomeren2012]\textemdash they are motivated both by emotional responses to grievance, and by a more rational calculation of individual-level risk and group-level efficacy.
 What is the likelihood that protest action can help resolve the grievance, and does the benefit of participating outweigh the risk of personal consequences to the individual for participating?
@@
 
-Finally, protest waves can have self-defeating and regime-sustaining effects.
-When unexpected post-election protests occur, trust in the government can increase,  especially among opposition supporters [@Frye2019].
-As the number of protests increases, support for the protesters' aims may fall; this effect is even larger for those who consume pro-government media [@tertytchnayaElectoralProtestsPolitical2020].
-These empirical results are in line with models proposed by Lohmann [-@lohmannDynamicsInformationalCascades1994] and Shadmehr and Bernhardt [-@shadmehrCollectiveActionUncertain2011]; they argue that when some citizens are too predisposed to protest, others become more inclined to stay home\textemdash dooming the first group's efforts to failure.
-These effects give reason to doubt that protest risk serves as a guardrail against election manipulation; rather, post-election protests may burn themselves out while leaving the public more approving of the government.
+In summary, fraudulent elections are not likely to pose a significant protest risk to incumbents.
+While opposition actors may seize upon fraudulent elections to initiate a protest, the relationship between this decision and the level of fraud is likely to be weak.
+Opposition groups are also incentivized to stage protests regardless of the level of fraud, given the highly salient moment afforded by the election itself [@cunhaSignalReceivedAuthoritarian2022].
+More importantly, even if a protest movement begins after the election, any useful information provided by the level of election manipulation quickly decays and is replaced by signals from the protests themselves [@lohmannDynamicsInformationalCascades1994].
+Protesters, bystanders, and elites can update their estimation of the relative strength of the protest coalition and the incumbent government by observing the size of crowds, the number of protests, and the government's reaction.
 
-In summary, fraudulent elections are not likely to pose a significant protest risk to incumbents for two reasons.
-First, protest initiation is unlikely even in fraudulent elections because it is a noisy signal of regime strength.
-Weak and strong governments alike may engage in election manipulation [@simpser2013], and acts of manipulation can be interpreted either as signs of strength or weakness [@Pop-Eleches2015].
-Second, even if a protest movement begins after the election, any useful information provided by the level of election manipulation quickly decays and is replaced by signals from the protests themselves [@lohmannDynamicsInformationalCascades1994].
-Protesters, bystanders, and elites can update their estimation of the relative strength of the protest coalition and the incumbent government by observing the size of crowds, the number of protests, and the government's reaction.
-Since the level of manipulation will no longer serve as an important indicator of regime strength once a challenge is taken to the streets, the number of protests will instead be related to other indicators of opposition and incumbent capacity.
-As identified in prior research, these factors include the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], the margin of victory in the election [@simpser2013], and whether or not the chief executive faces term limits [@hale_patronal_2014].
+Since the level of manipulation will no longer serve as an important indicator of regime strength once a challenge is taken to the streets, the number and size of protests will instead be related to other indicators of opposition and incumbent capacity.
+When opposition groups and civil society capacity are more organized and better resourced, it is likely that more protesters will take to the streets in more numerous events (see Chenoweth et al [-@chenowethStateRepressionNonviolent2017] for a review of factors making protest movements more resilient to repression).
+By contrast, when the government appears well-resourced and resolved, more 'passionate economists' should stay home.
+As identified in prior research, informative factors should be the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], the margin of victory in the election [@simpser2013], and whether or not the chief executive faces term limits [@hale_patronal_2014].
 Generally speaking, these structural variables indicate aspects of the political opportunity structure that make protest more likely to occur [@meyer_protest_2004].
 
->H5: Once protests begin, the severity of election fraud will not be correlated with the number of protest events.
+Here, this logic will be tested using two measures, both of which may indicate the strength of a protest movement, as well as the cost of protest to incumbent governments: the total number of protest events and the estimated number of protesters on the largest single day of protests.
+I assume that more numerous protests, and those that built to a higher peak participation, are costlier for governments to contain, all else held equal.
+If fraudulent elections reveal information about regime weakness or generate sustained grievance-motivated protest, as the deterrence logic predicts, more fraudulent elections should be associated with larger and more numerous protest waves.
+Prior theory expects these effects to be largest when margins of victory are smaller, or when economic grievances are high.
+
+However, the theory articulated above does **not** expect that the severity of fraud will be associated with the course of protest.
+Instead, it predicts that underlying indicators of opposition and ruling-party strength will influence the direction of protest movements.
+In particular, it expects the size, scope, and duration of protest waves will be sensitive to the repressive capacity of the state and to indicators of opposition strength.
+that high and low levels of repression will be associated with smaller protest movements (in number and size).
+
+>H5: The relationship between repressive capacity and the number of protest events and the maximum event size is curvilinear, with number and size likely to to first increase and then fall as repressive capacity grows.
+
+>H6: The number of protest events and the maximum event size are likely to increase with indicators of opposition and civil-society strength.
+
+As a final test of the mechanism outlined in the theory presented here\textemdash that election manipulation offers few clues to regime strength, and thus the course of electoral protests are determined by more structural factors\textemdash the same models will also be applied to the duration of post-election protest waves.
+Longer protest waves indicate a more even match between the resources and resiliency of the protesters and the capacity and cohesiveness of the regime.
+It is expected that more severe election fraud will *not* be associated with longer protest waves, as would be predicted if fraud were an indicator of state weakness.
+Instead, structural factors should again be more influential;   harsh repression should be associated with shorter protest waves, while indicators of opposition and civil-society strength should help keep protest waves alive.^[Note here about excluding successful movements.]
 
@@
 library(stargazer)
-library(cowplot)
-library(pscl)
+library(ggeffects)
+library(glmmTMB)
 library(modelsummary)
-library(ggeffects)
-library(MASS)
+library(huxtable)
+library(simPH)
 
-nelda_massmob <- read.csv(here::here("Data", "nelda_massmob_all.csv")) #nelda_complete.csv for only elex-related
+nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2023-05-22.csv"))
 
-nelda_massmob <- nelda_massmob %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
-nelda_massmob <- nelda_massmob %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))
+nelda2.sub <- nelda2.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
+nelda2.sub <- nelda2.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))
 
@@
 
-nelda_massmob <- nelda_massmob %>% rename(boycott = nelda14)
-nelda_massmob <- nelda_massmob %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
+nelda2.sub <- nelda2.sub %>% rename(boycott = nelda14)
+nelda2.sub <- nelda2.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                  NA)))
@@
 
-nelda_massmob <- nelda_massmob %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
-nelda_massmob <- nelda_massmob %>% mutate(econ.crisis = as.factor(econ.crisis))
+nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
+nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = as.factor(econ.crisis))
 
-nelda_massmob <- nelda_massmob %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
-nelda_massmob <- nelda_massmob %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))
+nelda2.sub <- nelda2.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
+nelda2.sub <- nelda2.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))
 
-nelda_massmob <- nelda_massmob %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))
-
-##Removing cases before 1990, when ECAV starts
-nelda_massmob <- nelda_massmob %>% filter(year >= 1990)
+nelda2.sub <- nelda2.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))
 
 ## Removing cases where inc. loses
-nelda_massmob <- nelda_massmob %>% filter(inc.lose == 0)
-
-##Nelda crisis as a factor
-
-nelda_massmob <- nelda_massmob %>% mutate(econ.crisis.nelda.1lag = as.factor(econ.crisis.nelda.1lag))
+#nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)  #See how this affects things
 
 ##Intmon as a factor
-nelda_massmob <- nelda_massmob %>% mutate(v2elintmon = as.factor(v2elintmon))
+nelda2.sub <- nelda2.sub %>% mutate(v2elintmon = as.factor(v2elintmon))
 
-##Presidential as a facotr
-nelda_massmob <- nelda_massmob %>% mutate(presidential = as.factor(presidential))
-
-nelda_massmob <- nelda_massmob %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))
+nelda2.sub <- nelda2.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))
 
@@
 
-nelda_massmob <- nelda_massmob %>% mutate(physintlag.sq = physinteg.1lag^2)
+nelda2.sub <- nelda2.sub %>% mutate(physintlag.sq = physinteg.1lag^2)
 
-#Crisis as factor
+## Categorical variable for incumbent status
 
-nelda_massmob <- nelda_massmob %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))
+nelda2.sub <- nelda2.sub %>% mutate(incumb.status = as.factor(ifelse(inc.termlimit == "yes" & chosen.successor == 1, "Chosen successor",
+                                                           ifelse(inc.termlimit == "yes" & chosen.successor == 0, "Contested exec. election", 
+                                                                  ifelse(inc.termlimit == "no", "Executive eligible", "Legislative election"))))) 
+nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Legislative election"))
+
+#Coding as factors
+
+nelda2.sub <- nelda2.sub %>% mutate(presidential = factor(presidential))
+nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = factor(inc.termlimit.num, exclude = NULL))
+nelda2.sub <- nelda2.sub %>% mutate(rentierstate = factor(rentierstate))
+nelda2.sub <- nelda2.sub %>% mutate(taxstate = factor(taxstate))
+nelda2.sub <- nelda2.sub %>% mutate(unscheduled_election = factor(unscheduled_election))
+nelda2.sub <- nelda2.sub %>% mutate(inc.lose = factor(inc.lose))
+nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))
 
@@
 
-nelda_massmob <- nelda_massmob %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))
+nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))
 
@@
 
-nelda_massmob <- nelda_massmob %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | elections.cancel.prot, 1, 0))
+nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))
 
@@
 
-#protest.cases <- nelda_massmob %>% filter(any.ecav.protest == 1)
+protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1)
 
@@
 
-nelda_massmob <- nelda_massmob %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))
+nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))
 
-## Changing participants from numeric category to text
-#nelda_massmob <- nelda_massmob %>% mutate(median.participants = round(median.participants))
-#nelda_massmob <- nelda_massmob %>% mutate(particpants_categorical = ifelse(is.na(median.participants)==T, NA, ifelse(median.participants == 1, "Less than 10",
-#                     ifelse(2, "10 to 99", ifelse(3, "100 to 999",
-#                                                  ifelse(4, "1000 to 9,999", ifelse(5, "Greater than 10,000", NA)))))))
-#nelda_massmob <- nelda_massmob %>% mutate(particpants_categorical = as.factor(particpants_categorical))
+## Replacing NA with 0 for protest size
 
-## Median particiants (MM) as factor
+nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(is.na(max_size)==T & n.events.post == 0, 0, max_size))
 
-nelda_massmob <- nelda_massmob %>% mutate(median.participants.mm.post = as.factor(median.participants.mm.post))
+## Fixing missing values of winner.share.upd for liberal democracies
 
-## Inverting the vote buying measure so it is not reverse coded
+nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(is.na(winner.share.upd)==T & v2x_regime == 3, winner.margin, winner.share.upd))
 
-nelda_massmob <- nelda_massmob %>% mutate(v2elvotbuy_inv = -1*v2elvotbuy)
+## Excluding liberal democracies
+
+nelda2.sub <- nelda2.sub %>% filter(v2x_regime <= 2) #To-do: Extent to include liberal dems in appendix
 
@@
 
-These hypotheses are tested using data from three datasets: NELDA [@hyde2012], V-Dem version 11.1 [@vdem11.1], and the Mass Mobilization Project dataset [@massmob].
+These hypotheses are tested using data from three datasets: NELDA [@hyde2012], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019].
 The unit of observation for this study is an election period as recorded in NELDA.
 This framework prevents selecting on the dependent variable, by capturing both protested and non-protested elections.
-NELDA also includes a variable indicating elections in which incumbents lost; as this study is focused on protest against ostensibly victorious incumbents, I use this variable to filter out elections which the opposition won outright.
-I further limit the sample to regimes coded as electoral autocracies or electoral democracies in V-Dem.
-This excludes closed autocracies (where no national elections are held) and liberal democracies (where electoral integrity and rule of law are both high).^[The specific variable used is 'v2x_regime' from V-Dem.
-This variable codes an electoral democracy, in part, if the country scores above a 2 on the measure of free and fair elections ('v2elfrfair_osp').
-This still permits election fraud to be detected in the electoral democracies; v2elfrfair can take on a value of 3 when "deficiencies and some degree of fraud and irregularities" occur.] Finally, the beginning of the date range is limited by the availability of data on protest from the Mass Mobilization dataset, as discussed below, which begins in 1990.
-The inclusion of a control variable from the Latent Judicial Independence dataset [@Linzer2015] limits the end of the time range to 2012.
+I limit the sample by excluding closed autocracies (where no national elections are held).
+Since the theory of backlash to electoral manipulation could in theory apply to liberal democracies, I include them in the main analysis (as coded by V-Dem).^[The specific variable used is 'v2x_regime'.] As a robustness check for this sampling decision, I also include the main models without liberal democracies in the appendix, with no substantial changes in the outcomes.
+Finally, the beginning of the date range is limited by the availability of data on protest from ECAV, as discussed below, which ranges from 1990 to 2012.
 The resulting dataset includes 647 total election-period observations across 109 countries.
 
-The Mass Mobilization dataset, from which study's measures of protest are drawn, attempts to record all anti-government protests drawing 50 or more participants.
-The data is in event-day format.
-I divide this data into a pre-election period (180 days prior to an election) to capture the number and size of pre-election protest, and a post-election period (90 days after the election) for the dependent variables.
-I include all protest events in the post-election period, rather than filtering based on key terms like "election" or "vote," since the conduct and outcome of the election may spark protests that do not explicitly refer back to the election in protesters' demands.
-This is an extension of both the strategic and grievance logics of post-election protest.
-An electoral outcome that signals weakness for the incumbent regime may lead to strategic protests by outsider groups pressing their demands, even if they do not specifically focus on the integrity of the election.
-Likewise, disaffected groups may use the election as a focal point for their socioeconomic demands, irrespective of the nature of the election.
-Excluding such cases would like bias the results as a consequence.
+ECAV records protests and other contentious actions related to an election, ranging from six months before to three months after the vote (or three months after the final round of a multi-round election).
+To focus the study on protest, I only include events labeled as protests, occupations, or blockades, excluding acts like shootings, bombings, and coups.
+ECAV data is the event-day format; to capture the *number of protests* before and after the election, I take the sum number of events recorded in ECAV before and after election day, with election-day protests being counted as post-election.
+To evaluate hypotheses regarding protest size, I convert ECAV's categorical data on number of participants per event-day to estimate numbers, and take the sum where multiple protests were held on a given day.
+The ECAV dataset codes protest size as a categorical variable with five levels, ranging from fewer than ten participants to more than 10,000.
+To deal with this data structure, where the top category has no upper limit, I take the median value of each category, with the exception of the top category.
+For this category, which has no upper bound, I take the minimum value of 10,000.
+The variable *maximum protest size* is found by taking the number of estimated protest participants who turned out on a single day.
 
-Two variables are derived from the Mass Mobilization data.
-First, I collect the *median size* of post-election protest.
-The Mass Moblization data records protest size using six categories.\footnote{This are: 50-99 participants, 100-999, 1,000-1,999, 2,000-4,999, 5,000-9,999, and greater than 10,000 participants.} I then take the median size of the pre-election set of protests per election period; the result is an ordered categorical variable.
-Second, I capture the  *number of protests* by taking the sum number of events recorded in the pre- and post-election data.
-The result is a count variable ranging from zero to 435 for the pre-election period and 0 to 385 for the post-election period.
-The large majority of elections experience no post-election protest: in 504 of the 647 elections, no protests are recorded.
-Figure \@ref(fig:elex-hist) reports the distribution of protest counts across the dataset, as well as winners' vote-shares.
+The large majority of elections experience no post-election protest: in 1168 of the 1477 elections, no protests are recorded.
+Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest number and maximum single-day protest size.
 
@@
 
-p.hist.events <- ggplot(nelda_massmob, aes(x=n.events.massmob.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()
+p.hist.events <- ggplot(nelda2.sub, aes(x=n.events.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()
 
-p.hist.fraud <- ggplot(nelda_massmob, aes(x=v2elirreg.inv)) + geom_histogram(bins=100) + xlab("Election fraud") + theme_bw()
+p.hist.size <- ggplot(nelda2.sub, aes(x=max_size)) + geom_histogram(bins=100) + xlab("Maximum protest size") + theme_bw()
+
+p.hist.fraud <- ggplot(nelda2.sub, aes(x = v2elirreg.inv)) + geom_histogram(bins=100) + xlab("Election fraud") + theme_bw()
 ```
 
-```{r elex-hist, fig.cap="Distributions of election fraud and post-election protest", fig.height=7, echo=FALSE, message=FALSE, warning = FALSE}
-plot.combined.hist <- cowplot::plot_grid(p.hist.fraud, p.hist.events, ncol=1)
+```{r elex-hist, fig.cap="Distributions of election fraud, post-election protest, and maximum protest size", fig.height=7, echo=FALSE, message=FALSE, warning = FALSE}
+plot.combined.hist <- cowplot::plot_grid(p.hist.fraud, p.hist.events, p.hist.size, ncol=1)
 plot.combined.hist
@@
 ```{r, include=F}
-incl.vars <- nelda_massmob %>% dplyr::select('Judicial independence' = jucon.1lag, 
+incl.vars <- nelda2.sub %>% dplyr::select('Judicial independence' = jucon.1lag, 
                                             'Legislative constraints' = legcon.1lag, 
@@
                                             'Incumbent vote-share' = winner.share.upd, 
-                                            'Median pre-election protest size' = median.participants.mm.pre,
+                                            'Median pre-election protest size' = median.participants.pre,
+                                            'Largest protest size' = max_size,
                                             'Off-schedule election' = unscheduled_election)
@@
 
-This distribution is in line with the theoretical proposition that two data-generating processes are at work: the factors that lead to the occurrence of protest may not be the same as those that sustain or intensify it.
+This distribution\textemdash with large numbers of zeroes\textemdash is in line with the theoretical proposition that two data-generating processes are at work: the factors that lead to the occurrence of protest may not be the same as those that sustain or intensify it.
 Consequently, I model the results using negative binomial hurdle regressions.
@@
 The logistic portion of the model identifies the factors that make *any protest* more likely to occur; the negative binomial regressions to model the count of protests given that protest has begun.
+Since protest size cannot take on a negative value, it is also modeled as a count variable with a negative binomial distribution in the hurdle models.
 
 Each model includes the explanatory variables proposed above.
-First, three proxy measures of economic grievance include an indicator of *public sector corruption* and the *lagged GDP growth rate*, both taken from V-Dem, and a binary measure of *economic crisis* from the NELDA dataset.
-Higher levels of corruption, negative GDP growth rates, and a positive value for *economic crisis* should each be expected to increase citizens' sense of grievance.
+First, as a proxy measures of economic grievance, I include the *lagged GDP growth rate*, taken from V-Dem.
+Lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance.
 To capture the *severity of election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than aggregate measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], it can fail to signal strength [@Harvey2020], and it can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021].
-This measure of election integrity should thus pose a strong test of the protest-oriented model.
+This measure of election integrity should thus serve as a strong test of the protest-oriented model.
 Finally, the winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia and Psephos.}
@@
 Several control variables are included, which could be common causes for both the severity of election manipulation and protest.
-To help control for the underlying mobilizational capacity, which likely affects incumbents' decision to manipulate and the latent risk of protest, I include lagged indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem) and the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs').
+To help control for the underlying mobilizational capacity, which likely affects incumbents' decision to manipulate and the latent risk of protest, I include lagged indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), and the available of *alternative sources of information*.
+Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019].
+Finally, all models include measures of *judicial constraints* and  *legislative constraints* on the executive, V-Dem variables which captures the extent to which the judicial and legislative bodies in the country act independently in practice to limit executive power.
+
 Likewise, the ability of the state to engage in repression is likely to embolden the ruling party and deter protests; I thus control for *physical integrity* using the 'physical violence index' from V-Dem lagged one year.
 Higher values for this variable indicate greater freedom from torture and political killing.
-Since intense repression may itself become a grievance that triggers further protest, I also include the variable's square term.
+I also include the variable's square term to account for the 'murder in the middle' framework.
+The physical integrity variables are expected to be explanatory in models protest, but are important control variables in the models of election fraud.
+Since repressive capacity helps influence protest in expectation [@Ritter2016], incumbents that are more confident of deterring protest in advance should be more willing to engage in fraud under the protest-deterrent model of manipulation.
+Controlling for repression removes this confounding bias and allows for an estimate of the effect of pre-election disquiet on fraud.
+As an especially useful test, the deterrent model might expect fraud to be most reduced in the context of high pre-election protest and low repressive capacity.
 
-Next, all models include three measures of the degree of political openness in the country.
-The first of these is an indicator variable for *electoral democracy*; the reference category is electoral autocracy.
-The second and third are *judicial constraints* and  *legislative constraints* on the executive, V-Dem variables which captures the extent to which the judicial and legislative bodies in the country act independently in practice to limit executive power.
-Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019].
+NELDA also includes a binary variable indicating elections in which *incumbents lost*; I expect that protest will be less likely when this variable takes a value of one.
 
@@
 
+Hypothesis 1 suggests that a pre-electoral environment characterized by greater protest risk will be associated with lower election-day fraud.
+While the true value of pre-election protest risk is unobservable, I proxy it using the *number of pre-election protest events*, the state's *repressive capacity* and the interaction of the two.
+More protest events in the pre-election period is prima facie evidence of mobilization capacity in civil society and/or the opposition, and may also indicate an elevated sense of grievance.
+
 The results of these models are not promising for argument that protest risk deters election manipulation.
-First, while Hypothesis 1 is supported in Model 1 in Table  \@ref(tab:manipdv) \textemdash the coefficient on *pre-election protest* is negative and significant\textemdash the substantive effect of the relationship is minuscule.
-The model predicts that each additional pre-election protest event is associated with a reduction in the fraud measure of 0.007.
-This is a very small effect, given that the dependent variable ranges from `r min(nelda_massmob$v2elirreg.inv, na.rm = T)` to `r max(nelda_massmob$v2elirreg.inv, na.rm = T)`.
-A very large increase in the degree of protest, from zero to the ninetieth percentile (`r as.numeric(quantile(nelda_massmob$n.events.massmob.pre, probs = .9))` events) is associated with a `r (as.numeric(quantile(nelda_massmob$n.events.massmob.pre, probs = .9)))*.007` reduction in election-day fraud.
-This represents less that ten percent of a standard deviation of the election fraud variable.
-So, while pre-election protest (an indicator of post-election protest risk) is indeed associated with reduced election-day fraud, this reduction is only likely to be noticeable in extreme circumstances.
-For example, pre-election protest in the 99th percentile is associated with a reduction of `r (as.numeric(quantile(nelda_massmob$n.events.massmob.pre, probs = .99)))*.007`, a bit less than half of a standard deviation.
-This suggests that the level of election manipulation is not highly responsive to changes in the pre-election protest environment.
+First, Hypothesis 1 is not supported in Models 1 and 2 in Table  \@ref(tab:manipdv) \textemdash the coefficient on *pre-election protest* is positive and significant.
+The model predicts that each additional pre-election protest event is associated with an **increase** in the fraud measure of 0.008.
+This is a very small effect, given that the dependent variable ranges from `r min(nelda2.sub$v2elirreg.inv, na.rm = T)` to `r max(nelda2.sub$v2elirreg.inv, na.rm = T)`.
+An increase from zero to five protest events results in a predicted increase in fraud of roughly three percent of a standard deviation.
+While this is a small effect, it runs counter to the expectations of the deterrence logic.
+And the effect may be sizable in extreme circumstances.
+For example, pre-election protest in the 99th percentile is associated with an increase of `r (as.numeric(quantile(nelda2.sub$n.events.pre, probs = .99)))*.008`, about one-fifth of a standard deviation.
+This suggests that the level of election manipulation, rather than being deterred by a highly contentious environment, may be deployed to make up for anticipated poor election results.
 
@@
 ```{r manipdv, echo=FALSE}
-m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   + presidential + inc.termlimit.num  +  libdem.1lag + urban.pct + rentierstate + taxstate + n.events.massmob.pre +  physinteg.1lag + physintlag.sq  + median.participants.mm.pre + unscheduled_election + (1 | stateid), data = nelda_massmob %>% filter(n.events.massmob.pre < 200)) #There are two extreme observations removed by this filter
+m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate + n.events.pre  +  unscheduled_election + regime_age + (1 | stateid), data = nelda2.sub)
 
@@
 
-modelsummary(m1, output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Random-intercept multilevel model of election-day fraud', coef_rename = 
+m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   + presidential   +  libdem.1lag + urban.pct + rentierstate + taxstate + physinteg.1lag +  unscheduled_election + regime_age+ (1 | stateid), data = nelda2.sub)
+
+#summary(m1b)
+
+m1c <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + unscheduled_election + regime_age + (1 | stateid), data = nelda2.sub)
+summary(m1c)
+
+m1d <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate + n.events.pre*physinteg.1lag +  unscheduled_election + regime_age + (1 | stateid), data = nelda2.sub)
+summary(m1d)
+
+modelsummary(c(m1, m1b, m1c, m1d), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Random-intercept multilevel model of election-day fraud', coef_rename = 
                c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
@@
 ```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
-model.n.novoteshare.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +     + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election , data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+model.n.novoteshare.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + inc.lose,
+                                ziformula=~.,
+                                data = nelda2.sub,
+                                family=truncated_nbinom2)
 summary(model.n.novoteshare.hurd.mm)
 
-model.n.base.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election| jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+model.n.base.hurd.mm <- glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + inc.lose,
+                               ziformula=~.,
+                              data = nelda2.sub,
+                                family=truncated_nbinom2)
 summary(model.n.base.hurd.mm)
 
-model.n.base2.hurd.mm <- hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election |  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+model.n.base2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + inc.lose,
+                                          zi = ~.,
+                                          family= nbinom2,
+                                          data = nelda2.sub)
 summary(model.n.base2.hurd.mm)
 
-model.n.griev.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election| jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, dist = "negbin", data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+model.n.griev.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age+ inc.lose,
+                                ziformula=~.,
+                                data = nelda2.sub,
+                                family=truncated_nbinom2)
 summary(model.n.griev.hurd.mm)
 
-model.n.griev2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election | jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, data = nelda_massmob, dist = "negbin")
+model.n.griev2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + inc.lose ,
+                                ziformula=~.,
+                                data = nelda2.sub,
+                                family=truncated_nbinom2)
 summary(model.n.griev2.hurd.mm)
 
-model.n.griev3.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election  | jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, data = nelda_massmob, dist = "negbin")
+model.n.griev3.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + inc.lose  ,
+                                ziformula=~.,
+                                data = nelda2.sub,
+                                family=truncated_nbinom2)
 summary(model.n.griev3.hurd.mm)
+
+model.n.repr.hurd <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv  +  physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + inc.lose,
+                                ziformula=~.,
+                                data = nelda2.sub,
+                                family=truncated_nbinom2)
+summary(model.n.repr.hurd)
+```
+
+```{r models-hurdle-duration, warning = FALSE, message = FALSE, include = F}
+model.d.novoteshare.hurd.mm <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + inc.lose,
+                                ziformula=~.,
+                                data = nelda2.sub,
+                                family=truncated_nbinom2)
+summary(model.d.novoteshare.hurd.mm)
+
+model.d.base2.hurd <- glmmTMB::glmmTMB(d.events.num ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age +  inc.lose ,
+                                          zi = ~.,
+                                          family= nbinom2,
+                                          data = nelda2.sub)
+summary(model.d.base2.hurd)
+
+model.d.griev.hurd <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + inc.lose,
+                                ziformula=~.,
+                                data = nelda2.sub,
+                                family=truncated_nbinom2)
+summary(model.d.griev.hurd)
+
+model.d.griev2.hurd <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + as.factor(inc.lose) ,
+                                ziformula=~.,
+                                data = nelda2.sub,
+                                family=truncated_nbinom2)
+summary(model.d.griev2.hurd)
+
+model.d.griev3.hurd <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + as.factor(inc.lose)  ,
+                                ziformula=~.,
+                                data = nelda2.sub,
+                                family=truncated_nbinom2)
+summary(model.d.griev3.hurd)
+
+```
+
+```{r, include=FALSE}
+library(survival)
+nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
+km <-  with(nelda2.sub, Surv(d.events.num, surv_status))
+
+cox1 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age, data = nelda2.sub %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+summary(cox1)
+
+cox2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age , data = nelda2.sub %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+summary(cox2)
+
+cox3 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age, data = nelda2.sub %>% filter(d.events.num > 0))
+summary(cox3)
+
+cox4 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age, data = nelda2.sub %>% filter(d.events.num > 0))
+summary(cox4)
+
+```
+
+```{r models-hurdle-size, warning = FALSE, message = FALSE, include = F}
+model.size.novoteshare <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose) ,
+                                ziformula=~.,
+                                data = nelda2.sub,
+                                family=truncated_nbinom2)
+summary(model.size.novoteshare)
+
+model.size.base <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   v2elirreg.inv + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose) ,
+                                ziformula=~.,
+                                data = nelda2.sub,
+                                family=truncated_nbinom2)
+summary(model.size.base)
+
+model.size.base2 <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose) ,
+                                ziformula=~.,
+                                data = nelda2.sub,
+                                family=truncated_nbinom2)
+summary(model.size.base2)
+
+model.size.griev <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose) ,
+                                ziformula=~.,
+                                data = nelda2.sub,
+                                family=truncated_nbinom2)
+summary(model.size.griev)
+
+```
+
+```{r zero-model-table, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+models_t1 <- list()
+models_t1[['Model 3']] <- model.n.base.hurd.mm
+models_t1[['Model 4']] <- model.n.base2.hurd.mm
+models_t1[['Model 5']] <- model.n.griev.hurd.mm
+#models_t1[['Model 5']] <- m_duration_count
+#models_t1[['Model 6']] <- model.d.base2.hurd
+#models_t1[['Model 7']] <- model.d.griev.hurd
+
+t1a <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, omit_coefs = c("rentierstate1", "taxstate1", "n.events.pre", "unscheduled_election1", "inc.lose1", "v2elirreg.inv:winner.share.upd", "gdpgro.1lag:v2elirreg.inv"), statistics = character(0)) %>% set_caption("Table 3: Zero-inflated models of protest initiation")
+
+theme_compact(t1a, header_rows = TRUE, header_cols = FALSE)
+
+t1b <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, omit_coefs = c("(Intercept)", "jucon.1lag" , "legcon.1lag" , "gdpgro.1lag" , "altinf.1lag" , "pubseccorrup.lag" , "osorg.1lag" , "econ.crisis.nelda.1lag1" , "natlpartyorg.lag" , "v2elirreg.inv" , "pubseccorrup.lag"   , "presidential1" , "inc.termlimit.num1"  ,  "physinteg.1lag" , "physintlag.sq" , "urban.pct")) %>% set_caption("Table 3 cont.: Zero-inflated models of protest initiation")
+
+theme_compact(t1b, header_rows = TRUE, header_cols = FALSE)
+
+models_t2 <- list()
+models_t2[['Model 3']] <- model.n.base.hurd.mm
+models_t2[['Model 4']] <- model.n.base2.hurd.mm
+models_t2[['Model 5']] <- model.n.griev.hurd.mm
+models_t2[['Model 6']] <- model.size.base
+models_t2[['Model 7']] <- model.size.base2
+models_t2[['Model 8']] <- model.size.griev
+
+t2a <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), omit_coefs = c("rentierstate1", "taxstate1", "n.events.pre", "unscheduled_election1", "inc.lose1", "v2elirreg.inv:winner.share.upd", "gdpgro.1lag:v2elirreg.inv"), statistics = character(0)) %>% set_caption("Table 4: Negative binomial models of number and duration of protests") %>% insert_row("", "", "Number", "", "", "Size", "", after = 0)
+
+theme_compact(t2a, header_rows = TRUE, header_cols = FALSE)
+
+t2b <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), omit_coefs = c("(Intercept)", "jucon.1lag" , "legcon.1lag" , "gdpgro.1lag" , "altinf.1lag" , "pubseccorrup.lag" , "osorg.1lag" , "econ.crisis.nelda.1lag1" , "natlpartyorg.lag" , "v2elirreg.inv" , "pubseccorrup.lag"   , "presidential1" , "inc.termlimit.num1"  ,  "physinteg.1lag" , "physintlag.sq" , "urban.pct")) %>% set_caption("Table 4 cont.: Negative binomial models of number and duration of protests") %>% insert_row("", "", "Number", "", "", "Size", "", after = 0)
+
+theme_compact(t2b, header_rows = TRUE, header_cols = FALSE)
+
+#modelsummary::modelsummary(models, shape = term + estimate + component ~ model, coef_omit = "count", gof_omit = "R2", stars = T, longtable=T,  title = 'Binary portion of hurdle models of post-election protest', coef_rename =
+#               c("zero_(Intercept)" = "Intercept", "zero_jucon.1lag" = "Judical independence (lag)", "zero_legcon.1lag" = "Leg. constraints (lag)", 
+#                 "zero_gdpgro.1lag" = "GDP growth rate (lag)", "zero_altinf.1lag" = "Alternative info. (lag)", 
+#                 "zero_osorg.1lag" = "Civil soc. openness (lag)", "zero_econ.crisis.nelda.1lag1" = "Economic crisis", 
+#                 "zero_natlpartyorg.lag" = "National party organization (lag)", "zero_presidential1" = "Presidential election",
+#                 "zero_inc.termlimit.num" = "Incumbent term-limited",  
+#                 "zero_urban.pct" = "Urbanization", "zero_rentierstate" = "Rentier state", "zero_taxstate" = "Tax state",
+#                 "zero_n.events.pre" = "Number of pre-election protests", "zero_pubseccorrup.lag" = "Public sector corruption (lag)", "zero_v2elirreg.inv" = "Election fraud",
+#                 "zero_physinteg.1lag" = "Physical integrity (lag)", "zero_physintlag.sq" = "Physical integrity (lag), squared", "zero_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "zero_unscheduled_election" = "Off-schedule election")) # %>%
+#  add_header_above(c(" " = 1, "Protest quantity" = 3, "Maximum protest size" = 3))
+
 ```
@@
 
-The second set of models also challenge the conventional wisdom that more severe electoral manipulation is associated with an increased risk of anti-government protest, except under extraordinary circumstances.
-Beginning with the binary portion of the hurdle models, which model the likelihood of any protest occurring after the election, the standalone coefficient for election fraud in Model 2 is not statistically significant.
-Model 3 interacts election fraud with the incumbent's official vote-share; here we do observe statistically significant relationships at the p < .05 level.
+The second set of models challenge the idea that more severe electoral manipulation is associated with an increased risk of anti-government protest, except under extraordinary circumstances.
+Beginning with the binary portion of the hurdle models, which model the likelihood of any protest occurring after the election, the standalone coefficient for election fraud is not significant in Models 3.
+Model 4 interacts election fraud with the incumbent's official vote-share; here we do observe statistically significant relationships at the p < .05 level.
 However, as the marginal effects plot in Figure \@ref(fig:vote-share-prob-fig) shows, this relationship is only statistically significant when comparing extreme scenarios.
@@
 
-```{r zero-model-table, echo=FALSE}
-
-models <- list()
-models[['Model 2']] <- model.n.base.hurd.mm
-models[['Model 3']] <- model.n.base2.hurd.mm
-#models[['Model 4']] <- model.n.griev.hurd.mm
-models[['Model 4']] <- model.n.griev2.hurd.mm
-models[['Model 5']] <- model.n.griev3.hurd.mm
-
-modelsummary::modelsummary(models, coef_omit = "count", gof_omit = "R2", stars = T, longtable=T,  title = 'Binary portion of hurdle models of post-election protest', coef_rename = 
-               c("zero_(Intercept)" = "Intercept", "zero_jucon.1lag" = "Judical independence (lag)", "zero_legcon.1lag" = "Leg. constraints (lag)", 
-                 "zero_gdpgro.1lag" = "GDP growth rate (lag)", "zero_altinf.1lag" = "Alternative info. (lag)", 
-                 "zero_osorg.1lag" = "Civil soc. openness (lag)", "zero_econ.crisis.nelda.1lag1" = "Economic crisis", 
-                 "zero_natlpartyorg.lag" = "National party organization (lag)", "zero_presidential1" = "Presidential election",
-                 "zero_inc.termlimit.num" = "Incumbent term-limited",  
-                 "zero_urban.pct" = "Urbanization", "zero_rentierstate" = "Rentier state", "zero_taxstate" = "Tax state",
-                 "zero_n.events.massmob.pre" = "Number of pre-election protests", "zero_pubseccorrup.lag" = "Public sector corruption (lag)", "zero_v2elirreg.inv" = "Election fraud",
-                 "zero_physinteg.1lag" = "Physical integrity (lag)", "zero_physintlag.sq" = "Physical integrity (lag), squared", "zero_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "zero_unscheduled_election" = "Off-schedule election"))
-
-```
+To-do: Fix binary models so that they report same as hurdles
 
 ```{r include=FALSE}
-nelda_massmob <- nelda_massmob %>% mutate(any_protest = ifelse(n.events.massmob.post > 0, 1, 0))
+nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(n.events.post > 0, 1, 0))
+nelda2.sub <- nelda2.sub %>% mutate(any_size = ifelse(max_size > 0, 1, 0))
 
-model.noint.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda_massmob %>% filter(gdpgro.1lag < 1))
-summary(model.noint.binary)
+model.margin.binary_n <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age, family = binomial(link = "logit"), data = nelda2.sub)
+summary(model.margin.binary_n)
 
-model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda_massmob %>% filter(gdpgro.1lag < 1))
-summary(model.margin.binary)
+model.margin.binary_size <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose), family = binomial(link = "logit"), data = nelda2.sub)
+summary(model.margin.binary_size)
 
-model.binary.griev1 <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda_massmob %>% filter(gdpgro.1lag < 1))
-summary(model.binary.griev1)
+model.binary.griev1_n <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose), family = binomial(link = "logit"), data = nelda2.sub)
+summary(model.binary.griev1_n)
 
-model.binary.griev2 <- glm(any_protest ~  jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, data = nelda_massmob, family = binomial(link = "logit"))
-summary(model.binary.griev2)
+model.binary.griev1_size <- glm(any_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose), family = binomial(link = "logit"), data = nelda2.sub)
+summary(model.binary.griev1_size)
 
-model.binary.griev3 <- glm(any_protest ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election, data = nelda_massmob, family = binomial(link = "logit"))
-summary(model.binary.griev3)
+model.binary.repr_n <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose), family = binomial(link = "logit"), data = nelda2.sub)
+summary(model.binary.repr_n)
 
@@
 
-plot_binary_margin_max <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-2.3, 3.2]"))
-plot_binary_margin_max <- plot_binary_margin_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()
+#plot_binary_margin_max <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-2.3, 3.2]"))
+#plot_binary_margin_max <- plot_binary_margin_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()
 
-plot_binary_margin_typ <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]"))
-plot_binary_margin_typ <- plot_binary_margin_typ +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "", color = "Election fraud") + theme_bw()
+plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary_n, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]"))
+plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "", color = "Election fraud") + theme_bw()
+
+plot_binary_margin_typ_size <- sjPlot::plot_model(model.margin.binary_size, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]"))
+plot_binary_margin_typ_size <- plot_binary_margin_typ_size +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "", color = "Election fraud") + theme_bw()
 
@@
 
-plot_binary_gdp_max <- sjPlot::plot_model(model.binary.griev1, type = "pred", terms = c("gdpgro.1lag [all]", "v2elirreg.inv [-2.3, 3.2]"))
-plot_binary_gdp_max <- plot_binary_gdp_max
+plot_binary_gdp_typ_n <- sjPlot::plot_model(model.binary.griev1_n, type = "pred", terms = c("gdpgro.1lag [all]", "v2elirreg.inv [-.45, 1.78]"))
+plot_binary_gdp_typ_n <- plot_binary_gdp_typ_n +  labs(x = "GDP growth (lagged)", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()
 
-plot_binary_gdp_typ <- sjPlot::plot_model(model.binary.griev1, type = "pred", terms = c("gdpgro.1lag [all]", "v2elirreg.inv [-.45, 1.78]"))
-plot_binary_gdp_typ <- plot_binary_gdp_typ
+plot_binary_gdp_typ_size <- sjPlot::plot_model(model.binary.griev1_size, type = "pred", terms = c("gdpgro.1lag [all]", "v2elirreg.inv [-.45, 1.78]"))
+plot_binary_gdp_typ_size <- plot_binary_gdp_typ_size +  labs(x = "GDP growth (lagged)", y = "Predicted probability", title = "Predicted probability of any size protest", color = "Election fraud") + theme_bw()
 
-###Crisis
+## Repression
 
-plot_binary_crisis_max <- sjPlot::plot_model(model.binary.griev2, type = "pred", terms = c("v2elirreg.inv [all]", "econ.crisis.nelda.1lag"))
-plot_binary_crisis_max <- plot_binary_crisis_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()
+plot_binary_repr_typ_n <- sjPlot::plot_model(model.binary.repr_n, type = "pred", terms = c("physinteg.1lag [all]", "v2elirreg.inv [-.45, 1.78]"))
+plot_binary_repr_typ_n <- plot_binary_repr_typ_n +  labs(x = "Physical integrity (lagged)", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()
 
-plot_binary_crisis_typ <- sjPlot::plot_model(model.binary.griev2, type = "pred", terms = c("v2elirreg.inv [all]", "econ.crisis.nelda.1lag"))
-plot_binary_crisis_typ <- plot_binary_crisis_typ +  labs(x = "Election fraud", y = "Predicted probability", title = "", color = "Economic crisis") + theme_bw()
-
-##Corruption
-
-plot_binary_corruption_max  <- sjPlot::plot_model(model.binary.griev3, type = "pred", terms = c("pubseccorrup.lag", "v2elirreg.inv [-2.3, 3.2]"))
-plot_binary_corruption_max <- plot_binary_corruption_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()
-
-plot_binary_corruption_typ <- sjPlot::plot_model(model.binary.griev3, type = "pred", terms = c("pubseccorrup.lag", "v2elirreg.inv [-.45, 1.78]"))
-plot_binary_corruption_typ <- plot_binary_corruption_typ +  labs(x = "Public sector corruption", y = "Predicted probability", title = "", color = "Election fraud") + theme_bw()
 ```
 
-```{r vote-share-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
-plot.combined.prob <- cowplot::plot_grid(plot_binary_margin_max, plot_binary_margin_typ, ncol=1)
+```{r n-protests-gdp, fig.cap="Fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE}
+dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "count") 
+p.n.margin <- plot(dat)
+p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Number of events", title = "", color = "Election fraud") + theme_bw()
+```
+
+```{r gdp-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
+plot.combined.prob <- cowplot::plot_grid(plot_binary_margin_typ_n, p.n.margin, ncol=1)
 plot.combined.prob
@@
 
-```{r crisis-prob-fig, fig.cap="Fraud, economic crisis, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
-plot.combined.crisis.prob <- cowplot::plot_grid(plot_binary_crisis_max, plot_binary_crisis_typ, ncol=1)
-plot.combined.crisis.prob
-```
-
-```{r corrupt-prob-fig, fig.cap="Fraud, public sector corruption, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
-plot.combined.corrupt.prob <- cowplot::plot_grid(plot_binary_corruption_max, plot_binary_corruption_typ, ncol=1)
-plot.combined.corrupt.prob
+```{r n-protests-margin, fig.cap="Fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE}
+dat <- ggpredict(model.d.base2.hurd, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "count") 
+p.d.margin <- plot(dat)
+p.d.margin <- p.d.margin +  labs(x = "Incumbent vote-share", y = "Duration of wave (days)", title = "", color = "Election fraud") + theme_bw() + coord_cartesian(ylim = c(0, 400))
 ```
@@
 
-```{r n-protests-margin, fig.cap="Fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd", "v2elirreg.inv [-.45, 1.64]")) 
-p.n.margin <- plot(dat)
-p.n.margin + coord_cartesian(ylim = c(0, 70)) + labs(x = "Incumbent vote-share", y = "Number of events", title = "Predicted counts of protest events", color = "Election fraud") + theme_bw()
+```{r n-protests-crisis, fig.cap="Fraud, economic crisis, and number of protest events", echo=FALSE, message=FALSE}
+newdata <- data.frame(
+   jucon.1lag =  0.50,
+          legcon.1lag =  0.50,
+           osorg.1lag =  0.90,
+          altinf.1lag =  0.65,
+     natlpartyorg.lag =  0.76,
+     pubseccorrup.lag = -0.86,
+         presidential =     0,
+    inc.termlimit.num =  0,
+       physinteg.1lag =  0.59,
+        physintlag.sq =  0.41,
+            urban.pct = 50.52,
+         rentierstate =     0,
+             taxstate =     1,
+         n.events.pre =  0.00,
+ unscheduled_election =     0,
+             inc.lose =     0,
+ v2elirreg.inv = c(-.45, -.45, 1.64, 1.64),
+ econ.crisis.nelda.1lag = c(0, 1, 0, 1)
+)
 
-```
+phat <- predict(model.n.griev2.hurd.mm, newdata, type = "response", se.fit = T) 
+newdata$phat <- phat$fit
+newdata$phatse <- phat$se.fit
+newdata$conflow <- newdata$phat - 1.96*(newdata$phatse)
+newdata$confhi <- newdata$phat + 1.96*(newdata$phatse)
 
-```{r n-protests-crisis, fig.cap="Fraud, economic crisis, and number of protest events", echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.griev2.hurd.mm, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-.45, 1.64]")) 
-p.n.crisis <- plot(dat)
-p.n.crisis + coord_cartesian(ylim = c(0, 75)) + labs(x = "Economic crisis", y = "Number of events", title = "Predicted counts of protest events", color = "Election fraud") + theme_bw()
+newdata$conflow_exp <- exp(newdata$conflow)
+newdata$confhi_exp <- exp(newdata$confhi)
+
+p.n.crisis <- ggplot(data=newdata, aes(x=jitter(econ.crisis.nelda.1lag), y = phat, color = v2elirreg.inv)) +
+  geom_point() +
+  geom_linerange(aes(ymin = conflow, ymax = confhi))
+p.n.crisis  + labs(x = "Economic crisis", y = "Number of events", title = "Predicted counts of protest events", color = "Election fraud")  + theme_bw()
 
@@
 ```{r n-protests-corrupt, fig.cap="Fraud, public sector corruption, and number of protest events", echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.griev3.hurd.mm, terms = c("pubseccorrup.lag", "v2elirreg.inv [-.45, 1.64]")) 
+dat <- ggpredict(model.n.griev3.hurd.mm, terms = c("pubseccorrup.lag", "v2elirreg.inv [-.45, 1.64]"), type = "fe") 
 p.corr.n.margin <- plot(dat)
-p.corr.n.margin + coord_cartesian(ylim = c(0, 120)) + labs(x = "Public sector corruption", y = "Number of events", title = "Predicted counts of protest events", color = "Election fraud") + theme_bw()
+p.corr.n.margin + coord_cartesian(ylim = c(0, 50)) + labs(x = "Public sector corruption", y = "Number of events", title = "Predicted counts of protest events", color = "Election fraud") + theme_bw()
 
@@
 ```{r zero-model-table-count, echo=FALSE}
-modelsummary::modelsummary(models, coef_omit = "zero", gof_omit = "R2", longtable=T, stars = T, title = 'Hurdle models of post-election protest (counts)',
-                           coef_rename = 
-               c("count_(Intercept)" = "Intercept", "count_jucon.1lag" = "Judical independence (lag)", "count_legcon.1lag" = "Leg. constraints (lag)", 
-                 "count_gdpgro.1lag" = "GDP growth rate (lag)", "count_altinf.1lag" = "Alternative info. (lag)", 
-                 "count_osorg.1lag" = "Civil soc. openness (lag)", "count_econ.crisis.nelda.1lag1" = "Economic crisis", 
-                 "count_natlpartyorg.lag" = "National party organization (lag)", "count_presidential1" = "Presidential election",
-                 "count_inc.termlimit.num" = "Incumbent term-limited",  
-                 "count_urban.pct" = "Urbanization", "count_rentierstate" = "Rentier state", "count_taxstate" = "Tax state",
-                 "count_n.events.massmob.pre" = "Number of pre-election protests", "count_pubseccorrup.lag" = "Public sector corruption (lag)", "count_v2elirreg.inv" = "Election fraud",
-                 "count_physinteg.1lag" = "Physical integrity (lag)", "count_physintlag.sq" = "Physical integrity (lag), squared", "count_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "count_unscheduled_election" = "Off-schedule election"))
+#modelsummary::modelsummary(models, coef_omit = "zero", gof_omit = "R2", longtable=T, stars = T, title = 'Hurdle models of post-election protest (counts)',
+#                           coef_rename = 
+ #              c("count_(Intercept)" = "Intercept", "count_jucon.1lag" = "Judical independence (lag)", "count_legcon.1lag" = "Leg. constraints (lag)", 
+#                 "count_gdpgro.1lag" = "GDP growth rate (lag)", "count_altinf.1lag" = "Alternative info. (lag)", 
+#                 "count_osorg.1lag" = "Civil soc. openness (lag)", "count_econ.crisis.nelda.1lag1" = "Economic crisis", #
+#                 "count_natlpartyorg.lag" = "National party organization (lag)", "count_presidential1" = "Presidential election",
+#                 "count_inc.termlimit.num" = "Incumbent term-limited",  
+#                 "count_urban.pct" = "Urbanization", "count_rentierstate" = "Rentier state", "count_taxstate" = "Tax state",
+#                 "count_n.events.pre" = "Number of pre-election protests", "count_pubseccorrup.lag" = "Public sector corruption (lag)", "count_v2elirreg.inv" = "Election fraud",
+#                 "count_physinteg.1lag" = "Physical integrity (lag)", "count_physintlag.sq" = "Physical integrity (lag), squared", "count_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "count_unscheduled_election" = "Off-schedule election"))
+
+```
+
+# A further test: protest wave duration
+
+To examine the proposition that the information value of election fraud is low, relative to other indicators of regime and opposition strength, I employ Cox proportional-hazards models of protest-wave duration.
+That data in these models is limited to those cases where protests are initiated; in this sense, they represent the cases where opposition groups were most optimistic about challenging the government (or most aggrieved).
+In Table XX, coefficients with a negative sign indicate factors that are associated with longer-lived protest movements; those with a positive sign indicate factors associated with a speedier demise.
+As the models in Table XX show, more severe election fraud is not associated with longer-lasting protest (which would indicate a more determined opposition and/or a weaker ruling party).
+Instead, we observe that repressive capacity is the variable most consistently associated with protest duration\textemdash as repressive capacity eases initially, protest waves last longer; as repressive capacity diminishes further, protest waves again become shorter.
+This suggests that more open, democratic states are more likely to offer concessions to protesters.
+
+```{r, echo=FALSE, warning=FALSE}
+models3 <- list()
+models3[['Model 5']] <- cox1
+models3[['Model 6']] <- cox2
+models3[['Model 7']] <- cox3
+models3[['Model 8']] <- cox4
+
+modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Cox proportional hazard models of protest duration')
 
@@
 
-Altogether, the findings challenge the idea that the fear of protest will deter election-manipulation by ruling parties.
+Altogether, the findings challenge the idea that the fear of protest will help deter election-manipulation by ruling parties.
 To the extent that fraud is a signal of incumbent strength, it is a noisy and unreliable one.
@@
 By overlooking the poor quality of election fraud as a signal, or by treating protest primarily as a binary event, the existing literature on electoral protest has overemphasized the risk to incumbents of engaging in electoral manipulation.
-This results presented here show that fraud has no bearing on how numerous protest events are, and only little influence on protest initiation.
+The results presented here show that fraud has no bearing on how numerous protest events are, and only little influence on protest initiation.
 While pre-election protest is associated with reduced fraud, the substantive size of this relationship is very small.
@@
 These findings indicate that, rather than being deterred by protest risk, governments have an incentive to manipulate as much as possible in order to achieve a large official vote-share.
-The evidence suggests that democratic breakthroughs in non-democratic elections are rare and unlikely to be driven by the severity of manipulation---a gloomy picture for supporters of democratization---and that researchers should consider other mechanisms beyond mass protest in their models of electoral manipulation.
+The absence of protest risk as a guardrail is felt in both electoral authoritarian regimes and in electoral democracies\textemdash those democracies that are most at risk of democratic erosion.
+The evidence thus suggests that mass demonstrations in favor of advancing (or restoring) democratic elections are rare and unlikely to be driven by the severity of manipulation---a gloomy picture for supporters of democratization---and that researchers should consider other mechanisms beyond mass protest in their models of electoral manipulation.
 

### 2023-05-30 Output/Submission files/title_page.Rmd
Submission round two updates

---
title: "A missing guardrail: The failure of protest risk to deter election manipulation in non-democracies"
output:
    bookdown::pdf_document2: default
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
indent: true
header-includes: 
  - \usepackage{booktabs}
  - \usepackage{setspace}
abstract: "The risk of mass protest figures prominently in many models of election manipulation, as a deterrent to tampering with elections. Yet election manipulation is common outside liberal democracies, while protest is rare (and successful protest rarer still). Using count data on protest events from 647 elections from 1990 to 2012, this paper argues that the risk to incumbents has been overstated in prior work for two reasons. First, information revealed by election fraud is ambiguous; second, any such information quickly decays once protests do begin. The results show that the level of fraud has no bearing on the number of protests after an election, and is associated with protest initiation only in rare circumstances. This pessimistic finding implies that democracies are more fragile, and non-democracies more robust, than prior work predicts."
---

\begin{center}
Word count: 9,498
\end{center}

### 2023-05-30 Output/appendix_v4_part2.Rmd
Submission round two updates

@@
 library(stargazer)
-library(cowplot)
-library(pscl)
+library(ggeffects)
+library(glmmTMB)
 library(modelsummary)
-library(ggeffects)
-library(MASS)
+library(huxtable)
+library(simPH)
+library(survival)
 
 nelda_massmob <- read.csv(here::here("Data", "nelda_massmob_all.csv")) #nelda_complete.csv for only elex-related
+nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2023-05-22.csv"))
+
+##Bringing nelda2.sub variables over to massmob dataset
+nelda2_select <- nelda2.sub %>% dplyr::select(NeldaID, regime_age, v2elirreg.inv.lag)
+nelda_massmob <- nelda_massmob %>% rename(NeldaID = electionid)
+
+nelda_massmob <- nelda_massmob %>% left_join(nelda2_select, by = "NeldaID")
 
@@
 ## Removing cases where inc. loses
-nelda_massmob <- nelda_massmob %>% filter(inc.lose == 0)
+#nelda_massmob <- nelda_massmob %>% filter(inc.lose == 0)
 
@@
 
-## Alternative measures of election manipulation
+# Robustness check using Mass Mobilization dataset
 
-To ensure that the results presented in the main text are not biased by the chosen measure of election fraud, this section replaces that measure with a broader indicator of election integrity from V-Dem.
-This variable is `v2xel_frefair,' which is a comprehensive indicator of quality of the electoral process, including measures of electoral management body autonomy and capacity, vote-buying, intimidation, fraud, and an overall assessment of the fairness of the election.
-This variable ranges from 0 to 1, with higher values indicating freer and fairer elections.
-The models presented below utilize this variable in place of the main measure of election fraud, while retaining all other control and explanatory variables, using the Mass Mobilization data on protest.
+Not including duration or size data since MassMob data includes all protest, not just election focused and has less reliable size data. Will still use number.
 
-```{r manipdv, echo=FALSE}
-m1 <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   + presidential + inc.termlimit.num  +  libdem.1lag + urban.pct + rentierstate + taxstate + n.events.massmob.pre + median.participants.mm.pre  + (1 | stateid), data = nelda_massmob %>% filter(n.events.massmob.pre < 200)) #There are two extreme observations removed by this filter
+```{r}
+m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate + n.events.massmob.pre +  physinteg.1lag + physintlag.sq    + unscheduled_election + regime_age + (1 | stateid), data = nelda_massmob)
 
@@
 
-modelsummary(m1, stars = T,  title = 'Random-intercept multilevel model of overall election integrity')
+m1b <- lmer(v2elirreg.inv ~ v2elirreg.inv.lag + jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   + presidential   +  libdem.1lag + urban.pct + rentierstate + taxstate + n.events.massmob.pre +  physinteg.1lag + physintlag.sq  + unscheduled_election + regime_age+ (1 | stateid), data = nelda_massmob)
+summary(m1b)
 ```
 
-The first model shows that the number of pre-election protests is weakly associated with improved election integrity, but just as in the main text this effect is extremely small.
-Each additional protest is associated with an improvement in election integrity equivalent to 0.2\% of the standard deviation of the dependent variable.
-Given that the mean number of pre-election protests is six, this is an extremely small influence on the overall degree of election integrity.
+```{r}
+nelda_massmob <- nelda_massmob %>% mutate(n.events.massmob.pre_any = ifelse(n.events.massmob.pre > 0, 1, 0)) #Including the raw value of this variable becomes almost perfectly predictive of the zero-inflation portion of the model
 
-```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
-model.n.novoteshare.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    + v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +     + v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
-summary(model.n.novoteshare.hurd.mm)
-
-model.n.base.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    winner.share.upd + v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    winner.share.upd + v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
-summary(model.n.base.hurd.mm)
-
-model.n.base2.hurd.mm <- hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre |  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+model.n.base2.hurd.mm <- glmmTMB::glmmTMB(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election + regime_age + inc.lose,
+                                          zi = ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre_any + unscheduled_election + regime_age + inc.lose,
+                                          family= nbinom2,
+                                          data = nelda_massmob)
 summary(model.n.base2.hurd.mm)
 
-model.n.griev.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   gdpgro.1lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   gdpgro.1lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, dist = "negbin", data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+model.n.griev.hurd.mm <- glmmTMB::glmmTMB(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election + regime_age+ inc.lose,
+                                ziformula=~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre_any + unscheduled_election + regime_age + inc.lose,
+                                data = nelda_massmob,
+                                family=truncated_nbinom2)
 summary(model.n.griev.hurd.mm)
 
-model.n.griev2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob, dist = "negbin")
+```
+
+```{r zero-model-table, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+models_t1 <- list()
+models_t1[['Model 4']] <- model.n.base2.hurd.mm
+models_t1[['Model 5']] <- model.n.griev.hurd.mm
+#models_t1[['Model 5']] <- m_duration_count
+#models_t1[['Model 6']] <- model.d.base2.hurd
+#models_t1[['Model 7']] <- model.d.griev.hurd
+
+t1a <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, omit_coefs = c("rentierstate1", "taxstate1", "n.events.pre", "unscheduled_election1", "inc.lose1", "v2elirreg.inv:winner.share.upd", "gdpgro.1lag:v2elirreg.inv"), statistics = character(0)) %>% set_caption("Table 3: Zero-inflated models of protest initiation")
+
+theme_compact(t1a, header_rows = TRUE, header_cols = FALSE)
+
+t1b <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, omit_coefs = c("(Intercept)", "jucon.1lag" , "legcon.1lag" , "gdpgro.1lag" , "altinf.1lag" , "pubseccorrup.lag" , "osorg.1lag" , "econ.crisis.nelda.1lag1" , "natlpartyorg.lag" , "v2elirreg.inv" , "pubseccorrup.lag"   , "presidential1" , "inc.termlimit.num1"  ,  "physinteg.1lag" , "physintlag.sq" , "urban.pct")) %>% set_caption("Table 3 cont.: Zero-inflated models of protest initiation")
+
+theme_compact(t1b, header_rows = TRUE, header_cols = FALSE)
+
+models_t2 <- list()
+models_t2[['Model 4']] <- model.n.base2.hurd.mm
+models_t2[['Model 5']] <- model.n.griev.hurd.mm
+
+t2a <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), omit_coefs = c("rentierstate1", "taxstate1", "n.events.pre", "unscheduled_election1", "inc.lose1", "v2elirreg.inv:winner.share.upd", "gdpgro.1lag:v2elirreg.inv"), statistics = character(0)) %>% set_caption("Table 4: Negative binomial models of number of protests")
+
+theme_compact(t2a, header_rows = TRUE, header_cols = FALSE)
+
+t2b <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), omit_coefs = c("(Intercept)", "jucon.1lag" , "legcon.1lag" , "gdpgro.1lag" , "altinf.1lag" , "pubseccorrup.lag" , "osorg.1lag" , "econ.crisis.nelda.1lag1" , "natlpartyorg.lag" , "v2elirreg.inv" , "pubseccorrup.lag"   , "presidential1" , "inc.termlimit.num1"  ,  "physinteg.1lag" , "physintlag.sq" , "urban.pct")) %>% set_caption("Table 4 cont.: Negative binomial models of number of protests")
+
+theme_compact(t2b, header_rows = TRUE, header_cols = FALSE)
+
+#modelsummary::modelsummary(models, shape = term + estimate + component ~ model, coef_omit = "count", gof_omit = "R2", stars = T, longtable=T,  title = 'Binary portion of hurdle models of post-election protest', coef_rename =
+#               c("zero_(Intercept)" = "Intercept", "zero_jucon.1lag" = "Judical independence (lag)", "zero_legcon.1lag" = "Leg. constraints (lag)", 
+#                 "zero_gdpgro.1lag" = "GDP growth rate (lag)", "zero_altinf.1lag" = "Alternative info. (lag)", 
+#                 "zero_osorg.1lag" = "Civil soc. openness (lag)", "zero_econ.crisis.nelda.1lag1" = "Economic crisis", 
+#                 "zero_natlpartyorg.lag" = "National party organization (lag)", "zero_presidential1" = "Presidential election",
+#                 "zero_inc.termlimit.num" = "Incumbent term-limited",  
+#                 "zero_urban.pct" = "Urbanization", "zero_rentierstate" = "Rentier state", "zero_taxstate" = "Tax state",
+#                 "zero_n.events.pre" = "Number of pre-election protests", "zero_pubseccorrup.lag" = "Public sector corruption (lag)", "zero_v2elirreg.inv" = "Election fraud",
+#                 "zero_physinteg.1lag" = "Physical integrity (lag)", "zero_physintlag.sq" = "Physical integrity (lag), squared", "zero_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "zero_unscheduled_election" = "Off-schedule election")) # %>%
+#  add_header_above(c(" " = 1, "Protest quantity" = 3, "Maximum protest size" = 3))
+
+```
+
+Figures \@ref(fig:n-protests-margin-max), \@ref(fig:n-protests-crisis-max), and \@ref(fig:n-protests-corrupt-max) illustrate the marginal effects of extreme values for election fraud on the expected number of post-election protests, based on Models 4 - 6 in the main text.
+They demonstrate that there is no significant relationship between the severity of fraud and the size of post-election protest waves for any of the moderator variables\textemdash incumbent vote-share, economic crisis, or public sector corruption.
+
+```{r, include=FALSE}
+model.n.base2.hurd.mm <- hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre |  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+summary(model.n.base2.hurd.mm)
+
+model.n.griev.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, dist = "negbin", data = nelda_massmob %>% filter(gdpgro.1lag < 1))
+summary(model.n.griev.hurd.mm)
+
+model.n.griev2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob, dist = "negbin")
 summary(model.n.griev2.hurd.mm)
 
-model.n.griev3.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   pubseccorrup.lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre  | jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   pubseccorrup.lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, data = nelda_massmob, dist = "negbin")
+model.n.griev3.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre  | jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, data = nelda_massmob, dist = "negbin")
 summary(model.n.griev3.hurd.mm)
@@
 
-Moving to the hurdle models of post-election protest, we do not observe much of a relationship between overall election integrity and protest.
-There is no statistically significant relationship between election integrity and protest occurrence in Model 2.
-Neither is there any significant relationship in interaction with the winner's vote-share, as shown in Figure \@ref(fig:b-protests-margin), or with economic crisis, as in Figure \@ref(fig:b-protests-crisis).
-The coefficient on the interaction term for public sector corruption and overall election integrity is statistically significant, though there is no significant effect at typical values for covariates and election integrity (as shown in Figure \@ref(fig:b-protests-corrupt)).
-
-In the corresponding results for the number of post-election protests, there is similarly muted evidence for a connection between the overall level of manipulation and protest risk to incumbents.
-There is no significant difference between the marginal effect of common levels of election integrity on the number of post-election protests for typical cases across any margin of victory (Figure \@ref(fig:n-protests-margin)).
-During an economic crisis, *cleaner* elections are associated with an increase in the number of protests (\@ref(fig:n-protests-crisis)).
-This is plausibly due to the crisis weakening both the government's ability to generate manipulation [@greene_why_2007] and its ability to co-opt or put down the protests.
-Finally, public sector corruption is associated with a significant increase in the marginal effect of poor-integrity elections on protest\textemdash but only when the corruption measure is at its maximum (Figure \@ref(fig:n-protests-corrupt)).
-
-As a whole, the results are in accordance with the theory proposed in the main paper.
-There is no consistent evidence that lower-quality elections are associated with either a higher risk of protest initiation, or with a larger number of post-election protests.
-
-```{r zero-model-table, echo=FALSE}
-
-models <- list()
-models[['Model 2']] <- model.n.base.hurd.mm
-models[['Model 3']] <- model.n.base2.hurd.mm
-#models[['Model 4']] <- model.n.griev.hurd.mm
-models[['Model 4']] <- model.n.griev2.hurd.mm
-models[['Model 5']] <- model.n.griev3.hurd.mm
-
-modelsummary::modelsummary(models, coef_omit = "count", gof_omit = "R2", stars = T, longtable=T,  title = 'Binary portion of hurdle models of post-election protest')
+```{r n-protests-margin-max, fig.cap="Fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE, out.width = '70%', fig.align='center'}
+dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd", "v2elirreg.inv [-2.3, 3.2]")) 
+p.n.margin <- plot(dat)
+p.n.margin + coord_cartesian(ylim = c(0, 70)) + labs(x = "Incumbent vote-share", y = "Number of events", title = "Predicted counts of protest events", color = "Election fraud") + theme_bw()
 
@@
 
-```{r zero-model-table-count, echo=FALSE}
-modelsummary::modelsummary(models, coef_omit = "zero", gof_omit = "R2", longtable=T, stars = T, title = 'Hurdle models of post-election protest (counts)')
+```{r n-protests-crisis-max, fig.cap="Fraud, economic crisis, and number of protest events", echo=FALSE, message=FALSE, out.width = '70%', fig.align='center'}
+dat <- ggpredict(model.n.griev2.hurd.mm, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-2.3, 3.2]")) 
+p.n.crisis <- plot(dat)
+p.n.crisis + coord_cartesian(ylim = c(0, 125)) + labs(x = "Economic crisis", y = "Number of events", title = "Predicted counts of protest events", color = "Election fraud") + theme_bw()
 
@@
 
-```{r, echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd", "v2xel_frefair [.2, .7]")) 
-p.n.margin <- plot(dat)
-p.n.margin <- p.n.margin  + theme_bw()
-
-dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd", "v2xel_frefair [.2, .7]"), type = "zi_prob") 
-p.b.margin <- plot(dat)
-p.b.margin <- p.b.margin  + theme_bw()
+```{r n-protests-corrupt-max, fig.cap="Fraud, public sector corruption, and number of protest events", echo=FALSE, message=FALSE, out.width = '70%', fig.align='center'}
+dat <- ggpredict(model.n.griev3.hurd.mm, terms = c("pubseccorrup.lag", "v2elirreg.inv [-2.3, 3.2]")) 
+p.corr.n.margin <- plot(dat)
+p.corr.n.margin + coord_cartesian(ylim = c(0, 120)) + labs(x = "Public sector corruption", y = "Number of events", title = "Predicted counts of protest events", color = "Election fraud") + theme_bw()
 
@@
 
-```{r n-protests-margin, fig.cap="Fraud, public sector corruption, and number of protest events"}
+\newpage
+
+# Removing controls to address multicollinearity
+
+The models in the main text include a suite of control variables aimed at controlling for common causes of election manipulation and protest, in order to minimize the risks of confounding and omitted variable bias.
+However, this approach runs the risk that individual coefficients may be highly sensitive to model specification due to multicollinearity.
+In particular, many of the control variables\textemdash such as *legislative constraints*, *judicial independence*, and *physical integrity* are likely to be correlated with pre-election protest.
+To address this risk, the models in Tables \@ref(tab:zero-model-table-small) and \@ref(tab:zero-model-table-count-small) exclude non-electoral control variables.
+That is, they include variables related to the election itself, the measure of *pre-election protest*, and the explanatory variables from the main models.
+
+The results echo those from the main text.
+Severity of election fraud is not associated with an increased number of protests, conditional on incumbent vote-share, economic crisis, or public-sector corruption.
+Marginal effects for interaction terms are presented in the following figures.
+
+```{r models-hurdlesmall-mm, warning = FALSE, message = FALSE, include = F}
+model.small.novoteshare.hurd.mm <- hurdle(n.events.massmob.post ~ v2elirreg.inv + presidential + inc.termlimit.num  + n.events.massmob.pre + unscheduled_election |  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.massmob.pre + unscheduled_election , data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+summary(model.small.novoteshare.hurd.mm)
+
+model.small.base.hurd.mm <- hurdle(n.events.massmob.post ~  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.massmob.pre + unscheduled_election + winner.share.upd |  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.massmob.pre + unscheduled_election + winner.share.upd , data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+summary(model.small.base.hurd.mm)
+
+model.small.base2.hurd.mm <- hurdle(n.events.massmob.post ~  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.massmob.pre + unscheduled_election + winner.share.upd + winner.share.upd*v2elirreg.inv | v2elirreg.inv + presidential + inc.termlimit.num  + n.events.massmob.pre + unscheduled_election + winner.share.upd  + winner.share.upd*v2elirreg.inv, data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
+summary(model.small.base2.hurd.mm)
+
+model.small.griev2.hurd.mm <- hurdle(n.events.massmob.post ~  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.massmob.pre + unscheduled_election + winner.share.upd + econ.crisis.nelda.1lag*v2elirreg.inv | v2elirreg.inv + presidential + inc.termlimit.num  + n.events.massmob.pre + unscheduled_election + winner.share.upd + econ.crisis.nelda.1lag*v2elirreg.inv , data = nelda_massmob, dist = "negbin")
+summary(model.small.griev2.hurd.mm)
+
+model.small.griev3.hurd.mm <- hurdle(n.events.massmob.post ~ v2elirreg.inv + presidential + inc.termlimit.num  + n.events.massmob.pre + unscheduled_election + winner.share.upd + pubseccorrup.lag*v2elirreg.inv   |   v2elirreg.inv + presidential + inc.termlimit.num  + n.events.massmob.pre + unscheduled_election + winner.share.upd + pubseccorrup.lag*v2elirreg.inv , data = nelda_massmob, dist = "negbin")
+summary(model.small.griev3.hurd.mm)
+```
+\newpage
+
+```{r zero-model-table-small, echo=FALSE}
+
+models <- list()
+models[['Model 11']] <- model.small.base.hurd.mm
+models[['Model 12']] <- model.small.base2.hurd.mm
+#models[['Model ']] <- model.n.griev.hurd.mm
+models[['Model 13']] <- model.small.griev2.hurd.mm
+models[['Model 14']] <- model.small.griev3.hurd.mm
+
+modelsummary::modelsummary(models, coef_omit = "count", gof_omit = "R2", stars = T, longtable=T,  title = 'Binary portion of hurdle models of post-election protest', coef_rename = 
+               c("zero_(Intercept)" = "Intercept", "zero_presidential1" = "Presidential election",
+                 "zero_inc.termlimit.num" = "Incumbent term-limited",  
+                                  "zero_n.events.massmob.pre" = "Number of pre-election protests", "zero_pubseccorrup.lag" = "Public sector corruption (lag)", "zero_v2elirreg.inv" = "Election fraud", "zero_unscheduled_election" = "Off-schedule election", "zero_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "winner.share.upd" = "Incumbent vote-share",
+                 "zero_econ.crisis.nelda.1lag1" = "Economic crisis", "econ.crisis.nelda.1lag1" = "Economic crisis",
+                 "pubseccorrup.lag" = "Public sector corruption"))
+
+```
+
+\newpage
+
+```{r zero-model-table-count-small, echo=FALSE}
+modelsummary::modelsummary(models, coef_omit = "zero", gof_omit = "R2", longtable=T, stars = T, title = 'Hurdle models of post-election protest (counts)',
+                           coef_rename = 
+              c("count_(Intercept)" = "Intercept", "count_presidential1" = "Presidential election",
+                 "count_inc.termlimit.num" = "Incumbent term-limited",  
+                                  "count_n.events.massmob.pre" = "Number of pre-election protests", "count_pubseccorrup.lag" = "Public sector corruption (lag)", "count_v2elirreg.inv" = "Election fraud", "count_unscheduled_election" = "Off-schedule election", "count_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "winner.share.upd" = "Incumbent vote-share",
+                 "count_econ.crisis.nelda.1lag1" = "Economic crisis", "econ.crisis.nelda.1lag1" = "Economic crisis",
+                 "pubseccorrup.lag" = "Public sector corruption"))
+
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+dat <- ggpredict(model.small.base2.hurd.mm, terms = c("winner.share.upd", "v2elirreg.inv [-.44, 1.77]")) 
+p.n.margin <- plot(dat)
+p.n.margin <- p.n.margin + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events") + coord_cartesian(ylim = c(0, 100)) + theme_bw()
+
+dat <- ggpredict(model.small.base2.hurd.mm, terms = c("winner.share.upd", "v2elirreg.inv [-.44, 1.77]"), type = "zi_prob") 
+p.b.margin <- plot(dat)
+p.b.margin <- p.b.margin + labs(x = "Incumbent vote-share", y = "Predicted probability of protest" , title = "Predicted probability of any protest")  + theme_bw()
+
+```
+
+```{r n-protests-margin-small, fig.cap="Election integrity, incumbent vote-share, and number of protest events", echo=FALSE, warning=FALSE, out.width = '70%', fig.align='center'}
 p.n.margin
@@
 
-```{r b-protests-margin, fig.cap="Fraud, public sector corruption, and likelihood of protest"}
+```{r b-protests-margin-small, fig.cap="Election integrity, incumbent vote-share, and likelihood of protest", echo=FALSE, warning=FALSE, out.width = '70%', fig.align='center'}
 p.b.margin
@@
 
-```{r, echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.griev2.hurd.mm, terms = c("econ.crisis.nelda.1lag", "v2xel_frefair [.15, .66]")) 
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+dat <- ggpredict(model.small.griev2.hurd.mm, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-.44, 1.77]")) 
 p.n.crisis <- plot(dat)
-p.n.crisis <- p.n.crisis  + theme_bw()
+p.n.crisis <- p.n.crisis + labs(x = "Economic crisis", y = "Predicted number of protests" , title = "Predicted count of protest events") + theme_bw()
 
-dat <- ggpredict(model.n.griev2.hurd.mm, terms = c("econ.crisis.nelda.1lag", "v2xel_frefair [.15, .66]"), type = "zi_prob") 
+dat <- ggpredict(model.small.griev2.hurd.mm, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-.44, 1.77]"), type = "zi_prob") 
 p.b.crisis <- plot(dat)
-p.b.crisis <- p.b.crisis +  theme_bw()
+p.b.crisis <- p.b.crisis + labs(x = "Economic crisis", y = "Predicted probability of protest" , title = "Predicted probability of any protest") +  theme_bw()
 
 ```
-```{r n-protests-crisis, fig.cap="Fraud, economic crisis, and number of protest events"}
+
+```{r n-protests-crisis-small, fig.cap="Election integrity, economic crisis, and number of protest events", echo=FALSE, warning=FALSE, out.width = '70%', fig.align='center'}
 p.n.crisis
@@
 
-```{r b-protests-crisis, fig.cap="Fraud, economic crisis, and likelihood of protest"}
+```{r b-protests-crisis-small, fig.cap="Election integrity, economic crisis, and likelihood of protest", warning = FALSE, echo=FALSE, out.width = '70%', fig.align='center'}
 p.b.crisis
@@
 
-```{r, echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.griev3.hurd.mm, terms = c("pubseccorrup.lag", "v2xel_frefair [.29, .66]")) 
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+dat <- ggpredict(model.small.griev3.hurd.mm, terms = c("pubseccorrup.lag", "v2elirreg.inv [-.44, 1.77]")) 
 p.corr.n.margin <- plot(dat)
-p.corr.n.margin <- p.corr.n.margin + coord_cartesian(ylim = c(0, 100)) + theme_bw()
+p.corr.n.margin <- p.corr.n.margin + coord_cartesian(ylim = c(0, 100)) + labs(x = "Public sector corruption", y = "Predicted number of protests" , title = "Predicted count of protest events") + theme_bw()
 
-dat <- ggpredict(model.n.griev3.hurd.mm, terms = c("pubseccorrup.lag", "v2xel_frefair [.29, .66]"), type = "zi_prob") 
+dat <- ggpredict(model.small.griev3.hurd.mm, terms = c("pubseccorrup.lag", "v2elirreg.inv [-.44, 1.77]"), type = "zi_prob") 
 p.corr.b.margin <- plot(dat)
-p.corr.b.margin <- p.corr.b.margin  + theme_bw()
+p.corr.b.margin <- p.corr.b.margin + labs(x = "Public sector corruption", y = "Predicted probability of protest" , title = "Predicted probability of any protest") + theme_bw()
 
@@
 
-```{r n-protests-corrupt, fig.cap="Fraud, public sector corruption, and number of protest events"}
+```{r n-protests-corrupt-small, fig.cap="Election integrity, public sector corruption, and number of protest events", echo=FALSE, warning=FALSE, out.width = '70%', fig.align='center'}
 p.corr.n.margin
@@
 
-```{r b-protests-corrupt, fig.cap="Fraud, public sector corruption, and likelihood of protest"}
+```{r b-protests-corrupt-small, fig.cap="Election integrity, public sector corruption, and likelihood of protest", echo=FALSE, warning=FALSE, out.width = '70%', fig.align='center'}
 p.corr.b.margin
 ```
+
+\newpage
+
+# References

### 2023-06-02 Output/Submission files/manuscript_anon_v2.Rmd
Revisions to text

@@
 In response, hundreds of thousands of Belarusians took to the streets over the course of several months.
-Though the protests put significant strain on the regime, a harsh campaign of repression ultimately quelled the protests and President Lukashenko was inaugurated into a new five-year term.
+Though the protests put significant strain on the regime, a harsh campaign of repression ultimately quelled the protests and Lukashenko was inaugurated into a new five-year term.
 The Belarusian case offers a useful illustration of a central puzzle in the study of election manipulation: why did the risk of costly protest fail to deter the incumbent from engaging in such widespread fraud?
@@
 For this reason, many scholars---especially those working in the formal modeling tradition---have identified the risk that election manipulation will spark protest as a major deterrent to regimes working to engineer election outcomes.
-Most often, though, non-democratic states can rely on superior resources, elite solidarity, and repressive capacity to quash electoral protests.
+Often, though, incumbents can rely on superior resources, elite solidarity, and repressive capacity to quash electoral protests.
 How threatening, then, are electoral protests really, and can that threat constrain electoral manipulation by governments?
@@
 In this paper, I argue that protest risk rarely serves as a deterrent to election manipulation for two reasons.
-First, election fraud is carried out both by strong, cohesive ruling parties, and by weakened parties seeking to cling to power.
+First, election fraud is not only carried out by the weak types assumed in some formal models, but also by strong, cohesive ruling parties.
 The ambiguity of the electoral environment outside liberal democracies, where political and civil liberties may not be fully protected, makes it difficult for opposition actors to correctly distinguish between the two.
@@
 In most cases, the risk of protest has both been conceptualized and operationalized as a binary outcome: protests occur or they do not.
-This approach obscures the importance of the number of electoral protests in a ruling party's calculation of the risk involved in manipulation an election.
-Small, short-lived protests may be easily contained by a well-resourced government, while a long-lasting movement with a large number of protest events poses a more significant challenge.
-
-Using cross-national data on 647 elections from 1990 to 2012 where incumbents claimed victory, combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Mass Mobilization dataset [@massmob], this paper finds little support for the hypothesis that more fraudulent elections are associated with a higher risk that any post-election protest event occurs.
-Very highly fraudulent elections are associated with increased risk of protest initiation when incumbent vote-shares are small, in keeping with the formal-modeling literature, which predicts that overreach in the form of election fraud indicates government weakness and makes opposition groups confident of winning a conflict in the streets.
-And, in keeping with the social-grievance literature, fraud is associated with protest initiation in societies with worse public sector corruption.
-However, these effects are largely confined to extreme scenarios, and there is no statistically significant association between the level of election fraud and the number of protest events that occur.
-This suggests a more complex relationship between election fraud and protest, undermining the argument that protest risk is a major deterrent to electoral malfeasance.
-
-Since fraud is not associated with the overall number of events in a protest wave, the risk to governments has likely been overemphasized.
+This approach obscures the importance of the number and size of electoral protests in a ruling party's calculation of the risk involved in manipulation an election.
+Small, narrow protests may be easily contained by a well-resourced government, while more numerous, larger protest events pose a more significant challenge.
+
+Using cross-national data on 1110 elections from 1989 to 2012 where incumbents claimed victory, combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Electoral Contention and Vioence dataset [@Daxecker2019], this paper finds little support for the hypothesis that more fraudulent elections are associated with a higher risk that any post-election protest event occurs.
+Furthermore, it shows that more severe election fraud is not associated with more numerous or larger post-election protests.
+Instead, these protest dynamics appear driven more by structural and institutional factors, including the size of the political opposition and whether or not the incumbent chief executive is term limited.
+
+Since fraud is not associated with the overall number or size of of events in a protest wave, the risk to governments has likely been overemphasized.
 That is, even if protests do break out in response to large-scale fraud, the government has many tools it may deploy to curtail them.
@@
 Broadly speaking, prior work has considered two major explanations for the severity of manipulated elections: the costs and benefits of manipulation to the ruling party and government (a demand-side framework) and the costs and benefits for the front-line actors who do the work of manipulating elections (a supply-side framework).
-The risk of costly protest is the central, though not sole^[In some studies, like [@birch2011], the costs of manipulation are attributed to a more diffuse loss of legitimacy; such costs **may** manifest as protest, but may also manifest as a reduction in general cooperation with the regime.
-However, it is worth noting the argument of [@gehlbach2015electoral] that election manipulation can **induce** cooperative behavior from some agents], factor which affects the demand-side cost of manipulation in most models.
+The risk of costly protest is the central, though not sole^[In some studies, like [@birch2011], the costs of manipulation are attributed to a more diffuse loss of legitimacy; such costs *may* manifest as protest, but may also manifest as a reduction in general cooperation with the regime.
+However, it is worth noting the argument of [@gehlbach2015electoral] that election manipulation can *induce* cooperative behavior from some agents], factor which affects the demand-side cost of manipulation in most models.
 This insight can be found, in related forms, in both the literature on self-enforcing democracy and on the conduct of non-democratic elections.
@@
 1662].
-Such arrangements constitute the large majority of regimes today, ranging from the electoral authoritarian to the liberal democratic.
-While many previous works argue that the risk of protest can mitigate election manipulation in regimes where elections are held\textemdash including established democracies [@millerRepublicIfYou2021], electoral authoritarian regimes [@Egorov2021], and any regime where elections are held but may be manipulated [@Little2015;@Luo2020]\textemdash, this body of work has yet to grapple with the empirical finding that election integrity is relatively poor across these regimes outside of a narrow club of 'consolidated' democracies.
-
-There are thought to be two mechanisms by which election manipulation leads to protest, and thereby restrains manipulation.
+Such arrangements are not limited to liberal democracies, and in fact constitute the large majority of regimes today.
+Many previous works argue that the risk of protest can mitigate election manipulation in regimes where elections are held\textemdash including established democracies [@millerRepublicIfYou2021], electoral authoritarian regimes [@Egorov2021], and any regime where elections are held but may be manipulated [@Little2015;@Luo2020]\textemdash but this body of work has yet to grapple with the empirical finding that election integrity is relatively poor across these regimes outside of a narrow club of 'consolidated' democracies.
+
+There are thought to be two mechanisms by which election manipulation leads to protest, and thereby might restrain manipulation.
 The first is through revealing information about the strength of the government or ruling party.
@@
 This logic is often seen in formal models [@Egorov2021;@chernykh2015third].
-Several prominent models of electoral manipulation argue that the risk of mass post-election protest helps restrain incumbents who might otherwise prefer larger levels of manipulation [@magaloni2010; @Little2015; @Little2012; @birch2011; @Fearon2011;@Luo2020;@Rozenas2016b].
+Several prominent models of electoral manipulation argue that the risk of mass post-election protest helps restrain incumbents who might otherwise prefer larger levels of manipulation [@magaloni2010; @Little2015; @Little2012; @birch2011; @fearon_self-enforcing_2011;@Luo2020;@Rozenas2016b].
 The risk of mass protest in these models is assumed to constitute a significant guardrail for election integrity in democracies, and a significant challenge for ruling parties in non-democracies.
@@
 Despite this consensus in the formal literature, there is relatively little empirical evidence to support the claim that manipulation per se indicates weakness, and thus increases protest risk.
-For example, looking at the 2011 election and protest cycle in Russia, Ananyev and Poyker [-@ananyevDictatorsSignalStrength2022] show that post-election protest was not associated with the severity of fraud using several measures.^[Though, it should be noted that Lankina and Skovoroda [-@Lankina2017] show that a different measure of fraud was associated with post-election protest in the subsequent presidential election.] In a cross-national study, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e.
+For example, looking at the 2011 election and protest cycle in Russia, Ananyev and Poyker [-@ananyevDictatorsSignalStrength2022] show that post-election protest was not associated with the severity of fraud using several measures.^[It should be noted that Lankina and Skovoroda [-@Lankina2017] show that a different measure of fraud was associated with post-election protest in the subsequent presidential election.] In a cross-national study, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e.
 whether opponents are allowed to run, whether the media environment is pro-regime), rather than by the election-day manipulation that usually figures in theories of election integrity and protest.
+This result is consistent with the finding that parties are more likely to protest elections after ex ante changes to the rules [@chernykh2014].
 
@@
 In this way, large-scale manipulation can contribute to a ruling party's ability to produce a large, stabilizing majority vote-share [@magaloni2006].
-Similarly, Luo and Rozenas [-@Luo2020] posit that ex post fraud is more likely to result in protest than ex ante manipulation, though this stands in contrast with the finding that parties are more likely to protest elections after ex ante changes to the rules [@chernykh2014].
-In sum, it is likely that manipulation *may* signal regime weakness, but empirical works suggests that this is more likely when margins of victory are narrow and when falsification is a major tool; by contrast, other work suggests that manipulation may at times signal strength.
+Similarly, Luo and Rozenas [-@Luo2020] posit that ex post fraud is more likely to result in protest than ex ante manipulation.
+In sum, election manipulation can be performed by strong regimes for reasons that go beyond 'mere winning.'
 
@@
 
-Without disputing that election manipulation conveys information, a second school of thought has emphasized the ways in which biased elections may predispose individuals to engage in protest [@kuntz2009].
-In this view, election manipulation acts as a shared grievance among individuals who prefer to see the ruling party lose---especially when it appears that fraud may have been decisive [@kuntz_more_2009].
+A second school of thought has emphasized the ways in which biased elections may predispose individuals to engage in protest [@kuntz_more_2009].
+In this view, election manipulation acts as a shared grievance among individuals who prefer to see the ruling party lose\textemdash especially when it appears that fraud may have been decisive [@kuntz_more_2009].
 Belief in a stolen election is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007], possibly initiating a protest cascade [@Kuran1991].
@@
 
-The grievance-based approach poses a challenge for the deterrence logic, in that protests hinge more on citizens' subjective beliefs about fraud than on the actual level of malfeasance.
-A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b], which suggests that emotions, moral condemnation, or grievances take precedence over strategic calculations for individuals.
+The grievance-based approach suggests that protests hinge more on citizens' subjective beliefs about fraud and the broader social context than on the actual level of malfeasance.
+A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b].
 Subjective belief in fraud was also shown to be correlated with approval of mass protests after Russia's 2011 election [@Chaisty2013].
 Furthermore, simply being on the losing side in an election is associated with reduced belief that the election was fair [@Cantu2015], and with increased support for electoral protest [@Anderson2006].
-These studies suggest that electoral protest can occur at low or high levels of fraud, contingent on other factors.
+These studies suggest that electoral protest can occur at low or high levels of fraud, contingent on other factors\textemdash including subjective, and perhaps mistaken, beliefs.
 
@@
 
-Despite widespread election manipulation, electoral protest is relatively rare; even in cases where it does occur, most efforts to overturn fraudulent election results through protest have failed [@kalandadze2009electoral].
+Finally, despite widespread election manipulation, electoral protest is relatively rare; even in cases where it does occur, most efforts to overturn fraudulent election results through protest have failed [@kalandadze2009electoral].
 The apparent failure of protest risk to deter election fraud underlines our limited understanding of non-democratic ruling parties' decision-making.
@@
 Elections in non-democracies are *generally* risky for incumbents, regardless of how free and fair they may be, as they create an opportunity for opposition collective action [@Knutsen2017] and reveal information about the government and society [@Pop-Eleches2015].
-Scholarly attention might turn more profitably toward supply-side explanations for manipulation, such as the organizational capacity of the state [@Simpser2011], the popularity of the government [@rundlett2016deliver], the risks faced by front-line agents [@Harvey2022].
+In most cases, however, election manipulation adds little to this risk.
 
@@
 
-## Predictions from prior work: Fraud and the risk of protest
+## Predictions from prior work: Fraud and the risk of protest initiation
 
 As noted above, much of the formal literature on election manipulation assumes that the risk of protest serves as a deterrent to election manipulation.
-Given that pre-election protest is likely to be a strong predictor of post-election protest\textemdash by indicating strong social networks [@mateo2022], pre-existing opposition groups and civil society capacity [@Bunce2006], and/or popular grievances [@Brancati2016]\textemdash the deterrent logic suggests that non-democratic governments will pull back on election-day fraud when the pre-election environment is characterized by protest [@Rozenas2016b].
-While it is possible that pre-election protest may provide an opportunity for incumbents to crack down, creating more space for manipulating the election, more active pre-election protest suggests that opposition capacity is more robust, all else held equal.
-Moreover, in general, incumbents tend to avoid mass repression in the pre-election phase in favor of targeted harassment of opposition activists [@Bhasin2013].
-The prior literature thus implies the following hypothesis.
-
->H1: Great pre-election protest risk will be associated with lower levels of election-day manipulation.
-
 In addition, the logic of deterrence suggests that increasingly severe election fraud should make post-election more likely, either unconditionally or when the incumbent's margin of victory is low.
@@
 
+>H1: Greater pre-election protest risk will be associated with lower levels of election-day manipulation.
+
 >H2: The likelihood of protest initiation will increase as the severity of election fraud increases.
@@
 >H4: The effect of election manipulation on the likelihood of protest initiation will increase as economic grievances rise.
-
-[My theory here:  Election manipulation (fraud?) cannot serve as a reliable indicator of strength or weakness, meaning that opposition groups cannot use it to behave strategically.
-Strong and weak alike engage in it (noted already).
-Also, election manipulation or its absence can be deployed by regimes for varying reasons.
-A strong regime may feel comfortable dialing back on fraud to gain more information (find cites for this); a weak one may not be able to muster the resources.
-A strong regime may generate lots of manipulation in order to flaunt its capacity; a weak regime may need to rely on fraud to boost otherwise anemic public support.
-So because it doesn't communicate information about strength, it doesn't influence protest; protests happen or not--and live or die--based on the balance of resources on either side.]
 
@@
 In either case, these factors create an incentive for opposition protest whether electoral fraud is minor or severe\textemdash a kind of case of 'crying wolf'\textemdash an outcome that is also posited in some of the formal modeling literature [@Little2015].
-As a result, it is likely that the severity of fraud is thus untethered from protest risk; what matters is citizens' underlying estimate of the regime's strength.
+As a result, it is likely that the severity of fraud has little bearing on protest risk; what matters is the underlying balance of strength between the incumbent regime and its opposition.
 
@@
 
-Finally, and perhaps most importantly, by emphasizing binary protest outcomes at the expense of trends over time, prior work has overemphasized the risk of popular backlash to incumbents.
+While it is useful to check prior predictions on protest initiation, by emphasizing binary protest outcomes at the expense of trends over time, prior work has overemphasized the risk of popular backlash to incumbents.
 While many incumbents may indeed prefer a quiet post-election period, the occurrence of protest can also have regime-sustaining benefits.
@@
 The initial emergence of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014].
-That election dates are known in advance creates opportunities for opposition mobilization; anticipating this, regimes can more easily engage in pre-emptive repression [@Truex2019;@Sullivan2016;@Bhasin2013] \textemdash Knutsen et al [-@Knutsen2017] argue that regimes become more effective at these sorts of behaviors over time.
+That election dates are known in advance creates opportunities for opposition mobilization; anticipating this, regimes can more easily engage in pre-emptive repression [@Truex2019;@Sullivan2016;@Bhasin2013].
 
@@
 Though such crackdowns risk provoking a backlash, they are aimed at demobilizing the opposition by increasing the costs of participation in overt dissent [@DeVogel2022].
-In other cases, the government may attempt to accommodate the opposition, by making concessions on policy or institutions [@Moore2000;@Hafner-Burton2016;@Piven1979].
+In other cases, the government may attempt to accommodate the opposition, by making concessions on policy or institutions [@mooreRepressionDissentSubstitution2000;@Hafner-Burton2016;@Piven1979].
 Alternatively, concessions may be aimed at the more moderate public, in order to keep them on the sidelines of the conflict [@Hummel2019].
-
 In addition, governments may learn from the experiences of other governments, just as opposition groups may learn from opposition.
@@
 High-capacity states, as a result, are more likely to be able to engage in successful repression [@tarrowPowerMovementSocial2011].
-Here I expect to see support for the so-called 'more murder in the middle' framework [@feinLifeIntegrityViolationsDemocracy1995], in which the regimes with the most demonstrated repressive capacity deter potential challengers from taking to the streets in the first place, and the least repressive states defuse protest through accommodation [@pierskallaProtestDeterrenceEscalation2010;@mooreRepressionDissentSubstitution2000].
 
@@
 
-In this struggle between the potential protest coalition and the incumbent government, protesters and their possible allies behave as "passionate economists" [@VanZomeren2012]\textemdash they are motivated both by emotional responses to grievance, and by a more rational calculation of individual-level risk and group-level efficacy.
+In this struggle between the potential protest coalition and the incumbent government, protesters and their possible allies behave as "passionate economists" whose balance of emotional versus strategic considerations can be updated over time [@VanZomeren2012].
 What is the likelihood that protest action can help resolve the grievance, and does the benefit of participating outweigh the risk of personal consequences to the individual for participating?
 A fraudulent election may heighten citizens' grievances and generate uncertainty around the underlying popularity of the regime, creating a moment of elevated protest risk.
-Crucially, though, this evaluation is dynamic and can be updated over time in response to new information and new events [@VanZomeren2012].
+Crucially, though, this evaluation is dynamic and can be updated in response to new information and new events [@VanZomeren2012].
 If protest movements have sufficient numbers and organizational resources to put the regime on the back foot, this can lead to more numerous protests as individuals' sense of group efficacy grows and their sense of risk declines [@Chenoweth_Stephan2011].
-On the other hand, if the regime is well-resourced, capable, and unified, its actions to end a protest movement (including repression, concessions, and co-optation) can lead to a negative re-evaluation of the movement's odds of success.
+On the other hand, if the regime is well-resourced, capable, and unified, its actions to end a protest movement can lead to a negative re-evaluation of the movement's odds of success.
 This does not, of course, eliminate citizens' sense of grievance.
@@
 Opposition groups are also incentivized to stage protests regardless of the level of fraud, given the highly salient moment afforded by the election itself [@cunhaSignalReceivedAuthoritarian2022].
-More importantly, even if a protest movement begins after the election, any useful information provided by the level of election manipulation quickly decays and is replaced by signals from the protests themselves [@lohmannDynamicsInformationalCascades1994].
+Fraud itself provides little information about regime strength, and even if a protest movement begins after the election, any useful information provided by the level of fraud is quickly replaced by signals from the protests themselves [@lohmannDynamicsInformationalCascades1994].
 Protesters, bystanders, and elites can update their estimation of the relative strength of the protest coalition and the incumbent government by observing the size of crowds, the number of protests, and the government's reaction.
 
-Since the level of manipulation will no longer serve as an important indicator of regime strength once a challenge is taken to the streets, the number and size of protests will instead be related to other indicators of opposition and incumbent capacity.
+The number and size of protests will instead be related to other indicators of opposition and incumbent capacity.
 When opposition groups and civil society capacity are more organized and better resourced, it is likely that more protesters will take to the streets in more numerous events (see Chenoweth et al [-@chenowethStateRepressionNonviolent2017] for a review of factors making protest movements more resilient to repression).
@@
 I assume that more numerous protests, and those that built to a higher peak participation, are costlier for governments to contain, all else held equal.
-If fraudulent elections reveal information about regime weakness or generate sustained grievance-motivated protest, as the deterrence logic predicts, more fraudulent elections should be associated with larger and more numerous protest waves.
+If fraudulent elections reveal information about regime weakness or generate large grievance-motivated protest, more fraudulent elections should be associated with larger and more numerous protest waves.
 Prior theory expects these effects to be largest when margins of victory are smaller, or when economic grievances are high.
 
-However, the theory articulated above does **not** expect that the severity of fraud will be associated with the course of protest.
+However, the theory articulated above does *not* expect that the severity of fraud will be associated with the course of protest.
 Instead, it predicts that underlying indicators of opposition and ruling-party strength will influence the direction of protest movements.
-In particular, it expects the size, scope, and duration of protest waves will be sensitive to the repressive capacity of the state and to indicators of opposition strength.
+In particular, it expects the size and number of protest events will be sensitive to the repressive capacity of the state and to indicators of opposition strength.
 that high and low levels of repression will be associated with smaller protest movements (in number and size).
@@
 
-nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2023-05-22.csv"))
+nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2023-05-30.csv"))
 
@@
 The resulting dataset includes 647 total election-period observations across 109 countries.
+
+## Dependent variables
+
+To capture the *severity of election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than aggregate measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], it can fail to signal strength [@Harvey2020], and it can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021].
+This measure of election integrity should thus serve as a strong test of the protest-oriented model.
+The variable is used as a dependent variable in the first set of models, and as a potential explanatory variable in the subsequent models of protest number, size, and duration.
 
@@
 
-Each model includes the explanatory variables proposed above.
-First, as a proxy measures of economic grievance, I include the *lagged GDP growth rate*, taken from V-Dem.
-Lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance.
-To capture the *severity of election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than aggregate measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], it can fail to signal strength [@Harvey2020], and it can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021].
-This measure of election integrity should thus serve as a strong test of the protest-oriented model.
-Finally, the winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia and Psephos.}
+## Explanatory variables
+
+Given that pre-election protest is likely to be a strong predictor of post-election protest\textemdash by indicating strong social networks [@mateo2022], pre-existing opposition groups and civil society capacity [@Bunce2006], and/or popular grievances [@Brancati2016]\textemdash the deterrent logic suggests that non-democratic governments will pull back on election-day fraud when the pre-election environment is characterized by protest [@Rozenas2016b].
+While it is possible that pre-election protest may provide an opportunity for incumbents to crack down, creating more space for manipulating the election, more active pre-election protest suggests that opposition capacity is more robust, all else held equal.
+Moreover, in general, incumbents tend to avoid mass repression in the pre-election phase in favor of targeted harassment of opposition activists [@Bhasin2013].
+
+The first set of models test Hypothesis 1, on the relationship between protest risk and fraud.
+Protest risk is operationalized in three ways.
+First, the *number of pre-election protests* in ECAV is taken as *prima facie* evidence of a more turbulent and risky election environment\textemdash it indicates greater mobilization capacity in civil society and/or the opposition, and may also indicate an elevated sense of grievance.
+Second, the repressive capacity of the regime is modeled using the lagged *physical integrity index* for the country-year.
+In one approach, higher values represent more permissive environments, and should be associated with greater protest risk (and thus reduced fraud).
+Alternatively, in keeping with the 'murder in the middle' framework, the highest protest risks (and most reduced fraud) might be found at middle values of repression.
+Finally, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby.
+This diffusion mechanism is modeled using each country's geopolitical region as coded in Teorell et al [-@qog_teorell] in conjunction with binary variables indicating a successful protest movement from NELDA.
+A 'successful protest' is recorded if a protest movement resulted in either a new election being run or the incumbent stepping down.
+The *regional diffusion* dummy variable takes on a value of one for all countries in the same region as a successful protest, for elections within one calendar of the successfully overturned election.^[For example, the variable takes on a positive value for elections held in the post-communist region, within one year of the September 24, 2000 election in Serbia that sparked the 'Bulldozer Revolution'.
+All told there are xx successful electoral protests in the dataset: India 1989, Bulgaria 1990, Bangladesh 1996, Mali 1997, Lesotho 1998, Cote d'Ivoire 2000 and 2010, Peru 2000, Serbia 2000, Georgia 2003, Ukraine 2003-2004, Azerbaijan 2005, Kyrgyzstan 2005, Haiti 2006, and Cameroon 2007]
+
+The models of protest quantity, size, and duration incorporate several potential explanatory variables, including the *election fraud* variable described above.
+The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance.
+The winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia and Psephos.} The deterrence model suggests that closer elections with more severe fraud should be more likely to generate protest.
+Finally, the theory proposed here suggests that structural conditions\textemdash in particular, the state's repressive capacity\textemdash should be more influential for protest dynamics.
+
+I measure repressive capacity using the *physical integrity* index from V-Dem, lagged one year.
+Higher values for this variable indicate greater freedom from torture and political killing.
+I also include the variable's square term to account for the 'murder in the middle' framework.
+The physical integrity variables are expected to be explanatory in models of protest, but are important control variables in the models of election fraud.
+Since repressive capacity helps influence protest in expectation [@Ritter2016], incumbents that are more confident of deterring protest in advance should be more willing to engage in fraud under the protest-deterrent model of manipulation.
+
+Here I expect to see support for the so-called 'more murder in the middle' framework [@feinLifeIntegrityViolationsDemocracy1995], in which the regimes with the most demonstrated repressive capacity deter potential challengers from taking to the streets in the first place, and the least repressive states defuse protest through accommodation [@pierskallaProtestDeterrenceEscalation2010;@mooreRepressionDissentSubstitution2000].
+
+## Control variables
 
@@
 Finally, all models include measures of *judicial constraints* and  *legislative constraints* on the executive, V-Dem variables which captures the extent to which the judicial and legislative bodies in the country act independently in practice to limit executive power.
-
-Likewise, the ability of the state to engage in repression is likely to embolden the ruling party and deter protests; I thus control for *physical integrity* using the 'physical violence index' from V-Dem lagged one year.
-Higher values for this variable indicate greater freedom from torture and political killing.
-I also include the variable's square term to account for the 'murder in the middle' framework.
-The physical integrity variables are expected to be explanatory in models protest, but are important control variables in the models of election fraud.
-Since repressive capacity helps influence protest in expectation [@Ritter2016], incumbents that are more confident of deterring protest in advance should be more willing to engage in fraud under the protest-deterrent model of manipulation.
-Controlling for repression removes this confounding bias and allows for an estimate of the effect of pre-election disquiet on fraud.
-As an especially useful test, the deterrent model might expect fraud to be most reduced in the context of high pre-election protest and low repressive capacity.
-
-NELDA also includes a binary variable indicating elections in which *incumbents lost*; I expect that protest will be less likely when this variable takes a value of one.
 
@@
 Figure 2 shows summary statistics for all of these variables.
+Lastly, a binary variable from NELDA indicates elections in which *incumbents lost*; I expect that protest will be less likely when this variable takes a value of one.
 To address concerns about multicollinearity, I present models in the appendix (page 25) that exclude control variables associated with pre-election protest.
@@
 ## Pre-election protest and election-day fraud
-
-Hypothesis 1 suggests that a pre-electoral environment characterized by greater protest risk will be associated with lower election-day fraud.
-While the true value of pre-election protest risk is unobservable, I proxy it using the *number of pre-election protest events*, the state's *repressive capacity* and the interaction of the two.
-More protest events in the pre-election period is prima facie evidence of mobilization capacity in civil society and/or the opposition, and may also indicate an elevated sense of grievance.
 
@@
 ```{r manipdv, echo=FALSE}
-m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate + n.events.pre  +  unscheduled_election + regime_age + (1 | stateid), data = nelda2.sub)
+m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate + n.events.pre  +  unscheduled_election  + (1 | stateid), data = nelda2.sub)
 
@@
 
-m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   + presidential   +  libdem.1lag + urban.pct + rentierstate + taxstate + physinteg.1lag +  unscheduled_election + regime_age+ (1 | stateid), data = nelda2.sub)
+m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + (1 | stateid), data = nelda2.sub)
+#summary(m1a)
+
+m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +  libdem.1lag + urban.pct + rentierstate + taxstate + physinteg.1lag +  unscheduled_election + (1 | stateid), data = nelda2.sub)
 
@@
 
-m1c <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + unscheduled_election + regime_age + (1 | stateid), data = nelda2.sub)
-summary(m1c)
-
-m1d <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate + n.events.pre*physinteg.1lag +  unscheduled_election + regime_age + (1 | stateid), data = nelda2.sub)
-summary(m1d)
-
-modelsummary(c(m1, m1b, m1c, m1d), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Random-intercept multilevel model of election-day fraud', coef_rename = 
+m1c <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + unscheduled_election  + (1 | stateid), data = nelda2.sub)
+#summary(m1c)
+
+modelsummary(c(m1, m1b, m1c, m1a), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Random-intercept multilevel model of election-day fraud', coef_rename = 
                c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
@@
 ```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
-model.n.novoteshare.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + inc.lose,
+model.n.novoteshare.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                                 ziformula=~.,
-                                data = nelda2.sub,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                 family=truncated_nbinom2)
@@
 
-model.n.base.hurd.mm <- glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + inc.lose,
+model.n.base.hurd.mm <- glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                ziformula=~.,
-                              data = nelda2.sub,
+                              data = nelda2.sub %>% filter(inc.lose == "0"),
                                 family=truncated_nbinom2)
@@
 
-model.n.base2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + inc.lose,
+model.n.base2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                           zi = ~.,
                                           family= nbinom2,
-                                          data = nelda2.sub)
+                                          data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.n.base2.hurd.mm)
 
-model.n.griev.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age+ inc.lose,
+model.n.griev.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,
-                                data = nelda2.sub,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                 family=truncated_nbinom2)
@@
 
-model.n.griev2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + inc.lose ,
+model.n.griev2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,
-                                data = nelda2.sub,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                 family=truncated_nbinom2)
@@
 
-model.n.griev3.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + inc.lose  ,
+model.n.griev3.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag  ,
                                 ziformula=~.,
-                                data = nelda2.sub,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                 family=truncated_nbinom2)
@@
 
-model.n.repr.hurd <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv  +  physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + inc.lose,
+model.n.repr.hurd <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv  +  physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                 ziformula=~.,
-                                data = nelda2.sub,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                 family=truncated_nbinom2)
@@
 
-```{r models-hurdle-duration, warning = FALSE, message = FALSE, include = F}
-model.d.novoteshare.hurd.mm <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + inc.lose,
-                                ziformula=~.,
-                                data = nelda2.sub,
-                                family=truncated_nbinom2)
-summary(model.d.novoteshare.hurd.mm)
-
-model.d.base2.hurd <- glmmTMB::glmmTMB(d.events.num ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age +  inc.lose ,
-                                          zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub)
-summary(model.d.base2.hurd)
-
-model.d.griev.hurd <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + inc.lose,
-                                ziformula=~.,
-                                data = nelda2.sub,
-                                family=truncated_nbinom2)
-summary(model.d.griev.hurd)
-
-model.d.griev2.hurd <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + as.factor(inc.lose) ,
-                                ziformula=~.,
-                                data = nelda2.sub,
-                                family=truncated_nbinom2)
-summary(model.d.griev2.hurd)
-
-model.d.griev3.hurd <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + as.factor(inc.lose)  ,
-                                ziformula=~.,
-                                data = nelda2.sub,
-                                family=truncated_nbinom2)
-summary(model.d.griev3.hurd)
-
-```
-
 ```{r, include=FALSE}
 library(survival)
+library(ggsurvfit)
 nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
-km <-  with(nelda2.sub, Surv(d.events.num, surv_status))
-
-cox1 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age, data = nelda2.sub %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+nelda2.sub <- nelda2.sub %>% mutate(fraud_level = ifelse(v2elirreg.inv >= (.98 + .97), "High",
+                                                            ifelse(v2elirreg.inv <= (.98 - .97), "Low", "Moderate")))
+
+nelda2.sub.km <- nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0)
+
+km <-  with(nelda2.sub.km, Surv(d.events.num, surv_status))
+
+survfit2(Surv(d.events.num, surv_status) ~ 1, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
+  ggsurvfit() +
+#   add_confidence_interval() +
+  labs(
+    x = "Days",
+    y = "Overall survival probability"
+  )
+
+survfit2(Surv(d.events.num, surv_status) ~ fraud_level, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
+  ggsurvfit() +
+#   add_confidence_interval() +
+  labs(
+    x = "Days",
+    y = "Overall survival probability"
+  )
+
+cox1 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag , data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
 summary(cox1)
 
-cox2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age , data = nelda2.sub %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+cz <- cox.zph(cox1, transform = "identity")
+print(cz)
+plot(cz) # The points on the plot are residuals of v2elirreg.inv (and other vars) based on hte predicted value of X and time t.
+The line shows the estimated value of the coefficient on variable X at time t compared to its baseline.
+You want it to be zero for consistent effects over time.
+V2elirreg.inv is positive for about two weeks early on.
+The substantive effect of this can be shown by taking two values of v2elirreg.inv (i.e.
+first and third quartile), taking the difference, and finding exp(coefficient_residual * difference).
+The result shows here that protests in states in the third quartile of fraud are exp(.75*1.4) = 2.86 times more likely to die than those in the first quartile during those couple weeks.
+So this suggests that if protests survive the first week or two, the state's capacity starts to kick in?
+
+sandbox_data <- survSplit(Surv(d.events.num, surv_status) ~ 
+                      jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + fraud_level  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+                      data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), cut = c(2), episode = "tgroup")
+
+cox1.2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + fraud_level*(as.factor(tgroup))  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = sandbox_data)
+summary(cox1.2)
+
+cox2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
 summary(cox2)
 
-cox3 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age, data = nelda2.sub %>% filter(d.events.num > 0))
+cox3 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag , data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
 summary(cox3)
 
-cox4 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age, data = nelda2.sub %>% filter(d.events.num > 0))
+cox4 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
 summary(cox4)
@@
 ```{r models-hurdle-size, warning = FALSE, message = FALSE, include = F}
-model.size.novoteshare <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose) ,
+model.size.novoteshare <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,
-                                data = nelda2.sub,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                 family=truncated_nbinom2)
@@
 
-model.size.base <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   v2elirreg.inv + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose) ,
+model.size.base <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   v2elirreg.inv + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,
-                                data = nelda2.sub,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                 family=truncated_nbinom2)
@@
 
-model.size.base2 <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose) ,
+model.size.base2 <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,
-                                data = nelda2.sub,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                 family=truncated_nbinom2)
@@
 
-model.size.griev <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose) ,
+model.size.griev <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                                 ziformula=~.,
-                                data = nelda2.sub,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                 family=truncated_nbinom2)
@@
 The second set of models challenge the idea that more severe electoral manipulation is associated with an increased risk of anti-government protest, except under extraordinary circumstances.
-Beginning with the binary portion of the hurdle models, which model the likelihood of any protest occurring after the election, the standalone coefficient for election fraud is not significant in Models 3.
-Model 4 interacts election fraud with the incumbent's official vote-share; here we do observe statistically significant relationships at the p < .05 level.
-However, as the marginal effects plot in Figure \@ref(fig:vote-share-prob-fig) shows, this relationship is only statistically significant when comparing extreme scenarios.
+Beginning with the zero-inflation portion of the hurdle models, which model the likelihood of *no* protest occurring after the election, the standalone coefficient for election fraud is not significant in Model XX, nor are the interaction coefficients in Models XX and XX.
+Marginal effects plots [include XX] show that there are no significant effects across values of *winner's vote share* or *lagged GDP growth*.
+The most consistently significant relationship is the negative coefficient for *presidential elections*; this indicates that protest initiation is more likely after a presidential election.
 
 While more fraudulent elections are associated with an increasing risk of protest as margins of victory shrink, this increase is not statistically  significant for typical values of manipulation.
-In the bottom panel of Figure  \@ref(fig:vote-share-prob-fig), typical values for low and high levels of fraud are used (one standard deviation below and above the mean, respectively).
+In the bottom panel of Figure  \@ref(fig:margin-prob-fig), typical values for low and high levels of fraud are used (one standard deviation below and above the mean, respectively).
 In this lower panel, the increased risk of protest in the high-fraud scenario is not statistically distinguishable from the low-fraud scenario.
@@
 
-A similar, but somewhat stronger effect, can also be seen in the relationship between fraud, public sector corruption, and protest.
-As shown in Figure \@ref(fig:corrupt-prob-fig), higher levels of fraud are associated with higher risk of protest when corruption is high.
-This effect is not statistically significant for typical values of fraud (one SD above and below the mean), but quickly becomes so when comparing more disparate levels of fraud.
-At the maximum values for fraud, as the top panel shows, the difference in risk is significant even for modestly elevated levels of corruption.
-This is at least partially in line with the predictions of the literature following Tucker [-@tucker2007], though the relationship is not significant when comparing more typical levels of fraud.
-Finally, there is no significant effect for *economic crisis*, at least under the typical covariate values used to create the marginal effects plot in Figure \@ref(fig:crisis-prob-fig).
-
 \newpage
@@
 
-To-do: Fix binary models so that they report same as hurdles
-
 ```{r include=FALSE}
-nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(n.events.post > 0, 1, 0))
-nelda2.sub <- nelda2.sub %>% mutate(any_size = ifelse(max_size > 0, 1, 0))
-
-model.margin.binary_n <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age, family = binomial(link = "logit"), data = nelda2.sub)
+nelda2.sub <- nelda2.sub %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))
+nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))
+
+model.margin.binary_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary_n)
 
-model.margin.binary_size <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose), family = binomial(link = "logit"), data = nelda2.sub)
+model.margin.binary_size <- glm(no_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary_size)
 
-model.binary.griev1_n <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose), family = binomial(link = "logit"), data = nelda2.sub)
+model.binary.griev1_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.binary.griev1_n)
 
-model.binary.griev1_size <- glm(any_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose), family = binomial(link = "logit"), data = nelda2.sub)
+model.binary.griev1_size <- glm(no_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.binary.griev1_size)
 
-model.binary.repr_n <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regime_age + as.factor(inc.lose), family = binomial(link = "logit"), data = nelda2.sub)
+model.binary.repr_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.binary.repr_n)
@@
 ```
-
-```{r n-protests-gdp, fig.cap="Fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE}
+```{r}
+plot_binary_gdp_typ_n
+```
+
+```{r n-protests-margin, fig.cap="Fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE}
 dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "count") 
@@
 
-```{r gdp-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
+```{r margin-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
 plot.combined.prob <- cowplot::plot_grid(plot_binary_margin_typ_n, p.n.margin, ncol=1)
 plot.combined.prob
-```
-
-```{r n-protests-margin, fig.cap="Fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE}
-dat <- ggpredict(model.d.base2.hurd, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "count") 
-p.d.margin <- plot(dat)
-p.d.margin <- p.d.margin +  labs(x = "Incumbent vote-share", y = "Duration of wave (days)", title = "", color = "Election fraud") + theme_bw() + coord_cartesian(ylim = c(0, 400))
 ```
@@
 
-```{r n-protests-crisis, fig.cap="Fraud, economic crisis, and number of protest events", echo=FALSE, message=FALSE}
-newdata <- data.frame(
-   jucon.1lag =  0.50,
-          legcon.1lag =  0.50,
-           osorg.1lag =  0.90,
-          altinf.1lag =  0.65,
-     natlpartyorg.lag =  0.76,
-     pubseccorrup.lag = -0.86,
-         presidential =     0,
-    inc.termlimit.num =  0,
-       physinteg.1lag =  0.59,
-        physintlag.sq =  0.41,
-            urban.pct = 50.52,
-         rentierstate =     0,
-             taxstate =     1,
-         n.events.pre =  0.00,
- unscheduled_election =     0,
-             inc.lose =     0,
- v2elirreg.inv = c(-.45, -.45, 1.64, 1.64),
- econ.crisis.nelda.1lag = c(0, 1, 0, 1)
-)
-
-phat <- predict(model.n.griev2.hurd.mm, newdata, type = "response", se.fit = T) 
-newdata$phat <- phat$fit
-newdata$phatse <- phat$se.fit
-newdata$conflow <- newdata$phat - 1.96*(newdata$phatse)
-newdata$confhi <- newdata$phat + 1.96*(newdata$phatse)
-
-newdata$conflow_exp <- exp(newdata$conflow)
-newdata$confhi_exp <- exp(newdata$confhi)
-
-p.n.crisis <- ggplot(data=newdata, aes(x=jitter(econ.crisis.nelda.1lag), y = phat, color = v2elirreg.inv)) +
-  geom_point() +
-  geom_linerange(aes(ymin = conflow, ymax = confhi))
-p.n.crisis  + labs(x = "Economic crisis", y = "Number of events", title = "Predicted counts of protest events", color = "Election fraud")  + theme_bw()
-
-```
-
-```{r n-protests-corrupt, fig.cap="Fraud, public sector corruption, and number of protest events", echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.griev3.hurd.mm, terms = c("pubseccorrup.lag", "v2elirreg.inv [-.45, 1.64]"), type = "fe") 
-p.corr.n.margin <- plot(dat)
-p.corr.n.margin + coord_cartesian(ylim = c(0, 50)) + labs(x = "Public sector corruption", y = "Number of events", title = "Predicted counts of protest events", color = "Election fraud") + theme_bw()
-
+```{r pred-size, fig.cap="Fraud, winning vote-share, and predicted max protest size", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
+
+p.gdplow <- sjPlot::plot_model(model.size.base2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(gdpgro.1lag = -.5))
+
+p.gdphigh <- sjPlot::plot_model(model.size.base2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(gdpgro.1lag = 3))
+
+cowplot::plot_grid(p.gdplow, p.gdphigh)
 ```
@@
 
+One objection to this framing may be that, if the level of manipulation and the risk of protest are endogenous to one another, perhaps the null result found for the number of protests is to be expected.
+That is, perhaps we are observing the equilibrium behavior predicted in many formal models: the threat of protest keeps fraud lower than it would be otherwise, resulting in limited (if any) protest.
+In some states, this equilibrium level of fraud is higher than in others, resulting in no correlation between fraud and observed protest.
+There are two reasons to be skeptical of this interpretation, one empirical and one substantive.
+
+Empirically, deterrence models would expect the equilibrium level of fraud to fall as protest risk increases, and for more severe protest to occur as a result of miscalculation by the regime.
+We can test this by examining the predicted intensity of protest from the models specified here.
+If the severity of fraud is held constant at a high level while underlying protest risk increases, we should see an increase in the predicted protest outcome.
+Taking protest size as an example, using Model XX, Figure \@ref(fig:pred-size) shows there is no evidence of such an effect.
+Both panels show the marginal effect of a relatively fraudulent election (*intentional irregularities* is set to one standard deviation above the mean) on the predicted maximum protest size, conditional on the closeness of the election result.
+All control variables have been set to their means or modes with the exception of lagged GDP growth rate.
+In the left panel, GDP growth has been set to -0.5, close to its minimum value; on the right, it has been set to 3.
+An equilibrium approach would suggest that high fraud is more dangerous in the troubled economic environment captured on the left, especially at low levels of *winner's vote-share*.
+However, we see practically no difference in the predicted maximum protest size across the different economic conditions modeled here\textemdash changing economic conditions do not appear to affect the 'appropriate' level of fraud, throwing cold water on the equilibrium argument.
+
+Substantively, even if one does not fully discount the claim that the null results shown here represent equilibrium behavior, it is difficult to dispute that the resulting outcome is weighted heavily in favor of governments.
+In the dataset of 1110 elections used here, only 19 elections in five countries are coded as having an *intentional irregularities* score of less than -2 (indicating extremely fraud-free elections).
+By contrast, 121 elections in 27 countries score higher than 2 (indicating highly fraudulent elections).
+The frequency of fraud, the rarity of protest, and the finding that fraud severity is not correlated with the scope of protest events all point to the same conclusion: protest risk is not a substantial constraint on election fraud in electoral authoritarian and electoral democratic regimes.
+
 # A further test: protest wave duration
@@
 models3[['Model 7']] <- cox3
-models3[['Model 8']] <- cox4
 
@@
 
-One objection to this framing may be that, if the level of manipulation and the risk of protest are endogenous to one another, perhaps the null result found for the number of protests is to be expected.
-That is, if incumbents choose a level of manipulation calculated to avoid large-scale protest---sometimes high and sometimes low, conditional on contextual factors---no correlation between fraud and protest quantity would be found.
-This paper does not dispute that decisions about manipulation and protest may be endogenous to one another, but it does argue that the resulting equilibrium is weighted heavily in favor of governments.
-In the dataset used here, only ten elections in four countries are coded as having an *intentional irregularities* score of less than -2 (indicating extremely fraud-free elections).
-By contrast, 103 elections in 27 countries score higher than 2 (indicating highly fraudulent elections).
-The frequency of fraud, the rarity of protest, and the finding that fraud severity is not correlated with the scope of protest events all point to the same conclusion: protest risk is not a substantial constraint on election fraud in non-democracies.
-
 These results are consistent with different specifications presented in an online appendix.
 In particular, the main results are supported when data from the Electoral Contention and Violence (ECAV) dataset [@Daxecker2019] are used in place of the Mass Mobilization data (appendix page 1), and when a measure of overall election integrity is used in place of the measure of election fraud (appendix page 15).
+
+Scholarly attention might turn more profitably toward supply-side explanations for manipulation, such as the organizational capacity of the state [@Simpser2011], the popularity of the government [@rundlett2016deliver], the risks faced by front-line agents [@Harvey2022].
 

### 2023-06-05 Output/Submission files/manuscript_anon_v2.Rmd
Updates to text + section on Cox models

@@
 
-As a final test of the mechanism outlined in the theory presented here\textemdash that election manipulation offers few clues to regime strength, and thus the course of electoral protests are determined by more structural factors\textemdash the same models will also be applied to the duration of post-election protest waves.
-Longer protest waves indicate a more even match between the resources and resiliency of the protesters and the capacity and cohesiveness of the regime.
-It is expected that more severe election fraud will *not* be associated with longer protest waves, as would be predicted if fraud were an indicator of state weakness.
-Instead, structural factors should again be more influential;   harsh repression should be associated with shorter protest waves, while indicators of opposition and civil-society strength should help keep protest waves alive.^[Note here about excluding successful movements.]
+To further test the mechanism outlined in the theory presented here\textemdash that election manipulation offers few clues to regime strength, and thus the course of electoral protests are determined by more structural factors\textemdash I examine the relationship between fraud and the survival of protest movements in cases where protests began but the incumbent eventually survived.
+Restricting the sample to these cases allows for a consistent interpretation of results: longer protest waves indicate a more even match between the resources and resiliency of the protesters and the capacity and cohesiveness of the regime, while shorter-lived protest movements are those which were more easily defeated.
+In this sub-sample, I would first expect the relationship between election fraud and protest survival to be non-linear with respect to time\textemdash the longer the protest movement survives, the less relevant the fraud signal (such as it is) becomes).
+Second, it is expected that more severe election fraud will *not* be associated with longer protest waves, as would be predicted if fraud were an indicator of state weakness.
+Instead, structural factors should again be more influential; indicators of opposition and civil-society strength should help keep protest waves alive.
 
@@
 nelda2.sub <- nelda2.sub %>% mutate(presidential = factor(presidential))
-nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = factor(inc.termlimit.num, exclude = NULL))
+#nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = factor(inc.termlimit.num))
 nelda2.sub <- nelda2.sub %>% mutate(rentierstate = factor(rentierstate))
@@
 I limit the sample by excluding closed autocracies (where no national elections are held).
-Since the theory of backlash to electoral manipulation could in theory apply to liberal democracies, I include them in the main analysis (as coded by V-Dem).^[The specific variable used is 'v2x_regime'.] As a robustness check for this sampling decision, I also include the main models without liberal democracies in the appendix, with no substantial changes in the outcomes.
-Finally, the beginning of the date range is limited by the availability of data on protest from ECAV, as discussed below, which ranges from 1990 to 2012.
-The resulting dataset includes 647 total election-period observations across 109 countries.
-
-## Dependent variables
-
-To capture the *severity of election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than aggregate measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], it can fail to signal strength [@Harvey2020], and it can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021].
+Since the theory of popular backlash to manipulation has largely been articulated in an electoral-democratic context, I exclude liberal democracies from the main analysis.^[The specific variable used is 'v2x_regime'; I limit the sample to those regime-years coded 1 or 2.] As a robustness check for this sampling decision, I also run the main models with data including liberal democracies in the appendix, with no substantial changes in the outcomes.
+The first set of models, focusing on Hypothesis 1, include all the remaining cases in the dataset.
+For the subsequent models, focusing on post-election protest, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the theoretical expectations of protest risk following an election in which the incumbent won, possibly through reliance on election fraud.
+Finally, the beginning of the date range is limited by the availability of data on protest from ECAV, as discussed below, which ranges from 1989 to 2012.
+The resulting dataset includes 1110 total election-period observations across 118 countries.
+
+## Dependent variables and modeling strategies
+
+To capture the severity of *election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than aggregate measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], it can fail to signal strength [@Harvey2020], and it can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021].
 This measure of election integrity should thus serve as a strong test of the protest-oriented model.
-The variable is used as a dependent variable in the first set of models, and as a potential explanatory variable in the subsequent models of protest number, size, and duration.
+The variable is used as a dependent variable in the first set of models that test H1, and as a potential explanatory variable in the subsequent models of protest number, size, and duration.
 
@@
 ECAV data is the event-day format; to capture the *number of protests* before and after the election, I take the sum number of events recorded in ECAV before and after election day, with election-day protests being counted as post-election.
-To evaluate hypotheses regarding protest size, I convert ECAV's categorical data on number of participants per event-day to estimate numbers, and take the sum where multiple protests were held on a given day.
-The ECAV dataset codes protest size as a categorical variable with five levels, ranging from fewer than ten participants to more than 10,000.
+To evaluate hypotheses regarding protest size, I convert ECAV's categorical data on number of participants per event-day to estimate numbers, and take the sum where multiple protests were held on a given day.^[This approach is necessary to avoid conflating days with, for example, five category 1 protests (totaling perhaps 50 participants) and days with a single category 5 protest (totaling more than 10,000).] The ECAV dataset codes protest size as a categorical variable with five levels, ranging from fewer than ten participants to more than 10,000.
 To deal with this data structure, where the top category has no upper limit, I take the median value of each category, with the exception of the top category.
@@
 
-The large majority of elections experience no post-election protest: in 1168 of the 1477 elections, no protests are recorded.
+The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0) %>% nrow()` of the `r nrow(nelda2.sub)` elections, no protests are recorded.
 Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest number and maximum single-day protest size.
+As the figure shows, while election fraud is common among this set of regimes, protests are rare, and usually small in number and maximum size when they do occur.
+This distribution\textemdash with large numbers of zeroes\textemdash is in line with the theoretical proposition that two data-generating processes are at work: the factors that lead to the occurrence of protest may not be the same as those that sustain or intensify it.
+Consequently, I model the number of protests using negative binomial hurdle regressions.
+Hurdle models are two-part regressions in which the variation in the dependent variable is assumed to follow two distributions: the probability of obtaining a zero or non-zero outcome is modeled as binomial logit function, while the distribution of the non-zero count variable is modeled (in this case) using the negative binomial function.
+The logistic portion of the model identifies the factors that make *any protest* more likely to occur; the negative binomial regressions model the count of protests given that protest has begun.
+Since protest size cannot take on a negative value, it is also modeled as a count variable with a negative binomial distribution in the hurdle models.
+*Election fraud*, which follows a more normal distribution, is modeled using multilevel models with random intercepts by country.
 
@@
                                             'Alternative information' = altinf.1lag,
+                                          'Liberal democracy index' = libdem.1lag,
                                             'GDP growth rate' = gdpgro.1lag, 
@@
                                             'Public sector corruption' = pubseccorrup.lag,
-                                            'Economic crisis' = econ.crisis.nelda.1lag,
                                             'National party organization' = natlpartyorg.lag, 
@@
                                             'Tax state' =  taxstate,
-                                            'Incumbent vote-share' = winner.share.upd, 
-                                            'Median pre-election protest size' = median.participants.pre,
-                                            'Largest protest size' = max_size,
-                                            'Off-schedule election' = unscheduled_election)
+                                            'Winning vote-share' = winner.share.upd, 
+                                            'Number of pre-election protests' = n.events.pre,
+                                            'Off-schedule election' = unscheduled_election,
+                                          'Regional protest diffusion' = regional_success_lag)
 #incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r summary-table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Summary statistics"}
 library(kableExtra)
@@
 \doublespace
-
-This distribution\textemdash with large numbers of zeroes\textemdash is in line with the theoretical proposition that two data-generating processes are at work: the factors that lead to the occurrence of protest may not be the same as those that sustain or intensify it.
-Consequently, I model the results using negative binomial hurdle regressions.
-Hurdle models are two-part regressions in which the variation in the dependent variable is assumed to follow two distributions: the probability of obtaining a zero or non-zero outcome is modeled as binomial logic function, while the distribution of the non-zero count variable is modeled (in this case) using the negative binomial function.
-The logistic portion of the model identifies the factors that make *any protest* more likely to occur; the negative binomial regressions to model the count of protests given that protest has begun.
-Since protest size cannot take on a negative value, it is also modeled as a count variable with a negative binomial distribution in the hurdle models.
 
@@
 In one approach, higher values represent more permissive environments, and should be associated with greater protest risk (and thus reduced fraud).
-Alternatively, in keeping with the 'murder in the middle' framework, the highest protest risks (and most reduced fraud) might be found at middle values of repression.
-Finally, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby.
-This diffusion mechanism is modeled using each country's geopolitical region as coded in Teorell et al [-@qog_teorell] in conjunction with binary variables indicating a successful protest movement from NELDA.
-A 'successful protest' is recorded if a protest movement resulted in either a new election being run or the incumbent stepping down.
-The *regional diffusion* dummy variable takes on a value of one for all countries in the same region as a successful protest, for elections within one calendar of the successfully overturned election.^[For example, the variable takes on a positive value for elections held in the post-communist region, within one year of the September 24, 2000 election in Serbia that sparked the 'Bulldozer Revolution'.
-All told there are xx successful electoral protests in the dataset: India 1989, Bulgaria 1990, Bangladesh 1996, Mali 1997, Lesotho 1998, Cote d'Ivoire 2000 and 2010, Peru 2000, Serbia 2000, Georgia 2003, Ukraine 2003-2004, Azerbaijan 2005, Kyrgyzstan 2005, Haiti 2006, and Cameroon 2007]
-
-The models of protest quantity, size, and duration incorporate several potential explanatory variables, including the *election fraud* variable described above.
-The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance.
-The winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia and Psephos.} The deterrence model suggests that closer elections with more severe fraud should be more likely to generate protest.
-Finally, the theory proposed here suggests that structural conditions\textemdash in particular, the state's repressive capacity\textemdash should be more influential for protest dynamics.
-
+Alternatively, in keeping with the 'murder in the middle' framework [@feinLifeIntegrityViolationsDemocracy1995], the highest protest risks (and most reduced fraud) might be found at middle values of repression; regimes with the most demonstrated repressive capacity deter potential challengers from taking to the streets in the first place, and the least repressive states defuse protest through accommodation [@pierskallaProtestDeterrenceEscalation2010;@mooreRepressionDissentSubstitution2000].
 I measure repressive capacity using the *physical integrity* index from V-Dem, lagged one year.
@@
 
-Here I expect to see support for the so-called 'more murder in the middle' framework [@feinLifeIntegrityViolationsDemocracy1995], in which the regimes with the most demonstrated repressive capacity deter potential challengers from taking to the streets in the first place, and the least repressive states defuse protest through accommodation [@pierskallaProtestDeterrenceEscalation2010;@mooreRepressionDissentSubstitution2000].
+Finally, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby.
+This diffusion mechanism is modeled using each country's geopolitical region as coded in Teorell et al [-@qog_teorell] in conjunction with binary variables indicating a successful protest movement from NELDA.
+A 'successful protest' is recorded if a protest movement resulted in either a new election being run or the incumbent stepping down.
+The *regional diffusion* dummy variable takes on a value of one for all countries in the same region as a successful protest, for elections within one calendar year of the successfully overturned election.^[For example, the variable takes on a positive value for elections held in the post-communist region, within one year of the September 24, 2000 election in Serbia that sparked the 'Bulldozer Revolution'.
+All told there are 15 successful electoral protests in the dataset: India 1989, Bulgaria 1990, Bangladesh 1996, Mali 1997, Lesotho 1998, Cote d'Ivoire 2000 and 2010, Peru 2000, Serbia 2000, Georgia 2003, Ukraine 2003-2004, Azerbaijan 2005, Kyrgyzstan 2005, Haiti 2006, and Cameroon 2007.]
+
+The models of protest quantity, size, and duration incorporate several potential explanatory variables, including the *election fraud* variable described above.
+The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance.
+The winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia and Psephos.} The deterrence model suggests that closer elections with more severe fraud should be more likely to generate protest.
+Finally, the theory proposed here suggests that structural conditions\textemdash in particular, the state's repressive capacity\textemdash should be more influential for protest dynamics.
 
@@
 Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019].
-Finally, all models include measures of *judicial constraints* and  *legislative constraints* on the executive, V-Dem variables which captures the extent to which the judicial and legislative bodies in the country act independently in practice to limit executive power.
+All models include measures of *judicial constraints* and  *legislative constraints* on the executive, V-Dem variables which captures the extent to which the judicial and legislative bodies in the country act independently in practice to limit executive power.
+To capture the overall level of pre-election protection for rights associated with liberal democracy, I include the lagged *liberal democracy index* from V-Dem.
 
@@
 Lastly, since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.
-Figure 2 shows summary statistics for all of these variables.
-Lastly, a binary variable from NELDA indicates elections in which *incumbents lost*; I expect that protest will be less likely when this variable takes a value of one.
+Figure \@ref(fig:summary-table) shows summary statistics for all of these variables.
 To address concerns about multicollinearity, I present models in the appendix (page 25) that exclude control variables associated with pre-election protest.
@@
 The results of these models are not promising for argument that protest risk deters election manipulation.
-First, Hypothesis 1 is not supported in Models 1 and 2 in Table  \@ref(tab:manipdv) \textemdash the coefficient on *pre-election protest* is positive and significant.
-The model predicts that each additional pre-election protest event is associated with an **increase** in the fraud measure of 0.008.
-This is a very small effect, given that the dependent variable ranges from `r min(nelda2.sub$v2elirreg.inv, na.rm = T)` to `r max(nelda2.sub$v2elirreg.inv, na.rm = T)`.
-An increase from zero to five protest events results in a predicted increase in fraud of roughly three percent of a standard deviation.
-While this is a small effect, it runs counter to the expectations of the deterrence logic.
-And the effect may be sizable in extreme circumstances.
-For example, pre-election protest in the 99th percentile is associated with an increase of `r (as.numeric(quantile(nelda2.sub$n.events.pre, probs = .99)))*.008`, about one-fifth of a standard deviation.
+First, Hypothesis 1 is not supported in the models shown in Table  \@ref(tab:manipdv).
+Models 1 through 5 apply different measures of pre-election protest risk; all of them behave contrary to the standard expectation in the literature.
+In Model 1, the number of pre-election protests is associated with an increase in election-day fraud; this effect is unchanged in Model 5, where *liberal democracy index* replaces other indicators of political openness.
+the coefficient on *pre-election protest* is positive and significant.
+While the effect is small, it runs counter to the expectations of the deterrence logic.
+A similar effect is found for *diffusion*; recent successful protest in a nearby country is also associated with a small increase in expected fraud.
 This suggests that the level of election manipulation, rather than being deterred by a highly contentious environment, may be deployed to make up for anticipated poor election results.
 
+Model 2 indicates that fraud declines as societies become less repressive; an intuitive finding.
+However, Model 3 shows that fraud first increases with physical integrity, but then declines.
+In other words, fraud is lowest on the extremes of repressive capacity\textemdash where repression is severe and where it is most limited.
+This is contrary to the expectation that a higher risk of protest is associated with reduced fraud.
+
 \newpage
@@
 ```{r manipdv, echo=FALSE}
-m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate + n.events.pre  +  unscheduled_election  + (1 | stateid), data = nelda2.sub)
+m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  +  unscheduled_election  + (1 | stateid), data = nelda2.sub)
 
@@
 
-m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + (1 | stateid), data = nelda2.sub)
+m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + (1 | stateid), data = nelda2.sub)
 #summary(m1a)
 
-m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +  libdem.1lag + urban.pct + rentierstate + taxstate + physinteg.1lag +  unscheduled_election + (1 | stateid), data = nelda2.sub)
+m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag +  unscheduled_election + (1 | stateid), data = nelda2.sub)
 
@@
 
-m1c <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + unscheduled_election  + (1 | stateid), data = nelda2.sub)
+m1c <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + unscheduled_election  + (1 | stateid), data = nelda2.sub)
 #summary(m1c)
 
-modelsummary(c(m1, m1b, m1c, m1a), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Random-intercept multilevel model of election-day fraud', coef_rename = 
+m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  +  unscheduled_election  + (1 | stateid), data = nelda2.sub)
+
+#summary(m1)
+
+m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + (1 | stateid), data = nelda2.sub)
+#summary(m1a)
+
+m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag +  unscheduled_election + (1 | stateid), data = nelda2.sub)
+
+#summary(m1b)
+
+m1d <- lmer(v2elirreg.inv ~  libdem.1lag + gdpgro.1lag +    presidential +  urban.pct + rentierstate + taxstate  +  unscheduled_election + n.events.pre  + (1 | stateid), data = nelda2.sub) 
+#summary(m1d)
+
+modelsummary(c(m1, m1b, m1c, m1a, m1d), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Random-intercept multilevel model of election-day fraud', coef_rename = 
                c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
@@
 ```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
-model.n.novoteshare.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
+model.n.novoteshare.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                                 ziformula=~.,
@@
 
-model.n.base.hurd.mm <- glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+model.n.base.hurd.mm <- glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                ziformula=~.,
@@
 
-model.n.base2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+model.n.base2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                           zi = ~.,
@@
 
-model.n.griev.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+model.n.griev.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,
@@
 
-model.n.griev2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+model.n.griev2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,
@@
 
-model.n.griev3.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag  ,
+model.n.griev3.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag  ,
                                 ziformula=~.,
@@
 
-model.n.repr.hurd <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv  +  physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+model.n.repr.hurd <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv  +  physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                 ziformula=~.,
@@
 
-```{r, include=FALSE}
-library(survival)
-library(ggsurvfit)
-nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
-nelda2.sub <- nelda2.sub %>% mutate(fraud_level = ifelse(v2elirreg.inv >= (.98 + .97), "High",
-                                                            ifelse(v2elirreg.inv <= (.98 - .97), "Low", "Moderate")))
-
-nelda2.sub.km <- nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0)
-
-km <-  with(nelda2.sub.km, Surv(d.events.num, surv_status))
-
-survfit2(Surv(d.events.num, surv_status) ~ 1, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
-  ggsurvfit() +
-#   add_confidence_interval() +
-  labs(
-    x = "Days",
-    y = "Overall survival probability"
-  )
-
-survfit2(Surv(d.events.num, surv_status) ~ fraud_level, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
-  ggsurvfit() +
-#   add_confidence_interval() +
-  labs(
-    x = "Days",
-    y = "Overall survival probability"
-  )
-
-cox1 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag , data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
-summary(cox1)
-
-cz <- cox.zph(cox1, transform = "identity")
-print(cz)
-plot(cz) # The points on the plot are residuals of v2elirreg.inv (and other vars) based on hte predicted value of X and time t.
-The line shows the estimated value of the coefficient on variable X at time t compared to its baseline.
-You want it to be zero for consistent effects over time.
-V2elirreg.inv is positive for about two weeks early on.
-The substantive effect of this can be shown by taking two values of v2elirreg.inv (i.e.
-first and third quartile), taking the difference, and finding exp(coefficient_residual * difference).
-The result shows here that protests in states in the third quartile of fraud are exp(.75*1.4) = 2.86 times more likely to die than those in the first quartile during those couple weeks.
-So this suggests that if protests survive the first week or two, the state's capacity starts to kick in?
-
-sandbox_data <- survSplit(Surv(d.events.num, surv_status) ~ 
-                      jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + fraud_level  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
-                      data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), cut = c(2), episode = "tgroup")
-
-cox1.2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + fraud_level*(as.factor(tgroup))  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = sandbox_data)
-summary(cox1.2)
-
-cox2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
-summary(cox2)
-
-cox3 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag , data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
-summary(cox3)
-
-cox4 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
-summary(cox4)
-
-```
-
 ```{r models-hurdle-size, warning = FALSE, message = FALSE, include = F}
-model.size.novoteshare <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+model.size.novoteshare <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,
@@
 
-model.size.base <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   v2elirreg.inv + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+model.size.base <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   v2elirreg.inv + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,
@@
 
-model.size.base2 <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+model.size.base2 <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,
@@
 
-model.size.griev <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
+model.size.griev <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                                 ziformula=~.,
@@
 models_t1 <- list()
-models_t1[['Model 3']] <- model.n.base.hurd.mm
-models_t1[['Model 4']] <- model.n.base2.hurd.mm
-models_t1[['Model 5']] <- model.n.griev.hurd.mm
+models_t1[['Model 6']] <- model.n.base.hurd.mm
+models_t1[['Model 7']] <- model.n.base2.hurd.mm
+models_t1[['Model 8']] <- model.n.griev.hurd.mm
 #models_t1[['Model 5']] <- m_duration_count
@@
 
-t1b <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, omit_coefs = c("(Intercept)", "jucon.1lag" , "legcon.1lag" , "gdpgro.1lag" , "altinf.1lag" , "pubseccorrup.lag" , "osorg.1lag" , "econ.crisis.nelda.1lag1" , "natlpartyorg.lag" , "v2elirreg.inv" , "pubseccorrup.lag"   , "presidential1" , "inc.termlimit.num1"  ,  "physinteg.1lag" , "physintlag.sq" , "urban.pct")) %>% set_caption("Table 3 cont.: Zero-inflated models of protest initiation")
+t1b <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, omit_coefs = c("(Intercept)", "jucon.1lag" , "legcon.1lag" , "gdpgro.1lag" , "altinf.1lag" , "libdem.1lag", "pubseccorrup.lag" , "osorg.1lag" , "econ.crisis.nelda.1lag1" , "natlpartyorg.lag" , "v2elirreg.inv" , "pubseccorrup.lag"   , "presidential1" , "inc.termlimit.num1"  ,  "physinteg.1lag" , "physintlag.sq" , "urban.pct")) %>% set_caption("Table 3 cont.: Zero-inflated models of protest initiation")
 
@@
 models_t2 <- list()
-models_t2[['Model 3']] <- model.n.base.hurd.mm
-models_t2[['Model 4']] <- model.n.base2.hurd.mm
-models_t2[['Model 5']] <- model.n.griev.hurd.mm
-models_t2[['Model 6']] <- model.size.base
-models_t2[['Model 7']] <- model.size.base2
-models_t2[['Model 8']] <- model.size.griev
+models_t2[['Model 6']] <- model.n.base.hurd.mm
+models_t2[['Model 7']] <- model.n.base2.hurd.mm
+models_t2[['Model 8']] <- model.n.griev.hurd.mm
+models_t2[['Model 9']] <- model.size.base
+models_t2[['Model 10']] <- model.size.base2
+models_t2[['Model 11']] <- model.size.griev
 
@@
 
-t2b <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), omit_coefs = c("(Intercept)", "jucon.1lag" , "legcon.1lag" , "gdpgro.1lag" , "altinf.1lag" , "pubseccorrup.lag" , "osorg.1lag" , "econ.crisis.nelda.1lag1" , "natlpartyorg.lag" , "v2elirreg.inv" , "pubseccorrup.lag"   , "presidential1" , "inc.termlimit.num1"  ,  "physinteg.1lag" , "physintlag.sq" , "urban.pct")) %>% set_caption("Table 4 cont.: Negative binomial models of number and duration of protests") %>% insert_row("", "", "Number", "", "", "Size", "", after = 0)
+t2b <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), omit_coefs = c("(Intercept)", "jucon.1lag" , "legcon.1lag" , "gdpgro.1lag" , "altinf.1lag" , "libdem.1lag",  "pubseccorrup.lag" , "osorg.1lag" , "econ.crisis.nelda.1lag1" , "natlpartyorg.lag" , "v2elirreg.inv" , "pubseccorrup.lag"   , "presidential1" , "inc.termlimit.num1"  ,  "physinteg.1lag" , "physintlag.sq" , "urban.pct")) %>% set_caption("Table 4 cont.: Negative binomial models of number and duration of protests") %>% insert_row("", "", "Number", "", "", "Size", "", after = 0)
 
@@
 
-The second set of models challenge the idea that more severe electoral manipulation is associated with an increased risk of anti-government protest, except under extraordinary circumstances.
-Beginning with the zero-inflation portion of the hurdle models, which model the likelihood of *no* protest occurring after the election, the standalone coefficient for election fraud is not significant in Model XX, nor are the interaction coefficients in Models XX and XX.
-Marginal effects plots [include XX] show that there are no significant effects across values of *winner's vote share* or *lagged GDP growth*.
-The most consistently significant relationship is the negative coefficient for *presidential elections*; this indicates that protest initiation is more likely after a presidential election.
-
-While more fraudulent elections are associated with an increasing risk of protest as margins of victory shrink, this increase is not statistically  significant for typical values of manipulation.
-In the bottom panel of Figure  \@ref(fig:margin-prob-fig), typical values for low and high levels of fraud are used (one standard deviation below and above the mean, respectively).
-In this lower panel, the increased risk of protest in the high-fraud scenario is not statistically distinguishable from the low-fraud scenario.
-It is only when comparing extreme values, as in the top panel where the minimum and maximum values for election fraud are used, that the increase in risk becomes significant.
-It is also worth noting that the inclusion of the interaction term does not improve the fit of the model compared to the baseline Model 2.
-In summary, it appears that though the strategic logic of protest can be detected\textemdash protest appears more likely when fraud is high and winning margins low\textemdash this pattern only emerges in unusual scenarios.
-This finding, where the relationship become fraud and protest is only visible at the margins, is similar to that observed in Model 1 above.
+The second set of models challenge the idea that more severe electoral manipulation is associated with an increased risk of anti-government protest.
+Beginning with the zero-inflation portion of the hurdle models in Table \@ref(tab:zero-model-table), which model the likelihood of *no* protest occurring after the election, the standalone coefficient for election fraud is not significant in Model 6.
+A marginal effects plot in the appendix [include XX] shows that there are no significant effect across values of *lagged GDP growth*.
+
+More fraudulent elections do appear associated with an increasing risk of protest initiation as margins of victory shrink, as shown in the top panel of Figure  \@ref(fig:margin-prob-fig).
+Here, typical values for low and high levels of fraud are used (one standard deviation below and above the mean, respectively).
+The top panel shows that more fraudulent elections are associated with a significantly higher risk of protest initiation, but only when winning vote-shares are very small.
 
@@
 
-model.margin.binary_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary_n)
 
-model.margin.binary_size <- glm(no_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary_size <- glm(no_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary_size)
 
-model.binary.griev1_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.binary.griev1_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.binary.griev1_n)
 
-model.binary.griev1_size <- glm(no_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.binary.griev1_size <- glm(no_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.binary.griev1_size)
 
-model.binary.repr_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.binary.repr_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.binary.repr_n)
@@
 plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary_n, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]"))
-plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "", color = "Election fraud") + theme_bw()
+plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw()
 
@@
 
-Turning to the models of number of protest events shown in Table \@ref(tab:zero-model-table-count), Model 2\textemdash which includes no interaction terms\textemdash shows no significant relationship between fraud and protest in the count portions of the hurdle regression.
-Similarly, none of the three models that include interaction terms yield a significant relationship between fraud, the conditional variable, and number of protest events.
-Model 3 shows no significant interaction effect for fraud and winning vote-share on quantity of protest.
-Figure \@ref(fig:n-protests-margin) shows the marginal effect of election fraud at varying levels of winning vote-share, holding other variables constant at their means (continuous variables) or modes (categorical variables).
-The high and low levels of fraud in the plot represent one standard deviation above and below the mean, respectively.
-As the figure shows, the predicted number of post-election protests is somewhat higher in the high fraud condition when winning margin is low, but this relationship is not statistically significant.
-
-Neither economic crises nor public sector corruption is associated with a significant marginal effect for typical levels of election fraud on protest quantity, as Figures \@ref(fig:n-protests-crisis) and \@ref(fig:n-protests-corrupt) show.\footnote{Figures demonstrating minimum and maximum values for fraud are shown in the appendix (page 5); they also indicate no significant relationship.} In fact, more fraudulent elections appear to be associated with *fewer* incidents of protest during periods of economic crisis, though this distinction is not statistically significant at the p < .05 level.
-
-Altogether, the argument that election fraud increases the risk of protest initiation is mixed.
-Hypothesis 1 is weakly supported: pre-election protest is associated with reduced election-day fraud, but the substantive effect of this relationship is minimal.
-Hypothesis 2 is also weakly supported; there is a significant and sizable association between election fraud and protest initiation when the ruling party's winning margin is low, but only at extremely low and high values of fraud.
-Support for Hypothesis 3 is mixed\textemdash fraud severity is not associated with increased protest risk during economic crises, though there is a significant and sizable increase in the marginal effect as public sector corruption increases.
-Election fraud appears to be associated with a higher risk of protest initiation, but generally in unusual scenarios\textemdash when comparing a highly fraudulent to a largely clean elections when the incumbent's margin of victory is quite low, for example\textemdash or with meager substantive effect.
-
-However, even this mixed finding likely presents too optimistic a picture for the protest-deterrent model of election.
-fraud.
-The count results from the hurdle regressions show no effect for election fraud across any of the models, alone or in interaction.
-Elections that were more fraudulent are not associated with more numerous protest.
+Turning to the models of number and size of protest events, conditional on protest initiation, shown in Table \@ref(tab:zero-model-table-count).
+Models 6 and 9\textemdash which includes no interaction terms\textemdash show no significant relationship between fraud and and the number or maximum size of protest events.
+The bottom panel of Figure \@ref(fig:n-protests-margin) shows the marginal effect of election fraud on the number of protests, given varying levels of winning vote-share, holding other variables constant at their means (continuous variables) or modes (categorical variables).
+As the figure shows, there is no statistically significant difference in the expected number of protests.
+No significant effects are found for fraud and GDP growth in Models 8 or 11.
+Nor is there a significant effect for fraud on protest size, conditional on margin of victory, as discussed below.
+
+Altogether, there scanty evidence to support the argument that protest risk deters election fraud.
+Models 1 through 5 show that various measures of pre-election protest risk are not associated with reduced fraud\textemdash in fact, some appear to increase the severity of fraud.
+Models 6 through 11 show that higher levels of fraud is not "punished" by more or larger protests.
+Hypothesis 2 is also weakly supported; there is a significant and sizable association between election fraud and protest initiation when the ruling party's winning margin is low, but only at extremely low winning margins.
+However, even this suppotive finding likely presents too optimistic a picture for the protest-deterrent model of election fraud.
+The number and size results from the hurdle regressions show no effect for election fraud across any of the models, alone or in interaction.
 Even when protests are initiated\textemdash a relatively rare event in the first place\textemdash the size of a protest wave is not influenced by the degree of fraud perpetrated by the ruling party.
@@
 
-\singlespace
 ```{r zero-model-table-count, echo=FALSE}
@@
 
-One objection to this framing may be that, if the level of manipulation and the risk of protest are endogenous to one another, perhaps the null result found for the number of protests is to be expected.
-That is, perhaps we are observing the equilibrium behavior predicted in many formal models: the threat of protest keeps fraud lower than it would be otherwise, resulting in limited (if any) protest.
+One objection to this framing may be that, if the level of manipulation and the risk of protest are endogenous to one another, perhaps the null result found for the number and size of protests is the result of equilibrium behavior predicted in many formal models.
+The threat of protest keeps fraud lower than it would be otherwise, resulting in limited (if any) protest.
 In some states, this equilibrium level of fraud is higher than in others, resulting in no correlation between fraud and observed protest.
-There are two reasons to be skeptical of this interpretation, one empirical and one substantive.
-
-Empirically, deterrence models would expect the equilibrium level of fraud to fall as protest risk increases, and for more severe protest to occur as a result of miscalculation by the regime.
+There are three reasons to be skeptical of this interpretation, two empirical and one substantive.
+
+First, such an objection does not account for the unexpected results found in Models 1 through 5, where indicators of pre-election protest risk were found to be associated with more severe election-day fraud.
+Second, deterrence models would expect the equilibrium level of fraud to fall as protest risk increases, and for more severe protest to occur as a result of miscalculation by the regime.
 We can test this by examining the predicted intensity of protest from the models specified here.
 If the severity of fraud is held constant at a high level while underlying protest risk increases, we should see an increase in the predicted protest outcome.
-Taking protest size as an example, using Model XX, Figure \@ref(fig:pred-size) shows there is no evidence of such an effect.
-Both panels show the marginal effect of a relatively fraudulent election (*intentional irregularities* is set to one standard deviation above the mean) on the predicted maximum protest size, conditional on the closeness of the election result.
+Taking protest size as an example, using Model 10, Figure \@ref(fig:pred-size) shows there is no evidence of such an effect.
+Both panels show the marginal effect of a relatively fraudulent election (*election fraud* is set to one standard deviation above the mean) on the predicted maximum protest size, conditional on the closeness of the election result.
 All control variables have been set to their means or modes with the exception of lagged GDP growth rate.
@@
 
-To examine the proposition that the information value of election fraud is low, relative to other indicators of regime and opposition strength, I employ Cox proportional-hazards models of protest-wave duration.
-That data in these models is limited to those cases where protests are initiated; in this sense, they represent the cases where opposition groups were most optimistic about challenging the government (or most aggrieved).
-In Table XX, coefficients with a negative sign indicate factors that are associated with longer-lived protest movements; those with a positive sign indicate factors associated with a speedier demise.
-As the models in Table XX show, more severe election fraud is not associated with longer-lasting protest (which would indicate a more determined opposition and/or a weaker ruling party).
-Instead, we observe that repressive capacity is the variable most consistently associated with protest duration\textemdash as repressive capacity eases initially, protest waves last longer; as repressive capacity diminishes further, protest waves again become shorter.
-This suggests that more open, democratic states are more likely to offer concessions to protesters.
-
-```{r, echo=FALSE, warning=FALSE}
+```{r, include=FALSE}
+library(survival)
+library(ggsurvfit)
+nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
+nelda2.sub <- nelda2.sub %>% mutate(fraud_level = ifelse(v2elirreg.inv >= (.98 + .97), "High",
+                                                            ifelse(v2elirreg.inv <= (.98 - .97), "Low", "Moderate")))
+
+nelda2.sub.km <- nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0)
+
+km <-  with(nelda2.sub.km, Surv(d.events.num, surv_status))
+
+survfit2(Surv(d.events.num, surv_status) ~ 1, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
+  ggsurvfit() +
+#   add_confidence_interval() +
+  labs(
+    x = "Days",
+    y = "Overall survival probability"
+  )
+
+survfit2(Surv(d.events.num, surv_status) ~ fraud_level, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
+  ggsurvfit() +
+#   add_confidence_interval() +
+  labs(
+    x = "Days",
+    y = "Overall survival probability"
+  )
+
+cox1 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+summary(cox1)
+cz1 <- cox.zph(cox1, transform = "identity")
+print(cz1)
+plot(cz1[8])
+abline(h=0)
+abline(v = 11)
+abline(v = 25)
+abline(v = 90)
+
+cox1t <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv + tt(v2elirreg.inv) + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag , data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(11, 25, 90),
+Boundary.knots = FALSE))
+summary(cox1t)
+
+cz <- cox.zph(cox1, transform = function(time) 1/(log(time + 1)))
+print(cz)
+plot(cz) # The points on the plot are residuals of v2elirreg.inv (and other vars) based on hte predicted value of X and time t.
+The line shows the estimated value of the coefficient on variable X at time t compared to its baseline.
+You want it to be zero for consistent effects over time.
+V2elirreg.inv is positive for about two weeks early on.
+The substantive effect of this can be shown by taking two values of v2elirreg.inv (i.e.
+first and third quartile), taking the difference, and finding exp(coefficient_residual * difference).
+The result shows here that protests in states in the third quartile of fraud are exp(.75*1.4) = 2.86 times more likely to die than those in the first quartile during those couple weeks.
+So this suggests that if protests survive the first week or two, the state's capacity starts to kick in?
+
+sandbox_data <- survSplit(Surv(d.events.num, surv_status) ~ 
+                      jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + fraud_level  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+                      data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), cut = c(2), episode = "tgroup")
+
+cox1.2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + fraud_level*(as.factor(tgroup))  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = sandbox_data)
+summary(cox1.2)
+
+cox2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+summary(cox2)
+
+cox3 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag , data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+summary(cox3)
+
+cox4 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
+summary(cox4)
+
+```
+
+This dataset can be used for a test of the idea that election fraud is not informative about the duration of protest outside of an initial window.
+To do so, I employ Cox proportional-hazards models of protest-wave duration after subsetting the data to those cases where protests are initiated but incumbents survived; in this sense, they represent the cases where opposition groups were most optimistic about challenging the government (or most aggrieved), but ultimately were overmatched.
+In Table \@ref(tab:cox-models-tab), coefficients with a negative sign indicate factors that are associated with longer-lived protest movements; those with a positive sign indicate factors associated with a speedier demise.
+
+In Model 12, all covariates' effects are modeled as time-invariant.
+However, diagnostics of the model suggest that this assumption may not be upheld for *election fraud*.^[Schoenfeld residuals for *election fraud* are not independent of time, as calculated using the `survival` package in R [@survival-package].] To adjust for this, Model 13 allows the coefficient for *election fraud* to take on different values at different points in time.
+Based on the diagnostics, I allow the value of the coefficient to vary for three periods: 1 - 11 days, 12 - 25 days, and 26 days to the maximum value.
+Model 13 shows the result of this analysis; in it, *election fraud* has a positive and significant coefficient for the earliest time period.
+This indicates that, for the first 11 days of a protest movement, states that engaged in more election fraud are more likely to be able to quash the protests (whether by repression or accommodation).
+After this period, there is no significant relationship between the level of fraud and the survival of the protest wave.
+That is, if a protest movement has survived longer than (roughly) a week and a half after the election, the severity of fraud does not influence its duration.
+
+These results indicate, first, that the influence of a fraudulent election on the duration of mass protest diminishes over time.
+It also suggests that most protests are failed bets.
+Upon taking to the streets, protesters are more likely to find themselves facing a capable incumbent regime the more fraudulent the election was.
+Finally, if protests survive beyond the short term, the severity of fraud no longer has any effect.
+Instead, structural indicators\textemdash the strength of the opposition in the legislature, the degree to which parties are organized across the country, and (to a lesser extent) the openness of civil society are all associated with longer-lived protest movements.
+Similarly, a more fractured incumbent regime, as indicated by a term-limited incumbent, also make protest survival more likely.
+All of this is supportive of the theoretical model laid out above.
+
+```{r cox-models-tab, echo=FALSE, warning=FALSE}
 models3 <- list()
-models3[['Model 5']] <- cox1
-models3[['Model 6']] <- cox2
-models3[['Model 7']] <- cox3
+models3[['Model 12']] <- cox1
+models3[['Model 13']] <- cox1t
+#models3[['Model 14']] <- cox3
 

### 2023-06-05 Scripts/data_cleaning.Rmd
Updates to text + section on Cox models

@@
 nelda <- nelda %>% rename(opp_leaders_prevented = nelda13)
+nelda <- nelda %>% rename(unscheduled_election = nelda6)
+nelda <- nelda %>% rename(inc.run = nelda21)
+nelda <- nelda %>% rename(chosen.successor = nelda22)
 
@@
 
+nelda <- nelda %>% mutate(unscheduled_election = ifelse(is.na(unscheduled_election)==T, NA, ifelse(unscheduled_election == "yes" | unscheduled_election == "N/A", 1, 0)))   #"N/A" here refers to no common expectation for election timing
+
 nelda <- nelda %>% mutate(econ.crisis.nelda = ifelse(econ.crisis.nelda == "yes", 1, ifelse(econ.crisis.nelda == "no", 0, NA)))
@@
 nelda <- nelda %>% mutate(new.elections.held = ifelse(new.elections.held == "yes", 1, 0))
+
+nelda <- nelda %>% mutate(chosen.successor = ifelse(inc.run == "no" & chosen.successor == "yes", 1, 0))
 
@@
 ```{r}
-nelda <- nelda %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "gwn", destination = "vdem"))
+nelda <- nelda %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "gwn", destination = "vdem", 
+                                                   custom_match = c("345" = "198"))) #Fixes Serbia; only non-match that needs fixing; others are microstates / not in V-Dem
  #Missing matches here look to be countries below the population threshold: some Carib. islands, Monaco, etc.
 lji <- lji %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "cown", 
-                                               destination = "vdem"))
+                                               destination = "vdem",
+                                               custom_match = c("345" = "198"))) #ID code is the same as above
 
@@
 
-ecav <- ecav %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "cown", destination = "vdem"))
+ecav <- ecav %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "cown", destination = "vdem",
+                                                 custom_match = c("345" = "198")))
 
@@
 
-```{r}
+```{r, warning=FALSE}
 vdem.recent <- vdem.data %>% filter(year >= 1944)
@@
 vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(reform_positive.lag = lag(reform_positive))
-vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(v2elirreg.inv.lag = lag(v2elirreg.inv))
 vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(elcompindex.m.1lag = lag(elcompindex.m))
@@
 
+## Getting lagged election fraud
+countryids <- data.frame(unique(vdem.recent$country_id))
+lag_with_ids <- data.frame(country_id_year = NA, v2elirreg.inv.lag = NA)
+
+for (i in 1:nrow(countryids)){
+  #lag_with_ids <- data.frame(country_id_year = NA, v2elirreg.inv.lag = NA)
+  country <- vdem.recent %>% filter(country_id == countryids[i,1])
+  country <- country %>% filter(is.na(v2elirreg.inv) == F)
+  country <- country %>% mutate(v2elirreg.inv.lag = lag(v2elirreg.inv))
+  lag_with_ids_temp <- country %>% dplyr::select(country_id_year, v2elirreg.inv.lag)
+  lag_with_ids <- bind_rows(lag_with_ids, lag_with_ids_temp)
+}
+lag_with_ids <- lag_with_ids %>% dplyr::select(-COWcode)
+vdem.recent <- vdem.recent %>% left_join(lag_with_ids, by = "country_id_year")
+
 ##Leading variables for protest after election years
@@
 
-nelda2 <- nelda2 %>% merge(vdem.recent %>% dplyr::select(v2elintmon.lag, v2elintmon, v2eltvrexo, v2elvotlrg, v2elvotsml, v2ellostlg, v2ellostsl, v2ellostsm, v2ellostss, v2ellovtlg, v2ellovtsm, v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, v2x_egaldem, v2x_freexp_altinf, v2x_api, v2x_mpi, v2x_frassoc_thick, v2x_suffr, v2xel_frefair, v2x_elecoff, v2x_liberal, v2x_jucon, v2x_partip, v2x_cspart, v2xeg_eqdr, v2psoppaut, v2jureform, v2jupurge, v2jupoatck, v2jupack, v2juaccnt, v2juhcind, v2juncind, v2mecenefm, e_migdppcln, e_migdpgro, e_miinflat, v2xlg_legcon, legcon.1lag, polyarchy.1lag, libdem.1lag, partipdem.1lag,
+nelda2 <- nelda2 %>% merge(vdem.recent %>% dplyr::select(v2elintmon.lag, v2elintmon, v2eltvrexo, v2elvotlrg, v2elvotsml, v2ellostlg, v2ellostsl, v2ellostsm, v2ellostss, v2ellovtlg, v2ellovtsm, v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, v2x_egaldem, v2x_freexp_altinf, v2x_api, v2x_mpi, v2x_frassoc_thick, v2x_suffr, v2xel_frefair, v2x_elecoff, v2x_liberal, v2x_jucon, v2x_partip, v2x_cspart, v2xeg_eqdr, v2psoppaut, v2jureform, v2jupurge, v2jupoatck, v2jupack, v2juaccnt, v2juhcind, v2juncind, v2mecenefm, e_migdppcln, e_migdpgro, e_miinflat, v2xlg_legcon, legcon.1lag, polyarchy.1lag, libdem.1lag, partipdem.1lag, v2elirreg.inv.lag,
                                                   delibdem.1lag,
@@
                                                   reform_positive.lag,
+                                                  e_regionpol,
                                                   e_mipopula,
@@
 urban <- read.csv(here::here("Data", "urbanization_worldbank.csv"))
+#urban <- urban %>% rename(Country.Name = X...Country.Name)
 test <- urban %>%
@@
 test <- test %>% rename(urban.pct = value)
-test <- test %>% dplyr::select(year, urban.pct, stateid)
 
-nelda2 <- nelda2 %>% mutate(year = as.character(year))
+test <- test %>% mutate(country_id = countrycode(sourcevar = Country.Code, origin = 'iso3c', destination = 'vdem'))
+test <- test %>% mutate(country_id_year = paste0(country_id, "_", year))
 
-nelda2 <- nelda2 %>% left_join(test, by = c("stateid", "year"))
+test <- test %>% dplyr::select(country_id_year, urban.pct)
+
+nelda2 <- nelda2 %>% left_join(test, by = "country_id_year")
 
@@
                                                                 ifelse(ViolenceInitiator == "2.1" & Target2Side == 1, 1, 0)))))))
+```
+
+Coding pro-gov violence from ecav
+
+```{r}
+## Coding pro-government initiation of violence
+
+ecav <- ecav %>% mutate(progov.violence = 
+                          ifelse(EventViolence == 0, 0,
+                                 ifelse(ViolenceInitiator == "-99", NA,
+                                                  ifelse(ViolenceInitiator == ".1" & Actor1Side == 0, 1, 
+                                                  ifelse(ViolenceInitiator == ".2" & Actor2Side == 0, 1,
+                                                         ifelse(ViolenceInitiator == "1.1" & Target1Side == 0, 1,
+                                                                ifelse(ViolenceInitiator == "2.1" & Target2Side == 0, 1, 0)))))))
 ```
@@
 
-## Violence
+## Anti-gov Violence
 violence_data_post <- ecav.post %>% group_by(NeldaID) %>% summarize(antigov.violence = sum(antigov.violence, na.rm = T))
 violence_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(antigov.violence.pre = sum(antigov.violence, na.rm = T))
+
+## Pro-gov Violence
+pg_violence_data_post <- ecav.post %>% group_by(NeldaID) %>% summarize(progov.violence = sum(progov.violence, na.rm = T))
+pg_violence_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(progov.violence.pre = sum(progov.violence, na.rm = T))
 
@@
 ecav_data_post <- left_join(ecav_data_post, violence_data_post, by = "NeldaID")
+ecav_data_post <- left_join(ecav_data_post, pg_violence_data_post, by = "NeldaID")
 
@@
 ecav_data_pre <- left_join(ecav_data_pre, violence_data_pre, by = "NeldaID")
+ecav_data_pre <- left_join(ecav_data_pre, pg_violence_data_pre, by = "NeldaID")
 
@@
 ```{r}
-nelda2.sub <- nelda2 %>% filter(v2x_regime == 1 | v2x_regime == 2) #Note that coding scheme for electoral democracies allows for "some degree of fraud and irregularities but these did not in the end affect the outcome" (v2elfrfair, which is a component of v2x_regime)
+nelda2.sub <- nelda2 %>% filter(v2x_regime > 0) #Note that coding scheme for electoral democracies allows for "some degree of fraud and irregularities but these did not in the end affect the outcome" (v2elfrfair, which is a component of v2x_regime)
 ```
@@
 nelda2.sub <- nelda2.sub %>% filter(election_mdy >= "1989-10-04")
-nelda2.test <- nelda2.sub %>% filter(election_mdy <= "2013-01-13")
+nelda2.sub <- nelda2.sub %>% filter(election_mdy <= "2013-01-13")
 ```
@@
 nelda2.sub <- nelda2.sub %>% mutate(antigov.violence = ifelse(is.na(antigov.violence)==T, 0, antigov.violence))
+nelda2.sub <- nelda2.sub %>% mutate(antigov.violence = ifelse(is.na(progov.violence)==T, 0, progov.violence))
+
 nelda2.sub <- nelda2.sub %>% mutate(n.events.post = ifelse(is.na(n.events.post)==T, 0, n.events.post))
@@
 nelda2.sub <- nelda2.sub %>% mutate(antigov.violence.pre = ifelse(is.na(antigov.violence.pre) == T, 0, antigov.violence.pre))
+nelda2.sub <- nelda2.sub %>% mutate(progov.violence.pre = ifelse(is.na(progov.violence.pre) == T, 0, antigov.violence.pre))
 
@@
 nelda2.sub <- nelda2.sub %>% mutate(antigov.violence = ifelse(is.na(antigov.violence) == T, 0, antigov.violence))
+nelda2.sub <- nelda2.sub %>% mutate(progov.violence = ifelse(is.na(progov.violence) == T, 0, progov.violence))
 
@@
  - Slow or no growth after fraud indicates opp believes state is strong
- - Quick decay after peak indicates state is strong / believed strong
+ - Quick decay after peak indicates state is strong / believed strong (or not--regime is weak and so quickly collapses)
  - Slow decay after peak indicates state is weak / believed weak
@@
 
+## Remove things other than protests
+ecav_prot <- ecav %>% filter(str_detect(EventName, "rotest") == T | str_detect(EventName, "ccupation") == T |
+                                  str_detect(EventName, "lockade") == T) #Only captures protests
+```
+
+Replacing NA protest size with average for that wave
+```{r}
 ## Set up holding df
 
-dv_df <- data.frame(NeldaID = NA, prot_intensity = NA, decay_intensity = NA)
+dv_df <- data.frame(NeldaID = NA, avg_size = NA)
 
-## Replace categorical with min value, replace -99s
+## Filter by election and replace -99 with NA for protest size
+
+ election_ids <- unique(ecav_prot$NeldaID)
+ecav_prot <- ecav_prot %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
+
+for(i in 1:length(election_ids)){ 
+  ecav_temp <- ecav_prot %>% filter(NeldaID == election_ids[i])  #There is a loop here
+
+    ## Remove protests before election
+  ecav_temp <- ecav_temp %>% filter(Date >= Electiondate)
+  if(nrow(ecav_temp) == 0){
+     new_row <- data.frame(NeldaID = election_ids[i], 
+                           avg_size = 0) #No post-election protest
+
+  dv_df[i,] <- new_row
+  print(paste0("No post-election protest for observation ", i, "."))
+  next
+  } else{
+
+  new_row <- data.frame(NeldaID = election_ids[i], avg_size = floor(mean(ecav_temp$ParticipantNumber, na.rm = T)))
+
+  dv_df[i,] <- new_row
+  print(paste0("Average size for observation ", i, " completed."))
+  }
+}
+
+## Merge dv_df back into main dataset
+
+ecav_prot <- ecav_prot %>% left_join(dv_df, by = "NeldaID")
+
+```
+
+Current to-do: Double-check the output of the size code
+
+```{r}
+## Set up holding df
+
+dv_df <- data.frame(NeldaID = NA, prot_intensity = NA, decay_intensity = NA, max_size = NA, day_max = NA)
+
+ecav_prot <- ecav_prot %>% mutate(ParticipantNumber_adj = ifelse(is.na(ParticipantNumber)==T, avg_size, ParticipantNumber))
+ecav_prot <- ecav_prot %>% mutate(ParticipantNumber_adj = ifelse(ParticipantNumber_adj == "NaN", 2, ParticipantNumber_adj)) #Assume NA size defaults to category 2
 
 #ecav_rus <- ecav %>% filter(country == "Russia")   # Filter by country
-ecav <- ecav %>% mutate(ParticipantNumber_num_min = ifelse(ParticipantNumber == 1, 2, 
-                                ifelse(ParticipantNumber == 2, 10, 
-                                       ifelse(ParticipantNumber == 3, 100, 
-                                              ifelse(ParticipantNumber == 4, 1000,
-                                                     ifelse(ParticipantNumber == 5, 10000, 10)))))) #Last 10 here hard-codes any missing participant size as 10
+ecav_prot <- ecav_prot %>% mutate(ParticipantNumber_num_min = ifelse(ParticipantNumber_adj == 1, 5, 
+                                ifelse(ParticipantNumber_adj == 2, 55, 
+                                       ifelse(ParticipantNumber_adj == 3, 550, 
+                                              ifelse(ParticipantNumber_adj == 4, 5500,
+                                                     ifelse(ParticipantNumber_adj == 5, 10000, NA))))))
 
@@
 
- election_ids <- unique(ecav$NeldaID)
+ election_ids <- unique(ecav_prot$NeldaID)
 
 for(i in 1:length(election_ids)){ 
- ecav_temp <- ecav %>% filter(NeldaID == election_ids[i])  #There is a loop here
+  ecav_temp <- ecav_prot %>% filter(NeldaID == election_ids[i])  #There is a loop here
 
@@
  ecav_temp_comb <- ecav_temp_comb %>% filter(date_col >= election_date_temp)
+
   if (nrow(ecav_temp_comb) == 0) {
-    new_row <- data.frame(election_ids[i], NA, NA)
+    new_row <- data.frame(election_ids[i], NA, NA, NA, NA)
     dv_df[i,] <- new_row
@@
 
-  if (nrow(max_data) > 1) {
+  # Assign NA to NA size
+ if (nrow(max_data)==0){
+    new_row <- data.frame(election_ids[i], NA, NA, NA, NA)
+    dv_df[i,] <- new_row
+    print(paste0("Election observation ", i, " completed. (Note: no size estimate."))
+    next
+ }
+
+  if (nrow(max_data) >= 1) {
     prot_intensity <- max_data$total[1] / (1 + as.numeric(ymd(max_data$date_col[1]) - ymd(election_date_temp))) #[1] gets the first peak
@@
     decay_intensity <-  last_peak$total / (1 + as.numeric(max(ecav_temp_comb$date_col) - (ymd(last_peak$date_col))))
-    new_row <- data.frame(election_ids[i], prot_intensity, decay_intensity)
-  #new_row <- new_row %>% rename(NeldaID = election_ids.i.)
+    day_max <- as.numeric(ymd(max_data$date_col[1]) - ymd(election_date_temp))
+    new_row <- data.frame(election_ids[i], prot_intensity, decay_intensity, max_data$total[1], day_max)
+  new_row <- new_row %>% rename(max_size = max_data.total.1.)
     dv_df[i,] <- new_row
@@
 
-  new_row <- data.frame(election_ids[i], prot_intensity, decay_intensity)
-  #new_row <- new_row %>% rename(NeldaID = election_ids.i.)
+  day_max <- as.numeric(ymd(max_data$date_col[1])) - as.numeric(ymd(election_date_temp))
+  new_row <- data.frame(election_ids[i], prot_intensity, decay_intensity, max_data$total, day_max)
+  new_row <- new_row %>% rename(max_size = max_data.total)
+
   dv_df[i,] <- new_row
@@
 ```  
+Pick up here with fixing this code:
+
+```{r, eval=FALSE}
+dv_df <- data.frame(NeldaID = NA, growth_coef = NA)
+for(i in 1:length(election_ids)){
+ ecav_temp <- ecav_prot %>% filter(NeldaID == election_ids[i])  #There is a loop here
+
+###Getting a regression slope for protest size and time
+  ecav_temp_comb <- ecav_temp %>% mutate(date_col = date(Date)) %>%
+    group_by(date_col) %>%
+     summarize(total = sum(ParticipantNumber_num_min))
+  max_date_reg <- ecav_temp_comb %>% filter(total == max(total)) %>%  
+   dplyr::select(date_col)
+
+ ecav_temp_reg <- ecav_temp %>% filter(as_date(Date) >= as_date(Electiondate))
+ ecav_temp_reg <- ecav_temp_reg %>% mutate(date_col = date(Date))
+ ecav_temp_reg <- ecav_temp_reg %>% left_join(ecav_temp_comb, by = "date_col")  #Gets totals by date
+
+ ecav_temp_reg <- ecav_temp_reg %>% filter(as_date(Date) <= max_date_reg[1,1])
+ ecav_temp_reg <- ecav_temp_reg %>% mutate(date_index = as.numeric(ymd(ecav_temp_reg$Date) - ymd(Electiondate)))
+ if(nrow(ecav_temp_reg) == 0) {
+   new_row <- data.frame(election_ids[i], NA)
+   dv_df[i,] <- new_row
+   print(paste0("No observations for election number ", i, "."))
+   next
+ }
+ if(length(unique(ecav_temp_reg$Date))==1){
+   growth_coef <-  ((ecav_temp_reg$total - 0 )/((ecav_temp_reg$date_index+1)-0))[1] #m = (y2 - y1)/(x2 - x1) #+1 in denominator is a kludge to avoid Inf values for day 0 protests
+   new_row <- data.frame(election_ids[i], growth_coef)
+   dv_df[i,] <- new_row
+   print(paste0("One date observed for election ", i, "."))
+   next
+ } else {
+
+ fit <- lm(total ~ 0 + date_index, ecav_temp_reg)
+ sum_table <- summary(fit)
+ growth_coef <- sum_table$coefficients[1]
+  new_row <- data.frame(election_ids[i], growth_coef)
+   dv_df[i,] <- new_row
+   print(paste0("Coefficient estimated for election ", i, "."))
+ }
+ rm(growth_coef)
+}
+
+## Merge dv_df back into main dataset
+
+nelda2.sub.b <- nelda2.sub.b %>% left_join( dv_df, by = "NeldaID")
+```
+
+```{r}
+### Getting a regression slope for protest decay over time
+# last_date_reg <- ecav_temp_reg %>% filter(Date == max(Date)) %>% slice_head(n=1) %>%
+#   dplyr::select(Date)
+# ecav_temp_reg_decay <- ecav_temp_reg %>% filter(Date >= as_date(max_date_reg$date_col))
+#  ecav_temp_reg_decay <- ecav_temp_reg_deacy %>% mutate(date_index_decay = as.numeric(as_date(ecav_temp_reg_decay$Date) - as_date(max_date_reg$date_col)))
+#  if(length(unique(ecav_temp_reg_deacy$Date)) == 1){
+#    decay_coef <-  ((0-ecav_temp_reg_decay$total)/(1-0))[1] #m = (y2 - y1)/(x2 - x1) =
+#  } else{
+#    fit_decay <- lm(total ~ 0 + date_index_decay, ecav_temp_reg_decay)
+# sum_table_decay <- summary(fit_decay)
+# decay_coef <- sum_table_decay$coefficients[1]
+#  }
+#}
+
+```
+
+## Coding for prior successful protest in political region
+
+```{r}
+success_cases <- nelda2.sub.b %>% filter(inc.replaced.prot == 1 | nelda35 == "yes") %>% dplyr::select(e_regionpol, year, election_mdy)
+
+## Coding rule here: A successful protest counts in its region for 1 year after the election which sparked it; the wave ends 1 year after the last successful protest in a chain
+
+nelda2.sub.b <- nelda2.sub.b %>% mutate(regional_success_lag = ifelse(e_regionpol == 1 & (election_mdy %within% interval(start = "1990-06-17", end = "1991-06-17") | election_mdy %within% interval(start = "2000-09-24", end = "2001-09-24") | election_mdy %within% interval(start = "2003-11-02", end = "2006-11-06")), 1,
+                                      ifelse(e_regionpol == 4 & (election_mdy %within% interval(start = "1997-04-13", end = "1998-04-13") | election_mdy %within% interval(start = "1998-05-23", end = "1999-05-23") | election_mdy %within% interval(start = "2000-10-22", end = "2001-10-22") | election_mdy %within% interval(start = "2007-07-22", end = "2008-07-22") | election_mdy %within% interval(start = "2010-11-28", end = "2011-11-28")), 1,
+                                             ifelse(e_regionpol == 2 & (election_mdy %within% interval(start = "2000-04-09", end = "2001-04-09") | election_mdy %within% interval(start = "2000-05-28", end = "2001-05-28") | election_mdy %within% interval(start = "2006-02-07", end = "2007-02-07")), 1,
+                                                    ifelse(e_regionpol == 8 & (election_mdy %within% interval(start = "1989-11-22", end = "1990-11-22")  | election_mdy %within% interval(start = "1996-02-15", end = "1997-02-15")), 1, 0)))))
+```
 
@@
 ```{r}
-write.csv(nelda2.sub, here::here("Data", "nelda2_sub_2023-04-19.csv"))
+write.csv(nelda2.sub.b, here::here("Data", "nelda2_sub_2023-05-30.csv"))
+```
+
+## Adding ERT data for regime duration
+
+```{r, eval=FALSE}
+nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2023-05-10.csv"))
+ert <- read.csv(here::here("Data", "ert.csv"))
+ert <- ert %>% filter(year > 1988)
+ert.sub <- ert %>% dplyr::select(year, country_id, country_name, reg_start_year)
+ert.sub <- ert.sub %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
+
+## ID numbers are the same because they both use V-dem format
+
+ert.sub <- ert.sub %>% dplyr::select(country_id_year, reg_start_year)
+
+nelda2.sub <- nelda2.sub %>% left_join(ert.sub, by = "country_id_year")
+
+## Getting age of regime
+
+nelda2.sub  <- nelda2.sub %>% mutate(regime_age = year - reg_start_year)
+
+write.csv(nelda2.sub, here::here("Data", "nelda2_sub_2023-05-30.csv"))
+
 ```

### 2023-06-09 Output/Submission files/manuscript_anon_v2.Rmd
Beginning cleaning up of appendices

@@
 The risk of mass protest in these models is assumed to constitute a significant guardrail for election integrity in democracies, and a significant challenge for ruling parties in non-democracies.
-Building on this logic, researchers have proposed that third-party institutions---like election monitors [@Little2012;@Hyde2014; @daxecker2012], civil society groups [@birch2017getting], and courts [@Chernykh2015]---can reduce manipulation and increase protest risk by revealing information about electoral malfeasance.
+Building on this logic, researchers have proposed that third-party institutions---like election monitors [@Little2012;@Hyde2014;@daxecker2012], civil society groups [@birch2017getting], and courts [@chernykh2015third]---can reduce manipulation and increase protest risk by revealing information about electoral malfeasance.
 Of course, electoral manipulation is not understood to be the sole cause of protest\textemdash protests occur for all sorts of reasons and across time\textemdash but the literature has drawn a close connection between the risk of specifically electoral protest and the conduct of elections.
@@
 The literature on protest as a deterrent to manipulation also often overlooks the functions that elections, even highly controlled elections, may serve in non-democratic settings.
-For instance, Simpser [-@Simpser2011] shows that high levels of manipulation can actually induce cooperation from opposition groups, elites, and other actors by revealing information about regime strength.
+For instance, Simpser [-@simpser2013] shows that high levels of manipulation can actually induce cooperation from opposition groups, elites, and other actors by revealing information about regime strength.
 Similarly, the type of manipulation employed can convey information about the ruling party's capacity to mobilize, with costlier forms of manipulation associated with less protest [@Harvey2020].
@@
 Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304].
-Several empirical studies have, accordingly, found links between election manipulation and binary measures of protest [@Brancati2016;@Hyde2014a;@Harvey2020;@Rod2019].
+Several empirical studies have, accordingly, found links between election manipulation and binary measures of protest [@Brancati2016;@Hyde2014;@Harvey2020;@Rod2019].
 
@@
 
+## Set max_size zero for elections with no protest
+
+nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))
+
 ## Excluding liberal democracies
@@
 
-These hypotheses are tested using data from three datasets: NELDA [@hyde2012], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019].
+These hypotheses are tested using data from three datasets: NELDA [@hyde_2012_marinov], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019].
 The unit of observation for this study is an election period as recorded in NELDA.
@@
                                             'Civil society openness' = osorg.1lag, 
-                                            'Public sector corruption' = pubseccorrup.lag,
+                                            #'Public sector corruption' = pubseccorrup.lag,
                                             'National party organization' = natlpartyorg.lag, 
@@
 
-```{r summary-table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Summary statistics"}
+```{r summary-table, echo=FALSE, message=FALSE, warning=FALSE}
 library(kableExtra)
-datasummary_skim(incl.vars, output = "kableExtra")  %>%
+datasummary_skim(incl.vars, output = "kableExtra", title = "Table 1: Summary statistics")  %>%
     kable_styling(font_size = 9, full_width = FALSE)
@@
 Finally, several binary control variables are included.
-First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@Simpser2011] and a greater risk of protest.
+First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest.
 Next, a variable from NELDA indicates if the chief executive is *term limited* during the election at hand.
@@
 Lastly, since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.
-Figure \@ref(fig:summary-table) shows summary statistics for all of these variables.
+Table \@ref(tab:summary-table) shows summary statistics for all of these variables.
 To address concerns about multicollinearity, I present models in the appendix (page 25) that exclude control variables associated with pre-election protest.
@@
 
-modelsummary(c(m1, m1b, m1c, m1a, m1d), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Random-intercept multilevel model of election-day fraud', coef_rename = 
+modelsummary(c(m1, m1b, m1c, m1a, m1d), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Table 2: Random-intercept multilevel model of election-day fraud', coef_omit = "SD", coef_rename =
                c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
@@
                  "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
-                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election")) %>% kable_styling(latex_options = "HOLD_position")
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
+                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
+integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion"
+                 )) %>% kable_styling(latex_options = "HOLD_position")
 
@@
 ```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
-model.n.novoteshare.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
+model.n.novoteshare.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                                 ziformula=~.,
@@
 
-model.n.base.hurd.mm <- glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+model.n.base.hurd.mm <- glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                ziformula=~.,
@@
 
-model.n.base2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+model.n.base2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                           zi = ~.,
@@
 
-model.n.griev.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+model.n.griev.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +  v2elirreg.inv*gdpgro.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,
@@
 
-model.n.griev2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0"),
-                                family=truncated_nbinom2)
-summary(model.n.griev2.hurd.mm)
-
-model.n.griev3.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag  ,
-                                ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0"),
-                                family=truncated_nbinom2)
-summary(model.n.griev3.hurd.mm)
-
-model.n.repr.hurd <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv  +  physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+model.n.repr.hurd <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +   + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv  +  physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                 ziformula=~.,
@@
 ```{r models-hurdle-size, warning = FALSE, message = FALSE, include = F}
-model.size.novoteshare <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+model.size.novoteshare <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv +  +   v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,
@@
 
-model.size.base <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   v2elirreg.inv + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+model.size.base <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv +  +   v2elirreg.inv + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,
@@
 
-model.size.base2 <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+model.size.base2 <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv +  +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,
@@
 
-model.size.griev <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
+model.size.griev <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +  v2elirreg.inv*gdpgro.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                 ziformula=~.,
@@
 
-```{r zero-model-table, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+```{r initiation-table-setup, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
 
@@
 
-t1a <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, omit_coefs = c("rentierstate1", "taxstate1", "n.events.pre", "unscheduled_election1", "inc.lose1", "v2elirreg.inv:winner.share.upd", "gdpgro.1lag:v2elirreg.inv"), statistics = character(0)) %>% set_caption("Table 3: Zero-inflated models of protest initiation")
+t1a <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, coefs = c("Intercept" = "(Intercept)", "Judicial ind.
+(lag)" = "jucon.1lag", "Liberal democracy index" = "libdem.1lag",
+            "Legislative constraints (lag)" = "legcon.1lag", "Alternative info.
+(lag)" = "altinf.1lag",                             "GDP growth rate (lag)" = "gdpgro.1lag",                               "Civil society openness (lag)" = "osorg.1lag", "National party organization (lag)" = "natlpartyorg.lag",                "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Presidential" = "presidential1", "Incumbent term-limited" = "inc.termlimit.num",
+            "Physical integrity (lag)" = "physinteg.1lag", "Phys. int. (lag) squared" = "physintlag.sq",
+            "Urban" = "urban.pct", "Rentier state" = "rentierstate1", "Tax state" = "taxstate1"),
+              statistics = character(0))  %>% set_caption("Table 3: Zero-inflated models of protest initiation")
 
@@
 
-t1b <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, omit_coefs = c("(Intercept)", "jucon.1lag" , "legcon.1lag" , "gdpgro.1lag" , "altinf.1lag" , "libdem.1lag", "pubseccorrup.lag" , "osorg.1lag" , "econ.crisis.nelda.1lag1" , "natlpartyorg.lag" , "v2elirreg.inv" , "pubseccorrup.lag"   , "presidential1" , "inc.termlimit.num1"  ,  "physinteg.1lag" , "physintlag.sq" , "urban.pct")) %>% set_caption("Table 3 cont.: Zero-inflated models of protest initiation")
+t1b <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, coefs = c("Number of pre-election protests" = "n.events.pre", "Off-schedule election" = "unscheduled_election1", "Regional protest diffusion" = "regional_success_lag", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd", "Election fraud : GDP growth (lag)" = "gdpgro.1lag:v2elirreg.inv")) %>% set_caption("Table 3 cont.: Zero-inflated models of protest initiation")
 
 theme_compact(t1b, header_rows = TRUE, header_cols = FALSE)
-
+```
+
+```{r conditional-table, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
 models_t2 <- list()
@@
 
-t2a <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), omit_coefs = c("rentierstate1", "taxstate1", "n.events.pre", "unscheduled_election1", "inc.lose1", "v2elirreg.inv:winner.share.upd", "gdpgro.1lag:v2elirreg.inv"), statistics = character(0)) %>% set_caption("Table 4: Negative binomial models of number and duration of protests") %>% insert_row("", "", "Number", "", "", "Size", "", after = 0)
+t2a <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), coefs = c("Intercept" = "(Intercept)", "Judicial ind.
+(lag)" = "jucon.1lag",
+            "Legislative constraints (lag)" = "legcon.1lag", "Alternative info.
+(lag)" = "altinf.1lag",                             "Liberal democracy index (lag)" = "libdem.1lag", "GDP growth rate (lag)" = "gdpgro.1lag",                               "Civil society openness (lag)" = "osorg.1lag", "National party organization (lag)" = "natlpartyorg.lag",                "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Presidential" = "presidential1", "Incumbent term-limited" = "inc.termlimit.num",
+            "Physical integrity (lag)" = "physinteg.1lag", "Phys. int. (lag) squared" = "physintlag.sq",
+            "Urban" = "urban.pct", "Rentier state" = "rentierstate1", "Tax state" = "taxstate1"), statistics = character(0)) %>% set_caption("Table 4: Negative binomial models of number and duration of protests") %>% insert_row("", "", "Number", "", "", "Size", "", after = 0)
 
@@
 
-t2b <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), omit_coefs = c("(Intercept)", "jucon.1lag" , "legcon.1lag" , "gdpgro.1lag" , "altinf.1lag" , "libdem.1lag",  "pubseccorrup.lag" , "osorg.1lag" , "econ.crisis.nelda.1lag1" , "natlpartyorg.lag" , "v2elirreg.inv" , "pubseccorrup.lag"   , "presidential1" , "inc.termlimit.num1"  ,  "physinteg.1lag" , "physintlag.sq" , "urban.pct")) %>% set_caption("Table 4 cont.: Negative binomial models of number and duration of protests") %>% insert_row("", "", "Number", "", "", "Size", "", after = 0)
+t2b <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), coefs = c("Number of pre-election protests" = "n.events.pre", "Off-schedule election" = "unscheduled_election1", "Regional protest diffusion" = "regional_success_lag", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd", "Election fraud : GDP growth (lag)" = "gdpgro.1lag:v2elirreg.inv")) %>% set_caption("Table 4 cont.: Negative binomial models of number and duration of protests") %>% insert_row("", "", "Number", "", "", "Size", "", after = 0)
 
@@
 The second set of models challenge the idea that more severe electoral manipulation is associated with an increased risk of anti-government protest.
-Beginning with the zero-inflation portion of the hurdle models in Table \@ref(tab:zero-model-table), which model the likelihood of *no* protest occurring after the election, the standalone coefficient for election fraud is not significant in Model 6.
-A marginal effects plot in the appendix [include XX] shows that there are no significant effect across values of *lagged GDP growth*.
+Beginning with the zero-inflation portion of the hurdle models in Table 3, which model the likelihood of *no* protest occurring after the election, the standalone coefficient for election fraud is not significant in Model 6.
+Figure \@ref(fig:margin-plot-gdp) shows that there are no significant effect across values of *lagged GDP growth*.
 
@@
 
-model.margin.binary_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary_n)
 
-model.margin.binary_size <- glm(no_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary_size <- glm(no_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary_size)
 
-model.binary.griev1_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.binary.griev1_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +  +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.binary.griev1_n)
 
-model.binary.griev1_size <- glm(no_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.binary.griev1_size <- glm(no_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +  +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.binary.griev1_size)
@@
 plot_binary_gdp_typ_n <- sjPlot::plot_model(model.binary.griev1_n, type = "pred", terms = c("gdpgro.1lag [all]", "v2elirreg.inv [-.45, 1.78]"))
-plot_binary_gdp_typ_n <- plot_binary_gdp_typ_n +  labs(x = "GDP growth (lagged)", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()
+plot_binary_gdp_typ_n <- plot_binary_gdp_typ_n +  labs(x = "GDP growth (lagged)", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw() + coord_cartesian(xlim = c(-.4, 3))
 
@@
 ```
-```{r}
+```{r margin-plot-gdp, echo=FALSE, message=F, warning=FALSE, fig.cap="Marginal effect of election fraud on protest initiation, conditional on lagged GDP growth"}
 plot_binary_gdp_typ_n
@@
 p.n.margin <- plot(dat)
-p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Number of events", title = "", color = "Election fraud") + theme_bw()
+p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Number of events", title = "Predicted number of protests", color = "Election fraud") + theme_bw()
 ```
@@
 
-Turning to the models of number and size of protest events, conditional on protest initiation, shown in Table \@ref(tab:zero-model-table-count).
+Turning to the models of number and size of protest events, conditional on protest initiation, shown in Table 4.
 Models 6 and 9\textemdash which includes no interaction terms\textemdash show no significant relationship between fraud and and the number or maximum size of protest events.
-The bottom panel of Figure \@ref(fig:n-protests-margin) shows the marginal effect of election fraud on the number of protests, given varying levels of winning vote-share, holding other variables constant at their means (continuous variables) or modes (categorical variables).
+The bottom panel of Figure \@ref(fig:margin-prob-fig) shows the marginal effect of election fraud on the number of protests, given varying levels of winning vote-share, holding other variables constant at their means (continuous variables) or modes (categorical variables).
 As the figure shows, there is no statistically significant difference in the expected number of protests.
@@
 
-p.gdplow <- sjPlot::plot_model(model.size.base2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(gdpgro.1lag = -.5))
-
-p.gdphigh <- sjPlot::plot_model(model.size.base2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(gdpgro.1lag = 3))
-
-cowplot::plot_grid(p.gdplow, p.gdphigh)
+p.gdplow <- sjPlot::plot_model(model.size.base2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(gdpgro.1lag = -.5)) + labs(x = "Incumbent vote-share", y = "Max.
+size", title = "Predicted max.
+size \n(GDP growth: -0.5)")
+
+p.gdphigh <- sjPlot::plot_model(model.size.base2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(gdpgro.1lag = 3)) + labs(x = "Incumbent vote-share", y = "", title = "Predicted max.
+size \n(GDP growth: 3)")
+
+cowplot::plot_grid(p.gdplow, p.gdphigh, nrow = 1)
 ```
@@
 
-cox1 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+cox1 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + v2elirreg.inv , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
 summary(cox1)
@@
 
-cox1t <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv + tt(v2elirreg.inv) + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag , data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(11, 25, 90),
+cox1t <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + v2elirreg.inv + tt(v2elirreg.inv), data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(11, 25, 90),
 Boundary.knots = FALSE))
@@
 To do so, I employ Cox proportional-hazards models of protest-wave duration after subsetting the data to those cases where protests are initiated but incumbents survived; in this sense, they represent the cases where opposition groups were most optimistic about challenging the government (or most aggrieved), but ultimately were overmatched.
-In Table \@ref(tab:cox-models-tab), coefficients with a negative sign indicate factors that are associated with longer-lived protest movements; those with a positive sign indicate factors associated with a speedier demise.
+In Table 5, coefficients with a negative sign indicate factors that are associated with longer-lived protest movements; those with a positive sign indicate factors associated with a speedier demise.
 
@@
 That is, if a protest movement has survived longer than (roughly) a week and a half after the election, the severity of fraud does not influence its duration.
+It is also noteworthy that this formulation is a much better fit to the data than the linear effect in Model 12, with a much lower root mean square error.
 
@@
 
-modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Cox proportional hazard models of protest duration')
+modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Table 5: Cox proportional hazard models of protest duration', coef_rename = 
+               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
+                 "osorg.1lag" = "Civil soc. openness (lag)", 
+                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
+                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
+integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "v2elirreg.inv" = "Election fraud",
+                 "tt(v2elirreg.inv)1" = "Election fraud (day 12-25)", "tt(v2elirreg.inv)2" = "Election fraud (day 26-90)"
+                 ))
 
@@
 To the extent that fraud is a signal of incumbent strength, it is a noisy and unreliable one.
-When an incumbent party wins by a narrow margin in a highly fraudulent election, the likelihood of protest initiation is substantially higher than when it wins more cleanly: roughly a 50 to 60\% chance in the former case versus 5 to 10\% in the latter as shown in Figure 2.
+When an incumbent party wins by a narrow margin in a highly fraudulent election, the likelihood of protest initiation is somewhat higher than when it wins more cleanly.
 However, since even strong incumbents commit fraud, the decision to take to the streets is often a losing gamble\textemdash anticipating a weak incumbent, they instead find one that is capable of repressing, coopting, or conciliating an emerging protest movement.
-Indeed, while there are 143 cases of protest out of the 647 elections in this dataset, only 9 elections resulted in either the incumbent being replaced or the election being re-run due to mass protests (according to NELDA).\footnote{This information is determined using the NELDA variables NELDA37 and NELDA41.
-The cases of successful protest are: Ukraine 2004, Azerbaijan 2005, Georgia 2003, Kyrgyzstan 2005, Bulgaria 1990, Bangladesh 1996, Mali 1997, Peru 2000, and Cote d'Ivoire 2000.} This suggests a `success rate' for post-election protest of about six percent *conditional on protest initiation*, and about one percent for all elections in electoral authoritarian and electoral democratic regimes.
+Indeed, while there are 194 cases of protest out of the 661 elections in this dataset where incumbents claimed victory, only 8 resulted in either the incumbent being replaced or the election being re-run due to mass protests (according to NELDA).\footnote{This information is determined using the NELDA variables NELDA37 and NELDA41.
+The cases of successful protest are: Ukraine 2004, Azerbaijan 2005, Georgia 2003, Kyrgyzstan 2005, Malaysia 2004, Bangladesh 1996, Peru 2000, and Cote d'Ivoire 2000.} This suggests a `success rate' for post-election protest of about four percent *conditional on protest initiation*, and about one percent for all elections in electoral authoritarian and electoral democratic regimes.
 
@@
 The rarity of post-election protest, especially successful protest, suggests that such events are highly contingent; they hinge on a multitude of factors that go beyond the regime's efforts to bias election outcomes.
-Election fraud is a complex phenomenon with multiple drivers, including resource costs, signaling, information-gathering, principal-agent dynamics, and more; these findings indicate that protest risk is not a significant factor under most circumstances.
+Both protest and election fraud are complex phenomena with multiple drivers, including resource costs, signaling, information-gathering, principal-agent dynamics, and more; these findings indicate that protest risk is not a significant factor under most circumstances.
 
@@
 By overlooking the poor quality of election fraud as a signal, or by treating protest primarily as a binary event, the existing literature on electoral protest has overemphasized the risk to incumbents of engaging in electoral manipulation.
-The results presented here show that fraud has no bearing on how numerous protest events are, and only little influence on protest initiation.
+The results presented here show that fraud has no bearing on how numerous or large protest events are.
 While pre-election protest is associated with reduced fraud, the substantive size of this relationship is very small.
-Moreover, contrary to some prior research, the risk of a large protest wave does not appear to be heightened by poor economic conditions [@Brancati2014a] or intensive corruption in the public sector.
+Moreover, contrary to some prior research, the risk of a large protest wave does not appear to be heightened by poor economic conditions [@Brancati2014a].
 As a result, circumstances that are to some extent outside of leaders' control, like a worsening macroeconomic climate, do not appear to make engaging in fraud any riskier.
@@
 But if protest risk is largely untethered from the severity of fraud, it is likely that other mechanisms are at work.
+Scholarly attention might turn more profitably toward supply-side explanations for manipulation, such as the organizational capacity of the state [@simpser2013], the popularity of the government [@rundlett2016deliver], and the risks faced by front-line agents [@Harvey2022].
 
 These results are consistent with different specifications presented in an online appendix.
-In particular, the main results are supported when data from the Electoral Contention and Violence (ECAV) dataset [@Daxecker2019] are used in place of the Mass Mobilization data (appendix page 1), and when a measure of overall election integrity is used in place of the measure of election fraud (appendix page 15).
-
-Scholarly attention might turn more profitably toward supply-side explanations for manipulation, such as the organizational capacity of the state [@Simpser2011], the popularity of the government [@rundlett2016deliver], the risks faced by front-line agents [@Harvey2022].
+In particular, the main results are supported when data from the Mass Mobilization dataset are used in place of the ECAV data (appendix page 1), and when a measure of overall election integrity is used in place of the measure of election fraud (appendix page 15).
 
@@
 This paper addresses these concerns by testing the relationship between election fraud, the risk that any protest occurs, and the number of protest events that occur after an election.
-Drawing on data from the V-Dem, NELDA, and Mass Mobilization datasets, it uses cross-national data on 647 elections to show that---while there is an elevated risk of some protest occurring in a fraudulent election with narrow margins of victory or in highly corrupt societies---election fraud has no significant relationship with the number of protests that occur in a protest wave.
+Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 1110 elections to show that---while there is an elevated risk of some protest occurring in a fraudulent election with narrow margins of victory or in highly corrupt societies---election fraud has no significant relationship with the number of protests that occur in a protest wave.
 These results imply that protest risk should not be considered a central deterrent to election fraud.
-Instead, most fraudulent elections go un-protested, and most post-election protests are strategic gambles gone wrong.
+Instead, most fraudulent elections go un-protested, and most post-election protests appear to be strategic gambles gone wrong.
 The initiation of protest more often reveals the incumbent to be strong enough to manage protest through repression, cooptation, or other concessions.
-These findings indicate that, rather than being deterred by protest risk, governments have an incentive to manipulate as much as possible in order to achieve a large official vote-share.
 The absence of protest risk as a guardrail is felt in both electoral authoritarian regimes and in electoral democracies\textemdash those democracies that are most at risk of democratic erosion.

### 2023-06-09 Output/appendix_v4.Rmd
Beginning cleaning up of appendices

---
title: "Appendix: A missing guardrail---The minimal effect of election fraud on protest intensity in non-democracies"
author: "Cole J. Harvey (Oklahoma State University)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    bookdown::pdf_document2: default
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
header-includes: \usepackage{booktabs}
---

```{r setup, include = FALSE}
library(tidyverse)
library(countrycode)
library(readxl)
library(vdemdata)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
library(stargazer)
library(ggeffects)
library(pscl)
library(survival)
library(modelsummary)
library(glmmTMB)
library(huxtable)
library(kableExtra)
library(simPH)

nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2023-05-30.csv"))

nelda2.sub <- nelda2.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda2.sub <- nelda2.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))

##Naming nelda variables

nelda2.sub <- nelda2.sub %>% rename(boycott = nelda14)
nelda2.sub <- nelda2.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                 NA)))

##Creating economic crisis variable

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = as.factor(econ.crisis))

nelda2.sub <- nelda2.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))

nelda2.sub <- nelda2.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))

##Removing cases before 1990, when ECAV starts
nelda2.sub <- nelda2.sub %>% filter(year >= 1990)

## Removing cases where inc. loses
#nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)

##Updated winner share
nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))

##Intmon as a factor
nelda2.sub <- nelda2.sub %>% mutate(v2elintmon = as.factor(v2elintmon))

nelda2.sub <- nelda2.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))

#Square of physint

nelda2.sub <- nelda2.sub %>% mutate(physintlag.sq = physinteg.1lag^2)

#Crisis as factor

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))

#Margins with updates

nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))

#Successful protest = new elections or new incumbent

nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))

#Protest cases only

protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1)

# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections

nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))

## Categorical variable for incumbent status

nelda2.sub <- nelda2.sub %>% mutate(incumb.status = as.factor(ifelse(inc.termlimit == "yes" & chosen.successor == 1, "Chosen successor",
                                                           ifelse(inc.termlimit == "yes" & chosen.successor == 0, "Contested exec. election", 
                                                                  ifelse(inc.termlimit == "no", "Executive eligible", "Legislative election"))))) 
nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Legislative election"))

## Coding max_size as zero for elections with no protest
nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))

## Loading massmob

massmob <- read.csv(here::here("Data", "nelda_massmob_all.csv"))
massmob <- massmob %>% rename("NeldaID" = "electionid")
massmob <- massmob %>% dplyr::select(NeldaID, n.events.massmob.pre, n.events.massmob.post)

nelda2.sub <- nelda2.sub %>% left_join(massmob, by = "NeldaID")

```

# Alternative protest data: Mass Mobilization dataset

```{r models-hurdle-mm, message = FALSE, echo=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(n.events.massmob.pre.scale = scale(n.events.massmob.pre))

model.n.base.hurd.mm <- pscl::hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +     winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre +  unscheduled_election + regional_success_lag,
                                         zero.dist = "binomial",
                                          dist= "negbin",
                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.n.base.hurd.mm)

model.n.base2.hurd.mm <- pscl::hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +     winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre +  unscheduled_election + regional_success_lag,
                                         zero.dist = "binomial",
                                          dist= "negbin",
                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.n.base2.hurd.mm)

model.n.griev.hurd.mm <- pscl::hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +     gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre +  unscheduled_election + regional_success_lag,
                                         zero.dist = "binomial",
                                          dist= "negbin",
                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.n.griev.hurd.mm)

```

The Mass Mobilization dataset [@massmob] aims to collect data on every protest event with greater than 50 participants in over 160 countries, from 1990 to 2014. Since the 50-participant threshold undercounts small protests, I do not use it in the main text and I do not use it here to test hypotheses on protest size. To match the ECAV format, I limit the window for protests to 180 days prior to the election and 90 days after. Unlike with ECAV, the counted protests need not be explicitly tied to the election. Using Mass Mobilization data generates results similar to those in the main text, as shown in Tables \@ref(tab:table-mm-binary) and \@ref(tab:table-mm-count). Fraud severity has no standalone effect on protest initiation or protest count. Election fraud does have a significant relationship to protest initiation in Model 2, at low incumbent vote-shares, but as Figure \@ref(fig:plot-margin-mm) indicates, this relationship is not apparent at typical levels of election fraud. As in the main text, the main drivers of protest count appear to be institutional and structural\textemdash repressive capacity and opposition organization.

```{r table-mm-binary, echo=FALSE, warning=FALSE, message=FALSE}
models_a1 <- list()
models_a1[['Model 1']] <- model.n.base.hurd.mm
models_a1[['Model 2']] <- model.n.base2.hurd.mm
models_a1[['Model 3']] <- model.n.griev.hurd.mm

modelsummary::modelsummary(models_a1, stars = T, longtable=T, coef_omit = "count", title = 'Binary portion of hurdle models of post-election protest', coef_rename = 
               c("zero_(Intercept)" = "Intercept", "zero_jucon.1lag" = "Judical independence (lag)", "zero_legcon.1lag" = "Leg. constraints (lag)", 
                 "zero_gdpgro.1lag" = "GDP growth rate (lag)", "zero_altinf.1lag" = "Alternative info. (lag)", 
                 "zero_osorg.1lag" = "Civil soc. openness (lag)", "zero_econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "zero_natlpartyorg.lag" = "National party organization (lag)", "zero_presidential" = "Presidential election", "zero_libdem.1lag" = "Liberal democracy index",
                 "zero_inc.termlimit.num" = "Incumbent term-limited",  "zero_regional_success_lag" = "Regional diffusion",
                 "zero_urban.pct" = "Urbanization", "zero_rentierstate" = "Rentier state", "zero_taxstate" = "Tax state",
                 "zero_n.events.massmob.pre" = "Number of pre-election protests", "zero_v2elirreg.inv" = "Election fraud",
                 "zero_physinteg.1lag" = "Physical integrity (lag)", "zero_physintlag.sq" = "Physical integrity (lag), squared", "zero_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "zero_unscheduled_election" = "Off-schedule election"))
```

\newpage

```{r table-mm-count, echo=FALSE, warning=FALSE, message=FALSE}
modelsummary::modelsummary(models_a1, stars = T, longtable=T, coef_omit = "zero", title = 'Count portion of hurdle models of post-election protest', coef_rename = 
               c("count_(Intercept)" = "Intercept", "count_jucon.1lag" = "Judical independence (lag)", "count_legcon.1lag" = "Leg. constraints (lag)", 
                 "count_gdpgro.1lag" = "GDP growth rate (lag)", "count_altinf.1lag" = "Alternative info. (lag)", 
                 "count_osorg.1lag" = "Civil soc. openness (lag)", "count_econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "count_natlpartyorg.lag" = "National party organization (lag)", "count_presidential" = "Presidential election", "count_libdem.1lag" = "Liberal democracy index",
                 "count_inc.termlimit.num" = "Incumbent term-limited",  "count_regional_success_lag" = "Regional diffusion",
                 "count_urban.pct" = "Urbanization", "count_rentierstate" = "Rentier state", "count_taxstate" = "Tax state",
                 "count_n.events.massmob.pre" = "Number of pre-election protests", "count_v2elirreg.inv" = "Election fraud",
                 "count_physinteg.1lag" = "Physical integrity (lag)", "count_physintlag.sq" = "Physical integrity (lag), squared", "count_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "count_unscheduled_election" = "Off-schedule election"))
```

```{r plot-margin-mm, echo=FALSE, message=FALSE, warning=F, fig.cap="Fraud, winning vote-share, and probability of protest initiation (Mass Mobilization data)", fig.height=7}
nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(is.na(n.events.massmob.post)==T, NA, ifelse(n.events.massmob.post > 0, 1, 0)))

model.margin.binary_n <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.margin.binary_n)

plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary_n, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.64, 1.55]"))
plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()

## Count portion

dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.64, 1.55]"), type = "count") 
p.n.margin <- plot(dat)
p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Number of events", title = "Predicted number of protests", color = "Election fraud") + theme_bw()

plot.combined.prob <- cowplot::plot_grid(plot_binary_margin_typ_n, p.n.margin, ncol=1)
plot.combined.prob

```

```{r, include=FALSE}
model.binary.griev1_n <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +  +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.binary.griev1_n)

plot_binary_gdp_typ_n <- sjPlot::plot_model(model.binary.griev1_n, type = "pred", terms = c("gdpgro.1lag [all]", "v2elirreg.inv [-.45, 1.78]"))
plot_binary_gdp_typ_n <- plot_binary_gdp_typ_n +  labs(x = "GDP growth (lagged)", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw() + coord_cartesian(xlim = c(-.4, 3))

## Count portion

dat <- ggpredict(model.n.griev.hurd.mm, terms = c("gdpgro.1lag", "v2elirreg.inv [-.64, 1.55]"), type = "count") 
p.n.gdp <- plot(dat)
p.n.gdp <- p.n.gdp +  labs(x = "GDP growth rate (lag)", y = "Number of events", title = "Predicted number of protests", color = "Election fraud") + theme_bw() + coord_cartesian(xlim = c(-.4, 3), ylim = c(0, 200))
```

## Repression and pre-election protest

The main analysis already controls for possible confounding by the state's repressive capacity; i.e. the possibility that in highly repressive states, protest is deterred and fraud can proceed unopposed. However, it may still be useful to check an alternative explanation\textemdash perhaps protest is deterred in highly repressive states with considerable pre-election unrest (that is, when protest appears to be exceeding the ability of the state to contain). To check this possibility, a variation of Model 1 from the main text is run, but with an interaction between the number of pre-election protests and the lagged *physical integrity* variable. If protest risk deters manipulation in this way, there should be a negative and significant effect for the number of pre-election protests in highly repressive states. Figure \@ref(fig:prot-rep-fraud) shows the absence of such an effect. In fact, there is no significant difference in the predicted level of fraud at lower levels of physical integrity, even when pre-election protest is widespread. A significant effect is seen at the high end of protection of physical integrity rights, but it is present only in extreme cases\textemdash the value of pre-election protests chosen here, 25, is above the 98th percentile for the variable\textemdash and is in the opposite direction predicted by this counterargument. That is, more numerous pre-election protest is associated with higher predicted levels of fraud in less repressive states than when the pre-election period is quiet. This robustness check shows that protest in a repressive environment does not appear to deter fraud, and supports the view that manipulation may be used to attempt to claim a large margin of victory or to prevent elite or bureaucratic splits in a challenging period.

```{r prot-rep-fraud, message=FALSE, warning=F, echo=FALSE, fig.cap="Marginal effect of pre-election protest on fraud, conditional on physical integrity"}
m1_rep <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre*physinteg.1lag  +  unscheduled_election  + (1 | stateid), data = nelda2.sub %>% filter(v2x_regime <= 2)) 
#summary(m1_rep)
p.m1rep <- sjPlot::plot_model(m1_rep, type = "pred", terms = c("physinteg.1lag", "n.events.pre [0, 25]"))
p.m1rep + labs(title = "Predicted values of election fraud", x = "Physical integrity (lag)", y = "Election fraud", color = "N. of pre-election protests") + theme_bw()
```

## Prior manipulation

Another explanation for post-election protest could hinge on changes over time; elections that are much more fraudulent than prior elections may exacerbate citizens' grievances or convince opposition groups that the incumbent is unexpectedly weak. This possibility is tested by replacing the *election fraud* variable with *fraud change*, the difference between the current election and the one immediately prior. Since the prior election's value is subtracted from the 'current' one, positive values mean that the current election was more fraudulent than its predecessor, while negative values indicate election fraud declined. Using this conceptualization of election fraud does result in some support for the deterrence thesis: an election that is markedly more fraudulent than the prior election does appear associated with more numerous and longer-lasting protest events at higher incumbent vote-shares. However, as in other examples, this effect only becomes statistically significant at the extremes. The values for change in fraud used in the marginal effects plots in Figures \@ref(fig:n-protests-ch) and \@ref(fig:d-protests-ch) represent very unusual changes, roughly between the 5th and 95th percentiles. In other words, the expected number and duration of protest events are only statistically significantly different when comparing elections that are far cleaner than the previous election and those that are far worse\textemdash a set of options that is unlikely to confront most incumbents.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(fraud.change = v2elirreg.inv - v2elirreg.inv.lag) #Positive values mean worsening fraud

model.n.base2.lag <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + fraud.change*winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub %>% filter(v2x_regime <= 2))
#summary(model.n.base2.lag)

model.d.base2.lag <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + fraud.change*winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag ,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub %>% filter(v2x_regime <= 2))
#summary(model.d.base2.lag)

model.size.base2.lag <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  fraud.change*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2),
                                family=truncated_nbinom2)
#summary(model.size.base2.lag)

```

```{r table-lagged-manip, results='asis', echo=FALSE}
models3 <- list()
models3[['Model 5']] <- model.n.base2.lag
models3[['Model 6']] <- model.d.base2.lag
models3[['Model 7']] <- model.size.base2.lag

t4a <- huxreg(models3, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Judicial ind. (lag)" = "jucon.1lag", "Liberal democracy index" = "libdem.1lag",
            "Legislative constraints (lag)" = "legcon.1lag", "Alternative info. (lag)" = "altinf.1lag",                             "GDP growth rate (lag)" = "gdpgro.1lag",                               "Civil society openness (lag)" = "osorg.1lag", "National party organization (lag)" = "natlpartyorg.lag", "Presidential" = "presidential", "Incumbent term-limited" = "inc.termlimit.num",
            "Physical integrity (lag)" = "physinteg.1lag", "Phys. int. (lag) squared" = "physintlag.sq", "Number of pre-election protests" = "n.events.pre", "Off-schedule election" = "unscheduled_election", "Regional protest diffusion" = "regional_success_lag"),
              statistics = character(0)) %>% set_caption("Table XX: Models of number, duration, and size of protests") %>% insert_row("", "Number", "Duration", "Size", after = 0)
theme_compact(t4a, header_rows = TRUE, header_cols = FALSE)

t4b <- huxreg(models3, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Change in election fraud" = "fraud.change", "Winner vote-share" = "winner.share.upd", 
            "Urban" = "urban.pct", "Rentier state" = "rentierstate", "Tax state" = "taxstate", "Change in fraud : Winner vote-share" = "fraud.change:winner.share.upd"),
              statistics = character(0)) %>% set_caption("Table XX: Models of number, duration, and size of protests") %>% insert_row("", "Number", "Duration", "Size", after = 0)
theme_compact(t4b, header_rows = TRUE, header_cols = FALSE)

```

```{r n-protests-ch, fig.cap="Change in fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE}
dat <- ggpredict(model.n.base2.lag, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "fraud.change [-1.12, 1.12]"), type = "count") 
p.n.margin.ch <- plot(dat)
p.n.margin.ch <- p.n.margin.ch +  labs(x = "Incumbent vote-share", y = "Number of events", title = "", color = "Change in fraud") + theme_bw()
p.n.margin.ch
```

```{r d-protests-ch, fig.cap="Change in fraud, incumbent vote-share, and duration of protest events", echo=FALSE, message=FALSE}
dat <- ggpredict(model.d.base2.lag, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "fraud.change [-1.22, 1.22]"), type = "count") 
p.d.margin.ch <- plot(dat)
p.d.margin.ch <- p.d.margin.ch +  labs(x = "Incumbent vote-share", y = "Duration of protest wave", title = "", color = "Change in fraud") + theme_bw()
p.d.margin.ch + coord_cartesian(ylim=c(0, 100))
```

## Chosen successors

```{r}
nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Executive eligible"))
nelda2.sub <- nelda2.sub %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))

model.n.base2.succ <- glmmTMB::glmmTMB(n.events.post ~ incumb.status + v2elirreg.inv + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd   +    physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag ,
                                         zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
summary(model.n.base2.succ)

model.n.base2.succ.bin <- glm(no_protest ~ incumb.status + v2elirreg.inv + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd   +    physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                              data = nelda2.sub  %>% filter(inc.lose == "0" & v2x_regime <= 2))
summary(model.n.base2.succ.bin)

plot_binary_margin_succ_n <- sjPlot::plot_model(model.n.base2.succ.bin, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]"))
plot_binary_margin_succ_n <- plot_binary_margin_succ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw()

####

model.d.base2.succ <- glmmTMB::glmmTMB(d.events.num ~ incumb.status + v2elirreg.inv + jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag  + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv+  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + inc.lose ,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub)
#summary(model.d.base2.succ)

model.size.base2.lmer <- lmer(log(max_size) ~  incumb.status + jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + as.factor(inc.lose) + ( 1 | stateid), data = nelda2.sub %>% filter(max_size > 0))
#summary(model.size.base2.lmer)

models4 <- list()
models4[['Model 5']] <- model.n.base2.hurd.mm
models4[['Model 6']] <- model.d.base2.hurd
models4[['Model 7']] <- model.size.base2.lmer

t5a <- huxreg(models4, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Table XX: Models of number, duration, and size of protests") %>% insert_row("", "Number", "Duration", "Size", after = 0)
theme_compact(t5a, header_rows = TRUE, header_cols = FALSE)

```

## Including liberal democracies

```{r}

model.n.base2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ v2elirreg.inv+ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag +  pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election ,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub)
#summary(model.n.base2.hurd.mm)

model.d.base2.hurd <- glmmTMB::glmmTMB(d.events.num ~  v2elirreg.inv + jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag  + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + inc.lose ,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub)
#summary(model.d.base2.hurd)

model.size.base2.lmer <- lmer(log(max_size) ~  v2elirreg.inv+ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + as.factor(inc.lose) + ( 1 | stateid), data = nelda2.sub %>% filter(max_size > 0))
#summary(model.size.base2.lmer)

models5 <- list()
models5[['Model 5']] <- model.n.base2.hurd.mm
models5[['Model 6']] <- model.d.base2.hurd
models5[['Model 7']] <- model.size.base2.lmer

t6a <- huxreg(models5, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Table XX: Models of number, duration, and size of protests") %>% insert_row("", "Number", "Duration", "Size", after = 0)
theme_compact(t6a, header_rows = TRUE, header_cols = FALSE)
```

# Alternative measures of election manipulation

To ensure that the results presented in the main text are not biased by the chosen measure of election fraud, this section replaces that measure with a broader indicator of election integrity from V-Dem. This variable is `v2xel_frefair,' which is a comprehensive indicator of quality of the electoral process, including measures of electoral management body autonomy and capacity, vote-buying, intimidation, fraud, and an overall assessment of the fairness of the election. This variable ranges from 0 to 1, with higher values indicating freer and fairer elections. The models presented below utilize this variable in place of the main measure of election fraud, while retaining all other control and explanatory variables, using the Mass Mobilization data on protest.

```{r manipdv-full, echo=FALSE}
m1 <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate + n.events.pre +  physinteg.1lag + physintlag.sq    + unscheduled_election  + (1 | stateid), data = nelda2.sub)

#summary(m1)

models <- list()
models[['Model 10']] <- m1

modelsummary(models, stars = T,  title = 'Random-intercept multilevel model of overall election integrity',
             coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests"))
```

The first model shows that the number of pre-election protests is weakly associated with improved election integrity, but just as in the main text this effect is extremely small. Each additional protest is associated with an improvement in election integrity equivalent to 0.2\% of the standard deviation of the dependent variable. Given that the mean number of pre-election protests is six, this is an extremely small influence on the overall degree of election integrity.

```{r models-hurdle-fulldv, warning = FALSE, message = FALSE, include = F}
model.n.base2.elfr <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election +  inc.lose,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub)
summary(model.n.base2.elfr)

model.size.base2.elfr <- lmer(log(max_size) ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election +  as.factor(inc.lose) + (1 | stateid), data = nelda2.sub %>% filter(max_size > 0))
summary(model.size.base2.elfr)

nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
km <-  with(nelda2.sub, Surv(d.events.num, surv_status))

cox2.elfr <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election +  as.factor(inc.lose), data = nelda2.sub %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
summary(cox2.elfr)

```

Moving to the hurdle models of post-election protest, we do not observe much of a relationship between overall election integrity and protest. There is no statistically significant relationship between election integrity and protest occurrence in Model 2. Neither is there any significant relationship in interaction with the winner's vote-share, as shown in Figure \@ref(fig:b-protests-margin), or with economic crisis, as in Figure \@ref(fig:b-protests-crisis). The coefficient on the interaction term for public sector corruption and overall election integrity is statistically significant, though there is no significant effect at typical values for covariates and election integrity (as shown in Figure \@ref(fig:b-protests-corrupt)).

In the corresponding results for the number of post-election protests, there is similarly muted evidence for a connection between the overall level of manipulation and protest risk to incumbents. There is no significant difference between the marginal effect of common levels of election integrity on the number of post-election protests for typical cases across any margin of victory (Figure \@ref(fig:n-protests-margin)). During an economic crisis, *cleaner* elections are associated with an increase in the number of protests (\@ref(fig:n-protests-crisis)). This is plausibly due to the crisis weakening both the government's ability to generate manipulation [@greene_why_2007] and its ability to co-opt or put down the protests. Finally, public sector corruption is associated with a significant increase in the marginal effect of poor-integrity elections on protest\textemdash but only when the corruption measure is at its maximum (Figure \@ref(fig:n-protests-corrupt)).

As a whole, the results are in accordance with the theory proposed in the main paper. There is no consistent evidence that lower-quality elections are associated with either a higher risk of protest initiation, or with a larger number of post-election protests.

```{r zero-model-table-fulldv, echo=FALSE, results='asis'}

models <- list()
#models[['Model XXa']] <- model.n.base2.elfr
models[['Model XXb']] <- model.size.base2.elfr
models[['Model XXc']] <- cox2.elfr

modelsummary::modelsummary(models, coef_omit = "(Intercept)", longtable=T, stars = T, title = 'Hurdle models of post-election protest (counts)',
                           coef_rename = 
               c("count_(Intercept)" = "Intercept", "count_jucon.1lag" = "Judical independence (lag)", "count_legcon.1lag" = "Leg. constraints (lag)", 
                 "count_gdpgro.1lag" = "GDP growth rate (lag)", "count_altinf.1lag" = "Alternative info. (lag)", 
                 "count_osorg.1lag" = "Civil soc. openness (lag)", "count_econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "count_natlpartyorg.lag" = "National party organization (lag)", "count_presidential1" = "Presidential election",
                 "count_inc.termlimit.num" = "Incumbent term-limited",  
                 "count_urban.pct" = "Urbanization", "count_rentierstate" = "Rentier state", "count_taxstate" = "Tax state",
                 "count_n.events.massmob.pre" = "Number of pre-election protests", "count_pubseccorrup.lag" = "Public sector corruption (lag)", "count_v2xel_frefair" = "Election integrity",
                 "count_physinteg.1lag" = "Physical integrity (lag)", "count_physintlag.sq" = "Physical integrity (lag), squared", "count_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud",  "winner.share.upd" = "Incumbent vote-share",  "v2xel_frefair" = "Election integrity"))

```

\newpage

```{r zero-model-table-count-fulldv, echo=FALSE}

t_count_general <- huxreg(model.n.base2.elfr, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, omit_coefs = c("rentierstate1", "taxstate1", "n.events.pre", "unscheduled_election1", "inc.lose1", "v2elirreg.inv:winner.share.upd", "gdpgro.1lag:v2elirreg.inv"), statistics = character(0)) %>% set_caption("Table 3: Zero-inflated models of protest initiation")

theme_compact(t_count_general, header_rows = TRUE, header_cols = FALSE)

```

```{r, echo=FALSE, message=FALSE}
dat <- ggpredict(model.n.base2.elfr, terms = c("winner.share.upd", "v2xel_frefair [-.41, 1.68]")) 
p.n.margin <- plot(dat)
p.n.margin <- p.n.margin + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events")  + theme_bw()

dat <- ggpredict(model.size.base2.elfr, terms = c("winner.share.upd", "v2xel_frefair [-.41, 1.68]")) 
p.size.margin <- plot(dat)
p.size.margin <- p.size.margin + labs(x = "Incumbent vote-share", y = "Predicted size of protests" , title = "Predicted maximum of protest events")  + theme_bw()

```

```{r n-protests-duration-sim, echo=FALSE, message=FALSE}
Sim1 <- coxsimInteract(cox2.elfr, b2 = "v2xel_frefair", b1 = "winner.share.upd", qi = "Hazard Ratio", X2 = c(-.4, 1.7),  X1 = seq(0.01, 1, by = .01), nsim = 2000)
p.duration.margin <- simGG(Sim1)
p.duration.margin <- p.duration.margin + labs(x = "Incumbent vote-share", y = "Hazard ratio" , title = "Hazard ratios for protest termination")  + theme_bw()
```

```{r n-protests-margin-fulldv, fig.cap="Election integrity, incumbent vote-share, and number of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
p.n.margin

```

```{r n-protests-size-fulldv, fig.cap="Election integrity, incumbent vote-share, and size of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
p.size.margin

```

```{r n-protests-duration-fulldv, fig.cap="Election integrity, incumbent vote-share, and duration of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
p.duration.margin #Hazard ratios > 1 indicate an increased risk of termination of protest wave. So worse elections are associated with quicker demise of protest waves among cases where protest occurred and incumbents were not replaced.
```

### 2023-06-15 Output/appendix_v4.Rmd
Updates to appendix: liberal democracies, chosen successors

@@
 
-```{r}
+The models in the main text control for the presence of term-limited executive incumbents, based on the logic in Hale [@hale_patronal_2014].
+However, it could be the case that the nature of a term-limited election affects the behavior of protesters and election manipulators alike; that is, perhaps when there is a designated successor, the heir apparent is better able to mobilize the resources of the state to manipulate elections and suppress protest.
+Using a variable from NELDA that codes for designated successors, the following models add a categorical variable for the incumbent's status, which can take on the following values: *executive eligible*, *chosen successor*, *contested executive election*, and *legislative election*.
+As Table \@ref(tab:successor-table) and Figure \@ref(fig:successor-fig) show, the results are largely similar to those in the main text.
+Higher levels of fraud does make the occurrence of protest more likely at low incumbent vote-shares, but this does not translate into a higher risk of more numerous or larger protest events.
+
+```{r successor-table, echo=FALSE, message=FALSE, warning=FALSE}
 nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Executive eligible"))
 nelda2.sub <- nelda2.sub %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))
+nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))
 
-model.n.base2.succ <- glmmTMB::glmmTMB(n.events.post ~ incumb.status + v2elirreg.inv + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd   +    physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag ,
+model.n.base2.succ <- glmmTMB::glmmTMB(n.events.post ~ incumb.status + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  +   physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                                          zi = ~.,
@@
                                           data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
-summary(model.n.base2.succ)
+#summary(model.n.base2.succ)
 
-model.n.base2.succ.bin <- glm(no_protest ~ incumb.status + v2elirreg.inv + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd   +    physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
+model.n.base2.succ.bin <- glm(no_protest ~ incumb.status + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  +   physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                               data = nelda2.sub  %>% filter(inc.lose == "0" & v2x_regime <= 2))
-summary(model.n.base2.succ.bin)
+#summary(model.n.base2.succ.bin)
 
@@
 
-model.d.base2.succ <- glmmTMB::glmmTMB(d.events.num ~ incumb.status + v2elirreg.inv + jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag  + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv+  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + inc.lose ,
+model.d.base2.succ <- glmmTMB::glmmTMB(d.events.num ~ incumb.status + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  +   physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                                           zi = ~.,
                                           family= nbinom2,
-                                          data = nelda2.sub)
+                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
 #summary(model.d.base2.succ)
 
-model.size.base2.lmer <- lmer(log(max_size) ~  incumb.status + jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + as.factor(inc.lose) + ( 1 | stateid), data = nelda2.sub %>% filter(max_size > 0))
-#summary(model.size.base2.lmer)
+model.size.base2 <- glmmTMB::glmmTMB(max_size ~ incumb.status + jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  v2elirreg.inv*winner.share.upd  +    physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+                                ziformula=~.,
+                                data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2),
+                                family=truncated_nbinom2)
+#summary(model.size.base2)
+
+model.size.base2.succ.bin <- glm(no_size ~ incumb.status + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  +   physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
+                              data = nelda2.sub  %>% filter(inc.lose == "0" & v2x_regime <= 2))
+#summary(model.size.base2.succ.bin)
+
+plot_binary_margin_succ_size <- sjPlot::plot_model(model.size.base2.succ.bin, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]"))
+plot_binary_margin_succ_size <- plot_binary_margin_succ_size +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw()
+
+### Table
 
 models4 <- list()
-models4[['Model 5']] <- model.n.base2.hurd.mm
-models4[['Model 6']] <- model.d.base2.hurd
-models4[['Model 7']] <- model.size.base2.lmer
+models4[['Model 5']] <- model.n.base2.succ
+#models4[['Model 6']] <- model.d.base2.hurd
+models4[['Model 6']] <- model.size.base2
 
-t5a <- huxreg(models4, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Table XX: Models of number, duration, and size of protests") %>% insert_row("", "Number", "Duration", "Size", after = 0)
+t5a <- huxreg(models4, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Models of number and size of protests") %>% insert_row("", "Number", "Size", after = 0)
 theme_compact(t5a, header_rows = TRUE, header_cols = FALSE)
@@
 
+```{r n-protests-margin, echo=FALSE, message=FALSE, warning=FALSE}
+dat <- ggpredict(model.n.base2.succ, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "count") 
+p.n.margin <- plot(dat)
+p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Number of events", title = "Predicted number of protests", color = "Election fraud") + theme_bw() + coord_cartesian(ylim = c(0, 40))
+```
+
+```{r size-protests-margin, echo=FALSE, message=FALSE, warning=FALSE}
+dat <- ggpredict(model.size.base2, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "count") 
+p.s.margin <- plot(dat)
+p.s.margin <- p.s.margin +  labs(x = "Incumbent vote-share", y = "Size of events", title = "Predicted size of protests", color = "Election fraud") + theme_bw() + coord_cartesian(ylim = c(0, 18000))
+```
+
+```{r successor-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
+plot.combined.prob <- cowplot::plot_grid(plot_binary_margin_succ_n, p.n.margin, p.s.margin, ncol=1)
+plot.combined.prob
+```
+
 ## Including liberal democracies
 
-```{r}
+The main text excludes liberal democracies from the analysis, on the grounds that the prior literature on protest and fraud typically confines itself to regimes that hold multiparty elections, and the liberal-democratic cases exhibit little variation in their (low) levels of election fraud.
+But perhaps it is in liberal democracies where the deterrent logic operates most effectively, settling into an equilibrium of high protest risk and low fraud.
+Including liberal democracies in the analysis results in little difference from the findings in the main text, however.
+Election fraud does not appear associated with a higher number of protest events, or larger protest sizes.
 
-model.n.base2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ v2elirreg.inv+ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag +  pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election ,
+```{r lib-table, echo=FALSE}
+
+model.n.base2.lib <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                           zi = ~.,
@@
                                           data = nelda2.sub)
-#summary(model.n.base2.hurd.mm)
+#summary(model.n.base2.lib)
 
-model.d.base2.hurd <- glmmTMB::glmmTMB(d.events.num ~  v2elirreg.inv + jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag  + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + inc.lose ,
+model.size.base2.lib <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                           zi = ~.,
@@
                                           data = nelda2.sub)
-#summary(model.d.base2.hurd)
-
-model.size.base2.lmer <- lmer(log(max_size) ~  v2elirreg.inv+ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + as.factor(inc.lose) + ( 1 | stateid), data = nelda2.sub %>% filter(max_size > 0))
-#summary(model.size.base2.lmer)
+#summary(model.size.base2.lib)
 
 models5 <- list()
-models5[['Model 5']] <- model.n.base2.hurd.mm
-models5[['Model 6']] <- model.d.base2.hurd
-models5[['Model 7']] <- model.size.base2.lmer
+models5[['Model 5']] <- model.n.base2.lib
+models5[['Model 7']] <- model.size.base2.lib
 
-t6a <- huxreg(models5, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Table XX: Models of number, duration, and size of protests") %>% insert_row("", "Number", "Duration", "Size", after = 0)
+t6a <- huxreg(models5, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Table XX: Models of number and size of protests, inclu.
+liberal democracies") %>% insert_row("", "Number", "Size", after = 0)
 theme_compact(t6a, header_rows = TRUE, header_cols = FALSE)

### 2023-07-01 Output/Submission files/manuscript_anon_v2.Rmd
Copyedits to appendix (e.g. shortening tables)

@@
 
-model.size.base2 <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv +  +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+model.size.base2 <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + gdpgro.1lag  +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                 ziformula=~.,

### 2023-07-01 Output/appendix_v4.Rmd
Copyedits to appendix (e.g. shortening tables)

@@
 
-## Repression and pre-election protest
+# Repression and pre-election protest
 
@@
 If protest risk deters manipulation in this way, there should be a negative and significant effect for the number of pre-election protests in highly repressive states.
+
 Figure \@ref(fig:prot-rep-fraud) shows the absence of such an effect.
@@
 
-## Prior manipulation
+# Prior manipulation
 
@@
 
-```{r table-lagged-manip, results='asis', echo=FALSE}
+```{r table-lagged-manip, results='asis', echo=FALSE, warning=FALSE}
 models3 <- list()
-models3[['Model 5']] <- model.n.base2.lag
-models3[['Model 6']] <- model.d.base2.lag
-models3[['Model 7']] <- model.size.base2.lag
+models3[['Model 4']] <- model.n.base2.lag
+models3[['Model 5']] <- model.d.base2.lag
+models3[['Model 6']] <- model.size.base2.lag
 
-t4a <- huxreg(models3, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Judicial ind.
-(lag)" = "jucon.1lag", "Liberal democracy index" = "libdem.1lag",
-            "Legislative constraints (lag)" = "legcon.1lag", "Alternative info.
-(lag)" = "altinf.1lag",                             "GDP growth rate (lag)" = "gdpgro.1lag",                               "Civil society openness (lag)" = "osorg.1lag", "National party organization (lag)" = "natlpartyorg.lag", "Presidential" = "presidential", "Incumbent term-limited" = "inc.termlimit.num",
-            "Physical integrity (lag)" = "physinteg.1lag", "Phys.
-int.
-(lag) squared" = "physintlag.sq", "Number of pre-election protests" = "n.events.pre", "Off-schedule election" = "unscheduled_election", "Regional protest diffusion" = "regional_success_lag"),
-              statistics = character(0)) %>% set_caption("Table XX: Models of number, duration, and size of protests") %>% insert_row("", "Number", "Duration", "Size", after = 0)
-theme_compact(t4a, header_rows = TRUE, header_cols = FALSE)
-
-t4b <- huxreg(models3, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Change in election fraud" = "fraud.change", "Winner vote-share" = "winner.share.upd",
-            "Urban" = "urban.pct", "Rentier state" = "rentierstate", "Tax state" = "taxstate", "Change in fraud : Winner vote-share" = "fraud.change:winner.share.upd"),
-              statistics = character(0)) %>% set_caption("Table XX: Models of number, duration, and size of protests") %>% insert_row("", "Number", "Duration", "Size", after = 0)
-theme_compact(t4b, header_rows = TRUE, header_cols = FALSE)
+t4a <- huxreg(models3, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Fraud change" = "fraud.change", "Incumbent vote-share" = "winner.share.upd", "Fraud change:Incumbent vote-share" = "fraud.change:winner.share.upd"),
+              statistics = character(0)) %>% set_caption("Models of number, duration, and size of protests") %>% insert_row("", "Number", "Duration", "Size", after = 0) %>% insert_row("Controls", "Yes", "Yes", "Yes", after = 10)
+theme_compact(t4a, header_rows = TRUE, header_cols = FALSE)  %>% set_bottom_border(10, everywhere)
 
@@
 
-## Chosen successors
+# Chosen successors
 
@@
 Using a variable from NELDA that codes for designated successors, the following models add a categorical variable for the incumbent's status, which can take on the following values: *executive eligible*, *chosen successor*, *contested executive election*, and *legislative election*.
-As Table \@ref(tab:successor-table) and Figure \@ref(fig:successor-fig) show, the results are largely similar to those in the main text.
+As Figure \@ref(fig:successor-fig) show, the results are largely similar to those in the main text.
 Higher levels of fraud does make the occurrence of protest more likely at low incumbent vote-shares, but this does not translate into a higher risk of more numerous or larger protest events.
@@
 models4 <- list()
-models4[['Model 5']] <- model.n.base2.succ
+models4[['Model 7']] <- model.n.base2.succ
 #models4[['Model 6']] <- model.d.base2.hurd
-models4[['Model 6']] <- model.size.base2
+models4[['Model 8']] <- model.size.base2
 
 t5a <- huxreg(models4, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Models of number and size of protests") %>% insert_row("", "Number", "Size", after = 0)
-theme_compact(t5a, header_rows = TRUE, header_cols = FALSE)
+#theme_compact(t5a, header_rows = TRUE, header_cols = FALSE)
 
@@
 
-## Including liberal democracies
+# Including liberal democracies
 
@@
 
-```{r lib-table, echo=FALSE}
+```{r lib-table, echo=FALSE, warning=FALSE}
 
@@
 
-t6a <- huxreg(models5, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Table XX: Models of number and size of protests, inclu.
+t6a <- huxreg(models5, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd"), statistics = character(0)) %>% set_caption("Models of number and size of protests, inclu.
 liberal democracies") %>% insert_row("", "Number", "Size", after = 0)
-theme_compact(t6a, header_rows = TRUE, header_cols = FALSE)
+theme_compact(t6a, header_rows = TRUE, header_cols = FALSE)  %>% insert_row("Controls", "Yes", "Yes", after = 10) %>% set_bottom_border(10, everywhere)
 ```
@@
 models <- list()
-models[['Model 10']] <- m1
+models[['Model 9']] <- m1
 
-modelsummary(models, stars = T,  title = 'Random-intercept multilevel model of overall election integrity',
+modelsummary(models, stars = T,  title = 'Random-intercept multilevel model of overall election integrity', gof_omit = "AIC|BIC",
              coef_rename = 
                c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
-                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", "presidential" = "Presidential", "n.events.pre" = "N. of pre-election protest",
+                 "physinteg.1lag" = "Physical integrity (lag)",
+                 "physintlag.sq" = "Phys. integ. (lag), squared", "unscheduled_election" = "Off-schedule election",
                  "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
@@
 ```{r models-hurdle-fulldv, warning = FALSE, message = FALSE, include = F}
-model.n.base2.elfr <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election +  inc.lose,
-                                          zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub)
+model.n.base2.elfr <- glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag +  + osorg.1lag  + natlpartyorg.lag +  v2xel_frefair*winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+                               ziformula=~.,
+                              data = nelda2.sub %>% filter(inc.lose == "0"),
+                                family=truncated_nbinom2)
 summary(model.n.base2.elfr)
@@
 
-nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
-km <-  with(nelda2.sub, Surv(d.events.num, surv_status))
+model.n.griev.elfr <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2xel_frefair +  v2xel_frefair*gdpgro.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+                                ziformula=~.,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
+                                family=truncated_nbinom2)
 
-cox2.elfr <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election +  as.factor(inc.lose), data = nelda2.sub %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
-summary(cox2.elfr)
+model.size.base2.elfr <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2xel_frefair + gdpgro.1lag  +   winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+                                ziformula=~.,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
+                                family=truncated_nbinom2)
+
+model.size.griev.elfr <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2xel_frefair +  v2xel_frefair*gdpgro.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+                                ziformula=~.,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
+                                family=truncated_nbinom2)
 
@@
 There is no statistically significant relationship between election integrity and protest occurrence in Model 2.
-Neither is there any significant relationship in interaction with the winner's vote-share, as shown in Figure \@ref(fig:b-protests-margin), or with economic crisis, as in Figure \@ref(fig:b-protests-crisis).
-The coefficient on the interaction term for public sector corruption and overall election integrity is statistically significant, though there is no significant effect at typical values for covariates and election integrity (as shown in Figure \@ref(fig:b-protests-corrupt)).
+Neither is there any significant relationship in interaction with the winner's vote-share, as shown in Figure \@ref(fig:b-protests-margin).
 
@@
 There is no significant difference between the marginal effect of common levels of election integrity on the number of post-election protests for typical cases across any margin of victory (Figure \@ref(fig:n-protests-margin)).
-During an economic crisis, *cleaner* elections are associated with an increase in the number of protests (\@ref(fig:n-protests-crisis)).
-This is plausibly due to the crisis weakening both the government's ability to generate manipulation [@greene_why_2007] and its ability to co-opt or put down the protests.
-Finally, public sector corruption is associated with a significant increase in the marginal effect of poor-integrity elections on protest\textemdash but only when the corruption measure is at its maximum (Figure \@ref(fig:n-protests-corrupt)).
-
 As a whole, the results are in accordance with the theory proposed in the main paper.
@@
 
-```{r zero-model-table-fulldv, echo=FALSE, results='asis'}
+```{r zero-model-table-fulldv, echo=FALSE, results='asis', warning=FALSE}
 
@@
 #models[['Model XXa']] <- model.n.base2.elfr
-models[['Model XXb']] <- model.size.base2.elfr
-models[['Model XXc']] <- cox2.elfr
+models[['Model 10']] <- model.n.base2.elfr
+models[['Model 11']] <- model.n.griev.elfr
+models[['Model 12']] <- model.size.base2.elfr
+models[['Model 13']] <- model.size.griev.elfr
 
-modelsummary::modelsummary(models, coef_omit = "(Intercept)", longtable=T, stars = T, title = 'Hurdle models of post-election protest (counts)',
-                           coef_rename = 
-               c("count_(Intercept)" = "Intercept", "count_jucon.1lag" = "Judical independence (lag)", "count_legcon.1lag" = "Leg. constraints (lag)", 
-                 "count_gdpgro.1lag" = "GDP growth rate (lag)", "count_altinf.1lag" = "Alternative info. (lag)", 
-                 "count_osorg.1lag" = "Civil soc. openness (lag)", "count_econ.crisis.nelda.1lag1" = "Economic crisis", 
-                 "count_natlpartyorg.lag" = "National party organization (lag)", "count_presidential1" = "Presidential election",
-                 "count_inc.termlimit.num" = "Incumbent term-limited",  
-                 "count_urban.pct" = "Urbanization", "count_rentierstate" = "Rentier state", "count_taxstate" = "Tax state",
-                 "count_n.events.massmob.pre" = "Number of pre-election protests", "count_pubseccorrup.lag" = "Public sector corruption (lag)", "count_v2xel_frefair" = "Election integrity",
-                 "count_physinteg.1lag" = "Physical integrity (lag)", "count_physintlag.sq" = "Physical integrity (lag), squared", "count_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud",  "winner.share.upd" = "Incumbent vote-share",  "v2xel_frefair" = "Election integrity"))
+t7 <-  huxreg(models, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,
+ statistics = character(0), coefs = c("Intercept" = "(Intercept)", "Electoral integrity" = "v2xel_frefair", "Incumbent vote-share" = "winner.share.upd", "GDP growth (lag)" = "gdpgro.1lag", "Electoral integrity:Incumbent vote-share" = "v2xel_frefair:winner.share.upd",  "Election integrity : GDP growth (lag)" = "gdpgro.1lag:v2xel_frefair")) %>% set_caption("Conditional hurlde models of protest number and size") %>% insert_row("", "Number", "", "Size", "", after = 0)
+theme_compact(t7, header_rows = TRUE, header_cols = FALSE) %>% insert_row("Controls", "Yes", "Yes", "Yes", "Yes", after = 14) %>% set_bottom_border(14, everywhere)
 
@@
 
-```{r n-protests-duration-sim, echo=FALSE, message=FALSE}
-Sim1 <- coxsimInteract(cox2.elfr, b2 = "v2xel_frefair", b1 = "winner.share.upd", qi = "Hazard Ratio", X2 = c(-.4, 1.7),  X1 = seq(0.01, 1, by = .01), nsim = 2000)
-p.duration.margin <- simGG(Sim1)
-p.duration.margin <- p.duration.margin + labs(x = "Incumbent vote-share", y = "Hazard ratio" , title = "Hazard ratios for protest termination")  + theme_bw()
-```
-
 ```{r n-protests-margin-fulldv, fig.cap="Election integrity, incumbent vote-share, and number of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
@@
 ```
-
-```{r n-protests-duration-fulldv, fig.cap="Election integrity, incumbent vote-share, and duration of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
-p.duration.margin #Hazard ratios > 1 indicate an increased risk of termination of protest wave.
-So worse elections are associated with quicker demise of protest waves among cases where protest occurred and incumbents were not replaced.
-```

### 2023-07-01 Scripts/data_cleaning_massmob.Rmd
Copyedits to appendix (e.g. shortening tables)

@@
 nelda <- nelda %>% dplyr::rename(first.mp.election = nelda2)
+nelda <- nelda %>% rename(unscheduled_election = nelda6)
 nelda <- nelda %>% rename(protests = nelda9)
@@
 nelda <- nelda %>% mutate(opp_leaders_prevented = ifelse(nelda3 == "No", "In general", ifelse(opp_leaders_prevented == "yes", "Specifically", ifelse(opp_leaders_prevented == "no", "No", NA))))
+
+nelda <- nelda %>% mutate(unscheduled_election = ifelse(is.na(unscheduled_election)==T, NA, ifelse(unscheduled_election == "yes" | unscheduled_election == "N/A", 1, 0)))   #"N/A" here refers to no common expectation for election timing
 
@@
 
-nelda_massmob <- nelda_massmob %>% filter(is.na(protest_mdy) == T | protest_mdy < election_mdy | (protest_mdy >= election_mdy & str_detect(notes_massmob, regex(pattern = "election*|vot*|ballot*", ignore_case = T))))
+#nelda_massmob <- nelda_massmob %>% filter(is.na(protest_mdy) == T | protest_mdy < election_mdy | (protest_mdy >= election_mdy & str_detect(notes_massmob, regex(pattern = "election*|vot*|ballot*", ignore_case = T)))) #Turn this on to focus only on protests related to the election
 ```
@@
 ```{r}
-write.csv(nelda_complete, here::here("Data", "nelda_complete.csv"))
+write.csv(nelda_complete, here::here("Data", "nelda_massmob_all.csv")) #File that only looks at election-related massmob was called nelda_complete
 ```

### 2023-07-10 Output/Submission files/manuscript_anon_v2.Rmd
Edits to election fraud choice and repression

@@
 
+Here, this logic will be tested using two measures, both of which may indicate the strength of a protest movement, as well as the cost of protest to incumbent governments: the total number of protest events and the estimated number of protesters on the largest single day of protests.
+I assume that more numerous protests, and those that built to a higher peak participation, are costlier for governments to contain, all else held equal.
+If fraudulent elections reveal information about regime weakness or generate large grievance-motivated protest, more fraudulent elections should be associated with larger and more numerous protest waves.
+Prior theory expects these effects to be largest when margins of victory are smaller, or when economic grievances are high.
+
 In summary, fraudulent elections are not likely to pose a significant protest risk to incumbents.
@@
 Protesters, bystanders, and elites can update their estimation of the relative strength of the protest coalition and the incumbent government by observing the size of crowds, the number of protests, and the government's reaction.
-
-The number and size of protests will instead be related to other indicators of opposition and incumbent capacity.
-When opposition groups and civil society capacity are more organized and better resourced, it is likely that more protesters will take to the streets in more numerous events (see Chenoweth et al [-@chenowethStateRepressionNonviolent2017] for a review of factors making protest movements more resilient to repression).
+Thus, I expect that Hypothesis 5\textemdash which might be assumed under other manipulation-oriented models of protest\textemdash to *not* be supported.
+
+>H5: The number and maximum size of protest events are likely to increase as the severity of election fraud increases.
+
+Once protests begin, the number and size of protest events is likely to be influenced by the organizational capacity of civil society and the opposition on the one hand, and on the ability of regime elites to remain united on the other.
+When opposition groups and civil society capacity are more organized and better resourced, it is likely that more protesters will take to the streets in more numerous events (see Chenoweth et al [-@chenowethStateRepressionNonviolent2017] for a review of factors making protest movements more resilient).
 By contrast, when the government appears well-resourced and resolved, more 'passionate economists' should stay home.
-As identified in prior research, informative factors should be the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], the margin of victory in the election [@simpser2013], and whether or not the chief executive faces term limits [@hale_patronal_2014].
-Generally speaking, these structural variables indicate aspects of the political opportunity structure that make protest more likely to occur [@meyer_protest_2004].
-
-Here, this logic will be tested using two measures, both of which may indicate the strength of a protest movement, as well as the cost of protest to incumbent governments: the total number of protest events and the estimated number of protesters on the largest single day of protests.
-I assume that more numerous protests, and those that built to a higher peak participation, are costlier for governments to contain, all else held equal.
-If fraudulent elections reveal information about regime weakness or generate large grievance-motivated protest, more fraudulent elections should be associated with larger and more numerous protest waves.
-Prior theory expects these effects to be largest when margins of victory are smaller, or when economic grievances are high.
-
-However, the theory articulated above does *not* expect that the severity of fraud will be associated with the course of protest.
-Instead, it predicts that underlying indicators of opposition and ruling-party strength will influence the direction of protest movements.
-In particular, it expects the size and number of protest events will be sensitive to the repressive capacity of the state and to indicators of opposition strength.
-that high and low levels of repression will be associated with smaller protest movements (in number and size).
-
->H5: The relationship between repressive capacity and the number of protest events and the maximum event size is curvilinear, with number and size likely to to first increase and then fall as repressive capacity grows.
-
->H6: The number of protest events and the maximum event size are likely to increase with indicators of opposition and civil-society strength.
+Much of this balance is mediated by political and social elites' decision(s) to remain supportive of the incumbent; protests can create division among the incumbent elite (e.g.
+among hardliners and softliners [@schmitter1986transitions]) and incentives for disaffected elites to move against the incumbent leadership [@kimPathways2019].
+In turn, the opportunity created by a divided elite\textemdash and their resources [@robertson_protest_2010]\textemdash can keep protest movements alive.
+
+Since information on elite and mass preferences can be difficult to observe outside liberal democracies, in part due to preference falsification [@Kuran1991], it is difficult to measure the risk of elite splits *a priori*.
+However, Hale [-@hale_patronal_2014] makes the case that elites become incentivized to turn against incumbent governments when they expect the chief executive's time in office has become limited.
+While the incumbent's age and popularity may be signals of such, a constitutional term limit can also serve this role.
+As such, Hypothesis 6 predicts that protest number and size will increase for elections where the incumbent faces a term limit.
+This hypothesis serves as useful counter-point to the traditional view on election manipulation.
+Unlike election fraud (as presented here), a constitutional term limit creates a clear, unambiguous signal that enables coordination among protesters and elites that can influence the course of protest.
+
+>H6: The number of protest events and the maximum event size are likely to increase following elections where the chief executive is.
 
@@
 This measure of election integrity should thus serve as a strong test of the protest-oriented model.
-The variable is used as a dependent variable in the first set of models that test H1, and as a potential explanatory variable in the subsequent models of protest number, size, and duration.
+It also matches the conception of election manipulation used in much of the relevant formal literature.^[For example, election manipulation is treated as post-hoc adjustment of the results in Little et al [-@Little2015], Little [-@little_elections_2012], Magaloni [-@magaloni2010], and Fearon [-@fearon_self-enforcing_2011], among others.
+Egorov and Sonin [-@Egorov2021] is an exception, which focuses on pre-election restrictions.] The variable is used as a dependent variable in the first set of models that test H1, and as a potential explanatory variable in the subsequent models of protest number, size, and duration.
+As an additional check in the appendix, a fuller measure of overall election integrity is used instead.
 
@@
 First, the *number of pre-election protests* in ECAV is taken as *prima facie* evidence of a more turbulent and risky election environment\textemdash it indicates greater mobilization capacity in civil society and/or the opposition, and may also indicate an elevated sense of grievance.
-Second, the repressive capacity of the regime is modeled using the lagged *physical integrity index* for the country-year.
-In one approach, higher values represent more permissive environments, and should be associated with greater protest risk (and thus reduced fraud).
-Alternatively, in keeping with the 'murder in the middle' framework [@feinLifeIntegrityViolationsDemocracy1995], the highest protest risks (and most reduced fraud) might be found at middle values of repression; regimes with the most demonstrated repressive capacity deter potential challengers from taking to the streets in the first place, and the least repressive states defuse protest through accommodation [@pierskallaProtestDeterrenceEscalation2010;@mooreRepressionDissentSubstitution2000].
-I measure repressive capacity using the *physical integrity* index from V-Dem, lagged one year.
-Higher values for this variable indicate greater freedom from torture and political killing.
-I also include the variable's square term to account for the 'murder in the middle' framework.
-The physical integrity variables are expected to be explanatory in models of protest, but are important control variables in the models of election fraud.
-Since repressive capacity helps influence protest in expectation [@Ritter2016], incumbents that are more confident of deterring protest in advance should be more willing to engage in fraud under the protest-deterrent model of manipulation.
 
@@
 
+The number and size of protests will instead be related to other indicators of opposition and incumbent capacity.
+As identified in prior research, informative factors should be the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], the margin of victory in the election [@simpser2013], and whether or not the chief executive faces term limits [@hale_patronal_2014].
+Generally speaking, these structural variables indicate aspects of the political opportunity structure that make protest more likely to occur [@meyer_protest_2004].
+
 Several control variables are included, which could be common causes for both the severity of election manipulation and protest.
@@
 
-Finally, several binary control variables are included.
+Several binary control variables are included.
 First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest.
@@
 Lastly, since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.
+
+Lastly, the repressive capacity of the state deserves extended mention, since a state with greater capacity for repression may be better able to deter protest and thus 'get away with' more manipulation [@Egorov2021,@Ritter2016].
+Omitting a control for repressive capacity would likely produce a biased finding, in which high fraud was associated with low levels of protest as a result.
+I measure repressive capacity using the *physical integrity* index from V-Dem, lagged one year.
+Higher values for this variable indicate greater freedom from torture and political killing.
+In one approach, higher values represent more permissive environments, and should be associated with greater protest risk (and thus reduced fraud).
+Alternatively, in keeping with the 'murder in the middle' framework [@feinLifeIntegrityViolationsDemocracy1995], the highest protest risks (and most reduced fraud) might be found at middle values of repression; regimes with the most demonstrated repressive capacity deter potential challengers from taking to the streets in the first place, and the least repressive states defuse protest through accommodation [@pierskallaProtestDeterrenceEscalation2010;@mooreRepressionDissentSubstitution2000].
+I thus include the variable's square term to account for the 'murder in the middle' framework.
+
 Table \@ref(tab:summary-table) shows summary statistics for all of these variables.

### 2023-07-11 Output/Submission files/manuscript_anon_v2.Rmd
Continuing efforts to address R comments

@@
 library(kableExtra)
-datasummary_skim(incl.vars, output = "kableExtra", title = "Table 1: Summary statistics")  %>%
+datasummary_skim(incl.vars, output = "kableExtra", title = "Table 1: Summary statistics and variable distributions ")  %>%
     kable_styling(font_size = 9, full_width = FALSE)

### 2023-07-11 Output/appendix_v4.Rmd
Continuing efforts to address R comments

@@
 
-```{r plot-margin-mm, echo=FALSE, message=FALSE, warning=F, fig.cap="Fraud, winning vote-share, and probability of protest initiation (Mass Mobilization data)", fig.height=7}
+```{r plot-margin-mm, echo=FALSE, message=FALSE, warning=F, fig.cap="Fraud, winning vote-share, and probability of protest initiation (Mass Mobilization data)", out.width = '85%', fig.pos="h!"}
 nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(is.na(n.events.massmob.post)==T, NA, ifelse(n.events.massmob.post > 0, 1, 0)))
@@
 
-```{r prot-rep-fraud, message=FALSE, warning=F, echo=FALSE, fig.cap="Marginal effect of pre-election protest on fraud, conditional on physical integrity"}
+```{r prot-rep-fraud, message=FALSE, warning=F, echo=FALSE, fig.cap="Marginal effect of pre-election protest on fraud, conditional on physical integrity", fig.pos="h!"}
 m1_rep <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre*physinteg.1lag  +  unscheduled_election  + (1 | stateid), data = nelda2.sub %>% filter(v2x_regime <= 2))
@@
 
-# Prior manipulation
+# Prior manipulation and expecations
 
 Another explanation for post-election protest could hinge on changes over time; elections that are much more fraudulent than prior elections may exacerbate citizens' grievances or convince opposition groups that the incumbent is unexpectedly weak.
-This possibility is tested by replacing the *election fraud* variable with *fraud change*, the difference between the current election and the one immediately prior.
+This possibility is tested in two ways.
+First, the *election fraud* variable is replaced with *fraud change*, the difference between the current election and the one immediately prior.
 Since the prior election's value is subtracted from the 'current' one, positive values mean that the current election was more fraudulent than its predecessor, while negative values indicate election fraud declined.
-Using this conceptualization of election fraud does result in some support for the deterrence thesis: an election that is markedly more fraudulent than the prior election does appear associated with more numerous and longer-lasting protest events at higher incumbent vote-shares.
+Using this conceptualization of election fraud does result in some support for the deterrence thesis: an election that is markedly more fraudulent than the prior election does appear associated with more numerous protest events at higher incumbent vote-shares.
 However, as in other examples, this effect only becomes statistically significant at the extremes.
-The values for change in fraud used in the marginal effects plots in Figures \@ref(fig:n-protests-ch) and \@ref(fig:d-protests-ch) represent very unusual changes, roughly between the 5th and 95th percentiles.
+The values for change in fraud used in the marginal effects plots in Figures \@ref(fig:n-protests-ch) and \@ref(fig:size-protests-ch) represent very unusual changes, roughly between the 5th and 95th percentiles.
 In other words, the expected number and duration of protest events are only statistically significantly different when comparing elections that are far cleaner than the previous election and those that are far worse\textemdash a set of options that is unlikely to confront most incumbents.
@@
 
-model.d.base2.lag <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + fraud.change*winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag ,
-                                          zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub %>% filter(v2x_regime <= 2))
+#model.d.base2.lag <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + fraud.change*winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag ,
+#                                          zi = ~.,
+#                                          family= nbinom2,
+#                                          data = nelda2.sub %>% filter(v2x_regime <= 2))
 #summary(model.d.base2.lag)
@@
 models3[['Model 4']] <- model.n.base2.lag
-models3[['Model 5']] <- model.d.base2.lag
-models3[['Model 6']] <- model.size.base2.lag
+#models3[['Model 5']] <- model.d.base2.lag
+models3[['Model 5']] <- model.size.base2.lag
 
 t4a <- huxreg(models3, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Fraud change" = "fraud.change", "Incumbent vote-share" = "winner.share.upd", "Fraud change:Incumbent vote-share" = "fraud.change:winner.share.upd"),
-              statistics = character(0)) %>% set_caption("Models of number, duration, and size of protests") %>% insert_row("", "Number", "Duration", "Size", after = 0) %>% insert_row("Controls", "Yes", "Yes", "Yes", after = 10)
+              statistics = character(0)) %>% set_caption("Models of number, duration, and size of protests") %>% insert_row("", "Number", "Size", after = 0) %>% insert_row("Controls", "Yes", "Yes", after = 10)
 theme_compact(t4a, header_rows = TRUE, header_cols = FALSE)  %>% set_bottom_border(10, everywhere)
@@
 
-```{r n-protests-ch, fig.cap="Change in fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE}
+```{r n-protests-ch, fig.cap="Change in fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE, fig.pos="h!"}
 dat <- ggpredict(model.n.base2.lag, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "fraud.change [-1.12, 1.12]"), type = "count") 
@@
 
-```{r d-protests-ch, fig.cap="Change in fraud, incumbent vote-share, and duration of protest events", echo=FALSE, message=FALSE}
-dat <- ggpredict(model.d.base2.lag, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "fraud.change [-1.22, 1.22]"), type = "count") 
-p.d.margin.ch <- plot(dat)
-p.d.margin.ch <- p.d.margin.ch +  labs(x = "Incumbent vote-share", y = "Duration of protest wave", title = "", color = "Change in fraud") + theme_bw()
-p.d.margin.ch + coord_cartesian(ylim=c(0, 100))
+```{r size-protests-ch, fig.cap="Change in fraud, incumbent vote-share, and size of protest events", echo=FALSE, message=FALSE, fig.pos="h!"}
+dat <- ggpredict(model.size.base2.lag, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "fraud.change [-1.22, 1.22]"), type = "count") 
+p.s.margin.ch <- plot(dat)
+p.s.margin.ch <- p.s.margin.ch +  labs(x = "Incumbent vote-share", y = "Size of protest wave", title = "", color = "Change in fraud") + theme_bw() + coord_cartesian(ylim = c(0, 70000))
+p.s.margin.ch 
+```
+
+Second, the NELDA dataset includes a binary variable noting whether "there was evidence of domestic or international concern that the election process was not going to be free or fair" in advance of the election (nelda11).
+While this is likely to be rough indicator, it still provides an opportunity to test the prediction that election fraud may be more likely trigger protest when it exceeds a baseline expectation.
+In particular, when fraud is high and there are no pre-election expectations of severe manipulation, this view might predict a higher likelihood and severity of protest.
+This is tested by interacting *intentional voting irregularities* with the binary variable from NELDA, called *pre-election expectations* here.
+
+```{r expectations-table, echo=FALSE, message=FALSE, warning=FALSE}
+model.n.concerns.hurd <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    preelection_concerns*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+                                          zi = ~.,
+                                          family= nbinom2,
+                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
+#summary(model.n.concerns.hurd)
+
+model.size.concerns.hurd <- glmmTMB::glmmTMB(max_size ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    preelection_concerns*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+                                          zi = ~.,
+                                          family= nbinom2,
+                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
+#summary(model.size.concerns.hurd)
+
+models8 <- list()
+models8[['Model 6']] <- model.n.concerns.hurd
+models8[['Model 7']] <- model.size.concerns.hurd
+
+t8a <- huxreg(models8, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Election fraud" = "v2elirreg.inv", "Pre-election concerns" = "preelection_concerns", "Election fraud : Pre-election concerns" = "v2elirreg.inv:preelection_concerns"), statistics = character(0)) %>% set_caption("Models of number and size of protests") %>% insert_row("", "Number", "Size", after = 0)
+theme_compact(t8a, header_rows = TRUE, header_cols = FALSE)  %>% insert_row("Controls", "Yes", "Yes", after = 10) %>% set_bottom_border(10, everywhere)
+```
+
+As Table \@ref(tab:expectations-table) and Figures \@ref(fig:n-protests-concerns) and \@ref(fig:protests-size-concerns) indicate, there is no apparent effect for high levels of fraud when there are no significant pre-election concerns about the freeness and fairness of the election (as coded by NELDA).
+Taken together, these results suggest that the alternative, 'exceeding expectations' framework for the protest model of election manipulation is not supported.
+
+```{r, echo=FALSE, message=FALSE}
+dat <- ggpredict(model.n.concerns.hurd, terms = c("v2elirreg.inv", "preelection_concerns"), type = "count") 
+p.n.concerns <- plot(dat)
+p.n.concerns <- p.n.concerns + labs(x = "Intentional irregularities", y = "Predicted number of protests" , title = "Predicted count of protest events", color = "Pre-election concerns")  + theme_bw()
+
+dat <- ggpredict(model.size.concerns.hurd, terms = c("v2elirreg.inv", "preelection_concerns")) 
+p.size.concerns <- plot(dat)
+p.size.concerns <- p.size.concerns +  labs(x = "Intentional irregularities", y = "Predicted size of protests" , title = "Predicted max.
+size of protest events", color = "Pre-election concerns")  + theme_bw() + coord_cartesian(ylim = c(0, 50000))
+
+```
+
+```{r n-protests-concerns, fig.cap="Election fraud, pre-election expectations, and number of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
+p.n.concerns
+
+```
+
+```{r protests-size-concerns, fig.cap="Election fraud, pre-election expectations, and size of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
+p.size.concerns
+
 ```
@@
 models4 <- list()
-models4[['Model 7']] <- model.n.base2.succ
+models4[['Model 8']] <- model.n.base2.succ
 #models4[['Model 6']] <- model.d.base2.hurd
-models4[['Model 8']] <- model.size.base2
+models4[['Model 9']] <- model.size.base2
 
@@
 plot.combined.prob
+```
+
+# Electoral authoritarian regimes only
+
+The models in the main text make use of observations from country-years coded as electoral authoritarian and electoral democratic by the 'v2x_regime' variable in V-Dem.
+This approach is taken because it reflects the widespread availability of election-manipulation techniques across this range of regime types.
+But it may be that considering the two groups simultaneously obscures some within-group patterns.
+For example, it may be that electoral protest is more likely to be observed in electoral authoritarian regimes following the logic in Tucker [-@tucker2007], and given several prominent examples of electoral revolution in that category (e.g.
+Ukraine 2004, Georgia 2003).
+The models in Table \@ref(tab:auth-table) show a statistically significant coefficient for the interaction of election fraud and winner's vote-share, but Figures \@ref(fig:n-protests-margin-auth) and \@ref(fig:size-protests-margin-auth) show that this effect is not apparent for typical values of election fraud in that subset of the population (the first and third quartiles, respectively).
+As a result, there is no strong evidence to suggest that election fraud has different effects on protest dynamics in electoral authoritarian vs.
+electoral democratic regimes.
+
+```{r, echo=FALSE}
+model.n.base2.auth <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+                                          zi = ~.,
+                                          family= nbinom2,
+                                          data = nelda2.sub %>% filter(v2x_regime == 1))
+#summary(model.n.base2.auth)
+
+model.size.base2.auth <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+                                          zi = ~.,
+                                          family= nbinom2,
+                                          data = nelda2.sub %>% filter(v2x_regime == 1))
+#summary(model.size.base2.auth)
+
+```
+
+```{r auth-table, echo=FALSE, warning=FALSE, message=FALSE}
+
+models5a <- list()
+models5a[['Model 10']] <- model.n.base2.auth
+models5a[['Model 11']] <- model.size.base2.auth
+
+t6b <- huxreg(models5a, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd"), statistics = character(0)) %>% set_caption("Models of number and size of protests, only electoral authoritarian regimes") %>% insert_row("", "Number", "Size", after = 0)
+theme_compact(t6b, header_rows = TRUE, header_cols = FALSE)  %>% insert_row("Controls", "Yes", "Yes", after = 10) %>% set_bottom_border(10, everywhere)
+```
+
+```{r, echo=FALSE, message=FALSE}
+dat <- ggpredict(model.n.base2.auth, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [.35, 1.86]"), type = "count") 
+p.n.margin.auth <- plot(dat)
+p.n.margin.auth <- p.n.margin.auth + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events")  + theme_bw()
+
+dat <- ggpredict(model.size.base2.auth, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [.35, 1.86]")) 
+p.size.margin.auth <- plot(dat)
+p.size.margin.auth <- p.size.margin.auth + labs(x = "Incumbent vote-share", y = "Predicted size of protests" , title = "Predicted maximum of protest events")  + theme_bw()
+
+```
+
+```{r n-protests-margin-auth, fig.cap="Election integrity, incumbent vote-share, and number of protest events in authoritarian regimes", echo=FALSE, out.width = '70%', fig.align='center'}
+p.n.margin.auth
+
+```
+
+```{r size-protests-margin-auth, fig.cap="Election integrity, incumbent vote-share, and maximum size of protest events in authoritarian regimes", echo=FALSE, out.width = '70%', fig.align='center'}
+p.size.margin.auth
+
 ```
@@
 models5 <- list()
-models5[['Model 5']] <- model.n.base2.lib
-models5[['Model 7']] <- model.size.base2.lib
+models5[['Model 10']] <- model.n.base2.lib
+models5[['Model 11']] <- model.size.base2.lib
 
@@
 This variable ranges from 0 to 1, with higher values indicating freer and fairer elections.
-The models presented below utilize this variable in place of the main measure of election fraud, while retaining all other control and explanatory variables, using the Mass Mobilization data on protest.
+The models presented below utilize this variable in place of the main measure of election fraud, while retaining all other control and explanatory variables, using the ECAV data on protest.
 
@@
 models <- list()
-models[['Model 9']] <- m1
+models[['Model 12']] <- m1
 
@@
 
-The first model shows that the number of pre-election protests is weakly associated with improved election integrity, but just as in the main text this effect is extremely small.
-Each additional protest is associated with an improvement in election integrity equivalent to 0.2\% of the standard deviation of the dependent variable.
-Given that the mean number of pre-election protests is six, this is an extremely small influence on the overall degree of election integrity.
+The first model shows that the number of pre-election protests is not associated with improved overall election integrity.
+This suggests that pre-election unrest is not associated with broad-scale improvements in election integrity across the electoral environment.
 
@@
 Moving to the hurdle models of post-election protest, we do not observe much of a relationship between overall election integrity and protest.
-There is no statistically significant relationship between election integrity and protest occurrence in Model 2.
-Neither is there any significant relationship in interaction with the winner's vote-share, as shown in Figure \@ref(fig:b-protests-margin).
-
-In the corresponding results for the number of post-election protests, there is similarly muted evidence for a connection between the overall level of manipulation and protest risk to incumbents.
-There is no significant difference between the marginal effect of common levels of election integrity on the number of post-election protests for typical cases across any margin of victory (Figure \@ref(fig:n-protests-margin)).
+There is no statistically significant relationship between election integrity and protest size or quantity in any of the models, either alone or interaction with winner vote-share and lagged GDP growth.
+Figure \@ref(fig:n-protests-margin-fulldv) shows the absence of any effect for the number of protest events, while Figure \@ref(fig:protests-size-fulldv) shows the same for protest size.
+The values for *overall election integrity* chosen for these marginal effects plots represent the first and third quartiles, respectively.
 As a whole, the results are in accordance with the theory proposed in the main paper.
@@
 #models[['Model XXa']] <- model.n.base2.elfr
-models[['Model 10']] <- model.n.base2.elfr
-models[['Model 11']] <- model.n.griev.elfr
-models[['Model 12']] <- model.size.base2.elfr
-models[['Model 13']] <- model.size.griev.elfr
+models[['Model 13']] <- model.n.base2.elfr
+models[['Model 14']] <- model.n.griev.elfr
+models[['Model 15']] <- model.size.base2.elfr
+models[['Model 16']] <- model.size.griev.elfr
 
@@
 
-```{r zero-model-table-count-fulldv, echo=FALSE}
+```{r, echo=FALSE, message=FALSE}
+dat <- ggpredict(model.n.base2.elfr, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2xel_frefair [.197, .61]"), type = "count") 
+p.n.margin <- plot(dat)
+p.n.margin <- p.n.margin + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events")  + theme_bw() + coord_cartesian(ylim=c(0, 75))
 
-t_count_general <- huxreg(model.n.base2.elfr, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, omit_coefs = c("rentierstate1", "taxstate1", "n.events.pre", "unscheduled_election1", "inc.lose1", "v2elirreg.inv:winner.share.upd", "gdpgro.1lag:v2elirreg.inv"), statistics = character(0)) %>% set_caption("Table 3: Zero-inflated models of protest initiation")
-
-theme_compact(t_count_general, header_rows = TRUE, header_cols = FALSE)
-
-```
-
-```{r, echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.base2.elfr, terms = c("winner.share.upd", "v2xel_frefair [-.41, 1.68]")) 
-p.n.margin <- plot(dat)
-p.n.margin <- p.n.margin + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events")  + theme_bw()
-
-dat <- ggpredict(model.size.base2.elfr, terms = c("winner.share.upd", "v2xel_frefair [-.41, 1.68]")) 
+dat <- ggpredict(model.size.base2.elfr, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2xel_frefair [.197, .61]")) 
 p.size.margin <- plot(dat)
-p.size.margin <- p.size.margin + labs(x = "Incumbent vote-share", y = "Predicted size of protests" , title = "Predicted maximum of protest events")  + theme_bw()
+p.size.margin <- p.size.margin + labs(x = "Incumbent vote-share", y = "Predicted size of protests" , title = "Predicted maximum of protest events")  + theme_bw() + coord_cartesian(ylim = c(0, 50000))
 
@@
 
-```{r n-protests-size-fulldv, fig.cap="Election integrity, incumbent vote-share, and size of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
+```{r protests-size-fulldv, fig.cap="Election integrity, incumbent vote-share, and size of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
 p.size.margin
@@
 ```
+
+\newpage
+
+# References

### 2024-01-19 Scripts/authoritarian_only.Rmd
New draft post-AJPS reject

--- 2023-07-01 Output/Submission files/manuscript_anon_v2.Rmd
@@
 On August 9, 2020, Belarus held an election in which President Aleksander Lukashenko ran for a sixth term.
-The official results credited Lukashenko with over 80\% of the vote, a result which was regarded by observers as wildly inflated [@GolosPlatformPresents2020].
+The official results credited Lukashenko with over 80\% of the vote, a result which observers regarded as wildly inflated [@GolosPlatformPresents2020].
 In response, hundreds of thousands of Belarusians took to the streets over the course of several months.
-Though the protests put significant strain on the regime, a harsh campaign of repression ultimately quelled the protests and Lukashenko was inaugurated into a new five-year term.
-The Belarusian case offers a useful illustration of a central puzzle in the study of election manipulation: why did the risk of costly protest fail to deter the incumbent from engaging in such widespread fraud?
-And why did the regime use fraud to claim such an implausibly high margin of victory?
+Though the protests put significant strain on the regime, a harsh campaign of repression (and support from neighboring Russia) ultimately quelled the protests; Lukashenko was inaugurated into a new five-year term.
+
+The Belarusian case is typical in some ways, and an outlier in others.
+Like many previous elections in Belarus, and in authoritarian regimes around the world, the contest was heavily managed and the results inflated by fraud.
+However, the sustained, massive protest wave that occurred post-election is extremely unusual.
+However, the regime's ultimate victory over the protest movement\textemdash while costly and never guaranteed\textemdash is also the typical outcome of post-election protest.
+
+These patterns highlight a puzzle generated by existing research on the relationship between election manipulation and popular protest.
+The topic has significant implications for authoritarian governance and democratization, and as such has been a very active subject of study.
+However, the various theoretical perspectives brought to bear on the topic have generated competing predictions.
+To briefly summarize: for some models, election fraud is a tool of weak incumbents.
+As a result, incumbents avoid fraud when it might be exposed and reveal their weakness to potential protesters\textemdash an auspicious prediction for democratization and ultimately the under-pinning of 'self-enforcing' democracy [@fearon2011].
+For others, fraud is committed by weak and strong incumbents alike, making protest a risky gamble the opposition undertakes for other reasons.
+In these models, fraud may or may not be deterred based on other exogenous factors: a dubious, though not completely dismissive, prospect for democratization and self-enforcing democracy.
+Lastly, a third school of thought holds that election fraud is an indicator of genuine incumbent strength.
+This approach, which is expanded upon below,
 
@@
 Testing this proposition requires looking beyond binary measures of protest, and toward measures of its scope.
-Research on the risk posed to non-democratic governments by electoral protest often acknowledges---whether explicitly or implicitly---the importance of broad-based protest campaigns, rather than one-off events [@Bunce2006;@tucker2007;@DAnieri2006;@Carnaghan2016;@bunce2010].
-As Brancati [-@Brancati2016] finds, approximately two-thirds of pro-democracy electoral protests ended within three days, posing little threat to incumbents.
+Research on the risk posed to governments by electoral protest often acknowledges---whether explicitly or implicitly---the importance of broad-based protest campaigns, rather than one-off events [@Bunce2006;@tucker2007;@DAnieri2006;@Carnaghan2016;@bunce2010].
 This implicit assumption has largely been overlooked in theories and empirical studies of the relationship between election manipulation and protest, however.
@@
 This approach obscures the importance of the number and size of electoral protests in a ruling party's calculation of the risk involved in manipulation an election.
-Small, narrow protests may be easily contained by a well-resourced government, while more numerous, larger protest events pose a more significant challenge.
-
-Using cross-national data on 1110 elections from 1989 to 2012 where incumbents claimed victory, combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Electoral Contention and Vioence dataset [@Daxecker2019], this paper finds little support for the hypothesis that more fraudulent elections are associated with a higher risk that any post-election protest event occurs.
+
+Using cross-national data on 1110 elections from 1989 to 2012 where incumbents claimed victory, combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019], this paper finds little support for the hypothesis that more fraudulent elections are associated with a higher risk that any post-election protest event occurs.
 Furthermore, it shows that more severe election fraud is not associated with more numerous or larger post-election protests.
-Instead, these protest dynamics appear driven more by structural and institutional factors, including the size of the political opposition and whether or not the incumbent chief executive is term limited.
+Instead, these protest dynamics appear driven more by structural and institutional factors, including in particular whether or not the incumbent chief executive is term limited.
 
@@
 These results help explain both the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory.
-They also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations.
-
-# Literature review
-
-Broadly speaking, prior work has considered two major explanations for the severity of manipulated elections: the costs and benefits of manipulation to the ruling party and government (a demand-side framework) and the costs and benefits for the front-line actors who do the work of manipulating elections (a supply-side framework).
-The risk of costly protest is the central, though not sole^[In some studies, like [@birch2011], the costs of manipulation are attributed to a more diffuse loss of legitimacy; such costs *may* manifest as protest, but may also manifest as a reduction in general cooperation with the regime.
-However, it is worth noting the argument of [@gehlbach2015electoral] that election manipulation can *induce* cooperative behavior from some agents], factor which affects the demand-side cost of manipulation in most models.
-This insight can be found, in related forms, in both the literature on self-enforcing democracy and on the conduct of non-democratic elections.
-In the former literature, generally free and fair elections are thought to be a self-enforcing equilibrium through the threat of mass protest by wronged groups if the incumbent government rigs the results.
-In many formal models, including those discussed below, the decision to protest or consent to the election is the play available to citizens or opposition parties; protesting imposes a cost on the regime and/or carries a risk that the incumbent is overthrown.
-
-In the major formulations of this argument, such as Fearon  [-@fearon_self-enforcing_2011] and Przeworski [-@przeworskiSelfenforcingDemocracy2006], 'democracy' is understood in procedural terms.
-For Przeworski, democracy requires that "authorization to rule is derived from elections" which designate winners and losers according to a set of rules [@przeworskiSelfenforcingDemocracy2006, p.
-312].
-Fearon adopts a similar definition for electoral democracy [@fearon_self-enforcing_2011, p.
-1662].
-Such arrangements are not limited to liberal democracies, and in fact constitute the large majority of regimes today.
-Many previous works argue that the risk of protest can mitigate election manipulation in regimes where elections are held\textemdash including established democracies [@millerRepublicIfYou2021], electoral authoritarian regimes [@Egorov2021], and any regime where elections are held but may be manipulated [@Little2015;@Luo2020]\textemdash but this body of work has yet to grapple with the empirical finding that election integrity is relatively poor across these regimes outside of a narrow club of 'consolidated' democracies.
-
-There are thought to be two mechanisms by which election manipulation leads to protest, and thereby might restrain manipulation.
-The first is through revealing information about the strength of the government or ruling party.
-This information, in turn, feeds into strategic decision-making by opposition groups.
-This logic is often seen in formal models [@Egorov2021;@chernykh2015third].
-Several prominent models of electoral manipulation argue that the risk of mass post-election protest helps restrain incumbents who might otherwise prefer larger levels of manipulation [@magaloni2010; @Little2015; @Little2012; @birch2011; @fearon_self-enforcing_2011;@Luo2020;@Rozenas2016b].
-The risk of mass protest in these models is assumed to constitute a significant guardrail for election integrity in democracies, and a significant challenge for ruling parties in non-democracies.
-Building on this logic, researchers have proposed that third-party institutions---like election monitors [@Little2012;@Hyde2014;@daxecker2012], civil society groups [@birch2017getting], and courts [@chernykh2015third]---can reduce manipulation and increase protest risk by revealing information about electoral malfeasance.
-Of course, electoral manipulation is not understood to be the sole cause of protest\textemdash protests occur for all sorts of reasons and across time\textemdash but the literature has drawn a close connection between the risk of specifically electoral protest and the conduct of elections.
-
-Despite this consensus in the formal literature, there is relatively little empirical evidence to support the claim that manipulation per se indicates weakness, and thus increases protest risk.
-For example, looking at the 2011 election and protest cycle in Russia, Ananyev and Poyker [-@ananyevDictatorsSignalStrength2022] show that post-election protest was not associated with the severity of fraud using several measures.^[It should be noted that Lankina and Skovoroda [-@Lankina2017] show that a different measure of fraud was associated with post-election protest in the subsequent presidential election.] In a cross-national study, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e.
-whether opponents are allowed to run, whether the media environment is pro-regime), rather than by the election-day manipulation that usually figures in theories of election integrity and protest.
-This result is consistent with the finding that parties are more likely to protest elections after ex ante changes to the rules [@chernykh2014].
-
-The literature on protest as a deterrent to manipulation also often overlooks the functions that elections, even highly controlled elections, may serve in non-democratic settings.
-For instance, Simpser [-@simpser2013] shows that high levels of manipulation can actually induce cooperation from opposition groups, elites, and other actors by revealing information about regime strength.
-Similarly, the type of manipulation employed can convey information about the ruling party's capacity to mobilize, with costlier forms of manipulation associated with less protest [@Harvey2020].
-In this way, large-scale manipulation can contribute to a ruling party's ability to produce a large, stabilizing majority vote-share [@magaloni2006].
-Similarly, Luo and Rozenas [-@Luo2020] posit that ex post fraud is more likely to result in protest than ex ante manipulation.
-In sum, election manipulation can be performed by strong regimes for reasons that go beyond 'mere winning.'
-
-<!--  One critique of this view, although still within the information/strategic framework, comes from Simpser's [-@Simpser2011] pathbreaking book.
-Simpser argues that when wanipulation efforts are excessive---well beyond what is needed simply to win---they demonstrate the organizational capacity, resources, and manpower of the ruling party [@Simpser2013].
-Large-scale manipulation is, thus, a signal of strength rather than weakness, and can deter protest.
-By contrast, partial manipulation efforts that fail to secure a large majority vote-share signals the government lacks both popular support and mobilizational capacity, making opposition protest more likely.
--->
-
-A second school of thought has emphasized the ways in which biased elections may predispose individuals to engage in protest [@kuntz_more_2009].
-In this view, election manipulation acts as a shared grievance among individuals who prefer to see the ruling party lose\textemdash especially when it appears that fraud may have been decisive [@kuntz_more_2009].
-Belief in a stolen election is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007], possibly initiating a protest cascade [@Kuran1991].
-This conception is closely connected to the idea that manipulated elections carry legitimacy costs for the ruling party [@birch2011].
-
-For example, ballot-stuffing and voter coercion have been shown to generate stronger feelings of anger than pre-election manipulation of ballot access [@Szakonyi2021], and access to information about manipulation appears to reduce support for the ruling party [@Reuter2021], increase support for protest and reduce the perceived legitimacy of the regime [@Williamson2021].
-This argument has found support in studies of the relationship between economic grievances, manipulated elections, and protest [@Rod2019;@Brancati2016].
-
-The grievance-based approach suggests that protests hinge more on citizens' subjective beliefs about fraud and the broader social context than on the actual level of malfeasance.
-A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b].
-Subjective belief in fraud was also shown to be correlated with approval of mass protests after Russia's 2011 election [@Chaisty2013].
-Furthermore, simply being on the losing side in an election is associated with reduced belief that the election was fair [@Cantu2015], and with increased support for electoral protest [@Anderson2006].
-These studies suggest that electoral protest can occur at low or high levels of fraud, contingent on other factors\textemdash including subjective, and perhaps mistaken, beliefs.
-
-The strategic and social-psychological explanations are, of course, interconnected.
-As Schedler writes, "Grievances provide moral justifications for collective protest: we need to do something.
-Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304].
-Several empirical studies have, accordingly, found links between election manipulation and binary measures of protest [@Brancati2016;@Hyde2014;@Harvey2020;@Rod2019].
-
-There are thus unresolved tensions in the literature on the relationship between electoral manipulation and electoral protest.
-One branch of the literature argues that the risk of mass protest reduces incumbents' incentive to engage in manipulation, resulting in cleaner elections.
-A second branch argues that mass protest hinges on citizens' subjective assessments of election integrity, their individual grievances, and whether or not the election is seen to be stolen.
-This view complicates the deterrence logic, since incumbents may minimize protest risk (and thus more freely manipulate) by reducing citizen grievances (i.e.
-intervening in the business cycle at election time, distributing club goods, etc.) or by controlling the information environment.
-Since citizens' subjective views of manipulation may diverge from the true level, incumbents may also face an incentive to manipulate the election severely in the hopes of achieving a large margin of victory, under the assumption that protest is likely even at low levels of true manipulation.
-Finally, while a few empirical studies do show evidence for the deterrence model, overall empirical support for the proposition is mixed at best and lacks a general test.
-
-Finally, despite widespread election manipulation, electoral protest is relatively rare; even in cases where it does occur, most efforts to overturn fraudulent election results through protest have failed [@kalandadze2009electoral].
-The apparent failure of protest risk to deter election fraud underlines our limited understanding of non-democratic ruling parties' decision-making.
-Below, it will be argued that election manipulation by its nature will rarely be associated with mass post-election protest.
-Elections in non-democracies are *generally* risky for incumbents, regardless of how free and fair they may be, as they create an opportunity for opposition collective action [@Knutsen2017] and reveal information about the government and society [@Pop-Eleches2015].
-In most cases, however, election manipulation adds little to this risk.
-
-# Theory
-
-The foregoing literature review sketches out prior models of the relationship between election manipulation and protest.
-That existing work has several implications that will be tested here.
-First, does a higher risk of protest deter subsequent election fraud?
-And second, do high levels of fraud increase the risk of subsequent protest?
-
-## Predictions from prior work: Fraud and the risk of protest initiation
-
-As noted above, much of the formal literature on election manipulation assumes that the risk of protest serves as a deterrent to election manipulation.
-In addition, the logic of deterrence suggests that increasingly severe election fraud should make post-election more likely, either unconditionally or when the incumbent's margin of victory is low.
-More extensive election-day manipulation, in these models, is generally taken as a signal of regime unpopularity.
-This in turn makes citizens [@Egorov2021;@Little2015] or opposition parties [@Luo2020;@Rozenas2016b;@magaloni2010] more likely to stage protests.
-Previous research has also suggested that a fraudulent election is more likely to trigger protest when social grievances are high [@Brancati2016;@tucker2007].
-With some exceptions, such as the study by Lankina and Skovoroda [-@Lankina2017], most research on election fraud and protest treats the latter as a binary outcome\textemdash protests either occur or not.
-This leads to the following hypotheses:
-
->H1: Greater pre-election protest risk will be associated with lower levels of election-day manipulation.
-
->H2: The likelihood of protest initiation will increase as the severity of election fraud increases.
-
->H3: The effect of election manipulation on the likelihood of protest initiation will increase as the incumbent's margin of victory falls.
-
->H4: The effect of election manipulation on the likelihood of protest initiation will increase as economic grievances rise.
-
-However, there are reasons to doubt the veracity of the deterrence logic.
-Election fraud and other forms of manipulation that do not directly mobilize voters (such as vote-buying) cannot be a reliable indicator of the strength or weakness of the regime; consequently the 'signal' provided by a fraudulent election will not be able to reliably generate strategic behavior by the opposition.
-The central problem is that election manipulation is a tool that can be used by weak and strong governments alike [@simpser2013].
-A strong government may allow for relatively clean elections, in order to gather more information about the opposition and society [@ananyevDictatorsSignalStrength2022;@trinhTeaLeafElections2022]; a weak government may be unable to muster the necessary resources [@greene_why_2007].
-The observation of a relatively clean election alone does not discriminate between the two.
-Likewise, a strong government may manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver].
-This is all in contrast to the typical deterrence model, in which a weaker government attempts to manipulate in order to obscure its weakness, but in keeping with Seeberg's [-@Seeberg2019] finding that higher-capacity states are weakly associated with more election fraud.
-For these reasons, acts of manipulation can be interpreted either as signs of regime strength or weakness [@Pop-Eleches2015].
-It is also in line with the formal model proposed by Little [-@Little2012], in which citizens\textemdash knowing the regime is capable of manipulating the results\textemdash discount an expected level of fraud from the results when deciding whether or not to protest.
-
-Additionally, in-depth investigations of the causes of the Color Revolutions (in particular) have emphasized the role of opposition agency and learning in staging post-election protest.
-In this view, the likelihood of mass electoral protest is increased by deliberate efforts by opposition parties to form a unified opposition coalition, to raise popular expectation of a stolen election, to prepare logistics, and to actively monitor the vote and publicize discrepancies [@bunce2010].
-Others have noted structural factors like the importance of ties to democratic Western states and the size and strength of the ruling party and security apparatus in influencing the course of the Color Revolutions [@LucanWay2008].
-In either case, these factors create an incentive for opposition protest whether electoral fraud is minor or severe\textemdash a kind of case of 'crying wolf'\textemdash an outcome that is also posited in some of the formal modeling literature [@Little2015].
-As a result, it is likely that the severity of fraud has little bearing on protest risk; what matters is the underlying balance of strength between the incumbent regime and its opposition.
-
-## The course of protest: do higher levels of fraud make protest waves more significant?
-
-While it is useful to check prior predictions on protest initiation, by emphasizing binary protest outcomes at the expense of trends over time, prior work has overemphasized the risk of popular backlash to incumbents.
-While many incumbents may indeed prefer a quiet post-election period, the occurrence of protest can also have regime-sustaining benefits.
-When unexpected post-election protests occur, trust in the government can increase, especially among opposition supporters [@Frye2019].
-As the number of protests increases, support for the protesters' aims may fall; this effect is even larger for those who consume pro-government media [@tertytchnayaElectoralProtestsPolitical2020].
-These empirical results are in line with models proposed by Lohmann [-@lohmannDynamicsInformationalCascades1994] and Shadmehr and Bernhardt [-@shadmehrCollectiveActionUncertain2011]; they argue that when some citizens are too predisposed to protest, others become more inclined to stay home\textemdash dooming the first group's efforts to failure.
-Such influences help offset the 'risk' of protest to incumbents, especially when protest movements are relatively small and/or isolated.
-
-What incumbents likely *do* seek to avoid are large, widespread, and long-lasting protest movements that strain the government's resources and put pressure on elites [@chenowethPhysicsDissentEffects2019].
-Here I argue that, once protests begin, the extent of electoral fraud is unlikely to affect their scope.
-The initial emergence of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014].
-That election dates are known in advance creates opportunities for opposition mobilization; anticipating this, regimes can more easily engage in pre-emptive repression [@Truex2019;@Sullivan2016;@Bhasin2013].
-
-In many cases, the government doubles down on repression, calling upon the security services to disperse protests, arrest leaders, and perhaps employ harsher methods including disappearance, torture, and execution [@Escriba_Folch2013;@Davenport2007].
-Though such crackdowns risk provoking a backlash, they are aimed at demobilizing the opposition by increasing the costs of participation in overt dissent [@DeVogel2022].
-In other cases, the government may attempt to accommodate the opposition, by making concessions on policy or institutions [@mooreRepressionDissentSubstitution2000;@Hafner-Burton2016;@Piven1979].
-Alternatively, concessions may be aimed at the more moderate public, in order to keep them on the sidelines of the conflict [@Hummel2019].
-In addition, governments may learn from the experiences of other governments, just as opposition groups may learn from opposition.
-Anticipating threats from below may encourage less democratic governments to invest more in the security services, or to reform political and legal institutions in order to stifle potential protests before they take shape [@hallAuthoritarianLearningConceptual2017a;@Morgenbesser2020].
-
-What results is a test of strength between the anti-incumbent protest coalition and the regime with its supporters [@mcadamDynamicsContention2001;@tillyRegimesRepertoires2010].
-In essence, as Beaulieu [@Beaulieu2014a] notes, the onset of electoral protest marks a case of bargaining failure.
-Just as in the international context, the ensuing conflict acts to reveal information on the strength and resolve of the two sides.
-In such a contest, the incumbent enjoys multiple advantages, though it may lack the support of a majority of the population.
-It controls the resources of the state, and can call upon the support of those elites who benefit from the status quo.
-High-capacity states, as a result, are more likely to be able to engage in successful repression [@tarrowPowerMovementSocial2011].
-
-Though a biased election may temporarily allow diverse opposition groups to set aide their differences [@trejo_ballot_2014], this internal diversity can enable the incumbent to divide the opposition with selective repression or concessions.
-While the regime may have its own splits---between reformists and hardliners, for example---the emergence of protest can act to solidify support for the ruling party among its supporters.
-This can be especially true non-democracies where elites have strong incentives to hold to power, in the form of access to rents and preservation of immunity for prosecution   [@baturoDemocracyDictatorshipTerm2014].
-
-In this struggle between the potential protest coalition and the incumbent government, protesters and their possible allies behave as "passionate economists" whose balance of emotional versus strategic considerations can be updated over time [@VanZomeren2012].
-What is the likelihood that protest action can help resolve the grievance, and does the benefit of participating outweigh the risk of personal consequences to the individual for participating?
-A fraudulent election may heighten citizens' grievances and generate uncertainty around the underlying popularity of the regime, creating a moment of elevated protest risk.
-Crucially, though, this evaluation is dynamic and can be updated in response to new information and new events [@VanZomeren2012].
-If protest movements have sufficient numbers and organizational resources to put the regime on the back foot, this can lead to more numerous protests as individuals' sense of group efficacy grows and their sense of risk declines [@Chenoweth_Stephan2011].
-On the other hand, if the regime is well-resourced, capable, and unified, its actions to end a protest movement can lead to a negative re-evaluation of the movement's odds of success.
-This does not, of course, eliminate citizens' sense of grievance.
-Instead, it leads them to abandon 'approach-oriented' forms of coping, to avoidance-oriented forms: demobilization and disengagement [@van_zomeren_put_2004;@Young2019].
-For example, while ruling-party supporters in Russia disapprove of election fraud [@Reuter2021], they are more likely to withdraw from politics after exposure to electoral protest than to support the opposition [@Tertytchnaya2020].
-
-In summary, fraudulent elections are not likely to pose a significant protest risk to incumbents.
-While opposition actors may seize upon fraudulent elections to initiate a protest, the relationship between this decision and the level of fraud is likely to be weak.
-Opposition groups are also incentivized to stage protests regardless of the level of fraud, given the highly salient moment afforded by the election itself [@cunhaSignalReceivedAuthoritarian2022].
-Fraud itself provides little information about regime strength, and even if a protest movement begins after the election, any useful information provided by the level of fraud is quickly replaced by signals from the protests themselves [@lohmannDynamicsInformationalCascades1994].
-Protesters, bystanders, and elites can update their estimation of the relative strength of the protest coalition and the incumbent government by observing the size of crowds, the number of protests, and the government's reaction.
-
-The number and size of protests will instead be related to other indicators of opposition and incumbent capacity.
-When opposition groups and civil society capacity are more organized and better resourced, it is likely that more protesters will take to the streets in more numerous events (see Chenoweth et al [-@chenowethStateRepressionNonviolent2017] for a review of factors making protest movements more resilient to repression).
-By contrast, when the government appears well-resourced and resolved, more 'passionate economists' should stay home.
-As identified in prior research, informative factors should be the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], the margin of victory in the election [@simpser2013], and whether or not the chief executive faces term limits [@hale_patronal_2014].
-Generally speaking, these structural variables indicate aspects of the political opportunity structure that make protest more likely to occur [@meyer_protest_2004].
-
-Here, this logic will be tested using two measures, both of which may indicate the strength of a protest movement, as well as the cost of protest to incumbent governments: the total number of protest events and the estimated number of protesters on the largest single day of protests.
-I assume that more numerous protests, and those that built to a higher peak participation, are costlier for governments to contain, all else held equal.
-If fraudulent elections reveal information about regime weakness or generate large grievance-motivated protest, more fraudulent elections should be associated with larger and more numerous protest waves.
-Prior theory expects these effects to be largest when margins of victory are smaller, or when economic grievances are high.
-
-However, the theory articulated above does *not* expect that the severity of fraud will be associated with the course of protest.
-Instead, it predicts that underlying indicators of opposition and ruling-party strength will influence the direction of protest movements.
-In particular, it expects the size and number of protest events will be sensitive to the repressive capacity of the state and to indicators of opposition strength.
-that high and low levels of repression will be associated with smaller protest movements (in number and size).
-
->H5: The relationship between repressive capacity and the number of protest events and the maximum event size is curvilinear, with number and size likely to to first increase and then fall as repressive capacity grows.
-
->H6: The number of protest events and the maximum event size are likely to increase with indicators of opposition and civil-society strength.
-
-To further test the mechanism outlined in the theory presented here\textemdash that election manipulation offers few clues to regime strength, and thus the course of electoral protests are determined by more structural factors\textemdash I examine the relationship between fraud and the survival of protest movements in cases where protests began but the incumbent eventually survived.
-Restricting the sample to these cases allows for a consistent interpretation of results: longer protest waves indicate a more even match between the resources and resiliency of the protesters and the capacity and cohesiveness of the regime, while shorter-lived protest movements are those which were more easily defeated.
-In this sub-sample, I would first expect the relationship between election fraud and protest survival to be non-linear with respect to time\textemdash the longer the protest movement survives, the less relevant the fraud signal (such as it is) becomes).
-Second, it is expected that more severe election fraud will *not* be associated with longer protest waves, as would be predicted if fraud were an indicator of state weakness.
-Instead, structural factors should again be more influential; indicators of opposition and civil-society strength should help keep protest waves alive.
-
-<!-- -Democracy leads to less repression: Bueno de Mesquita et al., 2005). Davenport & Armstrong (2004),  -->
-<!-- -Prior repression makes repression more likely (Davenport 2007 (Tyrannical peace)) -->
-<!-- -Resource wealth makes authoritarian survival more likely (Girod et al 2016) -->
-
-<!-- From E-F 2013: Nevertheless, violent repression may have pernicious consequences, since it may bring about a -->
-<!-- violent escalation or a radicalization of positions, thereby unleashing a violent response from some -->
-<!-- groups and even potentially provoking a coup (Gurr, 1970; Ortiz, 2007; Pierskalla, 2010). Thus, -->
-<!-- Opp and Ruehl (1990) suggest that repression may lead to micro-mobilization processes leading to -->
-<!-- further protest activities and unrest. Likewise, Ginkel and Smith suggest that ‘dissident activity is -->
-<!-- more likely to be effective in motivating rebellious action under highly repressive conditions’ -->
-<!-- (1999: 292). Some argue that when nonviolent actions are violently repressed, protesters may turn -->
-<!-- to violent activities (Lichbach, 1987; Moore, 1998), as the Syrian case illustrates. Similarly, -->
-<!-- Goodwin (2001) argues that revolutions are more likely in patrimonial and very repressive states. -->
-<!-- For instance, in Iran on Black Friday in September 1978, the regime’s security forces killed hun- -->
-<!-- dreds of anti-shah protesters in Tehran. After a brief decrease in activity, public dissent and strikes -->
-<!-- quickly resumed and escalated. The mobilizations that took place in December gathered hundreds -->
-<!-- of thousands of people calling for the shah’s rule to end. In January 1979, the shah left the -->
-<!-- country -->
-
-<!-- ...restrictions on civil rights can thwart the ability of some -->
-<!-- actors (especially the opposition) to organize, coordinate, and mobilize in the short run -->
-
-<!-- On uncertainty of repression and protest, from Girod et al: Yet the differences in demobilization of protest movements following repression remain largely unexplained.
-Government repression of protesters is strongly associated with quelling protest activity in some cases (Lyall, 2009; Schatzman, 2005), and triggering increased mass mobilization in others (Francisco, 1995; Khawaja, 1993; Mason and Krane, 1989).
-Complicating matters further, in some cases the relationship between repression and protest appears to change over time (Francisco, 1995; Inclan, 2009; Mason and Krane, 1989; Rasler, 1996; Tsebelis and Sprague, 1989), across levels of repression (DeNardo, 1985; Levitsky and Way, 2010; Muller and Weede, 1990), and based on the complex ways in which “threat” and “opportunity” combine to influence protest dynamics (Goldstone and Tilly, 2001).
-Because fine-grained global data on repression of mass movements have not been available until now (Davenport, 2005: viii; 2007a: 10), extant studies depend largely on only a few cases.
-In short, the literature has yet to identify cross-national patterns to explain when publics in an authoritarian regime respond to repression with greater collective action vs.
-disbanding.
--->
-
- <!-- Chernykh and Svolik, in particular, argue that by revealing information about manipulation, even moderately independent courts can increase the risk of protest against manipulated elections (thus deterring manipulation), while also reducing the risk that opposition groups mistakenly protest clean elections.
-This result is in keeping with the argument that electoral protests are the outcome of bargaining failure among incumbents and opposition groups, which are worsened in low-information environments [@Beaulieu2014a].
--->
-
-<!-- However, one of the central functions of independent judiciaries outside of liberal democracies is to create mechanisms through which citizens grievances can be addressed [@Whiting2017; @Solomon2007; @Hilbink2007d].
-Relatively impartial courts create an opportunity for citizens to resolve grievances against the state and against other individuals who may be more powerful or influential.
-For example, survey-experimental data show that citizens of several Middle Eastern non-democracies were more willing to engage in legal action to address their grievances when they believe courts are more independent and assertive [@Schaaf2021].
-Moreover, the belief that procedures are generally followed impartially can bolster the state's claim to legitimacy [@Gerschewski2013], increasing citizens' trust in authority and willingness to abide by political outcomes [@Grimes2006].
-Evidence from European democracies suggests that that de facto independence is key for such judgments [@Buhlmann2011].
-Taken together, greater de facto judicial independence ought to reduce protest risk, both by creating a channel for the state to address citizens' general grievances and by offering an alternative venue to the street for addressing specific electoral concerns.
--->
-
-<!-- These two functions for courts are at odds with regard to electoral disputes.
-As a result, the relationship between judicial independence and post-election protest is conceptually and empirically muddled---despite the centrality of protest risk to predominant models of electoral manipulation.
-The formal model proposed by Chernykh and Svolik [-@Chernykh2015] is not tested empirically, while Chernykh [-@chernykh2014] and Eisenstadt [-@eisenstadt_courting_2003] both find no effect for judicial independence on the likelihood of protest in post-communist Europe and Mexico, respectively.
-This calls into question the prediction that more independent courts make protest more likely in the presence of election manipulation, though to my knowledge the question has not been studied in a large-n cross-national context.
--->
+They also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation.
 
@@
 
-nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2023-05-30.csv"))
+nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2024-01-16.csv"))
 
@@
 
-#Margins with updates
-
-nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))
-
 #Successful protest = new elections or new incumbent
@@
 
+# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections
+
+nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))
+
+## Replacing NA with 0 for protest size
+
+nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(is.na(max_size)==T & n.events.post == 0, 0, max_size))
+
+## Fixing missing values of winner.share.upd for liberal democracies
+
+nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(is.na(winner.share.upd)==T & v2x_regime == 3, winner.margin, winner.share.upd))
+
+## Set max_size zero for elections with no protest
+
+nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))
+
+## Excluding liberal democracies
+
+nelda2.sub <- nelda2.sub %>% filter(v2x_regime == 1) #Includes only authoritarian
+
 #Protest cases only
@@
 
-# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections
-
-nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))
-
-## Replacing NA with 0 for protest size
-
-nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(is.na(max_size)==T & n.events.post == 0, 0, max_size))
-
-## Fixing missing values of winner.share.upd for liberal democracies
-
-nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(is.na(winner.share.upd)==T & v2x_regime == 3, winner.margin, winner.share.upd))
-
-## Set max_size zero for elections with no protest
-
-nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))
-
-## Excluding liberal democracies
-
-nelda2.sub <- nelda2.sub %>% filter(v2x_regime <= 2) #To-do: Extent to include liberal dems in appendix
-
-```
-
-These hypotheses are tested using data from three datasets: NELDA [@hyde_2012_marinov], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019].
-The unit of observation for this study is an election period as recorded in NELDA.
-This framework prevents selecting on the dependent variable, by capturing both protested and non-protested elections.
-I limit the sample by excluding closed autocracies (where no national elections are held).
-Since the theory of popular backlash to manipulation has largely been articulated in an electoral-democratic context, I exclude liberal democracies from the main analysis.^[The specific variable used is 'v2x_regime'; I limit the sample to those regime-years coded 1 or 2.] As a robustness check for this sampling decision, I also run the main models with data including liberal democracies in the appendix, with no substantial changes in the outcomes.
-The first set of models, focusing on Hypothesis 1, include all the remaining cases in the dataset.
-For the subsequent models, focusing on post-election protest, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the theoretical expectations of protest risk following an election in which the incumbent won, possibly through reliance on election fraud.
-Finally, the beginning of the date range is limited by the availability of data on protest from ECAV, as discussed below, which ranges from 1989 to 2012.
-The resulting dataset includes 1110 total election-period observations across 118 countries.
-
-## Dependent variables and modeling strategies
-
-To capture the severity of *election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than aggregate measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], it can fail to signal strength [@Harvey2020], and it can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021].
-This measure of election integrity should thus serve as a strong test of the protest-oriented model.
-The variable is used as a dependent variable in the first set of models that test H1, and as a potential explanatory variable in the subsequent models of protest number, size, and duration.
-
-ECAV records protests and other contentious actions related to an election, ranging from six months before to three months after the vote (or three months after the final round of a multi-round election).
-To focus the study on protest, I only include events labeled as protests, occupations, or blockades, excluding acts like shootings, bombings, and coups.
-ECAV data is the event-day format; to capture the *number of protests* before and after the election, I take the sum number of events recorded in ECAV before and after election day, with election-day protests being counted as post-election.
-To evaluate hypotheses regarding protest size, I convert ECAV's categorical data on number of participants per event-day to estimate numbers, and take the sum where multiple protests were held on a given day.^[This approach is necessary to avoid conflating days with, for example, five category 1 protests (totaling perhaps 50 participants) and days with a single category 5 protest (totaling more than 10,000).] The ECAV dataset codes protest size as a categorical variable with five levels, ranging from fewer than ten participants to more than 10,000.
-To deal with this data structure, where the top category has no upper limit, I take the median value of each category, with the exception of the top category.
-For this category, which has no upper bound, I take the minimum value of 10,000.
-The variable *maximum protest size* is found by taking the number of estimated protest participants who turned out on a single day.
-
-The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0) %>% nrow()` of the `r nrow(nelda2.sub)` elections, no protests are recorded.
-Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest number and maximum single-day protest size.
-As the figure shows, while election fraud is common among this set of regimes, protests are rare, and usually small in number and maximum size when they do occur.
-This distribution\textemdash with large numbers of zeroes\textemdash is in line with the theoretical proposition that two data-generating processes are at work: the factors that lead to the occurrence of protest may not be the same as those that sustain or intensify it.
-Consequently, I model the number of protests using negative binomial hurdle regressions.
-Hurdle models are two-part regressions in which the variation in the dependent variable is assumed to follow two distributions: the probability of obtaining a zero or non-zero outcome is modeled as binomial logit function, while the distribution of the non-zero count variable is modeled (in this case) using the negative binomial function.
-The logistic portion of the model identifies the factors that make *any protest* more likely to occur; the negative binomial regressions model the count of protests given that protest has begun.
-Since protest size cannot take on a negative value, it is also modeled as a count variable with a negative binomial distribution in the hurdle models.
-*Election fraud*, which follows a more normal distribution, is modeled using multilevel models with random intercepts by country.
+```
 
@@
 p.hist.fraud <- ggplot(nelda2.sub, aes(x = v2elirreg.inv)) + geom_histogram(bins=100) + xlab("Election fraud") + theme_bw()
+
+ ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= jitter(n.events.post))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("n.events.post (jittered)") + theme_bw()
+
+  ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= max_size)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("max size") + theme_bw()
 ```
@@
 
-Given that pre-election protest is likely to be a strong predictor of post-election protest\textemdash by indicating strong social networks [@mateo2022], pre-existing opposition groups and civil society capacity [@Bunce2006], and/or popular grievances [@Brancati2016]\textemdash the deterrent logic suggests that non-democratic governments will pull back on election-day fraud when the pre-election environment is characterized by protest [@Rozenas2016b].
-While it is possible that pre-election protest may provide an opportunity for incumbents to crack down, creating more space for manipulating the election, more active pre-election protest suggests that opposition capacity is more robust, all else held equal.
-Moreover, in general, incumbents tend to avoid mass repression in the pre-election phase in favor of targeted harassment of opposition activists [@Bhasin2013].
-
-The first set of models test Hypothesis 1, on the relationship between protest risk and fraud.
-Protest risk is operationalized in three ways.
-First, the *number of pre-election protests* in ECAV is taken as *prima facie* evidence of a more turbulent and risky election environment\textemdash it indicates greater mobilization capacity in civil society and/or the opposition, and may also indicate an elevated sense of grievance.
-Second, the repressive capacity of the regime is modeled using the lagged *physical integrity index* for the country-year.
-In one approach, higher values represent more permissive environments, and should be associated with greater protest risk (and thus reduced fraud).
-Alternatively, in keeping with the 'murder in the middle' framework [@feinLifeIntegrityViolationsDemocracy1995], the highest protest risks (and most reduced fraud) might be found at middle values of repression; regimes with the most demonstrated repressive capacity deter potential challengers from taking to the streets in the first place, and the least repressive states defuse protest through accommodation [@pierskallaProtestDeterrenceEscalation2010;@mooreRepressionDissentSubstitution2000].
-I measure repressive capacity using the *physical integrity* index from V-Dem, lagged one year.
-Higher values for this variable indicate greater freedom from torture and political killing.
-I also include the variable's square term to account for the 'murder in the middle' framework.
-The physical integrity variables are expected to be explanatory in models of protest, but are important control variables in the models of election fraud.
-Since repressive capacity helps influence protest in expectation [@Ritter2016], incumbents that are more confident of deterring protest in advance should be more willing to engage in fraud under the protest-deterrent model of manipulation.
-
-Finally, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby.
-This diffusion mechanism is modeled using each country's geopolitical region as coded in Teorell et al [-@qog_teorell] in conjunction with binary variables indicating a successful protest movement from NELDA.
-A 'successful protest' is recorded if a protest movement resulted in either a new election being run or the incumbent stepping down.
-The *regional diffusion* dummy variable takes on a value of one for all countries in the same region as a successful protest, for elections within one calendar year of the successfully overturned election.^[For example, the variable takes on a positive value for elections held in the post-communist region, within one year of the September 24, 2000 election in Serbia that sparked the 'Bulldozer Revolution'.
-All told there are 15 successful electoral protests in the dataset: India 1989, Bulgaria 1990, Bangladesh 1996, Mali 1997, Lesotho 1998, Cote d'Ivoire 2000 and 2010, Peru 2000, Serbia 2000, Georgia 2003, Ukraine 2003-2004, Azerbaijan 2005, Kyrgyzstan 2005, Haiti 2006, and Cameroon 2007.]
-
-The models of protest quantity, size, and duration incorporate several potential explanatory variables, including the *election fraud* variable described above.
-The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance.
-The winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia and Psephos.} The deterrence model suggests that closer elections with more severe fraud should be more likely to generate protest.
-Finally, the theory proposed here suggests that structural conditions\textemdash in particular, the state's repressive capacity\textemdash should be more influential for protest dynamics.
-
-## Control variables
-
-Several control variables are included, which could be common causes for both the severity of election manipulation and protest.
-To help control for the underlying mobilizational capacity, which likely affects incumbents' decision to manipulate and the latent risk of protest, I include lagged indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), and the available of *alternative sources of information*.
-Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019].
-All models include measures of *judicial constraints* and  *legislative constraints* on the executive, V-Dem variables which captures the extent to which the judicial and legislative bodies in the country act independently in practice to limit executive power.
-To capture the overall level of pre-election protection for rights associated with liberal democracy, I include the lagged *liberal democracy index* from V-Dem.
-
-Finally, several binary control variables are included.
-First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest.
-Next, a variable from NELDA indicates if the chief executive is *term limited* during the election at hand.
-This variable is included since incumbents running up against constitutional term limits can pose major challenges for the unity of the regime, possibly leading to fractured manipulation efforts and post-election protest [@hale_patronal_2014].
-Finally, I include dummy variables that provide a categorical evaluation of state capacity.
-These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap).
-If the V-Dem data codes a state as not capable of raising revenue or primarily relying on loans and foreign aid, I code it as a *weak state*.
-If the state is described as primarily relying on resource rents and expropriation, I code it is a *rentier state*.
-States that rely primarily on taxes on property or economic transactions are coded as *tax states*.
-This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018].
-Lastly, since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.
-Table \@ref(tab:summary-table) shows summary statistics for all of these variables.
-To address concerns about multicollinearity, I present models in the appendix (page 25) that exclude control variables associated with pre-election protest.
-
-# Results
-
-## Pre-election protest and election-day fraud
-
-The results of these models are not promising for argument that protest risk deters election manipulation.
-First, Hypothesis 1 is not supported in the models shown in Table  \@ref(tab:manipdv).
-Models 1 through 5 apply different measures of pre-election protest risk; all of them behave contrary to the standard expectation in the literature.
-In Model 1, the number of pre-election protests is associated with an increase in election-day fraud; this effect is unchanged in Model 5, where *liberal democracy index* replaces other indicators of political openness.
-the coefficient on *pre-election protest* is positive and significant.
-While the effect is small, it runs counter to the expectations of the deterrence logic.
-A similar effect is found for *diffusion*; recent successful protest in a nearby country is also associated with a small increase in expected fraud.
-This suggests that the level of election manipulation, rather than being deterred by a highly contentious environment, may be deployed to make up for anticipated poor election results.
-
-Model 2 indicates that fraud declines as societies become less repressive; an intuitive finding.
-However, Model 3 shows that fraud first increases with physical integrity, but then declines.
-In other words, fraud is lowest on the extremes of repressive capacity\textemdash where repression is severe and where it is most limited.
-This is contrary to the expectation that a higher risk of protest is associated with reduced fraud.
-
 \newpage
@@
 ```{r manipdv, echo=FALSE}
-m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  +  unscheduled_election  + (1 | stateid), data = nelda2.sub)
+m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  +  unscheduled_election + state.owned.1lag + patronage.appts.1lag + antisys.1lag + v2elembaut + (1 | stateid), data = nelda2.sub)
 
@@
 
-m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + (1 | stateid), data = nelda2.sub)
+m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag + antisys.1lag + regional_success_lag +  v2elembaut + (1 | stateid), data = nelda2.sub)
 #summary(m1a)
 
-m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag +  unscheduled_election + (1 | stateid), data = nelda2.sub)
+m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag + antisys.1lag + unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)
 
@@
 
-m1c <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + unscheduled_election  + (1 | stateid), data = nelda2.sub)
+m1c <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag + antisys.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub)
 #summary(m1c)
 
-m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  +  unscheduled_election  + (1 | stateid), data = nelda2.sub)
+m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + antisys.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub)
 
@@
 
-m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + (1 | stateid), data = nelda2.sub)
+m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag + antisys.1lag + regional_success_lag + v2elembaut +  (1 | stateid), data = nelda2.sub)
 #summary(m1a)
 
-m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag +  unscheduled_election + (1 | stateid), data = nelda2.sub)
+m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag + antisys.1lag +  unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)
 
@@
 
-m1d <- lmer(v2elirreg.inv ~  libdem.1lag + gdpgro.1lag +    presidential +  urban.pct + rentierstate + taxstate  +  unscheduled_election + n.events.pre  + (1 | stateid), data = nelda2.sub) 
+m1d <- lmer(v2elirreg.inv ~  libdem.1lag + gdpgro.1lag +    presidential +  urban.pct + rentierstate + taxstate  +  unscheduled_election + n.events.pre  + state.owned.1lag + patronage.appts.1lag + antisys.1lag + v2elembaut +  (1 | stateid), data = nelda2.sub)
 #summary(m1d)
 
-modelsummary(c(m1, m1b, m1c, m1a, m1d), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Table 2: Random-intercept multilevel model of election-day fraud', coef_omit = "SD", coef_rename =
+modelsummary(c(m1, m1b, m1c, m1a), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Table 2: Random-intercept multilevel model of election-day fraud', coef_omit = "SD", coef_rename =
                c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
@@
 integrity index (lag)",
-                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion"
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity"
                  )) %>% kable_styling(latex_options = "HOLD_position")
 
+##Note: state owned variable is inverse coded, so should be labeled as private ownership in economy
+##Note: Patronage appointments ialso inverse coded. Higher values = more professionalism
 ```
@@
 models_t1 <- list()
-models_t1[['Model 6']] <- model.n.base.hurd.mm
-models_t1[['Model 7']] <- model.n.base2.hurd.mm
-models_t1[['Model 8']] <- model.n.griev.hurd.mm
+models_t1[['Model 5']] <- model.n.base.hurd.mm
+models_t1[['Model 6']] <- model.n.base2.hurd.mm
+models_t1[['Model 7']] <- model.n.griev.hurd.mm
 #models_t1[['Model 5']] <- m_duration_count
@@
             "Urban" = "urban.pct", "Rentier state" = "rentierstate1", "Tax state" = "taxstate1"),
-              statistics = character(0))  %>% set_caption("Table 3: Zero-inflated models of protest initiation")
+              statistics = character(0))  %>% set_caption("Table 3: Zero-inflated models of no protest initiation")
 
@@
 
-t1b <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, coefs = c("Number of pre-election protests" = "n.events.pre", "Off-schedule election" = "unscheduled_election1", "Regional protest diffusion" = "regional_success_lag", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd", "Election fraud : GDP growth (lag)" = "gdpgro.1lag:v2elirreg.inv")) %>% set_caption("Table 3 cont.: Zero-inflated models of protest initiation")
+t1b <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, coefs = c("Number of pre-election protests" = "n.events.pre", "Off-schedule election" = "unscheduled_election1", "Regional protest diffusion" = "regional_success_lag", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd", "Election fraud : GDP growth (lag)" = "gdpgro.1lag:v2elirreg.inv")) %>% set_caption("Table 3 cont.: Zero-inflated models of no protest initiation")
+
+t1b <- huxtable::add_footnote(t1b, "Note: Negative coefficients indicate protest is more likely.")
 
@@
 models_t2 <- list()
-models_t2[['Model 6']] <- model.n.base.hurd.mm
-models_t2[['Model 7']] <- model.n.base2.hurd.mm
-models_t2[['Model 8']] <- model.n.griev.hurd.mm
-models_t2[['Model 9']] <- model.size.base
-models_t2[['Model 10']] <- model.size.base2
-models_t2[['Model 11']] <- model.size.griev
+models_t2[['Model 5']] <- model.n.base.hurd.mm
+models_t2[['Model 6']] <- model.n.base2.hurd.mm
+models_t2[['Model 7']] <- model.n.griev.hurd.mm
+models_t2[['Model 8']] <- model.size.base
+models_t2[['Model 9']] <- model.size.base2
+models_t2[['Model 10']] <- model.size.griev
 
@@
 The second set of models challenge the idea that more severe electoral manipulation is associated with an increased risk of anti-government protest.
-Beginning with the zero-inflation portion of the hurdle models in Table 3, which model the likelihood of *no* protest occurring after the election, the standalone coefficient for election fraud is not significant in Model 6.
+Beginning with the zero-inflation portion of the hurdle models in Table 3, which model the likelihood of *no* protest occurring after the election, the standalone coefficient for election fraud is not significant in Model 5.
 Figure \@ref(fig:margin-plot-gdp) shows that there are no significant effect across values of *lagged GDP growth*.
@@
 Here, typical values for low and high levels of fraud are used (one standard deviation below and above the mean, respectively).
-The top panel shows that more fraudulent elections are associated with a significantly higher risk of protest initiation, but only when winning vote-shares are very small.
+The top panel shows that more fraudulent elections are associated with a significantly higher risk of protest initiation, but only when winning vote-shares are extremely small.
 
@@
 
-plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary_n, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]"))
+plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary_n, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-1.5, 1.5]"), show.data = T)
 plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw()
 
-plot_binary_margin_typ_size <- sjPlot::plot_model(model.margin.binary_size, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]"))
-plot_binary_margin_typ_size <- plot_binary_margin_typ_size +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "", color = "Election fraud") + theme_bw()
-
 ###GDP
 
-plot_binary_gdp_typ_n <- sjPlot::plot_model(model.binary.griev1_n, type = "pred", terms = c("gdpgro.1lag [all]", "v2elirreg.inv [-.45, 1.78]"))
-plot_binary_gdp_typ_n <- plot_binary_gdp_typ_n +  labs(x = "GDP growth (lagged)", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw() + coord_cartesian(xlim = c(-.4, 3))
-
-plot_binary_gdp_typ_size <- sjPlot::plot_model(model.binary.griev1_size, type = "pred", terms = c("gdpgro.1lag [all]", "v2elirreg.inv [-.45, 1.78]"))
-plot_binary_gdp_typ_size <- plot_binary_gdp_typ_size +  labs(x = "GDP growth (lagged)", y = "Predicted probability", title = "Predicted probability of any size protest", color = "Election fraud") + theme_bw()
+plot_binary_gdp_typ_n <- sjPlot::plot_model(model.binary.griev1_n, type = "pred", terms = c("v2elirreg.inv", "gdpgro.1lag [-0.04, .08]"))
+plot_binary_gdp_typ_n <- plot_binary_gdp_typ_n +  labs(x = "Election fraud", y = "Predicted probability", title = "Predicted probability of no protest", color = "GDP growth (lagged)") + theme_bw()
 
@@
 
-plot_binary_repr_typ_n <- sjPlot::plot_model(model.binary.repr_n, type = "pred", terms = c("physinteg.1lag [all]", "v2elirreg.inv [-.45, 1.78]"))
-plot_binary_repr_typ_n <- plot_binary_repr_typ_n +  labs(x = "Physical integrity (lagged)", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()
-
-```
+plot_binary_repr_typ_n <- sjPlot::plot_model(model.binary.repr_n, type = "pred", terms = c("v2elirreg.inv", "physinteg.1lag [.25, .61]"))
+plot_binary_repr_typ_n <- plot_binary_repr_typ_n +  labs(x = "Election fraud", y = "Predicted probability", title = "Predicted probability of any protest", color = "Physical integrity (lagged)") + theme_bw()
+
+```
+
 ```{r margin-plot-gdp, echo=FALSE, message=F, warning=FALSE, fig.cap="Marginal effect of election fraud on protest initiation, conditional on lagged GDP growth"}
@@
 ```{r n-protests-margin, fig.cap="Fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "count") 
+dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-1.5, 1.5]"), type = "count") 
 p.n.margin <- plot(dat)
@@
 
-## Election-day fraud and number of post-election protests
+## The number and size of post-election protests
 
@@
 
-Turning to the models of number and size of protest events, conditional on protest initiation, shown in Table 4.
-Models 6 and 9\textemdash which includes no interaction terms\textemdash show no significant relationship between fraud and and the number or maximum size of protest events.
+Results from the models of number and size of protest events, conditional on protest initiation, are shown in Table 4.
+Models 5 and 8 show no significant relationship between fraud and and the number or maximum size of protest events.
 The bottom panel of Figure \@ref(fig:margin-prob-fig) shows the marginal effect of election fraud on the number of protests, given varying levels of winning vote-share, holding other variables constant at their means (continuous variables) or modes (categorical variables).
 As the figure shows, there is no statistically significant difference in the expected number of protests.
-No significant effects are found for fraud and GDP growth in Models 8 or 11.
-Nor is there a significant effect for fraud on protest size, conditional on margin of victory, as discussed below.
-
-Altogether, there scanty evidence to support the argument that protest risk deters election fraud.
-Models 1 through 5 show that various measures of pre-election protest risk are not associated with reduced fraud\textemdash in fact, some appear to increase the severity of fraud.
+No significant effects are found for fraud and GDP growth in Models 7 or 10.
+Nor is there a significant effect for fraud on protest size, conditional on margin of victory.
+However, Hypothesis 6 is consistently supported\textemdash incumbent term limits do seem associated with increased size and number of post-election protests.
+
+Altogether, there is scant evidence to support the argument that protest risk deters election fraud.
+Models 1 through 4 show that various measures of pre-election protest risk are not associated with reduced fraud\textemdash in fact, some appear to increase the severity of fraud.
 Models 6 through 11 show that higher levels of fraud is not "punished" by more or larger protests.
-Hypothesis 2 is also weakly supported; there is a significant and sizable association between election fraud and protest initiation when the ruling party's winning margin is low, but only at extremely low winning margins.
-However, even this suppotive finding likely presents too optimistic a picture for the protest-deterrent model of election fraud.
+Hypothesis 2 is supported, but very weakly; there is a significant association between election fraud and protest initiation, but only when the incumbent's winning vote-share is near its minimum.
+However, even this supportive finding likely presents too optimistic a picture for the protest-deterrent model of election fraud.
 The number and size results from the hurdle regressions show no effect for election fraud across any of the models, alone or in interaction.
@@
 An even more troubling implication would suggest that the protest-manipulation logic, rather than serving as a deterrent, creates perverse incentives to manipulate as much as possible.
-If ruling parties do indeed seek to avoid protest, but manipulating heavily does not come with a cost in the number of post-election protests, ruling parties have an incentive fraudulently inflate their official margins of victory in order to reduce the risk of protest initiation.
+If ruling parties do indeed seek to avoid protest initiation, but manipulating heavily does not entail more severe protest, ruling parties have an incentive fraudulently inflate their official margins of victory in order to reduce the risk of protest initiation.
+
+```{r pred-size, fig.cap="Fraud, winning vote-share, and predicted number of protests", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
+
+p.protestlow <- sjPlot::plot_model(model.n.base2.hurd.mm, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(n.events.pre = 0)) + labs(x = "Incumbent vote-share", y = "", title = "Predicted n of post-election protests (Pre-\nelection protest: 5th percentile)") + coord_cartesian(ylim = c(0, 20))
+
+p.protesthigh <- sjPlot::plot_model(model.n.base2.hurd.mm, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(n.events.pre = 8)) + labs(x = "Incumbent vote-share", y = "", title = "Predicted n of post-election protests (Pre-\nelection protest: 95th percentile)") + coord_cartesian(ylim = c(0, 20))
+
+cowplot::plot_grid(p.protestlow, p.protesthigh, nrow = 1)
+```
 
@@
 
-p.gdplow <- sjPlot::plot_model(model.size.base2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(gdpgro.1lag = -.5)) + labs(x = "Incumbent vote-share", y = "Max.
+p.protestlow <- sjPlot::plot_model(model.size.base2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(n.events.pre = 0)) + labs(x = "Incumbent vote-share", y = "Max.
 size", title = "Predicted max.
-size \n(GDP growth: -0.5)")
-
-p.gdphigh <- sjPlot::plot_model(model.size.base2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(gdpgro.1lag = 3)) + labs(x = "Incumbent vote-share", y = "", title = "Predicted max.
-size \n(GDP growth: 3)")
-
-cowplot::plot_grid(p.gdplow, p.gdphigh, nrow = 1)
+size (Pre-\nelection protest: 5th percentile)") + coord_cartesian(ylim = c(0, 8000))
+
+p.protesthigh <- sjPlot::plot_model(model.size.base2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(n.events.pre = 8)) + labs(x = "Incumbent vote-share", y = "", title = "Predicted max.
+size (Pre-\nelection protest: 95th percentile)") + coord_cartesian(ylim = c(0, 8000))
+
+cowplot::plot_grid(p.protestlow, p.protesthigh, nrow = 1)
 ```
@@
 
-One objection to this framing may be that, if the level of manipulation and the risk of protest are endogenous to one another, perhaps the null result found for the number and size of protests is the result of equilibrium behavior predicted in many formal models.
+One objection to this framing could be: perhaps the null result found for the number and size of protests is the result of equilibrium behavior predicted in many formal models.
 The threat of protest keeps fraud lower than it would be otherwise, resulting in limited (if any) protest.
@@
 
-First, such an objection does not account for the unexpected results found in Models 1 through 5, where indicators of pre-election protest risk were found to be associated with more severe election-day fraud.
+First, such an objection does not account for the unexpected results found in Models 1 through 4, where indicators of pre-election protest risk were found to be associated with more severe election-day fraud.
 Second, deterrence models would expect the equilibrium level of fraud to fall as protest risk increases, and for more severe protest to occur as a result of miscalculation by the regime.
@@
 If the severity of fraud is held constant at a high level while underlying protest risk increases, we should see an increase in the predicted protest outcome.
-Taking protest size as an example, using Model 10, Figure \@ref(fig:pred-size) shows there is no evidence of such an effect.
+Taking protest size as an example, using Model 9, Figure \@ref(fig:pred-size) shows there is no evidence of such an effect.
 Both panels show the marginal effect of a relatively fraudulent election (*election fraud* is set to one standard deviation above the mean) on the predicted maximum protest size, conditional on the closeness of the election result.
-All control variables have been set to their means or modes with the exception of lagged GDP growth rate.
-In the left panel, GDP growth has been set to -0.5, close to its minimum value; on the right, it has been set to 3.
-An equilibrium approach would suggest that high fraud is more dangerous in the troubled economic environment captured on the left, especially at low levels of *winner's vote-share*.
-However, we see practically no difference in the predicted maximum protest size across the different economic conditions modeled here\textemdash changing economic conditions do not appear to affect the 'appropriate' level of fraud, throwing cold water on the equilibrium argument.
+All control variables have been set to their means or modes with the exception of *pre-election protest*.
+In the left panel, the quantity of pre-election protest has been set to the fifth percentile for that variable (0 events), on the right, it has been set to the 95th percentile (8 events).
+An equilibrium approach would suggest that high fraud is more dangerous in the more active protest envionment captured on the right, especially at low levels of *winner's vote-share*.
+However, we see practically no difference in the predicted maximum protest size across the different conditions modeled here\textemdash changing protest conditions do not appear to affect the 'appropriate' level of fraud, throwing cold water on the equilibrium argument.
 
@@
 
-cox1 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + v2elirreg.inv , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+cox1 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + n.events.post + unscheduled_election + regional_success_lag + v2elirreg.inv , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
 summary(cox1)
@@
 print(cz1)
-plot(cz1[8])
+plot(cz1[19])
 abline(h=0)
-abline(v = 11)
-abline(v = 25)
-abline(v = 90)
-
-cox1t <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + v2elirreg.inv + tt(v2elirreg.inv), data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(11, 25, 90),
+abline(v = 21)
+abline(v = 61)
+
+cox1t <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + n.events.post + regional_success_lag + v2elirreg.inv + tt(v2elirreg.inv), data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(21, 61, 90),
 Boundary.knots = FALSE))
@@
 
-cz <- cox.zph(cox1, transform = function(time) 1/(log(time + 1)))
+cz <- cox.zph(cox1, transform = "identity") #function(time) 1/(log(time + 1)))
 print(cz)
@@
 sandbox_data <- survSplit(Surv(d.events.num, surv_status) ~ 
-                      jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + fraud_level  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+                      jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + fraud_level + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                       data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), cut = c(2), episode = "tgroup")
 
-cox1.2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + fraud_level*(as.factor(tgroup))  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = sandbox_data)
+cox1.2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + fraud_level*(as.factor(tgroup)) + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = sandbox_data)
 summary(cox1.2)
 
-cox2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+cox2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
 summary(cox2)
+sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.15, 2.27]")) + coord_cartesian(ylim = c(0, 10)) #This is the 10th and 90th quantiles
 
@@
 
-This dataset can be used for a test of the idea that election fraud is not informative about the duration of protest outside of an initial window.
-To do so, I employ Cox proportional-hazards models of protest-wave duration after subsetting the data to those cases where protests are initiated but incumbents survived; in this sense, they represent the cases where opposition groups were most optimistic about challenging the government (or most aggrieved), but ultimately were overmatched.
-In Table 5, coefficients with a negative sign indicate factors that are associated with longer-lived protest movements; those with a positive sign indicate factors associated with a speedier demise.
-
-In Model 12, all covariates' effects are modeled as time-invariant.
-However, diagnostics of the model suggest that this assumption may not be upheld for *election fraud*.^[Schoenfeld residuals for *election fraud* are not independent of time, as calculated using the `survival` package in R [@survival-package].] To adjust for this, Model 13 allows the coefficient for *election fraud* to take on different values at different points in time.
+```{r}
+library(survsim) #Looking at time invariance
+zph <- cox.zph(cox1)
+zph
+
+plot(zph[19], lwd =2)
+abline(0,0, col=1, lty = 3, lwd = 2)
+abline(h = cox1$coefficients[19], col=3, lwd=2, lty=2)
+```
+
+To further test the mechanism outlined in the theory above, I examine the relationship between fraud and the survival of protest movements in cases where protests began but the incumbent eventually survived.
+Restricting the sample to these cases allows for a consistent interpretation of results: longer protest waves indicate a more even match between the resources and resiliency of the protesters and the capacity and cohesiveness of the regime, while shorter-lived protest movements are those which were more easily defeated.
+In this sub-sample, the relationship between election fraud and protest survival should be non-linear with respect to time\textemdash the longer the protest movement survives, the less relevant the fraud signal (such as it is) becomes.
+That is, any effect of fraud on protest should decline in size and and become statistically insignificant as the election recedes.
+
+Table 5 reports the results of Cox proportional hazard models of protest duration; coefficients with a negative sign indicate factors that are associated with longer-lived protest movements; those with a positive sign indicate factors associated with a speedier demise.
+
+In Model 11, all covariates' effects are modeled as time-invariant.
+However, diagnostics of the model suggest that this assumption may not be upheld for *election fraud*.^[Schoenfeld residuals for *election fraud* are not independent of time, as calculated using the `survival` package in R [@survival-package].] To adjust for this, Model 12 allows the coefficient for *election fraud* to take on different values at different points in time.
 Based on the diagnostics, I allow the value of the coefficient to vary for three periods: 1 - 11 days, 12 - 25 days, and 26 days to the maximum value.
-Model 13 shows the result of this analysis; in it, *election fraud* has a positive and significant coefficient for the earliest time period.
+Model 12 shows the result of this analysis; in it, *election fraud* has a positive and significant coefficient for the earliest time period.
 This indicates that, for the first 11 days of a protest movement, states that engaged in more election fraud are more likely to be able to quash the protests (whether by repression or accommodation).
+This is contrary to the expectations of deterrence models.
 After this period, there is no significant relationship between the level of fraud and the survival of the protest wave.
-That is, if a protest movement has survived longer than (roughly) a week and a half after the election, the severity of fraud does not influence its duration.
-It is also noteworthy that this formulation is a much better fit to the data than the linear effect in Model 12, with a much lower root mean square error.
-
-These results indicate, first, that the influence of a fraudulent election on the duration of mass protest diminishes over time.
+That is, if a protest movement has survived longer than (roughly) a week and a half after the election, the severity of fraud does not influence its continued duration.
+It is also noteworthy that this formulation is a much better fit to the data than the linear effect in Model 11, with a much lower root mean square error.
+
+These results indicate that the influence of a fraudulent election on the duration of mass protest diminishes over time.
 It also suggests that most protests are failed bets.
@@
 models3 <- list()
-models3[['Model 12']] <- cox1
-models3[['Model 13']] <- cox1t
+models3[['Model 11']] <- cox1
+models3[['Model 12']] <- cox1t
 #models3[['Model 14']] <- cox3
@@
                  "osorg.1lag" = "Civil soc. openness (lag)", 
+                 "winner.share.upd" = "Incumbent vote-share",
                  "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
@@
                  "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "v2elirreg.inv" = "Election fraud",
-                 "tt(v2elirreg.inv)1" = "Election fraud (day 12-25)", "tt(v2elirreg.inv)2" = "Election fraud (day 26-90)"
+                 "tt(v2elirreg.inv)1" = "Election fraud (day 2-20)", "tt(v2elirreg.inv)2" = "Election fraud (day 21-90)"
                  ))
@@
 
-Altogether, the findings challenge the idea that the fear of protest will help deter election-manipulation by ruling parties.
-To the extent that fraud is a signal of incumbent strength, it is a noisy and unreliable one.
-When an incumbent party wins by a narrow margin in a highly fraudulent election, the likelihood of protest initiation is somewhat higher than when it wins more cleanly.
-However, since even strong incumbents commit fraud, the decision to take to the streets is often a losing gamble\textemdash anticipating a weak incumbent, they instead find one that is capable of repressing, coopting, or conciliating an emerging protest movement.
-Indeed, while there are 194 cases of protest out of the 661 elections in this dataset where incumbents claimed victory, only 8 resulted in either the incumbent being replaced or the election being re-run due to mass protests (according to NELDA).\footnote{This information is determined using the NELDA variables NELDA37 and NELDA41.
-The cases of successful protest are: Ukraine 2004, Azerbaijan 2005, Georgia 2003, Kyrgyzstan 2005, Malaysia 2004, Bangladesh 1996, Peru 2000, and Cote d'Ivoire 2000.} This suggests a `success rate' for post-election protest of about four percent *conditional on protest initiation*, and about one percent for all elections in electoral authoritarian and electoral democratic regimes.
-
-The results indicate that we should be cautious when building and interpreting theories of non-democratic politics that emphasize the risk of mass protest as a trip-wire to be avoided, and add empirical evidence to Przeworski's [-@przeworskiFormalModelsAuthoritarian2022] critique of such models.
+Altogether, the findings challenge the idea that the fear of protest helps deter election manipulation.
+To the extent that fraud is a signal of incumbent strength, it is an unreliable one.
+Since even strong incumbents commit fraud, the decision to take to the streets is often a losing gamble\textemdash anticipating a weak incumbent, they instead find one that is capable of repressing, coopting, or conciliating an emerging protest movement.
+Indeed, while there are 194 cases of protest out of the 661 elections in this dataset where incumbents claimed victory, only 8 resulted in either the incumbent being replaced or the election being re-run due to mass protests.\footnote{This information is determined using the NELDA variables NELDA37 and NELDA41.} This suggests a `success rate' for post-election protest of about four percent *conditional on protest initiation*, and about one percent for all elections in electoral authoritarian and electoral democratic regimes.
+
+The results indicate that we should be cautious when building and interpreting theories of politics outside liberal democratic regimes that emphasize the risk of mass protest as a trip-wire to be avoided, and add empirical evidence to Przeworski's [-@przeworskiFormalModelsAuthoritarian2022] critique of such models.
 The rarity of post-election protest, especially successful protest, suggests that such events are highly contingent; they hinge on a multitude of factors that go beyond the regime's efforts to bias election outcomes.
@@
 The results presented here show that fraud has no bearing on how numerous or large protest events are.
-While pre-election protest is associated with reduced fraud, the substantive size of this relationship is very small.
-Moreover, contrary to some prior research, the risk of a large protest wave does not appear to be heightened by poor economic conditions [@Brancati2014a].
-As a result, circumstances that are to some extent outside of leaders' control, like a worsening macroeconomic climate, do not appear to make engaging in fraud any riskier.
-Instead, the conditions that are correlated with more damaging protest are generally institutional factors, which are either more amenable to control or can be anticipated by non-democratic incumbents.
-Legislative constraints, an important predictor of the number of protests in all four models, can be adjusted through institutional reforms and cooptation of the opposition over time.
-In other words, election fraud only has a minor cost in terms of protest risk, a risk which can be further blunted to the extent that elections can often be held on incumbents' terms.
+Moreover, heightened protest risk appears to be associated with *more* election fraud, rather than less.
+Instead, the conditions that are correlated with more damaging protest are generally institutional factors, especially executive term limits, which are either more amenable to control or can be anticipated by incumbents.
 
@@
 That electoral protest is less risky than is often assumed, in turn, implies that governments and ruling parties have a freer hand to engage in electoral manipulation than many models posit.
-Since the severity of fraud has no apparent bearing on whether or not a major protest wave will emerge as a serious threat to the regime, as shown in Figure 3, government have an incentive to manipulate as much as possible to secure a large winning vote-share.
+Since the severity of fraud has no apparent bearing on whether or not a major protest wave will emerge as a serious threat to the regime, government have an incentive to manipulate as much as possible to secure a large winning vote-share.
 This perverse incentive undermines the case that free and fair elections in an electoral democracy represent a self-enforcing equilibrium backed up by the risk of citizen protest [@fearon_self-enforcing_2011], or that risk of popular mobilization can substantially moderate election manipulation in non-democracies.
@@
 This study also suggests implications for the study of election manipulation.
-First, it is clear that electoral protests charged by allegations of manipulation do occur at times.
-While the main argument of this paper is that the overall level of fraud does not predict these protests, there may still be connections between manipulation and protest to uncover.
-For example, while fraud was used as the main variable in this study, it could be that other forms of manipulation--perhaps intimidation or violence--are associated with higher risk of escalating protest.
-Perhaps some forms of electoral manipulation are particularly offensive to certain groups of people\textemdash young people, highly educated people, people with ideological commitments to democracy, etc.
-To the extent that countries vary according to the size of such populations, they may vary in their underlying protest risk conditional on the type of manipulation employed.
-Second, the study implies that supply-side theories of electoral manipulation\textemdash such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]\textemdash may be more important than protest risk for understanding election integrity, which may drive future research as well as policy interventions.
+First, the study implies that supply-side theories of electoral manipulation\textemdash such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]\textemdash may be more important than protest risk for understanding election integrity, which may drive future research as well as policy interventions.
 Third, and relatedly, it undermines some current understandings of the role of election monitors, courts, and other third parties in upholding election integrity.
@@
 
+Second, it suggests avenues for future research.
+It is clear that electoral protests charged by allegations of manipulation do occur at times.
+While the main argument of this paper is that election fraud itself does not predict these protests, there may still be connections between manipulation and protest to uncover.
+For example, it could be that other forms of manipulation\textemdash perhaps intimidation or violence\textemdash are associated with higher risk of escalating protest.
+Perhaps some forms of electoral manipulation are particularly offensive to certain groups\textemdash young people, highly educated people, people with ideological commitments to democracy, etc.
+To the extent that countries vary according to the size of such populations, they may vary in their underlying protest risk.
+
+As a final point, the results also speak to incidents of mobilization around unfounded allegations of election-rigging, such as those of January 6, 2020 in the United States or 2022-2023 in Brazil.
+Indeed, out of the 293 instances of post-election protest in this dataset, 18 occurred after very clean elections (those one standard deviation below the mean for election fraud)\textemdash a small but not minuscule percentage.
+These incidents demonstrate that, while crying fraud may sometimes be a useful tool for protest entrepreneurs, losing an election alone can be a catalyst for protest, irrespective of the actual level of election manipulation.
+This possibility tends to reduce the 'opportunity cost' to incumbents of engaging in real fraud, while also underlining the fact that the initiation and course of protest is often driven more by the relative resources and opportunities available to the disaffected and the incumbent government.
+
 These results are consistent with different specifications presented in an online appendix.
@@
 
-This paper addresses these concerns by testing the relationship between election fraud, the risk that any protest occurs, and the number of protest events that occur after an election.
-Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 1110 elections to show that---while there is an elevated risk of some protest occurring in a fraudulent election with narrow margins of victory or in highly corrupt societies---election fraud has no significant relationship with the number of protests that occur in a protest wave.
-These results imply that protest risk should not be considered a central deterrent to election fraud.
+This paper addresses these concerns by testing the relationship between election fraud and the number and maximum size of protest events that occur after an election.
+Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 1110 elections to show that election fraud has no significant relationship with the number or size of protests that occur in a protest wave.
+Additional tests suggest that this null finding is not the result of equilibrium behavior predicted in deterrence models.
 Instead, most fraudulent elections go un-protested, and most post-election protests appear to be strategic gambles gone wrong.
 The initiation of protest more often reveals the incumbent to be strong enough to manage protest through repression, cooptation, or other concessions.
+The results suggest that theories which treat protest risk as an unqualified deterrent to election manipulation should be reconsidered.
+
 The absence of protest risk as a guardrail is felt in both electoral authoritarian regimes and in electoral democracies\textemdash those democracies that are most at risk of democratic erosion.
@@
 
+```{r, include=FALSE}
+ggplot(data=nelda2.sub, aes(x=v2elirreg.inv, y = jitter(any.ecav.protest), color = as.factor(v2x_regime))) +
+  geom_point()
+
+```
+
+This is an attempt at modeling endogenous relationships using two-step residual inclusion regression:
+
+```{r}
+nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(n.events.post > 0, 1, 0))
+
+test_model <- OneSampleMR::tsri(any_protest ~ v2elirreg.inv | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +   osorg.1lag  + natlpartyorg.lag  +  presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = nelda2.sub, link = "logit")
+summary(test_model)
+
+test_model2 <- OneSampleMR::tsri(any_protest ~ winner.share.upd | v2elirreg.inv +  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +   osorg.1lag  + natlpartyorg.lag  +  presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, data = nelda2.sub, link = "logit")
+summary(test_model2)
+
+```
+
+Predicting post-election protest counts and getting residuals:
+
+```{r}
+test3 <- glm.nb(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag +  + osorg.1lag  + natlpartyorg.lag +   inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag, data = nelda2.sub)
+summary(test3)
+
+ggplot(mapping = aes(x=predict(test3, type = "response"), y= test3$model$n.events.post)) + geom_point() + coord_cartesian(xlim=c(0, 500))
+
+newdata <- modelr::add_predictions(nelda2.sub, test3)
+newdata <- newdata %>% mutate(pred_error = n.events.post - pred)
+  #because this is set up so that the predicted value is subtracted from the actual, underpredicting
+  # protest is a positrive value (i.e. actual 5 - 0 predicted = 5)
+
+test4 <- lm(pred_error ~ winner.share.upd + v2elirreg.inv + presidential, data = newdata)
+summary(test4)
+
+test5 <- lm(pred_error ~ winner.share.upd*v2elirreg.inv + presidential, data = newdata)
+summary(test5)
+
+sjPlot::plot_model(test5, type = "pred", terms = c("v2elirreg.inv", "winner.share.upd [.4, .7]"))
+
+plot(newdata$v2elirreg.inv, newdata$pred_error)
+```
+
+Predicting the level of election manipulation, and then seeing if actual values above prediction are associated with protest
+
+```{r}
+nelda2.sub <- nelda2.sub %>% mutate(inc.confident.num = ifelse(inc.confident == "N/A", NA, ifelse(inc.confident == "yes", 1, 0)))
+
+###For testing missingness
+subsample <- nelda2.sub %>% dplyr::select(v2elirreg.inv, jucon.1lag, legcon.1lag, gdpgro.1lag,  natlpartyorg.lag,  presidential, urban.pct, rentierstate,  taxstate,  unscheduled_election, regional_success_lag, physinteg.1lag, antisys.1lag, csenv.1lag, preelection_concerns, inc.termlimit.num, liberal_no_jucon, v2elirreg.inv.lag, patronage.appts.1lag,
+                                                  state.owned.1lag)
+summary(subsample)
+############################ Note: Drop legcon from analysis because of high missingness and high cor with lib_no_jucon
+
+test6 <- lmer(v2elirreg.inv ~ jucon.1lag +  gdpgro.1lag +  natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + physinteg.1lag + antisys.1lag + csenv.1lag + preelection_concerns + inc.termlimit.num   + liberal_no_jucon + v2elirreg.inv.lag +
+                patronage.appts.1lag +  state.owned.1lag +  v2elembaut + (1 | stateid), data = nelda2.sub) 
+summary(test6)
+
+testdata <- broom.mixed::augment(test6)
+  #This adds fitted values and residuals.
+  #Residuals are actual values minus predicted value, so
+  #over-predicting fraud will result in a negative number (actual fraud < predicted )
+  #Under-predicting fraud--what we are interested in--will result in a positive number (actual > predicted)
+
+## Need to add rownames to nelda2.sub, so that can merge variables from nelda2.sub into testdata
+nelda2.sub <- nelda2.sub %>% mutate(.rownames = rownames(nelda2.sub))
+neldavars <- nelda2.sub %>% dplyr::select(.rownames, n.events.post, inc.lose, winner.share.upd, country, year, diem_explicheat_max, diem_fraud_max, country_id_year, surv_status,  inc.replaced.prot, d.events.num,
+                                          legcon.1lag, osorg.1lag, altinf.1lag, libdem.1lag,  winner.share.upd, n.events.pre)
+testdata <- testdata %>% left_join(neldavars, by = ".rownames")
+
+model_protest <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd + .resid  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + .resid*winner.share.upd,
+                               ziformula=~.,
+                              data = testdata %>% filter(inc.lose == 0),
+                                family=truncated_nbinom2)
+summary(model_protest)
+
+model_protest2 <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd + .resid  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+                               ziformula=~.,
+                              data = testdata %>% filter(inc.lose == 0),
+                                family=truncated_nbinom2)
+summary(model_protest2)
+
+plot(testdata$winner.share.upd, testdata$.resid)
+
+testdata <- testdata %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))
+testdata <- as.data.frame(testdata)
+model_protest_binary <- glm(no_protest ~ .resid*winner.share.upd, family = binomial(link = "logit"),
+                             data = testdata %>% filter(inc.lose == 0))
+plot_binary_margin_typ_n <- sjPlot::plot_model(model_protest_binary, type = "pred", terms = c(".resid", "winner.share.upd [.2, .8]"))
+plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Excess election fraud", y = "Predicted probability", title = "Predicted probability of no protest", color = "Incumbent vote-share") + theme_bw()
+
+model_protest2_binary <- glm(no_protest ~ .resid + winner.share.upd, family = binomial(link = "logit"),
+                             data = testdata %>% filter(inc.lose == 0))
+
+sjPlot::plot_model(model_protest2_binary, type = "pred", terms = c(".resid", "winner.share.upd [.2, .8]"), show.data = T)
+sjPlot::plot_model(model_protest2_binary, type = "est", vline.color = "black") + theme_bw()
+
+### Seeing how residuals compare to DIEM data
+
+summary(lm(.resid ~ diem_explicheat_max + diem_fraud_max, data = testdata))
+plot(testdata$diem_explicheat_max, testdata$.resid)
+plot(testdata$diem_fraud_max, testdata$.resid)
+  ##This shows that the resids are basically noise with regard to election observer reports, which should be good; I think this helps push back against the idea that V-Dem coders are giving higher scores to protest elections.
+But it would be good to push deeper and look at the correlation only among the most up-coded resids.
+
+high_resids <- testdata %>% filter(.resid >= .222)
+summary(lm(.resid ~ diem_explicheat_max + diem_fraud_max, data = high_resids))
+
+ggplot(data = testdata, aes(x = jitter(diem_fraud_max), y = .resid)) + geom_point(aes(size = n.events.post))
+
+plot(high_resids$diem_explicheat_max, high_resids$.resid)
+plot(high_resids$diem_fraud_max, high_resids$.resid)
+  #We do see more outliers here, good to think deeper--there are two major outliers that are also big protest events
+
+### Duration
+
+cox1_b <- coxph(Surv(d.events.num, surv_status) ~ winner.share.upd  + .resid, data = testdata %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+summary(cox1_b)
+
+## DIEM and protest
+
+model_protest_diem <- glmmTMB(n.events.post ~   winner.share.upd +  diem_explicheat_max ,
+                               ziformula=~.,
+                              data = nelda2.sub %>% filter(inc.lose == 0),
+                                family=truncated_nbinom2)
+summary(model_protest_diem)
+
+```
+
+```{r}
+mydf <- ggpredict(model_protest2, type = "count", terms = ".resid [all]", back.transform = T)
+ggplot(mydf, aes(x, predicted)) +
+  geom_line() +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .1) + coord_cartesian(ylim = c(0, 100))
+
+library(pscl)
+model_protest2b <- zeroinfl(n.events.post ~  .resid + winner.share.upd | ., data = testdata %>% dplyr::select(n.events.post, .resid, winner.share.upd), dist = "negbin")
+summary(model_protest2b)
+
+newdata1 <- expand.grid((seq(from = -1.2, to = 1.3, by = .1)), mean(testdata$winner.share.upd, na.rm=T))
+colnames(newdata1) <- c(".resid", "winner.share.upd")
+newdata1$phat <- predict(model_protest2b, newdata1)
+
+ggplot(newdata1, aes(x = .resid, y = phat)) +
+  geom_point() +
+  geom_line() +
+   labs(x = "Fraud relative to baseline", y = "Predicted protest events")
+
+model_protest2c <- zeroinfl(n.events.post ~  .resid*winner.share.upd + .resid + winner.share.upd  | ., data = testdata %>% dplyr::select(n.events.post, .resid, winner.share.upd), dist = "negbin")
+summary(model_protest2c)
+
+newdata2 <- expand.grid((seq(from = -1.2, to = 1.3, by = .1)), seq(from = 0, to = 1, by =.25))
+colnames(newdata1) <- c(".resid", "winner.share.upd")
+newdata1$phat <- predict(model_protest2c, newdata1)
+
+ggplot(newdata2, aes(x = .resid, y = phat, colour = factor(winner.share.upd))) +
+  geom_point() +
+  geom_line() +
+   labs(x = "Fraud relative to baseline", y = "Predicted protest events")
+
+cox1 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + v2elirreg.inv , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+summary(cox1)
+
+```
+
+```{r}
+nelda2.sub <- nelda2.sub %>% group_by(country) %>% mutate(elirreg_mean = mean(v2elirreg.inv, na.rm=T))
+nelda2.sub <- nelda2.sub %>% mutate(elirreg_mean_diff = v2elirreg.inv - elirreg_mean)
+
+model_protest1 <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd*elirreg_mean_diff  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+                               ziformula=~.,
+                              data =nelda2.sub %>% filter(inc.lose == 0),
+                                family=truncated_nbinom2)
+summary(model_protest1)
+sjPlot::plot_model(model_protest1, type ="pred", terms = c("elirreg_mean_diff" , "winner.share.upd [.2, .8]")) + coord_cartesian(ylim = c(0, 50))
+
+nelda2.sub <- nelda2.sub %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))
+model_protest_binary <- glm(no_protest ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd*elirreg_mean_diff  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+                             data =nelda2.sub %>% filter(inc.lose == 0),
+                                family=binomial(link = "logit"))
+summary(model_protest_binary)
+sjPlot::plot_model(model_protest_binary, type="pred", terms = c("elirreg_mean_diff" , "winner.share.upd [.2, .8]"))
+
+model_protest2 <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd + elirreg_mean_diff  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
+                               ziformula=~.,
+                              data =nelda2.sub %>% filter(inc.lose == 0),
+                                family=truncated_nbinom2)
+summary(model_protest2)
+
+```
+
+```{r}
+#Lots of missinginess in winner share
+
+winner_missing <- nelda2.sub %>% filter(is.na(winner.share.upd) == T) %>% dplyr::select(country, year)
+```
+
+What I'm looking for here is weak government -> high fraud + narrow vote share -> protest. Not sure if that pattern reveals itself.
+```{r}
+ggplot(nelda2.sub %>% filter(presidential == 1), aes(x = winner.share.upd, y = v2elirreg.inv)) + geom_point(aes(size=n.events.post)) + geom_vline(xintercept = .5, linetype = 2) + geom_hline(yintercept = 0)
+```
+
 \newpage

### 2024-01-19 Scripts/authoritarian_only.Rmd
Updates to intro

@@
 Lastly, a third school of thought holds that election fraud is an indicator of genuine incumbent strength.
-This approach, which is expanded upon below,
+This approach, which is expanded upon below, is a pessimistic account for democracy.
+If this model is correct, it is authoritarian politics rather than democracy that is self-enforcing\textemdash the ability to commit election fraud helps win elections, signal strength, deter challengers under most conditions, and put down protest movements if they do begin.
+This paper argues that the third model best fits the overall pattern of the data, and is supported by specific hypothesis tests.
 
-Events like those in Belarus are puzzling because many theories of election manipulation expect that incumbents will work hard to avoid them.
-A large, sustained protest movement creates a moment of heightened risk for the regime; the uncertainty of the outcome creates opportunities for elite defection, splits in the security services, and other negative outcomes.
-On some occasions, like Ukraine's Orange Revolution or the People Power Revolution in the Philippines, these protests can result in the fall of an incumbent and a political opening.
-For this reason, many scholars---especially those working in the formal modeling tradition---have identified the risk that election manipulation will spark protest as a major deterrent to regimes working to engineer election outcomes.
-Often, though, incumbents can rely on superior resources, elite solidarity, and repressive capacity to quash electoral protests.
-How threatening, then, are electoral protests really, and can that threat constrain electoral manipulation by governments?
+In this paper, I argue that\textemdash in keeping with the idea of 'informational autocracy' (ADD CITES HERE)\textemdash an election outcome produces a signal of the relative strength of the incumbent.
+As Magaloni (2008) and others have noted, incumbents typically work to produce a higher margin of victory in the election, in order to demonstrate their strength and induce cooperation from other actors in society.
+In some theories, this only works if the votes for the incumbent are largely 'genuine' (i.e.
+not fabricated), or produced using the incumbent's organizational resources (i.e.
+vote-buying and intimidation).
+(ADD CITES HERE, INCLUDING TO HARVEY AND M).
+These two kinds of votes correspond to two of the three 'pillars of stability' identified by Gerschewski (2013) (ADD CITE): legitimation and cooptation.
+Fraud is considered a form of cheap talk in these models, which opposition actors can discount if they become aware of it.
 
-In this paper, I argue that protest risk rarely serves as a deterrent to election manipulation for two reasons.
-First, election fraud is not only carried out by the weak types assumed in some formal models, but also by strong, cohesive ruling parties.
-The ambiguity of the electoral environment outside liberal democracies, where political and civil liberties may not be fully protected, makes it difficult for opposition actors to correctly distinguish between the two.
-As a result, we observe a high proportion of 'failed' protest movements, and\textemdash most likely\textemdash instances where weak governments go un-protested.
-Second, if electoral protests do materialize, their success or failure quickly becomes untethered from the severity of manipulation itself.
-Instead, the noisy election-day signal of regime strength is replaced by much clearer signals in the contest with protesters.
-Consequently, there is little penalty for severe manipulation in terms of protest risk.
+However, these models overlook that engineering widespread fraud requires control over patronage resources and the criminal and judicial apparatus (ADD CITES to HARVEY).
+As a result, they overlook the fact that high levels of fraud are a sign of incumbent strength in cooptation and Gerschewski's third pillar: repression.
+Both of these tools can be useful for defusing or dismantling an emergent protest movement.
+As a result, effective fraudsters should also be effective protest-crushers.
+This helps explain the broad pattern, where most fraudulent elections go unprotested, and among those that do generate protest, the vast majority fail to dislodge the incumbent.
+Consequently, there is little penalty for severe manipulation in terms of protest risk, and little deterrent effect.
 
-Testing this proposition requires looking beyond binary measures of protest, and toward measures of its scope.
+Testing this proposition in more depth is difficult.
+requires looking beyond binary measures of protest, and toward measures of its scope.
 Research on the risk posed to governments by electoral protest often acknowledges---whether explicitly or implicitly---the importance of broad-based protest campaigns, rather than one-off events [@Bunce2006;@tucker2007;@DAnieri2006;@Carnaghan2016;@bunce2010].

### 2024-01-19 Scripts/authoritarian_only.Rmd
intro updates

@@
 ```
+Duration as zero-inflated count model (i.e. count of days that protest movement lasted)
+```{r}
+model.d.base <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv +  +   v2elirreg.inv + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+                                ziformula=~.,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
+                                family=truncated_nbinom2)
+summary(model.d.base)
+
+model.d.base2 <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + gdpgro.1lag  +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+                                ziformula=~.,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
+                                family=truncated_nbinom2)
+summary(model.d.base2)
+
+sjPlot::plot_model(model.d.base2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-1, 1.5]"))
+
+mydf <- ggpredict(model.d.base2, type = "re.zi", terms = c("winner.share.upd [.2, .3, .4, .5, .6, .7, .8, .9]", "v2elirreg.inv [-.25, 1.5]"))
+ggplot(mydf, aes(x, predicted, color = group)) +
+  geom_line() 
+```
 

### 2024-01-31 Scripts/authoritarian_only.Rmd
Edits through data and methods

--- 2024-01-19 Scripts/authoritarian_only.Rmd
@@
 The topic has significant implications for authoritarian governance and democratization, and as such has been a very active subject of study.
-However, the various theoretical perspectives brought to bear on the topic have generated competing predictions.
+However, the various theoretical perspectives brought to bear on the topic have been built on different assumptions, and have different theoretical implications.
 To briefly summarize: for some models, election fraud is a tool of weak incumbents.
-As a result, incumbents avoid fraud when it might be exposed and reveal their weakness to potential protesters\textemdash an auspicious prediction for democratization and ultimately the under-pinning of 'self-enforcing' democracy [@fearon2011].
-For others, fraud is committed by weak and strong incumbents alike, making protest a risky gamble the opposition undertakes for other reasons.
-In these models, fraud may or may not be deterred based on other exogenous factors: a dubious, though not completely dismissive, prospect for democratization and self-enforcing democracy.
+As a result, incumbents avoid fraud when it might be exposed and reveal their weakness to potential protesters\textemdash an auspicious prediction for democratization and ultimately the under-pinning of 'self-enforcing' democracy [@Fearon2011].
+For others, fraud is committed by weak and strong incumbents alike, making protest a risky gamble the opposition undertakes for exogenous reasons.
+In these models, fraud may or may not be deterred based on other factors: a dubious, though not completely dismissive, prospect for democratization and self-enforcing democracy.
 Lastly, a third school of thought holds that election fraud is an indicator of genuine incumbent strength.
@@
 However, these models overlook that engineering widespread fraud requires control over patronage resources and the criminal and judicial apparatus (ADD CITES to HARVEY).
-As a result, they overlook the fact that high levels of fraud are a sign of incumbent strength in cooptation and Gerschewski's third pillar: repression.
+As a result, they overlook the fact that high levels of fraud are a sign of incumbent strength in both cooptation and Gerschewski's third pillar: repression.
 Both of these tools can be useful for defusing or dismantling an emergent protest movement.
-As a result, effective fraudsters should also be effective protest-crushers.
+As a result, effective fraudsters should also be effective dismantlers of protest.
 This helps explain the broad pattern, where most fraudulent elections go unprotested, and among those that do generate protest, the vast majority fail to dislodge the incumbent.
-Consequently, there is little penalty for severe manipulation in terms of protest risk, and little deterrent effect.
+Consequently, in this view, there is little penalty for severe manipulation in terms of protest risk, and little deterrent effect.
 
 Testing this proposition in more depth is difficult.
-requires looking beyond binary measures of protest, and toward measures of its scope.
-Research on the risk posed to governments by electoral protest often acknowledges---whether explicitly or implicitly---the importance of broad-based protest campaigns, rather than one-off events [@Bunce2006;@tucker2007;@DAnieri2006;@Carnaghan2016;@bunce2010].
-This implicit assumption has largely been overlooked in theories and empirical studies of the relationship between election manipulation and protest, however.
-In most cases, the risk of protest has both been conceptualized and operationalized as a binary outcome: protests occur or they do not.
-This approach obscures the importance of the number and size of electoral protests in a ruling party's calculation of the risk involved in manipulation an election.
-
-Using cross-national data on 1110 elections from 1989 to 2012 where incumbents claimed victory, combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019], this paper finds little support for the hypothesis that more fraudulent elections are associated with a higher risk that any post-election protest event occurs.
-Furthermore, it shows that more severe election fraud is not associated with more numerous or larger post-election protests.
-Instead, these protest dynamics appear driven more by structural and institutional factors, including in particular whether or not the incumbent chief executive is term limited.
-
-Since fraud is not associated with the overall number or size of of events in a protest wave, the risk to governments has likely been overemphasized.
-That is, even if protests do break out in response to large-scale fraud, the government has many tools it may deploy to curtail them.
-Once protests begin, the noisy signal of regime strength provided by the election is replaced by a very clear signal: the force of the regime's response to the emerging protest movement in contest with the mobilizational capacity of the opposition.
-To the extent that larger protest waves are more damaging to the government than small ones, protest risk does not appear to be a substantial constraint on the severity of election fraud.
-Troublingly, these findings indicate that ruling parties should follow an 'all-in' strategy on election manipulation.
-More manipulation is not associated with a larger number of protests, but a large margin of victory is associated with a lower risk of protest initiation.
+If, for example, in equilibrium incumbents manipulate just the right amount to avoid mass protest\textemdash and are deterred from manipulating more by that risk\textemdash we would observe no statistical relationship between the severity of fraud and the likelihood of protest initiation.
+High and low levels of fraud would generally be associated with no protest.
+Beginning to untangle the relationship between election fraud and protest requires looking beyond standard binary measures of protest.
+Instead, this paper tests outcomes "off the equilibrium path," to use the language of formal modeling.
+Does greater risk to the incumbent in the pre-election period result in reduced election fraud?
+When protest movements do arise after an election, is fraud associated with weaker governments?
+Using data from 614 non-democratic elections across 97 countries, from 1989 to 2012, this paper finds that the answer to both questions is no.
+In fact, more severe election fraud is associated with a *faster* demise for protest waves.
+This indicates that, far from being a sign of weakness, in most cases the ability to generate significant electoral fraud is an indicator of durable regime strength.
+This finding holds even when incumbents claim small vote-shares, suggesting that electoral fraud capacity is an independent component of regime stability.
+
+Combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019],... [MOVE]
+
+Given these results, it is likely that the downside risk for incumbents that engage in election fraud in non-democracies has likely been overemphasized.
+Even in the unlikely event that protests do break out in response to large-scale fraud, the government has many tools it may deploy to curtail them.
+Fraud itself is a sign of the robustness of some of those tools\textemdash the control over resources that incentivizes fraud can be used to coopt the opposition or constrain elites, while control over the judicial and prosecutorial systems can be used to intensify repression during a crisis.
+
+[REWRITE LATER] One troubling implication of these results is that ruling parties should follow an 'all-in' strategy on election fraud, a process that entails consolidating control over patronage resources and further eroding the rule of law.
+More fraud is not associated with longer-lasting protest waves, but a large margin of victory is associated with a lower risk of protest initiation.
 These results help explain both the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory.
 They also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation.
+
+# Literature review: election manipulation and mass protest
+
+Prior research has often considered that electoral manipulation has both costs and benefits for ruling parties.
+The benefits include better odds of winning the official tally [@lehoucq2002], of course, as well as any signals of strength and stability that a larger vote-share [@magaloni2006] or display of power that manipulation might provide [@simpser2013;@gehlbach2015electoral].
+On the other side, election manipulation is generally characterized as having both operational costs (which the incumbent may be unwilling or unable to pay [@Greene2010]) and legitimacy costs which reduce other actors' level of cooperation with the incumbent [@birch2011].
+The most visible, and arguably most studied, form of such legitimacy costs is mass post-election protest against the incumbent.
+
+The possibility of post-election protest regularly figures as a cost that incumbents must weigh when determining whether (and how much) to inflate their vote share.
+In these approaches, the cost of protest be conceptualized as the risk that the incumbent may be unseated and/or the resource cost needed to repress or buy off the incumbent's opponents.
+Various theories model the decision to engage in post-election protest either as a strategic one by opposition groups [@magaloni2010;@Luo2020;@przeworski2015], or more spontaneous coordination by citizens expressing a shared grievance [@tucker2007;@Brancati2016].
+In either case, the risk of protest is a classic component of models of electoral manipulation, as a restraint on incumbents who would prefer to use fraud to increase their vote-shares, all else held equal.
+This argument can also be found in theories about the conduct of non-democratic elections, the focus here, as well as in the literature on self-enforcing democracy [@Fearon2011;@przeworskiSelfenforcingDemocracy2006].
+In the latter literature, generally free and fair elections are thought to be a self-enforcing equilibrium through the threat of mass protest by wronged groups if the incumbent government rigs the results.
+While the focus here is on elections in authoritarian cases, the implications for self-enforcing democracy are discussed at the close of the paper.
+
+<!-- Broadly, prior work has considered two major explanations for the severity of manipulated elections: the costs and benefits of manipulation to the ruling party and government (a demand-side framework) and the costs and benefits for the front-line actors who do the work of manipulating elections (a supply-side framework).
+The risk of costly protest is the central, though not sole^[In some studies, like [@birch2011], the costs of manipulation are attributed to a more diffuse loss of legitimacy; such costs *may* manifest as protest, but may also manifest as a reduction in general cooperation with the regime.
+However, it is worth noting the argument of [@gehlbach2015electoral] that election manipulation can *induce* cooperative behavior from some agents.], factor which affects the demand-side cost of manipulation in most models.
+-->
+
+<!-- In the major formulations of this argument, such as Fearon  [-@fearon_self-enforcing_2011] and Przeworski [-@przeworskiSelfenforcingDemocracy2006], 'democracy' is understood in procedural terms.
+For Przeworski, democracy requires that "authorization to rule is derived from elections" which designate winners and losers according to a set of rules [@przeworskiSelfenforcingDemocracy2006, p.
+312].
+Fearon adopts a similar definition for electoral democracy [@fearon_self-enforcing_2011, p.
+1662].
+Such arrangements are not limited to liberal democracies, and in fact constitute the large majority of regimes today.
+Many previous works argue that the risk of protest can mitigate election manipulation in regimes where elections are held\textemdash including established democracies [@millerRepublicIfYou2021], electoral authoritarian regimes [@Egorov2021], and any regime where elections are held but may be manipulated [@Little2015;@Luo2020].
+In most such models, protests are successful in overthrowing the government with some probability, but impose a cost on both the protesters and the government regardless of success (e.g.
+[@Luo2020]); the logic thus extends even to highly repressive and entrenched regimes.
+Nevertheless, this body of work has yet to grapple with the empirical finding that election integrity is relatively poor across these regimes outside of a narrow club of 'consolidated' democracies.
+-->
+
+Much of the work on this subject has been done in the formal modeling tradition.
+In game-theoretic research where the risk of protest is shown to deter fraud, the election is usually conceived as a signaling game.
+In such models, incumbents are characterized as 'weak' or 'strong' based on their support in the electorate; weak incumbents may need to resort to fraud in order to secure re-election.
+As a result, the opposition perceives election fraud as a reliable signal of incumbent weakness, and will stage protests in order to push out the weak incumbent.
+For example, the model developed by Little, Tucker, and LaGatta [-@Little2015] predicts that "Just as stronger election results for the incumbent lead to less protest because they indicate that she is popular, reports of more fraud lead to more protest because they indicate that the incumbent is, for a given election result, less popular" (p.
+1153).
+In equilibrium, this can lead weaker incumbents to step down rather than face protest due to massive fraud\textemdash a pro-democracy outcome.
+Similar results are developed by Chernykh and Svolik [-@chernykh2015third], Luo and Rozenas [-@Luo2020] and [-@Luo2017], and Magaloni [-@magaloni2010].
+These are normatively optimistic models: fraud can be kept in check, and ultimately reduced, through the development of election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].
+
+A second class of formal models holds that fraud operates as a distortion of the signal of the incumbent's true strength.
+In this type of model, such as those developed by Little [-@littleFraudMonitoringNoncompetitive2015] or Egorov and Sonin [-@Egorov2021], fraud may be used by incumbents to bluff their opponents into being more supportive than they would be if they knew the incumbent's true popularity.
+However, greater transparency makes such bluffs more costly.
+The typical result in these models is that protest risk can *partially* deter fraud.
+In equilibrium, the incumbent will generally choose the level of fraud anticipated by the opposition and no more\textemdash but also no less.
+High levels of fraud can be committed by strong or weak incumbents, and as such these models do not predict a relationship between the risk of protest and the level of fraud per se.
+As Egorov and Sonin [-@egorov2014incumbency] write when discussing their model of this type, "We see that even dictators who are sure to win may use fraud; the reason is that if they do not, citizens will believe that they are less popular and used fraud anyway [thus inviting protest]." [ADD PAGE] In these models, protest initiation is most likely when fraud is higher than expected and when the incumbent's vote-share is low (see, e.g., Little [-@little_elections_2012].
+
+Similar predictions emerge from another school of thought, which holds that electoral protest is driven by citizens' grievances rather than by a strategic, organized opposition [@Brancati2016].
+In this view, election manipulation can act as a shared grievance among individuals who prefer to see the ruling party lose\textemdash especially when it appears that fraud may have been decisive [@kuntz_more_2009].
+Belief in a stolen election is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007], possibly initiating a protest cascade [@Kuran1991].
+Just as above, these models tend to predict greater protest risk when margins of victory are low and/or grievances are high in the presence of fraud.
+
+Empirical work has found some support for the microfoundations of this theory.
+For example, ballot-stuffing and voter coercion have been shown to generate stronger feelings of anger than pre-election manipulation of ballot access [@Szakonyi2021], and access to information about manipulation appears to reduce support for the ruling party [@Reuter2021], increase support for protest and reduce the perceived legitimacy of the regime [@Williamson2021].
+This argument has found support in studies of the relationship between economic grievances, manipulated elections, and protest [@Rod2019;@Brancati2016].
+
+The grievance-based approach suggests that protests hinge more on citizens' subjective beliefs about fraud and the broader social context than on the actual level of malfeasance.
+A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b].
+Subjective belief in fraud was also shown to be correlated with approval of mass protests after Russia's 2011 election [@Chaisty2013].
+Furthermore, simply being on the losing side in an election is associated with reduced belief that the election was fair [@Cantu2015], and with increased support for electoral protest [@Anderson2006].
+These studies imply only a limited deterrent effect of protest on fraud in non-democracies.
+If the incumbent may pay the cost of protest whether or not it engages in fraud (due to subjective, and perhaps mistaken, beliefs about fraud and economic conditions), it may as well engage in manipulation in order to gain the benefits.
+
+The strategic and social-psychological explanations are, of course, interconnected.
+As Schedler writes, "Grievances provide moral justifications for collective protest: we need to do something.
+Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304].
+
+Finally, there is a third school of thought, which holds that election manipulation may indicate regime strength, especially if manipulation is blatant or produces large margins of victory [@simpser2013].
+In the next section, I expand on this school of thought in more detail.
+
+There is a general, though not complete, consensus in empirical work that election fraud is associated with increased risk of protest.
+Several cross-national studies have found links between election manipulation and binary measures of protest [@Brancati2016;@Hyde2014;@Harvey2020;@Rod2019].
+A study of African elections from 1997 to 2009 shows that election fraud is associated with an increased number of post-election conflicts [@daxecker2012].
+However, Hafner-Burton et al [-@Hafner-Burton2016] find that the positive association between election fraud and post-election protest falls short of conventional statistical significance.
+Likewise, looking at the 2011 election and protest cycle in Russia, Ananyev and Poyker [-@ananyevDictatorsSignalStrength2022] show that post-election protest was not associated with the severity of fraud using several measures^[Lankina and Skovoroda [-@Lankina2017] find that post-election protest was positively associated with a different measure of election fraud in that case, however.] Taking an alternative approach, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e.
+whether opponents are allowed to run), rather than by the election-day manipulation that usually figures in theoretical work.
+This result is consistent with the finding that parties are more likely to protest elections after ex ante changes to the rules [@chernykh2014].
+
+While these works have identified a connection between fraud and protest, in general, they cannot distinguish between the competing assumptions of the two theories discussed here, or the third possible approach discussed in below.
+All of the relevant theories predict that protest should be most likely when evidence of fraud is high and the incumbent's vote-share is low (or, relatedly, grievances are high).
+However, the implications of the theories are substantively distinct.
+Protest risk only acts as a deterrent to fraud in theories where fraud is a signal of regime weakness.
+So, it is necessary but insufficient to ask whether the occurrence of post-election protests is associated with higher fraud.
+It is also necessary to test the assumptioins underlying the different models.
+In particular, do post-election protest movements encounter weak incumbents as signaling models predict?
+If not, the deterrence model is undermined.
+
+<!--  One critique of this view, although still within the information/strategic framework, comes from Simpser's [-@Simpser2011] pathbreaking book.
+Simpser argues that when wanipulation efforts are excessive---well beyond what is needed simply to win---they demonstrate the organizational capacity, resources, and manpower of the ruling party [@Simpser2013].
+Large-scale manipulation is, thus, a signal of strength rather than weakness, and can deter protest.
+By contrast, partial manipulation efforts that fail to secure a large majority vote-share signals the government lacks both popular support and mobilizational capacity, making opposition protest more likely.
+-->
+
+[MOVE TO DISCUSSION AND REWRITE] Generally speaking, protest-based models predict that manipulation should decline as opposition groups become stronger relative to incumbents.
+This dynamic creates a sort of virtuous cycle, wherein strong opposition groups' credible threat of protest makes elections fairer, which in turn improves the strength of the opposition--leading ultimately to the ideal of self-enforcing democracy.
+The alternative school of thought, among other predictions, argues that opposition strength drives front-line agent behavior more than that of leaders, and causes a shift to less dispersed, less observable kinds of manipulation [@asunka2017electoral;@Harvey2019].
+Election integrity can thus remain low, even in competitive contexts.
+Putting the protest-deterrent logic to the test, then, helps resolve significant questions in the study of election manipulation and democratization.
+
+  Of course, electoral manipulation is not understood to be the sole cause of protest\textemdash protests occur for all sorts of reasons\textemdash but the literature has drawn a close connection between the risk of specifically electoral protest and the conduct of elections.
+
+# Theory
+
+As discussed above, one view of election fraud and protest views fraud as a tool of weak incumbents, while another views it as a tool that is employed by weak and strong alike.
+A third view, associated with Simpser [-@simpser2013], holds that election fraud demonstrates incumbent strength.
+In Simpser's original account, we should expect electoral manipulation to indicate strength when it is blatant\textemdash i.e.
+readily observable\textemdash or excessive, producing large margins of victory.
+A strong government may manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver].
+This argument is in keeping with Seeberg's [-@Seeberg2019] finding that higher-capacity states are weakly associated with more election fraud.
+
+[CONSIDER Ananyev and Poyker, Trinh 2022--do they fit in?]
+
+I build on Simpser's logic, and subsequent research, by arguing that that extensive election fraud itself requires the use of resources that are sources of incumbent strength, whether or not these are observed.
+Simpser's model is concerned with whether or or not other actors perceive the regime's signal of strength, leading to the focus on excessive or blatant manipulation.
+I argue here that even when large-scale election fraud is marginal (i.e.
+produces narrow margins of victory), it can indicate regime strength independent of popularity.
+This argument  draws on the recent strand of research on election manipulation that emphasizes the principal-agent problems inherent in such campaigns.
+Political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].
+
+control over the bulk of patronage resources in society , and the ability to shield agents from criminal and legal risks [@Harvey2019;].
+Each of these resources can be used to minimize protest waves if they do occur after the election.
+
+To the extent that election results in non-democracies send a signal, it is composed of a mix of three factors: the incumbent's genuine popularity, its ability to mobilize otherwise reluctant voters (through patronage, clientelism, and coercion), and its ability to control the bureaucratic apparatus (including the criminal and judicial systems) to generate fraud.
+Large-scale election fraud requires that the incumbent have consolidated control over patronage resources in society; this makes cooperation with the incumbent attractive for front-line agents, who are less willing to risk losing access to the incumbent's network [@Harvey2019].
+However, such efforts also require that incumbents be able to shield front-line agents from negative consequences if their illegal efforts are publicly exposed [@Harvey2022].
+In other words, the incumbent must have sufficient control over the police and the courts, such that ordinary individuals expect little punishment for committing criminal election fraud on their behalf.
+Each of these capacities is required to commit large-scale fraud, and both are useful for disarming protest movements.
+
+The initial emergence of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014].
+Concessions may be aimed at protesting parties themselves [@reuter2015;@mooreRepressionDissentSubstitution2000;@Hafner-Burton2016;@Piven1979] or the more moderate public, in order to keep them on the sidelines of the conflict [@Hummel2019].
+In either case, large-scale election fraud indicates control over considerable patronage resources, which in turn indicates the capacity to engage in such payments (or withhold them from clients in exchange for passive support [@Rosenfeld2017]).
+
+Ruling parties may also lean on repression, calling upon the security services to disperse protests, arrest leaders, and perhaps employ harsher methods including disappearance, torture, and execution [@Escriba_Folch2013;@Davenport2007].
+Though such crackdowns risk provoking a backlash, they are aimed at demobilizing the opposition by increasing the costs of participation in overt dissent [@DeVogel2022].
+The ability to protect regime agents from any legal consequences for carrying out orders, a requirement for successful election fraud, is also an asset when asking law enforcement agents to carry out possibly violent repression [@MAGALONI_RODRIGUEZ_2020;@stavro_police_2023;@abouharbFactoJudicialIndependence2013].
+
+This argument\textemdash that successful, large-scale fraud draws on the same resources a state may use to suppress protest\textemdash turns the deterrent logic on its head.
+Instead of signalling weakness, large-scale fraud indicates a well of resilience.
+Consequently, we should not expect incumbent regimes capable of committing mass fraud to be deterred from doing so by the risk of protest.
+This argument clearly echoes the seminal argument of Alberto Simpser [@simpser2013].
+It is distinct, however; for Simpser, election manipulation only communicates strength to other actors when it is excessive (generating high margins of victory) or blatant.
+As Simpser notes, "...the marginal versus excessive dimension of electoral manipulation does not fully overlap with the question of scale or extent: while excessive electoral manipulation is generally associated with large-scale manipulation, the marginal kind can result either from a low amount of manipulation (e.g., in a tight race) or from a large amount [due to an unpopular incumbent]" (p.
+7).
+
+All together, I argue that a large margin of victory combined with election fraud sends an unambiguous signal of ruling-party strength, with uncertainty around its components.
+That is, large margins of victory can be composed of a number of values of 'natural' popularity and election fraud.
+As margins of victory decline, the relative value of the signal becomes clearer.
+A narrow margin of victory combined with extensive fraud is a clear signal of low 'natural' popularity.
+While this may indicate an opportune moment for opposition actors, the large extent of fraud is nevertheless itself an independent indicator of regime strength.
+The combination of patronage resources and control over the criminal and judicial systems required for mass fraud can be turned to coopting or repressing protest movements.
+
+This theory makes predictions for protest initiation, protest duration, and the level of fraud employed.
+First, if large-scale election fraud indicates strength, we should expect that incumbents will respond to pre-election signs of opposition strength by increasing the level of fraud.
+This runs directly counter to the prediction of the classic models, which hold that a greater risk of post-election protest should deter fraud.
+The prediction of second-generation [USE THIS AS SHORTHAND?] models is embedded in the null hypothesis of H1, since these models hold that protest risk is not tethered to the level of fraud itself (e.g.
+[@egorov2014incumbency]).
+In those models, a greater risk of protest makes it more important to incumbents that they satisfy the opposition's expected level of fraud, which may be high or low.
+
+> H1: Pre-election environments that are more uncertain for the incumbent will be associated with increased election fraud.
+
+Likewise, if election fraud and incumbent vote-share are both indicator of incumbent strength, we should expect protest initiation to be most likely when both are low.
+A high vote-share in the absence of large-scale fraud indicates broader regime popularity, which dampens protest risk.
+As argued here, a high level of fraud can help compensate for a low vote-share by indicating independent sources of regime strength (i.e.
+resource capacity, agent loyalty, and biased rule of law).
+Only when both signals are low should we expect a high risk of protest.
+This hypothesis is consistent across all three schools of thought.
+
+> H2: The likelihood of protest initiation will be highest when incumbent vote-share and election fraud are low.
+
+Finally, if large-scale protest is indicative of regime resiliency, we should expect that election fraud will be associated with reduced protest duration among incumbents who successfully fended off protest movements.
+That is, among strong incumbents who were capable of defeating post-election protest movements, the resources associated with large-scale fraud should enable them to more swiftly quell the movement.
+Those who lack such resources are expected to take longer to dispatch the protest movement.
+The effect should run in the opposite direction for incumbents who were removed through post-election protest.
+These types, revealed ex post to be weak, should survive longer against the protest movement if they control the resources needed to carry out large-scale fraud.
+Those who lack such resources should be defeated more quickly.
+Again, these expectations are counter to the standard protest-deterrent logic and distinct from the null hypothesis expected in second-generation [USE?] models.
+
+> H3a: Among incumbents that survive post-election protest movements, more extensive election fraud should be associated with shorter protest waves.
+
+>H3b: Among incumbents that are defeated by poste-election protest movements, more extensive election should be associated with longer protest waves.
+
+That election dates are known in advance creates opportunities for opposition mobilization; anticipating this, regimes can more easily engage in pre-emptive repression [@Truex2019;@Sullivan2016;@Bhasin2013].
+
+In addition, governments may learn from the experiences of other governments, just as opposition groups may learn from international opposition movements.
+Anticipating threats from below may encourage less democratic governments to invest more in the security services, or to reform political and legal institutions in order to stifle potential protests before they take shape [@hallAuthoritarianLearningConceptual2017a;@Morgenbesser2020], further reducing the expected cost of protest.
+
+In particular,
 
@@
 library(simPH)
-
-nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2024-01-16.csv"))
+library(broom)
+library(broom.helpers)
+
+nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2024-01-30.csv"))
 
@@
 
-nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(is.na(max_size)==T & n.events.post == 0, 0, max_size))
+#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(is.na(max_size)==T & n.events.post == 0, 0, max_size))
 
@@
 
-nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))
+#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))
 
@@
 
-nelda2.sub <- nelda2.sub %>% filter(v2x_regime == 1) #Includes only authoritarian
+nelda2.sub <- nelda2.sub %>% filter(v2x_regime == 1 | regime.1lag == 1) #Includes only authoritarian, add e-dems with | v2x_regime == 2  #Current version includes lagged 1 in order to capture pre-election auth.
+
+## Remove outlier
+nelda2.sub <- nelda2.sub %>% filter(d.events.num <= 100)
 
@@
 
-```
+## Coding median size as 0 when n.events = 0
+
+nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(n.events.post == 0, 0, median.participants))
+
+```
+
+These hypotheses are tested using data from three datasets: NELDA [@hyde_2012_marinov], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019].
+The unit of observation for this study is an election period as recorded in NELDA.
+This framework prevents selecting on the dependent variable, by capturing both protested and non-protested elections.
+I limit the sample to authoritarian regimes which hold multiparty national elections.
+This excludes fully closed autocracies (where no national elections are held de jure), as well as electoral and liberal democracies.^[The specific variable used is 'v2x_regime'; I limit the sample to those regime-years coded 1.]
+
+While it is possible that the theoretical mechanisms considered here may be at work in democracies as well, democracies are excluded from this analysis for two reasons.
+First, the theory presented above is derived from prior work on authoritarian elections; the factors that drive the theory are expected to be weaker in democratic regimes.
+As regimes trend toward liberal democracy, patronage consolidation should decline and rule of law should increase by definition.
+Consequently, the theory expects that fraud should be rare in such cases, even in the absence of a protest deterrent: incumbents simply lack the tools.
+At the same time, the more liberal-democratic a state becomes, the more alternatives their are to protest as a means to check the executive.
+More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest\textemdash the electoral authoritarian regimes\textemdash where we should expect to see a more pronounced role for protest risk in curtailing election fraud.
+The case selection strategy here is thus one that both fits the theory and makes for a strong test of the classic fraud-protest model.
+
+Of course, treating regime type as a sharply defined binary indicator, rather than as a continuum, will introduce some bias around edge cases.
+As a robustness check for this case selection process, I also run the main analyses on a larger sample that adds electoral democratic regimes.
+While the strength of the findings is somewhat attenuated, as is expected, the results are generally upheld in the broader sample as shown in the appendix.
+
+Models where the dependent variable is the level of election fraud include data from all cases.
+For the models that focus on the initiation of post-election protest, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the theoretical expectations of protest risk following an election in which the incumbent won, possibly through reliance on election fraud.
+For models of protest-wave duration, only cases with a positive number of post-election protests are included.
+The date range is limited by the availability of data on protest from ECAV, as discussed below, which ranges from 1989 to 2012.
+The resulting dataset includes 614 total election-period observations across 97 countries.
+
+## Dependent variables and modeling strategies
+
+To capture the severity of *election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than alternative measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], may not mobilize real voters [@Harvey2020], and can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021].
+This measure of election integrity should thus serve as a strong test of the protest-oriented model.
+It also matches the conception of election manipulation used in much of the relevant formal literature.^[For example, election manipulation is treated as post-hoc adjustment of the results in Little et al [-@Little2015], Little [-@Little2012], Magaloni [-@magaloni2010], and Fearon [-@fearon_self-enforcing_2011], among others.] The variable is used as a dependent variable in the first set of models that test H1, and as a potential explanatory variable in the subsequent models of protest initiation and duration.
+As an additional check in the appendix, a fuller measure of overall election integrity is used instead.
+
+ECAV records protests and other contentious actions related to an election, ranging from six months before to three months after the vote (or three months after the final round of a multi-round election).
+To focus the study on protest, I only include events labeled as protests, occupations, or blockades, excluding acts like shootings, bombings, and coups.
+ECAV data is the event-day format; to capture the *number of protests* before the election, I take the sum number of events recorded in ECAV before election day.
+To evaluate hypotheses regarding *protest duration* I take the number of days between the last protest recorded in ECAV and the first, for each election.^[The duration of a protest wave with only one protest is taken to be one day.
+While the main analysis assumes that the last protest in the dataset marks the end of the protest movement, an alternative specification in the appendix assumes right-censoring for a portion of the observations due to the 90-day observation window.]
+
+The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0) %>% nrow()` of the `r nrow(nelda2.sub)` elections, no protests are recorded.
+Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest number and duration.
+As the figure shows, while election fraud is common among this set of regimes, protests are rare, and, when they do occur, are usually short in duration.
+This distribution\textemdash with large numbers of zeroes\textemdash is in line with the theoretical proposition that most fraudulent elections do not produce signals of weakness that prompt opposition protest.
+Still, approximately one-third of elections experienced some protest, and so protest initiation is modeled using standard logistic regression.
+*Election fraud*, which follows a more normal distribution, is modeled using multilevel models with random intercepts by country.
+Lastly, protest wave duration is modeled using Cox proportional hazards models.
 
@@
 
-p.hist.size <- ggplot(nelda2.sub, aes(x=max_size)) + geom_histogram(bins=100) + xlab("Maximum protest size") + theme_bw()
+p.hist.duration <- ggplot(nelda2.sub, aes(x=d.events.num)) + geom_histogram(bins=100) + xlab("Protest wave duration") + theme_bw()
 
@@
 
-  ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= max_size)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("max size") + theme_bw()
-```
-
-```{r elex-hist, fig.cap="Distributions of election fraud, post-election protest, and maximum protest size", fig.height=7, echo=FALSE, message=FALSE, warning = FALSE}
-plot.combined.hist <- cowplot::plot_grid(p.hist.fraud, p.hist.events, p.hist.size, ncol=1)
+#  ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= max_size)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("max size") + theme_bw()
+```
+
+```{r elex-hist, fig.cap="Distributions of election fraud, post-election protest, and protest duration", fig.height=7, echo=FALSE, message=FALSE, warning = FALSE}
+plot.combined.hist <- cowplot::plot_grid(p.hist.fraud, p.hist.duration, ncol=1)
 plot.combined.hist
@@
                                             'Off-schedule election' = unscheduled_election,
-                                          'Regional protest diffusion' = regional_success_lag)
+                                          'Regional protest diffusion' = regional_success_lag,
+                                         "Privatization" =  state.owned.1lag,
+                                         "Professional bureaucracy" = patronage.appts.1lag,
+                                        "Anti-system" = antisys.1lag,
+                                        "EMB autonomy"= v2elembaut)
 #incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))
+
+incl.vars <- incl.vars %>% mutate(`Presidential election` = as.numeric(`Presidential election`))
+incl.vars <- incl.vars %>% mutate(`Rentier state` = as.numeric(`Rentier state`))
+incl.vars <- incl.vars %>% mutate(`Tax state` = as.numeric(`Tax state`))
+incl.vars <- incl.vars %>% mutate(`Off-schedule election` = as.numeric(`Off-schedule election`))
 
@@
 
+```{r, include=FALSE}
+library(ggcorrplot)
+corr <- round(cor(incl.vars, use = "pair"), 2)
+
+ggcorrplot(corr, hc.order = TRUE, type = "lower")
+```
+
+To add: Beaulieu book; Daxecker 2019 et al (fraud perceptions), Daxecker 2012 (cost of exposing cheating), Donno 2013 (book) , Beaulieu 2014 book, hyde and marinov 2014, Morrison, Donno, Savun, and Davutoglu (2024)
+
 ## Explanatory variables
 
+The first set of models test Hypothesis 1, on the relationship between protest risk and fraud.
+Protest risk is operationalized in three ways.
+First, the *number of pre-election protests* in ECAV is taken as *prima facie* evidence of a more turbulent and risky election environment.
+Given that pre-election protest is likely to be a strong predictor of post-election protest\textemdash by indicating strong social networks [@mateo2022], pre-existing opposition groups and civil society capacity [@Bunce2006], and/or popular grievances [@Brancati2016]\textemdash the deterrent logic suggests that incumbents will pull back on election-day fraud when the pre-election environment is characterized by protest [@Rozenas2016b].
+On the other hand, H1 predicts that incumbents will generate more manipulation in this context.
+
+[MOVE] Moreover, in general, incumbents tend to avoid mass repression in the pre-election phase in favor of targeted harassment of opposition activists [@Bhasin2013].
+
+Second, since a state with greater capacity for repression may be better able to deter protest and thus 'get away with' more manipulation [@Egorov2021,@Ritter2016], I measure repressive capacity using the *physical integrity* index from V-Dem, lagged one year.
+Higher values for this variable indicate greater freedom from torture and political killing.
+In one approach, higher values represent more permissive environments, and should be associated with greater protest risk (and thus reduced fraud).
+Alternatively, in keeping with the 'murder in the middle' framework [@feinLifeIntegrityViolationsDemocracy1995], the highest protest risks (and thus most reduced fraud) might be found at middle values of repression; regimes with the most demonstrated repressive capacity deter potential challengers from taking to the streets in the first place, and the least repressive states defuse protest through accommodation [@pierskallaProtestDeterrenceEscalation2010;@mooreRepressionDissentSubstitution2000].
+I thus include the variable's square term to account for the 'murder in the middle' framework.
+
+Finally, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby.
+This diffusion mechanism is modeled using each country's geopolitical region as coded in Teorell et al [-@qog_teorell] in conjunction with binary variables indicating a successful protest movement from NELDA.
+A 'successful protest' is recorded if a protest movement resulted in either a new election being run or the incumbent stepping down.
+The *regional diffusion* dummy variable takes on a value of one for all countries in the same region as a successful protest, for elections within one calendar year of the successfully overturned election.^[For example, the variable takes on a positive value for elections held in the post-communist region, within one year of the September 24, 2000 election in Serbia that sparked the 'Bulldozer Revolution'.
+All told there are 15 successful electoral protests in the dataset: Bulgaria 1990, Lesotho 1998, Cote d'Ivoire 2000 and 2010, Peru 2000, Serbia 2000, Georgia 2003, Ukraine 2003-2004, Azerbaijan 2005, Kyrgyzstan 2005, Haiti 2006, and Cameroon 2007.]
+
+The models incorporate several potential explanatory variables, in addition to the *election fraud* variable described above.
+The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance.
+The winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia and Psephos.} Finally, whether or not the chief executive is *term limited* is taken from NELDA.
+
+## Control variables
+
+As identified in prior research, other factors that may influence both fraud and protest include the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), and the available of *alternative sources of information*.
+Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019].
+All models include *legislative constraints* on the executive, a V-Dem variable which captures the extent to which the legislative body in the country acts independently in practice to limit executive power.
+Additional V-Dem variables capture *privatization* of the economy ('v2clstown') and *bureaucratic professionalization* ('v2stcritrecadm'), both of which might give incumbents influence over election fraud and subsequent protest.
+To capture the overall level of pre-election protection for rights associated with liberal democracy, I include the lagged *liberal democracy index* from V-Dem.
+*Physical integrity* is used as a control here, to address the risk of selection bias (if, for example, high-repression states select into a high-fraud condition with the hope that the repressive apparatus will deter protest).
+
+Several binary control variables are also included.
+First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest.
+Finally, I include dummy variables that provide a categorical evaluation of state capacity.
+These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap).
+If the V-Dem data codes a state as not capable of raising revenue or primarily relying on loans and foreign aid, I code it as a *weak state*.
+If the state is described as primarily relying on resource rents and expropriation, I code it is a *rentier state*.
+States that rely primarily on taxes on property or economic transactions are coded as *tax states*.
+This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018].
+Lastly, since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.
+
+Table \@ref(tab:summary-table) shows summary statistics for all of these variables.
+To address concerns about multicollinearity, I present models in the appendix (page 25) that exclude control variables associated with pre-election protest.
+
 \newpage
@@
 ```{r manipdv, echo=FALSE}
-m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  +  unscheduled_election + state.owned.1lag + patronage.appts.1lag + antisys.1lag + v2elembaut + (1 | stateid), data = nelda2.sub)
+
+m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag +  v2elembaut + (1 | stateid), data = nelda2.sub)
+#summary(m1a)
+
+m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag  + unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)
+
+#summary(m1b)
+
+m1c <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub)
+#summary(m1c)
+
+m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub)
 
@@
 
-m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag + antisys.1lag + regional_success_lag +  v2elembaut + (1 | stateid), data = nelda2.sub)
+m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag + v2elembaut +  (1 | stateid), data = nelda2.sub)
 #summary(m1a)
 
-m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag + antisys.1lag + unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)
+m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag +  unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)
 
@@
 
-m1c <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag + antisys.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub)
-#summary(m1c)
-
-m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + antisys.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub)
-
-#summary(m1)
-
-m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag + antisys.1lag + regional_success_lag + v2elembaut +  (1 | stateid), data = nelda2.sub)
-#summary(m1a)
-
-m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag + antisys.1lag +  unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)
-
-#summary(m1b)
-
-m1d <- lmer(v2elirreg.inv ~  libdem.1lag + gdpgro.1lag +    presidential +  urban.pct + rentierstate + taxstate  +  unscheduled_election + n.events.pre  + state.owned.1lag + patronage.appts.1lag + antisys.1lag + v2elembaut +  (1 | stateid), data = nelda2.sub)
+m1d <- lmer(v2elirreg.inv ~  libdem.1lag + gdpgro.1lag +    presidential +  urban.pct + rentierstate + taxstate  +  unscheduled_election + n.events.pre  + state.owned.1lag + patronage.appts.1lag + v2elembaut +  (1 | stateid), data = nelda2.sub)
 #summary(m1d)
@@
 
+\doublespace
+
+# Results
+
 ## Election-day fraud and protest risk
 
-```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
-model.n.novoteshare.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
-                                ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0"),
-                                family=truncated_nbinom2)
-summary(model.n.novoteshare.hurd.mm)
-
-model.n.base.hurd.mm <- glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
-                               ziformula=~.,
-                              data = nelda2.sub %>% filter(inc.lose == "0"),
-                                family=truncated_nbinom2)
-summary(model.n.base.hurd.mm)
-
-model.n.base2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
-                                          zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.n.base2.hurd.mm)
-
-model.n.griev.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +  v2elirreg.inv*gdpgro.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0"),
-                                family=truncated_nbinom2)
-summary(model.n.griev.hurd.mm)
-
-model.n.repr.hurd <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +   + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv  +  physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
-                                ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0"),
-                                family=truncated_nbinom2)
-summary(model.n.repr.hurd)
-```
-
-```{r models-hurdle-size, warning = FALSE, message = FALSE, include = F}
-model.size.novoteshare <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv +  +   v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0"),
-                                family=truncated_nbinom2)
-summary(model.size.novoteshare)
-
-model.size.base <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv +  +   v2elirreg.inv + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0"),
-                                family=truncated_nbinom2)
-summary(model.size.base)
-
-model.size.base2 <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + gdpgro.1lag  +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0"),
-                                family=truncated_nbinom2)
-summary(model.size.base2)
-
-model.size.griev <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +  v2elirreg.inv*gdpgro.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
-                                ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0"),
-                                family=truncated_nbinom2)
-summary(model.size.griev)
-
-```
-
-```{r initiation-table-setup, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-models_t1 <- list()
-models_t1[['Model 5']] <- model.n.base.hurd.mm
-models_t1[['Model 6']] <- model.n.base2.hurd.mm
-models_t1[['Model 7']] <- model.n.griev.hurd.mm
-#models_t1[['Model 5']] <- m_duration_count
-#models_t1[['Model 6']] <- model.d.base2.hurd
-#models_t1[['Model 7']] <- model.d.griev.hurd
-
-t1a <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, coefs = c("Intercept" = "(Intercept)", "Judicial ind.
-(lag)" = "jucon.1lag", "Liberal democracy index" = "libdem.1lag",
-            "Legislative constraints (lag)" = "legcon.1lag", "Alternative info.
-(lag)" = "altinf.1lag",                             "GDP growth rate (lag)" = "gdpgro.1lag",                               "Civil society openness (lag)" = "osorg.1lag", "National party organization (lag)" = "natlpartyorg.lag",                "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Presidential" = "presidential1", "Incumbent term-limited" = "inc.termlimit.num",
-            "Physical integrity (lag)" = "physinteg.1lag", "Phys. int. (lag) squared" = "physintlag.sq",
-            "Urban" = "urban.pct", "Rentier state" = "rentierstate1", "Tax state" = "taxstate1"),
-              statistics = character(0))  %>% set_caption("Table 3: Zero-inflated models of no protest initiation")
-
-theme_compact(t1a, header_rows = TRUE, header_cols = FALSE)
-
-t1b <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, coefs = c("Number of pre-election protests" = "n.events.pre", "Off-schedule election" = "unscheduled_election1", "Regional protest diffusion" = "regional_success_lag", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd", "Election fraud : GDP growth (lag)" = "gdpgro.1lag:v2elirreg.inv")) %>% set_caption("Table 3 cont.: Zero-inflated models of no protest initiation")
-
-t1b <- huxtable::add_footnote(t1b, "Note: Negative coefficients indicate protest is more likely.")
-
-theme_compact(t1b, header_rows = TRUE, header_cols = FALSE)
-```
-
-```{r conditional-table, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-models_t2 <- list()
-models_t2[['Model 5']] <- model.n.base.hurd.mm
-models_t2[['Model 6']] <- model.n.base2.hurd.mm
-models_t2[['Model 7']] <- model.n.griev.hurd.mm
-models_t2[['Model 8']] <- model.size.base
-models_t2[['Model 9']] <- model.size.base2
-models_t2[['Model 10']] <- model.size.griev
-
-t2a <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), coefs = c("Intercept" = "(Intercept)", "Judicial ind.
-(lag)" = "jucon.1lag",
-            "Legislative constraints (lag)" = "legcon.1lag", "Alternative info.
-(lag)" = "altinf.1lag",                             "Liberal democracy index (lag)" = "libdem.1lag", "GDP growth rate (lag)" = "gdpgro.1lag",                               "Civil society openness (lag)" = "osorg.1lag", "National party organization (lag)" = "natlpartyorg.lag",                "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Presidential" = "presidential1", "Incumbent term-limited" = "inc.termlimit.num",
-            "Physical integrity (lag)" = "physinteg.1lag", "Phys. int. (lag) squared" = "physintlag.sq",
-            "Urban" = "urban.pct", "Rentier state" = "rentierstate1", "Tax state" = "taxstate1"), statistics = character(0)) %>% set_caption("Table 4: Negative binomial models of number and duration of protests") %>% insert_row("", "", "Number", "", "", "Size", "", after = 0)
-
-theme_compact(t2a, header_rows = TRUE, header_cols = FALSE)
-
-t2b <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), coefs = c("Number of pre-election protests" = "n.events.pre", "Off-schedule election" = "unscheduled_election1", "Regional protest diffusion" = "regional_success_lag", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd", "Election fraud : GDP growth (lag)" = "gdpgro.1lag:v2elirreg.inv")) %>% set_caption("Table 4 cont.: Negative binomial models of number and duration of protests") %>% insert_row("", "", "Number", "", "", "Size", "", after = 0)
-
-theme_compact(t2b, header_rows = TRUE, header_cols = FALSE)
-
-#modelsummary::modelsummary(models, shape = term + estimate + component ~ model, coef_omit = "count", gof_omit = "R2", stars = T, longtable=T,  title = 'Binary portion of hurdle models of post-election protest', coef_rename =
-#               c("zero_(Intercept)" = "Intercept", "zero_jucon.1lag" = "Judical independence (lag)", "zero_legcon.1lag" = "Leg. constraints (lag)", 
-#                 "zero_gdpgro.1lag" = "GDP growth rate (lag)", "zero_altinf.1lag" = "Alternative info. (lag)", 
-#                 "zero_osorg.1lag" = "Civil soc. openness (lag)", "zero_econ.crisis.nelda.1lag1" = "Economic crisis", 
-#                 "zero_natlpartyorg.lag" = "National party organization (lag)", "zero_presidential1" = "Presidential election",
-#                 "zero_inc.termlimit.num" = "Incumbent term-limited",  
-#                 "zero_urban.pct" = "Urbanization", "zero_rentierstate" = "Rentier state", "zero_taxstate" = "Tax state",
-#                 "zero_n.events.pre" = "Number of pre-election protests", "zero_pubseccorrup.lag" = "Public sector corruption (lag)", "zero_v2elirreg.inv" = "Election fraud",
-#                 "zero_physinteg.1lag" = "Physical integrity (lag)", "zero_physintlag.sq" = "Physical integrity (lag), squared", "zero_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "zero_unscheduled_election" = "Off-schedule election")) # %>%
-#  add_header_above(c(" " = 1, "Protest quantity" = 3, "Maximum protest size" = 3))
-
-```
-
-\doublespace
-
-The second set of models challenge the idea that more severe electoral manipulation is associated with an increased risk of anti-government protest.
+The results of the first set of multilevel models are not promising for argument that protest risk deters election manipulation.
+Hypothesis 1a is not supported in the models shown in Table  \@ref(tab:manipdv).
+Models 1 through 4 apply different measures of pre-election protest risk; all of them behave contrary to the standard expectation in the literature.
+In Model 1, the number of pre-election protests is associated with an increase in election-day fraud.
+While the effect is small\textemdash ten pre-election protests only tend to increase election fraud by about six percent of a standard deviation\textemdash it runs counter to the expectations of the deterrence logic.
+A similar effect is found for *diffusion* in Model 4; recent successful protest in a nearby country is also associated with a modest increase in expected fraud.
+This suggests that the level of election manipulation, rather than being deterred by a highly contentious environment, may be deployed to make up for anticipated poor election results and/or to signal organizational strength.
+
+Model 2 does not demonstrate a significant linear relationship between repression and election fraud.
+However, Model 3 shows that fraud first increases with physical integrity, but then declines.
+In other words, fraud is lowest on the extremes of repressive capacity\textemdash where repression is severe and where it is most limited.
+This is contrary to the expectation that a higher risk of protest is associated with reduced fraud.
+Under the expectation that protest risk is highest for these middle values [@pierskallaProtestDeterrenceEscalation2010], H1a is not supported.
+By contrast, H1b is supported by Models 1, 3, and 4: by each measure, fraud is higher when protest risk is higher.
+
+## Fraud and the likelihood of protest
+
+The second set of models challenge the idea that more severe electoral manipulation is associated with an increased risk of protest.
 Beginning with the zero-inflation portion of the hurdle models in Table 3, which model the likelihood of *no* protest occurring after the election, the standalone coefficient for election fraud is not significant in Model 5.
@@
 ```{r include=FALSE}
-nelda2.sub <- nelda2.sub %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))
-nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))
-
-model.margin.binary_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.margin.binary_n)
-
-model.margin.binary_size <- glm(no_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.margin.binary_size)
-
-model.binary.griev1_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +  +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.binary.griev1_n)
-
-model.binary.griev1_size <- glm(no_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +  +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.binary.griev1_size)
-
-model.binary.repr_n <- glm(no_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.binary.repr_n)
+nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
+#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))
+
+model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+summary(model.margin.binary)
+
+model.margin.binary_noc <- glm(any_protest ~ jucon.1lag + legcon.1lag +  altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+summary(model.margin.binary_noc)
 
@@
 
-plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary_n, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-1.5, 1.5]"), show.data = T)
-plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw()
-
-###GDP
-
-plot_binary_gdp_typ_n <- sjPlot::plot_model(model.binary.griev1_n, type = "pred", terms = c("v2elirreg.inv", "gdpgro.1lag [-0.04, .08]"))
-plot_binary_gdp_typ_n <- plot_binary_gdp_typ_n +  labs(x = "Election fraud", y = "Predicted probability", title = "Predicted probability of no protest", color = "GDP growth (lagged)") + theme_bw()
-
-## Repression
-
-plot_binary_repr_typ_n <- sjPlot::plot_model(model.binary.repr_n, type = "pred", terms = c("v2elirreg.inv", "physinteg.1lag [.25, .61]"))
-plot_binary_repr_typ_n <- plot_binary_repr_typ_n +  labs(x = "Election fraud", y = "Predicted probability", title = "Predicted probability of any protest", color = "Physical integrity (lagged)") + theme_bw()
-
-```
-
-```{r margin-plot-gdp, echo=FALSE, message=F, warning=FALSE, fig.cap="Marginal effect of election fraud on protest initiation, conditional on lagged GDP growth"}
-plot_binary_gdp_typ_n
-```
-
-```{r n-protests-margin, fig.cap="Fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-1.5, 1.5]"), type = "count") 
-p.n.margin <- plot(dat)
-p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Number of events", title = "Predicted number of protests", color = "Election fraud") + theme_bw()
-```
+plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-.05, 2]"), show.data = T)
+plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()
+
+```
+
+```{r logit-table, echo=FALSE}
+models2 <- list()
+models2[['Model 5']] <- model.margin.binary_noc
+models2[['Model 6']] <- model.margin.binary
+
+modelsummary(models2, output = "kableExtra",  stars = T, title = 'Table 3: Logit models of protest occurrence', coef_rename = 
+               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
+                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
+                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
+                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
+integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity"
+                 )) %>% kable_styling(latex_options = "HOLD_position")
+```
+
+\newpage
+\doublespace
+
+Table \@ref(tab:logit-table) shows the results for Hypothesis 2, which is expected to be supported across all three schools of thought.
+Model 5 is a base model, where incumbent margin is interacted with *intentional irregularities*; Model 6 includes the same interaction along with additional control variables that could be associated with both fraud and post-election protest.
+The expected, significant interaction effect is detected in both models, and it is in the expected direction: the occurrence of any protest is most likely when incumbent vote-share is low and fraud is high.
+
+Figure \@ref(fig:margin-prob-fig) illustrates just how tenuous this effect is, however.
+The figure shows the predicted probability of post-election protest for two specific values of election fraud, across all values of incumbent vote-share, holding all other variables constant at their means (continuous variables) or modes (categorical variables).
+The figure highlights that the predicted probability of protest is only statistically significant for the two values of fraud at extremely low levels of incumbent vote-share.
+This represents an exceedingly unlikely outcome along both variables.
+As the figure shows, there are only a small number of observations in the range where the predicted probabilities are statistically significant.
+At the same time, the values for *intentional irregularities* are also highly unusual.
+They represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively.
+
+While all three schools of thought are consistent with H2, the results here may be taken as circumstantial evidence in favor of the fraud-as-strength model.
+Fraud only appears significantly associated with increased protest risk when the incumbent's result in the election is very low and fraud is very high.
+This is suggestive of the model where opposition actors interpret the result as showing extremely low genuine support for the incumbent, perhaps enough that it outweighs the signal of strength sent by fraud for some oppositions.
+Middle values for vote-share combined with high values of fraud suggest durable regime strength, bringing expected protest close to zero.
+In the alternative view where fraud conveys weakness, we might expect the probability of protest to rise more sharply as incumbent vote-share declines, and to reach higher predicted probabilities.
+This pattern is only suggestive, of course, and for more definitive evidence I look to protest duration.
 
 ```{r margin-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
-plot.combined.prob <- cowplot::plot_grid(plot_binary_margin_typ_n, p.n.margin, ncol=1)
-plot.combined.prob
-```
-
-\newpage
-
-## The number and size of post-election protests
-
-\doublespace
-
-Results from the models of number and size of protest events, conditional on protest initiation, are shown in Table 4.
-Models 5 and 8 show no significant relationship between fraud and and the number or maximum size of protest events.
-The bottom panel of Figure \@ref(fig:margin-prob-fig) shows the marginal effect of election fraud on the number of protests, given varying levels of winning vote-share, holding other variables constant at their means (continuous variables) or modes (categorical variables).
-As the figure shows, there is no statistically significant difference in the expected number of protests.
-No significant effects are found for fraud and GDP growth in Models 7 or 10.
-Nor is there a significant effect for fraud on protest size, conditional on margin of victory.
-However, Hypothesis 6 is consistently supported\textemdash incumbent term limits do seem associated with increased size and number of post-election protests.
-
-Altogether, there is scant evidence to support the argument that protest risk deters election fraud.
-Models 1 through 4 show that various measures of pre-election protest risk are not associated with reduced fraud\textemdash in fact, some appear to increase the severity of fraud.
-Models 6 through 11 show that higher levels of fraud is not "punished" by more or larger protests.
-Hypothesis 2 is supported, but very weakly; there is a significant association between election fraud and protest initiation, but only when the incumbent's winning vote-share is near its minimum.
-However, even this supportive finding likely presents too optimistic a picture for the protest-deterrent model of election fraud.
-The number and size results from the hurdle regressions show no effect for election fraud across any of the models, alone or in interaction.
-Even when protests are initiated\textemdash a relatively rare event in the first place\textemdash the size of a protest wave is not influenced by the degree of fraud perpetrated by the ruling party.
-In turn, this suggests two pessimistic interpretations of fraud-protest dynamic.
-First, protest may simply be less damaging to ruling parties than is regularly assumed in theories of election manipulation.
-Faced with a nascent protest movement, ruling parties are generally able to contain, co-opt, or repress them\textemdash regardless of the severity of election fraud.
-An even more troubling implication would suggest that the protest-manipulation logic, rather than serving as a deterrent, creates perverse incentives to manipulate as much as possible.
-If ruling parties do indeed seek to avoid protest initiation, but manipulating heavily does not entail more severe protest, ruling parties have an incentive fraudulently inflate their official margins of victory in order to reduce the risk of protest initiation.
-
-```{r pred-size, fig.cap="Fraud, winning vote-share, and predicted number of protests", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
-
-p.protestlow <- sjPlot::plot_model(model.n.base2.hurd.mm, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(n.events.pre = 0)) + labs(x = "Incumbent vote-share", y = "", title = "Predicted n of post-election protests (Pre-\nelection protest: 5th percentile)") + coord_cartesian(ylim = c(0, 20))
-
-p.protesthigh <- sjPlot::plot_model(model.n.base2.hurd.mm, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(n.events.pre = 8)) + labs(x = "Incumbent vote-share", y = "", title = "Predicted n of post-election protests (Pre-\nelection protest: 95th percentile)") + coord_cartesian(ylim = c(0, 20))
-
-cowplot::plot_grid(p.protestlow, p.protesthigh, nrow = 1)
-```
-
-```{r pred-size, fig.cap="Fraud, winning vote-share, and predicted max protest size", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
-
-p.protestlow <- sjPlot::plot_model(model.size.base2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(n.events.pre = 0)) + labs(x = "Incumbent vote-share", y = "Max.
-size", title = "Predicted max.
-size (Pre-\nelection protest: 5th percentile)") + coord_cartesian(ylim = c(0, 8000))
-
-p.protesthigh <- sjPlot::plot_model(model.size.base2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [1.5]"), condition = c(n.events.pre = 8)) + labs(x = "Incumbent vote-share", y = "", title = "Predicted max.
-size (Pre-\nelection protest: 95th percentile)") + coord_cartesian(ylim = c(0, 8000))
-
-cowplot::plot_grid(p.protestlow, p.protesthigh, nrow = 1)
-```
-
-```{r zero-model-table-count, echo=FALSE}
-#modelsummary::modelsummary(models, coef_omit = "zero", gof_omit = "R2", longtable=T, stars = T, title = 'Hurdle models of post-election protest (counts)',
-#                           coef_rename = 
- #              c("count_(Intercept)" = "Intercept", "count_jucon.1lag" = "Judical independence (lag)", "count_legcon.1lag" = "Leg. constraints (lag)", 
-#                 "count_gdpgro.1lag" = "GDP growth rate (lag)", "count_altinf.1lag" = "Alternative info. (lag)", 
-#                 "count_osorg.1lag" = "Civil soc. openness (lag)", "count_econ.crisis.nelda.1lag1" = "Economic crisis", #
-#                 "count_natlpartyorg.lag" = "National party organization (lag)", "count_presidential1" = "Presidential election",
-#                 "count_inc.termlimit.num" = "Incumbent term-limited",  
-#                 "count_urban.pct" = "Urbanization", "count_rentierstate" = "Rentier state", "count_taxstate" = "Tax state",
-#                 "count_n.events.pre" = "Number of pre-election protests", "count_pubseccorrup.lag" = "Public sector corruption (lag)", "count_v2elirreg.inv" = "Election fraud",
-#                 "count_physinteg.1lag" = "Physical integrity (lag)", "count_physintlag.sq" = "Physical integrity (lag), squared", "count_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "count_unscheduled_election" = "Off-schedule election"))
-
-```
-
-One objection to this framing could be: perhaps the null result found for the number and size of protests is the result of equilibrium behavior predicted in many formal models.
-The threat of protest keeps fraud lower than it would be otherwise, resulting in limited (if any) protest.
-In some states, this equilibrium level of fraud is higher than in others, resulting in no correlation between fraud and observed protest.
-There are three reasons to be skeptical of this interpretation, two empirical and one substantive.
-
-First, such an objection does not account for the unexpected results found in Models 1 through 4, where indicators of pre-election protest risk were found to be associated with more severe election-day fraud.
-Second, deterrence models would expect the equilibrium level of fraud to fall as protest risk increases, and for more severe protest to occur as a result of miscalculation by the regime.
-We can test this by examining the predicted intensity of protest from the models specified here.
-If the severity of fraud is held constant at a high level while underlying protest risk increases, we should see an increase in the predicted protest outcome.
-Taking protest size as an example, using Model 9, Figure \@ref(fig:pred-size) shows there is no evidence of such an effect.
-Both panels show the marginal effect of a relatively fraudulent election (*election fraud* is set to one standard deviation above the mean) on the predicted maximum protest size, conditional on the closeness of the election result.
-All control variables have been set to their means or modes with the exception of *pre-election protest*.
-In the left panel, the quantity of pre-election protest has been set to the fifth percentile for that variable (0 events), on the right, it has been set to the 95th percentile (8 events).
-An equilibrium approach would suggest that high fraud is more dangerous in the more active protest envionment captured on the right, especially at low levels of *winner's vote-share*.
-However, we see practically no difference in the predicted maximum protest size across the different conditions modeled here\textemdash changing protest conditions do not appear to affect the 'appropriate' level of fraud, throwing cold water on the equilibrium argument.
+plot_binary_margin_typ_n
+```
+
+## Protest wave duration
+
+[CLEAN UP]
 
@@
 
-# A further test: protest wave duration
-
 ```{r, include=FALSE}
@@
 nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
-nelda2.sub <- nelda2.sub %>% mutate(fraud_level = ifelse(v2elirreg.inv >= (.98 + .97), "High",
-                                                            ifelse(v2elirreg.inv <= (.98 - .97), "Low", "Moderate")))
+nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
+                                                                       ifelse(d.events.num >= 62, 0, 1)))
 
@@
 
-survfit2(Surv(d.events.num, surv_status) ~ fraud_level, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
-  ggsurvfit() +
-#   add_confidence_interval() +
-  labs(
-    x = "Days",
-    y = "Overall survival probability"
-  )
-
-cox1 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + n.events.post + unscheduled_election + regional_success_lag + v2elirreg.inv , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
 summary(cox1)
@@
 print(cz1)
-plot(cz1[19])
+plot(cz1[16])
 abline(h=0)
@@
 
-cox1t <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + n.events.post + regional_success_lag + v2elirreg.inv + tt(v2elirreg.inv), data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(21, 61, 90),
+cox1.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+summary(cox1.cen)
+
+cox1t <- coxph(Surv(d.events.num, surv_status) ~ legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election +  regional_success_lag + n.events.post + v2elirreg.inv + tt(v2elirreg.inv), data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(15, 60, 90),
 Boundary.knots = FALSE))
@@
 
+cox1t2 <- coxph(Surv(d.events.num, surv_status) ~ legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election +  regional_success_lag + n.events.post + v2elirreg.inv + tt(n.events.post), data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(10, 70, 90),
+Boundary.knots = FALSE))
+summary(cox1t2)
+
 sandbox_data <- survSplit(Surv(d.events.num, surv_status) ~ 
@@
 
-cox2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
 summary(cox2)
-sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.15, 2.27]")) + coord_cartesian(ylim = c(0, 10)) #This is the 10th and 90th quantiles
+p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.05, 2]")) + coord_cartesian(ylim = c(0, 10)) #This is the 10th and 90th quantiles
+p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ration", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()
+
+cox2.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+summary(cox2.cen)
 
@@
 
-```{r}
-library(survsim) #Looking at time invariance
-zph <- cox.zph(cox1)
-zph
-
-plot(zph[19], lwd =2)
-abline(0,0, col=1, lty = 3, lwd = 2)
-abline(h = cox1$coefficients[19], col=3, lwd=2, lty=2)
-```
-
 To further test the mechanism outlined in the theory above, I examine the relationship between fraud and the survival of protest movements in cases where protests began but the incumbent eventually survived.
 Restricting the sample to these cases allows for a consistent interpretation of results: longer protest waves indicate a more even match between the resources and resiliency of the protesters and the capacity and cohesiveness of the regime, while shorter-lived protest movements are those which were more easily defeated.
-In this sub-sample, the relationship between election fraud and protest survival should be non-linear with respect to time\textemdash the longer the protest movement survives, the less relevant the fraud signal (such as it is) becomes.
-That is, any effect of fraud on protest should decline in size and and become statistically insignificant as the election recedes.
-
-Table 5 reports the results of Cox proportional hazard models of protest duration; coefficients with a negative sign indicate factors that are associated with longer-lived protest movements; those with a positive sign indicate factors associated with a speedier demise.
-
-In Model 11, all covariates' effects are modeled as time-invariant.
+If significant election fraud is indicative of regime resilience, the expectation is that more fraudulent elections will be associated with more quickly dispatched protest movements.^[Protest waves are defeated in the large majority of cases.
+There are 211 cases where protest occurs, out of which only 13 see new elections or the incumbent step down.]
+
+Table \@ref(tab:cox-models-tab) reports the results of Cox proportional hazard models of protest duration.
+In such models, coefficients with a positive sign indicate factors that are associated with shorter-lived; those with a negative sign indicate factors associated with a slower demise.
+The coefficient for *election fraud* is positive and significant in Model 7.
+Exponentiation of the coefficient for election fraud gives a hazard ratio of 3.3, indicating that for each one-unit increase in fraud, a protest movement becomes three times more likely to end at any given time.
+This supports H3b, and fails to support H3a.
+
+Further investigation of the relationship between fraud and protest duration yields additional insights.
+In Model 8, incumbent vote-share is interacted with *election fraud*.
+Figure \@ref(fig:margin-dur-fig) shows the marginal effect.
+As shown in the figure, for all incumbent vote-shares near and below fifty percent, large-scale fraud significantly increases the likelihood that a protest movement will end.
+This result also supports the fraud-as-strength model, and shows that only as vote-share increases above fifty percent does a government in a low-fraud election have the same likelihood of defeating a protest movement at a given time.
+
+```{r margin-dur-fig, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
+p.duration.margin
+```
+
+The relationship between the level of fraud and the survival of protest movements can also be demonstrated non-parametrically using Kaplan-Meier curves for varying levels of fraud.
+In Figure \@ref(km-fig), observations are divided into three groups.
+High fraud observations are those where *election fraud* is scored at one standard deviation above the mean or more, low fraud observations are those scored at one standard deviation below the mean or less, with moderate levels in between.
+The figure shows a clear difference between the low-fraud group and the other two.
+Low-fraud cases are much more likely to see their protest movements survive past the first day than their counterparts.
+For much of the range of the data, they remain roughly twice as likely to survive at any given date than protest movements in medium- or high-fraud cases.
+This is evidence in favor of the idea that high levels of fraud is an indication of regime resilience, and contradicts the claim that fraud either reveals weakness or galvanizes political opposition, as it is the low-fraud cases where governments struggle to put an end to protest movements on average.
+
+```{r km-fig, fig.cap = "Kaplan-Meier curves for protest duration and high, low, and medium levels of fraud", fig.height= = 7, echo=F, message=F, fig.pos='H'}
+nelda2.sub <- nelda2.sub %>% mutate(fraud_level = ifelse(v2elirreg.inv >= (.98 + .97), "High",
+                                                            ifelse(v2elirreg.inv <= (.98 - .97), "Low", "Moderate")))
+
+survfit2(Surv(d.events.num, surv_status) ~ fraud_level, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
+  ggsurvfit() +
+#   add_confidence_interval() +
+  labs(
+    x = "Days",
+    y = "Overall survival probability"
+  )
+```
+
+In Model 7, all covariates' effects are modeled as time-invariant.
 However, diagnostics of the model suggest that this assumption may not be upheld for *election fraud*.^[Schoenfeld residuals for *election fraud* are not independent of time, as calculated using the `survival` package in R [@survival-package].] To adjust for this, Model 12 allows the coefficient for *election fraud* to take on different values at different points in time.
@@
 models3 <- list()
-models3[['Model 11']] <- cox1
-models3[['Model 12']] <- cox1t
-#models3[['Model 14']] <- cox3
+models3[['Model 7']] <- cox1
+models3[['Model 8']] <- cox2
+models3[['Model 9']] <- cox1t
 
@@
 
-This is an attempt at modeling endogenous relationships using two-step residual inclusion regression:
-
-```{r}
-nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(n.events.post > 0, 1, 0))
-
-test_model <- OneSampleMR::tsri(any_protest ~ v2elirreg.inv | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +   osorg.1lag  + natlpartyorg.lag  +  presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = nelda2.sub, link = "logit")
-summary(test_model)
-
-test_model2 <- OneSampleMR::tsri(any_protest ~ winner.share.upd | v2elirreg.inv +  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +   osorg.1lag  + natlpartyorg.lag  +  presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, data = nelda2.sub, link = "logit")
-summary(test_model2)
-
-```
-
 Predicting post-election protest counts and getting residuals:
 
-```{r}
+```{r, include=FALSE}
 test3 <- glm.nb(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag +  + osorg.1lag  + natlpartyorg.lag +   inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag, data = nelda2.sub)
@@
 
-```{r}
+```{r, include=FALSE}
 nelda2.sub <- nelda2.sub %>% mutate(inc.confident.num = ifelse(inc.confident == "N/A", NA, ifelse(inc.confident == "yes", 1, 0)))
@@
 
-```{r}
-mydf <- ggpredict(model_protest2, type = "count", terms = ".resid [all]", back.transform = T)
-ggplot(mydf, aes(x, predicted)) +
-  geom_line() +
-  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .1) + coord_cartesian(ylim = c(0, 100))
-
-library(pscl)
-model_protest2b <- zeroinfl(n.events.post ~  .resid + winner.share.upd | ., data = testdata %>% dplyr::select(n.events.post, .resid, winner.share.upd), dist = "negbin")
-summary(model_protest2b)
-
-newdata1 <- expand.grid((seq(from = -1.2, to = 1.3, by = .1)), mean(testdata$winner.share.upd, na.rm=T))
-colnames(newdata1) <- c(".resid", "winner.share.upd")
-newdata1$phat <- predict(model_protest2b, newdata1)
-
-ggplot(newdata1, aes(x = .resid, y = phat)) +
-  geom_point() +
-  geom_line() +
-   labs(x = "Fraud relative to baseline", y = "Predicted protest events")
-
-model_protest2c <- zeroinfl(n.events.post ~  .resid*winner.share.upd + .resid + winner.share.upd  | ., data = testdata %>% dplyr::select(n.events.post, .resid, winner.share.upd), dist = "negbin")
-summary(model_protest2c)
-
-newdata2 <- expand.grid((seq(from = -1.2, to = 1.3, by = .1)), seq(from = 0, to = 1, by =.25))
-colnames(newdata1) <- c(".resid", "winner.share.upd")
-newdata1$phat <- predict(model_protest2c, newdata1)
-
-ggplot(newdata2, aes(x = .resid, y = phat, colour = factor(winner.share.upd))) +
-  geom_point() +
-  geom_line() +
-   labs(x = "Fraud relative to baseline", y = "Predicted protest events")
-
-cox1 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + v2elirreg.inv , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
-summary(cox1)
-
-```
-
-```{r}
+```{r, include=FALSE}
 nelda2.sub <- nelda2.sub %>% group_by(country) %>% mutate(elirreg_mean = mean(v2elirreg.inv, na.rm=T))
@@
 
-```{r}
-#Lots of missinginess in winner share
-
-winner_missing <- nelda2.sub %>% filter(is.na(winner.share.upd) == T) %>% dplyr::select(country, year)
-```
-
-What I'm looking for here is weak government -> high fraud + narrow vote share -> protest. Not sure if that pattern reveals itself.
-```{r}
-ggplot(nelda2.sub %>% filter(presidential == 1), aes(x = winner.share.upd, y = v2elirreg.inv)) + geom_point(aes(size=n.events.post)) + geom_vline(xintercept = .5, linetype = 2) + geom_hline(yintercept = 0)
-```
-
 \newpage

### 2024-01-31 Scripts/authoritarian_only.Rmd
Updates to Cox results and discussion

@@
 
-protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1)
+protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1 & inc.lose == 0)
 
@@
 nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
-                                                                       ifelse(d.events.num >= 62, 0, 1)))
+                                                                       ifelse(d.events.num >= 69, 0, 1)))
 
@@
 
-sandbox_data <- survSplit(Surv(d.events.num, surv_status) ~ 
-                      jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + fraud_level + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
-                      data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), cut = c(2), episode = "tgroup")
-
-cox1.2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + fraud_level*(as.factor(tgroup)) + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = sandbox_data)
-summary(cox1.2)
-
 cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
@@
 If significant election fraud is indicative of regime resilience, the expectation is that more fraudulent elections will be associated with more quickly dispatched protest movements.^[Protest waves are defeated in the large majority of cases.
-There are 211 cases where protest occurs, out of which only 13 see new elections or the incumbent step down.]
+There are 156 cases where protest occurs, out of which only 9 see new elections or the incumbent step down.]
 
 Table \@ref(tab:cox-models-tab) reports the results of Cox proportional hazard models of protest duration.
-In such models, coefficients with a positive sign indicate factors that are associated with shorter-lived; those with a negative sign indicate factors associated with a slower demise.
-The coefficient for *election fraud* is positive and significant in Model 7.
+In such models, coefficients with a positive sign indicate factors that are associated with shorter-lived protests; those with a negative sign indicate factors associated with a slower demise.
+The coefficient for *election fraud* is positive and significant in Model 7, indicating that more fraudulent elections are associated with more quickly dispatched protest movements.
 Exponentiation of the coefficient for election fraud gives a hazard ratio of 3.3, indicating that for each one-unit increase in fraud, a protest movement becomes three times more likely to end at any given time.
 This supports H3b, and fails to support H3a.
+
+One threat to inference in this design derives from the 90-day post-election window considered by ECAV.
+It is possible that some protest movements considered 'dead' in Models 7 and 8 are, in fact, 'alive;' it is just that subsequent protests occurred after 90 days and are censored as a result.
+As a test of the robustness of the conclusions, in Models 9 and 10, observations are assumed to be right-censored if a protest occurred within three-weeks of the 90-day cutoff.
+This amounts to 15 of the 156 observations in the sample.
+The results remain consistent with those from the uncensored models.
+While the coefficient on *election fraud* falls slightly short of conventional statistical significance in Model 9, it retains the same size and sign compared to Model 7.
+Similarly, the interaction effect in Model 10 increases in size, while retaining its direction and statistical significance.
+This suggests that the main models are not drawing overly incorrect inferences from longer-lasting protest movements.
 
@@
 
-In Model 7, all covariates' effects are modeled as time-invariant.
-However, diagnostics of the model suggest that this assumption may not be upheld for *election fraud*.^[Schoenfeld residuals for *election fraud* are not independent of time, as calculated using the `survival` package in R [@survival-package].] To adjust for this, Model 12 allows the coefficient for *election fraud* to take on different values at different points in time.
-Based on the diagnostics, I allow the value of the coefficient to vary for three periods: 1 - 11 days, 12 - 25 days, and 26 days to the maximum value.
-Model 12 shows the result of this analysis; in it, *election fraud* has a positive and significant coefficient for the earliest time period.
-This indicates that, for the first 11 days of a protest movement, states that engaged in more election fraud are more likely to be able to quash the protests (whether by repression or accommodation).
-This is contrary to the expectations of deterrence models.
-After this period, there is no significant relationship between the level of fraud and the survival of the protest wave.
-That is, if a protest movement has survived longer than (roughly) a week and a half after the election, the severity of fraud does not influence its continued duration.
-It is also noteworthy that this formulation is a much better fit to the data than the linear effect in Model 11, with a much lower root mean square error.
-
-These results indicate that the influence of a fraudulent election on the duration of mass protest diminishes over time.
-It also suggests that most protests are failed bets.
-Upon taking to the streets, protesters are more likely to find themselves facing a capable incumbent regime the more fraudulent the election was.
-Finally, if protests survive beyond the short term, the severity of fraud no longer has any effect.
-Instead, structural indicators\textemdash the strength of the opposition in the legislature, the degree to which parties are organized across the country, and (to a lesser extent) the openness of civil society are all associated with longer-lived protest movements.
-Similarly, a more fractured incumbent regime, as indicated by a term-limited incumbent, also make protest survival more likely.
-All of this is supportive of the theoretical model laid out above.
-
 ```{r cox-models-tab, echo=FALSE, warning=FALSE}
@@
 models3[['Model 8']] <- cox2
-models3[['Model 9']] <- cox1t
+models3[['Model 9']] <- cox1.cen
+models3[['Model 10']] <- cox2.cen
 
@@
 
-Altogether, the findings challenge the idea that the fear of protest helps deter election manipulation.
-To the extent that fraud is a signal of incumbent strength, it is an unreliable one.
-Since even strong incumbents commit fraud, the decision to take to the streets is often a losing gamble\textemdash anticipating a weak incumbent, they instead find one that is capable of repressing, coopting, or conciliating an emerging protest movement.
+Altogether, the findings are a challenge to several prior assumptions and findings in the literature on election fraud, non-democratic elections, and authoritarian stability.
+These results do not suggest that election fraud is used as a tool of weak incumbents, in fact just the opposite.
+When encountering a more precarious pre-election environment\textemdash due to higher levels of pre-election protest or a successful post-election protest in the region, for instance\textemdash governments appear to respond by producing higher levels of election fraud.
+On its own, this is a potentially significant result which suggests that governments are not deterred from engaging in election fraud by the risk of post-election protest, as numerous models predict.
+
+More compellingly, this project finds that when post-election protests *do* occur, they are shorter-lived in for cases where fraud was higher.
+This indicates that governments which produced high levels of fraud were more able to quash those protest movements which did emerge more quickly than governments that produced low levels of fraud.
+This is contrary to three prevailing wisdoms.
+The first is that weak governments manipulate more than strong, the second is that more election fraud is a grievance that mobilizes regime antagonists, and the third is that fraud can only deter protest if it is excessive or blatant.
+
 Indeed, while there are 194 cases of protest out of the 661 elections in this dataset where incumbents claimed victory, only 8 resulted in either the incumbent being replaced or the election being re-run due to mass protests.\footnote{This information is determined using the NELDA variables NELDA37 and NELDA41.} This suggests a `success rate' for post-election protest of about four percent *conditional on protest initiation*, and about one percent for all elections in electoral authoritarian and electoral democratic regimes.

### 2024-02-07 Output/appendix_v4.Rmd
Updates to appendix--removing MM

@@
 
-nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2023-05-30.csv"))
+nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2024-02-06.csv"))
 
@@
 
-##Removing cases before 1990, when ECAV starts
-nelda2.sub <- nelda2.sub %>% filter(year >= 1990)
-
 ## Removing cases where inc. loses
-#nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)
-
-##Updated winner share
-nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))
+#nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)  #See how this affects things
 
@@
 
-#Crisis as factor
-
-nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))
-
-#Margins with updates
-
-nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))
-
-#Successful protest = new elections or new incumbent
-
-nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))
-
-#Protest cases only
-
-protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1)
-
-# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections
-
-nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))
-
 ## Categorical variable for incumbent status
@@
 
-## Coding max_size as zero for elections with no protest
-nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))
+#Coding as factors
 
-## Loading massmob
+nelda2.sub <- nelda2.sub %>% mutate(presidential = factor(presidential))
+#nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = factor(inc.termlimit.num))
+nelda2.sub <- nelda2.sub %>% mutate(rentierstate = factor(rentierstate))
+nelda2.sub <- nelda2.sub %>% mutate(taxstate = factor(taxstate))
+nelda2.sub <- nelda2.sub %>% mutate(unscheduled_election = factor(unscheduled_election))
+nelda2.sub <- nelda2.sub %>% mutate(inc.lose = factor(inc.lose))
+nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))
 
-massmob <- read.csv(here::here("Data", "nelda_massmob_all.csv"))
-massmob <- massmob %>% rename("NeldaID" = "electionid")
-massmob <- massmob %>% dplyr::select(NeldaID, n.events.massmob.pre, n.events.massmob.post)
+#Successful protest = new elections or new incumbent
 
-nelda2.sub <- nelda2.sub %>% left_join(massmob, by = "NeldaID")
+nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))
 
-```
+# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections
 
-# Alternative protest data: Mass Mobilization dataset
+nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))
 
-```{r models-hurdle-mm, message = FALSE, echo=FALSE}
-nelda2.sub <- nelda2.sub %>% mutate(n.events.massmob.pre.scale = scale(n.events.massmob.pre))
+## Replacing NA with 0 for protest size
 
-model.n.base.hurd.mm <- pscl::hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +     winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre +  unscheduled_election + regional_success_lag,
-                                         zero.dist = "binomial",
-                                          dist= "negbin",
-                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
-#summary(model.n.base.hurd.mm)
+#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(is.na(max_size)==T & n.events.post == 0, 0, max_size))
 
-model.n.base2.hurd.mm <- pscl::hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +     winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre +  unscheduled_election + regional_success_lag,
-                                         zero.dist = "binomial",
-                                          dist= "negbin",
-                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
-#summary(model.n.base2.hurd.mm)
+## Fixing missing values of winner.share.upd for liberal democracies
 
-model.n.griev.hurd.mm <- pscl::hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +     gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre +  unscheduled_election + regional_success_lag,
-                                         zero.dist = "binomial",
-                                          dist= "negbin",
-                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
-#summary(model.n.griev.hurd.mm)
+nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(is.na(winner.share.upd)==T & v2x_regime == 3, winner.margin, winner.share.upd))
 
-```
+## Set max_size zero for elections with no protest
 
-The Mass Mobilization dataset [@massmob] aims to collect data on every protest event with greater than 50 participants in over 160 countries, from 1990 to 2014.
-Since the 50-participant threshold undercounts small protests, I do not use it in the main text and I do not use it here to test hypotheses on protest size.
-To match the ECAV format, I limit the window for protests to 180 days prior to the election and 90 days after.
-Unlike with ECAV, the counted protests need not be explicitly tied to the election.
-Using Mass Mobilization data generates results similar to those in the main text, as shown in Tables \@ref(tab:table-mm-binary) and \@ref(tab:table-mm-count).
-Fraud severity has no standalone effect on protest initiation or protest count.
-Election fraud does have a significant relationship to protest initiation in Model 2, at low incumbent vote-shares, but as Figure \@ref(fig:plot-margin-mm) indicates, this relationship is not apparent at typical levels of election fraud.
-As in the main text, the main drivers of protest count appear to be institutional and structural\textemdash repressive capacity and opposition organization.
+#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))
 
-```{r table-mm-binary, echo=FALSE, warning=FALSE, message=FALSE}
-models_a1 <- list()
-models_a1[['Model 1']] <- model.n.base.hurd.mm
-models_a1[['Model 2']] <- model.n.base2.hurd.mm
-models_a1[['Model 3']] <- model.n.griev.hurd.mm
+## Excluding liberal democracies
 
-modelsummary::modelsummary(models_a1, stars = T, longtable=T, coef_omit = "count", title = 'Binary portion of hurdle models of post-election protest', coef_rename = 
-               c("zero_(Intercept)" = "Intercept", "zero_jucon.1lag" = "Judical independence (lag)", "zero_legcon.1lag" = "Leg. constraints (lag)", 
-                 "zero_gdpgro.1lag" = "GDP growth rate (lag)", "zero_altinf.1lag" = "Alternative info. (lag)", 
-                 "zero_osorg.1lag" = "Civil soc. openness (lag)", "zero_econ.crisis.nelda.1lag1" = "Economic crisis", 
-                 "zero_natlpartyorg.lag" = "National party organization (lag)", "zero_presidential" = "Presidential election", "zero_libdem.1lag" = "Liberal democracy index",
-                 "zero_inc.termlimit.num" = "Incumbent term-limited",  "zero_regional_success_lag" = "Regional diffusion",
-                 "zero_urban.pct" = "Urbanization", "zero_rentierstate" = "Rentier state", "zero_taxstate" = "Tax state",
-                 "zero_n.events.massmob.pre" = "Number of pre-election protests", "zero_v2elirreg.inv" = "Election fraud",
-                 "zero_physinteg.1lag" = "Physical integrity (lag)", "zero_physintlag.sq" = "Physical integrity (lag), squared", "zero_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "zero_unscheduled_election" = "Off-schedule election"))
-```
+nelda2.sub <- nelda2.sub %>% filter(v2x_regime == 1 | regime.1lag == 1) #Includes only authoritarian, add e-dems with | v2x_regime == 2  #Current version includes lagged 1 in order to capture pre-election auth.
 
-\newpage
+## Remove outlier
+nelda2.sub <- nelda2.sub %>% filter(d.events.num <= 100)
 
-```{r table-mm-count, echo=FALSE, warning=FALSE, message=FALSE}
-modelsummary::modelsummary(models_a1, stars = T, longtable=T, coef_omit = "zero", title = 'Count portion of hurdle models of post-election protest', coef_rename = 
-               c("count_(Intercept)" = "Intercept", "count_jucon.1lag" = "Judical independence (lag)", "count_legcon.1lag" = "Leg. constraints (lag)", 
-                 "count_gdpgro.1lag" = "GDP growth rate (lag)", "count_altinf.1lag" = "Alternative info. (lag)", 
-                 "count_osorg.1lag" = "Civil soc. openness (lag)", "count_econ.crisis.nelda.1lag1" = "Economic crisis", 
-                 "count_natlpartyorg.lag" = "National party organization (lag)", "count_presidential" = "Presidential election", "count_libdem.1lag" = "Liberal democracy index",
-                 "count_inc.termlimit.num" = "Incumbent term-limited",  "count_regional_success_lag" = "Regional diffusion",
-                 "count_urban.pct" = "Urbanization", "count_rentierstate" = "Rentier state", "count_taxstate" = "Tax state",
-                 "count_n.events.massmob.pre" = "Number of pre-election protests", "count_v2elirreg.inv" = "Election fraud",
-                 "count_physinteg.1lag" = "Physical integrity (lag)", "count_physintlag.sq" = "Physical integrity (lag), squared", "count_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "count_unscheduled_election" = "Off-schedule election"))
-```
+#Protest cases only
 
-```{r plot-margin-mm, echo=FALSE, message=FALSE, warning=F, fig.cap="Fraud, winning vote-share, and probability of protest initiation (Mass Mobilization data)", out.width = '85%', fig.pos="h!"}
-nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(is.na(n.events.massmob.post)==T, NA, ifelse(n.events.massmob.post > 0, 1, 0)))
+protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1 & inc.lose == 0)
 
-model.margin.binary_n <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
-#summary(model.margin.binary_n)
+## Coding median size as 0 when n.events = 0
 
-plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary_n, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.64, 1.55]"))
-plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()
+nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(n.events.post == 0, 0, median.participants))
 
-## Count portion
-
-dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.64, 1.55]"), type = "count") 
-p.n.margin <- plot(dat)
-p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Number of events", title = "Predicted number of protests", color = "Election fraud") + theme_bw()
-
-plot.combined.prob <- cowplot::plot_grid(plot_binary_margin_typ_n, p.n.margin, ncol=1)
-plot.combined.prob
-
-```
-
-```{r, include=FALSE}
-model.binary.griev1_n <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +  +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
-#summary(model.binary.griev1_n)
-
-plot_binary_gdp_typ_n <- sjPlot::plot_model(model.binary.griev1_n, type = "pred", terms = c("gdpgro.1lag [all]", "v2elirreg.inv [-.45, 1.78]"))
-plot_binary_gdp_typ_n <- plot_binary_gdp_typ_n +  labs(x = "GDP growth (lagged)", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw() + coord_cartesian(xlim = c(-.4, 3))
-
-## Count portion
-
-dat <- ggpredict(model.n.griev.hurd.mm, terms = c("gdpgro.1lag", "v2elirreg.inv [-.64, 1.55]"), type = "count") 
-p.n.gdp <- plot(dat)
-p.n.gdp <- p.n.gdp +  labs(x = "GDP growth rate (lag)", y = "Number of events", title = "Predicted number of protests", color = "Election fraud") + theme_bw() + coord_cartesian(xlim = c(-.4, 3), ylim = c(0, 200))
 ```
@@
 the possibility that in highly repressive states, protest is deterred and fraud can proceed unopposed.
-However, it may still be useful to check an alternative explanation\textemdash perhaps protest is deterred in highly repressive states with considerable pre-election unrest (that is, when protest appears to be exceeding the ability of the state to contain).
+However, it may still be useful to check an alternative explanation\textemdash perhaps manipulation is deterred in highly repressive states with considerable pre-election unrest (that is, when protest appears to be exceeding the ability of the state to contain it).
 To check this possibility, a variation of Model 1 from the main text is run, but with an interaction between the number of pre-election protests and the lagged *physical integrity* variable.
@@
 Figure \@ref(fig:prot-rep-fraud) shows the absence of such an effect.
-In fact, there is no significant difference in the predicted level of fraud at lower levels of physical integrity, even when pre-election protest is widespread.
-A significant effect is seen at the high end of protection of physical integrity rights, but it is present only in extreme cases\textemdash the value of pre-election protests chosen here, 25, is above the 98th percentile for the variable\textemdash and is in the opposite direction predicted by this counterargument.
-That is, more numerous pre-election protest is associated with higher predicted levels of fraud in less repressive states than when the pre-election period is quiet.
-This robustness check shows that protest in a repressive environment does not appear to deter fraud, and supports the view that manipulation may be used to attempt to claim a large margin of victory or to prevent elite or bureaucratic splits in a challenging period.
+In fact, there is no significant change in the predicted level of fraud at lower levels of physical integrity, even when pre-election protest is widespread.
+We do observe a significant effect when comparing high- and low-repression regimes in highly extreme cases\textemdash there are only three cases in the dataset with more than 70 protest events\textemdash and is in the opposite direction predicted by this counterargument.
+That is, more numerous pre-election protest is associated with higher predicted levels of fraud in less repressive states, not harsher ones.
+The figure also shows that less repressive regimes appear to respond to greater civic unrest with more fraud.
+Altogether, this robustness check shows that pre-election protest in a repressive environment does not appear to deter fraud, and supports the view that manipulation may be used to attempt to claim a large margin of victory or to prevent elite or bureaucratic splits in a challenging period.
 
-```{r prot-rep-fraud, message=FALSE, warning=F, echo=FALSE, fig.cap="Marginal effect of pre-election protest on fraud, conditional on physical integrity", fig.pos="h!"}
-m1_rep <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre*physinteg.1lag  +  unscheduled_election  + (1 | stateid), data = nelda2.sub %>% filter(v2x_regime <= 2))
+```{r prot-rep-fraud, message=FALSE, warning=F, echo=FALSE, fig.cap="Marginal effect of pre-election protest on fraud, conditional on physical integrity", out.width = '75%', fig.align='center'}
+m1_rep <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + unscheduled_election  + v2elembaut + physinteg.1lag*n.events.pre + (1 | stateid), data = nelda2.sub)
 #summary(m1_rep)
-p.m1rep <- sjPlot::plot_model(m1_rep, type = "pred", terms = c("physinteg.1lag", "n.events.pre [0, 25]"))
-p.m1rep + labs(title = "Predicted values of election fraud", x = "Physical integrity (lag)", y = "Election fraud", color = "N. of pre-election protests") + theme_bw()
+p.m1rep <- sjPlot::plot_model(m1_rep, type = "pred", terms = c("n.events.pre", "physinteg.1lag [.25, .63]"))
+p.m1rep + labs(title = "Predicted values of election fraud", x = "Number of pre-election protests", y = "Election fraud", color = "Physical integrity (lag)") + theme_bw()
 ```
@@
 Since the prior election's value is subtracted from the 'current' one, positive values mean that the current election was more fraudulent than its predecessor, while negative values indicate election fraud declined.
-Using this conceptualization of election fraud does result in some support for the deterrence thesis: an election that is markedly more fraudulent than the prior election does appear associated with more numerous protest events at higher incumbent vote-shares.
-However, as in other examples, this effect only becomes statistically significant at the extremes.
-The values for change in fraud used in the marginal effects plots in Figures \@ref(fig:n-protests-ch) and \@ref(fig:size-protests-ch) represent very unusual changes, roughly between the 5th and 95th percentiles.
-In other words, the expected number and duration of protest events are only statistically significantly different when comparing elections that are far cleaner than the previous election and those that are far worse\textemdash a set of options that is unlikely to confront most incumbents.
+As in other examples, there is no statistically significant for typical values.
+The values for change in fraud used in the marginal effects plots in Figures \@ref(fig:n-protests-ch) and \@ref(fig:size-protests-ch) represent large, but not unusual changes in election fraud from election to election: roughly one standard deviation above and below the mean.
+As shown in Figure \@ref(fig:n-protests-ch), the expected number of protests does appear to decline for significantly cleaner elections when comparing very low vs.
+very high incumbent vote-shares.
+However, there is no significant change for more fraudulent elections\textemdash precisely where the deterrent logic would expect to see an effect.
+Moreover, the two types of elections show no significant difference between them at any level of incumbent vote-share.
 
@@
 
-model.n.base2.lag <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + fraud.change*winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
-                                          zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub %>% filter(v2x_regime <= 2))
+model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + fraud.change +  +    winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+summary(model.margin.binary)
+
 #summary(model.n.base2.lag)
@@
 
-model.size.base2.lag <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  fraud.change*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2),
-                                family=truncated_nbinom2)
-#summary(model.size.base2.lag)
+model.n.base2.lag <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + fraud.change*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+#summary(model.n.base2.lag)
 
@@
 t4a <- huxreg(models3, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Fraud change" = "fraud.change", "Incumbent vote-share" = "winner.share.upd", "Fraud change:Incumbent vote-share" = "fraud.change:winner.share.upd"),
-              statistics = character(0)) %>% set_caption("Models of number, duration, and size of protests") %>% insert_row("", "Number", "Size", after = 0) %>% insert_row("Controls", "Yes", "Yes", after = 10)
+              statistics = character(0)) %>% set_caption("Models of number and size of protests") %>% insert_row("", "Number", "Size", after = 0) %>% insert_row("Controls", "Yes", "Yes", after = 10)
 theme_compact(t4a, header_rows = TRUE, header_cols = FALSE)  %>% set_bottom_border(10, everywhere)
@@
 
-```{r n-protests-ch, fig.cap="Change in fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE, fig.pos="h!"}
-dat <- ggpredict(model.n.base2.lag, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "fraud.change [-1.12, 1.12]"), type = "count") 
+```{r n-protests-ch, fig.cap="Change in fraud, incumbent vote-share, and number of protest events", out.width = '75%', echo=FALSE, message=FALSE, fig.align='center'}
+dat <- ggpredict(model.n.base2.lag, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "fraud.change [-.58, .58]"), type = "count") 
 p.n.margin.ch <- plot(dat)
@@
 
-```{r size-protests-ch, fig.cap="Change in fraud, incumbent vote-share, and size of protest events", echo=FALSE, message=FALSE, fig.pos="h!"}
-dat <- ggpredict(model.size.base2.lag, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "fraud.change [-1.22, 1.22]"), type = "count") 
+```{r size-protests-ch, fig.cap="Change in fraud, incumbent vote-share, and size of protest events", out.width = '75%', echo=FALSE, message=FALSE, fig.align='center'}
+dat <- ggpredict(model.size.base2.lag, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "fraud.change [-.58, .58]"), type = "count") 
 p.s.margin.ch <- plot(dat)
-p.s.margin.ch <- p.s.margin.ch +  labs(x = "Incumbent vote-share", y = "Size of protest wave", title = "", color = "Change in fraud") + theme_bw() + coord_cartesian(ylim = c(0, 70000))
+p.s.margin.ch <- p.s.margin.ch +  labs(x = "Incumbent vote-share", y = "Size of protest wave", title = "", color = "Change in fraud") + theme_bw() + coord_cartesian(ylim = c(0, 20000))
 p.s.margin.ch 
@@
 
-```{r n-protests-concerns, fig.cap="Election fraud, pre-election expectations, and number of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
+```{r n-protests-concerns, fig.cap="Election fraud, pre-election expectations, and number of protest events", echo=FALSE, out.width = '75%', fig.align='center'}
 p.n.concerns
@@
 
-```{r protests-size-concerns, fig.cap="Election fraud, pre-election expectations, and size of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
+```{r protests-size-concerns, fig.cap="Election fraud, pre-election expectations, and size of protest events", echo=FALSE, out.width = '75%', fig.align='center'}
 p.size.concerns
@@
 Using a variable from NELDA that codes for designated successors, the following models add a categorical variable for the incumbent's status, which can take on the following values: *executive eligible*, *chosen successor*, *contested executive election*, and *legislative election*.
-As Figure \@ref(fig:successor-fig) show, the results are largely similar to those in the main text.
-Higher levels of fraud does make the occurrence of protest more likely at low incumbent vote-shares, but this does not translate into a higher risk of more numerous or larger protest events.
 
@@
 
-plot_binary_margin_succ_n <- sjPlot::plot_model(model.n.base2.succ.bin, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]"))
+plot_binary_margin_succ_n <- sjPlot::plot_model(model.n.base2.succ.bin, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]")) #To set condition for successor value, use , condition = c(incumb.status = "Contested exec.
+election")
 plot_binary_margin_succ_n <- plot_binary_margin_succ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw()
@@
 
-t5a <- huxreg(models4, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Models of number and size of protests") %>% insert_row("", "Number", "Size", after = 0)
-#theme_compact(t5a, header_rows = TRUE, header_cols = FALSE)
+t5a <- huxreg(models4, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, coefs = c("Legislative election" = "incumb.statusLegislative election", "Chosen successor" = "incumb.statusChosen successor", "Contested exec.
+election" = "incumb.statusContested exec.
+election", "Intentional irregularities" = "v2elirreg.inv", "Incumbent vote-share" = "winner.share.upd", "Intentional irregularities : Incumbent vote-share" = "v2elirreg.inv:winner.share.upd"),  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Models of number and size of protests") %>% insert_row("", "Number", "Size", after = 0) %>% insert_row("Controls", "Yes", "Yes", after = 14)
+theme_compact(t5a, header_rows = TRUE, header_cols = FALSE)
 
@@
 
-```{r successor-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
+As Figure \@ref(fig:successor-fig) show, the results are largely similar to those in the main text.
+For these figures, *incumbent status* was set to its modal value\textemdash an executive election in which the incumbent is eligible.
+In this model, higher levels of fraud reduce the likelihood of no protest following the election, at low levels of incumbent vote-share..
+This is in line with the expectation of the deterrence logic.
+However, this does not translate into a higher risk of more numerous or larger protest events by level of manipulation.
+Table \@ref(tab:successor-table), abridged here to show the key variables shows that there is no effect of successor type on the size of electoral protest.
+On the other hand, elections where there is a chosen successor appear to prompt more numerous protest than when the incumbent runs again.
+This could indicate dissatisfaction among the losing faction in the internal struggle for succession, and/or unhappiness among the democratic opposition surrounding a choreographed transition.
+
+\newpage
+
+```{r successor-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=6, echo=FALSE, message=FALSE, fig.align='center'}
 plot.combined.prob <- cowplot::plot_grid(plot_binary_margin_succ_n, p.n.margin, p.s.margin, ncol=1)
@@
 
-t6b <- huxreg(models5a, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd"), statistics = character(0)) %>% set_caption("Models of number and size of protests, only electoral authoritarian regimes") %>% insert_row("", "Number", "Size", after = 0)
-theme_compact(t6b, header_rows = TRUE, header_cols = FALSE)  %>% insert_row("Controls", "Yes", "Yes", after = 10) %>% set_bottom_border(10, everywhere)
+t6b <- huxreg(models5a, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Incumbent term-limited" = "inc.termlimit.num", "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd"), statistics = character(0)) %>% set_caption("Models of number and size of protests, only electoral authoritarian regimes") %>% insert_row("", "Number", "Size", after = 0)
+theme_compact(t6b, header_rows = TRUE, header_cols = FALSE)  %>% insert_row("Controls", "Yes", "Yes", after = 12) %>% set_bottom_border(12, everywhere)
 ```
@@
 p.n.margin.auth <- plot(dat)
-p.n.margin.auth <- p.n.margin.auth + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events")  + theme_bw()
+p.n.margin.auth <- p.n.margin.auth + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events", color = "Election fraud")  + theme_bw()
 
@@
 p.size.margin.auth <- plot(dat)
-p.size.margin.auth <- p.size.margin.auth + labs(x = "Incumbent vote-share", y = "Predicted size of protests" , title = "Predicted maximum of protest events")  + theme_bw()
+p.size.margin.auth <- p.size.margin.auth + labs(x = "Incumbent vote-share", y = "Predicted size of protests" , title = "Predicted maximum size of protest events", color = "Election fraud")  + theme_bw()
 
@@
 
-```{r n-protests-margin-auth, fig.cap="Election integrity, incumbent vote-share, and number of protest events in authoritarian regimes", echo=FALSE, out.width = '70%', fig.align='center'}
+```{r n-protests-margin-auth, fig.cap="Election fraud, incumbent vote-share, and number of protest events in authoritarian regimes", echo=FALSE, out.width = '75%', fig.align='center', fig.pos="h!"}
 p.n.margin.auth
@@
 
-```{r size-protests-margin-auth, fig.cap="Election integrity, incumbent vote-share, and maximum size of protest events in authoritarian regimes", echo=FALSE, out.width = '70%', fig.align='center'}
+```{r size-protests-margin-auth, fig.cap="Election fraud, incumbent vote-share, and maximum size of protest events in authoritarian regimes", echo=FALSE, out.width = '75%', fig.align='center', fig.pos="h!"}
 p.size.margin.auth
@@
 Election fraud does not appear associated with a higher number of protest events, or larger protest sizes.
+Results are presented in Table \@ref(tab:lib-table) and Figures \@ref(fig:n-protests-margin-lib) and \@ref(fig:size-protests-margin-lib)
 
@@
 
+model.n.griev.lib <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+                                          zi = ~.,
+                                          family= nbinom2,
+                                          data = nelda2.sub)
+#summary(model.n.base2.lib)
+
+model.size.griev.lib <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
+                                          zi = ~.,
+                                          family= nbinom2,
+                                          data = nelda2.sub)
+
 models5 <- list()
@@
 models5[['Model 11']] <- model.size.base2.lib
+models5[['Model 12']] <- model.n.griev.lib
+models5[['Model 13']] <- model.size.griev.lib
 
-t6a <- huxreg(models5, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd"), statistics = character(0)) %>% set_caption("Models of number and size of protests, inclu.
-liberal democracies") %>% insert_row("", "Number", "Size", after = 0)
-theme_compact(t6a, header_rows = TRUE, header_cols = FALSE)  %>% insert_row("Controls", "Yes", "Yes", after = 10) %>% set_bottom_border(10, everywhere)
+t6a <- huxreg(models5, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd",
+                                                                                                                                                             "Election fraud : Winner vote-share" = "winner.share.upd:v2elirreg.inv",
+                                                       "GDP growth (lag)" =  "gdpgro.1lag",
+                          "Election fraud : GDP growth (lag)"  =  "gdpgro.1lag:v2elirreg.inv"), statistics = character(0)) %>% set_caption("Models of number and size of protests, inclu.
+liberal democracies") %>% insert_row("", "Number", "Size", "Number", "Size", after = 0)
+theme_compact(t6a, header_rows = TRUE, header_cols = FALSE)  %>% insert_row("Controls", "Yes", "Yes", "Yes", "Yes", after = 14) %>% set_bottom_border(14, everywhere)
+```
+
+```{r, echo=FALSE, message=FALSE}
+dat <- ggpredict(model.n.base2.lib, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [.35, 1.86]"), type = "count") 
+p.n.margin.lib <- plot(dat)
+p.n.margin.lib <- p.n.margin.lib + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events", color = "Election fraud")   + theme_bw()
+
+dat <- ggpredict(model.size.base2.lib, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [.35, 1.86]")) 
+p.size.margin.lib <- plot(dat)
+p.size.margin.lib <- p.size.margin.lib + labs(x = "Incumbent vote-share", y = "Predicted size of protests" , title = "Predicted maximum of size protest events", color = "Election fraud")  + theme_bw()
+
+```
+
+```{r n-protests-margin-lib, fig.cap="Election integrity, incumbent vote-share, and number of protest events (including liberal democracies)", echo=FALSE, out.width = '75%', fig.align='center', fig.pos="h!"}
+p.n.margin.lib
+
+```
+
+```{r size-protests-margin-lib, fig.cap="Election integrity, incumbent vote-share, and size of protest events (including liberal democracies)", echo=FALSE, out.width = '75%', fig.align='center', fig.pos="h!"}
+p.size.margin.lib
+
 ```
@@
                                ziformula=~.,
-                              data = nelda2.sub %>% filter(inc.lose == "0"),
+                              data = nelda2.sub %>% filter(v2x_regime <= 2 & inc.lose == "0"),
                                 family=truncated_nbinom2)
@@
                                 ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0"),
+                                data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2),
                                 family=truncated_nbinom2)
@@
 t7 <-  huxreg(models, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,
- statistics = character(0), coefs = c("Intercept" = "(Intercept)", "Electoral integrity" = "v2xel_frefair", "Incumbent vote-share" = "winner.share.upd", "GDP growth (lag)" = "gdpgro.1lag", "Electoral integrity:Incumbent vote-share" = "v2xel_frefair:winner.share.upd",  "Election integrity : GDP growth (lag)" = "gdpgro.1lag:v2xel_frefair")) %>% set_caption("Conditional hurlde models of protest number and size") %>% insert_row("", "Number", "", "Size", "", after = 0)
-theme_compact(t7, header_rows = TRUE, header_cols = FALSE) %>% insert_row("Controls", "Yes", "Yes", "Yes", "Yes", after = 14) %>% set_bottom_border(14, everywhere)
+ statistics = character(0), coefs = c("Intercept" = "(Intercept)", "Incumbent term-limted" = "inc.termlimit.num", "Electoral integrity" = "v2xel_frefair", "Incumbent vote-share" = "winner.share.upd", "GDP growth (lag)" = "gdpgro.1lag", "Electoral integrity:Incumbent vote-share" = "v2xel_frefair:winner.share.upd",  "Election integrity : GDP growth (lag)" = "gdpgro.1lag:v2xel_frefair")) %>% set_caption("Conditional hurdle models of protest number and size") %>% insert_row("", "Number", "", "Size", "", after = 0)
+theme_compact(t7, header_rows = TRUE, header_cols = FALSE) %>% insert_row("Controls", "Yes", "Yes", "Yes", "Yes", after = 16) %>% set_bottom_border(16, everywhere)
 
@@
 p.n.margin <- plot(dat)
-p.n.margin <- p.n.margin + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events")  + theme_bw() + coord_cartesian(ylim=c(0, 75))
+p.n.margin <- p.n.margin + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events",  color = "Election integrity")  + theme_bw() + coord_cartesian(ylim=c(0, 75))
 
@@
 p.size.margin <- plot(dat)
-p.size.margin <- p.size.margin + labs(x = "Incumbent vote-share", y = "Predicted size of protests" , title = "Predicted maximum of protest events")  + theme_bw() + coord_cartesian(ylim = c(0, 50000))
+p.size.margin <- p.size.margin + labs(x = "Incumbent vote-share", y = "Predicted size of protests" , title = "Predicted maximum size of protest events", color = "Election integrity")  + theme_bw() + coord_cartesian(ylim = c(0, 50000))
 
@@
 
-```{r n-protests-margin-fulldv, fig.cap="Election integrity, incumbent vote-share, and number of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
+```{r n-protests-margin-fulldv, fig.cap="Election integrity, incumbent vote-share, and number of protest events", echo=FALSE, out.width = '75%', fig.align='center', fig.pos="h!"}
 p.n.margin
@@
 
-```{r protests-size-fulldv, fig.cap="Election integrity, incumbent vote-share, and size of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
+```{r protests-size-fulldv, fig.cap="Election integrity, incumbent vote-share, and size of protest events", echo=FALSE, out.width = '75%', fig.align='center', fig.pos="h!"}
 p.size.margin
+
+```
+
+# Proportional hazards diagnostics
+
+Cox proportional hazard models assume that the hazard ratio for each predictor is constant over time.
+A test of this assumption was developed by Schoenfeld [-@schoenfeld1982partial], in which risk-weighted covariate means are estimated for each point in time.
+Individual units' observed values for these covariates are then subtracted from the risk-weighted mean, giving the Schoenfeld residuals.
+As described by Grambsch and Therneau [-@grambsch1994proportional], these residuals can be plotted against event time with a smoothing spline to illustrate the general trend.
+The expected value for this relationship over time should be a horizontal line around zero, if the assumption of constant hazard ratio holds.
+Figure \@ref(fig:cox-diag) shows this relationship for *intentional voting irregularities*, with confidence bands of +/- two standard errors of the residuals at each time point.
+A horizontal line marks the zero line, while vertical lines indicate rough points where the curve of the smoothed line shifts, indicating cutpoints in the function.
+These cutpoints are used to estimate changes in the hazard function in Model 13 in the main text.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+library(survival)
+library(ggsurvfit)
+nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
+
+cox1 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + v2elirreg.inv , data = nelda2.sub %>% filter(v2x_regime <= 2 & inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+#summary(cox1)
+cz1 <- cox.zph(cox1, transform = "identity")
+#print(cz1)
+```
+
+```{r cox-diag, fig.cap="Schoenfeld residuals for election fraud by time (Model 12 in main text)", echo=FALSE, out.width = '85%', fig.align='center', warning=FALSE, message=FALSE, fig.pos="h!"}
+plot(cz1[18])
+abline(h=0)
+abline(v = 11)
+abline(v = 25)
+```
+
+# Removing controls to address multicollinearity
+
+The models in the main text include a suite of control variables aimed at controlling for common causes of election manipulation and protest, in order to minimize the risks of confounding and omitted variable bias.
+However, this approach runs the risk that individual coefficients may be highly sensitive to model specification due to multicollinearity.
+In particular, many of the control variables\textemdash such as *legislative constraints*, *judicial independence*, and *physical integrity* are likely to be correlated with pre-election protest.
+To address this risk, the models in Tables \@ref(tab:zero-model-table-small) and \@ref(tab:zero-model-table-count-small) exclude non-electoral control variables.
+That is, they include variables related to the election itself, the measure of *pre-election protest*, and the explanatory variables from the main models.
+
+The results echo those from the main text.
+Fraudulent elections are associated with a higher risk of protest initiation at low levels of ruling-party vote-share (Figure \@ref(fig:b-protests-margin)), though this does not translate into a higher risk of experiencing large protests (Figure \@ref(fig:n-protests-margin-small)).
+
+```{r models-hurdlesmall-mm, warning = FALSE, message = FALSE, include = F}
+model.small.novoteshare.hurd <- hurdle(n.events.post ~ v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election |  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election,
+                              ziformula=~.,
+                                data = nelda2.sub %>% filter(inc.lose == "0"),
+                                dist = "negbin")
+#summary(model.small.novoteshare.hurd)
+
+model.small.base.hurd <- hurdle(n.events.post ~  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election + winner.share.upd |  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election + winner.share.upd , data = nelda2.sub %>% filter(inc.lose == "0"), dist = "negbin")
+#summary(model.small.base.hurd)
+
+model.small.base2.hurd <- hurdle(n.events.post ~  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election + winner.share.upd + winner.share.upd*v2elirreg.inv | v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election + winner.share.upd  + winner.share.upd*v2elirreg.inv, data = nelda2.sub %>% filter(inc.lose == "0"), dist = "negbin")
+#summary(model.small.base2.hurd)
+
+model.small.griev.hurd <- hurdle(n.events.post ~  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election + gdpgro.1lag + gdpgro.1lag*v2elirreg.inv | v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election + gdpgro.1lag  + gdpgro.1lag*v2elirreg.inv, data = nelda2.sub %>% filter(inc.lose == "0"), dist = "negbin")
+#summary(model.small.griev.hurd)
+
+```
+\newpage
+
+```{r zero-model-table-small, echo=FALSE}
+
+models <- list()
+models[['Model 11']] <- model.small.base2.hurd
+models[['Model 12']] <- model.small.griev.hurd
+#models[['Model ']] <- model.n.griev.hurd.mm
+
+modelsummary::modelsummary(models, coef_omit = "count", gof_omit = "R2", stars = T, longtable=T,  title = 'Binary portion of hurdle models of post-election protest', coef_rename = 
+               c("zero_(Intercept)" = "Intercept", "zero_presidential1" = "Presidential election",
+                 "zero_inc.termlimit.num" = "Incumbent term-limited",  
+                                  "zero_n.events.pre" = "Number of pre-election protests", "zero_pubseccorrup.lag" = "Public sector corruption (lag)", "zero_v2elirreg.inv" = "Election fraud", "zero_unscheduled_election" = "Off-schedule election", "zero_winner.share.upd" = "Incumbent vote-share", "zero_gdpgro.1lag" = "GDP growth (lag)", "zero_presidential" = "Presidential",
+                 "zero_v2elirreg.inv:gdpgro.1lag" = "Election fraud : GDP growth (lag)", 
+                 "zero_v2elirreg.inv:winner.share.upd" = "Election fraud : Winner vote-share"))
 
@@
 
+```{r zero-model-table-count-small, echo=FALSE}
+modelsummary::modelsummary(models, coef_omit = "zero", gof_omit = "R2", longtable=T, stars = T, title = 'Hurdle models of post-election protest (counts)',
+                           coef_rename = 
+              c("count_(Intercept)" = "Intercept", "count_presidential1" = "Presidential election",
+                 "count_inc.termlimit.num" = "Incumbent term-limited",  
+                                  "count_n.events.pre" = "Number of pre-election protests", "count_pubseccorrup.lag" = "Public sector corruption (lag)", "count_v2elirreg.inv" = "Election fraud", "count_unscheduled_election" = "Off-schedule election", "count_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "winner.share.upd" = "Incumbent vote-share",
+                "count_gdpgro.1lag" = "GDP growth (lag)", "count_presidential" = "Presidential",
+                "count_v2elirreg.inv:gdpgro.1lag" = "Election fraud : GDP growth (lag)",
+                "count_v2elirreg.inv:winner.share.upd" = "Election fraud : Winner vote-share"))
+
+```
+
+```{r b-protests-margin, fig.cap="Fraud, incumbent vote-share, and protest initiation", out.width = '75%', echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}
+dat <- ggpredict(model.small.base2.hurd, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "zi_prob") 
+p.zi.margin <- plot(dat)
+p.zi.margin <- p.zi.margin +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest initiation", color = "Election fraud") + theme_bw() + coord_cartesian(y = c(0, 1))
+p.zi.margin
+```
+
+```{r n-protests-margin-small, fig.cap="Fraud, incumbent vote-share, and number of protest events", out.width = '75%', echo=FALSE, message=FALSE, fig.align='center'}
+dat <- ggpredict(model.small.base2.hurd, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "count") 
+p.n.margin <- plot(dat)
+p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Number of events", title = "Predicted number of protests", color = "Election fraud") + theme_bw() + coord_cartesian(y = c(0, 25))
+p.n.margin
+```
+
+\newpage
+
 # References

### 2024-02-07 Scripts/authoritarian_only.Rmd
Updates to appendix--removing MM

@@
 
-nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2024-01-30.csv"))
+nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2024-02-06.csv"))
 
@@
 Models where the dependent variable is the level of election fraud include data from all cases.
-For the models that focus on the initiation of post-election protest, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the theoretical expectations of protest risk following an election in which the incumbent won, possibly through reliance on election fraud.
-For models of protest-wave duration, only cases with a positive number of post-election protests are included.
+For the models that focus on the initiation of post-election protest, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the theoretical expectations of protest risk following an election in which the incumbent won, possibly through reliance on election fraud.^[Excluding cases where incumbents stepped down after losing an election is necessary to understand the initiation of protest against the alternative\textemdash where incumbents claimed victory.
+Howevever, this design does influence the interpretation of the results from these models.
+Threats to inference from this section are evaluated in more detail in the discussion.] For models of protest-wave duration, only cases with a positive number of post-election protests are included.
 The date range is limited by the availability of data on protest from ECAV, as discussed below, which ranges from 1989 to 2012.
@@
 Beginning with the zero-inflation portion of the hurdle models in Table 3, which model the likelihood of *no* protest occurring after the election, the standalone coefficient for election fraud is not significant in Model 5.
-Figure \@ref(fig:margin-plot-gdp) shows that there are no significant effect across values of *lagged GDP growth*.
 
@@
 
-model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary)
+
+model.margin.binary.int <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+summary(model.margin.binary.int)
 
@@
 models2 <- list()
-models2[['Model 5']] <- model.margin.binary_noc
-models2[['Model 6']] <- model.margin.binary
+models2[['Model 5']] <- model.margin.binary
+models2[['Model 6']] <- model.margin.binary.int
+models2[['Model 7']] <- model.margin.binary_noc
 
@@
 
-nelda2.sub.km <- nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0)
-
-km <-  with(nelda2.sub.km, Surv(d.events.num, surv_status))
-
-survfit2(Surv(d.events.num, surv_status) ~ 1, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
-  ggsurvfit() +
-#   add_confidence_interval() +
-  labs(
-    x = "Days",
-    y = "Overall survival probability"
-  )
-
 cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
@@
 The relationship between the level of fraud and the survival of protest movements can also be demonstrated non-parametrically using Kaplan-Meier curves for varying levels of fraud.
-In Figure \@ref(km-fig), observations are divided into three groups.
+In Figure \@ref(fig:km-fig), observations are divided into three groups.
 High fraud observations are those where *election fraud* is scored at one standard deviation above the mean or more, low fraud observations are those scored at one standard deviation below the mean or less, with moderate levels in between.
@@
 
-```{r km-fig, fig.cap = "Kaplan-Meier curves for protest duration and high, low, and medium levels of fraud", fig.height= = 7, echo=F, message=F, fig.pos='H'}
+```{r km-fig, fig.cap="Kaplan-Meier curves for protest duration across levels of fraud", fig.height= 7, echo=F, message=F, fig.pos='H'}
 nelda2.sub <- nelda2.sub %>% mutate(fraud_level = ifelse(v2elirreg.inv >= (.98 + .97), "High",
                                                             ifelse(v2elirreg.inv <= (.98 - .97), "Low", "Moderate")))
-
-survfit2(Surv(d.events.num, surv_status) ~ fraud_level, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
+nelda2.sub.km <- nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0)
+
+p.km <- survfit2(Surv(d.events.num, surv_status) ~ fraud_level, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
   ggsurvfit() +
@@
   )
+
+p.km
 ```
@@
 models3 <- list()
-models3[['Model 7']] <- cox1
-models3[['Model 8']] <- cox2
-models3[['Model 9']] <- cox1.cen
-models3[['Model 10']] <- cox2.cen
+models3[['Model 8']] <- cox1
+models3[['Model 9']] <- cox2
+models3[['Model 10']] <- cox1.cen
+models3[['Model 11']] <- cox2.cen
 
@@
 
-Altogether, the findings are a challenge to several prior assumptions and findings in the literature on election fraud, non-democratic elections, and authoritarian stability.
-These results do not suggest that election fraud is used as a tool of weak incumbents, in fact just the opposite.
-When encountering a more precarious pre-election environment\textemdash due to higher levels of pre-election protest or a successful post-election protest in the region, for instance\textemdash governments appear to respond by producing higher levels of election fraud.
-On its own, this is a potentially significant result which suggests that governments are not deterred from engaging in election fraud by the risk of post-election protest, as numerous models predict.
-
-More compellingly, this project finds that when post-election protests *do* occur, they are shorter-lived in for cases where fraud was higher.
-This indicates that governments which produced high levels of fraud were more able to quash those protest movements which did emerge more quickly than governments that produced low levels of fraud.
-This is contrary to three prevailing wisdoms.
-The first is that weak governments manipulate more than strong, the second is that more election fraud is a grievance that mobilizes regime antagonists, and the third is that fraud can only deter protest if it is excessive or blatant.
-
-Indeed, while there are 194 cases of protest out of the 661 elections in this dataset where incumbents claimed victory, only 8 resulted in either the incumbent being replaced or the election being re-run due to mass protests.\footnote{This information is determined using the NELDA variables NELDA37 and NELDA41.} This suggests a `success rate' for post-election protest of about four percent *conditional on protest initiation*, and about one percent for all elections in electoral authoritarian and electoral democratic regimes.
+The results can be summarized as follows.
+The severity of election fraud is not associated with protest initiation, except when combined with very low vote-shares for the incumbent.
+Indicators of pre-election instability are not associated with reduced fraud, but with higher levels.
+Finally, incumbents who have employed large-scale fraud appear better able to put an end to what protest movements do emerge than those who have presided over cleaner elections.
+This is true even when incumbent vote-shares are low.
+Altogether, the findings are a substantial challenge to several prior assumptions and findings in the literature on election fraud, non-democratic elections, and authoritarian stability.
+
+The results strongly contradict the assumptions of the classic school of thought, which holds that election fraud indicates regime weakness.
+Instead, across the three tests here, fraud appears to be an indicator of regime resilience.
+This is a substantively important distinction.
+Classical models hold that, since fraud indicates weakness, incumbents that would lose fair elections can be deterred from fraud by the risk of post-election protest.
+Such models predict that a democratic exchange of power can occur in such scenarios, backed up by the risk of protest.
+The results here would appear to overturn that model.
+
+The results are also challenging for the second school of thought, which proposes that weak and strong incumbents both engage in electoral manipulation, and protest occurs when ruling parties under- *or* over-shoot the opposition's expected level of fraud.
+Here we would not expect to see consistent patterns with regard either to fraud deterrence or to fraud and the survival of protest movements.
+Instead we see a consistent pattern that associates higher levels of fraud with more resilient incumbents.
+
+Third, these results are most consistent with the third view, which holds that even election fraud can be an indicator of regime strength.
+Where this paper goes beyond the traditional formulation of this view\textemdash apart from testing the underlying assumptions of the three models\textemdash is the observation that large-scale fraud can be indicative of incumbent strength even when it is not excessive.
+That is, even when ruling party vote-shares are low (perhaps especially when this is so), incumbents which showed themselves capable of large-scale fraud are more capable of ending protest movements than low-fraud incumbents.
+
+What does all this tell us?
+[FIX] First, these results may inform how researchers think about the causes of election manipulation.
+They suggest a that the importance of factors like the opposition's capacity to organize street protest in limiting election fraud has been overestimated.
+In fact, these results suggest that incumbents who have the resources will respond to such a scenario by *increasing* their production of fraud.
+Instead, the results point toward those factors that enable incumbents to both drive manipulation efforts and to quash protest movements: resource consolidation and protection for regime agents who act against the law in incumbents' interest.
+Further research on incumbents efforts to develop these capacities over time\textemdash and obstacles they face\textemdash may yield insights into election integrity.
+
+In turn, this revision to our understanding of election manipulation has implications for broader topics like authoritarian persistence and democratization.
+We observe that protest initiation is most likely when fraud is high and incumbent vote-share is low.
+This could be indicative of strategic behavior by the rational opposition actors found in most formal models (who take this outcome as best evidence of a weak incumbent), or by less strategic behavior by citizens en masse.
+For such citizens, election manipulation may act as a shared grievance, especially when it appears that fraud may have been decisive [@kuntz_more_2009;@tucker2007].
+A close election marked by fraud intensifies the connection between emotional and strategic protest incentives.
+As Schedler writes, "Grievances provide moral justifications for collective protest: we need to do something.
+Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304].
+Yet this highlights a dilemma for opposition supporters, since it is exactly these scenarios wherein post-election protest is most likely to fail quickly.
+
+There is a long-standing debate over the role of elections in democratization\textemdash whether they contribute to it or regime stability\textemdash of which the manipulation-protest nexus is a component [@Schedler2013;@Morgenbesser2019;@edgellWhenWhereElections2018].
+At the core, much of this debate revolves around the extent to which non-democratic elections affect the ruling party's cost of suppressing the opposition versus tolerating it [@dahlPolyarchyParticipationOpposition1971].
+Election manipulation, including fraud, directly reduces the cost of toleration, by making it less likely that opposition groups win elections.
+To the extent that it makes electoral protest more likely and/or more costly, however, it increases the cost of suppression.
+In this version, election fraud has the potential to lay the groundwork for democratization by incentivizing the ruling party to tolerate organized opposition which, one day, it may be unable to repress after an election.
+This logic can be found in studies of the color revolutions [@bunce2010], and in the observation that competitive authoritarian regimes are more likely to democratize after an election than hegemonic ones [@donno2013elections]
+
+By showing that election fraud does not make protest more likely (most of the time) and that the infrastructure of fraud can actually reduce the cost of post-election protest to the regime, this result flips that logic on its head.
+Instead, election fraud can both reduce the cost of toleration *and* the cost of suppression.
+This implies that electoral authoritarianism is something of a cul-de-sac: the capacity to generate severe election fraud makes it appealing for closed authoritarian regimes to introduce elections, while the manageable risk of protest makes it unappealing to democratize.
+This logic also creates incentives for ruling parties in weakly consolidated democracies to take measures that will enable them to engage in electoral manipulation; namely, the consolidation of patronage resources and erosion of rule of law.
+This provides additional theoretical support for understanding what some have called a 'wave of autocratization' or a 'democratic recession' in recent years [@luhrmannThirdWaveAutocratization2019;@diamondDemocraticRegressionComparative2021].
+If the same tools that produce broad election fraud also reduce the cost of protest, then electoral authoritarianism has a greater gravitational pull as a regime type.
+
+Indeed, while there are 164 cases of protest in elections where the incumbent claimed victory out of the 661 total non-democratic elections in this dataset, only 8 resulted in either the incumbent being replaced or the election being re-run due to mass protests.\footnote{This information is determined using the NELDA variables NELDA37 and NELDA41.} This suggests a `success rate' for post-election protest of about five percent *conditional on protest initiation*, and about one percent for all elections in electoral authoritarian and electoral democratic regimes.
+Moreover, the modal protest 'wave' in the dataset lasted only one day, with the median wave lasting three days.
+This is
+
+One challenge to the above interpretation considers the sample selection for the models which take protest initiation as the dependent variable.
+This paper is primarily concerned with whether or not election manipulation is indicative of regime weakness or resilience; its primary evidence is finding that fraud appears to increase with pre-election uncertainty and that fraud is associated with more resilient incumbents in the face of protest.
+Yet one of the key predictions of the classic models is that weaker incumbents will generate little fraud and voluntarily step down after losing elections.
+The models of protest initiation cannot evaluate this possibility, since cases where incumbents lose are excluded by necessity.
+Is there evidence to support this interpretation, and help salvage the protest-deterrent model?
+
+At first glance, the pattern appears promising.
+The mean estimate of *election fraud* in the 129 observations where incumbents lost is 0.322, much lower than the 1.20 average in the group where incumbents claimed victory.
+However, this observation alone does not discriminate between the two possibilities considered here: where these incumbents deterred from committing more severe fraud (despite being able to), or were they unable to muster an effective enough manipulation effort to 'win' the election?
+The overall pattern in the data leans toward the latter interpretation.
+Table \@ref(tab:anova-table) shows how different the two groups are along several dimensions that affect the availability of fraud as a strategy.
+For instance, among the cases where the incumbent lost, judicial independence is markedly higher.
+Similarly, the losing group has a substantially more privatized economy, a more professional bureaucracy, and more autonomous election management bodies.
+Additionally, the losing group also shows lower average economic growth.
+All of these should tend to reduce the availability of fraud.
+Of the factors considered here, only *number of pre-election protests* clearly behaves as the deterrence logic would predict, with the losing group subject to one more protest, on average, than the winning group.
+However, unlike the factors which point toward the alternative hypothesis, this difference is not statistically significant.
+Thus, while future research might test this aspect of the deterrence model more rigorously, the general pattern is supportive of the view put forth in this paper.
+
+```{r, include=F}
+## Anova
+anova_dat <- nelda2.sub %>% dplyr::select(inc.lose, v2elirreg.inv,  jucon.1lag, legcon.1lag, gdpgro.1lag,  osorg.1lag, natlpartyorg.lag,  urban.pct, n.events.pre, state.owned.1lag, patronage.appts.1lag, v2elembaut)
+
+# boxplots and t-tests for the 4 variables at once
+for (i in 2:ncol(anova_dat)) { # variables to compare are variables 2 to end
+  boxplot(anova_dat[, i] ~ anova_dat$inc.lose, # draw boxplots by group
+    ylab = names(anova_dat[i]), # rename y-axis with variable's name
+    xlab = "Incumbent loss"
+  )
+  print(paste(c("ANOVA for variable ", (names(anova_dat[i])))))
+  print(t.test(anova_dat[, i] ~ anova_dat$inc.lose)) # print results of t-test
+}
+
+aov_output <- aov(state.owned.1lag ~ inc.lose, data = nelda2.sub)
+summary(aov_output)
+```
+
+```{r anova-table, echo=FALSE}
+p.values <- c("NA", 3.009e-15, 6.014e-11, 6.149e-10, 0.01146,  3.758e-07,  0.6281, 0.003519,
+              0.1852, 0.007405, 0.003086, 2.2e-16)
+win_lose_diff <- c("NA", .87, -0.15, -0.16, .06, -.47, .04, 5.3, -1.46, -.265, -.257, -.893)
+variables <- names(anova_dat)
+anova_dat <- data.frame(cbind(variables,win_lose_diff, p.values))
+anova_dat <- anova_dat %>% mutate(win_lose_diff = as.numeric(win_lose_diff))
+anova_dat <- anova_dat %>% mutate(p.values = as.numeric(p.values))
+
+kable(anova_dat, digits = 3)  %>%
+    kable_styling(font_size = 9, full_width = FALSE)
+
+```
 
@@
 
+Ruling party turnover
+
+```{r, include=FALSE}
+model.margin.turnover <- glm(inc.lose ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +   + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + state.owned.1lag + patronage.appts.1lag, family = binomial(link = "logit"), data = nelda2.sub)
+summary(model.margin.turnover)
+
+m.inclose <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag*patronage.appts.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub %>% filter(inc.lose == 1))
+summary(m.inclose)
+sjPlot::plot_model(m.inclose, type = "int")
+
+m.incwin <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag*patronage.appts.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub %>% filter(inc.lose == 0))
+summary(m.incwin)
+sjPlot::plot_model(m.incwin, type = "int")
+
+### Z score
+
+z <- (-1.647988 - (-1.144192)) / sqrt((0.691932)^2 + (0.329801)^2) #B1 - B2 / sqrt((seB1)^2 + (seB2)^2)
+
+```
+
 \newpage

### 2024-02-07 Scripts/data_cleaning.Rmd
Updates to appendix--removing MM

@@
 nelda <- nelda %>% dplyr::rename(first.mp.election = nelda2)
-nelda <- nelda %>% rename(protests = nelda9)
+nelda <- nelda %>% rename(protests = nelda29)
+nelda <- nelda %>% rename(protest.fraud.allege = nelda30)
 nelda <- nelda %>% rename(inc.fav.cancel = nelda34)
@@
 nelda <- nelda %>% rename(econ.crisis.nelda = nelda18)
-nelda <- nelda %>% rename(opp.votegain = nelda7)
-nelda <- nelda %>% rename(nelda.protests.fraud = nelda30)
+nelda <- nelda %>% rename(inc.confident = nelda12)
+nelda <- nelda %>% rename(opp.votegain = nelda27)
 nelda <- nelda %>% rename(new.elections.held = nelda37)
@@
 nelda <- nelda %>% mutate(opp.votegain = ifelse(opp.votegain == "yes", 1, ifelse(opp.votegain == "no", 0, NA)))
-nelda <- nelda %>% mutate(nelda.protests.fraud = ifelse(protests == 0, 0, ifelse(nelda.protests.fraud == "yes", 1, ifelse(nelda.protests.fraud == "no", 0, NA))))
+nelda <- nelda %>% mutate(protest.fraud.allege = ifelse(protests == 0, 0, ifelse(protest.fraud.allege == "yes", 1, ifelse(protest.fraud.allege == "no", 0, NA))))
 nelda <- nelda %>% mutate(inc.replaced.prot = ifelse(inc.replaced.prot == "yes", 1, 0))
@@
 vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(legcon.1lag = lag(v2xlg_legcon))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(antisys.1lag = lag(v2csantimv))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(csenv.1lag = lag(v2csprtcpt))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(patronage.appts.1lag = lag(v2stcritrecadm))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(state.owned.1lag = lag(v2clstown))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(polariz.1lag = lag(v2cacamps))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(regime.1lag = lag(v2x_regime))
+vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(wb_inst.1lag = lag(e_wbgi_pve))
 
@@
                                                   v2ex_elechog,
+                                                  antisys.1lag,
+                                                  v2csantimv,
+                                                  csenv.1lag,
+                                                  v2csprtcpt,
+                                                  patronage.appts.1lag,
+                                                  v2stcritrecadm,
+                                                  state.owned.1lag,
+                                                  v2clstown,
+                                                  v2elembaut,
+                                                  v2cacamps,
+                                                  polariz.1lag,
+                                                  regime.1lag,
+                                                  wb_inst.1lag,
+
 egaldem.1lag ,
@@
 
-```
+##Amended vote-shares 2 followed the following coding rules:
+#- Seat shares instead of vote-shares where the latter are not provided
+#   -This is true of many post-communist and highly authoritarian countries (i.e. not missing at random)
+#- Independents used as largest party when there is no ruling party but parties are allowed
+#- Lebanon is excluded due to consociational structure w/ independents mixing with parties
+#- Iran is also excluded due to the absence of real parties (though there are factions)
 
-## Ideas for adding rate of growth and rate of decay to be used as DVs
+winner.shares2 <- read.csv(here::here("Data", "amended_vote_shares2.csv"))
+nelda2.sub <- nelda2.sub %>% left_join(winner.shares2, by = "NeldaID")
 
- - Quick growth after fraud indicates opp believes state is weak
- - Slow or no growth after fraud indicates opp believes state is strong
- - Quick decay after peak indicates state is strong / believed strong (or not--regime is weak and so quickly collapses)
- - Slow decay after peak indicates state is weak / believed weak
+nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))
 
-The idea here would be:
-  - Replace categorical measures with minimum number (average won't work because top category is top-less)
-  - Replace -99 with...
-  - Filter by election
-  - Group by date and sum
-  - Fit exponential function regression to find coefficient for 'r' (rate of growth)
-  - This becomes the DV for a regression on fraud
+nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(is.na(winner.share.upd)==F, winner.share.upd,
+                                                ifelse(types == "Executive", v2elvotlrg_amend_2/100, v2ellovtlg_amend_2/100)))
 
-Instead, because protests don't always follow the exponential distribution, going to create a simple measure of protest growth / decay intensity.
-
-The idea here would be:
-  - Replace categorical measures with minimum number (average won't work because top category is top-less)
-  - Replace -99 with...
-  - Filter by election
-  - Group by date and sum
-  - Find date of largest protest
-  - Calculate max_size / max_date - 0 and max_size / end_date - max_date
-  - Merge these values into main dataset by election (NeldaID for example)
-  - This becomes the DV for a regression on fraud
-
-```{r}
-# Example
-
-## Remove things other than protests
-ecav_prot <- ecav %>% filter(str_detect(EventName, "rotest") == T | str_detect(EventName, "ccupation") == T |
-                                  str_detect(EventName, "lockade") == T) #Only captures protests
-```
-
-Replacing NA protest size with average for that wave
-```{r}
-## Set up holding df
-
-dv_df <- data.frame(NeldaID = NA, avg_size = NA)
-
-## Filter by election and replace -99 with NA for protest size
-
- election_ids <- unique(ecav_prot$NeldaID)
-ecav_prot <- ecav_prot %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
-
-for(i in 1:length(election_ids)){ 
-  ecav_temp <- ecav_prot %>% filter(NeldaID == election_ids[i])  #There is a loop here
-
-    ## Remove protests before election
-  ecav_temp <- ecav_temp %>% filter(Date >= Electiondate)
-  if(nrow(ecav_temp) == 0){
-     new_row <- data.frame(NeldaID = election_ids[i], 
-                           avg_size = 0) #No post-election protest
-
-  dv_df[i,] <- new_row
-  print(paste0("No post-election protest for observation ", i, "."))
-  next
-  } else{
-
-  new_row <- data.frame(NeldaID = election_ids[i], avg_size = floor(mean(ecav_temp$ParticipantNumber, na.rm = T)))
-
-  dv_df[i,] <- new_row
-  print(paste0("Average size for observation ", i, " completed."))
-  }
-}
-
-## Merge dv_df back into main dataset
-
-ecav_prot <- ecav_prot %>% left_join(dv_df, by = "NeldaID")
-
-```
-
-Current to-do: Double-check the output of the size code
-
-```{r}
-## Set up holding df
-
-dv_df <- data.frame(NeldaID = NA, prot_intensity = NA, decay_intensity = NA, max_size = NA, day_max = NA)
-
-ecav_prot <- ecav_prot %>% mutate(ParticipantNumber_adj = ifelse(is.na(ParticipantNumber)==T, avg_size, ParticipantNumber))
-ecav_prot <- ecav_prot %>% mutate(ParticipantNumber_adj = ifelse(ParticipantNumber_adj == "NaN", 2, ParticipantNumber_adj)) #Assume NA size defaults to category 2
-
-#ecav_rus <- ecav %>% filter(country == "Russia")   # Filter by country
-ecav_prot <- ecav_prot %>% mutate(ParticipantNumber_num_min = ifelse(ParticipantNumber_adj == 1, 5, 
-                                ifelse(ParticipantNumber_adj == 2, 55, 
-                                       ifelse(ParticipantNumber_adj == 3, 550, 
-                                              ifelse(ParticipantNumber_adj == 4, 5500,
-                                                     ifelse(ParticipantNumber_adj == 5, 10000, NA))))))
-
-## Filter by election
-
- election_ids <- unique(ecav_prot$NeldaID)
-
-for(i in 1:length(election_ids)){ 
-  ecav_temp <- ecav_prot %>% filter(NeldaID == election_ids[i])  #There is a loop here
-
-## Group by date and sum number of participants by date:
-
- ecav_temp_comb <- ecav_temp %>% mutate(date_col = date(Date)) %>%
-    group_by(date_col) %>%
-     summarize(total = sum(ParticipantNumber_num_min))
-
- ## Get eleection date
-
- election_date_temp <- unique(ecav_temp$Electiondate)
-
- ## Remove protests before election
-
- ecav_temp_comb <- ecav_temp_comb %>% filter(date_col >= election_date_temp)
-
-  if (nrow(ecav_temp_comb) == 0) {
-    new_row <- data.frame(election_ids[i], NA, NA, NA, NA)
-    dv_df[i,] <- new_row
-    print(paste0("No post-election protests for obseveration ", i, ", election ID", election_ids[i], ","))
-    next}  #End loop if zero post-election protests
-
- ## Get date of largest protest
-
- max_data <-  ecav_temp_comb %>% filter(total == max(total, na.rm = T))
-
-  # Assign NA to NA size
- if (nrow(max_data)==0){
-    new_row <- data.frame(election_ids[i], NA, NA, NA, NA)
-    dv_df[i,] <- new_row
-    print(paste0("Election observation ", i, " completed. (Note: no size estimate."))
-    next
- }
-
-  if (nrow(max_data) >= 1) {
-    prot_intensity <- max_data$total[1] / (1 + as.numeric(ymd(max_data$date_col[1]) - ymd(election_date_temp))) #[1] gets the first peak
-    last_peak <- tail(max_data, n = 1)
-    decay_intensity <-  last_peak$total / (1 + as.numeric(max(ecav_temp_comb$date_col) - (ymd(last_peak$date_col))))
-    day_max <- as.numeric(ymd(max_data$date_col[1]) - ymd(election_date_temp))
-    new_row <- data.frame(election_ids[i], prot_intensity, decay_intensity, max_data$total[1], day_max)
-  new_row <- new_row %>% rename(max_size = max_data.total.1.)
-    dv_df[i,] <- new_row
-    print(paste0("Election observation ", i, " completed. (Note: multiple peaks."))
-    next
-  }
-
- ## Calculate DVs
-
- prot_intensity <- max_data$total / (1 + as.numeric(ymd(max_data$date_col) - ymd(election_date_temp)))
- decay_intensity <-  max_data$total / (1 + as.numeric(max(ecav_temp_comb$date_col) - (ymd(max_data$date_col)))) #The +1 is a kludge to prevent Inf values
-
- ##Add to holding df
-
-  day_max <- as.numeric(ymd(max_data$date_col[1])) - as.numeric(ymd(election_date_temp))
-  new_row <- data.frame(election_ids[i], prot_intensity, decay_intensity, max_data$total, day_max)
-  new_row <- new_row %>% rename(max_size = max_data.total)
-
-  dv_df[i,] <- new_row
-  print(paste0("Election observation ", i, " completed."))
-
- }
-
-  ## Merge dv_df back into main dataset
-
-nelda2.sub.b <- nelda2.sub %>% left_join( dv_df, by = "NeldaID")
-
-```  
-Pick up here with fixing this code:
-
-```{r, eval=FALSE}
-dv_df <- data.frame(NeldaID = NA, growth_coef = NA)
-for(i in 1:length(election_ids)){
- ecav_temp <- ecav_prot %>% filter(NeldaID == election_ids[i])  #There is a loop here
-
-###Getting a regression slope for protest size and time
-  ecav_temp_comb <- ecav_temp %>% mutate(date_col = date(Date)) %>%
-    group_by(date_col) %>%
-     summarize(total = sum(ParticipantNumber_num_min))
-  max_date_reg <- ecav_temp_comb %>% filter(total == max(total)) %>%  
-   dplyr::select(date_col)
-
- ecav_temp_reg <- ecav_temp %>% filter(as_date(Date) >= as_date(Electiondate))
- ecav_temp_reg <- ecav_temp_reg %>% mutate(date_col = date(Date))
- ecav_temp_reg <- ecav_temp_reg %>% left_join(ecav_temp_comb, by = "date_col")  #Gets totals by date
-
- ecav_temp_reg <- ecav_temp_reg %>% filter(as_date(Date) <= max_date_reg[1,1])
- ecav_temp_reg <- ecav_temp_reg %>% mutate(date_index = as.numeric(ymd(ecav_temp_reg$Date) - ymd(Electiondate)))
- if(nrow(ecav_temp_reg) == 0) {
-   new_row <- data.frame(election_ids[i], NA)
-   dv_df[i,] <- new_row
-   print(paste0("No observations for election number ", i, "."))
-   next
- }
- if(length(unique(ecav_temp_reg$Date))==1){
-   growth_coef <-  ((ecav_temp_reg$total - 0 )/((ecav_temp_reg$date_index+1)-0))[1] #m = (y2 - y1)/(x2 - x1) #+1 in denominator is a kludge to avoid Inf values for day 0 protests
-   new_row <- data.frame(election_ids[i], growth_coef)
-   dv_df[i,] <- new_row
-   print(paste0("One date observed for election ", i, "."))
-   next
- } else {
-
- fit <- lm(total ~ 0 + date_index, ecav_temp_reg)
- sum_table <- summary(fit)
- growth_coef <- sum_table$coefficients[1]
-  new_row <- data.frame(election_ids[i], growth_coef)
-   dv_df[i,] <- new_row
-   print(paste0("Coefficient estimated for election ", i, "."))
- }
- rm(growth_coef)
-}
-
-## Merge dv_df back into main dataset
-
-nelda2.sub.b <- nelda2.sub.b %>% left_join( dv_df, by = "NeldaID")
 ```
@@
 ```{r}
-success_cases <- nelda2.sub.b %>% filter(inc.replaced.prot == 1 | nelda35 == "yes") %>% dplyr::select(e_regionpol, year, election_mdy)
+success_cases <- nelda2.sub %>% filter(inc.replaced.prot == 1 | nelda35 == "yes") %>% dplyr::select(e_regionpol, year, election_mdy)
 
@@
 
-nelda2.sub.b <- nelda2.sub.b %>% mutate(regional_success_lag = ifelse(e_regionpol == 1 & (election_mdy %within% interval(start = "1990-06-17", end = "1991-06-17") | election_mdy %within% interval(start = "2000-09-24", end = "2001-09-24") | election_mdy %within% interval(start = "2003-11-02", end = "2006-11-06")), 1,
+nelda2.sub <- nelda2.sub %>% mutate(regional_success_lag = ifelse(e_regionpol == 1 & (election_mdy %within% interval(start = "1990-06-17", end = "1991-06-17") | election_mdy %within% interval(start = "2000-09-24", end = "2001-09-24") | election_mdy %within% interval(start = "2003-11-02", end = "2006-11-06")), 1,
                                       ifelse(e_regionpol == 4 & (election_mdy %within% interval(start = "1997-04-13", end = "1998-04-13") | election_mdy %within% interval(start = "1998-05-23", end = "1999-05-23") | election_mdy %within% interval(start = "2000-10-22", end = "2001-10-22") | election_mdy %within% interval(start = "2007-07-22", end = "2008-07-22") | election_mdy %within% interval(start = "2010-11-28", end = "2011-11-28")), 1,
@@
 
-## Writing
-
-```{r}
-write.csv(nelda2.sub.b, here::here("Data", "nelda2_sub_2023-05-30.csv"))
-```
-
 ## Adding ERT data for regime duration
@@
 ```{r, eval=FALSE}
-nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2023-05-10.csv"))
 ert <- read.csv(here::here("Data", "ert.csv"))
@@
 nelda2.sub  <- nelda2.sub %>% mutate(regime_age = year - reg_start_year)
+```
 
-write.csv(nelda2.sub, here::here("Data", "nelda2_sub_2023-05-30.csv"))
+## DIEM data for cross-checking export coding in VDEM
 
+```{r}
+load(here::here("Data", "diem_data.rda"))
+diem <- da31461.0001
+rm(da31461.0001)
+
+##Convert diem's COW identifiers to the vdem type used above, variable called country_id
+
+diem <- diem %>% mutate(country_id = countrycode(sourcevar = COW, origin = "cown", destination = "vdem")) #Needs fixes for a couple countries
+
+diem <- diem %>% rename(diem_fraud_max = MAXR211R)
+diem <- diem %>% rename(diem_explicheat_max = MAXR21CHEAT)
+diem <- diem %>% rename(presidential = EXELEC)
+diem <- diem %>% mutate(country_id_year = paste(country_id, YEAR, sep = "_"))
+
+#####Convert variables to numeric
+
+diem <- diem %>% mutate(diem_fraud_max = ifelse(is.na(diem_fraud_max)==T, NA, 
+                                                ifelse(diem_fraud_max == "(0) Good - no problems", 0,
+                                                       ifelse(diem_fraud_max == "(1) Low - minor problems only", 1, 
+                                                              ifelse(diem_fraud_max == "(2) Moderate - moderate problems", 2, 3)))))
+
+diem <- diem %>% mutate(diem_explicheat_max = ifelse(is.na(diem_explicheat_max)==T, NA, 
+                                                ifelse(diem_explicheat_max == "(0) Good - no problems", 0,
+                                                       ifelse(diem_explicheat_max == "(1) Low - minor problems only", 1, 
+                                                              ifelse(diem_explicheat_max == "(2) Moderate - moderate problems", 2, 3)))))
+
+diem <- diem %>% mutate(presidential = ifelse(presidential == "(1) Executive Election", 1, 0))
+
+#####
+
+## Find average of observer reports per election
+
+mean_max_fraud <- diem %>% group_by(NEWID) %>% summarise(mean(diem_fraud_max))
+mean_max_explicheat <- diem %>% group_by(NEWID) %>% summarise(mean(diem_explicheat_max))
+
+## Get month and year for each election
+
+diem <- diem %>% mutate(MONTH = ifelse(is.na(DATEEXEC)==T, DATELEG, DATEEXEC))
+
+diem <- diem %>% mutate(country_id_year_month = paste(country_id, YEAR, MONTH, sep = "_"))
+
+## For merging
+
+diem_sub <- diem %>% dplyr::select(country_id_year_month, diem_explicheat_max, diem_fraud_max)
+diem_sub <- diem_sub %>%
+  distinct(.keep_all = TRUE)
+
+##Get month separately for NELDA
+
+nelda2.sub <- nelda2.sub %>% mutate(MONTH = substr(mmdd, 1, 2))
+nelda2.sub <- nelda2.sub %>% mutate(country_id_year_month =  paste(country_id_year, MONTH, sep = "_"))
+
+nelda2.sub <- nelda2.sub %>% left_join(diem_sub, by = c("country_id_year_month")) ## Gotta figure out a way to merge by election--including month
+
+```
+
+## Adding in Mass Mob events
+
+```{r}
+massmob <- read.csv(here::here("Data", "mmALL_073119_csv.csv"))
+
+massmob <- massmob %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "p4n", 
+                                               destination = "vdem")) #Missing here tend to be communist
+massmob <- massmob %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
+
+#nelda <- nelda %>% mutate(mmdd = ifelse(mmdd < 1000, paste0(0, mmdd), mmdd))
+
+#nelda <- nelda %>% mutate(election_mdy = paste0(mmdd, year))
+#nelda <- nelda %>% mutate(election_mdy = mdy(election_mdy))
+
+massmob <- massmob %>% mutate(startday = ifelse(startday < 10, paste0(0, startday), startday))
+massmob <- massmob %>% mutate(startmonth = ifelse(startmonth < 10, paste0(0, startmonth), startmonth))
+
+massmob <- massmob %>% mutate(endday = ifelse(endday < 10, paste0(0, endday), endday))
+massmob <- massmob %>% mutate(endmonth = ifelse(endmonth < 10, paste0(0, endmonth), endmonth))
+
+massmob <- massmob %>% mutate(protest_mdy = ifelse(is.na(startday) == T | is.na(startmonth) == T, NA,  paste0(startmonth, startday, year)))
+massmob <- massmob %>% mutate(protest_mdy_end = ifelse(is.na(endday) == T | is.na(endmonth) == T, NA,  paste0(endmonth, endday, year)))
+
+massmob <- massmob %>% mutate(protest_mdy = mdy(protest_mdy))
+massmob <- massmob %>% mutate(protest_mdy_end = mdy(protest_mdy_end))
+
+massmob <- massmob %>% filter(is.na(protest_mdy) == F)
+
+massmob <- massmob %>% filter(is.na(country_id) == FALSE)
+
+## Getting the mass mob events
+
+nelda_selection <- nelda2.sub %>% dplyr::select(NeldaID, country_id.x, election_mdy)
+
+  election_id_temp <- nelda_selection$NeldaID[1]
+  country_temp <- nelda_selection$country_id[1]
+  election_date_temp <- nelda_selection$election_mdy[1] # Get the election date
+  election_row <- nelda_selection[1,]  # Get the election at hand (all Nelda data)
+  massmob_temp <- massmob %>% filter(country_id == country_temp) #Get all protests from the country at hand
+  massmob_temp <- massmob_temp %>% mutate(days_to_election = election_date_temp - protest_mdy) # Get the days between each protest and the election
+  massmob_temp <- massmob_temp %>% filter(days_to_election <= 180 & days_to_election >= -180) #Remove elections outside the window--upped to 180 days
+  massmob_temp <- massmob_temp %>% mutate(NeldaID = election_id_temp)
+  election_row <- election_row %>% left_join(massmob_temp, by = "NeldaID") #Merge Nelda data with remaining protests
+
+nelda_massmob <- election_row #Start the dataset
+
+for (i in 2:nrow(nelda2.sub)) {  #
+  election_id_temp <- nelda_selection$NeldaID[i]
+  country_temp <- nelda_selection$country_id[i]
+  election_date_temp <- nelda_selection$election_mdy[i]
+  election_row <- nelda_selection[i,]
+  massmob_temp <- massmob %>% filter(country_id == country_temp)
+  massmob_temp <- massmob_temp %>% mutate(days_to_election = election_date_temp - protest_mdy)
+  massmob_temp <- massmob_temp %>% filter(days_to_election <= 180 & days_to_election >= -180)
+
+  massmob_temp <- massmob_temp %>% mutate(NeldaID = election_id_temp)
+  election_row <- election_row %>% left_join(massmob_temp, by = "NeldaID")
+  nelda_massmob <- bind_rows(nelda_massmob, election_row)
+
+}
+
+## Remove non-election related protests
+nelda_massmob <- nelda_massmob %>% rename(notes_massmob = notes)
+
+nelda_massmob <- nelda_massmob %>% filter(is.na(protest_mdy) == T | protest_mdy < election_mdy | (protest_mdy >= election_mdy & str_detect(notes_massmob, regex(pattern = "elector*|election*|vot*|ballot*", ignore_case = T)))) #Turn this on to focus only on protests related to the election
+
+##Next get number and duration
+nelda_massmob <- nelda_massmob %>% mutate(protest_mdy_end = ifelse(is.na(protest_mdy_end) == T, protest_mdy, protest_mdy_end)) #This codes protests with no end date as lasting ending on same day they began
+
+nelda_massmob <- nelda_massmob %>% mutate(protest_days = as.Date(protest_mdy_end) - as.Date(protest_mdy))
+nelda_massmob <- nelda_massmob %>% mutate(protest_days = ifelse(is.na(protest_mdy)==F & is.na(protest_mdy_end)==T, 1, protest_days))
+
+nelda_massmob.pre <- nelda_massmob %>% filter(as.numeric(days_to_election) > 0) #protest_mdy < election_mdy
+nelda_massmob.post <- nelda_massmob %>% filter(as.numeric(days_to_election) <= 0)
+
+## Count
+count_data_mm_pre <- nelda_massmob.pre %>% group_by(NeldaID) %>% summarize(n.events.massmob.pre = n())
+count_data_mm_post <-nelda_massmob.post %>% group_by(NeldaID) %>% summarize(n.events.massmob.post = n())
+
+## Total duration
+
+dur_data_mm_pre <- nelda_massmob.pre %>% group_by(NeldaID) %>% summarize(mm_duration.pre = as.Date(max(protest_mdy_end, na.rm = T)) - as.Date(min(protest_mdy, na.rm=T)))
+dur_data_mm_post <-nelda_massmob.post %>% group_by(NeldaID) %>% summarize(mm_duration.post = as.Date(max(protest_mdy_end, na.rm = T)) - as.Date(min(protest_mdy, na.rm=T)))
+
+## Merge back into nelda
+
+##Combined post data together, then pre data, then both
+mm_data_post_comb <- left_join(count_data_mm_post, dur_data_mm_post, by = "NeldaID")
+mm_data_pre_comb <- left_join(count_data_mm_pre, dur_data_mm_pre, by = "NeldaID")
+
+mm_data_comb <- left_join(mm_data_pre_comb, mm_data_post_comb, by = "NeldaID")
+
+#mm_data_comb <- mm_data_comb %>% rename(electionid = NeldaID)
+
+## Merge back to Nelda classic
+nelda2.sub <- nelda2.sub %>% left_join(mm_data_comb, by = "NeldaID")
+
+#nelda_massmob <- nelda_massmob %>% left_join(mm_data_comb, by = "electionid")  ## Double check to make sure this joining works
+
+## Code NAs for number of protests as 0s (these reflect election years recorded in NELDA where there were no protests recorded by MM)
+
+nelda2.sub <- nelda2.sub %>% mutate(n.events.massmob.pre = ifelse(is.na(n.events.massmob.pre)==T, 0, n.events.massmob.pre))
+nelda2.sub <- nelda2.sub %>% mutate(n.events.massmob.post = ifelse(is.na(n.events.massmob.post)==T, 0, n.events.massmob.post))
+
+nelda2.sub <- nelda2.sub %>% mutate(mm_duration.pre = ifelse(is.na(mm_duration.pre)==T, 0, mm_duration.pre))
+nelda2.sub <- nelda2.sub %>% mutate(mm_duration.post = ifelse(is.na(mm_duration.post)==T, 0, mm_duration.post))
+
+```
+
+## Writing
+
+```{r}
+write.csv(nelda2.sub, here::here("Data", "nelda2_sub_2024-02-06.csv"))
 ```

### 2024-02-07 Output/appendix_v4.Rmd
Updates to appendix

@@
 library(ggeffects)
-library(pscl)
 library(survival)
@@
 
+## Any protest
+
+nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(n.events.post > 0, 1, 0))
+
+## Surv status
+library(survival)
+library(ggsurvfit)
+nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
+
 ```
@@
 Figure \@ref(fig:prot-rep-fraud) shows the absence of such an effect.
-In fact, there is no significant change in the predicted level of fraud at lower levels of physical integrity, even when pre-election protest is widespread.
-We do observe a significant effect when comparing high- and low-repression regimes in highly extreme cases\textemdash there are only three cases in the dataset with more than 70 protest events\textemdash and is in the opposite direction predicted by this counterargument.
+In fact, there is no significant change in the predicted level of fraud at lower levels of physical integrity, across all levels of pre-election protest.
+We do observe a significant effect for low repression regimes; however, it is in the opposite direction predicted by this counterargument.
 That is, more numerous pre-election protest is associated with higher predicted levels of fraud in less repressive states, not harsher ones.
 The figure also shows that less repressive regimes appear to respond to greater civic unrest with more fraud.
-Altogether, this robustness check shows that pre-election protest in a repressive environment does not appear to deter fraud, and supports the view that manipulation may be used to attempt to claim a large margin of victory or to prevent elite or bureaucratic splits in a challenging period.
+Altogether, this robustness check shows that pre-election protest in a repressive environment does not appear to deter fraud, and supports the view that manipulation may be used to indicate regime durability in a challenging period.
 
@@
 #summary(m1_rep)
-p.m1rep <- sjPlot::plot_model(m1_rep, type = "pred", terms = c("n.events.pre", "physinteg.1lag [.25, .63]"))
+p.m1rep <- sjPlot::plot_model(m1_rep, type = "pred", terms = c("n.events.pre", "physinteg.1lag [.2, .7]")) #.25 and .75 quantile
 p.m1rep + labs(title = "Predicted values of election fraud", x = "Number of pre-election protests", y = "Election fraud", color = "Physical integrity (lag)") + theme_bw()
@@
 nelda2.sub <- nelda2.sub %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))
-nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))
 
-model.n.base2.succ <- glmmTMB::glmmTMB(n.events.post ~ incumb.status + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  +   physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
-                                         zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
-#summary(model.n.base2.succ)
+## Protest binary
 
-model.n.base2.succ.bin <- glm(no_protest ~ incumb.status + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  +   physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
-                              data = nelda2.sub  %>% filter(inc.lose == "0" & v2x_regime <= 2))
-#summary(model.n.base2.succ.bin)
+model.margin.int.suc <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + incumb.status  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+summary(model.margin.int.suc)
 
-plot_binary_margin_succ_n <- sjPlot::plot_model(model.n.base2.succ.bin, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]")) #To set condition for successor value, use , condition = c(incumb.status = "Contested exec.
-election")
-plot_binary_margin_succ_n <- plot_binary_margin_succ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw()
+## Duration
 
-####
-
-model.d.base2.succ <- glmmTMB::glmmTMB(d.events.num ~ incumb.status + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  +   physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
-                                          zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
-#summary(model.d.base2.succ)
-
-model.size.base2 <- glmmTMB::glmmTMB(max_size ~ incumb.status + jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  v2elirreg.inv*winner.share.upd  +    physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2),
-                                family=truncated_nbinom2)
-#summary(model.size.base2)
-
-model.size.base2.succ.bin <- glm(no_size ~ incumb.status + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  +   physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
-                              data = nelda2.sub  %>% filter(inc.lose == "0" & v2x_regime <= 2))
-#summary(model.size.base2.succ.bin)
-
-plot_binary_margin_succ_size <- sjPlot::plot_model(model.size.base2.succ.bin, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]"))
-plot_binary_margin_succ_size <- plot_binary_margin_succ_size +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw()
+cox2.suc <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + incumb.status +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv*winner.share.upd  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+summary(cox2.suc)
 
@@
 models4 <- list()
-models4[['Model 8']] <- model.n.base2.succ
+models4[['Model 8']] <- model.margin.int.suc
 #models4[['Model 6']] <- model.d.base2.hurd
-models4[['Model 9']] <- model.size.base2
+models4[['Model 9']] <- cox2.suc
 
@@
 ```{r n-protests-margin, echo=FALSE, message=FALSE, warning=FALSE}
-dat <- ggpredict(model.n.base2.succ, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "count") 
+dat <- ggpredict(model.margin.int.suc, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]")) 
 p.n.margin <- plot(dat)
-p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Number of events", title = "Predicted number of protests", color = "Election fraud") + theme_bw() + coord_cartesian(ylim = c(0, 40))
+p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Probability", title = "Predicted probability of protests", color = "Election fraud") + theme_bw()
 ```
@@
 ```{r size-protests-margin, echo=FALSE, message=FALSE, warning=FALSE}
-dat <- ggpredict(model.size.base2, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "count") 
-p.s.margin <- plot(dat)
-p.s.margin <- p.s.margin +  labs(x = "Incumbent vote-share", y = "Size of events", title = "Predicted size of protests", color = "Election fraud") + theme_bw() + coord_cartesian(ylim = c(0, 18000))
+sjPlot::plot_model(cox2.suc, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.45, 1.64]"))
 ```
@@
 
-# Electoral authoritarian regimes only
+# Including liberal democracies
 
-The models in the main text make use of observations from country-years coded as electoral authoritarian and electoral democratic by the 'v2x_regime' variable in V-Dem.
-This approach is taken because it reflects the widespread availability of election-manipulation techniques across this range of regime types.
-But it may be that considering the two groups simultaneously obscures some within-group patterns.
-For example, it may be that electoral protest is more likely to be observed in electoral authoritarian regimes following the logic in Tucker [-@tucker2007], and given several prominent examples of electoral revolution in that category (e.g.
-Ukraine 2004, Georgia 2003).
-The models in Table \@ref(tab:auth-table) show a statistically significant coefficient for the interaction of election fraud and winner's vote-share, but Figures \@ref(fig:n-protests-margin-auth) and \@ref(fig:size-protests-margin-auth) show that this effect is not apparent for typical values of election fraud in that subset of the population (the first and third quartiles, respectively).
-As a result, there is no strong evidence to suggest that election fraud has different effects on protest dynamics in electoral authoritarian vs.
-electoral democratic regimes.
-
-```{r, echo=FALSE}
-model.n.base2.auth <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                          zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub %>% filter(v2x_regime == 1))
-#summary(model.n.base2.auth)
-
-model.size.base2.auth <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                          zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub %>% filter(v2x_regime == 1))
-#summary(model.size.base2.auth)
-
-```
-
-```{r auth-table, echo=FALSE, warning=FALSE, message=FALSE}
-
-models5a <- list()
-models5a[['Model 10']] <- model.n.base2.auth
-models5a[['Model 11']] <- model.size.base2.auth
-
-t6b <- huxreg(models5a, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Incumbent term-limited" = "inc.termlimit.num", "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd"), statistics = character(0)) %>% set_caption("Models of number and size of protests, only electoral authoritarian regimes") %>% insert_row("", "Number", "Size", after = 0)
-theme_compact(t6b, header_rows = TRUE, header_cols = FALSE)  %>% insert_row("Controls", "Yes", "Yes", after = 12) %>% set_bottom_border(12, everywhere)
-```
-
-```{r, echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.base2.auth, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [.35, 1.86]"), type = "count") 
-p.n.margin.auth <- plot(dat)
-p.n.margin.auth <- p.n.margin.auth + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events", color = "Election fraud")  + theme_bw()
-
-dat <- ggpredict(model.size.base2.auth, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [.35, 1.86]")) 
-p.size.margin.auth <- plot(dat)
-p.size.margin.auth <- p.size.margin.auth + labs(x = "Incumbent vote-share", y = "Predicted size of protests" , title = "Predicted maximum size of protest events", color = "Election fraud")  + theme_bw()
-
-```
-
-```{r n-protests-margin-auth, fig.cap="Election fraud, incumbent vote-share, and number of protest events in authoritarian regimes", echo=FALSE, out.width = '75%', fig.align='center', fig.pos="h!"}
-p.n.margin.auth
-
-```
-
-```{r size-protests-margin-auth, fig.cap="Election fraud, incumbent vote-share, and maximum size of protest events in authoritarian regimes", echo=FALSE, out.width = '75%', fig.align='center', fig.pos="h!"}
-p.size.margin.auth
-
-```
-
-# Including liberal democracies
+[MOVE TO SEPARATE CHILD FILE TO USE SEPARATE DATASET]
 
@@
 The models presented below utilize this variable in place of the main measure of election fraud, while retaining all other control and explanatory variables, using the ECAV data on protest.
+Since the dependent variable of these models is on an inverse scale, compared to the variable *intentional irregularities*, we should expect to see coefficients for key variables take on opposite-signed values from the models in the main text.
 
 ```{r manipdv-full, echo=FALSE}
-m1 <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate + n.events.pre +  physinteg.1lag + physintlag.sq    + unscheduled_election   + (1 | stateid), data = nelda2.sub)
+m1 <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + unscheduled_election  + v2elembaut +   (1 | stateid), data = nelda2.sub)
 
 #summary(m1)
+
+m1a <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag + v2elembaut +  (1 | stateid), data = nelda2.sub)
+#summary(m1a)
+
+m1b <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag +  unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)
+
+#summary(m1b)
+
+m1c <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub)
+#summary(m1c)
 
@@
 ```{r models-hurdle-fulldv, warning = FALSE, message = FALSE, include = F}
-model.n.base2.elfr <- glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag +  + osorg.1lag  + natlpartyorg.lag +  v2xel_frefair*winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
-                               ziformula=~.,
-                              data = nelda2.sub %>% filter(v2x_regime <= 2 & inc.lose == "0"),
-                                family=truncated_nbinom2)
-summary(model.n.base2.elfr)
 
-model.size.base2.elfr <- lmer(log(max_size) ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election +  as.factor(inc.lose) + (1 | stateid), data = nelda2.sub %>% filter(max_size > 0))
-summary(model.size.base2.elfr)
+model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd +  presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+summary(model.margin.binary)
 
-model.n.griev.elfr <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2xel_frefair +  v2xel_frefair*gdpgro.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2),
-                                family=truncated_nbinom2)
+model.margin.binary.int <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+summary(model.margin.binary.int)
 
-model.size.base2.elfr <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2xel_frefair + gdpgro.1lag  +   winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2),
-                                family=truncated_nbinom2)
-
-model.size.griev.elfr <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2xel_frefair +  v2xel_frefair*gdpgro.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
-                                ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2),
-                                family=truncated_nbinom2)
+model.margin.binary_noc <- glm(any_protest ~ jucon.1lag + legcon.1lag +  altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+summary(model.margin.binary_noc)
 
@@
 ```{r, echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.base2.elfr, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2xel_frefair [.197, .61]"), type = "count") 
-p.n.margin <- plot(dat)
-p.n.margin <- p.n.margin + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events",  color = "Election integrity")  + theme_bw() + coord_cartesian(ylim=c(0, 75))
-
-dat <- ggpredict(model.size.base2.elfr, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2xel_frefair [.197, .61]")) 
-p.size.margin <- plot(dat)
-p.size.margin <- p.size.margin + labs(x = "Incumbent vote-share", y = "Predicted size of protests" , title = "Predicted maximum size of protest events", color = "Election integrity")  + theme_bw() + coord_cartesian(ylim = c(0, 50000))
+sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c("winner.share.upd", "v2xel_frefair [.12, .46]"))
 
@@
 
-```{r n-protests-margin-fulldv, fig.cap="Election integrity, incumbent vote-share, and number of protest events", echo=FALSE, out.width = '75%', fig.align='center', fig.pos="h!"}
-p.n.margin
+```{r}
 
-```
+cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2xel_frefair  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+summary(cox1)
 
-```{r protests-size-fulldv, fig.cap="Election integrity, incumbent vote-share, and size of protest events", echo=FALSE, out.width = '75%', fig.align='center', fig.pos="h!"}
-p.size.margin
+cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2xel_frefair*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+summary(cox2)
+p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2xel_frefair [.14, .43]"))  #This is the 10th and 90th quantiles
+p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ration", title = "Hazard ratio for protest movement end", color = "Election free and fair") + theme_bw()
 

### 2024-02-09 Output/appendix_v4.Rmd
Changes to appendices and fixes in main text

@@
 
+# Alternative measures of election manipulation
+
+To ensure that the results presented in the main text are not biased by the chosen measure of election fraud, this section replaces that measure with a broader indicator of election integrity.
+This variable is `v2xel_frefair,' from V-Dem, which is a comprehensive indicator of quality of the electoral process, including measures of electoral management body autonomy and capacity, vote-buying, intimidation, fraud, and an overall assessment of the fairness of the election.
+This variable ranges from 0 to 1, with higher values indicating freer and fairer elections.
+The models presented below utilize this variable in place of the main measure of election fraud, while retaining all other control and explanatory variables, using the ECAV data on protest.
+Since the dependent variable of these models is on an inverse scale, compared to the variable *intentional irregularities*, we should expect to see coefficients for key variables take on opposite-signed values from the models in the main text.
+
+```{r manipdv-full, echo=FALSE}
+m1 <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + unscheduled_election  +   (1 | stateid), data = nelda2.sub)
+
+#summary(m1)
+
+m1a <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag  +  (1 | stateid), data = nelda2.sub)
+#summary(m1a)
+
+m1b <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag +  unscheduled_election  +  (1 | stateid), data = nelda2.sub)
+
+#summary(m1b)
+
+m1c <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + unscheduled_election   +  (1 | stateid), data = nelda2.sub)
+#summary(m1c)
+
+models <- list()
+models[['Model 1']] <- m1
+models[['Model 2']] <- m1a
+models[['Model 3']] <- m1b
+models[['Model 4']] <- m1c
+
+modelsummary(models, stars = T,  title = 'Random-intercept multilevel model of overall election integrity', gof_omit = "AIC|BIC",
+             coef_rename = 
+               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", "presidential" = "Presidential", "n.events.pre" = "N. of pre-election protest",
+                 "physinteg.1lag" = "Physical integrity (lag)",
+                 "physintlag.sq" = "Phys. integ. (lag), squared", "unscheduled_election1" = "Off-schedule election",
+                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
+                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", "rentierstate1" = "Rentier state", "taxstate1" = "Tax state",
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests",
+                 "state.owned.1lag" = "Private ownership (lag)",
+                 "patronage.appts.1lag" = "Professional bureaucracy (lag)",
+                 "regional_success_lag" = "Regional protest diffusion"))
+```
+
+As expected, the first model shows that the number of pre-election protests is negatively associated with overall election integrity\textemdash contrary to deterrence expectations and in line with the argument of the main text.
+We also see that overall election integrity is lowest at middle levels of repression; this is, again, in line with the expectations of the incumbent strength model and contrary to deterrence expectations.
+No significant effect on overall election integrity is observed for successful electoral protests in the region.
+
+```{r models-hurdle-fulldv, warning = FALSE, message = FALSE, include = F, echo=FALSE}
+
+model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd +  presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+#summary(model.margin.binary)
+
+model.margin.binary.int <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+#summary(model.margin.binary.int)
+
+model.margin.binary_noc <- glm(any_protest ~ jucon.1lag + legcon.1lag +  altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+#summary(model.margin.binary_noc)
+
+```
+
+Moving to the logit models of post-election protest, we observe the same interaction relationship between overall election integrity, incumbent vote-share, and protest as is shown for fraud in the main text.
+That is, less free-and-fair elections are only associated with a higher protest risk and low levels of incumbent vote-share.
+
+```{r logit-table, echo=FALSE}
+models2 <- list()
+models2[['Model 5']] <- model.margin.binary
+models2[['Model 6']] <- model.margin.binary.int
+models2[['Model 7']] <- model.margin.binary_noc
+
+modelsummary(models2, output = "kableExtra",  stars = T, title = 'Logit models of protest occurrence', coef_rename = 
+               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
+                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
+                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
+                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
+integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity",
+                 "v2xel_frefair" = "Overall election integrity",
+                 "winner.share.upd" = "Incumbent vote-share",
+                 "v2xel_frefair:winner.share.upd" = "Election integ. : incumbent vote-share"
+                 )) %>% kable_styling(latex_options = "HOLD_position")
+```
+
+\newpage
+
+```{r overall-binary, fig.cap="Overall election integrity, winning vote-share, and probability of protest initiation", fig.height=6, echo=FALSE, message=FALSE, fig.align='center'}
+sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c("winner.share.upd", "v2xel_frefair [.12, .46]"))  +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election integrity") + theme_bw()
+
+```
+
+```{r, echo=FALSE}
+
+cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2xel_frefair  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+#summary(cox1)
+
+cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2xel_frefair*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+#summary(cox2)
+p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2xel_frefair [.14, .43]"))  #This is the 10th and 90th quantiles
+p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election free and fair") + theme_bw()
+
+```
+
+```{r cox-models-tab-overall, echo=FALSE, warning=FALSE}
+models3 <- list()
+models3[['Model 8']] <- cox1
+models3[['Model 9']] <- cox2
+
+modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Cox proportional hazard models of protest duration', coef_rename = 
+               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
+                 "osorg.1lag" = "Civil soc. openness (lag)", 
+                 "winner.share.upd" = "Incumbent vote-share",
+                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
+                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
+integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "v2xel_frefair" = "Overall election integrity",
+                 "n.events.post" = "N. post-election protest"
+                 ))
+
+```
+
+```{r margin-dur-fig-overall, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
+p.duration.margin
+```
+
+Likewise, the models of protest movement survival are also in line with those in the main text.
+Table \@ref(tab:cox-models-tab-overall) shows that cleaner elections are associated with a slower end to protest movements (here, a negative coefficient).
+When elections are more unfair overall, incumbents are able to more quickly bring protest to an end.
+Figure \@ref(fig:margin-dur-fig-overall) shows that this effect is significant in interaction with low-to-moderate incumbent vote-shares.
+
+These three sets of models all replace a measure of *election fraud* with a broader measure of overall election integrity.
+In all three cases, the results broadly confirm the incumbent-strength model and are unsupportive of the deterrence model.
+
+\newpage
+
 # Repression and pre-election protest
@@
 
-# Prior manipulation and expecations
-
-Another explanation for post-election protest could hinge on changes over time; elections that are much more fraudulent than prior elections may exacerbate citizens' grievances or convince opposition groups that the incumbent is unexpectedly weak.
-This possibility is tested in two ways.
-First, the *election fraud* variable is replaced with *fraud change*, the difference between the current election and the one immediately prior.
-Since the prior election's value is subtracted from the 'current' one, positive values mean that the current election was more fraudulent than its predecessor, while negative values indicate election fraud declined.
-As in other examples, there is no statistically significant for typical values.
-The values for change in fraud used in the marginal effects plots in Figures \@ref(fig:n-protests-ch) and \@ref(fig:size-protests-ch) represent large, but not unusual changes in election fraud from election to election: roughly one standard deviation above and below the mean.
-As shown in Figure \@ref(fig:n-protests-ch), the expected number of protests does appear to decline for significantly cleaner elections when comparing very low vs.
-very high incumbent vote-shares.
-However, there is no significant change for more fraudulent elections\textemdash precisely where the deterrent logic would expect to see an effect.
-Moreover, the two types of elections show no significant difference between them at any level of incumbent vote-share.
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-nelda2.sub <- nelda2.sub %>% mutate(fraud.change = v2elirreg.inv - v2elirreg.inv.lag) #Positive values mean worsening fraud
-
-model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + fraud.change +  +    winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.margin.binary)
-
-#summary(model.n.base2.lag)
-
-#model.d.base2.lag <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + fraud.change*winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag ,
-#                                          zi = ~.,
-#                                          family= nbinom2,
-#                                          data = nelda2.sub %>% filter(v2x_regime <= 2))
-#summary(model.d.base2.lag)
-
-model.n.base2.lag <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + fraud.change*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-#summary(model.n.base2.lag)
-
-```
-
-```{r table-lagged-manip, results='asis', echo=FALSE, warning=FALSE}
-models3 <- list()
-models3[['Model 4']] <- model.n.base2.lag
-#models3[['Model 5']] <- model.d.base2.lag
-models3[['Model 5']] <- model.size.base2.lag
-
-t4a <- huxreg(models3, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Fraud change" = "fraud.change", "Incumbent vote-share" = "winner.share.upd", "Fraud change:Incumbent vote-share" = "fraud.change:winner.share.upd"),
-              statistics = character(0)) %>% set_caption("Models of number and size of protests") %>% insert_row("", "Number", "Size", after = 0) %>% insert_row("Controls", "Yes", "Yes", after = 10)
-theme_compact(t4a, header_rows = TRUE, header_cols = FALSE)  %>% set_bottom_border(10, everywhere)
-
-```
-
-```{r n-protests-ch, fig.cap="Change in fraud, incumbent vote-share, and number of protest events", out.width = '75%', echo=FALSE, message=FALSE, fig.align='center'}
-dat <- ggpredict(model.n.base2.lag, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "fraud.change [-.58, .58]"), type = "count") 
-p.n.margin.ch <- plot(dat)
-p.n.margin.ch <- p.n.margin.ch +  labs(x = "Incumbent vote-share", y = "Number of events", title = "", color = "Change in fraud") + theme_bw()
-p.n.margin.ch
-```
-
-```{r size-protests-ch, fig.cap="Change in fraud, incumbent vote-share, and size of protest events", out.width = '75%', echo=FALSE, message=FALSE, fig.align='center'}
-dat <- ggpredict(model.size.base2.lag, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "fraud.change [-.58, .58]"), type = "count") 
-p.s.margin.ch <- plot(dat)
-p.s.margin.ch <- p.s.margin.ch +  labs(x = "Incumbent vote-share", y = "Size of protest wave", title = "", color = "Change in fraud") + theme_bw() + coord_cartesian(ylim = c(0, 20000))
-p.s.margin.ch 
-```
-
-Second, the NELDA dataset includes a binary variable noting whether "there was evidence of domestic or international concern that the election process was not going to be free or fair" in advance of the election (nelda11).
-While this is likely to be rough indicator, it still provides an opportunity to test the prediction that election fraud may be more likely trigger protest when it exceeds a baseline expectation.
-In particular, when fraud is high and there are no pre-election expectations of severe manipulation, this view might predict a higher likelihood and severity of protest.
-This is tested by interacting *intentional voting irregularities* with the binary variable from NELDA, called *pre-election expectations* here.
-
-```{r expectations-table, echo=FALSE, message=FALSE, warning=FALSE}
-model.n.concerns.hurd <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    preelection_concerns*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
-                                          zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
-#summary(model.n.concerns.hurd)
-
-model.size.concerns.hurd <- glmmTMB::glmmTMB(max_size ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    preelection_concerns*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
-                                          zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
-#summary(model.size.concerns.hurd)
-
-models8 <- list()
-models8[['Model 6']] <- model.n.concerns.hurd
-models8[['Model 7']] <- model.size.concerns.hurd
-
-t8a <- huxreg(models8, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Election fraud" = "v2elirreg.inv", "Pre-election concerns" = "preelection_concerns", "Election fraud : Pre-election concerns" = "v2elirreg.inv:preelection_concerns"), statistics = character(0)) %>% set_caption("Models of number and size of protests") %>% insert_row("", "Number", "Size", after = 0)
-theme_compact(t8a, header_rows = TRUE, header_cols = FALSE)  %>% insert_row("Controls", "Yes", "Yes", after = 10) %>% set_bottom_border(10, everywhere)
-```
-
-As Table \@ref(tab:expectations-table) and Figures \@ref(fig:n-protests-concerns) and \@ref(fig:protests-size-concerns) indicate, there is no apparent effect for high levels of fraud when there are no significant pre-election concerns about the freeness and fairness of the election (as coded by NELDA).
-Taken together, these results suggest that the alternative, 'exceeding expectations' framework for the protest model of election manipulation is not supported.
-
-```{r, echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.concerns.hurd, terms = c("v2elirreg.inv", "preelection_concerns"), type = "count") 
-p.n.concerns <- plot(dat)
-p.n.concerns <- p.n.concerns + labs(x = "Intentional irregularities", y = "Predicted number of protests" , title = "Predicted count of protest events", color = "Pre-election concerns")  + theme_bw()
-
-dat <- ggpredict(model.size.concerns.hurd, terms = c("v2elirreg.inv", "preelection_concerns")) 
-p.size.concerns <- plot(dat)
-p.size.concerns <- p.size.concerns +  labs(x = "Intentional irregularities", y = "Predicted size of protests" , title = "Predicted max.
-size of protest events", color = "Pre-election concerns")  + theme_bw() + coord_cartesian(ylim = c(0, 50000))
-
-```
-
-```{r n-protests-concerns, fig.cap="Election fraud, pre-election expectations, and number of protest events", echo=FALSE, out.width = '75%', fig.align='center'}
-p.n.concerns
-
-```
-
-```{r protests-size-concerns, fig.cap="Election fraud, pre-election expectations, and size of protest events", echo=FALSE, out.width = '75%', fig.align='center'}
-p.size.concerns
-
-```
+\newpage
 
@@
 Using a variable from NELDA that codes for designated successors, the following models add a categorical variable for the incumbent's status, which can take on the following values: *executive eligible*, *chosen successor*, *contested executive election*, and *legislative election*.
+Including this variables requires the exclusion of the original variables *presidential election* and *incumbent term limited*.
 
@@
 model.margin.int.suc <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + incumb.status  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.margin.int.suc)
+#summary(model.margin.int.suc)
 
@@
 cox2.suc <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + incumb.status +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv*winner.share.upd  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
-summary(cox2.suc)
+#summary(cox2.suc)
 
@@
 models4 <- list()
-models4[['Model 8']] <- model.margin.int.suc
+models4[['Model 10']] <- model.margin.int.suc
 #models4[['Model 6']] <- model.d.base2.hurd
-models4[['Model 9']] <- cox2.suc
+models4[['Model 11']] <- cox2.suc
 
@@
 election" = "incumb.statusContested exec.
-election", "Intentional irregularities" = "v2elirreg.inv", "Incumbent vote-share" = "winner.share.upd", "Intentional irregularities : Incumbent vote-share" = "v2elirreg.inv:winner.share.upd"),  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Models of number and size of protests") %>% insert_row("", "Number", "Size", after = 0) %>% insert_row("Controls", "Yes", "Yes", after = 14)
+election", "Intentional irregularities" = "v2elirreg.inv", "Incumbent vote-share" = "winner.share.upd", "Intentional irregularities : Incumbent vote-share" = "v2elirreg.inv:winner.share.upd"),  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Models of number and size of protests") %>% insert_row("", "Protest initiation", "Protest end", after = 0) %>% insert_row("Controls", "Yes", "Yes", after = 14)
 theme_compact(t5a, header_rows = TRUE, header_cols = FALSE)
@@
 ```{r n-protests-margin, echo=FALSE, message=FALSE, warning=FALSE}
-dat <- ggpredict(model.margin.int.suc, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]")) 
+dat <- ggpredict(model.margin.int.suc, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), condition = c(incumb.status = "Executive eligible")) 
 p.n.margin <- plot(dat)
@@
 
-```{r size-protests-margin, echo=FALSE, message=FALSE, warning=FALSE}
-sjPlot::plot_model(cox2.suc, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.45, 1.64]"))
+```{r duration-protests-margin, echo=FALSE, message=FALSE, warning=FALSE}
+p.s.margin <- sjPlot::plot_model(cox2.suc, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.45, 1.64]")) +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio of protest end", color = "Election fraud") + theme_bw()
+
 ```
@@
 As Figure \@ref(fig:successor-fig) show, the results are largely similar to those in the main text.
-For these figures, *incumbent status* was set to its modal value\textemdash an executive election in which the incumbent is eligible.
-In this model, higher levels of fraud reduce the likelihood of no protest following the election, at low levels of incumbent vote-share..
-This is in line with the expectation of the deterrence logic.
-However, this does not translate into a higher risk of more numerous or larger protest events by level of manipulation.
-Table \@ref(tab:successor-table), abridged here to show the key variables shows that there is no effect of successor type on the size of electoral protest.
-On the other hand, elections where there is a chosen successor appear to prompt more numerous protest than when the incumbent runs again.
-This could indicate dissatisfaction among the losing faction in the internal struggle for succession, and/or unhappiness among the democratic opposition surrounding a choreographed transition.
+For these figures, *incumbent status* was set to an executive election in which the incumbent is eligible.
+In this model, higher levels of fraud increase the likelihood of protest following the election, at low levels of incumbent vote-share; this is in line with expectations of both the deterrence and incumbent-strenght models.
+However, the results for protest survival models support the incumbent-strength model.
 
-\newpage
+Table \@ref(tab:successor-table), abridged here to show the key variables, shows that there is no effect of successor type on the duration of electoral protest.
+There is, however, the same relationship between election fraud, incumbent vote-share, and protest survival that is found in the main text.
+That is, incumbents who report low vote-shares but make use of large-scale fraud are better able to end protest movements than those who produce less fraud.
 
 ```{r successor-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=6, echo=FALSE, message=FALSE, fig.align='center'}
-plot.combined.prob <- cowplot::plot_grid(plot_binary_margin_succ_n, p.n.margin, p.s.margin, ncol=1)
+plot.combined.prob <- cowplot::plot_grid(p.n.margin, p.s.margin, ncol=1)
 plot.combined.prob
@@
 
-# Including liberal democracies
-
-[MOVE TO SEPARATE CHILD FILE TO USE SEPARATE DATASET]
-
-The main text excludes liberal democracies from the analysis, on the grounds that the prior literature on protest and fraud typically confines itself to regimes that hold multiparty elections, and the liberal-democratic cases exhibit little variation in their (low) levels of election fraud.
-But perhaps it is in liberal democracies where the deterrent logic operates most effectively, settling into an equilibrium of high protest risk and low fraud.
-Including liberal democracies in the analysis results in little difference from the findings in the main text, however.
-Election fraud does not appear associated with a higher number of protest events, or larger protest sizes.
-Results are presented in Table \@ref(tab:lib-table) and Figures \@ref(fig:n-protests-margin-lib) and \@ref(fig:size-protests-margin-lib)
-
-```{r lib-table, echo=FALSE, warning=FALSE}
-
-model.n.base2.lib <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                          zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub)
-#summary(model.n.base2.lib)
-
-model.size.base2.lib <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                          zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub)
-#summary(model.size.base2.lib)
-
-model.n.griev.lib <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                          zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub)
-#summary(model.n.base2.lib)
-
-model.size.griev.lib <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
-                                          zi = ~.,
-                                          family= nbinom2,
-                                          data = nelda2.sub)
-
-models5 <- list()
-models5[['Model 10']] <- model.n.base2.lib
-models5[['Model 11']] <- model.size.base2.lib
-models5[['Model 12']] <- model.n.griev.lib
-models5[['Model 13']] <- model.size.griev.lib
-
-t6a <- huxreg(models5, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd",
-                                                                                                                                                             "Election fraud : Winner vote-share" = "winner.share.upd:v2elirreg.inv",
-                                                       "GDP growth (lag)" =  "gdpgro.1lag",
-                          "Election fraud : GDP growth (lag)"  =  "gdpgro.1lag:v2elirreg.inv"), statistics = character(0)) %>% set_caption("Models of number and size of protests, inclu.
-liberal democracies") %>% insert_row("", "Number", "Size", "Number", "Size", after = 0)
-theme_compact(t6a, header_rows = TRUE, header_cols = FALSE)  %>% insert_row("Controls", "Yes", "Yes", "Yes", "Yes", after = 14) %>% set_bottom_border(14, everywhere)
-```
-
-```{r, echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.base2.lib, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [.35, 1.86]"), type = "count") 
-p.n.margin.lib <- plot(dat)
-p.n.margin.lib <- p.n.margin.lib + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events", color = "Election fraud")   + theme_bw()
-
-dat <- ggpredict(model.size.base2.lib, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [.35, 1.86]")) 
-p.size.margin.lib <- plot(dat)
-p.size.margin.lib <- p.size.margin.lib + labs(x = "Incumbent vote-share", y = "Predicted size of protests" , title = "Predicted maximum of size protest events", color = "Election fraud")  + theme_bw()
-
-```
-
-```{r n-protests-margin-lib, fig.cap="Election integrity, incumbent vote-share, and number of protest events (including liberal democracies)", echo=FALSE, out.width = '75%', fig.align='center', fig.pos="h!"}
-p.n.margin.lib
-
-```
-
-```{r size-protests-margin-lib, fig.cap="Election integrity, incumbent vote-share, and size of protest events (including liberal democracies)", echo=FALSE, out.width = '75%', fig.align='center', fig.pos="h!"}
-p.size.margin.lib
-
-```
-
-# Alternative measures of election manipulation
-
-To ensure that the results presented in the main text are not biased by the chosen measure of election fraud, this section replaces that measure with a broader indicator of election integrity from V-Dem.
-This variable is `v2xel_frefair,' which is a comprehensive indicator of quality of the electoral process, including measures of electoral management body autonomy and capacity, vote-buying, intimidation, fraud, and an overall assessment of the fairness of the election.
-This variable ranges from 0 to 1, with higher values indicating freer and fairer elections.
-The models presented below utilize this variable in place of the main measure of election fraud, while retaining all other control and explanatory variables, using the ECAV data on protest.
-Since the dependent variable of these models is on an inverse scale, compared to the variable *intentional irregularities*, we should expect to see coefficients for key variables take on opposite-signed values from the models in the main text.
-
-```{r manipdv-full, echo=FALSE}
-m1 <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + unscheduled_election  + v2elembaut +   (1 | stateid), data = nelda2.sub)
-
-#summary(m1)
-
-m1a <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag + v2elembaut +  (1 | stateid), data = nelda2.sub)
-#summary(m1a)
-
-m1b <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag +  unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)
-
-#summary(m1b)
-
-m1c <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub)
-#summary(m1c)
-
-models <- list()
-models[['Model 12']] <- m1
-
-modelsummary(models, stars = T,  title = 'Random-intercept multilevel model of overall election integrity', gof_omit = "AIC|BIC",
-             coef_rename = 
-               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
-                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", "presidential" = "Presidential", "n.events.pre" = "N. of pre-election protest",
-                 "physinteg.1lag" = "Physical integrity (lag)",
-                 "physintlag.sq" = "Phys. integ. (lag), squared", "unscheduled_election" = "Off-schedule election",
-                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
-                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
-                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
-                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
-                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests"))
-```
-
-The first model shows that the number of pre-election protests is not associated with improved overall election integrity.
-This suggests that pre-election unrest is not associated with broad-scale improvements in election integrity across the electoral environment.
-
-```{r models-hurdle-fulldv, warning = FALSE, message = FALSE, include = F}
-
-model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd +  presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.margin.binary)
-
-model.margin.binary.int <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.margin.binary.int)
-
-model.margin.binary_noc <- glm(any_protest ~ jucon.1lag + legcon.1lag +  altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.margin.binary_noc)
-
-```
-
-Moving to the hurdle models of post-election protest, we do not observe much of a relationship between overall election integrity and protest.
-There is no statistically significant relationship between election integrity and protest size or quantity in any of the models, either alone or interaction with winner vote-share and lagged GDP growth.
-Figure \@ref(fig:n-protests-margin-fulldv) shows the absence of any effect for the number of protest events, while Figure \@ref(fig:protests-size-fulldv) shows the same for protest size.
-The values for *overall election integrity* chosen for these marginal effects plots represent the first and third quartiles, respectively.
-As a whole, the results are in accordance with the theory proposed in the main paper.
-There is no consistent evidence that lower-quality elections are associated with either a higher risk of protest initiation, or with a larger number of post-election protests.
-
-```{r zero-model-table-fulldv, echo=FALSE, results='asis', warning=FALSE}
-
-models <- list()
-#models[['Model XXa']] <- model.n.base2.elfr
-models[['Model 13']] <- model.n.base2.elfr
-models[['Model 14']] <- model.n.griev.elfr
-models[['Model 15']] <- model.size.base2.elfr
-models[['Model 16']] <- model.size.griev.elfr
-
-t7 <-  huxreg(models, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,
- statistics = character(0), coefs = c("Intercept" = "(Intercept)", "Incumbent term-limted" = "inc.termlimit.num", "Electoral integrity" = "v2xel_frefair", "Incumbent vote-share" = "winner.share.upd", "GDP growth (lag)" = "gdpgro.1lag", "Electoral integrity:Incumbent vote-share" = "v2xel_frefair:winner.share.upd",  "Election integrity : GDP growth (lag)" = "gdpgro.1lag:v2xel_frefair")) %>% set_caption("Conditional hurdle models of protest number and size") %>% insert_row("", "Number", "", "Size", "", after = 0)
-theme_compact(t7, header_rows = TRUE, header_cols = FALSE) %>% insert_row("Controls", "Yes", "Yes", "Yes", "Yes", after = 16) %>% set_bottom_border(16, everywhere)
-
-```
-
 \newpage
-
-```{r, echo=FALSE, message=FALSE}
-sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c("winner.share.upd", "v2xel_frefair [.12, .46]"))
-
-```
-
-```{r}
-
-cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2xel_frefair  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
-summary(cox1)
-
-cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2xel_frefair*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
-summary(cox2)
-p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2xel_frefair [.14, .43]"))  #This is the 10th and 90th quantiles
-p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ration", title = "Hazard ratio for protest movement end", color = "Election free and fair") + theme_bw()
-
-```
-
-# Proportional hazards diagnostics
-
-Cox proportional hazard models assume that the hazard ratio for each predictor is constant over time.
-A test of this assumption was developed by Schoenfeld [-@schoenfeld1982partial], in which risk-weighted covariate means are estimated for each point in time.
-Individual units' observed values for these covariates are then subtracted from the risk-weighted mean, giving the Schoenfeld residuals.
-As described by Grambsch and Therneau [-@grambsch1994proportional], these residuals can be plotted against event time with a smoothing spline to illustrate the general trend.
-The expected value for this relationship over time should be a horizontal line around zero, if the assumption of constant hazard ratio holds.
-Figure \@ref(fig:cox-diag) shows this relationship for *intentional voting irregularities*, with confidence bands of +/- two standard errors of the residuals at each time point.
-A horizontal line marks the zero line, while vertical lines indicate rough points where the curve of the smoothed line shifts, indicating cutpoints in the function.
-These cutpoints are used to estimate changes in the hazard function in Model 13 in the main text.
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-library(survival)
-library(ggsurvfit)
-nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
-
-cox1 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + v2elirreg.inv , data = nelda2.sub %>% filter(v2x_regime <= 2 & inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
-#summary(cox1)
-cz1 <- cox.zph(cox1, transform = "identity")
-#print(cz1)
-```
-
-```{r cox-diag, fig.cap="Schoenfeld residuals for election fraud by time (Model 12 in main text)", echo=FALSE, out.width = '85%', fig.align='center', warning=FALSE, message=FALSE, fig.pos="h!"}
-plot(cz1[18])
-abline(h=0)
-abline(v = 11)
-abline(v = 25)
-```
 
@@
 In particular, many of the control variables\textemdash such as *legislative constraints*, *judicial independence*, and *physical integrity* are likely to be correlated with pre-election protest.
-To address this risk, the models in Tables \@ref(tab:zero-model-table-small) and \@ref(tab:zero-model-table-count-small) exclude non-electoral control variables.
+To address this risk, the models in Tables \@ref(tab:logit-table-small) and \@ref(tab:cox-models-tab-small) exclude non-electoral control variables.
 That is, they include variables related to the election itself, the measure of *pre-election protest*, and the explanatory variables from the main models.
@@
 The results echo those from the main text.
-Fraudulent elections are associated with a higher risk of protest initiation at low levels of ruling-party vote-share (Figure \@ref(fig:b-protests-margin)), though this does not translate into a higher risk of experiencing large protests (Figure \@ref(fig:n-protests-margin-small)).
+Fraudulent elections are associated with a higher risk of protest initiation at low levels of ruling-party vote-share (Figure \@ref(fig:margin-prob-fig-small)), though these same conditions are associated with a higher likelihood that protest movements will fail quickly \@ref(fig:margin-dur-fig-small).
 
-```{r models-hurdlesmall-mm, warning = FALSE, message = FALSE, include = F}
-model.small.novoteshare.hurd <- hurdle(n.events.post ~ v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election |  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election,
-                              ziformula=~.,
-                                data = nelda2.sub %>% filter(inc.lose == "0"),
-                                dist = "negbin")
-#summary(model.small.novoteshare.hurd)
+```{r include=FALSE, echo=FALSE}
+nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
+#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))
 
-model.small.base.hurd <- hurdle(n.events.post ~  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election + winner.share.upd |  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election + winner.share.upd , data = nelda2.sub %>% filter(inc.lose == "0"), dist = "negbin")
-#summary(model.small.base.hurd)
+model.margin.binary <- glm(any_protest ~  v2elirreg.inv +  winner.share.upd  + presidential + inc.termlimit.num  +  n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+#summary(model.margin.binary)
 
-model.small.base2.hurd <- hurdle(n.events.post ~  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election + winner.share.upd + winner.share.upd*v2elirreg.inv | v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election + winner.share.upd  + winner.share.upd*v2elirreg.inv, data = nelda2.sub %>% filter(inc.lose == "0"), dist = "negbin")
-#summary(model.small.base2.hurd)
+model.margin.binary.int <- glm(any_protest ~ v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+#summary(model.margin.binary.int)
 
-model.small.griev.hurd <- hurdle(n.events.post ~  v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election + gdpgro.1lag + gdpgro.1lag*v2elirreg.inv | v2elirreg.inv + presidential + inc.termlimit.num  + n.events.pre + unscheduled_election + gdpgro.1lag  + gdpgro.1lag*v2elirreg.inv, data = nelda2.sub %>% filter(inc.lose == "0"), dist = "negbin")
-#summary(model.small.griev.hurd)
+## Plots
 
-```
-\newpage
+####Margin
 
-```{r zero-model-table-small, echo=FALSE}
+#plot_binary_margin_max <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-2.3, 3.2]"))
+#plot_binary_margin_max <- plot_binary_margin_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()
 
-models <- list()
-models[['Model 11']] <- model.small.base2.hurd
-models[['Model 12']] <- model.small.griev.hurd
-#models[['Model ']] <- model.n.griev.hurd.mm
-
-modelsummary::modelsummary(models, coef_omit = "count", gof_omit = "R2", stars = T, longtable=T,  title = 'Binary portion of hurdle models of post-election protest', coef_rename = 
-               c("zero_(Intercept)" = "Intercept", "zero_presidential1" = "Presidential election",
-                 "zero_inc.termlimit.num" = "Incumbent term-limited",  
-                                  "zero_n.events.pre" = "Number of pre-election protests", "zero_pubseccorrup.lag" = "Public sector corruption (lag)", "zero_v2elirreg.inv" = "Election fraud", "zero_unscheduled_election" = "Off-schedule election", "zero_winner.share.upd" = "Incumbent vote-share", "zero_gdpgro.1lag" = "GDP growth (lag)", "zero_presidential" = "Presidential",
-                 "zero_v2elirreg.inv:gdpgro.1lag" = "Election fraud : GDP growth (lag)", 
-                 "zero_v2elirreg.inv:winner.share.upd" = "Election fraud : Winner vote-share"))
+plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-.05, 2]"), show.data = T)
+plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()
 
@@
 
-```{r zero-model-table-count-small, echo=FALSE}
-modelsummary::modelsummary(models, coef_omit = "zero", gof_omit = "R2", longtable=T, stars = T, title = 'Hurdle models of post-election protest (counts)',
-                           coef_rename = 
-              c("count_(Intercept)" = "Intercept", "count_presidential1" = "Presidential election",
-                 "count_inc.termlimit.num" = "Incumbent term-limited",  
-                                  "count_n.events.pre" = "Number of pre-election protests", "count_pubseccorrup.lag" = "Public sector corruption (lag)", "count_v2elirreg.inv" = "Election fraud", "count_unscheduled_election" = "Off-schedule election", "count_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "winner.share.upd" = "Incumbent vote-share",
-                "count_gdpgro.1lag" = "GDP growth (lag)", "count_presidential" = "Presidential",
-                "count_v2elirreg.inv:gdpgro.1lag" = "Election fraud : GDP growth (lag)",
-                "count_v2elirreg.inv:winner.share.upd" = "Election fraud : Winner vote-share"))
+```{r logit-table-small, echo=FALSE}
+models2 <- list()
+models2[['Model 12']] <- model.margin.binary
+models2[['Model 13']] <- model.margin.binary.int
 
-```
-
-```{r b-protests-margin, fig.cap="Fraud, incumbent vote-share, and protest initiation", out.width = '75%', echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}
-dat <- ggpredict(model.small.base2.hurd, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "zi_prob") 
-p.zi.margin <- plot(dat)
-p.zi.margin <- p.zi.margin +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest initiation", color = "Election fraud") + theme_bw() + coord_cartesian(y = c(0, 1))
-p.zi.margin
-```
-
-```{r n-protests-margin-small, fig.cap="Fraud, incumbent vote-share, and number of protest events", out.width = '75%', echo=FALSE, message=FALSE, fig.align='center'}
-dat <- ggpredict(model.small.base2.hurd, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "count") 
-p.n.margin <- plot(dat)
-p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Number of events", title = "Predicted number of protests", color = "Election fraud") + theme_bw() + coord_cartesian(y = c(0, 25))
-p.n.margin
+modelsummary(models2, output = "kableExtra",  stars = T, title = 'Logit models of protest occurrence', coef_rename = 
+               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
+                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
+                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
+                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
+integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity",
+                 "v2elirreg.inv" = "Election fraud",
+                 "winner.share.upd" = "Incumbent vote-share"
+                 )) %>% kable_styling(latex_options = "HOLD_position")
 ```
@@
 
-# References
+```{r margin-prob-fig-small, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
+plot_binary_margin_typ_n
+```
+
+```{r, include=FALSE, echo=FALSE}
+library(survival)
+library(ggsurvfit)
+nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
+nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
+                                                                       ifelse(d.events.num >= 69, 0, 1)))
+
+cox1 <- coxph(Surv(d.events.num, surv_status) ~   winner.share.upd  + presidential + inc.termlimit.num  +   unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+#summary(cox1)
+
+cox2 <- coxph(Surv(d.events.num, surv_status) ~  v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +   unscheduled_election + regional_success_lag +  n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+#summary(cox2)
+p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.05, 2]"))  #This is the 10th and 90th quantiles
+p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()
+
+```
+
+```{r cox-models-tab-small, echo=FALSE, warning=FALSE}
+models3 <- list()
+models3[['Model 14']] <- cox1
+models3[['Model 15']] <- cox2
+
+modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Cox proportional hazard models of protest duration', coef_rename = 
+               c("winner.share.upd" = "Incumbent vote-share",
+                 "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "unscheduled_election" = "Off-schedule election",
+                 "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "regional_success_lag" = "Regional protest diffusion", "v2elirreg.inv" = "Election fraud",
+                 "n.events.post" = "N. post-election protests"
+                 ))
+
+```
+
+```{r margin-dur-fig-small, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
+p.duration.margin
+```
+
+\newpage

### 2024-02-09 Output/appendix_v4_part2.Rmd
Changes to appendices and fixes in main text

--- 2022-12-30 Output/appendix_v4_part2.Rmd
@@
 ---
+
 ```{r setup, include = FALSE}
@@
 library(stargazer)
-library(cowplot)
-library(pscl)
+library(ggeffects)
+library(survival)
 library(modelsummary)
-library(ggeffects)
-library(MASS)
+library(glmmTMB)
+library(huxtable)
+library(kableExtra)
+library(simPH)
 
-nelda_massmob <- read.csv(here::here("Data", "nelda_massmob_all.csv")) #nelda_complete.csv for only elex-related
+nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2024-02-06.csv"))
 
-nelda_massmob <- nelda_massmob %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
-nelda_massmob <- nelda_massmob %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))
+nelda2.sub <- nelda2.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
+nelda2.sub <- nelda2.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))
 
@@
 
-nelda_massmob <- nelda_massmob %>% mutate(physintlag.sq = physinteg.1lag^2)
+nelda2.sub <- nelda2.sub %>% mutate(physintlag.sq = physinteg.1lag^2)
 
-#Crisis as factor
+## Categorical variable for incumbent status
 
-nelda_massmob <- nelda_massmob %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))
+nelda2.sub <- nelda2.sub %>% mutate(incumb.status = as.factor(ifelse(inc.termlimit == "yes" & chosen.successor == 1, "Chosen successor",
+                                                           ifelse(inc.termlimit == "yes" & chosen.successor == 0, "Contested exec. election", 
+                                                                  ifelse(inc.termlimit == "no", "Executive eligible", "Legislative election"))))) 
+nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Legislative election"))
 
-#Margins with updates
+#Coding as factors
 
-nelda_massmob <- nelda_massmob %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))
+nelda2.sub <- nelda2.sub %>% mutate(presidential = factor(presidential))
+#nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = factor(inc.termlimit.num))
+nelda2.sub <- nelda2.sub %>% mutate(rentierstate = factor(rentierstate))
+nelda2.sub <- nelda2.sub %>% mutate(taxstate = factor(taxstate))
+nelda2.sub <- nelda2.sub %>% mutate(unscheduled_election = factor(unscheduled_election))
+nelda2.sub <- nelda2.sub %>% mutate(inc.lose = factor(inc.lose))
+nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))
 
@@
 
-nelda_massmob <- nelda_massmob %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | elections.cancel.prot, 1, 0))
+nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))
+
+# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections
+
+nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))
+
+## Replacing NA with 0 for protest size
+
+#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(is.na(max_size)==T & n.events.post == 0, 0, max_size))
+
+## Fixing missing values of winner.share.upd for liberal democracies
+
+nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(is.na(winner.share.upd)==T & v2x_regime == 3, winner.margin, winner.share.upd))
+
+## Set max_size zero for elections with no protest
+
+#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))
+
+## Excluding liberal democracies
+
+nelda2.sub <- nelda2.sub %>% filter(v2x_regime == 1 | regime.1lag == 1 | v2x_regime == 2 | regime.1lag == 2  ) #Includes only authoritarian, add e-dems with | v2x_regime == 2  #Current version includes lagged 1 in order to capture pre-election auth.
+
+## Remove outlier
+nelda2.sub <- nelda2.sub %>% filter(d.events.num <= 100)
 
@@
 
-#protest.cases <- nelda_massmob %>% filter(any.ecav.protest == 1)
+protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1 & inc.lose == 0)
 
-# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections
+## Coding median size as 0 when n.events = 0
 
-nelda_massmob <- nelda_massmob %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))
+nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(n.events.post == 0, 0, median.participants))
 
-## Changing participants from numeric category to text
-#nelda_massmob <- nelda_massmob %>% mutate(median.participants = round(median.participants))
-#nelda_massmob <- nelda_massmob %>% mutate(particpants_categorical = ifelse(is.na(median.participants)==T, NA, ifelse(median.participants == 1, "Less than 10",
-#                     ifelse(2, "10 to 99", ifelse(3, "100 to 999",
-#                                                  ifelse(4, "1000 to 9,999", ifelse(5, "Greater than 10,000", NA)))))))
-#nelda_massmob <- nelda_massmob %>% mutate(particpants_categorical = as.factor(particpants_categorical))
+## Any protest
 
-## Median particiants (MM) as factor
+nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(n.events.post > 0, 1, 0))
 
-nelda_massmob <- nelda_massmob %>% mutate(median.participants.mm.post = as.factor(median.participants.mm.post))
-
-## Inverting the vote buying measure so it is not reverse coded
-
-nelda_massmob <- nelda_massmob %>% mutate(v2elvotbuy_inv = -1*v2elvotbuy)
+## Surv status
+library(survival)
+library(ggsurvfit)
+nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
 
@@
 
-## Alternative measures of election manipulation
+# Including electoral democracies
 
-To ensure that the results presented in the main text are not biased by the chosen measure of election fraud, this section replaces that measure with a broader indicator of election integrity from V-Dem.
-This variable is `v2xel_frefair,' which is a comprehensive indicator of quality of the electoral process, including measures of electoral management body autonomy and capacity, vote-buying, intimidation, fraud, and an overall assessment of the fairness of the election.
-This variable ranges from 0 to 1, with higher values indicating freer and fairer elections.
-The models presented below utilize this variable in place of the main measure of election fraud, while retaining all other control and explanatory variables, using the Mass Mobilization data on protest.
+The theory articulated in the main text is primarily concerned with dynamics within authoritarian regimes, where patronage resources can be marshaled to support election-manipulation machines, and rule of law can be subverted to protect election-manipulating agents.
+These mechanisms should be expected to become attenuated, on average, as regimes become more democratic.
+As such, the main analysis is confined to authoritarian regimes (as coded in V-Dem) where multiparty elections are held.
+However, this approach imposes a binary cutoff on what is a continuum of regimes, and could introduce some bias by excluding edge cases.
 
-```{r manipdv, echo=FALSE}
-m1 <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   + presidential + inc.termlimit.num  +  libdem.1lag + urban.pct + rentierstate + taxstate + n.events.massmob.pre + median.participants.mm.pre  + (1 | stateid), data = nelda_massmob %>% filter(n.events.massmob.pre < 200)) #There are two extreme observations removed by this filter
+To check this possibility, the analysis here shows the results of models from the main text using a broader range of cases: including both the authoritarian regimes from the original analysis and electoral democracies (V-Dem variable v2x_regime = 2).
+I expect that changing the sample in this way will weaken the results somewhat, due to the inclusion of cases where theoretically important variables are weaker.
+For example, the average level of *judical constraints on the executive* in the original sample is 0.37, while the average in the electoral democracies in almost twice as high at 0.66.
+Similarly, the average incumbent vote-share in the original sample is 0.55, while in the electoral democracies it is 0.41.
+The liberal democracies have substantially more professionalized bureaucratic appointments on average (0.27) than the electoral authoritarian regimes (-0.33).
+Consequently I expect levels of electoral manipulation to be lower and less variable in the new sample, on average, if only because the electoral democracies have less developed tools to engage in large-scale fraud.
+
+Though the results are somewhat weaker, as expected, the general patterns hold.
+Table \@ref(tab:manipdv-ed) shows how levels of electoral manipulation respond to varying levels of pre-election risk to incumbents.
+As in the main text, the results show that more pre-election protest is associated with higher levels of fraud.
+We also see that higher levels of manipulation are seen at middle levels of repression, as in the main text.
+Two difference are apparent, however.
+Adding a number of higher-values cases reveals a straightforward, negative, statistically significant relationship between physical integrity protection and election fraud.
+However, goodness-of-fit measures indicate that the curvilinear relationship fits the data substantially better.
+Finally, while *regional protest success* has a positive coefficient, it is not statistically significant as it is in the main text.
+Despite the less theoretically appropriate sample, we still observe the predicted relationship in two models.
+In a a third, we see the coefficient is in the predicted direction, but loses its statistical significance.
+This is broadly supportive of the main argument.
+
+```{r manipdv-ed, echo=FALSE}
+
+m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag +  v2elembaut + (1 | stateid), data = nelda2.sub)
+#summary(m1a)
+
+m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag  + unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)
+
+#summary(m1b)
+
+m1c <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub)
+#summary(m1c)
+
+m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub)
 
@@
 
-modelsummary(m1, stars = T,  title = 'Random-intercept multilevel model of overall election integrity')
+m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag + v2elembaut +  (1 | stateid), data = nelda2.sub)
+#summary(m1a)
+
+m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag +  unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)
+
+#summary(m1b)
+
+models <- list()
+models[['Model 16']] <- m1
+models[['Model 17']] <- m1b
+models[['Model 18']] <- m1c
+models[['Model 19']] <- m1a
+
+modelsummary(models, output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Random-intercept multilevel model of election-day fraud', coef_omit = "SD", coef_rename = 
+               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
+                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
+                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
+                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
+integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity",
+                  "state.owned.1lag" = "Private ownership (lag)",
+                 "patronage.appts.1lag" = "Professional bureaucracy (lag)",
+                 "v2elembaut" = "Election management body autonomy"
+                 )) %>% kable_styling(latex_options = "HOLD_position")
+
+##Note: state owned variable is inverse coded, so should be labeled as private ownership in economy
+##Note: Patronage appointments ialso inverse coded. Higher values = more professionalism
 ```
 
-The first model shows that the number of pre-election protests is weakly associated with improved election integrity, but just as in the main text this effect is extremely small.
-Each additional protest is associated with an improvement in election integrity equivalent to 0.2\% of the standard deviation of the dependent variable.
-Given that the mean number of pre-election protests is six, this is an extremely small influence on the overall degree of election integrity.
+Turning to models of protest initiation, the findings from the main text are upheld.
+Higher levels of fraud are associated with a higher risk of protest, but only when incumbent vote-shares are low.
+Table \@ref(tab:logit-table-edem) reports these results, and Figure \@ref(fig:margin-prob-fig-edem) shows the predicted probabilities.
 
-```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
-model.n.novoteshare.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    + v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +     + v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
-summary(model.n.novoteshare.hurd.mm)
+```{r include=FALSE}
+nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
+#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))
 
-model.n.base.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    winner.share.upd + v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    winner.share.upd + v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
-summary(model.n.base.hurd.mm)
+model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+summary(model.margin.binary)
 
-model.n.base2.hurd.mm <- hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre |  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, data = nelda_massmob %>% filter(gdpgro.1lag < 1), dist = "negbin")
-summary(model.n.base2.hurd.mm)
+model.margin.binary.int <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+summary(model.margin.binary.int)
 
-model.n.griev.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   gdpgro.1lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   gdpgro.1lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, dist = "negbin", data = nelda_massmob %>% filter(gdpgro.1lag < 1))
-summary(model.n.griev.hurd.mm)
+model.margin.binary_noc <- glm(any_protest ~ jucon.1lag + legcon.1lag +  altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+summary(model.margin.binary_noc)
 
-model.n.griev2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre | jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre , data = nelda_massmob, dist = "negbin")
-summary(model.n.griev2.hurd.mm)
+## Plots
 
-model.n.griev3.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   pubseccorrup.lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre  | jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   pubseccorrup.lag*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre, data = nelda_massmob, dist = "negbin")
-summary(model.n.griev3.hurd.mm)
-```
+####Margin
 
-Moving to the hurdle models of post-election protest, we do not observe much of a relationship between overall election integrity and protest.
-There is no statistically significant relationship between election integrity and protest occurrence in Model 2.
-Neither is there any significant relationship in interaction with the winner's vote-share, as shown in Figure \@ref(fig:b-protests-margin), or with economic crisis, as in Figure \@ref(fig:b-protests-crisis).
-The coefficient on the interaction term for public sector corruption and overall election integrity is statistically significant, though there is no significant effect at typical values for covariates and election integrity (as shown in Figure \@ref(fig:b-protests-corrupt)).
+#plot_binary_margin_max <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-2.3, 3.2]"))
+#plot_binary_margin_max <- plot_binary_margin_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()
 
-In the corresponding results for the number of post-election protests, there is similarly muted evidence for a connection between the overall level of manipulation and protest risk to incumbents.
-There is no significant difference between the marginal effect of common levels of election integrity on the number of post-election protests for typical cases across any margin of victory (Figure \@ref(fig:n-protests-margin)).
-During an economic crisis, *cleaner* elections are associated with an increase in the number of protests (\@ref(fig:n-protests-crisis)).
-This is plausibly due to the crisis weakening both the government's ability to generate manipulation [@greene_why_2007] and its ability to co-opt or put down the protests.
-Finally, public sector corruption is associated with a significant increase in the marginal effect of poor-integrity elections on protest\textemdash but only when the corruption measure is at its maximum (Figure \@ref(fig:n-protests-corrupt)).
-
-As a whole, the results are in accordance with the theory proposed in the main paper.
-There is no consistent evidence that lower-quality elections are associated with either a higher risk of protest initiation, or with a larger number of post-election protests.
-
-```{r zero-model-table, echo=FALSE}
-
-models <- list()
-models[['Model 2']] <- model.n.base.hurd.mm
-models[['Model 3']] <- model.n.base2.hurd.mm
-#models[['Model 4']] <- model.n.griev.hurd.mm
-models[['Model 4']] <- model.n.griev2.hurd.mm
-models[['Model 5']] <- model.n.griev3.hurd.mm
-
-modelsummary::modelsummary(models, coef_omit = "count", gof_omit = "R2", stars = T, longtable=T,  title = 'Binary portion of hurdle models of post-election protest')
+plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-.05, 2]"), show.data = T)
+plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()
 
@@
 
-```{r zero-model-table-count, echo=FALSE}
-modelsummary::modelsummary(models, coef_omit = "zero", gof_omit = "R2", longtable=T, stars = T, title = 'Hurdle models of post-election protest (counts)')
+```{r logit-table-edem, echo=FALSE}
+models2 <- list()
+models2[['Model 20']] <- model.margin.binary
+models2[['Model 21']] <- model.margin.binary.int
+models2[['Model 22']] <- model.margin.binary_noc
+
+modelsummary(models2, output = "kableExtra",  stars = T, title = 'Logit models of protest occurrence', coef_rename = 
+               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
+                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
+                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
+                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
+integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity",
+                 "v2elirreg.inv" = "Election fraud"
+                 )) %>% kable_styling(latex_options = "HOLD_position")
+```
+
+```{r margin-prob-fig-edem, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
+plot_binary_margin_typ_n
+```
+
+Lastly, the survival models also support the findings of the main text, as Table \@ref(tab:cox-models-tab-edem) and Figure \@ref(fig:margin-dur-fig-edem) show.
+Higher levels of election fraud are associated with shorter-lived protest movements when incumbent vote-shares are low, even in the expanded sample.
+
+```{r, include=FALSE}
+library(survival)
+library(ggsurvfit)
+nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
+nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
+                                                                       ifelse(d.events.num >= 69, 0, 1)))
+
+cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+summary(cox1)
+
+cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+summary(cox2)
+p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.73, 1.79]")) + coord_cartesian(ylim = c(0, 10)) #This is the 10th and 90th quantiles
+p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ration", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw() #These are extreme values: 15 and 85th percentiles
 
@@
 
-```{r, echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd", "v2xel_frefair [.2, .7]")) 
-p.n.margin <- plot(dat)
-p.n.margin <- p.n.margin  + theme_bw()
+```{r cox-models-tab-edem, echo=FALSE, warning=FALSE}
+models3 <- list()
+models3[['Model 23']] <- cox1
+models3[['Model 24']] <- cox2
 
-dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd", "v2xel_frefair [.2, .7]"), type = "zi_prob") 
-p.b.margin <- plot(dat)
-p.b.margin <- p.b.margin  + theme_bw()
+modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Cox proportional hazard models of protest duration', coef_rename = 
+               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
+                 "osorg.1lag" = "Civil soc. openness (lag)", 
+                 "winner.share.upd" = "Incumbent vote-share",
+                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
+                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
+integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "v2elirreg.inv" = "Election fraud",
+                 "n.events.post" = "N. post-election protests"
+                 ))
 
@@
 
-```{r n-protests-margin, fig.cap="Fraud, public sector corruption, and number of protest events"}
-p.n.margin
-
+```{r margin-dur-fig-edem, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
+p.duration.margin
 ```
 
-```{r b-protests-margin, fig.cap="Fraud, public sector corruption, and likelihood of protest"}
-p.b.margin
+\newpage
 
-```
-
-```{r, echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.griev2.hurd.mm, terms = c("econ.crisis.nelda.1lag", "v2xel_frefair [.15, .66]")) 
-p.n.crisis <- plot(dat)
-p.n.crisis <- p.n.crisis  + theme_bw()
-
-dat <- ggpredict(model.n.griev2.hurd.mm, terms = c("econ.crisis.nelda.1lag", "v2xel_frefair [.15, .66]"), type = "zi_prob") 
-p.b.crisis <- plot(dat)
-p.b.crisis <- p.b.crisis +  theme_bw()
-
-```
-```{r n-protests-crisis, fig.cap="Fraud, economic crisis, and number of protest events"}
-p.n.crisis
-```
-
-```{r b-protests-crisis, fig.cap="Fraud, economic crisis, and likelihood of protest"}
-p.b.crisis
-```
-
-```{r, echo=FALSE, message=FALSE}
-dat <- ggpredict(model.n.griev3.hurd.mm, terms = c("pubseccorrup.lag", "v2xel_frefair [.29, .66]")) 
-p.corr.n.margin <- plot(dat)
-p.corr.n.margin <- p.corr.n.margin + coord_cartesian(ylim = c(0, 100)) + theme_bw()
-
-dat <- ggpredict(model.n.griev3.hurd.mm, terms = c("pubseccorrup.lag", "v2xel_frefair [.29, .66]"), type = "zi_prob") 
-p.corr.b.margin <- plot(dat)
-p.corr.b.margin <- p.corr.b.margin  + theme_bw()
-
-```
-
-```{r n-protests-corrupt, fig.cap="Fraud, public sector corruption, and number of protest events"}
-p.corr.n.margin
-```
-
-```{r b-protests-corrupt, fig.cap="Fraud, public sector corruption, and likelihood of protest"}
-p.corr.b.margin
-```
+# References

### 2024-02-09 Scripts/authoritarian_only.Rmd
Changes to appendices and fixes in main text

@@
 
-In this paper, I argue that\textemdash in keeping with the idea of 'informational autocracy' (ADD CITES HERE)\textemdash an election outcome produces a signal of the relative strength of the incumbent.
-As Magaloni (2008) and others have noted, incumbents typically work to produce a higher margin of victory in the election, in order to demonstrate their strength and induce cooperation from other actors in society.
+In this paper, I argue \textemdash in keeping with a growing literature on the informational aspects of elections in authoritarian regimes [@guriev2020;@smyth2020elections;@Geddes1989;@Miller2015a]\textemdash that an election outcome produces a signal of the relative strength of the incumbent.
+As Magaloni [-@magaloni2006] nd others have noted, incumbents typically work to produce a higher margin of victory in the election, in order to demonstrate their strength and induce cooperation from other actors in society.
 In some theories, this only works if the votes for the incumbent are largely 'genuine' (i.e.
 not fabricated), or produced using the incumbent's organizational resources (i.e.
-vote-buying and intimidation).
-(ADD CITES HERE, INCLUDING TO HARVEY AND M).
-These two kinds of votes correspond to two of the three 'pillars of stability' identified by Gerschewski (2013) (ADD CITE): legitimation and cooptation.
+vote-buying and intimidation) [@Harvey2020;@magaloni2006].
+These two kinds of votes correspond to two of the three 'pillars of stability' identified by Gerschewski [-@Gerschewski2013]: legitimation and cooptation.
 Fraud is considered a form of cheap talk in these models, which opposition actors can discount if they become aware of it.
 
-However, these models overlook that engineering widespread fraud requires control over patronage resources and the criminal and judicial apparatus (ADD CITES to HARVEY).
+However, these models overlook that engineering widespread fraud requires control over patronage resources and the criminal and judicial apparatus [@Harvey2019;@Harvey2022].
 As a result, they overlook the fact that high levels of fraud are a sign of incumbent strength in both cooptation and Gerschewski's third pillar: repression.
@@
 
-Combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019],... [MOVE]
-
 Given these results, it is likely that the downside risk for incumbents that engage in election fraud in non-democracies has likely been overemphasized.
@@
 
-[REWRITE LATER] One troubling implication of these results is that ruling parties should follow an 'all-in' strategy on election fraud, a process that entails consolidating control over patronage resources and further eroding the rule of law.
+One troubling implication of these results is that\textemdash if protest is indeed costly to incumbents\textemdash ruling parties should follow an 'all-in' strategy on election fraud, a process that entails consolidating control over patronage resources and further eroding the rule of law.
 More fraud is not associated with longer-lasting protest waves, but a large margin of victory is associated with a lower risk of protest initiation.
-These results help explain both the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory.
-They also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation.
+These results help explain several empirical patterns at once: the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory in conjunction with high levels of fraud.
+At a theoretical elvel, they also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation.
+Finally, they suggest a pessimistic outlook for democratization by election in authoritarian regimes.
 
@@
 -->
-
-[MOVE TO DISCUSSION AND REWRITE] Generally speaking, protest-based models predict that manipulation should decline as opposition groups become stronger relative to incumbents.
-This dynamic creates a sort of virtuous cycle, wherein strong opposition groups' credible threat of protest makes elections fairer, which in turn improves the strength of the opposition--leading ultimately to the ideal of self-enforcing democracy.
-The alternative school of thought, among other predictions, argues that opposition strength drives front-line agent behavior more than that of leaders, and causes a shift to less dispersed, less observable kinds of manipulation [@asunka2017electoral;@Harvey2019].
-Election integrity can thus remain low, even in competitive contexts.
-Putting the protest-deterrent logic to the test, then, helps resolve significant questions in the study of election manipulation and democratization.
-
-  Of course, electoral manipulation is not understood to be the sole cause of protest\textemdash protests occur for all sorts of reasons\textemdash but the literature has drawn a close connection between the risk of specifically electoral protest and the conduct of elections.
 
@@
 
-plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-.05, 2]"), show.data = T)
+plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-.05, 2]"), show.data = T)
 plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()
@@
 In particular, the main results are supported when data from the Mass Mobilization dataset are used in place of the ECAV data (appendix page 1), and when a measure of overall election integrity is used in place of the measure of election fraud (appendix page 15).
+[CHECK PAGES]
+
+[MOVE TO DISCUSSION AND REWRITE] Generally speaking, protest-based models predict that manipulation should decline as opposition groups become stronger relative to incumbents.
+This dynamic creates a sort of virtuous cycle, wherein strong opposition groups' credible threat of protest makes elections fairer, which in turn improves the strength of the opposition--leading ultimately to the ideal of self-enforcing democracy.
+The alternative school of thought, among other predictions, argues that opposition strength drives front-line agent behavior more than that of leaders, and causes a shift to less dispersed, less observable kinds of manipulation [@asunka2017electoral;@Harvey2019].
+Election integrity can thus remain low, even in competitive contexts.
 
@@
 
-```{r, include=FALSE}
-ggplot(data=nelda2.sub, aes(x=v2elirreg.inv, y = jitter(any.ecav.protest), color = as.factor(v2x_regime))) +
-  geom_point()
-
-```
-
-Predicting post-election protest counts and getting residuals:
-
-```{r, include=FALSE}
-test3 <- glm.nb(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag +  + osorg.1lag  + natlpartyorg.lag +   inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag, data = nelda2.sub)
-summary(test3)
-
-ggplot(mapping = aes(x=predict(test3, type = "response"), y= test3$model$n.events.post)) + geom_point() + coord_cartesian(xlim=c(0, 500))
-
-newdata <- modelr::add_predictions(nelda2.sub, test3)
-newdata <- newdata %>% mutate(pred_error = n.events.post - pred)
-  #because this is set up so that the predicted value is subtracted from the actual, underpredicting
-  # protest is a positrive value (i.e. actual 5 - 0 predicted = 5)
-
-test4 <- lm(pred_error ~ winner.share.upd + v2elirreg.inv + presidential, data = newdata)
-summary(test4)
-
-test5 <- lm(pred_error ~ winner.share.upd*v2elirreg.inv + presidential, data = newdata)
-summary(test5)
-
-sjPlot::plot_model(test5, type = "pred", terms = c("v2elirreg.inv", "winner.share.upd [.4, .7]"))
-
-plot(newdata$v2elirreg.inv, newdata$pred_error)
-```
-
-Predicting the level of election manipulation, and then seeing if actual values above prediction are associated with protest
-
-```{r, include=FALSE}
-nelda2.sub <- nelda2.sub %>% mutate(inc.confident.num = ifelse(inc.confident == "N/A", NA, ifelse(inc.confident == "yes", 1, 0)))
-
-###For testing missingness
-subsample <- nelda2.sub %>% dplyr::select(v2elirreg.inv, jucon.1lag, legcon.1lag, gdpgro.1lag,  natlpartyorg.lag,  presidential, urban.pct, rentierstate,  taxstate,  unscheduled_election, regional_success_lag, physinteg.1lag, antisys.1lag, csenv.1lag, preelection_concerns, inc.termlimit.num, liberal_no_jucon, v2elirreg.inv.lag, patronage.appts.1lag,
-                                                  state.owned.1lag)
-summary(subsample)
-############################ Note: Drop legcon from analysis because of high missingness and high cor with lib_no_jucon
-
-test6 <- lmer(v2elirreg.inv ~ jucon.1lag +  gdpgro.1lag +  natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + physinteg.1lag + antisys.1lag + csenv.1lag + preelection_concerns + inc.termlimit.num   + liberal_no_jucon + v2elirreg.inv.lag +
-                patronage.appts.1lag +  state.owned.1lag +  v2elembaut + (1 | stateid), data = nelda2.sub) 
-summary(test6)
-
-testdata <- broom.mixed::augment(test6)
-  #This adds fitted values and residuals.
-  #Residuals are actual values minus predicted value, so
-  #over-predicting fraud will result in a negative number (actual fraud < predicted )
-  #Under-predicting fraud--what we are interested in--will result in a positive number (actual > predicted)
-
-## Need to add rownames to nelda2.sub, so that can merge variables from nelda2.sub into testdata
-nelda2.sub <- nelda2.sub %>% mutate(.rownames = rownames(nelda2.sub))
-neldavars <- nelda2.sub %>% dplyr::select(.rownames, n.events.post, inc.lose, winner.share.upd, country, year, diem_explicheat_max, diem_fraud_max, country_id_year, surv_status,  inc.replaced.prot, d.events.num,
-                                          legcon.1lag, osorg.1lag, altinf.1lag, libdem.1lag,  winner.share.upd, n.events.pre)
-testdata <- testdata %>% left_join(neldavars, by = ".rownames")
-
-model_protest <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd + .resid  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + .resid*winner.share.upd,
-                               ziformula=~.,
-                              data = testdata %>% filter(inc.lose == 0),
-                                family=truncated_nbinom2)
-summary(model_protest)
-
-model_protest2 <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd + .resid  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
-                               ziformula=~.,
-                              data = testdata %>% filter(inc.lose == 0),
-                                family=truncated_nbinom2)
-summary(model_protest2)
-
-plot(testdata$winner.share.upd, testdata$.resid)
-
-testdata <- testdata %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))
-testdata <- as.data.frame(testdata)
-model_protest_binary <- glm(no_protest ~ .resid*winner.share.upd, family = binomial(link = "logit"),
-                             data = testdata %>% filter(inc.lose == 0))
-plot_binary_margin_typ_n <- sjPlot::plot_model(model_protest_binary, type = "pred", terms = c(".resid", "winner.share.upd [.2, .8]"))
-plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Excess election fraud", y = "Predicted probability", title = "Predicted probability of no protest", color = "Incumbent vote-share") + theme_bw()
-
-model_protest2_binary <- glm(no_protest ~ .resid + winner.share.upd, family = binomial(link = "logit"),
-                             data = testdata %>% filter(inc.lose == 0))
-
-sjPlot::plot_model(model_protest2_binary, type = "pred", terms = c(".resid", "winner.share.upd [.2, .8]"), show.data = T)
-sjPlot::plot_model(model_protest2_binary, type = "est", vline.color = "black") + theme_bw()
-
-### Seeing how residuals compare to DIEM data
-
-summary(lm(.resid ~ diem_explicheat_max + diem_fraud_max, data = testdata))
-plot(testdata$diem_explicheat_max, testdata$.resid)
-plot(testdata$diem_fraud_max, testdata$.resid)
-  ##This shows that the resids are basically noise with regard to election observer reports, which should be good; I think this helps push back against the idea that V-Dem coders are giving higher scores to protest elections.
-But it would be good to push deeper and look at the correlation only among the most up-coded resids.
-
-high_resids <- testdata %>% filter(.resid >= .222)
-summary(lm(.resid ~ diem_explicheat_max + diem_fraud_max, data = high_resids))
-
-ggplot(data = testdata, aes(x = jitter(diem_fraud_max), y = .resid)) + geom_point(aes(size = n.events.post))
-
-plot(high_resids$diem_explicheat_max, high_resids$.resid)
-plot(high_resids$diem_fraud_max, high_resids$.resid)
-  #We do see more outliers here, good to think deeper--there are two major outliers that are also big protest events
-
-### Duration
-
-cox1_b <- coxph(Surv(d.events.num, surv_status) ~ winner.share.upd  + .resid, data = testdata %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
-summary(cox1_b)
-
-## DIEM and protest
-
-model_protest_diem <- glmmTMB(n.events.post ~   winner.share.upd +  diem_explicheat_max ,
-                               ziformula=~.,
-                              data = nelda2.sub %>% filter(inc.lose == 0),
-                                family=truncated_nbinom2)
-summary(model_protest_diem)
-
-```
-
-```{r, include=FALSE}
-nelda2.sub <- nelda2.sub %>% group_by(country) %>% mutate(elirreg_mean = mean(v2elirreg.inv, na.rm=T))
-nelda2.sub <- nelda2.sub %>% mutate(elirreg_mean_diff = v2elirreg.inv - elirreg_mean)
-
-model_protest1 <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd*elirreg_mean_diff  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
-                               ziformula=~.,
-                              data =nelda2.sub %>% filter(inc.lose == 0),
-                                family=truncated_nbinom2)
-summary(model_protest1)
-sjPlot::plot_model(model_protest1, type ="pred", terms = c("elirreg_mean_diff" , "winner.share.upd [.2, .8]")) + coord_cartesian(ylim = c(0, 50))
-
-nelda2.sub <- nelda2.sub %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))
-model_protest_binary <- glm(no_protest ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd*elirreg_mean_diff  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
-                             data =nelda2.sub %>% filter(inc.lose == 0),
-                                family=binomial(link = "logit"))
-summary(model_protest_binary)
-sjPlot::plot_model(model_protest_binary, type="pred", terms = c("elirreg_mean_diff" , "winner.share.upd [.2, .8]"))
-
-model_protest2 <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd + elirreg_mean_diff  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
-                               ziformula=~.,
-                              data =nelda2.sub %>% filter(inc.lose == 0),
-                                family=truncated_nbinom2)
-summary(model_protest2)
-
-```
-
-Ruling party turnover
-
-```{r, include=FALSE}
-model.margin.turnover <- glm(inc.lose ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +   + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + state.owned.1lag + patronage.appts.1lag, family = binomial(link = "logit"), data = nelda2.sub)
-summary(model.margin.turnover)
-
-m.inclose <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag*patronage.appts.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub %>% filter(inc.lose == 1))
-summary(m.inclose)
-sjPlot::plot_model(m.inclose, type = "int")
-
-m.incwin <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag*patronage.appts.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub %>% filter(inc.lose == 0))
-summary(m.incwin)
-sjPlot::plot_model(m.incwin, type = "int")
-
-### Z score
-
-z <- (-1.647988 - (-1.144192)) / sqrt((0.691932)^2 + (0.329801)^2) #B1 - B2 / sqrt((seB1)^2 + (seB2)^2)
-
-```
-
 \newpage

### 2024-07-15 Scripts/authoritarian_only_latex.Rmd
Beginning post-BJPS updates

--- 2024-02-09 Scripts/authoritarian_only.Rmd
@@
 output:
-    bookdown::pdf_document2: default
+    pdf_document:
+      keep_tex: yes
 linkcolor: blue
@@
 indent: true
+abstract: "The risk of mass protest figures prominently in many models of election manipulation, as a deterrent to tampering with elections.
+Yet election manipulation is common outside liberal democracies, while protest is rare (and successful protest rarer still).
+Using data on election fraud and duration data for protest events from 647 elections from 1990 to 2012, this paper argues that the risk to incumbents has been overstated in prior work.
+It argues that incumbents who solve the principal-agent dilemmas associated with election fraud are likely to be highly resilient to protest.
+There are two main findings.
+First, a riskier pre-election environment for incumbents is associated with more fraud, not less.
+Second, when protests do occur, greater election-day fraud is associated with a faster demise for protest movements.
+This pessimistic finding implies that election fraud is less risky than much prior work assumes, and helps explain the long-term stability of electoral authoritarianism as a regime type."
 header-includes: 
@@
 
-On August 9, 2020, Belarus held an election in which President Aleksander Lukashenko ran for a sixth term.
-The official results credited Lukashenko with over 80\% of the vote, a result which observers regarded as wildly inflated [@GolosPlatformPresents2020].
-In response, hundreds of thousands of Belarusians took to the streets over the course of several months.
-Though the protests put significant strain on the regime, a harsh campaign of repression (and support from neighboring Russia) ultimately quelled the protests; Lukashenko was inaugurated into a new five-year term.
+The December 2011 elections to the Russian State Duma delivered a significant setback for the ruling party.
+It claimed only forty-nine percent of the vote, and while it retained the majority in the legislature, it lost more than 75 seats from the previous convocation.
+This narrow 'victory' occurred despite widespread election-day manipulation, detected in many cases by domestic election observers.
+The election triggered one of the most substantial protest movements in post-Soviet Russian history.
+More than 100,000 people turned out in protest on two occasions in Moscow, with tens of thousands more taking to the streets in regional cities [@Dollbaum2017].
+Protests continued regularly in advance of the March 2012 presidential election, declining afterward.
 
-The Belarusian case is typical in some ways, and an outlier in others.
-Like many previous elections in Belarus, and in authoritarian regimes around the world, the contest was heavily managed and the results inflated by fraud.
-However, the sustained, massive protest wave that occurred post-election is extremely unusual.
-However, the regime's ultimate victory over the protest movement\textemdash while costly and never guaranteed\textemdash is also the typical outcome of post-election protest.
+Despite the significant challenge, the ruling party was able to defuse the protest movement over a period of months.
+Through a combination of repression and concessions [@golosov2012], the ruling party was able to reassert itself and continue a long process of sidelining the political opposition in Russia [@gelman_2015_troubled].
+The post-2011 protest wave is one much-studied case of the phenomenon of electoral protest aimed at punishing electoral manipulation.
+Yet the Russian example raises questions for that broader phenomenon: does election fraud itself make protest more likely, or only in the context of a weak electoral showing for the ruling party?
+Given the ruling party's poor result and substantial electoral fraud, why was it able to dispatch the mobilized opposition with relative ease?
 
-These patterns highlight a puzzle generated by existing research on the relationship between election manipulation and popular protest.
-The topic has significant implications for authoritarian governance and democratization, and as such has been a very active subject of study.
-However, the various theoretical perspectives brought to bear on the topic have been built on different assumptions, and have different theoretical implications.
-To briefly summarize: for some models, election fraud is a tool of weak incumbents.
-As a result, incumbents avoid fraud when it might be exposed and reveal their weakness to potential protesters\textemdash an auspicious prediction for democratization and ultimately the under-pinning of 'self-enforcing' democracy [@Fearon2011].
-For others, fraud is committed by weak and strong incumbents alike, making protest a risky gamble the opposition undertakes for exogenous reasons.
-In these models, fraud may or may not be deterred based on other factors: a dubious, though not completely dismissive, prospect for democratization and self-enforcing democracy.
-Lastly, a third school of thought holds that election fraud is an indicator of genuine incumbent strength.
+These questions get to the heart of a debate in the study of electoral manipulation and protest: does election fraud signify regime weakness or strength, and in what ways?
+The various theoretical perspectives brought to bear on the topic have been built on different assumptions, and yield different implications for authoritarian governance and democratization.
+For some models, election fraud is a tool of weak incumbents.
+As a result, incumbents avoid fraud when it might be exposed and reveal their weakness to potential protesters\textemdash an auspicious prediction for democratization and ultimately the under-pinning of 'self-enforcing' democracy [@fearon_self-enforcing_2011].
+Another school of thought holds that election fraud can be an indicator of incumbent strength.
 This approach, which is expanded upon below, is a pessimistic account for democracy.
-If this model is correct, it is authoritarian politics rather than democracy that is self-enforcing\textemdash the ability to commit election fraud helps win elections, signal strength, deter challengers under most conditions, and put down protest movements if they do begin.
-This paper argues that the third model best fits the overall pattern of the data, and is supported by specific hypothesis tests.
+If this model is correct, it is authoritarian politics rather than democracy that is self-enforcing\textemdash the ability to commit large-scale election fraud helps win elections, signal strength, deter challengers under most conditions, and put down protest movements if they do begin.
+This paper argues that this pessimistic model best fits the overall pattern of the data, and is supported by specific, novel hypothesis tests.
 
-In this paper, I argue \textemdash in keeping with a growing literature on the informational aspects of elections in authoritarian regimes [@guriev2020;@smyth2020elections;@Geddes1989;@Miller2015a]\textemdash that an election outcome produces a signal of the relative strength of the incumbent.
-As Magaloni [-@magaloni2006] nd others have noted, incumbents typically work to produce a higher margin of victory in the election, in order to demonstrate their strength and induce cooperation from other actors in society.
-In some theories, this only works if the votes for the incumbent are largely 'genuine' (i.e.
+In this paper, I argue \textemdash in keeping with a growing literature on the informational aspects of elections in authoritarian regimes [@guriev2020;@smyth2020elections;@Geddes1989;@Miller2015a;@Pop-Eleches2015]\textemdash that an election outcome produces a signal of the relative strength of the incumbent.
+As Magaloni [-@magaloni2006] and others have noted, incumbents typically work to produce a higher margin of victory in the election, in order to demonstrate their strength and induce cooperation from other actors in society.
+In some theories, this works best if the votes for the incumbent are largely 'genuine' (i.e.
 not fabricated), or produced using the incumbent's organizational resources (i.e.
@@
 These two kinds of votes correspond to two of the three 'pillars of stability' identified by Gerschewski [-@Gerschewski2013]: legitimation and cooptation.
-Fraud is considered a form of cheap talk in these models, which opposition actors can discount if they become aware of it.
+Fraud is considered a form of cheap talk in these models, which opposition actors can discount if they become aware of it [@little_elections_2012].
 
-However, these models overlook that engineering widespread fraud requires control over patronage resources and the criminal and judicial apparatus [@Harvey2019;@Harvey2022].
-As a result, they overlook the fact that high levels of fraud are a sign of incumbent strength in both cooptation and Gerschewski's third pillar: repression.
+However, these models overlook the fact that engineering widespread fraud requires solving principal-agent problems through consolidated control over patronage resources and the criminal and judicial apparatus [@Harvey2019;@Harvey2022].
+As a result, high levels of fraud are a sign of incumbent strength in both cooptation and repression (Gerschewski's third pillar).
 Both of these tools can be useful for defusing or dismantling an emergent protest movement.
-As a result, effective fraudsters should also be effective dismantlers of protest.
-This helps explain the broad pattern, where most fraudulent elections go unprotested, and among those that do generate protest, the vast majority fail to dislodge the incumbent.
+As a result, effective fraudsters should also be more effective at defeating protest movements.
 Consequently, in this view, there is little penalty for severe manipulation in terms of protest risk, and little deterrent effect.
@@
 Testing this proposition in more depth is difficult.
-If, for example, in equilibrium incumbents manipulate just the right amount to avoid mass protest\textemdash and are deterred from manipulating more by that risk\textemdash we would observe no statistical relationship between the severity of fraud and the likelihood of protest initiation.
+If, for example, incumbents manipulate just the right amount to avoid mass protest in equilibrium\textemdash and are deterred from manipulating more by that risk\textemdash we would observe no statistical relationship between the severity of fraud and the likelihood of protest initiation.
 High and low levels of fraud would generally be associated with no protest.
@@
 Using data from 614 non-democratic elections across 97 countries, from 1989 to 2012, this paper finds that the answer to both questions is no.
-In fact, more severe election fraud is associated with a *faster* demise for protest waves.
+In fact, greater pre-election risk is associated with worse fraud, and more severe election fraud is associated with a *faster* demise for protest waves.
 This indicates that, far from being a sign of weakness, in most cases the ability to generate significant electoral fraud is an indicator of durable regime strength.
@@
 
-Given these results, it is likely that the downside risk for incumbents that engage in election fraud in non-democracies has likely been overemphasized.
-Even in the unlikely event that protests do break out in response to large-scale fraud, the government has many tools it may deploy to curtail them.
-Fraud itself is a sign of the robustness of some of those tools\textemdash the control over resources that incentivizes fraud can be used to coopt the opposition or constrain elites, while control over the judicial and prosecutorial systems can be used to intensify repression during a crisis.
+Given these results, it is likely that the downside risk for incumbents that engage in election fraud in non-democracies has been overemphasized.
+Fraud itself is a sign of the robustness of some of the same tools that can be used to co-opt the opposition, constrain elites, or intensify repression during a crisis; far from being a sign of weakness, it can be a sign of resilience.
+This can deter protest, especially in conjunction with high claimed vote-shares for the incumbent.
+Even in the unlikely event that protests do break out in response to large-scale fraud, a government that has engaged in widespread fraud has many tools it may deploy to curtail them.
 
-One troubling implication of these results is that\textemdash if protest is indeed costly to incumbents\textemdash ruling parties should follow an 'all-in' strategy on election fraud, a process that entails consolidating control over patronage resources and further eroding the rule of law.
-More fraud is not associated with longer-lasting protest waves, but a large margin of victory is associated with a lower risk of protest initiation.
 These results help explain several empirical patterns at once: the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory in conjunction with high levels of fraud.
-At a theoretical elvel, they also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation.
+At a theoretical level, they also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation.
 Finally, they suggest a pessimistic outlook for democratization by election in authoritarian regimes.
@@
 Prior research has often considered that electoral manipulation has both costs and benefits for ruling parties.
-The benefits include better odds of winning the official tally [@lehoucq2002], of course, as well as any signals of strength and stability that a larger vote-share [@magaloni2006] or display of power that manipulation might provide [@simpser2013;@gehlbach2015electoral].
+The benefits include better odds of winning the official tally [@lehoucq2002], of course, as well as any signals of strength and stability that a larger vote-share [@magaloni2006] or display of power might provide [@simpser2013;@gehlbach2015electoral].
 On the other side, election manipulation is generally characterized as having both operational costs (which the incumbent may be unwilling or unable to pay [@Greene2010]) and legitimacy costs which reduce other actors' level of cooperation with the incumbent [@birch2011].
@@
 
-The possibility of post-election protest regularly figures as a cost that incumbents must weigh when determining whether (and how much) to inflate their vote share.
-In these approaches, the cost of protest be conceptualized as the risk that the incumbent may be unseated and/or the resource cost needed to repress or buy off the incumbent's opponents.
+The cost of protest is conceptualized as the risk that the incumbent may be unseated and/or the resource cost needed to repress or buy off the incumbent's opponents.
 Various theories model the decision to engage in post-election protest either as a strategic one by opposition groups [@magaloni2010;@Luo2020;@przeworski2015], or more spontaneous coordination by citizens expressing a shared grievance [@tucker2007;@Brancati2016].
 In either case, the risk of protest is a classic component of models of electoral manipulation, as a restraint on incumbents who would prefer to use fraud to increase their vote-shares, all else held equal.
-This argument can also be found in theories about the conduct of non-democratic elections, the focus here, as well as in the literature on self-enforcing democracy [@Fearon2011;@przeworskiSelfenforcingDemocracy2006].
+This argument can be found in theories about the conduct of non-democratic elections, as well as in the literature on self-enforcing democracy [@fearon_self-enforcing_2011;@przeworskiSelfenforcingDemocracy2006].
 In the latter literature, generally free and fair elections are thought to be a self-enforcing equilibrium through the threat of mass protest by wronged groups if the incumbent government rigs the results.
-While the focus here is on elections in authoritarian cases, the implications for self-enforcing democracy are discussed at the close of the paper.
+While the focus here is on elections in authoritarian cases, some implications for self-enforcing democracy are discussed at the close of the paper.
 
@@
 Similar results are developed by Chernykh and Svolik [-@chernykh2015third], Luo and Rozenas [-@Luo2020] and [-@Luo2017], and Magaloni [-@magaloni2010].
-These are normatively optimistic models: fraud can be kept in check, and ultimately reduced, through the development of election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].
+These are normatively optimistic models upon which a sizable literature has been founded: fraud can be kept in check, and ultimately reduced, through the development institutions that reveal information about election integrity: election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].
 
-A second class of formal models holds that fraud operates as a distortion of the signal of the incumbent's true strength.
-In this type of model, such as those developed by Little [-@littleFraudMonitoringNoncompetitive2015] or Egorov and Sonin [-@Egorov2021], fraud may be used by incumbents to bluff their opponents into being more supportive than they would be if they knew the incumbent's true popularity.
-However, greater transparency makes such bluffs more costly.
-The typical result in these models is that protest risk can *partially* deter fraud.
-In equilibrium, the incumbent will generally choose the level of fraud anticipated by the opposition and no more\textemdash but also no less.
-High levels of fraud can be committed by strong or weak incumbents, and as such these models do not predict a relationship between the risk of protest and the level of fraud per se.
-As Egorov and Sonin [-@egorov2014incumbency] write when discussing their model of this type, "We see that even dictators who are sure to win may use fraud; the reason is that if they do not, citizens will believe that they are less popular and used fraud anyway [thus inviting protest]." [ADD PAGE] In these models, protest initiation is most likely when fraud is higher than expected and when the incumbent's vote-share is low (see, e.g., Little [-@little_elections_2012].
-
-Similar predictions emerge from another school of thought, which holds that electoral protest is driven by citizens' grievances rather than by a strategic, organized opposition [@Brancati2016].
+Another school of thought holds that electoral protest is driven by citizens' grievances rather than by a strategic, organized opposition [@Brancati2016].
 In this view, election manipulation can act as a shared grievance among individuals who prefer to see the ruling party lose\textemdash especially when it appears that fraud may have been decisive [@kuntz_more_2009].
 Belief in a stolen election is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007], possibly initiating a protest cascade [@Kuran1991].
-Just as above, these models tend to predict greater protest risk when margins of victory are low and/or grievances are high in the presence of fraud.
+Just as above, these models tend to predict greater protest risk when margins of victory are low, suggesting a stolen election, and/or grievances are high in the presence of fraud.
 
@@
 
-The strategic and social-psychological explanations are, of course, interconnected.
-As Schedler writes, "Grievances provide moral justifications for collective protest: we need to do something.
-Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304].
+This implication connects to a second class of formal models, which hold that fraud operates as a distortion of the signal of the incumbent's true strength.
+In this type of model, such as those developed by Little [-@littleFraudMonitoringNoncompetitive2015] or Egorov and Sonin [-@Egorov2021], fraud may be used by incumbents to bluff their opponents into being more supportive than they would be if they knew the incumbent's true popularity.
+However, greater transparency makes such bluffs more costly.
+The typical result in these models is that protest risk can *partially* deter fraud.
+In equilibrium, the incumbent will generally choose the level of fraud anticipated by the opposition and no more\textemdash but also no less.
 
@@
 However, Hafner-Burton et al [-@Hafner-Burton2016] find that the positive association between election fraud and post-election protest falls short of conventional statistical significance.
-Likewise, looking at the 2011 election and protest cycle in Russia, Ananyev and Poyker [-@ananyevDictatorsSignalStrength2022] show that post-election protest was not associated with the severity of fraud using several measures^[Lankina and Skovoroda [-@Lankina2017] find that post-election protest was positively associated with a different measure of election fraud in that case, however.] Taking an alternative approach, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e.
+Studies of one archetypal example, the 2011 election and protest cycle in Russia, have found a positive relationship between fraud and protest by some measures [@Lankina2017], and no relationship by others [@ananyevDictatorsSignalStrength2022].
+Taking an alternative approach, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e.
 whether opponents are allowed to run), rather than by the election-day manipulation that usually figures in theoretical work.
@@
 While these works have identified a connection between fraud and protest, in general, they cannot distinguish between the competing assumptions of the two theories discussed here, or the third possible approach discussed in below.
-All of the relevant theories predict that protest should be most likely when evidence of fraud is high and the incumbent's vote-share is low (or, relatedly, grievances are high).
-However, the implications of the theories are substantively distinct.
-Protest risk only acts as a deterrent to fraud in theories where fraud is a signal of regime weakness.
-So, it is necessary but insufficient to ask whether the occurrence of post-election protests is associated with higher fraud.
-It is also necessary to test the assumptioins underlying the different models.
+For instance, all three approaches predict that protest should be most likely when evidence of fraud is high and the incumbent's vote-share is low (or, relatedly, grievances are high).
+Such a scenario could occur due to 1) strategic protest against an incumbent revealed to be weak through reliance on fraud, 2) mass-led protest against a 'stolen' election, or 3) risky strategic protest against a low-popularity incumbent despite institutional reservoirs of strength.
+The implications of all three scenarios are substantively distinct, however.
+To untangle them, it is necessary to test the assumptions underlying the different models.
 In particular, do post-election protest movements encounter weak incumbents as signaling models predict?
@@
 
-As discussed above, one view of election fraud and protest views fraud as a tool of weak incumbents, while another views it as a tool that is employed by weak and strong alike.
-A third view, associated with Simpser [-@simpser2013], holds that election fraud demonstrates incumbent strength.
-In Simpser's original account, we should expect electoral manipulation to indicate strength when it is blatant\textemdash i.e.
-readily observable\textemdash or excessive, producing large margins of victory.
+As discussed above, one view of election fraud and protest views fraud as a tool of weak incumbents.
+Another view, associated with Simpser [-@simpser2013], holds that election fraud demonstrates incumbent strength.
+In Simpser's original account, we should expect electoral manipulation to signal strength when it is readily observable or excessive, producing large margins of victory.
 A strong government may manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver].
@@
 
-[CONSIDER Ananyev and Poyker, Trinh 2022--do they fit in?]
-
-I build on Simpser's logic, and subsequent research, by arguing that that extensive election fraud itself requires the use of resources that are sources of incumbent strength, whether or not these are observed.
-Simpser's model is concerned with whether or or not other actors perceive the regime's signal of strength, leading to the focus on excessive or blatant manipulation.
+I build on Simpser's logic, and on subsequent research, by arguing that that extensive election fraud itself requires the use of resources that are sources of incumbent strength.
+Simpser's model is concerned with whether or or not other actors can perceive the regime's signal of strength; hence the focus on excessive or blatant manipulation.
 I argue here that even when large-scale election fraud is marginal (i.e.
@@
 
-control over the bulk of patronage resources in society , and the ability to shield agents from criminal and legal risks [@Harvey2019;].
-Each of these resources can be used to minimize protest waves if they do occur after the election.
-
-To the extent that election results in non-democracies send a signal, it is composed of a mix of three factors: the incumbent's genuine popularity, its ability to mobilize otherwise reluctant voters (through patronage, clientelism, and coercion), and its ability to control the bureaucratic apparatus (including the criminal and judicial systems) to generate fraud.
+To the extent that election results in non-democracies send a signal, it is composed of a mix of three factors [@Pop-Eleches2015]: the incumbent's genuine popularity, its ability to mobilize otherwise reluctant voters (through patronage, clientelism, and coercion), and its ability to control the bureaucratic apparatus (including the criminal and judicial systems) to generate fraud.
 Large-scale election fraud requires that the incumbent have consolidated control over patronage resources in society; this makes cooperation with the incumbent attractive for front-line agents, who are less willing to risk losing access to the incumbent's network [@Harvey2019].
@@
 Each of these capacities is required to commit large-scale fraud, and both are useful for disarming protest movements.
+To summarize, the more extensive the election-fraud effort, the more successfully the incumbent has solved the associated principal-agent problems, and the more likely it will be able to use those tools to quash protest.
 
@@
 
+In addition, as Gehlbach and Simpser [-@gehlbach2015electoral] argue, effective electoral manipulation can induce regime agents to exert effort on behalf of the incumbent.
+Elite defections, which can presage real challenges for the incumbent during a crisis, are also rarer in less democratic regions, where election manipulation is likely more severe [@Reuter2019].
+Just as election manipulation has been held to solve a coordination problem for regime opponents, it may also solve a coordination problem for regime agents: when successful, a fraud campaign signals to agents that other agents are likely to stick with the incumbent due to its revealed capacities.
+
 This argument\textemdash that successful, large-scale fraud draws on the same resources a state may use to suppress protest\textemdash turns the deterrent logic on its head.
 Instead of signalling weakness, large-scale fraud indicates a well of resilience.
-Consequently, we should not expect incumbent regimes capable of committing mass fraud to be deterred from doing so by the risk of protest.
-This argument clearly echoes the seminal argument of Alberto Simpser [@simpser2013].
-It is distinct, however; for Simpser, election manipulation only communicates strength to other actors when it is excessive (generating high margins of victory) or blatant.
-As Simpser notes, "...the marginal versus excessive dimension of electoral manipulation does not fully overlap with the question of scale or extent: while excessive electoral manipulation is generally associated with large-scale manipulation, the marginal kind can result either from a low amount of manipulation (e.g., in a tight race) or from a large amount [due to an unpopular incumbent]" (p.
-7).
+Consequently, we should expect incumbent regimes to be aggressive in their use of election fraud, especially in periods of uncertainty, in order to reveal their capacity to resolve principal-agent problems and signal resilience despite low popularity or an assertive opposition.
 
-All together, I argue that a large margin of victory combined with election fraud sends an unambiguous signal of ruling-party strength, with uncertainty around its components.
-That is, large margins of victory can be composed of a number of values of 'natural' popularity and election fraud.
+Since the incumbents' claimed vote-share is part of the signal generated by the election, the effect of election fraud should only be considered in conjunction with the margin of victory.
+A large margin of victory combined with election fraud sends an unambiguous signal of ruling-party strength, with uncertainty around its components.
+That is, large margins of victory can be composed of a number of values of 'natural' popularity, mobilized voters, and election fraud.
+This makes large margins of victory difficult signals for opposition actors to interpret.
 As margins of victory decline, the relative value of the signal becomes clearer.
@@
 The combination of patronage resources and control over the criminal and judicial systems required for mass fraud can be turned to coopting or repressing protest movements.
+This argument is the mirror-image of the one present in Trinh [-@trinhTeaLeafElections2022], who argues that when incumbents manipulate heavily, they limit the quality of information that they can extract from the election.
+By the same token, a heavily manipulated election also confounds the signal received by opposition actors; a strong, fraudulent result is a signal of regime stability whose exact composition is uncertain.
+
+The argument can be summarized as follows.
+Incumbents will generally seek to manipulate elections as much as they can, to capitalize on both the direct and indirect benefits of electoral manipulation [@simpser2013].
+To do so, they must solve principal-agent problems by consolidating control over patronage resources and shielding agents from negative consequences of their actions.
+The election outcome sends an ambiguous signal of the incumbents' resources\textemdash including natural popularity, resource control, and control over the legal apparatus\textemdash to possible regime opponents.
+Strategic opposition actors, which may include both organized groups and individuals weighing costs and benefits of participation, interpret this signal with some prior expectation of the incumbent's capacity to manipulate.
+However, even when the incumbent's popularity is low, its ability to engage in election fraud is indicative of resources that can be used to terminate protest movements.
 
 This theory makes predictions for protest initiation, protest duration, and the level of fraud employed.
-First, if large-scale election fraud indicates strength, we should expect that incumbents will respond to pre-election signs of opposition strength by increasing the level of fraud.
+First, if large-scale election fraud indicates strength, we should expect that incumbents will respond to pre-election signs of opposition strength by increasing the level of fraud in order to send a more forceful signal.
 This runs directly counter to the prediction of the classic models, which hold that a greater risk of post-election protest should deter fraud.
-The prediction of second-generation [USE THIS AS SHORTHAND?] models is embedded in the null hypothesis of H1, since these models hold that protest risk is not tethered to the level of fraud itself (e.g.
-[@egorov2014incumbency]).
-In those models, a greater risk of protest makes it more important to incumbents that they satisfy the opposition's expected level of fraud, which may be high or low.
 
@@
 
-Likewise, if election fraud and incumbent vote-share are both indicator of incumbent strength, we should expect protest initiation to be most likely when both are low.
-A high vote-share in the absence of large-scale fraud indicates broader regime popularity, which dampens protest risk.
-As argued here, a high level of fraud can help compensate for a low vote-share by indicating independent sources of regime strength (i.e.
-resource capacity, agent loyalty, and biased rule of law).
-Only when both signals are low should we expect a high risk of protest.
-This hypothesis is consistent across all three schools of thought.
+Likewise, in an election where the incumbent claims victory, protest should be most likely when levels of fraud are high and claimed vote-shares are low.
+A low vote-share for the incumbent combined with large-scale fraud indicates that the incumbent's 'natural' popularity is very low.
+As claimed vote-share increases for a given level of fraud, the incumbent appears better supported and harder to challenge.^[Of course, the strongest signal of weakness would be one where vote-share and fraud are low, but in this case it is likely that the incumbent has accepted an election loss, rather than falsify a victory.
+This scenario is explored more in the appendix.] This hypothesis is consistent across all three schools of thought, but should nevertheless be evaluated.
 
@@
 
-Finally, if large-scale protest is indicative of regime resiliency, we should expect that election fraud will be associated with reduced protest duration among incumbents who successfully fended off protest movements.
+Finally, if large-scale protest is indicative of regime resiliency, we should expect that election fraud will be associated with reduced protest duration among those incumbents who successfully fend off protest movements.
 That is, among strong incumbents who were capable of defeating post-election protest movements, the resources associated with large-scale fraud should enable them to more swiftly quell the movement.
 Those who lack such resources are expected to take longer to dispatch the protest movement.
-The effect should run in the opposite direction for incumbents who were removed through post-election protest.
-These types, revealed ex post to be weak, should survive longer against the protest movement if they control the resources needed to carry out large-scale fraud.
-Those who lack such resources should be defeated more quickly.
-Again, these expectations are counter to the standard protest-deterrent logic and distinct from the null hypothesis expected in second-generation [USE?] models.
+Again, these expectations are counter to the standard protest-deterrent logic.
 
-> H3a: Among incumbents that survive post-election protest movements, more extensive election fraud should be associated with shorter protest waves.
-
->H3b: Among incumbents that are defeated by poste-election protest movements, more extensive election should be associated with longer protest waves.
-
-That election dates are known in advance creates opportunities for opposition mobilization; anticipating this, regimes can more easily engage in pre-emptive repression [@Truex2019;@Sullivan2016;@Bhasin2013].
-
-In addition, governments may learn from the experiences of other governments, just as opposition groups may learn from international opposition movements.
-Anticipating threats from below may encourage less democratic governments to invest more in the security services, or to reform political and legal institutions in order to stifle potential protests before they take shape [@hallAuthoritarianLearningConceptual2017a;@Morgenbesser2020], further reducing the expected cost of protest.
-
-In particular,
+> H3: Among incumbents that survive post-election protest movements, more extensive election fraud should be associated with shorter protest waves.
 
@@
 I limit the sample to authoritarian regimes which hold multiparty national elections.
-This excludes fully closed autocracies (where no national elections are held de jure), as well as electoral and liberal democracies.^[The specific variable used is 'v2x_regime'; I limit the sample to those regime-years coded 1.]
+This excludes fully closed autocracies (where no national elections are held de jure), as well as electoral and liberal democracies.^[The specific variable used is 'v2x_regime'; I limit the sample to those regime-years coded 1 for either the current or lagged year.
+This coding rule will include cases that are coded as electoral democratic after an election but electoral authoritarian before.]
 
@@
 At the same time, the more liberal-democratic a state becomes, the more alternatives their are to protest as a means to check the executive.
-More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest\textemdash the electoral authoritarian regimes\textemdash where we should expect to see a more pronounced role for protest risk in curtailing election fraud.
+More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest\textemdash the electoral authoritarian regimes\textemdash where we should expect to see a more pronounced role for protest risk in curtailing election fraud under the deterrence logic.
 The case selection strategy here is thus one that both fits the theory and makes for a strong test of the classic fraud-protest model.
@@
 Models where the dependent variable is the level of election fraud include data from all cases.
-For the models that focus on the initiation of post-election protest, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the theoretical expectations of protest risk following an election in which the incumbent won, possibly through reliance on election fraud.^[Excluding cases where incumbents stepped down after losing an election is necessary to understand the initiation of protest against the alternative\textemdash where incumbents claimed victory.
-Howevever, this design does influence the interpretation of the results from these models.
-Threats to inference from this section are evaluated in more detail in the discussion.] For models of protest-wave duration, only cases with a positive number of post-election protests are included.
+For the models that focus on the initiation of post-election protest, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the population of interest: protest following an election in which the incumbent won, possibly through reliance on election fraud.^[Excluding cases where incumbents stepped down after losing an election is necessary to avoid conflating pro- and anti-incumbent protest.
+However, this design does influence the interpretation of the results from these models.
+Threats to inference from this section are evaluated in more detail in the discussion and appendix.] For models of protest-wave duration, only cases with a positive number of post-election protests are included.
 The date range is limited by the availability of data on protest from ECAV, as discussed below, which ranges from 1989 to 2012.
@@
 It also matches the conception of election manipulation used in much of the relevant formal literature.^[For example, election manipulation is treated as post-hoc adjustment of the results in Little et al [-@Little2015], Little [-@Little2012], Magaloni [-@magaloni2010], and Fearon [-@fearon_self-enforcing_2011], among others.] The variable is used as a dependent variable in the first set of models that test H1, and as a potential explanatory variable in the subsequent models of protest initiation and duration.
-As an additional check in the appendix, a fuller measure of overall election integrity is used instead.
+Two alternative measures are used as robustness checks in the appendix.
 
-ECAV records protests and other contentious actions related to an election, ranging from six months before to three months after the vote (or three months after the final round of a multi-round election).
-To focus the study on protest, I only include events labeled as protests, occupations, or blockades, excluding acts like shootings, bombings, and coups.
-ECAV data is the event-day format; to capture the *number of protests* before the election, I take the sum number of events recorded in ECAV before election day.
+ECAV records protests and other events related to an election, ranging from six months before to three months after the vote (or three months after the final round of a multi-round election).
+To qualify, an action must be contentious; rallies in support of the incumbent are excluded.
+I only include events labeled as protests, occupations, or blockades, excluding acts like shootings, bombings, and coups.
+ECAV data is in event-day format; to capture the *number of protests* before the election, I take the sum number of events recorded in ECAV before election day.
 To evaluate hypotheses regarding *protest duration* I take the number of days between the last protest recorded in ECAV and the first, for each election.^[The duration of a protest wave with only one protest is taken to be one day.
@@
 The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0) %>% nrow()` of the `r nrow(nelda2.sub)` elections, no protests are recorded.
-Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest number and duration.
+Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest duration.
 As the figure shows, while election fraud is common among this set of regimes, protests are rare, and, when they do occur, are usually short in duration.
@@
 Lastly, protest wave duration is modeled using Cox proportional hazards models.
+
+In the survival models, the sample is restricted to the 156 cases where post-election protests occurred.
+If significant election fraud is indicative of regime resilience, the expectation is that more fraudulent elections will be associated with more quickly dispatched protest movements.
+As an additional robustness check, I subsequently exclude the 7 cases where incumbents were removed following protests.
+This allows me to exclude the possibility that incumbents who make greater use of fraud are *weaker*, and more quickly removed from office, than their more electorally honest peers.
+A brief further analysis of the seven cases in which incumbents were ousted is also included in the discussion.
 
@@
 
-```{r elex-hist, fig.cap="Distributions of election fraud, post-election protest, and protest duration", fig.height=7, echo=FALSE, message=FALSE, warning = FALSE}
+```{r elex-hist, fig.cap="Distributions of election fraud, post-election protest, and protest duration", fig.height=4, echo=FALSE, message=FALSE, warning = FALSE}
 plot.combined.hist <- cowplot::plot_grid(p.hist.fraud, p.hist.duration, ncol=1)
@@
                                          "Professional bureaucracy" = patronage.appts.1lag,
-                                        "Anti-system" = antisys.1lag,
                                         "EMB autonomy"= v2elembaut)
@@
 library(kableExtra)
-datasummary_skim(incl.vars, output = "kableExtra", title = "Table 1: Summary statistics and variable distributions ")  %>%
+datasummary_skim(incl.vars, output = "kableExtra", title = "Table 1: Summary statistics and variable distributions", histogram = F)  %>%
     kable_styling(font_size = 9, full_width = FALSE)
@@
 
-To add: Beaulieu book; Daxecker 2019 et al (fraud perceptions), Daxecker 2012 (cost of exposing cheating), Donno 2013 (book) , Beaulieu 2014 book, hyde and marinov 2014, Morrison, Donno, Savun, and Davutoglu (2024)
-
 ## Explanatory variables
@@
 
-[MOVE] Moreover, in general, incumbents tend to avoid mass repression in the pre-election phase in favor of targeted harassment of opposition activists [@Bhasin2013].
-
 Second, since a state with greater capacity for repression may be better able to deter protest and thus 'get away with' more manipulation [@Egorov2021,@Ritter2016], I measure repressive capacity using the *physical integrity* index from V-Dem, lagged one year.
 Higher values for this variable indicate greater freedom from torture and political killing.
-In one approach, higher values represent more permissive environments, and should be associated with greater protest risk (and thus reduced fraud).
-Alternatively, in keeping with the 'murder in the middle' framework [@feinLifeIntegrityViolationsDemocracy1995], the highest protest risks (and thus most reduced fraud) might be found at middle values of repression; regimes with the most demonstrated repressive capacity deter potential challengers from taking to the streets in the first place, and the least repressive states defuse protest through accommodation [@pierskallaProtestDeterrenceEscalation2010;@mooreRepressionDissentSubstitution2000].
+In keeping with the 'murder in the middle' framework [@feinLifeIntegrityViolationsDemocracy1995], the highest protest risks (and thus most reduced fraud) might be found at middle values of repression; regimes with the most demonstrated repressive capacity deter potential challengers from taking to the streets in the first place, and the least repressive states defuse protest through accommodation [@pierskallaProtestDeterrenceEscalation2010;@mooreRepressionDissentSubstitution2000].
 I thus include the variable's square term to account for the 'murder in the middle' framework.
 
-Finally, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby.
+Finally, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby [@Bunce2006a] (see Brancati and Lucardi [-@Brancati2019] for an alternative view, however).
 This diffusion mechanism is modeled using each country's geopolitical region as coded in Teorell et al [-@qog_teorell] in conjunction with binary variables indicating a successful protest movement from NELDA.
@@
 
-As identified in prior research, other factors that may influence both fraud and protest include the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), and the available of *alternative sources of information*.
+As identified in prior research, other factors that may influence both fraud and protest include the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), *judicial constraints* on the executive ('v2x_jucon') and the available of *alternative sources of information*.
 Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019].
@@
 Additional V-Dem variables capture *privatization* of the economy ('v2clstown') and *bureaucratic professionalization* ('v2stcritrecadm'), both of which might give incumbents influence over election fraud and subsequent protest.
-To capture the overall level of pre-election protection for rights associated with liberal democracy, I include the lagged *liberal democracy index* from V-Dem.
-*Physical integrity* is used as a control here, to address the risk of selection bias (if, for example, high-repression states select into a high-fraud condition with the hope that the repressive apparatus will deter protest).
+Lastly, a V-Dem measure the *autonomy of election managment bodies* ('v2elembaut') is included.
 
@@
 First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest.
-Finally, I include dummy variables that provide a categorical evaluation of state capacity.
+Another NELDA variable indicates if the incumbent executive is not running due to *term limits*.
+I include dummy variables that provide a categorical evaluation of state capacity.
 These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap).
@@
 Table \@ref(tab:summary-table) shows summary statistics for all of these variables.
-To address concerns about multicollinearity, I present models in the appendix (page 25) that exclude control variables associated with pre-election protest.
+To address concerns about multicollinearity, I present models in the appendix that exclude control variables associated with pre-election protest.
 
@@
 
-modelsummary(c(m1, m1b, m1c, m1a), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Table 2: Random-intercept multilevel model of election-day fraud', coef_omit = "SD", coef_rename =
+modelsummary(c(m1, m1a, m1c), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Table 2: Random-intercept multilevel model of election-day fraud', coef_omit = "SD", coef_rename = 
                c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
@@
 integrity index (lag)",
-                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity"
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "state.owned.1lag" = "Privatization",
+                 "patronage.appts.1lag" = "Professional bureaucracy",
+                 "v2elembaut" = "EMB autonomy"
                  )) %>% kable_styling(latex_options = "HOLD_position")
@@
 The results of the first set of multilevel models are not promising for argument that protest risk deters election manipulation.
-Hypothesis 1a is not supported in the models shown in Table  \@ref(tab:manipdv).
-Models 1 through 4 apply different measures of pre-election protest risk; all of them behave contrary to the standard expectation in the literature.
+Hypothesis 1 is supported in the models shown in Table  \@ref(tab:manipdv).
+Models 1 through 3 apply different measures of pre-election protest risk; all of them behave contrary to the standard expectation in the literature.
 In Model 1, the number of pre-election protests is associated with an increase in election-day fraud.
 While the effect is small\textemdash ten pre-election protests only tend to increase election fraud by about six percent of a standard deviation\textemdash it runs counter to the expectations of the deterrence logic.
-A similar effect is found for *diffusion* in Model 4; recent successful protest in a nearby country is also associated with a modest increase in expected fraud.
+A similar effect is found for *diffusion* in Model 2; recent successful protest in a nearby country is also associated with a modest increase in expected fraud.
 This suggests that the level of election manipulation, rather than being deterred by a highly contentious environment, may be deployed to make up for anticipated poor election results and/or to signal organizational strength.
 
-Model 2 does not demonstrate a significant linear relationship between repression and election fraud.
-However, Model 3 shows that fraud first increases with physical integrity, but then declines.
+Model 3 shows that fraud first increases with physical integrity, but then declines.
 In other words, fraud is lowest on the extremes of repressive capacity\textemdash where repression is severe and where it is most limited.
 This is contrary to the expectation that a higher risk of protest is associated with reduced fraud.
-Under the expectation that protest risk is highest for these middle values [@pierskallaProtestDeterrenceEscalation2010], H1a is not supported.
-By contrast, H1b is supported by Models 1, 3, and 4: by each measure, fraud is higher when protest risk is higher.
+Under the expectation that protest risk is highest for these middle values [@pierskallaProtestDeterrenceEscalation2010], H1 is supported by each measure; fraud is higher when protest risk is higher.
+Models with fixed effects by country, in the appendix, show that these findings holds even within cases.
 
 ## Fraud and the likelihood of protest
-
-The second set of models challenge the idea that more severe electoral manipulation is associated with an increased risk of protest.
-Beginning with the zero-inflation portion of the hurdle models in Table 3, which model the likelihood of *no* protest occurring after the election, the standalone coefficient for election fraud is not significant in Model 5.
-
-More fraudulent elections do appear associated with an increasing risk of protest initiation as margins of victory shrink, as shown in the top panel of Figure  \@ref(fig:margin-prob-fig).
-Here, typical values for low and high levels of fraud are used (one standard deviation below and above the mean, respectively).
-The top panel shows that more fraudulent elections are associated with a significantly higher risk of protest initiation, but only when winning vote-shares are extremely small.
-
-\newpage
 
@@
 
-model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary)
 
-model.margin.binary.int <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary.int <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary.int)
 
-model.margin.binary_noc <- glm(any_protest ~ jucon.1lag + legcon.1lag +  altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary_noc <- glm(any_protest ~ jucon.1lag + legcon.1lag +  altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary_noc)
@@
 models2 <- list()
-models2[['Model 5']] <- model.margin.binary
-models2[['Model 6']] <- model.margin.binary.int
-models2[['Model 7']] <- model.margin.binary_noc
+models2[['Model 4']] <- model.margin.binary
+models2[['Model 5']] <- model.margin.binary.int
+#models2[['Model 6']] <- model.margin.binary_noc
 
@@
                  "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
-                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
+                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state", "v2elirreg.inv" = "Election fraud",
                  "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
@@
 Table \@ref(tab:logit-table) shows the results for Hypothesis 2, which is expected to be supported across all three schools of thought.
-Model 5 is a base model, where incumbent margin is interacted with *intentional irregularities*; Model 6 includes the same interaction along with additional control variables that could be associated with both fraud and post-election protest.
-The expected, significant interaction effect is detected in both models, and it is in the expected direction: the occurrence of any protest is most likely when incumbent vote-share is low and fraud is high.
+Model 4 is a base model, where incumbent vote-share and *intentional irregularities* are standalone variables; Model 5 includes their interaction.
+The expected, significant interaction effect is detected in Model 5, and it is in the expected direction: the occurrence of any protest is most likely when incumbent vote-share is low and fraud is high.
 
@@
 The figure shows the predicted probability of post-election protest for two specific values of election fraud, across all values of incumbent vote-share, holding all other variables constant at their means (continuous variables) or modes (categorical variables).
-The figure highlights that the predicted probability of protest is only statistically significant for the two values of fraud at extremely low levels of incumbent vote-share.
+The values for election fraud represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively.
+The figure highlights that the predicted probability of protest is only statistically significant for these unusual values of fraud at extremely low levels of incumbent vote-share.
 This represents an exceedingly unlikely outcome along both variables.
 As the figure shows, there are only a small number of observations in the range where the predicted probabilities are statistically significant.
-At the same time, the values for *intentional irregularities* are also highly unusual.
-They represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively.
 
 While all three schools of thought are consistent with H2, the results here may be taken as circumstantial evidence in favor of the fraud-as-strength model.
-Fraud only appears significantly associated with increased protest risk when the incumbent's result in the election is very low and fraud is very high.
-This is suggestive of the model where opposition actors interpret the result as showing extremely low genuine support for the incumbent, perhaps enough that it outweighs the signal of strength sent by fraud for some oppositions.
+Model 4 shows that fraud alone has no bearing on protest occurrence; while the level of fraud only appears significant when the incumbent's result in the election is very low.
+This is suggestive of the model where opposition actors interpret the result as showing extremely low genuine support for the incumbent, perhaps enough that it outweighs the signal of strength sent by fraud.
 Middle values for vote-share combined with high values of fraud suggest durable regime strength, bringing expected protest close to zero.
@@
 
-```{r margin-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
+```{r margin-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
 plot_binary_margin_typ_n
@@
 ## Protest wave duration
-
-[CLEAN UP]
-
-Substantively, even if one does not fully discount the claim that the null results shown here represent equilibrium behavior, it is difficult to dispute that the resulting outcome is weighted heavily in favor of governments.
-In the dataset of 1110 elections used here, only 19 elections in five countries are coded as having an *intentional irregularities* score of less than -2 (indicating extremely fraud-free elections).
-By contrast, 121 elections in 27 countries score higher than 2 (indicating highly fraudulent elections).
-The frequency of fraud, the rarity of protest, and the finding that fraud severity is not correlated with the scope of protest events all point to the same conclusion: protest risk is not a substantial constraint on election fraud in electoral authoritarian and electoral democratic regimes.
 
@@
 
-cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0))
 summary(cox1)
-cz1 <- cox.zph(cox1, transform = "identity")
-print(cz1)
-plot(cz1[16])
-abline(h=0)
-abline(v = 21)
-abline(v = 61)
 
-cox1.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+cox1.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0))
 summary(cox1.cen)
 
-cox1t <- coxph(Surv(d.events.num, surv_status) ~ legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election +  regional_success_lag + n.events.post + v2elirreg.inv + tt(v2elirreg.inv), data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(15, 60, 90),
-Boundary.knots = FALSE))
-summary(cox1t)
+#cox1t2 <- coxph(Surv(d.events.num, surv_status) ~ legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election +  regional_success_lag + n.events.post + v2elirreg.inv + tt(n.events.post), data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(10, 70, 90),
+#Boundary.knots = FALSE))
+#summary(cox1t2)
 
-cz <- cox.zph(cox1, transform = "identity") #function(time) 1/(log(time + 1)))
-print(cz)
-plot(cz) # The points on the plot are residuals of v2elirreg.inv (and other vars) based on hte predicted value of X and time t.
-The line shows the estimated value of the coefficient on variable X at time t compared to its baseline.
-You want it to be zero for consistent effects over time.
-V2elirreg.inv is positive for about two weeks early on.
-The substantive effect of this can be shown by taking two values of v2elirreg.inv (i.e.
-first and third quartile), taking the difference, and finding exp(coefficient_residual * difference).
-The result shows here that protests in states in the third quartile of fraud are exp(.75*1.4) = 2.86 times more likely to die than those in the first quartile during those couple weeks.
-So this suggests that if protests survive the first week or two, the state's capacity starts to kick in?
-
-cox1t2 <- coxph(Surv(d.events.num, surv_status) ~ legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election +  regional_success_lag + n.events.post + v2elirreg.inv + tt(n.events.post), data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(10, 70, 90),
-Boundary.knots = FALSE))
-summary(cox1t2)
-
-cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
 summary(cox2)
 p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.05, 2]")) + coord_cartesian(ylim = c(0, 10)) #This is the 10th and 90th quantiles
-p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ration", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()
+p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()
 
-cox2.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+cox2.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
 summary(cox2.cen)
 
-cox3 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag , data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
-summary(cox3)
-
-cox4 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
-summary(cox4)
+cox3.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post + median.participants, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+summary(cox3.cen)
 
@@
 
-To further test the mechanism outlined in the theory above, I examine the relationship between fraud and the survival of protest movements in cases where protests began but the incumbent eventually survived.
-Restricting the sample to these cases allows for a consistent interpretation of results: longer protest waves indicate a more even match between the resources and resiliency of the protesters and the capacity and cohesiveness of the regime, while shorter-lived protest movements are those which were more easily defeated.
-If significant election fraud is indicative of regime resilience, the expectation is that more fraudulent elections will be associated with more quickly dispatched protest movements.^[Protest waves are defeated in the large majority of cases.
-There are 156 cases where protest occurs, out of which only 9 see new elections or the incumbent step down.]
-
 Table \@ref(tab:cox-models-tab) reports the results of Cox proportional hazard models of protest duration.
 In such models, coefficients with a positive sign indicate factors that are associated with shorter-lived protests; those with a negative sign indicate factors associated with a slower demise.
-The coefficient for *election fraud* is positive and significant in Model 7, indicating that more fraudulent elections are associated with more quickly dispatched protest movements.
+The coefficient for *election fraud* is positive and significant in Model 7, indicating that more fraudulent elections are associated with less resilient protest movements.
 Exponentiation of the coefficient for election fraud gives a hazard ratio of 3.3, indicating that for each one-unit increase in fraud, a protest movement becomes three times more likely to end at any given time.
-This supports H3b, and fails to support H3a.
+This supports H3.
 
 One threat to inference in this design derives from the 90-day post-election window considered by ECAV.
-It is possible that some protest movements considered 'dead' in Models 7 and 8 are, in fact, 'alive;' it is just that subsequent protests occurred after 90 days and are censored as a result.
+It is possible that some protest movements considered 'dead' in Models 7 and 8 are, in fact, 'alive,' if subsequent protests occurred after 90 days and are unobserved.
 As a test of the robustness of the conclusions, in Models 9 and 10, observations are assumed to be right-censored if a protest occurred within three-weeks of the 90-day cutoff.
@@
 While the coefficient on *election fraud* falls slightly short of conventional statistical significance in Model 9, it retains the same size and sign compared to Model 7.
-Similarly, the interaction effect in Model 10 increases in size, while retaining its direction and statistical significance.
-This suggests that the main models are not drawing overly incorrect inferences from longer-lasting protest movements.
+The interaction effect in Model 10 increases in size, while retaining its direction and statistical significance.
+This suggests that the main models are not drawing overly incorrect inferences through censoring of longer-lasting protest movements.
+Model 11 adds an additional control, the median number of participants at events in the protest wave.
+Larger protests might indicate relative opposition strength, however, since there is considerable missingness in the variable, I use it as a robustness check.
+Including this control variable reduces the size of the interaction coefficient, but it retains its direction and statistical significance.
 
@@
 As shown in the figure, for all incumbent vote-shares near and below fifty percent, large-scale fraud significantly increases the likelihood that a protest movement will end.
-This result also supports the fraud-as-strength model, and shows that only as vote-share increases above fifty percent does a government in a low-fraud election have the same likelihood of defeating a protest movement at a given time.
+This result also supports the fraud-as-strength model, and shows that only as vote-share increases above fifty percent\textemdash creating a stronger signal of popularity\textemdash does a government in a low-fraud election have the same likelihood of defeating a protest movement at a given time.
 
-```{r margin-dur-fig, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
+```{r margin-dur-fig, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
 p.duration.margin
@@
 In Figure \@ref(fig:km-fig), observations are divided into three groups.
-High fraud observations are those where *election fraud* is scored at one standard deviation above the mean or more, low fraud observations are those scored at one standard deviation below the mean or less, with moderate levels in between.
+High fraud observations are those where *election fraud* is one standard deviation above the mean or more, low fraud observations are those one standard deviation below the mean or less, with moderate levels in between.
 The figure shows a clear difference between the low-fraud group and the other two.
@@
 For much of the range of the data, they remain roughly twice as likely to survive at any given date than protest movements in medium- or high-fraud cases.
-This is evidence in favor of the idea that high levels of fraud is an indication of regime resilience, and contradicts the claim that fraud either reveals weakness or galvanizes political opposition, as it is the low-fraud cases where governments struggle to put an end to protest movements on average.
+This is evidence in favor of the idea that high levels of fraud is an indication of regime resilience, and contradicts the claim that fraud indicates weakness or galvanizes political opposition, as it is the low-fraud cases where governments struggle to put an end to protest movements on average.
 
-```{r km-fig, fig.cap="Kaplan-Meier curves for protest duration across levels of fraud", fig.height= 7, echo=F, message=F, fig.pos='H'}
+```{r km-fig, fig.cap="Kaplan-Meier curves for protest duration across levels of fraud", fig.height= 5, echo=F, message=F, fig.pos='H'}
 nelda2.sub <- nelda2.sub %>% mutate(fraud_level = ifelse(v2elirreg.inv >= (.98 + .97), "High",
@@
 p.km <- survfit2(Surv(d.events.num, surv_status) ~ fraud_level, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
-  ggsurvfit() +
+  ggsurvfit(linetype_aes = T) +
 #   add_confidence_interval() +
@@
 models3 <- list()
-models3[['Model 8']] <- cox1
-models3[['Model 9']] <- cox2
-models3[['Model 10']] <- cox1.cen
-models3[['Model 11']] <- cox2.cen
+models3[['Model 7']] <- cox1
+models3[['Model 8']] <- cox2
+models3[['Model 9']] <- cox1.cen
+models3[['Model 10']] <- cox2.cen
+models3[['Model 11']] <- cox3.cen
 
-modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Table 5: Cox proportional hazard models of protest duration', coef_rename = 
+modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Table 4: Cox proportional hazard models of protest duration', coef_rename = 
                c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
@@
                  "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
-                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
+                 "n.events.post" = "N. of post-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                  "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
 integrity index (lag)",
-                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "v2elirreg.inv" = "Election fraud",
-                 "tt(v2elirreg.inv)1" = "Election fraud (day 2-20)", "tt(v2elirreg.inv)2" = "Election fraud (day 21-90)"
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "v2elirreg.inv" = "Election fraud"
                  ))
@@
 The results can be summarized as follows.
-The severity of election fraud is not associated with protest initiation, except when combined with very low vote-shares for the incumbent.
-Indicators of pre-election instability are not associated with reduced fraud, but with higher levels.
-Finally, incumbents who have employed large-scale fraud appear better able to put an end to what protest movements do emerge than those who have presided over cleaner elections.
+Pre-existing risk to incumbents is positively, rather than negatively, associated with election fraud.
+The severity of election fraud itself is not associated with protest initiation, except when combined with very low vote-shares for the incumbent.
+Finally, incumbents who have employed large-scale fraud appear better able to put an end to those protest movements that do emerge, compared to those who have presided over cleaner elections.
 This is true even when incumbent vote-shares are low.
@@
 
-The results strongly contradict the assumptions of the classic school of thought, which holds that election fraud indicates regime weakness.
+The results strongly contradict the assumptions of models which hold that election fraud indicates regime weakness.
 Instead, across the three tests here, fraud appears to be an indicator of regime resilience.
 This is a substantively important distinction.
-Classical models hold that, since fraud indicates weakness, incumbents that would lose fair elections can be deterred from fraud by the risk of post-election protest.
+If fraud indicates weakness, incumbents that would lose fair elections can be deterred from stealing the election by the risk of post-election protest.
 Such models predict that a democratic exchange of power can occur in such scenarios, backed up by the risk of protest.
-The results here would appear to overturn that model.
+The results here would appear to cast doubt on those predictions.
 
-The results are also challenging for the second school of thought, which proposes that weak and strong incumbents both engage in electoral manipulation, and protest occurs when ruling parties under- *or* over-shoot the opposition's expected level of fraud.
-Here we would not expect to see consistent patterns with regard either to fraud deterrence or to fraud and the survival of protest movements.
-Instead we see a consistent pattern that associates higher levels of fraud with more resilient incumbents.
+These results are most consistent with the view that even election fraud can be an indicator of regime strength.
+Where this paper goes beyond the traditional formulation of this view\textemdash apart from testing the underlying assumptions\textemdash is the observation that large-scale fraud can be indicative of incumbent strength even when it is not excessive.
+That is, even when ruling party vote-shares are low (perhaps especially when this is so), incumbents which showed themselves capable of large-scale fraud are more capable of ending protest movements than low-fraud incumbents.
+This holds true despite controlling for indicators of other possible explanations of the relationship between low fraud and long-lived protest: civil society strength, media openness, and median protest size.
 
-Third, these results are most consistent with the third view, which holds that even election fraud can be an indicator of regime strength.
-Where this paper goes beyond the traditional formulation of this view\textemdash apart from testing the underlying assumptions of the three models\textemdash is the observation that large-scale fraud can be indicative of incumbent strength even when it is not excessive.
-That is, even when ruling party vote-shares are low (perhaps especially when this is so), incumbents which showed themselves capable of large-scale fraud are more capable of ending protest movements than low-fraud incumbents.
+A very brief consideration of the seven cases of successful electoral protest\textemdash Ukraine (2004), Georgia (2003), Kyrgyzstan (2005), Serbia (2000), Haiti (2006) and Cote d'Ivoire (2000 and 2010)\textemdash in comparison with the other cases can also help aid interpretation of the results.
+Table ...
+[CONSIDER ADDING TABLE] shows the average values for several key variables for three groups: incumbents defeated by protest, incumbents who survived protest, and cases where there was no electoral protest.
+The table shows that defeated incumbents did not hold elections that were noticeably more fraudulent than those in where incumbents survived protest: 1.13 for the survivors to 1.15 for the defeated.
+However, those that were defeated claimed substantially smaller vote-shares: 42 percent compared to 55 percent for the survivors.
+Meanwhile, the unprotested group average 55 percent vote-share and a fraud score of 0.87.
+This supports the view that (relatively) low fraud and high vote-shares are a reliable signal of strength (preventing protest), while high levels of fraud and low incumbent vote-shares indicate that the opposition's popularity might overcome the incumbent's structural advantages in a struggle in the street.
+This course of action is uncertain, however, as the unsuccessful protest group shows.
+Lastly,
 
 What does all this tell us?
-[FIX] First, these results may inform how researchers think about the causes of election manipulation.
-They suggest a that the importance of factors like the opposition's capacity to organize street protest in limiting election fraud has been overestimated.
+First, these results may inform how researchers think about the causes of election manipulation.
+They suggest that factors like the opposition's capacity to organize street protest in limiting election fraud have been overestimated.
 In fact, these results suggest that incumbents who have the resources will respond to such a scenario by *increasing* their production of fraud.
-Instead, the results point toward those factors that enable incumbents to both drive manipulation efforts and to quash protest movements: resource consolidation and protection for regime agents who act against the law in incumbents' interest.
-Further research on incumbents efforts to develop these capacities over time\textemdash and obstacles they face\textemdash may yield insights into election integrity.
+Consequently, theories of electoral manipulation that focus on factors other than protest risk\textemdash such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]\textemdash may be more important for understanding election integrity, which may drive future research as well as policy interventions.
+Additionally, it may change some current understandings of the role of election monitors, courts, and other third parties in upholding election integrity.
+As noted earlier, prior studies argue that these actors restrain manipulation by revealing information about fraud and intensifying protest risk.
+But if protest risk is largely untethered from the severity of fraud, it is likely that other mechanisms are at work.
+Scholarly attention might turn more profitably toward other explanations for manipulation, such as the organizational capacity of the state [@simpser2013], the popularity of the government [@rundlett2016deliver], and the risks faced by front-line agents [@Harvey2022].
 
@@
 We observe that protest initiation is most likely when fraud is high and incumbent vote-share is low.
-This could be indicative of strategic behavior by the rational opposition actors found in most formal models (who take this outcome as best evidence of a weak incumbent), or by less strategic behavior by citizens en masse.
+This could be indicative of strategic behavior by the rational opposition actors found in most formal models, or by less strategic behavior by citizens en masse.
 For such citizens, election manipulation may act as a shared grievance, especially when it appears that fraud may have been decisive [@kuntz_more_2009;@tucker2007].
@@
 
-There is a long-standing debate over the role of elections in democratization\textemdash whether they contribute to it or regime stability\textemdash of which the manipulation-protest nexus is a component [@Schedler2013;@Morgenbesser2019;@edgellWhenWhereElections2018].
+There is a long-standing debate over the role of elections in democratization, of which the manipulation-protest nexus is a component [@Schedler2013;@Morgenbesser2019;@edgellWhenWhereElections2018].
 At the core, much of this debate revolves around the extent to which non-democratic elections affect the ruling party's cost of suppressing the opposition versus tolerating it [@dahlPolyarchyParticipationOpposition1971].
@@
 To the extent that it makes electoral protest more likely and/or more costly, however, it increases the cost of suppression.
-In this version, election fraud has the potential to lay the groundwork for democratization by incentivizing the ruling party to tolerate organized opposition which, one day, it may be unable to repress after an election.
+One view holds that election fraud has the potential to lay the groundwork for democratization by incentivizing the ruling party to tolerate organized opposition which, one day, it may be unable to repress after an election.
 This logic can be found in studies of the color revolutions [@bunce2010], and in the observation that competitive authoritarian regimes are more likely to democratize after an election than hegemonic ones [@donno2013elections]
@@
 This provides additional theoretical support for understanding what some have called a 'wave of autocratization' or a 'democratic recession' in recent years [@luhrmannThirdWaveAutocratization2019;@diamondDemocraticRegressionComparative2021].
-If the same tools that produce broad election fraud also reduce the cost of protest, then electoral authoritarianism has a greater gravitational pull as a regime type.
+If the same tools that produce broad election fraud also reduce the cost of protest, then electoral authoritarianism exerts a sort of gravitational pull as a regime type.
 
-Indeed, while there are 164 cases of protest in elections where the incumbent claimed victory out of the 661 total non-democratic elections in this dataset, only 8 resulted in either the incumbent being replaced or the election being re-run due to mass protests.\footnote{This information is determined using the NELDA variables NELDA37 and NELDA41.} This suggests a `success rate' for post-election protest of about five percent *conditional on protest initiation*, and about one percent for all elections in electoral authoritarian and electoral democratic regimes.
-Moreover, the modal protest 'wave' in the dataset lasted only one day, with the median wave lasting three days.
-This is
+This paper is primarily concerned with whether or not election manipulation is indicative of regime weakness or resilience; its primary evidence is finding that fraud appears to increase with pre-election uncertainty and that fraud is associated with more resilient incumbents in the face of protest.
+Due to space constraints, two other predictions are not tested in depth here.
+First, one of the key predictions of the protest-deterrent approach is that weaker incumbents will generate little fraud and voluntarily step down after losing elections.
+There is indeed less election fraud observed in those election-years: the mean estimate of *election fraud* in the 129 observations where incumbents lost is 0.322, much lower than the 1.20 average in the group where incumbents claimed victory.
+Were these incumbents deterred from committing worse fraud?
+Or are they simply systematically more limited in their ability to generate election fraud?
 
-One challenge to the above interpretation considers the sample selection for the models which take protest initiation as the dependent variable.
-This paper is primarily concerned with whether or not election manipulation is indicative of regime weakness or resilience; its primary evidence is finding that fraud appears to increase with pre-election uncertainty and that fraud is associated with more resilient incumbents in the face of protest.
-Yet one of the key predictions of the classic models is that weaker incumbents will generate little fraud and voluntarily step down after losing elections.
-The models of protest initiation cannot evaluate this possibility, since cases where incumbents lose are excluded by necessity.
-Is there evidence to support this interpretation, and help salvage the protest-deterrent model?
-
-At first glance, the pattern appears promising.
-The mean estimate of *election fraud* in the 129 observations where incumbents lost is 0.322, much lower than the 1.20 average in the group where incumbents claimed victory.
-However, this observation alone does not discriminate between the two possibilities considered here: where these incumbents deterred from committing more severe fraud (despite being able to), or were they unable to muster an effective enough manipulation effort to 'win' the election?
-The overall pattern in the data leans toward the latter interpretation.
-Table \@ref(tab:anova-table) shows how different the two groups are along several dimensions that affect the availability of fraud as a strategy.
-For instance, among the cases where the incumbent lost, judicial independence is markedly higher.
-Similarly, the losing group has a substantially more privatized economy, a more professional bureaucracy, and more autonomous election management bodies.
-Additionally, the losing group also shows lower average economic growth.
-All of these should tend to reduce the availability of fraud.
-Of the factors considered here, only *number of pre-election protests* clearly behaves as the deterrence logic would predict, with the losing group subject to one more protest, on average, than the winning group.
-However, unlike the factors which point toward the alternative hypothesis, this difference is not statistically significant.
-Thus, while future research might test this aspect of the deterrence model more rigorously, the general pattern is supportive of the view put forth in this paper.
-
-```{r, include=F}
+```{r, include=F, warning=FALSE, message=FALSE, echo=FALSE}
 ## Anova
@@
 
-```{r anova-table, echo=FALSE}
-p.values <- c("NA", 3.009e-15, 6.014e-11, 6.149e-10, 0.01146,  3.758e-07,  0.6281, 0.003519,
+```{r anova-table, echo=FALSE, warning=FALSE, message=FALSE}
+p.values <- c(3.009e-15, 6.014e-11, 6.149e-10, 0.01146,  3.758e-07,  0.6281, 0.003519,
               0.1852, 0.007405, 0.003086, 2.2e-16)
-win_lose_diff <- c("NA", .87, -0.15, -0.16, .06, -.47, .04, 5.3, -1.46, -.265, -.257, -.893)
-variables <- names(anova_dat)
+win_lose_diff <- c(.87, -0.15, -0.16, .06, -.47, .04, 5.3, -1.46, -.265, -.257, -.893)
+variables <- c("Election fraud", "Judicial constraints", "Legislative constraints", "GDP growth (lag)", "Civil society openness (lag)", "Natl. party institutionalization (lag)", "Urban",
+               "N. pre-election protests", "Privatization", "Professional bureaucracy", "EMB autonomy")
 anova_dat <- data.frame(cbind(variables,win_lose_diff, p.values))
@@
 anova_dat <- anova_dat %>% mutate(p.values = as.numeric(p.values))
+anova_dat <- anova_dat %>% rename(Variables = variables)
+anova_dat <- anova_dat %>% rename("Win-Loss Difference" = win_lose_diff)
+anova_dat <- anova_dat %>% rename("P-value" = p.values)
 
-kable(anova_dat, digits = 3)  %>%
+kable(anova_dat, booktabs = T, digits = 3, caption = "Table 5: Mean variable values for winning incumbents minus losing incumbents, with ANOVA p-values")  %>%
     kable_styling(font_size = 9, full_width = FALSE)
@@
 
-The results indicate that we should be cautious when building and interpreting theories of politics outside liberal democratic regimes that emphasize the risk of mass protest as a trip-wire to be avoided, and add empirical evidence to Przeworski's [-@przeworskiFormalModelsAuthoritarian2022] critique of such models.
+The overall pattern in the data leans toward the latter interpretation.
+Table \@ref(tab:anova-table) shows the mean differences between the two groups, with ANOVA p-values for those differences.
+Negative values indicate that the losing incumbents had higher values for the variable, on average.
+For instance, among the cases where the incumbent lost, judicial independence, professionalization of the bureaucracy, and the autonomy of election management bodies are all markedly higher.
+Additionally, the losing group shows lower average economic growth.
+All of these should tend to reduce the availability of fraud.
+Of the factors considered here, only *number of pre-election protests* clearly behaves as the deterrence logic would predict, with the losing group subject to one more protest, on average, than the winning group.
+However, unlike the factors which point toward the alternative hypothesis, this difference is not statistically significant.
+Thus, while future research might test this aspect of the deterrence model more rigorously, the general pattern is supportive of the view put forth in this paper.
+Further checks are performed in the appendix, including models which show that predicted levels of fraud based on structural factors alone are strongly predictive of incumbent defeat, while pre-election measures of risk are not.
+
+The second question this paper does not closely examine is the role of election fraud in sustaining (or not) the small number of incumbents who are defeated by mass protest.
+These cases are: Peru 2000, Cote d'Ivoire 2000, Georgia 2003, Malaysia 2004, Ukraine 2004, Azerbaijan 2005, and Kyrgyzstan 2005.
+The causes of the post-communist cases, which account for most of the seven cases, has been much debated [@LucanWay2008;@Bunce2009].
+Yet for most of these cases, election fraud either held steady after the 'successful' revolution, or increased before partially declining again some years later.
+These cycles, identified as a feature of 'patronal politics' [@hale_patronal_2014] could indicate that in these cases, fraud patterns and protest were caused by more fragmented elites who failed to resolve the principal-agent problems needed to engage in mass protest and control the streets.
+A thorough qualitative comparative analysis of the cases would be informative, but a sketch can be drawn that supports the theory here.
+
+Ukraine's Orange Revolution, for example, combined very high fraud with a very low incumbent vote-share.
+The analysis of residuals conducted in the appendix finds that fraud in the 2004 election far exceeded expectations based on structural factors.
+Combined with a term-limited incumbent, a strategic opposition may have interpreted these signals as indicating a deeply unpopular ruling party which would struggle to put down a determined protest movement due to a fragmented elite.
+On the other hand, the 2005 Tulip Revolution in Kyrgyzstan came on the heels of moderate election fraud that fell well below structural expectations (i.e.
+large negative residuals).
+This struggle to generate election fraud may have been interpreted as a sign of a weak ruling party, leading to successful protest.
+
+Nevertheless, electoral protest is rare in non-democracies, and successful protest extremely so.
+The overall results indicate that we should be cautious when building and interpreting theories that emphasize the risk of mass protest as a trip-wire to be avoided in non-democracies, and add empirical evidence to Przeworski's [-@przeworskiFormalModelsAuthoritarian2022] critique of such models.
 The rarity of post-election protest, especially successful protest, suggests that such events are highly contingent; they hinge on a multitude of factors that go beyond the regime's efforts to bias election outcomes.
-Both protest and election fraud are complex phenomena with multiple drivers, including resource costs, signaling, information-gathering, principal-agent dynamics, and more; these findings indicate that protest risk is not a significant factor under most circumstances.
+Both protest and election fraud are complex phenomena with multiple drivers, including resource costs, signaling, information-gathering, principal-agent dynamics, and more; these findings indicate that protest risk is not a practical constraint on fraud under most circumstances.
 
-This has several implications for understanding election manipulation and regime resilience.
-By overlooking the poor quality of election fraud as a signal, or by treating protest primarily as a binary event, the existing literature on electoral protest has overemphasized the risk to incumbents of engaging in electoral manipulation.
-The results presented here show that fraud has no bearing on how numerous or large protest events are.
-Moreover, heightened protest risk appears to be associated with *more* election fraud, rather than less.
-Instead, the conditions that are correlated with more damaging protest are generally institutional factors, especially executive term limits, which are either more amenable to control or can be anticipated by incumbents.
+Finally, this paper does not examine the role of protest risk in insulating established democracies from democratic backsliding [@fearon_self-enforcing_2011].
+This question deserves further investigation, but it is possible to draw a few implications from this study.
+First, even if protest risk fails to serve as a deterrent in non-democracies, it may still do so in democracies.
+In non-democracies, protest initiation is most likely when fraud is very high and incumbent vote-shares very low.
+The likelihood of protest initiation in a democracy may well be higher, triggering at lower levels of fraud and higher incumbent vote-shares, since opposition actors are likely to be better resources in a liberal democracy (and the government more constrained).
+At the same time, if the theory described here is accurate, liberal democracies would be constrained against using large-scale election fraud even in the absence of protest risk: they lack the patronage resources and arbitrary enforcement of laws needed for such a scenario.
 
-These are pessimistic findings for proponents of democracy.
-That electoral protest is less risky than is often assumed, in turn, implies that governments and ruling parties have a freer hand to engage in electoral manipulation than many models posit.
-Since the severity of fraud has no apparent bearing on whether or not a major protest wave will emerge as a serious threat to the regime, government have an incentive to manipulate as much as possible to secure a large winning vote-share.
-This perverse incentive undermines the case that free and fair elections in an electoral democracy represent a self-enforcing equilibrium backed up by the risk of citizen protest [@fearon_self-enforcing_2011], or that risk of popular mobilization can substantially moderate election manipulation in non-democracies.
-In effect, it is the existence of an election itself [@lucardiStrengthExpectationElections2019], and the opportunity it offers for opposition mobilization [@howard2006], that creates protest risk for incumbents, giving them a relatively free hand to manipulate to the extent they can recruit agents to do the work.
+It is also worth noting that mass electoral protest can serve an illiberal purpose, as illustrated by the events of January 6, 2020 in the United States or 2022-2023 in Brazil.
+Indeed, out of the 156 instances of post-election protest in this dataset, 10 occurred after relatively clean elections (*intentional irregularities* < 0)\textemdash a small but not minuscule percentage.
+These incidents demonstrate that, while crying fraud may sometimes be a useful tool for protest entrepreneurs, losing an election alone can be a catalyst for protest, irrespective of the actual level of election manipulation.
+This possibility tends to reduce the 'opportunity cost' to incumbents of engaging in real fraud, while also underlining the fact that the initiation and course of protest is often driven more by the relative resources and opportunities available to the disaffected and the incumbent government.
 
-This study also suggests implications for the study of election manipulation.
-First, the study implies that supply-side theories of electoral manipulation\textemdash such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]\textemdash may be more important than protest risk for understanding election integrity, which may drive future research as well as policy interventions.
-Third, and relatedly, it undermines some current understandings of the role of election monitors, courts, and other third parties in upholding election integrity.
-As noted earlier, prior studies argue that these actors restrain manipulation by revealing information about fraud and intensifying protest risk.
-But if protest risk is largely untethered from the severity of fraud, it is likely that other mechanisms are at work.
-Scholarly attention might turn more profitably toward supply-side explanations for manipulation, such as the organizational capacity of the state [@simpser2013], the popularity of the government [@rundlett2016deliver], and the risks faced by front-line agents [@Harvey2022].
-
-Second, it suggests avenues for future research.
+<!-- This project suggests additional avenues for future research.
 It is clear that electoral protests charged by allegations of manipulation do occur at times.
@@
 To the extent that countries vary according to the size of such populations, they may vary in their underlying protest risk.
-
-As a final point, the results also speak to incidents of mobilization around unfounded allegations of election-rigging, such as those of January 6, 2020 in the United States or 2022-2023 in Brazil.
-Indeed, out of the 293 instances of post-election protest in this dataset, 18 occurred after very clean elections (those one standard deviation below the mean for election fraud)\textemdash a small but not minuscule percentage.
-These incidents demonstrate that, while crying fraud may sometimes be a useful tool for protest entrepreneurs, losing an election alone can be a catalyst for protest, irrespective of the actual level of election manipulation.
-This possibility tends to reduce the 'opportunity cost' to incumbents of engaging in real fraud, while also underlining the fact that the initiation and course of protest is often driven more by the relative resources and opportunities available to the disaffected and the incumbent government.
+-->
 
 These results are consistent with different specifications presented in an online appendix.
-In particular, the main results are supported when data from the Mass Mobilization dataset are used in place of the ECAV data (appendix page 1), and when a measure of overall election integrity is used in place of the measure of election fraud (appendix page 15).
-[CHECK PAGES]
-
-[MOVE TO DISCUSSION AND REWRITE] Generally speaking, protest-based models predict that manipulation should decline as opposition groups become stronger relative to incumbents.
-This dynamic creates a sort of virtuous cycle, wherein strong opposition groups' credible threat of protest makes elections fairer, which in turn improves the strength of the opposition--leading ultimately to the ideal of self-enforcing democracy.
-The alternative school of thought, among other predictions, argues that opposition strength drives front-line agent behavior more than that of leaders, and causes a shift to less dispersed, less observable kinds of manipulation [@asunka2017electoral;@Harvey2019].
-Election integrity can thus remain low, even in competitive contexts.
+In particular, the main results are supported when data from the Mass Mobilization dataset are used in place of the ECAV data, and when alternative measures of election integrity are used in place of the measure of election fraud.
+They are also supported when electoral democracies are included in the dataset.
+The relationship between fraud levels and incumbent defeat is tested.
+Additional robustness checks include models with country fixed-effects and alternative specifications for selected control variables.
 
@@
 
-Protest risk is often considered to be a major risk to incumbent governments aiming to rig elections.
-This risk has a central role in several prominent models of electoral manipulation, serving to deter manipulation and to incentivize the toleration of election monitors and other restraints on the ruling party.
-However, I argue that such models overlook the ambiguity inherent in election manipulation efforts, as well as the quick decay of even such limited information once protests begin.
-Empirical research on the subject has largely treated protest as a binary variable, an approach which obscures the distinction between small, one-off protests and much more costly protest waves.
+There are competing views over the nature of election fraud, incumbent strength, and protest.
+The risk of post-election protest has a central role in several prominent models of electoral manipulation, serving to deter manipulation and to incentivize the toleration of election monitors and other restraints on the ruling party.
+Others contend that election fraud can indicate incumbent strength.
+Building on the principal-agent literature of election fraud, I argue that large-scale fraud is indicative of incumbent resilience even when regime popularity (and thus vote-shares) are low.
+Conducting election fraud requires consolidated control over patronage resources and the ability to protect front-line agents from the consequences of illegal behavior; this capacity also helps insulate incumbents from the effects of mass protest.
+Resolving this tension in the literature has important implications for the study of election fraud and authoritarian stability.
 
-This paper addresses these concerns by testing the relationship between election fraud and the number and maximum size of protest events that occur after an election.
-Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 1110 elections to show that election fraud has no significant relationship with the number or size of protests that occur in a protest wave.
-Additional tests suggest that this null finding is not the result of equilibrium behavior predicted in deterrence models.
-Instead, most fraudulent elections go un-protested, and most post-election protests appear to be strategic gambles gone wrong.
-The initiation of protest more often reveals the incumbent to be strong enough to manage protest through repression, cooptation, or other concessions.
-The results suggest that theories which treat protest risk as an unqualified deterrent to election manipulation should be reconsidered.
+This paper attempts to test the assumptions of these two models by investigating the relationship between pre-election risk and election fraud, as well as between election fraud and protest occurrence and duration.
+Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 661 elections in electoral authoritarian regimes to show that 1) when pre-election risks to incumbents are higher, subsequent fraud increases rather than decreases, 2) protest occurrence is most likely when incumbent vote-shares are very low and election fraud is high, and 3) incumbents that engaged in larger-scale election fraud are able to more quickly defeat protest movements.
 
-The absence of protest risk as a guardrail is felt in both electoral authoritarian regimes and in electoral democracies\textemdash those democracies that are most at risk of democratic erosion.
-The evidence thus suggests that mass demonstrations in favor of advancing (or restoring) democratic elections are rare and unlikely to be driven by the severity of manipulation---a gloomy picture for supporters of democratization---and that researchers should consider other mechanisms beyond mass protest in their models of electoral manipulation.
+All together, these results suggest that protest risk largely fails to serve as a constraint on election manipulation in authoritarian regimes.
+The absence of protest risk as a guardrail suggests that mass demonstrations in favor of advancing (or restoring) democratic elections is unlikely to be driven by the severity of manipulation, and that electoral authoritarianism as a regime-type is likely to be a stubborn equilibrium.
+This is a gloomy picture for supporters of democratization, and suggests that researchers should consider other mechanisms beyond mass protest in their models of electoral manipulation.
 

### 2024-07-29 Scripts/authoritarian_only_latex_2.Rmd
Authoritarian only, before re-write to compare non-strat + strong vs. strategic and weak

--- 2024-07-15 Scripts/authoritarian_only_latex.Rmd
@@
 output:
-    pdf_document:
-      keep_tex: yes
+    bookdown::pdf_document2:
+    keep_tex: true
 linkcolor: blue
@@
   - \captionsetup[table]{labelformat=empty}
+  - \usepackage{multirow}
+  - \usepackage{lscape}
+  - \usepackage{pdflscape}
+
 ---
@@
 The post-2011 protest wave is one much-studied case of the phenomenon of electoral protest aimed at punishing electoral manipulation.
-Yet the Russian example raises questions for that broader phenomenon: does election fraud itself make protest more likely, or only in the context of a weak electoral showing for the ruling party?
-Given the ruling party's poor result and substantial electoral fraud, why was it able to dispatch the mobilized opposition with relative ease?
+Yet the Russian example highlights questions that get to the heart of a debate in the study of electoral manipulation and protest.
+When opposition movements protest manipulated election results, are they doing so because the incumbent\textemdash by relying on manufactured votes\textemdash has signaled its own weakness?
+This view is adopted by many researchers, especially in the formal modelling tradition.
+Or, does election manipulation indicate sources of strength for incumbents (independent of genuine popularity), as others have held?
 
-These questions get to the heart of a debate in the study of electoral manipulation and protest: does election fraud signify regime weakness or strength, and in what ways?
-The various theoretical perspectives brought to bear on the topic have been built on different assumptions, and yield different implications for authoritarian governance and democratization.
-For some models, election fraud is a tool of weak incumbents.
-As a result, incumbents avoid fraud when it might be exposed and reveal their weakness to potential protesters\textemdash an auspicious prediction for democratization and ultimately the under-pinning of 'self-enforcing' democracy [@fearon_self-enforcing_2011].
-Another school of thought holds that election fraud can be an indicator of incumbent strength.
-This approach, which is expanded upon below, is a pessimistic account for democracy.
-If this model is correct, it is authoritarian politics rather than democracy that is self-enforcing\textemdash the ability to commit large-scale election fraud helps win elections, signal strength, deter challengers under most conditions, and put down protest movements if they do begin.
+The stakes of this debate are significant.
+The former view is an optimistic one; incumbents avoid fraud when it might be exposed and reveal their weakness to strategic protesters\textemdash an auspicious prediction for democratization and ultimately the under-pinning of 'self-enforcing' democracy [@fearon_self-enforcing_2011].
+If the latter view holds, electoral protest may mainly occur non-strategically, and would typically fail to deter election manipulation as a result.
+Mass protest would, in this case, be a fragile guardrail against autocratization.
 This paper argues that this pessimistic model best fits the overall pattern of the data, and is supported by specific, novel hypothesis tests.
 
-In this paper, I argue \textemdash in keeping with a growing literature on the informational aspects of elections in authoritarian regimes [@guriev2020;@smyth2020elections;@Geddes1989;@Miller2015a;@Pop-Eleches2015]\textemdash that an election outcome produces a signal of the relative strength of the incumbent.
-As Magaloni [-@magaloni2006] and others have noted, incumbents typically work to produce a higher margin of victory in the election, in order to demonstrate their strength and induce cooperation from other actors in society.
-In some theories, this works best if the votes for the incumbent are largely 'genuine' (i.e.
-not fabricated), or produced using the incumbent's organizational resources (i.e.
-vote-buying and intimidation) [@Harvey2020;@magaloni2006].
-These two kinds of votes correspond to two of the three 'pillars of stability' identified by Gerschewski [-@Gerschewski2013]: legitimation and cooptation.
-Fraud is considered a form of cheap talk in these models, which opposition actors can discount if they become aware of it [@little_elections_2012].
+This paper contributes to this debate by drawing out the implications of three strands of the literature on election manipulation and protest.
+In short: there are two views on the nature of electoral protest which make the same prediction.
+If opposition groups strategically initiate protest when they believe the incumbent to be weak, they will be most likely to do so when the election results show a low vote-share for the incumbent combined with high levels of manipulation\textemdash a combination of factors which reveals the incumbent's very low genuine popularity.
+However, if protest movements are more organic and non-strategic, reflecting citizen grievances, we should expect protest initiation in the same circumstances.
+Here the combination of low incumbent vote-share and high manipulation indicates severe citizen grievances (low natural popularity) which are galvanized by the shared grievance of a 'stolen' election.
+Since both theories make the same prediction about protest initiation, such tests\textemdash which have been the dominant empirical measure in the literature\textemdash do not reveal whether protest is strategic and manipulation indicates weakness or manipulation indicates sources of strength and anti-manipulation protest is non-strategic.
 
-However, these models overlook the fact that engineering widespread fraud requires solving principal-agent problems through consolidated control over patronage resources and the criminal and judicial apparatus [@Harvey2019;@Harvey2022].
-As a result, high levels of fraud are a sign of incumbent strength in both cooptation and repression (Gerschewski's third pillar).
-Both of these tools can be useful for defusing or dismantling an emergent protest movement.
-As a result, effective fraudsters should also be more effective at defeating protest movements.
-Consequently, in this view, there is little penalty for severe manipulation in terms of protest risk, and little deterrent effect.
+To unravel this question, I make use of data on the duration of post-election protests.
+Shorter-lived protest movements that end with the incumbent still in power indicate a government that is relatively strong compared to the protesting opposition, while longer-lived protest movements indicate a more even balance of power between the two.
+The reverse is also true: shorter-lived protest movements that end in opposition victory indicate a weaker government, while governments that hold on longer before being defeated are relatively stronger.
+Using data from 614 non-democratic elections across 97 countries, from 1989 to 2012, this paper finds that 1) protest initiation is most likely when manipulation is severe but incumbent vote-share is low, as both theories predict, and 2) that governments that engage in more manipulation are likely to put down protests more quickly than governments that produce little manipulation, especially when incumbents claim small vote-shares.
 
-Testing this proposition in more depth is difficult.
-If, for example, incumbents manipulate just the right amount to avoid mass protest in equilibrium\textemdash and are deterred from manipulating more by that risk\textemdash we would observe no statistical relationship between the severity of fraud and the likelihood of protest initiation.
-High and low levels of fraud would generally be associated with no protest.
-Beginning to untangle the relationship between election fraud and protest requires looking beyond standard binary measures of protest.
-Instead, this paper tests outcomes "off the equilibrium path," to use the language of formal modeling.
-Does greater risk to the incumbent in the pre-election period result in reduced election fraud?
-When protest movements do arise after an election, is fraud associated with weaker governments?
-Using data from 614 non-democratic elections across 97 countries, from 1989 to 2012, this paper finds that the answer to both questions is no.
-In fact, greater pre-election risk is associated with worse fraud, and more severe election fraud is associated with a *faster* demise for protest waves.
-This indicates that, far from being a sign of weakness, in most cases the ability to generate significant electoral fraud is an indicator of durable regime strength.
-This finding holds even when incumbents claim small vote-shares, suggesting that electoral fraud capacity is an independent component of regime stability.
-
+This pattern in the results is consistent with the argument that, rather than being a sign of weakness, the ability to generate significant electoral fraud is an indicator of durable regime strength.
+It is also consistent with the argument that much post-election protest in non-democracies is non-strategic; i.e., the actions of a weak opposition against a strong incumbent.
 Given these results, it is likely that the downside risk for incumbents that engage in election fraud in non-democracies has been overemphasized.
-Fraud itself is a sign of the robustness of some of the same tools that can be used to co-opt the opposition, constrain elites, or intensify repression during a crisis; far from being a sign of weakness, it can be a sign of resilience.
-This can deter protest, especially in conjunction with high claimed vote-shares for the incumbent.
-Even in the unlikely event that protests do break out in response to large-scale fraud, a government that has engaged in widespread fraud has many tools it may deploy to curtail them.
-
-These results help explain several empirical patterns at once: the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory in conjunction with high levels of fraud.
+This interpretation helps explain several empirical patterns at once: the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory in conjunction with high levels of fraud.
 At a theoretical level, they also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation.
@@
 
-# Literature review: election manipulation and mass protest
+# Literature review: three views of election manipulation and mass protest
 
-Prior research has often considered that electoral manipulation has both costs and benefits for ruling parties.
-The benefits include better odds of winning the official tally [@lehoucq2002], of course, as well as any signals of strength and stability that a larger vote-share [@magaloni2006] or display of power might provide [@simpser2013;@gehlbach2015electoral].
-On the other side, election manipulation is generally characterized as having both operational costs (which the incumbent may be unwilling or unable to pay [@Greene2010]) and legitimacy costs which reduce other actors' level of cooperation with the incumbent [@birch2011].
-The most visible, and arguably most studied, form of such legitimacy costs is mass post-election protest against the incumbent.
+Three understandings of the relationship between election manipulation and post-election protest have developed in the literature.
+Each view is discussed in detail below, after a brief initial summary here.
+A first view assumes that incumbents and opposition groups are strategic actors who who evaluate the election period for signs of each other's relative likelihood of winning a power struggle.
+Researchers who model election protest this way generally treat the use of election manipulation as a sign of the incumbent's unpopularity, and therefore weakness.
+Protest occurs after fraudulent elections in response to this perceived weakness; this possibility can serve to deter incumbents from engaging in manipulation under some circumstances.
+This is a normatively optimistic theory.
 
-The cost of protest is conceptualized as the risk that the incumbent may be unseated and/or the resource cost needed to repress or buy off the incumbent's opponents.
-Various theories model the decision to engage in post-election protest either as a strategic one by opposition groups [@magaloni2010;@Luo2020;@przeworski2015], or more spontaneous coordination by citizens expressing a shared grievance [@tucker2007;@Brancati2016].
-In either case, the risk of protest is a classic component of models of electoral manipulation, as a restraint on incumbents who would prefer to use fraud to increase their vote-shares, all else held equal.
-This argument can be found in theories about the conduct of non-democratic elections, as well as in the literature on self-enforcing democracy [@fearon_self-enforcing_2011;@przeworskiSelfenforcingDemocracy2006].
-In the latter literature, generally free and fair elections are thought to be a self-enforcing equilibrium through the threat of mass protest by wronged groups if the incumbent government rigs the results.
-While the focus here is on elections in authoritarian cases, some implications for self-enforcing democracy are discussed at the close of the paper.
+A second worldview argues that election manipulation can indicate a well-resourced, well-organized, powerful incumbent (independent of its popularity).
+Under this view, extensive election manipulation can be a show of strength, which induces cooperation with the incumbent from other societal actors.
+Since this view holds that more manipulation can further entrench the incumbent, it is a normatively pessimistic view.
 
-<!-- Broadly, prior work has considered two major explanations for the severity of manipulated elections: the costs and benefits of manipulation to the ruling party and government (a demand-side framework) and the costs and benefits for the front-line actors who do the work of manipulating elections (a supply-side framework).
-The risk of costly protest is the central, though not sole^[In some studies, like [@birch2011], the costs of manipulation are attributed to a more diffuse loss of legitimacy; such costs *may* manifest as protest, but may also manifest as a reduction in general cooperation with the regime.
-However, it is worth noting the argument of [@gehlbach2015electoral] that election manipulation can *induce* cooperative behavior from some agents.], factor which affects the demand-side cost of manipulation in most models.
--->
+Finally, others have argued that electoral protest occurs unrelated to strategic concerns about the relative strength of the incumbent and the opposition.
+In these models, electoral protest occurs because manipulation galvanizes citizens who hold diverse grievances, because economic conditions are poor, or simply because holding an election creates a moment of opportunity.
+The central feature of these models is that protest decisions are generally made without consideration for the likelihood that protests will overthrow the incumbent.
+As will be discussed in depth below, these different assumptions yield different predictions for when electoral protest is more likely to be initiated, and how long it will last against an incumbent that refuses to give up power.
+Table 1 presents these predictions for ease of reference.
 
-<!-- In the major formulations of this argument, such as Fearon  [-@fearon_self-enforcing_2011] and Przeworski [-@przeworskiSelfenforcingDemocracy2006], 'democracy' is understood in procedural terms.
-For Przeworski, democracy requires that "authorization to rule is derived from elections" which designate winners and losers according to a set of rules [@przeworskiSelfenforcingDemocracy2006, p.
-312].
-Fearon adopts a similar definition for electoral democracy [@fearon_self-enforcing_2011, p.
-1662].
-Such arrangements are not limited to liberal democracies, and in fact constitute the large majority of regimes today.
-Many previous works argue that the risk of protest can mitigate election manipulation in regimes where elections are held\textemdash including established democracies [@millerRepublicIfYou2021], electoral authoritarian regimes [@Egorov2021], and any regime where elections are held but may be manipulated [@Little2015;@Luo2020].
-In most such models, protests are successful in overthrowing the government with some probability, but impose a cost on both the protesters and the government regardless of success (e.g.
-[@Luo2020]); the logic thus extends even to highly repressive and entrenched regimes.
-Nevertheless, this body of work has yet to grapple with the empirical finding that election integrity is relatively poor across these regimes outside of a narrow club of 'consolidated' democracies.
--->
+```{=latex}
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+% \usepackage{multirow}
+\begin{table}[]
+\begin{tabular}{@{}llll@{}}
+\toprule
+                                                                                                              & Manipulation signals: & \multicolumn{1}{c}{Strategic protest}                                                           & \multicolumn{1}{c}{Non-strategic protest}                                                  \\ \midrule
+\multicolumn{1}{l|}{\multirow{2}{*}{Initiation most likely}}                                                  & Weakness              & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Low margin + \\ high manipulation\end{tabular}}  & \multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}Low margin + \\ high manipulation\end{tabular}} \\ \cmidrule(lr){2-3}
+\multicolumn{1}{l|}{}                                                                                         & Strength              & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Low margin + \\ low manipulation\end{tabular}}   &                                                                                            \\ \cmidrule(l){2-4}
+\multicolumn{1}{l|}{\multirow{2}{*}{Initiation least likely}}                                                 & Weakness              & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High margin + \\ low manipulation\end{tabular}}  & \multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}High margin +\\ low manipulation\end{tabular}}  \\ \cmidrule(lr){2-3}
+\multicolumn{1}{l|}{}                                                                                         & Strength              & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High margin + \\ high manipulation\end{tabular}} &                                                                                            \\ \cmidrule(l){2-4}
+\multicolumn{1}{l|}{\multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}Duration when \\ incumbent wins\end{tabular}}} & Weakness              & \multicolumn{2}{c}{More manipulation: Longer protest}                                                                                                                                        \\
+\multicolumn{1}{l|}{}                                                                                         & Strength              & \multicolumn{2}{c}{More manipulation: Shorter protest}                                                                                                                                       \\ \cmidrule(l){2-4}
+\end{tabular}
+\caption{Predictions for protest initiation and duration under varying theoretical assumptions}
+\label{tab:predictions}
+\end{table}
+```
 
-Much of the work on this subject has been done in the formal modeling tradition.
-In game-theoretic research where the risk of protest is shown to deter fraud, the election is usually conceived as a signaling game.
-In such models, incumbents are characterized as 'weak' or 'strong' based on their support in the electorate; weak incumbents may need to resort to fraud in order to secure re-election.
+In classic models of election manipulation, manipulation is a short-term tool used instrumentally to secure victory in the election [@lehoucq2002].
+Much of the work in this tradition has been elaborated through formal models.
+Though there are, of course, many variations, in several such models incumbents are characterized as 'weak' or 'strong' based on their true support in the electorate; weak incumbents may need to resort to fraud in order to secure re-election.
 As a result, the opposition perceives election fraud as a reliable signal of incumbent weakness, and will stage protests in order to push out the weak incumbent.
@@
 Similar results are developed by Chernykh and Svolik [-@chernykh2015third], Luo and Rozenas [-@Luo2020] and [-@Luo2017], and Magaloni [-@magaloni2010].
+
+A related class of formal models holds that fraud operates as a distortion of the signal of the incumbent's true strength.
+In this type of model, such as those developed by Little [-@littleFraudMonitoringNoncompetitive2015] or Egorov and Sonin [-@Egorov2021], fraud may be used by incumbents to bluff their opponents into being more supportive than they would be if they knew the incumbent's true popularity.
+However, greater transparency makes such bluffs more costly.
+The typical result in these models is that protest risk can *partially* deter fraud.
+In equilibrium, the incumbent will generally choose the level of fraud anticipated by the opposition and no more\textemdash but also no less.
+
+Fundamentally, the set of models which assume that protest is a centralized, strategic decision by opposition groups predict that the highest risk of protest should obtain when incumbents use massive fraud to generate a low incumbent vote-share, indicating that the incumbent's true popularity is very low.
+Conversely, they predict that protest should be least likely when the incumbent's vote-share is high and manipulation low, indicating a high level of true popularity that deters opposition protest.
+
+The models described above conceive of opposition protest to election manipulation as a strategic action undertaken because the use of more manipulation increases the likelihood that the opposition will have support in the streets (due to the incumbent's low popularity).
 These are normatively optimistic models upon which a sizable literature has been founded: fraud can be kept in check, and ultimately reduced, through the development institutions that reveal information about election integrity: election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].
 
-Another school of thought holds that electoral protest is driven by citizens' grievances rather than by a strategic, organized opposition [@Brancati2016].
+A second set of models shares the assumption that protest is a strategic action motivated by how confident opposition groups are they will succeed in a conflict with the incumbent.
+It differs, however, in its conceptualization of the information revealed by election manipulation.
+Most persuasively articulated by Simpser [-@simpser2013], this view holds that election manipulation demonstrates incumbent strength.
+As he writes, an incumbent that demonstrates the ability to engage in widespread manipulation "is also likely to
+have the resources, capacities, and inclinations to overcome or punish opponents, reward supporters, and circumvent the law" [@simpser2013, p.
+6].
+A strong incumbent (in the language used by the first set of models, i.e.
+one that is relatively popular) may nonetheless manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver].
+By demonstrating otherwise difficult-to-observe capabilities, large-scale election manipulation can solve a coordination problem for the incumbent, securing the cooperation of elites [@simpser2013].
+This can, in turn, fix expectations that the incumbent will continue to rule in the future, further enhancing its staying power [@hale_patronal_2014].
+This argument is in keeping with Seeberg's [-@Seeberg2019] finding that higher-capacity states are associated with more election fraud.
+
+Since these views hold that large-scale election manipulation can reveal information about sources of incumbent strength other than true popularity, and can have indirect, positive effects on incumbent stability, they are normatively pessimistic.
+They predict that opposition protest will be most likely when incumbent vote-share is low and manipulation is low, indicating that the incumbent lacks both genuine popularity and the resources and capacities needed to manufacture votes.
+By contrast, protest should be least likely to occur, under this framework, when incumbent vote-share and manipulation are both high.
+
+Finally, a third school of thought holds that decisions to engage in electoral protest are non-centralized, driven by citizens' grievances rather than by a strategic, organized opposition [@Brancati2016].
 In this view, election manipulation can act as a shared grievance among individuals who prefer to see the ruling party lose\textemdash especially when it appears that fraud may have been decisive [@kuntz_more_2009].
 Belief in a stolen election is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007], possibly initiating a protest cascade [@Kuran1991].
-Just as above, these models tend to predict greater protest risk when margins of victory are low, suggesting a stolen election, and/or grievances are high in the presence of fraud.
+These models thus tend to predict that protests are most likely when margins of victory are low and election manipulation high, suggesting both a high level of grievance in society and a "stolen" election.
+This framework predicts protest is least likely when incumbent vote-share is high and manipulation low (i.e.
+low grievances and low likelihood of a stolen election).
 
@@
 
-The grievance-based approach suggests that protests hinge more on citizens' subjective beliefs about fraud and the broader social context than on the actual level of malfeasance.
-A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b].
-Subjective belief in fraud was also shown to be correlated with approval of mass protests after Russia's 2011 election [@Chaisty2013].
-Furthermore, simply being on the losing side in an election is associated with reduced belief that the election was fair [@Cantu2015], and with increased support for electoral protest [@Anderson2006].
-These studies imply only a limited deterrent effect of protest on fraud in non-democracies.
-If the incumbent may pay the cost of protest whether or not it engages in fraud (due to subjective, and perhaps mistaken, beliefs about fraud and economic conditions), it may as well engage in manipulation in order to gain the benefits.
+Table 1 summarizes these predictions.
+As can be seen in the table, if we focus only on protest occurrence as a dependent variable, we cannot distinguish between two possible data-generating processes: one in which electoral protest is strategic and manipulation signals weakness, and one in which electoral protest is largely un-centralized but manipulation indicates regime strengths.
+Which of these data-generating processes predominates is of significant theoretical importance of understanding authoritarianism and democratization, since the latter view suggests much dimmer prospects for democratization than the former.
 
-This implication connects to a second class of formal models, which hold that fraud operates as a distortion of the signal of the incumbent's true strength.
-In this type of model, such as those developed by Little [-@littleFraudMonitoringNoncompetitive2015] or Egorov and Sonin [-@Egorov2021], fraud may be used by incumbents to bluff their opponents into being more supportive than they would be if they knew the incumbent's true popularity.
-However, greater transparency makes such bluffs more costly.
-The typical result in these models is that protest risk can *partially* deter fraud.
-In equilibrium, the incumbent will generally choose the level of fraud anticipated by the opposition and no more\textemdash but also no less.
+This ambiguity may help account for the inconsistent relationship between election manipulation and protest occurrence found in the literature.
+Several cross-national studies have found links between election manipulation and binary measures of protest [@Brancati2016;@Harvey2020].
+For instance, Hyde and Marinov [-@Hyde2014] use the NELDA dataset to show that protest is more likely when international observers "seriously questioned the winner of the election or the legitimacy of the process," across 758 elections (p.
+343).
+While that study did not include data on incumbents' vote-shares, such a measure of manipulation comes closest to the 'stolen' election framework.
+A study of African elections from 1997 to 2009 shows that election fraud is associated with an increased number of post-election conflicts [@daxecker2012].
+Similar results are found by Rød [-@Rod2019], using a different measure of manipulation.
+Using an alternative measure of manipulation from the NELDA dataset, however, Hafner-Burton et al [-@Hafner-Burton2016] find that the association between election fraud and post-election protest falls short of conventional statistical significance.
+Studies of one archetypal example, the 2011 election and protest cycle in Russia, have found a positive relationship between fraud and protest by some measures [@Lankina2017], and no relationship by others [@ananyevDictatorsSignalStrength2022].
+As noted earlier, Simpser finds a negative relationship between the severity of election manipulation and the likelihood of protest [@simpser2013] and a positive relationship between 'excessive manipulation' (i.e.
+extensive manipulation and a large incumbent vote-share) and regime tenure.
 
-Finally, there is a third school of thought, which holds that election manipulation may indicate regime strength, especially if manipulation is blatant or produces large margins of victory [@simpser2013].
-In the next section, I expand on this school of thought in more detail.
-
-There is a general, though not complete, consensus in empirical work that election fraud is associated with increased risk of protest.
-Several cross-national studies have found links between election manipulation and binary measures of protest [@Brancati2016;@Hyde2014;@Harvey2020;@Rod2019].
-A study of African elections from 1997 to 2009 shows that election fraud is associated with an increased number of post-election conflicts [@daxecker2012].
-However, Hafner-Burton et al [-@Hafner-Burton2016] find that the positive association between election fraud and post-election protest falls short of conventional statistical significance.
-Studies of one archetypal example, the 2011 election and protest cycle in Russia, have found a positive relationship between fraud and protest by some measures [@Lankina2017], and no relationship by others [@ananyevDictatorsSignalStrength2022].
-Taking an alternative approach, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e.
-whether opponents are allowed to run), rather than by the election-day manipulation that usually figures in theoretical work.
-This result is consistent with the finding that parties are more likely to protest elections after ex ante changes to the rules [@chernykh2014].
-
-While these works have identified a connection between fraud and protest, in general, they cannot distinguish between the competing assumptions of the two theories discussed here, or the third possible approach discussed in below.
-For instance, all three approaches predict that protest should be most likely when evidence of fraud is high and the incumbent's vote-share is low (or, relatedly, grievances are high).
-Such a scenario could occur due to 1) strategic protest against an incumbent revealed to be weak through reliance on fraud, 2) mass-led protest against a 'stolen' election, or 3) risky strategic protest against a low-popularity incumbent despite institutional reservoirs of strength.
-The implications of all three scenarios are substantively distinct, however.
-To untangle them, it is necessary to test the assumptions underlying the different models.
+Since the empirical record is mixed, and two theories make similar predictions with regard to protest initiaiton, it is necessary to generate new hypotheses that can distinguish between the three models of election manipulation and protest that have been advanced.
 In particular, do post-election protest movements encounter weak incumbents as signaling models predict?
@@
 
-<!--  One critique of this view, although still within the information/strategic framework, comes from Simpser's [-@Simpser2011] pathbreaking book.
-Simpser argues that when wanipulation efforts are excessive---well beyond what is needed simply to win---they demonstrate the organizational capacity, resources, and manpower of the ruling party [@Simpser2013].
-Large-scale manipulation is, thus, a signal of strength rather than weakness, and can deter protest.
-By contrast, partial manipulation efforts that fail to secure a large majority vote-share signals the government lacks both popular support and mobilizational capacity, making opposition protest more likely.
--->
-
 # Theory
 
-As discussed above, one view of election fraud and protest views fraud as a tool of weak incumbents.
-Another view, associated with Simpser [-@simpser2013], holds that election fraud demonstrates incumbent strength.
-In Simpser's original account, we should expect electoral manipulation to signal strength when it is readily observable or excessive, producing large margins of victory.
-A strong government may manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver].
-This argument is in keeping with Seeberg's [-@Seeberg2019] finding that higher-capacity states are weakly associated with more election fraud.
+Along with the political benefits it may provide, election manipulation is generally characterized as having both operational costs (which the incumbent may be unwilling or unable to pay [@Greene2010]) and legitimacy costs which reduce other actors' level of cooperation with the incumbent [@birch2011].
+The most visible, and arguably most studied, form of such legitimacy costs is mass post-election protest against the incumbent.
+Here, I follow the broad consensus in the literature that authoritarian incumbents find protest undesirable, since they entail some risk of incumbent defeat and may require costly repression or concessions.
+I assume, as a result, that incumbents seek to demobilize protest; protest movements that survive longer indicate the incumbent is relatively weaker compared to the opposition.
+By contrast, when protest movements meet a quick demise, the incumbent can be considered relatively strong.^[In both cases, this assume that incumbent survives the protest movement, by far the modal outcome.
+When the opposition wins, an outcome considered in some of the empirical models below, the assumption is reversed: a quicker opposition victory indicates a weaker incumbent.] How might election manipulation be connected to the incumbent's ability to quickly defeat a protest movement?
 
-I build on Simpser's logic, and on subsequent research, by arguing that that extensive election fraud itself requires the use of resources that are sources of incumbent strength.
-Simpser's model is concerned with whether or or not other actors can perceive the regime's signal of strength; hence the focus on excessive or blatant manipulation.
-I argue here that even when large-scale election fraud is marginal (i.e.
-produces narrow margins of victory), it can indicate regime strength independent of popularity.
-This argument  draws on the recent strand of research on election manipulation that emphasizes the principal-agent problems inherent in such campaigns.
-Political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].
+Integrating Simpser's argument with recent research on the principal-agent dynamics inherent to election manipulation provides insights into this question.
+Where Simpser argues that manipulation should be blatant and excessive (that is, producing a large margin of victory) to send signals of strength, the principal-agent framework for understanding manipulation suggests otherwise.
+This research trend emphasizes the degree to which political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].
+Factors that minimize that principal-agent problem have been shown to facilitate greater election manipulation.
+These same factors\textemdash the incumbent's local popularity [@rundlett2016deliver], the incumbent's consolidated control over patronage resources [@Harvey2019@;@hale_patronal_2014], and low risk of punishment for pro-regime lawbreaking [@Harvey2022]\textemdash are also independently useful for suppressing post-election protest.
+Consequently, in an incumbent is able to manufacture a large numbers of votes, this is a signal of resource capacity, control over bureaucratic institutions, and perhaps deep wells of popularity in parts of the country.
+These signals remain independent sources of regime durability, even if the incumbent's margin of victory is low.
 
-To the extent that election results in non-democracies send a signal, it is composed of a mix of three factors [@Pop-Eleches2015]: the incumbent's genuine popularity, its ability to mobilize otherwise reluctant voters (through patronage, clientelism, and coercion), and its ability to control the bureaucratic apparatus (including the criminal and judicial systems) to generate fraud.
-Large-scale election fraud requires that the incumbent have consolidated control over patronage resources in society; this makes cooperation with the incumbent attractive for front-line agents, who are less willing to risk losing access to the incumbent's network [@Harvey2019].
+To the extent that election results in non-democracies send a signal, it is composed of a mix of three factors [@Pop-Eleches2015]: the incumbent's genuine popularity, its ability to mobilize otherwise reluctant voters (through patronage, clientelism, and coercion), and its ability to control the bureaucratic apparatus (including the criminal and judicial systems) to generate fraudulent votes.
+Large-scale election manipulation is enabled by consolidated control over patronage resources in society; this makes cooperation with the incumbent attractive for front-line agents, who are less willing to risk losing access to the incumbent's network [@Harvey2019].
 However, such efforts also require that incumbents be able to shield front-line agents from negative consequences if their illegal efforts are publicly exposed [@Harvey2022].
-In other words, the incumbent must have sufficient control over the police and the courts, such that ordinary individuals expect little punishment for committing criminal election fraud on their behalf.
-Each of these capacities is required to commit large-scale fraud, and both are useful for disarming protest movements.
-To summarize, the more extensive the election-fraud effort, the more successfully the incumbent has solved the associated principal-agent problems, and the more likely it will be able to use those tools to quash protest.
+In other words, the incumbent must have sufficient control over the police, the courts, and the media, such that ordinary individuals expect that punishment for illegal election manipulation is unlikely.
+At the same time, these capacities are useful for disarming protest movements, as discussed below.
 
@@
 Concessions may be aimed at protesting parties themselves [@reuter2015;@mooreRepressionDissentSubstitution2000;@Hafner-Burton2016;@Piven1979] or the more moderate public, in order to keep them on the sidelines of the conflict [@Hummel2019].
-In either case, large-scale election fraud indicates control over considerable patronage resources, which in turn indicates the capacity to engage in such payments (or withhold them from clients in exchange for passive support [@Rosenfeld2017]).
+In either case, large-scale election manipulation indicates control over considerable patronage resources [@@Harvey2019@;@hale_patronal_2014;@greene_why_2007], which in turn indicates the capacity to engage in such payments (or withhold them from clients in exchange for passive support [@Rosenfeld2017]).
 
@@
 Though such crackdowns risk provoking a backlash, they are aimed at demobilizing the opposition by increasing the costs of participation in overt dissent [@DeVogel2022].
-The ability to protect regime agents from any legal consequences for carrying out orders, a requirement for successful election fraud, is also an asset when asking law enforcement agents to carry out possibly violent repression [@MAGALONI_RODRIGUEZ_2020;@stavro_police_2023;@abouharbFactoJudicialIndependence2013].
+The ability to protect regime agents from any legal consequences for carrying out orders, a condition which helps enable illegal election manipulation [@Harvey2022], is also an asset when asking law enforcement agents to carry out possibly violent repression [@MAGALONI_RODRIGUEZ_2020;@stavro_police_2023;@abouharbFactoJudicialIndependence2013].
 
@@
 Elite defections, which can presage real challenges for the incumbent during a crisis, are also rarer in less democratic regions, where election manipulation is likely more severe [@Reuter2019].
-Just as election manipulation has been held to solve a coordination problem for regime opponents, it may also solve a coordination problem for regime agents: when successful, a fraud campaign signals to agents that other agents are likely to stick with the incumbent due to its revealed capacities.
+Just as election manipulation has been held to solve a coordination problem for regime opponents, it may also solve a coordination problem for regime agents: when successful, a campaign to manufacture votes signals to agents that other agents are likely to stick with the incumbent due to its revealed capacities.
 
-This argument\textemdash that successful, large-scale fraud draws on the same resources a state may use to suppress protest\textemdash turns the deterrent logic on its head.
-Instead of signalling weakness, large-scale fraud indicates a well of resilience.
-Consequently, we should expect incumbent regimes to be aggressive in their use of election fraud, especially in periods of uncertainty, in order to reveal their capacity to resolve principal-agent problems and signal resilience despite low popularity or an assertive opposition.
+This argument\textemdash that successful, large-scale election manipulation draws on the same resources a state may use to suppress protest\textemdash turns the deterrent logic on its head.
+Instead of signalling weakness, large-scale fraud indicates a well of resilience: control over resources, impunity for pro-regime lawbreaking, and loyal agents.
+As a result, election outcomes contain information about several possible resources the incumbent may draw upon to defuse protests.
 
-Since the incumbents' claimed vote-share is part of the signal generated by the election, the effect of election fraud should only be considered in conjunction with the margin of victory.
-A large margin of victory combined with election fraud sends an unambiguous signal of ruling-party strength, with uncertainty around its components.
-That is, large margins of victory can be composed of a number of values of 'natural' popularity, mobilized voters, and election fraud.
-This makes large margins of victory difficult signals for opposition actors to interpret.
-As margins of victory decline, the relative value of the signal becomes clearer.
-A narrow margin of victory combined with extensive fraud is a clear signal of low 'natural' popularity.
-While this may indicate an opportune moment for opposition actors, the large extent of fraud is nevertheless itself an independent indicator of regime strength.
-The combination of patronage resources and control over the criminal and judicial systems required for mass fraud can be turned to coopting or repressing protest movements.
-This argument is the mirror-image of the one present in Trinh [-@trinhTeaLeafElections2022], who argues that when incumbents manipulate heavily, they limit the quality of information that they can extract from the election.
-By the same token, a heavily manipulated election also confounds the signal received by opposition actors; a strong, fraudulent result is a signal of regime stability whose exact composition is uncertain.
+Prior models of electoral protest by strategic opposition actors treat election manipulation and genuine votes as substitute goods.
+For any given vote-share, an increase in manufactured votes reduces the number of genuine votes.
+For classic models, which consider manufactured votes to be merely illusions of strength, the higher the number of manufactured votes for any given vote total, the weaker the incumbent.
+A simplified version of this argument is presented in Figure \@ref(fig:classic).
+In this figure and subsequent one, dashed objects indicate variables that actors must estimate, solid items are observed, and ellipses indicate strategic actions.
+As the figure shows, opposition actors must estimate the proportion of manipulated versus genuine votes in the election outcome.
+The incumbent's genuine popularity drives both in the classic model, increasing genuine votes and reducing the need for manipulation.
+Since genuine popularity reduces the likelihood of a positive protest outcome for the opposition, genuine votes are taken as signals of incumbent strength.
+Manipulated votes are empty in the classic model; subtracting them from the incumbent's vote-total gives opposition an actors their estimate of the incumbent's ability to win a post-election standoff.
+
+Thus the predicted relationship between manipulation and protest likelihood in the classic model is negative: more manipulation makes protest more likely.
+It is possible that that the size of the relationship is conditional on the incumbent's vote-share, however.
+If the incumbent's total vote-share is 70 percent, of which the opposition estimates 15 percent to be manufactured, the incumbent's genuine popularity is estimated at a respectable 55 percent.
+By contrast, if the incumbent posts a 50 percent margin of victory with the same 15 percent estimated to be manufactured, the opposition may perceive that the incumbent is weak in its genuine support (35 percent).
+This produces a first, baseline hypothesis, derived from the classic model.
+
+> H1 (Strategic protest when manipulation signals weakness OR non-strategic protest): Election manipulation is positively associated with protest initiation, with the size of the effect increasing as incumbent vote-share declines.
+
+It is important to note that the same hypothesis is also generated by non-centralized, non-strategic models of election protest.
+The mechanism, however, is different.
+In non-strategic models, protest emerges organically among disaffected citizens.
+Here, the combination of low incumbent vote-share and high manipulation is indicative of high grievances and a 'stolen' election.
+These are conditions that should be most likely to produce protest [@tucker2007].
+
+```{r classic, echo=FALSE, fig.cap="A classic theory of election manipulation and protest", out.width="90%", fig.align='center'}
+knitr::include_graphics("classic_theory.drawio.png")
+
+```
+
+Alternatively, if strategic opposition groups recognize that large-scale election manipulation conveys information about regime durability, they should be deterred from initiating protest as the level of manipulation increases.
+This effect should be magnified by increasing vote-share, which suggests greater genuine popularity at any given level of manipulation.
+
+> H2 (Strategic protest when manipulation signals strength): Election manipulation is negatively associated with protest initiation, with the size of the effect increasing as incumbent vote-share increases
+
+Similarly, since the classic model holds that manipulated votes communicate little about incumbents' post-election strength, it implies that when incumbents rely heavily on election manipulation they will struggle to put down post-election protests.
+This effect should likewise be conditional on incumbent vote-share, since those incumbents who produce small margins of victory while relying heavily on manipulation will have limited ability to push back successfully against protest movements.
+Consequently, more election manipulation should be associated with longer-lived protest movements against surviving incumbents.^[In the appendix, I check the separate prediction that manipulation should be positively associated with incumbents being ousted by protest movements.]
+
+> H3 (manipulation signals weakness): Election manipulation is positively associated with protest movement duration, with the size of this effect decreasing with incumbent vote-share.
+
+```{r newtheory, echo=FALSE, fig.cap="Theory diagram", out.width="90%", fig.align='center'}
+knitr::include_graphics("theory_diagram.drawio.png")
+
+```
+
+Lastly, if election manipulation is an indicator of underlying capacities the incumbent may use to defeat protest movements, this theory predicts that more severe electoral manipulation will be associated with shorter-lived protest waves.
+Incumbents with the resources and institutions needed to produce widespread election manipulation can draw on those same capacities to buy off, repress, or otherwise quash post-election protest.
+
+> H4 (manipulation signals strength): Among incumbents that survive post-election protest movements, more extensive election fraud should be associated with shorter protest waves.
+
+Figure \@ref(fig:newtheory) illustrates how the present argument differs from the classic model.
+Manufactured votes are not negatively associated with genuine popularity, but positively associated with impunity for agents and control over patronage resources.
+Abundant resources and extensive impunity reduce the principal-agent problems that frustrate election manipulation efforts.
+These same resources, institutions, and loyal agents make protest outcomes less likely to be favorable for the opposition.
+As such, opposition actors interpret manufactured and genuine votes as information about separate channels of incumbent strength.
+This produces hypotheses opposite to those of the classic model.
+Manipulation should be associated with reduced likelihood of protest, since it is a signal of capabilities that make protest success less likely.
+This effect should increase as margins of victory increase for the incumbent, providing additional, distinct information about incumbent resilience.
 
@@
 The election outcome sends an ambiguous signal of the incumbents' resources\textemdash including natural popularity, resource control, and control over the legal apparatus\textemdash to possible regime opponents.
-Strategic opposition actors, which may include both organized groups and individuals weighing costs and benefits of participation, interpret this signal with some prior expectation of the incumbent's capacity to manipulate.
-However, even when the incumbent's popularity is low, its ability to engage in election fraud is indicative of resources that can be used to terminate protest movements.
-
-This theory makes predictions for protest initiation, protest duration, and the level of fraud employed.
-First, if large-scale election fraud indicates strength, we should expect that incumbents will respond to pre-election signs of opposition strength by increasing the level of fraud in order to send a more forceful signal.
-This runs directly counter to the prediction of the classic models, which hold that a greater risk of post-election protest should deter fraud.
-
-> H1: Pre-election environments that are more uncertain for the incumbent will be associated with increased election fraud.
-
-Likewise, in an election where the incumbent claims victory, protest should be most likely when levels of fraud are high and claimed vote-shares are low.
-A low vote-share for the incumbent combined with large-scale fraud indicates that the incumbent's 'natural' popularity is very low.
-As claimed vote-share increases for a given level of fraud, the incumbent appears better supported and harder to challenge.^[Of course, the strongest signal of weakness would be one where vote-share and fraud are low, but in this case it is likely that the incumbent has accepted an election loss, rather than falsify a victory.
-This scenario is explored more in the appendix.] This hypothesis is consistent across all three schools of thought, but should nevertheless be evaluated.
-
-> H2: The likelihood of protest initiation will be highest when incumbent vote-share and election fraud are low.
-
-Finally, if large-scale protest is indicative of regime resiliency, we should expect that election fraud will be associated with reduced protest duration among those incumbents who successfully fend off protest movements.
-That is, among strong incumbents who were capable of defeating post-election protest movements, the resources associated with large-scale fraud should enable them to more swiftly quell the movement.
-Those who lack such resources are expected to take longer to dispatch the protest movement.
-Again, these expectations are counter to the standard protest-deterrent logic.
-
-> H3: Among incumbents that survive post-election protest movements, more extensive election fraud should be associated with shorter protest waves.
+The more successfully the incumbent has solved the principal-agent problems associated with election manipulation, however, the more likely it will be able to use those tools to quash protest.
+As a result, more extensive manipulation should be associated with shorter-lived protest movements against incumbents.
 
@@
 library(broom.helpers)
+library(kableExtra)
 
@@
 
-nelda2.sub <- nelda2.sub %>% filter(v2x_regime == 1 | regime.1lag == 1) #Includes only authoritarian, add e-dems with | v2x_regime == 2  #Current version includes lagged 1 in order to capture pre-election auth.
+nelda2.sub <- nelda2.sub %>% filter(v2x_regime == 1 | (v2x_regime == 2 & regime.1lag == 1)) #Includes only authoritarian, add e-dems with | v2x_regime == 2  #Current version includes lagged 1 in order to capture pre-election auth.
 
@@
 nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(n.events.post == 0, 0, median.participants))
+
+## Adding other dimensions of manipulation
+
+nelda2.sub <- nelda2.sub %>% mutate(v2elintim.inv = v2elintim * -1)
+nelda2.sub <- nelda2.sub %>% mutate(v2elvotbuy.inv = v2elvotbuy * -1)
+nelda2.sub <- nelda2.sub %>% mutate(emb_control = v2elembaut * -1)
+
+nelda2.sub <- nelda2.sub %>% mutate(illegal_man_index = v2elintim.inv + v2elvotbuy.inv + v2elirreg.inv)
 
@@
 
-Models where the dependent variable is the level of election fraud include data from all cases.
-For the models that focus on the initiation of post-election protest, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the population of interest: protest following an election in which the incumbent won, possibly through reliance on election fraud.^[Excluding cases where incumbents stepped down after losing an election is necessary to avoid conflating pro- and anti-incumbent protest.
+In all models, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the population of interest: protest following an election in which the incumbent won, possibly through reliance on election fraud.^[Excluding cases where incumbents stepped down after losing an election is necessary to avoid conflating pro- and anti-incumbent protest.
 However, this design does influence the interpretation of the results from these models.
-Threats to inference from this section are evaluated in more detail in the discussion and appendix.] For models of protest-wave duration, only cases with a positive number of post-election protests are included.
-The date range is limited by the availability of data on protest from ECAV, as discussed below, which ranges from 1989 to 2012.
-The resulting dataset includes 614 total election-period observations across 97 countries.
+Threats to inference from this section are evaluated in more detail in the discussion and appendix.] The full dataset includes 445 election-year observations across 81 non-democratic countries.
+The date range is limited by the availability of data on protest from ECAV, which ranges from 1989 to 2012.
+
+For models of protest-wave duration, only observations where post-election protests were initiated can be included.
+For these models, therefore, I limit that analysis to only those cases with a positive number of post-election protests.
+Of the 445 elections in which the incumbent claimed victory, only 156 fall into this more limited dataset.
 
 ## Dependent variables and modeling strategies
-
-To capture the severity of *election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than alternative measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], may not mobilize real voters [@Harvey2020], and can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021].
-This measure of election integrity should thus serve as a strong test of the protest-oriented model.
-It also matches the conception of election manipulation used in much of the relevant formal literature.^[For example, election manipulation is treated as post-hoc adjustment of the results in Little et al [-@Little2015], Little [-@Little2012], Magaloni [-@magaloni2010], and Fearon [-@fearon_self-enforcing_2011], among others.] The variable is used as a dependent variable in the first set of models that test H1, and as a potential explanatory variable in the subsequent models of protest initiation and duration.
-Two alternative measures are used as robustness checks in the appendix.
 
@@
 Still, approximately one-third of elections experienced some protest, and so protest initiation is modeled using standard logistic regression.
-*Election fraud*, which follows a more normal distribution, is modeled using multilevel models with random intercepts by country.
 Lastly, protest wave duration is modeled using Cox proportional hazards models.
@@
 
-```{r, include=F}
-incl.vars <- nelda2.sub %>% dplyr::select('Judicial independence' = jucon.1lag, 
-                                            'Legislative constraints' = legcon.1lag, 
-                                            'Alternative information' = altinf.1lag,
-                                          'Liberal democracy index' = libdem.1lag,
-                                            'GDP growth rate' = gdpgro.1lag, 
-                                            'Civil society openness' = osorg.1lag, 
-                                            #'Public sector corruption' = pubseccorrup.lag,
-                                            'National party organization' = natlpartyorg.lag, 
-                                             'Election fraud' = v2elirreg.inv, 
-                                            'Presidential election' = presidential,
-                                            'Incumbent term-limited' = inc.termlimit.num,
-                                            'Physical integrity index' = physinteg.1lag,
-                                            'Urbanization' = urban.pct, 
-                                            'Rentier state' = rentierstate,
-                                            'Tax state' =  taxstate,
-                                            'Winning vote-share' = winner.share.upd, 
-                                            'Number of pre-election protests' = n.events.pre,
-                                            'Off-schedule election' = unscheduled_election,
-                                          'Regional protest diffusion' = regional_success_lag,
-                                         "Privatization" =  state.owned.1lag,
-                                         "Professional bureaucracy" = patronage.appts.1lag,
-                                        "EMB autonomy"= v2elembaut)
-#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))
-
-incl.vars <- incl.vars %>% mutate(`Presidential election` = as.numeric(`Presidential election`))
-incl.vars <- incl.vars %>% mutate(`Rentier state` = as.numeric(`Rentier state`))
-incl.vars <- incl.vars %>% mutate(`Tax state` = as.numeric(`Tax state`))
-incl.vars <- incl.vars %>% mutate(`Off-schedule election` = as.numeric(`Off-schedule election`))
-
-```
-
-\singlespace
-
-```{r summary-table, echo=FALSE, message=FALSE, warning=FALSE}
-library(kableExtra)
-datasummary_skim(incl.vars, output = "kableExtra", title = "Table 1: Summary statistics and variable distributions", histogram = F)  %>%
-    kable_styling(font_size = 9, full_width = FALSE)
-```
-\doublespace
-
-```{r, include=FALSE}
-library(ggcorrplot)
-corr <- round(cor(incl.vars, use = "pair"), 2)
-
-ggcorrplot(corr, hc.order = TRUE, type = "lower")
-```
-
 ## Explanatory variables
 
-The first set of models test Hypothesis 1, on the relationship between protest risk and fraud.
-Protest risk is operationalized in three ways.
-First, the *number of pre-election protests* in ECAV is taken as *prima facie* evidence of a more turbulent and risky election environment.
-Given that pre-election protest is likely to be a strong predictor of post-election protest\textemdash by indicating strong social networks [@mateo2022], pre-existing opposition groups and civil society capacity [@Bunce2006], and/or popular grievances [@Brancati2016]\textemdash the deterrent logic suggests that incumbents will pull back on election-day fraud when the pre-election environment is characterized by protest [@Rozenas2016b].
-On the other hand, H1 predicts that incumbents will generate more manipulation in this context.
-
-Second, since a state with greater capacity for repression may be better able to deter protest and thus 'get away with' more manipulation [@Egorov2021,@Ritter2016], I measure repressive capacity using the *physical integrity* index from V-Dem, lagged one year.
-Higher values for this variable indicate greater freedom from torture and political killing.
-In keeping with the 'murder in the middle' framework [@feinLifeIntegrityViolationsDemocracy1995], the highest protest risks (and thus most reduced fraud) might be found at middle values of repression; regimes with the most demonstrated repressive capacity deter potential challengers from taking to the streets in the first place, and the least repressive states defuse protest through accommodation [@pierskallaProtestDeterrenceEscalation2010;@mooreRepressionDissentSubstitution2000].
-I thus include the variable's square term to account for the 'murder in the middle' framework.
-
-Finally, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby [@Bunce2006a] (see Brancati and Lucardi [-@Brancati2019] for an alternative view, however).
-This diffusion mechanism is modeled using each country's geopolitical region as coded in Teorell et al [-@qog_teorell] in conjunction with binary variables indicating a successful protest movement from NELDA.
-A 'successful protest' is recorded if a protest movement resulted in either a new election being run or the incumbent stepping down.
-The *regional diffusion* dummy variable takes on a value of one for all countries in the same region as a successful protest, for elections within one calendar year of the successfully overturned election.^[For example, the variable takes on a positive value for elections held in the post-communist region, within one year of the September 24, 2000 election in Serbia that sparked the 'Bulldozer Revolution'.
-All told there are 15 successful electoral protests in the dataset: Bulgaria 1990, Lesotho 1998, Cote d'Ivoire 2000 and 2010, Peru 2000, Serbia 2000, Georgia 2003, Ukraine 2003-2004, Azerbaijan 2005, Kyrgyzstan 2005, Haiti 2006, and Cameroon 2007.]
-
-The models incorporate several potential explanatory variables, in addition to the *election fraud* variable described above.
-The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance.
-The winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia and Psephos.} Finally, whether or not the chief executive is *term limited* is taken from NELDA.
+To capture the severity of *election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than alternative measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], may not mobilize real voters [@Harvey2020], and can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021].
+This measure of election integrity should thus serve as a strong test of the protest-oriented model.
+It also matches the conception of election manipulation used in much of the relevant formal literature.^[For example, election manipulation is treated as post-hoc adjustment of the results in Little et al [-@Little2015], Little [-@Little2012], Magaloni [-@magaloni2010], and Fearon [-@fearon_self-enforcing_2011], among others.] Two alternative measures are used as robustness checks in the appendix.
 
 ## Control variables
+
+The models incorporate several potential confounding variables, in addition to the *election fraud* variable described above.
+The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance.
+The winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia, where links to official sources were available, and Psephos.} Finally, whether or not the chief executive is *term limited* is taken from NELDA.
 
@@
 
-Table \@ref(tab:summary-table) shows summary statistics for all of these variables.
-To address concerns about multicollinearity, I present models in the appendix that exclude control variables associated with pre-election protest.
-
-\newpage
-
-\singlespace
-
-```{r manipdv, echo=FALSE}
-
-m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag +  v2elembaut + (1 | stateid), data = nelda2.sub)
-#summary(m1a)
-
-m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag  + unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)
-
-#summary(m1b)
-
-m1c <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub)
-#summary(m1c)
-
-m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub)
-
-#summary(m1)
-
-m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag + v2elembaut +  (1 | stateid), data = nelda2.sub)
-#summary(m1a)
-
-m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag +  unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)
-
-#summary(m1b)
-
-m1d <- lmer(v2elirreg.inv ~  libdem.1lag + gdpgro.1lag +    presidential +  urban.pct + rentierstate + taxstate  +  unscheduled_election + n.events.pre  + state.owned.1lag + patronage.appts.1lag + v2elembaut +  (1 | stateid), data = nelda2.sub)
-#summary(m1d)
-
-modelsummary(c(m1, m1a, m1c), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Table 2: Random-intercept multilevel model of election-day fraud', coef_omit = "SD", coef_rename = 
-               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
-                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
-                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
-                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
-                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
-                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
-                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
-                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
-integrity index (lag)",
-                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "state.owned.1lag" = "Privatization",
-                 "patronage.appts.1lag" = "Professional bureaucracy",
-                 "v2elembaut" = "EMB autonomy"
-                 )) %>% kable_styling(latex_options = "HOLD_position")
-
-##Note: state owned variable is inverse coded, so should be labeled as private ownership in economy
-##Note: Patronage appointments ialso inverse coded. Higher values = more professionalism
-```
-
-\newpage
-
-\doublespace
+Summary statistics for all of these variables are shown in the appendix.
+To address concerns about multicollinearity, I also present models in the appendix that exclude control variables associated with pre-election protest.
 
@@
 
-## Election-day fraud and protest risk
-
-The results of the first set of multilevel models are not promising for argument that protest risk deters election manipulation.
-Hypothesis 1 is supported in the models shown in Table  \@ref(tab:manipdv).
-Models 1 through 3 apply different measures of pre-election protest risk; all of them behave contrary to the standard expectation in the literature.
-In Model 1, the number of pre-election protests is associated with an increase in election-day fraud.
-While the effect is small\textemdash ten pre-election protests only tend to increase election fraud by about six percent of a standard deviation\textemdash it runs counter to the expectations of the deterrence logic.
-A similar effect is found for *diffusion* in Model 2; recent successful protest in a nearby country is also associated with a modest increase in expected fraud.
-This suggests that the level of election manipulation, rather than being deterred by a highly contentious environment, may be deployed to make up for anticipated poor election results and/or to signal organizational strength.
-
-Model 3 shows that fraud first increases with physical integrity, but then declines.
-In other words, fraud is lowest on the extremes of repressive capacity\textemdash where repression is severe and where it is most limited.
-This is contrary to the expectation that a higher risk of protest is associated with reduced fraud.
-Under the expectation that protest risk is highest for these middle values [@pierskallaProtestDeterrenceEscalation2010], H1 is supported by each measure; fraud is higher when protest risk is higher.
-Models with fixed effects by country, in the appendix, show that these findings holds even within cases.
-
 ## Fraud and the likelihood of protest
-
-\singlespace
 
@@
 
-model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd + illegal_man_index  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary)
 
-model.margin.binary.int <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary.int <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary.int)
 
-model.margin.binary_noc <- glm(any_protest ~ jucon.1lag + legcon.1lag +  altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary_noc <- glm(any_protest ~ lji.lag + legcon.1lag +  altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary_noc)
 
-## Plots
+# Plots
 
@@
 
-plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-.05, 2]"), show.data = T)
+plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.32, 2.2]"), show.data = T)
 plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()
@@
 models2 <- list()
-models2[['Model 4']] <- model.margin.binary
-models2[['Model 5']] <- model.margin.binary.int
+models2[['Model 1']] <- model.margin.binary
+models2[['Model 2']] <- model.margin.binary.int
 #models2[['Model 6']] <- model.margin.binary_noc
 
-modelsummary(models2, output = "kableExtra",  stars = T, title = 'Table 3: Logit models of protest occurrence', coef_rename = 
-               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+modelsummary(models2, output = "default",  stars = T, title = 'Table 3: Logit models of protest occurrence', coef_rename = 
+               c("lji.lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                  "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
@@
                  "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity"
-                 )) %>% kable_styling(latex_options = "HOLD_position")
+                 ))
 ```
 
-\newpage
-\doublespace
-
-Table \@ref(tab:logit-table) shows the results for Hypothesis 2, which is expected to be supported across all three schools of thought.
-Model 4 is a base model, where incumbent vote-share and *intentional irregularities* are standalone variables; Model 5 includes their interaction.
-The expected, significant interaction effect is detected in Model 5, and it is in the expected direction: the occurrence of any protest is most likely when incumbent vote-share is low and fraud is high.
-
-Figure \@ref(fig:margin-prob-fig) illustrates just how tenuous this effect is, however.
-The figure shows the predicted probability of post-election protest for two specific values of election fraud, across all values of incumbent vote-share, holding all other variables constant at their means (continuous variables) or modes (categorical variables).
-The values for election fraud represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively.
-The figure highlights that the predicted probability of protest is only statistically significant for these unusual values of fraud at extremely low levels of incumbent vote-share.
-This represents an exceedingly unlikely outcome along both variables.
-As the figure shows, there are only a small number of observations in the range where the predicted probabilities are statistically significant.
-
-While all three schools of thought are consistent with H2, the results here may be taken as circumstantial evidence in favor of the fraud-as-strength model.
-Model 4 shows that fraud alone has no bearing on protest occurrence; while the level of fraud only appears significant when the incumbent's result in the election is very low.
-This is suggestive of the model where opposition actors interpret the result as showing extremely low genuine support for the incumbent, perhaps enough that it outweighs the signal of strength sent by fraud.
-Middle values for vote-share combined with high values of fraud suggest durable regime strength, bringing expected protest close to zero.
-In the alternative view where fraud conveys weakness, we might expect the probability of protest to rise more sharply as incumbent vote-share declines, and to reach higher predicted probabilities.
-This pattern is only suggestive, of course, and for more definitive evidence I look to protest duration.
+Table \@ref(tab:logit-table) shows the results for Hypotheses 1 and 2.
+Model 1 is a base model, where incumbent vote-share and *intentional irregularities* are standalone variables.
+In keeping with the ambiguous findings in the prior literature, there is no significant relationship between fraud severity and the likelihood of protest initiation.
+The incumbent's claimed margin of victory is negatively associated with protest initiation, though this relationship falls short of conventional statistical significance (p = .07).
+Several control variables behave as expected; for instance, protests are more likely to occur after presidential elections and when there have been prior electoral protests in the region.
+These relationships are largely consistent in Model 2, the key model which includes the interaction between *intentional irregularities* and incumbent vote-share.
+The interaction is significant, as Figure \@ref(fig:margin-prob-fig) illustrates.
+The direction of the effect supports Hypothesis 1, but not Hypothesis 2.
+That is, the occurrence of any protest is most likely when incumbent vote-share is low and fraud is high.
+This suggests that either manipulation is a sign of weakness and protest is centralized and strategic, *or* protest is often non-strategic.
+In the latter case, we cannot yet determine whether manipulation is a signal of strength or weakness.
+To make this determination, we turn to the models of protest duration.
 
@@
 library(ggsurvfit)
+
+nelda2.sub <- nelda2.sub %>% mutate(incumbent_strength = ifelse(d.events.num == 0, 92, ifelse(inc.replaced.prot == 1, -1*(92-d.events.num), 92 - d.events.num)))
+
 nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
+nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(inc.replaced.prot == 1, 0, surv_status))
 nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
@@
 summary(cox2)
-p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.05, 2]")) + coord_cartesian(ylim = c(0, 10)) #This is the 10th and 90th quantiles
+p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.33, 2.2]")) #This is the 10th and 90th quantiles
 p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()
@@
 In such models, coefficients with a positive sign indicate factors that are associated with shorter-lived protests; those with a negative sign indicate factors associated with a slower demise.
-The coefficient for *election fraud* is positive and significant in Model 7, indicating that more fraudulent elections are associated with less resilient protest movements.
-Exponentiation of the coefficient for election fraud gives a hazard ratio of 3.3, indicating that for each one-unit increase in fraud, a protest movement becomes three times more likely to end at any given time.
-This supports H3.
+Models 3 and 5 do not include interactions, while the remaining models interact *intentional irregularities* with incumbent vote-share.
+One threat to inference in this design derives from the 90-day post-election window considered by ECAV.
+It is possible that some protest movements considered 'dead' in the baseline models 3 and 4 are, in fact, 'alive,' if subsequent protests occurred after 90 days and are unobserved.
+To address this, observations are assumed to be right-censored if a protest occurred within three-weeks of the 90-day cutoff in Models 5, 6, and 7.
+This amounts to 15 of the 156 observations in the sample.
+Model 7 is the most restrictive model, which includes right censoring as well as a control variable for the median number of participants during the protest wave.
+There is considerable missingness in this variable.
 
-One threat to inference in this design derives from the 90-day post-election window considered by ECAV.
-It is possible that some protest movements considered 'dead' in Models 7 and 8 are, in fact, 'alive,' if subsequent protests occurred after 90 days and are unobserved.
-As a test of the robustness of the conclusions, in Models 9 and 10, observations are assumed to be right-censored if a protest occurred within three-weeks of the 90-day cutoff.
-This amounts to 15 of the 156 observations in the sample.
-The results remain consistent with those from the uncensored models.
-While the coefficient on *election fraud* falls slightly short of conventional statistical significance in Model 9, it retains the same size and sign compared to Model 7.
-The interaction effect in Model 10 increases in size, while retaining its direction and statistical significance.
-This suggests that the main models are not drawing overly incorrect inferences through censoring of longer-lasting protest movements.
-Model 11 adds an additional control, the median number of participants at events in the protest wave.
-Larger protests might indicate relative opposition strength, however, since there is considerable missingness in the variable, I use it as a robustness check.
-Including this control variable reduces the size of the interaction coefficient, but it retains its direction and statistical significance.
+The coefficient for *election fraud* is positive and significant in Model 3, indicating that more fraudulent elections are associated with less resilient protest movements.
+Exponentiation of the coefficient for election fraud gives a hazard ratio of 1.3, indicating that for each one-unit increase in fraud, a protest movement becomes thirty percent more likely to end at any given time.
+This supports H4 and fails to support H3, suggesting that incumbents with the capacity to manipulate extensively are also able to draw on their resources to curtail post-election protest.
 
 Further investigation of the relationship between fraud and protest duration yields additional insights.
-In Model 8, incumbent vote-share is interacted with *election fraud*.
+In Model 4, incumbent vote-share is interacted with *election fraud*.
 Figure \@ref(fig:margin-dur-fig) shows the marginal effect.
@@
 This result also supports the fraud-as-strength model, and shows that only as vote-share increases above fifty percent\textemdash creating a stronger signal of popularity\textemdash does a government in a low-fraud election have the same likelihood of defeating a protest movement at a given time.
+
+The results of the baseline models are consistent with those models which incorporate censoring.
+While the coefficient on *election fraud* falls slightly short of conventional statistical significance in Model 5 (p = .059), it retains the same size and sign compared to Model 3.
+The interaction effect in Model 6 increases in size, while retaining its direction and statistical significance.
+This suggests that the main models are not drawing overly incorrect inferences through censoring of longer-lasting protest movements.
+Model 7, adds the median number of participants at events in the protest wave as a control.
+Larger protests might indicate relative opposition strength, however, since there is considerable missingness in the variable, I use it as a robustness check.
+Including this control variable reduces the size of the interaction coefficient, but it retains its direction and statistical significance.
 
@@
 models3 <- list()
-models3[['Model 7']] <- cox1
-models3[['Model 8']] <- cox2
-models3[['Model 9']] <- cox1.cen
-models3[['Model 10']] <- cox2.cen
-models3[['Model 11']] <- cox3.cen
+models3[['Model 3']] <- cox1
+models3[['Model 4']] <- cox2
+models3[['Model 5']] <- cox1.cen
+models3[['Model 6']] <- cox2.cen
+models3[['Model 7']] <- cox3.cen
 
 modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Table 4: Cox proportional hazard models of protest duration', coef_rename = 
-               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+               c("lji.lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                  "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
@@
 
-\doublespace
-
 # Discussion
 
-The results can be summarized as follows.
-Pre-existing risk to incumbents is positively, rather than negatively, associated with election fraud.
-The severity of election fraud itself is not associated with protest initiation, except when combined with very low vote-shares for the incumbent.
-Finally, incumbents who have employed large-scale fraud appear better able to put an end to those protest movements that do emerge, compared to those who have presided over cleaner elections.
-This is true even when incumbent vote-shares are low.
-Altogether, the findings are a substantial challenge to several prior assumptions and findings in the literature on election fraud, non-democratic elections, and authoritarian stability.
+The duration analysis consistently shows that more election manipulation is associated with quicker incumbent victories over protest movements, both independently and in conjunction with low incumbent vote-share.
+This is a strong indication that governments which can deliver large-scale election manipulation have the resources and capabilities to put down protest more quickly; that is, such manipulation ought to convey information about incumbent strength.
 
-The results strongly contradict the assumptions of models which hold that election fraud indicates regime weakness.
-Instead, across the three tests here, fraud appears to be an indicator of regime resilience.
-This is a substantively important distinction.
-If fraud indicates weakness, incumbents that would lose fair elections can be deterred from stealing the election by the risk of post-election protest.
-Such models predict that a democratic exchange of power can occur in such scenarios, backed up by the risk of protest.
-The results here would appear to cast doubt on those predictions.
-
-These results are most consistent with the view that even election fraud can be an indicator of regime strength.
-Where this paper goes beyond the traditional formulation of this view\textemdash apart from testing the underlying assumptions\textemdash is the observation that large-scale fraud can be indicative of incumbent strength even when it is not excessive.
-That is, even when ruling party vote-shares are low (perhaps especially when this is so), incumbents which showed themselves capable of large-scale fraud are more capable of ending protest movements than low-fraud incumbents.
+These results are most consistent with the view that election fraud can be an indicator of regime strength.
+This finding supports and extends the line of research initiated by Simpser [@simpser2013].
+While that line of research holds that electoral manipulation in conjunction with a large margin of victory for the incumbent can induce pro-regime behavior in society, this paper shows that even when such cooperation fails\textemdash when post-election protests occur\textemdash governments that manipulate more are more likely to successfully weather the challenge.
+That is, even when ruling party vote-shares are low (perhaps especially when this is so), incumbents which show themselves capable of large-scale fraud are more capable of quickly ending protest movements than low-fraud incumbents.
 This holds true despite controlling for indicators of other possible explanations of the relationship between low fraud and long-lived protest: civil society strength, media openness, and median protest size.
 
-A very brief consideration of the seven cases of successful electoral protest\textemdash Ukraine (2004), Georgia (2003), Kyrgyzstan (2005), Serbia (2000), Haiti (2006) and Cote d'Ivoire (2000 and 2010)\textemdash in comparison with the other cases can also help aid interpretation of the results.
+Adopting this interpretation of election manipulation as a reflection of effective sources of regime strength comports with a growing literature on the principal-agent dynamics of manipulation [@Harvey2019;@rundlett2016deliver] and the importance of patronage resources and electoral processes in sustaining authoritarian regimes [@hale_patronal_2014].
+It also helps us make sense of secular patterns that are puzzling if election fraud indicates regime weakness, as others have held.
+The pattern depicted in Figure ...
+for example, which shows high levels of manipulation and low levels of protest is sensible in a world where the more powerful and durable incumbents produce greater levels of manipulation, and where incumbents who struggle to produce high levels of manipulation are ones who face resource scarcity, institutional barriers to impunity for their agents, and defections among erstwhile allies.
+It also helps explain why some incumbents manufacture their way to highly implausible margins of victory, reasoning that a demonstration of their extra-electoral power may suggest to allies and bystanders that the incumbent may survive protest despite a low natural popularity.
+Scenarios where incumbents use massive fraud to post margins of victory in the eighty and ninety percentage point range are harder to explain under a worldview where more manipulation indicates deeper regime weakness.
+
+The findings regarding protest initiation are puzzling, however, for any interpretation of regime strength as communicated by manipulation.
+Protest is most likely to be initiated under the same circumstances when it is likely to be defeated quickly, which undermines notions of strategic initiation based on incumbent weakness.
+At the same time, it is very unlikely to be initiated when both manipulation and incumbent vote-shares are low, which confounds the expectations of both views of strategic manipulation.^[If manipulation signals weakness, such an outcome provides an honest signal of the incumbent's low popularity, which should trigger strategic protest.
+If manipulation signals strength, this signal of low genuine popularity is further bolstered by an independent signal of organizational weakness (which should also spark strategic protest).] This finding raises more questions than answers, and it is beyond the scope of this paper to resolve those questions.
+However, taking the results in total suggests that 1) large-scale election manipulation indicates regime strength in most cases, and 2) electoral protest behavior is either non-strategic (not based on the likelihood of winning the contest in the streets) or involves a mix of strategic and non-strategic motives.
+
+```{=latex}
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+% \usepackage[normalem]{ulem}
+% \useunder{\uline}{\ul}{}
+\begin{table}[]
+\begin{tabular}{@{}l|lll|lll|l@{}}
+\toprule
+                                                                                      & \begin{tabular}[c]{@{}l@{}}Successful \\ protests\end{tabular} & \begin{tabular}[c]{@{}l@{}}Failed \\ protests\end{tabular} & \begin{tabular}[c]{@{}l@{}}P-value \\ (successful vs.
+\\ unsuccessful)\end{tabular} & Protested & Unprotested & \begin{tabular}[c]{@{}l@{}}P-value \\ (protested vs.
+\\ unprotested)\end{tabular} & \begin{tabular}[c]{@{}l@{}}SD \\ (full sample)\end{tabular} \\ \midrule
+\begin{tabular}[c]{@{}l@{}}Intent.
+\\ irregularities\end{tabular}                     & 1.55                                                           & 1.25                                                       & 0.26                                                                                & 1.27      & 1.16        & 0.24                                                                              & 0.93                                                        \\
+Inc.
+vote-share                                                                       & 0.50                                                           & 0.59                                                       & 0.08                                                                                & 0.58      & 0.62        & 0.07                                                                              & 0.21                                                        \\
+\begin{tabular}[c]{@{}l@{}}Pre-election \\ protests (n)\end{tabular}                  & 7.29                                                           & 6                                                          & 0.25                                                                                & 6         & 0           & 0                                                                                 & 8.4                                                         \\
+\begin{tabular}[c]{@{}l@{}}Pre-election protests\\ (median participants)\end{tabular} & 3.58                                                           & 3                                                          & 0.1                                                                                 & 3.2       & NA          & NA                                                                                & 0.89                                                        \\
+\begin{tabular}[c]{@{}l@{}}Civil society \\ openness\end{tabular}                     & 0.30                                                           & 0.32                                                       & 0.85                                                                                & 0.32      & 0.14        & 0.09                                                                              & 1.08                                                        \\
+Phys.
+int.
+rights                                                                     & 0.45                                                           & 0.43                                                       & 0.81                                                                                & 0.43      & 0.44        & 0.7                                                                               & 0.22                                                        \\ \bottomrule
+\end{tabular}
+\caption{Group means for key variables}
+\label{tab:group-means}
+\end{table}
+```
+
+A very brief consideration of the seven cases of successful electoral protest\textemdash Ukraine (2004), Georgia (2003), Azerbaijan (2005), Kyrgyzstan (2005), Malaysia (2004), Peru (2000) and Cote d'Ivoire (2000)\textemdash in comparison with the other cases can help aid interpretation of the results.
 Table ...
-[CONSIDER ADDING TABLE] shows the average values for several key variables for three groups: incumbents defeated by protest, incumbents who survived protest, and cases where there was no electoral protest.
-The table shows that defeated incumbents did not hold elections that were noticeably more fraudulent than those in where incumbents survived protest: 1.13 for the survivors to 1.15 for the defeated.
-However, those that were defeated claimed substantially smaller vote-shares: 42 percent compared to 55 percent for the survivors.
-Meanwhile, the unprotested group average 55 percent vote-share and a fraud score of 0.87.
-This supports the view that (relatively) low fraud and high vote-shares are a reliable signal of strength (preventing protest), while high levels of fraud and low incumbent vote-shares indicate that the opposition's popularity might overcome the incumbent's structural advantages in a struggle in the street.
-This course of action is uncertain, however, as the unsuccessful protest group shows.
-Lastly,
+[NUMBER] shows the mean values for several key variables for four groups: incumbents defeated by protest, incumbents who survived protest, all protested elections, and all unprotested elections.
+The table also shows p-values for t-tests comparing the means of the relevant groups.
+While the level of fraud is typically higher in the successful protest group than the unsuccessful group, and in the protested versus unprotested group, these means are not significantly different.
+By contrast, incumbent vote-shares are significantly lower in the successful-protest and protest groups than their counterparts (at the p < .1 level).
+This suggests that strategic opposition protest entrepreneurs may look primarily to the official vote total as a signal of the incumbent's ability to marshal all its resources, with the level of fraud being of secondary strategic concern.
+In essence, they may be gambling that, under pressure from mass protests and facing low genuine popularity, elites and state officials will split from the incumbent.
+In other words, they may hope that by initiating protest when the incumbent has demonstrated low popularity, the cooperation between principals and agents demonstrated by effective fraud will begin to fray [@Hale2019;@satoEliteCoordinationPopular2019].
+Unfortunately, the duration analysis suggests this is often a losing bet.
+
+The table also shows that the average number of pre-election protests is significantly different from protested versus unprotested elections, and that pre-election protests that drew larger crowds were more common in the successful-protest cases than the unsuccessful ones (at the p < .1 level).
+This suggests an explanation for the unexpected finding that protests are unlikely to be initiated when fraud and incumbent vote-shares are both low, involving the interconnection between strategic and non-strategic protest.
+When protest entrepreneurs are considering initiating protest, they must consider how likely it is that ordinary citizens will join them [@buenodemesquitaRegimeChangeRevolutionary2010].
+In the absence of large-scale election manipulation to serve as a coordination device [@kuntz_more_2009;@tucker2007], they may judge that protests will be unsuccessful despite the regime's apparent low popularity.
+As Schedler writes, "Grievances provide moral justifications for collective protest: we need to do something.
+Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304].
+Yet this highlights a dilemma for opposition supporters, since it is exactly these scenarios wherein post-election protest is most likely to fail quickly.
+
+A further glance at Figure \@ref(fig:margin-prob-fig) illustrates how protest initiation may blend strategic and non-strategic motivations.
+The figure shows the predicted probability of post-election protest for two specific values of election fraud, across all values of incumbent vote-share, holding all other variables constant at their means (continuous variables) or modes (categorical variables).
+The values for election fraud represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively.
+The figure highlights that the predicted probability of protest is only statistically significant when comparing unusual values of fraud at extremely low levels of incumbent vote-share.
+This represents an exceedingly unlikely outcome along both variables.
+As the figure shows, there are only a small number of observations in the range where the difference in predicted probabilities is statistically significant.
+This suggests that electoral protests occur, for the most part, at random with respect to incumbent vote-share and indicators of fraud\textemdash indicators that have often been used as measures of incumbent strength in classic models.
+The extremes where significant effects are found would then correspond to strategic, but risky, gambles by protest entrepreneurs hoping that mass manipulation will help pull regime opponents into the streets, and betting that they will not when manipulation is low.
+Thus the results find support for both the strategic and non-strategic literatures on protest.
+
+Figure \@ref(fig:corfig) shows some support for this idea.
+Protest occurrence is very weakly correlated with measures of opposition strength, like the openness of civil society or the absence of barriers to forming independent parties.
+It is also weakly correlated with measures of incumbent strength, like incumbent vote-share and\textemdash as this paper argues\textemdash election fraud.
+Post-election protest is also weakly correlated with possible indicators of mass grievance, like an economic crisis or corruption in the public sector.
+This could indicate that post-election protests hinge on particular combinations of factors that are difficult to predict a priori.
+It may also be the case that such protests are highly contingent.
+While not an electoral protest, the initiation of the Arab Spring protests offer such an illustration.
+[EXPLAIN MORE]
+
+```{r corfig, fig.cap="Correlation coefficients of selected variables", echo=FALSE, message=FALSE, warning=FALSE}
+df <- nelda2.sub %>% filter(inc.lose == "0") %>% dplyr::select(any_protest, osorg.1lag, partybarriers.lag, winner.share.upd, v2elirreg.inv, econ.crisis.nelda.1lag, pubseccorrup.lag, regional_success_lag)
+df <- df %>% mutate(econ.crisis.nelda.1lag = as.numeric(econ.crisis.nelda.1lag))
+df <- df %>% rename("Any protest" = any_protest,
+                    "Civil soc. open." = osorg.1lag,
+                    "Low party barriers" = partybarriers.lag,
+                    "Inc. vote-share" = winner.share.upd,
+                    "Election fraud" = v2elirreg.inv,
+                    "Econ. crisis" = econ.crisis.nelda.1lag,
+                    "Pub. sec. corrupt." = pubseccorrup.lag,
+                    "Reg. protest" = regional_success_lag)
+M <- cor(df, use = "pair")
+p.cor <- corrplot::corrplot(M, type = "upper", tl.col = 'black',
+         cl.ratio = 0.2, tl.srt = 45, addCoef.col = 'grey50')
+
+```
 
@@
 They suggest that factors like the opposition's capacity to organize street protest in limiting election fraud have been overestimated.
-In fact, these results suggest that incumbents who have the resources will respond to such a scenario by *increasing* their production of fraud.
+In fact, these results suggest that incumbents who have the resources will respond to such a scenario by *increasing* their production of fraud to secure a higher vote-share.
 Consequently, theories of electoral manipulation that focus on factors other than protest risk\textemdash such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]\textemdash may be more important for understanding election integrity, which may drive future research as well as policy interventions.
@@
 Scholarly attention might turn more profitably toward other explanations for manipulation, such as the organizational capacity of the state [@simpser2013], the popularity of the government [@rundlett2016deliver], and the risks faced by front-line agents [@Harvey2022].
-
-In turn, this revision to our understanding of election manipulation has implications for broader topics like authoritarian persistence and democratization.
-We observe that protest initiation is most likely when fraud is high and incumbent vote-share is low.
-This could be indicative of strategic behavior by the rational opposition actors found in most formal models, or by less strategic behavior by citizens en masse.
-For such citizens, election manipulation may act as a shared grievance, especially when it appears that fraud may have been decisive [@kuntz_more_2009;@tucker2007].
-A close election marked by fraud intensifies the connection between emotional and strategic protest incentives.
-As Schedler writes, "Grievances provide moral justifications for collective protest: we need to do something.
-Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304].
-Yet this highlights a dilemma for opposition supporters, since it is exactly these scenarios wherein post-election protest is most likely to fail quickly.
 
@@
 
-This paper is primarily concerned with whether or not election manipulation is indicative of regime weakness or resilience; its primary evidence is finding that fraud appears to increase with pre-election uncertainty and that fraud is associated with more resilient incumbents in the face of protest.
+ [MOVE TO APPENDIX? ->]
+This paper is primarily concerned with whether or not election manipulation is indicative of regime weakness or resilience; its primary evidence is the finding that fraud is associated with more resilient incumbents in the face of protest.
 Due to space constraints, two other predictions are not tested in depth here.
@@
 
+For discussion section:
+
 The second question this paper does not closely examine is the role of election fraud in sustaining (or not) the small number of incumbents who are defeated by mass protest.
@@
 
-<!-- This project suggests additional avenues for future research.
-It is clear that electoral protests charged by allegations of manipulation do occur at times.
-While the main argument of this paper is that election fraud itself does not predict these protests, there may still be connections between manipulation and protest to uncover.
-For example, it could be that other forms of manipulation\textemdash perhaps intimidation or violence\textemdash are associated with higher risk of escalating protest.
-Perhaps some forms of electoral manipulation are particularly offensive to certain groups\textemdash young people, highly educated people, people with ideological commitments to democracy, etc.
-To the extent that countries vary according to the size of such populations, they may vary in their underlying protest risk.
--->
-
 These results are consistent with different specifications presented in an online appendix.
@@
 # References
+
+### Leftover text
+
+Fraud itself is a sign of the robustness of some of the same tools that can be used to co-opt the opposition, constrain elites, or intensify repression during a crisis; even in the unlikely event that protests do break out in response to large-scale fraud, the incumbent has many tools it may deploy to curtail them.
+
+In this paper, I argue \textemdash in keeping with a growing literature on the informational aspects of elections in authoritarian regimes [@guriev2020;@smyth2020elections;@Geddes1989;@Miller2015a;@Pop-Eleches2015]\textemdash that an election outcome produces a signal of the relative strength of the incumbent.
+As Magaloni [-@magaloni2006] and others have noted, incumbents typically work to produce a higher margin of victory in the election, in order to demonstrate their strength and induce cooperation from other actors in society.
+In some theories, this works best if the votes for the incumbent are largely 'genuine' (i.e.
+not fabricated), or produced using the incumbent's organizational resources (i.e.
+vote-buying and intimidation) [@Harvey2020;@magaloni2006].
+These two kinds of votes correspond to two of the three 'pillars of stability' identified by Gerschewski [-@Gerschewski2013]: legitimation and cooptation.
+Fraud is considered a form of cheap talk in these models, which opposition actors can discount if they become aware of it [@little_elections_2012].
+
+However, these models overlook the fact that engineering widespread fraud requires solving principal-agent problems through control over patronage resources and the criminal and judicial apparatus [@Harvey2019;@Harvey2022].
+As a result, high levels of fraud are a sign of incumbent strength in both cooptation and repression (Gerschewski's third pillar).
+Both of these tools can be useful for defusing or dismantling an emergent protest movement.
+As a result, effective fraudsters should also be more effective at defeating protest movements.
+Consequently, in this view, there is little penalty for severe manipulation in terms of protest risk, and little deterrent effect.
+
+Testing this proposition in more depth is difficult.
+If, for example, incumbents manipulate just the right amount to avoid mass protest in equilibrium\textemdash and are deterred from manipulating more by that risk\textemdash we would observe no statistical relationship between the severity of fraud and the likelihood of protest initiation.
+High and low levels of fraud would generally be associated with no protest.
+Beginning to untangle the relationship between election fraud and protest requires looking beyond standard binary measures of protest.
+Instead, this paper tests outcomes "off the equilibrium path," to use the language of formal modeling.
+Does greater risk to the incumbent in the pre-election period result in reduced election fraud?
+When protest movements do arise after an election, is fraud associated with weaker governments?
+
+Taking an alternative approach, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e.
+whether opponents are allowed to run), rather than by the election-day manipulation that usually figures in theoretical work.
+This result is consistent with the finding that parties are more likely to protest elections after ex ante changes to the rules [@chernykh2014].
+
+The grievance-based approach suggests that protests hinge more on citizens' subjective beliefs about fraud and the broader social context than on the actual level of malfeasance.
+A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b].
+Subjective belief in fraud was also shown to be correlated with approval of mass protests after Russia's 2011 election [@Chaisty2013].
+Furthermore, simply being on the losing side in an election is associated with reduced belief that the election was fair [@Cantu2015], and with increased support for electoral protest [@Anderson2006].
+These studies imply only a limited deterrent effect of protest on fraud in non-democracies.
+If the incumbent may pay the cost of protest whether or not it engages in fraud (due to subjective, and perhaps mistaken, beliefs about fraud and economic conditions), it may as well engage in manipulation in order to gain the benefits.

### 2024-07-31 Scripts/authoritarian_only_latex_2.Rmd
Edits to condense and clarify

@@
 Mass protest would, in this case, be a fragile guardrail against autocratization.
-This paper argues that this pessimistic model best fits the overall pattern of the data, and is supported by specific, novel hypothesis tests.
+This debate cannot be advanced by focusing solely on protest occurrence.
+Instead, this paper contributes by drawing out implications from three strands of the literature on election manipulation and protest, which have (to my knowledge) not previously been explicated.
 
-This paper contributes to this debate by drawing out the implications of three strands of the literature on election manipulation and protest.
 In short: there are two views on the nature of electoral protest which make the same prediction.
 If opposition groups strategically initiate protest when they believe the incumbent to be weak, they will be most likely to do so when the election results show a low vote-share for the incumbent combined with high levels of manipulation\textemdash a combination of factors which reveals the incumbent's very low genuine popularity.
-However, if protest movements are more organic and non-strategic, reflecting citizen grievances, we should expect protest initiation in the same circumstances.
-Here the combination of low incumbent vote-share and high manipulation indicates severe citizen grievances (low natural popularity) which are galvanized by the shared grievance of a 'stolen' election.
-Since both theories make the same prediction about protest initiation, such tests\textemdash which have been the dominant empirical measure in the literature\textemdash do not reveal whether protest is strategic and manipulation indicates weakness or manipulation indicates sources of strength and anti-manipulation protest is non-strategic.
+[NOTE: ADD CITES HERE] However, if protest movements are more organic and non-strategic, reflecting citizen grievances, we should expect protest initiation in the same circumstances.
+Here the combination of low incumbent vote-share and high manipulation indicates severe citizen grievances (low natural popularity) which are galvanized by the shared grievance of a 'stolen' election [@tucker2007].
+Since both theories make the same prediction about protest initiation, such tests\textemdash which have been the dominant empirical measure in the literature\textemdash do not reveal whether protest is strategic and manipulation indicates weakness or if manipulation is a hallmark of strong incumbents [@simpser2013] and protest is non-strategic.
 
-To unravel this question, I make use of data on the duration of post-election protests.
+To unravel this question, I take this analysis one step further.
+In strategic-protest models, the opposition protests because it believes the probability of a pro-government outcome of the protest is low.
+I assume that the duration of protest movements, when they begin, is an indicator of this underlying likelihood of a pro-government outcome.
+The more quickly a government can put down protests, the greater the underlying (unobserved) distribution of power favors the incumbent over the opposition.
+By contrast, the longer it takes the government to restore control, the less favorable the distribution of power.^[The analysis treats the small number of cases where the opposition succeeds in ousting the government in several ways.
+They are excluded from the main analysis, but incorporated in multiple robustness checks.]
+
+I make use of data on the duration of post-election protests.
 Shorter-lived protest movements that end with the incumbent still in power indicate a government that is relatively strong compared to the protesting opposition, while longer-lived protest movements indicate a more even balance of power between the two.
@@
 # Literature review: three views of election manipulation and mass protest
-
-Three understandings of the relationship between election manipulation and post-election protest have developed in the literature.
-Each view is discussed in detail below, after a brief initial summary here.
-A first view assumes that incumbents and opposition groups are strategic actors who who evaluate the election period for signs of each other's relative likelihood of winning a power struggle.
-Researchers who model election protest this way generally treat the use of election manipulation as a sign of the incumbent's unpopularity, and therefore weakness.
-Protest occurs after fraudulent elections in response to this perceived weakness; this possibility can serve to deter incumbents from engaging in manipulation under some circumstances.
-This is a normatively optimistic theory.
-
-A second worldview argues that election manipulation can indicate a well-resourced, well-organized, powerful incumbent (independent of its popularity).
-Under this view, extensive election manipulation can be a show of strength, which induces cooperation with the incumbent from other societal actors.
-Since this view holds that more manipulation can further entrench the incumbent, it is a normatively pessimistic view.
-
-Finally, others have argued that electoral protest occurs unrelated to strategic concerns about the relative strength of the incumbent and the opposition.
-In these models, electoral protest occurs because manipulation galvanizes citizens who hold diverse grievances, because economic conditions are poor, or simply because holding an election creates a moment of opportunity.
-The central feature of these models is that protest decisions are generally made without consideration for the likelihood that protests will overthrow the incumbent.
-As will be discussed in depth below, these different assumptions yield different predictions for when electoral protest is more likely to be initiated, and how long it will last against an incumbent that refuses to give up power.
-Table 1 presents these predictions for ease of reference.
-
-```{=latex}
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-% \usepackage{multirow}
-\begin{table}[]
-\begin{tabular}{@{}llll@{}}
-\toprule
-                                                                                                              & Manipulation signals: & \multicolumn{1}{c}{Strategic protest}                                                           & \multicolumn{1}{c}{Non-strategic protest}                                                  \\ \midrule
-\multicolumn{1}{l|}{\multirow{2}{*}{Initiation most likely}}                                                  & Weakness              & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Low margin + \\ high manipulation\end{tabular}}  & \multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}Low margin + \\ high manipulation\end{tabular}} \\ \cmidrule(lr){2-3}
-\multicolumn{1}{l|}{}                                                                                         & Strength              & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Low margin + \\ low manipulation\end{tabular}}   &                                                                                            \\ \cmidrule(l){2-4}
-\multicolumn{1}{l|}{\multirow{2}{*}{Initiation least likely}}                                                 & Weakness              & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High margin + \\ low manipulation\end{tabular}}  & \multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}High margin +\\ low manipulation\end{tabular}}  \\ \cmidrule(lr){2-3}
-\multicolumn{1}{l|}{}                                                                                         & Strength              & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High margin + \\ high manipulation\end{tabular}} &                                                                                            \\ \cmidrule(l){2-4}
-\multicolumn{1}{l|}{\multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}Duration when \\ incumbent wins\end{tabular}}} & Weakness              & \multicolumn{2}{c}{More manipulation: Longer protest}                                                                                                                                        \\
-\multicolumn{1}{l|}{}                                                                                         & Strength              & \multicolumn{2}{c}{More manipulation: Shorter protest}                                                                                                                                       \\ \cmidrule(l){2-4}
-\end{tabular}
-\caption{Predictions for protest initiation and duration under varying theoretical assumptions}
-\label{tab:predictions}
-\end{table}
-```
 
@@
 
-Fundamentally, the set of models which assume that protest is a centralized, strategic decision by opposition groups predict that the highest risk of protest should obtain when incumbents use massive fraud to generate a low incumbent vote-share, indicating that the incumbent's true popularity is very low.
-Conversely, they predict that protest should be least likely when the incumbent's vote-share is high and manipulation low, indicating a high level of true popularity that deters opposition protest.
+Fundamentally, both set of models share the assumption that protest is a centralized, strategic decision by opposition groups, and that fraud is a tool employed by weak (i.e.
+unpopular) incumbents.
+As a result, they predict that the highest risk of protest should obtain when incumbents use massive fraud but still produce a low incumbent vote-share, indicating that the incumbent's true popularity is very low.
+This, in turn, indicates to opposition actors that protest is more likely to succeed in unseating the incumbent.
+These are normatively optimistic models upon which a sizable literature has been founded: fraud can be kept in check, and ultimately reduced, through the development of institutions that reveal information about election integrity: election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].
+These institutions improve election integrity, it is claimed, by increasing the likelihood that fraud will be observed and thus, incumbent weakness exposed.
 
-The models described above conceive of opposition protest to election manipulation as a strategic action undertaken because the use of more manipulation increases the likelihood that the opposition will have support in the streets (due to the incumbent's low popularity).
-These are normatively optimistic models upon which a sizable literature has been founded: fraud can be kept in check, and ultimately reduced, through the development institutions that reveal information about election integrity: election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].
+However, an alternative set of models have also been advanced.
+A second worldview argues that election manipulation can indicate a well-resourced, well-organized, powerful incumbent (independent of its popularity).
+Under this view, extensive election manipulation can be a show of strength, which induces cooperation with the incumbent from other societal actors.
+Since this view holds that more manipulation can further entrench the incumbent, it is a normatively pessimistic view.
 
-A second set of models shares the assumption that protest is a strategic action motivated by how confident opposition groups are they will succeed in a conflict with the incumbent.
-It differs, however, in its conceptualization of the information revealed by election manipulation.
 Most persuasively articulated by Simpser [-@simpser2013], this view holds that election manipulation demonstrates incumbent strength.
-As he writes, an incumbent that demonstrates the ability to engage in widespread manipulation "is also likely to
-have the resources, capacities, and inclinations to overcome or punish opponents, reward supporters, and circumvent the law" [@simpser2013, p.
+As he writes, an incumbent that demonstrates the ability to engage in widespread manipulation "is also likely to have the resources, capacities, and inclinations to overcome or punish opponents, reward supporters, and circumvent the law" [@simpser2013, p.
 6].
@@
 
-Since these views hold that large-scale election manipulation can reveal information about sources of incumbent strength other than true popularity, and can have indirect, positive effects on incumbent stability, they are normatively pessimistic.
-They predict that opposition protest will be most likely when incumbent vote-share is low and manipulation is low, indicating that the incumbent lacks both genuine popularity and the resources and capacities needed to manufacture votes.
-By contrast, protest should be least likely to occur, under this framework, when incumbent vote-share and manipulation are both high.
+For the set of theories that treat election manipulation as an indicator of sources of regime strength, manipulation can serve to deter protest under some circumstances.
+[NOTE ADD CITES HERE?] This leaves protest, when it does occur in the presence of large-scale manipulation, as something of a puzzle if protest behavior is largely strategic.
+For this reason, the manipulation-as-strength view comports more easily with another established view of the causes of electoral protest.
+Researchers in this tradition have argued that electoral protest occurs unrelated to strategic concerns about the relative strength of the incumbent and the opposition.
+In these models, electoral protest occurs because manipulation galvanizes citizens who hold diverse grievances, because economic conditions are poor, or simply because the bare fact of holding an election creates a moment of opportunity for opposition organization irrespective of the level of manipulation [@Pop-Eleches2015] [NOTE: ADD CITES].
+The central feature of these models is that protest decisions are generally made without consideration for the likelihood that protests will overthrow the incumbent.
+As will be discussed in depth below, these different assumptions yield different predictions for when electoral protest is more likely to be initiated, and how long it will last against an incumbent that refuses to give up power.
 
-Finally, a third school of thought holds that decisions to engage in electoral protest are non-centralized, driven by citizens' grievances rather than by a strategic, organized opposition [@Brancati2016].
+This school of thought holds that decisions to engage in electoral protest are non-centralized, driven by citizens' grievances rather than by a strategic, organized opposition [@Brancati2016].
 In this view, election manipulation can act as a shared grievance among individuals who prefer to see the ruling party lose\textemdash especially when it appears that fraud may have been decisive [@kuntz_more_2009].
-Belief in a stolen election is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007], possibly initiating a protest cascade [@Kuran1991].
+Belief in a stolen election\textemdash one the incumbent would have lost in the absence of manipulation\textemdash is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007], possibly initiating a protest cascade [@Kuran1991].
 These models thus tend to predict that protests are most likely when margins of victory are low and election manipulation high, suggesting both a high level of grievance in society and a "stolen" election.
-This framework predicts protest is least likely when incumbent vote-share is high and manipulation low (i.e.
-low grievances and low likelihood of a stolen election).
+This framework predicts protest is most likely when incumbent vote-share is low and manipulation high (i.e.
+elevated grievances and a high likelihood of a stolen election).
 
@@
 This argument has found support in studies of the relationship between economic grievances, manipulated elections, and protest [@Rod2019;@Brancati2016].
+
+```{=latex}
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+\begin{tabular}{@{}lllllll@{}}
+\toprule
+                     &              &                           & \multicolumn{2}{l}{Strategic protest; manipulation = weakness} & \multicolumn{2}{l}{Non-strategic protest; manipulation = strength} \\ \midrule
+Incumbent vote-share & Manipulation & True popularity inference & Initiation              & Prob. of pro-gov outcome             & Initiation                & Prob. of pro-gov. outcome              \\
+Low                  & Low          & Low                       & Less likely             & More likely                          & Less likely               & Less likely                            \\
+Low                  & High         & Very low                  & More likely             & Less likely                          & More likely               & More likely                            \\
+Citations            &              &                           &                         &                                      &                           &                                        \\ \bottomrule
+\end{tabular}
+\caption{}
+\label{tab:predictions}
+\end{table}
+```
 
@@
 
-Since the empirical record is mixed, and two theories make similar predictions with regard to protest initiaiton, it is necessary to generate new hypotheses that can distinguish between the three models of election manipulation and protest that have been advanced.
+Since the empirical record is mixed, and two theories make similar predictions with regard to protest initiation, it is necessary to generate new hypotheses that can distinguish between the two models of election manipulation and protest that have been advanced.
 In particular, do post-election protest movements encounter weak incumbents as signaling models predict?
@@
 This research trend emphasizes the degree to which political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].
-Factors that minimize that principal-agent problem have been shown to facilitate greater election manipulation.
-These same factors\textemdash the incumbent's local popularity [@rundlett2016deliver], the incumbent's consolidated control over patronage resources [@Harvey2019@;@hale_patronal_2014], and low risk of punishment for pro-regime lawbreaking [@Harvey2022]\textemdash are also independently useful for suppressing post-election protest.
-Consequently, in an incumbent is able to manufacture a large numbers of votes, this is a signal of resource capacity, control over bureaucratic institutions, and perhaps deep wells of popularity in parts of the country.
-These signals remain independent sources of regime durability, even if the incumbent's margin of victory is low.
-
-To the extent that election results in non-democracies send a signal, it is composed of a mix of three factors [@Pop-Eleches2015]: the incumbent's genuine popularity, its ability to mobilize otherwise reluctant voters (through patronage, clientelism, and coercion), and its ability to control the bureaucratic apparatus (including the criminal and judicial systems) to generate fraudulent votes.
-Large-scale election manipulation is enabled by consolidated control over patronage resources in society; this makes cooperation with the incumbent attractive for front-line agents, who are less willing to risk losing access to the incumbent's network [@Harvey2019].
-However, such efforts also require that incumbents be able to shield front-line agents from negative consequences if their illegal efforts are publicly exposed [@Harvey2022].
-In other words, the incumbent must have sufficient control over the police, the courts, and the media, such that ordinary individuals expect that punishment for illegal election manipulation is unlikely.
-At the same time, these capacities are useful for disarming protest movements, as discussed below.
+Factors that minimize principal-agent problems have been shown to facilitate greater election manipulation.
+These include the incumbent's local popularity [@rundlett2016deliver], the incumbent's ability to distribute patronage resources [@Harvey2019;@hale_patronal_2014], and low risk of punishment for pro-regime lawbreaking [@Harvey2022].
+As discussed below, these are also independently useful for suppressing post-election protest.
+Consequently, if an incumbent is able to manufacture a large numbers of votes, it is also likely to have the resource capacity, control over bureaucratic institutions, and perhaps deep wells of popularity in parts of the country that can serve as independent sources of regime durability, even if the incumbent's nationwide margin of victory is low.
 
@@
 Concessions may be aimed at protesting parties themselves [@reuter2015;@mooreRepressionDissentSubstitution2000;@Hafner-Burton2016;@Piven1979] or the more moderate public, in order to keep them on the sidelines of the conflict [@Hummel2019].
-In either case, large-scale election manipulation indicates control over considerable patronage resources [@@Harvey2019@;@hale_patronal_2014;@greene_why_2007], which in turn indicates the capacity to engage in such payments (or withhold them from clients in exchange for passive support [@Rosenfeld2017]).
+In either case, large-scale election manipulation indicates control over considerable patronage resources [@Harvey2019;@hale_patronal_2014;@greene_why_2007], which in turn indicates the capacity to engage in such payments (or withhold them from clients in exchange for passive support [@Rosenfeld2017]).
 
@@
 Elite defections, which can presage real challenges for the incumbent during a crisis, are also rarer in less democratic regions, where election manipulation is likely more severe [@Reuter2019].
-Just as election manipulation has been held to solve a coordination problem for regime opponents, it may also solve a coordination problem for regime agents: when successful, a campaign to manufacture votes signals to agents that other agents are likely to stick with the incumbent due to its revealed capacities.
+Just as election manipulation has been held to solve a coordination problem for regime opponents, it may also solve a coordination problem for regime agents: when successful, a campaign to manufacture votes signals to agents that other agents are likely to stick with the incumbent due to its revealed capacities [@hale_patronal_2014].
 
@@
 Instead of signalling weakness, large-scale fraud indicates a well of resilience: control over resources, impunity for pro-regime lawbreaking, and loyal agents.
-As a result, election outcomes contain information about several possible resources the incumbent may draw upon to defuse protests.
+As a result, governments that build and deploy an electoral manipulation apparatus are simultaneously innoculating themselves against the cost of protest.
 
@@
 For classic models, which consider manufactured votes to be merely illusions of strength, the higher the number of manufactured votes for any given vote total, the weaker the incumbent.
-A simplified version of this argument is presented in Figure \@ref(fig:classic).
-In this figure and subsequent one, dashed objects indicate variables that actors must estimate, solid items are observed, and ellipses indicate strategic actions.
-As the figure shows, opposition actors must estimate the proportion of manipulated versus genuine votes in the election outcome.
-The incumbent's genuine popularity drives both in the classic model, increasing genuine votes and reducing the need for manipulation.
-Since genuine popularity reduces the likelihood of a positive protest outcome for the opposition, genuine votes are taken as signals of incumbent strength.
-Manipulated votes are empty in the classic model; subtracting them from the incumbent's vote-total gives opposition an actors their estimate of the incumbent's ability to win a post-election standoff.
+This means that strategic opposition actors are more likely to initiate protest when they perceive the incumbent to be weaker, and expect to be more likely to succeed in a challenge.
+A simplified version of this argument is presented in Panel B of Figure \@ref(fig:theoryfig).
+In the figure, dashed objects indicate variables that actors must estimate, while solid items are observed.
 
-Thus the predicted relationship between manipulation and protest likelihood in the classic model is negative: more manipulation makes protest more likely.
-It is possible that that the size of the relationship is conditional on the incumbent's vote-share, however.
-If the incumbent's total vote-share is 70 percent, of which the opposition estimates 15 percent to be manufactured, the incumbent's genuine popularity is estimated at a respectable 55 percent.
+As the figure shows, in the classic model, the incumbent's (unknown) true popularity is positively related to its vote-share in the election, and negatively related to its use of fraud.
+This dynamic is the fulcrum upon which these models hinge: less popular incumbents (who are expected to be less likely to defeat an opposition protest) must manipulate more in order to win elections.
+Popular incumbents need manipulate less.
+Opposition actors gauge the probability of an pro-government protest outcome by estimating the incumbent's true popularity based on the degree of manipulation and the incumbent's vote-share in the election.
+The level of manipulation is, in effect, a mediator for the effect of vote-share.
+A high vote-share with low fraud indicates a more popular incumbent, likely deterring protest.
+A low vote-share with a high degree of fraud indicates a very weak incumbent, making strategic protest more likely.
+In other words, manipulated votes are empty in the classic model; subtracting them from the incumbent's vote-total gives opposition actors their estimate of the incumbent's ability to win a post-election standoff.
+
+Thus the predicted relationship between manipulation and protest likelihood in the classic model is positive: more manipulation makes protest more likely.
+However, as increasing vote-share increases the number of genuine votes for the incumbent at any fixed level of fraud, the effect of fraud should be conditional on the claimed vote-share [NOTE: ADD CITES].
+If the incumbent's total vote-share is 70 percent, for example, of which the opposition estimates 15 percent to be manufactured, the incumbent's genuine popularity is estimated at a respectable 55 percent.
 By contrast, if the incumbent posts a 50 percent margin of victory with the same 15 percent estimated to be manufactured, the opposition may perceive that the incumbent is weak in its genuine support (35 percent).
@@
 
-> H1 (Strategic protest when manipulation signals weakness OR non-strategic protest): Election manipulation is positively associated with protest initiation, with the size of the effect increasing as incumbent vote-share declines.
+> H1: Increasing election fraud makes protest more likely, with the size of this effect decreasing with incumbent vote-share.
 
@@
 These are conditions that should be most likely to produce protest [@tucker2007].
+By contrast, as vote-share increases for a set level of fraud, the incumbent's genuine popularity is higher and thus grievances are lower (reducing the likelihood of protest).
 
-```{r classic, echo=FALSE, fig.cap="A classic theory of election manipulation and protest", out.width="90%", fig.align='center'}
-knitr::include_graphics("classic_theory.drawio.png")
+```{r theoryfig, echo=FALSE, fig.cap="A classic theory of election manipulation and protest", out.width="90%", fig.align='center'}
+knitr::include_graphics("theory_comp.drawio.png")
 
@@
 
-Alternatively, if strategic opposition groups recognize that large-scale election manipulation conveys information about regime durability, they should be deterred from initiating protest as the level of manipulation increases.
-This effect should be magnified by increasing vote-share, which suggests greater genuine popularity at any given level of manipulation.
+The classic model holds that manipulated votes correlate negatively with incumbents' post-election strength, implying that when incumbents rely heavily on election manipulation they will struggle to put down post-election protests.
+This effect should likewise be conditional on incumbent vote-share, since those incumbents who produce small margins of victory while relying heavily on manipulation will have limited ability to push back successfully against protest movements.
+Panel B in Figure [NOTE: FIX REF] illustrates this; a higher likelihood of pro-government protest outcome reduces the likelihood that the opposition will initiate protest.
+Consequently, classic models predict that more election manipulation should be associated with longer-lived protest movements against surviving incumbents.^[In the appendix, I check the separate prediction that manipulation should be positively associated with incumbents being ousted by protest movements.]
 
-> H2 (Strategic protest when manipulation signals strength): Election manipulation is negatively associated with protest initiation, with the size of the effect increasing as incumbent vote-share increases
-
-Similarly, since the classic model holds that manipulated votes communicate little about incumbents' post-election strength, it implies that when incumbents rely heavily on election manipulation they will struggle to put down post-election protests.
-This effect should likewise be conditional on incumbent vote-share, since those incumbents who produce small margins of victory while relying heavily on manipulation will have limited ability to push back successfully against protest movements.
-Consequently, more election manipulation should be associated with longer-lived protest movements against surviving incumbents.^[In the appendix, I check the separate prediction that manipulation should be positively associated with incumbents being ousted by protest movements.]
-
-> H3 (manipulation signals weakness): Election manipulation is positively associated with protest movement duration, with the size of this effect decreasing with incumbent vote-share.
-
-```{r newtheory, echo=FALSE, fig.cap="Theory diagram", out.width="90%", fig.align='center'}
-knitr::include_graphics("theory_diagram.drawio.png")
-
-```
-
-Lastly, if election manipulation is an indicator of underlying capacities the incumbent may use to defeat protest movements, this theory predicts that more severe electoral manipulation will be associated with shorter-lived protest waves.
-Incumbents with the resources and institutions needed to produce widespread election manipulation can draw on those same capacities to buy off, repress, or otherwise quash post-election protest.
-
-> H4 (manipulation signals strength): Among incumbents that survive post-election protest movements, more extensive election fraud should be associated with shorter protest waves.
-
-Figure \@ref(fig:newtheory) illustrates how the present argument differs from the classic model.
+On the other hand, if election manipulation is an indicator of underlying capacities the incumbent may use to defeat protest movements, more severe electoral manipulation will be associated with shorter-lived protest waves.
+Panel A of Figure \@ref(fig:theoryfig) illustrates how the present argument differs from the classic model.
 Manufactured votes are not negatively associated with genuine popularity, but positively associated with impunity for agents and control over patronage resources.
+They are also positively associated with incumbent popularity [@rundlett2016deliver].
 Abundant resources and extensive impunity reduce the principal-agent problems that frustrate election manipulation efforts.
 These same resources, institutions, and loyal agents make protest outcomes less likely to be favorable for the opposition.
-As such, opposition actors interpret manufactured and genuine votes as information about separate channels of incumbent strength.
-This produces hypotheses opposite to those of the classic model.
-Manipulation should be associated with reduced likelihood of protest, since it is a signal of capabilities that make protest success less likely.
-This effect should increase as margins of victory increase for the incumbent, providing additional, distinct information about incumbent resilience.
+Note that, in this model, protest initiation is not tied to expected protest outcome; protest is driven by non-strategic concerns related to the incumbent's popularity (i.e.
+citizens' diverse grievances) and whether manipulation is perceived as having been decisive (i.e.
+a stolen election).
+Incumbents with the resources and institutions needed to produce widespread election manipulation can draw on those same capacities to buy off, repress, or otherwise quash post-election protest.
+This yields Hypothesis 2.
+
+> H2: Election manipulation is positively associated with protest movement duration, with the size of this effect decreasing with incumbent vote-share.
 
@@
 To do so, they must solve principal-agent problems by consolidating control over patronage resources and shielding agents from negative consequences of their actions.
-The election outcome sends an ambiguous signal of the incumbents' resources\textemdash including natural popularity, resource control, and control over the legal apparatus\textemdash to possible regime opponents.
-The more successfully the incumbent has solved the principal-agent problems associated with election manipulation, however, the more likely it will be able to use those tools to quash protest.
+The same resources that can be used to manipulate elections on a wide scale\textemdash including natural popularity, resource control, and control over the legal apparatus\textemdash also better enable incumbents to dismantle protest movements when they do emerge.
 As a result, more extensive manipulation should be associated with shorter-lived protest movements against incumbents.
+Much election-related protest, then, is unrelated to strategic concerns about the likelihood of protest success.
 
@@
 ECAV data is in event-day format; to capture the *number of protests* before the election, I take the sum number of events recorded in ECAV before election day.
-To evaluate hypotheses regarding *protest duration* I take the number of days between the last protest recorded in ECAV and the first, for each election.^[The duration of a protest wave with only one protest is taken to be one day.
-While the main analysis assumes that the last protest in the dataset marks the end of the protest movement, an alternative specification in the appendix assumes right-censoring for a portion of the observations due to the 90-day observation window.]
+To evaluate hypotheses regarding *protest duration* I take the number of days between the last protest recorded in ECAV and the first, for each election.^[The duration of a protest wave with only one protest is taken to be one day.]
 
@@
 Still, approximately one-third of elections experienced some protest, and so protest initiation is modeled using standard logistic regression.
-Lastly, protest wave duration is modeled using Cox proportional hazards models.
+Protest wave duration is modeled using Cox proportional hazards models.
 
@@
 
-To capture the severity of *election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than alternative measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], may not mobilize real voters [@Harvey2020], and can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021].
+To capture the severity of *election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." It is important to note that this measure captures illegal election-day manipulation broadly, and is not confined simply to falsification.
+
+I operationalize electoral manipulation using this variable, rather than alternative measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], may not mobilize real voters [@Harvey2020], and can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021].
 This measure of election integrity should thus serve as a strong test of the protest-oriented model.
@@
 The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance.
-The winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia, where links to official sources were available, and Psephos.} Finally, whether or not the chief executive is *term limited* is taken from NELDA.
+The winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia, where links to official sources were available, and Psephos.}
 
-As identified in prior research, other factors that may influence both fraud and protest include the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), *judicial constraints* on the executive ('v2x_jucon') and the available of *alternative sources of information*.
+As identified in prior research, other factors that may influence both fraud and protest include the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), de facto *judicial independence* [@Linzer2015] and the available of *alternative sources of information* ('v2x_freexp_altinf' from V-Dem).
 Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019].
@@
 
-model.margin.binary <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd + illegal_man_index  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary)
@@
 summary(model.margin.binary_noc)
+
+model.binary_noc <- glm(any_protest ~  v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+summary(model.binary_noc)
 
@@
 integrity index (lag)",
-                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity"
+                 "physintlag.sq" = "Phys.
+int.
+lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity", "winner.share.upd" = "Incumb.
+vote-share",
+                 "median.participants" = "Median protest participants"
                  ))
@@
 
-Table \@ref(tab:logit-table) shows the results for Hypotheses 1 and 2.
+Table \@ref(tab:logit-table) shows the results for Hypothesis 1.
 Model 1 is a base model, where incumbent vote-share and *intentional irregularities* are standalone variables.
@@
 The interaction is significant, as Figure \@ref(fig:margin-prob-fig) illustrates.
-The direction of the effect supports Hypothesis 1, but not Hypothesis 2.
+The direction of the effect supports Hypothesis 1.
 That is, the occurrence of any protest is most likely when incumbent vote-share is low and fraud is high.
@@
 Table \@ref(tab:cox-models-tab) reports the results of Cox proportional hazard models of protest duration.
-In such models, coefficients with a positive sign indicate factors that are associated with shorter-lived protests; those with a negative sign indicate factors associated with a slower demise.
+In these models, positive coefficients indicate factors that are associated with shorter-lived protests; those with a negative sign indicate factors associated longer-lasting protest.
 Models 3 and 5 do not include interactions, while the remaining models interact *intentional irregularities* with incumbent vote-share.
@@
 Exponentiation of the coefficient for election fraud gives a hazard ratio of 1.3, indicating that for each one-unit increase in fraud, a protest movement becomes thirty percent more likely to end at any given time.
-This supports H4 and fails to support H3, suggesting that incumbents with the capacity to manipulate extensively are also able to draw on their resources to curtail post-election protest.
+This supports H2, suggesting that incumbents with the capacity to manipulate extensively are also able to draw on their resources to curtail post-election protest.
 
@@
 In Model 4, incumbent vote-share is interacted with *election fraud*.
-Figure \@ref(fig:margin-dur-fig) shows the marginal effect.
+Figure \@ref(fig:margin-dur-fig) shows the predicted probabilities.
 As shown in the figure, for all incumbent vote-shares near and below fifty percent, large-scale fraud significantly increases the likelihood that a protest movement will end.
@@
 Model 7, adds the median number of participants at events in the protest wave as a control.
-Larger protests might indicate relative opposition strength, however, since there is considerable missingness in the variable, I use it as a robustness check.
-Including this control variable reduces the size of the interaction coefficient, but it retains its direction and statistical significance.
+Including this control variable reduces the sample size due to missingness, but the interaction coefficient retains its direction and statistical significance.
 
@@
 The duration analysis consistently shows that more election manipulation is associated with quicker incumbent victories over protest movements, both independently and in conjunction with low incumbent vote-share.
-This is a strong indication that governments which can deliver large-scale election manipulation have the resources and capabilities to put down protest more quickly; that is, such manipulation ought to convey information about incumbent strength.
+This is a strong indication that governments which can deliver large-scale election manipulation have the resources and capabilities to put down protest more quickly; that is, such manipulation covaries with sources of regime durability.
+In real contests in which opposition and incumbent strengths are tested against each other, better manipulators reveal themselves to be stronger rather than weaker.
 
-These results are most consistent with the view that election fraud can be an indicator of regime strength.
 This finding supports and extends the line of research initiated by Simpser [@simpser2013].
-While that line of research holds that electoral manipulation in conjunction with a large margin of victory for the incumbent can induce pro-regime behavior in society, this paper shows that even when such cooperation fails\textemdash when post-election protests occur\textemdash governments that manipulate more are more likely to successfully weather the challenge.
+While that view holds that electoral manipulation in conjunction with a large margin of victory for the incumbent can induce pro-regime behavior in society, this paper shows that even when such cooperation fails\textemdash when post-election protests occur\textemdash governments that manipulate more are more likely to successfully weather the challenge.
 That is, even when ruling party vote-shares are low (perhaps especially when this is so), incumbents which show themselves capable of large-scale fraud are more capable of quickly ending protest movements than low-fraud incumbents.
@@
 \begin{table}[]
-\begin{tabular}{@{}l|lll|lll|l@{}}
+\begin{tabular}{@{}l|lll|l@{}}
 \toprule
                                                                                       & \begin{tabular}[c]{@{}l@{}}Successful \\ protests\end{tabular} & \begin{tabular}[c]{@{}l@{}}Failed \\ protests\end{tabular} & \begin{tabular}[c]{@{}l@{}}P-value \\ (successful vs.
-\\ unsuccessful)\end{tabular} & Protested & Unprotested & \begin{tabular}[c]{@{}l@{}}P-value \\ (protested vs.
-\\ unprotested)\end{tabular} & \begin{tabular}[c]{@{}l@{}}SD \\ (full sample)\end{tabular} \\ \midrule
+\\ unsuccessful)\end{tabular} & \begin{tabular}[c]{@{}l@{}}SD \\ (full sample)\end{tabular} \\ \midrule
 \begin{tabular}[c]{@{}l@{}}Intent.
-\\ irregularities\end{tabular}                     & 1.55                                                           & 1.25                                                       & 0.26                                                                                & 1.27      & 1.16        & 0.24                                                                              & 0.93                                                        \\
+\\ irregularities\end{tabular}                     & 1.55                                                           & 1.25                                                       & 0.26                                                                                & 0.93                                                        \\
 Inc.
-vote-share                                                                       & 0.50                                                           & 0.59                                                       & 0.08                                                                                & 0.58      & 0.62        & 0.07                                                                              & 0.21                                                        \\
-\begin{tabular}[c]{@{}l@{}}Pre-election \\ protests (n)\end{tabular}                  & 7.29                                                           & 6                                                          & 0.25                                                                                & 6         & 0           & 0                                                                                 & 8.4                                                         \\
-\begin{tabular}[c]{@{}l@{}}Pre-election protests\\ (median participants)\end{tabular} & 3.58                                                           & 3                                                          & 0.1                                                                                 & 3.2       & NA          & NA                                                                                & 0.89                                                        \\
-\begin{tabular}[c]{@{}l@{}}Civil society \\ openness\end{tabular}                     & 0.30                                                           & 0.32                                                       & 0.85                                                                                & 0.32      & 0.14        & 0.09                                                                              & 1.08                                                        \\
+vote-share                                                                       & 0.50                                                           & 0.59                                                       & 0.08                                                                                & 0.21                                                        \\
+\begin{tabular}[c]{@{}l@{}}Pre-election \\ protests (n)\end{tabular}                  & 7.29                                                           & 6                                                          & 0.25                                                                                & 8.4                                                         \\
+\begin{tabular}[c]{@{}l@{}}Pre-election protests\\ (median participants)\end{tabular} & 3.58                                                           & 3                                                          & 0.1                                                                                 & 0.89                                                        \\
+\begin{tabular}[c]{@{}l@{}}Civil society \\ openness\end{tabular}                     & 0.30                                                           & 0.32                                                       & 0.85                                                                                & 1.08                                                        \\
 Phys.
 int.
-rights                                                                     & 0.45                                                           & 0.43                                                       & 0.81                                                                                & 0.43      & 0.44        & 0.7                                                                               & 0.22                                                        \\ \bottomrule
+rights                                                                     & 0.45                                                           & 0.43                                                       & 0.81                                                                                & 0.22                                                        \\ \bottomrule
 \end{tabular}
@@
 The extremes where significant effects are found would then correspond to strategic, but risky, gambles by protest entrepreneurs hoping that mass manipulation will help pull regime opponents into the streets, and betting that they will not when manipulation is low.
+Protests that occur outside of this window are more likely to be 'bottom-up,' spontaneous reflections of public grievance.
 Thus the results find support for both the strategic and non-strategic literatures on protest.
@@
 
+Maybe add some thoughts on Kuran 1991--the idea being that authoritarian regimes are low info and rife with preference falsification.
+So opp has no choice but to gamble--so strategic in motivation but not in outcome or estimation?
+
 The second question this paper does not closely examine is the role of election fraud in sustaining (or not) the small number of incumbents who are defeated by mass protest.
@@
 Others contend that election fraud can indicate incumbent strength.
-Building on the principal-agent literature of election fraud, I argue that large-scale fraud is indicative of incumbent resilience even when regime popularity (and thus vote-shares) are low.
+Building on the principal-agent literature of election fraud, I argue that large-scale fraud draws on the same resources that make incumbents resilient to protest, even when regime popularity (and thus vote-shares) are low.
 Conducting election fraud requires consolidated control over patronage resources and the ability to protect front-line agents from the consequences of illegal behavior; this capacity also helps insulate incumbents from the effects of mass protest.
@@
 This paper attempts to test the assumptions of these two models by investigating the relationship between pre-election risk and election fraud, as well as between election fraud and protest occurrence and duration.
-Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 661 elections in electoral authoritarian regimes to show that 1) when pre-election risks to incumbents are higher, subsequent fraud increases rather than decreases, 2) protest occurrence is most likely when incumbent vote-shares are very low and election fraud is high, and 3) incumbents that engaged in larger-scale election fraud are able to more quickly defeat protest movements.
+Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 661 elections in electoral authoritarian regimes to show that protest occurrence is most likely when incumbent vote-shares are very low and election fraud is high, and\textemdash most importantly\textemdash incumbents that engage in larger-scale election fraud are able to more quickly defeat protest movements.
 

### 2025-02-26 Scripts/authoritarian_only_latex_2.Rmd
Updates prior to CPS

@@
 indent: true
-abstract: "The risk of mass protest figures prominently in many models of election manipulation, as a deterrent to tampering with elections.
-Yet election manipulation is common outside liberal democracies, while protest is rare (and successful protest rarer still).
-Using data on election fraud and duration data for protest events from 647 elections from 1990 to 2012, this paper argues that the risk to incumbents has been overstated in prior work.
-It argues that incumbents who solve the principal-agent dilemmas associated with election fraud are likely to be highly resilient to protest.
-There are two main findings.
-First, a riskier pre-election environment for incumbents is associated with more fraud, not less.
-Second, when protests do occur, greater election-day fraud is associated with a faster demise for protest movements.
-This pessimistic finding implies that election fraud is less risky than much prior work assumes, and helps explain the long-term stability of electoral authoritarianism as a regime type."
+abstract: "Two models of the relationship between election manipulation and protest can be found in the literature.
+A classic approach holds that election fraud reveals regime weakness, which can lead to strategic opposition protest.
+To avoid costly protest, incumbents may hold cleaner elections.
+A more recent approach contends that election manipulation can be a sign of underlying regime strengths; under this framework protest is less likely to deter manipulation.
+Prior empirical tests do not cleanly distinguish between these models.
+Using event-day data on electoral protest in 445 authoritarian cases from 1990-2012, this paper shows that 1) fraud severity is rarely associated with protest initiation, and 2) governments which manipulate more also defeat protests more quickly.
+This goes against the classic, deterrent model, and helps explain why electoral authoritarianism is a sticky equilibrium amid ongoing democratic backsliding.
+It suggests researchers should look beyond protest risk in models of electoral manipulation."
 header-includes: 
@@
 This view is adopted by many researchers, especially in the formal modelling tradition.
-Or, does election manipulation indicate sources of strength for incumbents (independent of genuine popularity), as others have held?
+Or, does election manipulation indicate sources of strength for incumbents (independent of genuine popularity), and does protest occur for non-strategic reasons?
 
 The stakes of this debate are significant.
-The former view is an optimistic one; incumbents avoid fraud when it might be exposed and reveal their weakness to strategic protesters\textemdash an auspicious prediction for democratization and ultimately the under-pinning of 'self-enforcing' democracy [@fearon_self-enforcing_2011].
-If the latter view holds, electoral protest may mainly occur non-strategically, and would typically fail to deter election manipulation as a result.
-Mass protest would, in this case, be a fragile guardrail against autocratization.
+The former view is an optimistic one; in it, incumbents may restrain fraud when it might be exposed and reveal their weakness to strategic protesters\textemdash an auspicious prediction for democratization and ultimately the under-pinning of 'self-enforcing' democracy [@fearon_self-enforcing_2011].
+If the latter view holds, non-strategic electoral protest may fail to deter election manipulation.
+The risk of mass protest would, in this case, be a fragile guardrail against autocratization.
 This debate cannot be advanced by focusing solely on protest occurrence.
@@
 In short: there are two views on the nature of electoral protest which make the same prediction.
-If opposition groups strategically initiate protest when they believe the incumbent to be weak, they will be most likely to do so when the election results show a low vote-share for the incumbent combined with high levels of manipulation\textemdash a combination of factors which reveals the incumbent's very low genuine popularity.
-[NOTE: ADD CITES HERE] However, if protest movements are more organic and non-strategic, reflecting citizen grievances, we should expect protest initiation in the same circumstances.
+If opposition groups strategically initiate protest when they believe the incumbent to be weak, they will be most likely to do so when the election results show a low vote-share for the incumbent combined with high levels of manipulation\textemdash a combination of factors which reveals the incumbent's very low genuine popularity [@Luo2017;@Little2012;@Schedler2002;@Beaulieu2014a].
+However, if protest movements are more organic and non-strategic, reflecting citizen grievances, we should expect protest initiation in the same circumstances.
 Here the combination of low incumbent vote-share and high manipulation indicates severe citizen grievances (low natural popularity) which are galvanized by the shared grievance of a 'stolen' election [@tucker2007].
@@
 In strategic-protest models, the opposition protests because it believes the probability of a pro-government outcome of the protest is low.
-I assume that the duration of protest movements, when they begin, is an indicator of this underlying likelihood of a pro-government outcome.
-The more quickly a government can put down protests, the greater the underlying (unobserved) distribution of power favors the incumbent over the opposition.
+I assume that the duration of protest movements, when they occur, is an indicator of this latent likelihood of a pro-government outcome.
+The more quickly a government puts down protests, the greater the underlying (unobserved) distribution of power favors the incumbent over the opposition.
 By contrast, the longer it takes the government to restore control, the less favorable the distribution of power.^[The analysis treats the small number of cases where the opposition succeeds in ousting the government in several ways.
-They are excluded from the main analysis, but incorporated in multiple robustness checks.]
+They are excluded from the main analysis, but incorporated in a robustness check.]
 
-I make use of data on the duration of post-election protests.
-Shorter-lived protest movements that end with the incumbent still in power indicate a government that is relatively strong compared to the protesting opposition, while longer-lived protest movements indicate a more even balance of power between the two.
-The reverse is also true: shorter-lived protest movements that end in opposition victory indicate a weaker government, while governments that hold on longer before being defeated are relatively stronger.
-Using data from 614 non-democratic elections across 97 countries, from 1989 to 2012, this paper finds that 1) protest initiation is most likely when manipulation is severe but incumbent vote-share is low, as both theories predict, and 2) that governments that engage in more manipulation are likely to put down protests more quickly than governments that produce little manipulation, especially when incumbents claim small vote-shares.
+Using data from 445 non-democratic elections across 97 countries, from 1989 to 2012, this paper finds that 1) protest initiation is most likely when manipulation is severe but incumbent vote-share is low, as both theories predict, and 2) that governments that engage in more manipulation are likely to put down protests more quickly than governments that produce little manipulation, especially when incumbents claim small vote-shares.
 
@@
 It is also consistent with the argument that much post-election protest in non-democracies is non-strategic; i.e., the actions of a weak opposition against a strong incumbent.
-Given these results, it is likely that the downside risk for incumbents that engage in election fraud in non-democracies has been overemphasized.
-This interpretation helps explain several empirical patterns at once: the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory in conjunction with high levels of fraud.
+Given these results, it is likely that\textemdash while incumbents do not seek to provoke protest\textemdash the downside risk for incumbents that engage in election fraud in non-democracies has been overemphasized.
+This interpretation helps explain several empirical patterns at once: the relative paucity of electoral protests [@Brancati2016], the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory in conjunction with high levels of fraud.
 At a theoretical level, they also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation.
@@
 Much of the work in this tradition has been elaborated through formal models.
-Though there are, of course, many variations, in several such models incumbents are characterized as 'weak' or 'strong' based on their true support in the electorate; weak incumbents may need to resort to fraud in order to secure re-election.
+Though there are, of course, many variations, in several such models incumbents are characterized as 'weak' or 'strong' based on their true support in the electorate.
+This true level of support can be obscured prior to elections due to pervasive preference falsification [@Kuran1991] and the low level of political information common in authoritarian regimes [@guriev2020].
+However, low popularity (i.e.
+weakness, in these models) can be revealed when incumbents need to resort to fraud in order to secure re-election.
 As a result, the opposition perceives election fraud as a reliable signal of incumbent weakness, and will stage protests in order to push out the weak incumbent.
@@
 This, in turn, indicates to opposition actors that protest is more likely to succeed in unseating the incumbent.
-These are normatively optimistic models upon which a sizable literature has been founded: fraud can be kept in check, and ultimately reduced, through the development of institutions that reveal information about election integrity: election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].
+Incumbents in these models typically face a tradeoff between the short-term benefits of electoral manipulation for winning the election, and any weakness that relying on fraud reveals to the opposition [@Rozenas2016b].
+As Beaulieu [-@Beaulieu2014a, p.
+48] writes in this tradition, "As the opposition grows stronger, the amount of [manipulation] he will tolerate shrinks, and the incumbent who wishes to avoid protest will need to reduce the amount of manipulation she commits." These are normatively optimistic models upon which a sizable literature has been founded: fraud can be kept in check, and ultimately reduced, through the development of institutions that reveal information about election integrity: election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].
 These institutions improve election integrity, it is claimed, by increasing the likelihood that fraud will be observed and thus, incumbent weakness exposed.
@@
 6].
-A strong incumbent (in the language used by the first set of models, i.e.
-one that is relatively popular) may nonetheless manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver].
+A relatively popular incumbent ("strong" in the language used by the first set of models) may nonetheless manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver].
 By demonstrating otherwise difficult-to-observe capabilities, large-scale election manipulation can solve a coordination problem for the incumbent, securing the cooperation of elites [@simpser2013].
@@
 
-For the set of theories that treat election manipulation as an indicator of sources of regime strength, manipulation can serve to deter protest under some circumstances.
-[NOTE ADD CITES HERE?] This leaves protest, when it does occur in the presence of large-scale manipulation, as something of a puzzle if protest behavior is largely strategic.
+If election manipulation can be an indicator of regime strength, it may serve to deter protest under some circumstances, rather than the other way around [@simpser2013].
+This leaves protest, when it does occur in the presence of large-scale manipulation, as something of a puzzle if protest behavior is largely strategic.
 For this reason, the manipulation-as-strength view comports more easily with another established view of the causes of electoral protest.
-Researchers in this tradition have argued that electoral protest occurs unrelated to strategic concerns about the relative strength of the incumbent and the opposition.
-In these models, electoral protest occurs because manipulation galvanizes citizens who hold diverse grievances, because economic conditions are poor, or simply because the bare fact of holding an election creates a moment of opportunity for opposition organization irrespective of the level of manipulation [@Pop-Eleches2015] [NOTE: ADD CITES].
-The central feature of these models is that protest decisions are generally made without consideration for the likelihood that protests will overthrow the incumbent.
-As will be discussed in depth below, these different assumptions yield different predictions for when electoral protest is more likely to be initiated, and how long it will last against an incumbent that refuses to give up power.
+Researchers in this tradition have argued that electoral protest occurs largely unrelated to strategic concerns about the relative strength of the incumbent and the opposition.
+In these models, electoral protest occurs because manipulation galvanizes citizens who hold diverse grievances, because economic conditions are poor, or simply because the bare fact of holding an election creates a moment of opportunity for opposition organization irrespective of the level of manipulation [@Pop-Eleches2015].
+The central feature of these models is that protest decisions are generally made without consideration for the likelihood that protests will overthrow the incumbent.^[To be clear, this is the narrow sense in which 'strategic' and 'non-strategic' is used here and in protest-deterrent models.
+Protest behavior may still be rational in a broader sense without being narrowly tied to the expected likelihood of ousting the government.
+For instance, protesters may believe it is unlikely they will be individually punished for participating in protests, making it individually rational for them to voice their grievances.]
 
-This school of thought holds that decisions to engage in electoral protest are non-centralized, driven by citizens' grievances rather than by a strategic, organized opposition [@Brancati2016].
+This school of thought holds that decisions to engage in electoral protest are driven in large part by citizens' grievances rather than by a strategic opposition [@tucker2007].
 In this view, election manipulation can act as a shared grievance among individuals who prefer to see the ruling party lose\textemdash especially when it appears that fraud may have been decisive [@kuntz_more_2009].
@@
 
+Rational protest entrepreneurs might seek to initiate protest to take advantage of this popular discontent for non-strategic reasons, in order to press for improved competitiveness in the *next* election or to bargain for policy concessions [@chernykh2014;@robertson_protest_2010].
+They may stage protests in order to identify new participants, forge new social ties [@barberaModelProtestsRevolution2019] and test new tactics.
+They may seek to communicate information about the government's unpopularity, to set the stage for larger, future protests [@lohmannDynamicsInformationalCascades1994].
+Protests may be staged as part of internal factional rivalries within the opposition [@Buckles2017], or to drive media coverage of issues important to the opposition [@huangMediaProtestDiffusion2019;@jenningsStreetDemonstrationsMedia2019].
+Or, a vanguard of protesters may take to the streets simply because their distaste for the government outweighs any risk of sanction [@lohmannDynamicsInformationalCascades1994].
+These and other motivations for initiating protest are theoretically significant because their rationality (for protesters) does not hinge on the likelihood of overturning the government, as classic models assume.
+As a result, it is possible that we may observe the initiation of anti-government electoral protest even when the likelihood of such protest overturning the election (or the government itself) is very low.
+
 Empirical work has found some support for the microfoundations of this theory.
@@
 
-```{=latex}
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[]
-\begin{tabular}{@{}lllllll@{}}
-\toprule
-                     &              &                           & \multicolumn{2}{l}{Strategic protest; manipulation = weakness} & \multicolumn{2}{l}{Non-strategic protest; manipulation = strength} \\ \midrule
-Incumbent vote-share & Manipulation & True popularity inference & Initiation              & Prob. of pro-gov outcome             & Initiation                & Prob. of pro-gov. outcome              \\
-Low                  & Low          & Low                       & Less likely             & More likely                          & Less likely               & Less likely                            \\
-Low                  & High         & Very low                  & More likely             & Less likely                          & More likely               & More likely                            \\
-Citations            &              &                           &                         &                                      &                           &                                        \\ \bottomrule
-\end{tabular}
-\caption{}
-\label{tab:predictions}
-\end{table}
-```
-
-Table 1 summarizes these predictions.
-As can be seen in the table, if we focus only on protest occurrence as a dependent variable, we cannot distinguish between two possible data-generating processes: one in which electoral protest is strategic and manipulation signals weakness, and one in which electoral protest is largely un-centralized but manipulation indicates regime strengths.
+Taking this all together, it becomes clear that if researchers focus only on protest occurrence as a dependent variable, they cannot distinguish between two possible data-generating processes: one in which manipulation signals weakness and electoral protest is a strategic response, and one in which electoral manipulation indicates regime strengths and protest occurs for a variety of reasons with limited connection to the odds of removing the government.
+In the former, protest is predicted to occur most often when manipulation is high and the incumbent's vote-share is low, since this provides a strong signal that the incumbent's true popularity is low.
+In the latter, protest is most likely in the same condition because the incumbent's low popularity indicates widespread grievances, which the high level of manipulation serves to coordinate despite the incumbent's organizational strengths.
 Which of these data-generating processes predominates is of significant theoretical importance of understanding authoritarianism and democratization, since the latter view suggests much dimmer prospects for democratization than the former.
@@
 
-Integrating Simpser's argument with recent research on the principal-agent dynamics inherent to election manipulation provides insights into this question.
+Prior models of electoral protest by strategic opposition actors, discussed above, treat election manipulation and genuine votes as substitute goods.
+For any given vote-share, an increase in manufactured votes reduces the number of genuine votes.
+For classic models, which consider manufactured votes to be merely illusions of strength, the higher the number of manufactured votes for any given vote total, the weaker the incumbent.
+This means that strategic opposition actors are more likely to initiate protest when they perceive the incumbent to be weaker, and expect to be more likely to succeed in a challenge.
+
+In the classic model, the incumbent's (unknown) true popularity is positively related to its vote-share in the election, and negatively related to its use of fraud.
+This dynamic is the fulcrum upon which these models hinge: less popular incumbents (who are expected to be less likely to defeat an opposition protest) must manipulate more in order to win elections.
+Popular incumbents need manipulate less.
+Opposition actors gauge the probability of an pro-government protest outcome by estimating the incumbent's true popularity based on the degree of manipulation and the incumbent's vote-share in the election.
+A high vote-share with low fraud indicates a more popular incumbent, likely deterring protest.
+A low vote-share with a high degree of fraud indicates a very weak incumbent, making strategic protest more likely.
+In other words, manipulated votes are empty in the classic model; subtracting them from the incumbent's vote-total gives opposition actors their estimate of the incumbent's ability to win a post-election standoff.
+
+Thus the predicted relationship between manipulation and protest likelihood in the classic model is positive: more manipulation makes protest more likely.
+However, as increasing vote-share increases the number of genuine votes for the incumbent at any fixed level of fraud, the effect of fraud should be conditional on the claimed vote-share.
+As Rozenas [@Rozenas2016b] notes, if an incumbent wins an election with low manipulation, it generates a strong signal of its popularity; if it wins a heavily manipulated election, this signal is weaker.
+As an illustration, if the incumbent's total vote-share is 70 percent, for example, of which the opposition estimates 15 percent to be manufactured, the incumbent's genuine popularity is estimated at a respectable 55 percent.
+By contrast, if the incumbent posts a 50 percent margin of victory with the same 15 percent estimated to be manufactured, the opposition may perceive that the incumbent is weak in its genuine support (35 percent).
+This produces a first, baseline hypothesis, derived from the classic model.
+
+> H1: Increasing election fraud makes protest more likely, with the size of this effect decreasing with incumbent vote-share.
+
+It is important to note that the same hypothesis is also generated by non-centralized, non-strategic models of election protest.
+The mechanism, however, is different.
+Here, the combination of low incumbent vote-share and high manipulation is indicative of widespread grievances coordinated a 'stolen' election.
+These are conditions that should be most likely to produce protest [@tucker2007].
+By contrast, as vote-share increases for a set level of fraud, the incumbent's genuine popularity is higher and thus grievances are lower (reducing the likelihood of protest).
+
+Integrating Simpser's argument with recent research on the principal-agent dynamics inherent to election manipulation provides different insights.
 Where Simpser argues that manipulation should be blatant and excessive (that is, producing a large margin of victory) to send signals of strength, the principal-agent framework for understanding manipulation suggests otherwise.
@@
 
-Prior models of electoral protest by strategic opposition actors treat election manipulation and genuine votes as substitute goods.
-For any given vote-share, an increase in manufactured votes reduces the number of genuine votes.
-For classic models, which consider manufactured votes to be merely illusions of strength, the higher the number of manufactured votes for any given vote total, the weaker the incumbent.
-This means that strategic opposition actors are more likely to initiate protest when they perceive the incumbent to be weaker, and expect to be more likely to succeed in a challenge.
-A simplified version of this argument is presented in Panel B of Figure \@ref(fig:theoryfig).
-In the figure, dashed objects indicate variables that actors must estimate, while solid items are observed.
-
-As the figure shows, in the classic model, the incumbent's (unknown) true popularity is positively related to its vote-share in the election, and negatively related to its use of fraud.
-This dynamic is the fulcrum upon which these models hinge: less popular incumbents (who are expected to be less likely to defeat an opposition protest) must manipulate more in order to win elections.
-Popular incumbents need manipulate less.
-Opposition actors gauge the probability of an pro-government protest outcome by estimating the incumbent's true popularity based on the degree of manipulation and the incumbent's vote-share in the election.
-The level of manipulation is, in effect, a mediator for the effect of vote-share.
-A high vote-share with low fraud indicates a more popular incumbent, likely deterring protest.
-A low vote-share with a high degree of fraud indicates a very weak incumbent, making strategic protest more likely.
-In other words, manipulated votes are empty in the classic model; subtracting them from the incumbent's vote-total gives opposition actors their estimate of the incumbent's ability to win a post-election standoff.
-
-Thus the predicted relationship between manipulation and protest likelihood in the classic model is positive: more manipulation makes protest more likely.
-However, as increasing vote-share increases the number of genuine votes for the incumbent at any fixed level of fraud, the effect of fraud should be conditional on the claimed vote-share [NOTE: ADD CITES].
-If the incumbent's total vote-share is 70 percent, for example, of which the opposition estimates 15 percent to be manufactured, the incumbent's genuine popularity is estimated at a respectable 55 percent.
-By contrast, if the incumbent posts a 50 percent margin of victory with the same 15 percent estimated to be manufactured, the opposition may perceive that the incumbent is weak in its genuine support (35 percent).
-This produces a first, baseline hypothesis, derived from the classic model.
-
-> H1: Increasing election fraud makes protest more likely, with the size of this effect decreasing with incumbent vote-share.
-
-It is important to note that the same hypothesis is also generated by non-centralized, non-strategic models of election protest.
-The mechanism, however, is different.
-In non-strategic models, protest emerges organically among disaffected citizens.
-Here, the combination of low incumbent vote-share and high manipulation is indicative of high grievances and a 'stolen' election.
-These are conditions that should be most likely to produce protest [@tucker2007].
-By contrast, as vote-share increases for a set level of fraud, the incumbent's genuine popularity is higher and thus grievances are lower (reducing the likelihood of protest).
-
-```{r theoryfig, echo=FALSE, fig.cap="A classic theory of election manipulation and protest", out.width="90%", fig.align='center'}
-knitr::include_graphics("theory_comp.drawio.png")
-
-```
-
-The classic model holds that manipulated votes correlate negatively with incumbents' post-election strength, implying that when incumbents rely heavily on election manipulation they will struggle to put down post-election protests.
-This effect should likewise be conditional on incumbent vote-share, since those incumbents who produce small margins of victory while relying heavily on manipulation will have limited ability to push back successfully against protest movements.
-Panel B in Figure [NOTE: FIX REF] illustrates this; a higher likelihood of pro-government protest outcome reduces the likelihood that the opposition will initiate protest.
-Consequently, classic models predict that more election manipulation should be associated with longer-lived protest movements against surviving incumbents.^[In the appendix, I check the separate prediction that manipulation should be positively associated with incumbents being ousted by protest movements.]
-
-On the other hand, if election manipulation is an indicator of underlying capacities the incumbent may use to defeat protest movements, more severe electoral manipulation will be associated with shorter-lived protest waves.
-Panel A of Figure \@ref(fig:theoryfig) illustrates how the present argument differs from the classic model.
+If election manipulation is an indicator of underlying capacities the incumbent may use to defeat protest movements, more severe electoral manipulation will be associated with shorter-lived protest waves.
 Manufactured votes are not negatively associated with genuine popularity, but positively associated with impunity for agents and control over patronage resources.
@@
 > H2: Election manipulation is positively associated with protest movement duration, with the size of this effect decreasing with incumbent vote-share.
+
+```{r theoryfig, echo=FALSE, fig.cap="A classic theory of election manipulation and protest", out.width="90%", fig.align='center', include=FALSE}
+knitr::include_graphics("theory_comp.drawio.png")
+
+```
+
+The classic model holds that manipulated votes correlate negatively with incumbents' post-election strength, implying that when incumbents rely heavily on election manipulation they will struggle to put down post-election protests.
+This effect should likewise be conditional on incumbent vote-share, since those incumbents who produce small margins of victory while relying heavily on manipulation will have limited ability to push back successfully against protest movements.
+A higher likelihood of pro-government protest outcome reduces the likelihood that the opposition will initiate protest.
+Consequently, classic models predict that more election manipulation should be associated with longer-lived protest movements against surviving incumbents.^[In the appendix, I check the separate prediction that manipulation should be positively associated with incumbents being ousted by protest movements.]
 
@@
 I limit the sample to authoritarian regimes which hold multiparty national elections.
-This excludes fully closed autocracies (where no national elections are held de jure), as well as electoral and liberal democracies.^[The specific variable used is 'v2x_regime'; I limit the sample to those regime-years coded 1 for either the current or lagged year.
-This coding rule will include cases that are coded as electoral democratic after an election but electoral authoritarian before.]
+This excludes fully closed autocracies (where no national elections are held de jure), as well as electoral and liberal democracies.
+The specific variable used is 'v2x_regime'.
+I limit the sample to those regime-years coded 1 (electoral authoritarian) for the election year, or coded 2 (electoral democracy) for the election year but 1 for the lagged year.^[This coding rule will capture cases that are coded as electoral democratic after an election but electoral authoritarian before, in order to ensure that cases which experience a liberalization after the election are included.] This sample includes a range of electoral authoritarian cases, including those that might be labeled as 'hegemonic' and 'competitive' [@donno2013elections].^[According to the V-Dem codebook, the coding scheme includes cases where multiparty elections were manipulated to the extent that "irregularities...affected the outcome of the election," and ones in which it is "hard to determine whether the irregularities affected the outcome or not." Only when irregularities definitively did not affect the outcome would election-years be categorized as electoral democracy, and excluded from the main analysis.] For instance, the dataset includes election years from Egypt and Belarus, but also from Ukraine, and Mexico in the 1990s.^[Of course, treating regime type as a sharply defined binary indicator, rather than as a continuum, will introduce some bias around edge cases.
+As a robustness check for this case selection process, I also run the main analyses on a larger sample that adds electoral democratic regimes.
+While the strength of the findings is somewhat attenuated, as is expected, the results are generally upheld in the broader sample as shown in the appendix.]
 
@@
 
-Of course, treating regime type as a sharply defined binary indicator, rather than as a continuum, will introduce some bias around edge cases.
-As a robustness check for this case selection process, I also run the main analyses on a larger sample that adds electoral democratic regimes.
-While the strength of the findings is somewhat attenuated, as is expected, the results are generally upheld in the broader sample as shown in the appendix.
-
-In all models, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the population of interest: protest following an election in which the incumbent won, possibly through reliance on election fraud.^[Excluding cases where incumbents stepped down after losing an election is necessary to avoid conflating pro- and anti-incumbent protest.
+In all models, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the population of interest: protest following an election in which the incumbent claimed victory, possibly through reliance on election fraud.^[Excluding cases where incumbents stepped down after losing an election is necessary to avoid conflating pro- and anti-incumbent protest.
 However, this design does influence the interpretation of the results from these models.
-Threats to inference from this section are evaluated in more detail in the discussion and appendix.] The full dataset includes 445 election-year observations across 81 non-democratic countries.
-The date range is limited by the availability of data on protest from ECAV, which ranges from 1989 to 2012.
+Threats to inference from this section are evaluated in more detail in the discussion section.] The full dataset includes 445 election-year observations across 81 non-democratic countries.
+The date range is limited by the availability of data on protest from ECAV, which ranges from 1989 to 2012.^[Some additional tests in the appendix include cases where incumbents accepted defeat, bringing the total number of observations to 659.]
 
@@
 Of the 445 elections in which the incumbent claimed victory, only 156 fall into this more limited dataset.
+Brief further analyses of the seven cases in which incumbents were ousted, as well as cases where incumbents stepped down after the election, are also included in the discussion and appendix.
 
@@
 I only include events labeled as protests, occupations, or blockades, excluding acts like shootings, bombings, and coups.
-ECAV data is in event-day format; to capture the *number of protests* before the election, I take the sum number of events recorded in ECAV before election day.
-To evaluate hypotheses regarding *protest duration* I take the number of days between the last protest recorded in ECAV and the first, for each election.^[The duration of a protest wave with only one protest is taken to be one day.]
+If any post-election protests are recorded in the ECAV data for a particular election, the binary variable *protest initiation* is marked as 1.
+ECAV data is in event-day format; to capture *protest duration* I take the number of days between the last protest recorded in ECAV and the first, for each election.^[The duration of a protest wave with only one protest is taken to be one day.] For the survival models, protest movements that were ultimately defeated by the incumbent are considered 'dead' after the last day of protest.
+Protest movements that prevailed in unseating the incumbent are included, and are marked as 'alive.'
 
-The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0) %>% nrow()` of the `r nrow(nelda2.sub)` elections, no protests are recorded.
+The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0 & inc.lose == "0") %>% nrow()` of the 445 elections, no protests are recorded.
 Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest duration.
@@
 
-In the survival models, the sample is restricted to the 156 cases where post-election protests occurred.
-If significant election fraud is indicative of regime resilience, the expectation is that more fraudulent elections will be associated with more quickly dispatched protest movements.
-As an additional robustness check, I subsequently exclude the 7 cases where incumbents were removed following protests.
-This allows me to exclude the possibility that incumbents who make greater use of fraud are *weaker*, and more quickly removed from office, than their more electorally honest peers.
-A brief further analysis of the seven cases in which incumbents were ousted is also included in the discussion.
-
 ```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
 
-p.hist.events <- ggplot(nelda2.sub, aes(x=n.events.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()
+p.hist.events <- ggplot(nelda2.sub %>% filter(inc.lose == "0"), aes(x=n.events.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()
 
-p.hist.duration <- ggplot(nelda2.sub, aes(x=d.events.num)) + geom_histogram(bins=100) + xlab("Protest wave duration") + theme_bw()
+p.hist.duration <- ggplot(nelda2.sub %>% filter(inc.lose == "0"), aes(x=d.events.num)) + geom_histogram(bins=100) + xlab("Protest wave duration") + theme_bw()
 
-p.hist.fraud <- ggplot(nelda2.sub, aes(x = v2elirreg.inv)) + geom_histogram(bins=100) + xlab("Election fraud") + theme_bw()
+p.hist.fraud <- ggplot(nelda2.sub %>% filter(inc.lose == "0"), aes(x = v2elirreg.inv)) + geom_histogram(bins=100) + xlab("Election fraud") + theme_bw()
 
- ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= jitter(n.events.post))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("n.events.post (jittered)") + theme_bw()
+ ggplot(nelda2.sub %>% filter(inc.lose == "0"), aes(y = v2elirreg.inv, x= jitter(n.events.post))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("n.events.post (jittered)") + theme_bw()
 
@@
 
-```{r elex-hist, fig.cap="Distributions of election fraud, post-election protest, and protest duration", fig.height=4, echo=FALSE, message=FALSE, warning = FALSE}
+```{r elex-hist, fig.cap="Histograms of election fraud and protest duration", fig.height=4, echo=FALSE, message=FALSE, warning = FALSE}
 plot.combined.hist <- cowplot::plot_grid(p.hist.fraud, p.hist.duration, ncol=1)
@@
 This measure of election integrity should thus serve as a strong test of the protest-oriented model.
-It also matches the conception of election manipulation used in much of the relevant formal literature.^[For example, election manipulation is treated as post-hoc adjustment of the results in Little et al [-@Little2015], Little [-@Little2012], Magaloni [-@magaloni2010], and Fearon [-@fearon_self-enforcing_2011], among others.] Two alternative measures are used as robustness checks in the appendix.
+It also matches the conception of election manipulation used in much of the relevant formal literature.^[For example, election manipulation is treated as post-hoc adjustment of the results in Little et al [-@Little2015], Little [-@Little2012], Magaloni [-@magaloni2010], and Fearon [-@fearon_self-enforcing_2011], among others.] An alternative measure is used as a robustness check in the appendix.
 
@@
 
-As identified in prior research, other factors that may influence both fraud and protest include the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), de facto *judicial independence* [@Linzer2015] and the available of *alternative sources of information* ('v2x_freexp_altinf' from V-Dem).
+As identified in prior research, other factors that may influence both fraud and protest are included as controls variables.
+The repressive capacity of the government [@Davenport2019] is measured using the physical integrity index from V-Dem (v2x_clphy), and its square term.
+Indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), de facto *judicial independence* [@Linzer2015] and the available of *alternative sources of information* ('v2x_freexp_altinf' from V-Dem) are also employed.
 Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019].
@@
 First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest.
-Another NELDA variable indicates if the incumbent executive is not running due to *term limits*.
+Another NELDA variable indicates if the incumbent executive is not running due to *term limits* [@hale_patronal_2014].
 I include dummy variables that provide a categorical evaluation of state capacity.
@@
 This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018].
-Lastly, since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.
+Since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.
+
+Lastly, I include the *number of pre-election protests* counted in the ECAV dataset, which is likely to be an important possible confounder; more numerous pre-election protest may affect the level of election fraud (as the deterrence logic predicts) as well as being directly tied to the likelihood of post-election protest.
 
@@
 
-model.margin.binary <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   + v2elirreg.inv +  +    winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary)
 
-model.margin.binary.int <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary.int <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary.int)
 
-model.margin.binary_noc <- glm(any_protest ~ lji.lag + legcon.1lag +  altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.margin.binary_noc <- glm(any_protest ~ lji.lag + legcon.1lag +  altinf.1lag  +  + osorg.1lag   + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.margin.binary_noc)
 
-model.binary_noc <- glm(any_protest ~  v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+model.binary_noc <- glm(any_protest ~  v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential , family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 summary(model.binary_noc)
@@
 
-plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.32, 2.2]"), show.data = T)
+plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.32, 2.2]"), show.data = T, colors = "gs")
 plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()
@@
 models2[['Model 1']] <- model.margin.binary
-models2[['Model 2']] <- model.margin.binary.int
+models2[['Model 2']] <- model.binary_noc
+models2[['Model 3']] <- model.margin.binary.int
 #models2[['Model 6']] <- model.margin.binary_noc
 
-modelsummary(models2, output = "default",  stars = T, title = 'Table 3: Logit models of protest occurrence', coef_rename = 
-               c("lji.lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
-                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
-                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
-                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
-                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
-                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state", "v2elirreg.inv" = "Election fraud",
-                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
-                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
-integrity index (lag)",
-                 "physintlag.sq" = "Phys.
-int.
-lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity", "winner.share.upd" = "Incumb.
-vote-share",
-                 "median.participants" = "Median protest participants"
-                 ))
+modelsummary(models2, output = "default",  stars = T, title = 'Table 3: Logit models of protest occurrence', gof_omit = c("Log|RMSE"), coef_map = c("(Intercept)" = "(Intercept)",
+  "lji.lag" = "Judical independence (lag)",                                                                 "legcon.1lag" = "Leg.
+constraints (lag)",                                                                 "gdpgro.1lag" = "GDP growth rate (lag)",                                                                   "altinf.1lag" = "Alternative info.
+(lag)",
+                 "osorg.1lag" = "Civil soc. openness (lag)",
+                 "inc.termlimit.num" = "Incumbent term-limited", 
+                 "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", 
+                 "rentierstate1" = "Rentier state", 
+                 "taxstate1" = "Tax state",
+
+                 "unscheduled_election1" = "Off-schedule election",
+                 "n.events.pre" = "Number of pre-election protests", 
+                 "physinteg.1lag" = "Phys. integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", 
+                 "regional_success_lag" = "Regional protest diffusion",
+                                  "presidential1" = "Presidential election",
+
+                 "median.participants" = "Median protest participants",
+                  "winner.share.upd" = "Incumb. vote-share",
+                  "v2elirreg.inv" = "Election fraud",
+                "v2elirreg.inv:winner.share.upd"= "Election fraud:Incumb. vote-share")
+                 )
+
 ```
@@
 In keeping with the ambiguous findings in the prior literature, there is no significant relationship between fraud severity and the likelihood of protest initiation.
-The incumbent's claimed margin of victory is negatively associated with protest initiation, though this relationship falls short of conventional statistical significance (p = .07).
+The incumbent's claimed margin of victory is negatively associated with protest initiation, though this relationship falls short of conventional statistical significance.
 Several control variables behave as expected; for instance, protests are more likely to occur after presidential elections and when there have been prior electoral protests in the region.
-These relationships are largely consistent in Model 2, the key model which includes the interaction between *intentional irregularities* and incumbent vote-share.
+Model 2 includes the interaction between *intentional irregularities* and incumbent vote-share, with no controls other than a dummy variable for *presidential elections*.
+This allows for the largest possible sample size, and the results are consistent with those of the full model with controls.
+
+Model 3 is the main model, which includes the full set of controls.
 The interaction is significant, as Figure \@ref(fig:margin-prob-fig) illustrates.
+The figure shows the predicted probability of post-election protest for two specific values of election fraud, across all values of incumbent vote-share, holding all other variables constant at their means (continuous variables) or modes (categorical variables).
+The values for election fraud represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively.
 The direction of the effect supports Hypothesis 1.
@@
 
-nelda2.sub <- nelda2.sub %>% mutate(incumbent_strength = ifelse(d.events.num == 0, 92, ifelse(inc.replaced.prot == 1, -1*(92-d.events.num), 92 - d.events.num)))
-
 nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
@@
 
-cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0))
+cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" &  n.events.post > 0))
 summary(cox1)
 
-cox1.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0))
+cox1.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0))
 summary(cox1.cen)
@@
 
-cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
+cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag +  + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
 summary(cox2)
-p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.33, 2.2]")) #This is the 10th and 90th quantiles
+p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.33, 2.2]"), colors = "gs") #This is the 10th and 90th quantiles
 p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()
 
-cox2.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
+cox2.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
 summary(cox2.cen)
 
-cox3.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post + median.participants, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+cox3.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag +  v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post + median.participants, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
 summary(cox3.cen)
@@
 In these models, positive coefficients indicate factors that are associated with shorter-lived protests; those with a negative sign indicate factors associated longer-lasting protest.
-Models 3 and 5 do not include interactions, while the remaining models interact *intentional irregularities* with incumbent vote-share.
+Models 4 and 6 do not include interactions, while the remaining models interact *intentional irregularities* with incumbent vote-share.
 One threat to inference in this design derives from the 90-day post-election window considered by ECAV.
-It is possible that some protest movements considered 'dead' in the baseline models 3 and 4 are, in fact, 'alive,' if subsequent protests occurred after 90 days and are unobserved.
-To address this, observations are assumed to be right-censored if a protest occurred within three-weeks of the 90-day cutoff in Models 5, 6, and 7.
+It is possible that some protest movements considered 'dead' in the baseline models 4 and 5 are, in fact, 'alive,' if subsequent protests occurred after 90 days and are unobserved.
+To address this, observations are assumed to be right-censored if a protest occurred within three-weeks of the 90-day cutoff in Models 6, 7, and 8.
 This amounts to 15 of the 156 observations in the sample.
-Model 7 is the most restrictive model, which includes right censoring as well as a control variable for the median number of participants during the protest wave.
+Model 8 is the most restrictive model, which includes right censoring as well as a control variable for the median number of participants during the protest wave.
 There is considerable missingness in this variable.
 
-The coefficient for *election fraud* is positive and significant in Model 3, indicating that more fraudulent elections are associated with less resilient protest movements.
+The coefficient for *election fraud* is positive and significant in Model 4, indicating that more fraudulent elections are associated with less resilient protest movements.
 Exponentiation of the coefficient for election fraud gives a hazard ratio of 1.3, indicating that for each one-unit increase in fraud, a protest movement becomes thirty percent more likely to end at any given time.
@@
 Further investigation of the relationship between fraud and protest duration yields additional insights.
-In Model 4, incumbent vote-share is interacted with *election fraud*.
-Figure \@ref(fig:margin-dur-fig) shows the predicted probabilities.
-As shown in the figure, for all incumbent vote-shares near and below fifty percent, large-scale fraud significantly increases the likelihood that a protest movement will end.
+In Model 5, incumbent vote-share is interacted with *election fraud*.
+Figure \@ref(fig:margin-dur-fig) shows the predicted hazard ratios As shown in the figure, for all incumbent vote-shares near and below fifty percent, large-scale fraud significantly increases the likelihood that a protest movement will end.
 This result also supports the fraud-as-strength model, and shows that only as vote-share increases above fifty percent\textemdash creating a stronger signal of popularity\textemdash does a government in a low-fraud election have the same likelihood of defeating a protest movement at a given time.
+In effect, either high genuine popularity or the ability to produce high levels of fraud enables governments to more quickly end protests.
 
 The results of the baseline models are consistent with those models which incorporate censoring.
-While the coefficient on *election fraud* falls slightly short of conventional statistical significance in Model 5 (p = .059), it retains the same size and sign compared to Model 3.
-The interaction effect in Model 6 increases in size, while retaining its direction and statistical significance.
+While the coefficient on *election fraud* falls slightly short of conventional statistical significance in Model 6 (p = .059), it retains the same size and sign compared to Model 4.
+The interaction effect in Model 7 increases in size, while retaining its direction and statistical significance.
 This suggests that the main models are not drawing overly incorrect inferences through censoring of longer-lasting protest movements.
-Model 7, adds the median number of participants at events in the protest wave as a control.
+Model 8, adds the median number of participants at events in the protest wave as a control.
 Including this control variable reduces the sample size due to missingness, but the interaction coefficient retains its direction and statistical significance.
@@
 models3 <- list()
-models3[['Model 3']] <- cox1
-models3[['Model 4']] <- cox2
-models3[['Model 5']] <- cox1.cen
-models3[['Model 6']] <- cox2.cen
-models3[['Model 7']] <- cox3.cen
+models3[['Model 4']] <- cox1
+models3[['Model 5']] <- cox2
+models3[['Model 6']] <- cox1.cen
+models3[['Model 7']] <- cox2.cen
+models3[['Model 8']] <- cox3.cen
 
-modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Table 4: Cox proportional hazard models of protest duration', coef_rename = 
-               c("lji.lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Table 4: Cox proportional hazard models of protest duration', coef_map = 
+               c("(Intercept)" = "(Intercept)", "lji.lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                  "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                  "osorg.1lag" = "Civil soc. openness (lag)", 
-                 "winner.share.upd" = "Incumbent vote-share",
                  "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
@@
 integrity index (lag)",
-                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "v2elirreg.inv" = "Election fraud"
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", 
+                 "median.participants" = "Median protest size",
+                                  "winner.share.upd" = "Incumbent vote-share",
+                 "v2elirreg.inv" = "Election fraud",
+"v2elirreg.inv:winner.share.upd" = "Election fraud:Inc. vote-share"
                  ))
-
 ```
@@
 It also helps us make sense of secular patterns that are puzzling if election fraud indicates regime weakness, as others have held.
-The pattern depicted in Figure ...
-for example, which shows high levels of manipulation and low levels of protest is sensible in a world where the more powerful and durable incumbents produce greater levels of manipulation, and where incumbents who struggle to produce high levels of manipulation are ones who face resource scarcity, institutional barriers to impunity for their agents, and defections among erstwhile allies.
+The pattern depicted in Figure \@ref(fig:elex-hist) for example, which shows high levels of manipulation and low levels of protest is sensible in a world where the more powerful and durable incumbents produce greater levels of manipulation, and where incumbents who struggle to produce high levels of manipulation are ones who face resource scarcity, institutional barriers to impunity for their agents, and defections among erstwhile allies.
 It also helps explain why some incumbents manufacture their way to highly implausible margins of victory, reasoning that a demonstration of their extra-electoral power may suggest to allies and bystanders that the incumbent may survive protest despite a low natural popularity.
@@
 
-The findings regarding protest initiation are puzzling, however, for any interpretation of regime strength as communicated by manipulation.
-Protest is most likely to be initiated under the same circumstances when it is likely to be defeated quickly, which undermines notions of strategic initiation based on incumbent weakness.
-At the same time, it is very unlikely to be initiated when both manipulation and incumbent vote-shares are low, which confounds the expectations of both views of strategic manipulation.^[If manipulation signals weakness, such an outcome provides an honest signal of the incumbent's low popularity, which should trigger strategic protest.
-If manipulation signals strength, this signal of low genuine popularity is further bolstered by an independent signal of organizational weakness (which should also spark strategic protest).] This finding raises more questions than answers, and it is beyond the scope of this paper to resolve those questions.
-However, taking the results in total suggests that 1) large-scale election manipulation indicates regime strength in most cases, and 2) electoral protest behavior is either non-strategic (not based on the likelihood of winning the contest in the streets) or involves a mix of strategic and non-strategic motives.
+Protest is most likely to be initiated under the same circumstances when it is likely to be defeated quickly, which is puzzling if we assume that protests are initiated strategically (i.e.
+when opposition groups believe the incumbent is weak).
+This finding raises more questions than answers, and it is beyond the scope of this paper to resolve those questions.
+However, taking the results in total suggests that 1) large-scale election manipulation indicates regime strength in most cases, and 2) electoral protest behavior involves a mix of actors with strategic and non-strategic motives.
+Such motives are discussed in more detail in the theory section.
 
-```{=latex}
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-% \usepackage[normalem]{ulem}
-% \useunder{\uline}{\ul}{}
-\begin{table}[]
-\begin{tabular}{@{}l|lll|l@{}}
-\toprule
-                                                                                      & \begin{tabular}[c]{@{}l@{}}Successful \\ protests\end{tabular} & \begin{tabular}[c]{@{}l@{}}Failed \\ protests\end{tabular} & \begin{tabular}[c]{@{}l@{}}P-value \\ (successful vs.
-\\ unsuccessful)\end{tabular} & \begin{tabular}[c]{@{}l@{}}SD \\ (full sample)\end{tabular} \\ \midrule
-\begin{tabular}[c]{@{}l@{}}Intent.
-\\ irregularities\end{tabular}                     & 1.55                                                           & 1.25                                                       & 0.26                                                                                & 0.93                                                        \\
-Inc.
-vote-share                                                                       & 0.50                                                           & 0.59                                                       & 0.08                                                                                & 0.21                                                        \\
-\begin{tabular}[c]{@{}l@{}}Pre-election \\ protests (n)\end{tabular}                  & 7.29                                                           & 6                                                          & 0.25                                                                                & 8.4                                                         \\
-\begin{tabular}[c]{@{}l@{}}Pre-election protests\\ (median participants)\end{tabular} & 3.58                                                           & 3                                                          & 0.1                                                                                 & 0.89                                                        \\
-\begin{tabular}[c]{@{}l@{}}Civil society \\ openness\end{tabular}                     & 0.30                                                           & 0.32                                                       & 0.85                                                                                & 1.08                                                        \\
-Phys.
-int.
-rights                                                                     & 0.45                                                           & 0.43                                                       & 0.81                                                                                & 0.22                                                        \\ \bottomrule
-\end{tabular}
-\caption{Group means for key variables}
-\label{tab:group-means}
-\end{table}
+```{r group-means-table, echo=FALSE, warning=FALSE, message=FALSE}
+group_means <- read.csv(here::here("Data", "group_means_table.csv"))
+group_means <- group_means %>% rename("Variable" = "X")
+group_means <- group_means %>% rename("Successful protest" = "Successful.protests")
+group_means <- group_means %>% rename("Failed protest" = "Failed.protests")
+group_means <- group_means %>% rename("P-value" = "P.value")
+group_means <- group_means %>% rename("SD (full sample)" = "SD..full.sample.")
+
+kable(group_means, booktabs = T, digits = 3, caption = "Table 3: Group means for successful and failed protests, with t-test p-values")  %>%
+    kable_styling(latex_options = "scale_down")
 ```
 
-A very brief consideration of the seven cases of successful electoral protest\textemdash Ukraine (2004), Georgia (2003), Azerbaijan (2005), Kyrgyzstan (2005), Malaysia (2004), Peru (2000) and Cote d'Ivoire (2000)\textemdash in comparison with the other cases can help aid interpretation of the results.
-Table ...
-[NUMBER] shows the mean values for several key variables for four groups: incumbents defeated by protest, incumbents who survived protest, all protested elections, and all unprotested elections.
-The table also shows p-values for t-tests comparing the means of the relevant groups.
-While the level of fraud is typically higher in the successful protest group than the unsuccessful group, and in the protested versus unprotested group, these means are not significantly different.
-By contrast, incumbent vote-shares are significantly lower in the successful-protest and protest groups than their counterparts (at the p < .1 level).
-This suggests that strategic opposition protest entrepreneurs may look primarily to the official vote total as a signal of the incumbent's ability to marshal all its resources, with the level of fraud being of secondary strategic concern.
-In essence, they may be gambling that, under pressure from mass protests and facing low genuine popularity, elites and state officials will split from the incumbent.
-In other words, they may hope that by initiating protest when the incumbent has demonstrated low popularity, the cooperation between principals and agents demonstrated by effective fraud will begin to fray [@Hale2019;@satoEliteCoordinationPopular2019].
-Unfortunately, the duration analysis suggests this is often a losing bet.
-
-The table also shows that the average number of pre-election protests is significantly different from protested versus unprotested elections, and that pre-election protests that drew larger crowds were more common in the successful-protest cases than the unsuccessful ones (at the p < .1 level).
-This suggests an explanation for the unexpected finding that protests are unlikely to be initiated when fraud and incumbent vote-shares are both low, involving the interconnection between strategic and non-strategic protest.
-When protest entrepreneurs are considering initiating protest, they must consider how likely it is that ordinary citizens will join them [@buenodemesquitaRegimeChangeRevolutionary2010].
+Additionally, it is likely that when strategic protest entrepreneurs are considering initiating protest, they must consider how likely it is that ordinary citizens will join them [@buenodemesquitaRegimeChangeRevolutionary2010].
 In the absence of large-scale election manipulation to serve as a coordination device [@kuntz_more_2009;@tucker2007], they may judge that protests will be unsuccessful despite the regime's apparent low popularity.
@@
 Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304].
-Yet this highlights a dilemma for opposition supporters, since it is exactly these scenarios wherein post-election protest is most likely to fail quickly.
+In essence, they may be gambling that, under pressure from mass protests and facing low genuine popularity, elites and state officials will split from the incumbent.
+In other words, they may hope that by initiating protest when the incumbent has demonstrated low popularity, the cooperation between principals and agents demonstrated by effective fraud will begin to fray [@Hale2019;@satoEliteCoordinationPopular2019].
+This kind of gamble may be especially sensible in more deeply authoritarian contexts, where information on incumbent strengths and weaknesses is scarce [@Pop-Eleches2015] and preference falsification more widespread [@Kuran1991].
+Unfortunately, the duration analysis suggests this is often a losing bet, since it is exactly these scenarios wherein post-election protest is most likely to fail quickly.
+Untangling the relationship between strategic and non-strategic protest when election manipulation indicates sources of strength is an important direction for future research.
 
-A further glance at Figure \@ref(fig:margin-prob-fig) illustrates how protest initiation may blend strategic and non-strategic motivations.
-The figure shows the predicted probability of post-election protest for two specific values of election fraud, across all values of incumbent vote-share, holding all other variables constant at their means (continuous variables) or modes (categorical variables).
-The values for election fraud represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively.
-The figure highlights that the predicted probability of protest is only statistically significant when comparing unusual values of fraud at extremely low levels of incumbent vote-share.
-This represents an exceedingly unlikely outcome along both variables.
-As the figure shows, there are only a small number of observations in the range where the difference in predicted probabilities is statistically significant.
-This suggests that electoral protests occur, for the most part, at random with respect to incumbent vote-share and indicators of fraud\textemdash indicators that have often been used as measures of incumbent strength in classic models.
-The extremes where significant effects are found would then correspond to strategic, but risky, gambles by protest entrepreneurs hoping that mass manipulation will help pull regime opponents into the streets, and betting that they will not when manipulation is low.
-Protests that occur outside of this window are more likely to be 'bottom-up,' spontaneous reflections of public grievance.
-Thus the results find support for both the strategic and non-strategic literatures on protest.
+A very brief consideration of the seven cases of successful electoral protest\textemdash Ukraine (2004), Georgia (2003), Azerbaijan (2005), Kyrgyzstan (2005), Malaysia (2004), Peru (2000) and Cote d'Ivoire (2000)\textemdash in comparison with the other cases can help aid interpretation of the results.
+Table 3 shows the mean values for several key variables for two groups: cases where incumbents were defeated by protest, and cases where they survived protest.
+The table also shows p-values for t-tests comparing the means of the two groups.
+While the level of fraud is typically higher in the successful protest group than the unsuccessful group, this difference is not significant.
+By contrast, incumbent vote-shares are significantly lower in the successful-protest and protest groups than their counterparts (at the p < .1 level).
+Similarly, pre-election protests that drew larger crowds were more common in the successful-protest cases than the unsuccessful ones (at the p < .1 level).
+This supports the notion that strategic protest entrepreneurs consider the likelihood that citizens will respond positively to the call to the streets en masse.
 
-Figure \@ref(fig:corfig) shows some support for this idea.
-Protest occurrence is very weakly correlated with measures of opposition strength, like the openness of civil society or the absence of barriers to forming independent parties.
-It is also weakly correlated with measures of incumbent strength, like incumbent vote-share and\textemdash as this paper argues\textemdash election fraud.
-Post-election protest is also weakly correlated with possible indicators of mass grievance, like an economic crisis or corruption in the public sector.
-This could indicate that post-election protests hinge on particular combinations of factors that are difficult to predict a priori.
-It may also be the case that such protests are highly contingent.
-While not an electoral protest, the initiation of the Arab Spring protests offer such an illustration.
-[EXPLAIN MORE]
-
-```{r corfig, fig.cap="Correlation coefficients of selected variables", echo=FALSE, message=FALSE, warning=FALSE}
-df <- nelda2.sub %>% filter(inc.lose == "0") %>% dplyr::select(any_protest, osorg.1lag, partybarriers.lag, winner.share.upd, v2elirreg.inv, econ.crisis.nelda.1lag, pubseccorrup.lag, regional_success_lag)
-df <- df %>% mutate(econ.crisis.nelda.1lag = as.numeric(econ.crisis.nelda.1lag))
-df <- df %>% rename("Any protest" = any_protest,
-                    "Civil soc. open." = osorg.1lag,
-                    "Low party barriers" = partybarriers.lag,
-                    "Inc. vote-share" = winner.share.upd,
-                    "Election fraud" = v2elirreg.inv,
-                    "Econ. crisis" = econ.crisis.nelda.1lag,
-                    "Pub. sec. corrupt." = pubseccorrup.lag,
-                    "Reg. protest" = regional_success_lag)
-M <- cor(df, use = "pair")
-p.cor <- corrplot::corrplot(M, type = "upper", tl.col = 'black',
-         cl.ratio = 0.2, tl.srt = 45, addCoef.col = 'grey50')
-
-```
+The causes of the post-communist cases, which account for most of the seven cases, has been much debated [@LucanWay2008;@Bunce2009].
+Yet for most of these cases, election fraud either held steady after the 'successful' revolution, or increased before partially declining again some years later.
+These cycles, identified as a feature of 'patronal politics' [@hale_patronal_2014] could indicate that in these cases, fraud patterns and protest were caused by more fragmented elites who failed to resolve the principal-agent problems needed to engage in mass protest and control the streets.
+A thorough qualitative comparative analysis of the cases would be informative.
 
@@
 They suggest that factors like the opposition's capacity to organize street protest in limiting election fraud have been overestimated.
-In fact, these results suggest that incumbents who have the resources will respond to such a scenario by *increasing* their production of fraud to secure a higher vote-share.
+In fact, these results suggest that incumbents who have the resources may respond to such a scenario by *increasing* their production of fraud to secure a higher vote-share.
 Consequently, theories of electoral manipulation that focus on factors other than protest risk\textemdash such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]\textemdash may be more important for understanding election integrity, which may drive future research as well as policy interventions.
@@
 But if protest risk is largely untethered from the severity of fraud, it is likely that other mechanisms are at work.
-Scholarly attention might turn more profitably toward other explanations for manipulation, such as the organizational capacity of the state [@simpser2013], the popularity of the government [@rundlett2016deliver], and the risks faced by front-line agents [@Harvey2022].
+For instance, it may be that election observation makes it harder for front-line agents to avoid personal consequences for engaging in manipulation [@Harvey2022].
 
@@
 Election manipulation, including fraud, directly reduces the cost of toleration, by making it less likely that opposition groups win elections.
+This makes incumbents more willing to tolerate opposition in contested, albeit biased, elections.
 To the extent that it makes electoral protest more likely and/or more costly, however, it increases the cost of suppression.
@@
 
-By showing that election fraud does not make protest more likely (most of the time) and that the infrastructure of fraud can actually reduce the cost of post-election protest to the regime, this result flips that logic on its head.
-Instead, election fraud can both reduce the cost of toleration *and* the cost of suppression.
-This implies that electoral authoritarianism is something of a cul-de-sac: the capacity to generate severe election fraud makes it appealing for closed authoritarian regimes to introduce elections, while the manageable risk of protest makes it unappealing to democratize.
+By showing that the infrastructure of fraud can actually reduce the cost of post-election protest to the regime, this result reverses that logic.
+Instead, the ability to produce large-scale election fraud can both reduce the cost of toleration *and* the cost of suppression.
+This implies that electoral authoritarianism is something of a cul-de-sac: the capacity to generate severe election fraud makes it appealing for closed authoritarian regimes to introduce elections, while the manageable risk of protest makes it unappealing to further democratize.
 This logic also creates incentives for ruling parties in weakly consolidated democracies to take measures that will enable them to engage in electoral manipulation; namely, the consolidation of patronage resources and erosion of rule of law.
@@
 
- [MOVE TO APPENDIX? ->]
-This paper is primarily concerned with whether or not election manipulation is indicative of regime weakness or resilience; its primary evidence is the finding that fraud is associated with more resilient incumbents in the face of protest.
-Due to space constraints, two other predictions are not tested in depth here.
-First, one of the key predictions of the protest-deterrent approach is that weaker incumbents will generate little fraud and voluntarily step down after losing elections.
+One key prediction of the strategic protest model is not tested here directly, because the research design focuses on the strength of incumbents against protest movements as revealed by actual protest.
+However, the protest-deterrent model predicts that weaker incumbents will generate little fraud and voluntarily step down after losing elections.
 There is indeed less election fraud observed in those election-years: the mean estimate of *election fraud* in the 129 observations where incumbents lost is 0.322, much lower than the 1.20 average in the group where incumbents claimed victory.
@@
 
-kable(anova_dat, booktabs = T, digits = 3, caption = "Table 5: Mean variable values for winning incumbents minus losing incumbents, with ANOVA p-values")  %>%
+kable(anova_dat, booktabs = T, digits = 3, caption = "Table 4: Mean variable values for winning incumbents minus losing incumbents, with ANOVA p-values")  %>%
     kable_styling(font_size = 9, full_width = FALSE)
@@
 Negative values indicate that the losing incumbents had higher values for the variable, on average.
-For instance, among the cases where the incumbent lost, judicial independence, professionalization of the bureaucracy, and the autonomy of election management bodies are all markedly higher.
-Additionally, the losing group shows lower average economic growth.
-All of these should tend to reduce the availability of fraud.
+For instance, among the cases where the incumbent lost, judicial independence, professionalization of the bureaucracy, economic privatization, civil society openness, and the autonomy of election management bodies are all significantly higher.
+All of these should tend to reduce the availability of fraud as a strategy.
 Of the factors considered here, only *number of pre-election protests* clearly behaves as the deterrence logic would predict, with the losing group subject to one more protest, on average, than the winning group.
 However, unlike the factors which point toward the alternative hypothesis, this difference is not statistically significant.
+Additionally, the winning group is significantly more urbanized, which cuts against the deterrence logic given that more urban environments are more prone to mass protest [@Wallace2013a].
 Thus, while future research might test this aspect of the deterrence model more rigorously, the general pattern is supportive of the view put forth in this paper.
 Further checks are performed in the appendix, including models which show that predicted levels of fraud based on structural factors alone are strongly predictive of incumbent defeat, while pre-election measures of risk are not.
-
-For discussion section:
-
-Maybe add some thoughts on Kuran 1991--the idea being that authoritarian regimes are low info and rife with preference falsification.
-So opp has no choice but to gamble--so strategic in motivation but not in outcome or estimation?
-
-The second question this paper does not closely examine is the role of election fraud in sustaining (or not) the small number of incumbents who are defeated by mass protest.
-These cases are: Peru 2000, Cote d'Ivoire 2000, Georgia 2003, Malaysia 2004, Ukraine 2004, Azerbaijan 2005, and Kyrgyzstan 2005.
-The causes of the post-communist cases, which account for most of the seven cases, has been much debated [@LucanWay2008;@Bunce2009].
-Yet for most of these cases, election fraud either held steady after the 'successful' revolution, or increased before partially declining again some years later.
-These cycles, identified as a feature of 'patronal politics' [@hale_patronal_2014] could indicate that in these cases, fraud patterns and protest were caused by more fragmented elites who failed to resolve the principal-agent problems needed to engage in mass protest and control the streets.
-A thorough qualitative comparative analysis of the cases would be informative, but a sketch can be drawn that supports the theory here.
-
-Ukraine's Orange Revolution, for example, combined very high fraud with a very low incumbent vote-share.
-The analysis of residuals conducted in the appendix finds that fraud in the 2004 election far exceeded expectations based on structural factors.
-Combined with a term-limited incumbent, a strategic opposition may have interpreted these signals as indicating a deeply unpopular ruling party which would struggle to put down a determined protest movement due to a fragmented elite.
-On the other hand, the 2005 Tulip Revolution in Kyrgyzstan came on the heels of moderate election fraud that fell well below structural expectations (i.e.
-large negative residuals).
-This struggle to generate election fraud may have been interpreted as a sign of a weak ruling party, leading to successful protest.
-
-Nevertheless, electoral protest is rare in non-democracies, and successful protest extremely so.
-The overall results indicate that we should be cautious when building and interpreting theories that emphasize the risk of mass protest as a trip-wire to be avoided in non-democracies, and add empirical evidence to Przeworski's [-@przeworskiFormalModelsAuthoritarian2022] critique of such models.
-The rarity of post-election protest, especially successful protest, suggests that such events are highly contingent; they hinge on a multitude of factors that go beyond the regime's efforts to bias election outcomes.
-Both protest and election fraud are complex phenomena with multiple drivers, including resource costs, signaling, information-gathering, principal-agent dynamics, and more; these findings indicate that protest risk is not a practical constraint on fraud under most circumstances.
 
@@
 These incidents demonstrate that, while crying fraud may sometimes be a useful tool for protest entrepreneurs, losing an election alone can be a catalyst for protest, irrespective of the actual level of election manipulation.
-This possibility tends to reduce the 'opportunity cost' to incumbents of engaging in real fraud, while also underlining the fact that the initiation and course of protest is often driven more by the relative resources and opportunities available to the disaffected and the incumbent government.
+This possibility, which would be puzzling if fraud indicates weakness, highlights the often non-strategic nature of protest initiation while also underlining the fact that the course of protest is often driven more by the relative resources and opportunities available to the disaffected and the incumbent government.
 
 These results are consistent with different specifications presented in an online appendix.
-In particular, the main results are supported when data from the Mass Mobilization dataset are used in place of the ECAV data, and when alternative measures of election integrity are used in place of the measure of election fraud.
+In particular, the main results are supported when alternative measures of election integrity are used in place of the measure of election fraud.
 They are also supported when electoral democracies are included in the dataset.
-The relationship between fraud levels and incumbent defeat is tested.
-Additional robustness checks include models with country fixed-effects and alternative specifications for selected control variables.
+The relationship between fraud levels and incumbent defeat is also tested (along with incumbent ouster by protest).
+Additional robustness checks include models with data from election observer reports, with country fixed-effects, alternative specifications for selected control variables.
+
+In sum, the results indicate that we should be cautious when building and interpreting theories that emphasize the risk of mass protest as a trip-wire to be avoided in non-democracies, and add empirical evidence to Przeworski's [-@przeworskiFormalModelsAuthoritarian2022] critique of such models.
+The rarity of post-election protest, especially successful protest, suggests that such events are highly contingent; they hinge on a multitude of factors that go beyond the regime's efforts to bias election outcomes.
+These findings indicate that, while incumbents do seek to contain protest, more effective manipulators are also more capable of defeating protest movements.
 
@@
 Building on the principal-agent literature of election fraud, I argue that large-scale fraud draws on the same resources that make incumbents resilient to protest, even when regime popularity (and thus vote-shares) are low.
-Conducting election fraud requires consolidated control over patronage resources and the ability to protect front-line agents from the consequences of illegal behavior; this capacity also helps insulate incumbents from the effects of mass protest.
+Conducting election fraud is made easier when incumbents have control over patronage resources and the ability to protect front-line agents from the consequences of illegal behavior; this capacity also helps insulate incumbents from the effects of mass protest.
 Resolving this tension in the literature has important implications for the study of election fraud and authoritarian stability.
 
-This paper attempts to test the assumptions of these two models by investigating the relationship between pre-election risk and election fraud, as well as between election fraud and protest occurrence and duration.
-Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 661 elections in electoral authoritarian regimes to show that protest occurrence is most likely when incumbent vote-shares are very low and election fraud is high, and\textemdash most importantly\textemdash incumbents that engage in larger-scale election fraud are able to more quickly defeat protest movements.
+This paper attempts to test the assumptions of these two models by investigating the relationship between election fraud and protest occurrence and duration.
+Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 445 elections in electoral authoritarian regimes to show that protest occurrence is most likely when incumbent vote-shares are very low and election fraud is high, as both models predict.
+However, and most importantly, incumbents that engage in larger-scale election fraud are able to more quickly defeat protest movements.
 
-All together, these results suggest that protest risk largely fails to serve as a constraint on election manipulation in authoritarian regimes.
-The absence of protest risk as a guardrail suggests that mass demonstrations in favor of advancing (or restoring) democratic elections is unlikely to be driven by the severity of manipulation, and that electoral authoritarianism as a regime-type is likely to be a stubborn equilibrium.
+All together, these results suggest that protest risk is unlikely to serve as a constraint on election manipulation in authoritarian regimes\textemdash the most effective election riggers are also the most well-prepared to defeat protest movements.
+The frailty of protest risk as a guardrail suggests that electoral authoritarianism as a regime-type is a stubborn equilibrium.
 This is a gloomy picture for supporters of democratization, and suggests that researchers should consider other mechanisms beyond mass protest in their models of electoral manipulation.
+
+# Data availability statement
+
+The data analyzed in this study will be made available at the Harvard Dataverse upon acceptance of the paper.
 
@@
 # References
-
-### Leftover text
-
-Fraud itself is a sign of the robustness of some of the same tools that can be used to co-opt the opposition, constrain elites, or intensify repression during a crisis; even in the unlikely event that protests do break out in response to large-scale fraud, the incumbent has many tools it may deploy to curtail them.
-
-In this paper, I argue \textemdash in keeping with a growing literature on the informational aspects of elections in authoritarian regimes [@guriev2020;@smyth2020elections;@Geddes1989;@Miller2015a;@Pop-Eleches2015]\textemdash that an election outcome produces a signal of the relative strength of the incumbent.
-As Magaloni [-@magaloni2006] and others have noted, incumbents typically work to produce a higher margin of victory in the election, in order to demonstrate their strength and induce cooperation from other actors in society.
-In some theories, this works best if the votes for the incumbent are largely 'genuine' (i.e.
-not fabricated), or produced using the incumbent's organizational resources (i.e.
-vote-buying and intimidation) [@Harvey2020;@magaloni2006].
-These two kinds of votes correspond to two of the three 'pillars of stability' identified by Gerschewski [-@Gerschewski2013]: legitimation and cooptation.
-Fraud is considered a form of cheap talk in these models, which opposition actors can discount if they become aware of it [@little_elections_2012].
-
-However, these models overlook the fact that engineering widespread fraud requires solving principal-agent problems through control over patronage resources and the criminal and judicial apparatus [@Harvey2019;@Harvey2022].
-As a result, high levels of fraud are a sign of incumbent strength in both cooptation and repression (Gerschewski's third pillar).
-Both of these tools can be useful for defusing or dismantling an emergent protest movement.
-As a result, effective fraudsters should also be more effective at defeating protest movements.
-Consequently, in this view, there is little penalty for severe manipulation in terms of protest risk, and little deterrent effect.
-
-Testing this proposition in more depth is difficult.
-If, for example, incumbents manipulate just the right amount to avoid mass protest in equilibrium\textemdash and are deterred from manipulating more by that risk\textemdash we would observe no statistical relationship between the severity of fraud and the likelihood of protest initiation.
-High and low levels of fraud would generally be associated with no protest.
-Beginning to untangle the relationship between election fraud and protest requires looking beyond standard binary measures of protest.
-Instead, this paper tests outcomes "off the equilibrium path," to use the language of formal modeling.
-Does greater risk to the incumbent in the pre-election period result in reduced election fraud?
-When protest movements do arise after an election, is fraud associated with weaker governments?
-
-Taking an alternative approach, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e.
-whether opponents are allowed to run), rather than by the election-day manipulation that usually figures in theoretical work.
-This result is consistent with the finding that parties are more likely to protest elections after ex ante changes to the rules [@chernykh2014].
-
-The grievance-based approach suggests that protests hinge more on citizens' subjective beliefs about fraud and the broader social context than on the actual level of malfeasance.
-A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b].
-Subjective belief in fraud was also shown to be correlated with approval of mass protests after Russia's 2011 election [@Chaisty2013].
-Furthermore, simply being on the losing side in an election is associated with reduced belief that the election was fair [@Cantu2015], and with increased support for electoral protest [@Anderson2006].
-These studies imply only a limited deterrent effect of protest on fraud in non-democracies.
-If the incumbent may pay the cost of protest whether or not it engages in fraud (due to subjective, and perhaps mistaken, beliefs about fraud and economic conditions), it may as well engage in manipulation in order to gain the benefits.

### 2025-02-27 Scripts/authoritarian_only_latex_2.Rmd
Edits to intro to clarify contribution.

@@
 
+<!--
 The December 2011 elections to the Russian State Duma delivered a significant setback for the ruling party.
@@
 Or, does election manipulation indicate sources of strength for incumbents (independent of genuine popularity), and does protest occur for non-strategic reasons?
+!-->
 
-The stakes of this debate are significant.
+It can be a stirring sight: following an unfair election, thousands of people take to the streets to demand that new elections be held under free and fair conditions.
+In some cases, the scale of the protests and solidarity of their participants can precipitate the collapse of a non-democratic regime; the 2004 Orange Revolution succeeded in preventing the theft of a presidential election in Ukraine, for instance, while large protests in Peru helped prevent President Alberto Fujimori from claiming a third term.
+In other cases, despite protesters' determination, incumbents are able to remain in power through a combination of repression and concessions; failed protest movements in Russia in 2011 and Belarus in 2020 offer stark examples.
+
+Events like these figure prominently in theories of election manipulation and how it might be restrained, with broader implications for the stability of both democratic and authoritarian regimes.
+A classic view, influential in the formal modeling tradition, argues that the use of election manipulation by an incumbent is indicative of the incumbent's unpopularity.
+This lack of popular support, in turn, makes it less likely that an incumbent will be able to defeat a protest movement if the opposition chooses to launch one.
+Such models predict that there is a set of incumbents who require manipulation to reliably win an election, but\textemdash fearing costly protests\textemdash choose instead to gamble on relatively clean elections.
+This framework has normatively optimistic implications, by showing that stronger opposition movements\textemdash those that are more organized, better resourced, and more popular\textemdash will be better able to deter incumbents' from holding rigged elections.
+In turn, this dynamic makes democratic transition and consolidation more likely.
+I refer to this normatively optimistic view as the regime-weakness model of manipulation.
+
+An alternative conceptualization of electoral manipulation has developed in recent years.
+In this view, developed most prominently by Simpser [-@simpser2013], election manipulation is in fact a costly signal of the incumbent's resources and capacities.
+When incumbents manipulate elections, they demonstrate financial, patronage, and organizational resources; they also demonstrate the ability to bend or flout the law.
+In this view, incumbents that manipulate elections effectively do not betray weakness in the form of unpopularity; instead they display alternative sources of resilience that can be used to compensate for low popular approval\textemdash these resources may also be deployed to contain a protest movement after elections.
+I call this the regime-strength model.
+
+The stakes of this debate are significant for theories of democratization and autocratization.
 The former view is an optimistic one; in it, incumbents may restrain fraud when it might be exposed and reveal their weakness to strategic protesters\textemdash an auspicious prediction for democratization and ultimately the under-pinning of 'self-enforcing' democracy [@fearon_self-enforcing_2011].
-If the latter view holds, non-strategic electoral protest may fail to deter election manipulation.
-The risk of mass protest would, in this case, be a fragile guardrail against autocratization.
+If the latter view holds, the risk of electoral protest may fail to deter election manipulation; it would, in this case, be a fragile guardrail against autocratization.
 This debate cannot be advanced by focusing solely on protest occurrence.
@@
 
-In short: there are two views on the nature of electoral protest which make the same prediction.
-If opposition groups strategically initiate protest when they believe the incumbent to be weak, they will be most likely to do so when the election results show a low vote-share for the incumbent combined with high levels of manipulation\textemdash a combination of factors which reveals the incumbent's very low genuine popularity [@Luo2017;@Little2012;@Schedler2002;@Beaulieu2014a].
-However, if protest movements are more organic and non-strategic, reflecting citizen grievances, we should expect protest initiation in the same circumstances.
-Here the combination of low incumbent vote-share and high manipulation indicates severe citizen grievances (low natural popularity) which are galvanized by the shared grievance of a 'stolen' election [@tucker2007].
-Since both theories make the same prediction about protest initiation, such tests\textemdash which have been the dominant empirical measure in the literature\textemdash do not reveal whether protest is strategic and manipulation indicates weakness or if manipulation is a hallmark of strong incumbents [@simpser2013] and protest is non-strategic.
+At first glance, it would appear straightforward to test these competing models.
+The regime-weakness model typically predicts that increasing manipulation (at a constant level of true popularity) increases the likelihood that opposition groups will initiate post-election protest @Luo2017;@Little2012;@Schedler2002;@Beaulieu2014a.
+A low total vote-share amid severe manipulation is the strongest possible signal of incumbent weakness; it has had to rely on the 'cheap talk' of manipulation to produce paltry results, indicating that the incumbent's genuine popularity is very low and its likelihood of defeating the protest movement is likewise small.
+By contrast, the regime-strength model would yield the opposite prediction [@simpser2013]; that protest initiation decreases with the level of manipulation (holding vote-share constant).
+Under the regime-strength model, the most favorable  election outcome for protesters is one where the incumbent's vote-share and manipulation level are low, a combination which indicate low natural popularity and diminished resource capacity.
 
-To unravel this question, I take this analysis one step further.
-In strategic-protest models, the opposition protests because it believes the probability of a pro-government outcome of the protest is low.
-I assume that the duration of protest movements, when they occur, is an indicator of this latent likelihood of a pro-government outcome.
+There are two problems, however.
+First, the empirical record on the relationship between manipulation and protest been mixed, illustrating the limitations of current theory.
+[FILL IN CITES HERE] Second, the expected relationship between manipulation and protest under the regime-weakness and regime-strength models is complicated by the emergence of a grievance-based model of electoral protest generally.
+This view, which has considerable empirical support, finds that electoral protests are largely driven by citizen grievances\textemdash not the narrowly strategic calculations about the likelihood of overthrowing the government which are assumed in the regime-weakness model.
+Under the grievance framework, when citizens have many individual grievances against the incumbent, a highly manipulated election can serve as a coordination point for collective action [@tucker2007].
+
+This theoretical complexity has frustrated our ability to distinguish between the regime-strength and regime-weakness models because they can produce the same hypotheses regarding protest initiation depending on which model of protest is true.
+If protest movements are narrowly strategic and manipulation indicates weakness, we will expect that the likelihood of protest will increase with electoral manipulation.
+On the other hand, if manipulation indicates strength we will still expect greater manipulation to make protest more likely\textemdash if electoral protests are largely driven by shared grievance rather than strategic considerations.
+Since the same outcome could support two theoretical models, we are at an impasse.
+
+To move the debate forward, I model electoral protest *duration* as a dependent variable, in addition to initiation.
+By contrast to protest initiation, the regime-weakness and regime-strength models generate distinct predictions with regard to protest duration.
+Once protests begin, the 'strength' of an incumbent (its likelihood of outlasting a protest movement) is no longer latent.
+Instead, the incumbent's response\textemdash repression, concessions, or attrition\textemdash and its survival (or not) reveal the incumbents' true strength relative to the opposition.
 The more quickly a government puts down protests, the greater the underlying (unobserved) distribution of power favors the incumbent over the opposition.
@@
 
-Using data from 445 non-democratic elections across 97 countries, from 1989 to 2012, this paper finds that 1) protest initiation is most likely when manipulation is severe but incumbent vote-share is low, as both theories predict, and 2) that governments that engage in more manipulation are likely to put down protests more quickly than governments that produce little manipulation, especially when incumbents claim small vote-shares.
+[Somewhere cite this article: https://doi.org/10.1177/0032321716674024 on how 'toleration' is really just slow attrition of protest movements.]
 
-This pattern in the results is consistent with the argument that, rather than being a sign of weakness, the ability to generate significant electoral fraud is an indicator of durable regime strength.
-It is also consistent with the argument that much post-election protest in non-democracies is non-strategic; i.e., the actions of a weak opposition against a strong incumbent.
+Using data from 445 non-democratic elections across 97 countries, from 1989 to 2012, this paper finds that 1) protest initiation is most likely when manipulation is severe but incumbent vote-share is low, and 2) that governments that engage in more manipulation are likely to put down protests more quickly than governments that produce little manipulation, especially when incumbents claim small vote-shares.
+This pattern of results in inconsistent with the regime-weakness model of election manipulation, but consistent with a world in which election manipulation indicates sources of regime resilience and protests are largely motivated by grievance.
 Given these results, it is likely that\textemdash while incumbents do not seek to provoke protest\textemdash the downside risk for incumbents that engage in election fraud in non-democracies has been overemphasized.

### 2025-03-03 Scripts/authoritarian_only_latex_2.Rmd
Edits to lit review: deterrence logic based on 2 assumptions

@@
 
-# Literature review: three views of election manipulation and mass protest
+# Literature review: two assumptions about election manipulation and mass protest
 
-In classic models of election manipulation, manipulation is a short-term tool used instrumentally to secure victory in the election [@lehoucq2002].
-Much of the work in this tradition has been elaborated through formal models.
-Though there are, of course, many variations, in several such models incumbents are characterized as 'weak' or 'strong' based on their true support in the electorate.
+As sketched briefly above, a fruitful literature has examined the connections between election integrity and mass protest.
+Research in the regime-weakness model of manipulation typically concludes that as incumbents become increasingly fearful of street protest, they dial back on manipulation\textemdash a potentially pro-democracy outcome.
+These models hinge on two assumptions, as discussed below.
+First, election manipulation is conceptualized as a short-term tool used instrumentally to secure victory in the election [@lehoucq2002]; it communicates only that the incumbent could not count on winning the election cleanly.
+In such models incumbents are typically characterized as 'weak' or 'strong' based on their true support in the electorate: strong incumbents are those that are likely to defeat protest movements, while weak incumbents are those that are likely to be ousted.
+
 This true level of support can be obscured prior to elections due to pervasive preference falsification [@Kuran1991] and the low level of political information common in authoritarian regimes [@guriev2020].
@@
 1153).
-In equilibrium, this can lead weaker incumbents to step down rather than face protest due to massive fraud\textemdash a pro-democracy outcome.
+In equilibrium, this can lead weaker incumbents to step down rather than face protest due to massive fraud\textemdash a pro-democracy outcome in which stronger opposition groups deter manipulation by incumbents.
 Similar results are developed by Chernykh and Svolik [-@chernykh2015third], Luo and Rozenas [-@Luo2020] and [-@Luo2017], and Magaloni [-@magaloni2010].
 
-A related class of formal models holds that fraud operates as a distortion of the signal of the incumbent's true strength.
+Some more recent models treat election manipulation as a distortion of the electoral signal of the incumbent's true strength, but are still located within the regime-weakness framework.
 In this type of model, such as those developed by Little [-@littleFraudMonitoringNoncompetitive2015] or Egorov and Sonin [-@Egorov2021], fraud may be used by incumbents to bluff their opponents into being more supportive than they would be if they knew the incumbent's true popularity.
-However, greater transparency makes such bluffs more costly.
+However, greater transparency makes such bluffs more costly, by making it easier for opposition actors to distinguish signals of true strength (i.e.
+genuine popularity) from noise (i.e.
+manufacured votes).
 The typical result in these models is that protest risk can *partially* deter fraud.
 In equilibrium, the incumbent will generally choose the level of fraud anticipated by the opposition and no more\textemdash but also no less.
+In sum, regime-weakness models of fraud typically hold that the level of election manipulation carries no signal of the incumbent's ability to defeat protesters; on the contrary, it only carries signals of weakness.
 
-Fundamentally, both set of models share the assumption that protest is a centralized, strategic decision by opposition groups, and that fraud is a tool employed by weak (i.e.
-unpopular) incumbents.
-As a result, they predict that the highest risk of protest should obtain when incumbents use massive fraud but still produce a low incumbent vote-share, indicating that the incumbent's true popularity is very low.
-This, in turn, indicates to opposition actors that protest is more likely to succeed in unseating the incumbent.
+The second assumption in the regime-weakness framework concerns the nature of opposition protest.
+The framework is tightly linked to a view of electoral protest that is a centralized decision by rational actors pursuing narrowly strategic goals.
+As a result, they predict that the likelihood of protest should be highest when incumbents use massive fraud but still produce a low incumbent vote-share.
+This signal of weakness, in turn, indicates to opposition actors that protest is more likely to succeed in unseating the incumbent.
 Incumbents in these models typically face a tradeoff between the short-term benefits of electoral manipulation for winning the election, and any weakness that relying on fraud reveals to the opposition [@Rozenas2016b].
 As Beaulieu [-@Beaulieu2014a, p.
-48] writes in this tradition, "As the opposition grows stronger, the amount of [manipulation] he will tolerate shrinks, and the incumbent who wishes to avoid protest will need to reduce the amount of manipulation she commits." These are normatively optimistic models upon which a sizable literature has been founded: fraud can be kept in check, and ultimately reduced, through the development of institutions that reveal information about election integrity: election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].
+48] writes in this tradition, "As the opposition grows stronger, the amount of [manipulation] he will tolerate shrinks, and the incumbent who wishes to avoid protest will need to reduce the amount of manipulation she commits."
+
+These are normatively optimistic models upon which a sizable literature has been founded: manipulation can be kept in check, and ultimately reduced, through the development of institutions that reveal information about election integrity: election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].
 These institutions improve election integrity, it is claimed, by increasing the likelihood that fraud will be observed and thus, incumbent weakness exposed.
 
-However, an alternative set of models have also been advanced.
-A second worldview argues that election manipulation can indicate a well-resourced, well-organized, powerful incumbent (independent of its popularity).
-Under this view, extensive election manipulation can be a show of strength, which induces cooperation with the incumbent from other societal actors.
-Since this view holds that more manipulation can further entrench the incumbent, it is a normatively pessimistic view.
-
-Most persuasively articulated by Simpser [-@simpser2013], this view holds that election manipulation demonstrates incumbent strength.
-As he writes, an incumbent that demonstrates the ability to engage in widespread manipulation "is also likely to have the resources, capacities, and inclinations to overcome or punish opponents, reward supporters, and circumvent the law" [@simpser2013, p.
-6].
-A relatively popular incumbent ("strong" in the language used by the first set of models) may nonetheless manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver].
-By demonstrating otherwise difficult-to-observe capabilities, large-scale election manipulation can solve a coordination problem for the incumbent, securing the cooperation of elites [@simpser2013].
-This can, in turn, fix expectations that the incumbent will continue to rule in the future, further enhancing its staying power [@hale_patronal_2014].
-This argument is in keeping with Seeberg's [-@Seeberg2019] finding that higher-capacity states are associated with more election fraud.
-
-If election manipulation can be an indicator of regime strength, it may serve to deter protest under some circumstances, rather than the other way around [@simpser2013].
-This leaves protest, when it does occur in the presence of large-scale manipulation, as something of a puzzle if protest behavior is largely strategic.
-For this reason, the manipulation-as-strength view comports more easily with another established view of the causes of electoral protest.
-Researchers in this tradition have argued that electoral protest occurs largely unrelated to strategic concerns about the relative strength of the incumbent and the opposition.
-In these models, electoral protest occurs because manipulation galvanizes citizens who hold diverse grievances, because economic conditions are poor, or simply because the bare fact of holding an election creates a moment of opportunity for opposition organization irrespective of the level of manipulation [@Pop-Eleches2015].
-The central feature of these models is that protest decisions are generally made without consideration for the likelihood that protests will overthrow the incumbent.^[To be clear, this is the narrow sense in which 'strategic' and 'non-strategic' is used here and in protest-deterrent models.
-Protest behavior may still be rational in a broader sense without being narrowly tied to the expected likelihood of ousting the government.
-For instance, protesters may believe it is unlikely they will be individually punished for participating in protests, making it individually rational for them to voice their grievances.]
+However, there is also a body of research that challenges the two assumptions upon which the regime-weakness model is founded.
+The assumption on the nature of protest as centralized and narrowly strategic, in particular, has been challenged in the literature on electoral protest.
+Recent research in this tradition has argued that electoral protest occurs largely unrelated to strategic concerns about the relative strength of the incumbent and the opposition.
+In these models, electoral protest occurs because manipulation galvanizes citizens who hold diverse grievances [@tucker2007], because economic conditions are poor [@Brancati2016;@Beaulieu2014a], or simply because the bare fact of holding an election creates a moment of opportunity for opposition organization irrespective of the level of manipulation [@Pop-Eleches2015].
+The central feature of these models is that protest decisions are generally made without consideration for the likelihood that protests will overthrow the incumbent.
 
@@
 
-Rational protest entrepreneurs might seek to initiate protest to take advantage of this popular discontent for non-strategic reasons, in order to press for improved competitiveness in the *next* election or to bargain for policy concessions [@chernykh2014;@robertson_protest_2010].
+Rational protest entrepreneurs might seek to initiate protest to take advantage of this popular discontent for broadly strategic reasons.
+They may wish to pressure the incumbent in order to push for improved competitiveness in the *next* election or to bargain for policy concessions [@chernykh2014;@robertson_protest_2010].
 They may stage protests in order to identify new participants, forge new social ties [@barberaModelProtestsRevolution2019] and test new tactics.
@@
 Or, a vanguard of protesters may take to the streets simply because their distaste for the government outweighs any risk of sanction [@lohmannDynamicsInformationalCascades1994].
-These and other motivations for initiating protest are theoretically significant because their rationality (for protesters) does not hinge on the likelihood of overturning the government, as classic models assume.
+These and other motivations for initiating protest are theoretically significant because their rationality (for protesters) does not hinge on the likelihood of overturning the government, as regime-weakness models of manipulation assume.
 As a result, it is possible that we may observe the initiation of anti-government electoral protest even when the likelihood of such protest overturning the election (or the government itself) is very low.
 
-Empirical work has found some support for the microfoundations of this theory.
-For example, ballot-stuffing and voter coercion have been shown to generate stronger feelings of anger than pre-election manipulation of ballot access [@Szakonyi2021], and access to information about manipulation appears to reduce support for the ruling party [@Reuter2021], increase support for protest and reduce the perceived legitimacy of the regime [@Williamson2021].
-This argument has found support in studies of the relationship between economic grievances, manipulated elections, and protest [@Rod2019;@Brancati2016].
+Finally, an alternative theory of election manipulation challenges the assumption that an incumbent's ability manufacture votes communicates nothing about its ability to defeat a protest movement.
+This view, most persuasively articulated by Simpser [-@simpser2013], holds that increasing levels of election manipulation can indicate a well-resourced, well-organized, powerful incumbent (independent of its popularity); these resources and capacities, in turn, may help the incumbent weather a protest movement.
+Under this view, extensive election manipulation can be a show of strength, which induces cooperation with the incumbent from other societal actors and may serve to deter protest [@simpser2013].
+Since this view holds that more manipulation can further entrench the incumbent, it is a normatively pessimistic one.
 
-Taking this all together, it becomes clear that if researchers focus only on protest occurrence as a dependent variable, they cannot distinguish between two possible data-generating processes: one in which manipulation signals weakness and electoral protest is a strategic response, and one in which electoral manipulation indicates regime strengths and protest occurs for a variety of reasons with limited connection to the odds of removing the government.
-In the former, protest is predicted to occur most often when manipulation is high and the incumbent's vote-share is low, since this provides a strong signal that the incumbent's true popularity is low.
-In the latter, protest is most likely in the same condition because the incumbent's low popularity indicates widespread grievances, which the high level of manipulation serves to coordinate despite the incumbent's organizational strengths.
-Which of these data-generating processes predominates is of significant theoretical importance of understanding authoritarianism and democratization, since the latter view suggests much dimmer prospects for democratization than the former.
+As Simpser writes, an incumbent that demonstrates the ability to engage in widespread manipulation "is also likely to have the resources, capacities, and inclinations to overcome or punish opponents, reward supporters, and circumvent the law" [@simpser2013, p.
+6].
+A relatively popular incumbent ("strong" in the language used by the first set of models) may nonetheless manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver].
+By demonstrating otherwise difficult-to-observe capabilities, large-scale election manipulation can solve a coordination problem for the incumbent, securing the cooperation of elites [@simpser2013].
+This can, in turn, fix expectations that the incumbent will continue to rule in the future, further enhancing its staying power [@hale_patronal_2014].
+This argument is in keeping with Seeberg's [-@Seeberg2019] finding that higher-capacity states are associated with more election fraud.
+
+We can thus consider four possible states of the world, based on which of the foregoing assumptions are true in fact: one where both assumptions hold, one where neither holds, and two where only one assumption holds.
+The normatively optimistic view of the relationship between election manipulation and protest requires both assumptions to be true: electoral manipulation must communicate little information about incumbents' ability to outlast protest, and protest initiation must be narrowly strategic.
+If either one of these assumptions fails to hold, the deterrence logic is undermined.
+However, empirically distinguishing between these four possible states of the world is further complicated by the fact that work on this subject, whether empirical or purely formal, has largely utilized a binary measure of protest occurrence as the outcome of interest.
+Unfortunately, as Table 1 shows, three of four possible states of the world make the same prediction: that increasing manipulation will result in an increased likelihood of protest.
+
+```{=latex}
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+\begin{tabular}{@{}llll@{}}
+\toprule
+\multicolumn{1}{c}{\textbf{Protest type}} &
+  \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Likelihood\\ of \\ Protest\end{tabular}}} &
+  \multicolumn{1}{c}{\textbf{Mechanism}} &
+  \textbf{Example sources} \\ \midrule
+\multicolumn{1}{l|}{} &
+  \multicolumn{3}{c}{Regime-weakness model} \\ \cmidrule(l){2-4} 
+\multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Narrowly \\ strategic\end{tabular}} &
+  Increases &
+  \begin{tabular}[c]{@{}l@{}}More manipulation indicates \\ lower true popularity, and thus \\ lower likelihood that \\ incumbent will survive protests.\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Luo and Rozenas (2020); \\ Chernykh and Svolik (2015); \\ Little, Tucker, and \\ LaGatta (2015)\end{tabular} \\
+\multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Broadly \\ strategic\end{tabular}} &
+  Increases &
+  \begin{tabular}[c]{@{}l@{}}More manipulation employed to \\ compensate for deeper \\ grievances; 'stolen' election as \\ trigger.\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Beaulieu (2014); \\ Brancati (2016); \\ Tucker (2007)\end{tabular} \\ \midrule
+\multicolumn{1}{l|}{} &
+  \multicolumn{3}{c}{Regime-strength model} \\ \cmidrule(l){2-4} 
+\multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Narrowly \\ strategic\end{tabular}} &
+  Decreases &
+  \begin{tabular}[c]{@{}l@{}}More manipulation indicates \\ capacities that increase \\ likelihood incumbent survives\\ protest, offsetting low popularity.\end{tabular} &
+  Simpser (2013) \\
+\multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Broadly \\ strategic\end{tabular}} &
+  Increases &
+  \begin{tabular}[c]{@{}l@{}}More manipulation employed to\\ compensate for deeper \\ grievances; 'stolen' election as\\ trigger.\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Beaulieu (2014); \\ Brancati (2016);\\ Tucker (2007)\end{tabular}
+\end{tabular}
+\caption{Table 1: Predictions for protest initiation under the regime-weakness model of electoral manipulation}
+\label{tab:protest-expectation}
+\end{table}
+```
 
@@
 
-Since the empirical record is mixed, and two theories make similar predictions with regard to protest initiation, it is necessary to generate new hypotheses that can distinguish between the two models of election manipulation and protest that have been advanced.
-In particular, do post-election protest movements encounter weak incumbents as signaling models predict?
-If not, the deterrence model is undermined.
+Nevertheless, there are considerable theoretical and substantive stakes in knowing which of these four possible data-generating processes is most likely; in only one does increasing protest risk actually improve electoral integrity.
+Since the empirical record is mixed, and three of the four possible models make similar predictions with regard to protest initiation, it is necessary to generate new hypotheses.
+In particular, the remainder of this manuscript will attempt to distinguish between regime-strength and regime-weakness models of manipulation by developing hypotheses about protest *duration* under each framework.
+Specifically, I will argue that protest duration helps reveal the true relative balance of power between incumbents and opposition actors.
+
+<!-- NOT SURE WHERE THIS GOES:
+Empirical work has found some support for the microfoundations of this theory.
+For example, ballot-stuffing and voter coercion have been shown to generate stronger feelings of anger than pre-election manipulation of ballot access [@Szakonyi2021], and access to information about manipulation appears to reduce support for the ruling party [@Reuter2021], increase support for protest and reduce the perceived legitimacy of the regime [@Williamson2021].
+This argument has found support in studies of the relationship between economic grievances, manipulated elections, and protest [@Rod2019;@Brancati2016].
+-->
 

### 2025-03-05 Scripts/authoritarian_only_latex_2.Rmd
Good clarifications to theory section; end of major post CPS revisions.

@@
 
-Along with the political benefits it may provide, election manipulation is generally characterized as having both operational costs (which the incumbent may be unwilling or unable to pay [@Greene2010]) and legitimacy costs which reduce other actors' level of cooperation with the incumbent [@birch2011].
+Election manipulation is generally characterized as having legitimacy costs which reduce other actors' level of cooperation with the incumbent [@birch2011].
 The most visible, and arguably most studied, form of such legitimacy costs is mass post-election protest against the incumbent.
-Here, I follow the broad consensus in the literature that authoritarian incumbents find protest undesirable, since they entail some risk of incumbent defeat and may require costly repression or concessions.
-I assume, as a result, that incumbents seek to demobilize protest; protest movements that survive longer indicate the incumbent is relatively weaker compared to the opposition.
-By contrast, when protest movements meet a quick demise, the incumbent can be considered relatively strong.^[In both cases, this assume that incumbent survives the protest movement, by far the modal outcome.
-When the opposition wins, an outcome considered in some of the empirical models below, the assumption is reversed: a quicker opposition victory indicates a weaker incumbent.] How might election manipulation be connected to the incumbent's ability to quickly defeat a protest movement?
+Here, I do not challenge the broad consensus in the literature that authoritarian incumbents find sustained, sizable protests undesirable, since they entail some risk of incumbent defeat and may require costly repression or concessions.
+I assume, as a result, that incumbents generally seek to demobilize protest; protest movements that survive longer indicate the incumbent is relatively weaker compared to the opposition.^[The implications of protest toleration by incumbents will be discussed later.] By contrast, when protest movements meet a quick demise, the incumbent can be considered relatively strong.^[In both cases, this assume that incumbent survives the protest movement, by far the modal outcome.
+When the opposition wins, an outcome considered in some of the empirical models below, the assumption is reversed: a quicker opposition victory indicates a weaker incumbent.] As argued below, electoral manipulation can be demonstrative of capacities that may also be used to defeat protest movements.
 
-Prior models of electoral protest by strategic opposition actors, discussed above, treat election manipulation and genuine votes as substitute goods.
+Regime-weakness models of electoral manipulation, discussed above, treat election manipulation and genuine votes as substitute goods.
 For any given vote-share, an increase in manufactured votes reduces the number of genuine votes.
-For classic models, which consider manufactured votes to be merely illusions of strength, the higher the number of manufactured votes for any given vote total, the weaker the incumbent.
+In these models, which consider manufactured votes to be merely illusions of strength, the higher the number of manufactured votes for any given vote total, the weaker the incumbent.
 This means that strategic opposition actors are more likely to initiate protest when they perceive the incumbent to be weaker, and expect to be more likely to succeed in a challenge.
-
-In the classic model, the incumbent's (unknown) true popularity is positively related to its vote-share in the election, and negatively related to its use of fraud.
-This dynamic is the fulcrum upon which these models hinge: less popular incumbents (who are expected to be less likely to defeat an opposition protest) must manipulate more in order to win elections.
-Popular incumbents need manipulate less.
-Opposition actors gauge the probability of an pro-government protest outcome by estimating the incumbent's true popularity based on the degree of manipulation and the incumbent's vote-share in the election.
 A high vote-share with low fraud indicates a more popular incumbent, likely deterring protest.
@@
 
-Thus the predicted relationship between manipulation and protest likelihood in the classic model is positive: more manipulation makes protest more likely.
-However, as increasing vote-share increases the number of genuine votes for the incumbent at any fixed level of fraud, the effect of fraud should be conditional on the claimed vote-share.
-As Rozenas [@Rozenas2016b] notes, if an incumbent wins an election with low manipulation, it generates a strong signal of its popularity; if it wins a heavily manipulated election, this signal is weaker.
-As an illustration, if the incumbent's total vote-share is 70 percent, for example, of which the opposition estimates 15 percent to be manufactured, the incumbent's genuine popularity is estimated at a respectable 55 percent.
-By contrast, if the incumbent posts a 50 percent margin of victory with the same 15 percent estimated to be manufactured, the opposition may perceive that the incumbent is weak in its genuine support (35 percent).
-This produces a first, baseline hypothesis, derived from the classic model.
+Thus the predicted relationship between manipulation and protest likelihood in the regime-weakness model is positive: more manipulation makes protest more likely for a given vote-share.
+However, the inverse is also to be expected: as incumbent vote-share increases for any fixed level of fraud, the opposition should perceive a stronger incumbent and be less likely to protest.
+In sum, the effect of fraud should be conditional on the claimed vote-share [@Rozenas2016b].
+This produces a first, baseline hypothesis.
 
@@
 
-It is important to note that the same hypothesis is also generated by non-centralized, non-strategic models of election protest.
+It is important to note that the same hypothesis is also generated if we assume that protest initiation is broadly, rather than narrowly, strategic.
 The mechanism, however, is different.
-Here, the combination of low incumbent vote-share and high manipulation is indicative of widespread grievances coordinated a 'stolen' election.
-These are conditions that should be most likely to produce protest [@tucker2007].
-By contrast, as vote-share increases for a set level of fraud, the incumbent's genuine popularity is higher and thus grievances are lower (reducing the likelihood of protest).
+Here, the argument is that widespread grievances result in low natural popularity for the incumbent; at election time, the incumbent must compensate for this low popularity by manipulating the election.
+Protest initiation is made more likely both by the population's grievances alone [@Beaulieu2014a;@Brancati2016], and by the coordination device provided by the 'stolen' election [@tucker2007].
+As natural vote-share increases for a fixed level of manipulation, the likelihood of protest initiation should decline due to reduced grievances; for a fixed level of popularity, increasing manipulation should increase the likelihood of protest by raising the plausibility of a stolen election among dissatisfied citizens.
 
-Integrating Simpser's argument with recent research on the principal-agent dynamics inherent to election manipulation provides different insights.
-Where Simpser argues that manipulation should be blatant and excessive (that is, producing a large margin of victory) to send signals of strength, the principal-agent framework for understanding manipulation suggests otherwise.
+We should only expect H1 to be falsified in the event that protest is driven by narrowly strategic concerns *and* election manipulation is indicative of regime strength.
+In this case, opposition actors should be deterred from launching protest movements as increasing manipulation indicates deeper regime resilience (in the form of resources, weak rule of law, loyalty of agents, and so on).
+The regime-strength model, as I call it, is defended in more detail below.
+
+## Developing the regime-strength model of electoral manipulation
+
+Simpser's foundational argument in the regime-strength model holds that manipulated elections can send costly signals of the incumbent's "capacities, resources, [and] dispositions" that will help it retain power in the future [@simpser2013, p.
+80]\textemdash that is, of its political strength.
+Additional research focusing on the principal-agent dyanmics inherent in election manipulation reinforces the idea that manipulation carries such costly signals, such that more extensive electoral manipulation indicates greater regime strength.
 This research trend emphasizes the degree to which political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].
-Factors that minimize principal-agent problems have been shown to facilitate greater election manipulation.
-These include the incumbent's local popularity [@rundlett2016deliver], the incumbent's ability to distribute patronage resources [@Harvey2019;@hale_patronal_2014], and low risk of punishment for pro-regime lawbreaking [@Harvey2022].
-As discussed below, these are also independently useful for suppressing post-election protest.
+
+Electoral manipulation has been shown to be more extensive under conditions that minimize that minimize principal-agent problems (i.e.
+those that make low-level agents more likely to cooperate with the incumbent).
+These include: greater local popularity for the incumbent [@rundlett2016deliver], greater incumbent control over patronage resources [@Harvey2019;@hale_patronal_2014], and lower risk of punishment for pro-regime lawbreaking [@Harvey2022].
+As discussed below, these capacities and resources are also independently useful for surviving post-election protest.
 Consequently, if an incumbent is able to manufacture a large numbers of votes, it is also likely to have the resource capacity, control over bureaucratic institutions, and perhaps deep wells of popularity in parts of the country that can serve as independent sources of regime durability, even if the incumbent's nationwide margin of victory is low.
+Lastly, extensive manipulation also indicates that large number of agents assess that the incumbent will remain in power in the future [@simpser2013], which itself becomes a self-reinforcing equilibrium [@hale_patronal_2014].
 
-The initial emergence of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014].
+These resources and capacities improve the incumbent's ability to respond to a nascent protest movement through repression or concessions.^[Toleration, another possibility, is discussed later.] [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014].
 Concessions may be aimed at protesting parties themselves [@reuter2015;@mooreRepressionDissentSubstitution2000;@Hafner-Burton2016;@Piven1979] or the more moderate public, in order to keep them on the sidelines of the conflict [@Hummel2019].
@@
 The ability to protect regime agents from any legal consequences for carrying out orders, a condition which helps enable illegal election manipulation [@Harvey2022], is also an asset when asking law enforcement agents to carry out possibly violent repression [@MAGALONI_RODRIGUEZ_2020;@stavro_police_2023;@abouharbFactoJudicialIndependence2013].
+Additionally, more consolidated control over patronage resources can help incentivize elite unity [@Reuter2019;@hale_patronal_2014;@Arriola_2009], which can enable repression by the security forces [@hendrix_2017;@Frugé_2019;@Albrecht_Ohl_2016].
 
@@
 Instead of signalling weakness, large-scale fraud indicates a well of resilience: control over resources, impunity for pro-regime lawbreaking, and loyal agents.
-As a result, governments that build and deploy an electoral manipulation apparatus are simultaneously innoculating themselves against the cost of protest.
+As a result, governments that build and deploy an electoral manipulation apparatus are simultaneously inoculating themselves against the expected cost of protest.
 
-If election manipulation is an indicator of underlying capacities the incumbent may use to defeat protest movements, more severe electoral manipulation will be associated with shorter-lived protest waves.
-Manufactured votes are not negatively associated with genuine popularity, but positively associated with impunity for agents and control over patronage resources.
-They are also positively associated with incumbent popularity [@rundlett2016deliver].
-Abundant resources and extensive impunity reduce the principal-agent problems that frustrate election manipulation efforts.
-These same resources, institutions, and loyal agents make protest outcomes less likely to be favorable for the opposition.
-Note that, in this model, protest initiation is not tied to expected protest outcome; protest is driven by non-strategic concerns related to the incumbent's popularity (i.e.
-citizens' diverse grievances) and whether manipulation is perceived as having been decisive (i.e.
-a stolen election).
-Incumbents with the resources and institutions needed to produce widespread election manipulation can draw on those same capacities to buy off, repress, or otherwise quash post-election protest.
+Putting it all together: this theory assumes that authoritarian incumbents generally prefer order to unrest, and that protesters\textemdash who under either a narrow or broadly strategic view are aiming for some kind of concession from the incumbent\textemdash want to demonstrate the popularity of their cause through large, sustained protests.
+Electoral protests are a form of competition between the incumbent and opposition actors, in which both sides are trying to demonstrate their strength in order to win concessions from\textemdash or defeat outright\textemdash the other [@Beaulieu2014a].
+The outcome of this contest reveals the underlying strengths of the participants.
+Consequently, governments that quickly defeat protest movements can be considered stronger than those that take weeks or months to prevail, which in turn can be considered stronger than those governments which collapse in the face of mass protest.
+Incumbents with the resources and capacities needed to produce widespread election manipulation can draw on those same tools to buy off, repress, or otherwise quash post-election protest.
 This yields Hypothesis 2.
@@
 
-The classic model holds that manipulated votes correlate negatively with incumbents' post-election strength, implying that when incumbents rely heavily on election manipulation they will struggle to put down post-election protests.
-This effect should likewise be conditional on incumbent vote-share, since those incumbents who produce small margins of victory while relying heavily on manipulation will have limited ability to push back successfully against protest movements.
-A higher likelihood of pro-government protest outcome reduces the likelihood that the opposition will initiate protest.
-Consequently, classic models predict that more election manipulation should be associated with longer-lived protest movements against surviving incumbents.^[In the appendix, I check the separate prediction that manipulation should be positively associated with incumbents being ousted by protest movements.]
-
 The argument can be summarized as follows.
@@
 The same resources that can be used to manipulate elections on a wide scale\textemdash including natural popularity, resource control, and control over the legal apparatus\textemdash also better enable incumbents to dismantle protest movements when they do emerge.
-As a result, more extensive manipulation should be associated with shorter-lived protest movements against incumbents.
-Much election-related protest, then, is unrelated to strategic concerns about the likelihood of protest success.
+As a result, more extensive manipulation should be associated with shorter-lived protest movements against incumbents.^[In the appendix, I check the separate prediction that manipulation should be positively associated with incumbents being ousted by protest movements.]
+
+One challenge to this theory is the prospect that incumbents may choose to tolerate protests, rather than attempting to repress them or co-opt their participants.
+This concern should be modest for the cases being considered here.
+First, toleration is most likely when protests impose few costs on the regime and make only targeted demands [@kleinDynamicsPoliticalProtests2018].
+National elections are highly politically salient, and challenging their validity gets to the core of electoral authoritarian incumbent's claim to authority.
+As such, tolerating or ignoring protest is unlikely in this setting.
+In such cases, we should expect to see apparent 'toleration' displaced by 'attrition'\textemdash efforts by the incumbent to increase the cost of protest while avoiding open repression [@yuenNeitherRepressionConcession2017].
+In other words, even in the absence of explicit repression or concessions, we should still expect incumbents to attempt to defuse protest movements aimed at challenging the most recent election.
+
+Testing H1 and H2, together, will help to evaluate the regime-weakness model of electoral manipulation.
+If H1 is supported and H2 is not, it would provide evidence in favor of the regime-weakness model; manipulation would appear associated with incumbent weakness, and protest initiation would appear to take advantage of such.
+On the other hand, if both hypotheses are supported, it would endorse the argument that electoral manipulation can be associated with sources of incumbent strength in a world where the aims of electoral protest are broadly (rather than narrowly) strategic.
 

### 2025-03-10 Scripts/authoritarian_only_latex_2.Rmd
Major changes to binary models (includes incumbent losses now)

@@
 There are two problems, however.
-First, the empirical record on the relationship between manipulation and protest been mixed, illustrating the limitations of current theory.
-[FILL IN CITES HERE] Second, the expected relationship between manipulation and protest under the regime-weakness and regime-strength models is complicated by the emergence of a grievance-based model of electoral protest generally.
+First, the empirical record on the relationship between manipulation and protest been mixed, illustrating the limitations of current theory [@Hyde2014;@daxecker2012;@Hafner-Burton2016;@Lankina2017;@ananyevDictatorsSignalStrength2022].
+Second, the expected relationship between manipulation and protest under the regime-weakness and regime-strength models is complicated by the emergence of a grievance-based model of electoral protest generally.
 This view, which has considerable empirical support, finds that electoral protests are largely driven by citizen grievances\textemdash not the narrowly strategic calculations about the likelihood of overthrowing the government which are assumed in the regime-weakness model.
@@
 They are excluded from the main analysis, but incorporated in a robustness check.]
-
-[Somewhere cite this article: https://doi.org/10.1177/0032321716674024 on how 'toleration' is really just slow attrition of protest movements.]
 
@@
 
-> H1: Increasing election fraud makes protest more likely, with the size of this effect decreasing with incumbent vote-share.
+> H1: Increasing election fraud makes protest more likely, with the size of this effect decreasing with incumbent-party vote-share.
 
@@
 80]\textemdash that is, of its political strength.
-Additional research focusing on the principal-agent dyanmics inherent in election manipulation reinforces the idea that manipulation carries such costly signals, such that more extensive electoral manipulation indicates greater regime strength.
+Additional research focusing on the principal-agent dynamics inherent in election manipulation reinforces the idea that manipulation carries such costly signals, such that more extensive electoral manipulation indicates greater regime strength.
 This research trend emphasizes the degree to which political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].
@@
 
-protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1 & inc.lose == 0)
+protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1)
 
@@
 
-While it is possible that the theoretical mechanisms considered here may be at work in democracies as well, democracies are excluded from this analysis for two reasons.
-First, the theory presented above is derived from prior work on authoritarian elections; the factors that drive the theory are expected to be weaker in democratic regimes.
+While it is possible that the theoretical mechanisms considered here may be at work in democracies as well, it is likely that these mechanisms are weaken as democratization advances.
 As regimes trend toward liberal democracy, patronage consolidation should decline and rule of law should increase by definition.
-Consequently, the theory expects that fraud should be rare in such cases, even in the absence of a protest deterrent: incumbents simply lack the tools.
-At the same time, the more liberal-democratic a state becomes, the more alternatives their are to protest as a means to check the executive.
-More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest\textemdash the electoral authoritarian regimes\textemdash where we should expect to see a more pronounced role for protest risk in curtailing election fraud under the deterrence logic.
-The case selection strategy here is thus one that both fits the theory and makes for a strong test of the classic fraud-protest model.
+These changes simultaneously make it more difficult for incumbents to manipulate elections and easier for opponents to engage in electoral protest.
+At the same time, the more liberal-democratic a state becomes, the more alternatives there are to protest as a means to check the executive.
+More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest\textemdash the electoral authoritarian regimes\textemdash where we should expect the deterrence logic to operate under a regime-weakness model.
+Finally, as a matter of measurement, protest duration is likely to be an increasingly imperfect measure of incumbent strength in more liberal-democratic settings, where protest is more commonplace and less immediately threatening to the regime.
+The case selection strategy here, excluding electoral and liberal democracies, is thus one that both fits the theory and makes for a stronger test of the regime-weakness model.
+Nevertheless, an expanded sample that includes electoral democracies provides similar results in the appendix.
+In total, there are 659 elections included in the dataset, across 100 countries.
+The date range is limited by the availability of data on protest from ECAV, which ranges from 1989 to 2012.
 
-In all models, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the population of interest: protest following an election in which the incumbent claimed victory, possibly through reliance on election fraud.^[Excluding cases where incumbents stepped down after losing an election is necessary to avoid conflating pro- and anti-incumbent protest.
-However, this design does influence the interpretation of the results from these models.
-Threats to inference from this section are evaluated in more detail in the discussion section.] The full dataset includes 445 election-year observations across 81 non-democratic countries.
-The date range is limited by the availability of data on protest from ECAV, which ranges from 1989 to 2012.^[Some additional tests in the appendix include cases where incumbents accepted defeat, bringing the total number of observations to 659.]
-
-For models of protest-wave duration, only observations where post-election protests were initiated can be included.
-For these models, therefore, I limit that analysis to only those cases with a positive number of post-election protests.
-Of the 445 elections in which the incumbent claimed victory, only 156 fall into this more limited dataset.
-Brief further analyses of the seven cases in which incumbents were ousted, as well as cases where incumbents stepped down after the election, are also included in the discussion and appendix.
+For models of protest-wave duration, as discussed in more detail below, I include only observations where a positive number of protests occurred.
+Only 156 fall into this more limited dataset.
+Brief further analysis of the seven cases in which incumbents were successfully ousted is also included in the discussion and appendix.
 
@@
 
-The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0 & inc.lose == "0") %>% nrow()` of the 445 elections, no protests are recorded.
+The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0) %>% nrow()` of the `r nelda2.sub %>% nrow()` elections, no protests are recorded.
 Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest duration.
 As the figure shows, while election fraud is common among this set of regimes, protests are rare, and, when they do occur, are usually short in duration.
-This distribution\textemdash with large numbers of zeroes\textemdash is in line with the theoretical proposition that most fraudulent elections do not produce signals of weakness that prompt opposition protest.
+These distributions\textemdash extensive election fraud combined with large numbers of non-protest\textemdash provides some early indication that the regime-weakness model has limited support.
 Still, approximately one-third of elections experienced some protest, and so protest initiation is modeled using standard logistic regression.
@@
 
-p.hist.events <- ggplot(nelda2.sub %>% filter(inc.lose == "0"), aes(x=n.events.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()
+p.hist.events <- ggplot(nelda2.sub, aes(x=n.events.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()
 
-p.hist.duration <- ggplot(nelda2.sub %>% filter(inc.lose == "0"), aes(x=d.events.num)) + geom_histogram(bins=100) + xlab("Protest wave duration") + theme_bw()
+p.hist.duration <- ggplot(nelda2.sub, aes(x=d.events.num)) + geom_histogram(bins=100) + xlab("Protest wave duration") + theme_bw()
 
-p.hist.fraud <- ggplot(nelda2.sub %>% filter(inc.lose == "0"), aes(x = v2elirreg.inv)) + geom_histogram(bins=100) + xlab("Election fraud") + theme_bw()
+p.hist.fraud <- ggplot(nelda2.sub, aes(x = v2elirreg.inv)) + geom_histogram(bins=100) + xlab("Election fraud") + theme_bw()
 
- ggplot(nelda2.sub %>% filter(inc.lose == "0"), aes(y = v2elirreg.inv, x= jitter(n.events.post))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("n.events.post (jittered)") + theme_bw()
+ ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= jitter(n.events.post))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("n.events.post (jittered)") + theme_bw()
 
@@
 
+It is necessary to condition the models on whether or not the incumbent party claimed victory in the election, due to the structure of the vote-share data in V-Dem.
+Vote-share data is provided for the winning party without reference to its position as incumbent or challenger.
+From the point of view of those considering anti-incumbent protest, for instance, the information conveyed by a large opposition vote-share is much different from that conveyed by a large winning vote-share.
+Consequently, in all models, I condition for whether the incumbent party lost, based on coding in NELDA.
+In the binary models of Hypothesis 1, I control for incumbent loss as a standalone variable or in interaction with the main explanatory variables.
+
+For the models of protest duration, I exclude all cases where the incumbent lost the election outright.
+This limits the observations to those that fit the population of interest: protest following an election in which the incumbent claimed victory, possibly through reliance on election fraud.^[Excluding cases where incumbents stepped down after losing an election avoids introducing measurement error in the dependent variable, due to a quirk of the ECAV dataset, which codes the 'side' of the protest (pro- or anti-government) as missing during the period between an election and the swearing in of a new government in cases where the incumbent loses.
+Including these cases thus runs the risk of conflating pro- and anti-incumbent protest, along with cases where protesters goals are genuinely unknown.
+However, excluding them also influences the interpretation of the results from these models.
+Threats to inference from this section are evaluated in more detail in the discussion section.]
+
 The models incorporate several potential confounding variables, in addition to the *election fraud* variable described above.
 The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance.
-The winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia, where links to official sources were available, and Psephos.}
+The winning party's *vote-share* is also taken from the V-Dem dataset.^[This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia, where links to official sources were available, and Psephos.]
 
@@
 All models include *legislative constraints* on the executive, a V-Dem variable which captures the extent to which the legislative body in the country acts independently in practice to limit executive power.
-Additional V-Dem variables capture *privatization* of the economy ('v2clstown') and *bureaucratic professionalization* ('v2stcritrecadm'), both of which might give incumbents influence over election fraud and subsequent protest.
-Lastly, a V-Dem measure the *autonomy of election managment bodies* ('v2elembaut') is included.
 
-Several binary control variables are also included.
+In addition, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby [@Bunce2010;@Beissinger_2007].
+Such a dynamic could plausibly increase the likelihood of protest in a nearby country, while also reducing the incumbent's use of fraud.
+I control for this using each country's geopolitical region as coded in Teorell et al [-@qog_teorell], in conjunction with binary variables indicating a successful protest movement from NELDA.
+A 'successful protest' is recorded if a protest movement resulted in either a new election being run or the incumbent stepping down.
+The *regional diffusion* dummy variable takes on a value of one for all countries in the same region as a successful protest, for elections within one calendar year of the successfully overturned election.^[For example, the variable takes on a positive value for elections held in the post-communist region, within one year of the September 24, 2000 election in Serbia that sparked the 'Bulldozer Revolution'.
+All told there are 14 successful electoral protests in the dataset: Bulgaria 1990, Burkina Faso 1997, Lesotho 1998, Cote d'Ivoire 2000 and 2010, Peru 2000, Serbia 2000, Mali 2002, Georgia 2003, Ukraine 2004, Azerbaijan 2005, Kyrgyzstan 2005, Haiti 2006, and Cameroon 2007.]
+
+Several other binary control variables are also included.
 First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest.
@@
 
-```{r include=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
@@
 
-model.margin.binary <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   + v2elirreg.inv +  +    winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.margin.binary)
+model.binary_noint <- glm(any_protest ~ inc.lose + v2elirreg.inv + winner.share.upd, data = nelda2.sub, family = binomial(link = "logit"))
 
-model.margin.binary.int <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.margin.binary.int)
+model.binary.base <-  glm(any_protest ~ winner.share.upd*v2elirreg.inv*inc.lose + presidential, data = nelda2.sub, family = binomial(link = "logit"))
 
-model.margin.binary_noc <- glm(any_protest ~ lji.lag + legcon.1lag +  altinf.1lag  +  + osorg.1lag   + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.margin.binary_noc)
+model.binary.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   +    winner.share.upd*v2elirreg.inv*inc.lose   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre + rentierstate + taxstate + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub)
 
-model.binary_noc <- glm(any_protest ~  v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential , family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-summary(model.binary_noc)
-
-# Plots
-
-####Margin
-
-#plot_binary_margin_max <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-2.3, 3.2]"))
-#plot_binary_margin_max <- plot_binary_margin_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()
-
-plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.32, 2.2]"), show.data = T, colors = "gs")
-plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()
-
+plot_binary_margin_typ_n <- sjPlot::plot_model(model.binary.full, type = "pred", terms = c("v2elirreg.inv", "winner.share.upd [.35, .75]", "inc.lose"), show.data = T)
+plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Electoral manipulation", y = "Predicted probability", title = "Predicted probability of protest", color = "Winner's vote-share") + theme_bw()
 ```
@@
 models2 <- list()
-models2[['Model 1']] <- model.margin.binary
-models2[['Model 2']] <- model.binary_noc
-models2[['Model 3']] <- model.margin.binary.int
+models2[['Model 1']] <- model.binary_noint
+models2[['Model 2']] <- model.binary.base
+models2[['Model 3']] <- model.binary.full
 #models2[['Model 6']] <- model.margin.binary_noc
@@
 modelsummary(models2, output = "default",  stars = T, title = 'Table 3: Logit models of protest occurrence', gof_omit = c("Log|RMSE"), coef_map = c("(Intercept)" = "(Intercept)",
+             "inc.lose1" = "Incumbent loss",
+             "v2elirreg.inv" = "Election fraud",
+             "winner.share.upd" = "Winner vote-share",
+"winner.share.upd:v2elirreg.inv" = "Winner vote-share:Election fraud",
+"winner.share.upd:inc.lose1" = "Winner vote-share:Inc. loss",
+"v2elirreg.inv:inc.lose1" = "Election fraud:Inc. loss",
+"winner.share.upd:v2elirreg.inv:inc.lose1" = "Election fraud:Winner vote-share:Inc. loss",
   "lji.lag" = "Judical independence (lag)",                                                                 "legcon.1lag" = "Leg.
@@
                  "taxstate1" = "Tax state", 
-
+                "presidential1" = "Presidential",
                  "unscheduled_election1" = "Off-schedule election",
@@
                  "regional_success_lag" = "Regional protest diffusion",
-                                  "presidential1" = "Presidential election",
 
                  "median.participants" = "Median protest participants",
-                  "winner.share.upd" = "Incumb. vote-share",
-                  "v2elirreg.inv" = "Election fraud",
-                "v2elirreg.inv:winner.share.upd"= "Election fraud:Incumb. vote-share")
+
+                "v2elirreg.inv:winner.share.upd"= "Election fraud:Winner vote-share")
                  )
@@
 Table \@ref(tab:logit-table) shows the results for Hypothesis 1.
-Model 1 is a base model, where incumbent vote-share and *intentional irregularities* are standalone variables.
-In keeping with the ambiguous findings in the prior literature, there is no significant relationship between fraud severity and the likelihood of protest initiation.
-The incumbent's claimed margin of victory is negatively associated with protest initiation, though this relationship falls short of conventional statistical significance.
-Several control variables behave as expected; for instance, protests are more likely to occur after presidential elections and when there have been prior electoral protests in the region.
-Model 2 includes the interaction between *intentional irregularities* and incumbent vote-share, with no controls other than a dummy variable for *presidential elections*.
+Model 1 is a base model, where incumbent loss, winner vote-share, and *election fraud* are standalone variables.
+In keeping the general (though inconsistent) findings in the prior literature, there is a significant positive relationship between fraud severity and the likelihood of protest initiation.
+The winner's claimed margin of victory is negatively associated with protest initiation.
+Model 2 includes the three-way interaction between *election fraud*, incumbent loss, and winning vote-share, with no controls other than a dummy variable for *presidential elections*.
 This allows for the largest possible sample size, and the results are consistent with those of the full model with controls.

### 2025-10-31 Output/authoritarian_only_word_pop_rev.Rmd
Updates to lit review; structure + new content on 'strength'

--- 2025-03-10 Scripts/authoritarian_only_latex_2.Rmd
@@
 output:
-    bookdown::pdf_document2:
-    keep_tex: true
+  word_document:
 linkcolor: blue
@@
 abstract: "Two models of the relationship between election manipulation and protest can be found in the literature.
-A classic approach holds that election fraud reveals regime weakness, which can lead to strategic opposition protest.
-To avoid costly protest, incumbents may hold cleaner elections.
-A more recent approach contends that election manipulation can be a sign of underlying regime strengths; under this framework protest is less likely to deter manipulation.
-Prior empirical tests do not cleanly distinguish between these models.
-Using event-day data on electoral protest in 445 authoritarian cases from 1990-2012, this paper shows that 1) fraud severity is rarely associated with protest initiation, and 2) governments which manipulate more also defeat protests more quickly.
-This goes against the classic, deterrent model, and helps explain why electoral authoritarianism is a sticky equilibrium amid ongoing democratic backsliding.
+One approach holds that election fraud reveals an incumbent who is more vulnerable to protest; in this view, protest risk can result in cleaner elections.
+An alternative approach contends that election manipulation can be a sign of underlying regime strengths; under this framework, protest risk is less likely to deter manipulation.
+This article interrogates the assumptions underlying these models, and conducts a novel test of their implications.
+Using data on the duration of protest waves in 659 authoritarian cases from 1990-2012, this paper shows that governments which manipulate elections more also defeat protests more quickly.
+This supports the ‘regime-strength’ model, and helps explain why electoral authoritarianism is a sticky equilibrium amid ongoing democratic backsliding.
 It suggests researchers should look beyond protest risk in models of electoral manipulation."
@@
   - \usepackage{pdflscape}
+  - \usepackage{longtable}
 
@@
 To move the debate forward, I model electoral protest *duration* as a dependent variable, in addition to initiation.
-By contrast to protest initiation, the regime-weakness and regime-strength models generate distinct predictions with regard to protest duration.
-Once protests begin, the 'strength' of an incumbent (its likelihood of outlasting a protest movement) is no longer latent.
+Taken together, the two dependent variables offer distinct predictions under the regime-weakness and regime-strength frameworks.
+Once protests run their course, the 'strength' of an incumbent (its likelihood of outlasting a protest movement) is no longer latent.
 Instead, the incumbent's response\textemdash repression, concessions, or attrition\textemdash and its survival (or not) reveal the incumbents' true strength relative to the opposition.
 The more quickly a government puts down protests, the greater the underlying (unobserved) distribution of power favors the incumbent over the opposition.
-By contrast, the longer it takes the government to restore control, the less favorable the distribution of power.^[The analysis treats the small number of cases where the opposition succeeds in ousting the government in several ways.
-They are excluded from the main analysis, but incorporated in a robustness check.]
-
-Using data from 445 non-democratic elections across 97 countries, from 1989 to 2012, this paper finds that 1) protest initiation is most likely when manipulation is severe but incumbent vote-share is low, and 2) that governments that engage in more manipulation are likely to put down protests more quickly than governments that produce little manipulation, especially when incumbents claim small vote-shares.
+By contrast, the longer it takes the government to restore control, the less favorable the distribution of power.
+
+Using data from 659 non-democratic elections across 100 countries, from 1989 to 2012, this paper finds that 1) protest initiation is most likely when manipulation is severe but incumbent vote-share is low, and 2) that governments that engage in more manipulation are likely to put down protests more quickly than governments that produce little manipulation, especially when incumbents claim small vote-shares.
 This pattern of results in inconsistent with the regime-weakness model of election manipulation, but consistent with a world in which election manipulation indicates sources of regime resilience and protests are largely motivated by grievance.
-Given these results, it is likely that\textemdash while incumbents do not seek to provoke protest\textemdash the downside risk for incumbents that engage in election fraud in non-democracies has been overemphasized.
+Given these results, it is likely that---while incumbents do not seek to provoke protest---the downside risk for incumbents that engage in election fraud in non-democracies has been overemphasized.
 This interpretation helps explain several empirical patterns at once: the relative paucity of electoral protests [@Brancati2016], the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory in conjunction with high levels of fraud.
@@
 
+These are normatively optimistic models upon which a sizable literature has been founded: manipulation can be kept in check, and ultimately reduced, through the development of institutions that reveal information about election integrity: election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].
+These institutions improve election integrity, it is claimed, by increasing the likelihood that fraud will be observed and thus, incumbent weakness exposed.
+
 # Literature review: two assumptions about election manipulation and mass protest
 
-As sketched briefly above, a fruitful literature has examined the connections between election integrity and mass protest.
-Research in the regime-weakness model of manipulation typically concludes that as incumbents become increasingly fearful of street protest, they dial back on manipulation\textemdash a potentially pro-democracy outcome.
+Research in the regime-weakness model of manipulation typically concludes that as incumbents become increasingly fearful of street protest, they dial back on manipulation---a potentially pro-democracy outcome.
 These models hinge on two assumptions, as discussed below.
 First, election manipulation is conceptualized as a short-term tool used instrumentally to secure victory in the election [@lehoucq2002]; it communicates only that the incumbent could not count on winning the election cleanly.
-In such models incumbents are typically characterized as 'weak' or 'strong' based on their true support in the electorate: strong incumbents are those that are likely to defeat protest movements, while weak incumbents are those that are likely to be ousted.
-
-This true level of support can be obscured prior to elections due to pervasive preference falsification [@Kuran1991] and the low level of political information common in authoritarian regimes [@guriev2020].
-However, low popularity (i.e.
-weakness, in these models) can be revealed when incumbents need to resort to fraud in order to secure re-election.
+Second, protest is typically conceptualized as narrowly instrumental, aimed at overthrowing a weak incumbent rather than any broader or more expressive goals.
+These assumptions face several challenges, which are discussed below.
+First, however, it is necessary to clarify the meaning of incumbent strength.
+
+As sketched briefly above, a fruitful literature has examined the connections between election integrity and mass protest.
+This literature has grappled with the extent to which election manipulation reflects, and reveals, the incumbent's strength.
+Various authors have put forth different modalities of incumbent strength, with some focusing on popularity and others on coercive power.
+As Leventoğlu et al [-@leventogluSynthesizingTheoriesAuthoritarian2025] discuss in their excellent analysis, genuine popularity is but one potential source of incumbent strength.
+Popularity certainly helps an incumbent win an election and survive any challenges in its aftermath, but it is only one tool; incumbents may also draw on resource wealth, organizational capacity, support among elite groups, the military and security services, or help from foreign patrons.
+These latter factors may be weakly correlated, if at all, with the incumbent's genuine popularity.
+This paper takes a general approach, following Leventoğlu et al [-@leventogluSynthesizingTheoriesAuthoritarian2025]: a stronger incumbent is one who is better able to survive in power throughout the whole electoral cycle.
+This means the incumbent has both prevailed in the election itself, and managed to hold off any subsequent popular revolts or elite challenges [@svolik_politics_2012].
+
+## The first assumption: manipulation is uninformative about strength
+
+The characteristic assumption in regime-weakness models is that electoral manipulation communicates nothing about incumbent strength, which is closely tied to popularity.
+At worst, election manipulation reveals low popularity (i.e.
+weakness).
 As a result, the opposition perceives election fraud as a reliable signal of incumbent weakness, and will stage protests in order to push out the weak incumbent.
@@
 1153).
-In equilibrium, this can lead weaker incumbents to step down rather than face protest due to massive fraud\textemdash a pro-democracy outcome in which stronger opposition groups deter manipulation by incumbents.
+In equilibrium, this can lead weaker incumbents to step down rather than face protest due to massive fraud---a pro-democracy outcome in which stronger opposition groups deter manipulation by incumbents.
 Similar results are developed by Chernykh and Svolik [-@chernykh2015third], Luo and Rozenas [-@Luo2020] and [-@Luo2017], and Magaloni [-@magaloni2010].
 
-Some more recent models treat election manipulation as a distortion of the electoral signal of the incumbent's true strength, but are still located within the regime-weakness framework.
+Some more recent models treat election manipulation as a distortion of the electoral signal of the incumbent's true strength, rather than a direct indicator of weakness.
+These models are still located within the regime-weakness framework, however.
 In this type of model, such as those developed by Little [-@littleFraudMonitoringNoncompetitive2015] or Egorov and Sonin [-@Egorov2021], fraud may be used by incumbents to bluff their opponents into being more supportive than they would be if they knew the incumbent's true popularity.
@@
 The typical result in these models is that protest risk can *partially* deter fraud.
-In equilibrium, the incumbent will generally choose the level of fraud anticipated by the opposition and no more\textemdash but also no less.
+In equilibrium, the incumbent will generally choose the level of fraud anticipated by the opposition and no more---but also no less.
 In sum, regime-weakness models of fraud typically hold that the level of election manipulation carries no signal of the incumbent's ability to defeat protesters; on the contrary, it only carries signals of weakness.
+
+## A second assumption: the nature of protest
 
@@
 
-These are normatively optimistic models upon which a sizable literature has been founded: manipulation can be kept in check, and ultimately reduced, through the development of institutions that reveal information about election integrity: election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].
-These institutions improve election integrity, it is claimed, by increasing the likelihood that fraud will be observed and thus, incumbent weakness exposed.
-
-However, there is also a body of research that challenges the two assumptions upon which the regime-weakness model is founded.
-The assumption on the nature of protest as centralized and narrowly strategic, in particular, has been challenged in the literature on electoral protest.
-Recent research in this tradition has argued that electoral protest occurs largely unrelated to strategic concerns about the relative strength of the incumbent and the opposition.
-In these models, electoral protest occurs because manipulation galvanizes citizens who hold diverse grievances [@tucker2007], because economic conditions are poor [@Brancati2016;@Beaulieu2014a], or simply because the bare fact of holding an election creates a moment of opportunity for opposition organization irrespective of the level of manipulation [@Pop-Eleches2015].
-The central feature of these models is that protest decisions are generally made without consideration for the likelihood that protests will overthrow the incumbent.
-
-This school of thought holds that decisions to engage in electoral protest are driven in large part by citizens' grievances rather than by a strategic opposition [@tucker2007].
-In this view, election manipulation can act as a shared grievance among individuals who prefer to see the ruling party lose\textemdash especially when it appears that fraud may have been decisive [@kuntz_more_2009].
-Belief in a stolen election\textemdash one the incumbent would have lost in the absence of manipulation\textemdash is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007], possibly initiating a protest cascade [@Kuran1991].
-These models thus tend to predict that protests are most likely when margins of victory are low and election manipulation high, suggesting both a high level of grievance in society and a "stolen" election.
-This framework predicts protest is most likely when incumbent vote-share is low and manipulation high (i.e.
-elevated grievances and a high likelihood of a stolen election).
-
-Rational protest entrepreneurs might seek to initiate protest to take advantage of this popular discontent for broadly strategic reasons.
+Most fundamentally, regime-weakness models assume that opposition protest is *elite-led and instrumental*.
+The opposition is typically modeled as a unitary, strategic actor who stages protest when the expected benefits of successfully overthrowing the incumbent outweighs the expected cost of failure.
+Such abstraction helps keep formal models tractable, but this assumption is difficult to square with recent research on electoral protest and a broader understanding of political protest in authoritarian regimes.
+
+Electoral protest may occur because manipulation galvanizes citizens who hold diverse grievances [@tucker2007], because economic conditions are poor [@Brancati2016;@Beaulieu2014a], because electoral manipulation makes people angry [@Szakonyi2021], or simply because their party lost the election [@andersonLearningLoseElection2006a].
+Opposition elites, in turn, might seek to initiate protest to take advantage of this popular discontent for strategic, rather than instrumental, reasons.
 They may wish to pressure the incumbent in order to push for improved competitiveness in the *next* election or to bargain for policy concessions [@chernykh2014;@robertson_protest_2010].
@@
 Protests may be staged as part of internal factional rivalries within the opposition [@Buckles2017], or to drive media coverage of issues important to the opposition [@huangMediaProtestDiffusion2019;@jenningsStreetDemonstrationsMedia2019].
-Or, a vanguard of protesters may take to the streets simply because their distaste for the government outweighs any risk of sanction [@lohmannDynamicsInformationalCascades1994].
-These and other motivations for initiating protest are theoretically significant because their rationality (for protesters) does not hinge on the likelihood of overturning the government, as regime-weakness models of manipulation assume.
-As a result, it is possible that we may observe the initiation of anti-government electoral protest even when the likelihood of such protest overturning the election (or the government itself) is very low.
+Indeed, opposition elites may take advantage of the bare fact of an election to challenge the incumbent, irrespective of the level of manipulation or the incumbent's vote-share [@Pop-Eleches2015].
+
+In line with this larger body of research, I argue that electoral protest in authoritarian regimes is best understood as *dispersed* and consequently *multi-purpose*.
+In other words, protests may be initiated and sustained by a mix of actors with a variety of expressive and strategic objectives, some only distantly aimed at overthrowing the government.
+The theoretical distinction is important.
+If protest is understood to be primarily instrumental, opposition actors will only launch protests they think have a good chance of unseating the incumbent.
+If protest is dispersed and multipurpose, protest movements may initiate for reasons orthogonal to perceived incumbent strength and severity of electoral manipulation.
+In such a setting, incumbents cannot count on quiet streets simply by manipulating at a low level, or even holding clean elections.
+
+## Election manipulation and regime weakness
+
+<!-- The central feature of these models is that protest decisions are generally decentralized, and made with less consideration for the likelihood that protests will overthrow the incumbent.
+Belief in a stolen election---one the incumbent would have lost in the absence of manipulation---is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007;@kuntz_more_2009].
+These models thus tend to predict that protests are most likely when margins of victory are low and election manipulation high, suggesting both a high level of grievance in society and a "stolen" election.
+-->
+
+## Election manipulation and incumbent strength
 
@@
 This view, most persuasively articulated by Simpser [-@simpser2013], holds that increasing levels of election manipulation can indicate a well-resourced, well-organized, powerful incumbent (independent of its popularity); these resources and capacities, in turn, may help the incumbent weather a protest movement.
-Under this view, extensive election manipulation can be a show of strength, which induces cooperation with the incumbent from other societal actors and may serve to deter protest [@simpser2013].
+Under this view, extensive election manipulation can be a show of strength, which induces cooperation with the incumbent from other societal actors [@simpser2013].
 Since this view holds that more manipulation can further entrench the incumbent, it is a normatively pessimistic one.
@@
 This can, in turn, fix expectations that the incumbent will continue to rule in the future, further enhancing its staying power [@hale_patronal_2014].
-This argument is in keeping with Seeberg's [-@Seeberg2019] finding that higher-capacity states are associated with more election fraud.
 
@@
 \begin{table}[]
-\begin{tabular}{@{}llll@{}}
+\begin{tabular}{@{}rll@{}}
 \toprule
-\multicolumn{1}{c}{\textbf{Protest type}} &
-  \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Likelihood\\ of \\ Protest\end{tabular}}} &
-  \multicolumn{1}{c}{\textbf{Mechanism}} &
-  \textbf{Example sources} \\ \midrule
-\multicolumn{1}{l|}{} &
-  \multicolumn{3}{c}{Regime-weakness model} \\ \cmidrule(l){2-4} 
-\multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Narrowly \\ strategic\end{tabular}} &
-  Increases &
-  \begin{tabular}[c]{@{}l@{}}More manipulation indicates \\ lower true popularity, and thus \\ lower likelihood that \\ incumbent will survive protests.\end{tabular} &
-  \begin{tabular}[c]{@{}l@{}}Luo and Rozenas (2020); \\ Chernykh and Svolik (2015); \\ Little, Tucker, and \\ LaGatta (2015)\end{tabular} \\
-\multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Broadly \\ strategic\end{tabular}} &
-  Increases &
-  \begin{tabular}[c]{@{}l@{}}More manipulation employed to \\ compensate for deeper \\ grievances; 'stolen' election as \\ trigger.\end{tabular} &
-  \begin{tabular}[c]{@{}l@{}}Beaulieu (2014); \\ Brancati (2016); \\ Tucker (2007)\end{tabular} \\ \midrule
-\multicolumn{1}{l|}{} &
-  \multicolumn{3}{c}{Regime-strength model} \\ \cmidrule(l){2-4} 
-\multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Narrowly \\ strategic\end{tabular}} &
-  Decreases &
-  \begin{tabular}[c]{@{}l@{}}More manipulation indicates \\ capacities that increase \\ likelihood incumbent survives\\ protest, offsetting low popularity.\end{tabular} &
-  Simpser (2013) \\
-\multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Broadly \\ strategic\end{tabular}} &
-  Increases &
-  \begin{tabular}[c]{@{}l@{}}More manipulation employed to\\ compensate for deeper \\ grievances; 'stolen' election as\\ trigger.\end{tabular} &
-  \begin{tabular}[c]{@{}l@{}}Beaulieu (2014); \\ Brancati (2016);\\ Tucker (2007)\end{tabular}
+Incumbent's genuine popularity:                & \multicolumn{2}{c}{Low}                                               \\ \midrule
+\multicolumn{1}{r|}{Protest is assumed to be:} & \multicolumn{1}{c|}{Instrumental} & \multicolumn{1}{c|}{Multipurpose} \\ \midrule
+\multicolumn{1}{|l|}{Manipulation...}          &                                   & \multicolumn{1}{l|}{}             \\ \midrule
+\multicolumn{1}{|r|}{...obscures strength / indicates weakness} &
+  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High and increasing\\ \\ (Luo and Rozenas 2020; \\ Chernykh and Svolik 2015; \\ Little, Tucker, and LaGatta 2015)\end{tabular}} &
+  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High and increasing\\ \\ (Tucker 2007; \\ Beaulieu 2014; \\ Brancati 2016)\end{tabular}} \\ \midrule
+\multicolumn{1}{r|}{...reflects strength} &
+  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High and decreasing\\ \\ (Simpser 2013)\end{tabular}} &
+  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High and increasing\\ \\ (Tucker 2007;\\ Beaulieu 2014;\\ Brancati 2016)\end{tabular}} \\ \bottomrule
 \end{tabular}
-\caption{Table 1: Predictions for protest initiation under the regime-weakness model of electoral manipulation}
-\label{tab:protest-expectation}
+\caption{Likelihood of protest initiation and expected relationship with increasing election manipulation}
+\label{tab:initiation}
 \end{table}
@@
 
-This ambiguity may help account for the inconsistent relationship between election manipulation and protest occurrence found in the literature.
-Several cross-national studies have found links between election manipulation and binary measures of protest [@Brancati2016;@Harvey2020].
-For instance, Hyde and Marinov [-@Hyde2014] use the NELDA dataset to show that protest is more likely when international observers "seriously questioned the winner of the election or the legitimacy of the process," across 758 elections (p.
-343).
-While that study did not include data on incumbents' vote-shares, such a measure of manipulation comes closest to the 'stolen' election framework.
+This ambiguity may help account for the inconsistent relationship between election manipulation and protest found in the literature.
+Several cross-national studies have found links between election manipulation and binary measures of protest [@Brancati2016;@Harvey2020;@Hyde2014].
 A study of African elections from 1997 to 2009 shows that election fraud is associated with an increased number of post-election conflicts [@daxecker2012].
@@
 
-Nevertheless, there are considerable theoretical and substantive stakes in knowing which of these four possible data-generating processes is most likely; in only one does increasing protest risk actually improve electoral integrity.
+Because this paper attempts to systematize and test two competing models of electoral manipulation and protest, it is worth spelling out each expectation and its mechanisms in detail.
+To do so, I will describe each theory's expectations under two worlds: one where the incumbent's genuine popularity is low, and one where it is high.
+For each of these notional worlds, expectations for protest initiation and duration are explored under both sets of assumptions about the nature of protest and electoral manipulation.
+
+First, consider the a world in which the incumbent's true popularity is low; this is presumed to be the case for most authoritarian regimes by their nature.
+Next, let us assume that the regime-weakness model is correct: election manipulation indicates weakness (or obscures information about strength).
+Finally, assume that protest is elite-led and instrumental.
+This is the world of the classic regime-weakness framework: we should expect to see protest initiation become more likely as the severity of electoral manipulation increases.
+
+The same prediction holds if we assume that protest is largely multi-purpose, while maintaining the assumptions that true popularity is low and manipulation indicates weakness.
+This is the classic focal point example [@tucker2007].
+Here, the incumbent's low genuine popularity implies high levels of grievance among the population, which are increasingly likely to be focused into mass protest as manipulation increases (i.e.
+as the election becomes more likely to appear 'stolen').
+
+Now, while maintaining the assumption that the incumbent's popularity is low, let us assume that increasing electoral manipulation indicates an otherwise stronger incumbent.
+In this case, if protest is primarily instrumental, we should expect opposition groups to shy away from protest initiation as electoral manipulation increases.
+An unpopular incumbent who manipulates extensively is stronger, in this framework, than an unpopular incumbent who does not.
+As a result, we should expect the opposite prediction of the classic regime-weakness model.
+
+However, if the incumbent's popularity is low, manipulation indicates underlying strength, and protest is largely dispersed and multi-purpose, the expectation is reversed.
+Just as in the regime-weakness case, low popularity indicates high grievances.
+Increasing electoral manipulation helps focus those grievances and drives protest initiation.
+
+The key takeaway, then, is that for most authoritarian regimes (where genuine incumbent popularity is likely low), an examination of protest initiation alone cannot distinguish between three out of four possible data-generating processes.
+A pattern in which higher levels of manipulation are associated with increasing risk of protest initiation can be explained either by the regime-weakness model, or by a world in which manipulation indicates incumbent staying-power but protest is non-instrumental.
+
+There are considerable theoretical and substantive stakes in knowing which of these four possible data-generating processes is most likely; in only one does increasing protest risk actually improve electoral integrity.
 Since the empirical record is mixed, and three of the four possible models make similar predictions with regard to protest initiation, it is necessary to generate new hypotheses.
 In particular, the remainder of this manuscript will attempt to distinguish between regime-strength and regime-weakness models of manipulation by developing hypotheses about protest *duration* under each framework.
-Specifically, I will argue that protest duration helps reveal the true relative balance of power between incumbents and opposition actors.
 
@@
 
-# Theory
+# Theory: electoral manipulation, incumbent strength, and protest duration
+
+Theories in the 'regime-weakness' view of electoral manipulation are built on several assumptions about the nature of protest and the information conveyed by the incumbent's vote-share and manipulation efforts.
+Updating some of these assumptions, I argue, provides a model that more accurately reflects contemporary research on manipulation *and* makes distinct empirical predictions.
+Each of these assumptions, and how they are adjusted in the 'regime-strength' model of manipulation, is discussed below.
+
+Talk about protest duration here:
 
@@
 The most visible, and arguably most studied, form of such legitimacy costs is mass post-election protest against the incumbent.
-Here, I do not challenge the broad consensus in the literature that authoritarian incumbents find sustained, sizable protests undesirable, since they entail some risk of incumbent defeat and may require costly repression or concessions.
+There is a broad consensus in the literature that authoritarian incumbents find sustained, sizable protests undesirable, since they entail some risk of incumbent defeat and may require costly repression or concessions.
 I assume, as a result, that incumbents generally seek to demobilize protest; protest movements that survive longer indicate the incumbent is relatively weaker compared to the opposition.^[The implications of protest toleration by incumbents will be discussed later.] By contrast, when protest movements meet a quick demise, the incumbent can be considered relatively strong.^[In both cases, this assume that incumbent survives the protest movement, by far the modal outcome.
@@
 
+There is an alternative view of electoral manipulation, however; Simpser's foundational argument holds that manipulated elections can send costly signals of the incumbent's "capacities, resources, [and] dispositions" that will help it retain power in the future [@simpser2013, p.
+80]---that is, of its political strength.
+I argue that more recent research on the principal-agent dynamics inherent in election manipulation reinforces and expands this idea.
+This research trend emphasizes the degree to which political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].
+Incumbents that can solve these principal-agent problems at election-time, are also more likely to be able to survive post-election challenges.
+
+Electoral manipulation has been shown to be more extensive under conditions that minimize principal-agent problems between low-level agents and the incumbent.
+These include: greater local popularity for the incumbent [@rundlett2016deliver], greater incumbent control over patronage resources [@Harvey2019;@hale_patronal_2014], and lower risk of punishment for pro-regime lawbreaking [@Harvey2022].
+Lastly, extensive manipulation also indicates that large number of agents assess that the incumbent will remain in power in the future [@simpser2013], which itself becomes a self-reinforcing equilibrium [@hale_patronal_2014].
+
+These resources and capacities improve the incumbent's ability to respond to a nascent protest movement through repression or concessions [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014].
+Concessions may be aimed at protesting parties themselves [@reuter2015;@mooreRepressionDissentSubstitution2000;@Hafner-Burton2016;@Piven1979] or the more moderate public, in order to keep them on the sidelines of the conflict [@Hummel2019].
+In either case, large-scale election manipulation indicates control over considerable patronage resources [@Harvey2019;@hale_patronal_2014;@greene_why_2007], which in turn indicates the capacity to engage in such payments (or withhold them from clients in exchange for passive support [@Rosenfeld2017]).
+
+Ruling parties may also lean on repression, calling upon the security services to disperse protests, arrest leaders, and perhaps employ harsher methods including disappearance, torture, and execution [@Escriba_Folch2013;@Davenport2007].
+Though such crackdowns risk provoking a backlash, they are aimed at demobilizing the opposition by increasing the costs of participation in overt dissent [@DeVogel2022].
+The ability to protect regime agents from any legal consequences for carrying out orders, a condition which helps enable illegal election manipulation [@Harvey2022], is also an asset when asking law enforcement agents to carry out possibly violent repression [@MAGALONI_RODRIGUEZ_2020;@stavro_police_2023;@abouharbFactoJudicialIndependence2013].
+Additionally, more consolidated control over patronage resources can help incentivize elite unity [@Reuter2019;@hale_patronal_2014;@Arriola_2009], which can enable repression by the security forces [@hendrix_2017;@Frugé_2019;@Albrecht_Ohl_2016].
+
+In addition, as Gehlbach and Simpser [-@gehlbach2015electoral] argue, effective electoral manipulation can induce regime agents to exert effort on behalf of the incumbent.
+Elite defections, which can presage real challenges for the incumbent during a crisis, are also rarer in less democratic regions, where election manipulation is likely more severe [@Reuter2019].
+Just as election manipulation has been held to solve a coordination problem for regime opponents, it may also solve a coordination problem for regime agents: when successful, a campaign to manufacture votes signals to agents that other agents are likely to stick with the incumbent due to its revealed capacities [@hale_patronal_2014].
+
+This argument---that successful, large-scale election manipulation draws on the same resources a state may use to suppress protest---turns the deterrent logic on its head.
+Instead of signalling weakness, large-scale fraud indicates a well of resilience: control over resources, impunity for pro-regime lawbreaking, and loyal agents.
+As a result, governments that build and deploy an electoral manipulation apparatus are simultaneously inoculating themselves against the expected cost of protest.
+This yields Hypothesis 2.
+
+This true level of support can be obscured prior to elections due to pervasive preference falsification [@Kuran1991] and the low level of political information common in authoritarian regimes [@guriev2020].
+<- Keep this for risk of protests spiraling out of control
+
 Regime-weakness models of electoral manipulation, discussed above, treat election manipulation and genuine votes as substitute goods.
 For any given vote-share, an increase in manufactured votes reduces the number of genuine votes.
-In these models, which consider manufactured votes to be merely illusions of strength, the higher the number of manufactured votes for any given vote total, the weaker the incumbent.
+In these models, which consider manufactured votes to be mere illusions of strength, the higher the number of manufactured votes for any given vote total, the weaker the incumbent.
 This means that strategic opposition actors are more likely to initiate protest when they perceive the incumbent to be weaker, and expect to be more likely to succeed in a challenge.
-A high vote-share with low fraud indicates a more popular incumbent, likely deterring protest.
-A low vote-share with a high degree of fraud indicates a very weak incumbent, making strategic protest more likely.
 In other words, manipulated votes are empty in the classic model; subtracting them from the incumbent's vote-total gives opposition actors their estimate of the incumbent's ability to win a post-election standoff.
@@
 
-Simpser's foundational argument in the regime-strength model holds that manipulated elections can send costly signals of the incumbent's "capacities, resources, [and] dispositions" that will help it retain power in the future [@simpser2013, p.
-80]\textemdash that is, of its political strength.
-Additional research focusing on the principal-agent dynamics inherent in election manipulation reinforces the idea that manipulation carries such costly signals, such that more extensive electoral manipulation indicates greater regime strength.
-This research trend emphasizes the degree to which political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].
-
-Electoral manipulation has been shown to be more extensive under conditions that minimize that minimize principal-agent problems (i.e.
-those that make low-level agents more likely to cooperate with the incumbent).
-These include: greater local popularity for the incumbent [@rundlett2016deliver], greater incumbent control over patronage resources [@Harvey2019;@hale_patronal_2014], and lower risk of punishment for pro-regime lawbreaking [@Harvey2022].
-As discussed below, these capacities and resources are also independently useful for surviving post-election protest.
-Consequently, if an incumbent is able to manufacture a large numbers of votes, it is also likely to have the resource capacity, control over bureaucratic institutions, and perhaps deep wells of popularity in parts of the country that can serve as independent sources of regime durability, even if the incumbent's nationwide margin of victory is low.
-Lastly, extensive manipulation also indicates that large number of agents assess that the incumbent will remain in power in the future [@simpser2013], which itself becomes a self-reinforcing equilibrium [@hale_patronal_2014].
-
-These resources and capacities improve the incumbent's ability to respond to a nascent protest movement through repression or concessions.^[Toleration, another possibility, is discussed later.] [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014].
-Concessions may be aimed at protesting parties themselves [@reuter2015;@mooreRepressionDissentSubstitution2000;@Hafner-Burton2016;@Piven1979] or the more moderate public, in order to keep them on the sidelines of the conflict [@Hummel2019].
-In either case, large-scale election manipulation indicates control over considerable patronage resources [@Harvey2019;@hale_patronal_2014;@greene_why_2007], which in turn indicates the capacity to engage in such payments (or withhold them from clients in exchange for passive support [@Rosenfeld2017]).
-
-Ruling parties may also lean on repression, calling upon the security services to disperse protests, arrest leaders, and perhaps employ harsher methods including disappearance, torture, and execution [@Escriba_Folch2013;@Davenport2007].
-Though such crackdowns risk provoking a backlash, they are aimed at demobilizing the opposition by increasing the costs of participation in overt dissent [@DeVogel2022].
-The ability to protect regime agents from any legal consequences for carrying out orders, a condition which helps enable illegal election manipulation [@Harvey2022], is also an asset when asking law enforcement agents to carry out possibly violent repression [@MAGALONI_RODRIGUEZ_2020;@stavro_police_2023;@abouharbFactoJudicialIndependence2013].
-Additionally, more consolidated control over patronage resources can help incentivize elite unity [@Reuter2019;@hale_patronal_2014;@Arriola_2009], which can enable repression by the security forces [@hendrix_2017;@Frugé_2019;@Albrecht_Ohl_2016].
-
-In addition, as Gehlbach and Simpser [-@gehlbach2015electoral] argue, effective electoral manipulation can induce regime agents to exert effort on behalf of the incumbent.
-Elite defections, which can presage real challenges for the incumbent during a crisis, are also rarer in less democratic regions, where election manipulation is likely more severe [@Reuter2019].
-Just as election manipulation has been held to solve a coordination problem for regime opponents, it may also solve a coordination problem for regime agents: when successful, a campaign to manufacture votes signals to agents that other agents are likely to stick with the incumbent due to its revealed capacities [@hale_patronal_2014].
-
-This argument\textemdash that successful, large-scale election manipulation draws on the same resources a state may use to suppress protest\textemdash turns the deterrent logic on its head.
-Instead of signalling weakness, large-scale fraud indicates a well of resilience: control over resources, impunity for pro-regime lawbreaking, and loyal agents.
-As a result, governments that build and deploy an electoral manipulation apparatus are simultaneously inoculating themselves against the expected cost of protest.
-
-Putting it all together: this theory assumes that authoritarian incumbents generally prefer order to unrest, and that protesters\textemdash who under either a narrow or broadly strategic view are aiming for some kind of concession from the incumbent\textemdash want to demonstrate the popularity of their cause through large, sustained protests.
-Electoral protests are a form of competition between the incumbent and opposition actors, in which both sides are trying to demonstrate their strength in order to win concessions from\textemdash or defeat outright\textemdash the other [@Beaulieu2014a].
+<!--
+Putting it all together: this theory assumes that authoritarian incumbents generally prefer order to unrest, and that protesters---who under either a narrow or broadly strategic view are aiming for some kind of concession from the incumbent---want to demonstrate the popularity of their cause through large, sustained protests.
+Electoral protests are a form of competition between the incumbent and opposition actors, in which both sides are trying to demonstrate their strength in order to win concessions from---or defeat outright---the other [@Beaulieu2014a].
 The outcome of this contest reveals the underlying strengths of the participants.
@@
 This yields Hypothesis 2.
+-->
 
@@
 To do so, they must solve principal-agent problems by consolidating control over patronage resources and shielding agents from negative consequences of their actions.
-The same resources that can be used to manipulate elections on a wide scale\textemdash including natural popularity, resource control, and control over the legal apparatus\textemdash also better enable incumbents to dismantle protest movements when they do emerge.
+The same resources that can be used to manipulate elections on a wide scale---including natural popularity, resource control, and control over the legal apparatus---also better enable incumbents to dismantle protest movements when they do emerge.
 As a result, more extensive manipulation should be associated with shorter-lived protest movements against incumbents.^[In the appendix, I check the separate prediction that manipulation should be positively associated with incumbents being ousted by protest movements.]
@@
 As such, tolerating or ignoring protest is unlikely in this setting.
-In such cases, we should expect to see apparent 'toleration' displaced by 'attrition'\textemdash efforts by the incumbent to increase the cost of protest while avoiding open repression [@yuenNeitherRepressionConcession2017].
+In such cases, we should expect to see apparent 'toleration' displaced by 'attrition'---efforts by the incumbent to increase the cost of protest while avoiding open repression [@yuenNeitherRepressionConcession2017].
 In other words, even in the absence of explicit repression or concessions, we should still expect incumbents to attempt to defuse protest movements aimed at challenging the most recent election.
-
-Testing H1 and H2, together, will help to evaluate the regime-weakness model of electoral manipulation.
+As a further control, however, I include a measure of average protest size in one model of protest duration.
+
+Testing H1 and H2 together will help to evaluate the regime-weakness model of electoral manipulation.
 If H1 is supported and H2 is not, it would provide evidence in favor of the regime-weakness model; manipulation would appear associated with incumbent weakness, and protest initiation would appear to take advantage of such.
@@
 library(tidyverse)
+library(colorspace)
 library(countrycode)
@@
 library(kableExtra)
+library(flextable)
+library(gt)
 
@@
 
+## Inc lose
+nelda2.sub <- nelda2.sub %>% mutate(inc.lose_clean = as.factor(ifelse(inc.lose == "<NA>", NA, inc.lose)))
 ```
@@
 The unit of observation for this study is an election period as recorded in NELDA.
-This framework prevents selecting on the dependent variable, by capturing both protested and non-protested elections.
+This prevents selecting on the dependent variable, by capturing both protested and non-protested elections.
 I limit the sample to authoritarian regimes which hold multiparty national elections.
-This excludes fully closed autocracies (where no national elections are held de jure), as well as electoral and liberal democracies.
-The specific variable used is 'v2x_regime'.
-I limit the sample to those regime-years coded 1 (electoral authoritarian) for the election year, or coded 2 (electoral democracy) for the election year but 1 for the lagged year.^[This coding rule will capture cases that are coded as electoral democratic after an election but electoral authoritarian before, in order to ensure that cases which experience a liberalization after the election are included.] This sample includes a range of electoral authoritarian cases, including those that might be labeled as 'hegemonic' and 'competitive' [@donno2013elections].^[According to the V-Dem codebook, the coding scheme includes cases where multiparty elections were manipulated to the extent that "irregularities...affected the outcome of the election," and ones in which it is "hard to determine whether the irregularities affected the outcome or not." Only when irregularities definitively did not affect the outcome would election-years be categorized as electoral democracy, and excluded from the main analysis.] For instance, the dataset includes election years from Egypt and Belarus, but also from Ukraine, and Mexico in the 1990s.^[Of course, treating regime type as a sharply defined binary indicator, rather than as a continuum, will introduce some bias around edge cases.
-As a robustness check for this case selection process, I also run the main analyses on a larger sample that adds electoral democratic regimes.
-While the strength of the findings is somewhat attenuated, as is expected, the results are generally upheld in the broader sample as shown in the appendix.]
-
-While it is possible that the theoretical mechanisms considered here may be at work in democracies as well, it is likely that these mechanisms are weaken as democratization advances.
-As regimes trend toward liberal democracy, patronage consolidation should decline and rule of law should increase by definition.
-These changes simultaneously make it more difficult for incumbents to manipulate elections and easier for opponents to engage in electoral protest.
+This excludes fully closed autocracies (where no national elections are held), as well as electoral and liberal democracies.
+I limit the sample to those regime-years coded 1 (electoral authoritarian) on the V-Dem variable 'v2x_regime' for the election year, or coded 2 (electoral democracy) for the election year but 1 for the lagged year.^[This coding rule will capture cases that are coded as electoral democratic after an election but electoral authoritarian before, in order to ensure that cases which experience a liberalization after the election are included.] This sample includes a range of electoral authoritarian cases, including those that might be labeled as 'hegemonic' and 'competitive' [@donno2013elections].^[According to the V-Dem codebook, the coding scheme includes cases where multiparty elections were manipulated to the extent that "irregularities...affected the outcome of the election," and ones in which it is "hard to determine whether the irregularities affected the outcome or not." Only when irregularities definitively did not affect the outcome would election-years be categorized as electoral democracy, and excluded from the main analysis.] For instance, the dataset includes election years from Egypt and Belarus, but also from Ukraine, and Mexico in the 1990s.
+
+It is likely that the theoretical mechanisms described above weaken as regimes become more liberal-democratic; in liberal democracies, patronage consolidation should decline and rule of law should increase by definition.
+These changes should make it more difficult for incumbents to manipulate elections, irrespective of protest risk.
 At the same time, the more liberal-democratic a state becomes, the more alternatives there are to protest as a means to check the executive.
-More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest\textemdash the electoral authoritarian regimes\textemdash where we should expect the deterrence logic to operate under a regime-weakness model.
+More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest---the electoral authoritarian regimes---where we should expect the deterrence logic to operate under a regime-weakness model.
 Finally, as a matter of measurement, protest duration is likely to be an increasingly imperfect measure of incumbent strength in more liberal-democratic settings, where protest is more commonplace and less immediately threatening to the regime.
@@
 For models of protest-wave duration, as discussed in more detail below, I include only observations where a positive number of protests occurred.
-Only 156 fall into this more limited dataset.
-Brief further analysis of the seven cases in which incumbents were successfully ousted is also included in the discussion and appendix.
+Only 156 observations fall into this more limited dataset.
 
@@
 As the figure shows, while election fraud is common among this set of regimes, protests are rare, and, when they do occur, are usually short in duration.
-These distributions\textemdash extensive election fraud combined with large numbers of non-protest\textemdash provides some early indication that the regime-weakness model has limited support.
+These distributions---extensive election fraud combined with large numbers of non-protest---provides some early indication that the regime-weakness model has limited support.
 Still, approximately one-third of elections experienced some protest, and so protest initiation is modeled using standard logistic regression.
@@
 
- ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= jitter(n.events.post))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("n.events.post (jittered)") + theme_bw()
+nelda2.sub <- nelda2.sub %>% mutate(fraud_transform = v2elirreg.inv - min(v2elirreg.inv, na.rm=T) + 1)
+nelda2.sub <- nelda2.sub %>% mutate(vs_fraud_ratio = winner.share.upd / fraud_transform)
+
+ ggplot(nelda2.sub, aes(x = vs_fraud_ratio, y= jitter(any.ecav.protest))) + geom_point(aes(color = v2elirreg.inv)) + xlab("Ratio") + theme_bw() + scale_color_viridis_b()
+
+  ggplot(nelda2.sub, aes(x = v2elirreg.inv, y= jitter(any.ecav.protest))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Fraud") + theme_bw()
+
+   ggplot(nelda2.sub, aes(x = v2elirreg.inv, y= n.events.post)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Fraud") + theme_bw()
+
+  ggplot(nelda2.sub, aes(x = winner.share.upd, y= jitter(any.ecav.protest))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Winner share") + theme_bw()
+
+    ggplot(nelda2.sub, aes(x = winner.share.upd, y= n.events.post)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Winner share") + theme_bw()
+
+   ggplot(nelda2.sub, aes(x = vs_fraud_ratio, y= d.events.num)) + geom_point(aes(color = v2elirreg.inv)) + xlab("Election fraud") + theme_bw() + scale_color_viridis_b()
+
+  ggplot(nelda2.sub, aes(x = v2elirreg.inv, y= d.events.num)) + geom_point(aes(color = winner.share.upd)) + scale_color_continuous_divergingx(palette = "ArmyRose", mid = .5) +  xlab("Fraud") + theme_bw()
 
 #  ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= max_size)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("max size") + theme_bw()
+  library(colorspace)
+
+ ggplot(nelda2.sub %>% filter(inc.lose != 1) , aes(x=winner.share.upd, y=n.events.post, color = v2elirreg.inv)) +
+    geom_point() +
+    scale_color_continuous_divergingx(palette = "ArmyRose", n_interp = 11, cmax1 = 500, cmax2 = 500) +
+    theme_bw() +
+   coord_cartesian(ylim = c(0, 100))
+
+ ggplot(nelda2.sub %>% filter(inc.lose != 1), aes(x=winner.share.upd, y=jitter(any.ecav.protest), color = v2elirreg.inv)) +
+    geom_point() +
+    scale_color_continuous_divergingx(palette = "ArmyRose", n_interp = 11, cmax1 = 500, cmax2 = 500) +
+    theme_bw()
+
+  ggplot(nelda2.sub %>% filter(inc.lose != 1 & d.events.num > 0), aes(x=winner.share.upd, y=d.events.num, color = v2elirreg.inv)) +
+    geom_point(aes(shape = as.factor(protest.success))) +
+    scale_color_continuous_divergingx(palette = "ArmyRose", n_interp = 11, cmax1 = 500, cmax2 = 500) +
+    theme_bw()
+
+  library(latticeExtra)
+  # showing data points on the same color scale 
+levelplot(any.ecav.protest ~ v2elirreg.inv * winner.share.upd, nelda2.sub %>% filter( protest.success == 0 & inc.lose == 0), 
+          panel = panel.levelplot.points, cex = 1.2
+    ) + 
+    layer_(panel.2dsmoother(..., n = 300))
+
+  # showing data points on the same color scale 
+levelplot(d.events.num ~ v2elirreg.inv * winner.share.upd, nelda2.sub %>% filter(d.events.num > 0 & protest.success == 0 & inc.lose == 0), 
+          panel = panel.levelplot.points, cex = 1.2
+    ) + 
+    layer_(panel.2dsmoother(..., n = 300))
+
+  testlogit <- glm(protest.success ~ v2elirreg.inv*winner.share.upd + presidential + inc.termlimit.num + n.events.pre, data = nelda2.sub)
+summary(testlogit)
+sjPlot::plot_model(testlogit, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.08, 1.96]"))
 ```
@@
 
-Lastly, I include the *number of pre-election protests* counted in the ECAV dataset, which is likely to be an important possible confounder; more numerous pre-election protest may affect the level of election fraud (as the deterrence logic predicts) as well as being directly tied to the likelihood of post-election protest.
+I include the *number of pre-election protests* counted in the ECAV dataset, which is likely to be an important possible confounder; more numerous pre-election protest may affect the level of election fraud (as the deterrence logic predicts) as well as being directly tied to the likelihood of post-election protest.
+Finally, in the models of protest duration, I also include a measure of the *median size of protest*, calculated using ECAV data.^[ECAV data codes protest size as a categorical variable, ranging from 1 (< 10 participants) to 5 (> 10,000).
+I take the median size category of all the events in the protest wave as a measure of average protest size.] This is important as a control for the possibility that incumbents could tolerate smaller protests, as discussed above.
 
@@
 ## Fraud and the likelihood of protest
+
+Table \@ref(tab:logit-table) shows the results for Hypothesis 1.
+Model 1 is a base model, where incumbent loss, winner vote-share, and *election fraud* are standalone variables.
+In keeping with the general (though inconsistent) findings in the prior literature, there is a significant positive relationship between fraud severity and the likelihood of protest initiation.
+The winner's claimed margin of victory is negatively associated with protest initiation.
+Model 2 includes the three-way interaction between *election fraud*, incumbent loss, and winning vote-share, with no controls other than a dummy variable for *presidential elections*.
+This allows for the largest possible sample size, and shows that the coefficient signs are consistent with those of the full Model 3.
 
@@
 
-model.binary_noint <- glm(any_protest ~ inc.lose + v2elirreg.inv + winner.share.upd, data = nelda2.sub, family = binomial(link = "logit"))
-
-model.binary.base <-  glm(any_protest ~ winner.share.upd*v2elirreg.inv*inc.lose + presidential, data = nelda2.sub, family = binomial(link = "logit"))
-
-model.binary.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   +    winner.share.upd*v2elirreg.inv*inc.lose   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre + rentierstate + taxstate + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub)
-
-plot_binary_margin_typ_n <- sjPlot::plot_model(model.binary.full, type = "pred", terms = c("v2elirreg.inv", "winner.share.upd [.35, .75]", "inc.lose"), show.data = T)
-plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Electoral manipulation", y = "Predicted probability", title = "Predicted probability of protest", color = "Winner's vote-share") + theme_bw()
+model.binary_noint <- glm(any_protest ~ inc.lose_clean + v2elirreg.inv + winner.share.upd, data = nelda2.sub, family = binomial(link = "logit"))
+
+model.binary.base <-  glm(any_protest ~ winner.share.upd*v2elirreg.inv*inc.lose_clean + presidential, data = nelda2.sub, family = binomial(link = "logit"))
+
+model.binary.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   +    winner.share.upd*v2elirreg.inv*inc.lose_clean   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre + rentierstate + taxstate + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub)
+
+plot_binary_margin_typ_n <- sjPlot::plot_model(model.binary.full, type = "pred", terms = c("v2elirreg.inv", "winner.share.upd [.35, .78]", "inc.lose_clean"))
+#plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Winner's vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()
 ```
@@
 modelsummary(models2, output = "default",  stars = T, title = 'Table 3: Logit models of protest occurrence', gof_omit = c("Log|RMSE"), coef_map = c("(Intercept)" = "(Intercept)",
-             "inc.lose1" = "Incumbent loss",
+             "inc.lose_clean2" = "Incumbent loss",
              "v2elirreg.inv" = "Election fraud",
@@
 "winner.share.upd:v2elirreg.inv" = "Winner vote-share:Election fraud",
-"winner.share.upd:inc.lose1" = "Winner vote-share:Inc. loss",
-"v2elirreg.inv:inc.lose1" = "Election fraud:Inc. loss",
+"winner.share.upd:inc.lose_clean2" = "Winner vote-share:Inc. loss",
+"v2elirreg.inv:inc.lose_clean2" = "Election fraud:Inc. loss",
 "winner.share.upd:v2elirreg.inv:inc.lose1" = "Election fraud:Winner vote-share:Inc. loss",
@@
 
-```
-
-Table \@ref(tab:logit-table) shows the results for Hypothesis 1.
-Model 1 is a base model, where incumbent loss, winner vote-share, and *election fraud* are standalone variables.
-In keeping the general (though inconsistent) findings in the prior literature, there is a significant positive relationship between fraud severity and the likelihood of protest initiation.
-The winner's claimed margin of victory is negatively associated with protest initiation.
-Model 2 includes the three-way interaction between *election fraud*, incumbent loss, and winning vote-share, with no controls other than a dummy variable for *presidential elections*.
-This allows for the largest possible sample size, and the results are consistent with those of the full model with controls.
+#reg_table |> tab_options(
+#  latex.use_longtable = TRUE, 
+#  data_row.padding = px(5)
+#  ) 
+```
 
 Model 3 is the main model, which includes the full set of controls.
-The interaction is significant, as Figure \@ref(fig:margin-prob-fig) illustrates.
-The figure shows the predicted probability of post-election protest for two specific values of election fraud, across all values of incumbent vote-share, holding all other variables constant at their means (continuous variables) or modes (categorical variables).
+Given the complexity of the model, Figure \@ref(fig:margin-prob-fig) illustrates the effects of election fraud on protest risk across various levels of winning vote-share and incumbent loss^[All other variables constant at their means (continuous variables) or modes (categorical variables).].
 The values for election fraud represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively.
+
 The direction of the effect supports Hypothesis 1.
-That is, the occurrence of any protest is most likely when incumbent vote-share is low and fraud is high.
-This suggests that either manipulation is a sign of weakness and protest is centralized and strategic, *or* protest is often non-strategic.
-In the latter case, we cannot yet determine whether manipulation is a signal of strength or weakness.
-To make this determination, we turn to the models of protest duration.
+When an incumbent claims victory (left panel) and their vote-share is low, electoral protest is significantly likelier when fraud is elevated than when the election is relatively clean.
+Furthermore, protest against winning incumbents is significantly likelier when fraud is high and vote-shares are low, compared to an election with an equivalent level of fraud but high claimed vote-shares.
+As noted above, such a finding could be consistent with either the regime-weakness or regime-strength models of manipulation, under different conceptualizations of protest.
+To distinguish between the two, we turn to the models of protest duration.
 
@@
 
-cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" &  n.events.post > 0))
+cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose != "1" &  n.events.post > 0))
 summary(cox1)
@@
 summary(cox2)
-p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.33, 2.2]"), colors = "gs") #This is the 10th and 90th quantiles
+p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.33, 2.2]")) #This is the 10th and 90th quantiles
 p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()
@@
 There is considerable missingness in this variable.
+Lastly, in order to intrepret these findings, it is important to recall that these models include only cases where incumbents claimed victory and electoral protest occurred.
 
@@
 Figure \@ref(fig:margin-dur-fig) shows the predicted hazard ratios As shown in the figure, for all incumbent vote-shares near and below fifty percent, large-scale fraud significantly increases the likelihood that a protest movement will end.
-This result also supports the fraud-as-strength model, and shows that only as vote-share increases above fifty percent\textemdash creating a stronger signal of popularity\textemdash does a government in a low-fraud election have the same likelihood of defeating a protest movement at a given time.
-In effect, either high genuine popularity or the ability to produce high levels of fraud enables governments to more quickly end protests.
+This result also supports the regime-strength model, and shows that only as vote-share increases above fifty percent---creating a stronger signal of popularity---does a government in a low-fraud election have the same likelihood of defeating a protest movement at a given time.
+In effect, either high genuine popularity *or* the ability to produce high levels of fraud enables governments to more quickly end protests.
 
 The results of the baseline models are consistent with those models which incorporate censoring.
-While the coefficient on *election fraud* falls slightly short of conventional statistical significance in Model 6 (p = .059), it retains the same size and sign compared to Model 4.
+The coefficient for *election fraud* retains the same size, sign, and level of significance as in Model 4.
 The interaction effect in Model 7 increases in size, while retaining its direction and statistical significance.
 This suggests that the main models are not drawing overly incorrect inferences through censoring of longer-lasting protest movements.
-Model 8, adds the median number of participants at events in the protest wave as a control.
+Model 8, adds the median number of participants at events in the protest wave as a control for the potential perceived threat of the protest movement by the incumbent.
 Including this control variable reduces the sample size due to missingness, but the interaction coefficient retains its direction and statistical significance.
@@
 In real contests in which opposition and incumbent strengths are tested against each other, better manipulators reveal themselves to be stronger rather than weaker.
-
-This finding supports and extends the line of research initiated by Simpser [@simpser2013].
-While that view holds that electoral manipulation in conjunction with a large margin of victory for the incumbent can induce pro-regime behavior in society, this paper shows that even when such cooperation fails\textemdash when post-election protests occur\textemdash governments that manipulate more are more likely to successfully weather the challenge.
+This is consistent with Hypothesis 2.
+
+This finding supports and extends the line of research initiated by Simpser [@simpser2013], which I have called the regime-strength model of election manipulation.
+While Simpser's original argument holds that electoral manipulation in conjunction with a large margin of victory for the incumbent can induce pro-regime behavior in society, this paper shows that even when such cooperation fails---when post-election protests occur---governments that manipulate more are more likely to successfully weather the challenge.
 That is, even when ruling party vote-shares are low (perhaps especially when this is so), incumbents which show themselves capable of large-scale fraud are more capable of quickly ending protest movements than low-fraud incumbents.
@@
 
-Protest is most likely to be initiated under the same circumstances when it is likely to be defeated quickly, which is puzzling if we assume that protests are initiated strategically (i.e.
-when opposition groups believe the incumbent is weak).
-This finding raises more questions than answers, and it is beyond the scope of this paper to resolve those questions.
-However, taking the results in total suggests that 1) large-scale election manipulation indicates regime strength in most cases, and 2) electoral protest behavior involves a mix of actors with strategic and non-strategic motives.
+This study finds support for Hypothesis 1---that protest will be most likely to begin when fraud is high and incumbent vote-shares low---*and* for the notion that governments that manipulate more effectively will more quickly quash protests.
+In other words, protest is most likely to be initiated under the same circumstances when it is likely to be defeated quickly.
+This combination of findings is a blow to the regime-weakness model, and the associated idea that protest risk can effectively deter election manipulation.
+I suggests, instead, that 1) large-scale election manipulation indicates regime strength in most cases, and 2) electoral protest behavior involves a mix of actors with broadly strategic motives that go beyond ousting the government.
 Such motives are discussed in more detail in the theory section.
 
-```{r group-means-table, echo=FALSE, warning=FALSE, message=FALSE}
-group_means <- read.csv(here::here("Data", "group_means_table.csv"))
-group_means <- group_means %>% rename("Variable" = "X")
-group_means <- group_means %>% rename("Successful protest" = "Successful.protests")
-group_means <- group_means %>% rename("Failed protest" = "Failed.protests")
-group_means <- group_means %>% rename("P-value" = "P.value")
-group_means <- group_means %>% rename("SD (full sample)" = "SD..full.sample.")
-
-kable(group_means, booktabs = T, digits = 3, caption = "Table 3: Group means for successful and failed protests, with t-test p-values")  %>%
-    kable_styling(latex_options = "scale_down")
-```
-
+<!--
 Additionally, it is likely that when strategic protest entrepreneurs are considering initiating protest, they must consider how likely it is that ordinary citizens will join them [@buenodemesquitaRegimeChangeRevolutionary2010].
@@
 
-A very brief consideration of the seven cases of successful electoral protest\textemdash Ukraine (2004), Georgia (2003), Azerbaijan (2005), Kyrgyzstan (2005), Malaysia (2004), Peru (2000) and Cote d'Ivoire (2000)\textemdash in comparison with the other cases can help aid interpretation of the results.
-Table 3 shows the mean values for several key variables for two groups: cases where incumbents were defeated by protest, and cases where they survived protest.
+A very brief consideration of the cases of successful electoral protest in comparison with the failed cases can help aid interpretation of the results.
+Table 3 shows the mean values for several key variables for two groups: cases where protests resulted in incumbents' ouster, and cases where protest movements were quashed.
 The table also shows p-values for t-tests comparing the means of the two groups.
 While the level of fraud is typically higher in the successful protest group than the unsuccessful group, this difference is not significant.
-By contrast, incumbent vote-shares are significantly lower in the successful-protest and protest groups than their counterparts (at the p < .1 level).
-Similarly, pre-election protests that drew larger crowds were more common in the successful-protest cases than the unsuccessful ones (at the p < .1 level).
-This supports the notion that strategic protest entrepreneurs consider the likelihood that citizens will respond positively to the call to the streets en masse.
+By contrast, incumbent vote-shares are lower in the successful-protest and protest groups than their counterparts (though this difference is only significant at the p < .1 level).
+Only two covariates are different at the conventional level of significance.
+Cases where protest succeeds in topping governments tend to be more heavily urbanized (by 10 percentage points) and have an economy less dominated by the state (by sixty-four percentage points!).
+This is further evidence that a high degree of electoral manipulation is not indicative of regime weakness; more structural variables appear to drive regime strength, in keeping with the model outlined here.
+
+```{r, include=F, warning=FALSE, message=FALSE, echo=FALSE}
+## Anova for protest success
+anova_dat <- protest.cases %>% dplyr::select(protest.success, v2elirreg.inv,  jucon.1lag, legcon.1lag, gdpgro.1lag,  osorg.1lag, natlpartyorg.lag,  urban.pct, n.events.pre, state.owned.1lag, patronage.appts.1lag, v2elembaut, winner.share.upd)
+
+# boxplots and t-tests for the 4 variables at once
+for (i in 2:ncol(anova_dat)) { # variables to compare are variables 2 to end
+  boxplot(anova_dat[, i] ~ anova_dat$protest.success, # draw boxplots by group
+    ylab = names(anova_dat[i]), # rename y-axis with variable's name
+    xlab = "Incumbent loss"
+  )
+  print(paste(c("ANOVA for variable ", (names(anova_dat[i])))))
+  print(t.test(anova_dat[, i] ~ anova_dat$protest.success)) # print results of t-test
+}
+
+aov_output <- aov(state.owned.1lag ~ inc.lose, data = nelda2.sub)
+summary(aov_output)
+```
+
+```{r anova-table, echo=FALSE, warning=FALSE, message=FALSE}
+#p.values <- c(3.009e-15, 6.014e-11, 6.149e-10, 0.01146,  3.758e-07,  0.6281, 0.003519,
+#              0.1852, 0.007405, 0.003086, 2.2e-16)
+#win_lose_diff <- c(.87, -0.15, -0.16, .06, -.47, .04, 5.3, -1.46, -.265, -.257, -.893)
+#variables <- c("Election fraud", "Judicial constraints", "Legislative constraints", "GDP growth (lag)", "Civil society openness (lag)", "Natl. party institutionalization (lag)", "Urban",
+#               "N. pre-election protests", "Privatization", "Professional bureaucracy", "EMB autonomy")
+#anova_dat <- data.frame(cbind(variables,win_lose_diff, p.values))
+#anova_dat <- anova_dat %>% mutate(win_lose_diff = as.numeric(win_lose_diff))
+#anova_dat <- anova_dat %>% mutate(p.values = as.numeric(p.values))
+#anova_dat <- anova_dat %>% rename(Variables = variables)
+#anova_dat <- anova_dat %>% rename("Win-Loss Difference" = win_lose_diff)
+#anova_dat <- anova_dat %>% rename("P-value" = p.values)
+
+##Commented out data above is for election win / loss by incumbent
+# Below we have protest success / failure
+
+anova_dat <- read.csv(here::here("Data", "anova_protest_success.csv"))
+
+anova_dat <- anova_dat %>% rename("Success-Fail Difference" = Success.failure.diff.)
+anova_dat <- anova_dat %>% rename("P-value" = p.value)
+anova_dat <- anova_dat %>% rename("Protest Failure" = Protest.failure)
+anova_dat <- anova_dat %>% rename("Protest success" = Protest.success)
+
+protest_table <- flextable(anova_dat)
+protest_table <-  set_caption (protest_table, "Table 4: Mean variable values for failed and successful protests, with ANOVA p-values") 
+protest_table
+
+```
 
@@
 A thorough qualitative comparative analysis of the cases would be informative.
+-->
 
@@
 They suggest that factors like the opposition's capacity to organize street protest in limiting election fraud have been overestimated.
-In fact, these results suggest that incumbents who have the resources may respond to such a scenario by *increasing* their production of fraud to secure a higher vote-share.
-Consequently, theories of electoral manipulation that focus on factors other than protest risk\textemdash such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]\textemdash may be more important for understanding election integrity, which may drive future research as well as policy interventions.
-Additionally, it may change some current understandings of the role of election monitors, courts, and other third parties in upholding election integrity.
+Consequently, theories of electoral manipulation that emphasize other factors---such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]---should be further developed, and may yield new policy interventions.
+In particular, it may change some current understandings of the role of election monitors, courts, and other third parties in upholding election integrity.
 As noted earlier, prior studies argue that these actors restrain manipulation by revealing information about fraud and intensifying protest risk.
-But if protest risk is largely untethered from the severity of fraud, it is likely that other mechanisms are at work.
+However, if more extensive manipulation is associated with more quickly subduing protest movements, it may be that other mechanisms are at work.
 For instance, it may be that election observation makes it harder for front-line agents to avoid personal consequences for engaging in manipulation [@Harvey2022].
 
-There is a long-standing debate over the role of elections in democratization, of which the manipulation-protest nexus is a component [@Schedler2013;@Morgenbesser2019;@edgellWhenWhereElections2018].
-At the core, much of this debate revolves around the extent to which non-democratic elections affect the ruling party's cost of suppressing the opposition versus tolerating it [@dahlPolyarchyParticipationOpposition1971].
-Election manipulation, including fraud, directly reduces the cost of toleration, by making it less likely that opposition groups win elections.
-This makes incumbents more willing to tolerate opposition in contested, albeit biased, elections.
-To the extent that it makes electoral protest more likely and/or more costly, however, it increases the cost of suppression.
-One view holds that election fraud has the potential to lay the groundwork for democratization by incentivizing the ruling party to tolerate organized opposition which, one day, it may be unable to repress after an election.
-This logic can be found in studies of the color revolutions [@bunce2010], and in the observation that competitive authoritarian regimes are more likely to democratize after an election than hegemonic ones [@donno2013elections]
-
-By showing that the infrastructure of fraud can actually reduce the cost of post-election protest to the regime, this result reverses that logic.
-Instead, the ability to produce large-scale election fraud can both reduce the cost of toleration *and* the cost of suppression.
-This implies that electoral authoritarianism is something of a cul-de-sac: the capacity to generate severe election fraud makes it appealing for closed authoritarian regimes to introduce elections, while the manageable risk of protest makes it unappealing to further democratize.
-This logic also creates incentives for ruling parties in weakly consolidated democracies to take measures that will enable them to engage in electoral manipulation; namely, the consolidation of patronage resources and erosion of rule of law.
-This provides additional theoretical support for understanding what some have called a 'wave of autocratization' or a 'democratic recession' in recent years [@luhrmannThirdWaveAutocratization2019;@diamondDemocraticRegressionComparative2021].
-If the same tools that produce broad election fraud also reduce the cost of protest, then electoral authoritarianism exerts a sort of gravitational pull as a regime type.
+By showing that the infrastructure of fraud can actually reduce the cost of post-election protest to the regime, this result poses a challenge to theory of democratization by election [@Schedler2013;@Morgenbesser2019;@edgellWhenWhereElections2018].
+If the capacity to engage in widespread manipulation makes it easier to both tolerate the existence of an opposition and to suppress it during periods of contestation [@dahlPolyarchyParticipationOpposition1971], electoral authoritarianism is something of an institutional cul-de-sac: the capacity to generate severe election fraud makes it appealing for closed authoritarian regimes to introduce elections, while the manageable risk of protest makes it unappealing to further democratize.
 
 One key prediction of the strategic protest model is not tested here directly, because the research design focuses on the strength of incumbents against protest movements as revealed by actual protest.
-However, the protest-deterrent model predicts that weaker incumbents will generate little fraud and voluntarily step down after losing elections.
-There is indeed less election fraud observed in those election-years: the mean estimate of *election fraud* in the 129 observations where incumbents lost is 0.322, much lower than the 1.20 average in the group where incumbents claimed victory.
-Were these incumbents deterred from committing worse fraud?
-Or are they simply systematically more limited in their ability to generate election fraud?
+However, several additional tests are presented in the appendix which suggest that the limits on election manipulation are largely structural (e.g.
+state capacity, judicial independence) rather than narrowly strategic.
+In fact, these additional checks indicate that, if anything, incumbents tend to respond to elevated protest risk by *increasing* production of electoral manipulation.
+
+<!--
 
 ```{r, include=F, warning=FALSE, message=FALSE, echo=FALSE}
-## Anova
+## Anova for winning vs. losing
 anova_dat <- nelda2.sub %>% dplyr::select(inc.lose, v2elirreg.inv,  jucon.1lag, legcon.1lag, gdpgro.1lag,  osorg.1lag, natlpartyorg.lag,  urban.pct, n.events.pre, state.owned.1lag, patronage.appts.1lag, v2elembaut)
@@
 
-aov_output <- aov(state.owned.1lag ~ inc.lose, data = nelda2.sub)
-summary(aov_output)
-```
-
-```{r anova-table, echo=FALSE, warning=FALSE, message=FALSE}
-p.values <- c(3.009e-15, 6.014e-11, 6.149e-10, 0.01146,  3.758e-07,  0.6281, 0.003519,
-              0.1852, 0.007405, 0.003086, 2.2e-16)
-win_lose_diff <- c(.87, -0.15, -0.16, .06, -.47, .04, 5.3, -1.46, -.265, -.257, -.893)
-variables <- c("Election fraud", "Judicial constraints", "Legislative constraints", "GDP growth (lag)", "Civil society openness (lag)", "Natl. party institutionalization (lag)", "Urban",
-               "N. pre-election protests", "Privatization", "Professional bureaucracy", "EMB autonomy")
-anova_dat <- data.frame(cbind(variables,win_lose_diff, p.values))
-anova_dat <- anova_dat %>% mutate(win_lose_diff = as.numeric(win_lose_diff))
-anova_dat <- anova_dat %>% mutate(p.values = as.numeric(p.values))
-anova_dat <- anova_dat %>% rename(Variables = variables)
-anova_dat <- anova_dat %>% rename("Win-Loss Difference" = win_lose_diff)
-anova_dat <- anova_dat %>% rename("P-value" = p.values)
-
-kable(anova_dat, booktabs = T, digits = 3, caption = "Table 4: Mean variable values for winning incumbents minus losing incumbents, with ANOVA p-values")  %>%
-    kable_styling(font_size = 9, full_width = FALSE)
-
-```
-
-The overall pattern in the data leans toward the latter interpretation.
-Table \@ref(tab:anova-table) shows the mean differences between the two groups, with ANOVA p-values for those differences.
-Negative values indicate that the losing incumbents had higher values for the variable, on average.
-For instance, among the cases where the incumbent lost, judicial independence, professionalization of the bureaucracy, economic privatization, civil society openness, and the autonomy of election management bodies are all significantly higher.
-All of these should tend to reduce the availability of fraud as a strategy.
-Of the factors considered here, only *number of pre-election protests* clearly behaves as the deterrence logic would predict, with the losing group subject to one more protest, on average, than the winning group.
-However, unlike the factors which point toward the alternative hypothesis, this difference is not statistically significant.
-Additionally, the winning group is significantly more urbanized, which cuts against the deterrence logic given that more urban environments are more prone to mass protest [@Wallace2013a].
-Thus, while future research might test this aspect of the deterrence model more rigorously, the general pattern is supportive of the view put forth in this paper.
-Further checks are performed in the appendix, including models which show that predicted levels of fraud based on structural factors alone are strongly predictive of incumbent defeat, while pre-election measures of risk are not.
-
+```
+-->
 Finally, this paper does not examine the role of protest risk in insulating established democracies from democratic backsliding [@fearon_self-enforcing_2011].
 This question deserves further investigation, but it is possible to draw a few implications from this study.
-First, even if protest risk fails to serve as a deterrent in non-democracies, it may still do so in democracies.
+First, even if protest risk fails to serve as a deterrent in non-democracies, it may still do so in more democratic settings.
 In non-democracies, protest initiation is most likely when fraud is very high and incumbent vote-shares very low.
-The likelihood of protest initiation in a democracy may well be higher, triggering at lower levels of fraud and higher incumbent vote-shares, since opposition actors are likely to be better resources in a liberal democracy (and the government more constrained).
+The likelihood of protest initiation in a democracy may well be higher, triggering at lower levels of fraud and higher incumbent vote-shares, since opposition actors are likely to be better resourced in a liberal democracy (and the government more constrained).
 At the same time, if the theory described here is accurate, liberal democracies would be constrained against using large-scale election fraud even in the absence of protest risk: they lack the patronage resources and arbitrary enforcement of laws needed for such a scenario.
 
+<!--
 It is also worth noting that mass electoral protest can serve an illiberal purpose, as illustrated by the events of January 6, 2020 in the United States or 2022-2023 in Brazil.
-Indeed, out of the 156 instances of post-election protest in this dataset, 10 occurred after relatively clean elections (*intentional irregularities* < 0)\textemdash a small but not minuscule percentage.
+Indeed, out of the 156 instances of post-election protest in this dataset, 10 occurred after relatively clean elections (*intentional irregularities* < 0)---a small but not minuscule percentage.
 These incidents demonstrate that, while crying fraud may sometimes be a useful tool for protest entrepreneurs, losing an election alone can be a catalyst for protest, irrespective of the actual level of election manipulation.
 This possibility, which would be puzzling if fraud indicates weakness, highlights the often non-strategic nature of protest initiation while also underlining the fact that the course of protest is often driven more by the relative resources and opportunities available to the disaffected and the incumbent government.
+-->
 
@@
 
-There are competing views over the nature of election fraud, incumbent strength, and protest.
-The risk of post-election protest has a central role in several prominent models of electoral manipulation, serving to deter manipulation and to incentivize the toleration of election monitors and other restraints on the ruling party.
-Others contend that election fraud can indicate incumbent strength.
+There are competing views on the relationship between election fraud and protest.
+In one family of models, which I have called the regime-weakness model, incumbent's may hold cleaner elections when the risk of post-election protest is higher.
+Other models contend that election fraud can indicate incumbent strength.
 Building on the principal-agent literature of election fraud, I argue that large-scale fraud draws on the same resources that make incumbents resilient to protest, even when regime popularity (and thus vote-shares) are low.
 Conducting election fraud is made easier when incumbents have control over patronage resources and the ability to protect front-line agents from the consequences of illegal behavior; this capacity also helps insulate incumbents from the effects of mass protest.
-Resolving this tension in the literature has important implications for the study of election fraud and authoritarian stability.
-
-This paper attempts to test the assumptions of these two models by investigating the relationship between election fraud and protest occurrence and duration.
-Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 445 elections in electoral authoritarian regimes to show that protest occurrence is most likely when incumbent vote-shares are very low and election fraud is high, as both models predict.
+Resolving the tension between these views in the literature has important implications for the study of election fraud and authoritarian stability.
+
+This paper attempts to test the assumptions of these two models by investigating the relationship between election fraud and two protest variables: initiation and duration.
+Specifically, both the regime-weakness and regime-strength models may make the same predictions regarding protest onset; their predictions differ with regard to protest duration.
+Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 659 elections in electoral authoritarian regimes to show that protest occurrence is most likely when incumbent vote-shares are very low and election fraud is high, as both models predict.
 However, and most importantly, incumbents that engage in larger-scale election fraud are able to more quickly defeat protest movements.
 
-All together, these results suggest that protest risk is unlikely to serve as a constraint on election manipulation in authoritarian regimes\textemdash the most effective election riggers are also the most well-prepared to defeat protest movements.
+All together, these results suggest that protest risk is unlikely to serve as a constraint on election manipulation in authoritarian regimes in the way that the regime-weakness model predicts---the most effective election manipulators are also the most well-prepared to defeat protest movements.
 The frailty of protest risk as a guardrail suggests that electoral authoritarianism as a regime-type is a stubborn equilibrium.
-This is a gloomy picture for supporters of democratization, and suggests that researchers should consider other mechanisms beyond mass protest in their models of electoral manipulation.
+This is a gloomy picture for supporters of democratization, and suggests that researchers should further consider other mechanisms beyond mass protest in their models of electoral manipulation.
 
@@
 
-The data analyzed in this study will be made available at the Harvard Dataverse upon acceptance of the paper.
+The data analyzed in this study will be made available at the Dataverse upon acceptance of the paper.
 
@@
 # References
+
+# TEst stuff
+
+```{r include=FALSE, eval=FALSE}
+library(lavaan)
+library(dplyr)
+
+# Prepare country-year level data
+nelda2.sub <- nelda2.sub %>%
+  mutate(
+    # For country-years with protests: inverse of mean duration
+    # Higher values = shorter duration = stronger regime
+    protest_duration_inverse = if_else(
+      any.ecav.protest == 1,
+      1 / d.events.num,  # or -log(mean_duration)
+      NA_real_
+    )
+  )
+
+nelda2.sub <- nelda2.sub %>%
+  mutate(taxstate.num = as.numeric(taxstate))
+nelda2.sub <- nelda2.sub %>%
+  mutate(rentierstate.num = as.numeric(rentierstate))
+nelda2.sub <- nelda2.sub %>%
+  mutate(unscheduled_election.num = as.numeric(unscheduled_election))
+
+# MIMIC model
+mimic_spec <- '
+  # Latent variable
+  regime_strength =~  gdpgro.1lag +v2elirreg.inv +   physinteg.1lag + urban.pct + rentierstate.num + taxstate.num +  unscheduled_election.num + inc.termlimit.num + winner.share.upd
+
+  # Structural part: causes of regime strength
+  protest_duration_inverse ~ regime_strength  + winner.share.upd
+
+  # Allow correlation between indicators
+ # protest_duration_inverse ~~ any.ecav.protest
+ v2elirreg.inv ~~ physinteg.1lag 
+ v2elirreg.inv ~~ urban.pct
+ v2elirreg.inv ~~  inc.termlimit.num
+'
+
+# Fit with FIML to handle missing protest_duration_inverse
+fit <- sem(mimic_spec, 
+           data = nelda2.sub %>% filter(d.events.num >0 & inc.lose == 0),
+           missing = "fiml")  # Full Information Maximum Likelihood
+
+summary(fit, fit.measures = TRUE, standardized = TRUE)
+
+# Extract latent scores
+regime_scores <- lavPredict(fit, type = "lv")
+data_cy$regime_strength_estimated <- regime_scores[, 1]
+```
+
+```{r eval=FALSE}
+# Step 1: Create a composite measure of regime strength from protests
+# (for country-years with protests)
+
+nelda2.sub <- nelda2.sub %>%
+  mutate(
+    # Composite indicator (you can refine this)
+    strength_indicator = case_when(
+      any.ecav.protest == 0 ~ NA_real_,  # Missing for no protests
+      any.ecav.protest == 1 ~ -log(d.events.num) # Lower duration = higher strength
+    )
+  )
+
+# Step 2: SEM with measurement error correction
+sem_spec <- '
+  # Latent variable with single indicator
+  # (need to fix measurement error variance)
+  regime_strength =~ 1*strength_indicator
+
+  # Estimate measurement error (you can set this based on reliability)
+  strength_indicator ~~ error_var*strength_indicator
+
+  # Structural model
+  regime_strength ~  v2elirreg.inv + gdpgro.1lag +   physinteg.1lag
+'
+
+# Fit model
+fit2 <- sem(sem_spec, 
+            data = nelda2.sub,
+            missing = "fiml")
+
+summary(fit2, standardized = TRUE)
+```
+```{r eval=FALSE}
+# This is cleaner and avoids identification issues
+path_model <- '
+  # Protest occurrence predicted by covariates
+  any.ecav.protest ~ v2elirreg.inv + gdpgro.1lag +   physinteg.1lag
+
+  # Protest duration (inverse) predicted by same covariates
+  # Only estimated for observations with protests
+  protest_duration_inverse ~ v2elirreg.inv + gdpgro.1lag +   physinteg.1lag
+
+  # Allow correlation between equations
+  any.ecav.protest ~~ protest_duration_inverse
+'
+
+fit_path <- sem(path_model, 
+                data = nelda2.sub,
+                missing = "fiml")
+
+summary(fit_path, standardized = TRUE, fit.measures = TRUE)
+```
+
+```{r eval=FALSE}
+# Use only country-years WITH protests
+data_protests_only <- nelda2.sub %>%
+  filter(any.ecav.protest == 1)
+
+# Simple regression path model
+protest_model <- '
+  # Duration as function of covariates
+  protest_duration_inverse ~ legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv
+
+  # If you have multiple protests per country-year, 
+  # you might want to include number of protests as control
+  protest_duration_inverse ~ n.events.post
+'
+
+fit_protests <- sem(protest_model,
+                    data = data_protests_only)
+
+summary(fit_protests, standardized = TRUE, rsquare = TRUE)
+
+# This tells you:
+# 1. Which covariates predict shorter duration (stronger regimes)
+# 2. How much variance in duration is explained
+# 3. Specific effect of election fraud on regime strength
+```
+
+Structural equation modelling doesn't seem the way to go
+
+```{r eval=FALSE}
+
+df <- nelda2.sub %>% filter(inc.replaced.prot == 0) %>% dplyr::select(legcon.1lag, altinf.1lag,  osorg.1lag, gdpgro.1lag, inc.termlimit.num,  physinteg.1lag, physintlag.sq, urban.pct, rentierstate.num, taxstate.num ,  unscheduled_election.num, inc.termlimit.num, v2elirreg.inv, NeldaID)
+
+df <- df %>% filter(complete.cases(.))
+
+CFA.model <- ' regime_strength  =~  v2elirreg.inv + physinteg.1lag + physintlag.sq + altinf.1lag + osorg.1lag + legcon.1lag   + gdpgro.1lag + inc.termlimit.num  +  urban.pct + rentierstate.num + taxstate.num +  unscheduled_election.num + inc.termlimit.num'
+
+fit <- cfa(CFA.model, data = df, std.lv=TRUE)
+summary(fit)
+
+regime_scores <- lavPredict(fit)
+
+df <- df %>% mutate(rgm_st = regime_scores[,1])
+
+df <- df %>% left_join(nelda2.sub, by = "NeldaID")
+
+init_probit <- glm(any.ecav.protest ~ rgm_st*winner.share.upd + presidential, binomial(link = "probit"), data = df)
+df_probit <- init_probit$data
+df_probit <- df_probit %>%  dplyr::select(legcon.1lag.x, altinf.1lag.x,  osorg.1lag.x, gdpgro.1lag.x, inc.termlimit.num.x,  physinteg.1lag.x, physintlag.sq.x, urban.pct.x, rentierstate.num.x, taxstate.num.x ,  unscheduled_election.num.x, inc.termlimit.num.x, v2elirreg.inv.x, NeldaID, winner.share.upd, d.events.num, rgm_st)
+df_probit <- df_probit %>% filter(complete.cases(.))
+
+df_probit$IMR <- sampleSelection::invMillsRatio( init_probit )$IMR1
+
+library(survival)
+library(ggsurvfit)
+
+df <- df %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
+df_probit <- df_probit %>% mutate(surv_status = ifelse(inc.replaced.prot == 1, 0, surv_status))
+df_probit <- df_probit %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
+                                                                       ifelse(d.events.num >= 69, 0, 1)))
+
+cox1 <- coxph(Surv(d.events.num, surv_status) ~  rgm_st*winner.share.upd  +  median.participants + regional_success_lag, data = df)
+summary(cox1)
+
+## Unclear what's going on here. But the general idea is that election fraud negatively loads onto the 'regime strength' variable, which in turn has no apparent effect on duration.
+```
+
+```{=latex}
+\begin{table}[]
+\begin{tabular}{|l|l|l|l|l|}
+\hline
+\multicolumn{1}{|r|}{Protest is assumed to be:} &
+  \multicolumn{1}{c|}{Instrumental protest} &
+   &
+  \multicolumn{1}{c|}{Multipurpose protest} &
+   \\ \hline
+\multicolumn{1}{|r|}{Incumbent's genuine popularity:} & Low & High & Low & High \\ \hline
+Manipulation...                                       &     &      &     &      \\ \hline
+...obscures strength or indicates weakness &
+  (I) High and increasing &
+  (I): Low and increasing &
+  (I): High and increasing &
+  (I): Low and increasing \\ \hline
+...reflects strength &
+  (I): High and decreasing &
+  (I): Low and decreasing &
+  (I): High and increasing &
+  (I): Low and no effect \\ \hline
+\end{tabular}
+\caption{}
+\label{tab:my-table}
+\end{table}
+```

### 2025-11-02 Output/authoritarian_only_word_pop_rev.Rmd
Section on duration theory with citations

@@
 
+Mass protest is one of the two primary threats to authoritarian regimes [@svolik_politics_2012], and as such they often go to great lengths to prevent mass protest.
+These efforts include costly (and risky) investments in the repressive apparatus [@Escriba_Folch2013;@draguPreventiveRepressionTwo2019], co-optation of opposition parties [@reuter2015], and large-scale monitoring and censorship schemes [@King2013].
+And they often relentlessly hound would-be protest entrepreneurs to deter such behavior [@xuRepressCooptAuthoritarian2021a;@mehrlSecretPoliceOrganizations2024].
+Mass protest then, is clearly undesirable from the authoritarian incumbent's point of view, and we can plausibly assume that once an electoral protest movement begins, the incumbent will exert considerable effort to quash it.
+Governments that are stronger, relative to the protesting opposition, will defeat protest movements faster than will weaker governments.
+
+An important objection could be raised here.
+Might not a stronger incumbent, feeling secure in their power, be willing to tolerate opposition protest longer than a weak incumbent?
+It is true that incumbents do sometimes tolerate protest movements [@yuenNeitherRepressionConcession2017].
+However, there are two reasons this is unlikely to be the case systematically for the circumstances studied here.
+First, authoritarian governments are more likely to tolerate protests aimed at low-salience, local issues, and more likely to clamp down on protests that threaten the legitimacy of the regime.
+As electoral protests challenge the incumbent's ostensible victory and right to remain in office, they are directly threatening to the regime.
+Second, authoritarian leaders are working in an information-poor environment, in which mass protests can unexpectedly grow from small sparks to wildfires that quickly sweep away the regime [@Kuran1991].
+We should thus expect incumbents to work to defuse them as quickly as possible.
+
 Talk about protest duration here:

### 2025-11-03 Output/authoritarian_only_word_pop_rev.Rmd
Updates to theory section

@@
 Mass protest then, is clearly undesirable from the authoritarian incumbent's point of view, and we can plausibly assume that once an electoral protest movement begins, the incumbent will exert considerable effort to quash it.
-Governments that are stronger, relative to the protesting opposition, will defeat protest movements faster than will weaker governments.
+Governments that are stronger, relative to the protesting opposition, will defeat protest movements faster than will weaker governments.^[In both cases, this assume that incumbent survives the protest movement, by far the modal outcome.
+When the opposition wins, an outcome considered in some of the empirical models below, the assumption is reversed: a quicker opposition victory indicates a weaker incumbent.] As argued below, electoral manipulation can be demonstrative of capacities that may also be used to defeat protest movements.
 
@@
 However, there are two reasons this is unlikely to be the case systematically for the circumstances studied here.
-First, authoritarian governments are more likely to tolerate protests aimed at low-salience, local issues, and more likely to clamp down on protests that threaten the legitimacy of the regime.
-As electoral protests challenge the incumbent's ostensible victory and right to remain in office, they are directly threatening to the regime.
-Second, authoritarian leaders are working in an information-poor environment, in which mass protests can unexpectedly grow from small sparks to wildfires that quickly sweep away the regime [@Kuran1991].
-We should thus expect incumbents to work to defuse them as quickly as possible.
-
-Talk about protest duration here:
-
-Election manipulation is generally characterized as having legitimacy costs which reduce other actors' level of cooperation with the incumbent [@birch2011].
-The most visible, and arguably most studied, form of such legitimacy costs is mass post-election protest against the incumbent.
-There is a broad consensus in the literature that authoritarian incumbents find sustained, sizable protests undesirable, since they entail some risk of incumbent defeat and may require costly repression or concessions.
-I assume, as a result, that incumbents generally seek to demobilize protest; protest movements that survive longer indicate the incumbent is relatively weaker compared to the opposition.^[The implications of protest toleration by incumbents will be discussed later.] By contrast, when protest movements meet a quick demise, the incumbent can be considered relatively strong.^[In both cases, this assume that incumbent survives the protest movement, by far the modal outcome.
-When the opposition wins, an outcome considered in some of the empirical models below, the assumption is reversed: a quicker opposition victory indicates a weaker incumbent.] As argued below, electoral manipulation can be demonstrative of capacities that may also be used to defeat protest movements.
-
-There is an alternative view of electoral manipulation, however; Simpser's foundational argument holds that manipulated elections can send costly signals of the incumbent's "capacities, resources, [and] dispositions" that will help it retain power in the future [@simpser2013, p.
+First, authoritarian governments are more likely to tolerate protests aimed at low-salience, local issues [@lorentzenRegularizingRiotingPermitting2013], and more likely to clamp down on protests that have higher concession costs [@kleinDynamicsPoliticalProtests2018].
+As electoral protests challenge the incumbent's ostensible victory and right to remain in office, they are directly threatening to the regime; the concession costs of stepping down or holding new elections are very high.
+Second, authoritarian leaders are working in an information-poor environment [@guriev2020], in which mass protests can unexpectedly grow from small sparks to wildfires that quickly sweep away the regime [@Kuran1991].
+"Ignoring" protests can lead to further mobilization against authoritarian regimes [@Bishara2015], as can concessions in the absence of durable political reform [@leuschnerStateConcessionsProtest2024].
+We should thus expect incumbents to work to defuse electoral protests as quickly as possible.
+
+The central argument of this manuscript is that, in authoritarian regimes, incumbents who develop the tools necessary to manipulate elections extensively will also have the tools necessary to quickly prevail over protest movements.
+This proposition builds off Simpser's foundational argument that manipulated elections can send costly signals of the incumbent's "capacities, resources, [and] dispositions" that will help it retain power in the future [@simpser2013, p.
 80]---that is, of its political strength.
-I argue that more recent research on the principal-agent dynamics inherent in election manipulation reinforces and expands this idea.
+However, I depart from Simpser in two ways.
+First, by incorporating more recent research on the principal-agent dynamics inherent in election manipulation, I argue that incumbents who can produce large amounts of election manipulation necessarily have "capacities, resources and dispositions" helpful for defeating protests.
 This research trend emphasizes the degree to which political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].
-Incumbents that can solve these principal-agent problems at election-time, are also more likely to be able to survive post-election challenges.
+Second, mine is not a signalling argument; good manipulators are likely to have deeper wells of strength than weaker manipulators, whether those resources are successfully signaled to opposition actors or no.
 
@@
 As a result, governments that build and deploy an electoral manipulation apparatus are simultaneously inoculating themselves against the expected cost of protest.
-This yields Hypothesis 2.
-
-This true level of support can be obscured prior to elections due to pervasive preference falsification [@Kuran1991] and the low level of political information common in authoritarian regimes [@guriev2020].
-<- Keep this for risk of protests spiraling out of control
-
-Regime-weakness models of electoral manipulation, discussed above, treat election manipulation and genuine votes as substitute goods.
-For any given vote-share, an increase in manufactured votes reduces the number of genuine votes.
-In these models, which consider manufactured votes to be mere illusions of strength, the higher the number of manufactured votes for any given vote total, the weaker the incumbent.
-This means that strategic opposition actors are more likely to initiate protest when they perceive the incumbent to be weaker, and expect to be more likely to succeed in a challenge.
-In other words, manipulated votes are empty in the classic model; subtracting them from the incumbent's vote-total gives opposition actors their estimate of the incumbent's ability to win a post-election standoff.
-
-Thus the predicted relationship between manipulation and protest likelihood in the regime-weakness model is positive: more manipulation makes protest more likely for a given vote-share.
-However, the inverse is also to be expected: as incumbent vote-share increases for any fixed level of fraud, the opposition should perceive a stronger incumbent and be less likely to protest.
-In sum, the effect of fraud should be conditional on the claimed vote-share [@Rozenas2016b].
-This produces a first, baseline hypothesis.
+
+```{=latex}
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+\begin{tabular}{@{}rll@{}}
+\toprule
+Incumbent's genuine popularity:       & \multicolumn{2}{c}{Low}  \\ \midrule
+\multicolumn{1}{r|}{Protest is assumed to be:}                  & \multicolumn{1}{c|}{Instrumental} & \multicolumn{1}{c|}{Multipurpose} \\ \midrule
+\multicolumn{1}{|l|}{Manipulation...} &  & \multicolumn{1}{l|}{} \\ \midrule
+\multicolumn{1}{|r|}{...obscures strength / indicates weakness} & \multicolumn{1}{l|}{Increasing}   & \multicolumn{1}{l|}{Increasing}   \\ \midrule
+\multicolumn{1}{r|}{...reflects strength}                       & \multicolumn{1}{l|}{Decreasing}   & \multicolumn{1}{l|}{Decreasing}   \\ \bottomrule
+\end{tabular}
+\caption{Expected relationship between protest duration and increasing election manipulation}
+\label{tab:duration}
+\end{table}
+```
+
+How, then, would this view of electoral manipulation map onto the instrumental vs.
+multipurpose views of protest discussed above?
+While regime-weakness models have exclusively (to my knowledge) been framed around protest initiation, it is straightforward to extend their predictions to protest duration.
+Regime-weakness models of electoral manipulation treat election manipulation and genuine votes as substitute goods.
+For any given vote-share, an increase in manufactured votes reduces the number of genuine votes; that is, the higher the number of manufactured votes for any given vote total, the less likely the incumbent is to survive a challenge.
+In an instrumental mode of protest, strategic opposition actors perceive the incumbent's relative lack of popularity, and count on the majority of the public to back their challenge.
+Increasing manipulation indicates deeper unpopularity (holding vote-share constant), so we should expect a less popular incumbent to fail more quickly.
+The same logic holds, with slightly different mechanisms, if protests are assumed to be multipurpose.
+Here, greater manipulation is used to overcome deeper unpopularity (i.e.
+widespread grievances); more manipulation indicates an incumbent facing an increasingly powerful challenger.
+Thus, the regime-weakness model of manipulation should predict that the duration of unsuccessful protests will be increasing in election manipulation: more manipulation indicates a stronger challenger, which will take longer to defeat on average.
+
+For the regime-strength model of manipulation, the opposite relationship is expected, as laid out above.
+Comparing Table \@ref(tab:duration) and \@ref(tab:initiation) shows that the hypotheses generated by the preferred model--one in which manipulation indicates strength and protest is multipurpose are distinct from those in other quadrants.
+
+Need to discuss vote-share next:
 
 > H1: Increasing election fraud makes protest more likely, with the size of this effect decreasing with incumbent-party vote-share.
-
-It is important to note that the same hypothesis is also generated if we assume that protest initiation is broadly, rather than narrowly, strategic.
-The mechanism, however, is different.
-Here, the argument is that widespread grievances result in low natural popularity for the incumbent; at election time, the incumbent must compensate for this low popularity by manipulating the election.
-Protest initiation is made more likely both by the population's grievances alone [@Beaulieu2014a;@Brancati2016], and by the coordination device provided by the 'stolen' election [@tucker2007].
-As natural vote-share increases for a fixed level of manipulation, the likelihood of protest initiation should decline due to reduced grievances; for a fixed level of popularity, increasing manipulation should increase the likelihood of protest by raising the plausibility of a stolen election among dissatisfied citizens.
-
-We should only expect H1 to be falsified in the event that protest is driven by narrowly strategic concerns *and* election manipulation is indicative of regime strength.
-In this case, opposition actors should be deterred from launching protest movements as increasing manipulation indicates deeper regime resilience (in the form of resources, weak rule of law, loyalty of agents, and so on).
-The regime-strength model, as I call it, is defended in more detail below.
-
-## Developing the regime-strength model of electoral manipulation
-
-<!--
-Putting it all together: this theory assumes that authoritarian incumbents generally prefer order to unrest, and that protesters---who under either a narrow or broadly strategic view are aiming for some kind of concession from the incumbent---want to demonstrate the popularity of their cause through large, sustained protests.
-Electoral protests are a form of competition between the incumbent and opposition actors, in which both sides are trying to demonstrate their strength in order to win concessions from---or defeat outright---the other [@Beaulieu2014a].
-The outcome of this contest reveals the underlying strengths of the participants.
-Consequently, governments that quickly defeat protest movements can be considered stronger than those that take weeks or months to prevail, which in turn can be considered stronger than those governments which collapse in the face of mass protest.
-Incumbents with the resources and capacities needed to produce widespread election manipulation can draw on those same tools to buy off, repress, or otherwise quash post-election protest.
-This yields Hypothesis 2.
--->
 

### 2025-11-06 Output/authoritarian_only_pdf_pop_rev.Rmd
Near-final edits to lit review; removing redundant text and adjusting structure

--- 2025-11-03 Output/authoritarian_only_word_pop_rev.Rmd
@@
 output:
-  word_document:
+  pdf_document:
 linkcolor: blue
@@
 At first glance, it would appear straightforward to test these competing models.
-The regime-weakness model typically predicts that increasing manipulation (at a constant level of true popularity) increases the likelihood that opposition groups will initiate post-election protest @Luo2017;@Little2012;@Schedler2002;@Beaulieu2014a.
+The regime-weakness model typically predicts that increasing manipulation (at a constant level of true popularity) increases the likelihood that opposition groups will initiate post-election protest [@Luo2017;@Little2012;@Schedler2002;@Beaulieu2014a].
 A low total vote-share amid severe manipulation is the strongest possible signal of incumbent weakness; it has had to rely on the 'cheap talk' of manipulation to produce paltry results, indicating that the incumbent's genuine popularity is very low and its likelihood of defeating the protest movement is likewise small.
@@
 
+As sketched briefly above, a fruitful literature has examined the connections between election integrity and mass protest.
+This literature has grappled with the extent to which election manipulation reflects, and reveals, the incumbent's strength.
+Various authors have put forth different modalities of incumbent strength, with some focusing on popularity and others on coercive power.
+This paper takes a general approach, following Leventoğlu et al [-@leventogluSynthesizingTheoriesAuthoritarian2025]: a stronger incumbent is one who is better able to survive in power throughout the whole electoral cycle.
+This means the incumbent has both prevailed in the election itself, and managed to hold off any subsequent popular revolts or elite challenges [@svolik_politics_2012].
+As Leventoğlu et al [-@leventogluSynthesizingTheoriesAuthoritarian2025] discuss in their excellent analysis, genuine popularity is but one potential source of incumbent strength.
+Popularity certainly helps an incumbent win an election and survive any challenges in its aftermath, but it is only one tool; incumbents may also draw on resource wealth, organizational capacity, support among influential groups, loyal military and security services, or help from foreign patrons.
+These latter factors may be weakly correlated, if at all, with the incumbent's genuine popularity.
+
 Research in the regime-weakness model of manipulation typically concludes that as incumbents become increasingly fearful of street protest, they dial back on manipulation---a potentially pro-democracy outcome.
@@
 
-As sketched briefly above, a fruitful literature has examined the connections between election integrity and mass protest.
-This literature has grappled with the extent to which election manipulation reflects, and reveals, the incumbent's strength.
-Various authors have put forth different modalities of incumbent strength, with some focusing on popularity and others on coercive power.
-As Leventoğlu et al [-@leventogluSynthesizingTheoriesAuthoritarian2025] discuss in their excellent analysis, genuine popularity is but one potential source of incumbent strength.
-Popularity certainly helps an incumbent win an election and survive any challenges in its aftermath, but it is only one tool; incumbents may also draw on resource wealth, organizational capacity, support among elite groups, the military and security services, or help from foreign patrons.
-These latter factors may be weakly correlated, if at all, with the incumbent's genuine popularity.
-This paper takes a general approach, following Leventoğlu et al [-@leventogluSynthesizingTheoriesAuthoritarian2025]: a stronger incumbent is one who is better able to survive in power throughout the whole electoral cycle.
-This means the incumbent has both prevailed in the election itself, and managed to hold off any subsequent popular revolts or elite challenges [@svolik_politics_2012].
-
 ## The first assumption: manipulation is uninformative about strength
 
-The characteristic assumption in regime-weakness models is that electoral manipulation communicates nothing about incumbent strength, which is closely tied to popularity.
-At worst, election manipulation reveals low popularity (i.e.
-weakness).
-As a result, the opposition perceives election fraud as a reliable signal of incumbent weakness, and will stage protests in order to push out the weak incumbent.
+The characteristic assumption in regime-weakness models is that electoral manipulation communicates nothing about incumbent strength, which is closely identified with popularity.
+Since more manufactured votes indicate deeper unpopularity, the opposition perceives election fraud as a reliable signal of incumbent weakness, and will stage protests in order to push out the weak incumbent.
 For example, the model developed by Little, Tucker, and LaGatta [-@Little2015] predicts that "Just as stronger election results for the incumbent lead to less protest because they indicate that she is popular, reports of more fraud lead to more protest because they indicate that the incumbent is, for a given election result, less popular" (p.
@@
 
+This assumption, however, is contested---most prominently by Simpser [-@simpser2013]---who argues that increasing levels of election manipulation can indicate a well-resourced, well-organized, powerful incumbent (independent of its popularity); these resources and capacities, in turn, may help the incumbent weather a protest movement.
+Under this view, extensive election manipulation can be a show of strength, which induces cooperation with the incumbent from other societal actors [@simpser2013].
+
+As Simpser writes, an incumbent that demonstrates the ability to engage in widespread manipulation "is also likely to have the resources, capacities, and inclinations to overcome or punish opponents, reward supporters, and circumvent the law" [@simpser2013, p.
+6].
+A relatively popular incumbent ("strong" in the language used by the first set of models) may nonetheless manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver].
+By demonstrating otherwise difficult-to-observe capabilities, large-scale election manipulation can solve a coordination problem for the incumbent, securing the cooperation of elites [@simpser2013].
+This can, in turn, fix expectations that the incumbent will continue to rule in the future, further enhancing its staying power [@hale_patronal_2014].
+
+Prior research on this subject has typically treated protest initiation as the dependent variable.
+This approach is limited, however, since both the regime-weakness model and Simpser's alternative view can be made to make the same predictions for protest initiation--namely that increasing manipulation makes protest more likely.
+The key is in the second assumption of the regime-weakness framework, which holds that protest initiation will occur only when opponents perceive the incumbent as weak.
+A broader view of protest, well-supported in the literature, indicates that protests can emerge when manipulation is severe, even when incumbents are strong.
+
 ## A second assumption: the nature of protest
 
-The second assumption in the regime-weakness framework concerns the nature of opposition protest.
-The framework is tightly linked to a view of electoral protest that is a centralized decision by rational actors pursuing narrowly strategic goals.
-As a result, they predict that the likelihood of protest should be highest when incumbents use massive fraud but still produce a low incumbent vote-share.
-This signal of weakness, in turn, indicates to opposition actors that protest is more likely to succeed in unseating the incumbent.
+Regime-weakness models assume that opposition protest is *elite-led and instrumental*.
+The opposition is typically modeled as a unitary, strategic actor who stages protest when the expected benefits of successfully overthrowing the incumbent outweighs the expected cost of failure.
+The signal of weakness sent by election manipulation, in turn, indicates to opposition actors that protest is more likely to succeed in unseating the incumbent [@Luo2017;@Little2012;@Schedler2002;@Beaulieu2014a].
 Incumbents in these models typically face a tradeoff between the short-term benefits of electoral manipulation for winning the election, and any weakness that relying on fraud reveals to the opposition [@Rozenas2016b].
 As Beaulieu [-@Beaulieu2014a, p.
-48] writes in this tradition, "As the opposition grows stronger, the amount of [manipulation] he will tolerate shrinks, and the incumbent who wishes to avoid protest will need to reduce the amount of manipulation she commits."
-
-Most fundamentally, regime-weakness models assume that opposition protest is *elite-led and instrumental*.
-The opposition is typically modeled as a unitary, strategic actor who stages protest when the expected benefits of successfully overthrowing the incumbent outweighs the expected cost of failure.
-Such abstraction helps keep formal models tractable, but this assumption is difficult to square with recent research on electoral protest and a broader understanding of political protest in authoritarian regimes.
-
-Electoral protest may occur because manipulation galvanizes citizens who hold diverse grievances [@tucker2007], because economic conditions are poor [@Brancati2016;@Beaulieu2014a], because electoral manipulation makes people angry [@Szakonyi2021], or simply because their party lost the election [@andersonLearningLoseElection2006a].
-Opposition elites, in turn, might seek to initiate protest to take advantage of this popular discontent for strategic, rather than instrumental, reasons.
+48] writes in this tradition, "As the opposition grows stronger, the amount of [manipulation] he will tolerate shrinks, and the incumbent who wishes to avoid protest will need to reduce the amount of manipulation she commits." Such abstraction helps keep formal models tractable, but this assumption is difficult to square with recent research on electoral protest and a broader understanding of political protest in authoritarian regimes.
+
+Electoral protest may occur because manipulation galvanizes citizens who hold diverse grievances [@tucker2007], because economic conditions are poor [@Brancati2016;@Beaulieu2014a], because electoral manipulation makes people angry [@Szakonyi2021], or simply because their party lost the election [@andersonLearningLoseElection2006].
+Opposition elites, in turn, might seek to initiate---or amplify---protest to take advantage of this popular discontent for broad, rather than instrumental, reasons.
 They may wish to pressure the incumbent in order to push for improved competitiveness in the *next* election or to bargain for policy concessions [@chernykh2014;@robertson_protest_2010].
@@
 
-## Election manipulation and regime weakness
-
-<!-- The central feature of these models is that protest decisions are generally decentralized, and made with less consideration for the likelihood that protests will overthrow the incumbent.
-Belief in a stolen election---one the incumbent would have lost in the absence of manipulation---is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007;@kuntz_more_2009].
-These models thus tend to predict that protests are most likely when margins of victory are low and election manipulation high, suggesting both a high level of grievance in society and a "stolen" election.
--->
-
-## Election manipulation and incumbent strength
-
-Finally, an alternative theory of election manipulation challenges the assumption that an incumbent's ability manufacture votes communicates nothing about its ability to defeat a protest movement.
-This view, most persuasively articulated by Simpser [-@simpser2013], holds that increasing levels of election manipulation can indicate a well-resourced, well-organized, powerful incumbent (independent of its popularity); these resources and capacities, in turn, may help the incumbent weather a protest movement.
-Under this view, extensive election manipulation can be a show of strength, which induces cooperation with the incumbent from other societal actors [@simpser2013].
-Since this view holds that more manipulation can further entrench the incumbent, it is a normatively pessimistic one.
-
-As Simpser writes, an incumbent that demonstrates the ability to engage in widespread manipulation "is also likely to have the resources, capacities, and inclinations to overcome or punish opponents, reward supporters, and circumvent the law" [@simpser2013, p.
-6].
-A relatively popular incumbent ("strong" in the language used by the first set of models) may nonetheless manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver].
-By demonstrating otherwise difficult-to-observe capabilities, large-scale election manipulation can solve a coordination problem for the incumbent, securing the cooperation of elites [@simpser2013].
-This can, in turn, fix expectations that the incumbent will continue to rule in the future, further enhancing its staying power [@hale_patronal_2014].
-
-We can thus consider four possible states of the world, based on which of the foregoing assumptions are true in fact: one where both assumptions hold, one where neither holds, and two where only one assumption holds.
-The normatively optimistic view of the relationship between election manipulation and protest requires both assumptions to be true: electoral manipulation must communicate little information about incumbents' ability to outlast protest, and protest initiation must be narrowly strategic.
-If either one of these assumptions fails to hold, the deterrence logic is undermined.
-However, empirically distinguishing between these four possible states of the world is further complicated by the fact that work on this subject, whether empirical or purely formal, has largely utilized a binary measure of protest occurrence as the outcome of interest.
-Unfortunately, as Table 1 shows, three of four possible states of the world make the same prediction: that increasing manipulation will result in an increased likelihood of protest.
-
-```{=latex}
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[]
-\begin{tabular}{@{}rll@{}}
-\toprule
-Incumbent's genuine popularity:                & \multicolumn{2}{c}{Low}                                               \\ \midrule
-\multicolumn{1}{r|}{Protest is assumed to be:} & \multicolumn{1}{c|}{Instrumental} & \multicolumn{1}{c|}{Multipurpose} \\ \midrule
-\multicolumn{1}{|l|}{Manipulation...}          &                                   & \multicolumn{1}{l|}{}             \\ \midrule
-\multicolumn{1}{|r|}{...obscures strength / indicates weakness} &
-  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High and increasing\\ \\ (Luo and Rozenas 2020; \\ Chernykh and Svolik 2015; \\ Little, Tucker, and LaGatta 2015)\end{tabular}} &
-  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High and increasing\\ \\ (Tucker 2007; \\ Beaulieu 2014; \\ Brancati 2016)\end{tabular}} \\ \midrule
-\multicolumn{1}{r|}{...reflects strength} &
-  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High and decreasing\\ \\ (Simpser 2013)\end{tabular}} &
-  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High and increasing\\ \\ (Tucker 2007;\\ Beaulieu 2014;\\ Brancati 2016)\end{tabular}} \\ \bottomrule
-\end{tabular}
-\caption{Likelihood of protest initiation and expected relationship with increasing election manipulation}
-\label{tab:initiation}
-\end{table}
-```
-
-This ambiguity may help account for the inconsistent relationship between election manipulation and protest found in the literature.
+The varied motivations for protest may help account for the inconsistent relationship between election manipulation and protest initiation found in the literature.
 Several cross-national studies have found links between election manipulation and binary measures of protest [@Brancati2016;@Harvey2020;@Hyde2014].
@@
 extensive manipulation and a large incumbent vote-share) and regime tenure.
-
-Because this paper attempts to systematize and test two competing models of electoral manipulation and protest, it is worth spelling out each expectation and its mechanisms in detail.
-To do so, I will describe each theory's expectations under two worlds: one where the incumbent's genuine popularity is low, and one where it is high.
-For each of these notional worlds, expectations for protest initiation and duration are explored under both sets of assumptions about the nature of protest and electoral manipulation.
-
-First, consider the a world in which the incumbent's true popularity is low; this is presumed to be the case for most authoritarian regimes by their nature.
-Next, let us assume that the regime-weakness model is correct: election manipulation indicates weakness (or obscures information about strength).
-Finally, assume that protest is elite-led and instrumental.
-This is the world of the classic regime-weakness framework: we should expect to see protest initiation become more likely as the severity of electoral manipulation increases.
-
-The same prediction holds if we assume that protest is largely multi-purpose, while maintaining the assumptions that true popularity is low and manipulation indicates weakness.
-This is the classic focal point example [@tucker2007].
-Here, the incumbent's low genuine popularity implies high levels of grievance among the population, which are increasingly likely to be focused into mass protest as manipulation increases (i.e.
-as the election becomes more likely to appear 'stolen').
-
-Now, while maintaining the assumption that the incumbent's popularity is low, let us assume that increasing electoral manipulation indicates an otherwise stronger incumbent.
-In this case, if protest is primarily instrumental, we should expect opposition groups to shy away from protest initiation as electoral manipulation increases.
-An unpopular incumbent who manipulates extensively is stronger, in this framework, than an unpopular incumbent who does not.
-As a result, we should expect the opposite prediction of the classic regime-weakness model.
-
-However, if the incumbent's popularity is low, manipulation indicates underlying strength, and protest is largely dispersed and multi-purpose, the expectation is reversed.
-Just as in the regime-weakness case, low popularity indicates high grievances.
-Increasing electoral manipulation helps focus those grievances and drives protest initiation.
-
-The key takeaway, then, is that for most authoritarian regimes (where genuine incumbent popularity is likely low), an examination of protest initiation alone cannot distinguish between three out of four possible data-generating processes.
-A pattern in which higher levels of manipulation are associated with increasing risk of protest initiation can be explained either by the regime-weakness model, or by a world in which manipulation indicates incumbent staying-power but protest is non-instrumental.
 
@@
 Regime-weakness models of electoral manipulation treat election manipulation and genuine votes as substitute goods.
-For any given vote-share, an increase in manufactured votes reduces the number of genuine votes; that is, the higher the number of manufactured votes for any given vote total, the less likely the incumbent is to survive a challenge.
+For any given vote-share, an increase in manufactured votes reduces the number of genuine votes; that is, the higher the number of manufactured votes for any given vote total, the lower the incumbent's genuine popularity is and the less likely they are to survive a challenge.
 In an instrumental mode of protest, strategic opposition actors perceive the incumbent's relative lack of popularity, and count on the majority of the public to back their challenge.
@@
 
-Need to discuss vote-share next:
-
-> H1: Increasing election fraud makes protest more likely, with the size of this effect decreasing with incumbent-party vote-share.
-
-> H2: Election manipulation is positively associated with protest movement duration, with the size of this effect decreasing with incumbent vote-share.
+## Empirical implications: vote-share vs. genuine popularity
+
+Both models of election manipulation consider genuine popularity to be a source of regime strength, and assume that increasing electoral manipulation either distorts or weakens this indicator by mingling genuine votes with manufactured ones.
+The distinction between the two models is over whether or not manipulation simultaneously indicates additional sources of strength other than popularity.
+
+Of course, in most authoritarian cases, the incumbent's true popularity is unknown due to censorship, repression of opposition candidates, and preference falsification by citizens.
+As a result, scholars (and political actors in real time) are left to use the incumbent's vote-share and perceived level of manipulation as imperfect indicators of the incumbent's true popularity.
+Since genuine popularity is indicated by the mix of incumbent vote-share and level of manipulation, both indicators must be included in our hypotheses.
+Table \@ref(tab:durations) summarizes how both models of electoral manipulation characterize incumbent strength under varying conditions, each of which is laid out in more detail below.
+
+```{=latex}
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+% \usepackage{multirow}
+\begin{table}[]
+\begin{tabular}{@{}lllll@{}}
+ &  & \multicolumn{2}{c}{Manipulation} & Duration prediction of defeated protests \\
+ &  & Low            & High            &                                          \\
+\multicolumn{1}{c}{\multirow{2}{*}{Vote-share}} &
+  High &
+  \begin{tabular}[c]{@{}l@{}}Popularity: High\\ MS: Mixed\\ \\ \\ Predicted strength (RW): Strong\\ Predicted strength (RS): Unclear\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Popularity: Low (RW) or mixed (RS)\\ MS: High\\ \\ \\ Predicted strength (RW): Weak\\ Predicted strength (RS): Moderate\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}RW: Duration increases with manipulation\\ RS: No significant effect with manipulation\end{tabular} \\
+\multicolumn{1}{c}{} &
+  Low &
+  \begin{tabular}[c]{@{}l@{}}Popularity: Low\\ MS: Low\\ \\ \\ Predicted strength (RW): Weak\\ Predicted strength (RS): Weak\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Popularity: Very low\\ MS: High\\ \\ \\ Predicted strength (RW): Very weak\\ Predicted strength (RS): Moderate\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}RW: Duration increases with manipulation\\ RS: Duration decreases with manipulation\end{tabular}
+\end{tabular}
+\caption{Predicted incumbent strength, and associated hypotheses for protest duration, under regime-weakness (RW) and regime-strength (RS) models of manipulation as incumbent vote-share and level of electoral manipulation vary.
+'MS' indicates incumbent's underlying strength associated with manipulation under regime-strength model (i.e.
+'manipulation strength').
+Duration predictions for protests which the government ultimately defeats.}
+\label{tab:durations}
+\end{table}
+```
+
+The upper left cell in Table \@ref(tab:durations) illustrates a scenario where the incumbent's vote-share is high and manipulation is low.
+This combination indicates a relatively popular incumbent, meaning it should be characterized as 'strong' under the regime-weakness model of manipulation.
+However, cases in this cell are less informative about 'manipulation strength,' those capacities and resources that electoral manipulation draws on.
+This cell could include regimes that lack the capacity to manipulation, *or* regimes which have such capacities but choose not to employ them in a particular election.
+As a result, from the regime-strength perspective, this cell would likely include a mix of weaker and stronger regimes.
+There is thus no clear expected tendency for this group under that model.
+
+Moving to the right in Table \@ref(tab:durations), incumbent vote-share remains high but manipulation is also elevated.
+For the regime-weakness model, cases in this cell are clearly expected to be weaker than those in the top left; one's high vote-share mostly reflects genuine popularity, while the other's is largely manufactured.
+In the regime-weakness model, regimes should unambiguously become weaker as we move to the right in this row.
+As a result, holding vote-share contant at a high level, increasing manipulation should lead to longer-lasting protests under the regime-weakness model.
+
+A different prediction emerges from the regime-strength framework.
+Here, too, increasing manipulation may indicate broadly lower popularity (holding vote-share constant).
+But it may also indicate a popular incumbent relying on manipulation tools to boost turnout in a low-competition election [@Smyth2018;@martinezicomaElectionTurnoutAuthoritarian2020].
+In any event, a decrease in genuine popularity is expected to be offset by increased strength indicated by the ability to manipulate the election severely.
+As a result, overall strength is expected to be moderate.
+Where the regime-weakness model predicts a sharp decline in incumbent strength moving from the top left to the top right of Table \@ref(tab:durations), the regime-strength model suggests there should be no consistent effect on protest duration as manipulation increases, since decreases in popularity are offset by increases in other sources of power.
+
+Moving to the lower left, the incumbent's vote-share is now assumed to be low, as is the level of electoral manipulation.
+Both models of manipulation agree here: such incumbents are weak.
+They lack both genuine popularity, and the ability to produce significant manipulation.
+Under the regime-strength framework, such an incumbent would have had strong incentives to produce more manipulation if he could have; the low level of manipulation is thus an indicator of low capacity.
+
+Finally, moving to the lower right, the level of manipulation increases while vote-share remains low.
+This is the weakest of all incumbents in the regime-weakness framework.
+Holding vote-share constant at a low value, increasing manipulation indicates the incumbent's genuine popularity---and thus strength---is exceedingly low.
+For the regime-strength model, cases in this cell are expected to be stronger than those to the left, as their increasing capacity to manipulate elections indicates sources of strength beyond their (declining) popularity.
+
+Moving left-to-right along Table \@ref(tab:durations) shows two clear hypotheses; one direct hypothesis under the regime-weakness model and one interaction effect under the regime-strength model.
+First, the regime-weakness model should predict that protest duration will increase with manipulation when incumbent at any vote-share (as 'cheap-talk' manufactured votes displace genuine popularity in the vote-share).
+The first hypothesis is straightforward.
+
+>H1 (Regime-weakness): Increases in electoral manipulation will be associated with increasing duration of unsuccessful electoral protest.
+
+Second, the regime-strength model predicts that protest duration will *decrease* as electoral manipulation increases, but only when the incumbent's vote-share is relatively low.
+At higher vote-shares, the model predicts ambiguous relationships between vote-share, manipulation, and overall regime strength, suggesting that statistical correlations should weaken.
+
+>H2 (Regime-weakness): Increases in electoral manipulation will be associated with decreasing duration of unsuccessful electoral protest when incumbent vote-share is low.
 
@@
 Testing H1 and H2 together will help to evaluate the regime-weakness model of electoral manipulation.
-If H1 is supported and H2 is not, it would provide evidence in favor of the regime-weakness model; manipulation would appear associated with incumbent weakness, and protest initiation would appear to take advantage of such.
-On the other hand, if both hypotheses are supported, it would endorse the argument that electoral manipulation can be associated with sources of incumbent strength in a world where the aims of electoral protest are broadly (rather than narrowly) strategic.
+It is important to note that the theories make contradictory predictions.
+H1 predicts an unconditional *positive* relationship between manipulation and protest duration, while H2 predicts a *negative* relationship, conditional on vote-share.
+
+[Note that the type of protest that happens (instrumental vs. multipurpose) will be examined in the discussion / appendix. Consider also coming up with names for 'manipulative strength']
 
@@
 summary(cox2)
-p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.33, 2.2]")) #This is the 10th and 90th quantiles
+p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.29, 2.2]")) #This is the 10th and 90th quantiles
 p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()

### 2025-11-06 Output/authoritarian_only_pdf_pop_rev.Rmd
Cuts and re-ordering in theory section

@@
 
-Theories in the 'regime-weakness' view of electoral manipulation are built on several assumptions about the nature of protest and the information conveyed by the incumbent's vote-share and manipulation efforts.
-Updating some of these assumptions, I argue, provides a model that more accurately reflects contemporary research on manipulation *and* makes distinct empirical predictions.
-Each of these assumptions, and how they are adjusted in the 'regime-strength' model of manipulation, is discussed below.
+Considering how electoral manipulation may be tied to protest duration under the regime-weakness and regime-strength models reveals distinct predictions.
+Specifically, as discussed below, protest duration should decrease with manipulation in the regime-weakness model, but increase in the regime-strength model (conditional on incumbent vote-share).
+The logic discussed below hinges on the assumption that, on average, authoritarian governments will seek to quash protest movements quickly once they appear.
+How plausible is this assumption?
 
@@
 
-The central argument of this manuscript is that, in authoritarian regimes, incumbents who develop the tools necessary to manipulate elections extensively will also have the tools necessary to quickly prevail over protest movements.
+## The regime-strength theory of manipulation and protest duration
+
+The central argument of this manuscript is that, in authoritarian regimes, incumbents who develop the tools necessary to manipulate elections extensively will also have the tools necessary to more quickly prevail over protest movements.
 This proposition builds off Simpser's foundational argument that manipulated elections can send costly signals of the incumbent's "capacities, resources, [and] dispositions" that will help it retain power in the future [@simpser2013, p.
@@
 First, by incorporating more recent research on the principal-agent dynamics inherent in election manipulation, I argue that incumbents who can produce large amounts of election manipulation necessarily have "capacities, resources and dispositions" helpful for defeating protests.
-This research trend emphasizes the degree to which political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].
 Second, mine is not a signalling argument; good manipulators are likely to have deeper wells of strength than weaker manipulators, whether those resources are successfully signaled to opposition actors or no.
 
+Recent research emphasizes the degree to which political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].
 Electoral manipulation has been shown to be more extensive under conditions that minimize principal-agent problems between low-level agents and the incumbent.
@@
 
-```{=latex}
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[]
-\begin{tabular}{@{}rll@{}}
-\toprule
-Incumbent's genuine popularity:       & \multicolumn{2}{c}{Low}  \\ \midrule
-\multicolumn{1}{r|}{Protest is assumed to be:}                  & \multicolumn{1}{c|}{Instrumental} & \multicolumn{1}{c|}{Multipurpose} \\ \midrule
-\multicolumn{1}{|l|}{Manipulation...} &  & \multicolumn{1}{l|}{} \\ \midrule
-\multicolumn{1}{|r|}{...obscures strength / indicates weakness} & \multicolumn{1}{l|}{Increasing}   & \multicolumn{1}{l|}{Increasing}   \\ \midrule
-\multicolumn{1}{r|}{...reflects strength}                       & \multicolumn{1}{l|}{Decreasing}   & \multicolumn{1}{l|}{Decreasing}   \\ \bottomrule
-\end{tabular}
-\caption{Expected relationship between protest duration and increasing election manipulation}
-\label{tab:duration}
-\end{table}
-```
-
-How, then, would this view of electoral manipulation map onto the instrumental vs.
-multipurpose views of protest discussed above?
-While regime-weakness models have exclusively (to my knowledge) been framed around protest initiation, it is straightforward to extend their predictions to protest duration.
-Regime-weakness models of electoral manipulation treat election manipulation and genuine votes as substitute goods.
-For any given vote-share, an increase in manufactured votes reduces the number of genuine votes; that is, the higher the number of manufactured votes for any given vote total, the lower the incumbent's genuine popularity is and the less likely they are to survive a challenge.
-In an instrumental mode of protest, strategic opposition actors perceive the incumbent's relative lack of popularity, and count on the majority of the public to back their challenge.
-Increasing manipulation indicates deeper unpopularity (holding vote-share constant), so we should expect a less popular incumbent to fail more quickly.
-The same logic holds, with slightly different mechanisms, if protests are assumed to be multipurpose.
-Here, greater manipulation is used to overcome deeper unpopularity (i.e.
-widespread grievances); more manipulation indicates an incumbent facing an increasingly powerful challenger.
-Thus, the regime-weakness model of manipulation should predict that the duration of unsuccessful protests will be increasing in election manipulation: more manipulation indicates a stronger challenger, which will take longer to defeat on average.
-
-For the regime-strength model of manipulation, the opposite relationship is expected, as laid out above.
-Comparing Table \@ref(tab:duration) and \@ref(tab:initiation) shows that the hypotheses generated by the preferred model--one in which manipulation indicates strength and protest is multipurpose are distinct from those in other quadrants.
-
 ## Empirical implications: vote-share vs. genuine popularity
 
+Still, a full accounting of a regime's strength would include the incumbent's genuine popularity [@Gerschewski2013].
 Both models of election manipulation consider genuine popularity to be a source of regime strength, and assume that increasing electoral manipulation either distorts or weakens this indicator by mingling genuine votes with manufactured ones.

### 2025-11-07 Output/authoritarian_only_pdf_pop_rev.Rmd
Final major edits to lit review and theory section

@@
 However, there are two reasons this is unlikely to be the case systematically for the circumstances studied here.
-First, authoritarian governments are more likely to tolerate protests aimed at low-salience, local issues [@lorentzenRegularizingRiotingPermitting2013], and more likely to clamp down on protests that have higher concession costs [@kleinDynamicsPoliticalProtests2018].
+First, authoritarian governments are more likely to tolerate protests aimed at low-salience, local issues [@lorentzenRegularizingRiotingPermitting2013;@kleinDynamicsPoliticalProtests2018], and more likely to clamp down on protests that have higher concession costs [@kleinDynamicsPoliticalProtests2018].
 As electoral protests challenge the incumbent's ostensible victory and right to remain in office, they are directly threatening to the regime; the concession costs of stepping down or holding new elections are very high.
@@
 "Ignoring" protests can lead to further mobilization against authoritarian regimes [@Bishara2015], as can concessions in the absence of durable political reform [@leuschnerStateConcessionsProtest2024].
-We should thus expect incumbents to work to defuse electoral protests as quickly as possible.
+In such cases, we should expect to see apparent 'toleration' displaced by 'attrition'---efforts by the incumbent to increase the cost of protest while avoiding open repression [@yuenNeitherRepressionConcession2017].
+
+In other words, even in the absence of explicit repression or concessions, we should still expect incumbents to attempt to defuse protest movements aimed at challenging the most recent election.
+As a further control, however, I include a measure of average protest size in one model of protest duration.
 
@@
 
-This argument---that successful, large-scale election manipulation draws on the same resources a state may use to suppress protest---turns the deterrent logic on its head.
+As a short-hand, I refer to this as "control-based" strength---those sources of regime durability that derive from its ability to mobilize and discipline agents, to distribute patronage, and to deploy coercion.
+This contrasts with the legitimacy-based view of strength centered in the regime-weakness framework.
+This argument---that successful, large-scale election manipulation draws on some of the same resources a state may use to suppress protest---turns the deterrent logic on its head.
 Instead of signalling weakness, large-scale fraud indicates a well of resilience: control over resources, impunity for pro-regime lawbreaking, and loyal agents.
@@
 Both models of election manipulation consider genuine popularity to be a source of regime strength, and assume that increasing electoral manipulation either distorts or weakens this indicator by mingling genuine votes with manufactured ones.
-The distinction between the two models is over whether or not manipulation simultaneously indicates additional sources of strength other than popularity.
+The distinction between the two models is over whether or not manipulation demonstrates "control-based" strength.
 
@@
 Since genuine popularity is indicated by the mix of incumbent vote-share and level of manipulation, both indicators must be included in our hypotheses.
-Table \@ref(tab:durations) summarizes how both models of electoral manipulation characterize incumbent strength under varying conditions, each of which is laid out in more detail below.
+Table \@ref(tab:durations) summarizes how both models of electoral manipulation characterize incumbent strength under varying conditions, along with predictions for the duration of unsuccessful protest movements under each model.^[In the appendix, I check the separate prediction that manipulation should be positively associated with incumbents being ousted by protest movements in the regime-weakness framework.]
 
@@
   High &
-  \begin{tabular}[c]{@{}l@{}}Popularity: High\\ MS: Mixed\\ \\ \\ Predicted strength (RW): Strong\\ Predicted strength (RS): Unclear\end{tabular} &
-  \begin{tabular}[c]{@{}l@{}}Popularity: Low (RW) or mixed (RS)\\ MS: High\\ \\ \\ Predicted strength (RW): Weak\\ Predicted strength (RS): Moderate\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Popularity: High\\ CS: Mixed\\ \\ \\ Predicted strength (RW): Strong\\ Predicted strength (RS): Unclear\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Popularity: Low (RW) or mixed (RS)\\ CS: High\\ \\ \\ Predicted strength (RW): Weak\\ Predicted strength (RS): Moderate\end{tabular} &
   \begin{tabular}[c]{@{}l@{}}RW: Duration increases with manipulation\\ RS: No significant effect with manipulation\end{tabular} \\
@@
   Low &
-  \begin{tabular}[c]{@{}l@{}}Popularity: Low\\ MS: Low\\ \\ \\ Predicted strength (RW): Weak\\ Predicted strength (RS): Weak\end{tabular} &
-  \begin{tabular}[c]{@{}l@{}}Popularity: Very low\\ MS: High\\ \\ \\ Predicted strength (RW): Very weak\\ Predicted strength (RS): Moderate\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Popularity: Low\\ CS: Low\\ \\ \\ Predicted strength (RW): Weak\\ Predicted strength (RS): Weak\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Popularity: Very low\\ CS: High\\ \\ \\ Predicted strength (RW): Very weak\\ Predicted strength (RS): Moderate\end{tabular} &
   \begin{tabular}[c]{@{}l@{}}RW: Duration increases with manipulation\\ RS: Duration decreases with manipulation\end{tabular}
@@
 This combination indicates a relatively popular incumbent, meaning it should be characterized as 'strong' under the regime-weakness model of manipulation.
-However, cases in this cell are less informative about 'manipulation strength,' those capacities and resources that electoral manipulation draws on.
+However, cases in this cell are less informative about 'control-based' strength.
 This cell could include regimes that lack the capacity to manipulation, *or* regimes which have such capacities but choose not to employ them in a particular election.
@@
 But it may also indicate a popular incumbent relying on manipulation tools to boost turnout in a low-competition election [@Smyth2018;@martinezicomaElectionTurnoutAuthoritarian2020].
-In any event, a decrease in genuine popularity is expected to be offset by increased strength indicated by the ability to manipulate the election severely.
+In any event, a decrease in genuine popularity is expected to be offset by increased 'control-based' strength, demonstrated by the ability to manipulate the election severely.
 As a result, overall strength is expected to be moderate.
-Where the regime-weakness model predicts a sharp decline in incumbent strength moving from the top left to the top right of Table \@ref(tab:durations), the regime-strength model suggests there should be no consistent effect on protest duration as manipulation increases, since decreases in popularity are offset by increases in other sources of power.
-
-Moving to the lower left, the incumbent's vote-share is now assumed to be low, as is the level of electoral manipulation.
+Where the regime-weakness model predicts a sharp decline in incumbent strength moving from the top left to the top right of Table \@ref(tab:durations), the regime-strength model suggests there should be no consistent effect on protest duration as manipulation increases, since both cells in the upper row contain a mix of weaker and stronger regimes.
+
+In the lower left cell, the incumbent's vote-share is now assumed to be low, as is the level of electoral manipulation.
 Both models of manipulation agree here: such incumbents are weak.
 They lack both genuine popularity, and the ability to produce significant manipulation.
-Under the regime-strength framework, such an incumbent would have had strong incentives to produce more manipulation if he could have; the low level of manipulation is thus an indicator of low capacity.
+Under the regime-strength framework, such an incumbent would have had strong incentives to produce more manipulation if he could have; the low level of manipulation is thus an indicator of low control-based strength.
 
@@
 Holding vote-share constant at a low value, increasing manipulation indicates the incumbent's genuine popularity---and thus strength---is exceedingly low.
-For the regime-strength model, cases in this cell are expected to be stronger than those to the left, as their increasing capacity to manipulate elections indicates sources of strength beyond their (declining) popularity.
+For the regime-strength model, cases in this cell are expected to be stronger than those to the left, as their increasing capacity to manipulate elections demonstrates sources of strength beyond their (low) popularity.
 
 Moving left-to-right along Table \@ref(tab:durations) shows two clear hypotheses; one direct hypothesis under the regime-weakness model and one interaction effect under the regime-strength model.
-First, the regime-weakness model should predict that protest duration will increase with manipulation when incumbent at any vote-share (as 'cheap-talk' manufactured votes displace genuine popularity in the vote-share).
+First, the regime-weakness model should predict that protest movements will survive longer against governments that manipulate more, regardless of vote-share (as 'cheap-talk' manufactured votes displace genuine popularity).
 The first hypothesis is straightforward.
 
->H1 (Regime-weakness): Increases in electoral manipulation will be associated with increasing duration of unsuccessful electoral protest.
+>**H1 (Regime-weakness)**: Increases in electoral manipulation will be associated with increasing duration of unsuccessful electoral protest.
 
@@
 
->H2 (Regime-weakness): Increases in electoral manipulation will be associated with decreasing duration of unsuccessful electoral protest when incumbent vote-share is low.
+>**H2 (Regime-weakness)**: Increases in electoral manipulation will be associated with decreasing duration of unsuccessful electoral protest when incumbent vote-share is low.
 
@@
 ```
-
-The argument can be summarized as follows.
-Incumbents will generally seek to manipulate elections as much as they can, to capitalize on both the direct and indirect benefits of electoral manipulation [@simpser2013].
-To do so, they must solve principal-agent problems by consolidating control over patronage resources and shielding agents from negative consequences of their actions.
-The same resources that can be used to manipulate elections on a wide scale---including natural popularity, resource control, and control over the legal apparatus---also better enable incumbents to dismantle protest movements when they do emerge.
-As a result, more extensive manipulation should be associated with shorter-lived protest movements against incumbents.^[In the appendix, I check the separate prediction that manipulation should be positively associated with incumbents being ousted by protest movements.]
-
-One challenge to this theory is the prospect that incumbents may choose to tolerate protests, rather than attempting to repress them or co-opt their participants.
-This concern should be modest for the cases being considered here.
-First, toleration is most likely when protests impose few costs on the regime and make only targeted demands [@kleinDynamicsPoliticalProtests2018].
-National elections are highly politically salient, and challenging their validity gets to the core of electoral authoritarian incumbent's claim to authority.
-As such, tolerating or ignoring protest is unlikely in this setting.
-In such cases, we should expect to see apparent 'toleration' displaced by 'attrition'---efforts by the incumbent to increase the cost of protest while avoiding open repression [@yuenNeitherRepressionConcession2017].
-In other words, even in the absence of explicit repression or concessions, we should still expect incumbents to attempt to defuse protest movements aimed at challenging the most recent election.
-As a further control, however, I include a measure of average protest size in one model of protest duration.
 

### 2025-11-12 Output/authoritarian_only_pdf_pop_rev.Rmd
Reorder and edits to control variable section

@@
 
-For models of protest-wave duration, as discussed in more detail below, I include only observations where a positive number of protests occurred.
-Only 156 observations fall into this more limited dataset.
+For models of protest-wave duration, as discussed in more detail below, I include only observations where a positive number of protests occurred; there are 211 such cases.
 
@@
 
+```{r binarytab, echo=FALSE, message=FALSE, warning=FALSE}
+count_table <- table(cut(nelda2.sub$winner.share.upd, breaks = c(0, .45, .55, 1), 
+    include.lowest = TRUE, labels = c("Low", "Medium", "High")),
+      cut(nelda2.sub$v2elirreg.inv, breaks =quantile(nelda2.sub$v2elirreg.inv, probs = c(0, 1/3, 2/3, 1), na.rm=T), 
+    include.lowest = TRUE, labels = c("Low", "Medium", "High")),
+      nelda2.sub$any.ecav.protest)
+
+prop_count_table <- nelda2.sub %>% filter(inc.lose == 0) %>%
+  mutate(winnershare_bin = cut(winner.share.upd,  breaks = c(0, .45, .55, 1), 
+    include.lowest = TRUE, labels = c("Low (0-45%)", "Medium (45-55%)", "High (55-100%)")),
+         fraud_bin = cut(v2elirreg.inv,  breaks =quantile(nelda2.sub$v2elirreg.inv, probs = c(0, 1/3, 2/3, 1), na.rm=T), 
+    include.lowest = TRUE, labels = c("Low", "Medium", "High"))) %>%
+  group_by(winnershare_bin, fraud_bin) %>%
+  summarise(pct_positive = mean(any.ecav.protest) * 100)
+
+prop_count_table <- pivot_wider(prop_count_table, names_from = fraud_bin, values_from = pct_positive)
+prop_count_table <- prop_count_table[1:3, 1:4]
+prop_count_table <- round(prop_count_table[,2:4], 2)
+prop_count_table <- cbind('Incumbent vote-share' = c("Low", "Medium", "High"), prop_count_table)
+
+prop_count_table %>%
+  kable(format = "latex", booktabs = TRUE, caption = "Percentage of cases in which post-electoral protests occurred") %>%
+  add_header_above(c(" " = 1, "Electoral fraud" = 3)) %>%
+  kable_styling(latex_options = c("striped", "hold_position"))
+```
+
+As a plausibility check, I first examine patterns of protest initiation.
+The regime-weakness model expects initiation to increase with fraud, holding vote-share constant; while the regime-strength model expects a less consistent pattern.
+Table \@ref(tab:binarytab) categorizes elections into 3 x 3 bins based on incumbent vote-share and degree of electoral fraud, and reports the percentage of elections in each category that were followed by protest.
+For this table, vote-share is considered low if it less than 45%, medium between 45% and 55%, and high if greater than 55%.
+The middle category is intended to capture a 'close' election.
+Election fraud is divided into terciles.
+These categories can be thought of as relatively clean (*election fraud* is between -2.33 and 0.5), moderately manipulated (0.5 to 1.46) and heavily manipulated (1.46 to 3.2).
+Two patterns appear in the data.
+First, protest initiation is most likely when incumbent vote-share is low and fraud is high.
+This is expected under both models.
+For the regime-weakness model, this is taken as a sign that opposition actors launch protests instrumentally against the weakest (i.e.
+least popular) incumbents.
+For the regime-strength model, this pattern could emerge as electoral manipulation focuses mass grievances.
+However, we do not observe a consistent increase in protest likelihood as manipulation increases, as the regime-weakness model would predict.
+Instead, we see relatively consistent rates of protest initiation at medium vote-share across levels of manipulation.
+And, at high incumbent vote-shares, there is even a decrease in protest initiation as manipulation increases from moderate to high.
+This comports with the mixed evidence found in prior research on protest initiation and manipulation, and is evidence that the regime-strength model is on a plausible foundation (specifically, that protests can occur even when the incumbent is predicted to be 'strong' under either model).
+Consider the cell where vote-share is high and election fraud is low---this is the strongest (i.e.
+clearly popular) type of regime in the regime-weakness model.
+Nevertheless, more than a quarter of such cases are protested.
+This is not so different from the cell which is most likely to contain strong regimes in the regime-strength model (high vote-share, high manipulation), in which thirty-two percent of cases launch protest.
+While not dispositive, this result indicates that the regime-strength model and regime-weakness models are equally compatible with patterns of protest initiation, as expected.
+
 ```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
@@
 
- ggplot(nelda2.sub, aes(x = vs_fraud_ratio, y= jitter(any.ecav.protest))) + geom_point(aes(color = v2elirreg.inv)) + xlab("Ratio") + theme_bw() + scale_color_viridis_b()
-
-  ggplot(nelda2.sub, aes(x = v2elirreg.inv, y= jitter(any.ecav.protest))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Fraud") + theme_bw()
-
-   ggplot(nelda2.sub, aes(x = v2elirreg.inv, y= n.events.post)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Fraud") + theme_bw()
-
-  ggplot(nelda2.sub, aes(x = winner.share.upd, y= jitter(any.ecav.protest))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Winner share") + theme_bw()
-
-    ggplot(nelda2.sub, aes(x = winner.share.upd, y= n.events.post)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Winner share") + theme_bw()
-
-   ggplot(nelda2.sub, aes(x = vs_fraud_ratio, y= d.events.num)) + geom_point(aes(color = v2elirreg.inv)) + xlab("Election fraud") + theme_bw() + scale_color_viridis_b()
+# ggplot(nelda2.sub, aes(x = vs_fraud_ratio, y= jitter(any.ecav.protest))) + geom_point(aes(color = v2elirreg.inv)) + xlab("Ratio") + theme_bw() + scale_color_viridis_b()
+
+#  ggplot(nelda2.sub, aes(x = v2elirreg.inv, y= jitter(any.ecav.protest))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Fraud") + theme_bw()
+
+#   ggplot(nelda2.sub, aes(x = v2elirreg.inv, y= n.events.post)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Fraud") + theme_bw()
+
+#  ggplot(nelda2.sub, aes(x = winner.share.upd, y= jitter(any.ecav.protest))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Winner share") + theme_bw()
+
+#    ggplot(nelda2.sub, aes(x = winner.share.upd, y= n.events.post)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Winner share") + theme_bw()
+
+#   ggplot(nelda2.sub, aes(x = vs_fraud_ratio, y= d.events.num)) + geom_point(aes(color = v2elirreg.inv)) + xlab("Election fraud") + theme_bw() + scale_color_viridis_b()
 
@@
 
- ggplot(nelda2.sub %>% filter(inc.lose != 1) , aes(x=winner.share.upd, y=n.events.post, color = v2elirreg.inv)) +
-    geom_point() +
-    scale_color_continuous_divergingx(palette = "ArmyRose", n_interp = 11, cmax1 = 500, cmax2 = 500) +
-    theme_bw() +
-   coord_cartesian(ylim = c(0, 100))
-
- ggplot(nelda2.sub %>% filter(inc.lose != 1), aes(x=winner.share.upd, y=jitter(any.ecav.protest), color = v2elirreg.inv)) +
-    geom_point() +
-    scale_color_continuous_divergingx(palette = "ArmyRose", n_interp = 11, cmax1 = 500, cmax2 = 500) +
-    theme_bw()
-
-  ggplot(nelda2.sub %>% filter(inc.lose != 1 & d.events.num > 0), aes(x=winner.share.upd, y=d.events.num, color = v2elirreg.inv)) +
-    geom_point(aes(shape = as.factor(protest.success))) +
-    scale_color_continuous_divergingx(palette = "ArmyRose", n_interp = 11, cmax1 = 500, cmax2 = 500) +
-    theme_bw()
+# ggplot(nelda2.sub %>% filter(inc.lose != 1) , aes(x=winner.share.upd, y=n.events.post, color = v2elirreg.inv)) +
+#    geom_point() +
+#    scale_color_continuous_divergingx(palette = "ArmyRose", n_interp = 11, cmax1 = 500, cmax2 = 500) +
+#    theme_bw() +
+#   coord_cartesian(ylim = c(0, 100))
+
+# ggplot(nelda2.sub %>% filter(inc.lose != 1), aes(x=winner.share.upd, y=jitter(any.ecav.protest), color = v2elirreg.inv)) +
+#    geom_point() +
+#    scale_color_continuous_divergingx(palette = "ArmyRose", n_interp = 11, cmax1 = 500, cmax2 = 500) +
+#    theme_bw()
+
+#  ggplot(nelda2.sub %>% filter(inc.lose != 1 & d.events.num > 0), aes(x=winner.share.upd, y=d.events.num, color = v2elirreg.inv)) +
+#    geom_point(aes(shape = as.factor(protest.success))) +
+#    scale_color_continuous_divergingx(palette = "ArmyRose", n_interp = 11, cmax1 = 500, cmax2 = 500) +
+#    theme_bw()
 
@@
   # showing data points on the same color scale 
-levelplot(any.ecav.protest ~ v2elirreg.inv * winner.share.upd, nelda2.sub %>% filter( protest.success == 0 & inc.lose == 0), 
+levelplot(any.ecav.protest ~ winner.share.upd * v2elirreg.inv, nelda2.sub %>% filter( protest.success == 0 & inc.lose == 0), 
           panel = panel.levelplot.points, cex = 1.2
@@
   # showing data points on the same color scale 
-levelplot(d.events.num ~ v2elirreg.inv * winner.share.upd, nelda2.sub %>% filter(d.events.num > 0 & protest.success == 0 & inc.lose == 0), 
+levelplot(d.events.num ~ winner.share.upd * v2elirreg.inv, nelda2.sub %>% filter(d.events.num > 0 & protest.success == 0 & inc.lose == 0), 
           panel = panel.levelplot.points, cex = 1.2
@@
 
-  testlogit <- glm(protest.success ~ v2elirreg.inv*winner.share.upd + presidential + inc.termlimit.num + n.events.pre, data = nelda2.sub)
-summary(testlogit)
-sjPlot::plot_model(testlogit, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.08, 1.96]"))
+#  testlogit <- glm(protest.success ~ v2elirreg.inv*winner.share.upd + presidential + inc.termlimit.num + n.events.pre, data = nelda2.sub)
+#summary(testlogit)
+#sjPlot::plot_model(testlogit, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.08, 1.96]"))
 ```
@@
 
-It is necessary to condition the models on whether or not the incumbent party claimed victory in the election, due to the structure of the vote-share data in V-Dem.
-Vote-share data is provided for the winning party without reference to its position as incumbent or challenger.
-From the point of view of those considering anti-incumbent protest, for instance, the information conveyed by a large opposition vote-share is much different from that conveyed by a large winning vote-share.
-Consequently, in all models, I condition for whether the incumbent party lost, based on coding in NELDA.
-In the binary models of Hypothesis 1, I control for incumbent loss as a standalone variable or in interaction with the main explanatory variables.
-
-For the models of protest duration, I exclude all cases where the incumbent lost the election outright.
-This limits the observations to those that fit the population of interest: protest following an election in which the incumbent claimed victory, possibly through reliance on election fraud.^[Excluding cases where incumbents stepped down after losing an election avoids introducing measurement error in the dependent variable, due to a quirk of the ECAV dataset, which codes the 'side' of the protest (pro- or anti-government) as missing during the period between an election and the swearing in of a new government in cases where the incumbent loses.
-Including these cases thus runs the risk of conflating pro- and anti-incumbent protest, along with cases where protesters goals are genuinely unknown.
-However, excluding them also influences the interpretation of the results from these models.
-Threats to inference from this section are evaluated in more detail in the discussion section.]
-
-The models incorporate several potential confounding variables, in addition to the *election fraud* variable described above.
-The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance.
-The winning party's *vote-share* is also taken from the V-Dem dataset.^[This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia, where links to official sources were available, and Psephos.]
-
-As identified in prior research, other factors that may influence both fraud and protest are included as controls variables.
-The repressive capacity of the government [@Davenport2019] is measured using the physical integrity index from V-Dem (v2x_clphy), and its square term.
-Indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), de facto *judicial independence* [@Linzer2015] and the available of *alternative sources of information* ('v2x_freexp_altinf' from V-Dem) are also employed.
+The theory suggests that increasing election manipulation is indicative of regime strength, but an actor's strength---like its close analogue, power [@dahl_concept_1957]---can only be measured relative to that of another.
+An incumbent who produces a high level of election manipulation may be 'strong' when facing a fragmented and unpopular opposition, but would be less so if the opposition were more formidable.
+As a result, it is necessary to control for potential indicators of opposition strength.
+
+First, I include the *number of pre-election protests* counted in the ECAV dataset, which is likely to be an important possible confounder; more numerous pre-election protest may affect the level of election fraud (as the deterrence logic predicts) as well as being directly tied to the likelihood of post-election protest.
+Next, the models of protest duration include a measure of the *median size of protest*, calculated using ECAV data.^[ECAV data codes protest size as a categorical variable, ranging from 1 (< 10 participants) to 5 (> 10,000).
+I take the median size category of all the events in the protest wave as a measure of average protest size.] This is important as a control for opposition strength, as well as for the possibility that incumbents might tolerate smaller protests.
+Since non-democratic governments may be able to adjust the electoral calendar strategically, to disrupt opposition coordination [@nygardTimingMattersImpact2020], I include a variable from NELDA that indicates *off-schedule elections*.
+
+Since higher repressiveness may weaken or intensify opposition [@chenowethStateRepressionNonviolent2017], I include a lagged measure of physical integrity index from V-Dem (v2x_clphy), and its square term.
+Indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), *legislative constraints* on the executive (V-Dem variable 'v2xlg_legcon', and the availability of *alternative sources of information* ('v2x_freexp_altinf' from V-Dem) are also employed as straightforward indicators of opposition strength.
 Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019].
-All models include *legislative constraints* on the executive, a V-Dem variable which captures the extent to which the legislative body in the country acts independently in practice to limit executive power.
-
-In addition, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby [@Bunce2010;@Beissinger_2007].
-Such a dynamic could plausibly increase the likelihood of protest in a nearby country, while also reducing the incumbent's use of fraud.
+
+Other controls are included for factors that might separately cause variation in electoral manipulation and protest duration.
+First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest.
+Another NELDA variable indicates if the incumbent executive is not running due to *term limits*, which may produce elite fragmentation [@hale_patronal_2014] that affects both protest risk and electoral manipulation.
+De facto *judicial independence* [@Linzer2015] is included, since judicial independence can reduce manipulation [@Harvey2022] and protest risk [@chernykh2014].
+The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance (increasing protest risk, but potentially deterring manipulation under the regime-weakness framework).
+
+In addition, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby due to opposition learning [@Bunce2010;@Beissinger_2007].
+Such a dynamic could plausibly increase the likelihood and severity of protest in a nearby country, while also reducing the incumbent's use of fraud.
 I control for this using each country's geopolitical region as coded in Teorell et al [-@qog_teorell], in conjunction with binary variables indicating a successful protest movement from NELDA.
@@
 
-Several other binary control variables are also included.
-First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest.
-Another NELDA variable indicates if the incumbent executive is not running due to *term limits* [@hale_patronal_2014].
-I include dummy variables that provide a categorical evaluation of state capacity.
+<!-- I include dummy variables that provide a categorical evaluation of state capacity.
 These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap).
@@
 This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018].
-Since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.
-
-I include the *number of pre-election protests* counted in the ECAV dataset, which is likely to be an important possible confounder; more numerous pre-election protest may affect the level of election fraud (as the deterrence logic predicts) as well as being directly tied to the likelihood of post-election protest.
-Finally, in the models of protest duration, I also include a measure of the *median size of protest*, calculated using ECAV data.^[ECAV data codes protest size as a categorical variable, ranging from 1 (< 10 participants) to 5 (> 10,000).
-I take the median size category of all the events in the protest wave as a measure of average protest size.] This is important as a control for the possibility that incumbents could tolerate smaller protests, as discussed above.
+-->
+
+Lastly, the incumbent party's *vote-share* is also taken from the V-Dem dataset.^[This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia, where links to official sources were available, and Psephos.] The original V-Dem variable indicates the *winner's* vote-share.
+However, I exclude all cases where the incumbent lost the election outright.
+This limits the observations to those that fit the population of interest: protest following an election in which the incumbent claimed victory, possibly through reliance on election fraud; it also ensures that the vote-share variable captures the incumbent's position.
+
+Excluding cases where incumbents stepped down after losing an election avoids introducing measurement error in the dependent variable, due to a quirk of the ECAV dataset, which codes the 'side' of the protest (pro- or anti-government) as missing during the period between an election and the swearing in of a new government in cases where the incumbent loses.
+Including these cases thus runs the risk of conflating pro- and anti-incumbent protest, along with cases where protesters goals are genuinely unknown.
+However, excluding them also influences the interpretation of the results from these models.
+Threats to inference from this section are evaluated in more detail in the discussion section.
 
@@
 
-model.binary.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   +    winner.share.upd*v2elirreg.inv*inc.lose_clean   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre + rentierstate + taxstate + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub)
+model.binary.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   +    winner.share.upd*v2elirreg.inv*inc.lose_clean   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre +  unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub)
 
@@
 
-cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose != "1" &  n.events.post > 0))
+cox_base <- coxph(Surv(d.events.num, surv_status) ~   winner.share.upd*v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose != "1" &  n.events.post > 0))
+summary(cox_base)
+
+cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose != "1" &  n.events.post > 0))
 summary(cox1)
 
-cox1.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0))
+cox1.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0))
 summary(cox1.cen)

### 2025-11-13 Output/appendix_v4.Rmd
Updates to appendix section on successful protest

@@
 ---
-title: "Appendix: A missing guardrail---The minimal effect of election fraud on protest intensity in non-democracies"
+title: "Appendix: A fragile guardrail---The minimal effect of election fraud on protest intensity in non-democracies"
 author: "Cole J. Harvey (Oklahoma State University)"
@@
 csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
-header-includes: \usepackage{booktabs}
+header-includes: 
+  - \usepackage{booktabs}
+  - \usepackage{longtable}
 ---
@@
 library(simPH)
+library(broom)
+library(broom.helpers)
 
@@
 
+# Clean inc lose
+nelda2.sub <- nelda2.sub %>% mutate(inc.lose_clean = as.factor(ifelse(inc.lose == "<NA>", NA, inc.lose)))
+
 #Protest cases only
@@
 
+# Summary statistics
+
+Table \@ref(tab:summary-table) provides summary statistics for the variables used in the main text.
+
+```{r, include=F}
+incl.vars <- nelda2.sub %>% dplyr::select('Judicial independence' = jucon.1lag, 
+                                            'Legislative constraints' = legcon.1lag, 
+                                            'Alternative information' = altinf.1lag,
+                                          'Liberal democracy index' = libdem.1lag,
+                                            'GDP growth rate' = gdpgro.1lag, 
+                                            'Civil society openness' = osorg.1lag, 
+                                            #'Public sector corruption' = pubseccorrup.lag,
+                                            'National party organization' = natlpartyorg.lag, 
+                                             'Election fraud' = v2elirreg.inv, 
+                                            'Presidential election' = presidential,
+                                            'Incumbent term-limited' = inc.termlimit.num,
+                                            'Physical integrity index' = physinteg.1lag,
+                                            'Urbanization' = urban.pct, 
+                                            'Rentier state' = rentierstate,
+                                            'Tax state' =  taxstate,
+                                            'Winning vote-share' = winner.share.upd, 
+                                            'Number of pre-election protests' = n.events.pre,
+                                            'Off-schedule election' = unscheduled_election,
+                                          'Regional protest diffusion' = regional_success_lag,
+                                         "Privatization" =  state.owned.1lag,
+                                         "Professional bureaucracy" = patronage.appts.1lag,
+                                        "EMB autonomy"= v2elembaut)
+#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))
+
+incl.vars <- incl.vars %>% mutate(`Presidential election` = as.numeric(`Presidential election`))
+incl.vars <- incl.vars %>% mutate(`Rentier state` = as.numeric(`Rentier state`))
+incl.vars <- incl.vars %>% mutate(`Tax state` = as.numeric(`Tax state`))
+incl.vars <- incl.vars %>% mutate(`Off-schedule election` = as.numeric(`Off-schedule election`))
+
+```
+
+```{r summary-table, echo=FALSE, message=FALSE, warning=FALSE}
+library(kableExtra)
+datasummary_skim(incl.vars, output = "kableExtra", title = "Summary statistics and variable distributions", histogram = F)  %>%
+    kable_styling(font_size = 9, full_width = FALSE)
+```
+
+# Fraud and the likelihood of protest
+
+Table \@ref(tab:logit-table) shows the results for Hypothesis 1.
+Model 1 is a base model, where incumbent loss, winner vote-share, and *election fraud* are standalone variables.
+In keeping with the general (though inconsistent) findings in the prior literature, there is a significant positive relationship between fraud severity and the likelihood of protest initiation.
+The winner's claimed margin of victory is negatively associated with protest initiation.
+Model 2 includes the three-way interaction between *election fraud*, incumbent loss, and winning vote-share, with no controls other than a dummy variable for *presidential elections*.
+This allows for the largest possible sample size, and shows that the coefficient signs are consistent with those of the full Model 3.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
+#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))
+
+model.binary_noint <- glm(any_protest ~ inc.lose_clean + v2elirreg.inv + winner.share.upd, data = nelda2.sub, family = binomial(link = "logit"))
+
+model.binary.base <-  glm(any_protest ~ winner.share.upd*v2elirreg.inv*inc.lose_clean + presidential, data = nelda2.sub, family = binomial(link = "logit"))
+
+model.binary.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   +    winner.share.upd*v2elirreg.inv*inc.lose_clean   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre +  unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub)
+
+plot_binary_margin_typ_n <- sjPlot::plot_model(model.binary.full, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.08, 2.07]", "inc.lose_clean"))
+plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Winner's vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()
+```
+
+```{r logit-table, echo=FALSE}
+models2 <- list()
+models2[['Model 1']] <- model.binary_noint
+models2[['Model 2']] <- model.binary.base
+models2[['Model 3']] <- model.binary.full
+#models2[['Model 6']] <- model.margin.binary_noc
+
+modelsummary(models2, output = "default",  stars = T, title = 'Table 3: Logit models of protest occurrence', gof_omit = c("Log|RMSE"), coef_map = c("(Intercept)" = "(Intercept)",
+             "inc.lose_clean2" = "Incumbent loss",
+             "v2elirreg.inv" = "Election fraud",
+             "winner.share.upd" = "Winner vote-share",
+"winner.share.upd:v2elirreg.inv" = "Winner vote-share:Election fraud",
+"winner.share.upd:inc.lose_clean2" = "Winner vote-share:Inc. loss",
+"v2elirreg.inv:inc.lose_clean2" = "Election fraud:Inc. loss",
+"winner.share.upd:v2elirreg.inv:inc.lose1" = "Election fraud:Winner vote-share:Inc. loss",
+  "lji.lag" = "Judical independence (lag)",                                                                 "legcon.1lag" = "Leg.
+constraints (lag)",                                                                 "gdpgro.1lag" = "GDP growth rate (lag)",                                                                   "altinf.1lag" = "Alternative info.
+(lag)",
+                 "osorg.1lag" = "Civil soc. openness (lag)",
+                 "inc.termlimit.num" = "Incumbent term-limited", 
+                 "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", 
+                 "rentierstate1" = "Rentier state", 
+                 "taxstate1" = "Tax state", 
+                "presidential1" = "Presidential",
+                 "unscheduled_election1" = "Off-schedule election",
+                 "n.events.pre" = "Number of pre-election protests", 
+                 "physinteg.1lag" = "Phys. integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", 
+                 "regional_success_lag" = "Regional protest diffusion",
+
+                 "median.participants" = "Median protest participants",
+
+                "v2elirreg.inv:winner.share.upd"= "Election fraud:Winner vote-share")
+                 )
+
+#reg_table |> tab_options(
+#  latex.use_longtable = TRUE, 
+#  data_row.padding = px(5)
+#  ) 
+```
+
+Model 3 is the main model, which includes the full set of controls.
+Given the complexity of the model, Figure \@ref(fig:margin-prob-fig) illustrates the effects of election fraud on protest risk across various levels of winning vote-share and incumbent loss^[All other variables constant at their means (continuous variables) or modes (categorical variables).].
+The values for election fraud represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively.
+
+The direction of the effect supports Hypothesis 1.
+When an incumbent claims victory (left panel) and their vote-share is low, electoral protest is significantly likelier when fraud is elevated than when the election is relatively clean.
+Furthermore, protest against winning incumbents is significantly likelier when fraud is high and vote-shares are low, compared to an election with an equivalent level of fraud but high claimed vote-shares.
+As noted above, such a finding could be consistent with either the regime-weakness or regime-strength models of manipulation, under different conceptualizations of protest.
+To distinguish between the two, we turn to the models of protest duration.
+
+```{r margin-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
+plot_binary_margin_typ_n$data$facet <-  ifelse(plot_binary_margin_typ_n$data$facet == "1", "Incumbent won", "Incumbent lost")
+
+plot_binary_margin_typ_n
+```
+
 # Alternative measures of election manipulation
+
+## Overall election integrity
 
@@
 
-```{r manipdv-full, echo=FALSE}
-m1 <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + unscheduled_election  +   (1 | stateid), data = nelda2.sub)
-
-#summary(m1)
-
-m1a <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag  +  (1 | stateid), data = nelda2.sub)
-#summary(m1a)
-
-m1b <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag +  unscheduled_election  +  (1 | stateid), data = nelda2.sub)
-
-#summary(m1b)
-
-m1c <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + unscheduled_election   +  (1 | stateid), data = nelda2.sub)
-#summary(m1c)
-
-models <- list()
-models[['Model 1']] <- m1
-models[['Model 2']] <- m1a
-models[['Model 3']] <- m1b
-models[['Model 4']] <- m1c
-
-modelsummary(models, stars = T,  title = 'Random-intercept multilevel model of overall election integrity', gof_omit = "AIC|BIC",
-             coef_rename = 
-               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
-                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", "presidential" = "Presidential", "n.events.pre" = "N. of pre-election protest",
-                 "physinteg.1lag" = "Physical integrity (lag)",
-                 "physintlag.sq" = "Phys. integ. (lag), squared", "unscheduled_election1" = "Off-schedule election",
-                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
-                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
-                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
-                 "urban.pct" = "Urbanization", "rentierstate1" = "Rentier state", "taxstate1" = "Tax state",
-                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests",
-                 "state.owned.1lag" = "Private ownership (lag)",
-                 "patronage.appts.1lag" = "Professional bureaucracy (lag)",
-                 "regional_success_lag" = "Regional protest diffusion"))
-```
-
-As expected, the first model shows that the number of pre-election protests is negatively associated with overall election integrity\textemdash contrary to deterrence expectations and in line with the argument of the main text.
-We also see that overall election integrity is lowest at middle levels of repression; this is, again, in line with the expectations of the incumbent strength model and contrary to deterrence expectations.
-No significant effect on overall election integrity is observed for successful electoral protests in the region.
+The same interaction relationship between overall election integrity, incumbent vote-share, and protest as is shown for fraud in the main text.
+That is, less free-and-fair elections are only associated with a higher protest risk at low levels of incumbent vote-share.
 
@@
 
-model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd +  presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-#summary(model.margin.binary)
+model.margin.binary.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+#summary(model.margin.binary.full)
 
-model.margin.binary.int <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
-#summary(model.margin.binary.int)
+model.margin.binary.int.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd*v2xel_frefair   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+#summary(model.margin.binary.int.full)
 
-model.margin.binary_noc <- glm(any_protest ~ jucon.1lag + legcon.1lag +  altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
 #summary(model.margin.binary_noc)
@@
 
-Moving to the logit models of post-election protest, we observe the same interaction relationship between overall election integrity, incumbent vote-share, and protest as is shown for fraud in the main text.
-That is, less free-and-fair elections are only associated with a higher protest risk and low levels of incumbent vote-share.
+```{r logit-table2, echo=FALSE}
+models2 <- list()
+models2[['Model 1']] <- model.margin.binary.full
+models2[['Model 2']] <- model.margin.binary.int.full
 
-```{r logit-table, echo=FALSE}
-models2 <- list()
-models2[['Model 5']] <- model.margin.binary
-models2[['Model 6']] <- model.margin.binary.int
-models2[['Model 7']] <- model.margin.binary_noc
-
-modelsummary(models2, output = "kableExtra",  stars = T, title = 'Logit models of protest occurrence', coef_rename = 
-               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+modelsummary(models2,  stars = T, title = 'Logit models of protest occurrence', gof_omit = "Log|RMSE", coef_rename = 
+               c("lji.lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                  "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
@@
                  "winner.share.upd" = "Incumbent vote-share",
-                 "v2xel_frefair:winner.share.upd" = "Election integ. : incumbent vote-share"
+                 "v2xel_frefair:winner.share.upd" = "Election integ. : Incumbent vote-share"
                  )) %>% kable_styling(latex_options = "HOLD_position")
@@
 ```{r overall-binary, fig.cap="Overall election integrity, winning vote-share, and probability of protest initiation", fig.height=6, echo=FALSE, message=FALSE, fig.align='center'}
-sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c("winner.share.upd", "v2xel_frefair [.12, .46]"))  +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election integrity") + theme_bw()
+sjPlot::plot_model(model.margin.binary.int.full, type = "pred", terms = c("winner.share.upd", "v2xel_frefair [.12, .46]"))  +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election integrity") + theme_bw()
 
@@
 
-cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2xel_frefair  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
+nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
+nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(inc.replaced.prot == 1, 0, surv_status))
+nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
+                                                                       ifelse(d.events.num >= 69, 0, 1)))
+
+cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2xel_frefair   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" &  n.events.post > 0))
 #summary(cox1)
 
-cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2xel_frefair*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2xel_frefair*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
+#summary(cox2)
 #summary(cox2)
@@
 models3 <- list()
-models3[['Model 8']] <- cox1
-models3[['Model 9']] <- cox2
+models3[['Model 3']] <- cox1
+models3[['Model 4']] <- cox2
 
@@
 
-These three sets of models all replace a measure of *election fraud* with a broader measure of overall election integrity.
-In all three cases, the results broadly confirm the incumbent-strength model and are unsupportive of the deterrence model.
+The two sets of models all replace a measure of *election fraud* with a broader measure of overall election integrity.
+In all cases, the results broadly confirm the incumbent-strength model and are unsupportive of the deterrence model.
 
@@
 
-# Repression and pre-election protest
+## Election observers' reports of fraud
 
-The main analysis already controls for possible confounding by the state's repressive capacity; i.e.
-the possibility that in highly repressive states, protest is deterred and fraud can proceed unopposed.
-However, it may still be useful to check an alternative explanation\textemdash perhaps manipulation is deterred in highly repressive states with considerable pre-election unrest (that is, when protest appears to be exceeding the ability of the state to contain it).
-To check this possibility, a variation of Model 1 from the main text is run, but with an interaction between the number of pre-election protests and the lagged *physical integrity* variable.
-If protest risk deters manipulation in this way, there should be a negative and significant effect for the number of pre-election protests in highly repressive states.
+One obstacle to the validity of the findings could be systematic bias by the human coders utilized in the V-Dem project [@Little_Meng_2024].
+While there is some evidence that such bias is small [@weidmann2023], and election monitors can also be subject to biases [@kelley2012], it may be useful to check the V-Dem coders' assessments against election monitors' reports.
+To do so, I make use of the Dataset on International Election Monitoring [@kelley2011], or DIEM.
+This dataset codes the judgments of international election monitoring missions as categorical variables.
+I use the variable "Explicit cheating" as a parallel to the V-Dem "intentional irregularities" variable.
+The explicit cheating variable is categorical, ranging from the 0 for the cleanest elections to 3 for the most fraudulent.
+While this data is closely tied to the facts on the ground, as reported by election monitors, it covers a much smaller range of cases than V-Dem: of the 661 observations in the full dataset, 471 do not have DIEM values.
+This is, in part, due to differences in the time periods covered (DIEM ranges from 1980 to 2004, while the main dataset covers 1989 - 2012).
+This problem is compounded by the fact that most elections in the dataset go un-protested.
+Of the 156 cases where protest occurs, 110 lack data in DIEM.
+While this limits the usefulness of the DIEM variable, it can still provide some confirmatory information for the main analysis.
 
-Figure \@ref(fig:prot-rep-fraud) shows the absence of such an effect.
-In fact, there is no significant change in the predicted level of fraud at lower levels of physical integrity, across all levels of pre-election protest.
-We do observe a significant effect for low repression regimes; however, it is in the opposite direction predicted by this counterargument.
-That is, more numerous pre-election protest is associated with higher predicted levels of fraud in less repressive states, not harsher ones.
-The figure also shows that less repressive regimes appear to respond to greater civic unrest with more fraud.
-Altogether, this robustness check shows that pre-election protest in a repressive environment does not appear to deter fraud, and supports the view that manipulation may be used to indicate regime durability in a challenging period.
+Table \@ref(tab:diem-table-1) reports a model where the DIEM measure as the explanatory variable in a survival model of protest movements.
+Here the difficulty with missingness becomes apparent, with only 39 observations in the model.
+Given the small number of observations, it is even more reassuring that the positive relationship between fraud and a quick end to protest movements is apparent.
 
-```{r prot-rep-fraud, message=FALSE, warning=F, echo=FALSE, fig.cap="Marginal effect of pre-election protest on fraud, conditional on physical integrity", out.width = '75%', fig.align='center'}
-m1_rep <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + unscheduled_election  + v2elembaut + physinteg.1lag*n.events.pre + (1 | stateid), data = nelda2.sub)
-#summary(m1_rep)
-p.m1rep <- sjPlot::plot_model(m1_rep, type = "pred", terms = c("n.events.pre", "physinteg.1lag [.2, .7]")) #.25 and .75 quantile
-p.m1rep + labs(title = "Predicted values of election fraud", x = "Number of pre-election protests", y = "Election fraud", color = "Physical integrity (lag)") + theme_bw()
+```{r, echo=FALSE}
+
+cox1_diem <- coxph(Surv(d.events.num, surv_status) ~ legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + diem_explicheat_max   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" &  n.events.post > 0))
+#summary(cox1_diem)
+
 ```
 
+```{r diem-table-1, echo=FALSE, message=FALSE, warning=FALSE}
+models <- list()
+models[['Model 5']] <- cox1_diem
+
+modelsummary(models, output = "kableExtra",  stars = T, title = 'Models using DIEM measure of election fraud',  coef_rename = 
+               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
+                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
+                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
+                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
+integrity index (lag)",
+                 "physintlag.sq" = "Phys.
+int.
+lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity", "diem_explicheat_max" = "Explicit cheating (DIEM)", "n.events.post" = "N.
+post-election protests"
+                 )) %>% kable_styling(latex_options = "HOLD_position")
+```
+
+# Further tests of deterrence logic: predicted and residual levels of fraud
+
+The main text focuses on elections where the incumbent claims victory.
+However, classic deterrence models also make predictions for incumbents who will voluntarily avoid electoral manipulation and accept a loss at the ballot box.
+Fully investigating this possibility should be the focus of future work, but some initial analyses are provided here in support of the claims made in the main paper.
+First, using all the observations in the dataset\textemdash winning and losing incumbents\textemdash to find the predicted level of election fraud for each case, using only structural variables (i.e.
+professionalization of the bureaucracy, electoral management body autonomy, urbanization, etc.).
+Factors associated with deterrence, like the number of pre-election protests are excluded.
+The fitted values and residuals from this first model are then used to check the implications of the theory in subsequent analysis.
+This model is simple OLS regression.
+
+Model 7 takes the residuals from the structural model as a dependent variable.
+This model is used to check whether pre-election protest risk is associated with incumbents over- or under-performing their expected level of manipulation based on structural factors.
+The deterrence model, which the main paper argues against, suggests that when the risk of anti-incumbent mobilization is higher, incumbents should be more likely to restrain themselves and under-deliver on manipulation.
+Since the residuals which are used as a DV are continuous, this is also an OLS model.
+
+The theory proposed in the main text implies that structural factors enabling manipulation should be significant *negative* predictors of incumbent loss, while measures of pre-election risk should not.
+That is, the theory expects that losing incumbents are more likely to structurally limited in the level of manipulation they can generate, rather than being structurally unconstrained but deterred by a more robust opposition.
+Model 8 in Table \@ref(tab:fraud-loss) takes incumbent defeat in the election as the dependent variable, with the expected value of manipulation (derived from the first model) as an explanatory variable.
+Pre-election protest and successful protests in the region are included as controls.
+Due to the binary dependent variable, this is a logit model.
+
+```{r, include=F, echo=FALSE, message=FALSE, warning=F}
+## Anova
+anova_dat <- nelda2.sub %>% dplyr::select(inc.replaced.prot, v2elirreg.inv,  jucon.1lag, legcon.1lag, gdpgro.1lag,  osorg.1lag, natlpartyorg.lag,  urban.pct, n.events.pre, state.owned.1lag, patronage.appts.1lag, v2elembaut)
+
+# boxplots and t-tests for the 4 variables at once
+for (i in 2:ncol(anova_dat)) { # variables to compare are variables 2 to end
+  boxplot(anova_dat[, i] ~ anova_dat$inc.replaced.prot, # draw boxplots by group
+    ylab = names(anova_dat[i]), # rename y-axis with variable's name
+    xlab = "Incumbent replaced after protest"
+  )
+  print(paste(c("ANOVA for variable ", (names(anova_dat[i])))))
+  print(t.test(anova_dat[, i] ~ anova_dat$inc.replaced.prot)) # print results of t-test
+}
+
+aov_output <- aov(state.owned.1lag ~ inc.replaced.prot, data = nelda2.sub)
+summary(aov_output)
+```
+
+```{r, echo=FALSE, message=FALSE, include=FALSE, warning=FALSE}
+m.fraud.pred <- lm(v2elirreg.inv ~ jucon.1lag +  gdpgro.1lag + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + v2elembaut, data = nelda2.sub)
+
+summary(m.fraud.pred)
+
+nelda2.sub.fraudpred <- nelda2.sub %>% mutate(.rownames = rownames(nelda2.sub))
+
+new_frame <- augment(m.fraud.pred)
+new_frame <- new_frame %>% dplyr::select(.fitted, .rownames, .resid)
+
+nelda2.sub.fraudpred <- nelda2.sub.fraudpred %>% left_join(new_frame, by = ".rownames")
+
+model.lose <- glm(inc.lose ~ .fitted + regional_success_lag + n.events.pre, data = nelda2.sub.fraudpred, family = binomial(link="logit"))
+summary(model.lose)
+
+model.fraud.resid <- lm(.resid ~ n.events.pre +  regional_success_lag , data = nelda2.sub.fraudpred)
+summary(model.fraud.resid)
+```
+
+```{r fraud-loss, echo=FALSE}
+models2 <- list()
+models2[['Model 6']] <- m.fraud.pred
+models2[['Model 7']] <- model.fraud.resid
+models2[['Model 8']] <- model.lose
+
+row <- data.frame(cbind("", "Fraud", "Fraud residual", "Inc. loss"))
+attr(row, 'position') <- c(0)
+
+modelsummary(models2, output = "kableExtra",  stars = T, title = 'Models of manipulation and residuals (OLS) and incumbent loss (logit)', coef_omit = "as.factor", gof_omit = "RMSE|Log.", coef_rename =
+               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
+                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
+                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", "v2elirreg.inv" = "Election fraud",
+                 "winner.share.upd" = "Incumbent vote-share",
+                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
+                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
+integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity",
+                 "state.owned.1lag" = "Private ownership (lag)",
+                 "patronage.appts.1lag" = "Professional bureaucracy (lag)",
+                 "v2elembaut" = "EMB autonomy",
+                 ".fitted" = "Expected fraud level"
+                 ), add_rows = row) %>% kable_styling(latex_options = "HOLD_position")
+
+## Figure out how to Add a row beneath the model numbers, with the DV
+```
+
+Table \@ref(tab:fraud-loss) shows the results.
+Model 6 shows how structural factors are predictors of the level of observed manipulation.
+We observe some expected results; for instance, more autonomous EMBs and more professionalized bureacuracies are associated with lower fraud.
+Model 7 shows that indicators of protest risk\textemdash the number of pre-election protests, and the existence of a recent successful protest movement in the region\textemdash are not associated with lower fraud.
+In fact, just the opposite: incumbents appear to manipulate more heavily under riskier conditions.
+This is initial evidence against the classic deterrent approach.
+Model 8 takes the fitted values from the first model and employs them as an explanatory variable for incumbent loss, while controlling for pre-election protest.
+As the latter model shows, the predicted level of fraud based on the incumbent's structural advantages is negatively associated with incumbent loss.
+This is supportive of the argument in the main text, which holds that the capacity to manipulate heavily is associated with incumbent durability.
+Altogether, the results in this section are consistent with a world in which losing incumbents are not voluntarily surrendering power in the face of a committed opposition (on average), but are failing to produce sufficient election manipulation to win due to structural constraints.
+
 \newpage
+
+# Incumbent ousted by protest
+
+The classic model of election manipulation as an indicator of weakness would suggest that incumbents who rely more on manipulation will be more likely to fall when protest movements do emerge.
+The main text does not examine successful protests, which are a small minority of cases, in order to avoid measurement error in the duration models.
+Successful protest is, instead, examined here; there is little to suggest that incumbents who manipulate more are more vulnerable to defeat by protesters.
+
+First, Figure \@ref(fig:successplot) shows cases of 'successful' and unsuccessful protest at varying levels of election fraud.
+Protests are coded as successful if they lead to incumbent ouster or to new elections as recorded in NELDA.
+As the figure shows, the large majority of protests are unsuccessful by this measure.
+Moreover, the small number of successful cases appear spread over approximately the same range of election fraud as unsuccessful cases; we do not observe a clustering of successful protest at higher levels of fraud.
+
+Next, Table \@ref(tab:oust-mods) reports three logit regression models of successful protest; the dataset for these models includes all cases with a least one protest event.
+Model 9 is a minimal model that includes *intentional election irregularities*, *incumbent vote-share* and a control for *presidential elections*.
+Model 10 includes a larger number of control variables, while Model 11 adds the interaction term.
+As the results in the table show, greater election manipulation is not associated with a higher likelihood of incumbent defeat in any model.
+Figure \@ref(fig:oust-plot) shows no significant effect of election manipulation on protest success across all values of incumbent vote-share.
+Taken together, this is further disconfirmation of the fraud-as-weakness model.
+
+```{r successplot, fig.cap="Election fraud and successful protest", fig.height=6, echo=FALSE, message=FALSE, fig.align='center'}
+
+ggplot(protest.cases, aes(v2elirreg.inv, protest.success)) +
+  geom_jitter(width = 0, height = 0.1)+
+  xlab("Electoral fraud") +
+  ylab("Successful protest (jittered)") +
+  theme_bw()
+```
+
+```{r oust-mods, echo=FALSE, message=FALSE, warning=FALSE}
+
+model.oust3 <- glm(protest.success ~  osorg.1lag + gdpgro.1lag  + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +   unscheduled_election + taxstate + rentierstate + regional_success_lag + v2elirreg.inv , family = binomial(link = "logit"), data = protest.cases %>% filter(inc.lose_clean == "1"))
+summary(model.oust3)
+
+model.oust4 <- glm(protest.success ~  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +   unscheduled_election + taxstate + rentierstate + regional_success_lag +  v2elirreg.inv*winner.share.upd , family = binomial(link = "logit"),  data = protest.cases %>% filter(inc.lose_clean == "1"))
+summary(model.oust4)
+
+model.oust5 <- glm(protest.success ~  presidential +  v2elirreg.inv*winner.share.upd , family = binomial(link = "logit"),  data = protest.cases %>% filter(inc.lose_clean == "1"))
+summary(model.oust5)
+
+models2 <- list()
+models2[['Model 9']] <- model.oust5
+models2[['Model 10']] <- model.oust3
+models2[['Model 11']] <- model.oust4
+
+modelsummary(models2, output = "default",  stars = T, title = 'Logit models of incumbent ouster by protest', coef_omit = "SD", gof_omit = "AIC|BIC|Log|F|RMSE", coef_rename = 
+               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
+                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
+                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
+                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
+integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "state.owned.1lag" = "Privatization",
+                 "patronage.appts.1lag" = "Professional bureaucracy",
+                 "v2elirreg.inv" = "Election fraud",
+                 "winner.share.upd" = "Winner vote-share"
+                 ))
+```
+
+```{r oust-plot, fig.cap="Fraud, winning vote-share, and incumbent ouster", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
+sjPlot::plot_model(model.oust5, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.4, 1.88]")) + labs(x = "Winner vote-share", y = "Pred.
+probability of successful protest", color = "Election fraud") + theme_bw()
+```
 
@@
 models4 <- list()
-models4[['Model 10']] <- model.margin.int.suc
+models4[['Model 13']] <- model.margin.int.suc
 #models4[['Model 6']] <- model.d.base2.hurd
-models4[['Model 11']] <- cox2.suc
+models4[['Model 14']] <- cox2.suc
 
@@
 
-```{r successor-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=6, echo=FALSE, message=FALSE, fig.align='center'}
+```{r successor-fig, fig.cap="Fraud, winning vote-share, and protest initiation / defeat", fig.height=6, echo=FALSE, message=FALSE, fig.align='center'}
 plot.combined.prob <- cowplot::plot_grid(p.n.margin, p.s.margin, ncol=1)
@@
 models2 <- list()
-models2[['Model 12']] <- model.margin.binary
-models2[['Model 13']] <- model.margin.binary.int
+models2[['Model 15']] <- model.margin.binary
+models2[['Model 16']] <- model.margin.binary.int
 
@@
 models3 <- list()
-models3[['Model 14']] <- cox1
-models3[['Model 15']] <- cox2
+models3[['Model 17']] <- cox1
+models3[['Model 18']] <- cox2
 
@@
 
+\newpage
+
 ```{r margin-dur-fig-small, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
@@
 \newpage
+
+# Using country fixed effects
+
+Models that include fixed effects by country can be used to estimate the effects of explanatory variables within countries, rather than being driven by a subset of cases where fraud and protest both tend to be high (for example).
+Table \@ref(tab:logit-table-fe), where logit models of protest initiation are presented, shows results consistent with those in the main text.
+Fraud alone has no significant relationship with protest initiation, but it does in the context of low incumbent vote-share.
+Fixed effects models are not employed for the proportional hazards models of protest survival; due to the small number of cases for which multiple protest waves occur, there is insufficient within-country variation to make use of such models.
+
+```{r include=FALSE}
+nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
+#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))
+
+model.margin.binary.fe <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +  n.events.pre + unscheduled_election + regional_success_lag + as.factor(country), family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+#summary(model.margin.binary)
+
+model.margin.binary.int.fe <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +  n.events.pre + unscheduled_election + regional_success_lag +  as.factor(country), family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
+#summary(model.margin.binary.int)
+
+```
+
+```{r logit-table-fe, echo=FALSE}
+models2 <- list()
+models2[['Model 19']] <- model.margin.binary.fe
+models2[['Model 20']] <- model.margin.binary.int.fe
+
+modelsummary(models2, output = "kableExtra",  stars = T, title = 'Table 3: Logit models of protest occurrence', coef_omit = "as.factor", gof_omit = "RMSE|Log.", coef_rename = 
+               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
+                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
+                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
+                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
+                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", "v2elirreg.inv" = "Election fraud",
+                 "winner.share.upd" = "Incumbent vote-share",
+                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
+                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
+                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys.
+integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity"
+                 ), add_rows = data.frame(cbind("Country fixed effects", "Y", "Y"))) %>% kable_styling(latex_options = "HOLD_position")
+```

### 2025-11-13 Output/authoritarian_only_pdf_pop_rev.Rmd
Cuts in appendix; begin edits to discussion section in main

@@
 
->**H2 (Regime-weakness)**: Increases in electoral manipulation will be associated with decreasing duration of unsuccessful electoral protest when incumbent vote-share is low.
+>**H2 (Regime-strength)**: Increases in electoral manipulation will be associated with decreasing duration of unsuccessful electoral protest when incumbent vote-share is low.
 
@@
 Still, approximately one-third of elections experienced some protest, and so protest initiation is modeled using standard logistic regression.
-Protest wave duration is modeled using Cox proportional hazards models.
+Protest wave duration is modeled using Cox proportional hazards models, which have the following formula:
+
+$$h(t∣X)=h_0(t)* exp(β_fX_f+β_vX_v+⋯+βpXp)$$
+The formula models the risk that a protest movement will fail at time $t$ given covariates $X$ is a function of the baseline hazard function, the degree of electoral fraud ($X_f$), the incumbent's vote-share ($X_v$) and a set of control variables.
 
@@
 
-Lastly, the incumbent party's *vote-share* is also taken from the V-Dem dataset.^[This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia, where links to official sources were available, and Psephos.] The original V-Dem variable indicates the *winner's* vote-share.
+Lastly, the *incumbent party's vote-share* is also taken from the V-Dem dataset.^[This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia, where links to official sources were available, and Psephos.] The original V-Dem variable indicates the *winner's* vote-share.
 However, I exclude all cases where the incumbent lost the election outright.
@@
 
-## Fraud and the likelihood of protest
-
-Table \@ref(tab:logit-table) shows the results for Hypothesis 1.
-Model 1 is a base model, where incumbent loss, winner vote-share, and *election fraud* are standalone variables.
-In keeping with the general (though inconsistent) findings in the prior literature, there is a significant positive relationship between fraud severity and the likelihood of protest initiation.
-The winner's claimed margin of victory is negatively associated with protest initiation.
-Model 2 includes the three-way interaction between *election fraud*, incumbent loss, and winning vote-share, with no controls other than a dummy variable for *presidential elections*.
-This allows for the largest possible sample size, and shows that the coefficient signs are consistent with those of the full Model 3.
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
-#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))
-
-model.binary_noint <- glm(any_protest ~ inc.lose_clean + v2elirreg.inv + winner.share.upd, data = nelda2.sub, family = binomial(link = "logit"))
-
-model.binary.base <-  glm(any_protest ~ winner.share.upd*v2elirreg.inv*inc.lose_clean + presidential, data = nelda2.sub, family = binomial(link = "logit"))
-
-model.binary.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   +    winner.share.upd*v2elirreg.inv*inc.lose_clean   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre +  unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub)
-
-plot_binary_margin_typ_n <- sjPlot::plot_model(model.binary.full, type = "pred", terms = c("v2elirreg.inv", "winner.share.upd [.35, .78]", "inc.lose_clean"))
-#plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Winner's vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()
-```
-
-```{r logit-table, echo=FALSE}
-models2 <- list()
-models2[['Model 1']] <- model.binary_noint
-models2[['Model 2']] <- model.binary.base
-models2[['Model 3']] <- model.binary.full
-#models2[['Model 6']] <- model.margin.binary_noc
-
-modelsummary(models2, output = "default",  stars = T, title = 'Table 3: Logit models of protest occurrence', gof_omit = c("Log|RMSE"), coef_map = c("(Intercept)" = "(Intercept)",
-             "inc.lose_clean2" = "Incumbent loss",
-             "v2elirreg.inv" = "Election fraud",
-             "winner.share.upd" = "Winner vote-share",
-"winner.share.upd:v2elirreg.inv" = "Winner vote-share:Election fraud",
-"winner.share.upd:inc.lose_clean2" = "Winner vote-share:Inc. loss",
-"v2elirreg.inv:inc.lose_clean2" = "Election fraud:Inc. loss",
-"winner.share.upd:v2elirreg.inv:inc.lose1" = "Election fraud:Winner vote-share:Inc. loss",
-  "lji.lag" = "Judical independence (lag)",                                                                 "legcon.1lag" = "Leg.
-constraints (lag)",                                                                 "gdpgro.1lag" = "GDP growth rate (lag)",                                                                   "altinf.1lag" = "Alternative info.
-(lag)",
-                 "osorg.1lag" = "Civil soc. openness (lag)",
-                 "inc.termlimit.num" = "Incumbent term-limited", 
-                 "libdem.1lag" = "Liberal democracy index (lag)", 
-                 "urban.pct" = "Urbanization", 
-                 "rentierstate1" = "Rentier state", 
-                 "taxstate1" = "Tax state", 
-                "presidential1" = "Presidential",
-                 "unscheduled_election1" = "Off-schedule election",
-                 "n.events.pre" = "Number of pre-election protests", 
-                 "physinteg.1lag" = "Phys. integrity index (lag)",
-                 "physintlag.sq" = "Phys. int. lag squared", 
-                 "regional_success_lag" = "Regional protest diffusion",
-
-                 "median.participants" = "Median protest participants",
-
-                "v2elirreg.inv:winner.share.upd"= "Election fraud:Winner vote-share")
-                 )
-
-#reg_table |> tab_options(
-#  latex.use_longtable = TRUE, 
-#  data_row.padding = px(5)
-#  ) 
-```
-
-Model 3 is the main model, which includes the full set of controls.
-Given the complexity of the model, Figure \@ref(fig:margin-prob-fig) illustrates the effects of election fraud on protest risk across various levels of winning vote-share and incumbent loss^[All other variables constant at their means (continuous variables) or modes (categorical variables).].
-The values for election fraud represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively.
-
-The direction of the effect supports Hypothesis 1.
-When an incumbent claims victory (left panel) and their vote-share is low, electoral protest is significantly likelier when fraud is elevated than when the election is relatively clean.
-Furthermore, protest against winning incumbents is significantly likelier when fraud is high and vote-shares are low, compared to an election with an equivalent level of fraud but high claimed vote-shares.
-As noted above, such a finding could be consistent with either the regime-weakness or regime-strength models of manipulation, under different conceptualizations of protest.
-To distinguish between the two, we turn to the models of protest duration.
-
-```{r margin-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
-plot_binary_margin_typ_n
-```
-
-## Protest wave duration
-
 ```{r, include=FALSE}
@@
 
-cox_base <- coxph(Surv(d.events.num, surv_status) ~   winner.share.upd*v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose != "1" &  n.events.post > 0))
+cox_base <- coxph(Surv(d.events.num, surv_status) ~   winner.share.upd*v2elirreg.inv + presidential  + n.events.post , data = nelda2.sub %>% filter(inc.lose != "1" &  n.events.post > 0))
 summary(cox_base)
@@
 In these models, positive coefficients indicate factors that are associated with shorter-lived protests; those with a negative sign indicate factors associated longer-lasting protest.
+Recall that these models include data only from electoral protests that *did not succeed* in ousting incumbents.
+Extensions of these models that
+
 Models 4 and 6 do not include interactions, while the remaining models interact *intentional irregularities* with incumbent vote-share.
@@
 models3 <- list()
+models3[['Model 1']] <- cox_base
 models3[['Model 4']] <- cox1
@@
 The duration analysis consistently shows that more election manipulation is associated with quicker incumbent victories over protest movements, both independently and in conjunction with low incumbent vote-share.
-This is a strong indication that governments which can deliver large-scale election manipulation have the resources and capabilities to put down protest more quickly; that is, such manipulation covaries with sources of regime durability.
 In real contests in which opposition and incumbent strengths are tested against each other, better manipulators reveal themselves to be stronger rather than weaker.
-This is consistent with Hypothesis 2.
-
-This finding supports and extends the line of research initiated by Simpser [@simpser2013], which I have called the regime-strength model of election manipulation.
-While Simpser's original argument holds that electoral manipulation in conjunction with a large margin of victory for the incumbent can induce pro-regime behavior in society, this paper shows that even when such cooperation fails---when post-election protests occur---governments that manipulate more are more likely to successfully weather the challenge.
-That is, even when ruling party vote-shares are low (perhaps especially when this is so), incumbents which show themselves capable of large-scale fraud are more capable of quickly ending protest movements than low-fraud incumbents.
-This holds true despite controlling for indicators of other possible explanations of the relationship between low fraud and long-lived protest: civil society strength, media openness, and median protest size.
-
-Adopting this interpretation of election manipulation as a reflection of effective sources of regime strength comports with a growing literature on the principal-agent dynamics of manipulation [@Harvey2019;@rundlett2016deliver] and the importance of patronage resources and electoral processes in sustaining authoritarian regimes [@hale_patronal_2014].
+This holds true even when incumbent vote-share is low, when the regime-weakness model would make the opposite prediction---in that model, incumbents who are unpopular and manipulate heavily ought to be the weakest and struggle hardest to quell protest.
+It is therefore inconsistent with Hypothesis 1, casting doubt on the 'regime-weakness' model of manipulation.
+Further, the relationship holds despite controlling for indicators of other possible explanations of the relationship between low fraud and long-lived protest, including civil society strength, media openness, regional electoral protest waves, number of protests and median protest size.
+This is consistent with Hypothesis 2, and supportive of what I have called the 'regime-strength' model of electoral manipulation.
+
+This finding supports and extends the line of research initiated by Simpser [@simpser2013].
+Understanding election manipulation as indicative of effective sources of regime strength comports with a growing literature on the principal-agent dynamics of manipulation [@Harvey2019;@rundlett2016deliver;@Harvey2022] and the importance of patronage resources and electoral processes in sustaining authoritarian regimes [@hale_patronal_2014].
+This study offers empirical evidence that authoritarian governments who manipulate more are stronger than those who manipulate less (all else equal).
+In addition to offering a counter-argument to regime-weakness models, it also illustrates how the mechanisms which underpin Simpser's model of manipulation-as-strength---coordination effects among pro-regime agents and costly signaling of incumbent resources---can function even when such signals of strength have failed to 'deter' protest.
+
+Specifically, the results highlight patterns that might be considered puzzling, whichever model of electoral manipulation we prefer.
+For the regime-weakness model, it is potentially puzzling that most manipulated elections go un-protested.
+Regime-weakness models may explain this pattern by arguing that, most of the time, incumbents produce the 'right' amount of manipulation to avoid protest in equilibrium.
+Yet the results of this study challenge that interpretation, since it demonstrates that outside this equilibrium (i.e.
+when elections are rigged and the opposition protests), higher levels of manipulation are associated with greater regime strength, not less.
+
+In rejoinder, it might reasonably be wondered why protests occur after highly manipulated elections, if manipulation is a signal of strength?
+First, in my view, this interpretation unnecessarily exports the instrumental logic of protest from the regime-weakness model to the regime-strength model.
+As discussed earlier in the paper, instrumental calculations about overthrowing incumbents are but one component of a long and varied literature on protest, which has identified multiple other expressive and strategic motives for protest.
+These other motives may foster protest movements in the aftermath of rigged elections, even when severe electoral manipulation indicates considerable regime strength.
+Second, if manipulation does indeed indicate strength, incumbents have every incentive to build and deploy those strengths against stronger opposition movements (who will themselves be more capable of initiating protest).
+
+most regimes are operating at an equilibrium, wherein just the right amount of manipulation is produced that avoids protest.
+that  and---as this study shows---when electoral protest does occur, such movements are defeated in the large majority of cases.
+If manipulation signals weakness, why are more rigged elections not protested, and why are more protest movements not successful?
+The regime-weakness model proposes answers for each of these questions.
+
+While Simpser's original argument holds that electoral manipulation in conjunction with a large margin of victory for the incumbent can induce pro-regime behavior in society, this paper shows that even when such cooperation fails---when post-election protests occur---governments that manipulate more are better positioned to quickly defeat challengers.
+
  It also helps us make sense of secular patterns that are puzzling if election fraud indicates regime weakness, as others have held.

### 2025-11-14 Output/authoritarian_only_pdf_pop_rev.Rmd
Shaggy additions to research design section on endogeneity and omitted variables.

@@
 abstract: "Two models of the relationship between election manipulation and protest can be found in the literature.
-One approach holds that election fraud reveals an incumbent who is more vulnerable to protest; in this view, protest risk can result in cleaner elections.
+One approach holds that more severe election fraud reveals an incumbent who is more vulnerable to protest; in this view, protest risk can result in cleaner elections.
 An alternative approach contends that election manipulation can be a sign of underlying regime strengths; under this framework, protest risk is less likely to deter manipulation.
 This article interrogates the assumptions underlying these models, and conducts a novel test of their implications.
-Using data on the duration of protest waves in 659 authoritarian cases from 1990-2012, this paper shows that governments which manipulate elections more also defeat protests more quickly.
+Using data on the occurrence and duration of protest waves in 659 authoritarian cases from 1990-2012, this paper shows that governments which manipulate elections more also defeat protests more quickly.
 This supports the ‘regime-strength’ model, and helps explain why electoral authoritarianism is a sticky equilibrium amid ongoing democratic backsliding.
@@
 
+Modeling protest duration rather than initiation helps reduce concerns about endogeneity.
+For instance, some regime-weakness models expect that incumbents will calibrate the level of election fraud to minimize risk of protest initiation.
+In that case, strong incumbents would manipulate more and weak incumbents less, but neither would systematically generate protest.
+Unpredictable and likely statistically insignificant correlations between fraud and protest initiation would be expected.
+Focusing on duration partially avoids this risk.
+Once protests begin (whether instrumental or multi-purpose), the die is cast.
+Authoritarian incumbents have strong incentives to defeat a public challenge to their legitimacy, while committed opponents have similar incentives to stand firm.
+Both should be expected to put significant effort into defeating / maintaining the protest movements, as a general rule, and the process by which the challenge is settled helps reveal previously hidden information about each side.
+Whether the incumbent manipulated 'too much' or 'too little' is now beside the point, and it is possible to observe whether incumbents who produce extensive election fraud are stronger or weaker than those that do not.
+
+While strategic endogeneity is less of a concern for this research design, selection bias and confounding variables may still be.
+For instance, it could be the case that incumbents facing well-organized opposition-oriented civil society may manipulate less; at the same time, such well-organized opposition may be able to sustain protest movements longer.
+A similar logic might imply that incumbents should manipulate less when economic grievances are high (for fear of sparking protest), and that economic troubles might make protests more durable.
+Either case could produce the relationship expected in H2, but could be considered spurious.
+To reduce this risk, I include several potential controls for opposition strength, and for grievance.
+Controlling for these variables helps isolate the relationship between manipulation severity and protest duration, holding opposition strength and grievances constant.
+
+As such, the results should be considered causally identified only to the extent that unknown or unobservable variables are not confounding the relationship between fraud and protest duration.
+While this problem is difficult to surmount with observational data, this research design offers three defenses.
+First, I control for known possible confounders, including the occurrence of a successful electoral protest in a nearby country (which might frighten incumbents into manipulating less and inspire opponents to hold the line longer).
+Second, I present two analyses in the appendix, which show that increased risk of protest leads incumbents to *intensify* their manipulation efforts, rather than diminish them.
+Thus a negative relationship between fraud and protest duration should indicate the direct hypothesized effect, rather than one that is driven by relatively weaker incumbents avoiding manipulation.
+Third, and relatedly, the research design pits two incompatible hypotheses against one another, rather than a single hypothesis against the null.
+A null relationship between fraud and protest duration might not be strong evidence against the regime-weakness model, but a negative relationship should be (after accounting for necessary controls).
+
 ```{r binarytab, echo=FALSE, message=FALSE, warning=FALSE}
@@
 Regime-weakness models may explain this pattern by arguing that, most of the time, incumbents produce the 'right' amount of manipulation to avoid protest in equilibrium.
-Yet the results of this study challenge that interpretation, since it demonstrates that outside this equilibrium (i.e.
-when elections are rigged and the opposition protests), higher levels of manipulation are associated with greater regime strength, not less.
+Such careful calibration is difficult to ensure, however, when incumbents must rely on thousands of low-level agents to produce the 'right' amount of manipulation; these agents may have collective incentives to shirk and underproduce, or to demonstrate their loyalty by over-producing.
+[ADD CITES] The results of this study add an empirical challenge, since it demonstrates that when the deterrence equilibrium fails (i.e.
+when the incumbent rigs elections and the opposition protests), higher levels of manipulation are associated with greater regime strength, not less.
 
@@
 These other motives may foster protest movements in the aftermath of rigged elections, even when severe electoral manipulation indicates considerable regime strength.
-Second, if manipulation does indeed indicate strength, incumbents have every incentive to build and deploy those strengths against stronger opposition movements (who will themselves be more capable of initiating protest).
+Second, if manipulation does indeed indicate strength, incumbents have every incentive to build and deploy those strengths in reaction to stronger opposition movements (who will themselves be more capable of initiating protest).
+Evidence that incumbents do, indeed, appear to increase manipulation in the face of increased protest threat is presented in the supplementary appendix.
 

### 2025-11-14 Output/appendix_v4.Rmd
Change subsection title to 'defeat' from 'ouster'
Cuts in appendix; begin edits to discussion section in main
Restructuring sections; rewriting fitted vs. residuals section to clarify and focus on addressing endogeneity concern.

@@
 
-# Fraud and the likelihood of protest
+\newpage
 
-Table \@ref(tab:logit-table) shows the results for Hypothesis 1.
-Model 1 is a base model, where incumbent loss, winner vote-share, and *election fraud* are standalone variables.
-In keeping with the general (though inconsistent) findings in the prior literature, there is a significant positive relationship between fraud severity and the likelihood of protest initiation.
-The winner's claimed margin of victory is negatively associated with protest initiation.
-Model 2 includes the three-way interaction between *election fraud*, incumbent loss, and winning vote-share, with no controls other than a dummy variable for *presidential elections*.
-This allows for the largest possible sample size, and shows that the coefficient signs are consistent with those of the full Model 3.
+# Robustness check: Two alternative measures of election manipulation
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
-#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))
+To ensure that the results presented in the main text are not idiosyncratic to the chosen measure of election fraud, this section replaces that measure with two alternative indicators of election integrity.
+The first variable is `v2xel_frefair,' from V-Dem, which is a comprehensive indicator of quality of the electoral process, including measures of electoral management body autonomy and capacity, vote-buying, intimidation, fraud, and an overall assessment of the fairness of the election.
+This is a retroactively coded, expert assessment (as is the variable in the main text), but captures a wider array of possible abuses of electoral integrity.
 
-model.binary_noint <- glm(any_protest ~ inc.lose_clean + v2elirreg.inv + winner.share.upd, data = nelda2.sub, family = binomial(link = "logit"))
+The second measure reflects contemporaneous judgments by election observers, as reported in the Dataset on International Election Monitoring, or DIEM [@kelley2011].
+I use the variable "Explicit cheating," as a parallel to the V-Dem "intentional irregularities" variable from the main text.
+This variable is categorical, and so captures less variation than the original measure.
+However, it helps address the concern that the findings could be systematically biased by the human coders utilized in the V-Dem project [@Little_Meng_2024].
+While there is some evidence that such bias is small [@weidmann2023], it is still useful to check the V-Dem coders' assessments against election monitors' reports.
 
-model.binary.base <-  glm(any_protest ~ winner.share.upd*v2elirreg.inv*inc.lose_clean + presidential, data = nelda2.sub, family = binomial(link = "logit"))
-
-model.binary.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   +    winner.share.upd*v2elirreg.inv*inc.lose_clean   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre +  unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub)
-
-plot_binary_margin_typ_n <- sjPlot::plot_model(model.binary.full, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.08, 2.07]", "inc.lose_clean"))
-plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Winner's vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()
-```
-
-```{r logit-table, echo=FALSE}
-models2 <- list()
-models2[['Model 1']] <- model.binary_noint
-models2[['Model 2']] <- model.binary.base
-models2[['Model 3']] <- model.binary.full
-#models2[['Model 6']] <- model.margin.binary_noc
-
-modelsummary(models2, output = "default",  stars = T, title = 'Table 3: Logit models of protest occurrence', gof_omit = c("Log|RMSE"), coef_map = c("(Intercept)" = "(Intercept)",
-             "inc.lose_clean2" = "Incumbent loss",
-             "v2elirreg.inv" = "Election fraud",
-             "winner.share.upd" = "Winner vote-share",
-"winner.share.upd:v2elirreg.inv" = "Winner vote-share:Election fraud",
-"winner.share.upd:inc.lose_clean2" = "Winner vote-share:Inc. loss",
-"v2elirreg.inv:inc.lose_clean2" = "Election fraud:Inc. loss",
-"winner.share.upd:v2elirreg.inv:inc.lose1" = "Election fraud:Winner vote-share:Inc. loss",
-  "lji.lag" = "Judical independence (lag)",                                                                 "legcon.1lag" = "Leg.
-constraints (lag)",                                                                 "gdpgro.1lag" = "GDP growth rate (lag)",                                                                   "altinf.1lag" = "Alternative info.
-(lag)",
-                 "osorg.1lag" = "Civil soc. openness (lag)",
-                 "inc.termlimit.num" = "Incumbent term-limited", 
-                 "libdem.1lag" = "Liberal democracy index (lag)", 
-                 "urban.pct" = "Urbanization", 
-                 "rentierstate1" = "Rentier state", 
-                 "taxstate1" = "Tax state", 
-                "presidential1" = "Presidential",
-                 "unscheduled_election1" = "Off-schedule election",
-                 "n.events.pre" = "Number of pre-election protests", 
-                 "physinteg.1lag" = "Phys. integrity index (lag)",
-                 "physintlag.sq" = "Phys. int. lag squared", 
-                 "regional_success_lag" = "Regional protest diffusion",
-
-                 "median.participants" = "Median protest participants",
-
-                "v2elirreg.inv:winner.share.upd"= "Election fraud:Winner vote-share")
-                 )
-
-#reg_table |> tab_options(
-#  latex.use_longtable = TRUE, 
-#  data_row.padding = px(5)
-#  ) 
-```
-
-Model 3 is the main model, which includes the full set of controls.
-Given the complexity of the model, Figure \@ref(fig:margin-prob-fig) illustrates the effects of election fraud on protest risk across various levels of winning vote-share and incumbent loss^[All other variables constant at their means (continuous variables) or modes (categorical variables).].
-The values for election fraud represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively.
-
-The direction of the effect supports Hypothesis 1.
-When an incumbent claims victory (left panel) and their vote-share is low, electoral protest is significantly likelier when fraud is elevated than when the election is relatively clean.
-Furthermore, protest against winning incumbents is significantly likelier when fraud is high and vote-shares are low, compared to an election with an equivalent level of fraud but high claimed vote-shares.
-As noted above, such a finding could be consistent with either the regime-weakness or regime-strength models of manipulation, under different conceptualizations of protest.
-To distinguish between the two, we turn to the models of protest duration.
-
-```{r margin-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
-plot_binary_margin_typ_n$data$facet <-  ifelse(plot_binary_margin_typ_n$data$facet == "1", "Incumbent won", "Incumbent lost")
-
-plot_binary_margin_typ_n
-```
-
-# Alternative measures of election manipulation
+Both variables produce the same substantive results in the Cox proportional hazards models of protest duration; more election manipulation is associated with shorter protests.
 
@@
 
-To ensure that the results presented in the main text are not biased by the chosen measure of election fraud, this section replaces that measure with a broader indicator of election integrity.
-This variable is `v2xel_frefair,' from V-Dem, which is a comprehensive indicator of quality of the electoral process, including measures of electoral management body autonomy and capacity, vote-buying, intimidation, fraud, and an overall assessment of the fairness of the election.
-This variable ranges from 0 to 1, with higher values indicating freer and fairer elections.
-The models presented below utilize this variable in place of the main measure of election fraud, while retaining all other control and explanatory variables, using the ECAV data on protest.
-Since the dependent variable of these models is on an inverse scale, compared to the variable *intentional irregularities*, we should expect to see coefficients for key variables take on opposite-signed values from the models in the main text.
+The 'election integrity' variable from V-Dem (v2xel_frefair) ranges from 0 to 1, with higher values indicating freer and fairer elections.
+Note that the sign of this variable is the inverse of the variable used in the main text.
+The models presented below utilize this variable in place of the main measure of election fraud, while retaining all other control and explanatory variables.
 
 The same interaction relationship between overall election integrity, incumbent vote-share, and protest as is shown for fraud in the main text.
-That is, less free-and-fair elections are only associated with a higher protest risk at low levels of incumbent vote-share.
+Table \@ref(tab:cox-models-tab-overall) shows that cleaner elections are associated with a slower end to protest movements.
+When elections are more unfair overall, incumbents are able to more quickly bring protest to an end.
+Figure \@ref(fig:margin-dur-fig-overall) shows that this effect is significant in interaction with low-to-moderate incumbent vote-shares.
 
@@
 
-```{r logit-table2, echo=FALSE}
+```{r logit-table2, echo=FALSE, include = FALSE}
 models2 <- list()
@@
 
-```{r overall-binary, fig.cap="Overall election integrity, winning vote-share, and probability of protest initiation", fig.height=6, echo=FALSE, message=FALSE, fig.align='center'}
+```{r overall-binary, fig.cap="Overall election integrity, winning vote-share, and probability of protest initiation", fig.height=6, echo=FALSE, message=FALSE, fig.align='center', include=FALSE, warning=FALSE}
 sjPlot::plot_model(model.margin.binary.int.full, type = "pred", terms = c("winner.share.upd", "v2xel_frefair [.12, .46]"))  +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election integrity") + theme_bw()
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 
-```{r margin-dur-fig-overall, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
+```{r margin-dur-fig-overall, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H', warning=FALSE}
 p.duration.margin
 ```
-
-Likewise, the models of protest movement survival are also in line with those in the main text.
-Table \@ref(tab:cox-models-tab-overall) shows that cleaner elections are associated with a slower end to protest movements (here, a negative coefficient).
-When elections are more unfair overall, incumbents are able to more quickly bring protest to an end.
-Figure \@ref(fig:margin-dur-fig-overall) shows that this effect is significant in interaction with low-to-moderate incumbent vote-shares.
-
-The two sets of models all replace a measure of *election fraud* with a broader measure of overall election integrity.
-In all cases, the results broadly confirm the incumbent-strength model and are unsupportive of the deterrence model.
 
@@
 
-One obstacle to the validity of the findings could be systematic bias by the human coders utilized in the V-Dem project [@Little_Meng_2024].
-While there is some evidence that such bias is small [@weidmann2023], and election monitors can also be subject to biases [@kelley2012], it may be useful to check the V-Dem coders' assessments against election monitors' reports.
-To do so, I make use of the Dataset on International Election Monitoring [@kelley2011], or DIEM.
 This dataset codes the judgments of international election monitoring missions as categorical variables.
-I use the variable "Explicit cheating" as a parallel to the V-Dem "intentional irregularities" variable.
 The explicit cheating variable is categorical, ranging from the 0 for the cleanest elections to 3 for the most fraudulent.
@@
 
-Table \@ref(tab:diem-table-1) reports a model where the DIEM measure as the explanatory variable in a survival model of protest movements.
-Here the difficulty with missingness becomes apparent, with only 39 observations in the model.
-Given the small number of observations, it is even more reassuring that the positive relationship between fraud and a quick end to protest movements is apparent.
+Table \@ref(tab:diem-table-1) reports two model where the DIEM measure as the explanatory variable in a survival model of protest movements, both as standalone variable and in interaction with the incumbent's vote-share.
+Here the difficulty with missingness becomes apparent, with only 44 observations in the models.
+Figure \@ref(fig:margin-dur-fig-diem) shows that higher 'explicit cheating' has a significant effect on protest duration around the fifty-percent vote-share mark, though the effect is small.
+Given the small number of observations, it is even more reassuring that the expected relationships between fraud and a quick end to protest movements are detected.
 
@@
 
-cox1_diem <- coxph(Surv(d.events.num, surv_status) ~ legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + diem_explicheat_max   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" &  n.events.post > 0))
+cox1_diem <- coxph(Surv(d.events.num, surv_status) ~ legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +  unscheduled_election + regional_success_lag + diem_explicheat_max   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" &  n.events.post > 0))
 #summary(cox1_diem)
+
+cox2_diem <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +  unscheduled_election + regional_success_lag + winner.share.upd*diem_explicheat_max   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" &  n.events.post > 0))
+summary(cox2_diem)
 
@@
 models[['Model 5']] <- cox1_diem
+models[['Model 6']] <- cox2_diem
 
@@
 lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity", "diem_explicheat_max" = "Explicit cheating (DIEM)", "n.events.post" = "N.
-post-election protests"
+post-election protests", "winner.share.upd" = "Incumbent vote-share"
                  )) %>% kable_styling(latex_options = "HOLD_position")
@@
 
-# Further tests of deterrence logic: predicted and residual levels of fraud
+```{r margin-dur-fig-diem, fig.cap="Fraud, winning vote-share, and duration of protest (DIEM)", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H', warning=FALSE}
+p.duration.margin_diem <- sjPlot::plot_model(cox2_diem, type = "pred", terms = c("winner.share.upd", "diem_explicheat_max [0, 3]"))  #This is the 10th and 90th quantiles
+p.duration.margin_diem <- p.duration.margin_diem +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Explicit cheating (DIEM)") + theme_bw() + coord_cartesian(ylim = c(0, 20))
+```
+
+# Addressing concerns about selection effects
+
+## Predicted and residual levels of fraud
 
 The main text focuses on elections where the incumbent claims victory.
-However, classic deterrence models also make predictions for incumbents who will voluntarily avoid electoral manipulation and accept a loss at the ballot box.
-Fully investigating this possibility should be the focus of future work, but some initial analyses are provided here in support of the claims made in the main paper.
-First, using all the observations in the dataset\textemdash winning and losing incumbents\textemdash to find the predicted level of election fraud for each case, using only structural variables (i.e.
-professionalization of the bureaucracy, electoral management body autonomy, urbanization, etc.).
+However, classic regime-weakness models also predict that some incumbents who will voluntarily avoid electoral manipulation and accept a loss at the ballot box.
+If this is the case, it could introduce selection bias into the sample for the duration analyses in the main paper by sorting only the strongest incumbents into the sample that faces protest.
+
+Fully investigating this possibility should be the focus of future work, but some initial analyses are provided here to show that this is not the case.
+The analysis in this subsection proceeds in three steps.
+First, using all the observations in the dataset\textemdash winning and losing incumbents\textemdash I estimate the predicted level of election fraud for each case, using only structural variables (i.e.
+professionalization of the bureaucracy, electoral management body autonomy, urbanization, etc) in an OLS model.
 Factors associated with deterrence, like the number of pre-election protests are excluded.
-The fitted values and residuals from this first model are then used to check the implications of the theory in subsequent analysis.
-This model is simple OLS regression.
+This should produce an estimate of the level of manipulation the incumbent could produce in the absence of strategic deterrence.
 
-Model 7 takes the residuals from the structural model as a dependent variable.
-This model is used to check whether pre-election protest risk is associated with incumbents over- or under-performing their expected level of manipulation based on structural factors.
-The deterrence model, which the main paper argues against, suggests that when the risk of anti-incumbent mobilization is higher, incumbents should be more likely to restrain themselves and under-deliver on manipulation.
-Since the residuals which are used as a DV are continuous, this is also an OLS model.
+Second, the residuals from this first model are then used as a dependent variable in second model.
+The residuals reflect the degree to which an incumbent over- or under-produced manipulation, relative to their expected baseline.
+If the deterrence interpretation is correct (which may lead to selection effects), the indicators of protest risk should be negatively associated with these residuals.
+In other words, higher protest risk should lead incumbents to under-produce manipulation.
+Under the regime-strength logic, we might expect the opposite, as incumbents work hard to build and indicate strength in riskier environments.
+This model is also an OLS model.
 
-The theory proposed in the main text implies that structural factors enabling manipulation should be significant *negative* predictors of incumbent loss, while measures of pre-election risk should not.
-That is, the theory expects that losing incumbents are more likely to structurally limited in the level of manipulation they can generate, rather than being structurally unconstrained but deterred by a more robust opposition.
-Model 8 in Table \@ref(tab:fraud-loss) takes incumbent defeat in the election as the dependent variable, with the expected value of manipulation (derived from the first model) as an explanatory variable.
-Pre-election protest and successful protests in the region are included as controls.
-Due to the binary dependent variable, this is a logit model.
+Lastly, two logit models of incumbent loss are presented, using the fitted values from the first model and indicators of protest risk as predictors.
+The fitted values indicated the expected, structural level of electoral manipulation.
+Here, the deterrence logic would expect that indicators of pre-election protest risk will make incumbents more likely to lose, while the regime-strength model would expect that incumbent loss will be largely driven by structural factors.
+
+Table \@ref(tab:fraud-loss) shows the results.
+Model 6 shows how structural factors are predictors of the level of observed manipulation.
+We observe some expected results; for instance, more autonomous EMBs and more professionalized bureaucracies are associated with lower fraud.
+Model 7 shows that indicators of protest risk\textemdash the number of pre-election protests, and the existence of a recent successful protest movement in the region\textemdash are not associated with lower fraud.
+In fact, just the opposite: incumbents appear to manipulate more heavily under riskier conditions.
+This is a key finding, and is evidence against the classic deterrent approach; it also suggests that weaker governments are not voluntarily dialing back manipulation to avoid protest, mitigating concerns about selection effects in the main models.
+
+Models 8 and 9 take the fitted values from the first model and employ them as an explanatory variable for incumbent loss, while controlling for pre-election protest.
+As both models show, the more incumbents invest in the structural prerequisites of manipulation, the less likely they are to lose elections.
+By contrast, we observe little or no effect for pre-election protest risk.
+Model 8 shows no effect for regional protest diffusion on incumbent loss, while Model 9 shows a significant (p = 0.05) but substantively small effect for pre-election protest.
+
+All told, this analysis supports the argument in the main text, which holds that the capacity to manipulate heavily is associated with incumbent durability.
+The results in this section are consistent with a world in which losing incumbents are not voluntarily surrendering power in the face of a committed opposition (on average), but are failing to produce sufficient election manipulation to win due to structural constraints.
+This should reduce the concern that the main findings are driven by an endogenous process, in which incumbents who face more organized and capable opposition actors are strategically avoiding protest by producing less fraud.
 
@@
 
-model.lose <- glm(inc.lose ~ .fitted + regional_success_lag + n.events.pre, data = nelda2.sub.fraudpred, family = binomial(link="logit"))
-summary(model.lose)
+model.lose_reg <- glm(inc.lose ~ .fitted + regional_success_lag , data = nelda2.sub.fraudpred, family = binomial(link="logit"))
+summary(model.lose_reg)
+
+model.lose_events <- glm(inc.lose ~ .fitted + n.events.pre , data = nelda2.sub.fraudpred, family = binomial(link="logit"))
+summary(model.lose_events)
 
@@
 models2[['Model 7']] <- model.fraud.resid
-models2[['Model 8']] <- model.lose
+models2[['Model 8']] <- model.lose_reg
+models2[['Model 9']] <- model.lose_events
 
-row <- data.frame(cbind("", "Fraud", "Fraud residual", "Inc. loss"))
+row <- data.frame(cbind("", "Fraud", "Fraud residual", "Inc. loss", "Inc. loss"))
 attr(row, 'position') <- c(0)
@@
 
-Table \@ref(tab:fraud-loss) shows the results.
-Model 6 shows how structural factors are predictors of the level of observed manipulation.
-We observe some expected results; for instance, more autonomous EMBs and more professionalized bureacuracies are associated with lower fraud.
-Model 7 shows that indicators of protest risk\textemdash the number of pre-election protests, and the existence of a recent successful protest movement in the region\textemdash are not associated with lower fraud.
-In fact, just the opposite: incumbents appear to manipulate more heavily under riskier conditions.
-This is initial evidence against the classic deterrent approach.
-Model 8 takes the fitted values from the first model and employs them as an explanatory variable for incumbent loss, while controlling for pre-election protest.
-As the latter model shows, the predicted level of fraud based on the incumbent's structural advantages is negatively associated with incumbent loss.
-This is supportive of the argument in the main text, which holds that the capacity to manipulate heavily is associated with incumbent durability.
-Altogether, the results in this section are consistent with a world in which losing incumbents are not voluntarily surrendering power in the face of a committed opposition (on average), but are failing to produce sufficient election manipulation to win due to structural constraints.
-
 \newpage
 
-# Incumbent ousted by protest
+# Incumbent defeat by protest
 
@@
 ```
+
+# Fraud and the likelihood of protest
+
+Table \@ref(tab:logit-table) shows the results for Hypothesis 1.
+Model 1 is a base model, where incumbent loss, winner vote-share, and *election fraud* are standalone variables.
+In keeping with the general (though inconsistent) findings in the prior literature, there is a significant positive relationship between fraud severity and the likelihood of protest initiation.
+The winner's claimed margin of victory is negatively associated with protest initiation.
+Model 2 includes the three-way interaction between *election fraud*, incumbent loss, and winning vote-share, with no controls other than a dummy variable for *presidential elections*.
+This allows for the largest possible sample size, and shows that the coefficient signs are consistent with those of the full Model 3.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
+#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))
+
+model.binary_noint <- glm(any_protest ~ inc.lose_clean + v2elirreg.inv + winner.share.upd, data = nelda2.sub, family = binomial(link = "logit"))
+
+model.binary.base <-  glm(any_protest ~ winner.share.upd*v2elirreg.inv*inc.lose_clean + presidential, data = nelda2.sub, family = binomial(link = "logit"))
+
+model.binary.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   +    winner.share.upd*v2elirreg.inv*inc.lose_clean   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre +  unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub)
+
+plot_binary_margin_typ_n <- sjPlot::plot_model(model.binary.full, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.08, 2.07]", "inc.lose_clean"))
+plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Winner's vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()
+```
+
+```{r logit-table, echo=FALSE}
+models2 <- list()
+models2[['Model 1']] <- model.binary_noint
+models2[['Model 2']] <- model.binary.base
+models2[['Model 3']] <- model.binary.full
+#models2[['Model 6']] <- model.margin.binary_noc
+
+modelsummary(models2, output = "default",  stars = T, title = 'Table 3: Logit models of protest occurrence', gof_omit = c("Log|RMSE"), coef_map = c("(Intercept)" = "(Intercept)",
+             "inc.lose_clean2" = "Incumbent loss",
+             "v2elirreg.inv" = "Election fraud",
+             "winner.share.upd" = "Winner vote-share",
+"winner.share.upd:v2elirreg.inv" = "Winner vote-share:Election fraud",
+"winner.share.upd:inc.lose_clean2" = "Winner vote-share:Inc. loss",
+"v2elirreg.inv:inc.lose_clean2" = "Election fraud:Inc. loss",
+"winner.share.upd:v2elirreg.inv:inc.lose1" = "Election fraud:Winner vote-share:Inc. loss",
+  "lji.lag" = "Judical independence (lag)",                                                                 "legcon.1lag" = "Leg.
+constraints (lag)",                                                                 "gdpgro.1lag" = "GDP growth rate (lag)",                                                                   "altinf.1lag" = "Alternative info.
+(lag)",
+                 "osorg.1lag" = "Civil soc. openness (lag)",
+                 "inc.termlimit.num" = "Incumbent term-limited", 
+                 "libdem.1lag" = "Liberal democracy index (lag)", 
+                 "urban.pct" = "Urbanization", 
+                 "rentierstate1" = "Rentier state", 
+                 "taxstate1" = "Tax state", 
+                "presidential1" = "Presidential",
+                 "unscheduled_election1" = "Off-schedule election",
+                 "n.events.pre" = "Number of pre-election protests", 
+                 "physinteg.1lag" = "Phys. integrity index (lag)",
+                 "physintlag.sq" = "Phys. int. lag squared", 
+                 "regional_success_lag" = "Regional protest diffusion",
+
+                 "median.participants" = "Median protest participants",
+
+                "v2elirreg.inv:winner.share.upd"= "Election fraud:Winner vote-share")
+                 )
+
+#reg_table |> tab_options(
+#  latex.use_longtable = TRUE, 
+#  data_row.padding = px(5)
+#  ) 
+```
+
+Model 3 is the main model, which includes the full set of controls.
+Given the complexity of the model, Figure \@ref(fig:margin-prob-fig) illustrates the effects of election fraud on protest risk across various levels of winning vote-share and incumbent loss^[All other variables constant at their means (continuous variables) or modes (categorical variables).].
+The values for election fraud represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively.
+
+The direction of the effect supports Hypothesis 1.
+When an incumbent claims victory (left panel) and their vote-share is low, electoral protest is significantly likelier when fraud is elevated than when the election is relatively clean.
+Furthermore, protest against winning incumbents is significantly likelier when fraud is high and vote-shares are low, compared to an election with an equivalent level of fraud but high claimed vote-shares.
+As noted above, such a finding could be consistent with either the regime-weakness or regime-strength models of manipulation, under different conceptualizations of protest.
+To distinguish between the two, we turn to the models of protest duration.
+
+```{r margin-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
+plot_binary_margin_typ_n$data$facet <-  ifelse(plot_binary_margin_typ_n$data$facet == "1", "Incumbent won", "Incumbent lost")
+
+plot_binary_margin_typ_n
+```

### 2025-11-15 Output/appendix_v4.Rmd
Edits to section on multicollinearity and on incumbent defeat by protest

@@
 
-# Incumbent defeat by protest
+## Incumbent defeat by protest
 
-The classic model of election manipulation as an indicator of weakness would suggest that incumbents who rely more on manipulation will be more likely to fall when protest movements do emerge.
-The main text does not examine successful protests, which are a small minority of cases, in order to avoid measurement error in the duration models.
-Successful protest is, instead, examined here; there is little to suggest that incumbents who manipulate more are more vulnerable to defeat by protesters.
+A second possible source of selection bias in the main analysis sample is the decision to exclude cases where incumbents were defeated by protest movements.
+While this decision is necessary given the research design, it might still introduce bias if---for instance---many defeated incumbents manipulated heavily.
+The main text does not examine successful protests, which are a small minority of cases, but I include them in an analysis here.
+There is little to suggest that incumbents who manipulate more are more vulnerable to defeat by protesters.
 
@@
 Figure \@ref(fig:oust-plot) shows no significant effect of election manipulation on protest success across all values of incumbent vote-share.
-Taken together, this is further disconfirmation of the fraud-as-weakness model.
+Taken together, this is further evidence against the fraud-as-weakness model, and suggests the main results are not biased by excluding high-fraud cases where protesters toppled the government.
 
@@
 
-model.oust3 <- glm(protest.success ~  osorg.1lag + gdpgro.1lag  + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +   unscheduled_election + taxstate + rentierstate + regional_success_lag + v2elirreg.inv , family = binomial(link = "logit"), data = protest.cases %>% filter(inc.lose_clean == "1"))
+model.oust3 <- glm(protest.success ~  osorg.1lag + gdpgro.1lag  + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +   unscheduled_election +  regional_success_lag + v2elirreg.inv , family = binomial(link = "logit"), data = protest.cases %>% filter(inc.lose_clean == "1"))
 summary(model.oust3)
 
-model.oust4 <- glm(protest.success ~  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +   unscheduled_election + taxstate + rentierstate + regional_success_lag +  v2elirreg.inv*winner.share.upd , family = binomial(link = "logit"),  data = protest.cases %>% filter(inc.lose_clean == "1"))
+model.oust4 <- glm(protest.success ~  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +   unscheduled_election +  regional_success_lag +  v2elirreg.inv*winner.share.upd , family = binomial(link = "logit"),  data = protest.cases %>% filter(inc.lose_clean == "1"))
 summary(model.oust4)
@@
 
-# Chosen successors
+<!-- # Chosen successors
 
@@
 
+-->
+
 # Removing controls to address multicollinearity
@@
 
-```{r include=FALSE, echo=FALSE}
+```{r include=FALSE, echo=FALSE, include=FALSE}
 nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
@@
 
-```{r logit-table-small, echo=FALSE}
+```{r logit-table-small, echo=FALSE, include=FALSE}
 models2 <- list()
@@
 
-```{r margin-prob-fig-small, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
+```{r margin-prob-fig-small, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H', include=FALSE}
 plot_binary_margin_typ_n
@@
 
-```{r, include=FALSE, echo=FALSE}
+```{r, message=FALSE, echo=FALSE, warning=FALSE}
 library(survival)
@@
 cox2 <- coxph(Surv(d.events.num, surv_status) ~  v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +   unscheduled_election + regional_success_lag +  n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
-#summary(cox2)
+summary(cox2)
 p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.05, 2]"))  #This is the 10th and 90th quantiles
@@
 
+<!--
 # Using country fixed effects
@@
 ```
+
+-->
 

### 2025-11-16 Output/authoritarian_only_pdf_pop_rev.Rmd
Final edits to intro

@@
 Events like these figure prominently in theories of election manipulation and how it might be restrained, with broader implications for the stability of both democratic and authoritarian regimes.
-A classic view, influential in the formal modeling tradition, argues that the use of election manipulation by an incumbent is indicative of the incumbent's unpopularity.
-This lack of popular support, in turn, makes it less likely that an incumbent will be able to defeat a protest movement if the opposition chooses to launch one.
-Such models predict that there is a set of incumbents who require manipulation to reliably win an election, but\textemdash fearing costly protests\textemdash choose instead to gamble on relatively clean elections.
+A classic view, influential in the formal modeling tradition, argues that the use of election manipulation by an incumbent reveals their unpopularity, and thus that they are less likely to survive a post-election challenge.
+Knowing this, opposition movements are expected to be increasingly to initiate electoral protest against incumbents as election manipulation increases (all else equal).
+Such models predict that some incumbents, who require manipulation to win an election with high probability, choose instead to gamble on relatively clean elections out of fear of costly protest.
 This framework has normatively optimistic implications, by showing that stronger opposition movements\textemdash those that are more organized, better resourced, and more popular\textemdash will be better able to deter incumbents' from holding rigged elections.
 In turn, this dynamic makes democratic transition and consolidation more likely.
-I refer to this normatively optimistic view as the regime-weakness model of manipulation.
+Because fraud indicates incumbent weakness in this model, I refer to it as the regime-weakness model of electoral manipulation.
 
@@
 The stakes of this debate are significant for theories of democratization and autocratization.
-The former view is an optimistic one; in it, incumbents may restrain fraud when it might be exposed and reveal their weakness to strategic protesters\textemdash an auspicious prediction for democratization and ultimately the under-pinning of 'self-enforcing' democracy [@fearon_self-enforcing_2011].
+The regime-weakness view makes auspicious predictions for democratization and ultimately under-pins the notion of 'self-enforcing' democracy [@fearon_self-enforcing_2011].
 If the latter view holds, the risk of electoral protest may fail to deter election manipulation; it would, in this case, be a fragile guardrail against autocratization.
-This debate cannot be advanced by focusing solely on protest occurrence.
-Instead, this paper contributes by drawing out implications from three strands of the literature on election manipulation and protest, which have (to my knowledge) not previously been explicated.
+In addition, a sizable literature has been founded on the regime-weakness framework.
+Scholars working in this mode have argued that manipulation can be kept in check, and ultimately reduced, through the development of institutions that reveal information about election integrity: election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].
 
@@
 A low total vote-share amid severe manipulation is the strongest possible signal of incumbent weakness; it has had to rely on the 'cheap talk' of manipulation to produce paltry results, indicating that the incumbent's genuine popularity is very low and its likelihood of defeating the protest movement is likewise small.
-By contrast, the regime-strength model would yield the opposite prediction [@simpser2013]; that protest initiation decreases with the level of manipulation (holding vote-share constant).
-Under the regime-strength model, the most favorable  election outcome for protesters is one where the incumbent's vote-share and manipulation level are low, a combination which indicate low natural popularity and diminished resource capacity.
-
-There are two problems, however.
-First, the empirical record on the relationship between manipulation and protest been mixed, illustrating the limitations of current theory [@Hyde2014;@daxecker2012;@Hafner-Burton2016;@Lankina2017;@ananyevDictatorsSignalStrength2022].
-Second, the expected relationship between manipulation and protest under the regime-weakness and regime-strength models is complicated by the emergence of a grievance-based model of electoral protest generally.
-This view, which has considerable empirical support, finds that electoral protests are largely driven by citizen grievances\textemdash not the narrowly strategic calculations about the likelihood of overthrowing the government which are assumed in the regime-weakness model.
-Under the grievance framework, when citizens have many individual grievances against the incumbent, a highly manipulated election can serve as a coordination point for collective action [@tucker2007].
-
-This theoretical complexity has frustrated our ability to distinguish between the regime-strength and regime-weakness models because they can produce the same hypotheses regarding protest initiation depending on which model of protest is true.
-If protest movements are narrowly strategic and manipulation indicates weakness, we will expect that the likelihood of protest will increase with electoral manipulation.
-On the other hand, if manipulation indicates strength we will still expect greater manipulation to make protest more likely\textemdash if electoral protests are largely driven by shared grievance rather than strategic considerations.
-Since the same outcome could support two theoretical models, we are at an impasse.
-
-To move the debate forward, I model electoral protest *duration* as a dependent variable, in addition to initiation.
-Taken together, the two dependent variables offer distinct predictions under the regime-weakness and regime-strength frameworks.
+By contrast, it might be presumed that the regime-strength model would yield the opposite prediction [@simpser2013]; that protest initiation decreases with the level of manipulation (holding vote-share constant).
+But this expectation only holds if we adopt a limiting assumption on the nature of protest from the regime-weakness model.
+
+The predictions of the regime-weakness model hinge on the assumption that protest initiation is narrowly instrumental---driven by the expected benefit of overthrowing the incumbent.
+However, this assumption abstracts away the findings of a long literature on the motivations of protest in non-democracies (including electoral protest), which finds that protests are complex and multi-purpose, driven by diverse actors' goals and grievances.
+Under this more expansive view of protest, a highly manipulated election can serve as a coordination device for aggrieved citizens when incumbent popularity is low [@tucker2007]; the same scenario where the regime-weakness model predicts protests to be most likely, though for different reasons.
+Moreover, the empirical record on the relationship between manipulation and protest been mixed, illustrating the limitations of current theory [@Hyde2014;@daxecker2012;@Hafner-Burton2016;@Lankina2017;@ananyevDictatorsSignalStrength2022].
+
+To move the debate forward, I model electoral protest *duration* as a dependent variable, which offers distinct predictions under the regime-weakness and regime-strength frameworks.
 Once protests run their course, the 'strength' of an incumbent (its likelihood of outlasting a protest movement) is no longer latent.
@@
 The more quickly a government puts down protests, the greater the underlying (unobserved) distribution of power favors the incumbent over the opposition.
-By contrast, the longer it takes the government to restore control, the less favorable the distribution of power.
-
-Using data from 659 non-democratic elections across 100 countries, from 1989 to 2012, this paper finds that 1) protest initiation is most likely when manipulation is severe but incumbent vote-share is low, and 2) that governments that engage in more manipulation are likely to put down protests more quickly than governments that produce little manipulation, especially when incumbents claim small vote-shares.
-This pattern of results in inconsistent with the regime-weakness model of election manipulation, but consistent with a world in which election manipulation indicates sources of regime resilience and protests are largely motivated by grievance.
+By contrast, the longer it takes the government to restore control, the less favorable the distribution of power for the incumbent.
+
+Using data from 659 non-democratic elections across 100 countries, from 1989 to 2012 (of which 156 experienced protest), this paper finds that governments that manipulate elections more heavily are likely to put down protests more quickly, especially when incumbents claim small vote-shares.
+This is inconsistent with the regime-weakness model of election manipulation, but consistent with a world in which election manipulation indicates sources of regime resilience.
 Given these results, it is likely that---while incumbents do not seek to provoke protest---the downside risk for incumbents that engage in election fraud in non-democracies has been overemphasized.
@@
 Finally, they suggest a pessimistic outlook for democratization by election in authoritarian regimes.
-
-These are normatively optimistic models upon which a sizable literature has been founded: manipulation can be kept in check, and ultimately reduced, through the development of institutions that reveal information about election integrity: election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].
-These institutions improve election integrity, it is claimed, by increasing the likelihood that fraud will be observed and thus, incumbent weakness exposed.
 
@@
 These distributions---extensive election fraud combined with large numbers of non-protest---provides some early indication that the regime-weakness model has limited support.
-Still, approximately one-third of elections experienced some protest, and so protest initiation is modeled using standard logistic regression.
+Still, approximately one-third of elections experienced some protest.
 Protest wave duration is modeled using Cox proportional hazards models, which have the following formula:
 
-$$h(t∣X)=h_0(t)* exp(β_fX_f+β_vX_v+⋯+βpXp)$$
+$$h(t | X)=h_0(t)* exp(\beta_fX_f+\beta_vX_v+ ... +\beta_pXp_)$$
 The formula models the risk that a protest movement will fail at time $t$ given covariates $X$ is a function of the baseline hazard function, the degree of electoral fraud ($X_f$), the incumbent's vote-share ($X_v$) and a set of control variables.
@@
 
-most regimes are operating at an equilibrium, wherein just the right amount of manipulation is produced that avoids protest.
-that  and---as this study shows---when electoral protest does occur, such movements are defeated in the large majority of cases.
-If manipulation signals weakness, why are more rigged elections not protested, and why are more protest movements not successful?
-The regime-weakness model proposes answers for each of these questions.
-
-While Simpser's original argument holds that electoral manipulation in conjunction with a large margin of victory for the incumbent can induce pro-regime behavior in society, this paper shows that even when such cooperation fails---when post-election protests occur---governments that manipulate more are better positioned to quickly defeat challengers.
-
 It also helps us make sense of secular patterns that are puzzling if election fraud indicates regime weakness, as others have held.
@@
 Scenarios where incumbents use massive fraud to post margins of victory in the eighty and ninety percentage point range are harder to explain under a worldview where more manipulation indicates deeper regime weakness.
-
-This study finds support for Hypothesis 1---that protest will be most likely to begin when fraud is high and incumbent vote-shares low---*and* for the notion that governments that manipulate more effectively will more quickly quash protests.
-In other words, protest is most likely to be initiated under the same circumstances when it is likely to be defeated quickly.
-This combination of findings is a blow to the regime-weakness model, and the associated idea that protest risk can effectively deter election manipulation.
-I suggests, instead, that 1) large-scale election manipulation indicates regime strength in most cases, and 2) electoral protest behavior involves a mix of actors with broadly strategic motives that go beyond ousting the government.
-Such motives are discussed in more detail in the theory section.
 
@@
 First, these results may inform how researchers think about the causes of election manipulation.
-They suggest that factors like the opposition's capacity to organize street protest in limiting election fraud have been overestimated.
-Consequently, theories of electoral manipulation that emphasize other factors---such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]---should be further developed, and may yield new policy interventions.
+They suggest that the degree to which the opposition's capacity to organize street protest (and related factors) can limit election fraud have been overestimated.
+Consequently, theories of electoral manipulation that emphasize alternatives---such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]---should be further developed, and may yield new policy interventions.
 In particular, it may change some current understandings of the role of election monitors, courts, and other third parties in upholding election integrity.
@@
 If the capacity to engage in widespread manipulation makes it easier to both tolerate the existence of an opposition and to suppress it during periods of contestation [@dahlPolyarchyParticipationOpposition1971], electoral authoritarianism is something of an institutional cul-de-sac: the capacity to generate severe election fraud makes it appealing for closed authoritarian regimes to introduce elections, while the manageable risk of protest makes it unappealing to further democratize.
-
-One key prediction of the strategic protest model is not tested here directly, because the research design focuses on the strength of incumbents against protest movements as revealed by actual protest.
-However, several additional tests are presented in the appendix which suggest that the limits on election manipulation are largely structural (e.g.
-state capacity, judicial independence) rather than narrowly strategic.
-In fact, these additional checks indicate that, if anything, incumbents tend to respond to elevated protest risk by *increasing* production of electoral manipulation.
 
@@
 
+In sum, the results indicate that we should be cautious when building and interpreting theories that emphasize the risk of mass protest as a trip-wire to be avoided in non-democracies, and add empirical evidence to Przeworski's [-@przeworskiFormalModelsAuthoritarian2022] critique of such models.
+The rarity of post-election protest, especially successful protest, suggests that such events are highly contingent; they hinge on a multitude of factors that go beyond the regime's efforts to bias election outcomes.
+These findings indicate that, while incumbents do seek to contain protest, more effective manipulators are also more capable of defeating protest movements.
+
 These results are consistent with different specifications presented in an online appendix.
@@
 Additional robustness checks include models with data from election observer reports, with country fixed-effects, alternative specifications for selected control variables.
-
-In sum, the results indicate that we should be cautious when building and interpreting theories that emphasize the risk of mass protest as a trip-wire to be avoided in non-democracies, and add empirical evidence to Przeworski's [-@przeworskiFormalModelsAuthoritarian2022] critique of such models.
-The rarity of post-election protest, especially successful protest, suggests that such events are highly contingent; they hinge on a multitude of factors that go beyond the regime's efforts to bias election outcomes.
-These findings indicate that, while incumbents do seek to contain protest, more effective manipulators are also more capable of defeating protest movements.
 

### 2025-11-16 Output/authoritarian_only_pdf_pop_rev.Rmd
Near-final edits to theory. Will still need to fix table and its cross references

@@
 This literature has grappled with the extent to which election manipulation reflects, and reveals, the incumbent's strength.
-Various authors have put forth different modalities of incumbent strength, with some focusing on popularity and others on coercive power.
+Various authors have put forth different conceptions of incumbent strength, with some focusing on popularity and others on coercive power.
 This paper takes a general approach, following Leventoğlu et al [-@leventogluSynthesizingTheoriesAuthoritarian2025]: a stronger incumbent is one who is better able to survive in power throughout the whole electoral cycle.
 This means the incumbent has both prevailed in the election itself, and managed to hold off any subsequent popular revolts or elite challenges [@svolik_politics_2012].
-As Leventoğlu et al [-@leventogluSynthesizingTheoriesAuthoritarian2025] discuss in their excellent analysis, genuine popularity is but one potential source of incumbent strength.
+As Leventoğlu et al [-@leventogluSynthesizingTheoriesAuthoritarian2025] discuss in their analysis, genuine popularity is but one potential source of incumbent strength.
 Popularity certainly helps an incumbent win an election and survive any challenges in its aftermath, but it is only one tool; incumbents may also draw on resource wealth, organizational capacity, support among influential groups, loyal military and security services, or help from foreign patrons.
@@
 These assumptions face several challenges, which are discussed below.
-First, however, it is necessary to clarify the meaning of incumbent strength.
 
@@
 
-The characteristic assumption in regime-weakness models is that electoral manipulation communicates nothing about incumbent strength, which is closely identified with popularity.
-Since more manufactured votes indicate deeper unpopularity, the opposition perceives election fraud as a reliable signal of incumbent weakness, and will stage protests in order to push out the weak incumbent.
+The characteristic assumption in regime-weakness models is that electoral manipulation communicates nothing positive about incumbent strength, which is closely identified with popularity.
+At any given vote-share, more manufactured votes indicate deeper unpopularity; the opposition thus perceives election fraud as a signal of incumbent weakness, and will stage protests in order to push out a sufficiently weak incumbent.
 For example, the model developed by Little, Tucker, and LaGatta [-@Little2015] predicts that "Just as stronger election results for the incumbent lead to less protest because they indicate that she is popular, reports of more fraud lead to more protest because they indicate that the incumbent is, for a given election result, less popular" (p.
@@
 In equilibrium, the incumbent will generally choose the level of fraud anticipated by the opposition and no more---but also no less.
-In sum, regime-weakness models of fraud typically hold that the level of election manipulation carries no signal of the incumbent's ability to defeat protesters; on the contrary, it only carries signals of weakness.
+In sum, regime-weakness models of fraud typically hold that the level of election manipulation conveys no positive information about the incumbent's ability to defeat protesters.
 
@@
 
-There are considerable theoretical and substantive stakes in knowing which of these four possible data-generating processes is most likely; in only one does increasing protest risk actually improve electoral integrity.
-Since the empirical record is mixed, and three of the four possible models make similar predictions with regard to protest initiation, it is necessary to generate new hypotheses.
+There are considerable theoretical and substantive stakes in knowing which of these  data-generating processes is most likely.
+Since the empirical record is mixed, and the models make similar predictions with regard to protest initiation, it is necessary to generate new hypotheses.
 In particular, the remainder of this manuscript will attempt to distinguish between regime-strength and regime-weakness models of manipulation by developing hypotheses about protest *duration* under each framework.
@@
 Considering how electoral manipulation may be tied to protest duration under the regime-weakness and regime-strength models reveals distinct predictions.
-Specifically, as discussed below, protest duration should decrease with manipulation in the regime-weakness model, but increase in the regime-strength model (conditional on incumbent vote-share).
-The logic discussed below hinges on the assumption that, on average, authoritarian governments will seek to quash protest movements quickly once they appear.
+Specifically, when manipulation increases, protest movements should last longer under the regime-weakness model, but be defeated more quickly under the regime-strength model (conditional on incumbent vote-share).^[In both cases, this assume that incumbent survives the protest movement, by far the modal outcome.
+When the opposition wins, an outcome considered in the appendix, the assumption would be reversed: a quicker opposition victory indicates a weaker incumbent.] The logic hinges on the assumption that, on average, authoritarian governments will seek to quash protest movements quickly once they appear.
 How plausible is this assumption?
 
-Mass protest is one of the two primary threats to authoritarian regimes [@svolik_politics_2012], and as such they often go to great lengths to prevent mass protest.
+Mass protest is one of the two primary threats to authoritarian regimes [@svolik_politics_2012], and as such they often go to great lengths to prevent it.
 These efforts include costly (and risky) investments in the repressive apparatus [@Escriba_Folch2013;@draguPreventiveRepressionTwo2019], co-optation of opposition parties [@reuter2015], and large-scale monitoring and censorship schemes [@King2013].
@@
 Mass protest then, is clearly undesirable from the authoritarian incumbent's point of view, and we can plausibly assume that once an electoral protest movement begins, the incumbent will exert considerable effort to quash it.
-Governments that are stronger, relative to the protesting opposition, will defeat protest movements faster than will weaker governments.^[In both cases, this assume that incumbent survives the protest movement, by far the modal outcome.
-When the opposition wins, an outcome considered in some of the empirical models below, the assumption is reversed: a quicker opposition victory indicates a weaker incumbent.] As argued below, electoral manipulation can be demonstrative of capacities that may also be used to defeat protest movements.
+Governments that are stronger, relative to the protesting opposition, will defeat protest movements faster than will weaker governments.
+As argued below, electoral manipulation can be demonstrative of capacities that may also be used to defeat protest movements.
 
@@
 It is true that incumbents do sometimes tolerate protest movements [@yuenNeitherRepressionConcession2017].
-However, there are two reasons this is unlikely to be the case systematically for the circumstances studied here.
+However, there are two reasons this is unlikely to be the case systematically here.
 First, authoritarian governments are more likely to tolerate protests aimed at low-salience, local issues [@lorentzenRegularizingRiotingPermitting2013;@kleinDynamicsPoliticalProtests2018], and more likely to clamp down on protests that have higher concession costs [@kleinDynamicsPoliticalProtests2018].
@@
 In such cases, we should expect to see apparent 'toleration' displaced by 'attrition'---efforts by the incumbent to increase the cost of protest while avoiding open repression [@yuenNeitherRepressionConcession2017].
-
-In other words, even in the absence of explicit repression or concessions, we should still expect incumbents to attempt to defuse protest movements aimed at challenging the most recent election.
-As a further control, however, I include a measure of average protest size in one model of protest duration.
+As a control, however, I include a measure of average protest size in one model of protest duration, in the expectation that larger protests are harder to tolerate.
 
@@
 However, I depart from Simpser in two ways.
-First, by incorporating more recent research on the principal-agent dynamics inherent in election manipulation, I argue that incumbents who can produce large amounts of election manipulation necessarily have "capacities, resources and dispositions" helpful for defeating protests.
+First, by incorporating more recent research on the principal-agent dynamics inherent in election manipulation, I argue that incumbents who can produce large amounts of election manipulation necessarily have "capacities, resources and dispositions" helpful for defeating protests---a peripheral subject for Simpser.
 Second, mine is not a signalling argument; good manipulators are likely to have deeper wells of strength than weaker manipulators, whether those resources are successfully signaled to opposition actors or no.
@@
 
-Still, a full accounting of a regime's strength would include the incumbent's genuine popularity [@Gerschewski2013].
+Still, a full accounting of a regime's strength would include the incumbent's "control-based strength" *and* their genuine popularity [@Gerschewski2013].
 Both models of election manipulation consider genuine popularity to be a source of regime strength, and assume that increasing electoral manipulation either distorts or weakens this indicator by mingling genuine votes with manufactured ones.
@@
 
-Testing H1 and H2 together will help to evaluate the regime-weakness model of electoral manipulation.
 It is important to note that the theories make contradictory predictions.
 H1 predicts an unconditional *positive* relationship between manipulation and protest duration, while H2 predicts a *negative* relationship, conditional on vote-share.
-
-[Note that the type of protest that happens (instrumental vs. multipurpose) will be examined in the discussion / appendix. Consider also coming up with names for 'manipulative strength']
 

### 2025-11-16 Output/authoritarian_only_pdf_pop_rev.Rmd
Final edits to variable descriptions

@@
 
-Using data from 659 non-democratic elections across 100 countries, from 1989 to 2012 (of which 156 experienced protest), this paper finds that governments that manipulate elections more heavily are likely to put down protests more quickly, especially when incumbents claim small vote-shares.
+Using data from 659 non-democratic elections across 100 countries, from 1989 to 2012 (of which 211 experienced protest), this paper finds that governments that manipulate elections more heavily are likely to put down protests more quickly, especially when incumbents claim small vote-shares.
 This is inconsistent with the regime-weakness model of election manipulation, but consistent with a world in which election manipulation indicates sources of regime resilience.
@@
 As a result, governments that build and deploy an electoral manipulation apparatus are simultaneously inoculating themselves against the expected cost of protest.
+
+It is likely that the theoretical mechanisms described above weaken as regimes become more liberal-democratic; in liberal democracies, patronage consolidation should decline and rule of law should increase by definition.
+These changes should make it more difficult for incumbents to manipulate elections, irrespective of protest risk.
+At the same time, the more liberal-democratic a state becomes, the more alternatives there are to protest as a means to check the executive.
+More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest---the electoral authoritarian regimes---where we should expect the deterrence logic to operate under a regime-weakness model.
+Finally, as a matter of measurement, protest duration is likely to be an increasingly imperfect measure of incumbent strength in more liberal-democratic settings, where protest is more commonplace and less immediately threatening to the regime.
 
@@
 
-It is likely that the theoretical mechanisms described above weaken as regimes become more liberal-democratic; in liberal democracies, patronage consolidation should decline and rule of law should increase by definition.
-These changes should make it more difficult for incumbents to manipulate elections, irrespective of protest risk.
-At the same time, the more liberal-democratic a state becomes, the more alternatives there are to protest as a means to check the executive.
-More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest---the electoral authoritarian regimes---where we should expect the deterrence logic to operate under a regime-weakness model.
-Finally, as a matter of measurement, protest duration is likely to be an increasingly imperfect measure of incumbent strength in more liberal-democratic settings, where protest is more commonplace and less immediately threatening to the regime.
-The case selection strategy here, excluding electoral and liberal democracies, is thus one that both fits the theory and makes for a stronger test of the regime-weakness model.
-Nevertheless, an expanded sample that includes electoral democracies provides similar results in the appendix.
-In total, there are 659 elections included in the dataset, across 100 countries.
+The case selection strategy here, excluding electoral and liberal democracies, is one that both fits the theory and makes for a stronger test of the regime-weakness model.
+Nevertheless, an expanded sample that includes a set of electoral democracies provides similar results in the appendix.
+In total, there are 659 elections included in the dataset, across 100 countries, with 211 experiencing post-election protest.
 The date range is limited by the availability of data on protest from ECAV, which ranges from 1989 to 2012.
-
-For models of protest-wave duration, as discussed in more detail below, I include only observations where a positive number of protests occurred; there are 211 such cases.
 
@@
 I only include events labeled as protests, occupations, or blockades, excluding acts like shootings, bombings, and coups.
-If any post-election protests are recorded in the ECAV data for a particular election, the binary variable *protest initiation* is marked as 1.
-ECAV data is in event-day format; to capture *protest duration* I take the number of days between the last protest recorded in ECAV and the first, for each election.^[The duration of a protest wave with only one protest is taken to be one day.] For the survival models, protest movements that were ultimately defeated by the incumbent are considered 'dead' after the last day of protest.
-Protest movements that prevailed in unseating the incumbent are included, and are marked as 'alive.'
-
-The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0) %>% nrow()` of the `r nelda2.sub %>% nrow()` elections, no protests are recorded.
-Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest duration.
-As the figure shows, while election fraud is common among this set of regimes, protests are rare, and, when they do occur, are usually short in duration.
-These distributions---extensive election fraud combined with large numbers of non-protest---provides some early indication that the regime-weakness model has limited support.
-Still, approximately one-third of elections experienced some protest.
-Protest wave duration is modeled using Cox proportional hazards models, which have the following formula:
+ECAV data is in event-day format; to capture *protest duration* I take the number of days between the last protest recorded in ECAV and the first, for each election.^[The duration of a protest wave with only one protest is taken to be one day.] Protest wave duration is modeled using Cox proportional hazards models, which have the following formula:
 
@@
 The formula models the risk that a protest movement will fail at time $t$ given covariates $X$ is a function of the baseline hazard function, the degree of electoral fraud ($X_f$), the incumbent's vote-share ($X_v$) and a set of control variables.
+For the survival models, protest movements that were ultimately defeated by the incumbent are considered 'dead' after the last day of protest.
+Protest movements that prevailed in unseating the incumbent are included, and are marked as 'alive.'
 
@@
 Focusing on duration partially avoids this risk.
-Once protests begin (whether instrumental or multi-purpose), the die is cast.
-Authoritarian incumbents have strong incentives to defeat a public challenge to their legitimacy, while committed opponents have similar incentives to stand firm.
-Both should be expected to put significant effort into defeating / maintaining the protest movements, as a general rule, and the process by which the challenge is settled helps reveal previously hidden information about each side.
-Whether the incumbent manipulated 'too much' or 'too little' is now beside the point, and it is possible to observe whether incumbents who produce extensive election fraud are stronger or weaker than those that do not.
+Once protests begin (whether instrumental or multi-purpose), authoritarian incumbents have strong incentives to defeat a public challenge to their legitimacy, while committed opponents have similar incentives to stand firm.
+Both should be expected to put significant effort into defeating / maintaining the protest movements, as a general rule, and the process by which the challenge is settled helps reveal previously hidden information about each side; it is possible to observe whether incumbents who produce extensive election fraud are stronger or weaker than those that do not.
 
 While strategic endogeneity is less of a concern for this research design, selection bias and confounding variables may still be.
-For instance, it could be the case that incumbents facing well-organized opposition-oriented civil society may manipulate less; at the same time, such well-organized opposition may be able to sustain protest movements longer.
+For instance, it could be the case that incumbents who face well-organized opposition-oriented civil society may manipulate less; at the same time, such well-organized opposition may be able to sustain protest movements longer.
 A similar logic might imply that incumbents should manipulate less when economic grievances are high (for fear of sparking protest), and that economic troubles might make protests more durable.
 Either case could produce the relationship expected in H2, but could be considered spurious.
-To reduce this risk, I include several potential controls for opposition strength, and for grievance.
-Controlling for these variables helps isolate the relationship between manipulation severity and protest duration, holding opposition strength and grievances constant.
-
-As such, the results should be considered causally identified only to the extent that unknown or unobservable variables are not confounding the relationship between fraud and protest duration.
+I take several steps to reduce this risk.
+
+First, I control for measures of opposition strength and grievances, to help isolate the relationship between manipulation severity and protest duration.
+Still, it is possible that unknown or unobservable variables may confound the relationship between fraud and protest duration.
 While this problem is difficult to surmount with observational data, this research design offers three defenses.
-First, I control for known possible confounders, including the occurrence of a successful electoral protest in a nearby country (which might frighten incumbents into manipulating less and inspire opponents to hold the line longer).
-Second, I present two analyses in the appendix, which show that increased risk of protest leads incumbents to *intensify* their manipulation efforts, rather than diminish them.
-Thus a negative relationship between fraud and protest duration should indicate the direct hypothesized effect, rather than one that is driven by relatively weaker incumbents avoiding manipulation.
+Next, I present two analyses in the appendix, which show that increased risk of protest leads incumbents to *intensify* their manipulation efforts, rather than diminish them.
+Thus a negative relationship between fraud and protest duration should indicate the direct hypothesized effect, rather than one that is driven by relatively weaker incumbents selecting out of manipulation.
 Third, and relatedly, the research design pits two incompatible hypotheses against one another, rather than a single hypothesis against the null.
@@
 
-```{r binarytab, echo=FALSE, message=FALSE, warning=FALSE}
+```{r binarytab, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
 count_table <- table(cut(nelda2.sub$winner.share.upd, breaks = c(0, .45, .55, 1), 
@@
 
-As a plausibility check, I first examine patterns of protest initiation.
+<!-- As a plausibility check, I first examine patterns of protest initiation.
 The regime-weakness model expects initiation to increase with fraud, holding vote-share constant; while the regime-strength model expects a less consistent pattern.
@@
 
+-->
+
 ```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
@@
 This measure of election integrity should thus serve as a strong test of the protest-oriented model.
-It also matches the conception of election manipulation used in much of the relevant formal literature.^[For example, election manipulation is treated as post-hoc adjustment of the results in Little et al [-@Little2015], Little [-@Little2012], Magaloni [-@magaloni2010], and Fearon [-@fearon_self-enforcing_2011], among others.] An alternative measure is used as a robustness check in the appendix.
+It also matches the conception of election manipulation used in much of the relevant formal literature.^[For example, election manipulation is treated as post-hoc adjustment of the results in Little et al [-@Little2015], Little [-@Little2012], Magaloni [-@magaloni2010], and Fearon [-@fearon_self-enforcing_2011], among others.] Two alternative measures for manipulation are used as robustness checks in the appendix.
+
+Lastly, the *incumbent party's vote-share* is also taken from the V-Dem dataset.^[This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia, where links to official sources were available, and Psephos.] The original V-Dem variable indicates the *winner's* vote-share.
+However, I exclude all cases where the incumbent lost the election outright.
+This limits the observations to those that fit the population of interest: protest following an election in which the incumbent claimed victory, possibly through reliance on election fraud; it also ensures that the vote-share variable captures the incumbent's position.
+
+Excluding cases where incumbents stepped down after losing an election avoids introducing measurement error in the dependent variable, due to a quirk of the ECAV dataset, which codes the 'side' of the protest (pro- or anti-government) as missing during the period between an election and the swearing in of a new government in cases where the incumbent loses.
+Including these cases thus runs the risk of conflating pro- and anti-incumbent protest, along with cases where protesters goals are genuinely unknown.
+However, excluding them also influences the interpretation of the results from these models.
+Threats to inference from this section are evaluated in more detail in the discussion section.
 
@@
 
-First, I include the *number of pre-election protests* counted in the ECAV dataset, which is likely to be an important possible confounder; more numerous pre-election protest may affect the level of election fraud (as the deterrence logic predicts) as well as being directly tied to the likelihood of post-election protest.
-Next, the models of protest duration include a measure of the *median size of protest*, calculated using ECAV data.^[ECAV data codes protest size as a categorical variable, ranging from 1 (< 10 participants) to 5 (> 10,000).
+First, I include the *number of pre-election protests* counted in the ECAV dataset, which is likely to be an important possible confounder; more numerous pre-election protest may affect the level of election fraud (as the deterrence logic predicts) as well as being directly tied to the likelihood of post-election protests.
+Next, I control for the *median size of protest*, calculated using ECAV data.^[ECAV data codes protest size as a categorical variable, ranging from 1 (< 10 participants) to 5 (> 10,000).
 I take the median size category of all the events in the protest wave as a measure of average protest size.] This is important as a control for opposition strength, as well as for the possibility that incumbents might tolerate smaller protests.
@@
 
-Since higher repressiveness may weaken or intensify opposition [@chenowethStateRepressionNonviolent2017], I include a lagged measure of physical integrity index from V-Dem (v2x_clphy), and its square term.
+Since higher repressiveness may weaken *or* intensify opposition [@chenowethStateRepressionNonviolent2017], I include a lagged measure of *physical integrity* index from V-Dem (v2x_clphy), and its square term.
 Indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), *legislative constraints* on the executive (V-Dem variable 'v2xlg_legcon', and the availability of *alternative sources of information* ('v2x_freexp_altinf' from V-Dem) are also employed as straightforward indicators of opposition strength.
@@
 In addition, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby due to opposition learning [@Bunce2010;@Beissinger_2007].
-Such a dynamic could plausibly increase the likelihood and severity of protest in a nearby country, while also reducing the incumbent's use of fraud.
+Such a dynamic could plausibly increase the severity of protest in a nearby country, while also reducing the incumbent's use of fraud.
 I control for this using each country's geopolitical region as coded in Teorell et al [-@qog_teorell], in conjunction with binary variables indicating a successful protest movement from NELDA.
@@
 
-Lastly, the *incumbent party's vote-share* is also taken from the V-Dem dataset.^[This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia, where links to official sources were available, and Psephos.] The original V-Dem variable indicates the *winner's* vote-share.
-However, I exclude all cases where the incumbent lost the election outright.
-This limits the observations to those that fit the population of interest: protest following an election in which the incumbent claimed victory, possibly through reliance on election fraud; it also ensures that the vote-share variable captures the incumbent's position.
-
-Excluding cases where incumbents stepped down after losing an election avoids introducing measurement error in the dependent variable, due to a quirk of the ECAV dataset, which codes the 'side' of the protest (pro- or anti-government) as missing during the period between an election and the swearing in of a new government in cases where the incumbent loses.
-Including these cases thus runs the risk of conflating pro- and anti-incumbent protest, along with cases where protesters goals are genuinely unknown.
-However, excluding them also influences the interpretation of the results from these models.
-Threats to inference from this section are evaluated in more detail in the discussion section.
-
 Summary statistics for all of these variables are shown in the appendix.
@@
 # Results
+
+The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0) %>% nrow()` of the `r nelda2.sub %>% nrow()` elections, no protests are recorded.
+Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest duration.
+As the figure shows, while election fraud is common among this set of regimes, protests are rare, and, when they do occur, are usually short in duration.
+These distributions---extensive election fraud combined with large numbers of non-protest---provides some early indication that the regime-weakness model has limited support.
+Still, approximately one-third of elections experienced some protest.
 
@@
 
-cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose != "1" &  n.events.post > 0))
+cox1 <- coxph(Surv(d.events.num, surv_status) ~  lji.lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose != "1" &  n.events.post > 0))
 summary(cox1)
 
-cox1.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0))
+cox1.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~ lji.lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0))
 summary(cox1.cen)
@@
 
-cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag +  + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
+cox2 <- coxph(Surv(d.events.num, surv_status) ~ lji.lag + legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag +  + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
 summary(cox2)
@@
 
-cox2.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
+cox2.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~ lji.lag +  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
 summary(cox2.cen)
 
-cox3.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag +  v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post + median.participants, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
+cox3.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~ lji.lag +  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag +  v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post + median.participants, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
 summary(cox3.cen)

### 2025-11-16 Output/authoritarian_only_pdf_pop_rev.Rmd
Final edits to results section

@@
 Recall that these models include data only from electoral protests that *did not succeed* in ousting incumbents.
-Extensions of these models that
-
-Models 4 and 6 do not include interactions, while the remaining models interact *intentional irregularities* with incumbent vote-share.
+
+Model 1 is a baseline with minimal controls.
+Models 2 and 4 do not include interactions, while the remaining models interact *intentional irregularities* with incumbent vote-share.
 One threat to inference in this design derives from the 90-day post-election window considered by ECAV.
-It is possible that some protest movements considered 'dead' in the baseline models 4 and 5 are, in fact, 'alive,' if subsequent protests occurred after 90 days and are unobserved.
-To address this, observations are assumed to be right-censored if a protest occurred within three-weeks of the 90-day cutoff in Models 6, 7, and 8.
-This amounts to 15 of the 156 observations in the sample.
-Model 8 is the most restrictive model, which includes right censoring as well as a control variable for the median number of participants during the protest wave.
+It is possible that some protest movements considered 'dead' in Models 1, 2, and 3 are, in fact, 'alive,' if subsequent protests occurred after 90 days and are unobserved.
+To address this, observations are assumed to be right-censored if a protest occurred within three-weeks of the 90-day cutoff in Models 4, 5, and 6.
+This amounts to 15 of the observations in the sample.
+Model 6 is the most restrictive model, which includes right censoring as well as a control variable for the median number of participants during the protest wave.
 There is considerable missingness in this variable.
-Lastly, in order to intrepret these findings, it is important to recall that these models include only cases where incumbents claimed victory and electoral protest occurred.
-
-The coefficient for *election fraud* is positive and significant in Model 4, indicating that more fraudulent elections are associated with less resilient protest movements.
+
+The coefficient for *election fraud* is positive and significant in Models 1 and 2, indicating that more fraudulent elections are associated with less resilient protest movements.
 Exponentiation of the coefficient for election fraud gives a hazard ratio of 1.3, indicating that for each one-unit increase in fraud, a protest movement becomes thirty percent more likely to end at any given time.
-This supports H2, suggesting that incumbents with the capacity to manipulate extensively are also able to draw on their resources to curtail post-election protest.
-
-Further investigation of the relationship between fraud and protest duration yields additional insights.
-In Model 5, incumbent vote-share is interacted with *election fraud*.
-Figure \@ref(fig:margin-dur-fig) shows the predicted hazard ratios As shown in the figure, for all incumbent vote-shares near and below fifty percent, large-scale fraud significantly increases the likelihood that a protest movement will end.
-This result also supports the regime-strength model, and shows that only as vote-share increases above fifty percent---creating a stronger signal of popularity---does a government in a low-fraud election have the same likelihood of defeating a protest movement at a given time.
-In effect, either high genuine popularity *or* the ability to produce high levels of fraud enables governments to more quickly end protests.
+This is inconsistent with H1, suggesting that increasing election fraud is not indicative of regime weakness.
+
+In Model 3, incumbent vote-share is interacted with *election fraud*.
+Figure \@ref(fig:margin-dur-fig) shows the predicted hazard ratios.
+As shown in the figure, for all incumbent vote-shares near and below fifty percent, large-scale fraud significantly increases the likelihood that a protest movement will end.
+This result is in line with H2 and the regime-strength model, and shows that only as vote-share increases above fifty percent---creating a stronger signal of popularity---does a government in a low-fraud election have the same likelihood of defeating a protest movement at a given time.
 
 The results of the baseline models are consistent with those models which incorporate censoring.
-The coefficient for *election fraud* retains the same size, sign, and level of significance as in Model 4.
-The interaction effect in Model 7 increases in size, while retaining its direction and statistical significance.
+The coefficient for *election fraud* retains the same size, sign, and level of significance as in Model 2.
+The interaction effect in Model 5 increases in size, while retaining its direction and statistical significance.
 This suggests that the main models are not drawing overly incorrect inferences through censoring of longer-lasting protest movements.
-Model 8, adds the median number of participants at events in the protest wave as a control for the potential perceived threat of the protest movement by the incumbent.
+Model 6, adds the median number of participants at events in the protest wave as a control for the potential perceived threat of the protest movement by the incumbent.
 Including this control variable reduces the sample size due to missingness, but the interaction coefficient retains its direction and statistical significance.
@@
 For much of the range of the data, they remain roughly twice as likely to survive at any given date than protest movements in medium- or high-fraud cases.
-This is evidence in favor of the idea that high levels of fraud is an indication of regime resilience, and contradicts the claim that fraud indicates weakness or galvanizes political opposition, as it is the low-fraud cases where governments struggle to put an end to protest movements on average.
+This is evidence in favor of the idea that high levels of fraud indicate stronger regimes.
 

### 2025-11-16 Output/authoritarian_only_pdf_pop_rev.Rmd
Final edits to discussion

@@
 Further, the relationship holds despite controlling for indicators of other possible explanations of the relationship between low fraud and long-lived protest, including civil society strength, media openness, regional electoral protest waves, number of protests and median protest size.
-This is consistent with Hypothesis 2, and supportive of what I have called the 'regime-strength' model of electoral manipulation.
-
-This finding supports and extends the line of research initiated by Simpser [@simpser2013].
-Understanding election manipulation as indicative of effective sources of regime strength comports with a growing literature on the principal-agent dynamics of manipulation [@Harvey2019;@rundlett2016deliver;@Harvey2022] and the importance of patronage resources and electoral processes in sustaining authoritarian regimes [@hale_patronal_2014].
-This study offers empirical evidence that authoritarian governments who manipulate more are stronger than those who manipulate less (all else equal).
-In addition to offering a counter-argument to regime-weakness models, it also illustrates how the mechanisms which underpin Simpser's model of manipulation-as-strength---coordination effects among pro-regime agents and costly signaling of incumbent resources---can function even when such signals of strength have failed to 'deter' protest.
+While this study does not purport to show perfect causal identification, these robust correlations are presented as incompatible with the regime-weakness framework and supportive of the regime-strength model of manipulation.
 
@@
 For the regime-weakness model, it is potentially puzzling that most manipulated elections go un-protested.
-Regime-weakness models may explain this pattern by arguing that, most of the time, incumbents produce the 'right' amount of manipulation to avoid protest in equilibrium.
-Such careful calibration is difficult to ensure, however, when incumbents must rely on thousands of low-level agents to produce the 'right' amount of manipulation; these agents may have collective incentives to shirk and underproduce, or to demonstrate their loyalty by over-producing.
-[ADD CITES] The results of this study add an empirical challenge, since it demonstrates that when the deterrence equilibrium fails (i.e.
+Regime-weakness models may explain this pattern by arguing that, most of the time, incumbents produce the 'right' amount of manipulation to avoid protest in equilibrium [@Little2012, e.g.].
+Such careful calibration is difficult to ensure, however, when incumbents must rely on thousands of low-level agents to produce the 'right' amount of manipulation; these agents may have collective incentives to shirk and underproduce [@rundlett2016deliver], or to demonstrate their loyalty by over-producing [@choWhenAgentsDeliver2025].
+The results of this study add an empirical challenge, since it demonstrates that when the deterrence equilibrium fails (i.e.
 when the incumbent rigs elections and the opposition protests), higher levels of manipulation are associated with greater regime strength, not less.
 
-In rejoinder, it might reasonably be wondered why protests occur after highly manipulated elections, if manipulation is a signal of strength?
-First, in my view, this interpretation unnecessarily exports the instrumental logic of protest from the regime-weakness model to the regime-strength model.
+In rejoinder, it might reasonably be wondered why---if manipulation indicates strength---why protests occur after highly manipulated elections?
+In my view, this interpretation unnecessarily exports the instrumental logic of protest from the regime-weakness model to the regime-strength model.
 As discussed earlier in the paper, instrumental calculations about overthrowing incumbents are but one component of a long and varied literature on protest, which has identified multiple other expressive and strategic motives for protest.
@@
 It also helps us make sense of secular patterns that are puzzling if election fraud indicates regime weakness, as others have held.
-The pattern depicted in Figure \@ref(fig:elex-hist) for example, which shows high levels of manipulation and low levels of protest is sensible in a world where the more powerful and durable incumbents produce greater levels of manipulation, and where incumbents who struggle to produce high levels of manipulation are ones who face resource scarcity, institutional barriers to impunity for their agents, and defections among erstwhile allies.
+The pattern depicted in Figure \@ref(fig:elex-hist) for example, which shows high levels of manipulation and low levels of protest, is sensible in a world where the more powerful and durable incumbents produce greater levels of manipulation, and where incumbents who struggle to produce high levels of manipulation are ones who face resource scarcity, institutional barriers to impunity for their agents, and defections among erstwhile allies.
 It also helps explain why some incumbents manufacture their way to highly implausible margins of victory, reasoning that a demonstration of their extra-electoral power may suggest to allies and bystanders that the incumbent may survive protest despite a low natural popularity.
@@
 What does all this tell us?
-First, these results may inform how researchers think about the causes of election manipulation.
-They suggest that the degree to which the opposition's capacity to organize street protest (and related factors) can limit election fraud have been overestimated.
+First, these findings support and extend the line of research initiated by Simpser [@simpser2013].
+Understanding election manipulation as indicative of effective sources of regime strength comports with a growing literature on the principal-agent dynamics of manipulation [@Harvey2019;@rundlett2016deliver;@Harvey2022] and the importance of patronage resources and electoral processes in sustaining authoritarian regimes [@hale_patronal_2014].
+This study offers empirical evidence that authoritarian governments who manipulate more are stronger than those who manipulate less (all else equal).
+It further illustrates how the mechanisms which underpin Simpser's model of manipulation-as-strength---coordination effects among pro-regime agents and costly signaling of incumbent resources---can function even when such signals of strength have failed to 'deter' protest.
+
+Second, these results may inform how researchers think about the causes of election manipulation.
+They suggest that the degree to which the opposition's capacity to organize street protest can limit election fraud has been overestimated.
 Consequently, theories of electoral manipulation that emphasize alternatives---such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]---should be further developed, and may yield new policy interventions.
@@
 By showing that the infrastructure of fraud can actually reduce the cost of post-election protest to the regime, this result poses a challenge to theory of democratization by election [@Schedler2013;@Morgenbesser2019;@edgellWhenWhereElections2018].
-If the capacity to engage in widespread manipulation makes it easier to both tolerate the existence of an opposition and to suppress it during periods of contestation [@dahlPolyarchyParticipationOpposition1971], electoral authoritarianism is something of an institutional cul-de-sac: the capacity to generate severe election fraud makes it appealing for closed authoritarian regimes to introduce elections, while the manageable risk of protest makes it unappealing to further democratize.
+If the capacity to engage in widespread manipulation makes it easier to both tolerate the existence of an opposition and to suppress it during periods of contestation [@dahlPolyarchyParticipationOpposition1971], electoral authoritarianism may be something of an institutional cul-de-sac: the capacity to generate severe election fraud makes it appealing for closed authoritarian regimes to introduce elections, while the manageable risk of protest makes it unappealing to further democratize.
 
@@
 First, even if protest risk fails to serve as a deterrent in non-democracies, it may still do so in more democratic settings.
-In non-democracies, protest initiation is most likely when fraud is very high and incumbent vote-shares very low.
 The likelihood of protest initiation in a democracy may well be higher, triggering at lower levels of fraud and higher incumbent vote-shares, since opposition actors are likely to be better resourced in a liberal democracy (and the government more constrained).
@@
 
-In sum, the results indicate that we should be cautious when building and interpreting theories that emphasize the risk of mass protest as a trip-wire to be avoided in non-democracies, and add empirical evidence to Przeworski's [-@przeworskiFormalModelsAuthoritarian2022] critique of such models.
-The rarity of post-election protest, especially successful protest, suggests that such events are highly contingent; they hinge on a multitude of factors that go beyond the regime's efforts to bias election outcomes.
-These findings indicate that, while incumbents do seek to contain protest, more effective manipulators are also more capable of defeating protest movements.
-
-These results are consistent with different specifications presented in an online appendix.
-In particular, the main results are supported when alternative measures of election integrity are used in place of the measure of election fraud.
-They are also supported when electoral democracies are included in the dataset.
-The relationship between fraud levels and incumbent defeat is also tested (along with incumbent ouster by protest).
-Additional robustness checks include models with data from election observer reports, with country fixed-effects, alternative specifications for selected control variables.
+These results remain consistent under a several alternate specifications presented in an online appendix.
+The results are supported using two alternative measures of election manipulation, one that covers a broader array of electoral violations, and another that is drawn directly from election observer reports.
+Additional tests show that incumbents intensify their efforts to manipulate when protest risk is high, that higher capacity for manipulation makes incumbent election loss less likely, and that the small number of successful protests occur across the spectrum of manipulation.
+This helps rule out an interpretation of the main results in which weaker regimes choose not to manipulate and thus avoid protest.
+The results also remain consistent when the selection criteria are relaxed, and a set of electoral democracies are included in the dataset.
 

### 2025-11-17 Output/authoritarian_only_pdf_pop_rev.Rmd
Final edits to conclusion

@@
 There are competing views on the relationship between election fraud and protest.
-In one family of models, which I have called the regime-weakness model, incumbent's may hold cleaner elections when the risk of post-election protest is higher.
+In one family of models, which I have called the regime-weakness model, incumbents may hold cleaner elections when the risk of post-election protest is higher.
 Other models contend that election fraud can indicate incumbent strength.
@@
 
-This paper attempts to test the assumptions of these two models by investigating the relationship between election fraud and two protest variables: initiation and duration.
-Specifically, both the regime-weakness and regime-strength models may make the same predictions regarding protest onset; their predictions differ with regard to protest duration.
-Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 659 elections in electoral authoritarian regimes to show that protest occurrence is most likely when incumbent vote-shares are very low and election fraud is high, as both models predict.
-However, and most importantly, incumbents that engage in larger-scale election fraud are able to more quickly defeat protest movements.
+This paper tests the assumptions of these two models by investigating the relationship between election fraud and the duration of protest.
+Though both models may make the same predictions regarding protest onset; their predictions differ with regard to protest duration.
+Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 659 elections in electoral authoritarian regimes to show that incumbents that engage in larger-scale election fraud are able to more quickly defeat protest movements.
 
@@
 # References
-
-# TEst stuff
-
-```{r include=FALSE, eval=FALSE}
-library(lavaan)
-library(dplyr)
-
-# Prepare country-year level data
-nelda2.sub <- nelda2.sub %>%
-  mutate(
-    # For country-years with protests: inverse of mean duration
-    # Higher values = shorter duration = stronger regime
-    protest_duration_inverse = if_else(
-      any.ecav.protest == 1,
-      1 / d.events.num,  # or -log(mean_duration)
-      NA_real_
-    )
-  )
-
-nelda2.sub <- nelda2.sub %>%
-  mutate(taxstate.num = as.numeric(taxstate))
-nelda2.sub <- nelda2.sub %>%
-  mutate(rentierstate.num = as.numeric(rentierstate))
-nelda2.sub <- nelda2.sub %>%
-  mutate(unscheduled_election.num = as.numeric(unscheduled_election))
-
-# MIMIC model
-mimic_spec <- '
-  # Latent variable
-  regime_strength =~  gdpgro.1lag +v2elirreg.inv +   physinteg.1lag + urban.pct + rentierstate.num + taxstate.num +  unscheduled_election.num + inc.termlimit.num + winner.share.upd
-
-  # Structural part: causes of regime strength
-  protest_duration_inverse ~ regime_strength  + winner.share.upd
-
-  # Allow correlation between indicators
- # protest_duration_inverse ~~ any.ecav.protest
- v2elirreg.inv ~~ physinteg.1lag 
- v2elirreg.inv ~~ urban.pct
- v2elirreg.inv ~~  inc.termlimit.num
-'
-
-# Fit with FIML to handle missing protest_duration_inverse
-fit <- sem(mimic_spec, 
-           data = nelda2.sub %>% filter(d.events.num >0 & inc.lose == 0),
-           missing = "fiml")  # Full Information Maximum Likelihood
-
-summary(fit, fit.measures = TRUE, standardized = TRUE)
-
-# Extract latent scores
-regime_scores <- lavPredict(fit, type = "lv")
-data_cy$regime_strength_estimated <- regime_scores[, 1]
-```
-
-```{r eval=FALSE}
-# Step 1: Create a composite measure of regime strength from protests
-# (for country-years with protests)
-
-nelda2.sub <- nelda2.sub %>%
-  mutate(
-    # Composite indicator (you can refine this)
-    strength_indicator = case_when(
-      any.ecav.protest == 0 ~ NA_real_,  # Missing for no protests
-      any.ecav.protest == 1 ~ -log(d.events.num) # Lower duration = higher strength
-    )
-  )
-
-# Step 2: SEM with measurement error correction
-sem_spec <- '
-  # Latent variable with single indicator
-  # (need to fix measurement error variance)
-  regime_strength =~ 1*strength_indicator
-
-  # Estimate measurement error (you can set this based on reliability)
-  strength_indicator ~~ error_var*strength_indicator
-
-  # Structural model
-  regime_strength ~  v2elirreg.inv + gdpgro.1lag +   physinteg.1lag
-'
-
-# Fit model
-fit2 <- sem(sem_spec, 
-            data = nelda2.sub,
-            missing = "fiml")
-
-summary(fit2, standardized = TRUE)
-```
-```{r eval=FALSE}
-# This is cleaner and avoids identification issues
-path_model <- '
-  # Protest occurrence predicted by covariates
-  any.ecav.protest ~ v2elirreg.inv + gdpgro.1lag +   physinteg.1lag
-
-  # Protest duration (inverse) predicted by same covariates
-  # Only estimated for observations with protests
-  protest_duration_inverse ~ v2elirreg.inv + gdpgro.1lag +   physinteg.1lag
-
-  # Allow correlation between equations
-  any.ecav.protest ~~ protest_duration_inverse
-'
-
-fit_path <- sem(path_model, 
-                data = nelda2.sub,
-                missing = "fiml")
-
-summary(fit_path, standardized = TRUE, fit.measures = TRUE)
-```
-
-```{r eval=FALSE}
-# Use only country-years WITH protests
-data_protests_only <- nelda2.sub %>%
-  filter(any.ecav.protest == 1)
-
-# Simple regression path model
-protest_model <- '
-  # Duration as function of covariates
-  protest_duration_inverse ~ legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv
-
-  # If you have multiple protests per country-year, 
-  # you might want to include number of protests as control
-  protest_duration_inverse ~ n.events.post
-'
-
-fit_protests <- sem(protest_model,
-                    data = data_protests_only)
-
-summary(fit_protests, standardized = TRUE, rsquare = TRUE)
-
-# This tells you:
-# 1. Which covariates predict shorter duration (stronger regimes)
-# 2. How much variance in duration is explained
-# 3. Specific effect of election fraud on regime strength
-```
-
-Structural equation modelling doesn't seem the way to go
-
-```{r eval=FALSE}
-
-df <- nelda2.sub %>% filter(inc.replaced.prot == 0) %>% dplyr::select(legcon.1lag, altinf.1lag,  osorg.1lag, gdpgro.1lag, inc.termlimit.num,  physinteg.1lag, physintlag.sq, urban.pct, rentierstate.num, taxstate.num ,  unscheduled_election.num, inc.termlimit.num, v2elirreg.inv, NeldaID)
-
-df <- df %>% filter(complete.cases(.))
-
-CFA.model <- ' regime_strength  =~  v2elirreg.inv + physinteg.1lag + physintlag.sq + altinf.1lag + osorg.1lag + legcon.1lag   + gdpgro.1lag + inc.termlimit.num  +  urban.pct + rentierstate.num + taxstate.num +  unscheduled_election.num + inc.termlimit.num'
-
-fit <- cfa(CFA.model, data = df, std.lv=TRUE)
-summary(fit)
-
-regime_scores <- lavPredict(fit)
-
-df <- df %>% mutate(rgm_st = regime_scores[,1])
-
-df <- df %>% left_join(nelda2.sub, by = "NeldaID")
-
-init_probit <- glm(any.ecav.protest ~ rgm_st*winner.share.upd + presidential, binomial(link = "probit"), data = df)
-df_probit <- init_probit$data
-df_probit <- df_probit %>%  dplyr::select(legcon.1lag.x, altinf.1lag.x,  osorg.1lag.x, gdpgro.1lag.x, inc.termlimit.num.x,  physinteg.1lag.x, physintlag.sq.x, urban.pct.x, rentierstate.num.x, taxstate.num.x ,  unscheduled_election.num.x, inc.termlimit.num.x, v2elirreg.inv.x, NeldaID, winner.share.upd, d.events.num, rgm_st)
-df_probit <- df_probit %>% filter(complete.cases(.))
-
-df_probit$IMR <- sampleSelection::invMillsRatio( init_probit )$IMR1
-
-library(survival)
-library(ggsurvfit)
-
-df <- df %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
-df_probit <- df_probit %>% mutate(surv_status = ifelse(inc.replaced.prot == 1, 0, surv_status))
-df_probit <- df_probit %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
-                                                                       ifelse(d.events.num >= 69, 0, 1)))
-
-cox1 <- coxph(Surv(d.events.num, surv_status) ~  rgm_st*winner.share.upd  +  median.participants + regional_success_lag, data = df)
-summary(cox1)
-
-## Unclear what's going on here. But the general idea is that election fraud negatively loads onto the 'regime strength' variable, which in turn has no apparent effect on duration.
-```
-
-```{=latex}
-\begin{table}[]
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-\multicolumn{1}{|r|}{Protest is assumed to be:} &
-  \multicolumn{1}{c|}{Instrumental protest} &
-   &
-  \multicolumn{1}{c|}{Multipurpose protest} &
-   \\ \hline
-\multicolumn{1}{|r|}{Incumbent's genuine popularity:} & Low & High & Low & High \\ \hline
-Manipulation...                                       &     &      &     &      \\ \hline
-...obscures strength or indicates weakness &
-  (I) High and increasing &
-  (I): Low and increasing &
-  (I): High and increasing &
-  (I): Low and increasing \\ \hline
-...reflects strength &
-  (I): High and decreasing &
-  (I): Low and decreasing &
-  (I): High and increasing &
-  (I): Low and no effect \\ \hline
-\end{tabular}
-\caption{}
-\label{tab:my-table}
-\end{table}
-```
